### 2022-10-14 bibliography/succ.tex
Uploading to have some file on the git

\documentclass[times,envcountsame]{llncs}
\usepackage{a4wide}
\usepackage{times}
\usepackage{amsmath,amssymb,eufrak,theorem}%mnsymbol}
\usepackage{enumerate}
\usepackage{complexity}
\usepackage{gastex}

\usepackage{xspace,tabularx,multicol}

%\usepackage[thmmarks,standard,thref]{ntheorem}

\iffalse

\newtheorem{theorem}{{\bf Theorem}}[section]
\newtheorem{corollary}[theorem]{{\bf Corollary}}
\newtheorem{definition}[theorem]{{\bf Definition}}
\newtheorem{example}[theorem]{{\bf Example}}
\newtheorem{lemma}[theorem]{{\bf Lemma}}
\newtheorem{proposition}[theorem]{{\bf Proposition}}
\newtheorem{remark}[theorem]{{\bf Remark}}
\newtheorem{fact}[theorem]{Fact}

\fi

\newtheorem{fact}[theorem]{Fact}

\usepackage{stmaryrd,rotating,mathrsfs}

%\usepackage[active]{srcltx}

\iffalse
\newcommand{\problemx}[3]{
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf QUESTION: }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup}

\newcommand{\problemy}[3]{
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf OUTPUT: }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup}

\fi

\newcommand{\problemx}[3]{
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf QUESTION:\ }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup}

\newcommand{\problemy}[3]{
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf OUTPUT:\ }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup}

\renewcommand{\mod}{\text{mod }}
\newcommand{\modulo}{\mod}

%\renewcommand{\vec}{}

\def\AC{{\mathrm{AC}}}
\def\AG{{\mathsf{AG}}}
\def\AND{{\mathrm{AND}}}
\def\BIN{{\mathrm{BIN}}}
\def\CRR{{\mathrm{CRR}}}
\def\CTL{{\text{CTL}}}
\def\div{{\mathrm{div}}}
\def\dom{{\mathrm{dom}}}
\def\EF{{\mathsf{EF}}}
\def\AX{{\mathsf{AX}}}
\def\EXPTIME{{\mathsf{EXPTIME}}}
\def\inp{{\mathrm{in}}}
\def\leaf{{\mathrm{leaf}}}
\def\MAJ{{\mathrm{MAJ}}}
\def\NC{{\mathrm{NC}}}
\def\Omc{{\mathbb{O}}}
\def\OR{{\mathrm{OR}}}
\def\out{{\mathrm{out}}}
\def\Pmc{{\mathcal{P}}}
\def\PSPACE{{\mathsf{PSPACE}}}
\def\TC{{\mathrm{TC}}}
\def\U{{\mathsf{U}}}
\def\WU{{\mathsf{WU}}}
\def\X{{\mathsf{X}}}
\def\EX{{\mathsf{EX}}}
\def\E{{\mathsf{E}}}

\newcommand{\prop}{\rho}

\renewcommand{\L}{\mathsf{LOGSPACE}}

%% \usepackage[thmmarks]{ntheorem}
%% %%% Theorem environments
%% \newtheorem{lemma}{Lemma}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{proposition}{Proposition}
%% \newtheorem{corollary}{Corollary}

%% \theoremstyle{definition}
%% \theoremsymbol{\ensuremath{\lozenge}}
%% \theorembodyfont{\normalfont}
%% \newtheorem{definition}{Definition}

%% \theoremstyle{nonumverplain}
%% \theoremheaderfont{\normalfont\bf}
%% \theorembodyfont{\normalfont}
%% \theoremseparator{.}
%% \theoremsymbol{\ensuremath{_\blacksquare}}
%% \newtheorem{example}{Example}

%% \theoremstyle{nonumberplain}
%% \theoremheaderfont{\normalfont\itshape}
%% \theorembodyfont{\normalfont}
%% \theoremseparator{.}
%% \theoremsymbol{\ensuremath{_\Box}}
%% \newtheorem{proof}{Proof}

% Symbols
\newcommand{\naturals}{\ensuremath{\mathbf{N}}}
\newcommand{\integers}{\ensuremath{\mathbf{Z}}}

% Machines and their ingredients
\newcommand{\automaton}{\ensuremath{A}}
\newcommand{\clocations}{\ensuremath{Q}}
\newcommand{\iclocation}{\ensuremath{q_{\mathit{in}}}}
\newcommand{\fclocation}{\ensuremath{F}}
\newcommand{\transitionrel}{\ensuremath{\Delta}}
\newcommand{\transitionlabel}{\ensuremath{\lambda}}
\newcommand{\clocationlabel}{\ensuremath{\xi}}
\newcommand{\operations}{\ensuremath{Op}}
\newcommand{\operation}{\ensuremath{op}}
\newcommand{\add}[1]{\ensuremath{\mathsf{add}(#1)}}
\newcommand{\zero}{\ensuremath{\mathsf{zero}}}
\newcommand{\propositions}{\ensuremath{L}}
\newcommand{\configuration}{\ensuremath{c}}
\newcommand{\configurations}{\ensuremath{\mathcal{C}}}
\newcommand{\run}{\ensuremath{\pi}}
\newcommand{\trace}{\ensuremath{\tau}}
\newcommand{\pocatuple}{\ensuremath{\langle \clocations,
\iclocation, \fclocation, \transitionrel, \transitionlabel,
\clocationlabel \rangle}}

\newcommand{\circuit}{\ensuremath{\mathbb{C}}}

\newcommand{\tm}{\ensuremath{\mathcal{M}}}

%%% Names
\newcommand{\oca}{\textsc{OCA} }
\newcommand{\poca}{\textsc{POCA} }
\newcommand{\Poca}{\mathbb{P}}
\newcommand{\Soca}{\mathbb{S}}

%%% Problems
\newcommand{\stsat}{\textsc{Succinct 3-SAT} }
\newcommand{\tsat}{\textsc{3-SAT} }
\newcommand{\ltlpoca}{\textsc{LTL-POCA-MC} }

%%% Complexity classes
\newcommand{\nexptime}{\textsc{NExpTime} }
\newcommand{\conexptime}{\textsc{co-NExpTime} }
\newcommand{\pspace}{\textsc{PSpace} }

%%% LTL
\newcommand{\ltl}{\text{LTL}}
\newcommand{\ctl}{\text{CTL}}
%\newcommand{\until}{\ensuremath{\mathcal{~U~}}}
\newcommand{\until}{\ensuremath{\U}}
\newcommand{\eventually}{\ensuremath{\mathsf{F}}}
\newcommand{\globally}{\ensuremath{\mathsf{G}}}
\newcommand{\propalphabet}{\ensuremath{\Sigma}}
\newcommand{\pastarred}{\ensuremath{(\mathcal{P}(\propalphabet))^{\omega}}}

%%% Logic
\newcommand{\true}{\ensuremath{\mathtt{tt}}}
\newcommand{\false}{\ensuremath{\mathtt{ff}}}

%%% Control commands
\newcommand{\defemph}[1]{\emph{#1}}

%\newtheorem{fact}{Fact}

\newcommand{\LCM}{\text{LCM}}

\begin{document}

\title{Model Checking Succinct and Parametric One-Counter Automata}
\author{Stefan G\"oller\inst{1} \and
Christoph Haase\inst{2} \and
Jo\"el Ouaknine\inst{2} \and
James Worrell\inst{2}
}

\institute{
Universit\"at
Bremen, Institut f\"ur Informatik, Germany%\\
%\email{goeller@informatik.uni-bremen.de}
\and
Oxford University Computing Laboratory, UK
}

\maketitle

\begin{abstract}
  We investigate the decidability and complexity of various model
  checking problems over one-counter automata. More specifically, we
  consider \emph{succinct} one-counter automata, in which additive
  updates are encoded in binary, as well as \emph{parametric}
  one-counter automata, in which additive updates may be given as
  unspecified parameters. We fully determine the complexity of model
  checking these automata against CTL, LTL, and modal $\mu$-calculus
  specifications.
\end{abstract}

\section{Introduction}

Counter automata, which comprise a finite-state controller together
with a number of counter variables, are a fundamental and widely-studied
computational model. One of the earliest results about counter
automata, which appeared in a seminal paper of Minsky's five decades
ago, is the fact that two counters suffice to achieve Turing
completeness~\cite{Min61}.

Following Minsky's work, much research has been directed towards
studying restricted classes of counter automata and related
formalisms. Among others, we note the use of restrictions to a single
counter, on the kinds of allowable tests on the counters, on the
underlying topology of the finite controller (such as
flatness~\cite{CJ98-cav,LS08-atva}), and on the types of computations
considered (such as reversal-boundedness~\cite{ID06-tcs}). Counter
automata are also closely related to Petri nets and pushdown automata.

In Minsky's original formulation, counters were represented as integer
variables that could be incremented, decremented, or tested for
equality with zero by the finite-state controller. More recently,
driven by complexity-theoretic considerations on the one hand, and
potential applications on the other, researchers have investigated
additional primitive operations on counters, such as additive updates
encoded in binary~\cite{BHIMV06,LS08-atva} or even
in \emph{parametric} form, i.e., whose precise values depend on
parameters~\cite{BIL06-icalp,IJTW93-icalp}. We refer to such counter automata as
\emph{succinct} and \emph{parametric} respectively, the former being
viewed as a subclass of the latter. Natural applications of such
counter machines include the modeling of resource-bounded processes,
programs with lists, recursive or multi-threaded programs, and XML
query evaluation; see, e.g.,
\cite{ID06-tcs,BHIMV06,CR04}.

In most cases, investigations have centered around the decidability
and complexity of the \emph{reachability} problem, i.e., whether a
given control state can be reached starting from the initial
configuration of the counter automaton. Various instances of the
reachability problem for succinct and parametric counter automata are
examined, for example, in~\cite{DG-jlc09,HKOW10,IJTW93-icalp}.

The aim of the present paper is to study the decidability and
complexity of \emph{model checking} for succinct and parametric
one-counter automata. In view of Minsky's result, we restrict our
attention to \emph{succinct one-counter automata (SOCA)} and
\emph{parametric one-counter automata (POCA)}. On the specification
side, we focus on the three most prominent formalisms in the
literature, namely the temporal logics CTL and LTL, as well as the
modal $\mu$-calculus. For a counter automaton $\mathbb{A}$ and a
specification $\varphi$, we therefore consider the question of
deciding whether $\mathbb{A} \models \varphi$,
in case of POCA for all values of the parameters,
%\footnote{In the case of
%parametric automata, one asks whether the automaton satisfies the
%formula \emph{for all} legal values of the parameters.}
and investigate both the \emph{data} complexity (in which the formula
$\varphi$ is fixed) as well as the \emph{combined} complexity of this
problem. Our main results are summarized in Table \ref{T results}.
\begin{table}
\begin{center}
\begin{tabular}{ >{\centering\arraybackslash}m{3cm}
  >{\centering\arraybackslash}m{2cm} |
  >{\centering\arraybackslash}m{3.3cm} |
> {\centering\arraybackslash}m{3.3cm} |}\\
  &  &{\bf SOCA}&{\bf POCA}\\ \hline
{\raisebox{-2ex}[0ex]{{\bf CTL, $\mu$-calculus}}}&   {\bf data} &   & \\ \cline{2-2}
{\raisebox{-2ex}[-5ex]{{$\phantom{\text{CTL}}$}}}&  {\bf combined}&
{\raisebox{2ex}[0ex]{$\EXPSPACE$-complete}} &
{\raisebox{2ex}[0ex]{ $\Pi^0_1$-complete}}\\ \cline{1-4}
{\raisebox{-2ex}[0ex]{{\bf LTL}}}&  {\bf data} &
\multicolumn{2}{c|} {\raisebox{0ex}[0ex]{$\coNP$-complete}}
\\ \cline{2-4}
{\raisebox{-2ex}[-5ex]{{$\phantom{\text{CTL}}$}}}&  {\bf combined}&
$\PSPACE$-complete &
$\coNEXP$-complete\\ \cline{1-4}
\end{tabular}
\ \\[0.4cm]
\caption{The complexity of CTL, the modal $\mu$-calculus, and LTL on SOCA and
POCA.}{\label{T results}}
\end{center}
\end{table}

One of the motivations for our work was the recent discovery that
reachability is decidable and in fact $\NP$-complete for both SOCA and
POCA~\cite{HKOW10}. We were also influenced by the work of Demri and
Gascon on model checking extensions of LTL over non-succinct,
non-parametric one-counter automata~\cite{DG-jlc09}, as well as the
recent result of G\"oller and Lohrey establishing that model checking
CTL on such counter automata is $\PSPACE$-complete \cite{GoLo10}.

We note some interesting differences between our results and
corresponding questions regarding finite automata. For the latter, the
(combined) model checking problems for CTL, the $\mu$-calculus, and
LTL are respectively known to be $\mathsf{P}$-complete, in
$\mathsf{NP} \cap \coNP$, and $\PSPACE$-complete. Somewhat
surprisingly, for SOCA and POCA, the complexity ordering is reversed
and LTL becomes easier to model check than either CTL or the
$\mu$-calculus.

On a technical level, the most intricate result is the
$\EXPSPACE$-hardness of CTL model checking for SOCA, which requires
several steps. We first show that $\mathsf{EXPSPACE}$ is `exponentially
$\mathsf{LOGSPACE}$-serializable', adapting the known proof that
$\PSPACE$ is $\mathsf{LOGSPACE}$-serializable. Unfortunately, and in
contrast to~\cite{GoLo10}, this does not immediately provide
an $\EXPSPACE$ lower bound. In a subsequent delicate stage of the
proof, we show how to `split' the counter in order simultaneously to
perform $\PSPACE$ computations in the counter and manipulate numbers
of exponential size in a SOCA of polynomial size.

Our paper is organized as follows. In Section \ref{S Prelim} we
introduce general notations.
Section \ref{S CTL} deals with CTL and the modal $\mu$-calculus.
LTL model checking is content of Section \ref{S LTL}.
Finally, in Section \ref{S Concl} we give conclusions.

\section{Preliminaries}{\label{S Prelim}}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\renewcommand{\O}{\mathbb{O}}
%\renewcommand{\P}{\mathcal{P}}

\newcommand{\trans}[1]{\stackrel{#1}{\longrightarrow}}
\newcommand{\hook}[1]{\stackrel{#1}{\hookrightarrow}}
\renewcommand{\rule}[2]{\stackrel{#2}{\stackrel{#1}{\longrightarrow}}}

\renewcommand{\M}{\mathcal{M}}

\newcommand{\bit}{\text{bit}}
\renewcommand{\succ}{\text{succ}}

\newcommand{\Reach}{\text{Reach}}
\newcommand{\Moves}{\text{Moves}}
\newcommand{\bin}{\text{bin}}

By $\Z$ we denote the {\em integers} and by $\N=\{0,1,2,\ldots\}$ the
denote the {\em naturals}.  For each $i,j\in\Z$ we define
$[i,j]=\{k\in\Z\mid i\leq k\leq j\}$ and $[j]=[1,j]$.  For each
$i,n\in\N$, let $\bit_i(n)$ denote the $i^{\text{th}}$ least
significant bit of the binary representation of $n$. Hence
$n=\sum_{i\in\N}2^i\cdot\bit_i(n)$.  By $\bin_m(n)=\bit_0(n)\cdots
\bit_{m-1}(n)$ we denote the first $m$ least significant bits written
from {\em left to right}. When $m$ is not important we just write
$\bin(n)$.
%For any non-empty subset $M\subseteq\Z^+$, define $\gcd(M)$ to be the largest
%$k\geq 1$ that divides every element of $M$.
%We define $\gcd(\emptyset)=0$.
Let $p_i$ denote the $i^{\text{th}}$ prime number for each $i\geq 1$, i.e. $p_1=2$,
$p_2=3$ and so on. We define $\log(n)=\min\{i\geq 1\mid 2^i> n\}$, in other
words $\log(n)$ denotes the number of bits that are needed to represent $n$ in
binary.
All polynomials $p:\N\rightarrow\N$ that occur in this paper are assumed to satisfy
$p(n)\geq n$ for each $n\in\N$.
For each word $v=a_1\cdots a_n\in\Sigma^n$ over some finite alphabet $\Sigma$
and each $i,j\in[n]$ define $v[i,j]=a_i\cdots a_j$ and
$v(i)=v[i,i]$.

\bigskip

\renewcommand{\l}{\triangleright}
\renewcommand{\r}{\triangleleft}
\newcommand{\move}{\vdash}
\noindent
{\bf Turing machines and complexity theory:} In the following, we
introduce $f(n)$ space-bounded deterministic Turing machines (DTMs) in
a suitable way for proving lower bounds. These contain precisely one
input tape and one working tape.  In our setting, the working alphabet is
 assumed to be $\{0,1,\l,\r\}$,
where $\l$ is the {\em left marker}
and $\r$ is the {\em right marker}.  The working tape of the initial
configuration of such a DTM on an input $w\in\Sigma^n$ is
assumed to be $\l0^{f(n)}\r$, whereas its input tape is $\l w\r$.
Before we define DTMs, let $\Upsilon=\{-1,0,+1\}$ denote a set of
directions.

Formally, a {\em $f(n)$-space bounded deterministic Turing machine (DTM)} is a tuple
$\M=(S,\Sigma,s_0,F,\mu)$, where $S$ is a finite set of {\em states},
$\Sigma$ is a {\em finite input alphabet} with $\l,\r\not\in\Sigma$,
$s_0\in S$ is an {\em initial state}, $F\subseteq S$ is a set of {\em final
states}, and with $\Sigma_\M=\Sigma\uplus\{\l,\r\}$ we have that
$$\mu\quad:\quad S\times\Sigma_\M\times\{0,1,\l,\r\}\ \rightarrow\
S\times\Upsilon^2\times\{0,1,\l,\r\}$$
is the {\em transition function}, where $\mu(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$ means
that $\M$ is currently in state $s$, its input head reads $b_1$, its
working tape reads $b_2$, and $\M$ changes to state $s'$, moves its
input head in direction $\delta_1$,
moves its working head in direction $\delta_2$, and
writes the bit $b$.
%We require
%moreover that whenever $\M$'s input (resp. working) head reads $\l$, then $\M$
%must move its input (resp. working) head to the right, i.e. $+1$, and whenever
%$\M$'s input (resp.
%working) head reads $\r$, %it must hold that
%$\M$ moves its
%input (resp. working) head to the left, i.e. $-1$. Formally,
%whenever $\delta(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$, then
%for each $i\in\{1,2\}$ we have that $b_i=\l$ implies $\delta_i=+1$ and $b_i=\r$
%implies $\delta_i=-1$.
Moreover, we require that $\M$ writes a marker precisely when it reads a marker,
more formally: $b_2=m$ if and only if $b=m$ for each $m\in\{\l,\r\}$.
As expected, a {\em configuration of $\M$} is a tuple $(s,\l w\r,\l w'\r,i,j)$,
where $s\in S$ is the current state,
$w\in\Sigma^n$ is an input,
$w'\in\{0,1\}^{f(n)}$ is the content of the working tape of $\M$,
$i\in[0,n+1]$ is the current input head position, and
$j\in[0,f(n)+1]$ is the current working head position.
%For configurations $C_1$ and $C_2$, we write $C_1\move_\M C_2$ whenever $\M$ can
%move from $C_1$ to $C_2$ on one step.
We define the {\em language $L(\M)$ of $\M$} to consist of all words
$w\in\Sigma^*$ such that
the {\em initial configuration}
$(s_0,\l w\r,\l 0^{f(|w|)}\r,0,0)$ reaches a configuration
$(s_0,\l w\r,\l w'\r,i,j)$ satisfying $s\in F$.

By $\L$, $\PSPACE$, and $\EXPSPACE$ we denote the class of all
problems that can be decided by DTM that
is logarithmically, polynomially, exponentially space bounded,
respectively. Recall that $\Sigma^0_1$ (resp. $\Pi^0_1$) is the class of
all languages that are (resp. whose complements are) recursively enumerable.
\newcommand{\Prop}{\mathcal{P}}
\newcommand{\Op}{\text{Op}}
%\newcommand{\add}{\mathsf{add}}
\renewcommand{\A}{\mathbb{A}}
%\newcommand{\EF}{\mathsf{EF}}
%\newcommand{\CTL}{\mathsh{CTL}}
\bigskip

\noindent
\textbf{Transition systems: }
In the following, we fix a countable set of atomic propositions $\Prop$.
A {\em transition system} is a tuple $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, where
$S$ is a set of {\em states}, $S_\prop\subseteq S$ for each $\prop\in\Prop$ and
$S_\prop$ is
non-empty for finitely many $\prop\in\Prop$, and finally
$\rightarrow \subseteq S\times S$ is a set of {\em transitions}.
We prefer to use the infix notation $s_1\rightarrow s_2$ to abbreviate
$(s_1,s_2)\in\rightarrow$.
%A {\em finite path} of length $n$ is sequence of transitions
%$\pi=s_0\rightarrow s_1\cdots \rightarrow s_n$, where $n\in\N $.
 An {\em infinite path} is an infinite sequence
$\pi=s_0\rightarrow s_1\cdots$.
% We denote
%by $|\pi|$ the length of the path $\pi$, which we define to be $\omega$
%if $\pi$ is infinite.
For each infinite path
$\pi= s_0\rightarrow s_1 \rightarrow \cdots$ and
$i\in\N$, we denote by $\pi^i$ the path $s_i\rightarrow s_{i+1}\cdots$
and by $\pi(i)$ the state $s_i$.
Define the {\em trace of $\pi$} as $\trace(\pi):\N\rightarrow2^\Prop$ where
$\trace(i)=\{\prop\in\Prop\mid s_i\in S_\prop\}$ for each $i\in\N$.
\bigskip

\noindent
\textbf{Succinct and parametric one-counter automata: }
A {\em succinct one-counter automaton (SOCA)} is a tuple
 $\Soca=(Q,\{Q_\prop\mid \prop\in\Prop\},E,\lambda)$, where $Q$ is a finite set
of {\em control states}, $Q_\prop\subseteq Q$ for each $\prop\in\Prop$ and
$Q_\rho$ is non-empty for finitely many $\prop\in\Prop$,
$E\subseteq Q\times Q$ is a finite set of {\em transitions}, and
$\lambda:E\rightarrow \Z\cup\{\zero\}$ labels the edges with decrements,
increments, and zero tests.
A {\em parametric one-counter automaton (POCA)} is a tuple
$\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, where the first
three components are same as for a SOCA,
where $X$ is a finite set of {\em parameters over the natural numbers},
and where $\lambda:E\rightarrow
(\Z\cup\{\zero\}\cup\{\circ x\mid\circ\in\{+,-\}, x\in X\})$.
For each assignment $\sigma:X\rightarrow\N$  the
{\em induced SOCA} is defined as $\Poca^\sigma=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda')$
where $\lambda'(e)=\circ\sigma(x)$ whenever $\lambda(e)=\circ x$
for some $\circ\in\{+,-\}$ and $\lambda'(e)=\lambda(e)$ otherwise.
If $X=\{x\}$ is a singleton, then we also write $\Poca(x)$ for $\Poca(X)$.
The {\em size} of a POCA is defined as
$|\Poca|=|Q|+|X|+|E|\cdot\max\{\log(|a|)\mid a\in\lambda(E)\cap\Z\}$.
Hence, we represent each appearing
integer in binary.
The size of a SOCA is defined analogously.
A SOCA $\Soca=(Q,\{Q_\prop\mid \prop\in\Prop\},E,\lambda)$ describes a transition system $T(\Soca)=(Q\times\N,\{Q_\prop\times\N\mid
\prop\in\Prop\}, \rightarrow)$, where for each $q_1,q_2\in Q$ and each
$n_1,n_2\in\N$ we have $q_1(n_1)\rightarrow q_2(n_2)$ if and only
if either $\lambda(q_1,q_2)=n_2-n_1$, or
($n_1=n_2=0$ and $\lambda(q_1,q_2)=\zero)$.

\section{CTL Model Checking}{\label{S CTL}}

In Section \ref{S CTL syn} we introduce the syntax and semantics of $\ctl$.
Since the upper bounds for the modal $\mu$-calculus follow from
known results and since the lower bounds follow from CTL, we do not introduce
the modal $\mu$-calculus formally.
In Section \ref{S Upper} we state an $\EXPSPACE$ upper bound for the combined
complexity of $\ctl$ and the modal $\mu$-calculus on SOCA and a $\Pi^0_1$ upper bound
on POCA. In Section \ref{S Serial} we recall basic notions from complexity theory
such as serializability and results on Chinese remainder representation.
In Section \ref{S CTL hardness} we prove the main result of this section, namely
that the data complexity of $\ctl$ on SOCA is $\EXPSPACE$-hard.
Finally, we prove a $\Pi^0_1$ lower bound for the data complexity
of $\ctl$ on POCA in Section \ref{S CTL und}.

\subsection{$\ctl$: Syntax and Semantics}{\label{S CTL syn}}
\noindent
Formulas $\varphi$ of $\ctl$ are given by the following
grammar, where $\prop$ ranges over $\Prop$:

$$
\varphi\quad ::=\quad \prop\ \mid \neg\varphi\ \mid\ \varphi\wedge\varphi\ \mid\
\EX\varphi\ \mid \E (\varphi\U\varphi)\ \mid\
\E(\varphi\WU\varphi)
$$

\noindent
We have the following abbreviations:
$\varphi_1\vee\varphi_2=\neg(\neg\varphi_1\wedge\neg\varphi_2)$,
$\varphi_1\rightarrow\varphi_2=\neg\varphi_1\vee\varphi_2$,
$\true=\prop\vee\neg\prop$ for some atomic proposition $\prop\in\Prop$,
and $\EF\varphi=\E(\true\U\varphi)$.
The {\em size} of a $\ctl$ formula is inductively defined as follows:
$|\prop|=1$ for each $\prop\in\Prop$, $|\neg\varphi|=|\EX\varphi|=|\varphi|+1$,
$|\varphi_1\wedge\varphi_2|=|\E(\varphi_1\U\varphi_2)|=|\E(\varphi_1\WU\varphi_2)|=
|\varphi_1|+|\varphi_2|+1$.
Given a transition system $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, a state
$s\in S$, and some $\CTL$ formula $\varphi$, define $(T,s)\models\varphi$ by
induction on the structure of $\varphi$ as follows:

\begin{eqnarray*}
  (T,s)\models \prop &  \quad\iff\quad & s\in S_p\quad \text{ for each
}\prop\in\Prop\\
  (T,s)\models \varphi_1\wedge\varphi_2 &  \quad\iff\quad &
(T,s)\models\varphi_1\text{ and }
(T,s)\models \varphi_2 \\
  (T,s)\models \neg\varphi &  \quad\iff\quad & (T,s)\not\models\varphi\\
  (T,s)\models \EX\varphi &  \quad\iff\quad & (T,t)\models\varphi
\text{ for some } t\in S \text{ with }s\rightarrow t\\
  (T,s)\models \E(\varphi_1\U\varphi_2) &  \quad\iff\quad &
\exists s_0,\cdots,s_n\in S, n\geq 0: s_0=s, (T,s_n)\models\varphi_2 \text{ such
that}\\
&& \forall i\in[0,n-1]: (T,s_i)\models\varphi_1\text{ and } s_i\rightarrow
s_{i+1}\\
  (T,s)\models \E(\varphi_1\WU\varphi_2) &  \quad\iff\quad &
(T,s)\models\E(\varphi_1\U\varphi_2)\ \text{ or }\\
&& \exists s_0,s_1,\ldots\in S:\forall i\geq 0: (T,s_{i})\models\varphi_1\text{ and }
s_i\rightarrow s_{i+1}
\end{eqnarray*}

\noindent
Let us define the {\em $\ctl$ model checking problem} on SOCA and POCA
respectively.\\

\problemx{$\ctl$ model checking on SOCA}
{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
$\varphi$.}
{$(T(\Soca),q(n))\models\varphi$?}
\

\problemx{$\ctl$ model checking on POCA}
{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
$\varphi$.}
{$(T(\Poca^\sigma),q(n))\models\varphi$ for every $\sigma:X\rightarrow\N$?}

\subsection{Upper bounds}{\label{S Upper}}
Let us first state upper bounds on model
checking CTL and the modal $\mu$-calculus on SOCA.
 {\em One-counter automata (OCA)} are SOCA in which
 the numbers that occur in the transition labels are represented  in unary
 instead in binary.
%In this section we state the upper bounds on model checking $\ctl$ on SOCA and
%POCA. They follow from known results or are immediate.
The following theorem gives an upper bound on model checking the
{\em modal $\mu$-calculus} on OCA.
Consult \cite{ArnNiw01} for more details on the modal $\mu$-calculus.

\begin{theorem}[\cite{Serr06}]
The combined complexity of the modal $\mu$-calculus on OCA is in $\PSPACE$.
\end{theorem}

\noindent
Since every SOCA can be transformed into an exponentially larger OCA
 and since each $\ctl$ formula can be translated into an alternation-free
$\mu$-calculus formula with a linear blowup, the following corollary is
immediate.

\begin{corollary}{\label{C CTL up}}
The combined complexity of $\ctl$ and the modal $\mu$-calculus on SOCA is in $\EXPSPACE$.
\end{corollary}

\noindent
The following upper bound for $\ctl$ on POCA is straightforward.

\begin{corollary}
The combined complexity of $\ctl$ and the modal $\mu$-calculus on POCA is in $\Pi^0_1$.
\end{corollary}
\begin{proof}
Let $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$ be a POCA,
 $q\in Q$, and $n\in\N$, and $\varphi$ be
a formula.
From Corollary \ref{C CTL up} it follows that
for each given $\sigma:X\rightarrow\N$ the question
$
(T(\Poca^\sigma),q(n))\models\varphi
$
is a decidable predicate.
Hence, the question if
$
(T(\Poca^\sigma),q(n))\models\varphi
$
for each $\sigma:X\rightarrow\N$ is a $\Pi^0_1$-predicate since
we can encode the set of all such assignments into the naturals
via G\"odel encoding.
\qed
\end{proof}

\subsection{Serializability and Chinese remainder representation}{\label{S Serial}}
For a language $L\subseteq\Sigma^*$ let $\chi_L:\Sigma^*\rightarrow\{0,1\}$
denote the {\em characteristic function of $L$}, i.e.
$\chi_L(x)=1$ if $x\in L$ and $\chi_L(x)=0$ otherwise, for each $x\in\Sigma^*$.
We define the {\em lexicographic
order on $n$-bit strings} as $\preceq_n\subseteq\{0,1\}^n\times\{0,1\}^n$, where
$x\preceq y$ if and only if
$\bin_n(x)\leq\bin_n(y)$, e.g. $101\preceq_3 011$.

\renewcommand{\C}{\mathcal{C}}
\begin{definition}
Let $\C$ be a complexity class.
We say a language $L$ is {\em $\C$-serializable via some language
$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$, and some language
$U\in\C$ such that for all $x\in\{0,1\}^n$
$$x\in L\qquad\!\!\iff\qquad\!\!
\chi_U\left(x,0^{p(n)}\right)\quad \cdots\quad \chi_U\left(x,1^{p(n)}\right)\in
R,$$ where with `$\cdots$'  we
refer to $\preceq_{p(n)}$ in each of the second components.
\end{definition}

\noindent
% it was shown that $\PSPACE$ is $\mathsf{P}$-serializable;
%in \cite{HLSVW93} this result was sharpened to $\AC^0$-serializability,
%see also \cite{Vol98}. It is not stated in \cite{HLSVW93,Vol98} but easy to
%see from the proofs that \emph{logspace-uniform} $\AC^0$ suffices for serializing
%$\PSPACE$, see the appendix for more details.
The following theorem was proven in \cite{GoLo09}.

\begin{theorem}[Theorem 22 in \cite{GoLo09}]{\label{T Serial}}
For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
logspace-uniformly $\mathsf{AC}^0$-serializable via $R$.
\end{theorem}

\noindent
As an immediate consequence, we obtain the following corollary.

\begin{corollary}{\label{C Serial}}
For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
$\L$-serializable via $R$.
\end{corollary}

\noindent
We remark that our definition of serializability is adopted from \cite{GoLo09}
and differs slightly from the standard
notion as used in \cite{CaFu91,HLSVW93,Vol98}.
We aim at lifting Corollary \ref{C Serial} via an appropriate notion of
serializability to $\EXPSPACE$.

\begin{definition}
Let $\C$ be some complexity class.
A language $L$ is {\em exponentially $\C$-serializable via some language
$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$ and some language
$U\in\C$ such that for all $x\in\{0,1\}^n$
$$x\in L\qquad\!\!\iff\qquad\!\!
\chi_U\left(x,0^{2^{p(n)}}\right)\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right)\in
R,$$ where with `$\cdots$' we
refer to $\preceq_{2^{p(n)}}$ in each of the second components.
\end{definition}

\noindent
The following proposition is folklore and immediate.
\begin{proposition}{\label{P Padding}}
Let $L\subseteq\Sigma^*$ be in $\EXPSPACE$. Then there is some polynomial $q$
such that the padded language
$\{x\$^{2^{q(|x|)}}\mid x\in L\}\subseteq(\Sigma\uplus\{\$\})^*$ is in $\PSPACE$.
\end{proposition}

\begin{theorem}\label{T succ seri}
For every language $L$ in $\EXPSPACE$ there is some regular language $R$ such
that $L$ is exponentially $\L$-serializable via $R$.
\end{theorem}
\begin{proof}
Let $L\subseteq\Sigma^*$ be some language in $\EXPSPACE$.
Then by Proposition \ref{P Padding} for some polynomial $q$ the language
$L'=\{x\$^{2^{q(|x|)}}\mid x\in L\}$ is in $\PSPACE$.
Due to $\L$-serializability of $\PSPACE$ by Corollary \ref{C Serial}, there exists
some polynomial $p'$, some regular language $R'$ and some $U'\in\L$
 such that for each $y\in(\Sigma\uplus\{\$\})^*$ we have
\begin{eqnarray}
y\in L'\quad\iff\quad \chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
\chi_{U'}\left(y,1^{p'(|y|)}\right)\in R', \label{E Serial}
\end{eqnarray}
where `$\cdots$' refers to the lexicographic order $\preceq_{p'(|y|)}$.
To prove the lemma we have to find some regular language $R$, some
polynomial $p$ and some $U\in\L$
such that for all $x\in\Sigma^*$ we have
$$
x\in L\quad \iff\quad \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \cdots\
\chi_U\left(x,1^{2^{p(|x|)}}\right)\in R,
$$
where here `$\cdots$' refers to the lexicographic order
$\preceq_{2^{p(|x|)}}$.
For now, we choose $p$ sufficiently fast growing,
let us postpone this to the end of the proof.
Let us describe the language $U$ and consider inputs $(x,w)$
 with $x\in\Sigma^n$ and $w\in\{0,1\}^{2^{p(n)}}$;
moreover let $y=x\$^{2^{q(|x|)}}$ be the padding of $x$.
Our language $U$ consists of all such pairs $(x,w)$
such that
$w$ can be factorized as $w=bz0^l$ for some $b\in\{0,1\}$
and some $l\geq 0$ such that
$b=0$ or $(y,z)\in U'$.
Roughly speaking, the goal of $U$ is to simulate $U'$
on the corresponding padded word,
however the length of the second component
$w$ has to be a power of two, in other words the last $l$ bits of $w$
 contain redundant information.
We will be able to filter out this redundant information via
the bit $b$.
Let us make this more precise.
Let
\begin{eqnarray*}
\gamma' &\quad=\quad&
\chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
\chi_{U'}\left(y,1^{p'(|y|)}\right)\quad\text{and}\\
\gamma &\quad=\quad& \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \ \cdots\ \
\chi_U\left(x,1^{2^{p(|x|)}}\right).
\end{eqnarray*}

\noindent
Then by definition of $U$, the first $2\cdot p'(|y|)$ bits of $\gamma$ are
$1\gamma'(1)1\gamma'(2)\cdots 1\gamma'(p'(|y|))$, where
the remaining bits of $\gamma$ are all $0$.
Hence, when reading $\gamma$ in {\em pairs of bits},
precisely when the first of the two bits
is set to $1$ we read some relevant information.

We have to give some regular language $R$ such that $\gamma'\in R'$ if and only
if $\gamma\in R$.
Recall that regular languages are closed under shuffle product $|\!|$ and
homomorphisms.
We define $R=\varphi(R'|\!|\{a\}^*)$, where
$\varphi:\{0,1,a\}\rightarrow\{0,1\}^*$ is the following homomorphism:
$\varphi(a)=00$, $\varphi(0)=10$, and $\varphi(1)=11$.

As expected, we choose $p$ to be an arbitrary polynomial satisfying
$2^{p(n)}\geq 1+p'(2^{q(n)}+n)$ for all $n\geq0$.
Finally, let us show that $U$ is in $\L$.
Let $(x,w)$ be an input to $U$ and let again $y=x\$^{2^{q(|x|)}}$ be the padding
of $x$.
Then it is straightforward to decide in logarithmic space if $w$ can be factorized as
$w=bzv$ such that $b\in\{0,1\}$, $|z|=p'(2^{q(|x|)}+|x|)$ and
$v\in\{0\}^*$.
The only thing that might remain to be verified in logarithmic space is whether
$(y,z)\in U'$, which in turn boils down to simulating some logarithmic space bounded TM
for $U$ on input $(y,z)$.
But this is possible since $|y|+|z|\leq 2\cdot(|x|+|w|)$ as
$$
|y|+|z|=|x|+2^{q(|x|)}+p'(2^{q(|x|)+|x|})\leq 2\cdot p'(2^{q(|x|)}+|x|)\leq
2\cdot 2^{p(|x|)}\leq 2\cdot (|x|+|w|).
$$
\qed
\end{proof}

\renewcommand{\M}{\mathcal{M}}

\noindent
\textbf{Chinese remainder representation: }
 For every $m,M\in\N$ we denote by $\CRR_m(M)$ {\em the Chinese remainder
representation} of $M$ as the Boolean tuple
$(b_{i,c})_{i\in[m],0\leq c<p_i}$, where $b_{i,c}=1$ if $M\modulo p_i=c$
and $b_{i,c}=0$ otherwise.
The following theorem tells us that in logarithmic space we can
compute the binary representation of a natural number that is given in Chinese
remainder representation.
It is a consequence of the result that division is in logspace-uniform
$\mathsf{NC}^1$, proven in \cite{ChDaLi01}.

\begin{theorem}[\cite{ChDaLi01} Theorem 3.3]{\label{T CRRBIN}}
The following problem is in $\L$:

INPUT: $\CRR_m(M),j\in[m]$, $b\in\{0,1\}$.

QUESTION: $\bit_j(M\ \modulo 2^m)=b$?
\end{theorem}

\subsection{$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA}{\label{S CTL hardness}}

In the rest of this section, we give the proof of $\EXPSPACE$-hardness
of the data complexity of $\ctl$ on SOCA.  Let $L\subseteq\{0,1\}^*$ be an arbitrary
language in $\EXPSPACE$. Then by Theorem \ref{T succ seri}, there is
some regular language $R\subseteq\{0,1\}^*$ such that $L$ is
exponentially $\L$-serializable via $R$. Hence there is some language $U\in\L$
and some polynomial $p$ such that for all $x\in\{0,1\}^n$ we have

\begin{eqnarray}{\label{E EXPSPACE char}}
x\in L\qquad\iff\qquad \chi_{U}\left(x,0^{2^{p(n)}}\right)
\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right) \in R,
\end{eqnarray}
where '$\cdots$' refers to the lexicographic order $\preceq_{2^{p(n)}}$
on the bit strings on the right-hand side.

\noindent
For the rest of this section, let us fix an input $x_0\in\{0,1\}^n$.
Let $N=p(n)$ and let $A=(Q,\{0,1\},q_0,\delta,F)$ be some
deterministic finite automaton with $L(A)=R$.  Let us briefly recall
what $A$ consists of: $Q$ is a finite set of {\em states}, $\{0,1\}$
is the {\em input alphabet}, $q_0\in Q$ is the {\em initial state},
$\delta:Q\times\{0,1\}\rightarrow Q$ is the {\em transition function},
and $F\subseteq Q$ is the set of {\em final states}.  Let us describe
equivalence (\ref{E EXPSPACE char}) differently: We have $x_0\in L$ if
and only if the program in Figure \ref{F U} returns $\mathtt{true}$.

\begin{figure}
\normalsize

\ \hspace{5cm} $q\in Q$; $q:=q_0$;\\[-0.4cm]

\ \hspace{5cm} $d\in\N$; $d:=0$;

\ \hspace{5cm} $b\in\{0,1\}$;\\[-0.4cm]

\ \hspace{5cm} $\mathtt{while }\  d\not=2^{2^N}$\ $\mathtt{ loop}$\\[-0.4cm]

\ \hspace{6cm} $b:=\chi_U(x_0,\text{bin}_{2^N}(d))$;\\[-0.4cm]

\ \hspace{6cm} $q:=\delta(q,b)$;\\[-0.4cm]

\ \hspace{6cm} $d:=d+1$;\\[-0.4cm]

\ \hspace{5cm} $\mathtt{end loop}$\\[-0.4cm]

\ \hspace{5cm} $\mathtt{return}\ q\in F$;%\\[0.4cm]

\caption{A program that returns $\mathtt{true}$ if and only if $x_0\in L$}{\label{F U}}
\end{figure}

\noindent
Our goal is to mimic the execution of the program in Figure \ref{F U} by
a fixed $\ctl$ formula and by a SOCA that depends on
our input $x_0$.
Before we start with the reduction, let us discuss the obstacles
that arise:

\begin{enumerate}[{\bf(A)}]
\item We need some way of storing $d$ on the counter. Of course
there are a lot of ways to do this, but since we want to access
all bits of $d$ in the assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$, the most
natural way is probably to represent $d$ in binary.
However, for this $2^N$ bits are required.
 More problematically, we need to be able to check if $d$ is equal to
$2^{2^N}$. This cannot be achieved by a transition in a SOCA
 that subtracts $2^{2^N}$, since the representation of this number
requires exponentially many bits in $n$.
\item As in \cite{GoLo10}, a solution to obstacle {\bf (A)} is to store $d$ in
Chinese remainder representation with the first $2^N$ prime numbers.
A polynomial number of bits (in $n$) suffice
to represent each of the occurring prime numbers, but
we need exponentially many of them.
Thus, we cannot equip a polynomial size SOCA with transitions for
each prime number, simply because there are too many of them.
\item
The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ implies that we need to simulate on
the counter a logarithmically space bounded DTM for the language $U$
on an exponentially
large input (in $n$).
Speaking in terms of the input size $n$, this means that we need to provide
polynomially many bits on the counter that can be used to describe the
working tape for this DTM. However, we need to provide
some on-the-fly mechanism for reading the input.
\end{enumerate}

\iffalse

\begin{enumerate}[(1)]
\item For the representation of the variable $d$ we need $2^N+1$ bits
  on our counter. More challenging, we need to check if $d$ is
  equal to $2^{2^N}$. To check this, we cannot simply allow a
  transition in our counter automaton that subtracts $2^{2^N}$,
  since the representation of this number requires exponentially many
  bits in $n$.
\item As in \cite{GoLo10}, a solution to problem (1) is to store $d$
  in Chinese remainder representation with the first $2^N+1$ primes.
  Polynomially (in $n$) many bits suffice to represent each of the
  occurring prime numbers, but we need exponentially many different
  prime numbers, where each of the primes has up to exponentially many
  residue classes that we cannot simply write down. So we cannot equip
  the SOCA that we have to derive with transitions for each prime
  number, simply since there are too many of them.

\item The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ means that we need to simulate on
our counter a logarithmically space bounded DTM for $U$ on an exponential
sized input (in $n$).
Speaking in terms of our input size $n$, this means that we need to provide
polynomially many bits on our counter that can be used to describe the
working tape for this DTM. However, we need to provide
some on-the-fly mechanism for reading the input.
\end{enumerate}

\fi
\noindent
To achieve this, let us give a high-level description of how we proceed.
In a first step, we carefully design a data structure on the counter
and describe the intuition behind it.
In a second step, we give five queries which we aim at implementing
via fixed CTL formulas and by SOCA that can be computed from $x_0$
in logarithmic space.

\bigskip

\noindent
{\bf The data structure and how to access it: }
Let $K=n+2^N+1$ denote the number of bits that are required to store an input
for $U$ plus one.
Let $\alpha=\log K$ denote the number of bits that we require for
storing a pointer to an input for $U$ and let $\beta$
be the number of bits that suffice for storing the
$K^{\text{th}}$ prime. Hence $\alpha=O(N)$ and by
the Prime Number Theorem, it follows that
$\beta\in O(\log(K\log(K)))=O(N)$.
The number $\alpha$ and such a sufficiently large number $\beta$ can be computed
 from $x_0$ in logarithmic space.

Let us describe how we will interpret the counter in our reduction.
Assume that the counter value is $v\in\N$.  Instead of
treating $v$ as a natural number, we are interested only in the $l$
least significant bits $V$ of the binary representation of $v$, where $l$
is some number that is exponentially bounded in $n$; the precise value
of $l$ will be made clear below.  Assume
$V=\bit_0(v)\cdots\bit_{l-1}(v)$.  We read $V$ to be factorized into
blocks of bits
$$\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad V\quad=\quad I\ M\ C\ J\ X\ Y\ Z\ B,\qquad\qquad\qquad\qquad\qquad\qquad(\bigstar)$$ where
\begin{itemize}
\item $I\in\{0,1\}^{\alpha}$ represents a prime number index,
\item $M\in\{0,1\}^\beta$ represents $I^{\text{th}}$
prime number $p_I$,
\item $C\in\{0,1\}^\beta$ represents some residue class modulo $M$,
\item $J\in\{0,1\}^\alpha$ represents a pointer to some bit of $B$,
\item
each $X,Y$ and $Z$ consist of polynomially many bits (in $n$) and
represent the working tape of three space bounded
DTMs that we will comment on later in more detail, and
\item $B\in\{0,1\}^{n+2^N+1}$ with $B=xB'$ for some
$x\in\{0,1\}^n$ and some
$B'\in\{0,1\}^{2^N+1}$,
$B$ represents the current
input for $U$, where
$x$ is reserved to represent our input $x_0$
and
where $B'$ represents the counter $d$ from program in Figure \ref{F U} from above.
The block $B'$ consists of $2^N+1$ bits since we want to be able to test
 if $d=2^{2^{N}}$.
\end{itemize}

\renewcommand{\Re}{\Rightarrow}
\newcommand{\Li}{\Leftarrow}
\newcommand{\re}{\rightarrow}
\newcommand{\li}{\leftarrow}

\noindent
Let us introduce some more notation for addressing
leftmost (starting to count from $0$) and rightmost bit
positions in each of the above sequences of bits in $V$.
For each such sequence $\Theta$ let $\Theta_{\Li}$ and $\Theta_{\Re}$ denote the
respective positions of the leftmost and rightmost bits of $\Theta$ within the bit
string $V$, e.g. $I_\Re=\alpha-1$ and $C_\Li=\alpha+\beta$.

\medskip

\noindent
{\bf Important remark: } Throughout the rest of this section $v$ will
denote an arbitrary natural number, moreover $I,M,C,J,X,Y,Z$, and $B$
will implicitly be coupled with $v$.  Note that all of the bit strings
have polynomial length in $n$ except for $B$.  Moreover, we identify each
of the blocks with the natural number they represent.

\medskip

\noindent
A very simple but important gadget that we need is to decide, for each
$b\in\{0,1\}$ if the $i^{\text{th}}$ bit of $v$ is $b$.

\begin{lemma}{\label{L bit}}
For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\bit,b}$
such that the following is computable in logarithmic space:

INPUT: $i\in\N$ in unary.% and our input $x_0$.

OUTPUT: A SOCA $\Soca_{\bit,i}$ and some control state $q_{\bit,i}$ such that
$$
(T(\Soca_{\bit,i}),q_{\bit,i}(v))\models\varphi_{\bit,b}\quad\Longleftrightarrow\quad
\bit_i(v)=b.
$$
\end{lemma}
\begin{proof}
The SOCA $\Soca_{\bit,i}$ has the two atomic propositions $\tau_1$ and $\tau_2$ and is
depicted below:

\begin{center}
\begin{picture}(85,20)(-20,-5)
%   \put(-20,-28){\framebox(85,33){}}
  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
 % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
 % \node(A)(0,0){$q_0$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

\put(13,-2){\small $\tau_1$}
\put(41,-2){\small $\tau_2$}
  \node(A)(20,0){$q_{\bit,i}$}
  \node(B)(40,0){$\bullet$}
\drawloop(A){$-2^{i+1}$}
\drawedge(A,B){$-2^i$}
\end{picture}
\end{center}
The simple way to check if $\bit_i(v)=1$ (resp. $\bit_i(v)=0$) is to
repeatedly subtract $2^{i+1}$ from $v$ until no longer possible
and then to check if $2^i$ can (resp. cannot) be subtracted.
Hence we put
\begin{eqnarray*}
\varphi_{\bit,1}&\quad=\quad&
\tau_1\wedge
\EF(\tau_1\wedge\neg\EX \tau_1\wedge\EX
\tau_2)\qquad\text{ and}\\
\varphi_{\bit,0}&=&
\tau_1\wedge
\EF(\tau_1\wedge\neg\EX \tau_1\wedge\neg\EX
\tau_2).
\end{eqnarray*}
\qed
\end{proof}

\noindent
{\bf Queries that we need to implement: }
We will implement the following five queries Q1 to Q5 by instances
of the model checking problem, where each query builds on top
of its preceding queries.

\begin{enumerate}[(Q1)]
\item \fbox{When assuming $C<M$, does $B\equiv C\ \mod M$ hold?}
\item \fbox{Is $M$ the $I^{\text{th}}$ prime number, i.e. $M=p_I$?}
%\item Let $\CRR_{2^\alpha}(B)=(b_{i,r})_{i\in[2^\alpha],0\leq r<p_i}$.
%What is $b_{I,R}$? %resp. $b_{I,R}=1$ hold?
\item \fbox{What is $\bit_J(B)$?}
\item \fbox{Does $(B[1,n],B[n+1,n+2^N])\in U$ hold?}
%\item Does $B'=2^{2^N}$ hold?
\item \fbox{Does $x_0\in L$ hold?}
\end{enumerate}

\noindent
We implement each of the five queries by providing fixed CTL formulas
and SOCA that can be computed from $x_0$ in logarithmic space.
$\EXPSPACE$-hardness of the data complexity of CTL on SOCA will hence
follow from the implementation of Q5.
First, let us give an implementation of Q1.

\begin{lemma}{\label{L mod}} There exists some fixed $\ctl$ formula $\varphi_\mod$ such that we
can compute from $x_0$ in logarithmic space some SOCA $\Soca_\mod$ and some
control state $q_\mod$ such that
$(T(\Soca_{\mod}),q_\mod(v))\models\varphi_\mod$ if and only if $B\equiv C\ \mod M$.
\end{lemma}
\begin{proof}
The SOCA $\Soca_{\mod}$ contains the three atomic propositions $\prop_0,\prop_1$, and
$\prop_2$ and is depicted in Figure \ref{F Smod}.
The $\ctl$ formula $\varphi_\mod$ expresses that we
traverse the upper
sequence of diamonds and thereby repeatedly subtract $M$ from $B$.
The number of diamonds both in the first and the second row equals $\beta$,
the number of bits of $M$ and of $C$.
In the upper row, one diamond corresponds to one bit of $M$. In case bit the rightmost bit of $M$
(in other words bit $M_\Re$ of the counter) is $1$,
which we can verify by a transition to the initial control state of
the SOCA $\Soca_{\bit,M_\Re}$,
 we subtract $2^{B_\Li+\beta-1}$ from $B$, otherwise we do not modify the counter value.
After that, we move on to the second diamond, which represents the second
rightmost  bit of $M$, and so on.
Hence, traversing a cycle that starts and ends
in our initial control state $q_\mod $ will correspond to subtracting $M$ once.
 Then we traverse
the lower sequence of diamonds and similarly subtract $C$, but this
this time only once.
Finally, after having traversed the lower sequence of diamonds and seeing the atomic
proposition $\prop_2$, we check if $B=0$ by trying to subtract $2^{B_\Li}$.
\begin{figure}
\begin{center}
\begin{picture}(100,120)(5,-41)
%   \put(-20,-28){\framebox(85,33){}}
  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
 % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
 % \node(A)(0,0){$q_0$}
%\put(15,-5){$q(M_{\abs,\r})$}
%\put(43,-5){$p_2$}
\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)
\put(10,25){\Large$\Soca_{\bit,M_\Re}$}

\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(58,20)(58,40)(35,40)
\put(38,25){\Large$\Soca_{\bit,M_\Re-1}$}

\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,20)(108,20)(108,40)(85,40)
\put(90,25){\Large$\Soca_{\bit,M_\Li}$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

  \node(1)(0,60){$q_{\mod}$}
  \node(2O)(15,70){$\bullet$}
  \node(2U)(15,50){$\bullet$}
  \put(17,72){\small$\prop_1$}
  \put(17,48){\small$\prop_0$}

  \node(1A)(15,35){$q_{\bit,M_\Re}$}
  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}

  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}

  \node(7A)(95,35){$q_{\bit,M_\Li}$}

  \node(3)(30,60){$\bullet$}
  \node(4O)(45,70){$\bullet$}

  \put(47,72){\small$\prop_1$}
  \put(47,48){\small$\prop_0$}
  \node(4U)(45,50){$\bullet$}
  \node(5)(60,60){$\bullet$}

  \node(6)(80,60){$\bullet$}
  \node(7O)(95,70){$\bullet$}
  \put(97,72){\small$\prop_1$}
  \put(97,48){\small$\prop_0$}

  \node(7U)(95,50){$\bullet$}
  \node(8)(110,60){$\bullet$}

  \put(65,58){\huge$\cdots$}

%\drawedge[curvedepth=-70](8,1){$0$}

\drawedge(1,2O){$0$}
\drawedge[ELside=r](1,2U){$0$}
\drawedge[ELside=l,ELpos=60](2O,3){$-2^{B_\Li+\beta-1}$}
\drawedge(2U,3){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](2O,1A){$0$}
\drawedge(2U,1A){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](4O,4A){$0$}
\drawedge(4U,4A){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](7O,7A){$0$}
\drawedge(7U,7A){$0$}

\drawedge(3,4O){$0$}
\drawedge[ELside=r](3,4U){$0$}
\drawedge[ELside=l,ELpos=70](4O,5){$-2^{B_\Li+\beta-2}$}
\drawedge(4U,5){$0$}

\drawedge(6,7O){$0$}
\drawedge[ELside=r](6,7U){$0$}
\drawedge[ELside=l,ELpos=60](7O,8){$-2^{B_\Li}$}
\drawedge(7U,8){$0$}

\drawcurve(110,60)(112,70)(55,80)(0,70)(0,62)

%\drawedge[ELpos=70,curvedepth=1.5](8,9A){$0$}
%\drawedge[ELside=r,ELpos=60](8,9B){$0$}
  \put(60,81){$0$}
%\drawedge[ELpos=50](8,9C){$0$}
%\drawedge[ELside=r,ELpos=50](8,9D){$-2^{B_L}$}

%%%%COPY

\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,-40)(28,-40)(28,-20)(5,-20)
\put(10,-35){\Large$\Soca_{\bit,C_\Re}$}

\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,-40)(58,-40)(58,-20)(35,-20)
\put(38,-35){\Large$\Soca_{\bit,C_\Re-1}$}

\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,-40)(108,-40)(108,-20)(85,-20)
\put(90,-35){\Large$\Soca_{\bit,C_\Li}$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

  \node(1C)(0,0){$\bullet$}
  \drawedge[ELside=r](1,1C){$0$}
  \node(2OC)(15,10){$\bullet$}
  \node(2UC)(15,-10){$\bullet$}
  \put(17,12){\small$\prop_1$}
  \put(17,-12){\small$\prop_0$}

  \node(1AC)(15,-25){$q_{\bit,C_\Re}$}
  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}

  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}

  \node(7AC)(95,-25){$q_{\bit,C_\Li}$}

  \node(3C)(30,0){$\bullet$}
  \node(4OC)(45,10){$\bullet$}

  \put(47,12){\small$\prop_1$}
  \put(47,-12){\small$\prop_0$}
  \node(4UC)(45,-10){$\bullet$}
  \node(5C)(60,0){$\bullet$}

  \node(6C)(80,0){$\bullet$}
  \node(7OC)(95,10){$\bullet$}
  \put(97,12){\small$\prop_1$}
  \put(97,-12){\small$\prop_0$}

  \node(7UC)(95,-10){$\bullet$}
  \node(8C)(110,0){$\bullet$}

  \node(9C)(110,-20){$\bullet$}

\drawedge(8C,9C){$-2^{B_\Li}$}

  \put(112,-2){$\prop_2$}

  \put(65,-2){\huge$\cdots$}

%\drawedge[curvedepth=-70](8,1){$0$}

\drawedge(1C,2OC){$0$}
\drawedge[ELside=r](1C,2UC){$0$}
\drawedge[ELside=l,ELpos=60](2OC,3C){$-2^{B_\Li+\beta-1}$}
\drawedge(2UC,3C){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](2OC,1AC){$0$}
\drawedge(2UC,1AC){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](4OC,4AC){$0$}
\drawedge(4UC,4AC){$0$}

\drawedge[curvedepth=-6,ELpos=70,ELside=r](7OC,7AC){$0$}
\drawedge(7UC,7AC){$0$}

\drawedge(3C,4OC){$0$}
\drawedge[ELside=r](3C,4UC){$0$}
\drawedge[ELside=l,ELpos=70](4OC,5C){$-2^{B_\Li+\beta-2}$}
\drawedge(4UC,5C){$0$}

\drawedge(6C,7OC){$0$}
\drawedge[ELside=r](6C,7UC){$0$}
\drawedge[ELside=l,ELpos=60](7OC,8C){$-2^{B_\Li}$}
\drawedge(7UC,8C){$0$}

\end{picture}
\caption{The SOCA $\Soca_\mod$}{\label{F Smod}
for checking if $B\equiv C\ \mod M$.}
\end{center}
\end{figure}
Finally, let us give the formula $\varphi_\mod$:
$$
\varphi_\mod\quad=\quad\E \left(\bigwedge_{i\in\{0,1\}} \prop_i
\rightarrow \EX \varphi_{\bit,i} \right) \U (\prop_2\wedge\neg\EX\true)
$$
\qed
\end{proof}

\noindent
We need the following
proposition for implementing Q2. Its proof is straightforward.

\begin{proposition}{\label{P prime}}
The following problem is in $\PSPACE$:

INPUT: $\bin(i)\$\bin(m)$

QUESTION: $m=p_i$?
\end{proposition}

%\renewcommand{\prime}{\text{prime}}
\newcommand{\prim}{\text{prime}}

\begin{lemma}{\label{L prim}} There is some fixed $\ctl$ formula $\varphi_\prim$ such that
from  $x_0$ we can compute in logarithmic space some SOCA
$\Soca_\prim$ and some control state $q_\prim$ such that
$(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$ %\quad\Leftrightarrow\quad
if and only if $M=p_I$.
\end{lemma}
\begin{proof}
By Proposition \ref{P prime} there is some $q(n)$ space-bounded DTM
$\M=(S,\Sigma,s_0,F,\mu)$
over the alphabet $\Sigma=\{0,1,\$\}$ that decides,
given $\bin(i)\$\bin(m)$, whether $m=p_i$, where $q$ is some polynomial.
The idea is to simulate $\M$ on input $z=I\$M$ using $X$ from $(\bigstar)$ as a working tape.
For this, we define our block $X$ to consist of
$l=q(|I|+|M|+1)=q(\alpha+\beta+1)$ many bits.
Moreover, we assume w.l.o.g. that $\M$'s behaviour on input $I\$M$ is
independent of the content of its initial working tape: This can be
achieved by adding extra states to $\M$ that first write $0^l$ onto
the working tape.
Hence, the initial configuration of $\M$ on input $z$ is
$(s_0,\l z\r,\l w \r,0,0)$ for some $w\in\{0,1\}^l$.
Before we give our SOCA $\Soca_\prim$, we describe the computation of
$\M$ on input $I\$M$ as a pseudo program that terminates if and only if $M=p_I$:\\

\ \hspace{2cm} $s\in S$; $s:=s_0$;\quad  (current state of $\M$)\\[-0.4cm]

\ \hspace{2cm} $i\in[0,\alpha+\beta+2]$; $i:=0$;\quad (current input head position)

\ \hspace{2cm} $h\in[0,l+1]$; $h:=0$;\quad (current working head position)

\ \hspace{2cm} $a\in\{0,1,\$,\l,\r\}$; $a:=\l$;\quad (current input symbol)

\ \hspace{2cm} $b\in\{0,1,\l,\r\}$; $b:=\l$;\quad (current working tape
symbol)\\[0.05cm]

\ \hspace{2cm} $\mathtt{while }\  s\not\in F$\ $\mathtt{ loop}$\\[-0.4cm]

\ \hspace{3cm} $a:=\begin{cases} \l& \text{ if }i=0\\
\r &\text{ if }i=\alpha+\beta+2\\
\$ &\text{ if }i=\alpha+1\\
\bit_i(I) &\text{ if }i\in[1,\alpha]\\
\bit_{i-\alpha-1}(M) &\text{ otherwise }
\end{cases}$

\ \hspace{3cm} $b:=\begin{cases} \l& \text{ if }h=0\\
\r &\text{ if }h=l+1\\
\bit_{h}(X) &\text{ otherwise }
\end{cases}$

%\\[-0.4cm]

\ \hspace{3cm} Let $\mu(s,a,b)=(s',\delta_1,\delta_2,b')$.\\[-0.4cm]

\ \hspace{3cm} $\mathtt{if}\ h\in[1,l]\ \ \mathtt{then}\  \ \bit_h(X):=b';\ \ \mathtt{fi}$\\[-0.4cm]

\ \hspace{3cm} $s:=s '$;\\[-0.4cm]

\ \hspace{3cm} $i:=i+\delta_1$;\\[-0.4cm]

\ \hspace{3cm} $h:=h+\delta_2$;\\[-0.4cm]

\ \hspace{2cm} $\mathtt{end loop}$\\[-0.2cm]

\newcommand{\inpu}{\text{input}}
\newcommand{\inz}{\text{input}_0}
\newcommand{\ino}{\text{input}_1}
\newcommand{\work}{\text{work}}
\newcommand{\workz}{\text{work}_0}
\newcommand{\worko}{\text{work}_1}

\noindent
Let us describe our SOCA $\Soca_\prim$.
The control states $Q$ of $\Soca_\prim$ will contain
$$
S\times[0,\alpha+\beta+2]\times[0,l+2]\times
\{0,1,\$,\l,\r\}\times\{0,1,\l,\r\}, \qquad\text{where}
$$
the first component corresponds to the variable $s$,
the second component corresponds to the variable $i$,
the third component corresponds to the variable $h$,
the fourth component corresponds to the variable $a$, and
finally the fifth component corresponds to the variable $b$
of the above program.
Moreover, our SOCA will contain
the SOCA $\Soca_{\bit,I_\Li},\Soca_{\bit,I_\Li+1}\ldots,\Soca_{\bit,I_\Re}$ and
$\Soca_{\bit,X_\Li},\Soca_{\bit,X_\Li+1}\ldots,\Soca_{\bit,X_\Re}$ in order to test if
certain bits
in $I$ and $X$ are set correctly.
Hence $Q$ contains the control states
$q_{\bit,I_\Li},q_{\bit,I_\Li+1}\ldots,q_{\bit,I_\Re}$ and
$q_{\bit,X_\Li},q_{\bit,X_\Li+1}\ldots,q_{\bit,X_\Re}$ as well.

\noindent
We will provide the atomic propositions
$\Prop=\{F,\inpu,\work,\inz,\ino,\workz,\worko\}$,
where
\begin{eqnarray*}
Q_F&\ =\ &\{(s,i,h,a,b)\in Q\mid s\in F\},\\
Q_\inpu&\ =\ &\{q_{\bit,k}\mid k\in[I_\Li,I_\Re]\},\\
Q_\work&\ =\ &\{q_{\bit,k}\mid k\in[X_\Li,X_\Re]\},\\
Q_{\inz}&\ =\ &\{(s,i,h,a,b)\in Q\mid a=0\},\\
Q_{\ino}&\ =\ &\{(s,i,h,a,b)\in Q\mid a=1\},\\
Q_{\workz}&\ =\ &\{(s,i,h,a,b)\in Q\mid b=0\},\\
Q_{\worko}&\ =\ &\{(s,i,h,a,b)\in Q\mid b=1\}.\\
\end{eqnarray*}
We do not give all transitions of $\Soca_\prim$, but illustrate
some cases by way of example.
Let us give the outgoing transitions of the control state
$(s,i,h,0,1)$, where $i\in[I_\Li,I_\Re]$ and $h\in[X_\Li,X_\Re]$,
i.e. we currently scan the $i^{\text{th}}$ bit
of $I$ and do not read a marker on our working tape.
Moreover, let us assume $\mu(s,0,1)=(s',+1,-1,0)$, i.e. we change to state $s'$,
we move the input head to the right, the working head to the left,
and modify the current content on the working tape from $1$ to $0$.
We realize the latter by subtracting $2^{X_\Li+h}$ from the counter,
 however we allow transitions
to states $(s',i+1,h-1,a,b)$, {\em for each} $a,b\in\{0,1\}$, hence we guess
the input tape symbol and the working tape symbol of the
successor configuration.
The $\ctl$ formula $\varphi_\prim$ will guarantee that our guessing was correct by
accessing the control states $q_{\bit,k}$ where
$k\in[I_\Li,I_\Re]\cup[M_\Li,M_\Re]$.

\begin{center}
\begin{picture}(85,73)(-20,4)
\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

\put(-20,35){\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)}

\put(50,35){\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)}

\put(-12,68){\large$\Soca_{\bit,I_\Li+i}$}
\node(I)(-3,60){$q_{\bit,I_\Li+i}$}
  \put(-12,57){\tiny\bf$\inpu$}

\put(58,68){\large$\Soca_{\bit,X_\Li+h}$}
\node(X)(67,60){$q_{\bit,X_\Li+h}$}
  \put(72,57){\tiny\bf$\work$}

  \node(A)(30,40){$(s,i,h,0,1)$}
  \put(39,40){\tiny\bf$\inz,\worko$}
  \node(B1)(-35,10){$(s,i+1,h-1,0,0)$}
  \put(-40,6){\tiny\bf$\inz,\workz$}

  \put(0,6){\tiny\bf$\inz,\worko$}
  \put(40,6){\tiny\bf$\ino,\workz$}
  \put(80,6){\tiny\bf$\ino,\worko$}

  \node(B2)(5,10){$(s,i+1,h-1,0,1)$}
  \node(B3)(45,10){$(s,i+1,h-1,1,0)$}
  \node(B4)(85,10){$(s,i+1,h-1,1,1)$}
\drawedge[ELside=r](A,B1){$-2^{X_\Li+h}$}
\drawedge[ELside=r,ELpos=70](A,B2){$-2^{X_\Li+h}$}
\drawedge[ELside=r](A,B3){$-2^{X_\Li+h}$}
\drawedge[ELside=l](A,B4){$-2^{X_\Li+h}$}

\drawedge[ELside=l](A,I){$0$}
\drawedge[ELside=l](A,X){$0$}

\end{picture}
\end{center}
The other cases can be dealt with analogously.
We put $q_\prim=(s_0,0,0,\l,\l)$.
Our final formula is
$$
\varphi_\prim\quad=\quad\E
\left(
\bigwedge_{k\in\{0,1\}}
\left(\text{input}_k\rightarrow\EX(\inpu\wedge\varphi_{\bit,k})\right)
\wedge
\left(\text{work}_k\rightarrow\EX(\work\wedge\varphi_{\bit,k})\right)
\right)\U\ F.
$$
 Then
$(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$
if and only if
$I\$M\in L(\M)$ if and only if
$M=p_I$.
\qed
\end{proof}

\newcommand{\BIT}{\text{BIT}}

\noindent
Let us now give an implementation of query Q3.

\begin{lemma}{\label{L BIT}}
For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\BIT,b}$
such that from $x_0$
we can compute in logarithmic space some SOCA $\Soca_{\BIT}$ and some
control state $q_\BIT$ such that
$(T(\Soca_\BIT),q_\BIT(v))\models\varphi_{\BIT,b}$
if and only if $\bit_J(B)=b$.
\end{lemma}
\renewcommand{\C}{\mathcal{C}}
\renewcommand{\R}{\mathcal{R}}
\begin{proof}[Sketch]
By Lemma \ref{L prim}, there is a fixed $\ctl$
formula $\varphi_\prim$ such that we can compute from $x_0$ a
SOCA $\Soca_\prim$ that allows to check if $M=p_I$.
Moreover, by Lemma \ref{L mod} there is a fixed $\ctl$
formula $\varphi_\mod$ and a SOCA $\Soca_\mod$ that allows
to check whether $B\equiv C \mod\ M$.
In other words, we can deal with the Chinese remainder representation
of $B$, but our goal is to access the the $J^{\text{th}}$ bit in the
binary representation of $B$.
So let us assume that $\R=\CRR_{K}(B)=(b_{i,c})_{1\leq i\leq K,0\leq c<p_i}$
is the Chinese remainder representation of $B$.
Note that we do not have $\R$ stored anywhere on the counter.
However, the bit strings $I$ and $C$ serve as pointers to access the
bit $b_{I,C}$ of $\R$.
By Theorem \ref{T CRRBIN}, given $\R$ (on-the-fly by our pointers $I$ and $C$), our
bit string $J$, and the bit $b$,
we can decide if $\bit_J(B)=b$ in logarithmic space.
So let $\M=(S,\Sigma,s_0,F,\mu)$ be some $k\cdot\log(m)$ space-bounded
Turing machine machine for this, where $k\geq 1$ is some constant.
Hence, in order to decide if
$\bit_J(B)=b$ we need to simulate $\M$ on input $\langle \R, J, b\rangle$.
For this we store the space that $\M$ requires on our reserved
sequence of bits $Y$ from $(\bigstar)$, hence $Y$ consists of
$l=k\cdot\log(|\R Jb|)=O(\beta^2+\alpha+1)=O(N^2)$ many bits.
The definition of SOCA $\Soca_{\BIT,b}$ works analogously to the construction
of $\Soca_{\prim}$ in the proof of Lemma \ref{L prim}
i.e. we introduce in $\Soca_{\BIT,b}$ control states
that remember the current bit of the input head, the current bit of
the working head, and the position of the working head.
The only difference is that a pointer to the input head cannot be
completely stored
in the control states.
As mentioned above, for this we employ the bit blocks $I$ and $C$,
and also the sequence $M$ for storing the $I^{\text{th}}$ prime number.
In order to obtain bit $b_{I,C}$, we allow transitions to our SOCA $\Soca_\mod$
(which we can compute in logarithmic
space from $x_0$ by Lemma \ref{L mod}) and checking
(i) via the fixed formula $\varphi_\prim$ whether $M=p_I$, and
then (ii) via the fixed formula $\varphi_{\mod}$
whether $B\equiv C\ \mod M$.
Pointers to the remaining parts of the input for $\M$, namely $J$ and $b$,
can directly be handled by the control states of $\Soca_{\BIT,b}$,
in analogy to the proof of Lemma \ref{L prim}.
By way of example we explain the behavior of $\Soca_{\BIT}$ when the input head
of $\M$ currently scans bit $b_{I,C}$ of $\R$ and when getting
to the successor configuration requires moving the input head to the left.
Then we simply decrement $C$ by $1$ which corresponds to subtracting $2^{C_\Li}$
from the counter. If, however, $C$ currently equals $0$, we need to
decrement $I$ by $1$, overwrite $M$ with prime $p_{I-1}$, and finally
overwrite $C$ with $p_{I-1}-1$. Decrementing $I$ by $1$ can simply achieved by
subtracting $2^{I_\Li}$ from the counter. Overwriting $M$ with prime $p_{I-1}$
can be done by repeatedly subtracting $1$ from $M$ (i.e. subtracting $2^{M_\Li}$
from the counter) until $M$ equals $p_{I-1}$;
checking if $M=p_{I-1}$ can be done via the fixed $\ctl$ formula $\varphi_{\prim}$
and the SOCA $\Soca_{\prim}$ by Lemma \ref{L prim}.
The other cases work analogously.
\qed
\end{proof}

\noindent
The following lemma implements query Q4.

\begin{lemma}{\label{L U}}
For each $b\in\{0,1\}$ there is some fixed $\ctl$ formula $\varphi_{U,b}$ such that
from $x_0$
we can compute in logarithmic space some SOCA $\Soca_U$ and some
control state $q_U$ such that $(T(\Soca_U),q_U(v))\models\varphi_{U,b}$
if and only if $b=\chi_U(B[1,n],B[n+1,n+2^N])$.
\end{lemma}
\begin{proof}[sketch]
The proof is similar to the proof of Lemma \ref{L BIT}.
Since $U$ is in $\L$, there is some $k\cdot\log(m)$-space bounded
Turing machine that decides $U$, where $k\geq 1$ is some constant.
In order to decide if $b=\chi_U(B[1,n],B[n+1,n+2^N])$ we
need to simulate $\M$ on input $\langle B[1,n],B[n+1,n+2^N]\rangle$.
We store the space that $\M$ requires on its working tape
in the sequence of bits $Z$ from $(\bigstar)$, hence $Z$ consists of
$k\cdot\log(n+2^N+1)=O(N)$ many bits.
We use the bit sequence $J$ as a pointer for accessing bits of $B$.
For reading the $J^{\text{th}}$ bit of $B$ we make use of the $\ctl$ formula
$\varphi_{\BIT,b}$ and use the SOCA $\Soca_\BIT$ from Lemma \ref{L BIT}.
The rest of the proof follows along the same lines as the proof of
Lemma \ref{L prim}.
\qed
\end{proof}

\noindent
The following lemma implements query Q5 and concludes the
$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA.

\begin{lemma}
There is some fixed $\ctl$ formula $\varphi_L$ such that from $x_0$
we can compute in logarithmic space some SOCA $\Soca_L$ and some
control state $q_L$ such that $(T(\Soca_L),q_L(0))\models\varphi_L$
if and only if $x_0\in L$.
\end{lemma}
\begin{proof}
First note that our $\ctl$ formula $\varphi_L$ will be evaluated in state $q_L(0)$.
Recall that our bit sequence $B$ has length $n+2^N+1$ where $B$ is factorized
as $B=xB'$
for some $x\in\{0,1\}^n$ and some $B'\in\{0,1\}^{2^N+1}$.
The SOCA $\Soca_L$ and the $\ctl$ formula $\varphi_L$ will mimic the execution
of the program from Figure \ref{F U}.
In the bit string $x$ we store the value $x_0$.
The bit string $B'$ represents
the variable $d$ of the program, hence we will initialize
$B'$ with $0$.
Note that incrementing $B'$ by $1$ corresponds to adding $2^{B_\Li+n}$ to the counter.
Thus, checking when $d$ becomes $2^{2^N}$ for the first time boils down to checking
when the $n+2^N+1^{\text{st}}$ bit of $B$ becomes $1$ for the first time.
 By Lemma
\ref{L BIT} the $\ctl$ formula $\varphi_{\BIT,1}$ and the SOCA $\Soca_\BIT$ allow to test
if the $J^{\text{th}}$ bit of $B$ equals $1$.
Therefore we store in $J$ the number $n+2^N+1$.
The following claim tells us that we can test whether we have
initialized the counter correctly.
Its proof is simple and therefore omitted.
\newcommand{\init}{\text{init}}

\medskip

\noindent
{\em Claim:} There is a fixed $\ctl$ formula $\varphi_\init$ such that from
$x_0$ we can compute in logarithmic space some SOCA $\Soca_\init$ and
some control state $q_\init$ such that
$(T(\Soca_\init),q_\init(v))\models\varphi_\init$
if and only if $J=n+2^N+1$ and $B=x_00^{2^N+1}$.\\

\noindent
Recall that $A=(Q,\{0,1\},q_0,\delta,F)$ is the deterministic finite automaton
of the program in Figure \ref{F U} that needs to be simulated.
Let us, before giving $\Soca_L$, define the auxiliary SOCA $\Soca_A$
to be connected with the control state $q_U$ of $\Soca_U$ and with the control
state $q_\BIT$ of $\Soca_\BIT$ along with the additional
control states $S=Q\times\{0,1\}$ and
the following transitions:
$$
(q,b)\xrightarrow{+2^{B_\Li+n}}(\delta(q,b),b'),\qquad
(q,b)\xrightarrow{0}q_\BIT,\qquad\text{and}\qquad
(q,b)\xrightarrow{0}q_U\qquad\text{for each }
b,b'\in\{0,1\}.
$$

\noindent
Moreover, $\Soca_A$ contains the atomic propositions
$\{\bit_0,\bit_1,F\}$, where $S_{\bit_b}=Q\times\{b\}$ for each $b\in\{0,1\}$ and
$S_F=F\times\{0,1\}$.
Before giving $\varphi_L$, let us depict our SOCA $\Soca_L$ which has
two additional labels $\rho$ and $\init$:

\begin{center}
\begin{picture}(85,53)(-40,-12)
%   \put(-20,-28){\framebox(85,33){}}
  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
 % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
 % \node(A)(0,0){$q_0$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

\put(-13,0){
\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(58,20)(58,40)(35,40)
\put(48.5,28){\Large$\Soca_{\init}$}
}

\put(-51,-35){
\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(68,20)(68,40)(35,40)
\put(48.5,25){\Large$\Soca_A$}
}

\put(0,23){\small $\prop$}
\put(25,33){\small $\init$}
  \node(A)(-30,20){$q_L$}
  \node(B)(0,20){$\bullet$}
  \node(C)(30,30){$q_\init$}
  \node(D1)(-10,0){$(q_0,0)$}
  \node(D2)(10,0){$(q_0,1)$}
\drawloop(A){$+1$}
\drawedge(A,B){$0$}
\drawedge(B,C){$0$}
\drawedge[ELside=r](B,D1){$0$}
\drawedge(B,D2){$0$}
\end{picture}
\end{center}
Let us define the auxiliary formula $\chi=\EX\varphi_\BIT$ that allows
to test if $B'=2^{2^N}$.
Our final formula is
$$
\varphi_L\quad=\quad\EF\left(\rho\wedge \EX(\init\wedge\varphi_\init)\wedge
\E\left(\neg\chi\wedge\bigwedge_{b\in\{0,1\}}\bit_b\rightarrow
\EX\varphi_{U,b}\right)
 \U\ (\chi\wedge F)
\right)
$$
We have $x_0\in L$ if and only if
the program from Figure \ref{F U} returns $\mathtt{true}$ if and only if
$(T(\Soca_L),q_L(0))\models\varphi_L$.
\qed
\end{proof}

\begin{theorem}
The data and combined complexity of $\ctl$ and the modal $\mu$-calculus
 on SOCA is $\EXPSPACE$-complete.
\label{thm:ctl-poca}
\end{theorem}

\iffalse
\section{$\EF$ model checking is in $\PSPACE$}

For a finite and non-empty set of integers $S\subseteq \Z\setminus\{0\}$ not
containing $0$, we denote by {\em $\gcd(S)$} the greatest natural number that
divides all numbers $\{|a|: a\in S\}$.
A {\em linear equation} over $\Z$ is given by
$$
x\quad =\quad a_1\cdot x_1 + \cdots\ + a_k\cdot x_n,
$$
where $a_i\in\Z$ is a {\em coefficient} and $x_i$ is a {\em variable}
ranging over $\N$ for each $i\in[n]$.
We call $b\in\Z$ a {\em solution} to the above linear equation if there exists a
mapping $\sigma:\{x_1,\ldots,x_n\}\rightarrow\N$ such that
$b=\sum_{i\in[n]} a_i\cdot\sigma(x_i)$.
An {\em arithmetic progression} with {\em offset} $a\in\Z$ and {\em period
$b\in\Z$} is the set $a+b\N=\{a+b\cdot\N\mid n\in\N\}$.
The following lemma has been used in \cite{AnthonyIPL,Chrobak,Shallit}.

\begin{lemma}[\cite{lookup citation}]\label{L NT}
Let $N\in\N$ and let
\begin{eqnarray}
x\quad=\quad a_1\cdot x_1 + \cdots\ + a_n\cdot x_n,\label{L Eq}
\end{eqnarray}
be a linear equation such that $1<a_1<\cdots a_n\leq N$ and where
$b=\gcd(a_1,\ldots,a_n)$.
Then, the set of solutions to (\ref{L Eq}) is some set
$U\cup\ a+b\N$,
where $U\subseteq[1,N^2]$,
and
$a=\min\{y>N^2\mid y\equiv 0\textsl{ mod }g\}$.
\end{lemma}

\begin{lemma}
Let $N\in\N$ be some natural and let
\begin{eqnarray}
x\quad=\quad a_1\cdot x_1 + \cdots\ + a_n\cdot x_n,\label{L Eq neg}
\end{eqnarray}
be a linear equation
where $-N\leq a_1<a_2<\cdots <a_n\leq N$ and where $g=\gcd(a_1,\ldots,a_n)$.
Let $b=\gcd(a_1,\ldots,a_n)$ and let
$a=\min\{y>N^2\mid y\equiv 0\textsl{ mod }b\}$.
Then, the set of positive solutions to (\ref{L Eq neg}) is a set
$U\cup\  a+ b\N$, where $U\subseteq[1,N^2]$ and the
set of negative solutions to (\ref{L Eq neg}) is a set
$V\cup -a-b\N$, where $V\subseteq[-N^2,-1]$.
\end{lemma}
\begin{proof}
We only prove the lemma for the negative solutions.
Let $c<0$ be an arbitrary negative solution. Thus,
$c=d+a_1'\cdot m_1+\ldots +a_h'\cdot m_h$, where
$d\in[-N,0]$ and
$\{a_1',\ldots,a_h'\}\subseteq\{a_1,\ldots,a_n\}\cap\Z^-$ is a subset of
the negative coefficients.
Hence $c$ is the solution
to some linear equation $x=b_1\cdot x_1+\ldots +b_l\cdot x_l$, where
$\{b_1,\ldots,b_l\}\subseteq[-N,-1]$. By application of Lemma
\ref{L Eq} and switching from positive integers to negative integers, the lemma
follows.
\qed
\end{proof}

\begin{lemma}
  \label{lem:bounded-coefficients}
  Let $N,m$ be natural numbers and let $a_1,\ldots, a_n$ be integers
  such that $-N\le a_1 \le \ldots \le a_n\le N$, $m>N^3$ and
  \begin{equation*}
    m = a_1 \cdot x_1 + \cdots + a_n \cdot x_n
  \end{equation*}
  for some natural numbers $x_1, \ldots, x_n$. Then there exist
  natural numbers $y_1, \ldots, y_n$ such that
  \begin{equation*}
    m = a_1 \cdot y_1 + \cdots + a_n \cdot y_n
  \end{equation*}
  and $y_i\le N^2$ for $1\le i<n$.
\end{lemma}
\begin{proof}
  Let $a_\ell\in\{a_1,a_n\}$ be the $a_i$ with the largest
  \emph{absolute} value. Define $x_i' =x_i \text{ mod } a_\ell$ for
  $1\le i\le n, i\neq \ell$, and let $m'= \sum_{1\le i\le n, i\neq
    \ell} a_i\cdot m_i$. Clearly, we have $m'<N^3$ and $m'\equiv m
  \text{ mod } a_\ell$. Let $h=m-m'=a_\ell \cdot x_\ell'$ for some
  integer $x_\ell'$. In case $a_\ell=a_n$, the lemma follows by
  setting $y_i=x_i'$ for $1\le i\le n$. Otherwise, let $x_n^*= \lceil
  \frac{h}{|a_1a_n|} \rceil$. The lemma follows by setting $y_i=x_i'$
  for $1<i<n$, $y_1=\frac{x_n^*|a_1a_n|-h}{|a_1|}$ and $y_n=x_n'+
  x_n^*|a_1|$.
\end{proof}
\begin{remark}
  \label{rem:maximum-valley}
  It follows from the previous lemma that $\sum_{a_i<0} a_i\cdot
  y_i>-N^3$.
\end{remark}

\section{Periodicity analysis}

\newcommand{\cycles}{\textrm{cycles}}

Let us fix a SOCA $\O=(Q,\{Q_p\mid p\in\Prop\},\delta_0,\delta_{>0})$ for
 the rest of this section.

In the following, we will treat $\O$ as a weighted directed graph $G(\O)$, where
the weights are taken according to $\delta_{>0}$.
More precisely, we define $G(\O)=(Q,\delta_{>0})$
and write $q_1\stackrel{a}{\mapsto}q_2$ as an abbreviation for
$(q_1,a,q_2)\in\delta_{0}$ and simply $q_1\mapsto q_2$ if
$(q_1,a,q_2)\in\delta_{>0}$ for some $a\in\Z$.
Note that every non-zero path in the transition system $T(\O)$ corresponds to a unique path in
$G(\O)$.
A {\em strongly connected component} of $G(\O)$ is an
equivalence class of the equivalence relation $\mapsto^*$.
For every finite path $\pi$ in $G(\O)$, let $w(\pi)$ denote the sum of all
weights of edges of $\pi$.
For every strongly connected component $C$ in $G(\O)$, let $\cycles(C)$ denote
the set of chord-free cycles whose total weight does not summarize to $0$,
more precisely
the set of paths $\pi=q_0\stackrel{m_1}{\mapsto} q_2\cdots
\stackrel{m_l}{\mapsto} q_l$, where $l\geq1$,
$q_0=q_l$, such that $i\not= j$ implies $q_i\not=q_j$ for all
$i,j\in[l]$ and moreover $\sum_{i\in[l]}m_i\not=0$.

Let $M=\max\{|a|:\exists q_1,q_2\in Q. q_1\stackrel{a}{\mapsto}q_2\}$
denote the maximal absolute value of a weight that appears in $G(\O)$.
Moreover, let $g_C=\gcd(w(cycles(C)))$ denote the greatest common divisor
of all non-zero weights of chord-free cycles in $C$.
In the following define $$\Delta\ =\ \LCM\{g_C\mid C\text{ is a strongly
connected component of } G(\O)\}.$$
\noindent
Note that $\Delta\in\exp(|\O|)$ since firstly $g_C\leq (k-1)\cdot M$ for each strongly
connected component $C$ and secondly there are at most $|Q|$
strongly connected components in $G(\O)$.
Let $k=|Q|$.

\begin{lemma}{\label{L Cycles}}
Let $C$ be some strongly connected component of $G(\O)$, let $q\in Q$ be some
control location, and let $n\geq k^2\cdot M$ be some natural.
Then in $T(\O)$ there exists some non-zero path from $(q,n)$ of height $h$ with $h\equiv
0\text{ mod }g_C$ and $|h|\leq k\cdot(k-1)\cdot M$  that visits each
control location in $C$ at least once.
\end{lemma}
\begin{proof}
Since $C$ is a strongly connected component and $n\geq k^2\cdot M$, there exists some
non-zero path
$$
\pi\quad =\quad(q_1,n_1)\rightarrow(q_2,n_2)\ \cdots\ \rightarrow (q_l,n_l)
$$
in $T(\O)$ with $l\in[k(k-1)]$ that visits each control state in $C$ at least once
(i.e. for all $p\in C$ there is some $i\in[l]$ such that $p=q_i$) such that
moreover $q_1=q_l=q$ and $n_1=n$. Thus, it follows $|n_1-n_l|\leq k\cdot
(k-1)\cdot M$.
Observe that for every subpath $\pi[a,b]=(q_a,n_a)\cdots\rightarrow(q_b,n_b)$ of $\pi$ that is a chord-free
cycle, we have $n_a\equiv n_b\text{ mod } g_C$ by definition of $g_C$.
By successively cutting out subpaths of $\pi$ that correspond to chord-free
cycles, it is apparent that $n_1\equiv n_l\text{ mod }g_C$.
\qed
\end{proof}

Observe that Lemma \ref{L Cycles} implies that there is a non-zero path in
$T(\O)$ that visits every
control location in $C$ at least once from $(q,k^2\cdot M+\delta)$ with the
desired properties with respect to the height.

\begin{lemma}
  Let $q\in Q$ be a state of some SCC $C$ and let $m,m'$ be natural
  numbers such that $m-m'>k^2\cdot M+ k^3\cdot M^3$ and $m'>k\cdot
  M$. Then $(q,m)\rightarrow^*(q,m')$ iff
  $(q,m+\Delta)\rightarrow^*(q,m')$.
\end{lemma}
\begin{proof}
  $(\Rightarrow)$ Let $q_1,\ldots,q_n\in C, 1\le n\le k$ be states on
  loop-free cycles $\pi_{a_i}:q_i\rightarrow \ldots \rightarrow q_i$
  with weights $a_i=w(\pi_{a_i})$ such that
  $\text{gcd}(a_1,\ldots,a_n)=g_C$.  Let $\pi_i^1: q\rightarrow^* q_i$
  and $\pi_i^2: q_i\rightarrow^* q$ be disjoint loop-free paths from
  $q$ to $q_i$ respectively $q_i$ to $q$. Denote by
  $\pi_i=\pi_i^1\cdot \pi_i^2$ a loop-free cycle from $q$ to itself
  passing through $q_i, 1\le i\le n$.

  Let $h=m-m'-\sum_{1\le i\le n}w(\pi_i)+\Delta$. Similarly to the
  previous lemma, we find $g_C|m-m'$, $g_C|\sum_{1\le i\le n}w(\pi_i)$
  and hence $g_C|h$. Moreover, $h>k^3\cdot M^3$ and thus we can derive
  from Lemma \ref{lem:bounded-coefficients} that there are
  $y_1,\ldots, y_n$ such that $\sum_{1\le i\le n}a_i\cdot y_i=h$ and
  $y_i<k\cdot M$, $1\le i<n$.

  Now define $\pi_i'=\pi_i^1\cdot (\pi_{a_i})^{y_i} \cdot \pi_i^2$ and
  $\pi=\pi_1'\cdot\pi_2'\cdot\cdots\cdot\pi_n'$. Cleary,
  $w(\pi)=m+\Delta-m'$. Moreover, $m'> k\cdot M$ ensures that is a
  valid path. Hence $(q,m+\Delta)\rightarrow^*(q,m')$ is witnessed
  by $\pi$.

  $(\Leftarrow)$ The proof from the previous direction can directly
  be adopted to show the statement.
\end{proof}

The following lemma is important for our periodicity analysis later.

\begin{lemma}
Let $q,q'\in Q$ and $n,n'\in\N$ such that
$n-n'>4\cdot k^3\cdot M^2$ and
$n'>k\cdot M$.
If there is a path from $(q,n)$ to $(q',n')$ in $T(\O)$, then there
is a path from $(q,n+\Delta)$ to $(q',n')$ in $T(\O)$.
\end{lemma}
\begin{proof}
Let us fix some path
$$
\pi\quad=\quad(q_1,n_1)\rightarrow(q_2,n_2)\quad \cdots\quad \rightarrow (q_l,n_l),
$$
where $(q_1,n_1)=(q,n)$ and $(q_l,n_l)=(q',n')$.
Since $n_1-n_l>4\cdot k^3\cdot M^2$ and since there are at most $k$ strongly connected components in $G(\O)$,  there
exists a strongly connected component $C$ of $G(\O)$ and some subpath $\pi[s,t]$
of $\pi$ such that the following holds
\begin{enumerate}
\item  $n_s-n_t>4\cdot k^2\cdot M^2$,
\item $n_t\geq n'$, and
\item $q_i\in C$ for each $i\in[s,t]$
\end{enumerate}
by the pigeonhole principle.
Let $\delta=n_t-n_s<-4\cdot k^2\cdot M^2$ denote the height of the subpath $\pi[s,t]$.
Let us treat the path $\pi[s,t]$ as a path in $G(\O)$. Now think of consecutively cutting out
chord-free cycles in $\pi[s,t]$  until this is no longer possible, in a greedy
manner from beginning to the end, say. Let us denote the resulting
path (possibly empty) by $\rho$ and let  $h\in[-(k-1)\cdot M,(k-1)\cdot M]$ denote
its weight.
\begin{remark}{\label{R rho}}
The path $\rho$ can be seen as a non-zero path in $T(\O)$ of length at most
$k-1$ from $(q_s,n_s)$ to $(q_t,n_s+h)$.
\end{remark}

\noindent
The following claim states that the height $\delta$ of $\pi[s,t]$ lies in some
arithmetic progression of small offset and small period.

\smallskip

\noindent
{\em Claim A.} We have $\delta\in (c+h)-g_C\N$,
for some
$c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ with $c\equiv 0\text{ mod }g_C$.

\medskip

\noindent
Let us postpone the proof of Claim A and finish the proof of the lemma.
It remains to show that there is some non-zero path
$(q,n+\Delta)\rightarrow^*(q',n')$ in $T(\O)$.
Before doing this, we claim that the following paths exist in $T(\O)$:
\begin{enumerate}[(1)]
\item $(q,n+\Delta)\rightarrow^*(q_s,n_s+\Delta)$.
\item $(q_t,n_t)\rightarrow^*(q',n')$.
\item $(q_s,n_t-h)\rightarrow^*(q_t,n_t)$.
\end{enumerate}
Point (1) is clear since $\pi[1,s]+\Delta$ witnesses this.
Point (2) holds by factorization of $\pi$.
For Point (3), observe that $\rho$ is a non-zero path from $(q_s,n_s)$ to $(q_t,n_s+h)$
of height $h$ and of length at most $k-1$
by Remark \ref{R rho}.
Since $n_t\geq n'>(k-1)\cdot M\geq h$, there is a non-zero path
from $(q_t,n_t-h)$ to $(q_t,n_t)$, namely
$\rho+(-n_s+n_t-h)=\rho+(\delta-h)$.
 By combining Points (3) and (2) we obtain some non-zero path $(q_t,n-h)\rightarrow^*
(q',n')$.
Due to Point (1), it suffices to show that there is some
non-zero path from $(q_s,n_s+\Delta)$ to $(q_s,n_t-h)$.
In other words, we have to prove that in $G(\O)$ there is some path from
$q_s$ to $q_s$ of weight
\begin{eqnarray*}
w\quad=\quad n_t-h-(n_s+\Delta)\quad=\quad n_t-n_s-h-\Delta\quad=\quad\delta-h-\Delta
\end{eqnarray*}
such that moreover the weight of any prefix of this path is strictly bigger
than $-(n_s+\Delta)$.
By Claim A we have $w=\delta-h-\Delta=(c+h)-r\cdot g_C-h-\Delta=c-r\cdot
g_C-\Delta$ for
some $r\in \N$ and some $c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ with
$c\equiv 0\text{ mod } g_C$.
%Since too $\Delta\equiv 0\text{ mod } g_c$, there exists some  $d\in\N$ such
%that
Since moreover $\Delta\equiv 0\text{ mod }g_C$, we have $w\equiv 0\text{
mod }g_C$.
Furthermore, we have $w<-3\cdot k^2\cdot M^2$.
Since $n_s+\Delta>BLA>k^2\cdot M$, there is some non-zero path $\sigma$ of height
$h_\sigma$ with $|h_\sigma|\leq k\cdot(k-1)\cdot M$ and $h_\sigma\equiv 0\text{
mod } g_C$
from $(q_s,n_s+\Delta)$ that visits each control location in $C$ at
least one by Lemma \ref{L Cycles} .
When taking the path $\sigma\circ\sigma$, the remaining weight is congruent $0$
modulo $g_C$ and can be estimated
as $$w-2\cdot h_\sigma\ <\ -3\cdot k^2\cdot M^2-2\cdot h_\sigma\ <
-3\cdot k^2\cdot M^2+2\cdot k\cdot (k-1)\cdot M<
-k^2\cdot M^2.
$$
Hence by Lemma \ref{L Eq neg} there exist $m_1,\ldots, m_n\in\N$ such that the
remaining weight equals $\sum_{i=1}^n a_i\cdot m_i$.
To finally obtain a path from $q_s$ to $q_s$ of weight $w$ in $G(\O)$, we
modify the first traversal of $\sigma$ by choosing for each positive $a_i$ some
chord-free cycle of weight $a_i$ through which we loop precisely $m_i$ many times, followed by
the modification of the second traversal of $\sigma$, where we choose for each
negative $a_i$ some chord-free cycle of weight $a_i$ through which we loop
precisely $m_i$ many times.
This concludes the proof of the lemma.

\smallskip

\noindent
{\em Proof of Claim A.}\\
\noindent
Observe that $\delta=h+r$, where $r$ is some solution
of the linear equation
$$
x\ =\ a_1\cdot x_1+\ \ldots\ +a_m\cdot x_m,
$$
where $\{a_1,\ldots, a_m\}\subseteq w(cycles(C))$ are weights of the chord-free cycles that
were cut out from $\pi[s,t]$ in order to obtain $\rho$. %\subseteq[-k\cdot M,k\cdot M]$.
Let $d=\gcd(a_1,\ldots,a_m)$ which is a multiple of $g_C$ by definition.
Since we know that $\delta<-4\cdot k^2\cdot M^2$ and $h\in[-(k-1)\cdot M,(k-1)\cdot M]$, we have
$$
r=\delta-h\quad<\quad-4\cdot k^2\cdot M^2+(k-1)\cdot M\quad\leq\quad -3\cdot k^2\cdot M^2.
$$
Moreover, we have obviously $-\sqrt{3}\cdot k\cdot M<-k\cdot M\leq a_i\leq
k\cdot M<\sqrt{3}\cdot k\cdot M$ for each $i\in[m]$.
Now let us define $c=\max\{y<-3\cdot k^2\cdot M^2\mid y\equiv 0\text{ mod }
d\}$.
As $d\in[M]$, it follows $c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ as
required.
By applying Lemma \ref{L Eq neg} we have that $r\in c-d\N$ and hence
$\delta\in (c+h)-d\N$.
Since $d$ is a multiple of $g_C$
it follows $\delta\in (c+h)-g_C\N$.
\qed
\end{proof}

\begin{theorem}
Model checking EF over SOCA is $\PSPACE$-complete.
\end{theorem}

\fi

\newcommand{\Two}{\mathbb{A}}
\renewcommand{\add}{\mathsf{add}}

\subsection{$\Pi^0_1$-Hardness of the Data Complexity of $\CTL$ on POCA}{\label{S CTL und}}
In this section we show that there already
exists a fixed $\CTL$ formula for which model checking of
$\poca$ is $\Pi^0_1$-hard.
We reduce from the complement of the emptiness problem for
\emph{two-counter automata}, which is $\Sigma^0_1$-complete \cite{Min61}.
Similar to a
SOCA, a {\em two-counter automaton} is a tuple $\Two=(Q, E, \lambda)$,
where $Q$ is a finite set of {\em control states},
$E\subseteq Q\times Q$ is a set of {\em edges},
 however
$\lambda: E \rightarrow \{\zero_1,\zero_2\}\cup \{\add_j(a)\mid j\in \{1,2\}, a\in
\{-1,1\} \}$. The {\em configuration graph induced by $\Two$} is defined
to be $G(\A)=(Q\times \N\times \N, \rightarrow)$, where
$(p,i_1,i_2)\rightarrow (q,i_1+d_1,i_2+d_2)$ if and only if
the following conditions are all satisfied:
\begin{itemize}
\item $(p,q)\in E$,
%\item  $i_1,i_2,i_1+d_1,i_2+d_2\geq 0$,
\item $\lambda(p,q)=\add_j(a)$ implies $d_j=a$ and $d_{3-j}=0$ for each
$j\in\{1,2\}$, and
\item $\lambda(p,q)=\zero_j$ implies $d_j=i_j=0$ and $d_{3-j}=0$ for each
$j\in\{1,2\}$.
\end{itemize}

\noindent
Let us define the following problem.
\smallskip

\problemx{Emptiness for two-counter automata}{Two-counter automaton $\Two=(Q,E,\lambda)$, control states
$q_0,q_1\in Q$.}{Are there $m,n\in\N$ such that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ in
$G(\Two)$?}

\medskip

\noindent
The idea of our reduction is as follows: Given a two-counter automaton
$\Two$, we construct a $\poca$ $\Poca(x)$ with one parameter in such a way
that the two counters from $\A$ are encoded into the single counter
from $\Poca(x)$ as follows: for the counter value $n$ of $\Poca(x)$,
we have that $n\ \mod x$ encodes the value of the first and $n\ \div\ x$ encodes the
value of the second counter of $\Two$.
Hence, testing the first counter for zero corresponds to checking
whether $n\equiv 0\ \mod x$ while testing the second counter for
zero corresponds to checking  whether $n\geq x$.
Incrementation (resp. decrementation) on the first counter can be
mimicked by adding (resp. subtracting) $1$, whereas on counter two this
corresponds to adding (resp. subtracting) $x$.
Of course, we need to ensure that we do not overflow.
For example, if $n\equiv -1\ \mod x$ and we would simulate to increment the first counter
in the above manner, this
would correspond to setting the first counter to zero and simultaneously
 incrementing the second counter. However, if the emptiness problem is
solvable, then $x$ can be instantiated with a large enough
value such that such an overflow does not occur.
Our $\ctl$ formula will ensure that we do not overflow.
\newcommand{\test}{\mathsf{test}}
Before we give the reduction from the emptiness problem for two-counter
automata, let us introduce some gadgets that we need in our reduction.
\begin{lemma}{\label{L POCAs}}
  There exists some fixed $\ctl$ formula $\varphi_\test$
for which the following holds:
\begin{enumerate}[(1)]
\item
There exists some fixed $\poca$ $\Poca_{\not\equiv 0}(x)$ with some control state
$q_{\not\equiv 0}$ such that for each $\sigma:\{x\}\rightarrow\N$
 $$(\Poca_{\not\equiv 0}^\sigma,q_{\not\equiv0}(n)) \models
 \varphi_{\test} \quad\Longleftrightarrow\quad n\not
  \equiv 0 \textsl{ mod } \sigma(x).$$
\item There exists some fixed $\poca$ $\Poca_{\not\equiv-1}(x)$ with some control
state $q_{\not\equiv -1}$ such that for each $\sigma:\{x\}\rightarrow\N$
 $$(\Poca_{\not\equiv -1}^\sigma,q_{\not\equiv-1}(n)) \models
 \varphi_\test \quad\Longleftrightarrow\quad n\not
  \equiv -1 \textsl{ mod } \sigma(x).$$
\item There exists some fixed $\poca$ $\Poca_{<x}(x)$ with some control
state $q_{<x}$ such that for each $\sigma:\{x\}\rightarrow\N$
 $$(\Poca_{< x}^\sigma,q_{<x}(n)) \models
 \varphi_\test \quad\Longleftrightarrow\quad n< \sigma(x).$$

\end{enumerate}
\end{lemma}
\begin{proof}
We choose $\varphi_\test=\neg\EF\prop$, where $\prop$ is an atomic
proposition.
Below we depict $\Poca_{\not\equiv0}(x)$,
$\Poca_{\not\equiv-1}(x)$, and $\Poca_{<x}(x)$, respectively.
\begin{center}
\begin{picture}(85,20)(-20,-3.5)
%   \put(-20,-28){\framebox(85,33){}}
  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
 % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
 % \node(A)(0,0){$q_0$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

\put(-65,0){
\put(41,-2){\small $\prop$}
  \node(A)(20,0){$q_{\not\equiv0}$}
  \node(B)(40,0){$\bullet$}
\drawloop(A){$-x$}
\drawedge(A,B){$\zero$}
}
%%% AUTOMATON 2

\put(-27,0){
\put(61,-2){\small $\prop$}
  \node(A)(20,0){$q_{\not\equiv-1}$}
  \node(B)(40,0){$\bullet$}
  \node(C)(60,0){$\bullet$}
\drawloop(B){$-x$}
\drawedge(A,B){$+1$}
\drawedge(B,C){$\zero$}
}

%%% AUTOMATON 3

\put(32,0){
\put(61,-2){\small $\prop$}
  \node(A)(20,0){$q_{< x}$}
  \node(C)(40,0){$\bullet$}
  \node(D)(60,0){$\bullet$}
\drawloop(A){$-1$}
\drawedge(A,C){$-x$}
\drawedge(C,D){$\zero$}
%\drawedge[ELside=r](C,D){$\zero$}
}

\end{picture}
\end{center}
Correctness of the lemma is immediate.
\qed
\end{proof}

\newcommand{\up}{\mathsf{up}}

\noindent
Let us fix some instance of the emptiness problem for
 two-counter automata:
Let $\langle\Two,q_0,q_1\rangle$ be an instance to the
emptiness problem, where
$\Two=(Q,E,\lambda)$ is some two-counter automaton.
We will give a fixed $\ctl$ formula $\varphi$ such that from
$\langle\Two,q_0,q_1\rangle$ we can construct some $\poca$ $\Poca(x)$ such that
$(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ if and only if
$(\Poca^\sigma,q_0(0))\models\varphi$ for some $\sigma:\{x\}\rightarrow\N$.

The control states and transitions will of $\Poca(x)$ can be described by the
following graph transformation rule that maps each pair of control states
$(p,q)\in E$ of $\Two$ with $\lambda(p,q)=l$
 into corresponding control states in $\Poca(x)$, by
possibly accessing the different $\poca$ of Lemma \ref{L POCAs}.
\begin{center}
\begin{picture}(85,30)(30,-10)
%   \put(-20,-28){\framebox(85,33){}}
  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
 % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
 % \node(A)(0,0){$q_0$}

\gasset{Nframe=n,loopdiam=9,ELdist=.7}
\gasset{Nadjust=wh,Nadjustdist=1}
\gasset{curvedepth=0}

\put(60,6){\Huge$\hookrightarrow$}

%\put(41,-2){\small $\prop$}
  \node(A)(20,10){$p$}
  \node(B)(40,10){$q$}

\drawedge(A,B){$l$}

\put(109,12){\small$l$}

  \node(C)(90,10){$p$}
  \node(D)(110,10){$\bullet$}
  \node(E)(130,10){$q$}
  \node(F)(110,-10){$\test(l)$}

%\drawloop(A){$-x$}
\drawedge(C,D){$0$}
\drawedge(D,E){$\up(l)$}
\drawedge(D,F){$0$}
\end{picture}
\end{center}

\noindent
where
$$
\test(l)\quad =\quad \begin{cases} (\Poca_{\not\equiv 0}(x),q_{\not\equiv 0}) & \text{ if }\quad
l\in\{ \add_1(-1),\zero_1\}\\
(\Poca_{\not\equiv -1}(x),q_{\not\equiv -1})& \text{ if }\quad l=\add_1(+1)\\
(\Poca_{<x}(x),q_{<x})& \text{ if }\quad l=\zero_2,
\end{cases}
$$

\noindent
and where
$$
\up(l)\quad =\quad \begin{cases}  a & \text{ if }\quad
l=\add_1(a)\\
\circ x& \text{ if }\quad l=\add_2(\circ 1)
\text{ for some }\circ\in\{+,-\}\\
0& \text{ otherwise.}
\end{cases}
$$

\noindent
Let us moreover ensure that every control state of the kind
$q_{\tau}$ with $\tau\in\{\not\equiv 0,\not\equiv\! -1,\geq x\}$ has
the label $\rho_\tau$.
Before we give our final formula, let us introduce the following constraint
formulas that guarantee that an overflow never occurs:
\begin{eqnarray*}
\psi_1&\quad=\quad& \add_1(+1)\longrightarrow \EX(\rho_{\not\equiv-1}
\wedge\varphi_\test)\
\\
\psi_2&\quad=\quad&
\add_1(-1)\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\varphi_\test)\\
\psi_3&\quad=\quad&
\zero_2\longrightarrow\EX(\rho_{<x}\wedge\varphi_{\test})\\
\psi_4&\quad=\quad&
\zero_1\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\neg\varphi_\test)
\end{eqnarray*}

\noindent
Note that we do not need any constraint when changing the second counter.
Let us introduce a label $f$ at control state $q_1$.
Our final $\ctl$ formula $\varphi$ is
$$
\varphi\quad=\quad\E\left(\bigwedge_{i\in[4]}\psi_i\ \U\  f\right).
$$
\iffalse
The correctness of the reduction follows from the following equivalences:
\begin{itemize}
\item $(T(\Poca),(q_0,0))\models \neg \varphi$
\item for all $\sigma:\{x\}\rightarrow \N$ we have $(\Poca^\sigma(x),q_0(0))\not\models\varphi$
\item for some $\sigma:\{x\}\rightarrow\N$ we have $(\Poca^\sigma(x),q_0(0))\models\varphi$
\item for some $\exists\sigma:\{x\}\rightarrow\N$ there is some path
$r_0(0)\rightarrow r_1(n_1)\cdots r_k(n_k)$ such that
$r_0=q_0$, $r_k=q_1$, $r_k\in F$ and
$(T(\Poca^\sigma(x)),r_i(n_i)

%$(q_0,0)(q_1,n_1\mod\sigma(x),n_1\div\sigma(x))\cdots (q_k,n_k\mod\sigma(x),
%n_k\div \sigma(x))$ in $T(\Poca^\sigma(x))$ such that

\end{itemize}
Then there is an assignment
$\sigma$ such that $(T(\Poca^\sigma),(q_0,0))\models \varphi$. It is
easily verified that the witnessing path $(q,0)(q_1,n_1)\ldots
(q',n')$ in $T(\A'(\sigma))$ corresponds to the path
$(q,0,0)(q_1,n_1\mod \sigma(x), n_1\div \sigma(x))\ldots (q',n'\mod
\sigma(x), n'\div \sigma(x))$ in $T(\A)$. Conversely, if
$(q,0,0)\rightarrow^* (q,n_1,n_2)$ in $T(\A)$, then the first counter
of $\A$ does not exceed some value $u \in \naturals$. Consequently, by
setting $\sigma(x):=u$, $(T(\A(\sigma)),(q,0))\models \varphi$.
\fi
We have that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ in $G(\A)$  if and only
if
$(T(\Poca^\sigma(x)),q_0(0))\models\varphi$ for some
$\sigma:\{x\}\rightarrow\N$.
We obtain the following theorem.

\begin{theorem}
  The data and combined complexity of $\CTL$ and the modal $\mu$-calculus
 on POCA is $\Pi^0_1$-complete.
\end{theorem}

\section{$\ltl$ Model Checking}{\label{S LTL}}

%\subsection{Linear-time logic ($\ltl$)}
Formulas of $\ltl$ are inductively defined according to the
following grammar, where $\prop$ ranges over $\Prop$:
\begin{equation*}
  \varphi\quad ::=\quad \prop \mid \neg \varphi \mid \varphi \wedge \varphi \mid
  \X \varphi \mid \varphi \until \varphi
\end{equation*}
The Boolean abbreviations and the formula $\true$ are defined in the same way as
in CTL. The \emph{finally modality} $\eventually
\varphi$ is an abbreviation for $\true \until \varphi$ and the
\emph{globally modality} $\globally \varphi$ abbreviates $\neg
\eventually \neg \varphi$.

The semantics of LTL is given in terms of infinite paths in a
transition system. Let $T=(S,\{S_\prop \mid \prop \in
\Prop\},\rightarrow)$ be a transition system, let $\pi: s_0
\rightarrow s_1 \rightarrow \cdots$ be an infinite path in $T$, and
let $\varphi$ be an $\ltl$ formula, we define $(T,\pi)\models\varphi$
by induction of the structure of $\varphi$, where $\prop$ ranges over
$\Prop$:
\begin{align*}
  (T,\pi) \models  \prop &  \quad\iff\quad  \pi(0) \in S_\prop\\
  (T,\pi) \models  \neg \varphi &\quad\iff\quad (T,\pi) \not \models \varphi\\
  (T,\pi) \models \varphi_1 \wedge \varphi_2 &\quad\iff\quad (T,\pi)
  \models \varphi_1 \text{ and } (T,\pi) \models \varphi_2\\
  (T,\pi) \models \X \varphi & \quad\iff\quad (T,\pi^1) \models \varphi\\
  (T,\pi) \models  \varphi_1 \until \varphi_2 & \quad\iff\quad
  \exists j\ge 0: (T,\pi^j)\models \varphi_2
  \text{ and } \forall 0\le i<j: (T,\pi^i)\models \varphi_1
\end{align*}
Let us now define the $\ltl$ model checking problem on SOCA
and POCA.\\

\problemx{$\ltl$ model checking on SOCA}
{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary,
  and an $\ltl$ formula $\varphi$.}
{$(T(\Soca),\pi)\models\varphi$ for every infinite
path $\pi$ with $\pi(0)=q(n)$?}
\

\problemx{$\ltl$ model checking on POCA}
{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and an
$\ltl$ formula $\varphi$.}
{$(T(\Poca^\sigma),\pi)\models\varphi$ for every $\sigma:X\rightarrow\N$ and
every infinite path $\pi$ with $\pi(0)=q(n)$?}

\subsection{Upper bounds}

A standard approach to $\ltl$ model checking is the automata-based
approach, in which systems are modeled as non-deterministic B\"uchi
automata (NBA). Given an NBA $\automaton$ and an $\ltl$ formula
$\varphi$, the idea is to translate $\varphi$ into an NBA
$\automaton_{\neg \varphi}$ such that $\automaton_{\neg \varphi}$
accepts all words that violate $\varphi$.  Then, by checking for
emptiness of the product automaton
$\automaton'=\automaton\times\automaton_{\neg \varphi}$, it can be
decided whether or not $\automaton$ generates traces that violate
$\varphi$.  Emptiness can be decided by checking for recurrent
reachability of a control state in the transition system induced by
$\automaton'$, which can be performed non-deterministically in space logarithmically in the size of
$\automaton'$. Vardi and Wolper showed in \cite{VW86} that the size of
$\automaton_{\neg \varphi}$ is $2^{O(|\varphi|)}$, which yields that
checking for emptiness of $\automaton'$ is in $\PSPACE$. The $\PSPACE$
lower bound for finite state model checking was originally shown
in \cite{LP85-popl} and hence carries over to SOCA.

Constructing the product automaton of two NBA is a standard technique
in model checking, see e.g.\ \cite{BK08} for a detailed treatment of
this. Moreover, this construction can be adapted in a
straight-forward way into the setting of POCA by equipping a POCA with
a set of final states and defining emptiness with respect to B\"uchi
acceptance condition in the standard way. For brevity, we do not give
any further details. We then have that given a POCA $\Poca$, an $\ltl$
formula $\varphi$, a state $q$ and $n\in\N$ we can construct a
POCA $\Poca'$ as the product of $\Poca$ and $\automaton_{\neg\varphi}$
such that for all assignments $\sigma$ we have that $\Poca'$ is empty
with respect to B\"uchi acceptence condition if and only if
$(T(\Poca^\sigma),q(n))\models \varphi$. Moreover, the size of
$\Poca'$ is $2^{O(|\varphi|)}|\Poca|$.

It has been shown in \cite{HKOW10} that checking emptiness with
respect to B\"uchi acceptance condition is $\coNP$-complete for both
SOCA and POCA, and in \cite{DG-jlc09} that it is $\NL$-complete for
OCA. We use these results in order to establish upper bounds for our
$\ltl$ model checking problems. Let us first consider the case of a
fixed formula $\varphi$. In the terminology of the previous paragraph,
$|\Poca'|=O(|\Poca|)$ and hence the data complexity of $\ltl$ on SOCA
and POCA is in $\coNP$.  The matching $\coNP$ lower bound of the data
complexity of LTL on SOCA follows from the $\NP$-hardness of the
reachability problem for SOCA.
%We obtain $\coNP$-hardness from the fact that
%we can reduce deciding emptiness of some SOCA $\Soca$ with respect to
%B\"uchi acceptance condition to $\ltl$ model checking of a slightly
%altered SOCA $\Soca'$ against an easily constructed fixed $\ltl$
%formula.

\begin{proposition}
The data complexity of $\ltl$ model checking on SOCA and POCA is
  $\coNP$-complete.
\end{proposition}
As seen above, if both $\Poca$ and the formula $\varphi$ are part of
the input then $|\Poca'|=2^{O(|\varphi)}\cdot|\Poca|$, and hence $\coNEXP$
is an upper bound for the combined complexity of $\ltl$ model checking
on both SOCA and POCA. However, we can improve this upper bound in the
case of SOCA. Given a SOCA $\Soca$, let $m$ be the absolute value of
the maximal increment or decrement performed on the transitions in
$\Soca$ and let $\Soca'$ be the product automaton of $\Soca$ and
$\automaton_{\neg \varphi}$. Let $\Soca''$ be obtained from $\Soca'$
by replacing every transition labeled with $z$ with a sequence of
fresh transitions and control locations of length $z$, where each
transition is labeled with $+1$ respectively $-1$, depending on the
sign of $z$. We conclude that
$|\Soca''|=m\cdot|\Soca'|=m\cdot2^{O(|\varphi|)}\cdot|\Soca|$. Combining the result
from \cite{DG-jlc09} together with the fact that $\ltl$ model checking
of non-deterministic finite state automata is $\PSPACE$-hard
\cite{LP85-popl}, we obtain the following proposition.

\begin{proposition}
The combined complexity of $\ltl$ on SOCA is $\PSPACE$-complete.
\end{proposition}
It remains to prove that the combined complexity of $\ltl$ model
checking on POCA is $\coNEXP$-hard, which we will show in the next
section.

\subsection{Combined Complexity of $\ltl$ on POCA}

We are now going to show $\coNEXP$-hardness of $\ltl$ model checking
on POCA via a reduction from (the complement of) \textsf{Succinct
  3-SAT}, which is a $\NEXP$-complete problem \cite{Papa94}. An input
of \textsf{Succinct 3-SAT} is given by a Boolean circuit $\mathbb{C}$
that encodes a Boolean formula $\psi_\circuit$ in 3-CNF,
i.e.\ $\psi_\circuit=\bigwedge_{j\in [0,M]}(\ell^j_1\vee \ell^j_2\vee
\ell^j_3)$ for some $M=O(2^{|\circuit|})$, and $\psi_\circuit$ is free
over Boolean variables $y_1,\ldots,y_{N}$ for some
$N=O(2^{|\circuit|})$. Let $j\in [0,M]$ be the index of a clause
encoded in \emph{binary} and $k\in\{1,2,3\}$. On input $(j\cdot k)$,
the output of $\circuit$ is $(i\cdot b)$, where $i\in [N]$ is the
index of the Boolean variable that appears in literal $\ell_k^j$, and
where $b=0$ when $\ell_k^j$ is negative and $b=1$ when $\ell_k^j$ is
positive.
\vspace*{0.25cm}
\problemx{\textsf{Succinct 3-SAT}}
{Boolean circuit $\circuit$}
{Is $\psi_\circuit$ satisfiable?}
\vspace*{0.25cm}

%\medskip

\noindent
In order to establish $\coNEXP$-hardness for the combined complexity
of $\ltl$ model checking, given an input $\circuit$ of
$\textsf{Succinct 3-SAT}$, we construct a POCA $\Poca(x)$ and an
$\ltl$ formula $\varphi$ such that $\psi_\circuit$ is satisfiable if
and only if there is an assignment $\sigma$ such that
$(T(\Poca^\sigma),q_{\mathsf{start}}(0))\models\varphi$ for some
distinguished state $q_{\mathsf{start}}$ of $\Poca(x)$.
%There are two key
%ideas for the reduction.

First, let us provide a suitable encoding of truth assignments by
natural numbers. The encoding we use has also been employed for
establishing lower bounds for model checking OCA~\cite{JKMZ04-ic}.
Recall that $p_i$ denotes the $i^{\text{th}}$ prime number. Every
natural number $x$ defines a truth assignment
$\nu_x:\{y_1,\ldots,y_N\} \to \{ 0,1\}$ such that $\nu_x(y_i)=1$ iff
$p_i$ divides $x$.  By the Prime Number Theorem, $p_N=O(N \log N)$ and
hence $O(|\circuit|)$ bits are sufficient to represent $p_N$. Of
course, since we need exponentially many prime numbers they cannot be
hard-wired into $\Poca(x)$.

%% Second, we employ $\ltl$ in order to perform arbitrary computations of
%% space-bounded Turing machines on traces of runs of SOCA. However, we
%% cannot make use of the approach we employed in $\ctl$. Intuitively
%% speaking, $\ctl$ allows for performing tests on the counter value
%% without destroying its value and this property has no correspondant in
%% the $\ltl$ setting. We can however use atomic propositions to model
%% computations along traces.

\begin{figure}
  \begin{center}
    \begin{picture}(142,60)(0,0)
      %   \put(-20,-28){\framebox(85,33){}}
      %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
      % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
      % \node(A)(0,0){$q_0$}

      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
      \gasset{Nadjust=wh,Nadjustdist=1}
      \gasset{curvedepth=0}

      \put(0,25){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(6,7){$\Soca_{\mathsf{inc}}(j)$}
      }

      \put(35,25){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,8){$\Soca_{\circuit}(j\cdot 2)$}
        \put(10,4){$=(i\cdot b)$ }
      }
      \put(35,45){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,8){$\Soca_{\circuit}(j\cdot 1)$}
        \put(10,4){$=(i\cdot b)$ }
      }
      \put(35,5){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,8){$\Soca_{\circuit}(j\cdot 3)$}
        \put(10,4){$=(i\cdot b)$ }
      }

      \put(70,25){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,6){$\Soca_{\mathsf{prime}}(i)=p_i$ }
      }

      \put(110,35){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,6){$\Soca_{\mathsf{divides}}(p_i,x)$ }
        \put(8,11){$b=0?$}
      }

      \put(110,10){
        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
        \put(3,6){$\Soca_{\mathsf{not\_divides}}(p_i,x)$ }
        \put(8,11){$b=1?$}
      }

      \node(A1)(12,29){}
      \node(A2)(20,32){}

      \node(B11)(40,8){}
      \node(B12)(55,15){}

      \node(B21)(40,28){}
      \node(B22)(55,35){}

      \node(B31)(40,48){}
      \node(B32)(55,55){}

      \node(C1)(73,36){}
      \node(C2)(90,28){}

      \node(D11)(115,39){}
      \node(D12)(132,39){}

      \node(D21)(113,13){}
      \node(D22)(132,13){}

      \node(E)(140,0){$\bullet$}
      \node(F)(12,0){$q_{\mathsf{start}}$}

      \drawedge(F,A1){$+x$}

      \drawedge(A2,B11){$0$}
      \drawedge(A2,B21){$0$}
      \drawedge(A2,B31){$0$}

      \drawedge(B12,C1){$0$}
      \drawedge(B22,C1){$0$}
      \drawedge(B32,C1){$0$}

      \drawedge(C2,D11){$0$}
      \drawedge(C2,D21){$0$}

      \drawqbedge(D12,142,25,E){$0$}
      \drawedge(D22,E){$0$}

      \drawedge(E,F){$\mathsf{zero}$}
    \end{picture}

  \end{center}
  \caption{High-level description of the automaton $\Poca(x)$ used for
    the reduction.}
  \label{fig:high-level}
\end{figure}

Let us now take a look at $\Poca(x)$, which is sketched in Figure
\ref{fig:high-level}. It uses one parameter $x$ and employs several
gadgets. The only gadgets manipulating the counter are
$\Soca_{\mathsf{divides}}$ and $\Soca_{\mathsf{not\_divides}}$. The
remaining gadgets are designed such that they communicate via
designated propositional variables, and not as in Section \ref{S CTL
  hardness} with the help of the counter. First, $\Poca(x)$ loads the
value of the parameter $x$ on the counter.  Think of $x$ encoding a
truth assignment of $\psi_\circuit$.  Next, $\Poca(x)$ traverses
through $\Soca_{\mathsf{inc}}$, which initially chooses an arbitrary
index $j$ identifying a clause of $\psi_\circuit$. Every time
$\Soca_{\mathsf{inc}}$ is traversed afterwards, it increments $j$
modulo $M+1$ and hereby moves on to the next clause.  Now $\Poca(x)$
branches non-deterministically into a gadget $\Soca_\circuit$ in order
to compute $(i\cdot b)$ from $\circuit$ on input $(j\cdot 1), (j\cdot
2)$, resp.\ $(j\cdot 3)$. The index $i$ is then used as input to a
gadget $\Soca_{\mathsf{prime}}$, which computes $p_i$. Then if $b=0$,
it is checked in $\Soca_{\mathsf{not\_divides}}$ that $p_i$ does not
divide the value of $x$, and likewise in $\Soca_{\mathsf{divides}}$
that $p_i$ divides the value of $x$ if $b=1$. For checking the latter,
the counter needs to be modified. After the checks have been finished,
we restore the value $x$ on the counter and the process continues with
clause $j+1~ \mod M+1$.

It remains to show how the gadgets and the communication between them
can be realized. The first observation is that the computations of
$\Soca_{\mathsf{inc}}$, $\Soca_{\circuit}$ and
$\Soca_{\mathsf{prime}}$ can be realized by space bounded DTM using no
more than $p(|\circuit|)$ tape cells for their in- and output tape for
some polynomial $p$ that is fixed once $\circuit$ is provided. Indeed,
it is easily seen that incrementing modulo $M+1$, evaluating
$\circuit$ and computing the $i^{\text{th}}$ prime number can be done
by such a DTM. Thus, we now show how from a generic DTM $\M$, we can
construct in logarithmic space a SOCA $\Soca_\M$ and some LTL formulas
that mimic computations of $\M$ on traces of $\Soca_\M$.

\begin{figure}
  \begin{center}
    \begin{picture}(140,100)(0,-52)
      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
      \gasset{Nadjust=wh,Nadjustdist=1}
      \gasset{curvedepth=0}

      \node(DASH)(20,40){}
%      \node(STATEIN)(-5,25){}
      \node(START)(0,25){}
      \node(A)(10,25){$\bullet$}
      \node(B)(20,25){$q$}
      \node(C1)(30,45){$\bullet$}
      \node(C2)(30,35){$\bullet$}
      \node(C3)(30,25){$\vdots$}
      \node(C4)(30,15){$\bullet$}
      \node(C5)(30,5){$\bullet$}
      \node(D)(40,25){$\bullet$}
%      \node(STATEOUT)(70,25){}

%      \drawedge[dash={1.5}0](DASH,B){}
%      \drawedge(STATEIN,A){}
      \drawedge(START,A){}
      \drawedge(A,B){}
      \drawedge(B,C1){}
      \drawedge[ELside=r](B,C2){}
%     \drawedge(B,C3){$0$}
      \drawedge(B,C4){}
      \drawedge[ELside=r](B,C5){}
      \drawedge(C1,D){}
      \drawedge[ELside=r](C2,D){}
%     \drawedge(C3,D){$0$}
      \drawedge(C4,D){}
      \drawedge[ELside=r](C5,D){}
%      \drawedge(D,STATEOUT){}

      \put(9,21){$\zeta$}
      \put(18,21){$\alpha$}
      \put(29,47){$\mathsf{s}_0$}
      \put(29,37){$\mathsf{s}_1$}
      \put(27,12){$\mathsf{s}_{k-1}$}
      \put(29,2){$\mathsf{s}_k$}
%      \put(60,21){$\mathsf{s}_\r$}

%%%%%%%%%%%%%%%%%% ihead

%      \node(DASH)(20,40){}
%      \node(STATEIN)(-5,25){}
%      \node(A)(5,25){$\bullet$}
      \node(E)(50,25){$\bullet$}
      \node(F1)(60,45){$\bullet$}
      \node(F2)(60,35){$\bullet$}
      \node(F3)(60,25){$\vdots$}
      \node(F4)(60,15){$\bullet$}
      \node(F5)(60,5){$\bullet$}
      \node(G)(70,25){$\bullet$}
%      \node(STATEOUT)(70,25){}

      \drawedge(D,E){}

%     \drawedge[dash={1.5}0](DASH,B){}
%      \drawedge(STATEIN,A){}
%      \drawedge(A,B){$0$}
      \drawedge(E,F1){}
      \drawedge[ELside=r](E,F2){}
%     \drawedge(B,C3){$0$}
      \drawedge(E,F4){}
      \drawedge[ELside=r](E,F5){}
      \drawedge(F1,G){}
      \drawedge[ELside=r](F2,G){}
%     \drawedge(C3,D){$0$}
      \drawedge(F4,G){}
      \drawedge[ELside=r](F5,G){}
%      \drawedge(D,STATEOUT){}

%      \put(18,21){$\alpha$}
      \put(58,47){$\mathsf{ih}_0$}
      \put(58,37){$\mathsf{ih}_1$}
      \put(57,12){$\mathsf{ih}_{m}$}
      \put(56,2){$\mathsf{ih}_{m+1}$}
%      \put(60,21){$\mathsf{s}_\r$}

%%%%%%%%%%%%%%%%%% whead

%      \node(DASH)(20,40){}
%      \node(STATEIN)(-5,25){}
%      \node(A)(5,25){$\bullet$}
      \node(E)(80,25){$\bullet$}
      \drawedge(G,E){}
      \node(F1)(90,45){$\bullet$}
      \node(F2)(90,35){$\bullet$}
      \node(F3)(90,25){$\vdots$}
      \node(F4)(90,15){$\bullet$}
      \node(F5)(90,5){$\bullet$}
      \node(G)(100,25){$\bullet$}
      \node(H)(100,0){$\bullet$}

%     \drawedge[dash={1.5}0](DASH,B){}
%      \drawedge(STATEIN,A){}
%      \drawedge(A,B){$0$}
      \drawedge(E,F1){}
      \drawedge[ELside=r](E,F2){}
%     \drawedge(B,C3){$0$}
      \drawedge(E,F4){}
      \drawedge[ELside=r](E,F5){}
      \drawedge(F1,G){}
      \drawedge[ELside=r](F2,G){}
%     \drawedge(C3,D){$0$}
      \drawedge(F4,G){}
      \drawedge[ELside=r](F5,G){}
      \drawedge(G,H){}

%      \put(18,21){$\alpha$}
      \put(88,47){$\mathsf{wh}_0$}
      \put(88,37){$\mathsf{wh}_1$}
      \put(87,12){$\mathsf{wh}_{n}$}
      \put(86,2){$\mathsf{wh}_{n+1}$}
%      \put(60,21){$\mathsf{s}_\r$}

      \node(I)(0,0){$\bullet$}
      \drawedge(H,I){}

%%%%%%%%%%%%%%%% wtape

%      \node(WORKIN)(0,45){}
      \node(A)(0,-15){$\bullet$}
      \drawedge(I,A){}
      \node(B1)(15,-25){$\bullet$}
      \node(B2)(15,-5){$\bullet$}
      \node(C)(30,-15){$\bullet$}
      \node(D1)(45,-25){$\bullet$}
      \node(D2)(45,-5){$\bullet$}
      \node(E)(60,-15){$\bullet$}
      \node(F)(70,-15){$\ldots$}
      \node(G)(80,-15){$\bullet$}
      \node(H1)( 95,-25){$\bullet$}
      \node(H2)(95,-5){$\bullet$}
      \node(I)(110,-15){$\bullet$}

      \node(J)(110,-30){$\bullet$}
      \node(K)(0,-30){$\bullet$}
      \drawedge(I,J){}
      \drawedge(J,K){}
%      \node(J)(135,-15){$\bullet$}
%      \node(DASH)(120,-25){}
%      \node(WORKOUT)(145,-15){}

%      \drawedge(WORKIN,A){}
      \drawedge(A,B1){}
      \drawedge(A,B2){}
      \drawedge(B1,C){}
      \drawedge(B2,C){}
      \drawedge(C,D1){}
      \drawedge(C,D2){}
      \drawedge(D1,E){}
      \drawedge(D2,E){}
      \drawedge(G,H1){}
      \drawedge(G,H2){}
      \drawedge(H1,I){}
      \drawedge(H2,I){}
%      \drawedge(I,J){}
%      \drawedge[dash={1.5}0](I,DASH){}

%      \put(4,42){$\mathsf{w}_\l$}
      \put(13,-9){$\mathsf{i}_{1,1}$}
      \put(13,-29){$\mathsf{i}_{1,0}$}
      \put(43,-9){$\mathsf{i}_{2,1}$}
      \put(43,-29){$\mathsf{i}_{2,0}$}
      \put(93,-9){$\mathsf{i}_{m,1}$}
      \put(93,-29){$\mathsf{i}_{m,0}$}
%      \put(120,42){$\mathsf{w}_\r$}
%      \put(133,42){$\mathsf{yield}$}

%%%%%%%%%%%%%%% wtape
      \node(A)(0,-45){$\bullet$}

      \drawedge(K,A){}

      \node(B1)(15,-55){$\bullet$}
      \node(B2)(15,-35){$\bullet$}
      \node(C)(30,-45){$\bullet$}
      \node(D1)(45,-55){$\bullet$}
      \node(D2)(45,-35){$\bullet$}
      \node(E)(60,-45){$\bullet$}
      \node(F)(70,-45){$\ldots$}
      \node(G)(80,-45){$\bullet$}
      \node(H1)( 95,-55){$\bullet$}
      \node(H2)(95,-35){$\bullet$}
      \node(I)(110,-45){$\bullet$}
      \node(J)(120,-45){$\bullet$}
      \node(K)(120,50){$\bullet$}
      \node(L)(20,50){$\bullet$}
      \node(M)(20,25){}
      \node(N)(130,-45){$\bullet$}
      \node(P)(140,-45){}

      \drawedge(I,J){}
      \drawedge(J,K){}
      \drawedge(K,L){}
      \drawedge(L,M){}

      \drawedge(J,N){}
      \drawedge(N,P){}

%      \node(DASH)(120,-25){}
%      \node(WORKOUT)(145,-15){}

%      \drawedge(WORKIN,A){}
      \drawedge(A,B1){}
      \drawedge(A,B2){}
      \drawedge(B1,C){}
      \drawedge(B2,C){}
      \drawedge(C,D1){}
      \drawedge(C,D2){}
      \drawedge(D1,E){}
      \drawedge(D2,E){}
      \drawedge(G,H1){}
      \drawedge(G,H2){}
      \drawedge(H1,I){}
      \drawedge(H2,I){}
%      \drawedge(I,J){}
%      \drawedge[dash={1.5}0](I,DASH){}

%      \put(4,42){$\mathsf{w}_\l$}
      \put(12,-39){$\mathsf{w}_{1,1}$}
      \put(12,-58){$\mathsf{w}_{1,0}$}
      \put(42,-39){$\mathsf{w}_{2,1}$}
      \put(42,-58){$\mathsf{w}_{2,0}$}
      \put(92,-39){$\mathsf{w}_{n,1}$}
      \put(93,-58){$\mathsf{w}_{n,0}$}

      \put(119,-49){$\gamma$}
      \put(129,-49){$\beta$}

%      \put(120,42){$\mathsf{w}_\r$}
%      \put(133,42){$\mathsf{yield}$}

    \end{picture}
  \end{center}
  \caption{SOCA $\Soca_\M$ for the simulation of space bounded Turing
    machines.}
  \label{fig:tm}
\end{figure}

Let $\M=(S, \Sigma, s_0, F, \mu)$ be a DTM with a fixed input tape
with $m$ tape cells, and $n$ working tape cells, and let
$S=\{s_0,\ldots,s_k\}$. Figure \ref{fig:tm} shows the SOCA $\Soca_\M$
that we use for the simulation of $\M$. All transitions of $\Soca_\M$
are implicitly labelled with $0$.  A simulation starts when $\Soca_\M$
is entered at the location labeled with $\zeta$ and is finished when
the location labeled with $\beta$ is reached.

The sequence of propositions occurring on a trace starting from and
ending in $q$ encodes a configuration of $\M$.  In detail,
$\mathsf{s}_i$ indicates that $\M$ is in state $s_i$; $\mathsf{ih}_i$
that the input tape head scans cell $i$; $\mathsf{wh}_i$ that the
working tape head scans cell $i$; $\mathsf{i}_{i,b}$ that the
$i^{\text{th}}$ bit of the input tape is set to $b$; and
$\mathsf{w}_{i,b}$ that the $i^{\text{th}}$ bit of the working tape is
set to $b$, where $i$ is in the respective range and $b\in \{0,1\}$.

\newcommand{\state}{\mathsf{state}}
\newcommand{\work}{\mathsf{work}}
\newcommand{\iNp}{\mathsf{input}}
\newcommand{\inh}{\mathsf{inhead}}
\newcommand{\woh}{\mathsf{wohead}}
\renewcommand{\succ}{\mathsf{succ}}

Let us introduce some LTL formulas that allow for testing properties
of the current configuration. Think of all of them as being evaluated
in $q$. The formula $\state_i=\X \mathsf{s}_i$ for each $i\in[0,k]$
expresses that the current state is $s_i$.  By the formula
$\inh_{i}=\X\X\X\X \mathsf{ih}_i$ we express that the input head is at
position $i$, where $i\in[m]$. Similarly, define the formulas
$\woh_i$, $\work_{j,b}$, and $\iNp_{i,b}$ for expressing that the
working head is at position $i$, that the $i^{\text{th}}$ bit of the
input tape is $b$, and that the $j^{\text{th}}$ bit of the working
tape is $b$, respectively, where $i\in[m]$,$j\in[n]$, and
$b\in\{0,1\}$.

The LTL formula below, assumed to be evaluated in $\alpha$, ensures
that the transition function is correctly encoded into traces of
$\Soca_\M$ for states $s\in S\setminus F$ whenever the input
respectively working tape head does not scan a start ($\l$)
respectively end marker ($\r$):

\begin{multline*}
\bigwedge_{s_l\in S\setminus F}
\bigwedge_{i\in[m]}
\bigwedge_{j\in[n]}
\bigwedge_{b_1,b_2\in\{0,1\}}
\state_l\wedge\inh_i\wedge\iNp_{i,b_1}\wedge\woh_j\wedge
\work_{j,b_2} \longrightarrow\\
\longrightarrow
\left(
\left(\X(\neg\alpha \wedge \neg \beta)
\U(\alpha\wedge \succ(s_l,i,j,b_1,b_2)\right)
\wedge
\bigwedge_{j'\not= j}\bigwedge_{b\in\{0,1\}}
\left(\work_{j',b}\leftrightarrow
(\X\neg\alpha\U(\alpha\wedge\work_{j',b}))\right)
\right).
\end{multline*}
Here, whenever $\mu(s_l,b_1,b_2)=(s_h,d_1,d_2,b)$, the formula
\begin{align*}
\succ(s_l,i,j,b_1,b_2)\quad =\quad\state_h\wedge\inh_{i+d_1}\wedge\woh_{j+d_2}\wedge
\work_{j,b}
\end{align*}
guarantees that the correct bit is ``written'' to the working tape and
that the state, the input head position, and the working tape position
of the next configuration seen indeed match the successor
configuration. A similar formula can be constructed for the case when
one or both of the input or working heads point to a start
respectively end marker. Once we have reached a final state $s_i\in
F$, we require that $\Soca_\M$ is left which is expressed by the
following formula when evaluated in $q$:
\begin{align*}
   \bigwedge_{s_i\in F}\state_i \longrightarrow (\neg \alpha \until
   \beta).
\end{align*}
It is now easily seen that we can construct a formula $\Phi$ that is
derived from a conjunction of the formulas from above such that the
formula $\globally (\alpha \rightarrow \Phi)$ constraints paths
through $\Soca_\M$ in such a way that their traces yield the encoding
of a valid computation of $\M$.

Let us now address towards ensuring that once we enter $\Soca_\M$ we
initially traverse it in such a way that the trace corresponds to an
initial configuration of $\M$. The formula
\begin{align*}
  \globally\left(\zeta \longrightarrow \X(\state_0 \wedge \inh_1 \wedge
  \woh_1\wedge \bigwedge_{1\le j\le n}\mathsf{work}_{j,0} ) \right)
\end{align*}
makes sure that the heads of the input and working tape point to the
first tape cell, that the working tape is filled with $0$s and that we
are in the initial state. In case the input tape can be initialized
with an arbitrary content, we are done. Otherwise, suppose that we
want to transfer the first $j$ bits of the output of a TM $\M'$ from
its corresponding SOCA $\Soca_{\M'}$ to the input of $\Soca_\M$.  For
$b\in \{0,1\}$, let $\bar{b}=0$ if $b=1$ and $\bar{b}=1$ otherwise,
and suppose that all atomic propositions are primed in $\Soca_{\M'}$.
The formula
\begin{align*}
  \bigwedge_{1\le i\le j\atop{b\in \{0,1\}}}
  \globally
   ((\mathsf{w}'_{i,b} \wedge (\neg \alpha' \until \beta'))
  \longrightarrow (\neg \mathsf{i}_{i,\bar{b}}\until \gamma)
  %\right)
\end{align*}
guarantees that we traverse through the first $j$ bits of the
component of $\Soca_\M$ representing the input tape of $\M$ in the
same way as we traverse the first $j$ bits of the working tape
component of $\M'$ in $\Soca_{\M'}$ when a computation has finished.
Coming back to Figure \ref{fig:high-level}, we have thus seen how the
SOCA $\Soca_{\mathsf{inc}}, \Soca_{\circuit}$ and
$\Soca_{\mathsf{prime}}$ and the communication between them can be
realized. The only major question left open is how we can perform a
divisibility respectively non-divisibility test of the counter value
with a prime number computed in $\Soca_{\mathsf{prime}}$. For this,
let us consider the SOCA $\Soca_{\mathsf{divides}}$ from Figure
\ref{fig:soca-divisibility}.

\begin{figure}[t]
  \begin{center}
    \begin{picture}(140,30)(0,0)
      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
      \gasset{Nadjust=wh,Nadjustdist=1}
      \gasset{curvedepth=0}

      %% \node(WORKIN)(0,45){}
      %% \node(A)(10,45){$\bullet$}
      %% \node(B1)(25,55){$\bullet$}
      %% \node(B2)(25,35){$\bullet$}
      %% \node(C)(40,45){$\bullet$}
      %% \node(D1)(55,55){$\bullet$}
      %% \node(D2)(55,35){$\bullet$}
      %% \node(E)(70,45){$\bullet$}
      %% \node(F)(80,45){$\ldots$}
      %% \node(G)(90,45){$\bullet$}
      %% \node(H1)(105,55){$\bullet$}
      %% \node(H2)(105,35){$\bullet$}
      %% \node(I)(120,45){$\bullet$}
      %% \node(J)(135,45){$\bullet$}
      %% \node(DASH)(120,55){}
      %% \node(WORKOUT)(145,45){}

      %% \drawedge(WORKIN,A){}
      %% \drawedge(A,B1){$0$}
      %% \drawedge(A,B2){$0$}
      %% \drawedge(B1,C){$0$}
      %% \drawedge(B2,C){$0$}
      %% \drawedge(C,D1){$0$}
      %% \drawedge(C,D2){$0$}
      %% \drawedge(D1,E){$0$}
      %% \drawedge(D2,E){$0$}
      %% \drawedge(G,H1){$0$}
      %% \drawedge(G,H2){$0$}
      %% \drawedge(H1,I){$0$}
      %% \drawedge(H2,I){$0$}
      %% \drawedge(I,J){$0$}
      %% \drawedge[dash={1.5}0](I,DASH){}

      %% \put(4,42){$\mathsf{work}_\l$}
      %% \put(21,57){$\mathsf{work_{1,1}}$}
      %% \put(21,32){$\mathsf{work_{1,0}}$}
      %% \put(51,57){$\mathsf{work_{2,1}}$}
      %% \put(51,32){$\mathsf{work_{2,0}}$}
      %% \put(101,57){$\mathsf{work}_{N,1}$}
      %% \put(101,32){$\mathsf{work}_{N,0}$}
      %% \put(120,42){$\mathsf{work}_\r$}
      %% \put(133,42){$\mathsf{yield}$}

      \node(DIVIDESIN)(0,15){}
      \node(A)(10,15){$\bullet$}
      \node(B1)(25,25){$\bullet$}
      \node(B2)(25,5){$\bullet$}
      \node(C)(40,15){$\bullet$}
      \node(D1)(55,25){$\bullet$}
      \node(D2)(55,5){$\bullet$}
      \node(E)(70,15){$\bullet$}
      \node(F)(80,15){$\ldots$}
      \node(G)(90,15){$\bullet$}
      \node(H1)(105,25){$\bullet$}
      \node(H2)(105,5){$\bullet$}
      \node(I)(120,15){$\bullet$}
      \node(J)(135,15){$\bullet$}
      \node(DIVIDESOUT)(145,15){}

      \drawedge(DIVIDESIN,A){}
      \drawedge(A,B1){$0$}
      \drawedge(A,B2){$0$}
      \drawedge(B1,C){$-2^0$}
      \drawedge(B2,C){$0$}
      \drawedge(C,D1){$0$}
      \drawedge(C,D2){$0$}
      \drawedge(D1,E){$-2^1$}
      \drawedge(D2,E){$0$}
      \drawedge(G,H1){$0$}
      \drawedge(G,H2){$0$}
      \drawedge(H1,I){$-2^l$}
      \drawedge(H2,I){$0$}
      \drawedge(I,J){$\mathsf{\zero}$}
      \drawbpedge[ELside=r](I,290,28,A,250,28){$0$}

%      \put(4,12){$\mathsf{work}_\l$}
      \put(22,27){$\prop_{0,1}$}
      \put(22,2){$\prop_{0,0}$}
      \put(52,27){$\prop_{1,1}$}
      \put(52,2){$\prop_{1,0}$}
      \put(102,27){$\prop_{l,1}$}
      \put(102,2){$\prop_{l,0}$}
%      \put(120,12){$\mathsf{work}_\r$}
      \put(134,12){$\xi$}

    \end{picture}
  \end{center}
  \caption{The SOCA $\Soca_{\mathsf{divides}}$  for testing the
    counter for divisibility with some natural number of
    bit length $l+1$.}
  \label{fig:soca-divisibility}
\end{figure}

One cycle through $\Soca_{\mathsf{divides}}$ subtracts some natural
number of bit length $l+1$ from the counter. However, in order to test
for divisibility we need to make sure that we remain on the same path
in every cycle. In the $\CTL$ setting, this problem was resolved by
branching into the additional SOCA $\Soca_{\mathsf{bit}}$. In
contrast, in $\ltl$ we cannot branch, but use the propositions
$\prop_{j,b_j}, j\in [0,l], b\in \{0,1\}$ in order to stay on
precisely one path in every cycle. Assuming that the number $p$ for
which we want to test for divisibility with the current counter value
is encoded as a sequence of propositions $\mathsf{w}_{j,b_j}$ of some
SOCA $\Soca_\M$, the subsequent formula enforces that we always
subtract $p$ in cycles of $\Soca_{\mathsf{divides}}$:
\begin{align*}
  \globally \left( \bigwedge_{0\le j\le l \atop{b_{j}\in \{0,1\}}}
  ((\mathsf{w}_{j,b_j} \wedge (\neg \alpha \until \beta) ) \longrightarrow
  ( \neg \prop_{j,\bar{b_j}} \until \xi)) \right).
\end{align*}
It is straight forward to derive a similar SOCA
$\Soca_{\mathsf{not\_divides}}$ and an appropriate LTL formula for
testing non-divisibility of the counter value with a previously
computed prime number. Finally, we can also adopt these techniques in
order to correctly handle the branching on $b$ performed in Figure
\ref{fig:high-level}.

In summary, by taking the disjoint union of all the gadgets from
Figure \ref{fig:high-level}, their appendent $\ltl$ formula that we
described in this section, connecting the gadgets correctly and taking
the conjunction of the relevant $\ltl$ formulas, we can construct
$\Poca(x)$ and an $\ltl$ formula $\varphi$ in logarithmic space such
that there is an assignment $\sigma$ assigning a natural number to $x$
such that $(T(\Poca^\sigma), q_{\mathsf{start}}(0))\not\models
\varphi$ if and only if $\psi_\circuit$ given by an input $\circuit$
of $\textsf{Succinct 3-SAT}$ is satisfiable.

\begin{theorem}
  The combined complexity of $\ltl$ model checking on POCA is
  $\coNEXP$-complete.
\end{theorem}

\section{Conclusion}{\label{S Concl}}

In this paper, we have settled the computational complexity of model
checking CTL and LTL on SOCA and POCA with respect to data and
combined complexity. Our proofs for providing lower bounds have
introduced some involved concepts and techniques, which we believe may
be of independent interest for providing lower bounds for decision
problems in the verification of infinite state systems.

An interesting aspect of future work could be the consideration of
\emph{synthesis problems} for POCA. Given a POCA $\Poca(X)$ and an LTL
formula $\varphi$, a natural question to pose is whether there
\emph{exists an assignment} $\sigma$ such that $(T(\Poca^\sigma),
\pi)\models \varphi$ for all infinite paths $\pi$ starting in some
state of $T(\Poca^\sigma)$.  For CTL, such a problem is undecidable by
Theorem \ref{thm:ctl-poca}, but we claim that it might be decidable
for LTL. Moreover, the CTL fragment EF seems to be a good candidate of
a branching-time logic for which model checking on POCA could be
decidable, but this remains subject to further investigations.

\bibliography{bib}
\bibliographystyle{plain}

%\thebibliography

\iffalse

\begin{thebibliography}{1}

\bibitem{GoLo10}
S.~G\"oller, M.~Lohrey.
\newblock Branchning-time model checking of one-counter processes.
\newblock In Proceedings of STACS 2010, to appear
\newblock {\em Arxiv:0909.1102v1}

\end{thebibliography}
\fi

\end{document}

### 2022-10-14 definitions.tex
Uploading to have some file on the git

\section{Definitions}\label{section definitions}

In this section, we introduce general notations and preliminary definitions.

\problemx{$k$-Resilience problem}
{A state $s$ of a WSTS $(S,,)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
{$\forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}

\problemx{Bounded Resilience problem}
{A state $s$ of a WSTS $(S,,)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}

### 2022-10-14 definitions.tex
Definitions written up a bit now need to smooth them out

@@
 
-\problemx{$k$-Resilience problem}
-{A state $s$ of a WSTS $(S,,)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
+The model we are interested in is (S)WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
+
+\textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
+
+\subsection{Transition systems}
+
+\begin{definition}
+A {\em labeled transition system} (LTS for short) is a tuple $T=(S, \Lambda, \rightarrow)$ where 
+$S$ is a set of {\em configurations}, $ \Lambda$ is a set of {\em labels}, and 
+${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
+ternary relation,
+denoted as the set of {\em labeled transitions}. 
+\end{definition}
+
+We
+ prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
+       $s  \xrightarrow{a}  s'$
+to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
+
+\noindent
+Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
+
+A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
+to a {\em target configuration} $s_n$ is a sequence 
+$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
+We define the {\em concatenation} $ \pi_1 \pi_2$ of 
+two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
+as expected.
+The {\em length} of 
+$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
+is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
+For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
+
+An {\em infinite path} is an infinite sequence
+ $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
+For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
+(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
+and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
+by $\pi[i,j]$ the path 
+$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
+ and by $\pi[i]$ the configuration $s_i$.
+As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
+a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
+Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
+$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
+
+\newcommand{\pred}{\textsc{Pred}}
+\newcommand{\succ}{\textsc{Succ}}
+
+\noindent
+The set of {\em successors} of a configuration $s \in S$ is deﬁned as
+$\succ(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
+A configuration without successors is called a {\em dead end}.  \\
+The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
+$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.
+
+A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
+$a \in \Lambda$,
+ $ s_1  \xrightarrow{a} s_{2}$ and  
+ $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
+
+\begin{definition}
+An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
+{\em configurations} and  
+$ {\rightarrow} \subseteq S \times S$ is a
+binary relation 
+on
+the set of configurations, denoted as the set of {\em transitions}. 
+\end{definition}
+
+We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from configuration $s$ to configuration $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
+Note 
+ that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
+Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
+then defined as expected.
+
+\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
+
+We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
+for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\succ$ and $\pred$...
+
+\textcolor{red}{This makes sense for TS but not so much for LTS ...}
+
+A transition system is {\em finitely branching} if all $\succ(s)$ are finite. 
+\textcolor{blue}{We restrict our attention to finitely branching TSs.}
+
+\textcolor{red}{Not sure that TS induced by TA are finitely branching.
+Actually I believe they are not, i.e.
+for instance for a TA with one clock $x$, from a state $q$ and clock $x$ set at $0$, if there is a transition e.g.
+$(q, x \geq 3, \emptyset, q')$ then the set of successors of
+$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
+
+\subsection{Well-quasi-orderings}
+
+A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$.
+
+We abbreviate $x \leq y \not\leq x$ by $x < y$.
+
+% A partial ordering (a po) is an antisymmetric qo
+
+Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
+% and gives rise to a po between the equivalence classes ?
+% 	do we need po ?
+
+We now recall a few results from the theory of well-orderings (add reference [...]).
+
+\begin{definition}
+ A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0 ; x_1 ; x_2 ; : : : $ in $X$, there exist indexes $i \leq j$ with
+$x_i \leq  x_j$.
+\end{definition}
+
+Notice that a wqo is well-founded, i.e. it admits no infinite strictly decreasing sequence
+$x_0 > x_1 > x_2 > \cdots$
+
+\textcolor{red}{Add lemma about infinite increasing subsequences ?}
+
+Given $\leq$ a quasi-ordering over some set $X$, an {\em upward-closed set} is any set $I \subseteq X$ such that if $y \geq x$ and $x \in I$ then $y \in I $.
+A {\em downward-closed set} is any set $I \subseteq X$ such that if $y \leq x$ and $x \in I$ then $y \in I $. 
+To any $A \subseteq X$ we associate
+the {\em upward-closure of $A$} 
+ $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
+ and the 
+ {\em downward-closure of $A$} 
+ $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
+We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
+as $\uparrow x$ (resp. $\downarrow x$).
+
+\begin{lemma}{(Higman [40])} 
+If $\leq$ is a wqo; then any upward-closed $I$ has a
+basis.
+\end{lemma}
+
+\textcolor{red}{Expliquer ce que c'est une base d'abords.}
+
+\begin{proof}
+The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
+only contains a finite number of non-equivalent elements otherwise they would make
+an infinite sequence contradicting the wqo assumption.
+\end{proof}
+
+\textcolor{red}{Un lemme je pense c'est important de le noter, peut-être pas comme ça peut être noter différement faudra voir}
+
+\begin{lemma}  
+If $\leq$ is a wqo; any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
+upward-closed sets eventually stabilizes; i.e. there is a $k \in N$ such that 
+$I_k = I_{k+1} = I_{k+2} = \cdots $.
+\end{lemma}
+
+\begin{proof}
+Assue we have a counter-example ...
+\end{proof}
+
+\textcolor{red}{Define WSTS}
+
+\textcolor{red}{Define SWSTS - may be necessary}
+
+\subsection{Well-structured transition systems}
+
+\begin{definition}
+A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. SWSTS) 
+is a TS
+$(S, \rightarrow, \leq)$
+equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
+%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
+the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_1 ≤ t_2$  and $s_1 \rightarrow s_2$ , there exists 
+	$t_2 \in S$ with 
+	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
+				(resp. $t_1 \rightarrow^{1} t_2$ ).
+\end{definition}
+
+Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering.
+
+\textcolor{red}{For one-counter automata, in case the only tests are zero tests then 
+I supposed $\leq$ is $\leq$ for non-zero integers, and I'll have to look-up/think for what to do with the zero element for instance. For TA it seem kind of nontrivial (since they allow $<c$ tests).}
+
+\textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and
+'decidable $\leq$' being independant from the WSTS definition}
+
+\begin{definition}
+A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
+any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
+\end{definition}
+
+\textcolor{red}{Define what an Ideal is.
+Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
+Anti-ideal just downward closed so again just not that helpful a notation.
+Maybe have a }
+
+\begin{definition}
+An {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
+$\uparrow I = I = \downarrow I$.
+\end{definition}
+
+A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
+$s \in J$.
+
+"Bi-ideals often represent “control states” as in [cf \%]. "
+% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
+
+\textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
+
+Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
+
+\textcolor{red}{Do we still demand this ?}
+
+\begin{claim}{(stability of ideals)}
+Let $I, J \subseteq S$ be upward-closed. Then the sets 
+$\pred(I)$, $I \cup J$, and $I \cap J$ are upward-closed.
+\end{claim}
+
+% \begin{claim}
+% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
+% \end{claim}
+
+\subsection{Defining resilience}
+
+\subsubsection{TS resilience}
+
+\textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
+
+We ask whether we can reach a state 
+%which satisfies
+in 
+%
+$SAFE$  in a reasonable amount of time whenever we reach a state 
+% which satisfies
+in
+%
+$BAD$. 
+From this we formulate two resilience problems. First consider the case where the recovery time
+is bound by a given natural number $k \geq 0$, i.e., the explicit resilience problem for TS.
+
+\problemx{TS $k$-resilience problem}
+{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
+{$\forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+
+% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).
+
+We can also ask whether there exists such a bound $k$. We call this problem the bounded resilience problem for TS.
+
+\problemx{TS bounded resilience problem}
+{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
+{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+
+\subsubsection{WSTS resilience}
+
+Properties in well-structured transition systems are often given as upward- or downward closed sets [references].
+%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
+%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
+Transfering the abstract resilience problems into this framework,
+it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
+upward-closed or 
+downward-closed sets.
+
+We assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set.
+
+\textcolor{red}{Seems like a reasonable assumption to me.}
+
+From these considerations, we formulate instances of the abstract resilience problems for well-
+structured transition systems.
+
+Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
+
+\problemx{WSTS $k$-resilience problem}
+{A state $s$ of a WSTS $(S,\rightarrow)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
 {$\forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
 
-\problemx{Bounded Resilience problem}
+Analogously, we formulate the bounded resilience problem for WSTSs.
+
+\problemx{WSTS bounded resilience problem}
 {A state $s$ of a WSTS $(S,,)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
 {$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+
+\textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
+
+\subsection{Timed Automata}
+
+\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
+
+\subsection{One-Counter Automata}
+
+\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
+
+\subsection{Vector Addition System with States}
+
+\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}

### 2022-10-17 definitions.tex
up to date

@@
 \newcommand{\pred}{\textsc{Pred}}
-\newcommand{\succ}{\textsc{Succ}}
+\renewcommand{\succ}{\textsc{Succ}}
 
@@
 $s_1, t_1 , s_2 \in S$
-	with $s_1 ≤ t_2$  and $s_1 \rightarrow s_2$ , there exists 
+	with $s_1 \leq t_2$  and $s_1 \rightarrow s_2$ , there exists 
 	$t_2 \in S$ with 
@@
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+{$\forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 \problemx{WSTS $k$-resilience problem}
-{A state $s$ of a WSTS $(S,\rightarrow)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
-{$\forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
+{$\forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 \problemx{WSTS bounded resilience problem}
-{A state $s$ of a WSTS $(S,,)$, an ideal $I$ with a given basis, a decidable anti-ideal $J$.}
-{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rigtarrow^{\leq k} s''$ ?\newline}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
+{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 

### 2022-10-17 definitions.tex
Added some definitions for timed automata at the end

@@
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \to \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 
+\renewcommand{\A}{\mathcal{A}}
+\newcommand{\B}{\mathcal{B}}
+\renewcommand{\C}{\mathcal{C}}
+\newcommand{\Const}{\mathsf{Consts}}
+\newcommand{\Conf}{\mathsf{Conf}}
+\newcommand{\guards}{{\textsc{Guards}}}
+
+% \subsubsection{Guards, Clocks}
+
+A {\em guard} over a finite set of clocks $\Omega$ 
+is a comparison of the form
+$\omega \bowtie c$, where $ \omega \in \Omega$, $c \in \N$,
+and $\bowtie\in\{<,\leq,=,\geq,>\}$.
+%
+We denote by $\guards(\Omega)$ the {\em set of guards} over the set of 
+clocks $\Omega$.
+The {\em size} %$|g|$
+ of a guard 
+$g=\omega \bowtie c$ is defined as %:
+$|g|=\log(c)$.
+A {\em clock valuation} is a function from $\Omega$ to $\N$;
+we write $\vec{0}$ to denote the clock valuation $\omega \mapsto 0$
+whenever the set $\Omega$ is clear from the context.
+For each clock valuation $v$ and each $t\in\N$ we denote
+by $v+t$ the clock valuation $\omega \mapsto v(\omega)+t$.
+%
+For each guard $g=\omega \bowtie c$ with $c\in\N$,
+we write $v\models g$ if $v(\omega)\bowtie c$.
+
+\iffalse
+We define an {\em empty guard} $g_\epsilon$ over a non-empty finite set of clocks
+$\Omega$ and to be of the form $\omega \geq 0$ for some 
+$\omega \in \Omega$. In particular, we
+defined $g_\epsilon$ such that for all $v \in \N^\Omega$ 
+we have
+$v \models g_\epsilon$, hence $g_\epsilon$ can be used as a guard that is always true. 
+\fi
+
+% \subsubsection{Timed automata}
+
+A timed automaton is a finite automaton extended with a finite set of clocks $\Omega$ that all progress at the same rate and that can individually be reset to zero.
+Moreover, every transition is labeled by a guard over
+$\Omega$  and by a set of clocks to be reset. \\
+
+\par\noindent\ignorespacesafterend
+Formally, a {\em timed automaton} ({\em TA} for short) is a tuple
+$\A=(Q,\Omega,R,q_{init},F)$, where
+\begin{samepage}
+\begin{itemize}
+	\item $Q$ is a non-empty finite {\em set of states}, 
+	\item $\Omega$ is a non-empty finite {\em set of clocks},
+	\item $R \subseteq Q\times\G(\Omega)\times \mathscr{P}(\Omega) \times Q$
+	is a finite {\em set of  rules},
+	\item $q_{init}\in Q$ is an {\em initial  state}, and 
+	\item $F\subseteq Q$ is a {\em set of final states}.
+\end{itemize}
+\end{samepage}
+
+\par\noindent\ignorespacesafterend
+We also refer to $\A$ as an $n$-TA if $|\Omega| = n$. 
+The {\em size} of $\A$ is defined as%:
+$$
+|\A| \ = \ |Q|+|\Omega|+|R|+\sum_{(q,g,U,q')\in R}|g|.
+$$
+Let 
+$\Const(\A) = \{ c\in\N \mid \exists(q,g,U,q')\in R, \ \exists \omega \in \Omega, \bowtie\in\{<,\leq,=,\geq,>\} : g = \omega \bowtie c \}$ denote the 
+set of constants that appear in the guards of the rules of $\A$.
+
+By $\Conf(\A)=Q\times\N^\Omega$ we denote the set of
+{\em configurations} of $\A$. 
+%We prefer however to denote a configuration in $\Const(\A)$ by $q(v)$ instead of $(q,v)$.\\
+We prefer however to abbreviate a configuration	%	 in $\Conf(\A)$	
+$(q,v)$ by $q(v)$.
+
+\begin{samepage}
+%\begin{definition}
+A TA $\A=(Q,\Omega,R,q_{init},F)$ induces the labeled transition system 
+$T_{\A} =  (\Conf(\A), \Lambda_{\A}, \rightarrow_{\A})$
+where $ \Lambda_{\A} = R \times \N $
+and 
+where $ \rightarrow_{\A}$ is defined such that, 
+for all $(\delta,t)\in R\times\N$ with  	$\delta = (q,g,U,q')\in R$,
+for all $q(v), q'(v') \in \Conf(\A)$,
+$q(v)\xrightarrow{\delta,t}_{\A} q'(v')$ if
+	$v+t\models g$, 
+	 $v'(u)=0$ for all $u \in U$ and $v'(\omega)=v(\omega)+t$ for all 
+	$\omega \in \Omega \setminus U$.
+%\end{definition}
+\end{samepage}
+
+A {\em run} from $q_0(v_0)$ to $q_n(v_n)$ in $\A$ is a path in the transition system $T_{\A}$, that is,
+a sequence 
+$\pi = q_0(v_0)\xrightarrow{\delta_1,t_1}_{\A}q_1(v_1)\cdots\xrightarrow{\delta_n,t_n}_{\A}q_n(v_n)$;
+it is called {\em reset-free} if for all $i \in \{1,\ldots,n\}$,
+ $\delta_i = (g_i,\emptyset)$ for some guard $g_i$.
+
+We say $\pi$ is {\em accepting} if $q_0(v_0) = q_{init}(\vec{0})$ and $q_n \in F$. 
+\iffalse
+We say {\em reachability holds} for the TA $\A$ 
+if there exists an accepting run. %
+% if there is a run in $\A$ from $q_{init}(\vec{0})$   to some configuration $q(v)$ for some $q\in F$, and $v\in\N^\Omega$.
+%We refer to Figure~\ref{example pta} for an instance of a PTA for which reachability holds.
+\fi
+
+It is worth mentioning that there are further modes of time valuations and guards which exist in the literature, we refer
+to \cite{Andre19} for a recent overview. 
+%
+% \mh{Comment on difference between continuous and discrete time}
+Notably, we consider in this article only the case of timed automata over discrete time. It is worth mentioning that in
+the case of timed automata over continuous time (i.e. with clocks having values in $\R_{\geq 0}$),
+% However, for parametric timed automata with closed (i.e., non-strict) clock constraints and parameters restricted to ranging over integers, 1 standard digitisation techniques apply [HMP92, OW03], reducing the reachability problem over dense time to discrete (integer) time.
+techniques~\cite{HenzingerMP92,OuaknineW03} exist for reducing the reachability problem to discrete time in the case of closed (i.e. non-strict) clock constraints ranging over integers. \\
+
+\problemx{TA $k$-resilience problem}
+{A state $q$ of a TA $(Q, X, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
+{$\forall q' \in BAD \forall v,v' \in \N^X ~ (q(v) \rightarrow^* q'(v')) \implies \exists q'' \in SAFE \exists v'' \in \N^X ~ q'(v') \rightarrow^{\leq k} q''(v'')$ ?\newline}
+
+Analogously, we formulate the bounded resilience problem for WSTSs.
+
+\problemx{TA bounded resilience problem}
+{A state $q$ of a TA $(Q, X, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
+{$\exists k \geq 0 ~ \forall q' \in BAD \forall v,v' \in \N^X ~ (q(v) \rightarrow^* q'(v')) \implies \exists q'' \in SAFE \exists v'' \in \N^X ~ q'(v') \rightarrow^{\leq k} q''(v'')$ ?\newline}
+
+\textcolor{red}{I think there can be a discussion to be had here about how to quantify on the clock valuations}
+
+\textcolor{red}{Here one thing that could be interesting to try to formalize is: how to enforce that the time that passes is less than $k$, rather than the number of transitions.
+This is tricky to deal with I find but it should be more doable if for instance we use one counter automata, where the counter effect of the sequence can be quantified more explicitly I suppose ?
+But here you could also use a kinda special clock $x$ that is reset when you enter $BAD$ and is not reset between a state in $BAD$ and a state in $SAFE$, you could check that $x < k$.}
+
+\textcolor{red}{...
+I guess if you use $0/1$-TA then the problems become closer one to another ?
+Also of note is that $0/1$-TA induces transition systems with bounded branching, so I guess it may be interesting to investigate these first ?}
+
+A {\em $0/1$ timed automaton } ({\em $0/1$-TA} for short) is a tuple
+$$\B=(Q,X, \Delta_0, \Delta_1, q_{init}, F),$$
+\par\noindent\ignorespacesafterend
+ where
+$\B_i=(Q,X, R_i, q_{init}, F)$ is a TA for all $i \in \{0,1\}$.
+For simplicity we define its {\em size}
+as $|\B|=|\B_0|+|\B_1|$.
+We analogously denote the constants of $\B$ 
+by $\Const(\B)$ and its configurations by  $\Conf(\B)$.
+
+\begin{samepage}
+%\begin{definition}
+A $0/1$ timed automaton $\B=(Q,X,R_0,R_1,q_{init},F)$ 
+induces the labeled transition system 
+$T_{\B} = (\Conf(\B), \lambda_{\B}, \rightarrow_{\B}) $
+where $ \lambda_{\B} = (R_0 \cup R_1) \times \{ 0,1\}$
+	and where $ \rightarrow_{\B}$
+	is defined such that
+	for all $q(z), q'(z') \in \Conf(\B)$, 
+	for all $(\delta,i) \in \lambda_{\B}$
+	with $\delta  = (q,g,U,q')\in R_i$
+	$q(v)\xrightarrow{\delta,i}_{\B} q'(v')$ if
+	$v+i \models g$, 
+	$v'(u)=0$ for all $u \in U$ and $v'(\omega)=v(\omega)+ i$ for all $\omega \in \Omega
+	\setminus U$. 
+%\end{definition}
+\end{samepage}
+
+As expected, we write $q(v)\xrightarrow{\delta,i}_{\B}q'(v')$ if 
+$q(v)\xrightarrow{\delta,i}_{\B}q'(v')$ for some 
+$i\in\{0,1\}$, and some $\delta \in R_i$.
+
 \subsection{One-Counter Automata}
@@
 
+\problemx{OCA $k$-resilience problem}
+{A state $q$ of a OCA $(Q, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
+{$\forall q' \in BAD \forall n,n' \in \N ~ (q(n) \rightarrow^* q'(n')) \implies \exists q'' \in SAFE \exists n'' \in \N ~ q'(n') \rightarrow^{\leq k} q''(n'')$ ?\newline}
+
+\problemx{OCA bounded resilience problem}
+{A state $q$ of a OCA $(Q, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
+{$\exists k \geq 0 ~ \forall q' \in BAD \forall n,n' \in \N ~ (q(n) \rightarrow^* q'(n')) \implies \exists q'' \in SAFE \exists n'' \in \N ~ q'(n') \rightarrow^{\leq k} q''(n'')$ ?\newline}
+
 \subsection{Vector Addition System with States}

### 2022-10-17 definitions.tex
début

@@
 
+\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
 \textcolor{red}{Not sure that TS induced by TA are finitely branching.
@@
 \begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0 ; x_1 ; x_2 ; : : : $ in $X$, there exist indexes $i \leq j$ with
+ A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
 $x_i \leq  x_j$.
@@
 If $\leq$ is a wqo; then any upward-closed $I$ has a
-basis.
+finite basis.
 \end{lemma}
@@
 \end{proof}
+\textcolor{blue}{Alain: non ceci est vrai seulement si le quasi ordre est un ordre cad antisymétrique. Mais ce n'est pas un pb. Relis mon papier de 2016 sur les well abstracted...}
 

### 2022-10-19 application.tex
Separating sections, writing a tentative sketch of a proof for k-resiliency

--- 2022-10-17 definitions.tex
@@
-\section{Definitions}\label{section definitions}
-
-In this section, we introduce general notations and preliminary definitions.
-
-The model we are interested in is (S)WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
-
-\textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
-
-\subsection{Transition systems}
-
-\begin{definition}
-A {\em labeled transition system} (LTS for short) is a tuple $T=(S, \Lambda, \rightarrow)$ where 
-$S$ is a set of {\em configurations}, $ \Lambda$ is a set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
-ternary relation,
-denoted as the set of {\em labeled transitions}. 
-\end{definition}
-
-We
- prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
-       $s  \xrightarrow{a}  s'$
-to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
-
-\noindent
-Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
-
-A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
-to a {\em target configuration} $s_n$ is a sequence 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
-We define the {\em concatenation} $ \pi_1 \pi_2$ of 
-two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
-as expected.
-The {\em length} of 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
-is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
-For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
-
-An {\em infinite path} is an infinite sequence
- $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
-For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
-(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
-and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
-by $\pi[i,j]$ the path 
-$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
- and by $\pi[i]$ the configuration $s_i$.
-As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
-a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
-Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
-$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
-
-\newcommand{\pred}{\textsc{Pred}}
-\renewcommand{\succ}{\textsc{Succ}}
-
-\noindent
-The set of {\em successors} of a configuration $s \in S$ is deﬁned as
-$\succ(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
-A configuration without successors is called a {\em dead end}.  \\
-The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.
-
-A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
-$a \in \Lambda$,
- $ s_1  \xrightarrow{a} s_{2}$ and  
- $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
-
-\begin{definition}
-An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
-{\em configurations} and  
-$ {\rightarrow} \subseteq S \times S$ is a
-binary relation 
-on
-the set of configurations, denoted as the set of {\em transitions}. 
-\end{definition}
-
-We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from configuration $s$ to configuration $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
-Note 
- that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
-Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
-then defined as expected.
-
-\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
-
-We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
-for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\succ$ and $\pred$...
-
-\textcolor{red}{This makes sense for TS but not so much for LTS ...}
-
-A transition system is {\em finitely branching} if all $\succ(s)$ are finite. 
-\textcolor{blue}{We restrict our attention to finitely branching TSs.}
-
-\textcolor{red}{Not sure that TS induced by TA are finitely branching.
-Actually I believe they are not, i.e.
-for instance for a TA with one clock $x$, from a state $q$ and clock $x$ set at $0$, if there is a transition e.g.
-$(q, x \geq 3, \emptyset, q')$ then the set of successors of
-$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
-
-\subsection{Well-quasi-orderings}
-
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$.
-
-We abbreviate $x \leq y \not\leq x$ by $x < y$.
-
-% A partial ordering (a po) is an antisymmetric qo
-
-Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
-% and gives rise to a po between the equivalence classes ?
-% 	do we need po ?
-
-We now recall a few results from the theory of well-orderings (add reference [...]).
-
-\begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0 ; x_1 ; x_2 ; : : : $ in $X$, there exist indexes $i \leq j$ with
-$x_i \leq  x_j$.
-\end{definition}
-
-Notice that a wqo is well-founded, i.e. it admits no infinite strictly decreasing sequence
-$x_0 > x_1 > x_2 > \cdots$
-
-\textcolor{red}{Add lemma about infinite increasing subsequences ?}
-
-Given $\leq$ a quasi-ordering over some set $X$, an {\em upward-closed set} is any set $I \subseteq X$ such that if $y \geq x$ and $x \in I$ then $y \in I $.
-A {\em downward-closed set} is any set $I \subseteq X$ such that if $y \leq x$ and $x \in I$ then $y \in I $. 
-To any $A \subseteq X$ we associate
-the {\em upward-closure of $A$} 
- $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
- and the 
- {\em downward-closure of $A$} 
- $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
-as $\uparrow x$ (resp. $\downarrow x$).
-
-\begin{lemma}{(Higman [40])} 
-If $\leq$ is a wqo; then any upward-closed $I$ has a
-basis.
-\end{lemma}
-
-\textcolor{red}{Expliquer ce que c'est une base d'abords.}
-
-\begin{proof}
-The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
-only contains a finite number of non-equivalent elements otherwise they would make
-an infinite sequence contradicting the wqo assumption.
-\end{proof}
-
-\textcolor{red}{Un lemme je pense c'est important de le noter, peut-être pas comme ça peut être noter différement faudra voir}
-
-\begin{lemma}  
-If $\leq$ is a wqo; any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
-upward-closed sets eventually stabilizes; i.e. there is a $k \in N$ such that 
-$I_k = I_{k+1} = I_{k+2} = \cdots $.
-\end{lemma}
-
-\begin{proof}
-Assue we have a counter-example ...
-\end{proof}
-
-\textcolor{red}{Define WSTS}
-
-\textcolor{red}{Define SWSTS - may be necessary}
-
-\subsection{Well-structured transition systems}
-
-\begin{definition}
-A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. SWSTS) 
-is a TS
-$(S, \rightarrow, \leq)$
-equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
-%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
-$s_1, t_1 , s_2 \in S$
-	with $s_1 \leq t_2$  and $s_1 \rightarrow s_2$ , there exists 
-	$t_2 \in S$ with 
-	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
-				(resp. $t_1 \rightarrow^{1} t_2$ ).
-\end{definition}
-
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering.
-
-\textcolor{red}{For one-counter automata, in case the only tests are zero tests then 
-I supposed $\leq$ is $\leq$ for non-zero integers, and I'll have to look-up/think for what to do with the zero element for instance. For TA it seem kind of nontrivial (since they allow $<c$ tests).}
-
-\textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and
-'decidable $\leq$' being independant from the WSTS definition}
-
-\begin{definition}
-A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-\end{definition}
-
-\textcolor{red}{Define what an Ideal is.
-Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
-Anti-ideal just downward closed so again just not that helpful a notation.
-Maybe have a }
-
-\begin{definition}
-An {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
-$\uparrow I = I = \downarrow I$.
-\end{definition}
-
-A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in J$.
-
-"Bi-ideals often represent “control states” as in [cf \%]. "
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-\textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
-
-Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
-
-\textcolor{red}{Do we still demand this ?}
-
-\begin{claim}{(stability of ideals)}
-Let $I, J \subseteq S$ be upward-closed. Then the sets 
-$\pred(I)$, $I \cup J$, and $I \cap J$ are upward-closed.
-\end{claim}
-
-% \begin{claim}
-% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \end{claim}
-
-\subsection{Defining resilience}
-
-\subsubsection{TS resilience}
-
-\textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$SAFE$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$BAD$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the explicit resilience problem for TS.
-
-\problemx{TS $k$-resilience problem}
-{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).
-
-We can also ask whether there exists such a bound $k$. We call this problem the bounded resilience problem for TS.
-
-\problemx{TS bounded resilience problem}
-{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
-{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-\subsubsection{WSTS resilience}
-
-Properties in well-structured transition systems are often given as upward- or downward closed sets [references].
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-Transfering the abstract resilience problems into this framework,
-it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
-upward-closed or 
-downward-closed sets.
-
-We assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set.
-
-\textcolor{red}{Seems like a reasonable assumption to me.}
-
-From these considerations, we formulate instances of the abstract resilience problems for well-
-structured transition systems.
-
-Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
-
-\problemx{WSTS $k$-resilience problem}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-Analogously, we formulate the bounded resilience problem for WSTSs.
-
-\problemx{WSTS bounded resilience problem}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-\textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
+\section{Applications section}
 

### 2022-10-19 definitions.tex
Adding definitions to prepare for the merge
Separating sections, writing a tentative sketch of a proof for k-resiliency

--- 2022-10-17 definitions.tex
@@
 \begin{definition}
-An {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
+A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
 $\uparrow I = I = \downarrow I$.
@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \to \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
 \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
-
-\subsection{Timed Automata}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
-
-\subsection{One-Counter Automata}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
-
-\subsection{Vector Addition System with States}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}

### 2022-10-19 proof.tex
Separating sections, writing a tentative sketch of a proof for k-resiliency

\section{Decidability}

\begin{proof}{sketch}

Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:

\begin{lemma}
Let $A \subseteq S$ be a set, $J \subseteq S$ upward-closed and $I' \subseteq S$ downward-closed. 
Then $A \cap J \subseteq I' \leftrightarrow (\downarrow  A) \cap J \subseteq I'$.
\end{lemma}

Et une fois que l'on a ça, on peut écrire
$ post^*(s)\cap J \subseteq I' \leftrightarrow (\downarrow  post^*(s)) \cap J \subseteq I'$ 
et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.

La question deviens: quel $k$ pour que
$\downarrow post^*(s) \cap J \subseteq I^k$.
Ou alors, à $k$ fixé, est-ce que
$\downarrow post^*(s) \cap J \subseteq I^k$.

Il faut aussi une condition pour s'assurer que $I^k$ soit bien downward-closed.

Où
$I^{k+1} = I \cup pre(I^k )$.

On vas imaginer temporairement qu'on arriver à s'assurer $I^k$ downward-closed. 
(Peut-être on doit faire la supposition de SWSTS strongly downward compatible à la place, pour avoir ça ?)

On reviens sur $\downarrow post^*(s) \cap J \subseteq I^k$.

C'est là peut être qu'on peut s'inspirer de l'algo forward.

Enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq X$ such that $\downarrow Post(D_i ) \subseteq D_i$. [...]
Every inductive invariant $D_i$ is an “over-approximation” of $post^*(x)$ if it contains $x$.

On a un nombre fini de bornes supérieures pour $D_i$ on peut considèrer qu’elles consistent en une base de $D_i$ ? On énumère les $D_i$ en énumérant leurs bases directement ?

Ensuite, si jamais	$Base(D_1)  \cap J \subseteq I^k$.

alors

$D_1  \cap J \subseteq I^k$	(par le lemme)

et donc

$\downarrow post*(s) \cap J \subseteq D_1  \cap J \subseteq I^k$

(puisque $D_1$ est une sur-approximation).

Ce qui voudrais dire que on a bien la propriété de résilience.

Le problème évidement c'est que cette procédure ne termine peut être pas enfin on n'a aucune guarrantie à priori. À ce moment là il faudrait sans doute faire une autre procédure en parallèle, comme dans l'algo forward, et qui elle terminerais si on a {\em pas} la propriété de résilience.

 La propriété de $k$-résilience peut être résumée par la formule: $post^*(s) \cap J \subseteq I^k$.

Donc la question ce serait est-ce que, à $k$ fixé ($k$ fixé pour l'instant on vas dire) on aurais une procédure qui termine pour montrer que ce n'est pas le cas ?

Ce serait une procédure qui pourrait par exemple calculer $post^m(s) \cap J $ jusqu'à trouver un élément pas dans $I^k$ ?

Càd on calcule les éléments de $post^m(s)$ un à un, pour chacun on vérifie s’il est dans $J$ puis s’il y est on vérifie s’il est dans $I^k$ ?

Il faut aussi une base de $J$ et que $I^k$ soit décidable, mais ça a l'air faisable j'ai l'impression...

\textcolor{red}{Ça me perturbe un peu quand même de devoir inverser toutes les propriétés et avoir SWSTS downward-compatible, je pense ça doit être possible de faire les choses mieux...}

\end{proof}

### 2022-11-03 proof.tex
Up

@@
 \section{Decidability}
+
+\textcolor{red}{Pour l'instant je regarde la décidabilité du problème "symmétrique" c'est à dire que j'ai inversé toutes les propriétés et demandé à ce que SWSTS soit downward-compatible.
+Je pense ça doit être possible de faire les choses autrement ...}
+
+\begin{theorem}
+{\sc SWSTS $k$-resilience} is decidable.
+\end{theorem}
 
@@
 
-Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
+Assume  $(S, \rightarrow, \leq)$ is a SWSTS with downward compatibility, $I$ is a decidable downward-closed subset of $S$, and $J$ is an upward-closed set with a given basis.
+
+We define inductively
+$I^{k+1} = I \cup pre(I^k )$. Note that for all $k \in \N$, $I^k$ is downward-closed due to
+the strongly downward compatibility of $(S, \rightarrow, \leq)$.
+
+% Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
+
+The $k$-resilience property can be expressed as the formula
+$ post^*(s) \cap J \subseteq I^k$. In order to decide whether the inclusion holds, we execute two procedures in parallel, one trying to prove $ post^*(s)\cap J \subseteq I^k$ 
+and one looking for a counter example.
+
+In order to certify inclusion in $I^k$, we need to work with finite representations.
+The next lemma uses that $I'$ and $J$ are downward- and upward-closed, respectively.
 
@@
 
-Et une fois que l'on a ça, on peut écrire
-$ post^*(s)\cap J \subseteq I' \leftrightarrow (\downarrow  post^*(s)) \cap J \subseteq I'$ 
-et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
+% Et une fois que l'on a ça, on peut écrire
 
-La question deviens: quel $k$ pour que
-$\downarrow post^*(s) \cap J \subseteq I^k$.
-Ou alors, à $k$ fixé, est-ce que
-$\downarrow post^*(s) \cap J \subseteq I^k$.
+\begin{corollary}
+For all $k \in \N$,
+$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\downarrow  post^*(s)) \cap J \subseteq I^k$. 
+% et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
+\end{corollary}
 
-Il faut aussi une condition pour s'assurer que $I^k$ soit bien downward-closed.
+% La question deviens: quel $k$ pour que
+% $\downarrow post^*(s) \cap J \subseteq I^k$.
+% Ou alors, à $k$ fixé, est-ce que
+% $\downarrow post^*(s) \cap J \subseteq I^k$.
 
-Où
-$I^{k+1} = I \cup pre(I^k )$.
+Assume $k$ is fixed for now.
 
-On vas imaginer temporairement qu'on arriver à s'assurer $I^k$ downward-closed. 
-(Peut-être on doit faire la supposition de SWSTS strongly downward compatible à la place, pour avoir ça ?)
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow Post(D_i ) \subseteq D_i$. 
+Every inductive invariant $D_i$ is an “over-approximation” of $post^*(s)$ if it contains $s$.
+(on énumère des sur-approximations de la cloture par le bas de $post*(s)$ par leur bases finies).
+Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\downarrow post^*(s)$ is such an inductive invariant and may
+eventually be found.
 
-On reviens sur $\downarrow post^*(s) \cap J \subseteq I^k$.
+Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
+$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and we know the basis for $J$, we can
+directly compute $b(D)  \cap J$.
+% We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
+Due to the Lemma, 
+$b(D)  \cap J \subseteq I^k$ implies
+$D  \cap J \subseteq I^k$.
+Hence
+$b(D)  \cap J \subseteq I^k$ implies
+$\downarrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
+(since $D$ contains $post^*(s)$).
 
-C'est là peut être qu'on peut s'inspirer de l'algo forward.
+The second procedure iteratively computes
+$post^{\leq n}(s) \cap J$
+until it finds an element
+not in $ I^k$.
 
-Enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq X$ such that $\downarrow Post(D_i ) \subseteq D_i$. [...]
-Every inductive invariant $D_i$ is an “over-approximation” of $post^*(x)$ if it contains $x$.
+% Ce serait une procédure qui pourrait par exemple calculer $post^m(s) \cap J $ jusqu'à trouver un élément pas dans $I^k$ ?
 
-On a un nombre fini de bornes supérieures pour $D_i$ on peut considèrer qu’elles consistent en une base de $D_i$ ? On énumère les $D_i$ en énumérant leurs bases directement ?
+% Càd on calcule les éléments de $post^m(s)$ un à un, pour chacun on vérifie s’il est dans $J$ puis s’il y est on vérifie s’il est dans $I^k$ ?
 
-Ensuite, si jamais	$Base(D_1)  \cap J \subseteq I^k$.
+% Il faut aussi une base de $J$ et que $I^k$ soit décidable, mais ça a l'air faisable j'ai l'impression...
 
-alors
+\begin{figure}
+\fbox{\parbox[t][3.cm][c]{6cm}{
+$\phantom{a}$\\
+(1)\phantom{aaaaa} $i \leftarrow 0$\\
+(2)\phantom{aaaaa}\textbf{while} $\neg( \downarrow post^*(D_i) \subseteq D_i $
+			 and $ s \in D_i$
+			 and $ b(D)  \cap J \subseteq I^k  )$ \textbf{loop} \newline
+	(3)\phantom{aaaaaa}$\phantom{aaaa} i \leftarrow i +1$ \newline
+(4)\phantom{aaaaa}\textbf{end loop} \newline
+(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
+	}
+}
+	\caption{\textbf{Procedure 1:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
+\end{figure}
 
-$D_1  \cap J \subseteq I^k$	(par le lemme)
+\begin{figure}
+\fbox{\parbox[t][3.cm][c]{6cm}{
+$\phantom{a}$\\
+(1)\phantom{aaaaa} $D \leftarrow \{ s \} $\\
+(2)\phantom{aaaaa}\textbf{while} $D \cap J \subseteq I^k $ \textbf{loop} \newline
+	(3)\phantom{aaaaaa}$\phantom{aaaa} D \leftarrow D \cup post(D)$ \newline
+(4)\phantom{aaaaa}\textbf{end loop} \newline
+(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
+	}
+}
+	\caption{\textbf{Procedure 2:} searches for a non-inclusion certificate.}\label{procedure2}
+\end{figure}
 
-et donc
+\newpage
 
-$\downarrow post*(s) \cap J \subseteq D_1  \cap J \subseteq I^k$
-
-(puisque $D_1$ est une sur-approximation).
-
-Ce qui voudrais dire que on a bien la propriété de résilience.
-
-Le problème évidement c'est que cette procédure ne termine peut être pas enfin on n'a aucune guarrantie à priori.
-À ce moment là il faudrait sans doute faire une autre procédure en parallèle, comme dans l'algo forward, et qui elle terminerais si on a {\em pas} la propriété de résilience.
-
- La propriété de $k$-résilience peut être résumée par la formule: $post^*(s) \cap J \subseteq I^k$.
-
-Donc la question ce serait est-ce que, à $k$ fixé ($k$ fixé pour l'instant on vas dire) on aurais une procédure qui termine pour montrer que ce n'est pas le cas ?
-
-Ce serait une procédure qui pourrait par exemple calculer $post^m(s) \cap J $ jusqu'à trouver un élément pas dans $I^k$ ?
-
-Càd on calcule les éléments de $post^m(s)$ un à un, pour chacun on vérifie s’il est dans $J$ puis s’il y est on vérifie s’il est dans $I^k$ ?
-
-Il faut aussi une base de $J$ et que $I^k$ soit décidable, mais ça a l'air faisable j'ai l'impression...
-
-\textcolor{red}{Ça me perturbe un peu quand même de devoir inverser toutes les propriétés et avoir SWSTS downward-compatible, je pense ça doit être possible de faire les choses mieux...}
+We show that these two procedures are correct:
 

### 2022-11-14 definitions.tex
Update avec une preuve presque terminée

@@
 
-\textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
+% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
 
@@
 
-\textcolor{red}{Add lemma about infinite increasing subsequences ?}
+% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
+
+\begin{lemma}
+(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
+\end{lemma}
+
+\iffalse
+\begin{proof}
+Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
+be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
+hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
+increasing subsequence.
+\end{proof}
+\fi
 
@@
 
+A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \bigcup_{x \in I_b} \uparrow x$. 
+% Higman investigated ordered sets with the finite basis property.
+
 \begin{lemma}{(Higman [40])} 
@@
 
-\textcolor{red}{Expliquer ce que c'est une base d'abords.}
+% \textcolor{red}{Expliquer ce que c'est une base d'abords.}
 
@@
 
-\textcolor{red}{Un lemme je pense c'est important de le noter, peut-être pas comme ça peut être noter différement faudra voir}
-
 \begin{lemma}  
@@
 \begin{proof}
-Assue we have a counter-example ...
+Assue we have a counter-example.
+We extract an infinite subsequence where
+inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$.
+Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$.
+The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
+contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
+closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
 \end{proof}
 
-\textcolor{red}{Define WSTS}
+% \textcolor{red}{Define WSTS}
 
-\textcolor{red}{Define SWSTS - may be necessary}
+% \textcolor{red}{Define SWSTS -  necessary}
 
@@
 
-\textcolor{red}{For one-counter automata, in case the only tests are zero tests then 
-I supposed $\leq$ is $\leq$ for non-zero integers, and I'll have to look-up/think for what to do with the zero element for instance. For TA it seem kind of nontrivial (since they allow $<c$ tests).}
+% \textcolor{red}{Define 'has effective pred-basis'.
+Maybe it should be included in WSTS definition, maybe it can be another def.
+I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
 
-\textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and
-'decidable $\leq$' being independant from the WSTS definition}
+\begin{proposition}
+If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
+closed.
+\end{proposition}
+
+\iffalse
+Proof.
+Assume $s \in \pred^* (I )$.
+Then $s \rightarrow^* t$ for some $t \in I $.
+If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$.
+Then $t' \in I$ and $s' \in \pred^*(I )$.
+\fi
+
+\begin{proposition}
+If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
+\end{proposition}
+
+\iffalse
+Proof.
+Assume $s \in \pred (I )$.
+Then $s \rightarrow t$ for some $t \in I $.
+If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
+Then $t' \in I$ and $s' \in \pred(I )$.
+\fi
 
@@
 
-\textcolor{red}{Define what an Ideal is.
+% \textcolor{red}{Define what an Ideal is.
 Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
@@
 
+\iffalse
 \begin{definition}
@@
 
-A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in J$.
-
 "Bi-ideals often represent “control states” as in [cf \%]. "
@@
 
-\textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
+% \textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
+\fi
 
+A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
+$s \in J$. 
 Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
 
-\textcolor{red}{Do we still demand this ?}
+% \textcolor{red}{Do we still demand this ?}
 
 \begin{claim}{(stability of ideals)}
-Let $I, J \subseteq S$ be upward-closed. Then the sets 
-$\pred(I)$, $I \cup J$, and $I \cap J$ are upward-closed.
+Let $I, J \subseteq S$ be upward-closed. Then the sets $I \cup J$, and $I \cap J$ are upward-closed.
 \end{claim}
@@
 % \end{claim}
+
+\begin{fact}
+% Fact 3 ([1]). 
+% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
+
+(i) For every upward-closed set $I \subseteq S$, there exists a finite basis $B$ of $I$.
+
+(ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
+\end{fact}
+
+\begin{definition}{ (index)}. 
+For an upward-closed set $I \subseteq S$ and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} pre^j(I)$.
+The {\em index} $k(I)$ is the
+smallest $k_0$ s.t. $I^k = I^{k_0}$ for all $k \geq k_0$.
+\end{definition}
+
+\begin{fact}
+% Fact 4 (stop condition). 
+Let $I \subseteq S$ be an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
+$k(I) \leq k$. This also implies that $pre^* (I) = I^k $.
+\end{fact}
+
+\begin{lemma}
+% Lemma 3 ([1]) 
+% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
+ Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of a strongly well-structured transition
+system, we can decide whether we can reach $I$ from $s$.
+\end{lemma}
+
+\begin{proof}
+We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
+Then the
+decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
+We have
+$$I^{k+1} = I \cup pre(I^k ) = I \cup
+\bigcup_{s' \in B}
+pre(\uparrow \{s' \}).$$
+
+Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
+Fact 3, we can compute a basis of $I^{k+1}$.
+\end{proof}
 
@@
 
-\textcolor{red}{Seems like a reasonable assumption to me.}
+% \textcolor{red}{Seems like a reasonable assumption to me.}
 
@@
 
-\textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
+% \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}

### 2022-11-14 proof.tex
Update avec une preuve presque terminée

@@
 \section{Decidability}
-
-\textcolor{red}{Pour l'instant je regarde la décidabilité du problème "symmétrique" c'est à dire que j'ai inversé toutes les propriétés et demandé à ce que SWSTS soit downward-compatible.
-Je pense ça doit être possible de faire les choses autrement ...}
 
@@
 
-\begin{proof}{sketch}
+\begin{proof}
 
-Assume  $(S, \rightarrow, \leq)$ is a SWSTS with downward compatibility, $I$ is a decidable downward-closed subset of $S$, and $J$ is an upward-closed set with a given basis.
+Assume  $(S, \rightarrow, \leq)$ is a SWSTS with upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
 
 We define inductively
-$I^{k+1} = I \cup pre(I^k )$. Note that for all $k \in \N$, $I^k$ is downward-closed due to
-the strongly downward compatibility of $(S, \rightarrow, \leq)$.
+$I^{k} = \bigcup_{0 \leq j \leq k} pre^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
+the strongly upward compatibility of $(S, \rightarrow, \leq)$.
 
@@
 In order to certify inclusion in $I^k$, we need to work with finite representations.
-The next lemma uses that $I'$ and $J$ are downward- and upward-closed, respectively.
+The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.
 
 \begin{lemma}
-Let $A \subseteq S$ be a set, $J \subseteq S$ upward-closed and $I' \subseteq S$ downward-closed. 
-Then $A \cap J \subseteq I' \leftrightarrow (\downarrow  A) \cap J \subseteq I'$.
+Let $A \subseteq S$ be a set, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
+Then $A \cap J \subseteq I \leftrightarrow (\uparrow  A) \cap J \subseteq I$.
 \end{lemma}
@@
 For all $k \in \N$,
-$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\downarrow  post^*(s)) \cap J \subseteq I^k$. 
+$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
 % et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
@@
 
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow Post(D_i ) \subseteq D_i$. 
-Every inductive invariant $D_i$ is an “over-approximation” of $post^*(s)$ if it contains $s$.
-(on énumère des sur-approximations de la cloture par le bas de $post*(s)$ par leur bases finies).
-Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\downarrow post^*(s)$ is such an inductive invariant and may
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
+Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
+(on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
+Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
 eventually be found.
@@
 Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
-$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and we know the basis for $J$, we can
+$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and $J$ is decidable, we can
 directly compute $b(D)  \cap J$.
 % We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
+We can compute a basis
+of $I^{k+1}$ if we have a basis of $I^k$. %This follows by the proof of Lemma 3.
 Due to the Lemma, 
@@
 $b(D)  \cap J \subseteq I^k$ implies
-$\downarrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
-(since $D$ contains $post^*(s)$).
+$\uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
+(since $D$ contains $ \uparrow post^*(s)$).
 
@@
 (1)\phantom{aaaaa} $i \leftarrow 0$\\
-(2)\phantom{aaaaa}\textbf{while} $\neg( \downarrow post^*(D_i) \subseteq D_i $
+(2)\phantom{aaaaa}\textbf{while} $\neg( \uparrow post^*(D_i) \subseteq D_i $
 			 and $ s \in D_i$
@@
 
+\begin{enumerate}
+
+\item $k$-resilience holds if, and only if, Procedure 1 terminates.
+
+\item $k$-resilience do not hold if, and only if, Procedure 2 terminates.
+
+\end{enumerate}
+
+Proof:
+
+\begin{enumerate}
+
+\item By a simple induction, it can be shown that $\uparrow post^*(D) \subseteq D$ for every inductive invariant $D$. 
+If Procedure 1 terminates, then
+$post^*(s) \cap J \subseteq \uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$
+which implies that $k$-resilience holds.
+
+It remains to show that Procedure 1 terminates whenever $k$-resilience holds. To do so, it suffices to prove that $\uparrow post^*(s)$ is an inductive invariant. Indeed, this implies that
+$\uparrow post^*(s)$ is eventyally found by Procedure 1 when $k$-resilience holds.
+
+Formally, let us show that $\uparrow post(\uparrow post^*(s)) \subseteq \uparrow post^*(s)$.
+Let $b \in \uparrow post(\uparrow post^*(s))$ 
+there exists $a', a, b$ such that
+$s \rightarrow^* a'$,
+$a' \leq a$,
+$a \rightarrow b'$,
+and
+$b' \leq b$.
+% By monotonicity, there exists b'' ≥ b' such that a → b'' . Therefore, x → ∗ b'' and b' ≥ b, hence b ∈ ↓ Post ∗ (x).
+\textcolor{red}{By downward compatibility 
+there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
+
+\item Procedure 2 computes
+ $post^{\leq n}(s) \cap J$
+ until it finds an element not in $ I^k$.
+
+If Procedure 2 terminates, then
+$k$-resilience does not hold.
+It remains to show that Procedure 2 terminates whenever $k$-resilience does not hold.
+
+\end{enumerate}
+
 \end{proof}
+
+\begin{theorem}
+{\sc SWSTS resilience} is decidable.
+\end{theorem}
+
+\begin{proof}{sketch}
+
+Iteratively check whether $k$-resilience holds. 
+If this is the case, return $k_{min} = k$. Otherwise check whether 
+$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
+continue.
+The stop condition is decidable
+and by Fact 4 also sufficient. 
+\end{proof}

### 2022-11-14 definitions.tex
details

@@
 \begin{lemma}{(Higman [40])} 
-If $\leq$ is a wqo; then any upward-closed $I$ has a
+If $\leq$ is a wqo then any upward-closed $I$ has a
 finite basis.
@@
 \end{proof}
-\textcolor{blue}{Alain: non ceci est vrai seulement si le quasi ordre est un ordre cad antisymétrique. Mais ce n'est pas un pb. Relis mon papier de 2016 sur les well abstracted...}
 
 \begin{lemma}  
-If $\leq$ is a wqo; any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
+If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
 upward-closed sets eventually stabilizes; i.e. there is a $k \in N$ such that 

### 2022-11-14 proof.tex
Adding Facts

@@
 It remains to show that Procedure 2 terminates whenever $k$-resilience does not hold.
-
+Assume $ post^*(s) \cap J \not\subseteq I^k$, then there exists $a \in post^*(s) \cap J$ such that $a \not\in I^k$. Since $post^*(s) = \bigcup_{n} post^{\leq n}(s)$, then 
+$a \in post^*(s)$
+implies
+there exists
+$n_a$
+such that
+$a \in post^{\leq n_a}(s)$.
+Hence,  $post^{\leq n_a}(s) \cap J$ contains an element not in 
+$I^k$,
+and Procedure 2 terminates.
 \end{enumerate}

### 2022-11-14 definitions.tex

@@
 If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
-upward-closed sets eventually stabilizes; i.e. there is a $k \in N$ such that 
+upward-closed sets eventually stabilizes, i.e. there is a $k \in N$ such that 
 $I_k = I_{k+1} = I_{k+2} = \cdots $.

### 2022-11-15 proof.tex
Highlited argument that required downward compatibility (in addition to strong upward one) in  red
details

@@
 (on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
-Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
-eventually be found.
+Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. \textcolor{red}{Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
+eventually be found.}
 
@@
 there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
-
+\textcolor{green}{où est cette hypothèse downward compatibility  ?}
 \item Procedure 2 computes

### 2022-11-16 definitions.tex
details

@@
 \begin{definition}
-A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. SWSTS) 
+A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. S-WSTS) 
 is a TS
@@
 $s_1, t_1 , s_2 \in S$
-	with $s_1 \leq t_2$  and $s_1 \rightarrow s_2$ , there exists 
+	with $s_1 \leq t_1$  and $s_1 \rightarrow s_2$ , there exists 
 	$t_2 \in S$ with 
@@
 \end{proposition}
-
+\alain{étrange ces deux propositions}
 \iffalse
@@
 (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
+\alain{A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
 \end{fact}
@@
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of a strongly well-structured transition
+ Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of an effective strongly well-structured transition
 system, we can decide whether we can reach $I$ from $s$.
@@
 pre(\uparrow \{s' \}).$$
-
+\alain{n'est-ce pas $I^{k+1} = I^k \cup pre(I^k ) = I^k \cup
+\bigcup_{s' \in B}
+pre(\uparrow \{s' \}).$ ?}
 Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
@@
 
-\subsubsection{TS resilience}
+\subsubsection{Resilience}
 
@@
 
-\problemx{TS $k$-resilience problem}
+\problemx{$k$-resilience problem}
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
@@
 
-\problemx{TS bounded resilience problem}
+\problemx{Resilience problem}
 {A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
@@
 
-\subsubsection{WSTS resilience}
+\subsubsection{Resilience}
 
@@
 
-\problemx{WSTS $k$-resilience problem}
+\problemx{$k$-resilience problem for WSTS}
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
@@
 
-Analogously, we formulate the bounded resilience problem for WSTSs.
+Analogously, we formulate the resilience problem for WSTSs.
 
-\problemx{WSTS bounded resilience problem}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\exists k \geq 0 ~ \forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+\problemx{Resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+{$\exists k \geq 0 ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 

### 2022-11-16 proof.tex
details

@@
 \begin{theorem}
-{\sc SWSTS $k$-resilience} is decidable.
+{\sc $k$-resilience} is decidable for effective WSTS with strong monotony.
 \end{theorem}
@@
 there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
-\textcolor{green}{où est cette hypothèse downward compatibility  ?}
+\alain{où est cette hypothèse downward compatibility  ?}
 \item Procedure 2 computes
@@
 \begin{theorem}
-{\sc SWSTS resilience} is decidable.
+{\sc $k$-resilience} is undecidable for effective WSTS ???
+\end{theorem}
+
+\begin{theorem}
+{\sc resilience} is decidable for WSTS with strong monotony.
 \end{theorem}
@@
 and by Fact 4 also sufficient. 
+\alain{Fact 4 don't exist}
 \end{proof}

### 2022-11-16 definitions.tex
details

@@
 \end{proposition}
-\alain{étrange ces deux propositions}
+\alain{étrange ces deux propositions, l'une suffirait}
 \iffalse
@@
 (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-\alain{A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
+\alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
 \end{fact}
@@
 \end{proof}
-
+\alain{pas de Fact 3}
 \subsection{Defining resilience}
 
-\subsubsection{Resilience}
+\subsubsection{Resilience for general transition systems}
 
@@
 \problemx{$k$-resilience problem}
-{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
+{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
 {$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
@@
 
-We can also ask whether there exists such a bound $k$. We call this problem the bounded resilience problem for TS.
+We can also ask whether there exists such a bound $k$. We call this problem the resilience problem.
 
 \problemx{Resilience problem}
-{A state $s$ of a TS $(S,\rightarrow)$, two disjoints subset of $S$ $SAFE$ and $BAD$.}
+{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
 {$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
-\subsubsection{Resilience}
+\subsubsection{Resilience for WSTS}
 
@@
 
+\alain{dire juste que SAFE = upward closed et BAD = downward closed.
+envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
+4 resilience problem (U,D), (U,U), (D,U), (D,D)}
 % \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}

### 2022-11-16 proof.tex
details

@@
 \begin{lemma}
-Let $A \subseteq S$ be a set, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
+Let $A \subseteq S$, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
 Then $A \cap J \subseteq I \leftrightarrow (\uparrow  A) \cap J \subseteq I$.
@@
 eventually be found.}
+\alain{attention D est clos par le bas et $\uparrow post^*(s)$ est clos par le haut, je ne vois pas pourquoi $\uparrow post^*(s) \subseteq D$}
 

### 2022-12-02 proof.tex
details

@@
 \section{Decidability}
-
-\begin{theorem}
-{\sc $k$-resilience} is decidable for effective WSTS with strong monotony.
-\end{theorem}
-
-\begin{proof}
-
-Assume  $(S, \rightarrow, \leq)$ is a SWSTS with upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
-
-We define inductively
-$I^{k} = \bigcup_{0 \leq j \leq k} pre^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
-the strongly upward compatibility of $(S, \rightarrow, \leq)$.
-
-% Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
-
-The $k$-resilience property can be expressed as the formula
-$ post^*(s) \cap J \subseteq I^k$. In order to decide whether the inclusion holds, we execute two procedures in parallel, one trying to prove $ post^*(s)\cap J \subseteq I^k$ 
-and one looking for a counter example.
-
-In order to certify inclusion in $I^k$, we need to work with finite representations.
-The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.
 
@@
 \end{lemma}
-
-% Et une fois que l'on a ça, on peut écrire
 
@@
 not in $ I^k$.
+
+This result is not in xxx and don't use the problematic hypothsis on $post^*$.
+
+\begin{theorem}
+{\sc $k$-resilience} is decidable for effective WSTS with downward/upward compatibility.
+%	with strong monotony.
+\end{theorem}
+
+\begin{proof}
+
+Assume  $(S, \rightarrow, \leq)$ is a WSTS with downward/upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
+
+We define inductively
+$I^{k} = \bigcup_{0 \leq j \leq k} pre^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
+the strongly upward compatibility of $(S, \rightarrow, \leq)$.
+
+% Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
+
+The $k$-resilience property can be expressed as the formula
+$ post^*(s) \cap J \subseteq I^k$. In order to decide whether the inclusion holds, we execute two procedures in parallel, one trying to prove $ post^*(s)\cap J \subseteq I^k$ 
+and one looking for a counter example.
+
+In order to certify inclusion in $I^k$, we need to work with finite representations.
+The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.
+
+% Et une fois que l'on a ça, on peut écrire
 
@@
 there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
-\alain{où est cette hypothèse downward compatibility  ?}
+%\alain{où est cette hypothèse downward compatibility  ?}
 \item Procedure 2 computes
@@
 \begin{theorem}
-{\sc $k$-resilience} is undecidable for effective WSTS ???
-\end{theorem}
-
-\begin{theorem}
-{\sc resilience} is decidable for WSTS with strong monotony.
+{\sc resilience} is decidable for WSTS with downward/upward compatibility.
 \end{theorem}
@@
 \end{proof}
+
+{\sc resilience} is decidable for WSTS with strong upward compatibility and k-resilience.
+
+jouer avec les hypothèses downward/upward compatibility
+
+\begin{theorem}
+{\sc $k$-resilience} is undecidable for effective WSTS ???
+\end{theorem}
+
+\begin{proof}
+totot
+\end{proof}

### 2022-12-02 definitions.tex
details

@@
 
+\alain{examiner les cas I,J un élement, cloture, tout, ...}
+
 \alain{dire juste que SAFE = upward closed et BAD = downward closed.
@@
 % \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
+
+citer
+
+\begin{theorem}(ref)
+{\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
+\end{theorem}
+
+Il y a des pbs
+
+\begin{proposition}
+The hypothesis $post^*$ is undecidable for WSTS with strong upward compatibility.
+\end{proposition}
+
+\begin{proof}
+$post^*$ donne l'accessibilité or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
+\end{proof}
+
+\begin{proposition}
+{\sc $k$-resilience}, hence {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility.
+\end{proposition}
+
+\begin{proof}
+totot
+\end{proof}

### 2022-12-02 definitions.tex
details

@@
 
-\problemx{Resilience problem for WSTS}
+\problemx{bounded Resilience problem for WSTS}
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
@@
 
-\alain{examiner les cas I,J un élement, cloture, tout, ...}
+\problemx{Resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+{$ ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{*} s''$ ?\newline}
+
+\alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
 
@@
 % \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
+
+\alain{regarder les modèles aut à pile...}
 
@@
 \begin{proposition}
-{\sc $k$-resilience}, hence {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility.
+{\sc $k$-resilience}, hence {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ?
 \end{proposition}

### 2022-12-02 definitions.tex
details

@@
 \begin{theorem}(ref)
+{\sc bounded resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
+\end{theorem}
+
+We may generalyse thier result:
+
+\begin{theorem}(ref)
 {\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
 \end{theorem}
+
+\begin{proof}
+totot
+\end{proof}
 
@@
 
+remarque si on enlève la strong monotony:
+
+\begin{theorem}(ref)
+{\sc resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
+\end{theorem}
+
 \begin{proof}
@@
 \begin{proposition}
-{\sc $k$-resilience}, hence {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ?
+{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ?
 \end{proposition}

### 2022-12-02 definitions.tex
details

@@
 
+\problemx{$k$-resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
+{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
+max des I,J possibles \\
+
+J=tout \\
+\problemx{$k$-resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
+{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
 Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
@@
 \alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
+
+\begin{remark}
+\begin{itemize}
+\item J=tout et I=$\uparrow x$ alors resilience = couverture.
+\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
+\end{itemize}
+\end{remark}
 
@@
 
+\begin{proof}
+$post^*$ donne l'accessibilité or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
+\end{proof}
+
 remarque si on enlève la strong monotony:
@@
 
-\begin{proof}
-$post^*$ donne l'accessibilité or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-\end{proof}
+si on enlève effective basis of $post^*$:
 

### 2022-12-02 proof.tex
details

@@
 \begin{theorem}
-{\sc $k$-resilience} is decidable for effective WSTS with downward/upward compatibility.
+{\sc $k$-resilience} is decidable for effective WSTS with downward/strong upward compatibility.
 %	with strong monotony.
@@
 \begin{theorem}
-{\sc resilience} is decidable for WSTS with downward/upward compatibility.
+{\sc bounded resilience} and {\sc resilience} is decidable for WSTS with downward/strong upward compatibility.
 \end{theorem}
@@
 jouer avec les hypothèses downward/upward compatibility
-
-\begin{theorem}
-{\sc $k$-resilience} is undecidable for effective WSTS ???
-\end{theorem}
-
-\begin{proof}
-totot
-\end{proof}

### 2022-12-02 definitions.tex
details

--- 2022-12-02 definitions.tex
@@
 
+\problemx{(I,J)-$k$-resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
+{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
+max des I,J possibles \\
+
+J=tout \\
+\problemx{general $k$-resilience problem for WSTS}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
+{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
 Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
@@
 \alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
+
+\begin{remark}
+\begin{itemize}
+\item expliquer toutes les variantes de la résilience
+\item J=tout et I=$\uparrow x$ alors resilience = couverture.
+\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
+\end{itemize}
+\end{remark}
 

### 2022-12-02 definitions.tex
details

@@
 
-We may generalyse thier result:
+We may generalyse thier result by weakiening the result
 
-\begin{theorem}(ref)
+\begin{corollary}
 {\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
-\end{theorem}
+\end{corollary}
 
@@
 \begin{theorem}(ref)
+{\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
+\end{theorem}
+
+\begin{theorem}(ref)
 {\sc resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
 \end{theorem}
+
+\begin{proof}
+on choisit $J=S$ ou $Reach(S)$. {$\forall s  \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}. il existe un chemin infini passant par inft par $q$, liveness undecidable LCS, Reset, WSTS.
+\end{proof}
 
@@
 \begin{proposition}
-{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ?
+{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ????
 \end{proposition}

### 2022-12-02 definitions.tex
details

@@
 
-remarque si on enlève la strong monotony:
+remarque si on enlève la strong monotony mais en gardant with effective basis of $post^*$:
 
-\begin{theorem}(ref)
+\begin{theorem}
 {\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
@@
 
-\begin{theorem}(ref)
-{\sc resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
+si on garde strong mais on enlève effective basis of $post^*$:
+
+\begin{proposition}
+{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ????
+\end{proposition}
+
+\begin{theorem}
+{\sc resilience} is undecidable for WSTS with upward compatibility ???
 \end{theorem}
@@
 
-si on enlève effective basis of $post^*$:
-
-\begin{proposition}
-{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ????
-\end{proposition}
+and with effective basis of $post^*$ ???
 

### 2022-12-02 definitions.tex
details

@@
 
+\begin{proof}
+totot
+\end{proof}
+
 si on garde strong mais on enlève effective basis of $post^*$:
@@
 \begin{proposition}
-{\sc $k$-resilience}, hence {\sc bounded resilience} and {\sc resilience}, is undecidable for effective WSTS with strong upward compatibility ????
+({\sc $k$-resilience},) hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility ????
 \end{proposition}
+
+\begin{proof}
+totot
+\end{proof}
 

### 2022-12-02 definitions.tex
details

@@
 \begin{proof}
-$post^*$ donne l'accessibilité or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
+$\uparrow post^*$ donne l'accessibilité de 0 sur les compteurs or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
+$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
 \end{proof}
@@
 \begin{proof}
-totot
+leur algorithme termine mais ne marche pas car l'intersection de post et J car ce n'est pas pertinent. et donc ???
 \end{proof}
@@
 \begin{proposition}
-({\sc $k$-resilience},) hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility ????
+{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility ????
 \end{proposition}
@@
 \begin{proof}
-totot
+pas de strong.
 \end{proof}
@@
 \begin{theorem}
-{\sc resilience} is undecidable for WSTS with upward compatibility ???
+{\sc resilience} is undecidable for WSTS (with upward compatibility) ???
 \end{theorem}
@@
 \begin{proof}
-on choisit $J=S$ ou $Reach(S)$. {$\forall s  \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}. il existe un chemin infini passant par inft par $q$, liveness undecidable LCS, Reset, WSTS.
+on choisit $J=Reach(S)$.
+resilience remains to {$\forall s \in S,  \exists s'' \in I ~ s' \rightarrow^{*} s''$ ?\newline}.
+il existe un chemin infini passant par inft par $q$, liveness undecidable LCS (Reach=$\downarrow Reach$), Reset, WSTS.
 \end{proof}

### 2022-12-02 definitions.tex
details

@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in J  \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$\exists k \geq 0 ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 \begin{proof}
-pas de strong.
+pas de effective basis of $post^*$.
 \end{proof}

### 2022-12-02 proof.tex
details
up

@@
-\section{Decidability}
+\section{Decidability for WSTS with downward/strong upward compatibility}
 
@@
 \end{lemma}
+
+\begin{proof}
+
+If 
+$A \cap J \subseteq I$,
+then
+${\uparrow (A \cap J)} \subseteq {\uparrow I} = I$.
+However
+$({\uparrow A}) \cap J \subseteq {\uparrow (A \cap J)}$.
+Indeed if $x$ in $({\uparrow A} \cap J)$,
+$x$ is in $\uparrow A$
+so there exists $a \in A$ such that $x \geq a$.
+However $x$ is also in $J$, meaning $a$ is in $J$.
+Meaning $a$ is in $(A \cap J)$
+and $x \in { \uparrow (A \cap J)}$.
+
+In the other direction,
+since $A \subseteq {\uparrow A}$,
+$({\uparrow  A}) \cap J \subseteq I$ implies
+$A \cap J \subseteq ({\uparrow  A}) \cap J \subseteq I$.
+\end{proof}
 

### 2022-12-02 definitions.tex
details
up

@@
 \end{definition}
+
+A \emph{downward-WSTS} is ... \\
 
@@
 
+applications: PN, les automates a pile sont WSTS avec $post^*$ calculable donc resilience calculable. Mais pas ResetPN.
+
 We may generalyse thier result by weakiening the result
@@
 
-\begin{theorem}
+\begin{theorem}(à prouver)
 {\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
@@
 
-\begin{proposition}
+\begin{proposition}(à prouver)
 {\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility ????
@@
 \begin{proof}
-pas de effective basis of $post^*$.
+pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?
 \end{proof}
@@
 \begin{theorem}
-{\sc resilience} is undecidable for WSTS (with upward compatibility) ???
+{\sc resilience} is undecidable for WSTS (with upward compatibility).
 \end{theorem}

### 2022-12-02 definitions.tex
details

@@
 
+Unfortunately, resilience is undecidable for general WSTS even with strong monotony.
+
 \begin{theorem}
-{\sc resilience} is undecidable for WSTS (with upward compatibility).
+{\sc resilience} is undecidable for WSTS with strong monotony.
 \end{theorem}
@@
 \begin{proof}
-on choisit $J=Reach(S)$.
-resilience remains to {$\forall s \in S,  \exists s'' \in I ~ s' \rightarrow^{*} s''$ ?\newline}.
-il existe un chemin infini passant par inft par $q$, liveness undecidable LCS (Reach=$\downarrow Reach$), Reset, WSTS.
+Let $J=\downarrow Reach(S)$.
+Now resilience remains to the following property: $\forall s \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
+This property is a liveness property that would allow to decide LTL but LTL has been proved undecidable for Lossy Channel Systems (LCS) for which Reach=$\downarrow Reach$.
+Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
 \end{proof}
 
+Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets but $Reach \neq \downarrow Reach$.
+
 and with effective basis of $post^*$ ???
-
-\begin{proof}
-totot
-\end{proof}

### 2022-12-02 definitions.tex
up
details

@@
 \begin{proof}
-Let $J=\downarrow Reach(S)$.
+Let $J=\downarrow Reach(S)$ and $I = \uparrow t$ where $t \in S$.
 Now resilience remains to the following property: $\forall s \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-This property is a liveness property that would allow to decide LTL but LTL has been proved undecidable for Lossy Channel Systems (LCS) for which Reach=$\downarrow Reach$.
+Let us translate this property for LCS for which Reach=$\downarrow Reach$: $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
+This property can be used to express the liveness property that there is an infinite run passing infinitely often through control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
 Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.

### 2022-12-05 definitions.tex
details

@@
 Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
-Fact 3, we can compute a basis of $I^{k+1}$.
+Fact 1, we can compute a basis of $I^{k+1}$.
 \end{proof}
@@
 
+\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
+avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
 \alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
@@
 
-\begin{theorem}(ref)
-{\sc bounded resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
+\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
+{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and the existence of an algorithm to compute a finite basis of $\uparrow post^*(S)$.
 \end{theorem}
 
-applications: PN, les automates a pile sont WSTS avec $post^*$ calculable donc resilience calculable. Mais pas ResetPN.
-
-We may generalyse thier result by weakiening the result
+\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because a finite basis of  $\uparrow post^*$ can be computed.
 
 \begin{corollary}
-{\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $post^*$.
+{\sc resilience} is decidable for context-free grammars and VASS.
+\end{corollary}
+
+Context-free grammars and VASS are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.
+
+We may generalyse this last result by strenghthenin to \emph{unbounded} resilience.
+
+\begin{corollary}
+{\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $\uparrow post^*$.
 \end{corollary}
@@
 \begin{proof}
-totot
+The proof is essentially the same than the previous one.
 \end{proof}
@@
 \begin{proposition}
-The hypothesis $post^*$ is undecidable for WSTS with strong upward compatibility.
+Given a WSTS with strong upward compatibility, an effective basis of $\uparrow post^*$ is not computable.
 \end{proposition}
@@
 Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets but $Reach \neq \downarrow Reach$.
+Another proof could be a reduction of liveness to resilience and liveness is undecidable for many models like reset PN,...
 

### 2022-12-05 definitions.tex
details

@@
 
-\begin{definition}
-A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. S-WSTS) 
-is a TS
-$(S, \rightarrow, \leq)$
+\begin{definition}\cite{DBLP:journals/tcs/FinkelS01}
+A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
+is a transition system $(S, \rightarrow)$
 equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
 %  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
+the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
 $s_1, t_1 , s_2 \in S$
@@
 	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
-				(resp. $t_1 \rightarrow^{1} t_2$ ).
+				(resp. $t_1 \rightarrow t_2$ ).
 \end{definition}
@@
 
-\begin{proposition}
+\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
 If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
@@
 
-\begin{proposition}
-If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
+\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
+If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
 \end{proposition}
@@
 
-\begin{definition}
+\begin{definition}\cite{DBLP:journals/iandc/AbdullaCJT00}
 A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
@@
 Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
-
+\alain{remarque pas claire}
 % \textcolor{red}{Do we still demand this ?}
@@
 From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the explicit resilience problem for TS.
+is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
 
-\problemx{$k$-resilience problem}
-{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+\problemx{$k$-Resilience problem or Explicit Resilience problem}
+{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
+{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
+If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
 
@@
 
-We can also ask whether there exists such a bound $k$. We call this problem the resilience problem.
+We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
 
-\problemx{Resilience problem}
-{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+\problemx{Bounded Resilience problem}
+{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
+{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+
+\alain{repetition...} plus court: equivaut à demander si $\exists k \geq 0$ tel que $S$ est $k$-resilient.
+
+\problemx{Bounded Resilience problem}
+{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
+{$\exists k \geq 0  S$ is $k$-resilient. ?\newline}
+
+pour $k,k'$ donnés, chercher si $SAFE_{max}$ et $BAD_{max}$ ont un sens ?
+on peut faire l'union des SAFE, SAFE' et des BAD, BAD' prendre le max des k.
+Fixer k, chercher et calculer $SAFE_{max}$ et $BAD_{max}$.
 
@@
 
-\begin{remark}
+\begin{proposition}
+coverability, liveness and reachability reduce to resilience.
+\end{proposition}
+
+\begin{proof}
+
 \begin{itemize}
-\item expliquer toutes les variantes de la résilience
 \item J=tout et I=$\uparrow x$ alors resilience = couverture.
 \item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
+\item J=tout sauf  exemple des LCS....
 \end{itemize}
-\end{remark}
+Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
+Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
+Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
+This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
+Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+\end{proof}
+
+expliquer toutes les variantes de la résilience
 
@@
 % \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
-
-\alain{regarder les modèles aut à pile...}
 
@@
 
-\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because a finite basis of  $\uparrow post^*$ can be computed.
+\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because VASS are a finite basis of  $\uparrow post^*$ can be computed.
 
 \begin{corollary}
-{\sc resilience} is decidable for context-free grammars and VASS.
+{\sc resilience} is decidable for context-free grammars.
 \end{corollary}
 
-Context-free grammars and VASS are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.
+Context-free grammars are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.
 
@@
 \begin{proposition}
-Given a WSTS with strong upward compatibility, an effective basis of $\uparrow post^*$ is not computable.
+Given a class of WSTS with strong upward compatibility, there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
 \end{proposition}
@@
 
-\begin{proof}
-Let $J=\downarrow Reach(S)$ and $I = \uparrow t$ where $t \in S$.
-Now resilience remains to the following property: $\forall s \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-Let us translate this property for LCS for which Reach=$\downarrow Reach$: $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
-This property can be used to express the liveness property that there is an infinite run passing infinitely often through control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+\begin{proof} \alain{à relire/refaire}
+
 \end{proof}
 
-Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets but $Reach \neq \downarrow Reach$.
-Another proof could be a reduction of liveness to resilience and liveness is undecidable for many models like reset PN,...
+Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
+Another proof could be a reduction of liveness to resilience and since liveness is undecidable for many models like reset PN,...resilience is also undecidable for WSTS with strong monotony.
 
 and with effective basis of $post^*$ ???
+
+\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}

### 2022-12-05 definitions.tex
details

@@
 
-\problemx{(I,J)-$k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
-{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-max des I,J possibles \\
-
-J=tout \\
-\problemx{general $k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
-{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%
+%\problemx{(I,J)-$k$-resilience problem for WSTS}
+%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
+%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%
+%\problemx{general $k$-resilience problem for WSTS}
+%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
+%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 
-Analogously, we formulate the resilience problem for WSTSs.
+%Analogously, we formulate the resilience problem for WSTSs.
 
@@
 
-\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
-avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
+compute the sets of $(I,J,k)$ st S is resilient.\\
+max des I,J possibles \\
+remark: $k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  AVERIFIER !!!
+
 \alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
 
-\begin{proposition}
-coverability, liveness and reachability reduce to resilience.
+\begin{proposition}\label{reductions}
+coverability, liveness, home state and reachability reduce to resilience.
 \end{proposition}
@@
 \begin{itemize}
-\item J=tout et I=$\uparrow x$ alors resilience = couverture.
-\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
-\item J=tout sauf  exemple des LCS....
+\item $J=\uparrow s$ et $I=\uparrow x$ then this particular resilience is coverability of $x$ from $s$.
+\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
+\item $J=S = \downarrow S$
 \end{itemize}
@@
 Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+
+Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
+
+and with effective basis of $post^*$ ???
+
+\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
+avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
 \end{proof}
@@
 
-citer
+Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
 
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and the existence of an algorithm to compute a finite basis of $\uparrow post^*(S)$.
+{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
 \end{theorem}
 
-\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because VASS are a finite basis of  $\uparrow post^*$ can be computed.
+We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
 
 \begin{corollary}
-{\sc resilience} is decidable for context-free grammars.
+{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
 \end{corollary}
 
-Context-free grammars are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.
-
-We may generalyse this last result by strenghthenin to \emph{unbounded} resilience.
-
-\begin{corollary}
-{\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $\uparrow post^*$.
-\end{corollary}
-
-\begin{proof}
-The proof is essentially the same than the previous one.
-\end{proof}
-
-Il y a des pbs
+However, the hypothesis that $\uparrow post^*$ is computable cannot be tested in the WSTS framework.
 
 \begin{proposition}
-Given a class of WSTS with strong upward compatibility, there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
+There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
 \end{proposition}
@@
 \begin{proof}
-$\uparrow post^*$ donne l'accessibilité de 0 sur les compteurs or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
+Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow post^*$ is computable.
+Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS.
+Then, on ewould be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$.
+But reachability of $0$ in a counter $i$ is undecidable for Reset VASS.
+sont WSTS with strong compatibility.
+et l'accessibilité est undecidable.
 $\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
@@
 
-Unfortunately, resilience is undecidable for general WSTS even with strong monotony.
+Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.
 
@@
 \begin{proof} \alain{à relire/refaire}
-
+Since reachability or liveness is undecidable for reset PN that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
 \end{proof}
 
-Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
-Another proof could be a reduction of liveness to resilience and since liveness is undecidable for many models like reset PN,...resilience is also undecidable for WSTS with strong monotony.
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded resilience is decidable for VASS, hence resilience is decidable for VASS.
+\alain{resilience decidable for LCS ?
+LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
 
-and with effective basis of $post^*$ ???
+We may consider context-free grammars.
+
+\begin{corollary}
+{\sc resilience} is decidable for context-free grammars.
+\end{corollary}
+
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
+resilience decidable.
 

### 2022-12-15 definitions.tex
details

@@
 
+justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
+
+et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
+
 compute the sets of $(I,J,k)$ st S is resilient.\\
@@
 
+\begin{proof}
+\textcolor{red}{à faire Mathieu}
+\end{proof}
+
 However, the hypothesis that $\uparrow post^*$ is computable cannot be tested in the WSTS framework.
@@
 \begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
+There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow post^*$.
 \end{proposition}
@@
 $\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
+\textcolor{red}{à relire par Mathieu}
 \end{proof}
+
+il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
 
@@
 
+\begin{proposition}(à prouver)
+ leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $post^*$ ?
+\end{proposition}
+
+\begin{proof}
+leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
+ \textcolor{red}{à faire Mathieu}
+\end{proof}
+
 \begin{theorem}(à prouver)
-{\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility and effective basis of $post^*$ ???
+{\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility, not necessarly strong, and an effective basis of $post^*$ ???
 \end{theorem}
@@
 \begin{proof}
-leur algorithme termine mais ne marche pas car l'intersection de post et J car ce n'est pas pertinent. et donc ???
+ \textcolor{red}{à faire Mathieu}
 \end{proof}
@@
 \begin{proposition}(à prouver)
-{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility ????
+{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility but without basis of post* ????
 \end{proposition}
@@
 \begin{proof}
-pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?
+pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{à faire Mathieu}
 \end{proof}
@@
 \begin{proof} \alain{à relire/refaire}
-Since reachability or liveness is undecidable for reset PN that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
+Since liveness is undecidable for reset PN that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
+  \textcolor{red}{à faire Mathieu}
 \end{proof}
+
+Let us deduce that effective basis of $post^*$ ex reset PN ?
 
@@
 \begin{corollary}
-{\sc resilience} is decidable for context-free grammars.
+{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
 \end{corollary}
 
+\begin{proof}
+
+  \textcolor{red}{à faire Mathieu}
 Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
 resilience decidable.
+\end{proof}
 

### 2023-01-10 definitions.tex
Some small update

@@
 $s_1, t_1 , s_2 \in S$
-	with $s_1 \leq t_1$  and $s_1 \rightarrow s_2$ , there exists 
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$ , there exists 
 	$t_2 \in S$ with 
-	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
-				(resp. $t_1 \rightarrow t_2$ ).
+	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$ 			
+				(resp. $s_2 \rightarrow t_2$ ).
 \end{definition}
@@
 
-\begin{fact}
+\begin{fact}\label{fact basis}
 % Fact 3 ([1]). 
@@
 \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
+
 \end{fact}
@@
 
-\begin{fact}
+\begin{fact}\label{stop condition}
 % Fact 4 (stop condition). 
@@
 pre(\uparrow \{s' \}).$ ?}
+
 Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
-Fact 1, we can compute a basis of $I^{k+1}$.
+Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
 \end{proof}
-\alain{pas de Fact 3}
 \subsection{Defining resilience}
@@
 \begin{proof}
-\textcolor{red}{à faire Mathieu}
+From Fact~\ref{stop condition}, there exists $k \in \N$ such that
+$pre^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
+$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
+returning $k_m$ if that is the case.
+Then, because $k_m$-resilience is decidable, 
+checking $\uparrow post^*(s) \cap J \subseteq I^k = pre^*(I)$ is,
+and resilience is decidable.
+% \textcolor{red}{à faire Mathieu}
 \end{proof}

### 2023-01-10 proof.tex
up
Some small update

@@
 Let $A \subseteq S$, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
-Then $A \cap J \subseteq I \leftrightarrow (\uparrow  A) \cap J \subseteq I$.
+Then $A \cap J \subseteq I$  iff $ (\uparrow  A) \cap J \subseteq I$.
 \end{lemma}
@@
 For all $k \in \N$,
-$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
+$ post^*(s)\cap J \subseteq I^k $  iff $ (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
 % et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
@@
 
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. upward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
 Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
 (on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
-Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. \textcolor{red}{Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
-eventually be found.}
-\alain{attention D est clos par le bas et $\uparrow post^*(s)$ est clos par le haut, je ne vois pas pourquoi $\uparrow post^*(s) \subseteq D$}
+Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
+eventually be found.
+% \alain{attention D est clos par le bas et $\uparrow post^*(s)$ est clos par le haut, je ne vois pas pourquoi $\uparrow post^*(s) \subseteq D$}
+% \mathieu{Oui effectivement les D devraient être clos par le haut, je corrige ça à la prochaine update.}
 
@@
 % By monotonicity, there exists b'' ≥ b' such that a → b'' . Therefore, x → ∗ b'' and b' ≥ b, hence b ∈ ↓ Post ∗ (x).
-\textcolor{red}{By downward compatibility 
-there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}
+By downward compatibility 
+there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.
 %\alain{où est cette hypothèse downward compatibility  ?}
@@
 The stop condition is decidable
-and by Fact 4 also sufficient. 
-\alain{Fact 4 don't exist}
+and by Fact~\ref{stop condition} also sufficient. 
 \end{proof}

### 2023-01-10 definitions.tex
Some small update
Up

@@
 
-\begin{lemma}  
+\begin{lemma}  \label{upward-closed stablizes}
 If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
@@
 \begin{definition}{ (index)}. 
-For an upward-closed set $I \subseteq S$ and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} pre^j(I)$.
+If $S$ is a WSTS with strong monotony and $I \subseteq S$  is an upward-closed set and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} pre^j(I)$.
 The {\em index} $k(I)$ is the
@@
 
+If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
+every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
+Remark that $I^{k+1} = I \cup \pred(I^k)$, 
+%thus $I^{k+1} = I^k$ implies $k(I) \leq k$.
+which ensures the following.
+
 \begin{fact}\label{stop condition}
 % Fact 4 (stop condition). 
-Let $I \subseteq S$ be an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
-$k(I) \leq k$. This also implies that $pre^* (I) = I^k $.
+If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
+$k(I) \leq k$. This also implies that $\pred^*(I) = I^k$.
 \end{fact}

### 2023-01-10 definitions.tex
Harmonizing notations Pred Pre & pre

@@
 \begin{definition}{ (index)}. 
-If $S$ is a WSTS with strong monotony and $I \subseteq S$  is an upward-closed set and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} pre^j(I)$.
+If $S$ is a WSTS with strong monotony and $I \subseteq S$  is an upward-closed set and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} \pred^j(I)$.
 The {\em index} $k(I)$ is the
@@
 We have
-$$I^{k+1} = I \cup pre(I^k ) = I \cup
+$$I^{k+1} = I \cup \pred(I^k ) = I \cup
 \bigcup_{s' \in B}
-pre(\uparrow \{s' \}).$$
-\alain{n'est-ce pas $I^{k+1} = I^k \cup pre(I^k ) = I^k \cup
+\pred(\uparrow \{s' \}).$$
+\alain{n'est-ce pas $I^{k+1} = I^k \cup \pred(I^k ) = I^k \cup
 \bigcup_{s' \in B}
-pre(\uparrow \{s' \}).$ ?}
+\pred(\uparrow \{s' \}).$ ?}
 
-Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
+Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
 Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
@@
 \alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
-avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
+avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
 \end{proof}
@@
 From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$pre^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
+$\pred^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
 $I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
@@
 Then, because $k_m$-resilience is decidable, 
-checking $\uparrow post^*(s) \cap J \subseteq I^k = pre^*(I)$ is,
+checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
 and resilience is decidable.

### 2023-01-11 definitions.tex
Remrking that Ozkan and all have put down a condition for how to ensure a basis of up post*(s) is computable. The goal with a proof not using such a basis would then be to have a lesser condition.

@@
 
-\newcommand{\pred}{\textsc{Pred}}
-\renewcommand{\succ}{\textsc{Succ}}
+\newcommand{\pred}{\textsf{pred}}
+\newcommand{\post}{\textsf{post}}
+% \renewcommand{\succ}{\textsf{Succ}}
 
@@
 The set of {\em successors} of a configuration $s \in S$ is deﬁned as
-$\succ(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
+$\post(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
 A configuration without successors is called a {\em dead end}.  \\
@@
 We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
-for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\succ$ and $\pred$...
+for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...
 
@@
 
-A transition system is {\em finitely branching} if all $\succ(s)$ are finite. 
+A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
 \textcolor{blue}{We restrict our attention to finitely branching TSs.}
@@
 every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
-Remark that $I^{k+1} = I \cup \pred(I^k)$, 
-%thus $I^{k+1} = I^k$ implies $k(I) \leq k$.
-which ensures the following.
+
+\begin{remark}
+$I^{k+1} $ can be rewritten $I^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
+I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
+I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
+=  I \cup \pred(I^k)$.
+\end{remark}
+
+This ensures the following.
 
@@
 \pred(\uparrow \{s' \}).$ ?}
+\mathieu{À strictement parler, c'est  $I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
+I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
+I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
+=  I \cup \pred(I^k)$. Je vais rajouter une remarque au dessus.}
 
@@
 From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\pred^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
+$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
 $I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
@@
 
-However, the hypothesis that $\uparrow post^*$ is computable cannot be tested in the WSTS framework.
+\"Ozkan goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
+given a state $s$ of a WSTS
+% in the regarded class 
+with strong monotonicity 
+and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$ (Proposition~1 in \cite{DBLP:conf/gg/Ozkan22}).
+The characterization is used to show that Petri nets are $\post^*$ -effective.
+It is well-known that Petri nets constitute WSTS with storng monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS0}).
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the WSTS framework.
 
 \begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow post^*$.
+There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
 \end{proposition}
@@
 \begin{proof}
-Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow post^*$ is computable.
+Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow \post^*$ is computable.
 Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS.

### 2023-01-11 definitions.tex
Searching for counter-examples

@@
 The characterization is used to show that Petri nets are $\post^*$ -effective.
-It is well-known that Petri nets constitute WSTS with storng monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS0}).
+It is well-known that Petri nets constitute WSTS with storng monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the WSTS framework.
@@
 Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS.
-Then, on ewould be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$.
+Then, one would be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$.
 But reachability of $0$ in a counter $i$ is undecidable for Reset VASS.
@@
  \textcolor{red}{à faire Mathieu}
+
+\mathieu{En train de chercher un contre exemple pour le Lemme~\ref{Lemma intersection}.}
+
 \end{proof}

### 2023-01-12 definitions.tex
up

@@
 
-\textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
+% \textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
 
@@
 
-Properties in well-structured transition systems are often given as upward- or downward closed sets [references].
+Properties in well-structured transition systems are often given as upward- or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
 %  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
@@
 
+\textcolor{blue}{
 justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
+}
 
+\textcolor{blue}{
 et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
+}
 
+\textcolor{blue}{
 compute the sets of $(I,J,k)$ st S is resilient.\\
@@
 remark: $k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  AVERIFIER !!!
+}
 
@@
 
-\"Ozkan goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
+The author in \cite{DBLP:conf/gg/Ozkan22}
+%\"Ozkan
+ goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
 given a state $s$ of a WSTS
@@
 and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$ (Proposition~1 in \cite{DBLP:conf/gg/Ozkan22}). 
-The characterization is used to show that Petri nets are $\post^*$ -effective.
-It is well-known that Petri nets constitute WSTS with storng monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the WSTS framework.
+The characterization is used to show that Petri nets are $\post^*$-effective.
+It is well-known that Petri nets constitute WSTS with strong monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
 
@@
 But reachability of $0$ in a counter $i$ is undecidable for Reset VASS.
-sont WSTS with strong compatibility.
-et l'accessibilité est undecidable.
+% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
 $\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???

### 2023-01-25 definitions.tex
Up

@@
 
+\begin{remark} 
+We require $U$ to be upward-closed, as, in the general case if we lift restrictions on $U$ and $D$, then reachability reduces itself to resilience.
+Indeed, the particular resilience problem where we take $D = \{s\}$ and $U = \{x\}$ is reachability of $x$ from $s$, and thus generally undecidable for WSTS.
+\end{remark}
+
 \textcolor{blue}{
@@
 max des I,J possibles \\
-remark: $k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  AVERIFIER !!!
 }
+
+\begin{remark} 
+$k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  \textcolor{red}{AVERIFIER !!}
+\end{remark}
 
@@
 \begin{proposition}\label{reductions}
-coverability, liveness, home state and reachability reduce to resilience.
+coverability, liveness, home state 
+%and reachability
+ reduce to resilience.
 \end{proposition}
@@
 \begin{itemize}
-\item $J=\uparrow s$ et $I=\uparrow x$ then this particular resilience is coverability of $x$ from $s$.
-\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
-\item $J=S = \downarrow S$
+\item $J=S= \downarrow S$ et $I=\uparrow x$ then this particular resilience is coverability of $x$ from $s$.
+%\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
+
+\item $t$-liveness c'est la résilience avec $J=S$ et $I = \uparrow \pred(t)$.
 \end{itemize}
-Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
+
+\textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
 Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
@@
 This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
+\mathieu{À refaire}
 
@@
 avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
+
 \end{proof}
@@
   \textcolor{red}{à faire Mathieu}
+
+Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
+Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
+Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
+\textcolor{red}{This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$ (à prouver dans les deux sens!)}; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
+Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+
+\mathieu{(Sketch)
+$\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ 
+expresses that there is an infinite
+run passing infinitely often through the control-state $q$
+
+d'abords direction 1
+
+$\forall s' \in S,  \exists w \in \Sigma^* $ $ ~ s' \rightarrow^{*} (q,w)$      $\implies$   there is an infinite run passing infinitely often through ...
+
+$(q,w) \to s \to^* (q,w')$
+
+globalement
+
+sinon
+
+l'autre direction
+
+si il y a un run infini
+
+c'est là que ça bloquait si je me souviens bien quand on avait travaillé dessus au tableau.
+
+Sinon il faut prouver que si on n'a pas le truc alors on n'a pas le run infini
+
+Si on n'a pas le truc
+
+On a $s'$ tel que si on arrive en $s'$ alors on ne peut plus retourner en $q$.
+
+Et $s'$ est accessible
+
+MAIS ça veut pas dire qu’il n’y a pas de run infini
+
+Plutôt que le Recurrent State Problem (RSP) on peut se concentrer plutôt du coup sur la $t$-liveness telle que décrite plus haut
+
+$t$-liveness s'exprime par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
+avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$
+
+$t$-liveness c'est la résilience avec $J=S$ et $I = \uparrow \pred(t)$.
+
+Vérifier que la $t$-liveness c'est indécidable pour les LCS.
+}
+
 \end{proof}

### 2023-01-27 definitions.tex
details

@@
 \end{definition}
-
-A \emph{downward-WSTS} is ... \\
 
@@
 
-\begin{corollary}
+\begin{corollary}\label{postcomputable}
 {\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.

### 2023-01-27 proof.tex
Harmonizing notations Pred Pre & pre
Searching for counter-examples
details

@@
-\section{Decidability for WSTS with downward/strong upward compatibility}
+\section{Decidability of resilience for downward-WSTS}
 
-\begin{lemma}
+Let us recall the definition of downward-WSTS.
+
+\begin{definition} \cite{DBLP:journals/tcs/FinkelS01}
+A \emph{downward-WSTS} is an ordered transition system $(S, \rightarrow, \leq)$
+such that $\leq$ is a wqo and
+%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
+the transition relation $ \rightarrow$ is downward-compatible with $\leq$, i.e., for all  transition $s_1 \rightarrow t_1$
+with $s_1, t_1 , s_2 \in S$
+	and $s_2 \leq s_1$, there exists a transition $s_2 \rightarrow^{*} t_2$ with $t_2 \leq t_1$. 		
+\end{definition}
+
+%		enlever l'hypothèse finitely branching ?
+A downward-WSTS has the \emph{reflexive} compatibility if for all transition  $s_1 \rightarrow t_1$ with
+$s_1, t_1 , s_2 \in S$ and $s_2 \leq s_1$, either $t_1 \leq s_2$ or there exists a transition $s_2 \rightarrow t_2$ with $t_2 \leq t_1$.
+
+Downward-WSTS enjoy a nice reachability property.
+
+\begin{theorem}[Proposition 5.4 in \cite{DBLP:journals/tcs/FinkelS01}]
+%	Since by Proposition 5.4 in \cite{DBLP:journals/tcs/FinkelS01}, we know that  
+For finitely branching downward-WSTS $(S, \rightarrow, \leq)$, with reflexive compatibility, decidable $\leq$ and effective $post$, a finite basis of $\uparrow post^*(s_0)$ is computable for any $s_0 \in S$.
+\end{theorem}
+
+%	Let us say that a downward-WSTS is \emph{effective}
+Using Corollary \ref{postcomputable}, we deduce that resilience is decidable for finitely branching downward-WSTS with reflexive compatibility and with strong upward compatibility, with a decidable $\leq$ and an effective $post$.
+
+%%%%%
+\alain{peut-on faire mieux ?}
+
+\begin{lemma}\label{Lemma intersection}
 Let $A \subseteq S$, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
@@
 This result is not in xxx and don't use the problematic hypothsis on $post^*$.
+\alain{ben si car downward-WSTS + qqhypothèses (voir plus haut) ont un $ \uparrow post^*(s)$ calculable !
+peut-on faire sauter qq hypothèses ?  finitely branching ??? with reflexive compatibility ????}
 
@@
 We define inductively
-$I^{k} = \bigcup_{0 \leq j \leq k} pre^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
+$I^{k} = \bigcup_{0 \leq j \leq k} \pred^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
 the strongly upward compatibility of $(S, \rightarrow, \leq)$.
@@
 \begin{theorem}
-{\sc bounded resilience} and {\sc resilience} is decidable for WSTS with downward/strong upward compatibility.
+{\sc bounded resilience} and {\sc resilience} are decidable for WSTS with downward/strong upward compatibility.
 \end{theorem}

### 2023-01-27 definitions.tex
details

@@
 \end{proof}
+
+%%%
+\iffalse
+%%%%
+%%%%
 \subsection{Defining resilience}
@@
 
-\problemx{$k$-Resilience problem or Explicit Resilience problem}
+\problemx{$k$-Resilience problem }
 {a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
@@
 
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-\alain{repetition...} plus court: equivaut à demander si $\exists k \geq 0$ tel que $S$ est $k$-resilient.
+%
+%\problemx{Bounded Resilience problem}
+%{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
+%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%
 
@@
 {$\exists k \geq 0  S$ is $k$-resilient. ?\newline}
+%%
+\fi
+%%%
 
@@
 
-\subsubsection{Resilience for WSTS}
+\section{Resilience for WSTS}
 
@@
 \problemx{$k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $I$ with a given basis, a decidable downward-closed set $J$.}
-{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in J  \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%
+plus simple : la $(D,U)$-resilience: $D \rightarrow^{\leq k} U$ ?
+\alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
+c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ?
+sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}
 
@@
 \problemx{bounded Resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}
 

### 2023-01-27 proof.tex
details

@@
-\section{Decidability of resilience for downward-WSTS}
+\section{Resilience for downward-WSTS}
 
@@
 A downward-WSTS has the \emph{reflexive} compatibility if for all transition  $s_1 \rightarrow t_1$ with
-$s_1, t_1 , s_2 \in S$ and $s_2 \leq s_1$, either $t_1 \leq s_2$ or there exists a transition $s_2 \rightarrow t_2$ with $t_2 \leq t_1$.
+$s_1, t_1 , s_2 \in S$ and $s_2 \leq s_1$, either $t_1 \leq s_2$ or there exists a transition $s_2 \rightarrow t_2$ with $t_2 \leq t_1$.\\
 
-Downward-WSTS enjoy a nice reachability property.
+Downward-WSTS with reflexive compatibility enjoy a nice reachability property.
 
@@
 %%%%%
-\alain{peut-on faire mieux ?}
+\alain{peut-on faire mieux ? enlever des hypothèses ?}
+
+Let us establish a general Lemma on downward-closed and upward-closed sets.
 
 \begin{lemma}\label{Lemma intersection}
-Let $A \subseteq S$, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
+Let $A \subseteq S$, $J \subseteq S$ be a downward-closed set and $I \subseteq S$ be an upward-closed set. 
 Then $A \cap J \subseteq I$  iff $ (\uparrow  A) \cap J \subseteq I$.
@@
 \begin{proof}
-
-If 
-$A \cap J \subseteq I$,
-then
-${\uparrow (A \cap J)} \subseteq {\uparrow I} = I$.
-However
-$({\uparrow A}) \cap J \subseteq {\uparrow (A \cap J)}$.
-Indeed if $x$ in $({\uparrow A} \cap J)$,
-$x$ is in $\uparrow A$
-so there exists $a \in A$ such that $x \geq a$.
-However $x$ is also in $J$, meaning $a$ is in $J$.
-Meaning $a$ is in $(A \cap J)$
-and $x \in { \uparrow (A \cap J)}$.
-
+Let us suppose that $A \cap J \subseteq I$. Then ${\uparrow (A \cap J)} \subseteq {\uparrow I} = I$.
+Show that $({\uparrow A}) \cap J \subseteq {\uparrow (A \cap J)}$.
+Let $x \in ({\uparrow A}) \cap J$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in J$ and $J$ is downward-closed, we also have $a \in J$.
+Hence $a \in A \cap J$ and then $x \in { \uparrow (A \cap J)}$.
 In the other direction,
-since $A \subseteq {\uparrow A}$,
+since $A \subseteq {\uparrow A}$, the inclusion
 $({\uparrow  A}) \cap J \subseteq I$ implies
@@
 
+Let us apply the previous Lemma with $A=post^*(s)$.
+
 \begin{corollary}
-For all $k \in \N$,
+For all $k \in \N$ and $s \in S$,
 $ post^*(s)\cap J \subseteq I^k $  iff $ (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
@@
 
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. upward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
+{\bf Explanations on the two procedures.}
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , .
+.
+.
+, i.e.
+upward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$.
+\alain{write U for upward closed subset and not D as for downward...}
 Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
@@
 of $I^{k+1}$ if we have a basis of $I^k$. %This follows by the proof of Lemma 3.
-Due to the Lemma, 
+Due to the Lemma ????, 
 $b(D)  \cap J \subseteq I^k$ implies
@@
 $\uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
-(since $D$ contains $ \uparrow post^*(s)$).
+(since $ \uparrow post^*(s)  \subseteq D$).
 
@@
 peut-on faire sauter qq hypothèses ?  finitely branching ??? with reflexive compatibility ????}
-
-\begin{theorem}
-{\sc $k$-resilience} is decidable for effective WSTS with downward/strong upward compatibility.
-%	with strong monotony.
-\end{theorem}
-
-\begin{proof}
-
-Assume  $(S, \rightarrow, \leq)$ is a WSTS with downward/upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
-
-We define inductively
-$I^{k} = \bigcup_{0 \leq j \leq k} \pred^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
-the strongly upward compatibility of $(S, \rightarrow, \leq)$.
-
-% Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
-
-The $k$-resilience property can be expressed as the formula
-$ post^*(s) \cap J \subseteq I^k$. In order to decide whether the inclusion holds, we execute two procedures in parallel, one trying to prove $ post^*(s)\cap J \subseteq I^k$ 
-and one looking for a counter example.
-
-In order to certify inclusion in $I^k$, we need to work with finite representations.
-The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.
 
@@
 }
-	\caption{\textbf{Procedure 1:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
+	\caption{\textbf{Procedure RES:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
 \end{figure}
@@
 }
-	\caption{\textbf{Procedure 2:} searches for a non-inclusion certificate.}\label{procedure2}
+	\caption{\textbf{Procedure NOT-RES:} searches for a non-inclusion certificate.}\label{procedure2}
 \end{figure}
 
-\newpage
+We show that these two procedures allow to solve the $k$-resilience.
 
-We show that these two procedures are correct:
+\begin{proposition}
+The procedures $RES$ and $NOT-RES$ have the following property.
+\begin{itemize}
+%
+\item $k$-resilience holds if and only if Procedure $RES$ terminates.
+%
+\item $k$-resilience do not hold if and only if Procedure $NOT-RES$ terminates.
+%
+\end{itemize}
+\end{proposition}
 
-\begin{enumerate}
-
-\item $k$-resilience holds if, and only if, Procedure 1 terminates.
-
-\item $k$-resilience do not hold if, and only if, Procedure 2 terminates.
-
-\end{enumerate}
-
-Proof:
+\begin{proof}
 
@@
 \begin{theorem}
+{\sc $k$-resilience} is decidable for effective \alain{pas défini}WSTS with downward/strong upward compatibility.
+%	with strong monotony.
+\end{theorem}
+\alain{les resultatts et preuves ne sont pas faciles à lire: faire un seul niveau de parentheses}
+\begin{proof}
+%			\begin{proof}
+
+Assume  $(S, \rightarrow, \leq)$ is a WSTS with downward/upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.
+We define inductively
+$I^{k} = \bigcup_{0 \leq j \leq k} \pred^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
+the strongly upward compatibility of $(S, \rightarrow, \leq)$.
+
+% Si on inverse les types de propriétés (downward closed pour $I = SAFE$, upward closed pour $J= BAD$) alors on peut écrire un lemme symmétrique du Lemme $4$:
+
+The $k$-resilience property can be expressed as the formula
+$ post^*(s) \cap J \subseteq I^k$. In order to decide whether the inclusion holds, we execute two procedures in parallel, one trying to prove $ post^*(s)\cap J \subseteq I^k$ 
+and one looking for a counter example.
+
+%    In order to certify inclusion in $I^k$, we need to work with finite representations.
+%    The next lemma \alain{where ? the previous lemmas ? numbers ?} 
+%    uses that $I$ and $J$ are upward- and downward-closed, respectively.
+\end{proof}
+
+\begin{theorem}
 {\sc bounded resilience} and {\sc resilience} are decidable for WSTS with downward/strong upward compatibility.
+\alain{+effective hypothesis}
 \end{theorem}
@@
 \end{proof}
-
-{\sc resilience} is decidable for WSTS with strong upward compatibility and k-resilience.
-
-jouer avec les hypothèses downward/upward compatibility
+%
+%{\sc resilience} is decidable for WSTS with strong upward compatibility and k-resilience.
+%
+%jouer avec les hypothèses downward/upward compatibility

### 2023-01-27 definitions.tex
details

@@
 
-Properties in well-structured transition systems are often given as upward- or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
+The $k$-resilience problem is to decide whether from a bad state, is it possible to find a path of length smaller than a given $k$ to reach a safe state.
+Let us define two sets BAD and SAFE.
+From all states $b \in BAD$ does there exist a path of length smaller than $k$ that reaches a state $s \in SAFE$ ?
+
+It can be formally stated as follows :
+
+\problemx{$k$-resilience problem}
+{A transition system $(S,\rightarrow)$, two sets $SAFE, BAD \subseteq S$.}
+{$BAD \longrightarrow^{\leq k} SAFE$ ?\newline}
+%
+
+For most of transition systems, this problem is decidable if both SAFE and BAD are finite.
+When one of the two sets is infinite, it generally becomes undecidable.
+So we restrict the framework to WSTS and sets SAFE and BAD will be upward or downward closed.
+In \cite{DBLP:conf/gg/Ozkan22} BAD is downward closed and SAFE is upward closed.
+But there are other possible cases: for instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
+Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE.
+%
+
+Most decidable properties in well-structured transition systems are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
 %  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
@@
 it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
-upward-closed or 
-downward-closed sets.
+upward-closed or downward-closed sets.
 
-We assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set.
+\begin{tabular}{ l c r }
+   BAD/SAFE & Up & Down \\
+   Up & dec ? & dec ? \\
+   Down & dec & ? \\
+ \end{tabular}
 
+We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
+\alain{très discutable. pourquoi pas supposer des conditions plus gnérales ? perd-on la décidabilité ?}
 % \textcolor{red}{Seems like a reasonable assumption to me.}
 
-From these considerations, we formulate instances of the abstract resilience problems for well-
-structured transition systems.
+%From these considerations, we formulate instances of the abstract resilience problems for well-
+%structured transition systems.
 
@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+{$D \longrightarrow^{\leq k} U$ ?\newline}
+%
 {$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 %
-plus simple : la $(D,U)$-resilience: $D \rightarrow^{\leq k} U$ ?
  \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
@@
 sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}
+
+ou plus simple : $D \rightarrow^{\leq k} U$ ?
 
@@
 
-\textcolor{blue}{
-justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
-}
+%\textcolor{blue}{
+%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
+%}
 

### 2023-01-30 definitions.tex
Cleaning up a bit

@@
 \begin{proposition}\label{reductions}
-coverability, liveness, home state 
+coverability, $t$-liveness, home state 
 %and reachability
@@
 
-\item $t$-liveness c'est la résilience avec $J=S$ et $I = \uparrow \pred(t)$.
+\item $t$-liveness  can be expressed as the following formula
+$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
+where
+$U_t=\uparrow \pred(t)$
+is the upward closure of the preconditions to use transition $t$.  
+The problem reduces itself to the resilience problem
+where $J=S$ and $I = \uparrow \pred(t)$.
+
+\item the home state problem can be expressed as $\ldots$
 \end{itemize}
 
-\textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
+%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
+%  \textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
 Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
@@
 This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
-\mathieu{À refaire}
+% Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
+%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
+\end{proof}
 
@@
 
-\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
+% \alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
 avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
-
-\end{proof}
 
@@
 
-\begin{proof} \alain{à relire/refaire}
-Since liveness is undecidable for reset PN that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
-  \textcolor{red}{à faire Mathieu}
+\begin{proof}
 
-Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
-Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
-\textcolor{red}{This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$ (à prouver dans les deux sens!)}; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
 
-\mathieu{(Sketch)
-$\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ 
-expresses that there is an infinite
-run passing infinitely often through the control-state $q$
+\end{proof}
 
-d'abords direction 1
+\begin{corollary}
+{\sc Resilience} is undecidable for WSTS.
+\end{corollary}
 
-$\forall s' \in S,  \exists w \in \Sigma^* $ $ ~ s' \rightarrow^{*} (q,w)$      $\implies$   there is an infinite run passing infinitely often through ...
+\begin{theorem}
+In WSTS with strong monotony, 
+the {\sc bounded resilience} problem
+is equivalent to the
+{\sc resilience} problem.
+\end{theorem}
 
-$(q,w) \to s \to^* (q,w')$
+\begin{proof}
+strong WSTS
 
-globalement
+$\pred^*(I) = \pred^{k}(I)$
 
-sinon
+\end{proof}
 
-l'autre direction
+\begin{corollary}
+Bounded Resilience is undecidable for WSTS with strong monotony.
+\end{corollary}
 
-si il y a un run infini
-
-c'est là que ça bloquait si je me souviens bien quand on avait travaillé dessus au tableau.
-
-Sinon il faut prouver que si on n'a pas le truc alors on n'a pas le run infini
-
-Si on n'a pas le truc
-
-On a $s'$ tel que si on arrive en $s'$ alors on ne peut plus retourner en $q$.
-
-Et $s'$ est accessible
-
-MAIS ça veut pas dire qu’il n’y a pas de run infini
-
-Plutôt que le Recurrent State Problem (RSP) on peut se concentrer plutôt du coup sur la $t$-liveness telle que décrite plus haut
-
-$t$-liveness s'exprime par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
-avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$
-
-$t$-liveness c'est la résilience avec $J=S$ et $I = \uparrow \pred(t)$.
-
-Vérifier que la $t$-liveness c'est indécidable pour les LCS.
-}
-
+\begin{proof} 
+  From
+  cf
+the two  theorems above. 
 \end{proof}

### 2023-01-31 definitions.tex
details

@@
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$D \longrightarrow^{\leq k} U$ ?\newline}
+{$D \longrightarrow^{\leq k} U$ ?  ou $D \cap post^*(s) \longrightarrow^{\leq k} U$ ?  \newline}
 %
-{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%  		{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 %
+$D \longrightarrow^{\leq k} U$: $\forall s' \in D \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\\
+%
+$D \cap post^*(s) \longrightarrow^{\leq k} U$ ?
+%
+$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?
  \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
@@
 
-ou plus simple : $D \rightarrow^{\leq k} U$ ?
+%ou plus simple : $D \rightarrow^{\leq k} U$ ?
 

### 2023-01-31 definitions.tex
details

@@
 
-\problemx{$k$-resilience problem}
+\problemx{uniform $k$-resilience problem}
 {A transition system $(S,\rightarrow)$, two sets $SAFE, BAD \subseteq S$.}
@@
 %
+\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
 

### 2023-02-02 definitions.tex
up

--- 2023-01-30 definitions.tex
@@
 \fi
+\mathieu{On peut probablement enlever la Proposition $8$ alors.
+La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
 
@@
 $s \in J$. 
-Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
+\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
 \alain{remarque pas claire}
 % \textcolor{red}{Do we still demand this ?}
+\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
 
@@
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-(i) For every upward-closed set $I \subseteq S$, there exists a finite basis $B$ of $I$.
-
-(ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-\alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
-
+For every upward-closed set $I \subseteq S$, there exists a finite basis $B$ of $I$. 
+% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
+% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
 \end{fact}

### 2023-02-02 definitions.tex
details

@@
 
-\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
+%\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
 
@@
 
-\textcolor{red}{This makes sense for TS but not so much for LTS ...}
+%\textcolor{red}{This makes sense for TS but not so much for LTS ...}
 
@@
 
-\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
-\textcolor{red}{Not sure that TS induced by TA are finitely branching.
+%\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
+%\textcolor{red}{Not sure that TS induced by TA are finitely branching.
 Actually I believe they are not, i.e.
@@
 $(q, x \geq 3, \emptyset, q')$ then the set of successors of
-$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
+%$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
 

### 2023-02-02 definitions.tex
details
Up

@@
 A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
-\textcolor{blue}{We restrict our attention to finitely branching TSs.}
+%\textcolor{blue}{We restrict our attention to finitely branching TSs.}
 
@@
 \end{proposition}
-\alain{étrange ces deux propositions, l'une suffirait}
+%\alain{étrange ces deux propositions, l'une suffirait}
 \iffalse
@@
 \textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
-\alain{remarque pas claire}
 % \textcolor{red}{Do we still demand this ?}
@@
 
+\iffalse
+\begin{remark}
+$I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
+I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
+I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
+=  I \cup \pred(I^k)$.
+\end{remark}
+\fi
+
 \begin{lemma}
@@
 \pred(\uparrow \{s' \}).$$
-\alain{n'est-ce pas $I^{k+1} = I^k \cup \pred(I^k ) = I^k \cup
-\bigcup_{s' \in B}
-\pred(\uparrow \{s' \}).$ ?}
-\mathieu{À strictement parler, c'est  $I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$. Je vais rajouter une remarque au dessus.}
 
@@
 
-\section{Resilience for WSTS}
+\section{Uniform resilience for WSTS}
 
@@
 
+\section{Resilience for WSTS}
+
 \begin{remark} 
@@
 \begin{proposition}\label{reductions}
-coverability, $t$-liveness, home state 
-%and reachability
- reduce to resilience.
+coverability are $t$-liveness reduceable to resilience.
+%	home state and reachability
 \end{proposition}
@@
 \begin{itemize}
-\item $J=S= \downarrow S$ et $I=\uparrow x$ then this particular resilience is coverability of $x$ from $s$.
+\item Let us choose $J=S= \downarrow S$ and $I=\uparrow x$ then resilience is equivalent to coverability of $x$ from $s$.
 %\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
@@
 
-\item the home state problem can be expressed as $\ldots$
+%\item the home state problem can be expressed as $\ldots$
 \end{itemize}

### 2023-02-02 definitions.tex
details

@@
 
+\alain{dire juste que SAFE = upward closed et BAD = downward closed.
+envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
+4 resilience problem (U,D), (U,U), (D,U), (D,D)}
+
 \section{Uniform resilience for WSTS}
@@
 \begin{remark} 
-$k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  \textcolor{red}{AVERIFIER !!}
+$k$-resilienc $\leq$ bounded Resilience 
+%	$\leq$ Resilience.  
+\textcolor{red}{oui si strong monootony, non ? sinon}
 \end{remark}
 
-\alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}
-
 \begin{proposition}\label{reductions}
-coverability are $t$-liveness reduceable to resilience.
+coverability and $t$-liveness are reduceable to resilience.
 %	home state and reachability
@@
 
-\alain{dire juste que SAFE = upward closed et BAD = downward closed.
-envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
-4 resilience problem (U,D), (U,U), (D,U), (D,D)}
 % \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
@@
 %\"Ozkan
+\begin{proposition}
 goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
@@
 and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$ (Proposition~1 in \cite{DBLP:conf/gg/Ozkan22}). 
+\end{proposition}
+
 The characterization is used to show that Petri nets are $\post^*$-effective.
@@
 \begin{proof}
-Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow \post^*$ is computable.
+Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
 Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS.
@@
 % sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
+%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
 \textcolor{red}{à relire par Mathieu}
@@
 
-il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
+%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
 

### 2023-02-02 definitions.tex
details

@@
 
-The author in \cite{DBLP:conf/gg/Ozkan22}
 %\"Ozkan
-\begin{proposition}
+\begin{proposition} \cite{DBLP:conf/gg/Ozkan22}
 goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
@@
 It is well-known that Petri nets constitute WSTS with strong monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
+Hence Resilience is decidable for Petri nets and VASS.
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
@@
 
-\begin{theorem}(à prouver)
-{\sc bounded resilience} is undecidable ??? for WSTS with upward compatibility, not necessarly strong, and an effective basis of $post^*$ ???
-\end{theorem}
-
-\begin{proof}
- \textcolor{red}{à faire Mathieu}
-\end{proof}
-
-si on garde strong mais on enlève effective basis of $post^*$:
-
 \begin{proposition}(à prouver)
-{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility but without basis of post* ????
+{\sc bounded resilience} is undecidable for WSTS with upward compatibility withd an effective basis of $post^*$.
 \end{proposition}
@@
 \begin{proof}
-pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{à faire Mathieu}
+ \textcolor{red}{pas certain ! à faire Mathieu}
 \end{proof}
+
+%
+%si on garde strong mais on enlève effective basis of $post^*$:
+%
+%\begin{proposition}(à prouver)
+%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
+%\end{proposition}
+%
+%\begin{proof}
+%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
+%\end{proof}
 
@@
 
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, and since reachability $\leq$ t-liveness,  we deduce that resilience is undecidable for WSTS with strong monotony.
 

### 2023-02-02 definitions.tex
details

@@
 
+\begin{theorem}
+A finite basis of $ \pred^*(U)$ is computable for any WSTS and any upward closed set $U$.
+\end{theorem}
+
 %%%
@@
 %
-\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
+%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
 
@@
 
-\begin{corollary}
-{\sc Resilience} is undecidable for WSTS.
-\end{corollary}
-
 \begin{theorem}

### 2023-02-02 definitions.tex
details

@@
 \begin{definition}
-A {\em labeled transition system} (LTS for short) is a tuple $T=(S, \Lambda, \rightarrow)$ where 
-$S$ is a set of {\em configurations}, $ \Lambda$ is a set of {\em labels}, and 
+A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Lambda, \rightarrow)$ where 
+$S$ is a set of {\em states}, $ \Lambda$ is a set of {\em labels}, and 
 ${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
@@
 An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
-{\em configurations} and  
+{\em states} and  
 $ {\rightarrow} \subseteq S \times S$ is a
@@
 on
-the set of configurations, denoted as the set of {\em transitions}. 
+the set of states, denoted as the set of {\em transitions}. 
 \end{definition}
 
-We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from configuration $s$ to configuration $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
+We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
 Note 
@@
 for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...
+
+Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
+We denote $X \longrightarrow^{*} Y$ (resp.
+$X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
+of length smaller than $k$) that reaches a state $y \in Y$.
 
@@
 \begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
+If $\mathscr{S}$ is an WSTS and $U \subseteq S$ is an upward-closed set of states, then $\pred^*(U )$ is upward-
 closed.
@@
 Proof.
-Assume $s \in \pred^* (I )$.
-Then $s \rightarrow^* t$ for some $t \in I $.
+Assume $s \in \pred^* (U )$.
+Then $s \rightarrow^* t$ for some $t \in U $.
 If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$.
-Then $t' \in I$ and $s' \in \pred^*(I )$.
+Then $t' \in U$ and $s' \in \pred^*(U )$.
 \fi
@@
 \begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set of states, then $\pred(U )$ is upward-closed.
 \end{proposition}
@@
 \begin{definition}\cite{DBLP:journals/iandc/AbdullaCJT00}
-A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
+A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
@@
 $s \in J$. 
-\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
+%\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
 % \textcolor{red}{Do we still demand this ?}
-\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
+%\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
 
 \begin{claim}{(stability of ideals)}
-Let $I, J \subseteq S$ be upward-closed. Then the sets $I \cup J$, and $I \cap J$ are upward-closed.
+Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
 \end{claim}
@@
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-For every upward-closed set $I \subseteq S$, there exists a finite basis $B$ of $I$. 
+For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$. 
 % (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
@@
 \begin{definition}{ (index)}. 
-If $S$ is a WSTS with strong monotony and $I \subseteq S$  is an upward-closed set and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} \pred^j(I)$.
-The {\em index} $k(I)$ is the
-smallest $k_0$ s.t. $I^k = I^{k_0}$ for all $k \geq k_0$.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is an upward-closed set and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
+The {\em index} $k(U)$ is the
+smallest $k_0$ s.t. $U^k = U^{k_0}$ for all $k \geq k_0$.
 \end{definition}
 
-If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
+If $\mathscr{S}$ is a WSTS with strong compatibility and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
 every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
@@
 % Fact 4 (stop condition). 
-If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
+If $\mathscr{S}$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
 $k(I) \leq k$. This also implies that $\pred^*(I) = I^k$.
@@
 \begin{theorem}
-A finite basis of $ \pred^*(U)$ is computable for any WSTS and any upward closed set $U$.
+A finite basis of $ \pred^*(U)$ is computable for any WSTS $\mathscr{S}$ and any upward closed set $U \subseteq S$.
 \end{theorem}
@@
 {a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\exists k \geq 0  S$ is $k$-resilient. ?\newline}
+{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
 %%
@@
 
-The $k$-resilience problem is to decide whether from a bad state, is it possible to find a path of length smaller than a given $k$ to reach a safe state.
-Let us define two sets BAD and SAFE.
-From all states $b \in BAD$ does there exist a path of length smaller than $k$ that reaches a state $s \in SAFE$ ?
+The resilience problem (resp.
+the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
+a path of length smaller than $k$) that reaches a safe state.
+Let us formalize these properties.
 
-It can be formally stated as follows :
-
-\problemx{uniform $k$-resilience problem}
-{A transition system $(S,\rightarrow)$, two sets $SAFE, BAD \subseteq S$.}
-{$BAD \longrightarrow^{\leq k} SAFE$ ?\newline}
+\problemx{uniform resilience problem}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
+{$BAD \longrightarrow^{*} SAFE$ ?\newline}
 %
@@
 
-For most of transition systems, this problem is decidable if both SAFE and BAD are finite.
-When one of the two sets is infinite, it generally becomes undecidable.
-So we restrict the framework to WSTS and sets SAFE and BAD will be upward or downward closed.
-In \cite{DBLP:conf/gg/Ozkan22} BAD is downward closed and SAFE is upward closed.
-But there are other possible cases: for instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
+\problemx{uniform $k$-resilience problem}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $SAFE, BAD \subseteq S$.}
+{$BAD \longrightarrow^{\leq k} SAFE$ ?\newline}
+
+We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
+%
+\problemx{uniform bounded resilience problem}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
+%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-resilient ?\newline}
+
+These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
+So we restrict our framework to the class of infinite-state WSTS.
+Since most of decidable properties in WSTS are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.
+
+For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
 Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE.
+In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that BAD is always downward closed and SAFE is always upward closed.
 %
 
-Most decidable properties in well-structured transition systems are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}.
 %  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
 %  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-Transfering the abstract resilience problems into this framework,
-it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
-upward-closed or downward-closed sets.
+
+%Transfering the abstract resilience problems into this framework,
+%it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
+%upward-closed or downward-closed sets.
 
@@
 
-We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-\alain{très discutable. pourquoi pas supposer des conditions plus gnérales ? perd-on la décidabilité ?}
+%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
 % \textcolor{red}{Seems like a reasonable assumption to me.}
@@
 %structured transition systems.
+
+\section{Resilience for WSTS}
 
@@
 }
-
-\section{Resilience for WSTS}
 

### 2023-02-02 definitions.tex
details

@@
 
-\problemx{uniform resilience problem}
+\problemx{uniform resilience problem (URP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
@@
 
-\problemx{uniform $k$-resilience problem}
+\problemx{uniform $k$-resilience problem (UkRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $SAFE, BAD \subseteq S$.}
@@
 %
-\problemx{uniform bounded resilience problem}
+\problemx{uniform bounded resilience problem (UBRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
@@
 %		
+In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I(D)$ satisfying $D=\downarrow D =I(D) \cap S$and that an upward closed set $U$ is given by its finite set of minimal elements, $b(U)$.
 
@@
 %upward-closed or downward-closed sets.
+
+\begin{proposition}\label{down-up}
+Let $\mathscr{S}$ be an effective WSTS with effective predbasis and let the two finite sets $I(BAD)$ and  $b(SAFE)$ representing $\downarrow BAD$ and $\uparrow SAFE$.
+The uniform resilience problem (URP), the uniform bounded resilience problem (UBRP) and the uniform $k$-resilience problem (UkRP) are decidable
+\end{proposition}
+
+\begin{proof}
+The uniform resilience problem (URP) reduces to decide coverability of the finite set $b(SAFE)$ from the infinite set $BAD$; but it is sufficient to decide coverability from the finite set of maximal elements of $BAD$ which are ideals but not elements in $S$.
+We use the fact that y is coverable from an $x \in BAD$ iff $y$ is coverable from $x' \in I(BAD)$.
+
+Hence it is decidable since
+uniform bounded resilience problem (UBRP) is equivalent to the uniform resilience problem (UBRP) for WSTS with...
+The uniform $k$-resilience problem (UkRP) is also decidable by computing the least $n$ such that $\pred^n(SAFE)=\pred^*(SAFE)$ and then to test whether $n \leq k$.
+\end{proof}
 

### 2023-02-03 definitions.tex
details

@@
 %		
-In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I(D)$ satisfying $D=\downarrow D =I(D) \cap S$and that an upward closed set $U$ is given by its finite set of minimal elements, $b(U)$.
+In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I(D)=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = S \cap \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set of minimal elements, $b(U)$.
 
@@
 Let $\mathscr{S}$ be an effective WSTS with effective predbasis and let the two finite sets $I(BAD)$ and  $b(SAFE)$ representing $\downarrow BAD$ and $\uparrow SAFE$.
+Let us suppose that is completion-post-effective class of WSTS.
 The uniform resilience problem (URP), the uniform bounded resilience problem (UBRP) and the uniform $k$-resilience problem (UkRP) are decidable
@@
 \begin{proof}
-The uniform resilience problem (URP) reduces to decide coverability of the finite set $b(SAFE)$ from the infinite set $BAD$; but it is sufficient to decide coverability from the finite set of maximal elements of $BAD$ which are ideals but not elements in $S$.
-We use the fact that y is coverable from an $x \in BAD$ iff $y$ is coverable from $x' \in I(BAD)$.
+Let $I(BAD)=\{J_1, J_2,...,J_n\}$ representing $BAD$ and  $b(SAFE)=\{b_1,b_2,...,b_m\}$ representing $SAFE$.
+The uniform resilience problem (URP) reduces to decide infinite instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ st $b_j$ is coverable from $x$.
+Let us reduce this infinite coverability questions by $n$ instances of the coverability problem from ideals but not elements in $S$: for all $i$, does there exist an $j$ st $b_j$ is coverable from $J_i$
+
+We use the fact that y is coverable from an $x \in BAD$ iff $\downarrow y$ is coverable from $\downarrow x$ in $\hat{S}$
+%' \in I(BAD)$.
+\alain{à prouver}
 
@@
 \end{proof}
+
+URP is decidable for lossy counter machines (LCM) as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 

### 2023-02-03 definitions.tex
details

--- 2023-02-02 definitions.tex
@@
 %
-\problemx{uniform bounded resilience problem}
+\problemx{uniform bounded resilience problem (UBRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-resilient ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}
 
@@
 %		
+In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
 
@@
 %upward-closed or downward-closed sets.
+
+The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
+$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
+Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
+
+Let us recall two other results in \cite{BFM-ic17}.
+Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+
+\begin{proposition}\label{down-up}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{SAFE}$.
+The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
+The uniform $k$-resilience problem (UkRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
+\end{proposition}
+
+\begin{proof}
+Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{SAFE}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $SAFE$.
+The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+
+%Proposition 30 in \cite{BFM-icalp14} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{*} y'$ in $\mathscr{S}$.
+
+Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ iff $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+
+Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
+From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $b_j$ is covered from $\downarrow x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
+%	for every $y \in J$, 
+there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
+
+%
+% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
+%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
+%' \in I(BAD)$.
+
+To decide the (uniform) bounded resilience problem (UBRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is uniformely $K$-resilient.
+Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the uniform resilience problem (UBRP) and the uniform bounded resilience problem (UBRP) are decidable.
+
+To decide the UkRP, we may compare the input $k$ with $K$.
+If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
+if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $\pred^n(SAFE)=\pred^*(SAFE)$ and then to test whether $n \leq k$.
+S is uniformely $k$-resilient iff $n \leq k$ and then UkRP is decidable.
+\end{proof}
+
+URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 

### 2023-02-03 definitions.tex
details

@@
 
-\alain{dire juste que SAFE = upward closed et BAD = downward closed.
-envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
-4 resilience problem (U,D), (U,U), (D,U), (D,D)}
+%
+%\alain{dire juste que SAFE = upward closed et BAD = downward closed.
+%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
+%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
 
@@
 
+\alain{trouver les bad et safe maximum tels que S est resilient.
+est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
+on pourrait décider que safe=complement de bad et donc on aurait un seul ensemble (et son complement)}
+
 These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.
+Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.
+\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
@@
 %		
+URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
+%
 In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
+
+Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
+All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
+
+\begin{proposition}\label{general}
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient iff $bad \subseteq \pred^*(safe)$.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-$k$-resilient iff $bad \subseteq \pred^k(safe)$.
+\end{proposition}
+
+\begin{proof}
+Resilient means that 
+\end{proof}
+
+\subsection{Case: $SAFE=\uparrow SAFE$.}
+
+\begin{theorem}\label{up-down}
+The three uniform resilience problems are decidable for WSTS with effective predbasis and $safe=\uparrow safe$ and bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
+Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient.
+Since $safe=\uparrow safe$, and WSTS, $\pred^*(safe)=\uparrow \pred^*(safe)$, so $\pred^*(safe)$ admits a finite basis $B_{\pred^*(safe)}$.
+since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(safe)$ with the backward coverability algorithm.
+%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(safe)}$ which remais to compare, in both cases, two finite sets. 
+if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(safe)}$ such that $s \leq b$.
+similar reasonning for the other case.
+hence the two uniform resilience problems are decidable.
+
+Let us examine the uniform $k$-resilience problem (UkRP).
+we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
+S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
+\end{proof}
+
+%One begins to test whether $Bad \cap \pred^*(safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans safe.\\
+
+If $Bad \cap \pred^*(safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(safe)$.
+resilient iff $Bad \subseteq \pred^*(safe)$  (eq $Bad \cap \pred^*(safe) = Bad$)
+
+$Bad \cap \pred^*(safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(safe)$ et c'est (Bad',safe)-resilient.
+
+\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\uparrow SAFE$.}
+
+%
+%		cas étudié par les deux articles sur les graphes
+%
 
@@
 If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $\pred^n(SAFE)=\pred^*(SAFE)$ and then to test whether $n \leq k$.
-S is uniformely $k$-resilient iff $n \leq k$ and then UkRP is decidable.
+if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
+S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
 \end{proof}
 
-URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+%
+
+\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\uparrow SAFE$.}
+
+%%
+
+\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\downarrow SAFE$.}
+
+%%%%%%%%
+\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\downarrow SAFE$.}
+%
+%		cas exclusion mutuelle
+%
+je prends les minimaux de bad et les idéaux de safe, pas sur que ce soit decidable....$pred(safe)$
 

### 2023-02-03 definitions.tex
details

@@
 All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
+%
+\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
@@
 \begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $safe=\uparrow safe$ and bad is upward closed or downward closed.
+The three uniform resilience problems are decidable for WSTS with effective predbasis and $safe=\uparrow safe$.
+%	and bad is upward closed or downward closed.
 \end{theorem}
@@
 Let us examine the uniform $k$-resilience problem (UkRP).
-we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
+we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
 S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
@@
 
-\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\uparrow SAFE$.}
+Let us show that the predbasis hypothesis can be replaced by other hypotheses.
+
+\subsection{Case: $SAFE=\uparrow SAFE$ and $BAD=\downarrow BAD$.}
 
@@
 
-\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\uparrow SAFE$.}
+\subsection{Case: $SAFE=\uparrow SAFE$ and $BAD=\uparrow BAD$.}
 
@@
 
-\subsection{Case: $BAD=\downarrow BAD$ and $SAFE=\downarrow SAFE$.}
+\subsection{Case: $SAFE=\downarrow SAFE$ and $BAD=\downarrow BAD$.}
 
 %%%%%%%%
-\subsection{Case: $BAD=\uparrow BAD$ and $SAFE=\downarrow SAFE$.}
+\subsection{Case: $SAFE=\downarrow SAFE$ and $BAD=\uparrow BAD$.}
 %

### 2023-02-06 definitions.tex
Cleaned up some

@@
 
-\begin{remark} 
-$k$-resilienc $\leq$ bounded Resilience 
-%	$\leq$ Resilience.  
-\textcolor{red}{oui si strong monootony, non ? sinon}
-\end{remark}
-
 \begin{proposition}\label{reductions}
@@
 
-Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
+% Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
 
-and with effective basis of $post^*$ ???
+% and with effective basis of $post^*$ ???
 
-% \alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ?
-avec $U_t=\uparrow \pred(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
+% expliquer toutes les variantes de la résilience
 
-expliquer toutes les variantes de la résilience
-
-% \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}
-
-Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
+% Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
 
@@
 
-\begin{theorem}
-In WSTS with strong monotony, 
-the {\sc bounded resilience} problem
-is equivalent to the
-{\sc resilience} problem.
-\end{theorem}
+\begin{proposition}
+In WSTS with strong monotony, the bounded resilience problem is
+reducible to the $k$-resilience problem.
+\end{proposition}
+
+\begin{proof}
+strong WSTS
+$\pred^*(I) = \pred^{k}(I)$
+Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
+\end{proof}
+
+\begin{proposition}
+In WSTS with strong monotony, the bounded resilience problem is
+equivalent to the resilience problem.
+\end{proposition}
 

### 2023-02-06 definitions.tex
Reset PN liveness undecidable confirmed

@@
 Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.
+This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
+
+\begin{theorem}{Adjusting theorem from ...}
+The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
+\end{theorem}
+
+\begin{proof}
+If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
+reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
+reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
+$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
+
+The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
+and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
+
+We first modify all transitions of T_1 to include r_1 as both an input and an output.
+The initial marking $\mu_2$ will include a token in $r_1$.
+
+The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
+Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
+
+Transition $s$ is defined to have $r_1$ as an input and a null output.
+This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
+and "freezing" the marking of $P_1$.
+
+Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
+
+The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
+
+Now we need to see if that marking is zero.
+
+We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
+
+If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
+$s$ fires in the marking zero.
+
+If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
+
+In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
+We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
+This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
+Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
+
+Now if $0$ is reachable
+in $R(C_1, \mu_1)$ 
+then the net $C_2$ can also reach this marking
+in the place of $P_1$ by executing the same sequence of transition firings.
+Then $s$ can fire, freezing the $C_1$ subset.
+Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
+Thus if $0$ is reachable then $C_2$ is not live.
+
+Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
+$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
+(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
+% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
+% et r_2 commence sans token)
+If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
+Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
+
+\end{proof}
+
+\begin{corollary}
+Reset Petri Net liveness is undecidable 
+\end{corolalry}
+
+\begin{corollary}
+Reset Petri Net $t$-liveness is undecidable 
+\end{corolalry}
+
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, and since Reset Petri Net $t$-liveness is undecidable,  we deduce that resilience is undecidable for WSTS with strong monotony.
 
@@
 \end{theorem}
-
-\begin{proof}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, and since reachability $\leq$ t-liveness,  we deduce that resilience is undecidable for WSTS with strong monotony.
-
-\end{proof}
 

### 2023-02-06 definitions.tex
Reset PN liveness undecidable confirmed, correcting some typos
up

@@
 
-\begin{theorem}{Adjusting theorem from ...}
+\begin{theorem}[Adjusting theorem from ...]
 The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
@@
 
-We first modify all transitions of T_1 to include r_1 as both an input and an output.
+We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
 The initial marking $\mu_2$ will include a token in $r_1$.
@@
 Reset Petri Net liveness is undecidable 
-\end{corolalry}
+\end{corollary}
 
@@
 Reset Petri Net $t$-liveness is undecidable 
-\end{corolalry}
+\end{corollary}
 

### 2023-02-06 definitions.tex
up

@@
 
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions}, and since Reset Petri Net $t$-liveness is undecidable,  we deduce that resilience is undecidable for WSTS with strong monotony.
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that resilience is undecidable for WSTS with strong monotony.
 

### 2023-02-07 definitions.tex
cleaning up prop 23

@@
 %\"Ozkan
-\begin{proposition} \cite{DBLP:conf/gg/Ozkan22}
-goes on to prove that for finite-branching WSTS with strong monotonicity, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
+\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
+For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
 given a state $s$ of a WSTS
 % in the regarded class 
-with strong monotonicity 
-and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$ (Proposition~1 in \cite{DBLP:conf/gg/Ozkan22}). 
+with strong upward compatibility 
+and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$. 
 \end{proposition}
@@
 The characterization is used to show that Petri nets are $\post^*$-effective.
-It is well-known that Petri nets constitute WSTS with strong monotonicity (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
+It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
 Hence Resilience is decidable for Petri nets and VASS.

### 2023-02-07 definitions.tex
prop 24 changing single-place zero-reachability by zero-reachability which I am more definitely convinced of in terms of indecidability

@@
 Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS.
-Then, one would be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$.
-But reachability of $0$ in a counter $i$ is undecidable for Reset VASS.
+Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
+%in counter $i$ 
+by examining if there is %some vector $v$ in the basis such that $v(i)=0$
+$0$ in the basis%
+. But reachability of $0$  %in a counter $i$ 
+is undecidable for Reset VASS. 
 % sont WSTS with strong compatibility. et l'accessibilité est undecidable.
 %	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
-\textcolor{red}{à relire par Mathieu}
 \end{proof}

### 2023-02-07 definitions.tex
Up
Trying to write down thoughts

@@
 
+\mathieu{
+
+On considère un LCS
+
+On prends:
+
+$(q, baaaaaa)	\to \ldots $
+
+$(q, b) 	\to (q,bb)$
+
+$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
+
+MAIS
+
+C'est pas "en une étape".
+
+Dans ce cas là, du coup, $pred( \{(q,bb)\} ) $ est pas clos par le haut.
+
+MAIS
+
+$\{(q,bb)\}$ c'est pas clos par le haut non plus.
+
+$\uparrow \{(q,bb)\}$ l'est.
+
+Et $pred(\uparrow \{(q,bb)\}$ n'est pas clos par le haut.
+
+Revenons au Corollaire
+
+On voudrais avoir
+
+$post^* (s) \cap J \subseteq I^k$
+
+MAIS pas
+
+$ (\uparrow post^* (s)) \cap J \subseteq I^k$
+
+Ça serait un contre exemple de quoi ?
+
+Ça serait un cas où leurs algorithme vas dire 
+"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
+
+Alors qu'on aurait la $k$-résilience (i.e. $post^∗ (s) \cap J \subseteq I^k$)
+
+Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,bb)\})$ qui n'est pas clos par le haut.
+
+On voudrait
+
+$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,bb)\})$
+
+Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
+$\uparrow post^∗ (s)$ being too big
+donc autant pas se restreindre du tout sur $J$
+
+$post^*(s) \subseteq pred(\uparrow \{ (q,bb)\})$
+
+si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
+$b$ ? (et c'est tout, disons)
+
+Alors le problème, c'est que depuis $(q,b)$ on peut AUSSI perdre le $b$ (parce qu'on est lossy).
+
+Donc $(q,\epsilon)$ serait dans $post^*((q,b))$
+
+Mais pas dans $pred(\uparrow \{ (q,bb)\})$
+
+Et vu que dans $post^*((q,b)) \cap J$, $J$ est downward-closed alors on ne peut pas
+se servir de $J$ pour "enlever" $(q,\epsilon)$ 
+(sinon on enleve tout).
+
+Après, si je reprends mon exemple...
+
+Mais je change le $I$
+
+Disons, je prends $pred(\uparrow \{ (q,\epsilon)\})$
+
+Bon à ce moment là malheureusement on a transformé le set pour que il y ait tout les mots dedans quoi.
+
+Et si je prends $pred(\uparrow \{ (q,b)\})$ ?
+
+Il y a $(q,\epsilon)$ dedans parce que $(q,\epsilon) \to (q,b)$.
+
+Donc j'ai
+
+$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
+
+Ah mais à ce moment là ça deviens compliqué parce que on a rajouté tout les pred de $(q,baa \cdots a)$ aussi, alors qu'avant on avait que les pred de $(q,bbaa \cdots a)$ ...
+
+Mais c'est bon ... c'est bon parce que
+
+Parce que $(q,aaaaaa)$ est dans 
+$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$
+
+Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
+(même si il est dans $pred^*(\uparrow \{ (q,b)\})$)
+
+Attends ... Attends ça dépends si on autorise à perdre des lettres de tout les cotés.}
+
 \end{proof}
@@
 
-\begin{theorem}[Adjusting theorem from ...]
+\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
 The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.

### 2023-02-07 definitions.tex
Trying to write down thoughts

@@
 
-Alors qu'on aurait la $k$-résilience (i.e. $post^∗ (s) \cap J \subseteq I^k$)
+Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
 
@@
 Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
-$\uparrow post^∗ (s)$ being too big
+$\uparrow post^* (s)$ being too big
 donc autant pas se restreindre du tout sur $J$
@@
 
-Attends ... Attends ça dépends si on autorise à perdre des lettres de tout les cotés.}
+}
 

### 2023-02-07 definitions.tex
Trying to write down thoughts

--- 2023-02-07 definitions.tex
@@
 
-\mathieu{En train de chercher un contre exemple pour le Lemme~\ref{Lemma intersection}.}
+\mathieu{En train de chercher un contre exemple pour le corollaire du Lemme~\ref{Lemma intersection}.}
+
+\mathieu{
+
+On considère un LCS, où on autorise à effacer des messages (lossyness) mais effacer un message prends une étape (contrairement à [ref], où ils font une sorte de cloture de effacer-transition-effacer).
+On n'a alors pas la strong compatibility:
+
+On prends: $(q, b) 	\to (q,bb)$ par exemple et $(q, baaaaaa) \geq (q,b) $, alors
+$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
+MAIS
+
+C'est pas "en une étape".
+
+(On considère -il faut le mentionner/vérifier- que on ne peut effaçer que d'un coté ? et pas au milieu?)
+
+Dans ce cas là, du coup, $pred(\uparrow \{(q,b)\}$ n'est pas clos par le haut. \\
+
+Revenons au Corollaire
+
+On voudrais avoir
+
+$post^* (s) \cap J \subseteq I^k$
+
+MAIS pas
+
+$ (\uparrow post^* (s)) \cap J \subseteq I^k$ \\
+
+Ça serait un cas où leurs algorithme vas dire 
+"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
+
+Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
+
+Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,b)\})$ qui n'est pas clos par le haut.
+
+On voudrait
+
+$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,b)\})$
+
+Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
+$\uparrow post^* (s)$ being too big
+donc autant pas se restreindre du tout sur $J$
+
+$post^*(s) \subseteq pred(\uparrow \{ (q,b)\})$
+
+Si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
+$b$ ? et évidement en perdant le dernier message (parce qu'on est lossy). Et c'est tout, disons.
+
+Déjà par lossiness $(q,\epsilon)$ serait dans $post^*((q,b))$.
+
+Et aussi dans $pred(\uparrow \{ (q,b)\})$ puisque $(q,\epsilon) \to (q,b)$.
+
+Donc on aurait:
+
+$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
+
+Et on aurait $(q,aaaaaa)$ dans 
+$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$.
+
+Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
+(même si il est dans $pred^*(\uparrow \{ (q,b)\})$).
+
+}
 

### 2023-02-07 definitions.tex
Up

@@
 
-remarque si on enlève la strong monotony mais en gardant with effective basis of $post^*$:
+remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:
 
 \begin{proposition}(à prouver)
- leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $post^*$ ?
+ leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $\uparrow post^*$ ?
 \end{proposition}

### 2023-02-09 definitions.tex
details

@@
 %\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, BAD \subseteq S$ and $Safe \cap BAD = \emptyset$.}
+%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 %
@@
 \problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
+{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, Bad \subseteq S$ and $Safe \cap Bad = \emptyset$.}
 {$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
@@
 
-pour $k,k'$ donnés, chercher si $SAFE_{max}$ et $BAD_{max}$ ont un sens ?
-on peut faire l'union des SAFE, SAFE' et des BAD, BAD' prendre le max des k.
-Fixer k, chercher et calculer $SAFE_{max}$ et $BAD_{max}$.
+pour $k,k'$ donnés, chercher si $Safe_{max}$ et $Bad_{max}$ ont un sens ?
+on peut faire l'union des Safe, Safe' et des Bad, Bad' prendre le max des k.
+Fixer k, chercher et calculer $Safe_{max}$ et $Bad_{max}$.
 
 %
-%\alain{dire juste que SAFE = upward closed et BAD = downward closed.
+%\alain{dire juste que Safe = upward closed et BAD = downward closed.
 %envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
@@
 the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a safe state.
+a path of length smaller than $k$) that reaches a Safe state.
 Let us formalize these properties.
@@
 \problemx{uniform resilience problem (URP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
-{$BAD \longrightarrow^{*} SAFE$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
+{$Bad \longrightarrow^{*} Safe$ ?\newline}
 %
@@
 \problemx{uniform $k$-resilience problem (UkRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $SAFE, BAD \subseteq S$.}
-{$BAD \longrightarrow^{\leq k} SAFE$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
+{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}
 
@@
 \problemx{uniform bounded resilience problem (UBRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $SAFE, BAD \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
@@
 
-\alain{trouver les bad et safe maximum tels que S est resilient.
+\alain{trouver les Bad et Safe maximum tels que S est resilient.
 est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que safe=complement de bad et donc on aurait un seul ensemble (et son complement)}
+on pourrait décider que Safe=complement de Bad et donc on aurait un seul ensemble (et son complement)}
 
@@
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets SAFE and BAD.
-\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
+Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets Safe and Bad.
+\alain{on peut penser à des ensembles Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
 For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $SAFE =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $BAD =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of SAFE.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that BAD is always downward closed and SAFE is always upward closed.
+Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe.
+In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
 %		
-URP is decidable for lossy counter machines (LCM) with SAFE and BAD semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+URP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 \begin{proposition}\label{general}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient iff $bad \subseteq \pred^*(safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-$k$-resilient iff $bad \subseteq \pred^k(safe)$.
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
 \end{proposition}
@@
 
-\subsection{Case: $SAFE=\uparrow SAFE$.}
+\subsection{Case: $Safe=\uparrow Safe$.}
 
 \begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $safe=\uparrow safe$.
+The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
 %	and bad is upward closed or downward closed.
@@
 \begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,safe)-resilient.
-Since $safe=\uparrow safe$, and WSTS, $\pred^*(safe)=\uparrow \pred^*(safe)$, so $\pred^*(safe)$ admits a finite basis $B_{\pred^*(safe)}$.
-since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(safe)$ with the backward coverability algorithm.
-%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(safe)}$ which remais to compare, in both cases, two finite sets. 
-if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(safe)}$ such that $s \leq b$.
+Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.
+Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
+since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm.
+%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
+if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$.
 similar reasonning for the other case.
@@
 Let us examine the uniform $k$-resilience problem (UkRP).
-we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
+we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
 S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
@@
 
-%One begins to test whether $Bad \cap \pred^*(safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans safe.\\
+%One begins to test whether $Bad \cap \pred^*(Safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans Safe.\\
 
-If $Bad \cap \pred^*(safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(safe)$.
-resilient iff $Bad \subseteq \pred^*(safe)$  (eq $Bad \cap \pred^*(safe) = Bad$)
+If $Bad \cap \pred^*(Safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(Safe)$.
+resilient iff $Bad \subseteq \pred^*(Safe)$  (eq $Bad \cap \pred^*(Safe) = Bad$)
 
-$Bad \cap \pred^*(safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(safe)$ et c'est (Bad',safe)-resilient.
+$Bad \cap \pred^*(Safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(Safe)$ et c'est (Bad',Safe)-resilient.
 
@@
 
-\subsection{Case: $SAFE=\uparrow SAFE$ and $BAD=\downarrow BAD$.}
+\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\downarrow BAD$.}
 
@@
 %Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, SAFE and BAD, are given by 
+%it is therefore reasonable to demand that both propositions, Safe and BAD, are given by 
 %upward-closed or downward-closed sets.
@@
 \begin{proposition}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{SAFE}$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
 The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
@@
 \begin{proof}
-Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{SAFE}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $SAFE$.
+Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
 The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
@@
 If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(SAFE)=  \uparrow \pred^*(SAFE)$.
+if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
 S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
@@
 
-\subsection{Case: $SAFE=\uparrow SAFE$ and $BAD=\uparrow BAD$.}
+\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
 
@@
 
-\subsection{Case: $SAFE=\downarrow SAFE$ and $BAD=\downarrow BAD$.}
+\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
 
 %%%%%%%%
-\subsection{Case: $SAFE=\downarrow SAFE$ and $BAD=\uparrow BAD$.}
+\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\uparrow BAD$.}
 %
@@
 %
-je prends les minimaux de bad et les idéaux de safe, pas sur que ce soit decidable....$pred(safe)$
+je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$
 
 \begin{tabular}{ l c r }
-   BAD/SAFE & Up & Down \\
+   BAD/Safe & Up & Down \\
    Up & dec ? & dec ? \\
@@
 
+Uniform resilience is a strong property since it implies that from bad there must exist a path to Safe.
+But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap Bad$, there must exist a path to Safe.
+However, this condition seems more difficult since one also must decide whether a state $s \in Bad$ is reachable from $s_0$.
 %

### 2023-02-09 definitions.tex
details

@@
 
-\begin{definition}\cite{DBLP:journals/tcs/FinkelS01}
+\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
 A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
@@
 I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
-
+\iffalse
 \begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
@@
 \end{proposition}
-
-\iffalse
+%
 Proof.
@@
 \begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set of states, then $\pred(U )$ is upward-closed.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ and $\pred^*(U )$ are upward-closed.
 \end{proposition}

### 2023-02-09 definitions.tex
details

@@
 Then $t' \in I$ and $s' \in \pred(I )$.
-\fi
+
 \mathieu{On peut probablement enlever la Proposition $8$ alors.
 La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
+\fi
 
-\begin{definition}\cite{DBLP:journals/iandc/AbdullaCJT00}
+\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
 A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
@@
 
-\begin{claim}{(stability of ideals)}
+\begin{claim}{(stability of upward-closed sets)}
 Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
@@
 \begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is an upward-closed set and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
 The {\em index} $k(U)$ is the

### 2023-02-17 definitions.tex
Thinking Resilience (not bounded resilience but resilience itself) is decidable for WSTS with post* effectiveness. Have written a proof, needs to re-read. Will keep trying to see if Bounded is too or could be proven to be undecidable

@@
 
+\begin{theorem}
+{\sc Resilience} is decidable for (not necessarily strong) WSTS with effective $\uparrow$ $\post^*$ basis.
+\end{theorem}
+
+\begin{proof}
+Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_I$ a basis of $I$
+and $J$ decidable downward-closed.
+By applying Lemma~\ref{Lemma intersection} twice, we obtain
+
+$$ post^*(s) \cap J \subseteq \pred^*(I) \equiv B_p \cap J \subseteq \pred^*(I)$$
+
+Since $B_p$ is finite and $J$ is decidable, we can directly compute $ B_p \cap J$.
+% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
+We can compute a basis of $\pred^*(I)$ from $B_I$, and hence check that $B_p \cap J \setminus \pred^*(I) = \emptyset$. 
+\end{proof}
+
+% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
+% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
+% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
+%
+% Well, maybe their version of resilience but resilience with * seems fine.
+%
+
 Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.

### 2023-03-03 definitions.tex
details

@@
 
-\problemx{uniform resilience problem (URP)}
+\problemx{resilience problem (URP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
@@
 
-\problemx{uniform $k$-resilience problem (UkRP)}
+\problemx{$k$-resilience problem (UkRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
@@
 %
-\problemx{uniform bounded resilience problem (UBRP)}
+\problemx{bounded resilience problem (UBRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}

### 2023-03-06 reachable-resilience.tex
Split different sections about (resp. uniform) resilience into different files

--- 2023-03-03 definitions.tex
@@
-\section{Definitions}\label{section definitions}
-
-In this section, we introduce general notations and preliminary definitions.
-
-The model we are interested in is (S)WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
-
-% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
-
-\subsection{Transition systems}
-
-\begin{definition}
-A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Lambda, \rightarrow)$ where 
-$S$ is a set of {\em states}, $ \Lambda$ is a set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
-ternary relation,
-denoted as the set of {\em labeled transitions}. 
-\end{definition}
-
-We
- prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
-       $s  \xrightarrow{a}  s'$
-to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
-
-\noindent
-Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
-
-A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
-to a {\em target configuration} $s_n$ is a sequence 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
-We define the {\em concatenation} $ \pi_1 \pi_2$ of 
-two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
-as expected.
-The {\em length} of 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
-is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
-For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
-
-An {\em infinite path} is an infinite sequence
- $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
-For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
-(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
-and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
-by $\pi[i,j]$ the path 
-$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
- and by $\pi[i]$ the configuration $s_i$.
-As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
-a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
-Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
-$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
-
-\newcommand{\pred}{\textsf{pred}}
-\newcommand{\post}{\textsf{post}}
-% \renewcommand{\succ}{\textsf{Succ}}
-
-\noindent
-The set of {\em successors} of a configuration $s \in S$ is deﬁned as
-$\post(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
-A configuration without successors is called a {\em dead end}.  \\
-The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.
-
-A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
-$a \in \Lambda$,
- $ s_1  \xrightarrow{a} s_{2}$ and  
- $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
-
-\begin{definition}
-An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
-{\em states} and  
-$ {\rightarrow} \subseteq S \times S$ is a
-binary relation 
-on
-the set of states, denoted as the set of {\em transitions}. 
-\end{definition}
-
-We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
-Note 
- that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
-Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
-then defined as expected.
-
-%\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
-
-We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
-for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...
-
-Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
-We denote $X \longrightarrow^{*} Y$ (resp.
-$X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
-of length smaller than $k$) that reaches a state $y \in Y$.
-
-%\textcolor{red}{This makes sense for TS but not so much for LTS ...}
-
-A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
-%\textcolor{blue}{We restrict our attention to finitely branching TSs.}
-
-%\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
-%\textcolor{red}{Not sure that TS induced by TA are finitely branching.
-Actually I believe they are not, i.e.
-for instance for a TA with one clock $x$, from a state $q$ and clock $x$ set at $0$, if there is a transition e.g.
-$(q, x \geq 3, \emptyset, q')$ then the set of successors of
-%$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
-
-\subsection{Well-quasi-orderings}
-
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$.
-
-We abbreviate $x \leq y \not\leq x$ by $x < y$.
-
-% A partial ordering (a po) is an antisymmetric qo
-
-Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
-% and gives rise to a po between the equivalence classes ?
-% 	do we need po ?
-
-We now recall a few results from the theory of well-orderings (add reference [...]).
-
-\begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
-$x_i \leq  x_j$.
-\end{definition}
-
-Notice that a wqo is well-founded, i.e. it admits no infinite strictly decreasing sequence
-$x_0 > x_1 > x_2 > \cdots$
-
-% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
-
-\begin{lemma}
-(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
-\end{lemma}
-
-\iffalse
-\begin{proof}
-Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
-be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
-hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
-increasing subsequence.
-\end{proof}
-\fi
-
-Given $\leq$ a quasi-ordering over some set $X$, an {\em upward-closed set} is any set $I \subseteq X$ such that if $y \geq x$ and $x \in I$ then $y \in I $.
-A {\em downward-closed set} is any set $I \subseteq X$ such that if $y \leq x$ and $x \in I$ then $y \in I $. 
-To any $A \subseteq X$ we associate
-the {\em upward-closure of $A$} 
- $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
- and the 
- {\em downward-closure of $A$} 
- $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
-as $\uparrow x$ (resp. $\downarrow x$).
-
-A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \bigcup_{x \in I_b} \uparrow x$. 
-% Higman investigated ordered sets with the finite basis property.
-
-\begin{lemma}{(Higman [40])} 
-If $\leq$ is a wqo then any upward-closed $I$ has a
-finite basis.
-\end{lemma}
-
-% \textcolor{red}{Expliquer ce que c'est une base d'abords.}
-
-\begin{proof}
-The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
-only contains a finite number of non-equivalent elements otherwise they would make
-an infinite sequence contradicting the wqo assumption.
-\end{proof}
-
-\begin{lemma}  \label{upward-closed stablizes}
-If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
-upward-closed sets eventually stabilizes, i.e. there is a $k \in N$ such that 
-$I_k = I_{k+1} = I_{k+2} = \cdots $.
-\end{lemma}
-
-\begin{proof}
-Assue we have a counter-example.
-We extract an infinite subsequence where
-inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$.
-Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$.
-The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
-contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
-closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
-\end{proof}
-
-% \textcolor{red}{Define WSTS}
-
-% \textcolor{red}{Define SWSTS -  necessary}
-
-\subsection{Well-structured transition systems}
-
-\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
-A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
-is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
-%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
-$s_1, t_1 , s_2 \in S$
-	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$ , there exists 
-	$t_2 \in S$ with 
-	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$ 			
-				(resp. $s_2 \rightarrow t_2$ ).
-\end{definition}
-
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering.
-
-% \textcolor{red}{Define 'has effective pred-basis'.
-Maybe it should be included in WSTS definition, maybe it can be another def.
-I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
-\iffalse
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is an WSTS and $U \subseteq S$ is an upward-closed set of states, then $\pred^*(U )$ is upward-
-closed.
-\end{proposition}
-%
-Proof.
-Assume $s \in \pred^* (U )$.
-Then $s \rightarrow^* t$ for some $t \in U $.
-If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$.
-Then $t' \in U$ and $s' \in \pred^*(U )$.
-\fi
-
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ and $\pred^*(U )$ are upward-closed.
-\end{proposition}
-%\alain{étrange ces deux propositions, l'une suffirait}
-\iffalse
-Proof.
-Assume $s \in \pred (I )$.
-Then $s \rightarrow t$ for some $t \in I $.
-If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
-Then $t' \in I$ and $s' \in \pred(I )$.
-
-\mathieu{On peut probablement enlever la Proposition $8$ alors.
-La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
-\fi
-
-\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-\end{definition}
-
-% \textcolor{red}{Define what an Ideal is.
-Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
-Anti-ideal just downward closed so again just not that helpful a notation.
-Maybe have a }
-
-\iffalse
-\begin{definition}
-A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
-$\uparrow I = I = \downarrow I$.
-\end{definition}
-
-"Bi-ideals often represent “control states” as in [cf \%]. "
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-% \textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
-\fi
-
-A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in J$. 
-%\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
-% \textcolor{red}{Do we still demand this ?}
-%\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
-
-\begin{claim}{(stability of upward-closed sets)}
-Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
-\end{claim}
-
-% \begin{claim}
-% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \end{claim}
-
-\begin{fact}\label{fact basis}
-% Fact 3 ([1]). 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$. 
-% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
-\end{fact}
-
-\begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
-The {\em index} $k(U)$ is the
-smallest $k_0$ s.t. $U^k = U^{k_0}$ for all $k \geq k_0$.
-\end{definition}
-
-If $\mathscr{S}$ is a WSTS with strong compatibility and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
-every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
-
-\begin{remark}
-$I^{k+1} $ can be rewritten $I^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
-\end{remark}
-
-This ensures the following.
-
-\begin{fact}\label{stop condition}
-% Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
-$k(I) \leq k$. This also implies that $\pred^*(I) = I^k$.
-\end{fact}
-
-\iffalse
-\begin{remark}
-$I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
-\end{remark}
-\fi
-
-\begin{lemma}
-% Lemma 3 ([1]) 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of an effective strongly well-structured transition
-system, we can decide whether we can reach $I$ from $s$.
-\end{lemma}
-
-\begin{proof}
-We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
-Then the
-decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
-We have
-$$I^{k+1} = I \cup \pred(I^k ) = I \cup
-\bigcup_{s' \in B}
-\pred(\uparrow \{s' \}).$$
-
-Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
-Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
-\end{proof}
-
-\begin{theorem}
-A finite basis of $ \pred^*(U)$ is computable for any WSTS $\mathscr{S}$ and any upward closed set $U \subseteq S$.
-\end{theorem}
-
-%%%
-\iffalse
-%%%%
-%%%%
-\subsection{Defining resilience}
-
-\subsubsection{Resilience for general transition systems}
-
-% \textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$SAFE$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$BAD$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
-
-\problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).
-
-We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
-
-%
-%\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, BAD \subseteq S$ and $Safe \cap BAD = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, Bad \subseteq S$ and $Safe \cap Bad = \emptyset$.}
-{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
-%%
-\fi
-%%%
-
-pour $k,k'$ donnés, chercher si $Safe_{max}$ et $Bad_{max}$ ont un sens ?
-on peut faire l'union des Safe, Safe' et des Bad, Bad' prendre le max des k.
-Fixer k, chercher et calculer $Safe_{max}$ et $Bad_{max}$.
-
-%
-%\alain{dire juste que Safe = upward closed et BAD = downward closed.
-%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
-%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
-
-\section{Uniform resilience for WSTS}
-
-The resilience problem (resp.
-the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a Safe state.
-Let us formalize these properties.
-
-\problemx{resilience problem (URP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{*} Safe$ ?\newline}
-%
-%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
-
-\problemx{$k$-resilience problem (UkRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}
-
-We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
-\problemx{bounded resilience problem (UBRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}
-
-\alain{trouver les Bad et Safe maximum tels que S est resilient.
-est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que Safe=complement de Bad et donc on aurait un seul ensemble (et son complement)}
-
-These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
-So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets Safe and Bad.
-\alain{on peut penser à des ensembles Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
-%		
-URP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
-%
-In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
-
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
-%
-\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-\begin{proposition}\label{general}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
-\end{proposition}
-
-\begin{proof}
-Resilient means that 
-\end{proof}
-
-\subsection{Case: $Safe=\uparrow Safe$.}
-
-\begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
-%	and bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.
-Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
-since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm.
-%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
-if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$.
-similar reasonning for the other case.
-hence the two uniform resilience problems are decidable.
-
-Let us examine the uniform $k$-resilience problem (UkRP).
-we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
-\end{proof}
-
-%One begins to test whether $Bad \cap \pred^*(Safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans Safe.\\
-
-If $Bad \cap \pred^*(Safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(Safe)$.
-resilient iff $Bad \subseteq \pred^*(Safe)$  (eq $Bad \cap \pred^*(Safe) = Bad$)
-
-$Bad \cap \pred^*(Safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(Safe)$ et c'est (Bad',Safe)-resilient.
-
-Let us show that the predbasis hypothesis can be replaced by other hypotheses.
-
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\downarrow BAD$.}
-
-%
-%		cas étudié par les deux articles sur les graphes
-%
-
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-
-%Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, Safe and BAD, are given by 
-%upward-closed or downward-closed sets.
-
-The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
-
-Let us recall two other results in \cite{BFM-ic17}.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-
-\begin{proposition}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
-The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
-The uniform $k$-resilience problem (UkRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
-\end{proposition}
-
-\begin{proof}
-Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
-The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-%Proposition 30 in \cite{BFM-icalp14} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{*} y'$ in $\mathscr{S}$.
-
-Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ iff $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-
-Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $b_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-
-%
-% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
-%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
-%' \in I(BAD)$.
-
-To decide the (uniform) bounded resilience problem (UBRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is uniformely $K$-resilient.
-Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the uniform resilience problem (UBRP) and the uniform bounded resilience problem (UBRP) are decidable.
-
-To decide the UkRP, we may compare the input $k$ with $K$.
-If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
-\end{proof}
-
-%
-
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
-
-%%
-
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
-
-%%%%%%%%
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\uparrow BAD$.}
-%
-%		cas exclusion mutuelle
-%
-je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$
-
-\begin{tabular}{ l c r }
-   BAD/Safe & Up & Down \\
-   Up & dec ? & dec ? \\
-   Down & dec & ? \\
- \end{tabular}
-
-%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-% \textcolor{red}{Seems like a reasonable assumption to me.}
-
-%From these considerations, we formulate instances of the abstract resilience problems for well-
-%structured transition systems.
-
 \section{Resilience for WSTS}

### 2023-03-06 resilience.tex
Split different sections about (resp. uniform) resilience into different files

--- 2023-02-09 definitions.tex
@@
-\section{Definitions}\label{section definitions}
-
-In this section, we introduce general notations and preliminary definitions.
-
-The model we are interested in is (S)WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
-
-% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
-
-\subsection{Transition systems}
-
-\begin{definition}
-A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Lambda, \rightarrow)$ where 
-$S$ is a set of {\em states}, $ \Lambda$ is a set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
-ternary relation,
-denoted as the set of {\em labeled transitions}. 
-\end{definition}
-
-We
- prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
-       $s  \xrightarrow{a}  s'$
-to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
-
-\noindent
-Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
-
-A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
-to a {\em target configuration} $s_n$ is a sequence 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
-We define the {\em concatenation} $ \pi_1 \pi_2$ of 
-two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
-as expected.
-The {\em length} of 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
-is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
-For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
-
-An {\em infinite path} is an infinite sequence
- $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
-For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
-(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
-and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
-by $\pi[i,j]$ the path 
-$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
- and by $\pi[i]$ the configuration $s_i$.
-As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
-a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
-Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
-$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
-
-\newcommand{\pred}{\textsf{pred}}
-\newcommand{\post}{\textsf{post}}
-% \renewcommand{\succ}{\textsf{Succ}}
-
-\noindent
-The set of {\em successors} of a configuration $s \in S$ is deﬁned as
-$\post(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
-A configuration without successors is called a {\em dead end}.  \\
-The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.
-
-A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
-$a \in \Lambda$,
- $ s_1  \xrightarrow{a} s_{2}$ and  
- $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
-
-\begin{definition}
-An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
-{\em states} and  
-$ {\rightarrow} \subseteq S \times S$ is a
-binary relation 
-on
-the set of states, denoted as the set of {\em transitions}. 
-\end{definition}
-
-We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
-Note 
- that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
-Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
-then defined as expected.
-
-%\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
-
-We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
-for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...
-
-Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
-We denote $X \longrightarrow^{*} Y$ (resp.
-$X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
-of length smaller than $k$) that reaches a state $y \in Y$.
-
-%\textcolor{red}{This makes sense for TS but not so much for LTS ...}
-
-A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
-%\textcolor{blue}{We restrict our attention to finitely branching TSs.}
-
-%\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
-%\textcolor{red}{Not sure that TS induced by TA are finitely branching.
-Actually I believe they are not, i.e.
-for instance for a TA with one clock $x$, from a state $q$ and clock $x$ set at $0$, if there is a transition e.g.
-$(q, x \geq 3, \emptyset, q')$ then the set of successors of
-%$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
-
-\subsection{Well-quasi-orderings}
-
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$.
-
-We abbreviate $x \leq y \not\leq x$ by $x < y$.
-
-% A partial ordering (a po) is an antisymmetric qo
-
-Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
-% and gives rise to a po between the equivalence classes ?
-% 	do we need po ?
-
-We now recall a few results from the theory of well-orderings (add reference [...]).
-
-\begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
-$x_i \leq  x_j$.
-\end{definition}
-
-Notice that a wqo is well-founded, i.e. it admits no infinite strictly decreasing sequence
-$x_0 > x_1 > x_2 > \cdots$
-
-% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
-
-\begin{lemma}
-(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
-\end{lemma}
-
-\iffalse
-\begin{proof}
-Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
-be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
-hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
-increasing subsequence.
-\end{proof}
-\fi
-
-Given $\leq$ a quasi-ordering over some set $X$, an {\em upward-closed set} is any set $I \subseteq X$ such that if $y \geq x$ and $x \in I$ then $y \in I $.
-A {\em downward-closed set} is any set $I \subseteq X$ such that if $y \leq x$ and $x \in I$ then $y \in I $. 
-To any $A \subseteq X$ we associate
-the {\em upward-closure of $A$} 
- $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
- and the 
- {\em downward-closure of $A$} 
- $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
-as $\uparrow x$ (resp. $\downarrow x$).
-
-A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \bigcup_{x \in I_b} \uparrow x$. 
-% Higman investigated ordered sets with the finite basis property.
-
-\begin{lemma}{(Higman [40])} 
-If $\leq$ is a wqo then any upward-closed $I$ has a
-finite basis.
-\end{lemma}
-
-% \textcolor{red}{Expliquer ce que c'est une base d'abords.}
-
-\begin{proof}
-The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
-only contains a finite number of non-equivalent elements otherwise they would make
-an infinite sequence contradicting the wqo assumption.
-\end{proof}
-
-\begin{lemma}  \label{upward-closed stablizes}
-If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
-upward-closed sets eventually stabilizes, i.e. there is a $k \in N$ such that 
-$I_k = I_{k+1} = I_{k+2} = \cdots $.
-\end{lemma}
-
-\begin{proof}
-Assue we have a counter-example.
-We extract an infinite subsequence where
-inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$.
-Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$.
-The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
-contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
-closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
-\end{proof}
-
-% \textcolor{red}{Define WSTS}
-
-% \textcolor{red}{Define SWSTS -  necessary}
-
-\subsection{Well-structured transition systems}
-
-\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
-A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
-is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
-%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
-$s_1, t_1 , s_2 \in S$
-	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$ , there exists 
-	$t_2 \in S$ with 
-	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$ 			
-				(resp. $s_2 \rightarrow t_2$ ).
-\end{definition}
-
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering.
-
-% \textcolor{red}{Define 'has effective pred-basis'.
-Maybe it should be included in WSTS definition, maybe it can be another def.
-I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
-\iffalse
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is an WSTS and $U \subseteq S$ is an upward-closed set of states, then $\pred^*(U )$ is upward-
-closed.
-\end{proposition}
-%
-Proof.
-Assume $s \in \pred^* (U )$.
-Then $s \rightarrow^* t$ for some $t \in U $.
-If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$.
-Then $t' \in U$ and $s' \in \pred^*(U )$.
-\fi
-
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ and $\pred^*(U )$ are upward-closed.
-\end{proposition}
-%\alain{étrange ces deux propositions, l'une suffirait}
-\iffalse
-Proof.
-Assume $s \in \pred (I )$.
-Then $s \rightarrow t$ for some $t \in I $.
-If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
-Then $t' \in I$ and $s' \in \pred(I )$.
-
-\mathieu{On peut probablement enlever la Proposition $8$ alors.
-La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
-\fi
-
-\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-\end{definition}
-
-% \textcolor{red}{Define what an Ideal is.
-Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
-Anti-ideal just downward closed so again just not that helpful a notation.
-Maybe have a }
-
-\iffalse
-\begin{definition}
-A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
-$\uparrow I = I = \downarrow I$.
-\end{definition}
-
-"Bi-ideals often represent “control states” as in [cf \%]. "
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-% \textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
-\fi
-
-A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in J$. 
-%\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
-% \textcolor{red}{Do we still demand this ?}
-%\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
-
-\begin{claim}{(stability of upward-closed sets)}
-Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
-\end{claim}
-
-% \begin{claim}
-% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \end{claim}
-
-\begin{fact}\label{fact basis}
-% Fact 3 ([1]). 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$. 
-% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
-\end{fact}
-
-\begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
-The {\em index} $k(U)$ is the
-smallest $k_0$ s.t. $U^k = U^{k_0}$ for all $k \geq k_0$.
-\end{definition}
-
-If $\mathscr{S}$ is a WSTS with strong compatibility and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
-every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
-
-\begin{remark}
-$I^{k+1} $ can be rewritten $I^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
-\end{remark}
-
-This ensures the following.
-
-\begin{fact}\label{stop condition}
-% Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
-$k(I) \leq k$. This also implies that $\pred^*(I) = I^k$.
-\end{fact}
-
-\iffalse
-\begin{remark}
-$I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
-\end{remark}
-\fi
-
-\begin{lemma}
-% Lemma 3 ([1]) 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of an effective strongly well-structured transition
-system, we can decide whether we can reach $I$ from $s$.
-\end{lemma}
-
-\begin{proof}
-We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
-Then the
-decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
-We have
-$$I^{k+1} = I \cup \pred(I^k ) = I \cup
-\bigcup_{s' \in B}
-\pred(\uparrow \{s' \}).$$
-
-Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
-Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
-\end{proof}
-
-\begin{theorem}
-A finite basis of $ \pred^*(U)$ is computable for any WSTS $\mathscr{S}$ and any upward closed set $U \subseteq S$.
-\end{theorem}
-
 %%%
@@
 %%
-\fi
+
 %%%
@@
 
+\fi
+
 \section{Uniform resilience for WSTS}
@@
 %structured transition systems.
-
-\section{Resilience for WSTS}
-
-Uniform resilience is a strong property since it implies that from bad there must exist a path to Safe.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap Bad$, there must exist a path to Safe.
-However, this condition seems more difficult since one also must decide whether a state $s \in Bad$ is reachable from $s_0$.
-%
-%\problemx{(I,J)-$k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
-%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-%\problemx{general $k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
-%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
-
-\problemx{$k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$D \longrightarrow^{\leq k} U$ ?  ou $D \cap post^*(s) \longrightarrow^{\leq k} U$ ?  \newline}
-%
-%  		{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-$D \longrightarrow^{\leq k} U$: $\forall s' \in D \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\\
-%
-$D \cap post^*(s) \longrightarrow^{\leq k} U$ ?
-%
-$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?
- \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
-c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ?
-sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}
-
-%ou plus simple : $D \rightarrow^{\leq k} U$ ?
-
-%Analogously, we formulate the resilience problem for WSTSs.
-
-\problemx{bounded Resilience problem for WSTS}
-{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}
-
-\problemx{Resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$ ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{*} s''$ ?\newline}
-
-%\textcolor{blue}{
-%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
-%}
-
-\begin{remark} 
-We require $U$ to be upward-closed, as, in the general case if we lift restrictions on $U$ and $D$, then reachability reduces itself to resilience.
-Indeed, the particular resilience problem where we take $D = \{s\}$ and $U = \{x\}$ is reachability of $x$ from $s$, and thus generally undecidable for WSTS.
-\end{remark}
-
-\textcolor{blue}{
-et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
-}
-
-\textcolor{blue}{
-compute the sets of $(I,J,k)$ st S is resilient.\\
-max des I,J possibles \\
-}
-
-\begin{proposition}\label{reductions}
-coverability and $t$-liveness are reduceable to resilience.
-%	home state and reachability
-\end{proposition}
-
-\begin{proof}
-
-\begin{itemize}
-\item Let us choose $J=S= \downarrow S$ and $I=\uparrow x$ then resilience is equivalent to coverability of $x$ from $s$.
-%\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the resilience problem
-where $J=S$ and $I = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
-
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
-%  \textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
-Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
-This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-% Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
-\end{proof}
-
-% Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
-
-% and with effective basis of $post^*$ ???
-
-% expliquer toutes les variantes de la résilience
-
-% Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
-
-\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
-\end{theorem}
-
-We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
-
-\begin{corollary}\label{postcomputable}
-{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
-\end{corollary}
-
-\begin{proof}
-From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
-$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
-returning $k_m$ if that is the case.
-Then, because $k_m$-resilience is decidable, 
-checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
-and resilience is decidable.
-% \textcolor{red}{à faire Mathieu}
-\end{proof}
-
-%\"Ozkan
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
-For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
-given a state $s$ of a WSTS
-% in the regarded class 
-with strong upward compatibility 
-and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$. 
-\end{proposition}
-
-The characterization is used to show that Petri nets are $\post^*$-effective.
-It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Hence Resilience is decidable for Petri nets and VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
-
-\begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}
-Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
-%in counter $i$ 
-by examining if there is %some vector $v$ in the basis such that $v(i)=0$
-$0$ in the basis%
-. But reachability of $0$  %in a counter $i$ 
-is undecidable for Reset VASS. 
-% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
-\end{proof}
-
-%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
-
-remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:
-
-\begin{proposition}(à prouver)
- leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $\uparrow post^*$ ?
-\end{proposition}
-
-\begin{proof}
-leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
- \textcolor{red}{à faire Mathieu}
-
-\mathieu{En train de chercher un contre exemple pour le corollaire du Lemme~\ref{Lemma intersection}.}
-
-\mathieu{
-
-On considère un LCS, où on autorise à effacer des messages (lossyness) mais effacer un message prends une étape (contrairement à [ref], où ils font une sorte de cloture de effacer-transition-effacer).
-On n'a alors pas la strong compatibility:
-
-On prends: $(q, b) 	\to (q,bb)$ par exemple et $(q, baaaaaa) \geq (q,b) $, alors
-$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
-MAIS
-
-C'est pas "en une étape".
-
-(On considère -il faut le mentionner/vérifier- que on ne peut effaçer que d'un coté ? et pas au milieu?)
-
-Dans ce cas là, du coup, $pred(\uparrow \{(q,b)\}$ n'est pas clos par le haut. \\
-
-Revenons au Corollaire
-
-On voudrais avoir
-
-$post^* (s) \cap J \subseteq I^k$
-
-MAIS pas
-
-$ (\uparrow post^* (s)) \cap J \subseteq I^k$ \\
-
-Ça serait un cas où leurs algorithme vas dire 
-"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
-
-Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
-
-Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,b)\})$ qui n'est pas clos par le haut.
-
-On voudrait
-
-$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,b)\})$
-
-Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
-$\uparrow post^* (s)$ being too big
-donc autant pas se restreindre du tout sur $J$
-
-$post^*(s) \subseteq pred(\uparrow \{ (q,b)\})$
-
-Si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
-$b$ ? et évidement en perdant le dernier message (parce qu'on est lossy). Et c'est tout, disons.
-
-Déjà par lossiness $(q,\epsilon)$ serait dans $post^*((q,b))$.
-
-Et aussi dans $pred(\uparrow \{ (q,b)\})$ puisque $(q,\epsilon) \to (q,b)$.
-
-Donc on aurait:
-
-$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
-
-Et on aurait $(q,aaaaaa)$ dans 
-$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$.
-
-Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
-(même si il est dans $pred^*(\uparrow \{ (q,b)\})$).
-
-}
-
-\end{proof}
-
-\begin{proposition}(à prouver)
-{\sc bounded resilience} is undecidable for WSTS with upward compatibility withd an effective basis of $post^*$.
-\end{proposition}
-
-\begin{proof}
- \textcolor{red}{pas certain ! à faire Mathieu}
-\end{proof}
-
-%
-%si on garde strong mais on enlève effective basis of $post^*$:
-%
-%\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
-%\end{proposition}
-%
-%\begin{proof}
-%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
-%\end{proof}
-
-Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.
-This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
-
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem}
-
-\begin{proof}
-If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
-reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
-reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
-$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
-
-The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
-and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
-
-We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $r_1$.
-
-The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
-
-Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
-and "freezing" the marking of $P_1$.
-
-Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
-The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
-
-Now we need to see if that marking is zero.
-
-We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
-
-If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
-$s$ fires in the marking zero.
-
-If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
-
-In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
-Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-
-Now if $0$ is reachable
-in $R(C_1, \mu_1)$ 
-then the net $C_2$ can also reach this marking
-in the place of $P_1$ by executing the same sequence of transition firings.
-Then $s$ can fire, freezing the $C_1$ subset.
-Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $0$ is reachable then $C_2$ is not live.
-
-Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
-$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
-(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
-% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
-% et r_2 commence sans token)
-If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
-
-\end{proof}
-
-\begin{corollary}
-Reset Petri Net liveness is undecidable 
-\end{corollary}
-
-\begin{corollary}
-Reset Petri Net $t$-liveness is undecidable 
-\end{corollary}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that resilience is undecidable for WSTS with strong monotony.
-
-\begin{theorem}
-{\sc resilience} is undecidable for WSTS with strong monotony.
-\end{theorem}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-reducible to the $k$-resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-$\pred^*(I) = \pred^{k}(I)$
-Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-
-$\pred^*(I) = \pred^{k}(I)$
-
-\end{proof}
-
-\begin{corollary}
-Bounded Resilience is undecidable for WSTS with strong monotony.
-\end{corollary}
-
-\begin{proof} 
-  From
-  cf
-the two  theorems above. 
-\end{proof}
-
-Let us deduce that effective basis of $post^*$ ex reset PN ?
-
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded resilience is decidable for VASS, hence resilience is decidable for VASS.
-\alain{resilience decidable for LCS ?
-LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-We may consider context-free grammars.
-
-\begin{corollary}
-{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
-\end{corollary}
-
-\begin{proof}
-
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
-\end{proof}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}

### 2023-03-09 resilience.tex
Changed name of Uniform resilience to remove 'Uniform'

@@
 
-\section{Uniform resilience for WSTS}
+\section{Resilience for WSTS}
 
@@
 
-\problemx{resilience problem (URP)}
+\problemx{resilience problem (RP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
@@
 
-\problemx{$k$-resilience problem (UkRP)}
+\problemx{$k$-resilience problem (kRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
@@
 %
-\problemx{bounded resilience problem (UBRP)}
+\problemx{bounded resilience problem (BRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
+ $k$-resilient ?\newline}
 
@@
 %		
-URP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+RP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 \begin{proposition}\label{general}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(bad,Safe)-resilient.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
 \end{proposition}
@@
 \begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
+The three %uniform 
+resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
 %	and bad is upward closed or downward closed.
@@
 \begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.
+Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(bad,Safe)-resilient.
 Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
@@
 similar reasonning for the other case.
-hence the two uniform resilience problems are decidable.
+hence the two %uniform
+ resilience problems are decidable.
 
-Let us examine the uniform $k$-resilience problem (UkRP).
-we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
+Let us examine the %uniform 
+$k$-resilience problem (kRP). we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
+$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
 \end{proof}
@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
-The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
-The uniform $k$-resilience problem (UkRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
+The %uniform 
+resilience problem (RP) and the %uniform 
+bounded resilience problem (BRP) are decidable.
+The %uniform
+ $k$-resilience problem (kRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
 \end{proposition}
@@
 Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
-The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+The %uniform 
+resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
 This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
@@
 
-To decide the (uniform) bounded resilience problem (UBRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is uniformely $K$-resilient.
-Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the uniform resilience problem (UBRP) and the uniform bounded resilience problem (UBRP) are decidable.
+To decide the %(uniform)
+ bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
+  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
+  resilience problem (BRP) and the %uniform 
+  bounded resilience problem (BRP) are decidable.
 
-To decide the UkRP, we may compare the input $k$ with $K$.
-If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
+To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
+$k$-resilient; if $k < K$ we cannot answer.
 if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
+$S$ is %uniformely
+ $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
 \end{proof}

### 2023-03-09 resilience.tex
Up

@@
 
-\subsection{Case: $Safe=\uparrow Safe$.}
+\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
 
-\begin{theorem}\label{up-down}
+\begin{theorem}\label{up-up}
 The three %uniform 
-resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
+resilience problems are decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
+and $BAD=\uparrow BAD$.
 %	and bad is upward closed or downward closed.
@@
 if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$.
+\iffalse
 similar reasonning for the other case.
+\fi
 hence the two %uniform
@@
 Let us examine the %uniform 
-$k$-resilience problem (kRP). we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
-$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
+$k$-resilience problem (kRP). We may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
+$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
+\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$, mais il faut traiter aussi les $k$ plus petits.}
+\mathieu{Si on a $\uparrow \pred^k(Safe) = \pred^k(Safe)$ on résoud comme ci-dessus.}
+\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(Safe) \neq \pred^k(Safe)$.}
+\mathieu{... et alors, c'est plus compliqué.}
+\mathieu{L'important c'est d'avoir le $n$ minimal tel que $bad \subseteq \pred^n(Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.}and then kRP is decidable.
 \end{proof}
@@
 
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
+\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
 
-%%
-
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
+If $Safe = \downarrow 0$ and $Bad = S$ . . .
 

### 2023-03-13 resilience.tex
Up
up

@@
 
+% $Bad$ still necessarily contains $0$, and $0 \rightarrow^* 0$.
+
+$Bad = J_1 \cup J_2 \cup \ldots \cup J_m$
+
+$Safe = I_1 \cup I_2 \cup \ldots \cup I_n$
+
+Question is: can we reach one of the $I_i$ from one of the $J_j$ ?
+
+If the ideal reachability problem is decidable then of course the question here should be decidable as well.
+
+But it may be possible to do more i.e.
+
+If we can express $\downarrow Post( J_j)$ then the question become:
+is $Safe \cap \cup_{j} \downarrow Post( J_j) \neq \emptyset$ ?
+
+(we need to reach an element of $Safe$ but not every possible element)
+
 %%%%%%%%
@@
 je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$
+
+If $Safe = \downarrow 0$ and $Bad = \uparrow s$ . . .
+
+Then resilience kinds of becomes "can we reach $0$ from $\uparrow s$?".
+
+In a reset petri-net, if $Bad = \uparrow q_0(\mu)$ for some $\mu \neq 0$,
+and $Safe = \downarrow q(0)$,
+then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.
+
+\mathieu{Ici}
 

### 2023-03-13 resilience.tex
up

@@
 
-If we can express $\downarrow Post( J_j)$ then the question become:
-is $Safe \cap \cup_{j} \downarrow Post( J_j) \neq \emptyset$ ?
+If we can express $\downarrow Post^*( J_j)$ then the question become:
+is $Safe \cap \cup_{j} \downarrow Post^*( J_j) \neq \emptyset$ ?
 
 (we need to reach an element of $Safe$ but not every possible element)
+
+Rather, the question is still
+"is $Safe \cap \cup_{j}  Post^*( J_j) \neq \emptyset$ ?"
+
+But maybe we can use $\downarrow Post^*( J_j)$ around here.
 

### 2023-03-14 resilience.tex
Separating theorem in three

@@
 \begin{theorem}\label{up-up}
-The three %uniform 
-resilience problems are decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
+The % three %uniform 
+resilience problems is decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
 and $BAD=\uparrow BAD$.
@@
 \fi
-hence the two %uniform
- resilience problems are decidable.
+hence the %two
+ %uniform
+ resilience problems is decidable.
+\end{proof}
 
+\begin{corollary}
+The bounded resilience problems is decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
+and $BAD=\uparrow BAD$.
+\end{corollary}
+
+\begin{theorem}\label{up-up}
+The % three %uniform 
+$k$-resilience problems is decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
+and $BAD=\uparrow BAD$.
+%	and bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
 Let us examine the %uniform 

### 2023-03-14 resilience.tex
up

@@
 
-In a reset petri-net, if $Bad = \uparrow q_0(\mu)$ for some $\mu \neq 0$,
-and $Safe = \downarrow q(0)$,
+In a reset petri-net, if $Bad = \uparrow \mu$ for some $\mu \neq 0$,
+and $Safe = \downarrow 0$,
 then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.

### 2023-03-23 resilience.tex
small up

--- 2023-03-13 resilience.tex
@@
 
-% \textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
+% \textcolor{red}{Ask the question why use a set of propositions for $\Safe$ and $\Bad$ rather than use subsets of the set of configurations ? }
 
@@
 %
-$SAFE$  in a reasonable amount of time whenever we reach a state 
+$\Safe$  in a reasonable amount of time whenever we reach a state 
 % which satisfies
@@
 %
-$BAD$. 
+$\Bad$. 
 From this we formulate two resilience problems. First consider the case where the recovery time
@@
 \problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
+{$\forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 
@@
 
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).
+% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(\Bad → 0≤ j≤k EX j \Safe).
 
@@
 %\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, BAD \subseteq S$ and $Safe \cap BAD = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+%{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
+%{$\exists k \geq 0 ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 %
@@
 \problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, Bad \subseteq S$ and $Safe \cap Bad = \emptyset$.}
+{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
 {$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
@@
 
-pour $k,k'$ donnés, chercher si $Safe_{max}$ et $Bad_{max}$ ont un sens ?
-on peut faire l'union des Safe, Safe' et des Bad, Bad' prendre le max des k.
-Fixer k, chercher et calculer $Safe_{max}$ et $Bad_{max}$.
+pour $k,k'$ donnés, chercher si $\Safe_{max}$ et $\Bad_{max}$ ont un sens ?
+on peut faire l'union des \Safe, \Safe' et des \Bad, \Bad' prendre le max des k.
+Fixer k, chercher et calculer $\Safe_{max}$ et $\Bad_{max}$.
 
 %
-%\alain{dire juste que Safe = upward closed et BAD = downward closed.
-%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
+%\alain{dire juste que \Safe = upward closed et \Bad = downward closed.
+%envisager les 3 autres cas. souvent \Bad = upward closed (par ex l'exclusion mutuelle).
 %4 resilience problem (U,D), (U,U), (D,U), (D,D)}
@@
 \fi
+
+\newcommand{\Bad}{\textsf{Bad}}
+\newcommand{\Safe}{\textsf{Safe}}
 
@@
 the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a Safe state.
+a path of length smaller than $k$) that reaches a \Safe state.
 Let us formalize these properties.
@@
 \problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{*} Safe$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
 %
@@
 \problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
+{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
 
@@
 \problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
@@
 
-\alain{trouver les Bad et Safe maximum tels que S est resilient.
+\alain{trouver les \Bad et \Safe maximum tels que S est resilient.
 est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que Safe=complement de Bad et donc on aurait un seul ensemble (et son complement)}
+on pourrait décider que \Safe=complement de \Bad et donc on aurait un seul ensemble (et son complement)}
 
@@
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets Safe and Bad.
-\alain{on peut penser à des ensembles Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
+Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets \Safe and \Bad.
+\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
 For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
+Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of \Safe.
+In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that \Bad is always downward closed and \Safe is always upward closed.
 %		
-RP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+RP is decidable for lossy counter machines (LCM) with \Safe and \Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 %
-\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
+\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
+(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(bad,Safe)-resilient.\\
+(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(\Bad,\Safe)-resilient.\\
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
+(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
 \end{proposition}
@@
 
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
 
 \begin{theorem}\label{up-up}
-The three %uniform 
-resilience problems are decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
-and $BAD=\uparrow BAD$.
-%	and bad is upward closed or downward closed.
+The % three %uniform 
+resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+and $\Bad=\uparrow \Bad$.
+%	and \Bad is upward closed or downward closed.
 \end{theorem}
@@
 Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(bad,Safe)-resilient.
-Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
-since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm.
-%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
-if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$.
+(\Bad,\Safe)-resilient.
+Since $\Safe=\uparrow \Safe$, and WSTS, $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$, so $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+%Then if \Bad is upward closed or downward closed, we may decide the inclusion $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ which remais to compare, in both cases, two finite sets. 
+if \Bad  is upward closed, $\Bad = \uparrow B_{\Bad}$ and $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
 \iffalse
@@
 \fi
-hence the two %uniform
- resilience problems are decidable.
-
-Let us examine the %uniform 
-$k$-resilience problem (kRP). We may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
-$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
-\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$, mais il faut traiter aussi les $k$ plus petits.}
-\mathieu{Si on a $\uparrow \pred^k(Safe) = \pred^k(Safe)$ on résoud comme ci-dessus.}
-\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(Safe) \neq \pred^k(Safe)$.}
-\mathieu{... et alors, c'est plus compliqué.}
-\mathieu{L'important c'est d'avoir le $n$ minimal tel que $bad \subseteq \pred^n(Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.}and then kRP is decidable.
+hence the %two
+ %uniform
+ resilience problem is decidable.
 \end{proof}
 
-%One begins to test whether $Bad \cap \pred^*(Safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans Safe.\\
+\begin{corollary}
+The bounded resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+and $\Bad=\uparrow \Bad$.
+\end{corollary}
 
-If $Bad \cap \pred^*(Safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(Safe)$.
-resilient iff $Bad \subseteq \pred^*(Safe)$  (eq $Bad \cap \pred^*(Safe) = Bad$)
+\begin{theorem}\label{up-up}
+The % three %uniform 
+$k$-resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+and $\Bad=\uparrow \Bad$.
+%	and \Bad is upward closed or downward closed.
+\end{theorem}
 
-$Bad \cap \pred^*(Safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(Safe)$ et c'est (Bad',Safe)-resilient.
+\begin{proof}
+Let us examine the %uniform 
+$k$-resilience problem (kRP). We may compute the least $n$ such that $\Bad \subseteq \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. $S$ is %uniformely 
+$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
+\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$, mais il faut traiter aussi les $k$ plus petits.}
+\mathieu{Si on a $\uparrow \pred^k(\Safe) = \pred^k(\Safe)$ on résoud comme ci-dessus.}
+\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(\Safe) \neq \pred^k(\Safe)$.}
+\mathieu{... et alors, c'est plus compliqué.}
+\mathieu{L'important c'est d'avoir le $n$ minimal tel que $\Bad \subseteq \pred^n(\Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.}and then kRP is decidable.
+\end{proof}
+
+%One begins to test whether $\Bad \cap \pred^*(\Safe) = \emptyset$ then never resilient car une fois dans \Bad j'y reste sans pouvoir aller dans \Safe.\\
+
+If $\Bad \cap \pred^*(\Safe) \neq \emptyset$, One computes the minimal elements $A$ of $\Bad \cap \pred^*(\Safe)$.
+resilient iff $\Bad \subseteq \pred^*(\Safe)$  (eq $\Bad \cap \pred^*(\Safe) = \Bad$)
+
+$\Bad \cap \pred^*(\Safe) \neq \Bad$ alors, on peut se limiter à $\Bad' = \Bad \cap \pred^*(\Safe)$ et c'est (\Bad',\Safe)-resilient.
 
@@
 
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\downarrow BAD$.}
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}
 
@@
 %Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, Safe and BAD, are given by 
+%it is therefore reasonable to demand that both propositions, \Safe and \Bad, are given by 
 %upward-closed or downward-closed sets.
@@
 \begin{proposition}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
 The %uniform 
@@
 bounded resilience problem (BRP) are decidable.
-The %uniform
- $k$-resilience problem (kRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
 \end{proposition}
@@
 \begin{proof}
-Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
+Let $I_{\Bad}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
 The %uniform 
-resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
 This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
@@
 %Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
-%' \in I(BAD)$.
+%' \in I(\Bad)$.
 
@@
   bounded resilience problem (BRP) are decidable.
+\end{proof}
 
+\begin{proposition}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
+The %uniform
+ $k$-resilience problem (kRP) is decidable.
+ \end{proposition}
+
+\begin{proof}
 To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
 $k$-resilient; if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
+if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
 $S$ is %uniformely
@@
 
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
+\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
 
-If $Safe = \downarrow 0$ and $Bad = S$ . . .
+% In the case \Safe and \Bad are both downward closed, they can both be finite.
 
-% $Bad$ still necessarily contains $0$, and $0 \rightarrow^* 0$.
+\begin{proposition}
+The resilience problem is undecidable for WSTS with \textcolor{red}{more than one minimal element}, $\Safe=\downarrow \Safe$
+and $\Bad=\downarrow \Bad$.
+\end{proposition}
 
-$Bad = J_1 \cup J_2 \cup \ldots \cup J_m$
+\begin{proof}
+In the case of a WSTS with more than one minimal element $0_1$ and $0_2$, then the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem.  
+\end{proof}
 
-$Safe = I_1 \cup I_2 \cup \ldots \cup I_n$
-
-Question is: can we reach one of the $I_i$ from one of the $J_j$ ?
-
-If the ideal reachability problem is decidable then of course the question here should be decidable as well.
-
-But it may be possible to do more i.e.
-
-If we can express $\downarrow Post^*( J_j)$ then the question become:
-is $Safe \cap \cup_{j} \downarrow Post^*( J_j) \neq \emptyset$ ?
-
-(we need to reach an element of $Safe$ but not every possible element)
-
-Rather, the question is still
-"is $Safe \cap \cup_{j}  Post^*( J_j) \neq \emptyset$ ?"
-
-But maybe we can use $\downarrow Post^*( J_j)$ around here.
+In the case of a WSTS with only one minimal element,  
+% (ex: Petri Nets and their «Zero») 
+then by downward closure the minimal element is present in both \Safe and \Bad and hence resilience always hold.
 
 %%%%%%%%
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\uparrow BAD$.}
+\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
 %
@@
 %
-je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$
+je prends les minimaux de $\Bad$ et les idéaux de $\Safe$, pas sur que ce soit decidable....$pred(\Safe)$
 
-If $Safe = \downarrow 0$ and $Bad = \uparrow s$ . . .
+If $\Safe = \downarrow 0$ and $\Bad = \uparrow s$ . . .
 
@@
 
-In a reset petri-net, if $Bad = \uparrow q_0(\mu)$ for some $\mu \neq 0$,
-and $Safe = \downarrow q(0)$,
+In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
+and $\Safe = \downarrow 0$,
 then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.
 
-\mathieu{Ici}
+\mathieu{So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.}
 
 \begin{tabular}{ l c r }
-   BAD/Safe & Up & Down \\
-   Up & dec ? & dec ? \\
-   Down & dec & ? \\
+\hline   \Bad/\Safe & Up & Down \\ \hline
+   Up & dec ? & dec ? \\ \hline
+   Down & dec & ? \\ \hline
  \end{tabular}

### 2023-03-23 definitions.tex
details

@@
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-resilient ?\newline}
 
@@
 \begin{proposition}\label{general}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
+$\mathscr{S}=(S,\rightarrow,\leq)$ is (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is (bad,Safe)-resilient.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is  (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
 \end{proposition}
@@
 
-\subsection{Case: $Safe=\uparrow Safe$.}
+\subsection{Case: $Safe=\uparrow Safe$ and $Bad=\uparrow Bad$.}
 
 \begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
+The three resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
 %	and bad is upward closed or downward closed.
@@
 \begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.
+Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (bad,Safe)-resilient.
 Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
@@
 similar reasonning for the other case.
-hence the two uniform resilience problems are decidable.
+hence the two resilience problems are decidable.
 
-Let us examine the uniform $k$-resilience problem (UkRP).
+Let us examine the $k$-resilience problem (UkRP).
 we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
+S is $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
 \end{proof}

### 2023-03-23 definitions.tex
up for merge

--- 2023-03-03 definitions.tex
@@
 \end{theorem}
-
-%%%
-\iffalse
-%%%%
-%%%%
-\subsection{Defining resilience}
-
-\subsubsection{Resilience for general transition systems}
-
-% \textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$SAFE$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$BAD$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
-
-\problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
-{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).
-
-We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
-
-%
-%\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, BAD \subseteq S$ and $Safe \cap BAD = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, Bad \subseteq S$ and $Safe \cap Bad = \emptyset$.}
-{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
-%%
-\fi
-%%%
-
-pour $k,k'$ donnés, chercher si $Safe_{max}$ et $Bad_{max}$ ont un sens ?
-on peut faire l'union des Safe, Safe' et des Bad, Bad' prendre le max des k.
-Fixer k, chercher et calculer $Safe_{max}$ et $Bad_{max}$.
-
-%
-%\alain{dire juste que Safe = upward closed et BAD = downward closed.
-%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
-%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
-
-\section{Uniform resilience for WSTS}
-
-The resilience problem (resp.
-the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a Safe state.
-Let us formalize these properties.
-
-\problemx{resilience problem (URP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{*} Safe$ ?\newline}
-%
-%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
-
-\problemx{$k$-resilience problem (UkRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
-{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}
-
-We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
-\problemx{bounded resilience problem (UBRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}
-
-\alain{trouver les Bad et Safe maximum tels que S est resilient.
-est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que Safe=complement de Bad et donc on aurait un seul ensemble (et son complement)}
-
-These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
-So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets Safe and Bad.
-\alain{on peut penser à des ensembles Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
-%		
-URP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
-%
-In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
-
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
-%
-\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-\begin{proposition}\label{general}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
-\end{proposition}
-
-\begin{proof}
-Resilient means that 
-\end{proof}
-
-\subsection{Case: $Safe=\uparrow Safe$.}
-
-\begin{theorem}\label{up-down}
-The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
-%	and bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.
-Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$.
-since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm.
-%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
-if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$.
-similar reasonning for the other case.
-hence the two uniform resilience problems are decidable.
-
-Let us examine the uniform $k$-resilience problem (UkRP).
-we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
-\end{proof}
-
-%One begins to test whether $Bad \cap \pred^*(Safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans Safe.\\
-
-If $Bad \cap \pred^*(Safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(Safe)$.
-resilient iff $Bad \subseteq \pred^*(Safe)$  (eq $Bad \cap \pred^*(Safe) = Bad$)
-
-$Bad \cap \pred^*(Safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(Safe)$ et c'est (Bad',Safe)-resilient.
-
-Let us show that the predbasis hypothesis can be replaced by other hypotheses.
-
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\downarrow BAD$.}
-
-%
-%		cas étudié par les deux articles sur les graphes
-%
-
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-
-%Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, Safe and BAD, are given by 
-%upward-closed or downward-closed sets.
-
-The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
-
-Let us recall two other results in \cite{BFM-ic17}.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-
-\begin{proposition}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
-The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
-The uniform $k$-resilience problem (UkRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
-\end{proposition}
-
-\begin{proof}
-Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
-The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-%Proposition 30 in \cite{BFM-icalp14} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{*} y'$ in $\mathscr{S}$.
-
-Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ iff $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-
-Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $b_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-
-%
-% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
-%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
-%' \in I(BAD)$.
-
-To decide the (uniform) bounded resilience problem (UBRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is uniformely $K$-resilient.
-Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the uniform resilience problem (UBRP) and the uniform bounded resilience problem (UBRP) are decidable.
-
-To decide the UkRP, we may compare the input $k$ with $K$.
-If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.
-S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
-\end{proof}
-
-%
-
-\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}
-
-%%
-
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}
-
-%%%%%%%%
-\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\uparrow BAD$.}
-%
-%		cas exclusion mutuelle
-%
-je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$
-
-\begin{tabular}{ l c r }
-   BAD/Safe & Up & Down \\
-   Up & dec ? & dec ? \\
-   Down & dec & ? \\
- \end{tabular}
-
-%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-% \textcolor{red}{Seems like a reasonable assumption to me.}
-
-%From these considerations, we formulate instances of the abstract resilience problems for well-
-%structured transition systems.
-
-\section{Resilience for WSTS}
-
-Uniform resilience is a strong property since it implies that from bad there must exist a path to Safe.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap Bad$, there must exist a path to Safe.
-However, this condition seems more difficult since one also must decide whether a state $s \in Bad$ is reachable from $s_0$.
-%
-%\problemx{(I,J)-$k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
-%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-%\problemx{general $k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
-%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
-
-\problemx{$k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$D \longrightarrow^{\leq k} U$ ?  ou $D \cap post^*(s) \longrightarrow^{\leq k} U$ ?  \newline}
-%
-%  		{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-$D \longrightarrow^{\leq k} U$: $\forall s' \in D \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\\
-%
-$D \cap post^*(s) \longrightarrow^{\leq k} U$ ?
-%
-$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?
- \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
-c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ?
-sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}
-
-%ou plus simple : $D \rightarrow^{\leq k} U$ ?
-
-%Analogously, we formulate the resilience problem for WSTSs.
-
-\problemx{bounded Resilience problem for WSTS}
-{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}
-
-\problemx{Resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$ ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{*} s''$ ?\newline}
-
-%\textcolor{blue}{
-%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
-%}
-
-\begin{remark} 
-We require $U$ to be upward-closed, as, in the general case if we lift restrictions on $U$ and $D$, then reachability reduces itself to resilience.
-Indeed, the particular resilience problem where we take $D = \{s\}$ and $U = \{x\}$ is reachability of $x$ from $s$, and thus generally undecidable for WSTS.
-\end{remark}
-
-\textcolor{blue}{
-et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
-}
-
-\textcolor{blue}{
-compute the sets of $(I,J,k)$ st S is resilient.\\
-max des I,J possibles \\
-}
-
-\begin{proposition}\label{reductions}
-coverability and $t$-liveness are reduceable to resilience.
-%	home state and reachability
-\end{proposition}
-
-\begin{proof}
-
-\begin{itemize}
-\item Let us choose $J=S= \downarrow S$ and $I=\uparrow x$ then resilience is equivalent to coverability of $x$ from $s$.
-%\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the resilience problem
-where $J=S$ and $I = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
-
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
-%  \textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
-Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
-This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-% Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
-\end{proof}
-
-% Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
-
-% and with effective basis of $post^*$ ???
-
-% expliquer toutes les variantes de la résilience
-
-% Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
-
-\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
-\end{theorem}
-
-We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
-
-\begin{corollary}\label{postcomputable}
-{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
-\end{corollary}
-
-\begin{proof}
-From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
-$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
-returning $k_m$ if that is the case.
-Then, because $k_m$-resilience is decidable, 
-checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
-and resilience is decidable.
-% \textcolor{red}{à faire Mathieu}
-\end{proof}
-
-%\"Ozkan
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
-For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
-given a state $s$ of a WSTS
-% in the regarded class 
-with strong upward compatibility 
-and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$. 
-\end{proposition}
-
-The characterization is used to show that Petri nets are $\post^*$-effective.
-It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Hence Resilience is decidable for Petri nets and VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
-
-\begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}
-Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
-%in counter $i$ 
-by examining if there is %some vector $v$ in the basis such that $v(i)=0$
-$0$ in the basis%
-. But reachability of $0$  %in a counter $i$ 
-is undecidable for Reset VASS. 
-% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
-\end{proof}
-
-%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
-
-remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:
-
-\begin{proposition}(à prouver)
- leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $\uparrow post^*$ ?
-\end{proposition}
-
-\begin{proof}
-leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
- \textcolor{red}{à faire Mathieu}
-
-\mathieu{En train de chercher un contre exemple pour le corollaire du Lemme~\ref{Lemma intersection}.}
-
-\mathieu{
-
-On considère un LCS, où on autorise à effacer des messages (lossyness) mais effacer un message prends une étape (contrairement à [ref], où ils font une sorte de cloture de effacer-transition-effacer).
-On n'a alors pas la strong compatibility:
-
-On prends: $(q, b) 	\to (q,bb)$ par exemple et $(q, baaaaaa) \geq (q,b) $, alors
-$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
-MAIS
-
-C'est pas "en une étape".
-
-(On considère -il faut le mentionner/vérifier- que on ne peut effaçer que d'un coté ? et pas au milieu?)
-
-Dans ce cas là, du coup, $pred(\uparrow \{(q,b)\}$ n'est pas clos par le haut. \\
-
-Revenons au Corollaire
-
-On voudrais avoir
-
-$post^* (s) \cap J \subseteq I^k$
-
-MAIS pas
-
-$ (\uparrow post^* (s)) \cap J \subseteq I^k$ \\
-
-Ça serait un cas où leurs algorithme vas dire 
-"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
-
-Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
-
-Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,b)\})$ qui n'est pas clos par le haut.
-
-On voudrait
-
-$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,b)\})$
-
-Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
-$\uparrow post^* (s)$ being too big
-donc autant pas se restreindre du tout sur $J$
-
-$post^*(s) \subseteq pred(\uparrow \{ (q,b)\})$
-
-Si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
-$b$ ? et évidement en perdant le dernier message (parce qu'on est lossy). Et c'est tout, disons.
-
-Déjà par lossiness $(q,\epsilon)$ serait dans $post^*((q,b))$.
-
-Et aussi dans $pred(\uparrow \{ (q,b)\})$ puisque $(q,\epsilon) \to (q,b)$.
-
-Donc on aurait:
-
-$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
-
-Et on aurait $(q,aaaaaa)$ dans 
-$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$.
-
-Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
-(même si il est dans $pred^*(\uparrow \{ (q,b)\})$).
-
-}
-
-\end{proof}
-
-\begin{proposition}(à prouver)
-{\sc bounded resilience} is undecidable for WSTS with upward compatibility withd an effective basis of $post^*$.
-\end{proposition}
-
-\begin{proof}
- \textcolor{red}{pas certain ! à faire Mathieu}
-\end{proof}
-
-%
-%si on garde strong mais on enlève effective basis of $post^*$:
-%
-%\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
-%\end{proposition}
-%
-%\begin{proof}
-%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
-%\end{proof}
-
-\begin{theorem}
-{\sc Resilience} is decidable for (not necessarily strong) WSTS with effective $\uparrow$ $\post^*$ basis.
-\end{theorem}
-
-\begin{proof}
-Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_I$ a basis of $I$
-and $J$ decidable downward-closed.
-By applying Lemma~\ref{Lemma intersection} twice, we obtain
-
-$$ post^*(s) \cap J \subseteq \pred^*(I) \equiv B_p \cap J \subseteq \pred^*(I)$$
-
-Since $B_p$ is finite and $J$ is decidable, we can directly compute $ B_p \cap J$.
-% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(I)$ from $B_I$, and hence check that $B_p \cap J \setminus \pred^*(I) = \emptyset$. 
-\end{proof}
-
-% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
-% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
-% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
-%
-% Well, maybe their version of resilience but resilience with * seems fine.
-%
-
-Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.
-This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
-
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem}
-
-\begin{proof}
-If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
-reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
-reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
-$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
-
-The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
-and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
-
-We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $r_1$.
-
-The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
-
-Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
-and "freezing" the marking of $P_1$.
-
-Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
-The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
-
-Now we need to see if that marking is zero.
-
-We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
-
-If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
-$s$ fires in the marking zero.
-
-If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
-
-In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
-Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-
-Now if $0$ is reachable
-in $R(C_1, \mu_1)$ 
-then the net $C_2$ can also reach this marking
-in the place of $P_1$ by executing the same sequence of transition firings.
-Then $s$ can fire, freezing the $C_1$ subset.
-Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $0$ is reachable then $C_2$ is not live.
-
-Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
-$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
-(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
-% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
-% et r_2 commence sans token)
-If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
-
-\end{proof}
-
-\begin{corollary}
-Reset Petri Net liveness is undecidable 
-\end{corollary}
-
-\begin{corollary}
-Reset Petri Net $t$-liveness is undecidable 
-\end{corollary}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that resilience is undecidable for WSTS with strong monotony.
-
-\begin{theorem}
-{\sc resilience} is undecidable for WSTS with strong monotony.
-\end{theorem}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-reducible to the $k$-resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-$\pred^*(I) = \pred^{k}(I)$
-Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-
-$\pred^*(I) = \pred^{k}(I)$
-
-\end{proof}
-
-\begin{corollary}
-Bounded Resilience is undecidable for WSTS with strong monotony.
-\end{corollary}
-
-\begin{proof} 
-  From
-  cf
-the two  theorems above. 
-\end{proof}
-
-Let us deduce that effective basis of $post^*$ ex reset PN ?
-
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded resilience is decidable for VASS, hence resilience is decidable for VASS.
-\alain{resilience decidable for LCS ?
-LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-We may consider context-free grammars.
-
-\begin{corollary}
-{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
-\end{corollary}
-
-\begin{proof}
-
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
-\end{proof}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}

### 2023-03-23 introduction.tex
details

\section{Introduction}\label{section introduction}

{\bf Context.} In the quite informal paper \cite{DBLP:journals/corr/PrasadZ16}, Prasad and Zuck studied resilience of a process to an adversary and they used the framework of WSTS enjoying both upward and downward monotonies to prove the decidability of a kind of resilience.

MFCS 2023 \\
Abstract submission deadline:    	April 24th (AoE) \\
Paper submission deadline:    	April 28th (AoE)

\begin{itemize}

\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)

\item safe clos haut : resilience indecidable for WSTS with strong.

\end{itemize}

### 2023-03-23 resilience.tex
details
up
Up

@@
 the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a \Safe state.
+a path of length smaller than $k$) that reaches a safe state.
 Let us formalize these properties.
@@
 
-\alain{trouver les \Bad et \Safe maximum tels que S est resilient.
+\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient.
 est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que \Safe=complement de \Bad et donc on aurait un seul ensemble (et son complement)}
+on pourrait décider que \Safe=complement de \Bad~ et donc on aurait un seul ensemble (et son complement)}
 
@@
 \begin{theorem}\label{up-up}
-The % three %uniform 
-resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+The resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.
-%	and \Bad is upward closed or downward closed.
+
 \end{theorem}
@@
 \begin{proof}
-Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient.
-Since $\Safe=\uparrow \Safe$, and WSTS, $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$, so $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-%Then if \Bad is upward closed or downward closed, we may decide the inclusion $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ which remais to compare, in both cases, two finite sets. 
-if \Bad  is upward closed, $\Bad = \uparrow B_{\Bad}$ and $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-\iffalse
-similar reasonning for the other case.
-\fi
-hence the %two
- %uniform
- resilience problem is decidable.
+% Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (\Bad,\Safe)-resilient. 
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
+Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$,
+hence the resilience problem is decidable.
 \end{proof}
@@
 
-\begin{theorem}\label{up-up}
+\begin{theorem}\label{k-up-up}
 The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+$k$-resilience problem is decidable for WSTS with effective predbasis, strong upward-compatibility, $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.
@@
 \begin{proof}
-Let us examine the %uniform 
-$k$-resilience problem (kRP). We may compute the least $n$ such that $\Bad \subseteq \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. $S$ is %uniformely 
-$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
-\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$, mais il faut traiter aussi les $k$ plus petits.}
-\mathieu{Si on a $\uparrow \pred^k(\Safe) = \pred^k(\Safe)$ on résoud comme ci-dessus.}
-\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(\Safe) \neq \pred^k(\Safe)$.}
-\mathieu{... et alors, c'est plus compliqué.}
-\mathieu{L'important c'est d'avoir le $n$ minimal tel que $\Bad \subseteq \pred^n(\Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.}and then kRP is decidable.
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward-compatibility,  $\pred^k(\Safe)=\uparrow \pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
 \end{proof}
-
-%One begins to test whether $\Bad \cap \pred^*(\Safe) = \emptyset$ then never resilient car une fois dans \Bad j'y reste sans pouvoir aller dans \Safe.\\
-
-If $\Bad \cap \pred^*(\Safe) \neq \emptyset$, One computes the minimal elements $A$ of $\Bad \cap \pred^*(\Safe)$.
-resilient iff $\Bad \subseteq \pred^*(\Safe)$  (eq $\Bad \cap \pred^*(\Safe) = \Bad$)
-
-$\Bad \cap \pred^*(\Safe) \neq \Bad$ alors, on peut se limiter à $\Bad' = \Bad \cap \pred^*(\Safe)$ et c'est (\Bad',\Safe)-resilient.
-
-Let us show that the predbasis hypothesis can be replaced by other hypotheses.
 
@@
 
-\begin{proposition}\label{down-up}
+\begin{theorem}\label{down-up}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
-The %uniform 
-resilience problem (RP) and the %uniform 
-bounded resilience problem (BRP) are decidable.
-\end{proposition}
+The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
+\end{theorem}
 
@@
 
-To decide the %(uniform)
- bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
-  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
-  resilience problem (BRP) and the %uniform 
+To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is   $K$-resilient.
+Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the
+  resilience problem (BRP) and the  
   bounded resilience problem (BRP) are decidable.
@@
 
-\begin{proposition}
+\begin{theorem}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
-The %uniform
- $k$-resilience problem (kRP) is decidable.
- \end{proposition}
+The  $k$-resilience problem (kRP) is decidable.
+ \end{theorem}
 
@@
 
-\begin{tabular}{ l c r }
-\hline   \Bad/\Safe & Up & Down \\ \hline
-   Up & dec ? & dec ? \\ \hline
-   Down & dec & ? \\ \hline
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
+   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & ? Undecidable & Undecidable \\ \hline
+   BRP & Decidable & ? Decidable (Theorem~\ref{down-up}) & ? & Undecidable \\ \hline
+      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable & ? & Undecidable \\ \hline
  \end{tabular}

### 2023-03-23 resilience.tex
up

@@
 
-\begin{proposition}\label{general}
+\begin{proposition}[Reformulation]\label{reformulation}
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
 (\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient.\\
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
@@
 
+\begin{proposition}\label{general}
+\textcolor{red}{
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(\Bad,\Safe)-resilient}
+\end{proposition}
+
 \begin{proof}
-Resilient means that 
+Resilient means that ...
 \end{proof}
@@
 
-\begin{corollary}
-The bounded resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+\begin{proposition}
+In WSTS with strong compatibility and effective predbasis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$,
+and there exists $K \in \N$ such that 
+$\pred^k(\Safe) = \uparrow \pred^k(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}\label{B-up-up}
+The bounded resilience problem is decidable for WSTS with effective predbasis,
+strong compatibility
+ $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.
@@
 The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective predbasis, strong upward-compatibility, $\Safe=\uparrow \Safe$
+$k$-resilience problem is decidable for WSTS with effective predbasis, strong %upward-
+compatibility, $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.
@@
 Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward-compatibility,  $\pred^k(\Safe)=\uparrow \pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
+From a similar reasoning as above, the $k$-resilience problem is decidable.
 \end{proof}
@@
 
-\begin{theorem}
+\begin{theorem}\label{k-down-up}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
@@
 
-\begin{proposition}
+\begin{theorem}\label{down-down}
 The resilience problem is undecidable for WSTS with \textcolor{red}{more than one minimal element}, $\Safe=\downarrow \Safe$
 and $\Bad=\downarrow \Bad$.
-\end{proposition}
+\end{theorem}
 
@@
 
+\begin{center}
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & ? Undecidable & Undecidable \\ \hline
-   BRP & Decidable & ? Decidable (Theorem~\ref{down-up}) & ? & Undecidable \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable & ? & Undecidable \\ \hline
+\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
+   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
+   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & Undecidable & ?? \\ \hline
+      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & Undecidable & ?? \\ \hline
  \end{tabular}
+\end{center}
+
+\mathieu{En rouge quand il s'agit de conjectures}
 

### 2023-03-23 reachable-resilience.tex
Changed name of Uniform resilience to remove 'Uniform'
up

@@
-\section{Resilience for WSTS}
+\section{Resilience from a single-state for WSTS}
 
-Uniform resilience is a strong property since it implies that from bad there must exist a path to Safe.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap Bad$, there must exist a path to Safe.
-However, this condition seems more difficult since one also must decide whether a state $s \in Bad$ is reachable from $s_0$.
+Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$.
+But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap \Bad$, there must exist a path to $\Safe$.
+However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
 %
@@
 \problemx{$k$-resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$D \longrightarrow^{\leq k} U$ ?  ou $D \cap post^*(s) \longrightarrow^{\leq k} U$ ?  \newline}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
+{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
 %
-%  		{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-$D \longrightarrow^{\leq k} U$: $\forall s' \in D \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\\
-%
-$D \cap post^*(s) \longrightarrow^{\leq k} U$ ?
-%
-$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?
- \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$.
-c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ?
-sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}
-
-%ou plus simple : $D \rightarrow^{\leq k} U$ ?
 

### 2023-03-23 resilience.tex
up

@@
 \subsubsection{Resilience for general transition systems}
-
-% \textcolor{red}{Ask the question why use a set of propositions for $\Safe$ and $\Bad$ rather than use subsets of the set of configurations ? }
 
@@
 (\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient}
+(\Bad,\Safe)-resilient (CONJECTURE FOR NOW)
+}
 \end{proposition}
@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
-The  $k$-resilience problem (kRP) is decidable.
+The  $k$-resilience problem (kRP) is decidable. \textcolor{red}{CONJECTURE for now}
  \end{theorem}
@@
 \begin{theorem}\label{down-down}
-The resilience problem is undecidable for WSTS with \textcolor{red}{more than one minimal element}, $\Safe=\downarrow \Safe$
+The resilience problem is undecidable for WSTS with more than one minimal element, 
+$\Safe=\downarrow \Safe$
 and $\Bad=\downarrow \Bad$.
@@
 \begin{proof}
-In the case of a WSTS with more than one minimal element $0_1$ and $0_2$, then the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem.  
+In the case of a WSTS with at least two minimal elements $0_1$ and $0_2$, the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem, hence its undecidability.  
 \end{proof}
@@
 % (ex: Petri Nets and their «Zero») 
-then by downward closure the minimal element is present in both \Safe and \Bad and hence resilience always hold.
+then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
 
@@
 and $\Safe = \downarrow 0$,
-then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.
+then the resilience problem is similar to the single-state zero-reachability problem, which is undecidable.
 
-\mathieu{So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.}
+So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.
+
+\subsection{Case Synthesis}
+
+\mathieu{En rouge quand il s'agit de conjectures}
 
@@
    RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
-   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & Undecidable & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & Undecidable & ?? \\ \hline
+   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
+      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & ?? & ?? \\ \hline
  \end{tabular}
 \end{center}
-
-\mathieu{En rouge quand il s'agit de conjectures}
 

### 2023-03-23 definitions.tex
up def
Up

@@
 
-A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in J$. 
+A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
+$s \in D$. 
 %\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
@@
 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $I \subseteq S$ is an upward-closed set, the sets $\pred(I )$, and $\pred^{\geq k}(I )$ for
-every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(I)$.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
+every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
 
 \begin{remark}
-$I^{k+1} $ can be rewritten $I^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
+$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
+U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
+U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
+=  U \cup \pred(U^k)$.
 \end{remark}
@@
 % Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
-$k(I) \leq k$. This also implies that $\pred^*(I) = I^k$.
+If $\mathscr{S}$ is a WSTS with strong monotony and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
+$k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
 \end{fact}
-
-\iffalse
-\begin{remark}
-$I^{k+1} = \bigcup_{0 \leq j \leq k+1} \pred^j(I) = 
-I \cup \bigcup_{1 \leq j \leq k+1} \pred^j(I) =
-I \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(I))
-=  I \cup \pred(I^k)$.
-\end{remark}
-\fi
 
@@
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of an effective strongly well-structured transition
-system, we can decide whether we can reach $I$ from $s$.
+ Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly well-structured transition
+system, we can decide whether we can reach $U$ from $s$.
 \end{lemma}
@@
 \begin{proof}
-We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
+We have to show that we can compute a basis of $U^{k+1}$ if we are given a basis of $U^k $. 
 Then the
-decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
+decidability of the stop condition follows directly. Let $B$ be a basis of $U^k$. 
 We have
-$$I^{k+1} = I \cup \pred(I^k ) = I \cup
+$$U^{k+1} = U \cup \pred(U^k ) = U \cup
 \bigcup_{s' \in B}
@@
 
-Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
-Fact~\ref{fact basis}, we can compute a basis of $I^{k+1}$.
+Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
+Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
 \end{proof}

### 2023-03-23 reachable-resilience.tex
up

@@
 
-\problemx{$k$-resilience problem for WSTS}
+\problemx{$k$-reachable-resilience problem for WSTS (kRRP)}
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
@@
 
-%Analogously, we formulate the resilience problem for WSTSs.
+Analogously, we formulate bounded and general variations of the problem.
 
-\problemx{bounded Resilience problem for WSTS}
-{A state $s$ of a WSTS $W=(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
+\problemx{bounded-reachable-resilience problem for WSTS (BRRP)}
+{A state $s$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $W$ is $k$-resilient. ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-reachable-resilient. ?\newline}
 
-\problemx{Resilience problem for WSTS}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
-{$ ~ \forall s' \in D ~ (s \rightarrow^* s') \implies \exists s'' \in U ~ s' \rightarrow^{*} s''$ ?\newline}
+\problemx{Rachable-resilience problem for WSTS (RRP)}
+{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
+{$ ~ \forall s' \in \Bad ~ (s \rightarrow^* s') \implies \exists s'' \in \Safe ~ s' \rightarrow^{*} s''$ ?\newline}
 
@@
 
-\begin{remark} 
-We require $U$ to be upward-closed, as, in the general case if we lift restrictions on $U$ and $D$, then reachability reduces itself to resilience.
-Indeed, the particular resilience problem where we take $D = \{s\}$ and $U = \{x\}$ is reachability of $x$ from $s$, and thus generally undecidable for WSTS.
-\end{remark}
+\iffalse
 
@@
 
+\fi
+
 \begin{proposition}\label{reductions}
-coverability and $t$-liveness are reduceable to resilience.
+Coverability and $t$-liveness are reduceable to the reachable resilience problem.
 %	home state and reachability
@@
 \begin{itemize}
-\item Let us choose $J=S= \downarrow S$ and $I=\uparrow x$ then resilience is equivalent to coverability of $x$ from $s$.
-%\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
+\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
 
@@
 is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the resilience problem
-where $J=S$ and $I = \uparrow \pred(t)$.
+The problem reduces itself to the reachable resilience problem
+where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
 
@@
 
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
-%  \textcolor{red}{  Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$.
-Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$.
-Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$.
-This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
-% Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.}
-%	\mathieu{Ne marche pas (un des deux sens est invalide imho)}
 \end{proof}
@@
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
+{\sc bounded-reachable-resilience} and {\sc $k$-reachable-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*$ is computable.
 \end{theorem}
 
-We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
+We may immediately generalyse this last result by strengthening to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
 
 \begin{corollary}\label{postcomputable}
-{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
+{\sc reachable-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*$ is computable.
 \end{corollary}
@@
 From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\post^*(I) = I^{k_m}$. We can compute this $k_m$ by iteratively computing
-$I^{k+1}$ from $I^k$, checking $I^{k+1}=I^k$, 
+$\pred^*(\Safe) =  \Safe^{k_m}$. We can compute this $k_m$ by iteratively computing
+$\Safe^{k+1}$ from $\Safe^k$, checking $\Safe^{k+1}=\Safe^k$, 
 returning $k_m$ if that is the case.
-Then, because $k_m$-resilience is decidable, 
-checking $\uparrow post^*(s) \cap J \subseteq I^k = \pred^*(I)$ is,
-and resilience is decidable.
-% \textcolor{red}{à faire Mathieu}
+Then, because $k_m$-reachable-resilience is decidable, 
+checking $\uparrow post^*(s) \cap \Bad \subseteq \Safe^k = \pred^*(\Safe)$ is,
+and reachable-resilience is decidable.
 \end{proof}
@@
 with strong upward compatibility 
-and a decidable downward-closed set $J$, it can be decided whether $\exists s' \in J ~ s \to^* s'$. 
+and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
 \end{proposition}
@@
 It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Hence Resilience is decidable for Petri nets and VASS.
+Hence reachable-resilience is decidable for Petri nets and VASS.
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
@@
 Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
+Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
 %in counter $i$ 
@@
 
-remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:
+\mathieu{remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:}
 
@@
 leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
- \textcolor{red}{à faire Mathieu}
-
-\mathieu{En train de chercher un contre exemple pour le corollaire du Lemme~\ref{Lemma intersection}.}
-
 \mathieu{
@@
 
-\begin{proposition}(à prouver)
-{\sc bounded resilience} is undecidable for WSTS with upward compatibility withd an effective basis of $post^*$.
-\end{proposition}
-
-\begin{proof}
- \textcolor{red}{pas certain ! à faire Mathieu}
-\end{proof}
-
 %
@@
 
+\begin{lemma}\label{Lemma intersection}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
+Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
+Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\uparrow A}$, the inclusion
+$({\uparrow  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
+\end{proof}
+
 \begin{theorem}
-{\sc Resilience} is decidable for (not necessarily strong) WSTS with effective $\uparrow$ $\post^*$ basis.
+{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
+$\uparrow$ $\post^*$ basis.
 \end{theorem}
@@
 \begin{proof}
-Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_I$ a basis of $I$
-and $J$ decidable downward-closed.
+Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
+and $\Bad$ decidable downward-closed.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
 
-$$ post^*(s) \cap J \subseteq \pred^*(I) \equiv B_p \cap J \subseteq \pred^*(I)$$
+$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$
 
-Since $B_p$ is finite and $J$ is decidable, we can directly compute $ B_p \cap J$.
+Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
 % For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(I)$ from $B_I$, and hence check that $B_p \cap J \setminus \pred^*(I) = \emptyset$. 
+We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
 \end{proof}
@@
 %
+
+\begin{proposition}
+\textcolor{red}{
+{\sc bounded resilience} is undecidable for WSTS with upward compatibility with an effective basis of $post^*$.
+CONJECTURE
+}
+\end{proposition}
+
+\begin{proof}
+ \textcolor{red}{pas certain ! à faire Mathieu}
+\end{proof}
 

### 2023-03-24 reachable-resilience.tex
Up

@@
 
-\mathieu{remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:}
+%
+%si on garde strong mais on enlève effective basis of $post^*$:
+%
+%\begin{proposition}(à prouver)
+%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
+%\end{proposition}
+%
+%\begin{proof}
+%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
+%\end{proof}
 
-\begin{proposition}(à prouver)
- leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $\uparrow post^*$ ?
+Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general reachable-resilience problem.
+
+\begin{lemma}\label{Lemma intersection}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
+Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
+Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\uparrow A}$, the inclusion
+$({\uparrow  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
+\end{proof}
+
+\begin{theorem}
+{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
+$\uparrow$ $\post^*$ basis.
+\end{theorem}
+
+\begin{proof}
+Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
+and $\Bad$ decidable downward-closed.
+By applying Lemma~\ref{Lemma intersection} twice, we obtain
+
+$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$
+
+Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
+% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
+We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
+\end{proof}
+
+% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
+% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
+% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
+%
+% Well, maybe their version of resilience but resilience with * seems fine.
+%
+
+However when removing strong compatibility, precision is lost and the algorithm do not work anymore.
+
+\begin{proposition}
+The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-reachable-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
 \end{proposition}
 
-\begin{proof}
-leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
+\begin{proof}{Sketch}
+Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
+For instance with a rule $\delta(q,b) = \text{ add some }b $,
+we have $(q,b) \to (q,bb)$, $(q, baaa) \geq (q,b)$ and
+$(q, baaa) \to^* (q,bb)$ however it takes more than a single step.
+
+In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
+$post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$
+do not imply
+$\uparrow post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$ for a downward-closed set $D$.
+In particular by lossiness $(q, \epsilon)$ would belong to
+$post^* (s)$,
+hence $(q,a)$, $(q,aa)$, etc. would belong to $\uparrow post^* (s)$ although not necessarily to
+$post^* (s)$. $(q,aaa)$ not being in $pred(\uparrow \{ (q,b)\})$ means the algorithm 
+from \cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold in this example when $\Safe=\uparrow \{(q,b)\}$ and $\Bad =S$,
+because $\uparrow post^* (s) \cap S \not\subseteq \pred(\uparrow \{(q,b)\})$,
+despite having 
+$post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.
+
+\iffalse
 \mathieu{
@@
 }
+\fi
 
 \end{proof}
-
-%
-%si on garde strong mais on enlève effective basis of $post^*$:
-%
-%\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
-%\end{proposition}
-%
-%\begin{proof}
-%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
-%\end{proof}
-
-\begin{lemma}\label{Lemma intersection}
-Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
-\end{lemma}
-
-\begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
-Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
-Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\uparrow A}$, the inclusion
-$({\uparrow  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
-\end{proof}
-
-\begin{theorem}
-{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
-$\uparrow$ $\post^*$ basis.
-\end{theorem}
-
-\begin{proof}
-Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
-and $\Bad$ decidable downward-closed.
-By applying Lemma~\ref{Lemma intersection} twice, we obtain
-
-$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$
-
-Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
-% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
-\end{proof}
-
-% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
-% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
-% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
-%
-% Well, maybe their version of resilience but resilience with * seems fine.
-%
 

### 2023-03-24 resilience.tex
Up

@@
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets \Safe and \Bad.
-\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
+Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
+\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
 For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of \Safe.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that \Bad is always downward closed and \Safe is always upward closed.
+Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
+In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is always downward closed and $\Safe$ is always upward closed.
 %		
-RP is decidable for lossy counter machines (LCM) with \Safe and \Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
-%
-\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
@@
 %
-je prends les minimaux de $\Bad$ et les idéaux de $\Safe$, pas sur que ce soit decidable....$pred(\Safe)$
 
-If $\Safe = \downarrow 0$ and $\Bad = \uparrow s$ . . .
+\begin{theorem}\label{down-up}
+\textcolor{red}{
+The resilience problem is undecidable for WSTS with more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$.
+}
+\end{theorem}
 
-Then resilience kinds of becomes "can we reach $0$ from $\uparrow s$?".
-
-In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
-and $\Safe = \downarrow 0$,
-then the resilience problem is similar to the single-state zero-reachability problem, which is undecidable.
-
-So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.
+\begin{proof}
+TBD
+\end{proof}
 

### 2023-03-24 reachable-resilience.tex
up

@@
 
-Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.
+Unfortunately, reachable-resilience is undecidable for (general) WSTS even with strong monotony.
 This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
@@
 
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that resilience is undecidable for WSTS with strong monotony.
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that reachable-resilience is undecidable for WSTS with strong monotony.
 
 \begin{theorem}
-{\sc resilience} is undecidable for WSTS with strong monotony.
+{\sc reachable-resilience} is undecidable for WSTS with strong monotony.
 \end{theorem}
@@
 \begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-reducible to the $k$-resilience problem.
+In WSTS with strong monotony, the bounded reachable-resilience problem is
+reducible to the $k$-reachable-resilience problem.
 \end{proposition}
@@
 \begin{proposition}
-In WSTS with strong monotony, the bounded resilience problem is
-equivalent to the resilience problem.
+In WSTS with strong monotony, the bounded reachable resilience problem is
+equivalent to the reachable-resilience problem.
 \end{proposition}
@@
 \begin{corollary}
-Bounded Resilience is undecidable for WSTS with strong monotony.
+Bounded reachable-resilience and 
+$k$ reachable-resilience are undecidable for WSTS with strong monotony.
 \end{corollary}
@@
 
-Let us deduce that effective basis of $post^*$ ex reset PN ?
+% Let us deduce that effective basis of $post^*$ ex reset PN ?
 
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded resilience is decidable for VASS, hence resilience is decidable for VASS.
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded reachable-resilience is decidable for VASS, hence reachable-resilience is decidable for VASS.
 \alain{resilience decidable for LCS ?

### 2023-03-24 reachable-resilience.tex
Up

@@
 It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded reachable-resilience is decidable for VASS.
 Hence reachable-resilience is decidable for Petri nets and VASS.
@@
 
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded reachable-resilience is decidable for VASS, hence reachable-resilience is decidable for VASS.
-\alain{resilience decidable for LCS ?
-LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
+ \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
 

### 2023-03-24 resilience.tex
Up
details

@@
 
+\iffalse
+
 \begin{proposition}\label{general}
@@
 \end{proof}
+
+\fi
 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$,
-and there exists $K \in \N$ such that 
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $k \in \N$ \alain{K pas utilisé} such that 
 $\pred^k(\Safe) = \uparrow \pred^k(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
@@
 compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
-From a similar reasoning as above, the $k$-resilience problem is decidable.
+From a similar reasoning as above, the $k$-resilience problem is decidable.\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
 \end{proof}
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$ \alain{Min ou B ?}.
 The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
@@
 To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is   $K$-resilient.
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?}.
 Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the
@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
-The  $k$-resilience problem (kRP) is decidable. \textcolor{red}{CONJECTURE for now}
+The  $k$-resilience problem (kRP) is decidable for $k \geq K$. \textcolor{red}{CONJECTURE for now}
  \end{theorem}

### 2023-03-24 resilience.tex
details

@@
 To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?}.
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?
+il pourrait exister un K' pour lequel $\mathscr{S}$ est K'-resilient....}.
 Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the

### 2023-03-24 resilience.tex
details

--- 2023-03-24 resilience.tex
@@
 To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?}.
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ \alain{comment trouver les $k_a$ ?} and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?
+il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}.
 Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the

### 2023-03-24 resilience.tex
up

@@
 compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $k \in \N$ \alain{K pas utilisé} such that 
-$\pred^k(\Safe) = \uparrow \pred^k(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
 Hence the equivalence.

### 2023-03-24 resilience.tex
up

@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$ \alain{Min ou B ?}.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $\Bad$ and $\Safe$.
 The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
@@
 \begin{proof}
-Let $I_{\Bad}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
+Let $\Bad=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
 The %uniform 
@@
 To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ \alain{comment trouver les $k_a$ ?} and we deduce that  $\mathscr{S}$ is $K$-resilient \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ?
-il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}.
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$
+ % \alain{comment trouver les $k_a$ ?} 
+ by iteratively checking the different possibles values of every $k_a$ until we find the right one,
+ and we deduce that  $\mathscr{S}$ is $K$-resilient,
+%  \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ? il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}. 
+else if some of these $n \times m$ coverability questions are negative then resilience
+do not hold and bounded resilience do not hold either.
  Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the 

### 2023-03-24 resilience.tex
up

@@
 \begin{theorem}\label{k-down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ and $\Safe$.
 The  $k$-resilience problem (kRP) is decidable for $k \geq K$. \textcolor{red}{CONJECTURE for now}

### 2023-03-24 resilience.tex
details

@@
  Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the 
-  resilience problem (BRP) and the  
+  resilience problem (RP) and the  
   bounded resilience problem (BRP) are decidable.

### 2023-03-24 resilience.tex
up
details

@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $\Bad$ and $\Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite \alain{non, confusion entre ensemble et base} sets: $\Bad$ and $\Safe$.
 The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
@@
 \begin{proof}
-Let $\Bad=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
+Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
 The %uniform 
 resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+This infinite set of coverability questions can be reduced to a finite number of instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
 
@@
 To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,m $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
+\alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
  % \alain{comment trouver les $k_a$ ?} 

### 2023-03-24 resilience.tex
details

@@
 there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-
+%
+This proves that resilience and bounded resilience are equivalent since, here, resilience implies bounded resilience (bounded resilience always implies resilience).
 %
@@
  % \alain{comment trouver les $k_a$ ?} 
- by iteratively checking the different possibles values of every $k_a$ until we find the right one,
+ by iteratively checking the different possibles values of every $k_{i,j}$ until we find the right one,
  and we deduce that  $\mathscr{S}$ is $K$-resilient,
@@
 To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
-$k$-resilient; if $k < K$ we cannot answer.
+$k$-resilient; 
+if $k < K$ we cannot answer.
 if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.

### 2023-03-24 resilience.tex
Up

@@
 \begin{theorem}\label{k-down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ and $\Safe$.
-The  $k$-resilience problem (kRP) is decidable for $k \geq K$. \textcolor{red}{CONJECTURE for now}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility% and the predbasis hypothesis
+, and two finite sets: $\Bad$ and $\Safe$.
+The  $k$-resilience problem (kRP) is decidable for $k \geq K$. 
+% \textcolor{red}{CONJECTURE for now}
  \end{theorem}
@@
 \begin{proof}
-To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
-$k$-resilient; 
-if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
-$S$ is %uniformely
- $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
+To decide the kRP, we may compare the input $k$ with $K=max(k_a \mid a=1,..,m \times n)$.
+	If $k \geq K$, we conclude that 
+$k$-resilience holds.
+	If $k < K$, then we can answer the question by checking every paths of length smaller
+	than $K$ from the $J_i$ in the completion.
 \end{proof}

### 2023-03-24 resilience.tex
Up
up

@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite \alain{non, confusion entre ensemble et base} sets: $\Bad$ and $\Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two 
+%finite \alain{non, confusion entre ensemble et base}
+ sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
 The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
@@
 	If $k < K$, then we can answer the question by checking every paths of length smaller
-	than $K$ from the $J_i$ in the completion.
+	than $K$ from the ideals of the decomposition of $\Bad$ in the completion.
 \end{proof}

### 2023-03-24 resilience.tex
details

--- 2023-03-24 resilience.tex
@@
 If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,m $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
-\alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
+\alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve puisqu'on l'utilise ailleurs (théorème suivant) comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
  % \alain{comment trouver les $k_a$ ?} 
- by iteratively checking the different possibles values of every $k_{i,j}$ until we find the right one,
+ by iteratively checking the different possibles values of every $k_{i,j}$ until we find the minimal one,
  and we deduce that  $\mathscr{S}$ is $K$-resilient,
@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ and $\Safe$.
-The  $k$-resilience problem (kRP) is decidable for $k \geq K$. \textcolor{red}{CONJECTURE for now}
+The  $k$-resilience problem (kRP) is decidable for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. \textcolor{red}{CONJECTURE for now}
  \end{theorem}
@@
 \begin{proof}
-To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
+We begin to compute $K$ and $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
+If $K=\infty$ then $\mathscr{S}$ is not resilient for Safe and Bad.
+Now, if $k \geq min(K,n)$, we conclude that is %uniformely 
 $k$-resilient; 
-if $k < K$ we cannot answer.
-if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
-$S$ is %uniformely
- $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
+%	if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, 
+ $S$ is $k$-resilient.
+%
+ \alain{hum...mais il peut y avoir des chemins plus courts}
+%	and then kRP is decidable.
 \end{proof}

### 2023-03-24 application.tex
details
up

@@
 \section{Applications section}
+
+\subsection{VASS, 2-VASS, semilinear VASS, Lossy Channel Systems, Integer et Continuous VASS, Regular Channel Systems}
+
+\subsection{LCM}
+
+\begin{theorem}{se déduit du papier Schnoebelen RP 2010, page 8}
+Resilience is decidable for LCM when $\Safe$ and $\Bad$ are semilinear sets.
+\end{theorem}
 
@@
 
-\subsection{Vector Addition System with States}
+\subsection{Vector Addition System with States or PN}
 

### 2023-03-24 resilience.tex
rm

@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ and $\Safe$.
-The  $k$-resilience problem (kRP) is decidable for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. \textcolor{red}{CONJECTURE for now}
+The  $k$-resilience problem (kRP) is decidable 
+%	for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. 
+\textcolor{red}{CONJECTURE for now}
  \end{theorem}
@@
 If $K=\infty$ then $\mathscr{S}$ is not resilient for Safe and Bad.
-Now, if $k \geq min(K,n)$, we conclude that is %uniformely 
-$k$-resilient; 
+Now, if $k \geq min(K,n)$, we conclude that is $k$-resilient; 
 %	if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, 
- $S$ is $k$-resilient.
-%
- \alain{hum...mais il peut y avoir des chemins plus courts}
 %	and then kRP is decidable.
+	If $k <  min(K,n)$, then we can answer the question by checking every paths of length smaller
+	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
 \end{proof}

### 2023-03-24 resilience.tex
up

@@
 resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to a finite number of instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+This infinite set of coverability questions can be reduced to a \emph{finite} number of instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
 
@@
 
-Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ iff $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
 
@@
 %
-This proves that resilience and bounded resilience are equivalent since, here, resilience implies bounded resilience (bounded resilience always implies resilience).
+
+This reduction to a finite number of questions implies that resilience and bounded resilience are equivalent since, here, resilience implies bounded resilience (bounded resilience always implies resilience): if it is resilient then it is $k$-resilient for $k$ the maximum of lengths of covering sequences.
 %
@@
 
-To decide the  bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
+To decide both the resilience and the bounded resilience, we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
 If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,m $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
@@
 
-\begin{theorem}\label{down-up}
+\begin{theorem}\label{up-down}
 \textcolor{red}{

### 2023-03-24 resilience.tex
up

--- 2023-03-24 resilience.tex
@@
 Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $b_j$ is covered from $\downarrow x$.
+From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $\downarrow b_j$ is covered from $\downarrow x$.
 Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 

### 2023-03-24 resilience.tex
details

--- 2023-03-24 resilience.tex
@@
 
-To decide both the resilience and the bounded resilience, we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,m $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
+To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
 \alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve puisqu'on l'utilise ailleurs (théorème suivant) comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
@@
 %	and then kRP is decidable.
-	If $k <  min(K,n)$, then we can answer the question by checking every paths of length smaller
+	If $k <  min(K,n)$, then we check every path of length smaller
 	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.

### 2023-03-24 resilience.tex
details

@@
 \begin{theorem}\label{down-down}
-The resilience problem is undecidable for WSTS with more than one minimal element, 
+The resilience problem is undecidable for WSTS with
+%	with more than one minimal element, 
 $\Safe=\downarrow \Safe$
@@
 \textcolor{red}{
-The resilience problem is undecidable for WSTS with more than one minimal element, 
+The resilience problem is undecidable for WSTS with 
+%	more than one minimal element, 
 $\Safe=\downarrow \Safe$
@@
    BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & ?? & ?? \\ \hline
+      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{k-down-up}) & ?? & ?? \\ \hline
  \end{tabular}

### 2023-03-24 definitions.tex
details

@@
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly well-structured transition
-system, we can decide whether we can reach $U$ from $s$.
+ Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly WSTS, we can decide whether $U  \xrightarrow {*}{} s$.
+%	we can reach $U$ from $s$.
 \end{lemma}
@@
 \begin{theorem}
-A finite basis of $ \pred^*(U)$ is computable for any WSTS $\mathscr{S}$ and any upward closed set $U \subseteq S$.
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
 \end{theorem}

### 2023-03-24 resilience.tex
details
up

@@
 a path of length smaller than $k$) that reaches a safe state.
+$\Safe, \Bad \subseteq S$ and we suppose that $\Safe \cap  \Bad = \emptyset$.
+A usual case is $Safe = S \backslash Bad$ and then Safe is downward-closed iff Bad is upward-closed.
 Let us formalize these properties.
@@
 
+%We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
+%
+
 \problemx{$k$-resilience problem (kRP)}
@@
 
-We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
 \problemx{bounded resilience problem (BRP)}
@@
 \subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
+
+example: ?
 
@@
 
+example: ?
 %
@@
  sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
-The resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
+The resilience problem (RP), the bounded resilience problem (BRP)
+and the $k$-resilience problem (kRP) are decidable.
 \end{theorem}
@@
   bounded resilience problem (BRP) are decidable.
+
+\iffalse
 	\end{proof}
@@
 	\begin{theorem}\label{k-down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ and $\Safe$.
+	Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS 
+	with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ 
+	and $\Safe$.
 	The  $k$-resilience problem (kRP) is decidable 
@@
 	\begin{proof}
+\fi
+
 We begin to compute $K$ and $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
@@
 
+example: ?
+
 % In the case \Safe and \Bad are both downward closed, they can both be finite.
@@
 \begin{proof}
-In the case of a WSTS with at least two minimal elements $0_1$ and $0_2$, the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem, hence its undecidability.  
+In the case of a WSTS with at least two minimal elements $m_1$ and $m_2$, the problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem, hence its undecidability.  
+\mathieu{dévellopper un peu la preuve, $m_2$ minimal de $\Safe$, $m_1$ minimal de $\Bad$...}
 \end{proof}
@@
 %
+example: mutual exclusion.
 

### 2023-03-25 resilience.tex
details

@@
 \begin{theorem}\label{up-up}
-The resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
+The resilience problem is decidable for WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
 Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
@@
 \begin{proposition}
-In WSTS with strong compatibility and effective predbasis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
 \end{proposition}
@@
 \begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective predbasis,
+The bounded resilience problem is decidable for WSTS with effective pred-basis,
 strong compatibility
@@
 The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective predbasis, strong %upward-
+$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
 compatibility, $\Safe=\uparrow \Safe$
@@
    BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{k-down-up}) & ?? & ?? \\ \hline
+      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
  \end{tabular}

### 2023-03-26 definitions.tex
details

@@
 % Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong monotony and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
 $k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
@@
 
-Since $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
+Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
 Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.

### 2023-03-26 resilience.tex
details

@@
 
-The resilience problem (resp.
-the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a safe state.
-$\Safe, \Bad \subseteq S$ and we suppose that $\Safe \cap  \Bad = \emptyset$.
-A usual case is $Safe = S \backslash Bad$ and then Safe is downward-closed iff Bad is upward-closed.
-Let us formalize these properties.
+In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$.
+%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
+The \emph{resilience problem} (resp.
+the \emph{$k$-resilience problem}) is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp.
+a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
+We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp.
+$\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.
+$\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
+In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite basis.
+%	If $\Safe = S \backslash \Bad$, then $\Safe$ is downward-closed iff $\Bad$ is upward-closed. 
+Let us formalize three resilience problems.
 
 \problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
 {$\Bad \longrightarrow^{*} \Safe$ ?\newline}
@@
 \problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
 {$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
@@
 \problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
@@
 
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient.
-est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que \Safe=complement de \Bad~ et donc on aurait un seul ensemble (et son complement)}
-
 These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
@@
 Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
 
@@
 Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is always downward closed and $\Safe$ is always upward closed.
+
+In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
 %		
@@
 %
-In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
+In what follows, we will always suppose that a downward closed set $D=\downarrow D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
-From a similar reasoning as above, the $k$-resilience problem is decidable.\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$, and $\pred^n(\Safe)$ admits a finite basis $B_{\pred^n(\Safe)}$.
+From a similar reasoning as above, the $k$-resilience problem is decidable.
+\alain{écrire une preuve}
+%	\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
 \end{proof}
@@
 \begin{proof}
-Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
+Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$.
 The %uniform 
-resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to a \emph{finite} number of instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+resilience problem (RP) can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
 
@@
 
-Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-
+Let us prove that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+%
 Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
@@
 there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
+
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow b_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
 %
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
+Let us now define $K_{\mathscr{S}}(\Safe,\Bad)=max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+%	(if all $k_{i,j}$ are finite) else $K=\infty$.
+We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe,\Bad)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe,\Bad)$-resilient with $K_{\mathscr{S}}(\Safe,\Bad)$ finite.
 
-This reduction to a finite number of questions implies that resilience and bounded resilience are equivalent since, here, resilience implies bounded resilience (bounded resilience always implies resilience): if it is resilient then it is $k$-resilient for $k$ the maximum of lengths of covering sequences.
+This implies that resilience and bounded resilience are equivalent to coverability.
 %
@@
 %' \in I(\Bad)$.
-
-To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
-\alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve puisqu'on l'utilise ailleurs (théorème suivant) comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
+%
+%To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $\downarrow b_j$ coverable from ideal $J_i$ ?
+If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $\downarrow b_j$ from $J_i$.
+%
  % \alain{comment trouver les $k_a$ ?} 
- by iteratively checking the different possibles values of every $k_{i,j}$ until we find the minimal one,
- and we deduce that  $\mathscr{S}$ is $K$-resilient,
+%
 %  \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ? il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}. 
-else if some of these $n \times m$ coverability questions are negative then resilience
-do not hold and bounded resilience do not hold either.
- Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the 
-  resilience problem (RP) and the  
-  bounded resilience problem (BRP) are decidable.
+%else if some of these $n \times m$ coverability questions are negative then resilience
+%do not hold and bounded resilience do not hold either.
+Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we deduce that both the 
+  resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
 
@@
 	\begin{proof}
-\fi
 
@@
 %	and then kRP is decidable.
-	If $k <  min(K,n)$, then we check every path of length smaller
-	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
+\fi
+
+Let us now show that the $k$-resilience problem (kRP), with $k \in \mathbb{N}$, is also decidable.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
+If $\downarrow b_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow b_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+
+Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
+
+%	If $k <  min(K,n)$, then we check every path of length smaller
+%	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
 \end{proof}
@@
 
+\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
+%
+\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
+
 %We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 

### 2023-03-26 resilience.tex
details

@@
 RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+We deduce that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.
+
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
+All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive \alain{???}.
 
@@
 \begin{theorem}\label{up-up}
-The resilience problem is decidable for WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
 and $\Bad=\uparrow \Bad$.

### 2023-03-27 resilience-safe-down.tex
Splitting resilience

--- 2023-03-25 resilience.tex
@@
-%%%
-\iffalse
-%%%%
-%%%%
-\subsection{Defining resilience}
-
-\subsubsection{Resilience for general transition systems}
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$\Safe$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$\Bad$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
-
-\problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(\Bad → 0≤ j≤k EX j \Safe).
-
-We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
-
-%
-%\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
-%%
-
-%%%
-
-pour $k,k'$ donnés, chercher si $\Safe_{max}$ et $\Bad_{max}$ ont un sens ?
-on peut faire l'union des \Safe, \Safe' et des \Bad, \Bad' prendre le max des k.
-Fixer k, chercher et calculer $\Safe_{max}$ et $\Bad_{max}$.
-
-%
-%\alain{dire juste que \Safe = upward closed et \Bad = downward closed.
-%envisager les 3 autres cas. souvent \Bad = upward closed (par ex l'exclusion mutuelle).
-%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
-
-\fi
-
-\newcommand{\Bad}{\textsf{Bad}}
-\newcommand{\Safe}{\textsf{Safe}}
-
-\section{Resilience for WSTS}
-
-The resilience problem (resp.
-the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp.
-a path of length smaller than $k$) that reaches a safe state.
-$\Safe, \Bad \subseteq S$ and we suppose that $\Safe \cap  \Bad = \emptyset$.
-A usual case is $Safe = S \backslash Bad$ and then Safe is downward-closed iff Bad is upward-closed.
-Let us formalize these properties.
-
-\problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
-%
-%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
-
-%We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
-
-\problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
-
-\problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
- $k$-resilient ?\newline}
-
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient.
-est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?
-on pourrait décider que \Safe=complement de \Bad~ et donc on aurait un seul ensemble (et son complement)}
-
-These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
-So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is always downward closed and $\Safe$ is always upward closed.
-%		
-RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
-%
-In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
-
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
-
-\begin{proposition}[Reformulation]\label{reformulation}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
-\end{proposition}
-
-\iffalse
-
-\begin{proposition}\label{general}
-\textcolor{red}{
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient (CONJECTURE FOR NOW)
-}
-\end{proposition}
-
-\begin{proof}
-Resilient means that ...
-\end{proof}
-
-\fi
-
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
-
-example: ?
-
-\begin{theorem}\label{up-up}
-The resilience problem is decidable for WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-
-\end{theorem}
-
-\begin{proof}
-% Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (\Bad,\Safe)-resilient. 
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
-Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$,
-hence the resilience problem is decidable.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-%	and \Bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^k(\Safe)= \uparrow~\pred^k(\Safe)$ for all $k \in \N$, and $\pred^k(\Safe)$ admits a finite basis $B_{\pred^k(\Safe)}$.
-From a similar reasoning as above, the $k$-resilience problem is decidable.\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
-\end{proof}
-
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}
-
-example: ?
-%
-%		cas étudié par les deux articles sur les graphes
-%
-
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-
-%Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, \Safe and \Bad, are given by 
-%upward-closed or downward-closed sets.
-
-The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
-
-Let us recall two other results in \cite{BFM-ic17}.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-
-\begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two 
-%finite \alain{non, confusion entre ensemble et base}
- sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
-The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable.
-\end{theorem}
-
-\begin{proof}
-Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
-The %uniform 
-resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-This infinite set of coverability questions can be reduced to a \emph{finite} number of instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-%Proposition 30 in \cite{BFM-icalp14} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{*} y'$ in $\mathscr{S}$.
-
-Let us show that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-
-Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $\downarrow b_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-%
-
-This reduction to a finite number of questions implies that resilience and bounded resilience are equivalent since, here, resilience implies bounded resilience (bounded resilience always implies resilience): if it is resilient then it is $k$-resilient for $k$ the maximum of lengths of covering sequences.
-%
-% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
-%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
-%' \in I(\Bad)$.
-
-To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $b_j$ from $J_i$.
-\alain{peut-être definir $K_{ \mathscr{S}}(Safe,Bad)$ en dehors de la preuve puisqu'on l'utilise ailleurs (théorème suivant) comme suit: If all these $n \times m$ coverability questions are positive then  $K(Safe,Bad)=max(k_{i,j})$ else $K=\infty$}
- % \alain{comment trouver les $k_a$ ?} 
- by iteratively checking the different possibles values of every $k_{i,j}$ until we find the minimal one,
- and we deduce that  $\mathscr{S}$ is $K$-resilient,
-%  \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ? il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}. 
-else if some of these $n \times m$ coverability questions are negative then resilience
-do not hold and bounded resilience do not hold either.
- Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the 
-  resilience problem (RP) and the  
-  bounded resilience problem (BRP) are decidable.
-
-\iffalse
-	\end{proof}
-
-	\begin{theorem}\label{k-down-up}
-	Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS 
-	with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ 
-	and $\Safe$.
-	The  $k$-resilience problem (kRP) is decidable 
-	%	for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow 	\pred^*(\Safe)$. 
-	\textcolor{red}{CONJECTURE for now}
-	 \end{theorem}
-
-	\begin{proof}
-\fi
-
-We begin to compute $K$ and $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
-If $K=\infty$ then $\mathscr{S}$ is not resilient for Safe and Bad.
-Now, if $k \geq min(K,n)$, we conclude that is $k$-resilient; 
-%	if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, 
-%	and then kRP is decidable.
-	If $k <  min(K,n)$, then we check every path of length smaller
-	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
-\end{proof}
-
-%
-
 \subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
@@
 \end{proof}
-
-\subsection{Case Synthesis}
-
-\mathieu{En rouge quand il s'agit de conjectures}
-
-\begin{center}
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
-   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
- \end{tabular}
-\end{center}
-
-%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-% \textcolor{red}{Seems like a reasonable assumption to me.}
-
-%From these considerations, we formulate instances of the abstract resilience problems for well-
-%structured transition systems.

### 2023-03-27 resilience-safe-up.tex
Splitting resilience

--- 2023-03-26 resilience.tex
@@
-%%%
-\iffalse
-%%%%
-%%%%
-\subsection{Defining resilience}
-
-\subsubsection{Resilience for general transition systems}
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$\Safe$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$\Bad$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
-
-\problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(\Bad → 0≤ j≤k EX j \Safe).
-
-We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
-
-%
-%\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
-%%
-
-%%%
-
-pour $k,k'$ donnés, chercher si $\Safe_{max}$ et $\Bad_{max}$ ont un sens ?
-on peut faire l'union des \Safe, \Safe' et des \Bad, \Bad' prendre le max des k.
-Fixer k, chercher et calculer $\Safe_{max}$ et $\Bad_{max}$.
-
-%
-%\alain{dire juste que \Safe = upward closed et \Bad = downward closed.
-%envisager les 3 autres cas. souvent \Bad = upward closed (par ex l'exclusion mutuelle).
-%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
-
-\fi
-
-\newcommand{\Bad}{\textsf{Bad}}
-\newcommand{\Safe}{\textsf{Safe}}
-
-\section{Resilience for WSTS}
-
-In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$.
-%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
-The \emph{resilience problem} (resp.
-the \emph{$k$-resilience problem}) is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp.
-a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
-We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp.
-$\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.
-$\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
-In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite basis.
-%	If $\Safe = S \backslash \Bad$, then $\Safe$ is downward-closed iff $\Bad$ is upward-closed. 
-Let us formalize three resilience problems.
-
-\problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
-%
-%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
-
-%We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
-
-\problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
-
-\problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
- $k$-resilient ?\newline}
-
-These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
-So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
-
-For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
-
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
-%		
-RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-We deduce that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.
-
-In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
-%
-In what follows, we will always suppose that a downward closed set $D=\downarrow D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
-
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive \alain{???}.
-
-\begin{proposition}[Reformulation]\label{reformulation}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
-\end{proposition}
-
-\iffalse
-
-\begin{proposition}\label{general}
-\textcolor{red}{
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient (CONJECTURE FOR NOW)
-}
-\end{proposition}
-
-\begin{proof}
-Resilient means that ...
-\end{proof}
-
-\fi
-
 \subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
@@
 %
-
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
-
-example: ?
-
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
-
-\begin{theorem}\label{down-down}
-The resilience problem is undecidable for WSTS with
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-\begin{proof}
-In the case of a WSTS with at least two minimal elements $m_1$ and $m_2$, the problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem, hence its undecidability.  
-\mathieu{dévellopper un peu la preuve, $m_2$ minimal de $\Safe$, $m_1$ minimal de $\Bad$...}
-\end{proof}
-
-In the case of a WSTS with only one minimal element,  
-% (ex: Petri Nets and their «Zero») 
-then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
-
-%%%%%%%%
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
-%
-%		cas exclusion mutuelle
-%
-example: mutual exclusion.
-
-\begin{theorem}\label{up-down}
-\textcolor{red}{
-The resilience problem is undecidable for WSTS with 
-%	more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
-}
-\end{theorem}
-
-\begin{proof}
-TBD
-\end{proof}
-
-\subsection{Case Synthesis}
-
-\mathieu{En rouge quand il s'agit de conjectures}
-
-\begin{center}
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
-   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
- \end{tabular}
-\end{center}
-
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
-%
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-% \textcolor{red}{Seems like a reasonable assumption to me.}
-
-%From these considerations, we formulate instances of the abstract resilience problems for well-
-%structured transition systems.

### 2023-03-27 resilience.tex
Splitting resilience

@@
 
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
+\input{resilience-safe-up}
 
-example: ?
-
-\begin{theorem}\label{up-up}
-The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-
-\end{theorem}
-
-\begin{proof}
-% Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (\Bad,\Safe)-resilient. 
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
-Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$,
-hence the resilience problem is decidable.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
-%	and \Bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$, and $\pred^n(\Safe)$ admits a finite basis $B_{\pred^n(\Safe)}$.
-From a similar reasoning as above, the $k$-resilience problem is decidable.
-\alain{écrire une preuve}
-%	\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
-\end{proof}
-
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}
-
-example: ?
-%
-%		cas étudié par les deux articles sur les graphes
-%
-
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-
-%Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, \Safe and \Bad, are given by 
-%upward-closed or downward-closed sets.
-
-The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
-
-Let us recall two other results in \cite{BFM-ic17}.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-
-\begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two 
-%finite \alain{non, confusion entre ensemble et base}
- sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
-The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable.
-\end{theorem}
-
-\begin{proof}
-Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$.
-The %uniform 
-resilience problem (RP) can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
-Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-%Proposition 30 in \cite{BFM-icalp14} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{*} y'$ in $\mathscr{S}$.
-
-Let us prove that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-%
-Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $\downarrow b_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-
-Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow b_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
-%
-Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
-Let us now define $K_{\mathscr{S}}(\Safe,\Bad)=max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
-%	(if all $k_{i,j}$ are finite) else $K=\infty$.
-We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe,\Bad)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe,\Bad)$-resilient with $K_{\mathscr{S}}(\Safe,\Bad)$ finite.
-
-This implies that resilience and bounded resilience are equivalent to coverability.
-%
-% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
-%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
-%' \in I(\Bad)$.
-%
-%To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $\downarrow b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $\downarrow b_j$ from $J_i$.
-%
- % \alain{comment trouver les $k_a$ ?} 
-%
-%  \alain{else $\mathscr{S}$ is not $K$-resilient...et alors qu'en déduit-on ? il pourrait exister un $K' \geq K$ pour lequel $\mathscr{S}$ est K'-resilient....}. 
-%else if some of these $n \times m$ coverability questions are negative then resilience
-%do not hold and bounded resilience do not hold either.
-Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we deduce that both the 
-  resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
-
-\iffalse
-	\end{proof}
-
-	\begin{theorem}\label{k-down-up}
-	Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS 
-	with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ 
-	and $\Safe$.
-	The  $k$-resilience problem (kRP) is decidable 
-	%	for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow 	\pred^*(\Safe)$. 
-	\textcolor{red}{CONJECTURE for now}
-	 \end{theorem}
-
-	\begin{proof}
-
-We begin to compute $K$ and $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
-If $K=\infty$ then $\mathscr{S}$ is not resilient for Safe and Bad.
-Now, if $k \geq min(K,n)$, we conclude that is $k$-resilient; 
-%	if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, 
-%	and then kRP is decidable.
-\fi
-
-Let us now show that the $k$-resilience problem (kRP), with $k \in \mathbb{N}$, is also decidable.
-Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
-If $\downarrow b_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow b_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
-
-Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
-
-%	If $k <  min(K,n)$, then we check every path of length smaller
-%	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
-\end{proof}
-
-%
-
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
-
-example: ?
-
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
-
-\begin{theorem}\label{down-down}
-The resilience problem is undecidable for WSTS with
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-\begin{proof}
-In the case of a WSTS with at least two minimal elements $m_1$ and $m_2$, the problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem, hence its undecidability.  
-\mathieu{dévellopper un peu la preuve, $m_2$ minimal de $\Safe$, $m_1$ minimal de $\Bad$...}
-\end{proof}
-
-In the case of a WSTS with only one minimal element,  
-% (ex: Petri Nets and their «Zero») 
-then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
-
-%%%%%%%%
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
-%
-%		cas exclusion mutuelle
-%
-example: mutual exclusion.
-
-\begin{theorem}\label{up-down}
-\textcolor{red}{
-The resilience problem is undecidable for WSTS with 
-%	more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
-}
-\end{theorem}
-
-\begin{proof}
-TBD
-\end{proof}
+\input{resilience-safe-down}
 

### 2023-03-27 resilience-safe-up.tex
up

@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$, and $\pred^n(\Safe)$ admits a finite basis $B_{\pred^n(\Safe)}$.
-From a similar reasoning as above, the $k$-resilience problem is decidable.
-\alain{écrire une preuve}
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$, and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$. 
+% From a similar reasoning as above, the $k$-resilience problem is decidable.
+% \alain{écrire une preuve}
 %	\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
+Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
+Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$,
+hence the resilience problem is decidable.
 \end{proof}

### 2023-03-27 resilience.tex
details

@@
 Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive \alain{???}.
+All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.
+%	 and ideals are recursive 
+\alain{???}.
 

### 2023-03-27 reachable-resilience.tex
up
Downward reachability reduction to reachable resilience

@@
 
-\problemx{Rachable-resilience problem for WSTS (RRP)}
+\problemx{Reachable-resilience problem for WSTS (RRP)}
 {A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
@@
 \begin{proposition}\label{reductions}
-Coverability and $t$-liveness are reduceable to the reachable resilience problem.
+Coverability and $t$-liveness are reducible to the reachable resilience problem.
 %	home state and reachability
@@
 
+\end{proof}
+
+\problemx{Downward reachability}
+{A state $s$, a downward-closed set $D$.}
+{$s \to^* D$? \newline}
+
+\begin{theorem}
+The downward reachability problem is reducible to the reachable resilience problem.
+\end{theorem}
+
+\begin{proof}
+Take $\Safe = \emptyset$ and $\Bad = D$.
+The reachable resilience problem then asks
+whether $post^*(s) \cap D \subseteq \emptyset$.
+If the answer is negative then $s \to^* D$,
+else $s \not\to^* D$.
 \end{proof}
@@
 
-\begin{proof}{Sketch}
+\begin{proof}[Sketch]
 Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 

### 2023-03-28 definitions.tex
details

@@
 
+In what follows, a downward closed set $D=\downarrow D$ is represented by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is represented by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
+
+Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
+All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.
+
 % \textcolor{red}{Define WSTS}

### 2023-03-28 resilience.tex
details

@@
 %
-In what follows, we will always suppose that a downward closed set $D=\downarrow D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
 
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.
 %	 and ideals are recursive 

### 2023-03-28 resilience-safe-down.tex
Brouillon 3point4

@@
 \end{proof}
+
+\begin{theorem}
+The resilience problem is decidable for WSTS with 
+%	more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$
+and
+\textcolor{red}{
+additional hypothesis.
+}
+\end{theorem}
+
+\begin{proof}
+Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Safe$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
+
+Let's assume for now for simplicity's sake that $\Bad = \uparrow \{b\}$
+
+As an hypothesis, we assume that $\pred^*(\Safe)$ is downward closed as well.
+And let $\{J'_1, J'_2,...,J'_n\}$ be the ideal decomposition of $\pred^*(\Safe)$.
+
+Resilience asks whether 
+$ \uparrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$.
+
+Of note is that $ \downarrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$
+if and only if
+there exists some $i$ such that
+$\downarrow \{b\} \subseteq J'_i$.
+
+\mathieu{Try and find a condition under which "$ \uparrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$?" is decidable and/or find an undecidable problem that reduce to a variation on 
+"$\uparrow \{b\} \subseteq J'_i$ ?"}
+\end{proof}

### 2023-03-28 definitions.tex
details

@@
+\newcommand{\pred}{\textsf{pred}}
+\newcommand{\post}{\textsf{post}}
+% \renewcommand{\succ}{\textsf{Succ}}
+
 \section{Definitions}\label{section definitions}
@@
 
-The model we are interested in is (S)WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
+The model we are interested in is WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
 
@@
 \begin{definition}
-A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Lambda, \rightarrow)$ where 
-$S$ is a set of {\em states}, $ \Lambda$ is a set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
+An {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
+{\em states} and  
+$ {\rightarrow} \subseteq S \times S$ is a
+binary relation 
+on
+the set of states, denoted as the set of {\em transitions}. A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
+$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
+${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
 ternary relation,
@@
 
-We
- prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
-       $s  \xrightarrow{a}  s'$
-to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
-
-\noindent
-Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
-
-A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
-to a {\em target configuration} $s_n$ is a sequence 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
-We define the {\em concatenation} $ \pi_1 \pi_2$ of 
-two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
-as expected.
-The {\em length} of 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
-is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
-For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
-
-An {\em infinite path} is an infinite sequence
- $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
-For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
-(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
-and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
-by $\pi[i,j]$ the path 
-$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
- and by $\pi[i]$ the configuration $s_i$.
-As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
-a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
-Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
-$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
-
-\newcommand{\pred}{\textsf{pred}}
-\newcommand{\post}{\textsf{post}}
-% \renewcommand{\succ}{\textsf{Succ}}
-
-\noindent
-The set of {\em successors} of a configuration $s \in S$ is deﬁned as
-$\post(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
-A configuration without successors is called a {\em dead end}.  \\
-The set of {\em predecessor} of a configuration $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid \exists a \in \Lambda ~ s' \xrightarrow{a} s\}$.
-
-A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
-$a \in \Lambda$,
- $ s_1  \xrightarrow{a} s_{2}$ and  
- $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
-
-\begin{definition}
-An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
-{\em states} and  
-$ {\rightarrow} \subseteq S \times S$ is a
-binary relation 
-on
-the set of states, denoted as the set of {\em transitions}. 
-\end{definition}
+Note 
+ that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.
 
 We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
-Note 
- that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
+
+%
+\alain{enlever ce qui est inutile}
+%
 Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
@@
 
+%
+%\begin{definition}
+%A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
+%$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
+%${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
+%ternary relation,
+%denoted as the set of {\em labeled transitions}. 
+%\end{definition}
+
+We
+ prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
+       $s  \xrightarrow{a}  s'$
+to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
+
+\noindent
+%Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
+
+A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
+to a {\em target configuration} $s_n$ is a sequence 
+$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$.
+
+We define the {\em concatenation} $ \pi_1 \pi_2$ of 
+two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
+as expected.
+The {\em length} of 
+$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
+is defined as $|\pi|=n$ and we say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
+For all $w \in \Sigma^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
+
+An {\em infinite path} is an infinite sequence
+ $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
+For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
+(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
+and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
+by $\pi[i,j]$ the path 
+$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
+ and by $\pi[i]$ the configuration $s_i$.
+As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
+a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
+Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
+$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
+
+\noindent
+The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as
+$\post(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s \xrightarrow{a} s'\}$.
+A state without successors is called a {\em dead end}.  \\
+The set of {\em (immediate) predecessor} of a state $s \in S$ is deﬁned as
+$\pred(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s' \xrightarrow{a} s\}$.
+
+A labeled transition system $(S, \Sigma, \rightarrow)$ is {\em deterministic} if for all states $s_1, s_2, s_3 \in S$ and all
+$a \in \Sigma$,
+ $ s_1  \xrightarrow{a} s_{2}$ and  
+ $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
+
+%
+\alain{enlever ce qui est inutile}
+%
+
 \subsection{Well-quasi-orderings}
 
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$.
+A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$.
 
@@
 \begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $\leq$ (over some set $X$ ) such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
+ A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
 $x_i \leq  x_j$.
@@
 
-Notice that a wqo is well-founded, i.e. it admits no infinite strictly decreasing sequence
-$x_0 > x_1 > x_2 > \cdots$
+There exists a stronger formulation.
+$(X,\leq)$ is a wqo iff any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
 
 % \textcolor{red}{Add lemma about infinite increasing subsequences ?}
-
+%
+\iffalse
+%
 \begin{lemma}
@@
 
-\iffalse
 \begin{proof}
@@
 \end{proof}
+%
 \fi
+%
 
-Given $\leq$ a quasi-ordering over some set $X$, an {\em upward-closed set} is any set $I \subseteq X$ such that if $y \geq x$ and $x \in I$ then $y \in I $.
-A {\em downward-closed set} is any set $I \subseteq X$ such that if $y \leq x$ and $x \in I$ then $y \in I $. 
-To any $A \subseteq X$ we associate
-the {\em upward-closure of $A$} 
+Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
+A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+To any subset $A \subseteq X$, we may associate
+its {\em upward-closure},
  $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
- and the 
- {\em downward-closure of $A$} 
+ and its 
+ {\em downward-closure},
  $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
@@
 
-A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \bigcup_{x \in I_b} \uparrow x$. 
+A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$. 
 % Higman investigated ordered sets with the finite basis property.
+
+Wqo admits many other equivalent formulations. For instance: 
+$(X,\leq)$ is a wqo if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
+upward-closed subsets in $X$ eventually stabilizes, 
+%      i.e. there is a $k \in N$ such that $U_k = U_{k+1} = U_{k+2} = \cdots $ 
+if and only if any upward-closed subset $U \subseteq X$ has a
+finite basis if and only if $(X,\leq)$ is well-founded (i.e.
+it admits no infinite strictly decreasing sequence) and without infinite antichains (i.e.
+there is no infinite subset in $X$ of mutually incomparable elements).
+%
+\iffalse
+%
 
@@
 \end{proof}
+%
+\fi
+%
 

### 2023-03-28 definitions.tex
details

@@
 if and only if any upward-closed subset $U \subseteq X$ has a
-finite basis if and only if $(X,\leq)$ is well-founded (i.e.
-it admits no infinite strictly decreasing sequence) and without infinite antichains (i.e.
-there is no infinite subset in $X$ of mutually incomparable elements).
+finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
 %
@@
 
-In what follows, a downward closed set $D=\downarrow D$ is represented by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is represented by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.
+Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo.
+In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_i$ (an ideal is a directed downward closed subset): $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
 
-Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.
+In what follows, a downward closed set is represented by its finite set of ideals 
+%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
+and an upward closed set $U$ is represented by its finite set minimal elements.
+% $M_U$ of its  satisfying $U=\uparrow M_U$.
+
+\alain{est-ce utile et correct ?
+Let $I,J$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
+All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.}
 

### 2023-03-28 definitions.tex
details

@@
 Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo.
-In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_i$ (an ideal is a directed downward closed subset): $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
+In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
 
@@
 %		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
-and an upward closed set $U$ is represented by its finite set minimal elements.
+and an upward closed set $U$ is represented by its finite set of minimal elements.
 % $M_U$ of its  satisfying $U=\uparrow M_U$.
@@
 \begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
-A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
+A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
 is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
+equipped with a wqo ${\leq} \subseteq S \times S$ such that  
 %  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
+the transition relation $ \rightarrow$ is compatible with $\leq$, i.e., for all 
 $s_1, t_1 , s_2 \in S$
-	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$ , there exists 
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
 	$t_2 \in S$ with 
-	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$ 			
-				(resp. $s_2 \rightarrow t_2$ ).
+	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
 \end{definition}
+
+We say that a WSTS $\mathscr{S}$ has the \emph{strong compatibility} when moreover for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+	$t_2 \in S$ with 
+	$t_1 \leq t_2$ and $s_2 \rightarrow t_2$.
 
@@
 \begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ and $\pred^*(U )$ are upward-closed.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed.
 \end{proposition}
@@
 
+%
+\iffalse
+%
 \begin{claim}{(stability of upward-closed sets)}
@@
 \end{claim}
+%
+\fi
+%
 
@@
 
-\begin{fact}\label{fact basis}
+%			\begin{fact}\label{fact basis}
 % Fact 3 ([1]). 
 % Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$. 
+
+%For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$.
+
 % (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
 % \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
-\end{fact}
+%			\end{fact}
 
+For a WSTS $\mathscr{S}$ and an upward closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$.
+When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)$ converges.
+Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
+$U_k = U_{k_0}$ for all $k \geq k_0$.
+We then have:  $\pred^*(U) = U_{k_0}$.
+
+\alain{ me semble maintenant inutile
+%
 \begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U^k= \bigcup_{0 \leq j \leq k} \pred^j(U)$.
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
 The {\em index} $k(U)$ is the
-smallest $k_0$ s.t. $U^k = U^{k_0}$ for all $k \geq k_0$.
+smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
 \end{definition}
-
+%
+%
 If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
 every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
-
+%
 \begin{remark}
@@
 Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
-\end{proof}
+\end{proof}}
 
-\begin{theorem}
-A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
+\begin{theorem}(ref)
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $pb(U)$.
 \end{theorem}

### 2023-03-28 definitions.tex
details

@@
 
-\section{Definitions}\label{section definitions}
+\section{Well-structured transition systems}\label{section definitions}
 
@@
 on
-the set of states, denoted as the set of {\em transitions}. A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
-$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
-ternary relation,
-denoted as the set of {\em labeled transitions}. 
+the set of states, denoted as the set of {\em transitions}.
+
+%A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
+%$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
+%${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
+%ternary relation,
+%denoted as the set of {\em labeled transitions}. 
 \end{definition}
 
-Note 
- that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.
+%Note 
+% that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.
 
-We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
+We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
+We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
+for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
+Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
+We denote $X \longrightarrow^{*} Y$ (resp.
+$X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
+of length smaller than $k$) that reaches a state $y \in Y$.
+\noindent
+The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$.
+A state without successors is called a {\em dead end}.
+The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
+$\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$.
 
@@
 \alain{enlever ce qui est inutile}
+%
+We use similar notation for $\post$ and $\pred$...
 %
@@
 %\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
-
-We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
-for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...
-
-Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
-We denote $X \longrightarrow^{*} Y$ (resp.
-$X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
-of length smaller than $k$) that reaches a state $y \in Y$.
 
@@
 
-\noindent
-The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as
-$\post(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s \xrightarrow{a} s'\}$.
-A state without successors is called a {\em dead end}.  \\
-The set of {\em (immediate) predecessor} of a state $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s' \xrightarrow{a} s\}$.
-
 A labeled transition system $(S, \Sigma, \rightarrow)$ is {\em deterministic} if for all states $s_1, s_2, s_3 \in S$ and all
@@
 We abbreviate $x \leq y \not\leq x$ by $x < y$.
-
 % A partial ordering (a po) is an antisymmetric qo
-
-Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
+\alain{Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).}
 % and gives rise to a po between the equivalence classes ?
@@
 \begin{definition}
- A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
-$x_i \leq  x_j$.
+ A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
 \end{definition}

### 2023-03-29 resilience-safe-up.tex
details
Adjusting 3.1

@@
 \subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
+
+\alain{2 clos par le haut ont une intersection non vide....}
 
@@
 The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
+and $\Bad=\uparrow \Bad \setminus \Safe$.
 
@@
 \begin{proof}
-% Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is (\Bad,\Safe)-resilient. 
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{\uparrow \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
 Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
-Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$,
+Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\Bad}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = \uparrow \Bad \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
 hence the resilience problem is decidable.

### 2023-03-29 resilience-safe-up.tex
Adjusting 3.1
up

@@
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad \setminus \Safe$.}
 
@@
 \begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{\uparrow \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
 Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
 Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\Bad}$.
+Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\uparrow \Bad}$.
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.

### 2023-03-29 resilience-safe-up.tex
up

@@
 compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
+and $\Bad=\uparrow \Bad \setminus \Safe$.
 %	and \Bad is upward closed or downward closed.
@@
 \begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$, and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$. 
-% From a similar reasoning as above, the $k$-resilience problem is decidable.
-% \alain{écrire une preuve}
-%	\alain{réserver k pour le k-résilience, il y a d'autres lettres...n, m, p,...}
-Since $\Bad$  is upward closed, there exists a finite basis $B_{\Bad}$ such that $\Bad = \uparrow B_{\Bad}$.
-Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$,
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
+Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\uparrow \Bad}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = \uparrow \Bad \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
 hence the resilience problem is decidable.

### 2023-03-29 resilience-safe-up.tex
up

@@
  $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
+and $\Bad=\uparrow \Bad \setminus \Safe$.
 \end{corollary}

### 2023-03-29 resilience.tex
details

@@
 The \emph{resilience problem} (resp.
-the \emph{$k$-resilience problem}) is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp.
+the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe,\Bad)$ is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp.
 a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
@@
 $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
-In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite basis.
+In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
 %	If $\Safe = S \backslash \Bad$, then $\Safe$ is downward-closed iff $\Bad$ is upward-closed. 
@@
 
-These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
+\alain{dessous: à dire mais pas sous forme de proposition}
+\begin{proposition}[Reformulation]\label{reformulation}
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
+$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
+\end{proposition}
+
+These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
+Since most of decidable properties in WSTS rely on the computation of upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
 
@@
 \alain{???}.
-
-\begin{proposition}[Reformulation]\label{reformulation}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
-\end{proposition}
 

### 2023-03-29 resilience-safe-down.tex
details

@@
 }
+\alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ?
+si l'ordre est omega2 alors closbas+inclusion est wqo donc il faudrait juste que la suite croissante $pred^n(D)+le reste$ converge...}
 \end{theorem}
@@
 
-Let's assume for now for simplicity's sake that $\Bad = \uparrow \{b\}$
+Let's assume for now for simplicity's sake that $\Bad = \uparrow b$
 
@@
 Resilience asks whether 
-$ \uparrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$.
+$ \uparrow b \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$.
 
-Of note is that $ \downarrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$
+Of note is that $ \downarrow b \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$
 if and only if
 there exists some $i$ such that
-$\downarrow \{b\} \subseteq J'_i$.
+$\downarrow b \subseteq J'_i$.
 

### 2023-03-30 resilience-safe-up.tex
\Bad = up \Bad \ßetminus \Safe replaced with an U, small parazgraph about motivation need to be expanded upon slightly

--- 2023-03-29 resilience-safe-up.tex
@@
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad \setminus \Safe$.}
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad = ... $.}
 
 \alain{2 clos par le haut ont une intersection non vide....}
+
+Since we impose $\Bad \cap \Safe = \emptyset$, we cannot have as an hypothesis
+both $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$. To approximate a case
+of double upward closure, we consider a case where $\Bad$ is as upward-closed
+as it can be without intersecting with $\Safe$,
+i.e. 
+such that there exists an upward closed set $U$
+such that $\Bad = U \setminus \Safe$.
+Morally speaking ... [].
 
@@
 The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad \setminus \Safe$.
+and $\Bad= U \setminus \Safe$.
 
@@
 Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\uparrow \Bad}$.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = \uparrow \Bad \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
 % si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
@@
  $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad$.
+and $\Bad=U \setminus \Safe$.
 \end{corollary}
@@
 compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=\uparrow \Bad \setminus \Safe$.
+and $\Bad=U \setminus \Safe$.
 %	and \Bad is upward closed or downward closed.
@@
 % Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $\uparrow \Bad$  is upward closed, there exists a finite basis $B_{\uparrow \Bad}$ such that $\uparrow \Bad = \uparrow B_{\uparrow \Bad}$.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = \uparrow \Bad \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
 % si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe

### 2023-03-30 resilience-safe-up.tex
up

@@
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
 % si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
@@
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
 % si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe

### 2023-03-30 resilience-safe-down.tex
Currently writing proof

--- 2023-03-27 resilience-safe-down.tex
@@
+%%%%%%%%
+\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
+%
+%		cas exclusion mutuelle
+%
+example: mutual exclusion.
+
+\begin{theorem}
+The resilience problem is decidable for ideal-effective WSTS with 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$
+and
+the additional hypothesis that
+for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed.
+% \alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ?
+si l'ordre est omega2 alors closbas+inclusion est wqo donc il faudrait juste que la suite croissante $pred^n(D)+le reste$ converge...}
+\end{theorem}
+
+\begin{proof}
+By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
+Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
+The resilience problem can be reformulated as 
+$\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed subsets.
+Hence we can compute the intersection of
+$\uparrow \{b_1,b_2,...,b_m\}$
+and
+$S \setminus \pred^*(\Safe)$.
+Since
+$\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
+can be reformulated as
+$\uparrow \{b_1,b_2,...,b_m\} \cap (S \setminus \pred^*(\Safe)) = \emptyset$
+the resilience problem is decidable.
+\end{proof}
+
+\begin{corollary}
+The resilience problem is decidable for ideal-effective downward-compatible WSTS with 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$.
+\end{corollary}
+
+\begin{proof}
+
+\end{proof}
+
+% \begin{proposition}
+In the case
+
 \subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
@@
 then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
-
-%%%%%%%%
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
-%
-%		cas exclusion mutuelle
-%
-example: mutual exclusion.
-
-\begin{theorem}\label{up-down}
-\textcolor{red}{
-The resilience problem is undecidable for WSTS with 
-%	more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
-}
-\end{theorem}
-
-\begin{proof}
-TBD
-\end{proof}

### 2023-03-30 resilience-safe-down.tex
up

@@
 \begin{proof}
+Downward compatibility of $\mathscr{S}=(S,\rightarrow, \leq)$ 
+means that
+for all $s_1, s_2, t_1 \in S$ with
+$s_2 \leq s_1$ and $s_1 \to t_1$
+there
+exists
+$t_2 \leq t_1$ such that $s_2 \to^* t_2$.
 
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \downarrow \pred^*(D)$.
+By design there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$
+and $d \in D$ such that
+$y \to a_1 \to a_2 \to \cdots \to a_m \to d$.
+
+By downward compatibility
+
+$y \to a_1$
+
+implies
+
+$x \to^* a'_1$
+
+with 
+$a'_1 \leq a_1$.
+
+Then
+
+$a_i \to a_{i+1}$
+
+implies the existence of $a'_{i+1} \leq a_{i+1}$ such that
+
+$a'_i \to^* a'_{i+1}$,
+
+and by induction
+
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* d'$
+
+where
+
+$d' \leq d$ belongs to $D$ by downward closure of $D$.
+
+Hence $x \in \pred^*(D)$.
 \end{proof}

### 2023-03-30 resilience-safe-down.tex
up

--- 2023-03-30 resilience-safe-down.tex
@@
 By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
-Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
+%  Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
 The resilience problem can be reformulated as 
-$\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
+% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
+$\Bad \subseteq  \pred^*(\Safe)$.
 Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed subsets.
 Hence we can compute the intersection of
-$\uparrow \{b_1,b_2,...,b_m\}$
+% $\uparrow \{b_1,b_2,...,b_m\}$
+$\Bad$
 and
-$S \setminus \pred^*(\Safe)$.
+$S \setminus \pred^*(\Safe)$,
+which are both upward-closed.
 Since
-$\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
+% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
+$\Bad \subseteq \pred^*(\Safe)$
 can be reformulated as
-$\uparrow \{b_1,b_2,...,b_m\} \cap (S \setminus \pred^*(\Safe)) = \emptyset$
+$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
 the resilience problem is decidable.
@@
 \begin{proof}
+Downward compatibility of $\mathscr{S}=(S,\rightarrow, \leq)$ 
+means that
+for all $s_1, s_2, t_1 \in S$ with
+$s_2 \leq s_1$ and $s_1 \to t_1$
+there
+exists
+$t_2 \leq t_1$ such that $s_2 \to^* t_2$.
 
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \downarrow \pred^*(D)$.
+By design there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$
+and $d \in D$ such that
+$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+
+By downward compatibility $a_0 \to a_1$
+implies there exists $a'_1 \leq a_1$ with
+$x \to^* a'_1$.
+More generally $a_i \to a_{i+1}$ and
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \leq a_{i+1}$ such that
+$a'_i \to^* a'_{i+1}$,
+and, by induction,
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* d'$
+where
+$d' \leq d$ belongs to $D$ by downward closure of $D$.
+Hence $x \in \pred^*(D)$.
 \end{proof}
@@
 In the case
+of a ideal-effective WSTS 
+where
+the additional hypothesis that
+for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed
+is not met,
+the above construction
+can provide a proof
+of non-resilience
+i.e. when
+$\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) \neq \emptyset$
+then
+$\Bad \not\subseteq \downarrow\pred^*(\Safe)$
+and hence
+$\Bad \not\subseteq \pred^*(\Safe)$.
+When $\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) = \emptyset$
+however
+it is not enough to conclude.
 

### 2023-03-30 resilience-safe-down.tex
up

@@
 \begin{theorem}\label{down-down}
-The resilience problem is undecidable for WSTS with
+The resilience problem is undecidable for strong effective WSTS with
 %	with more than one minimal element, 
@@
 \begin{proof}
-In the case of a WSTS with at least two minimal elements $m_1$ and $m_2$, the problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem, hence its undecidability.  
-\mathieu{dévellopper un peu la preuve, $m_2$ minimal de $\Safe$, $m_1$ minimal de $\Bad$...}
+Let us consider the case of a WSTS with at least two minimal elements $m_1$ and $m_2$.
+The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
+By undecidability of the reachability problem we conclude.
 \end{proof}
@@
 In the case of a WSTS with only one minimal element,  
-% (ex: Petri Nets and their «Zero») 
-then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
+then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.

### 2023-03-30 resilience-safe-down.tex
up

@@
 there
-exists
-$t_2 \leq t_1$ such that $s_2 \to^* t_2$.
+exists $t_2 \in S$ with
+$t_2 \leq t_1$ and $s_2 \to^* t_2$.
 

### 2023-03-30 resilience-safe-down.tex
up

@@
 
-\begin{theorem}
+\begin{theorem}\label{up-down}
 The resilience problem is decidable for ideal-effective WSTS with 

### 2023-03-30 resilience.tex
up

@@
 
-\mathieu{En rouge quand il s'agit de conjectures}
+% \mathieu{En rouge quand il s'agit de conjectures}
 
@@
 \hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
+   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & Decidable (Theorem~\ref{up-down}) \\ \hline
    BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline

### 2023-03-30 resilience.tex
up

@@
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & Decidable (Theorem~\ref{up-down}) \\ \hline
+\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
+   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up})  & Decidable (Theorem~\ref{up-down}) & Undecidable (Theorem~\ref{down-down}) \\ \hline
    BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline

### 2023-03-30 reachable-resilience.tex
details

--- 2023-03-24 reachable-resilience.tex
@@
-\section{Resilience from a single-state for WSTS}
+\section{State resilience}
 
 Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap \Bad$, there must exist a path to $\Safe$.
-However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
+But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap post^*(s_0)$, there must exist a path to $\Safe$.
+%
+%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
 %
@@
 
-Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
+%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
 
-\problemx{$k$-reachable-resilience problem for WSTS (kRRP)}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
+The three previous problems become:
+
+\problemx{State-resilience problem (SRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
+% 		an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
+% 		{$ ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe, ~ s' \rightarrow^{*} s''$ ?\newline}
+{$\Bad \cap \post^*(s_0)  \rightarrow^{*} \Safe $ ? \newline}
+
+\problemx{$k$-state-resilience problem (kSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
 { $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
@@
 
-Analogously, we formulate bounded and general variations of the problem.
-
-\problemx{bounded-reachable-resilience problem for WSTS (BRRP)}
-{A state $s$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
+\problemx{bounded-state-resilience problem (BSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-reachable-resilient. ?\newline}
-
-\problemx{Rachable-resilience problem for WSTS (RRP)}
-{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
-{$ ~ \forall s' \in \Bad ~ (s \rightarrow^* s') \implies \exists s'' \in \Safe ~ s' \rightarrow^{*} s''$ ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient. ?\newline}
 
@@
 
+%
 \iffalse
-
+%
 \textcolor{blue}{
@@
 }
-
+%
 \textcolor{blue}{
@@
 }
-
+%
 \fi
 
+Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
+
+\alain{mal placé}
 \begin{proposition}\label{reductions}
-Coverability and $t$-liveness are reduceable to the reachable resilience problem.
+In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
 %	home state and reachability
@@
 \item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ 
+$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
 where
@@
 
+\problemx{Downward reachability problem}
+{A state $s$, a downward-closed set $D$.}
+{$s \to^* D$? \newline}
+
+\begin{theorem}
+The downward reachability problem is reducible to the state resilience problem.
+\end{theorem}
+
+\begin{proof}
+Let $s \in S$, and a downward-closed set $D$.
+Take $\Safe = \emptyset$ and $\Bad = D$.
+The state resilience problem on $\Safe,\Bad$ asks
+whether $\Bad \cap \post^*(s) \subseteq \emptyset$.
+If the answer is negative then $s \to^* \Bad$,
+else $s \not\to^* \Bad$.
+\end{proof}
+
+\alain{mal placé}
+
 % Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
@@
 
+Let us recall a result about state-resilience (called resilieence in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded-reachable-resilience} and {\sc $k$-reachable-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*$ is computable.
+{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
 \end{theorem}
@@
 \begin{corollary}\label{postcomputable}
-{\sc reachable-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*$ is computable.
+{\sc state-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
 \end{corollary}
+
+\alain{definir downward reachability problem.....
+downward-closed problem given a state $s$ of a WSTS
+% in the regarded class 
+with strong upward compatibility 
+and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
 
@@
 From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\pred^*(\Safe) =  \Safe^{k_m}$. We can compute this $k_m$ by iteratively computing
+$\pred^*(\Safe) =  \Safe^{k_m}$. We can compute this $k_m$ by iteratively computing \alain{mauvaises notations: $k, k_m, Safe^k,...$ n'a aucun sens...}
 $\Safe^{k+1}$ from $\Safe^k$, checking $\Safe^{k+1}=\Safe^k$, 
@@
 
+What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
 %\"Ozkan
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
-For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every given state $s$ iff the downward-closed problem is decidable, i.e.
-given a state $s$ of a WSTS
+For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+%i 			.e. given a state $s$ of a WSTS
 % in the regarded class 
-with strong upward compatibility 
-and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
+%with strong upward compatibility 
+%and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
 \end{proposition}
+\alain{rappeler l'idée de la preuve}
 
 The characterization is used to show that Petri nets are $\post^*$-effective.
-It is well-known that Petri nets constitute WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded reachable-resilience is decidable for VASS.
-Hence reachable-resilience is decidable for Petri nets and VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.
+It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
+Hence state-resilience is decidable for Petri nets and VASS.
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:
 
@@
 \begin{proof}
-Let us show that Reset VASS don't enjoy the property that $\uparrow \post^*$ is computable.
+Let us show that Reset VASS, that are WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
 Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 

### 2023-03-30 resilience-safe-down.tex
details

--- 2023-03-30 resilience-safe-down.tex
@@
+\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
+
+example: ?
+
+% In the case \Safe and \Bad are both downward closed, they can both be finite.
+
+\begin{theorem}\label{down-down}
+The resilience problem is undecidable for effective WSTS with strong compatibility
+%	with more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\downarrow \Bad$.
+\end{theorem}
+
+\begin{proof}
+In the case of a WSTS with at least two minimal elements $m_1$ and $m_2$, the problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem, hence its undecidability.  
+\mathieu{dévellopper un peu la preuve, $m_2$ minimal de $\Safe$, $m_1$ minimal de $\Bad$...}
+\end{proof}
+
+In the case of a WSTS with only one minimal element,  
+% (ex: Petri Nets and their «Zero») 
+then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.
+
 %%%%%%%%
@@
 
+\begin{theorem}\label{up-down}
+\textcolor{red}{
+The resilience problem is undecidable for WSTS with 
+%	more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$.
+}
+\end{theorem}
+
+\begin{proof}
+TBD
+\end{proof}
+
 \begin{theorem}
-The resilience problem is decidable for ideal-effective WSTS with 
+The resilience problem is decidable for WSTS with 
+%	more than one minimal element, 
 $\Safe=\downarrow \Safe$
@@
 and
-the additional hypothesis that
-for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed.
-% \alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ?
+\textcolor{red}{
+additional hypothesis such that 
+}
+\alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ?
 si l'ordre est omega2 alors closbas+inclusion est wqo donc il faudrait juste que la suite croissante $pred^n(D)+le reste$ converge...}
@@
 \begin{proof}
-By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
-%  Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
-The resilience problem can be reformulated as 
-% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
-$\Bad \subseteq  \pred^*(\Safe)$.
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed subsets.
-Hence we can compute the intersection of
-% $\uparrow \{b_1,b_2,...,b_m\}$
-$\Bad$
-and
-$S \setminus \pred^*(\Safe)$,
-which are both upward-closed.
-Since
-% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
-$\Bad \subseteq \pred^*(\Safe)$
-can be reformulated as
-$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
-the resilience problem is decidable.
+Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Safe$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
+
+Let's assume for now for simplicity's sake that $\Bad = \uparrow b$
+
+As an hypothesis, we assume that $\pred^*(\Safe)$ is downward closed as well.
+And let $\{J'_1, J'_2,...,J'_n\}$ be the ideal decomposition of $\pred^*(\Safe)$.
+
+Resilience asks whether 
+$ \uparrow b \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$.
+
+Of note is that $ \downarrow b \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$
+if and only if
+there exists some $i$ such that
+$\downarrow b \subseteq J'_i$.
+
+\mathieu{Try and find a condition under which "$ \uparrow \{b\} \subseteq J'_1 \cup J'_2 \cup ... \cup J'_n$?" is decidable and/or find an undecidable problem that reduce to a variation on 
+"$\uparrow \{b\} \subseteq J'_i$ ?"}
 \end{proof}
-
-\begin{corollary}
-The resilience problem is decidable for ideal-effective downward-compatible WSTS with 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
-\end{corollary}
-
-\begin{proof}
-Downward compatibility of $\mathscr{S}=(S,\rightarrow, \leq)$ 
-means that
-for all $s_1, s_2, t_1 \in S$ with
-$s_2 \leq s_1$ and $s_1 \to t_1$
-there
-exists
-$t_2 \leq t_1$ such that $s_2 \to^* t_2$.
-
-Let $D$ be a downward-closed subset of $S$
-and let $x \in \downarrow \pred^*(D)$.
-By design there exists
-$y \in \pred^*(D)$ 
-such that $x \leq y$
-and $d \in D$ such that
-$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
-
-By downward compatibility $a_0 \to a_1$
-implies there exists $a'_1 \leq a_1$ with
-$x \to^* a'_1$.
-More generally $a_i \to a_{i+1}$ and
-$a'_i\leq a_i$ implies the existence of $a'_{i+1} \leq a_{i+1}$ such that
-$a'_i \to^* a'_{i+1}$,
-and, by induction,
- $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* d'$
-where
-$d' \leq d$ belongs to $D$ by downward closure of $D$.
-Hence $x \in \pred^*(D)$.
-\end{proof}
-
-% \begin{proposition}
-In the case
-of a ideal-effective WSTS 
-where
-the additional hypothesis that
-for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed
-is not met,
-the above construction
-can provide a proof
-of non-resilience
-i.e. when
-$\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) \neq \emptyset$
-then
-$\Bad \not\subseteq \downarrow\pred^*(\Safe)$
-and hence
-$\Bad \not\subseteq \pred^*(\Safe)$.
-When $\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) = \emptyset$
-however
-it is not enough to conclude.
-
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
-
-example: ?
-
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
-
-\begin{theorem}\label{down-down}
-The resilience problem is undecidable for strong effective WSTS with
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-\begin{proof}
-Let us consider the case of a WSTS with at least two minimal elements $m_1$ and $m_2$.
-The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
-By undecidability of the reachability problem we conclude.
-\end{proof}
-
-In the case of a WSTS with only one minimal element,  
-then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.

### 2023-03-30 resilience-safe-up.tex
swapping subsections around
details

@@
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad = ... $.}
-
-\alain{2 clos par le haut ont une intersection non vide....}
-
-Since we impose $\Bad \cap \Safe = \emptyset$, we cannot have as an hypothesis
-both $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$. To approximate a case
-of double upward closure, we consider a case where $\Bad$ is as upward-closed
-as it can be without intersecting with $\Safe$,
-i.e. 
-such that there exists an upward closed set $U$
-such that $\Bad = U \setminus \Safe$.
-Morally speaking ... [].
-
-example: ?
-
-\begin{theorem}\label{up-up}
-The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad= U \setminus \Safe$.
-
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$.
-%	and \Bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
-% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
 \subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}
@@
 %
+
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad = ... $.}
+
+\alain{2 clos par le haut ont une intersection non vide....}
+
+Since we impose $\Bad \cap \Safe = \emptyset$, we cannot have as an hypothesis
+both $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$. To approximate a case
+of double upward closure, we consider a case where $\Bad$ is as upward-closed
+as it can be without intersecting with $\Safe$,
+i.e. 
+such that there exists an upward closed set $U$
+such that $\Bad = U \setminus \Safe$.
+Morally speaking ... [].
+
+example: ?
+
+\begin{theorem}\label{up-up}
+The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+and $\Bad= U \setminus \Safe$ with $U=\uparrow U$.
+
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}\label{B-up-up}
+The bounded resilience problem is decidable for WSTS with effective pred-basis,
+strong compatibility
+ $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$.
+\end{corollary}
+
+\begin{theorem}\label{k-up-up}
+The % three %uniform 
+$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
+compatibility, $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$.
+%	and \Bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}

### 2023-03-30 resilience.tex
details

--- 2023-03-29 resilience.tex
@@
 \problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
 {$\Bad \longrightarrow^{*} \Safe$ ?\newline}
@@
 \problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
 {$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
@@
 \problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two finite encodings of sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}

### 2023-03-30 resilience.tex

@@
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
+\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
+   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up})  & Decidable (Theorem~\ref{up-down}) & Undecidable (Theorem~\ref{down-down}) \\ \hline
    BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline

### 2023-03-30 resilience-safe-up.tex
up

--- 2023-03-30 resilience-safe-up.tex
@@
 %
+
+\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad = \uparrow U \setminus \Safe $.}
+
+% \alain{2 clos par le haut ont une intersection non vide....}
+
+Since we impose $\Bad \cap \Safe = \emptyset$, we cannot have as an hypothesis
+both $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$. To approximate a case
+of double upward closure, we consider a case where $\Bad$ is as upward-closed
+as it can be without intersecting with $\Safe$,
+i.e. 
+such that there exists an upward-closed set $U$
+such that $\Bad = U \setminus \Safe$.
+Morally speaking ... [].
+
+example: ?
+
+\begin{theorem}\label{up-up}
+The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
+
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}\label{B-up-up}
+The bounded resilience problem is decidable for WSTS with effective pred-basis,
+strong compatibility
+ $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
+\end{corollary}
+
+\begin{theorem}\label{k-up-up}
+The % three %uniform 
+$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
+compatibility, $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
+%	and \Bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}

### 2023-03-30 resilience.tex
Theorem -> Thm in the tableau

@@
 \hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
-   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up})  & Decidable (Theorem~\ref{up-down}) & Undecidable (Theorem~\ref{down-down}) \\ \hline
-   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Theorem~\ref{k-up-up}) & Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
+   RP & Decidable (Thm~\ref{up-up}) & Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{up-down}) & Undecidable (Thm~\ref{down-down}) \\ \hline
+   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Thm~\ref{down-up}) & ?? & ?? \\ \hline
+      kRP & Decidable (Thm~\ref{k-up-up}) & Decidable (Thm~\ref{down-up}) & ?? & ?? \\ \hline
  \end{tabular}

### 2023-03-30 resilience.tex
up

@@
 
-\alain{dessous: à dire mais pas sous forme de proposition}
-\begin{proposition}[Reformulation]\label{reformulation}
+% \alain{dessous: à dire mais pas sous forme de proposition}
+\begin{remark}[Reformulation]\label{reformulation}
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
-\end{proposition}
+(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
+\end{remark}
 
@@
 RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-We deduce that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.
+We deduce \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
+
+\mathieu{changer ça}
 

### 2023-03-30 reachable-resilience.tex
details

@@
 \problemx{Downward reachability problem}
-{A state $s$, a downward-closed set $D$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D \subseteq S$.}
 {$s \to^* D$? \newline}
@@
 \begin{proof}
-Let $s \in S$, and a downward-closed set $D$.
+Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
 Take $\Safe = \emptyset$ and $\Bad = D$.
 The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \subseteq \emptyset$.
+whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
 If the answer is negative then $s \to^* \Bad$,
@@
 
-We may immediately generalyse this last result by strengthening to \emph{unbounded} resilience. The proof is essentially the same than the previous one.
+We may immediately generalyse this last result by strengthening to \emph{unbounded} state-resilience. The proof is essentially the same than the previous one.
 
@@
 
-What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+The computability of $\uparrow \post^*(s)$ seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
 %\"Ozkan

### 2023-03-30 application.tex
details

--- 2022-10-19 application.tex
@@
 \section{Applications section}
+
+\subsection{VASS, 2-VASS, semilinear VASS, Lossy Channel Systems, Integer et Continuous VASS, Regular Channel Systems}
+
+\begin{theorem}{}
+Resilience is decidable for semilinear VASS, LCM, Integer et Continuous VASS when $\Safe$ and $\Bad$ are semilinear sets.
+\end{theorem}
+
+se déduit du papier Schnoebelen RP 2010, page 8
+
+\begin{theorem}{}
+Resilience is decidable for Regular Channel Systems when $\Safe$ and $\Bad$ are regular languages.
+\end{theorem}
+
+open for VASS ???
 
@@
 
-\subsection{Vector Addition System with States}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
+%
+%
+%\subsection{Vector Addition System with States or PN}
+%
+%\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there} 
+%

### 2023-03-30 reachable-resilience.tex
l'idée de la preuve ...

@@
 
+The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
+$B_{\post^*(s)} \cap D = \emptyset$, equivalent to $\post^*(s)\cap D = \emptyset$ by
+Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed
+$U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
+Decidability of downward reachability leads to the decidability of the following stop condition:
+asking whether $S \setminus U_n$ is reachable from $s$.
+
 The characterization is used to show that Petri nets are $\post^*$-effective.
@@
 
-\begin{theorem}
+\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}
 {\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 

### 2023-03-30 resilience-safe-down.tex
up
details

--- 2023-03-30 resilience-safe-down.tex
@@
 %%%%%%%%
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
+\subsection{Case: $\Bad=\uparrow \Bad$.}
 %
@@
 %
+\subsubsection{Subcase: $\Safe=\downarrow \Safe$}
+
 example: mutual exclusion.
@@
 
-\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}
+\subsubsection{Case: $\Safe=\uparrow \Safe$.}
+
+% \alain{2 clos par le haut ont une intersection non vide....}
+
+Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward closed sets have not an empty intersection.
+We then consider the case where $\Bad$ is as upward-closed
+as it can be without intersecting with $\Safe$,
+i.e. 
+such that there exists an upward-closed set $U$
+such that $\Bad = U \setminus \Safe$.
+
+\alain{à quoi ça sert ici ?: Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
+
+and $\Bad = \uparrow U \setminus \Safe $
 
@@
 
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
-
-\begin{theorem}\label{down-down}
-The resilience problem is undecidable for strong effective WSTS with
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
+\begin{theorem}\label{up-up}
+The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
 \end{theorem}
@@
 \begin{proof}
-Let us consider the case of a WSTS with at least two minimal elements $m_1$ and $m_2$.
-The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
-By undecidability of the reachability problem we conclude.
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
 \end{proof}
 
-In the case of a WSTS with only one minimal element,  
-then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}\label{B-up-up}
+The bounded resilience problem is decidable for WSTS with effective pred-basis,
+strong compatibility
+ $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
+\end{corollary}
+
+\begin{theorem}\label{k-up-up}
+The % three %uniform 
+$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
+compatibility, $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
+%	and \Bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}

### 2023-03-30 resilience-safe-up.tex
up
details

--- 2023-03-30 resilience-safe-up.tex
@@
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad = ... $.}
+\subsection{Case: $\Bad=\downarrow \Bad$.}
 
-\alain{2 clos par le haut ont une intersection non vide....}
+%
+%		cas étudié par les deux articles sur les graphes
+%
 
-Since we impose $\Bad \cap \Safe = \emptyset$, we cannot have as an hypothesis
-both $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$. To approximate a case
-of double upward closure, we consider a case where $\Bad$ is as upward-closed
-as it can be without intersecting with $\Safe$,
-i.e. 
-such that there exists an upward closed set $U$
-such that $\Bad = U \setminus \Safe$.
-Morally speaking ... [].
+\subsubsection{Subcase: $\Safe=\downarrow \Safe$}
 
@@
 
-\begin{theorem}\label{up-up}
-The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad= U \setminus \Safe$.
+% In the case \Safe and \Bad are both downward closed, they can both be finite.
 
+\begin{theorem}\label{down-down}
+The resilience problem is undecidable for  effective WSTS with  strong  compatibility such that
+%	with more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{theorem}
@@
 \begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
+If the set $S$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ has an unique minimal element $m$, then $m$ belongs to both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.
+So let us consider the case of a set $S$ with at least two minimal elements $m_1$ and $m_2$.
+The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
+By undecidability of the reachability problem for effective WSTS with strong compatibility we conclude.
 \end{proof}
 
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
-\end{proposition}
+%%%%
 
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$.
-%	and \Bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
-% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
-\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}
+\subsubsection{Subcase: $\Safe=\uparrow \Safe$}
 
 example: ?
-%
-%		cas étudié par les deux articles sur les graphes
-%
 

### 2023-03-30 resilience.tex
up
details

@@
 
-In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$.
+In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$ and $\Safe ,  \Bad \neq \emptyset$.
 %	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
@@
 %		
-RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-We deduce \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
-
-\mathieu{changer ça}
+RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. We deduce 
+% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
+% \mathieu{changer ça}
+decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed as seen in Section~\ref{synthesis}.
 
@@
 
-\subsection{Case Synthesis}
+{\bf Synthesis}\label{synthesis}
 
@@
 
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
-%
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
 %We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 

### 2023-03-30 definitions.tex
details

@@
 
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering.
+Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
+Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
 
@@
 
+There is a simple condition that insures that a finite basis of $\pred^*(U )$ is computable.
+
 \begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
@@
 \end{definition}
+
+\begin{theorem}(ref)
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
+\end{theorem}
 
@@
 
-A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
-$s \in D$. 
+%         A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
+%         $s \in D$.
+
 %\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
@@
 \end{proof}}
-
-\begin{theorem}(ref)
-A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $pb(U)$.
-\end{theorem}

### 2023-03-31 reachable-resilience.tex
Starting rearranging things

@@
 
-\alain{mal placé}
-\begin{proposition}\label{reductions}
-In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
-%	home state and reachability
-\end{proposition}
+\subsection{Case: $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$}
 
-\begin{proof}
-
-\begin{itemize}
-\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the reachable resilience problem
-where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
-
-\end{proof}
-
-\problemx{Downward reachability problem}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D \subseteq S$.}
-{$s \to^* D$? \newline}
-
-\begin{theorem}
-The downward reachability problem is reducible to the state resilience problem.
-\end{theorem}
-
-\begin{proof}
-Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
-Take $\Safe = \emptyset$ and $\Bad = D$.
-The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
-If the answer is negative then $s \to^* \Bad$,
-else $s \not\to^* \Bad$.
-\end{proof}
-
-\alain{mal placé}
-
-% Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.
-
-% and with effective basis of $post^*$ ???
-
-% expliquer toutes les variantes de la résilience
-
-% Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.
+\mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
 
@@
 
+\begin{proof}
+%From Fact~\ref{stop condition},
+Since $\mathscr{S}$ is a WSTS there exists $n_0 \in \N$ such that
+$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute this $n_0$ by iteratively computing 
+%\alain{mauvaises notations: $k, k_m, Safe^k,...$ n'a aucun sens...}
+$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
+$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
+returning $n$ if that is the case.
+Then, because $n_0$-state-resilience is decidable, 
+checking $\uparrow post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+and state-resilience is decidable.
+\end{proof}
+
+\iffalse
 \alain{definir downward reachability problem.....
@@
 and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
-
-\begin{proof}
-From Fact~\ref{stop condition}, there exists $k \in \N$ such that
-$\pred^*(\Safe) =  \Safe^{k_m}$. We can compute this $k_m$ by iteratively computing \alain{mauvaises notations: $k, k_m, Safe^k,...$ n'a aucun sens...}
-$\Safe^{k+1}$ from $\Safe^k$, checking $\Safe^{k+1}=\Safe^k$, 
-returning $k_m$ if that is the case.
-Then, because $k_m$-reachable-resilience is decidable, 
-checking $\uparrow post^*(s) \cap \Bad \subseteq \Safe^k = \pred^*(\Safe)$ is,
-and reachable-resilience is decidable.
-\end{proof}
+\fi
 
 The computability of $\uparrow \post^*(s)$ seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is the WSTS for which the following problem is decidable.
+
+\problemx{Downward reachability problem}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
+\subseteq S$.}
+{$\exists s' \in D ~ s \to^* s'$? \newline}
+
 %\"Ozkan
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
-For finite-branching WSTS with strong upward compatibility, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For finite-branching WSTS%with strong upward compatibility
+, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 %i 			.e. given a state $s$ of a WSTS
@@
 \end{proposition}
-\alain{rappeler l'idée de la preuve}
 
+% \alain{rappeler l'idée de la preuve}
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
@@
 
+\alain{mal placé}
+\begin{proposition}\label{reductions}
+In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
+%	home state and reachability
+\end{proposition}
+
+\begin{proof}
+
+\begin{itemize}
+\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
+
+\item $t$-liveness  can be expressed as the following formula
+$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
+where
+$U_t=\uparrow \pred(t)$
+is the upward closure of the preconditions to use transition $t$.  
+The problem reduces itself to the reachable resilience problem
+where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
+
+%\item the home state problem can be expressed as $\ldots$
+\end{itemize}
+
+\end{proof}
+
+\iffalse
+
+\begin{theorem}
+The downward reachability problem is reducible to the state resilience problem.
+\end{theorem}
+
+\begin{proof}
+Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
+Take $\Safe = \emptyset$ and $\Bad = D$.
+The state resilience problem on $\Safe,\Bad$ asks
+whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
+If the answer is negative then $s \to^* \Bad$,
+else $s \not\to^* \Bad$.
+\end{proof}
+
+\fi
+
+\alain{mal placé}
+
 Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that reachable-resilience is undecidable for WSTS with strong monotony.
@@
 \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
+
+\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}

### 2023-03-31 reachable-resilience.tex
up

@@
 
-Let us recall a result about state-resilience (called resilieence in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+Let us recall a result about state-resilience (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
 
@@
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
-$B_{\post^*(s)} \cap D = \emptyset$, equivalent to $\post^*(s)\cap D = \emptyset$ by
+$B_{\uparrow \post^*(s)} \cap D = \emptyset$, equivalent to $\post^*(s)\cap D = \emptyset$ by
 Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed
@@
 \begin{proposition}
-In WSTS with strong monotony, the bounded reachable resilience problem is
-equivalent to the reachable-resilience problem.
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
 \end{proposition}
@@
 \begin{proof}
-strong WSTS
-
-$\pred^*(I) = \pred^{k}(I)$
-
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
 \end{proof}

### 2023-03-31 reachable-resilience.tex
up

@@
 
-%
-\iffalse
-%
-\textcolor{blue}{
-et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
-}
-%
-\textcolor{blue}{
-compute the sets of $(I,J,k)$ st S is resilient.\\
-max des I,J possibles \\
-}
-%
-\fi
+Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
 
-Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
+As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downard closed and upward closed.
 
@@
 The computability of $\uparrow \post^*(s)$ seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
-Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is the WSTS for which the following problem is decidable.
+Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
 
@@
 %\end{proof}
+
+Unfortunately, state-resilience is undecidable for (general) WSTS even with strong monotony.
+This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
+
+\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
+The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
+\end{theorem}
+
+\begin{proof}
+If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
+reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
+reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
+$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
+
+The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
+and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
+
+We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
+The initial marking $\mu_2$ will include a token in $r_1$.
+
+The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
+Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
+
+Transition $s$ is defined to have $r_1$ as an input and a null output.
+This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
+and "freezing" the marking of $P_1$.
+
+Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
+
+The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
+
+Now we need to see if that marking is zero.
+
+We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
+
+If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
+$s$ fires in the marking zero.
+
+If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
+
+In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
+We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
+This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
+Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
+
+Now if $0$ is reachable
+in $R(C_1, \mu_1)$ 
+then the net $C_2$ can also reach this marking
+in the place of $P_1$ by executing the same sequence of transition firings.
+Then $s$ can fire, freezing the $C_1$ subset.
+Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
+Thus if $0$ is reachable then $C_2$ is not live.
+
+Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
+$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
+(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
+% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
+% et r_2 commence sans token)
+If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
+Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
+
+\end{proof}
+
+\begin{corollary}
+Reset Petri Net liveness is undecidable 
+\end{corollary}
+
+\begin{corollary}
+Reset Petri Net $t$-liveness is undecidable 
+\end{corollary}
+
+\alain{mal placé}
+\begin{proposition}\label{reductions}
+In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
+%	home state and reachability
+\end{proposition}
+
+\begin{proof}
+
+\begin{itemize}
+\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
+
+\item $t$-liveness  can be expressed as the following formula
+$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
+where
+$U_t=\uparrow \pred(t)$
+is the upward closure of the preconditions to use transition $t$.  
+The problem reduces itself to the reachable resilience problem
+where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
+
+%\item the home state problem can be expressed as $\ldots$
+\end{itemize}
+
+\end{proof}
+
+\iffalse
+
+\begin{theorem}
+The downward reachability problem is reducible to the state resilience problem.
+\end{theorem}
+
+\begin{proof}
+Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
+Take $\Safe = \emptyset$ and $\Bad = D$.
+The state resilience problem on $\Safe,\Bad$ asks
+whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
+If the answer is negative then $s \to^* \Bad$,
+else $s \not\to^* \Bad$.
+\end{proof}
+
+\fi
+
+\alain{mal placé}
+
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong monotony.
+
+\begin{theorem}
+{\sc State-resilience} is undecidable for WSTS with strong monotony.
+\end{theorem}
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis, the bounded state-resilience problem is
+reducible to the $k$-state-resilience problem.
+\end{proposition}
+
+\begin{proof}
+% strong WSTS
+% $\pred^*(I) = \pred^{k}(I)$
+% Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+We
+\end{proof}
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}
+Bounded reachable-resilience and 
+$k$ reachable-resilience are undecidable for WSTS with strong monotony.
+\end{corollary}
+
+\begin{proof} 
+  From
+  cf
+the two  theorems above. 
+\end{proof}
 
@@
 
-Unfortunately, reachable-resilience is undecidable for (general) WSTS even with strong monotony.
-This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
-
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem}
-
-\begin{proof}
-If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
-reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
-reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
-$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
-
-The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
-and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
-
-We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $r_1$.
-
-The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
-
-Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
-and "freezing" the marking of $P_1$.
-
-Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
-The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
-
-Now we need to see if that marking is zero.
-
-We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
-
-If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
-$s$ fires in the marking zero.
-
-If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
-
-In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
-Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-
-Now if $0$ is reachable
-in $R(C_1, \mu_1)$ 
-then the net $C_2$ can also reach this marking
-in the place of $P_1$ by executing the same sequence of transition firings.
-Then $s$ can fire, freezing the $C_1$ subset.
-Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $0$ is reachable then $C_2$ is not live.
-
-Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
-$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
-(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
-% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
-% et r_2 commence sans token)
-If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
-
-\end{proof}
-
-\begin{corollary}
-Reset Petri Net liveness is undecidable 
-\end{corollary}
-
-\begin{corollary}
-Reset Petri Net $t$-liveness is undecidable 
-\end{corollary}
-
-\alain{mal placé}
-\begin{proposition}\label{reductions}
-In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
-%	home state and reachability
-\end{proposition}
-
-\begin{proof}
-
-\begin{itemize}
-\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the reachable resilience problem
-where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
-
-\end{proof}
-
-\iffalse
-
-\begin{theorem}
-The downward reachability problem is reducible to the state resilience problem.
-\end{theorem}
-
-\begin{proof}
-Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
-Take $\Safe = \emptyset$ and $\Bad = D$.
-The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
-If the answer is negative then $s \to^* \Bad$,
-else $s \not\to^* \Bad$.
-\end{proof}
-
-\fi
-
-\alain{mal placé}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that reachable-resilience is undecidable for WSTS with strong monotony.
-
-\begin{theorem}
-{\sc reachable-resilience} is undecidable for WSTS with strong monotony.
-\end{theorem}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded reachable-resilience problem is
-reducible to the $k$-reachable-resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-$\pred^*(I) = \pred^{k}(I)$
-Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}
-Bounded reachable-resilience and 
-$k$ reachable-resilience are undecidable for WSTS with strong monotony.
-\end{corollary}
-
-\begin{proof} 
-  From
-  cf
-the two  theorems above. 
-\end{proof}
-
 % Let us deduce that effective basis of $post^*$ ex reset PN ?

### 2023-03-31 appendix.tex
up

--- 2023-03-31 reachable-resilience.tex
@@
-\section{State resilience}
+\section{Proof of Theorem~\ref{liveness reset}}\label{appendix}
 
-Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap post^*(s_0)$, there must exist a path to $\Safe$.
-%
-%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
-%
-%\problemx{(I,J)-$k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
-%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-%\problemx{general $k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
-%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+Recall Theorem~\ref{liveness reset}.
 
-%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
+\begin{theorem*}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
+The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
+\end{theorem*}
 
-The three previous problems become:
-
-\problemx{State-resilience problem (SRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-% 		an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
-% 		{$ ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe, ~ s' \rightarrow^{*} s''$ ?\newline}
-{$\Bad \cap \post^*(s_0)  \rightarrow^{*} \Safe $ ? \newline}
-
-\problemx{$k$-state-resilience problem (kSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-%
-
-\problemx{bounded-state-resilience problem (BSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient. ?\newline}
-
-%\textcolor{blue}{
-%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
-%}
-
-%
-\iffalse
-%
-\textcolor{blue}{
-et encore: renforcer sur l'hypothèses sur I, J tous les deux clos par le haut+bas.
-}
-%
-\textcolor{blue}{
-compute the sets of $(I,J,k)$ st S is resilient.\\
-max des I,J possibles \\
-}
-%
-\fi
-
-Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
-
-\subsection{Case: $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$}
-
-\mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
-
-Let us recall a result about state-resilience (called resilieence in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
-
-\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
-\end{theorem}
-
-We may immediately generalyse this last result by strengthening to \emph{unbounded} state-resilience. The proof is essentially the same than the previous one.
-
-\begin{corollary}\label{postcomputable}
-{\sc state-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
-\end{corollary}
-
-\begin{proof}
-%From Fact~\ref{stop condition},
-Since $\mathscr{S}$ is a WSTS there exists $n_0 \in \N$ such that
-$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute this $n_0$ by iteratively computing 
-%\alain{mauvaises notations: $k, k_m, Safe^k,...$ n'a aucun sens...}
-$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
-$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
-returning $n$ if that is the case.
-Then, because $n_0$-state-resilience is decidable, 
-checking $\uparrow post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-and state-resilience is decidable.
-\end{proof}
-
-\iffalse
-\alain{definir downward reachability problem.....
-downward-closed problem given a state $s$ of a WSTS
-% in the regarded class 
-with strong upward compatibility 
-and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
-\fi
-
-The computability of $\uparrow \post^*(s)$ seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
-Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is the WSTS for which the following problem is decidable.
-
-\problemx{Downward reachability problem}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
-\subseteq S$.}
-{$\exists s' \in D ~ s \to^* s'$? \newline}
-
-%\"Ozkan
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
-For finite-branching WSTS%with strong upward compatibility
-, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-%i 			.e. given a state $s$ of a WSTS
-% in the regarded class 
-%with strong upward compatibility 
-%and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
-\end{proposition}
-
-% \alain{rappeler l'idée de la preuve}
-The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
-$B_{\post^*(s)} \cap D = \emptyset$, equivalent to $\post^*(s)\cap D = \emptyset$ by
-Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed
-$U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
-Decidability of downward reachability leads to the decidability of the following stop condition:
-asking whether $S \setminus U_n$ is reachable from $s$.
-
-The characterization is used to show that Petri nets are $\post^*$-effective.
-It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
-Hence state-resilience is decidable for Petri nets and VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:
-
-\begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}
-Let us show that Reset VASS, that are WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
-%in counter $i$ 
-by examining if there is %some vector $v$ in the basis such that $v(i)=0$
-$0$ in the basis%
-. But reachability of $0$  %in a counter $i$ 
-is undecidable for Reset VASS. 
-% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
-\end{proof}
-
-%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
-
-%
-%si on garde strong mais on enlève effective basis of $post^*$:
-%
-%\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
-%\end{proposition}
-%
-%\begin{proof}
-%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
-%\end{proof}
-
-Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general reachable-resilience problem.
-
-\begin{lemma}\label{Lemma intersection}
-Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
-\end{lemma}
-
-\begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
-Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
-Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\uparrow A}$, the inclusion
-$({\uparrow  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
-\end{proof}
-
-\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}
-{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
-$\uparrow$ $\post^*$ basis.
-\end{theorem}
-
-\begin{proof}
-Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
-and $\Bad$ decidable downward-closed.
-By applying Lemma~\ref{Lemma intersection} twice, we obtain
-
-$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$
-
-Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
-% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
-\end{proof}
-
-% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
-% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
-% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
-%
-% Well, maybe their version of resilience but resilience with * seems fine.
-%
-
-However when removing strong compatibility, precision is lost and the algorithm do not work anymore.
-
-\begin{proposition}
-The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-reachable-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}[Sketch]
-Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
-For instance with a rule $\delta(q,b) = \text{ add some }b $,
-we have $(q,b) \to (q,bb)$, $(q, baaa) \geq (q,b)$ and
-$(q, baaa) \to^* (q,bb)$ however it takes more than a single step.
-
-In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
-$post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$
-do not imply
-$\uparrow post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$ for a downward-closed set $D$.
-In particular by lossiness $(q, \epsilon)$ would belong to
-$post^* (s)$,
-hence $(q,a)$, $(q,aa)$, etc. would belong to $\uparrow post^* (s)$ although not necessarily to
-$post^* (s)$. $(q,aaa)$ not being in $pred(\uparrow \{ (q,b)\})$ means the algorithm 
-from \cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold in this example when $\Safe=\uparrow \{(q,b)\}$ and $\Bad =S$,
-because $\uparrow post^* (s) \cap S \not\subseteq \pred(\uparrow \{(q,b)\})$,
-despite having 
-$post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.
-
-\iffalse
-\mathieu{
-
-On considère un LCS, où on autorise à effacer des messages (lossyness) mais effacer un message prends une étape (contrairement à [ref], où ils font une sorte de cloture de effacer-transition-effacer).
-On n'a alors pas la strong compatibility:
-
-On prends: $(q, b) 	\to (q,bb)$ par exemple et $(q, baaaaaa) \geq (q,b) $, alors
-$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
-MAIS
-
-C'est pas "en une étape".
-
-(On considère -il faut le mentionner/vérifier- que on ne peut effaçer que d'un coté ? et pas au milieu?)
-
-Dans ce cas là, du coup, $pred(\uparrow \{(q,b)\}$ n'est pas clos par le haut. \\
-
-Revenons au Corollaire
-
-On voudrais avoir
-
-$post^* (s) \cap J \subseteq I^k$
-
-MAIS pas
-
-$ (\uparrow post^* (s)) \cap J \subseteq I^k$ \\
-
-Ça serait un cas où leurs algorithme vas dire 
-"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
-
-Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
-
-Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,b)\})$ qui n'est pas clos par le haut.
-
-On voudrait
-
-$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,b)\})$
-
-Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
-$\uparrow post^* (s)$ being too big
-donc autant pas se restreindre du tout sur $J$
-
-$post^*(s) \subseteq pred(\uparrow \{ (q,b)\})$
-
-Si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
-$b$ ? et évidement en perdant le dernier message (parce qu'on est lossy). Et c'est tout, disons.
-
-Déjà par lossiness $(q,\epsilon)$ serait dans $post^*((q,b))$.
-
-Et aussi dans $pred(\uparrow \{ (q,b)\})$ puisque $(q,\epsilon) \to (q,b)$.
-
-Donc on aurait:
-
-$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
-
-Et on aurait $(q,aaaaaa)$ dans 
-$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$.
-
-Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
-(même si il est dans $pred^*(\uparrow \{ (q,b)\})$).
-
-}
-\fi
-
-\end{proof}
-
-\begin{proposition}
-\textcolor{red}{
-{\sc bounded resilience} is undecidable for WSTS with upward compatibility with an effective basis of $post^*$.
-CONJECTURE
-}
-\end{proposition}
-
-\begin{proof}
- \textcolor{red}{pas certain ! à faire Mathieu}
-\end{proof}
-
-Unfortunately, reachable-resilience is undecidable for (general) WSTS even with strong monotony.
-This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
-
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem}
+Let us describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri}.
 
@@
 
-\begin{corollary}
-Reset Petri Net liveness is undecidable 
-\end{corollary}
-
-\begin{corollary}
-Reset Petri Net $t$-liveness is undecidable 
-\end{corollary}
-
-\alain{mal placé}
-\begin{proposition}\label{reductions}
-In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
-%	home state and reachability
-\end{proposition}
-
-\begin{proof}
-
-\begin{itemize}
-\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the reachable resilience problem
-where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
-
-\end{proof}
-
-\iffalse
-
-\begin{theorem}
-The downward reachability problem is reducible to the state resilience problem.
-\end{theorem}
-
-\begin{proof}
-Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
-Take $\Safe = \emptyset$ and $\Bad = D$.
-The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
-If the answer is negative then $s \to^* \Bad$,
-else $s \not\to^* \Bad$.
-\end{proof}
-
-\fi
-
-\alain{mal placé}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that reachable-resilience is undecidable for WSTS with strong monotony.
-
-\begin{theorem}
-{\sc reachable-resilience} is undecidable for WSTS with strong monotony.
-\end{theorem}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded reachable-resilience problem is
-reducible to the $k$-reachable-resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-$\pred^*(I) = \pred^{k}(I)$
-Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong monotony, the bounded reachable resilience problem is
-equivalent to the reachable-resilience problem.
-\end{proposition}
-
-\begin{proof}
-strong WSTS
-
-$\pred^*(I) = \pred^{k}(I)$
-
-\end{proof}
-
-\begin{corollary}
-Bounded reachable-resilience and 
-$k$ reachable-resilience are undecidable for WSTS with strong monotony.
-\end{corollary}
-
-\begin{proof} 
-  From
-  cf
-the two  theorems above. 
-\end{proof}
-
-% Let us deduce that effective basis of $post^*$ ex reset PN ?
-
- \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-We may consider context-free grammars.
-
-\begin{corollary}
-{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
-\end{corollary}
-
-\begin{proof}
-
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
-\end{proof}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
-
-\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}
+\section{More Appendix thing if necessary}\label{appendix}

### 2023-03-31 reachable-resilience.tex
up

@@
 
+\mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
+
 \subsection{Case: $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$}
-
-\mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
 
@@
 
-Unfortunately, state-resilience is undecidable for (general) WSTS even with strong monotony.
+Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
 This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
 
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
+\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
 The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
@@
 
+\iffalse
 \begin{proof}
@@
 \end{proof}
+\fi
+
+\mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
+
+Since reachability of $0$ is undecidable in Reset Petri-Nets, the reduction implies the following:
 
 \begin{corollary}
-Reset Petri Net liveness is undecidable 
+Reset Petri Net liveness is undecidable.
 \end{corollary}
 
+Since moreover liveness can be reduced to 
+$t$-liveness, we deduce:
+
 \begin{corollary}
-Reset Petri Net $t$-liveness is undecidable 
+Reset Petri Net $t$-liveness is undecidable.
 \end{corollary}
 
-\alain{mal placé}
+We now reduce $t$-liveness to the state resilience problem.
+
 \begin{proposition}\label{reductions}
-In WSTS, coverability and $t$-liveness are reducible to the state resilience problem.
+In WSTS, $t$-liveness is reducible to the state resilience problem.
 %	home state and reachability
@@
 \begin{proof}
-
-\begin{itemize}
-\item Let us choose $\Bad=S= \downarrow S$ and $\Safe=\uparrow x$ then reachable resilience is equivalent to coverability of $x$ from $s$.
-
-\item $t$-liveness  can be expressed as the following formula
-$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$ 
+We reformulate $t$-liveness as the following formula
+% $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
+$$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$  
 where
@@
 is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the reachable resilience problem
-where $\Bad=S$ and $\Safe = \uparrow \pred(t)$.
-
-%\item the home state problem can be expressed as $\ldots$
-\end{itemize}
+The problem reduces itself to the state resilience problem
+where $\Safe = \uparrow \pred(t)$ and $\Bad = S \setminus \Safe$.
 
@@
 
-\alain{mal placé}
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong monotony, from Proposition \ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong monotony.
+Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong upward compatibility, from Proposition \ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
 
 \begin{theorem}
-{\sc State-resilience} is undecidable for WSTS with strong monotony.
+{\sc State-resilience} is undecidable for strongly compatible WSTS with effective pred-basis.
 \end{theorem}
@@
 Bounded reachable-resilience and 
-$k$ reachable-resilience are undecidable for WSTS with strong monotony.
+$k$ reachable-resilience are undecidable for strongly compatible WSTS with effective pred-basis.
 \end{corollary}

### 2023-03-31 reachable-resilience.tex
Starting section 4 case Safe downward closed

@@
 
+This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
 \begin{proposition}
@@
 $\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-We
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
+We compute 
+$n_0$, then iteratively check whether $k$-state resilience hold for $k$ from $0$ to $n_0$.  
 \end{proof}
@@
 \begin{corollary}
-Bounded reachable-resilience and 
-$k$ reachable-resilience are undecidable for strongly compatible WSTS with effective pred-basis.
+Bounded state-resilience and 
+$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis.
 \end{corollary}
@@
 \begin{proof} 
-  From
-  cf
-the two  theorems above. 
+  From the two reductions above. 
 \end{proof}
@@
 
+\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}
+
+The authors from \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
+have only been concerned with the case $\Safe = \uparrow \Safe$
+and $\Bad =\downarrow \Bad$, however as mentioned in Section~3
+the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
+Hence it makes sense to consider this symmetrical case as well. \\
+
+Undecidability in general: $\Safe = 0$, $\Bad = S \setminus 0$ ?
+
+LOREM IPSUM \\
+
+\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap U \subseteq D$  iff $ (\downarrow  A) \cap U \subseteq D$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap U \subseteq D$. Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$.
+Let us show that $({\downarrow A}) \cap U \subseteq {\downarrow (A \cap U)}$.
+Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$.
+Since $x \in U$ and $U$ is downward-closed, we also have $a \in U$.
+Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
+In the other direction,
+since $A \subseteq {\downarrow A}$, the inclusion
+$({\downarrow  A}) \cap U \subseteq D$ implies
+$A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
+\end{proof}
+
+Lemma~\ref{Lemma intersection 2} can be used to show that
+$\post^*(s) \cap \Bad \subseteq D$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq D$.
+
+If, as in Section~XX, we have the \textcolor{red}{added hypothesis}
+that
+$\pred^*(\Safe)$ is downward closed this means that
+$\post^*(s) \cap \Bad \subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$.
+
+And hence decidability follows.
+
+\subsection{Other}
+
+\mathieu{Everything below should be moved to the applications section no?}
+
 % Let us deduce that effective basis of $post^*$ ex reset PN ?
@@
 \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
-
-\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}

### 2023-03-31 reachable-resilience.tex
up
tableau
Ref added to tables

@@
 
-\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
+\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
 {\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
@@
 
-\begin{theorem}
+\begin{theorem}\label{srp up down}
 {\sc State-resilience} is undecidable for strongly compatible WSTS with effective pred-basis.
@@
 
-\begin{corollary}
+\begin{corollary}\label{bsrp up down}
 Bounded state-resilience and 
@@
 
-\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}
+\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
 {\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
@@
 
+But what if we are in the case of $\uparrow \post^*$ effectiveness?
+
 LOREM IPSUM \\
@@
 
+\mathieu{Should try to put all results in a single table}
+
+With only strong compatibility hypothesis:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | r |}
+\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
+   SRP & Undecidable (Thm~\ref{srp up down}) & Undecidable  \\ \hline
+   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  \\ \hline
+      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? \\ \hline
+ \end{tabular}
+\end{center}
+
+With $\uparrow \post^*$ effectiveness hypothesis:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | r |}
+\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
+   SRP & Decidable (Thm~\ref{post srp}) & ??  \\ \hline
+   BSRP & ?? &  ??  \\ \hline
+      kSRP & ?? & ?? \\ \hline
+ \end{tabular}
+\end{center}
+
+With $\uparrow \post^*$ effectiveness and strong compatibility hypothesis:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | r |}
+\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
+   SRP & Decidable (Thm~\ref{postcomputable})& ??  \\ \hline
+   BSRP & Decidable (Thm~\ref{ref ozkan})&  ??  \\ \hline
+      kSRP & Decidable (Thm~\ref{ref ozkan}) & ?? \\ \hline
+ \end{tabular}
+\end{center}
+
 \mathieu{Everything below should be moved to the applications section no?}

### 2023-03-31 reachable-resilience.tex
Addind huyphothesis on safe and bad in theorems in Sect 4.1

@@
 
-As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downard closed and upward closed.
+As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downard closed and upward closed. We will consider here the cases where 
+$\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
+and where 
+$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 
-\mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
+% \mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
 
@@
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
-{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
+{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{theorem}
@@
 \begin{corollary}\label{postcomputable}
-{\sc state-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$.
+{\sc state-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{corollary}
@@
 \begin{theorem}\label{srp up down}
-{\sc State-resilience} is undecidable for strongly compatible WSTS with effective pred-basis.
+{\sc State-resilience} is undecidable for strongly compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{theorem}
@@
 Bounded state-resilience and 
-$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis.
+$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{corollary}
@@
 {\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
-$\uparrow$ $\post^*$ basis.
+$\uparrow$ $\post^*$ basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
 \end{theorem}
@@
 
-Undecidability in general: $\Safe = 0$, $\Bad = S \setminus 0$ ?
+The problem is however unfortunately undecidable in general.
+
+\begin{theorem}
+The resilience problem is undecidable for effective WSTS with  strong  compatibility 
+when
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$.
+\end{theorem}
+
+\begin{proof}
+Taking $\Safe = 0$, $\Bad = S \setminus 0$, it follows the state resilience problem is equivalent
+to reachability of $0$.
+\end{proof}
 

### 2023-03-31 reachable-resilience.tex
up

@@
 
-The problem is however unfortunately undecidable in general.
+Unfortunately however the problem is undecidable in general.
 
@@
 \begin{proof}
-Taking $\Safe = 0$, $\Bad = S \setminus 0$, it follows the state resilience problem is equivalent
-to reachability of $0$.
+Taking $\Safe = \{ 0 \}$, $\Bad = S \setminus 0$, it follows the state resilience problem is equivalent
+to reachability of $0$. Since reachability of $0$ is undecidable for Reset VASS, we conclude.
 \end{proof}

### 2023-03-31 reachable-resilience.tex
up

@@
 \begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
-{\sc Reachable-resilience} is decidable for (not necessarily strong) WSTS with effective 
+{\sc state-resilience} is decidable for (not necessarily strong) WSTS with effective 
 $\uparrow$ $\post^*$ basis
@@
 \begin{proposition}
-The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-reachable-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
+The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-state-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
 \end{proposition}

### 2023-03-31 resilience-safe-up.tex
details

@@
 %
+Let us choose $\Bad$ to be the set of states from which there is no infinite run.
+If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward closed.
 
@@
 
-example: ?
+The well-known mutual exclusion property can be modelized, in a $d$-counters machine, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.
+%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
 
@@
 
-example: ?
+example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e.
+from which it is always possible to fire a transition.
+This set of states is upward closed and $\Safe=S-\Bad=\uparrow \Safe$.
 
@@
 
-The \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
 The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.

### 2023-03-31 resilience.tex
details

@@
 
-For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one.
-Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
-
 In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
@@
 %	 and ideals are recursive
-\alain{???}.
 

### 2023-03-31 resilience-safe-up.tex
details

@@
 
-The well-known mutual exclusion property can be modelized, in a $d$-counters machine, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
+The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
 Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.

### 2023-03-31 resilience-safe-down.tex
details

@@
 %
-%		cas exclusion mutuelle
+%		cas unbounded counters
 %
+
+let us choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$.
+%			at least one counter is unbounded. 
+This set is upward-closed by upward compatibility in VASS.
+
 \subsubsection{Subcase: $\Safe=\downarrow \Safe$}
 
-example: mutual exclusion.
+As an example, let us choose $\Safe=S-\Bad$ then $\Safe=\downarrow \Safe$.
 
@@
 the additional hypothesis that
-for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed.
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
 % \alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ?

### 2023-03-31 reachable-resilience.tex
up

@@
 
-Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general reachable-resilience problem.
+Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
 

### 2023-03-31 reachable-resilience.tex
tryin be less ellyptic, proof needs a bit more robustness checking out tho

@@
 \begin{proof}
-Taking $\Safe = \{ 0 \}$, $\Bad = S \setminus 0$, it follows the state resilience problem is equivalent
-to reachability of $0$. Since reachability of $0$ is undecidable for Reset VASS, we conclude.
+Consider Reset VASS
+in the case $\Safe$ contains only the minimal element %$0$ 
+and $\Bad = S \setminus \Safe$, 
+it follows the state resilience problem is equivalent 
+to the problem of whether $0$ can be visited infinitely often. 
+%
+% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
+Hence the undecidability.
 \end{proof}

### 2023-04-03 reachable-resilience.tex
moved things to application section

@@
 \end{center}
-
-\mathieu{Everything below should be moved to the applications section no?}
-
-% Let us deduce that effective basis of $post^*$ ex reset PN ?
-
- \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-We may consider context-free grammars.
-
-\begin{corollary}
-{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
-\end{corollary}
-
-\begin{proof}
-
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
-\end{proof}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}

### 2023-04-03 reachable-resilience.tex
up
only one tabular now

@@
 
-\mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
+% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
+\begin{proof}
+See Appendix~\ref{appendix} for the complete proof.
+\end{proof}
 
-Since reachability of $0$ is undecidable in Reset Petri-Nets, the reduction implies the following:
+Since the Reset Petri Net zero reachability problem is undecidable in Reset Petri-Nets, the reduction implies the following:
 
@@
 
-\subsection{Other}
+\subsection{Synthesis}
 
+\iffalse
 \mathieu{Should try to put all results in a single table}
@@
 \end{center}
+
+\fi
+
+Results in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | c | c |}
+\hline  Hypothesis & strong compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
+   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
+   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+ \end{tabular}
+\end{center}

### 2023-04-03 reachable-resilience.tex
up

@@
 \begin{proof}[Sketch]
+\mathieu{ (Formalism should be made more explicit so that the proof can be really pertinent)}
+\textcolor{red}{
 Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
@@
 $(q, baaa) \to^* (q,bb)$ however it takes more than a single step. 
-
 In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
@@
 $post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.
+}
 
@@
 
+\mathieu{In case we have  effective basis of $post^*$ and upward compatibility (not strong) then we don't know the decidability status of k-state-resilience and bounded-state-resilience.
+Not necessarily a big deal, but should write a comment or two about this.}
+
+\iffalse
 \begin{proposition}
@@
 \end{proof}
+\fi
 

### 2023-04-03 reachable-resilience.tex
up

@@
 The characterization is used to show that Petri nets are $\post^*$-effective.
+
+\begin{proposition}
+Let $P = (N,M,W)$ be a Petri net and $s \in M$ a configuration of the Petri net, then one can compute a basis of $\uparrow \post^*(s)$.
+\end{proposition}
+
 It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
@@
 
-LOREM IPSUM \\
+\mathieu{Écrire que c'est globalement symmétrique au cas d'avant - je pense pas nécessairement qu'on arrive à faire beaucoup mieux que de juste dire ça.}
 

### 2023-04-03 reachable-resilience.tex
up

@@
 \begin{proposition}
-Let $P = (N,M,W)$ be a Petri net and $s \in M$ a configuration of the Petri net, then one can compute a basis of $\uparrow \post^*(s)$.
+Let $P = (N,M,W)$ be a Petri net and $s \in M$ a configuration of $P$, then one can compute a basis of $\uparrow \post^*(s)$.
 \end{proposition}
@@
 It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
-Hence state-resilience is decidable for Petri nets and VASS.
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
+Hence state-resilience is decidable for both Petri nets and VASS.
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:

### 2023-04-03 reachable-resilience.tex
up

@@
 
-The characterization is used to show that Petri nets are $\post^*$-effective.
+The characterization is used to show that Petri nets are $\post^*$-effective. Since the downward reachability problem is decidable for Petri-nets, we deduce the following:
 
@@
 Let us show that Reset VASS, that are WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
+Suppose that one is able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
 %in counter $i$ 
@@
 have only been concerned with the case $\Safe = \uparrow \Safe$
-and $\Bad =\downarrow \Bad$, however as mentioned in Section~3
+and $\Bad =\downarrow \Bad$, 
+\mathieu{pas tout à fait vrai: dans le technical report \cite{DBLP:conf/gg/Ozkan22} il considère aussi le cas où Bad et Safe sont tous les deux clos par le haut mais aussi le cas Safe clos par le bas Bad clos par le haut.}
+ however as mentioned in Section~3
 the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
 Hence it makes sense to consider this symmetrical case as well. \\
+
+\mathieu{rappeler résultats de \cite{DBLP:conf/gg/Ozkan22}}
+
+\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
+The bounded-state-resilience problem is decidable for \textcolor{red}{fin-marked} 
+strongly compatible WSTS
+ which are 
+ {\em \textcolor{red}{lossy} } 
+ and 
+ {\em \textcolor{red}{$\bot$-bounded}  }
+ if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
+\end{theorem}
 

### 2023-04-03 reachable-resilience.tex
added results from [9] on safe clos par le bas bad clos par le haut mais je trouve c'est bizare

@@
 
+\begin{definition}{marked WSTS} 
+A {\em marked} WSTS is a tuple $(S, \leq, \to, INIT)$
+where $(S, \leq, \to)$ is a WSTS and $INIT \subseteq S$. 
+If $INIT$ is finite, we call it {\em fin-marked}.
+\end{definition}
+
+\begin{definition}
+A marked WSTS is
+\begin{itemize}
+\item lossy if $\downarrow \post^* (INIT) = \post^* (INIT)$,
+\item $\bot$-bounded (bottom-bounded) if there is a natural number 
+$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
+$s_B$ of a basis of $S$ with $s \geq s_B$.
+\end{itemize}
+\end{definition}
+
+\mathieu{}
+
 \begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}

### 2023-04-03 reachable-resilience.tex
added results from [9] on safe clos par le bas bad clos par le haut mais je trouve c'est bizare
Proof of prop22 temporary

@@
 %\"Ozkan
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
+\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
 For finite-branching WSTS%with strong upward compatibility
@@
 \end{proposition}
+
+\begin{proof}
+By Proposition~\ref{post*}, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+Let us now show that the downward-reachability problem is decidable for Petri nets.
+Let us consider a downward closed set $D$ in a Petri net.
+$D$ is  defined by ... states where your number of counter is bounded by a constant, and others where it is unbounded. You have $m$ counters where it is bounded by $n$
+then you have $m \times n$ possibilities for the values in these counters
+you have unlimited possibilities for the values of the other counters,
+but that is where the submarking problem comes in handy,
+i.e. you ask the submarking problem for all of the $m \times n$ possibilities.
+
+So downward-set reachability is decidable for PN.
+\end{proof}
 
@@
 
-But what if we are in the case of $\uparrow \post^*$ effectiveness?
+\mathieu{But what if we are in the case of $\uparrow \post^*$ effectiveness?}
 
@@
 \end{center}
+
+Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:

### 2023-04-03 reachable-resilience.tex
up

@@
 Let us consider a downward closed set $D$ in a Petri net.
-$D$ is  defined by ... states where your number of counter is bounded by a constant, and others where it is unbounded. You have $m$ counters where it is bounded by $n$
+Let us consider $B$ a basis of the upward closed set $S \setminus D$.
+
+$D$ is  defined by ...
+states where your number of counter is bounded by a constant (deduce this from ...
+having to be smaller than elements of the basis $B$), and others where it is unbounded.
+You have $m$ counters where it is bounded by $n$
 then you have $m \times n$ possibilities for the values in these counters
@@
 \begin{theorem}
-The resilience problem is undecidable for effective WSTS with  strong  compatibility 
+The state resilience problem is undecidable for effective WSTS with  strong  compatibility 
 when

### 2023-04-03 reachable-resilience.tex
removed the proof until I have a cleaner one
up

@@
 \begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
+\hline  Hypothesis & strong upward compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline

### 2023-04-03 reachable-resilience.tex
up

@@
 \begin{proof}[Sketch]
-\mathieu{ (Formalism should be made more explicit so that the proof can be really pertinent)}
+Since $\pred(\uparrow \Safe)$ is not necessarily upward closed, it is possible to have 
+ $\uparrow post^* (s) \cap S \not\subseteq \pred(\Safe)$,
+despite having 
+$post^* (s) \cap S \subseteq \pred( \Safe)$.
+In such a case the algorithm from
+\cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold,
+which is incorrect.
+An example of this is ...
+\mathieu{ (Lossy counter machine whose Formalism should be made more explicit so that the proof can be really pertinent)}
+\iffalse
 \textcolor{red}{
@@
 }
-
-\iffalse
-\mathieu{
-
-On considère un LCS, où on autorise à effacer des messages (lossyness) mais effacer un message prends une étape (contrairement à [ref], où ils font une sorte de cloture de effacer-transition-effacer).
-On n'a alors pas la strong compatibility:
-
-On prends: $(q, b) 	\to (q,bb)$ par exemple et $(q, baaaaaa) \geq (q,b) $, alors
-$(s, baaaaaa) \to^* (s,b) \to (s,bb)$
-MAIS
-
-C'est pas "en une étape".
-
-(On considère -il faut le mentionner/vérifier- que on ne peut effaçer que d'un coté ? et pas au milieu?)
-
-Dans ce cas là, du coup, $pred(\uparrow \{(q,b)\}$ n'est pas clos par le haut. \\
-
-Revenons au Corollaire
-
-On voudrais avoir
-
-$post^* (s) \cap J \subseteq I^k$
-
-MAIS pas
-
-$ (\uparrow post^* (s)) \cap J \subseteq I^k$ \\
-
-Ça serait un cas où leurs algorithme vas dire 
-"On n'a pas $(\uparrow post^* (s)) \cap J \subseteq I^k$ donc on n'a pas la $k$-résilience "
-
-Alors qu'on aurait la $k$-résilience (i.e. $post^* (s) \cap J \subseteq I^k$)
-
-Pour $I^k$, le contre exemple prendrait du coup $pred(\uparrow \{ (q,b)\})$ qui n'est pas clos par le haut.
-
-On voudrait
-
-$post^*(s) \cap J \subseteq pred(\uparrow \{ (q,b)\})$
-
-Je propose de prendre $J$ = ensemble des configurations je pense le truc ici doit venir de
-$\uparrow post^* (s)$ being too big
-donc autant pas se restreindre du tout sur $J$
-
-$post^*(s) \subseteq pred(\uparrow \{ (q,b)\})$
-
-Si $s=(q,b)$ par exemple, et que on peut aller de $q$ vers $q$ en écrivant un 
-$b$ ? et évidement en perdant le dernier message (parce qu'on est lossy). Et c'est tout, disons.
-
-Déjà par lossiness $(q,\epsilon)$ serait dans $post^*((q,b))$.
-
-Et aussi dans $pred(\uparrow \{ (q,b)\})$ puisque $(q,\epsilon) \to (q,b)$.
-
-Donc on aurait:
-
-$post^*((q,b))  \subseteq pred(\uparrow \{ (q,b)\})$
-
-Et on aurait $(q,aaaaaa)$ dans 
-$\uparrow post^* ((q,b))$ puisque $(q,\epsilon)$ est dans $post^* ((q,b))$.
-
-Hors $(q,aaaaa)$ est pas dans $pred(\uparrow \{ (q,b)\})$
-(même si il est dans $pred^*(\uparrow \{ (q,b)\})$).
-
-}
 \fi
-
 \end{proof}
@@
 Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
+
+\mathieu{by symmetry} for downward compatible WSTS:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | c | c |}
+\hline  Hypothesis & strong downward compatibility ~ & $\downarrow post^*$ effective & both  \\ \hline
+   SRP & Undecidable  & Decidable   & Decidable \\ \hline
+   BSRP & Undecidable  &  ??  & Decidable \\ \hline
+      kSRP & Undecidable  & ?? & Decidable  \\ \hline
+ \end{tabular}
+\end{center}

### 2023-04-03 reachable-resilience.tex
up
details

--- 2023-04-03 reachable-resilience.tex
@@
 
-\problemx{Downward reachability problem}
+\problemx{Downward-reachability problem}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
@@
 %\"Ozkan
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
+\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]
 For finite-branching WSTS%with strong upward compatibility
@@
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
-$B_{\uparrow \post^*(s)} \cap D = \emptyset$, equivalent to $\post^*(s)\cap D = \emptyset$ by
-Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed
+$B_{\uparrow \post^*(s)} \cap D = \emptyset$, that is equivalent to $\post^*(s)\cap D = \emptyset$ by
+Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed sets
 $U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
@@
 
-The characterization is used to show that Petri nets are $\post^*$-effective. Since the downward reachability problem is decidable for Petri-nets, we deduce the following:
-
-\begin{proposition}
-Let $P = (N,M,W)$ be a Petri net and $s \in M$ a configuration of $P$, then one can compute a basis of $\uparrow \post^*(s)$.
-\end{proposition}
-
-\begin{proof}
-By Proposition~\ref{post*}, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-Let us now show that the downward-reachability problem is decidable for Petri nets.
-Let us consider a downward closed set $D$ in a Petri net.
-Let us consider $B$ a basis of the upward closed set $S \setminus D$.
-
-$D$ is  defined by ...
-states where your number of counter is bounded by a constant (deduce this from ...
-having to be smaller than elements of the basis $B$), and others where it is unbounded.
-You have $m$ counters where it is bounded by $n$
-then you have $m \times n$ possibilities for the values in these counters
-you have unlimited possibilities for the values of the other counters,
-but that is where the submarking problem comes in handy,
-i.e. you ask the submarking problem for all of the $m \times n$ possibilities.
-
-So downward-set reachability is decidable for PN.
-\end{proof}
-
+The characterization is used to show that Petri nets are $\post^*$-effective.
 It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}).
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
-Hence state-resilience is decidable for both Petri nets and VASS.
+Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
+Hence state-resilience is decidable for Petri nets and VASS.
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:
@@
 \begin{proof}
-Let us show that Reset VASS, that are WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one is able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
+Let us show that Reset VASS, that are effective WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
+Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether an element $m \in Min(\uparrow \post^*)$ is reachable.
 %in counter $i$ 
@@
 
-% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
-\begin{proof}
-See Appendix~\ref{appendix} for the complete proof.
-\end{proof}
+\mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
 
-Since the Reset Petri Net zero reachability problem is undecidable in Reset Petri-Nets, the reduction implies the following:
+Since reachability of $0$ is undecidable in Reset Petri-Nets, the reduction implies the following:
 
@@
 
-Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
+Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general reachable-resilience problem.
 
@@
 \begin{proof}[Sketch]
-\mathieu{ (Formalism should be made more explicit so that the proof can be really pertinent)}
-\textcolor{red}{
 Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
@@
 $(q, baaa) \to^* (q,bb)$ however it takes more than a single step.
+
 In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
@@
 $post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.
-}
 
@@
 
-\mathieu{In case we have  effective basis of $post^*$ and upward compatibility (not strong) then we don't know the decidability status of k-state-resilience and bounded-state-resilience.
-Not necessarily a big deal, but should write a comment or two about this.}
-
-\iffalse
 \begin{proposition}
@@
 \end{proof}
-\fi
 
@@
 have only been concerned with the case $\Safe = \uparrow \Safe$
-and $\Bad =\downarrow \Bad$, 
-\mathieu{pas tout à fait vrai: dans le technical report \cite{DBLP:conf/gg/Ozkan22} il considère aussi le cas où Bad et Safe sont tous les deux clos par le haut mais aussi le cas Safe clos par le bas Bad clos par le haut.}
- however as mentioned in Section~3
+and $\Bad =\downarrow \Bad$, however as mentioned in Section~3
 the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
 Hence it makes sense to consider this symmetrical case as well. \\
-
-\mathieu{rappeler résultats de \cite{DBLP:conf/gg/Ozkan22}}
-
-\begin{definition}{marked WSTS} 
-A {\em marked} WSTS is a tuple $(S, \leq, \to, INIT)$
-where $(S, \leq, \to)$ is a WSTS and $INIT \subseteq S$. 
-If $INIT$ is finite, we call it {\em fin-marked}.
-\end{definition}
-
-\begin{definition}
-A marked WSTS is
-\begin{itemize}
-\item lossy if $\downarrow \post^* (INIT) = \post^* (INIT)$,
-\item $\bot$-bounded (bottom-bounded) if there is a natural number 
-$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
-$s_B$ of a basis of $S$ with $s \geq s_B$.
-\end{itemize}
-\end{definition}
-
-\mathieu{}
-
-\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
-The bounded-state-resilience problem is decidable for \textcolor{red}{fin-marked} 
-strongly compatible WSTS
- which are 
- {\em \textcolor{red}{lossy} } 
- and 
- {\em \textcolor{red}{$\bot$-bounded}  }
- if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
-\end{theorem}
 
@@
 \begin{theorem}
-The state resilience problem is undecidable for effective WSTS with  strong  compatibility 
+The resilience problem is undecidable for effective WSTS with  strong  compatibility 
 when
@@
 \begin{proof}
-Consider Reset VASS
-in the case $\Safe$ contains only the minimal element %$0$ 
-and $\Bad = S \setminus \Safe$, 
-it follows the state resilience problem is equivalent 
-to the problem of whether $0$ can be visited infinitely often. 
-%
-% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
-Hence the undecidability.
+Taking $\Safe = \{ 0 \}$, $\Bad = S \setminus 0$, it follows the state resilience problem is equivalent
+to reachability of $0$. Since reachability of $0$ is undecidable for Reset VASS, we conclude.
 \end{proof}
 
-\mathieu{But what if we are in the case of $\uparrow \post^*$ effectiveness?}
+But what if we are in the case of $\uparrow \post^*$ effectiveness?
 
-\mathieu{Écrire que c'est globalement symmétrique au cas d'avant - je pense pas nécessairement qu'on arrive à faire beaucoup mieux que de juste dire ça.}
+LOREM IPSUM \\
 
@@
 
-\subsection{Synthesis}
+\subsection{Other}
 
-\iffalse
 \mathieu{Should try to put all results in a single table}
@@
 
-\fi
+\mathieu{Everything below should be moved to the applications section no?}
 
-Results in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
+% Let us deduce that effective basis of $post^*$ ex reset PN ?
 
-\begin{center}
-\begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
-   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
-   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
-      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
- \end{tabular}
-\end{center}
+ \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
 
-Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
+We may consider context-free grammars.
+
+\begin{corollary}
+{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
+\end{corollary}
+
+\begin{proof}
+
+  \textcolor{red}{à faire Mathieu}
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
+resilience decidable.
+\end{proof}
+
+\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}

### 2023-04-03 application.tex
details

@@
 open for VASS ???
+
+ \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
+
+We may consider context-free grammars.
+
+\begin{corollary}
+{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
+\end{corollary}
+
+\begin{proof}
+
+  \textcolor{red}{à faire Mathieu}
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
+resilience decidable.
+\end{proof}
+
+\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
 
@@
 
+\section{conclusion}
+
+\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
+%
+\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
+
 %

### 2023-04-03 reachable-resilience.tex
Reworking section 4.1

--- 2023-04-03 reachable-resilience.tex
@@
 % 		{$ ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe, ~ s' \rightarrow^{*} s''$ ?\newline}
-{$\Bad \cap \post^*(s_0)  \rightarrow^{*} \Safe $ ? \newline}
+{$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
 
@@
 
-Let us recall a result about state-resilience (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
+This stems from the fact that $t$-liveness is undecidable for WSTS and 
+is reducible to the state resilience problem.
+
+\begin{proposition}\label{reductions}
+In WSTS, $t$-liveness is reducible to the state resilience problem.
+%	home state and reachability
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness as the following formula
+% $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
+$$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$  
+where
+$U_t=\uparrow \pred(t)$
+is the upward closure of the preconditions to use transition $t$.  
+The problem reduces itself to the state resilience problem
+where $\Safe = U_t$ and $\Bad = S \setminus \Safe$.
+
+\end{proof}
+
+To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS.
+
+\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
+The  Reset VASS zero reachability problem can be reduced to the $t$-liveness problem.
+\end{theorem}
+
+% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
+\begin{proof}
+See Appendix~\ref{appendix} for the complete proof.
+\end{proof}
+
+Since the Reset VASS zero reachability problem is undecidable, the reduction implies 
+% the following:
+%
+\iffalse
+\begin{corollary}
+Reset VASS liveness is undecidable.
+\end{corollary}
+
+Since moreover liveness can be reduced to 
+$t$-liveness, we deduce:
+\fi
+%
+%\begin{corollary}
+$t$-liveness is undecidable for Reset VASS.
+%\end{corollary}
+
+\iffalse
+
+\begin{theorem}
+The downward reachability problem is reducible to the state resilience problem.
+\end{theorem}
+
+\begin{proof}
+Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
+Take $\Safe = \emptyset$ and $\Bad = D$.
+The state resilience problem on $\Safe,\Bad$ asks
+whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
+If the answer is negative then $s \to^* \Bad$,
+else $s \not\to^* \Bad$.
+\end{proof}
+
+\fi
+
+Since $t$-liveness is undecidable for Reset VASS that are WSTS with strong upward compatibility, from Proposition~\ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
+This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+\begin{theorem}\label{srp up down}
+{\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for strongly compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
+\end{theorem}
+
+\begin{proof}
+{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since $t$-liveness is undecidable.
+
+% \begin{proposition}
+% \end{proof}
+
+% \begin{proposition}
+In WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
+reducible to {\sc $k$-state-resilience}.
+% \end{proposition}
+% 
+% \begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+We compute 
+$n_0$, then iteratively check whether $k$-state resilience hold for $k$ from $0$ to $n_0$.  
+% \end{proof}
+Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem,
+% \end{proposition}
+%
+% \begin{proof}
+since 
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the %equivalence, and
+ undecidability of {\sc Bounded-state-resilience}
+and 
+ {\sc $k$-state-resilience}.
+
+\iffalse
+% \begin{corollary}\label{bsrp up down}
+Bounded state-resilience and 
+$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$.
+% \end{corollary}
+\fi
+
+%\begin{proof} 
+%  From the two reductions above. 
+\end{proof}
+
+On the positive side, let us recall a result about state-resilience decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
 
@@
 
-The characterization is used to show that Petri nets are $\post^*$-effective. Since the downward reachability problem is decidable for Petri-nets, we deduce the following:
+The characterization is used to show that VASS are $\uparrow \post^*$-effective. Since the downward-reachability problem is decidable for VASS, we deduce the following:
 
 \begin{proposition}
-Let $P = (N,M,W)$ be a Petri net and $s \in M$ a configuration of $P$, then one can compute a basis of $\uparrow \post^*(s)$.
+Let $V= (Q,T)$ be a $d$-VASS and $v \in \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(v)$.
 \end{proposition}
+%\alain{mauvaises notations, la tradition est que P est l'ensemble des places...les états/markings en minuscules,....}
 
@@
 Let us consider a downward closed set $D$ in a Petri net.
-Let us consider $B$ a basis of the upward closed set $S \setminus D$.
+Let us consider $B$ a basis of the upward-closed set $S \setminus D$.
 
@@
 
-It is well-known that Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
-Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
-Hence state-resilience is decidable for both Petri nets and VASS.
+It is well-known that 
+% Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
+% Since 
+VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
+Hence state-resilience is decidable for %both Petri nets and 
+VASS.
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:
@@
 \begin{proposition}
-There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
+There exist classes of WSTSs with strong (upward) compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
 \end{proposition}
@@
 \begin{proof}
-Let us show that Reset VASS, that are WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one is able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. Then, one would be able to decide whether $0$ is reachable 
+Let us show that Reset VASS, that are effective WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
+Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
+\iffalse
+Then, one would be able to decide whether an element $m \in Min(\uparrow \post^*)$ is reachable.
 %in counter $i$ 
@@
 %	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
+\fi
+Then, state-resilience would be decidable for Reset VASS.
 \end{proof}
@@
 %\end{proof}
-
-Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
-This stems from the fact that $t$-liveness is undecidable for reset Petri-Nets.
-
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem}
-
-\iffalse
-\begin{proof}
-If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
-reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
-reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
-$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
-
-The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
-and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
-
-We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $r_1$.
-
-The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
-
-Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
-and "freezing" the marking of $P_1$.
-
-Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
-The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
-
-Now we need to see if that marking is zero.
-
-We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
-
-If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
-$s$ fires in the marking zero.
-
-If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
-
-In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
-Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-
-Now if $0$ is reachable
-in $R(C_1, \mu_1)$ 
-then the net $C_2$ can also reach this marking
-in the place of $P_1$ by executing the same sequence of transition firings.
-Then $s$ can fire, freezing the $C_1$ subset.
-Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $0$ is reachable then $C_2$ is not live.
-
-Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
-$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
-(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
-% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
-% et r_2 commence sans token)
-If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero.
-
-\end{proof}
-\fi
-
-% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
-\begin{proof}
-See Appendix~\ref{appendix} for the complete proof.
-\end{proof}
-
-Since the Reset Petri Net zero reachability problem is undecidable in Reset Petri-Nets, the reduction implies the following:
-
-\begin{corollary}
-Reset Petri Net liveness is undecidable.
-\end{corollary}
-
-Since moreover liveness can be reduced to 
-$t$-liveness, we deduce:
-
-\begin{corollary}
-Reset Petri Net $t$-liveness is undecidable.
-\end{corollary}
-
-We now reduce $t$-liveness to the state resilience problem.
-
-\begin{proposition}\label{reductions}
-In WSTS, $t$-liveness is reducible to the state resilience problem.
-%	home state and reachability
-\end{proposition}
-
-\begin{proof}
-We reformulate $t$-liveness as the following formula
-% $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
-$$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$  
-where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
-The problem reduces itself to the state resilience problem
-where $\Safe = \uparrow \pred(t)$ and $\Bad = S \setminus \Safe$.
-
-\end{proof}
-
-\iffalse
-
-\begin{theorem}
-The downward reachability problem is reducible to the state resilience problem.
-\end{theorem}
-
-\begin{proof}
-Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
-Take $\Safe = \emptyset$ and $\Bad = D$.
-The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
-If the answer is negative then $s \to^* \Bad$,
-else $s \not\to^* \Bad$.
-\end{proof}
-
-\fi
-
-Since $t$-liveness is undecidable for reset Petri-Nets that are WSTS with strong upward compatibility, from Proposition \ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
-
-\begin{theorem}\label{srp up down}
-{\sc State-resilience} is undecidable for strongly compatible WSTS with effective pred-basis
-when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis, the bounded state-resilience problem is
-reducible to the $k$-state-resilience problem.
-\end{proposition}
-
-\begin{proof}
-% strong WSTS
-% $\pred^*(I) = \pred^{k}(I)$
-% Check all $k$ until $\pred^*(I) = \pred^{k}(I)$.
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-We compute 
-$n_0$, then iteratively check whether $k$-state resilience hold for $k$ from $0$ to $n_0$.  
-\end{proof}
-
-\begin{corollary}\label{bsrp up down}
-Bounded state-resilience and 
-$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
-when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{corollary}
-
-\begin{proof} 
-  From the two reductions above. 
-\end{proof}
 
@@
 \end{center}
+
+\mathieu{Montrer que leurs hypothèses Lossy + $\bot$-bounded + strong upward compatibility c'est plus fort que just strong downward compatibility et $\downarrow \post^*$ effective ?}

### 2023-04-03 reachable-resilience.tex
up

@@
 
-Since $t$-liveness is undecidable for Reset VASS that are WSTS with strong upward compatibility, from Proposition~\ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
+Since $t$-liveness is undecidable for Reset VASS that are WSTS with strong compatibility, from Proposition~\ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
 This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
@@
 In WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
-reducible to {\sc $k$-state-resilience}.
+reducible to {\sc $k$-state-resilience}:
 % \end{proposition}
@@
 % \begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
+since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
 compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
@@
 % \end{proof}
-Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded-state-resilience problem is equivalent to the state-resilience problem,
+Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
 % \end{proposition}
@@
 since 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
 Hence the %equivalence, and
@@
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
-{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
 when
@@
 \begin{corollary}\label{postcomputable}
-{\sc state-resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc state-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
 when
@@
 % in the regarded class 
-with strong upward compatibility 
+with strong compatibility 
 and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For finite-branching WSTS%with strong upward compatibility
+For finite-branching WSTS%with strong compatibility
 , a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
@@
 % in the regarded class 
-%with strong upward compatibility 
+%with strong compatibility 
 %and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
@@
 It is well-known that 
-% Petri nets are WSTS with strong upward compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
+% Petri nets are WSTS with strong  compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
 % Since 
-VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
 Hence state-resilience is decidable for %both Petri nets and 
@@
 \begin{proof}
-Let us show that Reset VASS, that are effective WSTSs with strong upward compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
+Let us show that Reset VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
 Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
@@
 \fi
-Then, state-resilience would be decidable for Reset VASS.
+Then, state-resilience would be decidable for Reset VASS, which lead to a contradiction.
 \end{proof}
@@
 %\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong upward compatibility.
+%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong compatibility.
 %\end{proposition}
@@
 
-Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
+Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
 
@@
 
-However when removing strong compatibility, precision is lost and the algorithm do not work anymore.
-
-\begin{proposition}
-The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-state-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}[Sketch]
+However when removing strong compatibility, precision is lost.
 Since $\pred(\uparrow \Safe)$ is not necessarily upward closed, it is possible to have 
@@
 An example of this is ...
-\mathieu{ (Lossy counter machine whose Formalism should be made more explicit so that the proof can be really pertinent)}
-\iffalse
-\textcolor{red}{
-Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
-For instance with a rule $\delta(q,b) = \text{ add some }b $,
-we have $(q,b) \to (q,bb)$, $(q, baaa) \geq (q,b)$ and
-$(q, baaa) \to^* (q,bb)$ however it takes more than a single step. 
-In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
-$post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$
-do not imply
-$\uparrow post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$ for a downward-closed set $D$.
-In particular by lossiness $(q, \epsilon)$ would belong to
-$post^* (s)$,
-hence $(q,a)$, $(q,aa)$, etc. would belong to $\uparrow post^* (s)$ although not necessarily to
-$post^* (s)$. $(q,aaa)$ not being in $pred(\uparrow \{ (q,b)\})$ means the algorithm 
-from \cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold in this example when $\Safe=\uparrow \{(q,b)\}$ and $\Bad =S$,
-because $\uparrow post^* (s) \cap S \not\subseteq \pred(\uparrow \{(q,b)\})$,
-despite having 
-$post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.
-}
-\fi
-\end{proof}
 
-\mathieu{In case we have  effective basis of $post^*$ and upward compatibility (not strong) then we don't know the decidability status of k-state-resilience and bounded-state-resilience.
-Not necessarily a big deal, but should write a comment or two about this.}
+Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of $k$-state-resilience and bounded-state-resilience.
 
-\iffalse
-\begin{proposition}
-\textcolor{red}{
-{\sc bounded resilience} is undecidable for WSTS with upward compatibility with an effective basis of $post^*$.
-CONJECTURE
-}
-\end{proposition}
-
-\begin{proof}
- \textcolor{red}{pas certain ! à faire Mathieu}
-\end{proof}
-\fi
+\mathieu{Until here}
 

### 2023-04-03 reachable-resilience.tex
up

@@
 \begin{proposition}
-There exist classes of WSTSs with strong (upward) compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
+There exist classes of WSTSs with strong 
+%(upward)
+ compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
 \end{proposition}
@@
 \fi
-Then, state-resilience would be decidable for Reset VASS, which lead to a contradiction.
+Then, {\sc State-resilience} would be decidable for Reset VASS, which leads to a contradiction.
 \end{proof}

### 2023-04-03 application.tex
details

@@
 
-\subsection{VASS, 2-VASS, semilinear VASS, Lossy Channel Systems, Integer et Continuous VASS, Regular Channel Systems}
+\subsection{Pushdown automata}
+
+%We may consider context-free grammars.
+
+\begin{corollary}
+{\sc resilience} is decidable for pushdown automata.
+\end{corollary}
+
+\begin{proof}
+pushdown automata $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ???? resilience decidable.
+
+  \textcolor{red}{à faire Mathieu}
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
+resilience decidable.
+\end{proof}
+%
+%
+\subsection{LCM, VASS, semilinear VASS, Integer et Continuous VASS}
 
 \begin{theorem}{}
-Resilience is decidable for semilinear VASS, LCM, Integer et Continuous VASS when $\Safe$ and $\Bad$ are semilinear sets.
+Resilience is decidable for LCM, semilinear VASS, Integer et Continuous VASS when $\Safe$ and $\Bad$ are semilinear sets.
 \end{theorem}
 
-se déduit du papier Schnoebelen RP 2010, page 8
+se déduit du papier Schnoebelen RP 2010, page 8.
+RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+
+open for VASS ???
+
+ \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
+
+\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
+
+\subsection{Regular Channel Systems}
 
@@
 
-open for VASS ???
+corollary: Lossy Channel Systems.
 
- \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-We may consider context-free grammars.
-
-\begin{corollary}
-{\sc resilience} is decidable for finite automata and context-free grammars (seen as transition systems)
-\end{corollary}
-
-\begin{proof}
-
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
-\end{proof}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
-
+%
+\iffalse
+%
 \subsection{Timed Automata}
@@
 
+%
+\fi
+%
+
 \section{conclusion}

### 2023-04-03 resilience.tex
details

@@
 
+Surprinsingly, the decidability of resilience problems was not studied in the previous papers.
+
 In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
 %		
-RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. We deduce 
+We deduce 
 % \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
@@
 
-{\bf Synthesis}\label{synthesis}
+{\bf Synthesis of the main results}\label{synthesis}
 

### 2023-04-03 application.tex
details

@@
-\section{Applications section}
+\section{Resilience in counter and fifo channel machines}
 
@@
 
-\begin{corollary}
-{\sc resilience} is decidable for pushdown automata.
-\end{corollary}
+\begin{theorem}
+The three {\sc resilience} problems are decidable for pushdown automata with $\Safe$ and $\Bad$ regular languages.
+\end{theorem}
 
 \begin{proof}
-pushdown automata $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ???? resilience decidable.
+Recall that resilience is equivalent to  $\Bad \subseteq \pred^*(\Safe)$.
+For every pushdown automaton $A=(Q,\Sigma,...)$ and every regular language $L \subseteq Q \times \Sigma^*$, the set $\pred^*(L)$ is a computable regular language \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
+%
+%			Moreover, $post^*(\Bad)$ is also a computable regular language \alain{pourquoi ?}   $post^*(L)$ and
+so the inclusion $\Bad \subseteq \pred^*(\Safe)$ is decidable and resilience too.
+%
+BRP
+%
+kRP
+%
+\end{proof}
+%
+Let us remark that $\uparrow post^*(s)$ is also a computable regular language, but we cannot directly apply Theorem \ref{} since PA are not WSTS for the usual orderings.
 
-  \textcolor{red}{à faire Mathieu}
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
+%  \textcolor{red}{à faire Mathieu}
+%Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
 resilience decidable.
-\end{proof}
+
 %
@@
 
-se déduit du papier Schnoebelen RP 2010, page 8.
-RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+\begin{proof}
+LCM se déduit du papier de Pachl (1986) that uses a regular inductive invariant to obtain a witness of non-reachability.
+
+Schnoebelen RP 2010, page 8. RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
+
+semilinear VASS
+
+Integer et Continuous VASS
+\end{proof}
 
 open for VASS ???
-
- \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
-\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
 
@@
 
+Another proof is possible for LCS which have a (non-computable) regular reachability set. LCS are WSTS with a computable $\uparrow post^*(S)$.
+
 corollary: Lossy Channel Systems.
+
+ \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
 

### 2023-04-03 reachable-resilience.tex
up
details

@@
 \begin{proof}
-Let $B_p$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
-and $\Bad$ decidable downward-closed.
+Let $B_s$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
+and $\Bad$ a downward-closed set given by its finite set of (maximal) ideals.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
 
-$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_p \cap \Bad \subseteq \pred^*(\Safe)$$
+$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_s \cap \Bad \subseteq \pred^*(\Safe)$$
 
-Since $B_p$ is finite and $\Bad$ is decidable, we can directly compute $ B_p \cap \Bad$.
+Since $B_s$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
 % For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_p \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
+We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_s \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
 \end{proof}
@@
 However when removing strong compatibility, precision is lost.
-Since $\pred(\uparrow \Safe)$ is not necessarily upward closed, it is possible to have 
+Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
  $\uparrow post^* (s) \cap S \not\subseteq \pred(\Safe)$,
@@
 which is incorrect.
-An example of this is ...
+% An example of this is ...
 

### 2023-04-04 application.tex
details

@@
 
-\subsection{Regular Channel Systems}
+\subsection{Regular Channel Machines}
+
+Fifo automata and systems of communicating finite-state machines (CFSMs) are essentially finite automata that communicate through fifo channels (or queues).
+In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set that gives a semi-algorithm to test non-reachability.
+As reachability is recursively enumerable, this provides an algorithm to solve reachability.
 
 \begin{theorem}{}
-Resilience is decidable for Regular Channel Systems when $\Safe$ and $\Bad$ are regular languages.
+Resilience is decidable for regular fifo automata and for systems of CFSMs when $\Safe$ and $\Bad$ are regular languages.
 \end{theorem}
 
-Another proof is possible for LCS which have a (non-computable) regular reachability set. LCS are WSTS with a computable $\uparrow post^*(S)$.
+Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.
 
-corollary: Lossy Channel Systems.
+\begin{corollary}
+Resilience is decidable for lossy channel systems.
+\end{corollary}
 
+Another proof is possible for LCS since LCS are WSTS with a computable $\uparrow post^*(S)$.
+%
  \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}

### 2023-04-04 reachable-resilience.tex
a first look at how 4.2 should be changed

@@
 
-The authors from \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
-have only been concerned with the case $\Safe = \uparrow \Safe$
-and $\Bad =\downarrow \Bad$, 
-\mathieu{pas tout à fait vrai: dans le technical report \cite{DBLP:conf/gg/Ozkan22} il considère aussi le cas où Bad et Safe sont tous les deux clos par le haut mais aussi le cas Safe clos par le bas Bad clos par le haut.}
- however as mentioned in Section~3
-the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
-Hence it makes sense to consider this symmetrical case as well. \\
-
-\mathieu{rappeler résultats de \cite{DBLP:conf/gg/Ozkan22}}
-
-\begin{definition}{marked WSTS} 
-A {\em marked} WSTS is a tuple $(S, \leq, \to, INIT)$
-where $(S, \leq, \to)$ is a WSTS and $INIT \subseteq S$. 
-If $INIT$ is finite, we call it {\em fin-marked}.
-\end{definition}
-
-\begin{definition}
-A marked WSTS is
-\begin{itemize}
-\item lossy if $\downarrow \post^* (INIT) = \post^* (INIT)$,
-\item $\bot$-bounded (bottom-bounded) if there is a natural number 
-$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
-$s_B$ of a basis of $S$ with $s \geq s_B$.
-\end{itemize}
-\end{definition}
-
-\mathieu{}
-
-\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
-The bounded-state-resilience problem is decidable for \textcolor{red}{fin-marked} 
-strongly compatible WSTS
- which are 
- {\em \textcolor{red}{lossy} } 
- and 
- {\em \textcolor{red}{$\bot$-bounded}  }
- if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
-\end{theorem}
-
-Unfortunately however the problem is undecidable in general.
+Unfortunately the problem is undecidable in general.
 
@@
 
-\mathieu{But what if we are in the case of $\uparrow \post^*$ effectiveness?}
-
-\mathieu{Écrire que c'est globalement symmétrique au cas d'avant - je pense pas nécessairement qu'on arrive à faire beaucoup mieux que de juste dire ça.}
+Despite this, "there is still hope". Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
+%
+For instance one can write the following lemma:
 
@@
 
+In the case of a WSTS with downward compatiblitity, not necessarily strong,
+then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
 Lemma~\ref{Lemma intersection 2} can be used to show that
-$\post^*(s) \cap \Bad \subseteq D$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq D$.
-
-If, as in Section~XX, we have the \textcolor{red}{added hypothesis}
-that
-$\pred^*(\Safe)$ is downward closed this means that
+if $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$,
+then
 $\post^*(s) \cap \Bad \subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$.
 
-And hence decidability follows.
+\begin{theorem}
+{\sc state-resilience} is decidable for WSTS with downward compatibility,
+$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+\end{theorem}
+
+\mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of 
+$\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}
+
+\begin{proof}
+Enumerate the downard-closed sets that contain $s$ and are closed under $\post$.
+For ‘each of them’
+check
+intersection with $\Bad \subseteq \pred^*(\Safe)$.
+If $(\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$
+then we know we will find some 
+member of the enumeration such that
+subseteq hold;
+Conversely
+compute
+
+$$ \post(s), ~ \downarrow \post(s), \downarrow \post^2(s) \ldots $$
+
+and ...
+
+check whether at one point we encounter some element in $\Bad$ but not in $\pred^*(\Safe)$.
+
+\end{proof}
+
+Strong downward compatibility implies furthermore the decidability
+of $k$-state-resilience and bounded-state-resilience.
+
+\iffalse	\mathieu{}
+
+ The result from \cite{DBLP:journals/corr/abs-2108-00889}
+only concerns with the case $\Safe = \uparrow \Safe$
+and $\Bad =\downarrow \Bad$, 
+ however as mentioned in Section~3
+the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
+Hence it makes sense to consider this symmetrical case as well. 
+ \cite{DBLP:conf/gg/Ozkan22} considers 
+ the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$ however the results are ...
+
+\begin{definition}[marked WSTS] 
+A {\em marked} WSTS is a tuple $(S, \leq, \to, s_0)$
+where $(S, \leq, \to)$ is a WSTS and $s_0 \in S$. 
+% If $INIT$ is finite, we call it {\em fin-marked}.
+\end{definition}
+
+\begin{definition}
+A marked WSTS is
+\begin{itemize}
+\item lossy if $\downarrow \post^* (s_0) = \post^* (s_0)$,
+\item $\bot$-bounded (bottom-bounded) if there is a natural number 
+$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
+$s_B$ of a basis of $S$ with $s \geq s_B$.
+\end{itemize}
+\end{definition}
+
+\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
+The bounded-state-resilience problem is decidable for marked
+strongly compatible WSTS
+ which are 
+ {\em \textcolor{red}{lossy} } 
+ and 
+ {\em \textcolor{red}{$\bot$-bounded}  }
+ if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
+\end{theorem}
+
+\mathieu{}	\fi
 

### 2023-04-04 reachable-resilience.tex
a first look at how 4.2 should be changed

@@
 \begin{theorem}
-{\sc state-resilience} is decidable for WSTS with downward compatibility,
+{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
@@
 \begin{proof}
-Enumerate the downard-closed sets that contain $s$ and are closed under $\post$.
-For ‘each of them’
-check
-intersection with $\Bad \subseteq \pred^*(\Safe)$.
-If $(\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$
-then we know we will find some 
-member of the enumeration such that
-subseteq hold;
-Conversely
-compute
 
-$$ \post(s), ~ \downarrow \post(s), \downarrow \post^2(s) \ldots $$
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
+Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
+Notice that, by standard monotonicity, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
 
-and ...
+Procedure 1 stops when it finds an invariant $D$ such that
+$D  \cap \Bad \subseteq \pred^*(\Safe)$. 
+Indeed
+$D  \cap \Bad \subseteq  \pred^*(\Safe)$ implies
+$\downarrow \post^*(s) \cap \Bad \subseteq  \pred^*(\Safe)$
+since $ \downarrow \post^*(s)  \subseteq D$.
 
-check whether at one point we encounter some element in $\Bad$ but not in $\pred^*(\Safe)$.
-
+The second procedure iteratively computes
+$post^{\leq n}(s) \cap \Bad$
+until it finds an element
+not in $\pred^*(\Safe)$.
 \end{proof}
@@
 
-\mathieu{by symmetry} for downward compatible WSTS:
+\mathieu{by symmetry} for downward compatible WSTS (but not necessarily upward compatible):
 
@@
 \mathieu{Montrer que leurs hypothèses Lossy + $\bot$-bounded + strong upward compatibility c'est plus fort que just strong downward compatibility et $\downarrow \post^*$ effective ?}
+
+\begin{center}
+\begin{tabular}{ | l | c | c | c | c |}
+\hline  Hypothesis & $\emptyset $ & downward compatible & strongly downward compatible  \\ \hline
+   SRP & Undecidable  & Decidable   & Decidable \\ \hline
+   BSRP & Undecidable  &  ??  & Decidable \\ \hline
+      kSRP & Undecidable  & ?? & Decidable  \\ \hline
+ \end{tabular}
+\end{center}

### 2023-04-04 application.tex
details

@@
 Recall that resilience is equivalent to  $\Bad \subseteq \pred^*(\Safe)$.
-For every pushdown automaton $A=(Q,\Sigma,...)$ and every regular language $L \subseteq Q \times \Sigma^*$, the set $\pred^*(L)$ is a computable regular language \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
+For every pushdown automaton $A=(Q,\Sigma,...)$ and every regular language $L \subseteq Q \times \Sigma^*$, the set $\pred^*(L)$ is a computable regular language \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}.
+Since $\Bad$ and $\Safe$ are both regular, we deduce that
 %
 %			Moreover, $post^*(\Bad)$ is also a computable regular language \alain{pourquoi ?}   $post^*(L)$ and
-so the inclusion $\Bad \subseteq \pred^*(\Safe)$ is decidable and resilience too.
+ the inclusion $\Bad \subseteq \pred^*(\Safe)$ is decidable and resilience too.
 %
@@
 %
-Let us remark that $\uparrow post^*(s)$ is also a computable regular language, but we cannot directly apply Theorem \ref{} since PA are not WSTS for the usual orderings.
+Let us remark that $\uparrow post^*(q,w)$ is also a computable regular language, but we cannot directly apply Theorem \ref{} since PA are not WSTS for the usual orderings (subword, prefix).
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow post^*$ is computable but context-free grammars are not strongly compatible.
 
 %  \textcolor{red}{à faire Mathieu}
-%Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ????
-resilience decidable.
 

### 2023-04-04 application.tex
details

@@
-\section{Resilience in counter and fifo channel machines}
+\section{Solving resilience with invariants}
 
@@
 %
-Let us remark that $\uparrow post^*(q,w)$ is also a computable regular language, but we cannot directly apply Theorem \ref{} since PA are not WSTS for the usual orderings (subword, prefix).
+Let us remark that $\uparrow post^*(q,w)$ is also a computable regular language, but we cannot directly apply previous results on WSTS since pushdown automata are not WSTS for the usual orderings (subword, prefix).
 Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow post^*$ is computable but context-free grammars are not strongly compatible.
@@
 %
-\subsection{LCM, VASS, semilinear VASS, Integer et Continuous VASS}
+
+\subsection{Regular fifo automata}
+
+Fifo automata and systems of communicating finite-state machines (CFSMs) are essentially finite automata that communicate through fifo channels (or queues).
+For the sake of simplicity, we consider automata with an unique fifo channel (that have the power of turing machines).
+In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability with a regular inductive invariant to obtain a witness of non-reachability Let $(q,w)$ be an initial state and $(q',w')$ be a state.
+Since $post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $post(L) \subseteq L$ (inductive invariant).
+We know that if $(q',w') \not\in post^*(q,w)$ then there exists, at least, a regular language $L$ such that $post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $post^*(q,w)$ is such inductive invariant.
+This strategy provides a semi-algorithm for non-reachability.
+As reachability is recursively enumerable, this provides an algorithm to solve reachability.
 
 \begin{theorem}{}
-Resilience is decidable for LCM, semilinear VASS, Integer et Continuous VASS when $\Safe$ and $\Bad$ are semilinear sets.
+State-resilience is decidable for regular fifo automata and for systems of CFSMs when $\Safe$ and $\Bad$ are regular languages.
 \end{theorem}
@@
 \begin{proof}
-LCM se déduit du papier de Pachl (1986) that uses a regular inductive invariant to obtain a witness of non-reachability.
-
-Schnoebelen RP 2010, page 8. RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}.
-
-semilinear VASS
-
-Integer et Continuous VASS
+$\Bad \subseteq \pred^*(\Safe)$ is decidable because $\pred^*(\Safe)$ is a computable regular language.
 \end{proof}
-
-open for VASS ???
-
-\subsection{Regular Channel Machines}
-
-Fifo automata and systems of communicating finite-state machines (CFSMs) are essentially finite automata that communicate through fifo channels (or queues).
-In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set that gives a semi-algorithm to test non-reachability.
-As reachability is recursively enumerable, this provides an algorithm to solve reachability.
-
-\begin{theorem}{}
-Resilience is decidable for regular fifo automata and for systems of CFSMs when $\Safe$ and $\Bad$ are regular languages.
-\end{theorem}
-
+%
 Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.
@@
  \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
+
+%%%
+\subsection{LCM, VASS, semilinear VASS, Integer et Continuous VASS}
+
+\begin{theorem}{}
+Resilience is decidable for LCM, semilinear VASS, Integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
+\end{theorem}
+
+\begin{proof}
+Resilience is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $Pre^*(\Safe)$ is a computable semilinear set, hence the inclusion between the two semilinear sets  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
+
+%		semilinear VASS.
+ \alain{le plus simple serait que $\pred^*(L)$ is semilinear because $post^*(s)$ is semilinear , peut-être en inversant le VASS V en V'on aurait pred*(V)=post*(V') hence $post^*(L)$ is also semilinear ?
+hence pred ????}
+
+%		Integer VASS
+Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
+
+\end{proof}
+
+We also deduce immediately that resilience is decidable for $2$-VASS that have a semilinear reachability relation.
+
+\begin{theorem}{}
+Resilience is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
+the rationals with addition and order.
+\end{theorem}
+
+\begin{proof}
+The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
+the rationals with addition and order whose complexity is EXPSPACE.
+Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE.
+\alain{and may be in coNP-complete.
+Recall that reachability is P-complete for continuous VASS}
+\end{proof}
+
+open for VASS ???
 

### 2023-04-04 application.tex
details

@@
 \section{Solving resilience with invariants}
+
+In this section, we study the general resilience problem.
 
@@
 %
-Let us remark that $\uparrow post^*(q,w)$ is also a computable regular language, but we cannot directly apply previous results on WSTS since pushdown automata are not WSTS for the usual orderings (subword, prefix).
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow post^*$ is computable but context-free grammars are not strongly compatible.
+Let us remark that $\uparrow \post^*(q,w)$ is also a computable regular language, but we cannot directly apply previous results on WSTS since pushdown automata are not WSTS for the usual orderings (subword, prefix).
+Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow \post^*$ is computable but context-free grammars are not strongly compatible.
 
@@
 For the sake of simplicity, we consider automata with an unique fifo channel (that have the power of turing machines).
-In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability with a regular inductive invariant to obtain a witness of non-reachability Let $(q,w)$ be an initial state and $(q',w')$ be a state.
-Since $post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $post(L) \subseteq L$ (inductive invariant).
-We know that if $(q',w') \not\in post^*(q,w)$ then there exists, at least, a regular language $L$ such that $post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $post^*(q,w)$ is such inductive invariant.
+In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability (with a regular inductive invariant to obtain a witness of non-reachability).
+Let $(q,w)$ be an initial state and $(q',w')$ be a state.
+Since $\post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $\post(L) \subseteq L$ (inductive invariant).
+We know that if $(q',w') \not\in \post^*(q,w)$ then there exists, at least, a regular language $L$ such that $\post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $\post^*(q,w)$ is such inductive invariant.
 This strategy provides a semi-algorithm for non-reachability.
-As reachability is recursively enumerable, this provides an algorithm to solve reachability.
+As reachability is recursively enumerable, this provides an algorithm to solve reachability for regular fifo automata.
+
+Let us say that a fifo automaton is \emph{post-regular} (resp.
+\emph{pred-regular}) if $\post^*(q,w)$ (resp.
+$\pred^*(q,w)$ ) is regular for all $(q,w) \in Q \times \Sigma^*$.
+The  fifo automaton is \emph{effectively} post-regular (resp.
+pred-regular) if the regular language $\post^*(q,w)$ (resp.
+$\pred^*(q,w)$) is computable.
+Reachability is (immediately) decidable for both effective post-regular and effective pred-regular fifo automata.
+\alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ?
+contre-exemples ?
+c'est le cas pour les LCM où pred ok mais pas post}
+%
+%
+%Let us say that a fifo automaton is \emph{regular} if the reachability relation is regular for all $(q,w)$; in this case, the fifo automaton is also both post-regular and pred-regular.
+%
+Now, we observe that $\Bad \subseteq \pred^*(\Safe)$ is decidable when $\pred^*(\Safe)$ is a computable regular language and $\Bad$ a regular language:
 
 \begin{theorem}{}
-State-resilience is decidable for regular fifo automata and for systems of CFSMs when $\Safe$ and $\Bad$ are regular languages.
+Resilience is decidable for effective pred-regular fifo automata when $\Safe$ and $\Bad$ are regular languages.
 \end{theorem}
 
-\begin{proof}
-$\Bad \subseteq \pred^*(\Safe)$ is decidable because $\pred^*(\Safe)$ is a computable regular language.
-\end{proof}
+%		\begin{proof}
+%		$\Bad \subseteq \pred^*(\Safe)$ is decidable because $\pred^*(\Safe)$ is a computable regular language.
+%		\alain{prover que $\pred^*(\Safe)$ is a computable regular language}
+%		\end{proof}
 %
-Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.
+%			Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.
+
+Since for  lossy channel systems $\pred^*(L)$ with $L=\uparrow L$ is a computable (upward-closed) regular language, we deduce :
 
 \begin{corollary}
-Resilience is decidable for lossy channel systems.
+Resilience is decidable for lossy channel systems with $\Safe=\uparrow \Safe$ and any regular language $\Bad$.
 \end{corollary}
 
-Another proof is possible for LCS since LCS are WSTS with a computable $\uparrow post^*(S)$.
+Another proof is possible for LCS since LCS are WSTS with a computable $\uparrow \post^*(S)$ \alain{à prouver et vérifier les autres hypothèses}
 %
- \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}
-
 %%%
@@
 \begin{proof}
-Resilience is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $Pre^*(\Safe)$ is a computable semilinear set, hence the inclusion between the two semilinear sets  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
+Resilience is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $Pre^*(\Safe)$ is a computable semilinear set.
+Hence since the inclusion between two semilinear setsis decidable, we deduce that  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
 

### 2023-04-05 application.tex
details

@@
 %
-BRP
-%
-kRP
+The BRP and the kRP are also decidable.
+The algorithm that computes $ \pred^*(\Safe)$ is polynomial.
+If $\Bad \subseteq \pred^*(\Safe)$ then one computes an integer $k$ such that the maximum of all minimal pathes that go from $\Bad$ to $\Safe$ is equal to $k$.\alain{à prouver}
 %
@@
 \begin{theorem}{}
-Resilience is decidable for LCM, semilinear VASS, Integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
+Resilience is decidable for  lossy counter machines, semilinear VASS, Integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
 \end{theorem}
@@
 \begin{proof}
-Resilience is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $Pre^*(\Safe)$ is a computable semilinear set.
-Hence since the inclusion between two semilinear setsis decidable, we deduce that  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
+Resilience is decidable for lossy counter machines with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $\pred^*(\Safe)$ is a computable semilinear set.
+Hence since the inclusion between two semilinear sets is decidable, we deduce that  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
 
 %		semilinear VASS.
- \alain{le plus simple serait que $\pred^*(L)$ is semilinear because $post^*(s)$ is semilinear , peut-être en inversant le VASS V en V'on aurait pred*(V)=post*(V') hence $post^*(L)$ is also semilinear ?
+ \alain{le plus simple serait que $\pred^*(L)$ is semilinear because $\post^*(s)$ is semilinear , peut-être en inversant le VASS V en V'on aurait pred*(V)=post*(V') hence $post^*(L)$ is also semilinear ?
 hence pred ????}

### 2023-04-05 application.tex
details

@@
 %
-Now, we observe that $\Bad \subseteq \pred^*(\Safe)$ is decidable when $\pred^*(\Safe)$ is a computable regular language and $\Bad$ a regular language:
+Now, we just observe that $\Bad \subseteq \pred^*(\Safe)$ is decidable when $\pred^*(\Safe)$ is a computable regular language and $\Bad$ a regular language:
 

### 2023-04-05 reachable-resilience.tex
up

@@
 where $\Safe = U_t$ and $\Bad = S \setminus \Safe$.
-
 \end{proof}
@@
 To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS.
+The following theorem is a adjustment ofTheorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.
 
@@
 
+\iffalse
 % \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
@@
  {\sc $k$-state-resilience}.
-
+%
+%
 \iffalse
@@
 \fi
-
+%
 %\begin{proof} 
@@
 
+The proof of Theorem~\ref{ref ozkan} rely on the computability of $\uparrow \post^*(s)$ and on the following lemma.
+
+\begin{lemma}\label{Lemma intersection}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
+Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
+Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\uparrow A}$, the inclusion
+$({\uparrow  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
+\end{proof}
+
 \iffalse
@@
 
-The computability of $\uparrow \post^*(s)$ seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
 Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
@@
 Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
-
-\begin{lemma}\label{Lemma intersection}
-Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
-\end{lemma}
-
-\begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
-Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
-Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\uparrow A}$, the inclusion
-$({\uparrow  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
-\end{proof}
 
@@
 
+Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt
+
 \begin{center}

### 2023-04-05 resilience.tex
details
up

@@
 
-In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
 %		
@@
 
+\alain{simplifier le tableau en enlevant les deux dernières colonnes}
 \begin{center}

### 2023-04-05 reachable-resilience.tex
up

@@
 \begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
-{\sc state-resilience} is decidable for (not necessarily strong) WSTS with effective 
+{\sc state-resilience} is decidable for 
+%(not necessarily strong)
+ WSTS with effective 
 $\uparrow$ $\post^*$ basis
@@
 
-Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of $k$-state-resilience and bounded-state-resilience.
-
-\mathieu{Until here}
+Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
+{\sc bounded-state-resilience}.
 
@@
 \begin{theorem}
-The state resilience problem is undecidable for effective WSTS with  strong  compatibility 
+{\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
 when
@@
 Consider Reset VASS
-in the case $\Safe$ contains only the minimal element %$0$ 
+in the case $\Safe$ contains only the minimal element $Min(S) = 0$ %$0$ 
 and $\Bad = S \setminus \Safe$, 
@@
 to the problem of whether $0$ can be visited infinitely often. 
-%
 % Since reachability of $0$ is undecidable for Reset VASS, we conclude.
@@
 
-Despite this, "there is still hope". Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
 is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
@@
 \begin{proof}
-Let us suppose that $A \cap U \subseteq D$. Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$.
+Let us suppose that $A \cap U \subseteq D$.
+Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$.
 Let us show that $({\downarrow A}) \cap U \subseteq {\downarrow (A \cap U)}$.
@@
 Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
-In the other direction,
-since $A \subseteq {\downarrow A}$, the inclusion
-$({\downarrow  A}) \cap U \subseteq D$ implies
-$A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
+In the other direction, since $A \subseteq {\downarrow A}$, the inclusion $({\downarrow  A}) \cap U \subseteq D$ implies $A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
 \end{proof}
@@
 
-\mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of 
-$\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}
-
 \begin{proof}
-
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
+In order to decide whether state-resilience hold, we execute two procedures in parallel,
+one looking for a resilience certificate and one looking for a non-resilience certificate.
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
 Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
@@
 
+% \mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of $\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}
+
 Strong downward compatibility implies furthermore the decidability
-of $k$-state-resilience and bounded-state-resilience.
+of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.
 
-\iffalse	\mathieu{}
-
- The result from \cite{DBLP:journals/corr/abs-2108-00889}
-only concerns with the case $\Safe = \uparrow \Safe$
-and $\Bad =\downarrow \Bad$, 
- however as mentioned in Section~3
-the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
-Hence it makes sense to consider this symmetrical case as well. 
- \cite{DBLP:conf/gg/Ozkan22} considers 
- the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$ however the results are ...
-
-\begin{definition}[marked WSTS] 
-A {\em marked} WSTS is a tuple $(S, \leq, \to, s_0)$
-where $(S, \leq, \to)$ is a WSTS and $s_0 \in S$. 
-% If $INIT$ is finite, we call it {\em fin-marked}.
-\end{definition}
-
-\begin{definition}
-A marked WSTS is
-\begin{itemize}
-\item lossy if $\downarrow \post^* (s_0) = \post^* (s_0)$,
-\item $\bot$-bounded (bottom-bounded) if there is a natural number 
-$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
-$s_B$ of a basis of $S$ with $s \geq s_B$.
-\end{itemize}
-\end{definition}
-
-\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
-The bounded-state-resilience problem is decidable for marked
-strongly compatible WSTS
- which are 
- {\em \textcolor{red}{lossy} } 
- and 
- {\em \textcolor{red}{$\bot$-bounded}  }
- if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
-\end{theorem}
-
-\mathieu{}	\fi
+\begin{proposition}
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
+$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+\end{proposition}
 
 \subsection{Synthesis}
-
-\iffalse
-\mathieu{Should try to put all results in a single table}
-
-With only strong compatibility hypothesis:
-
-\begin{center}
-\begin{tabular}{ | l | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
-   SRP & Undecidable (Thm~\ref{srp up down}) & Undecidable  \\ \hline
-   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  \\ \hline
-      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? \\ \hline
- \end{tabular}
-\end{center}
-
-With $\uparrow \post^*$ effectiveness hypothesis:
-
-\begin{center}
-\begin{tabular}{ | l | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
-   SRP & Decidable (Thm~\ref{post srp}) & ??  \\ \hline
-   BSRP & ?? &  ??  \\ \hline
-      kSRP & ?? & ?? \\ \hline
- \end{tabular}
-\end{center}
-
-With $\uparrow \post^*$ effectiveness and strong compatibility hypothesis:
-
-\begin{center}
-\begin{tabular}{ | l | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
-   SRP & Decidable (Thm~\ref{postcomputable})& ??  \\ \hline
-   BSRP & Decidable (Thm~\ref{ref ozkan})&  ??  \\ \hline
-      kSRP & Decidable (Thm~\ref{ref ozkan}) & ?? \\ \hline
- \end{tabular}
-\end{center}
-
-\fi
 
@@
 
+\noindent
 Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
-
-\mathieu{by symmetry} for downward compatible WSTS (but not necessarily upward compatible):
+\iffalse
+(by symmetry) for downward compatible WSTS (but not necessarily upward compatible):
 
@@
 Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt
+\fi
 

### 2023-04-05 reachable-resilience.tex
up

@@
 
-\begin{theorem}
+\begin{theorem}\label{srp down up}
 {\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
@@
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
-   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
-      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
  \end{tabular}
@@
 \hline  Hypothesis & $\emptyset $ & downward compatible & strongly downward compatible  \\ \hline
-   SRP & Undecidable  & Decidable   & Decidable \\ \hline
-   BSRP & Undecidable  &  ??  & Decidable \\ \hline
-      kSRP & Undecidable  & ?? & Decidable  \\ \hline
+   SRP & Undecidable (Thm~\ref{srp down up})  & Decidable   & Decidable \\ \hline
+   BSRP & ??  &  ??  & Decidable \\ \hline
+      kSRP & ??  & ?? & Decidable  \\ \hline
  \end{tabular}

### 2023-04-05 reachable-resilience.tex
up

@@
 
-\begin{theorem}
+\begin{theorem}\label{downward srp}
 {\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,
@@
 
-\begin{proposition}
+\begin{proposition}\label{downward brp}
 {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
@@
 \hline  Hypothesis & $\emptyset $ & downward compatible & strongly downward compatible  \\ \hline
-   SRP & Undecidable (Thm~\ref{srp down up})  & Decidable   & Decidable \\ \hline
-   BSRP & ??  &  ??  & Decidable \\ \hline
-      kSRP & ??  & ?? & Decidable  \\ \hline
+   SRP & Undecidable (Thm~\ref{srp down up})  & Decidable (Thm~\ref{downward srp})  & Decidable (Thm~\ref{downward srp}) \\ \hline
+   BSRP & ??  &  ??  & Decidable (Prop~\ref{downward brp}) \\ \hline
+      kSRP & ??  & ?? & Decidable (Prop~\ref{downward brp}) \\ \hline
  \end{tabular}

### 2023-04-06 resilience.tex
Adding examples to sec3 for now

@@
-%%%
-\iffalse
-%%%%
-%%%%
-\subsection{Defining resilience}
-
-\subsubsection{Resilience for general transition systems}
-
-We ask whether we can reach a state 
-%which satisfies
-in 
-%
-$\Safe$  in a reasonable amount of time whenever we reach a state 
-% which satisfies
-in
-%
-$\Bad$. 
-From this we formulate two resilience problems. First consider the case where the recovery time
-is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.
-
-\problemx{$k$-Resilience problem }
-{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
-If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.
-
-% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(\Bad → 0≤ j≤k EX j \Safe).
-
-We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.
-
-%
-%\problemx{Bounded Resilience problem}
-%{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-%{$\exists k \geq 0 ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-
-\problemx{Bounded Resilience problem}
-{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
-{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
-%%
-
-%%%
-
-pour $k,k'$ donnés, chercher si $\Safe_{max}$ et $\Bad_{max}$ ont un sens ?
-on peut faire l'union des \Safe, \Safe' et des \Bad, \Bad' prendre le max des k.
-Fixer k, chercher et calculer $\Safe_{max}$ et $\Bad_{max}$.
-
-%
-%\alain{dire juste que \Safe = upward closed et \Bad = downward closed.
-%envisager les 3 autres cas. souvent \Bad = upward closed (par ex l'exclusion mutuelle).
-%4 resilience problem (U,D), (U,U), (D,U), (D,D)}
-
-\fi
-
-\newcommand{\Bad}{\textsf{Bad}}
-\newcommand{\Safe}{\textsf{Safe}}
-
 \section{Resilience for WSTS}
@@
 
+\begin{example}
+% Avec Safe = up( q_1(3,0) ) et Bad = q_0 (N,N) on aurait la 1-résilience et donc la résilience par exemple
+% mais avec Safe = up(q_0(3,0)) et Bad = q_1(N,N) on ne l'aurait pas à cause de q_1(0,0) 
+Let $V = (\{q_0,q_1\}, \{ (q_0, (3,0), q_1), (q_1, (-1,1), q_0)\})$ be a vector addition system with states (VASS),
+and thus a WSTS.
+Case A) is $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.
+Case B) is $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience do not hold because of $q_1(0,0) \in \Bad$.
+% (si on veut se rajouter une hypothèse comme pas de dead-end on peut rajouter un troisième état q_2 duquel on puisse partir depuis q_1 en faisant par exemple +1,+1 et qui loop sur lui-même) 
+\end{example}
+
+\begin{center}
+	\begin{figure}
+			\hspace{3.75cm}
+\includegraphics[width=0.2\textwidth]{FigureA}
+	\label{example_1}
+	\end{figure}
+\end{center}
+
+\begin{example}
+% Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
+$\mathds{Z}$-VASS, one counter, $\Safe = \uparrow 0$, $\Bad = \downarrow -1$, transition: $+1$,
+then resilience hold but not bounded resilience. 
+\end{example}
+
 \input{resilience-safe-up}

### 2023-04-06 resilience.tex
up

@@
 Case A) is $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.
-Case B) is $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience do not hold because of $q_1(0,0) \in \Bad$.
+Case B) is $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience does not hold because of $q_1(0,0) \in \Bad$.
 % (si on veut se rajouter une hypothèse comme pas de dead-end on peut rajouter un troisième état q_2 duquel on puisse partir depuis q_1 en faisant par exemple +1,+1 et qui loop sur lui-même) 
@@
 % Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
-$\mathds{Z}$-VASS, one counter, $\Safe = \uparrow 0$, $\Bad = \downarrow -1$, transition: $+1$,
-then resilience hold but not bounded resilience. 
+Let us consider now a $\mathds{Z}$-VASS with only one state, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
+then resilience hold (one always ends up in the positive integers) but not bounded resilience
+(for every bound $b$ there is an element of $\Bad$ e.g. $q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$). 
 \end{example}

### 2023-04-06 resilience.tex
up

@@
 Let $V = (\{q_0,q_1\}, \{ (q_0, (3,0), q_1), (q_1, (-1,1), q_0)\})$ be a vector addition system with states (VASS),
-and thus a WSTS.
-Case A) is $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.
-Case B) is $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience does not hold because of $q_1(0,0) \in \Bad$.
+and thus a WSTS (see Figure~\ref{example_1}).
+In the case $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.
+On the other hand in the case $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience does not hold because of $q_1(0,0) \in \Bad$.
 % (si on veut se rajouter une hypothèse comme pas de dead-end on peut rajouter un troisième état q_2 duquel on puisse partir depuis q_1 en faisant par exemple +1,+1 et qui loop sur lui-même) 
@@
 			\hspace{3.75cm}
+				\label{example_1}
 \includegraphics[width=0.2\textwidth]{FigureA}
-	\label{example_1}
+	\caption{}
 	\end{figure}

### 2023-04-06 resilience.tex
up

@@
 % Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
-Let us consider now a $\mathds{Z}$-VASS with only one state, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
+Let us consider now a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
 then resilience hold (one always ends up in the positive integers) but not bounded resilience

### 2023-04-06 resilience.tex
up

@@
 Let $V = (\{q_0,q_1\}, \{ (q_0, (3,0), q_1), (q_1, (-1,1), q_0)\})$ be a vector addition system with states (VASS),
-and thus a WSTS (see Figure~\ref{example_1}).
+and thus a WSTS (see Figure~1).
 In the case $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.

### 2023-04-06 reachable-resilience.tex
details
up

@@
 \subseteq S$.}
-{$\exists s' \in D ~ s \to^* s'$? \newline}
+% {$\exists s' \in D ~ s \to^* s'$? \newline}\alain{ecrire $s  \to^* D$}
+{$s  \to^* D$? \newline}
 
@@
 \begin{proposition}
-Let $V= (Q,T)$ be a $d$-VASS and $v \in \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(v)$.
+Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.
 \end{proposition}

### 2023-04-07 application.tex
details

@@
 open for VASS ???
+
+\section{Computing the largest safe WSTS}
+
+Ici $\Safe$ clos haut.
+ici on veut rester toujours dans $\Safe$.
+on calcule les élements minimaux de $\pred^*(\Safe)=\{m_1,m_2,...,m_p\}$ puis on ajoute les préconditions $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ à toutes les transitions $s \rightarrow s'$ assurant que l'état atteint est au dessus des minimaux, ainsi on sera toujours dans $\Safe$.
+Il faut juste l'hypothèse raisonnable qu'on peut exprimer $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ comme une garde d'une transition (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given a WSTS and $\Safe$  upward-closed, the largest sub TS, that stays in $\Safe$ , is a computable WSTS.
+\end{theorem}
+
+Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(\Safe)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $\Safe$ .
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
+Dans ce cas, ça permettrait de résoudre autrement les 4 pbs dans \cite{DBLP:conf/rp/BozzelliG11} avec une complexité 2-exp plutôt qu'ackerman
+
+\begin{theorem}{}
+Given a VASS and $\Safe$ upward-closed, the largest sub TS, that  stays in $\Safe$, is a computable VASS of size 2-exp.
+\end{theorem}
 

### 2023-04-07 application.tex
details

@@
 Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(\Safe)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $\Safe$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
-Dans ce cas, ça permettrait de résoudre autrement les 4 pbs dans \cite{DBLP:conf/rp/BozzelliG11} avec une complexité 2-exp plutôt qu'ackerman
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
+Dans ce cas, ça permettrait de résoudre autrement les 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman
 

### 2023-04-07 application.tex
details

@@
 Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
-Dans ce cas, ça permettrait de résoudre autrement les 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman
+Dans ce cas, ça permettrait de résoudre autrement au moins les 3 premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman
 

### 2023-04-07 application.tex
details

--- 2023-04-07 application.tex
@@
 Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
-Dans ce cas, ça permettrait de résoudre autrement les 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman
+Dans ce cas, ça permettrait de résoudre autrement les 3 premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
+problèmes: marquage T-bloqué, dead, borné , T-continual.
+Je pense qu'au moins
 

### 2023-04-07 definitions.tex
details
up

@@
+\newcommand{\Bad}{\textsf{Bad}}
+\newcommand{\Safe}{\textsf{Safe}}
+
 \newcommand{\pred}{\textsf{pred}}
@@
 
-\subsection{Transition systems}
+\noindent
+{\bf Transition systems}
 
 \begin{definition}
-An {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
+A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
 {\em states} and  
@@
 The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$.
+$\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. \\
 
 %
-\alain{enlever ce qui est inutile}
-%
-We use similar notation for $\post$ and $\pred$...
-%
-Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
-then defined as expected.
-
-%\textcolor{red}{Thinking about whether or not it is pertinent to have LTS and not only TS. LTS can be usefull for TA because of the use of the guards/time as labels but it may be unecessary.}
-
-%\textcolor{red}{This makes sense for TS but not so much for LTS ...}
-
-A transition system is {\em finitely branching} if all $\post(s)$ are finite. 
-%\textcolor{blue}{We restrict our attention to finitely branching TSs.}
-
-%\textcolor{blue}{Alain: the forward coverability algorithm for infinitely branching TSs.; the backward cov algo may work for essentially finitely branching TSs.}
-%\textcolor{red}{Not sure that TS induced by TA are finitely branching.
-Actually I believe they are not, i.e.
-for instance for a TA with one clock $x$, from a state $q$ and clock $x$ set at $0$, if there is a transition e.g.
-$(q, x \geq 3, \emptyset, q')$ then the set of successors of
-%$(q,0)$ is $\{q'\} \times \{3, 4, 5, \ldots \}$. Need to check where finitely branching appears as an assumption/requirement.}
-
-%
-%\begin{definition}
-%A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
-%$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
-%${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
-%ternary relation,
-%denoted as the set of {\em labeled transitions}. 
-%\end{definition}
-
-We
- prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
-       $s  \xrightarrow{a}  s'$
-to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\
 
 \noindent
-%Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.
-
-A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
-to a {\em target configuration} $s_n$ is a sequence 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$.
-
-We define the {\em concatenation} $ \pi_1 \pi_2$ of 
-two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
-as expected.
-The {\em length} of 
-$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
-is defined as $|\pi|=n$ and we say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
-For all $w \in \Sigma^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.
-
-An {\em infinite path} is an infinite sequence
- $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
-For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
-(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
-and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
-by $\pi[i,j]$ the path 
-$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
- and by $\pi[i]$ the configuration $s_i$.
-As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
-a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
-Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
-$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.
-
-A labeled transition system $(S, \Sigma, \rightarrow)$ is {\em deterministic} if for all states $s_1, s_2, s_3 \in S$ and all
-$a \in \Sigma$,
- $ s_1  \xrightarrow{a} s_{2}$ and  
- $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\
-
-%
-\alain{enlever ce qui est inutile}
-%
-
-\subsection{Well-quasi-orderings}
+{\bf Well-quasi-orderings}
 
@@
 
-\subsection{Well-structured transition systems}
+\noindent
+{\bf Well-structured transition systems}
 
@@
 
+\iffalse
 \alain{ me semble maintenant inutile
@@
 \end{proof}}
+
+\fi

### 2023-04-07 resilience.tex
up

@@
 
-\alain{simplifier le tableau en enlevant les deux dernières colonnes}
+% \alain{simplifier le tableau en enlevant les deux dernières colonnes}
+
 \begin{center}
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ & $\downarrow$~ $\downarrow$~ \\ \hline
-   RP & Decidable (Thm~\ref{up-up}) & Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{up-down}) & Undecidable (Thm~\ref{down-down}) \\ \hline
-   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Thm~\ref{down-up}) & ?? & ?? \\ \hline
-      kRP & Decidable (Thm~\ref{k-up-up}) & Decidable (Thm~\ref{down-up}) & ?? & ?? \\ \hline
+\hline   \Safe~\Bad %& $\uparrow$~ $\uparrow$~ 
+		& $\uparrow$~ $\downarrow$~ 
+		 & $\downarrow$~ $\uparrow$~ 
+		 %& $\downarrow$~ $\downarrow$~
+ \\ \hline
+   RP %& Decidable (Thm~\ref{up-up}) 
+   	& Decidable (Thm~\ref{down-up})  
+   		 & Decidable (Thm~\ref{up-down}) 
+   	%	 & Undecidable (Thm~\ref{down-down})
+    \\ \hline
+   BRP %& Decidable (Corollary~\ref{B-up-up}) 
+   &  Decidable (Thm~\ref{down-up}) 
+   		 & ?? 
+   	%	 & ??
+    \\ \hline
+      kRP %& Decidable (Thm~\ref{k-up-up}) 
+      & Decidable (Thm~\ref{down-up}) 
+      		& ?? 
+      	%	& ??
+       \\ \hline
  \end{tabular}

### 2023-04-09 reachable-resilience.tex
details
up

@@
 \begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & $\emptyset $ & downward compatible & strongly downward compatible  \\ \hline
+\hline  Hypothesis &  none  & downward compatible & strongly downward compatible  \\ \hline
    SRP & Undecidable (Thm~\ref{srp down up})  & Decidable (Thm~\ref{downward srp})  & Decidable (Thm~\ref{downward srp}) \\ \hline
@@
 \end{center}
+% \alain{une hypothèse n'est pas l'ensemble vide}

### 2023-04-10 reachable-resilience.tex
details

@@
 
-As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downard closed and upward closed. We will consider here the cases where 
+As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
 $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
@@
 This stems from the fact that $t$-liveness is undecidable for WSTS and 
-is reducible to the state resilience problem.
+is reducible to the state resilience problem.\alain{definir t-liveness, dire ce qu'est une transition $t$ dans un WSTS étiqueté ou pas ?...}
+
+Let us define $pre(t)=\{ u \in S \mid \exists u' \in S$ in a labeled WSTS such that $u \xrightarrow{t} u'$\}
+
+\alain{ou alors $t=(u,u')$ such that $u \rightarrow u'$ mais alors c'est bizarre de nommer par la même lettre $t$ des couples différents $(u,u'), (v,v'),...$ dans un WSTS...ou alors on appelle transition un couple $(u,u')$ such that $u \rightarrow u'$ mais plus de lettre $t$...ou on définit un WSTS par un nombre fini de (shéma de ) transition $t_i$ qui sont compatibles par compatibilité mais ça contraint à un nombre fini (comme dans les VASS) ou alors on ne parle ici que des VASS et pas des WSTS mais ça suffira puisque state resilience pour VASS implique pour WSTS BREF il faut réfléchir à ce qu'est une transition...}
 
@@
 \begin{proof}
-We reformulate $t$-liveness as the following formula
+We reformulate $t$-liveness in a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ as the following formula
 % $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
@@
 where
-$U_t=\uparrow \pred(t)$
-is the upward closure of the preconditions to use transition $t$.  
+$U_t=\uparrow \pred(t)$ \alain{$\pred(t)$ n'est pas défini et $\pred$ est une mauvaise notation pour ça: plutôt: $pre(t)=\{ u \in S \mid \exists u' \in S$  such that $u \xrightarrow{t} u'$\} }
+is the upward closure of the preconditions to fire transition $t$.  
 The problem reduces itself to the state resilience problem
@@
 
+\alain{tu ne définis pas, même informellement et/ou rapidement, assez les objets que tu utilises: zero reachability problem, Reset VASS}
+
 \begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
-The  Reset VASS zero reachability problem can be reduced to the $t$-liveness problem.
+The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
 \end{theorem}
@@
 
-Since the Reset VASS zero reachability problem is undecidable, the reduction implies 
+Since the zero reachability problem is undecidable for Reset VASS, the reduction implies 
 % the following:
@@
 \begin{corollary}
-Reset VASS liveness is undecidable.
+Liveness is undecidable in Reset VASS.
 \end{corollary}
@@
 We compute 
-$n_0$, then iteratively check whether $k$-state resilience hold for $k$ from $0$ to $n_0$.  
+$n_0$, then iteratively check whether $k$-state resilience \alain{$k$-state resilience ou $k$-state-resilience ?}
+hold for $k$ from $0$ to $n_0$.  
 % \end{proof}
@@
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
-$B_{\uparrow \post^*(s)} \cap D = \emptyset$, that is equivalent to $\post^*(s)\cap D = \emptyset$ by
+$B_{\uparrow \post^*(s)} \cap D = \emptyset$, \alain{rapeler $B_{\uparrow \post^*(s)}$}  that is equivalent to $\post^*(s)\cap D = \emptyset$ by
 Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed sets
 $U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
-Decidability of downward reachability leads to the decidability of the following stop condition:
+Decidability of downward-reachability leads to the decidability of the following stop condition:
 asking whether $S \setminus U_n$ is reachable from $s$.
 
-The characterization is used to show that VASS are $\uparrow \post^*$-effective. Since the downward-reachability problem is decidable for VASS, we deduce the following:
+The characterization is used to show that VASS are $\uparrow \post^*$-effective.
+Since the downward-reachability problem is decidable for VASS \alain{ref ?}, we \alain{we ou Ozkan ?
+qui a fait la preuve ?} deduce the following:
 
 \begin{proposition}
-Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.
+Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.\alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
 \end{proposition}
@@
 \begin{proof}
-Let $B_s$ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
+Let $B_s$ \alain{notation fluctuante} be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
 and $\Bad$ a downward-closed set given by its finite set of (maximal) ideals.
@@
 $$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_s \cap \Bad \subseteq \pred^*(\Safe)$$
+\alain{$\equiv $  ???}
 
@@
 Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
- $\uparrow post^* (s) \cap S \not\subseteq \pred(\Safe)$,
+ $\uparrow \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
 despite having 
-$post^* (s) \cap S \subseteq \pred( \Safe)$.
+$\post^* (s) \cap S \subseteq \pred( \Safe)$.
 In such a case the algorithm from
@@
 
-Unfortunately the problem is undecidable in general.
+Unfortunately the {\sc State-resilience} problem is undecidable for WSTS.
 
@@
 \begin{proof}
-Consider Reset VASS
-in the case $\Safe$ contains only the minimal element $Min(S) = 0$ %$0$ 
-and $\Bad = S \setminus \Safe$, 
-it follows the state resilience problem is equivalent 
-to the problem of whether $0$ can be visited infinitely often. 
+Consider Reset $d$-VASS
+in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} \in \mathbb{N}^d$ 
+%			contains only the minimal element $Min(S) = 0$ %$0$ 
+\alain{tu n'as pas le doit d'écrire $0$ sans explication}
+and $\Bad = \mathbb{N}^d \setminus \Safe$, 
+it follows the state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}is equivalent 
+to the problem of whether $0$ can be visited infinitely often. \alain{incompréhensible, explique}
 % Since reachability of $0$ is undecidable for Reset VASS, we conclude.
@@
 Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$.
-Since $x \in U$ and $U$ is downward-closed, we also have $a \in U$.
+Since $x \in U$ and $U$ is downward-closed, we also have $a \in U$ \alain{faux}.
 Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
@@
 
-In the case of a WSTS with downward compatiblitity, not necessarily strong,
+In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
 then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
@@
 \begin{theorem}\label{downward srp}
-{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,
+{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,\alain{a-t-on aussi la upward compatibility ?}
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
@@
 one looking for a resilience certificate and one looking for a non-resilience certificate.
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward-closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
 Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
-Notice that, by standard monotonicity, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
+Notice that, by standard \alain{upward ? downward ?}compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
 
@@
 The second procedure iteratively computes
-$post^{\leq n}(s) \cap \Bad$
+$\post^{\leq n}(s) \cap \Bad$
 until it finds an element
@@
 \fi
-
+%
+\alain{tableau trop vide. retirer le dernier tableau et mettre le 1er dans la section correspondate, supprimer section 4.3}
+%
 \begin{center}

### 2023-04-10 resilience-safe-down.tex
details

@@
 %			at least one counter is unbounded. 
-This set is upward-closed by upward compatibility in VASS.
+This set is upward-closed by (upward) compatibility in VASS.
 
@@
 
-As an example, let us choose $\Safe=S-\Bad$ then $\Safe=\downarrow \Safe$.
-
+As an example, let us choose $\Safe=S-\Bad$ then since $\Bad=\uparrow \Bad$, we have: $\Safe=\downarrow \Safe$.
+%
+\alain{definir  ideal-effective WSTS}
+%
 \begin{theorem}\label{up-down}
@@
 $\Bad \subseteq  \pred^*(\Safe)$.
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed subsets.
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed \alain{non définis} subsets.
 Hence we can compute the intersection of
@@
 and let $x \in \downarrow \pred^*(D)$.
-By design there exists
+By design \alain{?} there exists
 $y \in \pred^*(D)$ 
@@
 where
-$d' \leq d$ belongs to $D$ by downward closure of $D$.
+$d' \leq d$ \alain{qui appartient à qui ?}belongs to $D$ by downward closure of $D$.
 Hence $x \in \pred^*(D)$.
@@
 the additional hypothesis that
-for all $D \subseteq S$ downward-closed, $\pred^*(D)$ is downward-closed
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed
 is not met,
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
 and there exists $n_0 \in \N$ such that 

### 2023-04-11 resilience-safe-down.tex
details

@@
 
-\alain{à quoi ça sert ici ?: Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
+\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
 
@@
 Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
 Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
@@
 \begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong upward compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong (upward) compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
 % Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.

### 2023-04-11 definitions.tex
up
Up

@@
 
-The model we are interested in is WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
+The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
 
@@
 
-%A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
-%$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
-%${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
-%ternary relation,
-%denoted as the set of {\em labeled transitions}. 
+\mathieu{ Maybe add this back ?
+A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
+$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
+${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
+ternary relation,
+denoted as the set of {\em labeled transitions}. }
 \end{definition}
 
-%Note 
-% that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.
-
+% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$.
 We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
@@
 \noindent
-The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$.
+The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
+ $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
+% $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
 A state without successors is called a {\em dead end}. 
 The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
-$\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. \\
+ $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
+% $\pred(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s' \xrightarrow{a} s\}$. 
+By iterating $\pred$ and $\post$ we obtain
+$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
+and
+$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
+However, we are generally more interested in
+$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$
+and
+$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$. \\
 
@@
 % A partial ordering (a po) is an antisymmetric qo
-\alain{Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).}
+	% \alain{
+Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
+	% }
 % and gives rise to a po between the equivalence classes ?
@@
 
-In what follows, a downward closed set is represented by its finite set of ideals 
+In what follows, a downward closed set $D$ is represented by its finite set of ideals 
 %		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
-and an upward closed set $U$ is represented by its finite set of minimal elements.
+and an upward closed set $U$ is represented by its finite set of minimal elements. \\
 % $M_U$ of its  satisfying $U=\uparrow M_U$.
-
-\alain{est-ce utile et correct ?
-Let $I,J$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition.
-All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable.}
-
-% \textcolor{red}{Define WSTS}
-
-% \textcolor{red}{Define SWSTS -  necessary}
 
@@
 % (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $min(A)$ est la base minimale}
+% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
 %			\end{fact}
@@
 We then have:  $\pred^*(U) = U_{k_0}$.
+\\
 
@@
 \fi
+
+\noindent
+{\bf Vector addition system with states}
+
+% Définir VASS
+
+\begin{definition}
+A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\N^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em states} of $V$, and where $T \subseteq Q \times \N^d \times Q$ will be referred to as {\em transitions} of $V$.
+% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
+$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
+\end{definition}
+
+Subsetquently, $Q \times \N^d$ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z}$.
+
+\mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
+
+When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+
+\problemx{$d$-VASS reachability}
+{A $d$-Vass $V=(Q,T)$, and configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
+{$s \to^* s'$ ? \\}
+
+\problemx{$d$-VASS zero reachability}
+{Vass $V=(Q,T)$, $q(\textbf{v}) \in Q \times \N^d$}
+{$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
+
+\noindent
+{\bf Reset VASS}
+
+A {\em reset ... }
+Reset VASS are VASS extended with reset transitions.
+
+\problemx{Reset $d$-VASS reachability}
+{A reset $d$-Vass $d$-Vass $V=(Q,T)$, $s,s' \in Q \times \N^d$}
+{$s \to^* s'$ ? \\}
+
+\problemx{Reset $d$-VASS zero reachability}
+{A reset $d$-Vass $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
+{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}

### 2023-04-11 reachable-resilience.tex
Up

@@
 
+A transition $t$ is live in a configuration $s$ if for each $s' \in \post^*(s)$ there exists a sequence $s''$ with $s' \to^* s''$ such that $t$ is enabled in $s''$.
+
+\problemx{$t$-liveness}
+{Vass $V=(Q,T)$, $t \in T$}
+{Is $t$ live ?}
+
 \begin{proposition}\label{reductions}
-In WSTS, $t$-liveness is reducible to the state resilience problem.
+In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
 %	home state and reachability
@@
 % \alain{rappeler l'idée de la preuve}
-The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, we check whether
-$B_{\uparrow \post^*(s)} \cap D = \emptyset$, \alain{rapeler $B_{\uparrow \post^*(s)}$}  that is equivalent to $\post^*(s)\cap D = \emptyset$ by
-Lemma~\ref{Lemma intersection}. For the converse direction, we compute the sequence of upward-closed sets
+The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
+$B_{\uparrow \post^*(s)} \cap D = \emptyset$, 
+where $B_{\uparrow \post^*(s)}$ is a basis of $\uparrow \post^*(s)$,
+%\alain{rapeler $B_{\uparrow \post^*(s)}$} 
+ that is equivalent to $\post^*(s)\cap D = \emptyset$ by
+Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
 $U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
@@
 \iffalse
-Then, one would be able to decide whether an element $m \in Min(\uparrow \post^*)$ is reachable.
+Then, one would be able to decide whether an element $m \in \min(\uparrow \post^*)$ is reachable.
 %in counter $i$ 
@@
 \begin{proof}
-Let $B_s$ \alain{notation fluctuante} be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
+Let $B_{\uparrow \post^*(s)}$ 
+%\alain{notation fluctuante}
+ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
 and $\Bad$ a downward-closed set given by its finite set of (maximal) ideals.
@@
 
-$$ post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \equiv B_s \cap \Bad \subseteq \pred^*(\Safe)$$
-\alain{$\equiv $  ???}
+$$ \post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} \cap \Bad \subseteq \pred^*(\Safe)$$
+% \alain{$\equiv $  ???}
 
-Since $B_s$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
+Since $B_{\uparrow \post^*(s)}$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
 % For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_s \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
+We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_{\uparrow \post^*(s)} \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
 \end{proof}
@@
 
+Results synthesis in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
+
+\begin{center}
+\begin{tabular}{ | l | c | c | c | c |}
+\hline  Hypothesis & strong upward compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
+   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
+   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+ \end{tabular}
+\end{center}
+
 \subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}
@@
 in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} \in \mathbb{N}^d$ 
-%			contains only the minimal element $Min(S) = 0$ %$0$ 
+%			contains only the minimal element $\min(S) = 0$ %$0$ 
 \alain{tu n'as pas le doit d'écrire $0$ sans explication}
@@
 Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$.
-Since $x \in U$ and $U$ is downward-closed, we also have $a \in U$ \alain{faux}.
+Since $x \in U$ and $U$ is upward-closed, we also have $a \in U$.
 Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
@@
 \end{proposition}
-
-\subsection{Synthesis}
-
-Results in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
-
-\begin{center}
-\begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong upward compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
-   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
-   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
-      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
- \end{tabular}
-\end{center}
-
-\noindent
-Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
-\iffalse
-(by symmetry) for downward compatible WSTS (but not necessarily upward compatible):
-
-\begin{center}
-\begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong downward compatibility ~ & $\downarrow post^*$ effective & both  \\ \hline
-   SRP & Undecidable  & Decidable   & Decidable \\ \hline
-   BSRP & Undecidable  &  ??  & Decidable \\ \hline
-      kSRP & Undecidable  & ?? & Decidable  \\ \hline
- \end{tabular}
-\end{center}
-
-\mathieu{Montrer que leurs hypothèses Lossy + $\bot$-bounded + strong upward compatibility c'est plus fort que just strong downward compatibility et $\downarrow \post^*$ effective ?}
-
-Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt
-\fi
-%
-\alain{tableau trop vide. retirer le dernier tableau et mettre le 1er dans la section correspondate, supprimer section 4.3}
-%
-\begin{center}
-\begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis &  none  & downward compatible & strongly downward compatible  \\ \hline
-   SRP & Undecidable (Thm~\ref{srp down up})  & Decidable (Thm~\ref{downward srp})  & Decidable (Thm~\ref{downward srp}) \\ \hline
-   BSRP & ??  &  ??  & Decidable (Prop~\ref{downward brp}) \\ \hline
-      kSRP & ??  & ?? & Decidable (Prop~\ref{downward brp}) \\ \hline
- \end{tabular}
-\end{center}
-% \alain{une hypothèse n'est pas l'ensemble vide}

### 2023-04-11 resilience-safe-up.tex
details
Up

@@
 Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-
+%
+\alain{definir completion-post-effective. Voir mon papier well behaved....}
+%
 \begin{theorem}\label{down-up}
@@
 Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
-Let us now define $K_{\mathscr{S}}(\Safe,\Bad)=max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+Let us now define $K_{\mathscr{S}}(\Safe,\Bad)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+%
+
+%
 %	(if all $k_{i,j}$ are finite) else $K=\infty$.
@@
 %To decide the bounded resilience, we decide $n \times m$ coverability questions: is state $\downarrow b_j$ coverable from ideal $J_i$ ?
-If all these $n \times m$ coverability questions are positive then we compute $K=max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $\downarrow b_j$ from $J_i$.
+If all these $n \times m$ coverability questions are positive then we compute $K=\max(k_{i,j} \mid i=1,..,n $ and $j= 1,..m)$ where $k_{i,j}$ is the least length of a sequence that covers  $\downarrow b_j$ from $J_i$.
 %
@@
 	The  $k$-resilience problem (kRP) is decidable 
-	%	for $k \geq min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow 	\pred^*(\Safe)$. 
+	%	for $k \geq \min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow 	\pred^*(\Safe)$. 
 	\textcolor{red}{CONJECTURE for now}
@@
 
-Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
+Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
 
-%	If $k <  min(K,n)$, then we check every path of length smaller
-%	than $min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
+%	If $k <  \min(K,n)$, then we check every path of length smaller
+%	than $\min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
 \end{proof}

### 2023-04-11 application.tex

@@
 ici on veut rester toujours dans $\Safe$.
-on calcule les élements minimaux de $\pred^*(\Safe)=\{m_1,m_2,...,m_p\}$ puis on ajoute les préconditions $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ à toutes les transitions $s \rightarrow s'$ assurant que l'état atteint est au dessus des minimaux, ainsi on sera toujours dans $\Safe$.
-Il faut juste l'hypothèse raisonnable qu'on peut exprimer $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ comme une garde d'une transition (facile dans les VASS et modèles habituels).
+on calcule les élements minimaux de $\pred^*(\Safe)=\{m_1,m_2,...,m_p\}$ puis on ajoute les préconditions $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ à toutes les transitions $s \rightarrow s'$ assurant que l'état atteint $s'$ est au dessus des minimaux de $\pred^*(\Safe)$, ainsi par (upward) compatibility on sera toujours dans $\Safe$.
+Il faut juste l'hypothèse raisonnable qu'on peut exprimer $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ comme une garde d'une transition dans les WSTS concernés (facile dans les VASS et modèles habituels).
 
@@
 Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(\Safe)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $\Safe$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman...à bien vérifier que $\pred^*(\Safe)$ n'est pas ackerman...en lisant en détail \cite{DBLP:conf/rp/BozzelliG11}.
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
 Dans ce cas, ça permettrait de résoudre autrement les 3 premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
 problèmes: marquage T-bloqué, dead, borné , T-continual.
-Je pense qu'au moins
 
@@
 \end{theorem}
+
+%		\section{Control of a WSTS to be resilient}
 

### 2023-04-11 reachable-resilience.tex
up

@@
 
+\mathieu{En train de penser que le mieux serait de définir la $t$-liveness pas pour les WSTS mais pour les (Reset) VASS uniquement et ensuite dire SRP indécidable dans WSTS car indécidable pour les Reset VASS.}
+
 A transition $t$ is live in a configuration $s$ if for each $s' \in \post^*(s)$ there exists a sequence $s''$ with $s' \to^* s''$ such that $t$ is enabled in $s''$.
 
-\problemx{$t$-liveness}
-{Vass $V=(Q,T)$, $t \in T$}
+\problemx{(Reset) VASS $t$-liveness}
+{(Reset) VASS $V=(Q,T)$, $t \in T$}
 {Is $t$ live ?}
@@
 Since the downward-reachability problem is decidable for VASS \alain{ref ?}, we \alain{we ou Ozkan ?
-qui a fait la preuve ?} deduce the following:
+qui a fait la preuve ?}
+\mathieu{le papier d'Ozkan: 
+[Petri nets (equipped with any
+finite set of initial states) are $\uparrow \post^*$-effective by above Prop.: Reachability for Petri nets is decidable and recursively equivalent to submarking reachability.
+This corresponds to the anti-ideal reachability problem for Petri nets.]}
+deduce the following:
 
@@
 \end{proposition}
-%\alain{mauvaises notations, la tradition est que P est l'ensemble des places...les états/markings en minuscules,....}
 
@@
 Consider Reset $d$-VASS
-in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} \in \mathbb{N}^d$ 
+in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 
 %			contains only the minimal element $\min(S) = 0$ %$0$ 
@@
 and $\Bad = \mathbb{N}^d \setminus \Safe$, 
-it follows the state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}is equivalent 
+it follows the 
+state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}
+is equivalent 
 to the problem of whether $0$ can be visited infinitely often. \alain{incompréhensible, explique}
@@
 {\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,\alain{a-t-on aussi la upward compatibility ?}
+\mathieu{oui}
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
@@
 Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
-Notice that, by standard \alain{upward ? downward ?}compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
+Notice that, by 
+%standard \alain{upward ? downward ?}
+upward compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
 

### 2023-04-11 resilience-safe-down.tex
Up
up

@@
 $\Bad \subseteq  \pred^*(\Safe)$.
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed \alain{non définis} subsets.
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of upward- or downward-closed 
+%\alain{non définis}
+ subsets.
 Hence we can compute the intersection of
@@
 and let $x \in \downarrow \pred^*(D)$.
-By design \alain{?} there exists
+By downward closure, there exists
 $y \in \pred^*(D)$ 
-such that $x \leq y$
-and $d \in D$ such that
+such that $x \leq y$.
+By definition of $\pred^*(D)$ then there exists 
+$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
 $y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
@@
 By downward compatibility $a_0 \to a_1$
-implies there exists $a'_1 \leq a_1$ with
+implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
 $x \to^* a'_1$.
 More generally $a_i \to a_{i+1}$ and
-$a'_i\leq a_i$ implies the existence of $a'_{i+1} \leq a_{i+1}$ such that
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
 $a'_i \to^* a'_{i+1}$,
 and, by induction,
- $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* d'$
-where
-$d' \leq d$ \alain{qui appartient à qui ?}belongs to $D$ by downward closure of $D$.
-Hence $x \in \pred^*(D)$.
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+with $d' \leq d$.
+Since
+$d'$ 
+% \alain{qui appartient à qui ?}
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$.
 \end{proof}
@@
 \alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
-
-and $\Bad = \uparrow U \setminus \Safe $
+\mathieu{Pas convaincu que ça serve, je vais enlever}
 
@@
 Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
+compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
 and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
 Hence the equivalence.
@@
 Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong (upward) compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong (upward) compatibility,  $\pred^{\leq n}(\Safe)=\uparrow \pred^{\leq n}(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
 % Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.

### 2023-04-11 definitions.tex
up

@@
 
+\iffalse
 \mathieu{ Maybe add this back ?
@@
 denoted as the set of {\em labeled transitions}. }
+\fi
 \end{definition}
@@
 \problemx{$d$-VASS reachability}
-{A $d$-Vass $V=(Q,T)$, and configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
+{A $d$-VASS $V=(Q,T)$, and configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
 {$s \to^* s'$ ? \\}
@@
 \problemx{$d$-VASS zero reachability}
-{Vass $V=(Q,T)$, $q(\textbf{v}) \in Q \times \N^d$}
+{VASS $V=(Q,T)$, $q(\textbf{v}) \in Q \times \N^d$}
 {$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
@@
 
-A {\em reset ... }
-Reset VASS are VASS extended with reset transitions.
+Reset VASS extend the basic VASS model with special “reset
+arcs” that resets (set to $0$) some coordinates in the vector. \\
 
-\problemx{Reset $d$-VASS reachability}
-{A reset $d$-Vass $d$-Vass $V=(Q,T)$, $s,s' \in Q \times \N^d$}
-{$s \to^* s'$ ? \\}
+More Formally {\em reset ... }
 
+\begin{samepage}
 \problemx{Reset $d$-VASS zero reachability}
-{A reset $d$-Vass $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
+{A reset $d$-VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
 {$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
+\end{samepage}

### 2023-04-12 definitions.tex
up

@@
 A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
 To any subset $A \subseteq X$, we may associate
@@
 
-% \textcolor{red}{Define what an Ideal is.
-Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion.
-Anti-ideal just downward closed so again just not that helpful a notation.
-Maybe have a }
-
-\iffalse
-\begin{definition}
-A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
-$\uparrow I = I = \downarrow I$.
-\end{definition}
-
-"Bi-ideals often represent “control states” as in [cf \%]. "
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-% \textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
-\fi
-
 %         A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
@@
 
+\begin{definition}{\cite{BFM-ic17}} 
+A class $C$ of WSTS is {\em ideally effective} if
+\begin{itemize}
+\item the function mapping the encoding of a configuration $s$ of an ordered transition system
+to the encoding of the ideal $\downarrow s$ is computable;
+\item inclusion of ideals is decidable;
+\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is com-
+putable from the ideal $I$.
+\end{itemize}
+\end{definition}
+
+\mathieu{I'm using the notion of ideally effective-ness for WBTS here}
+
 \noindent

### 2023-04-12 resilience-safe-up.tex
up

@@
 \alain{definir completion-post-effective. Voir mon papier well behaved....}
+
+\begin{definition}
+A class $C$ of WSTS is {\em completion-post-effective} if
+the class $\hat{C} = \{ \hat{S} : S \in C\}$ is post-effective, i.e. if it is effective,
+and 
+if there exists a Turing machine $M$ that computes $\post_{C(i)}(X)$ on input $(i,x)$, with $i \in \N$ and $x$ a configuration of $C(i)$.
+\end{definition}
+
+By extension, we say that a WSTS $S$ is {\em completion-post-effective} if the 
+class $\{S\}$ is completion-post-effective.
+
+\mathieu{Je vois que dans l'article il y a deux définitions completion-post-effective vs completion-effective et je me demandais si on avait besoin des deux ici, même si à priori j'avais plutôt l'impression on n'en avait besoin que d'une seule ?}
+
 %

### 2023-04-12 definitions.tex
up
Starting

@@
 \begin{definition}
-A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\N^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em states} of $V$, and where $T \subseteq Q \times \N^d \times Q$ will be referred to as {\em transitions} of $V$.
+A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ will be referred to as {\em control transitions} of $V$.
 % The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
@@
 Subsetquently, $Q \times \N^d$ denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z}$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z}$.
 
@@
 
+Op is the set of actions
+
+% If $\gamma \in \{1, \ldots, d\}$, we write $\overline{\gamma}$ for the $\{1, \ldots, d\} \setminus \{\gamma\}$
+
+\begin{definition}
+A {\em reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
+% $\mathds{Z}^d$
+labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where 
+%$T \subseteq Q \times \mathds{Z}^d \times Q$
+$T \subseteq Q \times Op \times Q$
+ will be referred to as {\em control transitions} of $V$. 
+where $Op = \{ add(\textbf{z}) | \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) | i \in \{1,\ldots,d\} \}$.
+\end{definition}
+
+% $(Q, Op, \rightarrow)$
+
+% $(S, Op, \rightarrow)$
+
+Again $Q \times \N^d$
+ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition
+
+and the set of transitions is
+
+\begin{itemize}
+
+\item if $(q,\textbf{z},q') \in T_1$ then for all $\textbf{u} \in \N^d$ such that $\textbf{u}+\textbf{z} \geq 0$
+ $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+
+\item if $(q,\gamma,q') \in T_2$ then for all $\textbf{u} \in \N^d$ 
+$q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where $\textbf{u}'[\gamma] = 0$ and $\textbf{u}'[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+
+\end{itemize}
+
 \begin{samepage}

### 2023-04-12 reachable-resilience.tex
Starting

@@
 %\end{corollary}
-
-\iffalse
-
-\begin{theorem}
-The downward reachability problem is reducible to the state resilience problem.
-\end{theorem}
-
-\begin{proof}
-Let $s \in S$, and a downward-closed set $D$. \alain{je ne comprends pas cette "preuve"}
-Take $\Safe = \emptyset$ and $\Bad = D$.
-The state resilience problem on $\Safe,\Bad$ asks
-whether $\Bad \cap \post^*(s) \rightarrow^* \Safe$  that is equivalent to $\Bad \cap \post^*(s) \subseteq \emptyset$.
-If the answer is negative then $s \to^* \Bad$,
-else $s \not\to^* \Bad$.
-\end{proof}
-
-\fi
 

### 2023-04-12 definitions.tex
up

@@
 \begin{definition}
-A {\em reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
+A {\em Reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
 % $\mathds{Z}^d$
@@
 \begin{samepage}
-\problemx{Reset $d$-VASS zero reachability}
-{A reset $d$-VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
+\problemx{$d$-dim Reset VASS zero reachability}
+{A $d$-dim Reset VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
 {$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}

### 2023-04-12 definitions.tex
up

@@
 \item inclusion of ideals is decidable;
-\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is com-
-putable from the ideal $I$.
+\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
 \end{itemize}
@@
 \begin{definition}
-A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ will be referred to as {\em control transitions} of $V$.
+A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ will be referred to as the {\em control transitions} of $V$.
 % The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
@@
 
-More Formally {\em reset ... }
+% More Formally {\em reset ... }
 
-Op is the set of actions
-
-% If $\gamma \in \{1, \ldots, d\}$, we write $\overline{\gamma}$ for the $\{1, \ldots, d\} \setminus \{\gamma\}$
+% Op is the set of actions
 
@@
 $T \subseteq Q \times Op \times Q$
- will be referred to as {\em control transitions} of $V$. 
-where $Op = \{ add(\textbf{z}) | \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) | i \in \{1,\ldots,d\} \}$.
+ will be referred to as the {\em control transitions} of $V$,
+where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
 \end{definition}
-
-% $(Q, Op, \rightarrow)$
-
-% $(S, Op, \rightarrow)$
 
@@
  denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition
-
-and the set of transitions is
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when
 
@@
 
-\item if $(q,\textbf{z},q') \in T_1$ then for all $\textbf{u} \in \N^d$ such that $\textbf{u}+\textbf{z} \geq 0$
- $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+\item  $t = (q,add(\textbf{z}),q') \in T$
+% then for all $\textbf{u} \in \N^d$ such that  
+% $\textbf{u}+\textbf{z} \geq 0$
+% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
 
-\item if $(q,\gamma,q') \in T_2$ then for all $\textbf{u} \in \N^d$ 
-$q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where $\textbf{u}'[\gamma] = 0$ and $\textbf{u}'[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-
+\item $t = (q,reset(\gamma),q') \in T$ 
+% then for all $\textbf{u} \in \N^d$ 
+% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
 \end{itemize}

### 2023-04-12 resilience-safe-down.tex
up

@@
 %
-\alain{definir  ideal-effective WSTS}
+% \alain{definir  ideal-effective WSTS}
 %
@@
 
+\iffalse
 \alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
 \mathieu{Pas convaincu que ça serve, je vais enlever}
+% hop
+\fi
 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
-compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
+compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+% \alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
 and there exists $n_0 \in \N$ such that 

### 2023-04-12 definitions.tex
up

@@
 {$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
-
-\noindent
-{\bf Reset VASS}
-
-Reset VASS extend the basic VASS model with special “reset
-arcs” that resets (set to $0$) some coordinates in the vector. \\
-
-% More Formally {\em reset ... }
-
-% Op is the set of actions
-
-\begin{definition}
-A {\em Reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
-% $\mathds{Z}^d$
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where 
-%$T \subseteq Q \times \mathds{Z}^d \times Q$
-$T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control transitions} of $V$,
-where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
-
-Again $Q \times \N^d$
- denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t$ we write
-$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when
-
-\begin{itemize}
-
-\item  $t = (q,add(\textbf{z}),q') \in T$
-% then for all $\textbf{u} \in \N^d$ such that  
-% $\textbf{u}+\textbf{z} \geq 0$
-% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
-and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-
-\item $t = (q,reset(\gamma),q') \in T$ 
-% then for all $\textbf{u} \in \N^d$ 
-% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
-and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize}
-
-\begin{samepage}
-\problemx{$d$-dim Reset VASS zero reachability}
-{A $d$-dim Reset VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
-{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
-\end{samepage}

### 2023-04-12 reachable-resilience.tex
up

@@
 Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
-This stems from the fact that $t$-liveness is undecidable for WSTS and 
-is reducible to the state resilience problem.\alain{definir t-liveness, dire ce qu'est une transition $t$ dans un WSTS étiqueté ou pas ?...}
+This stems from the fact that it is undecidable in the particular case of Reset VASS,
+where  $t$-liveness is both undecidable and 
+ reducible to the state resilience problem.
 
-Let us define $pre(t)=\{ u \in S \mid \exists u' \in S$ in a labeled WSTS such that $u \xrightarrow{t} u'$\}
+% \noindent
+% {\bf Reset VASS}
 
+Reset VASS extend the basic VASS model with special “reset
+transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.\\
+
+% More Formally {\em reset ... }
+
+\begin{definition}
+A {\em Reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
+% $\mathds{Z}^d$
+labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where 
+%$T \subseteq Q \times \mathds{Z}^d \times Q$
+$T \subseteq Q \times Op \times Q$
+ will be referred to as the {\em control transitions} of $V$,
+where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
+\end{definition}
+
+Again $Q \times \N^d$
+ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when
+
+\begin{itemize}
+
+\item  $t = (q,add(\textbf{z}),q') \in T$
+% then for all $\textbf{u} \in \N^d$ such that  
+% $\textbf{u}+\textbf{z} \geq 0$
+% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
+
+\item $t = (q,reset(\gamma),q') \in T$ 
+% then for all $\textbf{u} \in \N^d$ 
+% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+\end{itemize}
+
+It is well known that $d$-dim Reset VASS are WSTS [ref ? relire WSTS everywhere]. 
+Moreover, we are interested in the following undecidable [ref again] decision problem:
+
+\begin{samepage}
+\problemx{$d$-dim Reset VASS zero reachability}
+{A $d$-dim Reset VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
+{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
+\end{samepage}
+
+We reduce this problem to the problem of deciding whether a control transition is live, that is, whether it is always ,
+which we then reduce to the state-resilience problem.
+A control transition $t$ of a $d$-dim Reset VASS is called {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole $d$-dim Reset VASS is live if all its control transitions are
+live. This leads to the following problem.
+
+\problemx{$d$-dim Reset VASS $t$-liveness}
+{$d$-dim Reset VASS $V=(Q,T)$, $t \in T$}
+{Is $t$ live ? \\}
+
+To better discuss $t$-liveness let us define 
+% $pre(t)=\{ u \in S \mid \exists u' \in S$ in a labeled WSTS such that $u \xrightarrow{t} u'$\} 
+$pre(t)=\{  \}$ in a $d$-dim Reset VASS such that .
+
+\iffalse
 \alain{ou alors $t=(u,u')$ such that $u \rightarrow u'$ mais alors c'est bizarre de nommer par la même lettre $t$ des couples différents $(u,u'), (v,v'),...$ dans un WSTS...ou alors on appelle transition un couple $(u,u')$ such that $u \rightarrow u'$ mais plus de lettre $t$...ou on définit un WSTS par un nombre fini de (shéma de ) transition $t_i$ qui sont compatibles par compatibilité mais ça contraint à un nombre fini (comme dans les VASS) ou alors on ne parle ici que des VASS et pas des WSTS mais ça suffira puisque state resilience pour VASS implique pour WSTS BREF il faut réfléchir à ce qu'est une transition...}
@@
 
-A transition $t$ is live in a configuration $s$ if for each $s' \in \post^*(s)$ there exists a sequence $s''$ with $s' \to^* s''$ such that $t$ is enabled in $s''$.
-
-\problemx{(Reset) VASS $t$-liveness}
-{(Reset) VASS $V=(Q,T)$, $t \in T$}
-{Is $t$ live ?}
+\fi
 
@@
 \begin{proof}
-Consider Reset $d$-VASS
+Consider $d$-dim Reset VASS
 in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 

### 2023-04-12 reachable-resilience.tex
up

@@
 \problemx{$d$-dim Reset VASS $t$-liveness}
-{$d$-dim Reset VASS $V=(Q,T)$, $t \in T$}
-{Is $t$ live ? \\}
+{$d$-dim Reset VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{Is $t$ live in $s_0$ ? \\}
 
-To better discuss $t$-liveness let us define 
+To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
 % $pre(t)=\{ u \in S \mid \exists u' \in S$ in a labeled WSTS such that $u \xrightarrow{t} u'$\} 
-$pre(t)=\{  \}$ in a $d$-dim Reset VASS such that .
+$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
 
@@
 \begin{proposition}\label{reductions}
-In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
-%	home state and reachability
+% In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
+{\sc $d$-dim Reset VASS $t$-liveness} is reducible to {\sc state-resilience}.
 \end{proposition}
@@
 \begin{proof}
-We reformulate $t$-liveness in a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ as the following formula
-% $$ ~ \forall s' \in S, ~ s \rightarrow^* s' \implies \exists s'' \in U_t, ~ s' \rightarrow^{*} s''$$
-$$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$  
+We reformulate $t$-liveness 
+in a 
+% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
+$d$-dim Reset VASS $(Q,T)$ 
+with initial configuration $s_0$
+as the following formula
+% $$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$ 
+$$ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})$$  
 where
-$U_t=\uparrow \pred(t)$ \alain{$\pred(t)$ n'est pas défini et $\pred$ est une mauvaise notation pour ça: plutôt: $pre(t)=\{ u \in S \mid \exists u' \in S$  such that $u \xrightarrow{t} u'$\} }
+$U_t=\uparrow pre(t)$ 
+% \alain{$\pred(t)$ n'est pas défini et $\pred$ est une mauvaise notation pour ça: plutôt: $pre(t)=\{ u \in S \mid \exists u' \in S$  such that $u \xrightarrow{t} u'$\} }
 is the upward closure of the preconditions to fire transition $t$.  
 The problem reduces itself to the state resilience problem
-where $\Safe = U_t$ and $\Bad = S \setminus \Safe$.
+where $\Safe = U_t$ and $\Bad = Q \times \N^d \setminus \Safe$.
 \end{proof}
 
-To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS.
-The following theorem is a adjustment ofTheorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.
+% To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. 
+The following theorem is a adjustment of Theorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.
 
-\alain{tu ne définis pas, même informellement et/ou rapidement, assez les objets que tu utilises: zero reachability problem, Reset VASS}
+% \alain{tu ne définis pas, même informellement et/ou rapidement, assez les objets que tu utilises: zero reachability problem, Reset VASS}
 
@@
 
-Since the zero reachability problem is undecidable for Reset VASS, the reduction implies 
+Since the {$d$-dim Reset VASS zero reachability} is undecidable, the reduction implies 
 % the following:
@@
 %\begin{corollary}
-$t$-liveness is undecidable for Reset VASS.
+{$d$-dim Reset VASS $t$-liveness} is undecidable.
 %\end{corollary}
 
-Since $t$-liveness is undecidable for Reset VASS that are WSTS with strong compatibility, from Proposition~\ref{reductions},  we deduce that state-resilience is undecidable for WSTS with strong upward-compatibility.
+Since {$d$-dim Reset VASS $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc state-resilience} is undecidable  for $d$-dim Reset VASS, which are WSTS with strong compatibility.
+Hence {\sc state-resilience} is undecidable  for WSTS with strong upward-compatibility.
 This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
@@
 \begin{proof}
-{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since $t$-liveness is undecidable.
+{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since {$d$-dim Reset VASS $t$-liveness} is undecidable.
 
@@
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
-compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,
+compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
 and there exists $n_0 \in \N$ such that 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
 We compute 
-$n_0$, then iteratively check whether $k$-state resilience \alain{$k$-state resilience ou $k$-state-resilience ?}
+$n_0$, then iteratively check whether $k$-state-resilience 
+% \alain{$k$-state resilience ou $k$-state-resilience ?}
 hold for $k$ from $0$ to $n_0$.  
@@
 since 
-$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
 Hence the %equivalence, and

### 2023-04-12 reachable-resilience.tex
up

@@
 and $\Bad = \mathbb{N}^d \setminus \Safe$, 
-it follows the 
-state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}
+it follows
+{\sc state-resilience} \alain{écris toujours pareil les expressions choisies comme state-resilience}
 is equivalent 

### 2023-04-12 reachable-resilience.tex
details

--- 2023-04-12 reachable-resilience.tex
@@
 \begin{theorem}\label{downward srp}
-{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,\alain{a-t-on aussi la upward compatibility ?}
-\mathieu{oui}
+{\sc state-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities,
+%\alain{a-t-on aussi la upward compatibility ?}
+%\mathieu{oui}
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.

### 2023-04-12 definitions.tex
details

@@
 
-We say that a WSTS $\mathscr{S}$ has the \emph{strong compatibility} when moreover for all 
+We say that a WSTS $\mathscr{S}$ has the \emph{strong (upward) compatibility} when moreover for all 
 $s_1, t_1 , s_2 \in S$

### 2023-04-12 definitions.tex
details

@@
 
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed.
-\end{proposition}
+The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
+
 %\alain{étrange ces deux propositions, l'une suffirait}
@@
 
-There is a simple condition that insures that a finite basis of $\pred^*(U )$ is computable.
+Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
 

### 2023-04-12 application.tex
details

@@
 %
-The BRP and the kRP are also decidable.
-The algorithm that computes $ \pred^*(\Safe)$ is polynomial.
-If $\Bad \subseteq \pred^*(\Safe)$ then one computes an integer $k$ such that the maximum of all minimal pathes that go from $\Bad$ to $\Safe$ is equal to $k$.\alain{à prouver}
+The kRP is decidable: let us compute $ \pred^*(\Safe)$ (in polynomial time).
+If $\Bad \subseteq \pred^*(\Safe)$ then one computes the integer $k_0$ such that the maximum of all minimal pathes that go from $\Bad$ to $\Safe$ is equal to $k$.
+Then $A=(Q,\Sigma,...)$ is $p$-resilient for all $p \geq k_0$ and it is not $p$-resilient for all $p \leq k_0-1$.
+
+\alain{à prouver la BRP...}
 %

### 2023-04-12 definitions.tex
details

@@
 
-In what follows, a downward closed set $D$ is represented by its finite set of ideals 
+In what follows, a downward closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
 %		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
@@
 \begin{definition}
-A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ will be referred to as the {\em control transitions} of $V$.
+A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control transitions} of $V$.
 % The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
@@
 Subsetquently, $Q \times \N^d$ denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z}$.
+For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
 

### 2023-04-12 definitions.tex
details

@@
 {$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
+
+\subsection{Computing the largest safe WSTS}
+
+Ici $\Safe$ clos haut.
+ici on veut rester toujours dans $\Safe$.
+on calcule les élements minimaux de $\pred^*(\Safe)=\{m_1,m_2,...,m_p\}$ puis on ajoute les préconditions $s' \geq m_1  	\vee s' \geq m_2  	\vee...	\vee s' \geq m_p$ à toutes les transitions $s \rightarrow s'$ assurant que l'état atteint $s'$ est au dessus des minimaux de $\pred^*(\Safe)$, ainsi par (upward) compatibility on sera toujours dans $\Safe$.
+Il faut juste l'hypothèse raisonnable qu'on peut exprimer $s' \geq m_1  	\lor s' \geq m_2  	\lor...	\lor s' \geq m_p$ comme une garde d'une transition dans les WSTS concernés (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given a WSTS and $\Safe$  upward-closed, the largest sub TS, that stays in $\Safe$ , is a computable WSTS.
+\end{theorem}
+
+Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(\Safe)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $\Safe$ .
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
+Dans ce cas, ça permettrait de résoudre autrement les 3 premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
+problèmes: marquage T-bloqué, dead, borné , T-continual.
+
+\begin{theorem}{}
+Given a VASS and $\Safe$ upward-closed, the largest sub TS, that  stays in $\Safe$, is a computable VASS of size 2-exp.
+\end{theorem}

### 2023-04-12 reachable-resilience.tex
up

@@
 \begin{proposition}
-Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.\alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
+Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.
+% \alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
 \end{proposition}

### 2023-04-12 application.tex

@@
 open for VASS ???
-
-\section{Computing the largest safe WSTS}
-
-Ici $\Safe$ clos haut.
-ici on veut rester toujours dans $\Safe$.
-on calcule les élements minimaux de $\pred^*(\Safe)=\{m_1,m_2,...,m_p\}$ puis on ajoute les préconditions $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ à toutes les transitions $s \rightarrow s'$ assurant que l'état atteint $s'$ est au dessus des minimaux de $\pred^*(\Safe)$, ainsi par (upward) compatibility on sera toujours dans $\Safe$.
-Il faut juste l'hypothèse raisonnable qu'on peut exprimer $s' \geq m_1  	\wedge s' \geq m_2  	\wedge...	\wedge s' \geq m_p$ comme une garde d'une transition dans les WSTS concernés (facile dans les VASS et modèles habituels).
-
-\begin{theorem}{}
-Given a WSTS and $\Safe$  upward-closed, the largest sub TS, that stays in $\Safe$ , is a computable WSTS.
-\end{theorem}
-
-Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(\Safe)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $\Safe$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
-Dans ce cas, ça permettrait de résoudre autrement les 3 premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
-problèmes: marquage T-bloqué, dead, borné , T-continual.
-
-\begin{theorem}{}
-Given a VASS and $\Safe$ upward-closed, the largest sub TS, that  stays in $\Safe$, is a computable VASS of size 2-exp.
-\end{theorem}
 

### 2023-04-15 definitions.tex
details

--- 2023-04-12 definitions.tex
@@
 {$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
+
+\noindent
+{\bf Computing the maximal $X$-live system}
+
+Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be an (ordered transition) system with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s$ (from $s_0$), there exists (at least) a run that goes into $X$: $s \xrightarrow{*} X$.
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live.
+The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....
+
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guarded WSTSs} where every transition $s \rightarrow s'$ of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is now guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since these formula are upward-compatible, the constructed transition system $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is still a WSTS.
+%
+For example, VASS allows such guards.
+  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal $X$-live subsystem is a computable WSTS.
+\end{theorem}
+
+\begin{proof}
+Since $X$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+%
+%
+Then we add the following guards $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ satisfies the property to be $\pred^*(X)$-live and then it is the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
+
+%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
+\end{proof}
+%
+Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
+Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
+problèmes: marquage T-bloqué, dead, T-continual ?.
+
+\begin{theorem}{}
+Given a VASS and $X$ upward-closed, the maximal $X$-live subsystem is a computable VASS of size 2-exp.
+\end{theorem}
+
+applications to many upward-closed sets $X$:
+$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
+$m$ is dead if the reachability tree from $m$ is finite.
+$m$ is bounded if the reachability set from $m$ is finite.
+$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.

### 2023-04-15 definitions.tex
details

@@
 
-In this section, we introduce general notations and preliminary definitions.
+%   In this section, we introduce general notations and preliminary definitions.
 
@@
 The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
-
 %\alain{étrange ces deux propositions, l'une suffirait}
@@
 \fi
-
+%
 Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
@@
 \begin{theorem}(ref)
-A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
 \end{theorem}
@@
 
-For a WSTS $\mathscr{S}$ and an upward closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$.
-When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)$ converges.
+For a WSTS $\mathscr{S}$ and an upward-closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$.
+When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
 Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
@@
 \\
+%
+
+\noindent
+\subsection{Computing the maximal $X$-live system}
+
+Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be an (ordered transition) system with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s$ (from $s_0$), there exists (at least) a run that goes into $X$: $s \xrightarrow{*} X$.
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live.
+The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....
+
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guarded WSTSs} where every transition $s \rightarrow s'$ of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is now guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since these formula are upward-compatible, the constructed transition system $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is still a WSTS.
+%
+For example, VASS allows such guards.
+  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal $X$-live subsystem is a computable WSTS.
+\end{theorem}
+
+\begin{proof}
+Since $X$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+%
+%
+Then we add the following guards $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ satisfies the property to be $\pred^*(X)$-live and then it is the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
+
+%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
+\end{proof}
+%
+%
+%\noindent
+%{\bf Vector addition system with states}
+
+% Définir VASS
+
+Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
+$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
+
+Subsetquently, $Q \times \N^d$ is the set of configurations of $V$.
+For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+
+\mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
+
+When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+
+Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
+Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
+problèmes: marquage T-bloqué, dead, T-continual ?.
+
+\begin{theorem}{}
+Given a VASS and $X$ upward-closed, the maximal $X$-live subsystem is a computable VASS of size 2-exp.
+\end{theorem}
+
+applications to many upward-closed sets $X$:
+$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
+$m$ is dead if the reachability tree from $m$ is finite.
+$m$ is bounded if the reachability set from $m$ is finite.
+$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
 
@@
 
-\begin{definition}{\cite{BFM-ic17}} 
-A class $C$ of WSTS is {\em ideally effective} if
-\begin{itemize}
-\item the function mapping the encoding of a configuration $s$ of an ordered transition system
-to the encoding of the ideal $\downarrow s$ is computable;
-\item inclusion of ideals is decidable;
-\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-\end{itemize}
-\end{definition}
+\alain{pas ici}
+\problemx{$d$-VASS reachability}
+{A $d$-VASS $V=(Q,T)$, and two configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
+{$p(\textbf{u}) \to^* q(\textbf{v})$? \\}
 
-\mathieu{I'm using the notion of ideally effective-ness for WBTS here}
+\problemx{$d$-VASS zero-reachability}
+{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
+{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\}
 
-\noindent
-{\bf Vector addition system with states}
-
-% Définir VASS
-
-\begin{definition}
-A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control transitions} of $V$.
-% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
-$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
-\end{definition}
-
-Subsetquently, $Q \times \N^d$ denotes the set of configurations of $V$.
-For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
-
-\mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
-
-When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
-
-\problemx{$d$-VASS reachability}
-{A $d$-VASS $V=(Q,T)$, and configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
-{$s \to^* s'$ ? \\}
-
-\problemx{$d$-VASS zero reachability}
-{VASS $V=(Q,T)$, $q(\textbf{v}) \in Q \times \N^d$}
-{$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}
-
-\noindent
-{\bf Computing the maximal $X$-live system}
-
-Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be an (ordered transition) system with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
-We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s$ (from $s_0$), there exists (at least) a run that goes into $X$: $s \xrightarrow{*} X$.
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live.
-The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....
-
-Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guarded WSTSs} where every transition $s \rightarrow s'$ of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is now guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
-Since these formula are upward-compatible, the constructed transition system $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is still a WSTS.
-%
-For example, VASS allows such guards.
-  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
-
-\begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal $X$-live subsystem is a computable WSTS.
-\end{theorem}
-
-\begin{proof}
-Since $X$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
-%
-%
-Then we add the following guards $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ satisfies the property to be $\pred^*(X)$-live and then it is the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
-
-%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
-\end{proof}
-%
-Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
-Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
-problèmes: marquage T-bloqué, dead, T-continual ?.
-
-\begin{theorem}{}
-Given a VASS and $X$ upward-closed, the maximal $X$-live subsystem is a computable VASS of size 2-exp.
-\end{theorem}
-
-applications to many upward-closed sets $X$:
-$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
-$m$ is dead if the reachability tree from $m$ is finite.
-$m$ is bounded if the reachability set from $m$ is finite.
-$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
+The two previous problems are well-known decidable.

### 2023-04-15 reachable-resilience.tex
details

--- 2023-04-12 reachable-resilience.tex
@@
-\section{State resilience}
+\section{State-resilience}
 
@@
 \begin{definition}
-A {\em Reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
+A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
 % $\mathds{Z}^d$
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where 
+labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
 %$T \subseteq Q \times \mathds{Z}^d \times Q$
 $T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control transitions} of $V$,
+ will be referred to as the {\em control-transitions} of $V$,
 where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
@@
  denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t$ we write
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
 $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when
@@
 
+\alain{ce qui suit est inutile}
+
 \begin{samepage}
-\problemx{$d$-dim Reset VASS zero reachability}
-{A $d$-dim Reset VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
+\problemx{Zero-reachability}
+{A Reset $d$-VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
 {$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
 \end{samepage}
-
-We reduce this problem to the problem of deciding whether a control transition is live, that is, whether it is always ,
+%
+\alain{ce qui précède est inutile}
+%
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
 which we then reduce to the state-resilience problem.
-A control transition $t$ of a $d$-dim Reset VASS is called {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole $d$-dim Reset VASS is live if all its control transitions are
+A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a \alain{mélange de s et (p,u)}
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
 live. This leads to the following problem.
 
-\problemx{$d$-dim Reset VASS $t$-liveness}
-{$d$-dim Reset VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+\problemx{$t$-liveness}
+{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
 {Is $t$ live in $s_0$ ? \\}
@@
 % In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
-{\sc $d$-dim Reset VASS $t$-liveness} is reducible to {\sc state-resilience}.
+{\sc $t$-liveness} is reducible to {\sc state-resilience} in Reset $d$-VASS.
 \end{proposition}

### 2023-04-15 resilience-safe-up.tex
details

@@
 
+%
+
+\alain{pas ici}
+\begin{definition}{\cite{BFM-ic17}} 
+A class $C$ of WSTS is {\em ideally effective} if
+\begin{itemize}
+\item the function mapping the encoding of a configuration $s$ of a WSTS
+to the encoding of the ideal $\downarrow s$ is computable;
+\item inclusion of ideals is decidable;
+\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
+\end{itemize}
+\end{definition}
+
+\mathieu{I'm using the notion of ideally effective-ness for WBTS here}
+%
 Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.

### 2023-04-15 resilience.tex
details

@@
 
-Surprinsingly, the decidability of resilience problems was not studied in the previous papers.
+Surprinsingly, we have not found decidability results about the resilience problems for WSTS.
 
@@
 Let us consider now a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
-then resilience hold (one always ends up in the positive integers) but not bounded resilience
-(for every bound $b$ there is an element of $\Bad$ e.g. $q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$). 
+then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $b$ there is an element of $\Bad$ e.g.
+$q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$.
 \end{example}

### 2023-04-17 reachable-resilience.tex
Removed subsequent instances of 'd-dim', adjusted notations in t-liveness def

@@
 
-It is well known that $d$-dim Reset VASS are WSTS [ref ? relire WSTS everywhere]. 
-Moreover, we are interested in the following undecidable [ref again] decision problem:
-
+It is well known that Reset $d$-VASS are WSTS [ref ? relire WSTS everywhere]. 
+Moreover, we are interested in the 
+%following
+ undecidable [ref again] decision problem%:
+\iffalse
 \alain{ce qui suit est inutile}
@@
 \alain{ce qui précède est inutile}
+\fi
+~ of zero-reachability in Reset $d$-VASS.
 %
@@
 which we then reduce to the state-resilience problem.
-A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a \alain{mélange de s et (p,u)}
+A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+%\alain{mélange de s et (p,u)}
  $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
@@
 % WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
-$d$-dim Reset VASS $(Q,T)$ 
+Reset $d$-VASS $(Q,T)$ 
 with initial configuration $s_0$
@@
 
-Since the {$d$-dim Reset VASS zero reachability} is undecidable, the reduction implies 
+Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
 % the following:
@@
 %\begin{corollary}
-{$d$-dim Reset VASS $t$-liveness} is undecidable.
+{\sc %$ Reset $d$-VASS 
+$t$-liveness} is undecidable.
 %\end{corollary}
 
-Since {$d$-dim Reset VASS $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc state-resilience} is undecidable  for $d$-dim Reset VASS, which are WSTS with strong compatibility.
+Since {\sc %$Reset $d$-VASS 
+$t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc state-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
 Hence {\sc state-resilience} is undecidable  for WSTS with strong upward-compatibility.
@@
 \begin{proof}
-{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since {$d$-dim Reset VASS $t$-liveness} is undecidable.
+{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since 
+{\sc %Reset $d$-VASS 
+$t$-liveness} is undecidable.
 
@@
 \begin{proof}
-Consider $d$-dim Reset VASS
+Consider Reset $d$-VASS
 in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 

### 2023-04-17 definitions.tex
details

@@
 However, we are generally more interested in
-$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$
+$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$, $\post^*(s)= \bigcup_{1 \leq i} \post^i(s)$
 and
-$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$. \\
+$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. \\
 
@@
 %  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
-the transition relation $ \rightarrow$ is compatible with $\leq$, i.e., for all 
+the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
 $s_1, t_1 , s_2 \in S$
@@
 
-The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
+We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
 %\alain{étrange ces deux propositions, l'une suffirait}
@@
 %
-Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
 
 \begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
+A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
@@
 
+With the effective pred-basis hypothese, we obtain:
+
 \begin{theorem}(ref)
-A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
 \end{theorem}
@@
 %			\end{fact}
-
-For a WSTS $\mathscr{S}$ and an upward-closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$.
-When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
+For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
+When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
 Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
 $U_k = U_{k_0}$ for all $k \geq k_0$.
-We then have:  $\pred^*(U) = U_{k_0}$.
+We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
 \\
@@
 \noindent
-\subsection{Computing the maximal $X$-live system}
+\subsection{Computing the maximal $X$-live subsystem}
 
-Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be an (ordered transition) system with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
-We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s$ (from $s_0$), there exists (at least) a run that goes into $X$: $s \xrightarrow{*} X$.
+Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be WSTS with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s \in \post^*(s_0)$, $s \xrightarrow{*} X$ is satisfied.
 We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live.
@@
 
-Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guarded WSTSs} where every transition $s \rightarrow s'$ of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is now guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
-Since these formula are upward-compatible, the constructed transition system $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is still a WSTS.
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
+Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ are still WSTS.
 %

### 2023-04-18 reachable-resilience.tex
up

@@
 
-Unfortunately, state-resilience is undecidable for (general) WSTS even with strong upward-compatibility.
+Unfortunately, {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
 This stems from the fact that it is undecidable in the particular case of Reset VASS,
 where  $t$-liveness is both undecidable and 
- reducible to the state resilience problem.
+ reducible to {\sc State-resilience}.
 
@@
 We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
-which we then reduce to the state-resilience problem.
+which we then reduce to {\sc State-resilience}.
 A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
@@
 % In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
-{\sc $t$-liveness} is reducible to {\sc state-resilience} in Reset $d$-VASS.
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
 \end{proposition}
@@
 is the upward closure of the preconditions to fire transition $t$.  
-The problem reduces itself to the state resilience problem
+The problem reduces itself to {\sc State-resilience}
 where $\Safe = U_t$ and $\Bad = Q \times \N^d \setminus \Safe$.
@@
 Since {\sc %$Reset $d$-VASS 
-$t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc state-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
-Hence {\sc state-resilience} is undecidable  for WSTS with strong upward-compatibility.
+$t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
+Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
 This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
@@
 
-On the positive side, let us recall a result about state-resilience decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+On the positive side, let us recall a result about {\sc Bounded-state-resilience} decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
 
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
-{\sc bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
 when
@@
 
-We may immediately generalyse this last result by strengthening to \emph{unbounded} state-resilience. The proof is essentially the same than the previous one.
+We may immediately generalyse this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
 
 \begin{corollary}\label{postcomputable}
-{\sc state-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc State-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
 when
@@
 returning $n$ if that is the case.
-Then, because $n_0$-state-resilience is decidable, 
+Then, because {\sc $n_0$-state-resilience} is decidable, 
 checking $\uparrow post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-and state-resilience is decidable.
+and {\sc State-resilience} is decidable.
 \end{proof}
@@
 % Since 
-VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded state-resilience is decidable for VASS.
-Hence state-resilience is decidable for %both Petri nets and 
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded state-resilience} is decidable for VASS.
+Hence {\sc State-resilience} is decidable for %both Petri nets and 
 VASS.
@@
 
-Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general state-resilience problem.
+Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience }.
 
 \begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
-{\sc state-resilience} is decidable for 
+{\sc State-resilience} is decidable for 
 %(not necessarily strong)
@@
 Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
-{\sc bounded-state-resilience}.
+{\sc Bounded-state-resilience}.
 
@@
 Consider Reset $d$-VASS
-in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 
-%			contains only the minimal element $\min(S) = 0$ %$0$ 
-\alain{tu n'as pas le doit d'écrire $0$ sans explication}
+with $\Safe$ % where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 
+	containing only the
+		configurations of the form $Q \times \{ \textbf{0} \}$
+	where $\textbf{0} = \min(\mathbb{N}^d)$ %$0$ 
+% \alain{tu n'as pas le doit d'écrire $0$ sans explication}
 and $\Bad = \mathbb{N}^d \setminus \Safe$, 
 it follows
-{\sc state-resilience} \alain{écris toujours pareil les expressions choisies comme state-resilience}
+{\sc State-resilience} 
+% \alain{écris toujours pareil les expressions choisies comme state-resilience}
 is equivalent 
-to the problem of whether $0$ can be visited infinitely often. \alain{incompréhensible, explique}
+to the problem of whether 
+% $0$  can be visited infinitely often. 
+it is possible to visit infinitely often configurations with every vector having its value equal to $0$.
+\alain{incompréhensible, explique}
+Indeed, from such a configuration, then the next step is either also such a configuration, either it is in $\Bad$, in which case, we know it is possible to reach such a configuration again.
+Hence we construct a sequence
+$(q_i ({\textbf{v}_{i}}) )_{i\in\N}$ 
+such that for all $i\in\N$ 
+$\exists j \in \N$, $j>i$, s.t. ${\textbf{v}_j} = \textbf{0}$.
 % Since reachability of $0$ is undecidable for Reset VASS, we conclude.
@@
 \begin{theorem}\label{downward srp}
-{\sc state-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities,
+{\sc State-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities,
 %\alain{a-t-on aussi la upward compatibility ?}
@@
 \begin{proof}
-In order to decide whether state-resilience hold, we execute two procedures in parallel,
+In order to decide whether $\post^{\leq n}(s) \cap \Bad \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
 one looking for a resilience certificate and one looking for a non-resilience certificate.

### 2023-04-18 reachable-resilience.tex
up

@@
 	containing only the
-		configurations of the form $Q \times \{ \textbf{0} \}$
+		configurations in $Q \times \{ \textbf{0} \}$
 	where $\textbf{0} = \min(\mathbb{N}^d)$ %$0$ 
 % \alain{tu n'as pas le doit d'écrire $0$ sans explication}
-and $\Bad = \mathbb{N}^d \setminus \Safe$, 
-it follows
+and $\Bad = \mathbb{N}^d \setminus \Safe$.
+It follows in this case that
 {\sc State-resilience} 

### 2023-04-18 definitions.tex
section 2

@@
 \begin{proof}
-Since $X$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
 %
 %
-Then we add the following guards $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ satisfies the property to be $\pred^*(X)$-live and then it is the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-live and then it is also the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
 
@@
 \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
-
+\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$}
 When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

### 2023-04-19 reachable-resilience.tex
up

@@
 % \begin{corollary}\label{bsrp up down}
-Bounded state-resilience and 
+Bounded-state-resilience and 
 $k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
@@
 
-The characterization is used to show that VASS are $\uparrow \post^*$-effective.
-Since the downward-reachability problem is decidable for VASS \alain{ref ?}, we \alain{we ou Ozkan ?
-qui a fait la preuve ?}
-\mathieu{le papier d'Ozkan: 
-[Petri nets (equipped with any
-finite set of initial states) are $\uparrow \post^*$-effective by above Prop.: Reachability for Petri nets is decidable and recursively equivalent to submarking reachability.
-This corresponds to the anti-ideal reachability problem for Petri nets.]}
-deduce the following:
+\mathieu{Quelque chose de plus à dire sur la charactérisation?}
 
-\begin{proposition}
+VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
+\iffalse
+
+\begin{proposition}\cite{DBLP:journals/corr/abs-2108-00889}
 Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.
@@
 
-\iffalse
-\begin{proof}
-By Proposition~\ref{post*}, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-Let us now show that the downward-reachability problem is decidable for Petri nets.
-Let us consider a downward closed set $D$ in a Petri net.
-Let us consider $B$ a basis of the upward-closed set $S \setminus D$.
-
-$D$ is  defined by ...
-states where your number of counter is bounded by a constant (deduce this from ...
-having to be smaller than elements of the basis $B$), and others where it is unbounded.
-You have $m$ counters where it is bounded by $n$
-then you have $m \times n$ possibilities for the values in these counters
-you have unlimited possibilities for the values of the other counters,
-but that is where the submarking problem comes in handy,
-i.e. you ask the submarking problem for all of the $m \times n$ possibilities.
-
-So downward-set reachability is decidable for PN.
-\end{proof}
 \fi
-
 It is well-known that 
@@
 % Since 
-VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded state-resilience} is decidable for VASS.
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
 Hence {\sc State-resilience} is decidable for %both Petri nets and 

### 2023-04-19 definitions.tex
up

@@
 For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
-
-\mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
-\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$}
+%
+\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
+\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
 When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

### 2023-04-19 introduction.tex
details
section 2

@@
 
-MFCS 2023 \\
-Abstract submission deadline:    	April 24th (AoE) \\
-Paper submission deadline:    	April 28th (AoE)
+{\bf MFCS 2023
+Abstract submission deadline:    	 April 24th (AoE) 
+Paper submission deadline:    	April 28th (AoE)} \\
+%
+{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
+Submission Deadline 	May 2, 2023}\\
+%
+{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
+Notification to authors: September 16, 2022}
+
+trop d'hypotheses trop fortes sur les WSTS
+
+{\bf Our contributions}
+We introduce a more general definition of resilience generalising the two given in \cite{DBLP:conf/gg/Ozkan22}.
+Surprinsingly, the general undecidability statements were not known neither proved.
 

### 2023-04-19 application.tex
up
section 2

@@
 Then $A=(Q,\Sigma,...)$ is $p$-resilient for all $p \geq k_0$ and it is not $p$-resilient for all $p \leq k_0-1$.
-
-\alain{à prouver la BRP...}
+\mathieu{Pas convaincu qu'un tel $k_0$ existe systématiquement: par exemple $q \times a$ et 
+$q \times a b^*$ sont tout les deux réguliers et il n'y a pas de borne maximum sur la taille d'un chemin de $q \times a b^*$ vers $q \times a$.}
+% This also solves the BRP as $A=(Q,\Sigma,...)$ is $k_0$-resilient.
+% \alain{à prouver la BRP...}
 %
@@
 $\pred^*(q,w)$) is computable.
-Reachability is (immediately) decidable for both effective post-regular and effective pred-regular fifo automata.
+Reachability is (immediately) decidable for both effective post-regular or effective pred-regular fifo automata.
 \alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ?

### 2023-04-19 application.tex
up

--- 2023-04-12 application.tex
@@
 %
-The kRP is decidable: let us compute $ \pred^*(\Safe)$ (in polynomial time).
-If $\Bad \subseteq \pred^*(\Safe)$ then one computes the integer $k_0$ such that the maximum of all minimal pathes that go from $\Bad$ to $\Safe$ is equal to $k$.
-Then $A=(Q,\Sigma,...)$ is $p$-resilient for all $p \geq k_0$ and it is not $p$-resilient for all $p \leq k_0-1$.
-
-\alain{à prouver la BRP...}
-%
+%The kRP is decidable: % let us compute $ \pred^{\leq k}(\Safe)$ (in polynomial time). 
+Similarly, since $ \pred^{\leq k}(\Safe)$ is a computable regular language,
+we deduce 
+that the inclusion $\Bad \subseteq \pred^{\leq k}(\Safe)$ is decidable and $k$-resilience too.
+Since $k$-resilience is decidable, bounded-resilience also is;
+we can iteratively check whether $\Bad \subseteq \pred^{\leq k}(\Safe)$,
+if this is the case, $k$-resilience holds, and we stop with a positive,
+else we check whether $ \pred^{\leq k}(\Safe) =  \pred^{*}(\Safe)$,
+if this is the case, then bounded-resilience do not hold,
+else we continue.
 \end{proof}
@@
 %		\section{Control of a WSTS to be resilient}
+
+\mathieu{Tableau résumé de la situation pour les VASS}
+
+\begin{center}
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \Safe~\Bad 
+		& $\uparrow$~ $\downarrow$~ 
+		 & $\downarrow$~ $\uparrow$~
+
+ \\ \hline
+   RP  
+   	& Decidable (Thm~\ref{down-up})  
+   		 & Decidable (Thm~\ref{up-down})
+
+    \\ \hline
+   BRP  
+   &  Decidable (Thm~\ref{down-up}) 
+   		 & ??
+
+    \\ \hline
+      kRP  
+      & Decidable (Thm~\ref{down-up}) 
+      		& ??
+
+       \\ \hline
+   sRP  
+   	& Decidable (Thm~\ref{post srp})
+   		 & Decidable (Thm~\ref{downward srp})
+
+    \\ \hline
+   BsRP  
+   &  Decidable (Thm~\ref{ref ozkan})
+   		 & ??
+
+    \\ \hline
+      ksRP   
+      & Decidable (Thm~\ref{ref ozkan})
+      		& ??
+
+       \\ \hline
+
+\end{tabular}
+\end{center}
 

### 2023-04-19 definitions.tex
section 2
up

@@
 and
-$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. \\
+$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. % \\
+
+The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$, whether $s \to^* t$. It can be reformulated as asking whether 
+$s \in \pred^*(t)$ or asking whether $t \in \post^*(s)$. \\
+
+	\iffalse
+\problemx{Reachability}
+{A transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$}
+{$s \to^* t$? \\}
+	\fi
 
@@
 
+A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of Turing machines
+($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
+$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
+$M_\leq$ the ordering relation “$\leq$”. 
+We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
+exists an additional Turing machine that computes $|\post(x)| \in \N \cup \{\infty\}$ 
+on input $x$,
+with $x \in X_i $. 
+Such a Turing machine, in combination with $M_\rightarrow$, allows computing
+$\post(x)$ whenever the latter is finite. We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
+and if there exists an additional Turing machine that computes a finite basis of 
+$\uparrow \pred (\uparrow x)$
+on input $x$, where $x$ is a state of $\mathscr{S}$. 
+We say that $\mathscr{S}$
+is {\em ideally effective} if
+\begin{itemize}
+\item the function mapping the encoding of a configuration $s$
+to the encoding of the ideal $\downarrow s$ is computable;
+\item inclusion of ideals is decidable;
+\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
+\end{itemize}
+
+\iffalse
+\alain{pas ici}
+\begin{definition}{\cite{BFM-ic17}} 
+A class $C$ of WSTS is {\em ideally effective} if
+\begin{itemize}
+\item the function mapping the encoding of a configuration $s$ of a WSTS
+to the encoding of the ideal $\downarrow s$ is computable;
+\item inclusion of ideals is decidable;
+\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
+\end{itemize}
+\end{definition}
+
+\fi
+
 We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
@@
 Then $t' \in I$ and $s' \in \pred(I )$.
-
-\mathieu{On peut probablement enlever la Proposition $8$ alors.
-La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
 \fi
@@
 
-\alain{pas ici}
-\problemx{$d$-VASS reachability}
-{A $d$-VASS $V=(Q,T)$, and two configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
-{$p(\textbf{u}) \to^* q(\textbf{v})$? \\}
-
-\problemx{$d$-VASS zero-reachability}
-{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
-{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\}
-
 The two previous problems are well-known decidable.

### 2023-04-19 definitions.tex
section 2

@@
 $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
-
-Subsetquently, $Q \times \N^d$ is the set of configurations of $V$.
-For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+%
+Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
+For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
 %

### 2023-04-20 definitions.tex
up

@@
 The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....
-
+\mathieu{ne pas oublier de mettre une définition ici}
 Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.

### 2023-04-20 introduction.tex
up

@@
 \end{itemize}
+
+\begin{center}
+	\begin{figure}
+			\hspace{0.75cm}
+				\label{air control}
+\includegraphics[width=0.85\textwidth]{FigureB}
+	\caption{A channel system with three automata and four channels.
+Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+	\end{figure}
+\end{center}
+
+Consider the channel system in Figure~\ref{air control}.
+It models a scenario in which two aircrafts wants to land in an airport at the same time.
+Safety requires that only one aircraft at a time attempt to land.
+The role of the control tower is to ensure this;
+there are two possible choices: aircraft A waits for aircraft B to land
+or vice-versa. We want ...
+
+% contrôle de trafic aérien, perturbation dans le planning → Bad (par exemple, trop d’avions qui cherchent à atterir en même temps et qui doivent être mis en stand-by au dessus de l’aéroport?) → Safe (trafic fluide à nouveau).
+K-résilience important car les avions peuvent pas rester indéfiniment en stand-by (re : le carburant est un réactif limitant)

### 2023-04-20 state-resilience.tex
up

--- 2023-04-19 reachable-resilience.tex
@@
 
+\iffalse
+\problemx{$d$-VASS zero-reachability}
+{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
+{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\} 
+\fi
+
 It is well known that Reset $d$-VASS are WSTS [ref ? relire WSTS everywhere]. 
@@
 \fi
-~ of zero-reachability in Reset $d$-VASS.
+~ of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
+Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
 %
@@
 
-\mathieu{Quelque chose de plus à dire sur la charactérisation?}
+% \mathieu{Quelque chose de plus à dire sur la charactérisation?}
 
+More concretely,
 VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
@@
 		configurations in $Q \times \{ \textbf{0} \}$
-	where $\textbf{0} = \min(\mathbb{N}^d)$ %$0$ 
+	where $\textbf{0} = \min(\mathbb{N}^d)$, %$0$ 
 % \alain{tu n'as pas le doit d'écrire $0$ sans explication}
@@
 % $0$  can be visited infinitely often. 
-it is possible to visit infinitely often configurations with every vector having its value equal to $0$.
-\alain{incompréhensible, explique}
+it is possible to visit infinitely often configurations with every vector having all values equal to $0$.
+% \alain{incompréhensible, explique}
 Indeed, from such a configuration, then the next step is either also such a configuration, either it is in $\Bad$, in which case, we know it is possible to reach such a configuration again.
-Hence we construct a sequence
+From this we know there exists a sequence
 $(q_i ({\textbf{v}_{i}}) )_{i\in\N}$ 

### 2023-04-21 application.tex
section 5

@@
-\section{Solving resilience with invariants}
+\section{Solving resilience}
+
+\noindent
+\subsection{Computing the maximal resilient subsystem}
+
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp.
+$(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp.
+$\post^*(s_0) \xrightarrow{*} X$) is satisfied.
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
+The \emph{maximal $X$-resilient subsystem} of $(S,\rightarrow,\leq)$ is ....
+\mathieu{ne pas oublier de mettre une définition ici}
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
+Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
+%
+For example, VASS allows such guards.
+  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
+\end{theorem}
+
+\begin{proof}
+Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+%
+%
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
+
+%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
+\end{proof}
+%
+%
+%\noindent
+%{\bf Vector addition system with states}
+
+% Définir VASS
+
+Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
+$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
+%
+Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
+For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+%
+\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
+\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
+When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+
+Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
+Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
+Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
+problèmes: marquage T-bloqué, dead, T-continual ?.
+
+\begin{theorem}{}
+Given a VASS and $X$ upward-closed, the maximal $X$-resilient subsystem is a computable VASS of size 2-exp.
+\end{theorem}
+
+applications to many upward-closed sets $X$:
+$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
+$m$ is dead if the reachability tree from $m$ is finite.
+$m$ is bounded if the reachability set from $m$ is finite.
+$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
+
+\iffalse
+\alain{ me semble maintenant inutile
+%
+\begin{definition}{ (index)}. 
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
+The {\em index} $k(U)$ is the
+smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
+\end{definition}
+%
+%
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
+every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
+%
+\begin{remark}
+$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
+U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
+U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
+=  U \cup \pred(U^k)$.
+\end{remark}
+
+This ensures the following.
+
+\begin{fact}\label{stop condition}
+% Fact 4 (stop condition). 
+If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
+$k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
+\end{fact}
+
+\begin{lemma}
+% Lemma 3 ([1]) 
+% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
+ Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly WSTS, we can decide whether $U  \xrightarrow {*}{} s$.
+%	we can reach $U$ from $s$.
+\end{lemma}
+
+\begin{proof}
+We have to show that we can compute a basis of $U^{k+1}$ if we are given a basis of $U^k $. 
+Then the
+decidability of the stop condition follows directly. Let $B$ be a basis of $U^k$. 
+We have
+$$U^{k+1} = U \cup \pred(U^k ) = U \cup
+\bigcup_{s' \in B}
+\pred(\uparrow \{s' \}).$$
+
+Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
+Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
+\end{proof}}
+
+\fi
+
+The two previous problems are well-known decidable.
+
+\alain{probablement enlever toute la suite et garder seulement un point sur les VASS}
 

### 2023-04-21 definitions.tex
section 5

--- 2023-04-19 definitions.tex
@@
 %
-
-\noindent
-\subsection{Computing the maximal $X$-live subsystem}
-
-Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be WSTS with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
-We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s \in \post^*(s_0)$, $s \xrightarrow{*} X$ is satisfied.
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live.
-The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....
-
-Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
-Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
-Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ are still WSTS.
-%
-For example, VASS allows such guards.
-  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
-
-\begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal $X$-live subsystem is a computable WSTS.
-\end{theorem}
-
-\begin{proof}
-Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
-%
-%
-Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-live and then it is also the maximal $X$-live sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$.
-
-%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
-\end{proof}
-%
-%
-%\noindent
-%{\bf Vector addition system with states}
-
-% Définir VASS
-
-Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
-% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
-$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
-
-Subsetquently, $Q \times \N^d$ is the set of configurations of $V$.
-For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
-%
-\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
-\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
-When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
-
-Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
-Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
-problèmes: marquage T-bloqué, dead, T-continual ?.
-
-\begin{theorem}{}
-Given a VASS and $X$ upward-closed, the maximal $X$-live subsystem is a computable VASS of size 2-exp.
-\end{theorem}
-
-applications to many upward-closed sets $X$:
-$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
-$m$ is dead if the reachability tree from $m$ is finite.
-$m$ is bounded if the reachability set from $m$ is finite.
-$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
-
-\iffalse
-\alain{ me semble maintenant inutile
-%
-\begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
-The {\em index} $k(U)$ is the
-smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
-\end{definition}
-%
-%
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
-every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
-%
-\begin{remark}
-$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
-U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
-U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
-=  U \cup \pred(U^k)$.
-\end{remark}
-
-This ensures the following.
-
-\begin{fact}\label{stop condition}
-% Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
-$k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
-\end{fact}
-
-\begin{lemma}
-% Lemma 3 ([1]) 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly WSTS, we can decide whether $U  \xrightarrow {*}{} s$.
-%	we can reach $U$ from $s$.
-\end{lemma}
-
-\begin{proof}
-We have to show that we can compute a basis of $U^{k+1}$ if we are given a basis of $U^k $. 
-Then the
-decidability of the stop condition follows directly. Let $B$ be a basis of $U^k$. 
-We have
-$$U^{k+1} = U \cup \pred(U^k ) = U \cup
-\bigcup_{s' \in B}
-\pred(\uparrow \{s' \}).$$
-
-Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
-Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
-\end{proof}}
-
-\fi
-
-The two previous problems are well-known decidable.

### 2023-04-27 state-resilience.tex
up

@@
 \begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong upward compatibility ~ & $\uparrow post^*$ effective & both  \\ \hline
+\hline  Hypothesis & strong %upward 
+			compatibility ~ & $\uparrow post^*$ effective 
+			& strong %upward 
+					compatibility \textit{and} $\uparrow post^*$ effective \\ \hline
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline

### 2023-04-28 introduction.tex
up

@@
 
+Some systems are subjects at frequent intervals to accidents, attacks or changes.
+Think for instance of a supply chain, or an airport’s air trafic control.
+In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack [pushed it towards some kind of ‘error state’] and, if it can, whether it can perform the return in a satisfactory timeframe.
+In the exemple of air trafic control, an accident i.e.
+for instance a delay in the arrival of a plance can lead to conflicts of resolution (i.e.
+landing) which lead to prompting planes to remain on standby.
+Since their fuel quantity is limited it is hence pretty important to be able to know that trafic will return to its normal pace quickly enough.
+This is the resilience problem.
+
+Resilience is a concept of growing interest [references].
+Both Okzan and Würdemann [ref] and Prasad and Zuck [ref] have started to consider [casting the problem into] the framework of Well-Structured Transition Systems.
+
 \begin{center}
@@
 there are two possible choices: aircraft A waits for aircraft B to land
-or vice-versa. We want ...
+or vice-versa. 
+We assume that the channel are lossy, i.e. an arbitrary number of messages may be lost from the channels, but that they are still such that out of ten messages, at least one is not lost.
+In this example, the safe, desirable state would be one where the aircrafts have both landed. 
+Resilience here expresses that the aircrafts state will both land, while explicit resilience or $k$-resilience asks whether they will do so in at most $k$ steps. 
+Explicit resilience is of interest since in general we want to ensure the safe state will
+be reached in
+the least amount of step possible. 
+In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time.
+With one out of ten messages guarranteed to not be lost, $40$-resilience hold in the example.
 

### 2023-04-28 resilience.tex
up

@@
 
+% \mathieu{En fait de plus en plus j'ai l'impression que on peut se passer des neutres: même dans une situation où on aurait des configurations neutre j'aurais tendance à dire qu'on peut se ramener à un cas safe complémentaire de bad en duplicants les neutres en deux parties, une avec "on a vu un very bad récement il faut aller vers safe" (et qui dans le nouveau modèle seraient donc bad) et une "on a pas vu de very bad tout vas bien" (et qui dans le nouveau modèle seraient donc safe)
+% $=>$ Rédiger ça plus proprement}
+While considering a sequence of configurations, every prefix is always 
+either in the case that the latest visited $\Bad$ configuration has already been followed by a $\Safe$ configuration
+or in the case that it still has to. 
+Thus, in context, there are no true ``neutral'' configurations: either a neutral configuration is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ configuration, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ configuration and has, accordingly, to reach a $\Safe$ one.
+Hence we will consider not only $\Bad$ and $\Safe$ disjoint but
+% w.l.o.g.
+	also
+$\Bad$ complement of $\Safe$.
+
 \input{resilience-safe-up}

### 2023-04-28 application.tex
up

@@
    	& Decidable (Thm~\ref{down-up})  
-   		 & Decidable (Thm~\ref{up-down})
+   		 & ??
 
@@
    &  Decidable (Thm~\ref{down-up}) 
-   		 & ??
+   		 & Decidable (never hold ?)
 
@@
       & Decidable (Thm~\ref{down-up}) 
-      		& ??
+      		& Decidable (never hold ?)
 
@@
    	& Decidable (Thm~\ref{post srp})
-   		 & Decidable (Thm~\ref{downward srp})
+   		 & ??
 
@@
 \end{center}
+
+\begin{proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+\end{proposition}
+
+\begin{proof}
+Consider a given $k \in \N$.
+Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
+Consider a given Vector addition system, which has a finite number of transitions.
+Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.
+
+$\Bad$ admits a finite basis $B_\Bad$.
+Consider $\textbf{v}_k$ obtained by summing all members of the basis of $\Bad$ and then consider 
+$\textbf{u}_k = \textbf{v}_k + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+
+All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_k $
+and thus, are in $\Bad$, by upward-closedness.
+
+Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
+Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold.
+\end{proof}
+
+\mathieu{
+En train d'essayer de réfléchir voir si ça s'adapte aussi aux VASS.}
 

### 2023-04-28 state-resilience.tex
up

@@
 and $\Bad = \mathbb{N}^d \setminus \Safe$.
-It follows in this case that
-{\sc State-resilience} 
-% \alain{écris toujours pareil les expressions choisies comme state-resilience}
-is equivalent 
-to the problem of whether 
-% $0$  can be visited infinitely often. 
-it is possible to visit infinitely often configurations with every vector having all values equal to $0$.
-% \alain{incompréhensible, explique}
-Indeed, from such a configuration, then the next step is either also such a configuration, either it is in $\Bad$, in which case, we know it is possible to reach such a configuration again.
-From this we know there exists a sequence
-$(q_i ({\textbf{v}_{i}}) )_{i\in\N}$ 
-such that for all $i\in\N$ 
-$\exists j \in \N$, $j>i$, s.t. ${\textbf{v}_j} = \textbf{0}$.
-% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
-Hence the undecidability.
+
+\mathieu{refaire la preuve}
 \end{proof}

### 2023-04-28 application.tex
up

@@
 $\Bad$ admits a finite basis $B_\Bad$.
-Consider $\textbf{v}_k$ obtained by summing all members of the basis of $\Bad$ and then consider 
-$\textbf{u}_k = \textbf{v}_k + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
 
-All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_k $
+All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
 and thus, are in $\Bad$, by upward-closedness.

### 2023-04-28 application.tex
up

@@
    &  Decidable (Thm~\ref{down-up}) 
-   		 & Decidable (never hold ?)
+   		 & Decidable  ?
 
@@
       & Decidable (Thm~\ref{down-up}) 
-      		& Decidable (never hold ?)
+      		& Decidable ?
 
@@
 \mathieu{
-En train d'essayer de réfléchir voir si ça s'adapte aussi aux VASS.}
+resilience never hold in the case of VAS. 
+What happens when you add states ? \\
+So you consider for instance $q_0(u_k)$ instead of just $u_k$. \\
+Everything reachable from $q_0(u_k)$ is still of the form $q(u)$ with u greater than $v_{\Bad}$
+(we'd have $v_{\Bad}$ sum of the coordinate of the elements from the basis of $\Bad$ when not considering states).
+So now, let us fix some state $q$.
+Either there is an element of the basis of $\Bad$ with state $q$.
+Then $q(u)$ is necessarily in $\Bad$ (If this hold for all $q \in Q$, then again $k$-resilience does not hold).
+Either there is none.
+If there is none, then $q(\omega)$ is in the complement of $\Bad$, i.e. $\Safe$.
+Then, of course there are probably other elements in $\Safe$ but from
+the perspective of $\Bad$ and the extremely high elements $q_i(u_k)$  these ‘smaller’ elements are all but unreachable. 
+And the question becomes rather `can we reach $q(\omega)$ from every element of $\Bad$ save from maybe the few smaller ones ?’
+Since $q(\omega)$ here is upward-closed this is decidable i.e. it is asking coverability questions. 
+It’s not 100\% the question «asking, for every element in the basis of $\Bad$, whether or not $q(0)$ is coverable from it», since you could have, for smaller elements in $\Bad$, other elements of $\Safe$ you can reach, but it seem close.
+This would lead to a decision scheme.
+Need to write this more formally/properly.}
 

### 2023-05-01 appendix.tex
up
moved uninteresting cases to appendix

@@
+\section{Case: $\Bad=\downarrow \Bad$ and $\Safe=\downarrow \Safe$}\label{case down down}
+
+The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.
+%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
+
+% In the case \Safe and \Bad are both downward closed, they can both be finite.
+
+\begin{theorem}\label{down-down}
+The resilience problem is undecidable for  effective WSTS with  strong  compatibility such that
+%	with more than one minimal element, 
+$\Safe=\downarrow \Safe$
+and $\Bad=\downarrow \Bad$.
+\end{theorem}
+
+\begin{proof}
+If the set $S$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ has an unique minimal element $m$, then $m$ belongs to both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.
+So let us consider the case of a set $S$ with at least two minimal elements $m_1$ and $m_2$.
+The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
+By undecidability of the reachability problem for effective WSTS with strong compatibility we conclude.
+\end{proof}
+
+\section{Case: $\Bad=\uparrow \Bad$ and $\Safe=\uparrow \Safe$.}\label{case up up}
+
+% \alain{2 clos par le haut ont une intersection non vide....}
+
+Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward closed sets have not an empty intersection.
+We then consider the case where $\Bad$ is as upward-closed
+as it can be without intersecting with $\Safe$,
+i.e. 
+such that there exists an upward-closed set $U$
+such that $\Bad = U \setminus \Safe$.
+
+\iffalse
+\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
+\mathieu{Pas convaincu que ça serve, je vais enlever}
+% hop
+\fi
+
+example: ?
+
+\begin{theorem}\label{up-up}
+The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
+Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}
+
+\begin{proposition}
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+\end{proposition}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
+compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+% \alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
+and there exists $n_0 \in \N$ such that 
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+Hence the equivalence.
+\end{proof}
+
+\begin{corollary}\label{B-up-up}
+The bounded resilience problem is decidable for WSTS with effective pred-basis,
+strong compatibility
+ $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
+\end{corollary}
+
+\begin{theorem}\label{k-up-up}
+The % three %uniform 
+$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
+compatibility, $\Safe=\uparrow \Safe$
+and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
+%	and \Bad is upward closed or downward closed.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong (upward) compatibility,  $\pred^{\leq n}(\Safe)=\uparrow \pred^{\leq n}(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
+% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
+Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
+Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
+% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
+hence the resilience problem is decidable.
+\end{proof}
+
 \section{Proof of Theorem~\ref{liveness reset}}\label{appendix}
@@
 Recall Theorem~\ref{liveness reset}.
+
+\mathieu{À refaire avec des Reset VASS}
 
@@
 
-\section{More Appendix thing if necessary}\label{appendix}
+\section{More Appendix thing if necessary}\label{appendix B}

### 2023-05-01 resilience-safe-down.tex
moved uninteresting cases to appendix

@@
 %%%%%%%%
-\subsection{Case: $\Bad=\uparrow \Bad$.}
+\subsection{Case: $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.}
 %
@@
 This set is upward-closed by (upward) compatibility in VASS.
-
-\subsubsection{Subcase: $\Safe=\downarrow \Safe$}
 
@@
 it is not enough to conclude.
-
-\subsubsection{Case: $\Safe=\uparrow \Safe$.}
-
-% \alain{2 clos par le haut ont une intersection non vide....}
-
-Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward closed sets have not an empty intersection.
-We then consider the case where $\Bad$ is as upward-closed
-as it can be without intersecting with $\Safe$,
-i.e. 
-such that there exists an upward-closed set $U$
-such that $\Bad = U \setminus \Safe$.
-
-\iffalse
-\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
-\mathieu{Pas convaincu que ça serve, je vais enlever}
-% hop
-\fi
-
-example: ?
-
-\begin{theorem}\label{up-up}
-The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
-compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
-% \alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
-and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
-%	and \Bad is upward closed or downward closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong (upward) compatibility,  $\pred^{\leq n}(\Safe)=\uparrow \pred^{\leq n}(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
-% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}

### 2023-05-01 resilience-safe-up.tex
up
moved uninteresting cases to appendix

--- 2023-04-11 resilience-safe-up.tex
@@
-\subsection{Case: $\Bad=\downarrow \Bad$.}
+\subsection{Case: $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$.}
 
@@
 
-\subsubsection{Subcase: $\Safe=\downarrow \Safe$}
-
-The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
-Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.
-%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
-
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
-
-\begin{theorem}\label{down-down}
-The resilience problem is undecidable for  effective WSTS with  strong  compatibility such that
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-\begin{proof}
-If the set $S$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ has an unique minimal element $m$, then $m$ belongs to both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.
-So let us consider the case of a set $S$ with at least two minimal elements $m_1$ and $m_2$.
-The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
-By undecidability of the reachability problem for effective WSTS with strong compatibility we conclude.
-\end{proof}
-
 %%%%
-
-\subsubsection{Subcase: $\Safe=\uparrow \Safe$}
 
@@
 
+%
+
+% \mathieu{I'm using the notion of ideally effective-ness for WBTS here}
+%
 Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
@@
 $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
+A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
+it is post-effective, and there exists a Turing Machine $M_\downarrow$ that computes, on input $s \in S$, % some $\hat(e) \in \hat( ... )$ such that
+ $\downarrow s$, and some Turing Machine $M_{\uparrow^C}$ that computes, on input
+ $\{s_1, s_2, \ldots, s_m\} \in S^m$, 
+ the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$.
 Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
@@
 %
-\alain{definir completion-post-effective. Voir mon papier well behaved....}
+
 %

### 2023-05-01 resilience.tex
up
moved uninteresting cases to appendix

@@
 	also
-$\Bad$ complement of $\Safe$.
+$\Bad$ complement of $\Safe$,
+in which case
+if $\Bad$ is downward closed then $\Safe$ is upward closed and vice-versa.
+Cases where $\Bad$ or $\Safe$ share closure properties i.e. they are both downward
+closed
+have also been studied, \textcolor{red}{but are less interesting}, 
+see 
+Appendix~\ref{case down down}
+and
+Appendix~\ref{case up up} for results.
 

### 2023-05-02 application.tex
up

@@
 
+\begin{proposition}
+{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$ is decidable
+\end{proposition}
+
+\begin{proof}
+Consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+% $\Safe \subseteq Q \times \N^d$
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Consider a given Vector addition system with states, which has a finite number of transitions.
+Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.
+
+$\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $q \in Q$, all configurations reachable from $q(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q'(\textbf{v}')$ with $ \textbf{v}' > \textbf{v}_{\Bad}$.
+
+For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
+then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then $q'(\omega)$ is in the complement of $\Bad$, i.e.
+$\Safe$.
+
+Let us assume from now on $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+all subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+$\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
+$\pred^{\leq k}(q_{j_\ell}(\omega))$.
+
+We now consider the set $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+\mathieu{Is it actually possible to check whether or not it is finite ?}
+If it is finite then we can check for every element in it whether or not it is possible to reach from it $\Safe$ in $k$ or less steps.
+If it is infinite then
+$k$-resilience does not hold.
+\mathieu{...
+because it can only be infinite if there is some kind of unbounded $q(BIG)$ that is not in any of the $\pred^{\leq k}(q_{j_1}(\omega))$ and since from this $q(BIG)$ it's not possible to reach the `smaller' elements of $\Safe$ in $k$ or less steps then it's not possible at all to reach $\Safe$ in $k$ or less steps ergo $k$-resilience does not hold.}
+\end{proof}
+
 %

### 2023-05-02 application.tex
up

@@
 We now consider the set $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
-\mathieu{Is it actually possible to check whether or not it is finite ?}
-If it is finite then we can check for every element in it whether or not it is possible to reach from it $\Safe$ in $k$ or less steps.
+% \mathieu{Is it actually possible to check whether or not it is finite ?}
+% If it is finite then w
+We can check for the elements in it whether or not it is possible to reach from it $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have checked for every element.
 If it is infinite then
-$k$-resilience does not hold.
-\mathieu{...
-because it can only be infinite if there is some kind of unbounded $q(BIG)$ that is not in any of the $\pred^{\leq k}(q_{j_1}(\omega))$ and since from this $q(BIG)$ it's not possible to reach the `smaller' elements of $\Safe$ in $k$ or less steps then it's not possible at all to reach $\Safe$ in $k$ or less steps ergo $k$-resilience does not hold.}
+$k$-resilience does not hold 
+and there exists an element of the form $q(u)$
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(q_{j_i}(\omega))$
+for which it is not possible to reach the `smaller' elements of $\Safe$ in $k$ or less steps and thus it is not possible at all to reach $\Safe$ in $k$ or less steps.
 \end{proof}

### 2023-05-02 application.tex
up

@@
 
+\iffalse
 \mathieu{
@@
 Need to write this more formally/properly.}
+\fi
 
@@
 \end{proof}
+
+\mathieu{À priori on devrais avoir la bounded resilience aussi.
+En fait, comme on a strong monotony alors
+ $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$
+ converge vers
+  $\Bad \setminus (\pred^{*}(q_{j_1}(\omega)) \cup \ldots \pred^{*}(q_{j_\ell}(\omega)) )$.
+  Une fois que le set est stable alors, encore une fois, si il est infini on n'a pas la $k$-résilience
+(pour tout $k$ à partir de ce rang là, du coup)
+et si il est fini alors on peut vérifier, pour chaque élément, si il est possible d'atteindre
+$\Safe$.
+Si il est possible pour chaque élément alors on prends le maximum des chemins depuis chacun de ces éléments et on fait le max de ça et le $k$ qui permet d'atteindre $\pred^{\leq k} = \pred^*$ pour les $q(\omega)$ ci-dessus et ça nous donne un $k_{\max}$ pour lequel on a la $k_{\max}$-résilience.
+Si il est pas possible pour chaque élément alors ça veut dire il y a un élément qui atteindra jamais $\Safe$ et on n'a pas la bounded resilience ni la résilience d'ailleurs. 
+Il faut vérifier que c'est possible pour une configuration de vérifier qu'elle est capable d'atteindre $\Safe$ downward closed. 
+Normalement c'est le cas d'après Okzan (ils disent c'est équiv du submarking pb).}
 

### 2023-05-03 application.tex
up

@@
 Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
-Consider a given Vector addition system, which has a finite number of transitions.
-Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.
+Consider a given vector addition system $V$, which has a finite number of transitions.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
 
-$\Bad$ admits a finite basis $B_\Bad$.
+The set $\Bad$ admits a finite basis $B_\Bad$.
 Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
@@
 Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
-Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold.
+Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
 \end{proof}
@@
 \begin{proposition}
-{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$ is decidable
+{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
 \end{proposition}
@@
  downward-closed.
-Consider a given Vector addition system with states, which has a finite number of transitions.
-Let us call $c_{\max}$ the maximal constant in a coordinate of a transition.
+Consider a given VASS $V$, which has a finite number of transitions.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
 
-$\Bad$ admits a finite basis 
+The set $\Bad$ admits a finite basis 
 $B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
@@
 then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then $q'(\omega)$ is in the complement of $\Bad$, i.e.
-$\Safe$.
+If there is none, then
+% $q'(\omega)$
+$\uparrow q'(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
 
-Let us assume from now on $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+Let us assume from now on 
+% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+ $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
 all subsets of $\Safe$.
 Because these subsets are upward-closed, we can compute a basis of
-$\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
-$\pred^{\leq k}(q_{j_\ell}(\omega))$.
+% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
+% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
 
-We now consider the set $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+We now consider the set 
+% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
 % \mathieu{Is it actually possible to check whether or not it is finite ?}
 % If it is finite then w
-We can check for the elements in it whether or not it is possible to reach from it $\Safe$ in $k$ or less steps.
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
 If the set is finite then we stop the procedure once we have checked for every element.
@@
 $k$-resilience does not hold 
-and there exists an element of the form $q(u)$
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(q_{j_i}(\omega))$
-for which it is not possible to reach the `smaller' elements of $\Safe$ in $k$ or less steps and thus it is not possible at all to reach $\Safe$ in $k$ or less steps.
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
 \end{proof}
 
+\iffalse
 \mathieu{À priori on devrais avoir la bounded resilience aussi.
@@
 Normalement c'est le cas d'après Okzan (ils disent c'est équiv du submarking pb).}
+\fi
+
+\begin{proposition}
+{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
+\end{proposition}
+
+\begin{proof}
+$\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
+then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+% $q'(\omega)$
+$\uparrow q'(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+ $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
+% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+
+There exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards ...
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+\end{proof}
 

### 2023-05-04 application.tex
up

@@
 Consider a given vector addition system $V$, which has a finite number of transitions.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
 
@@
 $\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $q \in Q$, all configurations reachable from $q(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q'(\textbf{v}')$ with $ \textbf{v}' > \textbf{v}_{\Bad}$.
+For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
 
-For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
-then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
 If there is none, then
 % $q'(\omega)$
-$\uparrow q'(\textbf{0})$
+$\uparrow q(\textbf{0})$
  is in the complement of $\Bad$, i.e. $\Safe$.
@@
  $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$.
+all subsets of $\Safe$ and that their union contain all upward-closed supsets of $\Safe$.
 Because these subsets are upward-closed, we can compute a basis of
@@
 q_{i_m}(\textbf{v}_{i_m}) \}$.
-For all $q' \in Q$, either there is an element of the basis of $\Bad$ with state $q'$ \----
-then $q'(\textbf{v}')$ with $\textbf{v}' > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
 If there is none, then
 % $q'(\omega)$
-$\uparrow q'(\textbf{0})$
+$\uparrow q(\textbf{0})$
  is in the complement of $\Bad$, i.e. $\Safe$.

### 2023-05-04 application.tex
sections
up

@@
 $\pred^*(q,w)$) is computable.
+Lossy channel systems are effectively pred-regular and post-regular but they are not effectively post-regular.
+
 Reachability is (immediately) decidable for both effective post-regular or effective pred-regular fifo automata.
@@
 
+\mathieu{
+VASS donc la base de $\uparrow post^*(s_0)$ est calculable. 
+Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward closed donc ça se fait
+					de décider si c'est  $\subseteq \pred^{\leq k}(\Safe)$
+	(re: $k$-résilience décidable voir Prop. 34)
+Problème: on pourrait avoir $\uparrow post^*(s_0) \neq  post^*(s_0)$.
+Si $ \Bad \cap \uparrow \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$ alors on a aussi
+$ \Bad \cap  \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$.
+Si non, alors
+notre procédure décisionnelle pour vérifier
+ça
+donne un témoin.
+Le témoin, on peut vérifier si il est bien dans $\post^*(s_0)$
+(c'est juste la question de est-ce que on peut l'atteindre depuis $s_0$
+et l'accessibilité est décidable).
+Si il est effectivement dans $\post^*(s_0)$ alors c'est bon on a aussi un contre exemple
+qui prouve que
+$ \Bad \cap  \post^*(s_0) \nsubseteq  \pred^{\leq k}(\Safe)$.
+Si le témoin n'est pas dans $\post^*(s_0) \ldots$ alors
+$\ldots$
+j'imagine on peut reprendre la procédure décisionnelle sur 
+$ \Bad \cap \uparrow \post^*(s_0)$
+mais en excluant ce premier témoin,
+et ou bien on a que c'est bon si on enlève ce témoins alors on a bien 
+$\subseteq \pred^{\leq k}(\Safe)$
+ou bien ... on obtiens un nouveau témoin ... pour lui aussi on peut
+vérifier si il est effectivement dans $\post^*(s_0)$, et ainsi de suite.
+Si jamais la $k$-résilience n'est pas vérifiée, on vas bien finir par
+trouver un témoin que ça n'est pas vérifiée.
+Si jamais la $k$-résilience est vérifiée par contre ...
+$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
+doit être fini, on a dit.
+Est-ce qu'il y aurais un moyen de calculer ça plutôt ?
+Par exemple,
+on regarde
+$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$,
+comme on a le $\uparrow$ 
+devant $\uparrow \post^*(s_0)$
+alors
+deux choses se produisent,
+d'abords les "trous" entre les éléments de $ \post^*(s_0)$ sont comblés.
+Sauf que si un "trou" est un contre exemple et qu'on a qu'un nombre fini de
+trou alors "on voit bien" que c'est un trou en testant que
+ça appartiens ou non à $ \post^*(s_0)$ et "ça va" on peut s'en débarasser.
+L'autre chose qui se produit, c'est que on a les droites qui vont vers plus l'infini...
+Ça c'est plus difficile à gèrer parce que si
+il y a qu'un nombre fini d'éléments dans $ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
+et pas dans
+$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
+ça se fait,
+mais si il y a un nombre infini d'éléments dans
+$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
+potentiellement pas dans le plus petit,
+alors c'est plus compliqué de faire notre procédure décisionnelle.
+Il nous faudrait pouvoir décider si $\post^*(s_0)$ est fini ou non.
+Décider si $\post^*(s_0)$ est borné ou non.
+Est-ce que ça peut se réduire à un submarking problem ?
+Il faudrait prendre le complémentaire peut-être ...}
+
 %

### 2023-05-15 application.tex
sections

@@
 %%%
-\subsection{LCM, VASS, semilinear VASS, Integer et Continuous VASS}
+\subsection{VASS and variations of VASS}
 
 \begin{theorem}{}
-Resilience is decidable for  lossy counter machines, semilinear VASS, Integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
+Resilience is decidable for VASS when $\Safe$ and $\Bad$ are semilinear sets.
+\end{theorem}
+
+\begin{proof}
+\alain{faire d'abord la preuve pour les VASS.}
+\end{proof}
+
+\begin{theorem}{}
+Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
 \end{theorem}
@@
 Hence since the inclusion between two semilinear sets is decidable, we deduce that  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
-
-%		semilinear VASS.
- \alain{le plus simple serait que $\pred^*(L)$ is semilinear because $\post^*(s)$ is semilinear , peut-être en inversant le VASS V en V'on aurait pred*(V)=post*(V') hence $post^*(L)$ is also semilinear ?
-hence pred ????}
-
-%		Integer VASS
-Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
-
 \end{proof}
 
-We also deduce immediately that resilience is decidable for $2$-VASS that have a semilinear reachability relation.
+\begin{theorem}{}
+Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
+\end{theorem}
+
+\begin{proof}
+		Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
+\end{proof}
 
@@
 \end{proof}
-
-open for VASS ???
-
-%		\section{Control of a WSTS to be resilient}
 

### 2023-05-15 state-resilience.tex
sections

@@
 			& strong %upward 
-					compatibility \textit{and} $\uparrow post^*$ effective \\ \hline
+					compatibility \textit{+} $\uparrow post^*$ effective \\ \hline
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline

### 2023-05-15 application.tex
sections

@@
-\section{Solving resilience}
-
-\noindent
-\subsection{Computing the maximal resilient subsystem}
+\section{Computing the maximal resilient subsystem}
 
@@
 
-\alain{probablement enlever toute la suite et garder seulement un point sur les VASS}
-
-In this section, we study the general resilience problem.
+\section{Deciding resilience for extended automata}
 
@@
 
-Since for  lossy channel systems $\pred^*(L)$ with $L=\uparrow L$ is a computable (upward-closed) regular language, we deduce :
+Since for lossy channel systems $(Q, \Sigma,d,...)$ with $d$ fifo channels and a set $L \subseteq Q \times (\Sigma^*)^d$, the set $\pred^*(\uparrow L)$ is a computable upward-closed regular language, we deduce :
 
@@
 \begin{proof}
-\alain{faire d'abord la preuve pour les VASS.}
+\alain{faire la preuve pour les VASS.}
 \end{proof}
@@
 \begin{proof}
-		Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
+	\alain{à faire}	Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
 \end{proof}
@@
 The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE.
-Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE.
-\alain{and may be in coNP-complete.
-Recall that reachability is P-complete for continuous VASS}
+the rationals with addition and order whose complexity is EXPSPACE. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE \cite{ref}. 
+%			\alain{and may be in coNP-complete. Recall that reachability is P-complete for continuous VASS}
 \end{proof}
@@
 Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
-Consider a given vector addition system $V$, which has a finite number of transitions.
+Consider a given vector addition system $V$, \alain{bizarre de dire ça "which has a finite number of transitions."}
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
@@
 {\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
-\end{proposition}
+\end{proposition}\alain{quelle complexité ?}
 
 \begin{proof}
-Consider a given $k \in \N$.
+Let a given $d$-VASS $V$ and consider a given $k \in \N$.
 Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
@@
  downward-closed.
-Consider a given VASS $V$, which has a finite number of transitions.
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
@@
 \begin{proposition}
-{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
+{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable. \alain{quelle complexité ?}
 \end{proposition}

### 2023-05-16 application.tex
sections

@@
 
-Since for lossy channel systems $(Q, \Sigma,d,...)$ with $d$ fifo channels and a set $L \subseteq Q \times (\Sigma^*)^d$, the set $\pred^*(\uparrow L)$ is a computable upward-closed regular language, we deduce :
+Since for lossy channel systems $(Q, \Sigma,d,...)$ with $d$ fifo channels and a set of states $L \subseteq Q \times (\Sigma^*)^d$, the set $\pred^*(\uparrow L)$ is a computable upward-closed regular language, we deduce :
 
@@
 
+Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logics.
+
 \begin{theorem}{}
@@
 \begin{proof}
-Resilience is decidable for lossy counter machines with $\Safe$ and $\Bad$ semilinear sets as a consequence of Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that implies that $\pred^*(\Safe)$ is a computable semilinear set.
-Hence since the inclusion between two semilinear sets is decidable, we deduce that  $\Bad \subseteq \pred^*(\Safe)$ is decidable.
+We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
+Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
 \end{proof}
@@
 \begin{proof}
-	\alain{à faire}	Recall that reachability is NP-complete for integer VASS. We consider semilinear sets in $\mathbb{Z}^d$.
+The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\alain{à faire}	Recall that reachability is NP-complete for integer VASS.
 \end{proof}

### 2023-05-16 definitions.tex
sections

@@
-\newcommand{\Bad}{\textsf{Bad}}
-\newcommand{\Safe}{\textsf{Safe}}
-
 \newcommand{\pred}{\textsf{pred}}
@@
 % \renewcommand{\succ}{\textsf{Succ}}
+
+\newcommand{\Bad}{\textsf{Bad}}
+\newcommand{\Safe}{\textsf{Safe}}
 
@@
 
-The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
+%The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
 
@@
 \noindent
-{\bf Transition systems}
-
-\begin{definition}
+Let us recall thz well-known framework of transition systems. 
  A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
@@
  the set of states, denoted as the set of {\em transitions}. 
-
+%
 \iffalse
@@
 \fi
-\end{definition}
-
+%
 % We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$.
-We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
+We use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
 We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
@@
 \noindent
-{\bf Well-quasi-orderings}
-
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$.
-
-We abbreviate $x \leq y \not\leq x$ by $x < y$.
+A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. We abbreviate $x \leq y \not\leq x$ by $x < y$.
 % A partial ordering (a po) is an antisymmetric qo
@@
 % 	do we need po ?
-
-We now recall a few results from the theory of well-orderings (add reference [...]).
+Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
+A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
+To any subset $A \subseteq X$, we may associate
+its {\em upward-closure},
+ $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
+ and its 
+ {\em downward-closure},
+ $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
+We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
+as $\uparrow x$ (resp. $\downarrow x$).
+%
+A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$.
 
@@
 \end{definition}
-
-There exists a stronger formulation.
-$(X,\leq)$ is a wqo iff any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
-
+%
 % \textcolor{red}{Add lemma about infinite increasing subsequences ?}
@@
 
-Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
-A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
-It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
-To any subset $A \subseteq X$, we may associate
-its {\em upward-closure},
- $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
- and its 
- {\em downward-closure},
- $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
-as $\uparrow x$ (resp. $\downarrow x$).
-
-A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$. 
 % Higman investigated ordered sets with the finite basis property.
@@
 Wqo admits many other equivalent formulations. For instance: 
-$(X,\leq)$ is a wqo if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
+$(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$)
+if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
 upward-closed subsets in $X$ eventually stabilizes, 
@@
 
+\noindent
 Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo.
@@
 \noindent
-{\bf Well-structured transition systems}
+Let us now recall the most general definition of well-structured transition systems.
 
@@
 $M_\leq$ the ordering relation “$\leq$”.
+
 We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
 exists an additional Turing machine that computes $|\post(x)| \in \N \cup \{\infty\}$ 
-on input $x$,
-with $x \in X_i $. 
+on input $x$, with $x \in X_i $. 
 Such a Turing machine, in combination with $M_\rightarrow$, allows computing
-$\post(x)$ whenever the latter is finite. We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
-and if there exists an additional Turing machine that computes a finite basis of 
-$\uparrow \pred (\uparrow x)$
-on input $x$, where $x$ is a state of $\mathscr{S}$. 
+$\post(x)$ whenever the latter is finite.
+
+We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
+%\alain{étrange ces deux propositions, l'une suffirait}
+
+\iffalse
+Proof.
+Assume $s \in \pred (I )$.
+Then $s \rightarrow t$ for some $t \in I $.
+If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
+Then $t' \in I$ and $s' \in \pred(I )$.
+\fi
+%
+
+Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+
+\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
+A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} if there exists an algorithm accepting
+any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
+\end{definition}
+
+For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
+When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
+Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
+$U_k = U_{k_0}$ for all $k \geq k_0$.
+We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
+\\
+
+With the effective pred-basis hypothese, we obtain:
+
+\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
+A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+Hence coverability is decidable.
+\end{theorem}
+
+%We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
+%and if there exists an additional Turing machine that computes a finite basis of 
+%$\uparrow \pred (\uparrow x)$
+%on input $x$, where $x$ is a state of $\mathscr{S}$.
+
 We say that $\mathscr{S}$
-is {\em ideally effective} if
-\begin{itemize}
-\item the function mapping the encoding of a configuration $s$
-to the encoding of the ideal $\downarrow s$ is computable;
-\item inclusion of ideals is decidable;
-\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-\end{itemize}
-
+is {\em ideally effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
+to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
+%
 \iffalse
@@
 \end{definition}
-
-\fi
-
-We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
-%\alain{étrange ces deux propositions, l'une suffirait}
-\iffalse
-Proof.
-Assume $s \in \pred (I )$.
-Then $s \rightarrow t$ for some $t \in I $.
-If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
-Then $t' \in I$ and $s' \in \pred(I )$.
+%
 \fi
 %
-Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+%
+With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
 
-\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-\end{definition}
-
-With the effective pred-basis hypothese, we obtain:
-
-\begin{theorem}(ref)
-A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+\begin{theorem}\cite{BFM-ic17}
+Coverability is decidable for any ideally effective ordered compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
 \end{theorem}
@@
 %			\end{fact}
-For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
-When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
-Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
-$U_k = U_{k_0}$ for all $k \geq k_0$.
-We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
-\\
 %

### 2023-05-16 introduction.tex
up
sections

@@
 {\bf Context.} In the quite informal paper \cite{DBLP:journals/corr/PrasadZ16}, Prasad and Zuck studied resilience of a process to an adversary and they used the framework of WSTS enjoying both upward and downward monotonies to prove the decidability of a kind of resilience.
-
-{\bf MFCS 2023
-Abstract submission deadline:    	 April 24th (AoE) 
-Paper submission deadline:    	April 28th (AoE)} \\
 %
-{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
-Submission Deadline 	May 2, 2023}\\
+%{\bf MFCS 2023
+%Abstract submission deadline:    	 April 24th (AoE) 
+%Paper submission deadline:    	April 28th (AoE)} \\
+%%
+%{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
+%Submission Deadline 	May 2, 2023}\\
 %
@@
 
+Resilience is an instance of the home space problem: let $L_1,L_2$ be two sets of states, the home space for  $L_1,L_2$ is the following question:		
+	$\post^*(L_1)	\subseteq \pred^*(L_2)$
+\alain{ne trouve pas la macro pred et post}
+
 \begin{center}
@@
 			\hspace{0.75cm}
-				\label{air control}
 \includegraphics[width=0.85\textwidth]{FigureB}
@@
 Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+					\label{air control}
 	\end{figure}
@@
 
+\alain{il faut mieux expliquer le modèle, le rôle des 4 canaux, des messages, remarquer que 2 canaux sont des compteurs,....}
+
 % contrôle de trafic aérien, perturbation dans le planning → Bad (par exemple, trop d’avions qui cherchent à atterir en même temps et qui doivent être mis en stand-by au dessus de l’aéroport?) → Safe (trafic fluide à nouveau).

### 2023-05-16 application.tex
sections

@@
 Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp.
-$(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp.
-$\post^*(s_0) \xrightarrow{*} X$) is satisfied.
+
+\begin{definition}
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
+\end{definition}
+
 We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
@@
 Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
-
+\end{proof}
 %		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
-\end{proof}
 %
@@
 %{\bf Vector addition system with states}
-
-% Définir VASS
-
+%
 Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
@@
 
-Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
+Let $X \subseteq \mathds{N}^d$ be an upward-closed set and $B_X$ its minimal basis: $X=\uparrow B_X$.
+The number of elements of $min(\pred^*(\uparrow B_X))$ and the size of the minimal elements of $min(\pred^*(X))$ have been studied in
+\cite{DBLP:conf/rp/BozzelliG11}: both are bounded by $2^{2^n}$ where $n$ is the size of the VASS.
+
+ pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
 Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
@@
 \begin{proof}
-\alain{faire la preuve pour les VASS.}
+\alain{faire la preuve pour les VASS.} From \cite{DBLP:journals/corr/abs-2207-02697}...
 \end{proof}

### 2023-05-16 application.tex
sections

@@
 %
-Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
 % The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
@@
 
-Let $X \subseteq \mathds{N}^d$ be an upward-closed set and $B_X$ its minimal basis: $X=\uparrow B_X$.
-The number of elements of $min(\pred^*(\uparrow B_X))$ and the size of the minimal elements of $min(\pred^*(X))$ have been studied in
-\cite{DBLP:conf/rp/BozzelliG11}: both are bounded by $2^{2^n}$ where $n$ is the size of the VASS.
+\noindent
 
- pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
-Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman....
-Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman.
-problèmes: marquage T-bloqué, dead, T-continual ?.
+Let  $V$ be a $d$-VASS, $X \subseteq \mathds{N}^d$ be an upward-closed set and $B_X$ its minimal basis: $X=\uparrow B_X$.
+The number of elements of $min(\pred^*(\uparrow B_X))$ and the size of the minimal elements in $min(\pred^*(X))$ have been studied in
+\cite{DBLP:conf/rp/BozzelliG11}: both sizes are bounded by $2^{2^n}$ where $n$ is the size of $V$.
 
 \begin{theorem}{}
-Given a VASS and $X$ upward-closed, the maximal $X$-resilient subsystem is a computable VASS of size 2-exp.
+Let $V$ be a VASS and $X$ be an upward-closed set. The maximal $X$-resilient subsystem of $V$ is a computable VASS of size $2^{2^n}$ where $n$ is the size of $V$.
 \end{theorem}
 
-applications to many upward-closed sets $X$:
-$m$ is $\hat{T}$-blocked if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
-$m$ is dead if the reachability tree from $m$ is finite.
-$m$ is bounded if the reachability set from $m$ is finite.
-$m$ is $\hat{T}$-continual if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
+We may use the previous Theorem with different upward-closed sets $X$. Let us first recall the definition of 4 types of configurations used in \cite{DBLP:journals/acta/ValkJ85}.
 
+\begin{definition}
+Let $V = (Q,T)$ be a $d$-VASS, $\hat{T} \subseteq T$, and $m$ a configuration in $Q \times \N^d$.
+\begin{itemize}
+\item  $m$ is \em{$\hat{T}$-blocked} if no $t \in \hat{T}$ is fireable in $\post^*(m)$.
+\item $m$ is \em{dead} if the reachability tree from $m$ is finite.
+\item $m$ is \em{bounded} if the reachability set from $m$ is finite.
+\item $m$ is \em{$\hat{T}$-continual} if there is an unfinite run from $m$ such that all transitions in $\hat{T}$ appear infinitely often.
+\end{itemize}
+\end{definition}
+
+Each of these 4 questions is decidable: $m$ is $\hat{T}$-blocked iff the preconditions of each transition in $\hat{T}$ is not coverable from $m$.
+$m$ is dead iff $V$ with $m$ as initial configuration terminates. $m$ is bounded iff $V$ with $m$ as initial configuration is bounded.
+The first two problems reduce to coverability which is EXPSPACE.
+The third problem is also known in EXSPACE.
+The fourth problem can be decided by analysing circuits in the coverability graph \cite{DBLP:journals/acta/ValkJ85} but it can be shown in EXPSPACE.
+
+.
+Let us consider the following four associated upward-closed sets.
+
+\begin{itemize}
+\item  NotBlocked($\hat{T})= \{ m \in Q \times \N^d \mid m$ is not $\hat{T}$-blocked\}.
+\item NotDead = $\{ m \in Q \times \N^d \mid m$ is not dead\}.
+\item UnBounded = $\{ m \in Q \times \N^d \mid m$ is not bounded\}.
+\item Continual($\hat{T}$) = $\{ m \in Q \times \N^d \mid m$ is $\hat{T}$-continual\}.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \iffalse
@@
 \fi
-
-The two previous problems are well-known decidable.
+%%%%%%%%%%%%%%%%%%%
 

### 2023-05-16 application.tex
sections
up

@@
 
-.
 Let us consider the following four associated upward-closed sets.
@@
 \end{itemize}
+
+\cite{DBLP:journals/acta/ValkJ85} showed from a VASS $V$ how to construct another VASS $V'$ that has the property...
 
@@
 
+\begin{theorem}
+{\sc Resilience} is decidable for VASS when $\Safe$ is semilinear
+and $\Bad = S \setminus \Safe$.
+\end{theorem}
+
+\begin{proof}
+Hypothesis: $\Safe$ is semilinear, $\Bad = S \setminus \Safe$.
+Then $\Bad$ is semilinear too.
+{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
+We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
+Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
+$$\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).$$
+Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
+$\Bad$ and $\Safe$ are both semilinear \cite{DBLP:journals/corr/abs-2207-02697}, 
+{\sc Resilience} is decidable.
+\end{proof}
+
+\begin{proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+\end{proposition}
+
 \begin{theorem}{}
@@
 \end{center}
-
-\begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
-\end{proposition}
 

### 2023-05-17 application.tex
sections

@@
 Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logics.
+Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
 
@@
 
+Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.
+
 \begin{theorem}{}
@@
 The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
-\alain{à faire}	Recall that reachability is NP-complete for integer VASS.
+\alain{à faire}
+%		Recall that reachability is NP-complete for integer VASS.
 \end{proof}
+
+Continuous VASS \cite{DBLP:journals/tocl/BlondinFHH17} are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
 
@@
 The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE \cite{ref}. 
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE. 
 %			\alain{and may be in coNP-complete. Recall that reachability is P-complete for continuous VASS}

### 2023-05-17 application.tex
up
sections

@@
 For the sake of simplicity, we consider automata with an unique fifo channel (that have the power of turing machines).
-In 1986, Pachl introduced in \cite{} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability (with a regular inductive invariant to obtain a witness of non-reachability).
+In 1982, Pachl introduced in \cite{DBLP:journals/corr/cs-LO-0306121} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability (with a regular inductive invariant to obtain a witness of non-reachability).
 Let $(q,w)$ be an initial state and $(q',w')$ be a state.
-Since $\post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $\post(L) \subseteq L$ (inductive invariant).
+When $\post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $\post(L) \subseteq L$ (inductive invariant).
 We know that if $(q',w') \not\in \post^*(q,w)$ then there exists, at least, a regular language $L$ such that $\post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $\post^*(q,w)$ is such inductive invariant.
@@
 $\pred^*(q,w)$) is computable.
-Lossy channel systems are effectively pred-regular and post-regular but they are not effectively post-regular.
+Lossy channel systems are post-regular (since $\post^*(q,w)$ is downward-closed) but they are not effectively post-regular; they are also effectively pred-(upward-closed) regular.
 
 Reachability is (immediately) decidable for both effective post-regular or effective pred-regular fifo automata.
-\alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ?
+%\alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ?
 contre-exemples ?
@@
 %%%
-\subsection{VASS and variations of VASS}
+\subsection{Resilience for VASS}
 
@@
 The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also decidable in EXPSPACE. 
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also in EXPSPACE. 
 %			\alain{and may be in coNP-complete. Recall that reachability is P-complete for continuous VASS}
@@
 
+\alain{c'est la preuve de quoi en dessous ?}
 \begin{proof}

### 2023-05-17 application.tex
sections

@@
 
-\cite{DBLP:journals/acta/ValkJ85} showed from a VASS $V$ how to construct another VASS $V'$ that has the property...
+\cite{DBLP:journals/acta/ValkJ85} showed from a VASS $V$ and one of the 4 properties P, how to construct the maximal VASS $V'$ that has the property P which are exactly the maximal $X$-resilient VASS.
 
@@
 \end{center}
+
+\begin{proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+\end{proposition}
 

### 2023-05-17 application.tex
up

@@
 
-\alain{c'est la preuve de quoi en dessous ?}
 \begin{proof}
-Consider a given $k \in \N$.
-Consider $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed.
-Consider a given vector addition system $V$, \alain{bizarre de dire ça "which has a finite number of transitions."}
+Consider a given $k \in \N$,
+ $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
+consider a given vector addition system $V$.
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.

### 2023-05-17 resilience-safe-down.tex
sections

@@
 
+Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is downward compatible if
+for all $s_1, s_2, t_1 \in S$ with
+$s_2 \leq s_1$ and $s_1 \to t_1$
+there
+exists $t_2 \in S$ with
+$t_2 \leq t_1$ and $s_2 \to^* t_2$.
+
 \begin{corollary}
@@
 \begin{proof}
-Downward compatibility of $\mathscr{S}=(S,\rightarrow, \leq)$ 
-means that
-for all $s_1, s_2, t_1 \in S$ with
-$s_2 \leq s_1$ and $s_1 \to t_1$
-there
-exists $t_2 \in S$ with
-$t_2 \leq t_1$ and $s_2 \to^* t_2$.
 

### 2023-05-17 introduction.tex
up

@@
 	$\post^*(L_1)	\subseteq \pred^*(L_2)$
-\alain{ne trouve pas la macro pred et post}
 
@@
 
-Consider the channel system in Figure~\ref{air control}.
-It models a scenario in which two aircrafts wants to land in an airport at the same time.
-Safety requires that only one aircraft at a time attempt to land.
-The role of the control tower is to ensure this;
+Consider the channel system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time.
+
+The automata on the left and right represents the aircrafts attempting to land, while the middle one represent the control tower.
+The control tower can send messages to the aircrafts using channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using respectively channel $c_3$ and $c_4$.
+Remark channels $c_3$ and $c_4$ have a unary alphabet
+in the exemple and therefore can be seen as counters.
+
+Safety requires that only one aircraft at a time attempt to land. The role of the control tower is to ensure this;
 there are two possible choices: aircraft A waits for aircraft B to land
-or vice-versa. 
-We assume that the channel are lossy, i.e. an arbitrary number of messages may be lost from the channels, but that they are still such that out of ten messages, at least one is not lost.
+or vice-versa. To make its choice known to the aircrafts, the control tower will send them messages using the channels. 
+We assume that the channel are
+lossy \-- i.e. an arbitrary number of messages may be lost from the channels \--  but that they are still such that out of ten messages, at least one is not lost.
 In this example, the safe, desirable state would be one where the aircrafts have both landed. 

### 2023-05-17 application.tex
sections

@@
 We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
-The \emph{maximal $X$-resilient subsystem} of $(S,\rightarrow,\leq)$ is ....
+The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
 \mathieu{ne pas oublier de mettre une définition ici}

### 2023-05-22 appendix.tex
up

@@
 The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
-Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward closed.
-%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed.
+%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
 
-% In the case \Safe and \Bad are both downward closed, they can both be finite.
+% In the case \Safe and \Bad are both downward-closed, they can both be finite.
 
@@
 
-Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward closed sets have not an empty intersection.
+Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward-closed sets have not an empty intersection.
 We then consider the case where $\Bad$ is as upward-closed
@@
 \begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
+Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
 Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
 Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
+Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
 % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
@@
 and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
-%	and \Bad is upward closed or downward closed.
+%	and \Bad is upward-closed or downward-closed.
 \end{theorem}

### 2023-05-22 application.tex
up

@@
 The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
-\mathieu{ne pas oublier de mettre une définition ici}
+
 Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
@@
 VASS donc la base de $\uparrow post^*(s_0)$ est calculable. 
-Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward closed donc ça se fait
+Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward-closed donc ça se fait
 					de décider si c'est  $\subseteq \pred^{\leq k}(\Safe)$

### 2023-05-22 definitions.tex
changed macro definnition position
up

@@
-\newcommand{\pred}{\textsf{pred}}
-\newcommand{\post}{\textsf{post}}
-% \renewcommand{\succ}{\textsf{Succ}}
-
-\newcommand{\Bad}{\textsf{Bad}}
-\newcommand{\Safe}{\textsf{Safe}}
-
 \section{Well-structured transition systems}\label{section definitions}
@@
 Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo.
-In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
+In such \emph{Finite Antichains} sets $(X,\leq)$, every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward-closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
 
-In what follows, a downward closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
+In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
 %		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
-and an upward closed set $U$ is represented by its finite set of minimal elements. \\
+and an upward-closed set $U$ is represented by its finite set of minimal elements. \\
 % $M_U$ of its  satisfying $U=\uparrow M_U$.

### 2023-05-22 resilience.tex
up

--- 2023-04-28 resilience.tex
@@
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS rely on the computation of upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
+Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
 
@@
 
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward closed and $\Safe$ is upward closed.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
 %		
 We deduce 
-% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed sets.}
+% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
 % \mathieu{changer ça}
-decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward closed and upward closed as seen in Section~\ref{synthesis}.
+decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
 
@@
 	also
-$\Bad$ complement of $\Safe$.
+$\Bad$ complement of $\Safe$,
+in which case
+if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
+Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. they are both downward
+closed or both upward-closed,
+have also been studied, 
+see 
+Appendix~\ref{case down down}
+and
+Appendix~\ref{case up up} for results.
 

### 2023-05-22 resilience-safe-up.tex
up
downward closed and upward closed are now consistently written with a '-'

@@
 Let us choose $\Bad$ to be the set of states from which there is no infinite run.
-If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward closed.
+If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward-closed.
 
@@
 from which it is always possible to fire a transition.
-This set of states is upward closed and $\Safe=S-\Bad=\uparrow \Safe$.
+This set of states is upward-closed and $\Safe=S-\Bad=\uparrow \Safe$.
 

### 2023-05-22 resilience.tex
downward closed and upward closed are now consistently written with a '-'

@@
 if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
-Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. they are both downward
-closed or both upward-closed,
+Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. they are both downward-closed or both upward-closed,
 have also been studied, 

### 2023-05-22 application.tex
up

@@
 
-\begin{theorem}
-{\sc Resilience} is decidable for VASS when $\Safe$ is semilinear
-and $\Bad = S \setminus \Safe$.
-\end{theorem}
-
-\begin{proof}
-Hypothesis: $\Safe$ is semilinear, $\Bad = S \setminus \Safe$.
-Then $\Bad$ is semilinear too.
-{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
-We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
-Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
-$$\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).$$
-Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
-$\Bad$ and $\Safe$ are both semilinear \cite{DBLP:journals/corr/abs-2207-02697}, 
-{\sc Resilience} is decidable.
-\end{proof}
-
 \begin{proposition}
@@
 \alain{faire la preuve pour les VASS.} From \cite{DBLP:journals/corr/abs-2207-02697}...
+\mathieu{En train de penser que la meilleure stratégie ici est de réduire "w.l.o.g" au cas où ils sont complémentaires encore faut-il montrer que ce serait without loss of generality.
+}
+
+\mathieu{...}
+Without loss of generality hence let us consider only the case when $\Safe$ is semilinear
+and $\Bad = S \setminus \Safe$ is semilinear too.
+{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
+We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
+Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
+$$\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).$$
+Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
+$\Bad$ and $\Safe$ are both semilinear \cite{DBLP:journals/corr/abs-2207-02697}, 
+{\sc Resilience} is decidable.
 \end{proof}

### 2023-05-22 resilience.tex
up

@@
 if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
-Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. they are both downward-closed or both upward-closed,
+Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. are both downward-closed or both upward-closed,
 have also been studied, 

### 2023-05-22 application.tex
up

@@
 }
-
+When asking about resilience, in context, there are no true ``neutral'' configurations.
+Hence from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
+$V$, $\Safe$ and $\Bad$ if and only if {\sc Resilience} holds for
+$V'$, $\Safe'$ and $\Bad'$.
+For the construction, $V'$ will have two copies of the control states of $V$; $Q_{A}$ and $Q_B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp.
+$\Bad$.
+From a configuration with state in $Q_{A}$ you go to $Q_B$ when you are in a $\Bad$ place
+and from a configuration with state in $Q_{B}$ you go to $Q_{A}$ when you are in a $\Safe$ place.
+$\Safe'$ is $Q_A \times \N^d$ and $\Bad'$ is $Q_B \times N^d$.
 \mathieu{...}

### 2023-05-22 application.tex
up

@@
 $V'$, $\Safe'$ and $\Bad'$.
-For the construction, $V'$ will have two copies of the control states of $V$; $Q_{A}$ and $Q_B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp.
+For the construction, $V'$ will have two copies of the control states of $V$; $Q_{A}$ and $Q_{B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp.
 $\Bad$.

### 2023-05-22 application.tex
up

@@
 \end{proposition}\alain{quelle complexité ?}
+\mathieu{Alors.
+On doit calculer $c_{max}$ ça ça se fait facilement c'est temps poly en la taille de l'entrée.
+Ensuite, on regarde la base de $\Bad$ et on regarde quels sont les états $q$ n'ayant pas de représentant dedans, ça se fait en la taille de la base de $\Bad$.
+Ensuite, il faut calculer la base de
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$. Ça c'est plus gourmant en temps.
+La complexité exacte de ce calcul est de ... 
+Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$. Aussi. 
+La complexité exacte de ce calcul est de ... 
+Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre 
+$\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$. }
 
@@
  $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed supsets of $\Safe$.
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
 Because these subsets are upward-closed, we can compute a basis of

### 2023-05-22 application.tex
up

@@
 $\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$. Ça c'est plus gourmant en temps.
-La complexité exacte de ce calcul est de ... 
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Ça se fait en appliquant $k$ fois l'algorithme de pred-basis.
 Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$. Aussi. 
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps. 
 La complexité exacte de ce calcul est de ... 

### 2023-05-23 application.tex
up

@@
 $\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$. }
+\mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Mais le set n'est pas nécessairement borné.
+Mais si il est ``trop grand'' c'est ``pas grave'' parce qu'on vas rapidement tomber sur un contre exemple, en fait, dont la composante vectorielle serait plus grande que $\textbf{u}_k$.
+En fait, ce $\textbf{u}_k$ nous donne une borne sur la taille du search-space dans lequel il est nécessaire de chercher un contre exemple.
+$c_{\max}$ et les vecteurs qu'on additionne pour faire $v_{\Bad}$ sont représentés par $n$ bits où $n$ et la taille de l'entrée et donc bornés par $2^n$ ...
+et globalement oui, du coup on a un search space exponentiel en la taille de l'entrée et on est en $\sf{EXPSPACE}$}
 

### 2023-05-23 application.tex
sections

@@
 
-Each of these 4 questions is decidable: $m$ is $\hat{T}$-blocked iff the preconditions of each transition in $\hat{T}$ is not coverable from $m$.
+These 4 questions are decidable: $m$ is $\hat{T}$-blocked iff the preconditions of each transition in $\hat{T}$ is not coverable from $m$.
 $m$ is dead iff $V$ with $m$ as initial configuration terminates. $m$ is bounded iff $V$ with $m$ as initial configuration is bounded.
-The first two problems reduce to coverability which is EXPSPACE.
-The third problem is also known in EXSPACE.
-The fourth problem can be decided by analysing circuits in the coverability graph \cite{DBLP:journals/acta/ValkJ85} but it can be shown in EXPSPACE.
+Hence, the first two problems reduce to coverability which is EXPSPACE.
+The third problem is also well-known in EXPSPACE.
+The fourth problem can be decided by analysing circuits in the coverability graph \cite{DBLP:journals/acta/ValkJ85} but it can be also shown in EXPSPACE following techniques exposed in \cite{DBLP:journals/jcss/Demri13}.
 
@@
 
-\cite{DBLP:journals/acta/ValkJ85} showed from a VASS $V$ and one of the 4 properties P, how to construct the maximal VASS $V'$ that has the property P which are exactly the maximal $X$-resilient VASS.
+Let a VASS $V$ and any of the 4 previous properties denoted $X$.
+Valk and Jantzen showed that the maximal subsystem of $V$ that has the property $X$ is a VASS $V_X$ that can be constructed with the Ackermanian coverability graph reaching a enormous complexity  \cite{DBLP:journals/acta/ValkJ85}.
+Let us remark that these VASS $V_X$ are the maximal $X$-resilient VASS of $V$.
+Morever, the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$.
 

### 2023-05-24 application.tex
up

@@
 \begin{proposition}
-{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
-\end{proposition}\alain{quelle complexité ?}
-\mathieu{Alors.
+{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is in $\EXPSPACE$.
+\end{proposition}
+% \alain{quelle complexité ?}
+% \mathieu{Alors.
 On doit calculer $c_{max}$ ça ça se fait facilement c'est temps poly en la taille de l'entrée.
 Ensuite, on regarde la base de $\Bad$ et on regarde quels sont les états $q$ n'ayant pas de représentant dedans, ça se fait en la taille de la base de $\Bad$.
-Ensuite, il faut calculer la base de
-$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Ensuite, il faut calculer la base de $\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
 Ça se fait en appliquant $k$ fois l'algorithme de pred-basis.
-Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps. 
-La complexité exacte de ce calcul est de ... 
-Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre 
-$\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$. }
-\mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps, mais on notera plus bas qu'il n'est pas nécessaire de tout calculer.
+Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre  $\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$.
+}
+% \mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
 Mais le set n'est pas nécessairement borné.
@@
 \begin{proof}
-Let a given $d$-VASS $V$ and consider a given $k \in \N$.
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
 Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
@@
 We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have checked for every element.
+If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
 If it is infinite then
@@
 $\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus lead to the $\EXPSPACE$ upper bound.
 \end{proof}

### 2023-05-24 application.tex
up

@@
 \begin{proof}
-\alain{faire la preuve pour les VASS.} From \cite{DBLP:journals/corr/abs-2207-02697}...
-\mathieu{En train de penser que la meilleure stratégie ici est de réduire "w.l.o.g" au cas où ils sont complémentaires encore faut-il montrer que ce serait without loss of generality.
-}
 When asking about resilience, in context, there are no true ``neutral'' configurations.
@@
 $\Safe'$ is $Q_A \times \N^d$ and $\Bad'$ is $Q_B \times N^d$.
-\mathieu{...}
+
 Without loss of generality hence let us consider only the case when $\Safe$ is semilinear
@@
 $ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus lead to the $\EXPSPACE$ upper bound.
+thus leads to the $\EXPSPACE$ upper bound.
 \end{proof}
@@
 \begin{proposition}
-{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable. \alain{quelle complexité ?}
+{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is in $\EXPSPACE$.
 \end{proposition}

### 2023-05-24 application.tex
up

@@
 \begin{proof}
-When asking about resilience, in context, there are no true ``neutral'' configurations.
-Hence from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
+When asking about {\sc Resilience}, in context, there are no true ``neutral'' configurations.
+Indeed, from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
 $V$, $\Safe$ and $\Bad$ if and only if {\sc Resilience} holds for

### 2023-05-24 application.tex
up

@@
 \begin{proposition}
-{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is in $\EXPSPACE$.
+{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
 \end{proposition}
+\mathieu{Besoin de vérifier la complexité détaillée}
 

### 2023-05-25 application.tex
up

@@
 Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
-$\Bad$ and $\Safe$ are both semilinear \cite{DBLP:journals/corr/abs-2207-02697}, 
+$\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
 {\sc Resilience} is decidable.
@@
 %		Recall that reachability is NP-complete for integer VASS. 
+\mathieu{Comparé à la preuve précédente, il faut vérifier que le Home space problem est décidable dans les $\mathbb{Z}$-VASS}
+\mathieu{Proposition 2, Proposition 3 dans [9] valent aussi bien pour les VASS que les 
+ $\mathbb{Z}$-VASS à mon sens}
+ \mathieu{Pour le Lemma 4, par contre, c'est peut-être moins évident.
+DC (can decrease the token) et NDC (cannot decrease the token) sont plus difficile à calculer (par exemple, MIN(DC) n'est pas calculable).}
 \end{proof}

### 2023-05-26 application.tex
sections

--- 2023-05-24 application.tex
@@
 
-Let a VASS $V$ and any of the 4 previous properties denoted $X$.
-Valk and Jantzen showed that the maximal subsystem of $V$ that has the property $X$ is a VASS $V_X$ that can be constructed with the Ackermanian coverability graph reaching a enormous complexity  \cite{DBLP:journals/acta/ValkJ85}.
-Let us remark that these VASS $V_X$ are the maximal $X$-resilient VASS of $V$.
-Morever, the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$.
+Let $V$ be a VASS and $X$ be any of the 4 previous sets.
+Valk and Jantzen showed that the maximal subsystem of $V$ such that its reachability set $\post^*(V)$ is included in $X$ is still a VASS $V_X$ that can be constructed with the coverability graph \cite{DBLP:journals/acta/ValkJ85} whose complexity is Ackermanian.
+Let us remark that these VASS $V_X$ are the maximal $X$-resilient VASS of $V$ and the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$.
 
@@
 \begin{proof}
-When asking about {\sc Resilience}, in context, there are no true ``neutral'' configurations.
-Indeed, from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
+\alain{faire la preuve pour les VASS.} From \cite{DBLP:journals/corr/abs-2207-02697}...
+\mathieu{En train de penser que la meilleure stratégie ici est de réduire "w.l.o.g" au cas où ils sont complémentaires encore faut-il montrer que ce serait without loss of generality.
+}
+When asking about resilience, in context, there are no true ``neutral'' configurations.
+Hence from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
 $V$, $\Safe$ and $\Bad$ if and only if {\sc Resilience} holds for
@@
 $\Safe'$ is $Q_A \times \N^d$ and $\Bad'$ is $Q_B \times N^d$.
-
+\mathieu{...}
 Without loss of generality hence let us consider only the case when $\Safe$ is semilinear
@@
 \begin{proposition}
-{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is in $\EXPSPACE$.
-\end{proposition}
-% \alain{quelle complexité ?}
-% \mathieu{Alors.
+{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
+\end{proposition}\alain{quelle complexité ?}
+\mathieu{Alors.
 On doit calculer $c_{max}$ ça ça se fait facilement c'est temps poly en la taille de l'entrée.
 Ensuite, on regarde la base de $\Bad$ et on regarde quels sont les états $q$ n'ayant pas de représentant dedans, ça se fait en la taille de la base de $\Bad$.
-Ensuite, il faut calculer la base de $\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Ensuite, il faut calculer la base de
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
 Ça se fait en appliquant $k$ fois l'algorithme de pred-basis.
-Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps, mais on notera plus bas qu'il n'est pas nécessaire de tout calculer.
-Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre  $\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$.
-}
-% \mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps. 
+La complexité exacte de ce calcul est de ... 
+Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre 
+$\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$. }
+\mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
 Mais le set n'est pas nécessairement borné.
@@
 \begin{proof}
-Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Let a given $d$-VASS $V$ and consider a given $k \in \N$.
 Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
@@
 We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If the set is finite then we stop the procedure once we have checked for every element.
 If it is infinite then
@@
 $\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
-$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to the $\EXPSPACE$ upper bound.
 \end{proof}
@@
 \begin{proposition}
-{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
+{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable. \alain{quelle complexité ?}
 \end{proposition}
-\mathieu{Besoin de vérifier la complexité détaillée}
 

### 2023-05-26 application.tex
sections

--- 2023-05-25 application.tex
@@
 The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
-
+\alain{vérifier que la plus grande existe}
 Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
@@
 
-Let a VASS $V$ and any of the 4 previous properties denoted $X$.
-Valk and Jantzen showed that the maximal subsystem of $V$ that has the property $X$ is a VASS $V_X$ that can be constructed with the Ackermanian coverability graph reaching a enormous complexity  \cite{DBLP:journals/acta/ValkJ85}.
-Let us remark that these VASS $V_X$ are the maximal $X$-resilient VASS of $V$.
-Morever, the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$.
+Let $(V,v_0)$ be a VASS with an initial state and $X$ be any of the 4 previous sets.
+Valk and Jantzen showed that the maximal subsystem of $(V,v_0)$ such that its reachability set $\post^*(V,v_0)$ is included in $X$ is still a VASS $(V_X,v_0)$ that can be constructed with the coverability graph \cite{DBLP:journals/acta/ValkJ85} whose complexity is Ackermanian.
+Let us remark that the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$, the maximal $X$-resilient VASS of $V$ is also a computable VASS $V_X$ and that $V_X$ is independant of the initial state $v_0$.
 
@@
 
+We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...
+
 \alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}

### 2023-05-26 state-resilience.tex
sections

@@
 
-\begin{theorem}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
+\begin{proposition}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
 The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
-\end{theorem}
+\end{proposition}
 
@@
 Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience }.
+Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889}:
 
-\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
+\begin{theorem}\label{post srp}
 {\sc State-resilience} is decidable for 
@@
 
-\begin{proposition}\label{downward brp}
+\begin{corollary}\label{downward brp}
 {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
-\end{proposition}
+\end{corollary}

### 2023-05-26 definitions.tex
sections

@@
 
-A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of Turing machines
+{\bf On effectivity}
+A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of algorithms
 ($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
@@
 
-We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
-exists an additional Turing machine that computes $|\post(x)| \in \N \cup \{\infty\}$ 
-on input $x$, with $x \in X_i $. 
-Such a Turing machine, in combination with $M_\rightarrow$, allows computing
-$\post(x)$ whenever the latter is finite.
+A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
 
-We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
 %\alain{étrange ces deux propositions, l'une suffirait}
-
+%%%%%%%%%%%%%%%%%
 \iffalse
+%%%%%%%%%%%%%
 Proof.
@@
 Then $t' \in I$ and $s' \in \pred(I )$.
+%%%%%%%%%%%
 \fi
-%
+%%%%%%%%%%
 
@@
 
-\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-\end{definition}
+We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
 
@@
 \iffalse
+%%%%%%%%%
+
+%%%%%%%%%
+
 \alain{pas ici}

### 2023-05-26 resilience-safe-up.tex
sections

@@
 If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward-closed.
+
+We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
+exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
+on input $x$, with $x \in X_i $. 
+%Such a Turing machine, in combination with $M_\rightarrow$, allows computing
+%$\post(x)$ whenever the latter is finite.
 
@@
  the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$.
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.
+Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).
 

### 2023-05-26 state-resilience.tex
sections

@@
 
-\begin{proposition}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]\label{liveness reset}
+Adjusting Theorem~5.5 from \cite{peterson1981petri}:
+
+\begin{proposition}\label{liveness reset}
 The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.

### 2023-05-26 definitions.tex
sections

@@
 {\bf On effectivity}
-A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of algorithms
+Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
 ($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
@@
 
-A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
+
+We say that $\mathscr{S}$
+is {\em ideally effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
+to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
 
@@
 %on input $x$, where $x$ is a state of $\mathscr{S}$. 
-
-We say that $\mathscr{S}$
-is {\em ideally effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
-to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
 %

### 2023-05-26 definitions.tex
sections

@@
 
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
@@
 
-Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
 

### 2023-05-26 application.tex
up

@@
 
-\begin{proposition}
-{\sc $k$-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is in $\EXPSPACE$.
-\end{proposition}
-% \alain{quelle complexité ?}
-% \mathieu{Alors.
-On doit calculer $c_{max}$ ça ça se fait facilement c'est temps poly en la taille de l'entrée.
-Ensuite, on regarde la base de $\Bad$ et on regarde quels sont les états $q$ n'ayant pas de représentant dedans, ça se fait en la taille de la base de $\Bad$.
-Ensuite, il faut calculer la base de $\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
-Ça se fait en appliquant $k$ fois l'algorithme de pred-basis.
-Calculer "petit à petit" $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$, c'est plus gourmant en temps, mais on notera plus bas qu'il n'est pas nécessaire de tout calculer.
-Ensuite, pour chaque élément il faut vérifier si on est capable d'atteindre  $\Safe$ en au plus $k$ étape, ça se fait en $\leq$ nombre de transition puissance $k$.
-}
-% \mathieu{En fait, on cherche à determiner la taille de notre recherche dans $\Bad \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots $\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
-Mais le set n'est pas nécessairement borné.
-Mais si il est ``trop grand'' c'est ``pas grave'' parce qu'on vas rapidement tomber sur un contre exemple, en fait, dont la composante vectorielle serait plus grande que $\textbf{u}_k$.
-En fait, ce $\textbf{u}_k$ nous donne une borne sur la taille du search-space dans lequel il est nécessaire de chercher un contre exemple.
-$c_{\max}$ et les vecteurs qu'on additionne pour faire $v_{\Bad}$ sont représentés par $n$ bits où $n$ et la taille de l'entrée et donc bornés par $2^n$ ...
-et globalement oui, du coup on a un search space exponentiel en la taille de l'entrée et on est en $\sf{EXPSPACE}$}
+\begin{theorem}
+{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable, and moreover
+{\sc $k$-resilience } is in $\EXPSPACE$.
+\end{theorem}
 
@@
 thus leads to the $\EXPSPACE$ upper bound.
-\end{proof}
 
-\iffalse
-\mathieu{À priori on devrais avoir la bounded resilience aussi.
-En fait, comme on a strong monotony alors
- $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$
- converge vers
-  $\Bad \setminus (\pred^{*}(q_{j_1}(\omega)) \cup \ldots \pred^{*}(q_{j_\ell}(\omega)) )$.
-  Une fois que le set est stable alors, encore une fois, si il est infini on n'a pas la $k$-résilience
-(pour tout $k$ à partir de ce rang là, du coup)
-et si il est fini alors on peut vérifier, pour chaque élément, si il est possible d'atteindre
-$\Safe$.
-Si il est possible pour chaque élément alors on prends le maximum des chemins depuis chacun de ces éléments et on fait le max de ça et le $k$ qui permet d'atteindre $\pred^{\leq k} = \pred^*$ pour les $q(\omega)$ ci-dessus et ça nous donne un $k_{\max}$ pour lequel on a la $k_{\max}$-résilience.
-Si il est pas possible pour chaque élément alors ça veut dire il y a un élément qui atteindra jamais $\Safe$ et on n'a pas la bounded resilience ni la résilience d'ailleurs. 
-Il faut vérifier que c'est possible pour une configuration de vérifier qu'elle est capable d'atteindre $\Safe$ downward closed. 
-Normalement c'est le cas d'après Okzan (ils disent c'est équiv du submarking pb).}
-\fi
-
-\begin{proposition}
-{\sc Bounded-resilience } in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ is decidable.
-\end{proposition}
-\mathieu{Besoin de vérifier la complexité détaillée}
-
-\begin{proof}
-$\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-% $q'(\omega)$
-$\uparrow q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-Let us assume from now on 
-% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
- $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
-Because these subsets are upward-closed, we can compute a basis of
-% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
-% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
-$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
-
-We now consider the set 
-% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
-
-There exists some $k_0 \in \N$ for which
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
 and hence
@@
 
-From $k_0$ onwards ...
+From $k_0$ onwards,
 $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 

### 2023-05-26 resilience.tex
sections
up

@@
 Let us formalize three resilience problems.
+
+\alain{faire le lien avec home space, et les formules dans Schnoebelen}
 
@@
 \begin{example}
-% Avec Safe = up( q_1(3,0) ) et Bad = q_0 (N,N) on aurait la 1-résilience et donc la résilience par exemple
-% mais avec Safe = up(q_0(3,0)) et Bad = q_1(N,N) on ne l'aurait pas à cause de q_1(0,0) 
-Let $V = (\{q_0,q_1\}, \{ (q_0, (3,0), q_1), (q_1, (-1,1), q_0)\})$ be a vector addition system with states (VASS),
-and thus a WSTS (see Figure~1).
-In the case $\Safe = \uparrow q_1(3,0)$ and $\Bad = \uparrow q_0(0,0)$, $1$-resilience holds.
-On the other hand in the case $\Safe = \uparrow q_0(3,0)$ and $\Bad = \uparrow q_1(0,0)$, resilience does not hold because of $q_1(0,0) \in \Bad$.
-% (si on veut se rajouter une hypothèse comme pas de dead-end on peut rajouter un troisième état q_2 duquel on puisse partir depuis q_1 en faisant par exemple +1,+1 et qui loop sur lui-même) 
-\end{example}
-
-\begin{center}
-	\begin{figure}
-			\hspace{3.75cm}
-				\label{example_1}
-\includegraphics[width=0.2\textwidth]{FigureA}
-	\caption{}
-	\end{figure}
-\end{center}
-
-\begin{example}
-% Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
-Let us consider now a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
+Let us consider a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
 then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $b$ there is an element of $\Bad$ e.g.

### 2023-05-27 introduction.tex
sections

@@
 
-{\bf Context.} In the quite informal paper \cite{DBLP:journals/corr/PrasadZ16}, Prasad and Zuck studied resilience of a process to an adversary and they used the framework of WSTS enjoying both upward and downward monotonies to prove the decidability of a kind of resilience.
-%
-%{\bf MFCS 2023
-%Abstract submission deadline:    	 April 24th (AoE) 
-%Paper submission deadline:    	April 28th (AoE)} \\
-%%
-%{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
-%Submission Deadline 	May 2, 2023}\\
-%
-{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
-Notification to authors: September 16, 2022}
-
-trop d'hypotheses trop fortes sur les WSTS
-
-{\bf Our contributions}
-We introduce a more general definition of resilience generalising the two given in \cite{DBLP:conf/gg/Ozkan22}.
-Surprinsingly, the general undecidability statements were not known neither proved.
-
-\begin{itemize}
-
-\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
-
-\item safe clos haut : resilience indecidable for WSTS with strong.
-
-\end{itemize}
-
+{\bf Context.} 
+Resilience is a key notion for improving safety of systems.
+Resilience Engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain).
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
@@
 This is the resilience problem.
+Resilience is the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states.
 
-Resilience is a concept of growing interest [references].
-Both Okzan and Würdemann [ref] and Prasad and Zuck [ref] have started to consider [casting the problem into] the framework of Well-Structured Transition Systems.
+%
+%{\bf MFCS 2023
+%Abstract submission deadline:    	 April 24th (AoE) 
+%Paper submission deadline:    	April 28th (AoE)} \\
+%%
+%{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
+%Submission Deadline 	May 2, 2023}\\
 
-Resilience is an instance of the home space problem: let $L_1,L_2$ be two sets of states, the home space for  $L_1,L_2$ is the following question:		
-	$\post^*(L_1)	\subseteq \pred^*(L_2)$
+{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
+Notification to authors: September 16, 2022}
+
+trop d'hypotheses trop fortes sur les WSTS
+
+{\bf Our contributions}
+We introduce  .
+Surprinsingly, the general undecidability statements were not known neither proved.
+
+\begin{itemize}
+
+\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience.
+The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$.
+We remark that it is an instance of the home space problem defined in the framework of VASS in \cite{DBLP:journals/corr/abs-2207-02697}.
+State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is ...
+
+\item We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
+
+\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
+
+\item The resilience problem (RP), the bounded resilience problem (BRP)
+and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
+
+\item The resilience problem is decidable for ideal-effective WSTS with 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$
+and
+the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+
+\item Unfortunately, {\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for strongly (upward) compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$. We made a reduction of zero-reachability in reset-VASS to state-resilience in reset-VASS that are WSTS with strong upward-compatibility and with effective pred-basis.
+
+\item We generalize the main theorem of \cite{} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
+ WSTS with effective 
+$\uparrow$ $\post^*$ basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
+$\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
+But we found another effective hypothesis {\sc State-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities.
+and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility.
+
+\item VASS
+\end{itemize}
+
+{\bf Related works.} 
+In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward monotonies to give an idea of a proof of the decidability of two kinds of resilience.
+In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} modeled systems with graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+All these papers consider particular Safe and Bad: $\Bad$ is downward-closed and $\Safe$ is upward-closed.
 
@@
 
-\alain{il faut mieux expliquer le modèle, le rôle des 4 canaux, des messages, remarquer que 2 canaux sont des compteurs,....}
+%\alain{il faut mieux expliquer le modèle, le rôle des 4 canaux, des messages, remarquer que 2 canaux sont des compteurs,....}
 

### 2023-05-27 definitions.tex
up
sections

@@
 \noindent
-Let us recall thz well-known framework of transition systems. 
+%Let us recall the well-known framework of transition systems. 
  A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
@@
 % We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$.
-We use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
+We write $s \rightarrow s'$ to denote $ (s,s') \in  {\rightarrow} $.
 We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
@@
 % $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
-A state without successors is called a {\em dead end}. 
+%A state without successors is called a {\em dead end}. 
 The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
@@
 By iterating $\pred$ and $\post$ we obtain  
-$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
+$\post^n(s) = \{ s' \in S \mid  ~ s \xrightarrow{}^n s'\}$
 and
@@
 $\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. % \\
+\alain{supprimer ce qui est inutile}
 
 The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$, whether $s \to^* t$. It can be reformulated as asking whether 
-$s \in \pred^*(t)$ or asking whether $t \in \post^*(s)$. \\
+$s \in \pred^*(t)$ or asking whether $t \in \post^*(s)$. 
+\alain{où est-ce utilisé ?}
 
@@
 \noindent
-A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. We abbreviate $x \leq y \not\leq x$ by $x < y$.
+A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
+%\alain{utilisé ? We abbreviate $x \leq y \not\leq x$ by $x < y$.}
 % A partial ordering (a po) is an antisymmetric qo
@@
 Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
-	% }
+ \alain{utilisé ?}
+%
 % and gives rise to a po between the equivalence classes ?
@@
 \noindent
-Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo.
-In such \emph{Finite Antichains} sets $(X,\leq)$, every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward-closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
+Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward-closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
 
@@
 \noindent
-Let us now recall the most general definition of well-structured transition systems.
+Let us now recall the (most general) definition of well-structured transition systems.
 
@@
 
+\noindent
 {\bf On effectivity}
@@
 $M_\leq$ the ordering relation “$\leq$”.
-
+We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
+exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
+on input $x$, with $x \in X_i $. 
 We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-
+%
 We say that $\mathscr{S}$
@@
 \begin{theorem}\cite{BFM-ic17}
-Coverability is decidable for any ideally effective ordered compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
+Coverability is decidable for any ideally effective ordered upward-compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
 \end{theorem}

### 2023-05-27 introduction.tex
sections

@@
 {\bf Context.} 
-Resilience is a key notion for improving safety of systems.
-Resilience Engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain).
+Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain).
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
@@
 {\bf Our contributions}
-We introduce  .
-Surprinsingly, the general undecidability statements were not known neither proved.
 
@@
 
-\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience.
+\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
+and the $k$-(state-)resilience problem (kRP).
 The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$.
@@
 
-\item We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
+\item Surprinsingly, the general undecidability statements were not known neither proved. We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
 
-\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
+%\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
 
-\item The resilience problem (RP), the bounded resilience problem (BRP)
+\item We find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
 and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
@@
 {\sc $k$-state-resilience}
-are undecidable for strongly (upward) compatible WSTS with effective pred-basis
+are undecidable for strongly upward-compatible WSTS with effective pred-basis
 when
@@
 
-\item VASS
+\item We study the resilience for VASS and most of problems are shown decidable.
 \end{itemize}
@@
 {\bf Related works.} 
-In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward monotonies to give an idea of a proof of the decidability of two kinds of resilience.
-In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} modeled systems with graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-All these papers consider particular Safe and Bad: $\Bad$ is downward-closed and $\Safe$ is upward-closed.
+In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience.
+Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+All these papers consider particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed.
 

### 2023-05-27 resilience-safe-up.tex
sections

@@
 If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward-closed.
+\alain{ce qui précède est faux et pas convaincant: un vass qui seulement décroit ses compteurs}
 
-We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
-exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
-on input $x$, with $x \in X_i $. 
 %Such a Turing machine, in combination with $M_\rightarrow$, allows computing
@@
 from which it is always possible to fire a transition.
-This set of states is upward-closed and $\Safe=S-\Bad=\uparrow \Safe$.
+This set of states is upward-closed and $\Safe=S-\Bad=\uparrow \Safe$.\alain{mal rédigé}
 
@@
 
+\alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem (RP), the bounded resilience problem (BRP)
+and the $k$-resilience problem (kRP) are undecidable for WSTS with strong compatibility}
 %

### 2023-05-27 resilience.tex
sections

@@
 
+\begin{example}
+Let us consider a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
+then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g.
+$q(-k-1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$.
+\end{example}
+
 % \alain{dessous: à dire mais pas sous forme de proposition}
@@
 \end{remark}
+\alain{parfois $\mathscr{S}=(S,\rightarrow,\leq)$ is
+(\Bad,\Safe)-resilient parfois $\mathscr{S}=(S,\rightarrow,\leq)$ is 
+(\Safe,\Bad)-resilient , rendre cohérent avec $\mathscr{S}=(S,\rightarrow,\leq)$ is
+(\Bad,\Safe)-resilient}
 
@@
 Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
 
@@
 
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
 %		
-We deduce 
+We \alain{deduce} 
 % \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
@@
 %
-
+\alain{il manque un résultat d'indecidabilité pour les 3 resilience pour les WSTS+...}
 %	 and ideals are recursive
@@
 \fi
-
-\begin{example}
-Let us consider a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
-then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $b$ there is an element of $\Bad$ e.g.
-$q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$.
-\end{example}
 

### 2023-05-29 application.tex
up

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
-\alain{ me semble maintenant inutile
-%
-\begin{definition}{ (index)}. 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
-The {\em index} $k(U)$ is the
-smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
-\end{definition}
-%
-%
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
-every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
-%
-\begin{remark}
-$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
-U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
-U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
-=  U \cup \pred(U^k)$.
-\end{remark}
-
-This ensures the following.
-
-\begin{fact}\label{stop condition}
-% Fact 4 (stop condition). 
-If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
-$k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
-\end{fact}
-
-\begin{lemma}
-% Lemma 3 ([1]) 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
- Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly WSTS, we can decide whether $U  \xrightarrow {*}{} s$.
-%	we can reach $U$ from $s$.
-\end{lemma}
-
-\begin{proof}
-We have to show that we can compute a basis of $U^{k+1}$ if we are given a basis of $U^k $. 
-Then the
-decidability of the stop condition follows directly. Let $B$ be a basis of $U^k$. 
-We have
-$$U^{k+1} = U \cup \pred(U^k ) = U \cup
-\bigcup_{s' \in B}
-\pred(\uparrow \{s' \}).$$
-
-Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
-Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
-\end{proof}}
-
-\fi
-%%%%%%%%%%%%%%%%%%%
-
-\section{Deciding resilience for extended automata}
-
-\subsection{Pushdown automata}
-
-%We may consider context-free grammars.
-
-\begin{theorem}
-The three {\sc resilience} problems are decidable for pushdown automata with $\Safe$ and $\Bad$ regular languages.
-\end{theorem}
-
-\begin{proof}
-Recall that resilience is equivalent to  $\Bad \subseteq \pred^*(\Safe)$.
-For every pushdown automaton $A=(Q,\Sigma,...)$ and every regular language $L \subseteq Q \times \Sigma^*$, the set $\pred^*(L)$ is a computable regular language \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}.
-Since $\Bad$ and $\Safe$ are both regular, we deduce that
-%
-%			Moreover, $post^*(\Bad)$ is also a computable regular language \alain{pourquoi ?}   $post^*(L)$ and
- the inclusion $\Bad \subseteq \pred^*(\Safe)$ is decidable and resilience too.
-%
-%The kRP is decidable: % let us compute $ \pred^{\leq k}(\Safe)$ (in polynomial time). 
-Similarly, since $ \pred^{\leq k}(\Safe)$ is a computable regular language,
-we deduce 
-that the inclusion $\Bad \subseteq \pred^{\leq k}(\Safe)$ is decidable and $k$-resilience too.
-Since $k$-resilience is decidable, bounded-resilience also is;
-we can iteratively check whether $\Bad \subseteq \pred^{\leq k}(\Safe)$,
-if this is the case, $k$-resilience holds, and we stop with a positive,
-else we check whether $ \pred^{\leq k}(\Safe) =  \pred^{*}(\Safe)$,
-if this is the case, then bounded-resilience do not hold,
-else we continue.
-\end{proof}
-%
-Let us remark that $\uparrow \post^*(q,w)$ is also a computable regular language, but we cannot directly apply previous results on WSTS since pushdown automata are not WSTS for the usual orderings (subword, prefix).
-Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow \post^*$ is computable but context-free grammars are not strongly compatible.
-
-%  \textcolor{red}{à faire Mathieu}
-
-%
-%
-
-\subsection{Regular fifo automata}
-
-Fifo automata and systems of communicating finite-state machines (CFSMs) are essentially finite automata that communicate through fifo channels (or queues).
-For the sake of simplicity, we consider automata with an unique fifo channel (that have the power of turing machines).
-In 1982, Pachl introduced in \cite{DBLP:journals/corr/cs-LO-0306121} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability (with a regular inductive invariant to obtain a witness of non-reachability).
-Let $(q,w)$ be an initial state and $(q',w')$ be a state.
-When $\post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $\post(L) \subseteq L$ (inductive invariant).
-We know that if $(q',w') \not\in \post^*(q,w)$ then there exists, at least, a regular language $L$ such that $\post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $\post^*(q,w)$ is such inductive invariant.
-This strategy provides a semi-algorithm for non-reachability.
-As reachability is recursively enumerable, this provides an algorithm to solve reachability for regular fifo automata.
-
-Let us say that a fifo automaton is \emph{post-regular} (resp.
-\emph{pred-regular}) if $\post^*(q,w)$ (resp.
-$\pred^*(q,w)$ ) is regular for all $(q,w) \in Q \times \Sigma^*$.
-The  fifo automaton is \emph{effectively} post-regular (resp.
-pred-regular) if the regular language $\post^*(q,w)$ (resp.
-$\pred^*(q,w)$) is computable.
-Lossy channel systems are post-regular (since $\post^*(q,w)$ is downward-closed) but they are not effectively post-regular; they are also effectively pred-(upward-closed) regular.
-
-Reachability is (immediately) decidable for both effective post-regular or effective pred-regular fifo automata.
-%\alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ?
-contre-exemples ?
-c'est le cas pour les LCM où pred ok mais pas post}
-%
-%
-%Let us say that a fifo automaton is \emph{regular} if the reachability relation is regular for all $(q,w)$; in this case, the fifo automaton is also both post-regular and pred-regular.
-%
-Now, we just observe that $\Bad \subseteq \pred^*(\Safe)$ is decidable when $\pred^*(\Safe)$ is a computable regular language and $\Bad$ a regular language:
-
-\begin{theorem}{}
-Resilience is decidable for effective pred-regular fifo automata when $\Safe$ and $\Bad$ are regular languages.
-\end{theorem}
-
-%		\begin{proof}
-%		$\Bad \subseteq \pred^*(\Safe)$ is decidable because $\pred^*(\Safe)$ is a computable regular language.
-%		\alain{prover que $\pred^*(\Safe)$ is a computable regular language}
-%		\end{proof}
-%
-%			Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.
-
-Since for lossy channel systems $(Q, \Sigma,d,...)$ with $d$ fifo channels and a set of states $L \subseteq Q \times (\Sigma^*)^d$, the set $\pred^*(\uparrow L)$ is a computable upward-closed regular language, we deduce :
-
-\begin{corollary}
-Resilience is decidable for lossy channel systems with $\Safe=\uparrow \Safe$ and any regular language $\Bad$.
-\end{corollary}
-
-Another proof is possible for LCS since LCS are WSTS with a computable $\uparrow \post^*(S)$ \alain{à prouver et vérifier les autres hypothèses}
-%
-%%%
-\subsection{Resilience for VASS}
+\section{Resilience for VASS}
 

### 2023-05-29 definitions.tex
up

@@
 \alain{où est-ce utilisé ?}
-
-	\iffalse
-\problemx{Reachability}
-{A transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$}
-{$s \to^* t$? \\}
-	\fi
 
@@
 
-%We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
-%and if there exists an additional Turing machine that computes a finite basis of 
-%$\uparrow \pred (\uparrow x)$
-%on input $x$, where $x$ is a state of $\mathscr{S}$. 
-%
-\iffalse
-%%%%%%%%%
-
-%%%%%%%%%
-
-\alain{pas ici}
-\begin{definition}{\cite{BFM-ic17}} 
-A class $C$ of WSTS is {\em ideally effective} if
-\begin{itemize}
-\item the function mapping the encoding of a configuration $s$ of a WSTS
-to the encoding of the ideal $\downarrow s$ is computable;
-\item inclusion of ideals is decidable;
-\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-\end{itemize}
-\end{definition}
-%
-\fi
-%
-%
 With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
@@
 \end{theorem}
-
-%         A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
-%         $s \in D$.
-
-%\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
-% \textcolor{red}{Do we still demand this ?}
-%\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}
-
-%
-\iffalse
-%
-\begin{claim}{(stability of upward-closed sets)}
-Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
-\end{claim}
-%
-\fi
-%
-
-% \begin{claim}
-% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \end{claim}
-
-%			\begin{fact}\label{fact basis}
-% Fact 3 ([1]). 
-% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-
-%For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$.
-
-% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
-% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
-%			\end{fact}
-%

### 2023-05-29 state-resilience.tex
up

@@
 % To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. 
-The following theorem is a adjustment of Theorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.
-
-% \alain{tu ne définis pas, même informellement et/ou rapidement, assez les objets que tu utilises: zero reachability problem, Reset VASS}
-
-Adjusting Theorem~5.5 from \cite{peterson1981petri}:
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS, whose proof can be seen in Appendix~\ref{appendix}.
 
@@
 
-\iffalse
-% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
-\begin{proof}
-See Appendix~\ref{appendix} for the complete proof.
-\end{proof}
-\fi
-
 Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
@@
 %
-\iffalse
-\begin{corollary}
-Liveness is undecidable in Reset VASS.
-\end{corollary}
-
-Since moreover liveness can be reduced to 
-$t$-liveness, we deduce:
-\fi
 %
-%\begin{corollary}
 {\sc %$ Reset $d$-VASS 
 $t$-liveness} is undecidable.
-%\end{corollary}
 

### 2023-05-29 application.tex
up

@@
 The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
-\alain{à faire}
-%		Recall that reachability is NP-complete for integer VASS. 
-\mathieu{Comparé à la preuve précédente, il faut vérifier que le Home space problem est décidable dans les $\mathbb{Z}$-VASS}
-\mathieu{Proposition 2, Proposition 3 dans [9] valent aussi bien pour les VASS que les 
- $\mathbb{Z}$-VASS à mon sens}
- \mathieu{Pour le Lemma 4, par contre, c'est peut-être moins évident.
-DC (can decrease the token) et NDC (cannot decrease the token) sont plus difficile à calculer (par exemple, MIN(DC) n'est pas calculable).}
+% For Z-VAS we observe that we are asking for inclusion between linear sets. 
+% Again let us without loss of generality hence let us consider only the case when $\Safe$ is semilinear and $\Bad = S \setminus \Safe$ is semilinear too.
+\mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
 \end{proof}

### 2023-05-30 application.tex
up

@@
 \begin{theorem}
-{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable, and moreover
-{\sc $k$-resilience } is in $\EXPSPACE$.
+{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
+.
 \end{theorem}
@@
 $ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to the $\EXPSPACE$ upper bound.
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience }.
 

### 2023-05-30 application.tex
up

@@
 Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
-$$\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).$$
+\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).\]
 Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
@@
 The {\em size} of $\A$ is defined as%:
-$$
+\[
 |\A| \ = \ |Q|+|\Omega|+|R|+\sum_{(q,g,U,q')\in R}|g|.
-$$
+\]
 Let 
@@
 A {\em $0/1$ timed automaton } ({\em $0/1$-TA} for short) is a tuple
-$$\B=(Q,X, \Delta_0, \Delta_1, q_{init}, F),$$
+\[
+\B=(Q,X, \Delta_0, \Delta_1, q_{init}, F),
+\]
 \par\noindent\ignorespacesafterend

### 2023-06-02 resilience.tex
up

@@
 
-% \alain{dessous: à dire mais pas sous forme de proposition}
 \begin{remark}[Reformulation]\label{reformulation}
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
+(\Safe, \Bad)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
 $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
+(\Safe, \Bad)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
 \end{remark}
-\alain{parfois $\mathscr{S}=(S,\rightarrow,\leq)$ is
-(\Bad,\Safe)-resilient parfois $\mathscr{S}=(S,\rightarrow,\leq)$ is 
-(\Safe,\Bad)-resilient , rendre cohérent avec $\mathscr{S}=(S,\rightarrow,\leq)$ is
-(\Bad,\Safe)-resilient}
 

### 2023-06-02 state-resilience.tex
up
Lipics 5678...5678  changed to \[ ... \] for conformity
Replaced remaining 'post^*' with '\post^*' for consistency

@@
 Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$.
-But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap post^*(s_0)$, there must exist a path to $\Safe$.
+However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
 %
@@
 %following
- undecidable [ref again] decision problem%:
+ undecidable \cite{araki1976PN} decision problem%:
 \iffalse
@@
 as the following formula
-% $$ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'$$ 
-$$ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})$$  
+% \[ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'\] 
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
 where
@@
 Then, because {\sc $n_0$-state-resilience} is decidable, 
-checking $\uparrow post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+checking $\uparrow \post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
 and {\sc State-resilience} is decidable.
@@
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \uparrow post^{\leq n}(s)$ until it becomes stationnary. 
+$U_n = \uparrow \post^{\leq n}(s)$ until it becomes stationnary. 
 Decidability of downward-reachability leads to the decidability of the following stop condition:
@@
 % sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-%	$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
+%	$\uparrow \post^*$ n'est pas calculable pour les LCS non plus car $\uparrow \post^*= \uparrow \downarrow \post^*$ n'est pas calculable ???
 \fi
@@
 
-%il existe des modèles où on peut calculer $\uparrow post^*$: inserted channel systems: on sait calculer $post^*$ qui est rationnel donc on sait calculer $\uparrow post^*$.
+%il existe des modèles où on peut calculer $\uparrow \post^*$: inserted channel systems: on sait calculer $\post^*$ qui est rationnel donc on sait calculer $\uparrow \post^*$.
 
 %
-%si on garde strong mais on enlève effective basis of $post^*$:
+%si on garde strong mais on enlève effective basis of $\post^*$:
 %
@@
 %\begin{proof}
-%pas de effective basis of $post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
+%pas de effective basis of $\post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
 %\end{proof}
@@
 
-$$ \post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} \cap \Bad \subseteq \pred^*(\Safe)$$
+\[ \post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} \cap \Bad \subseteq \pred^*(\Safe)\]
 % \alain{$\equiv $  ???}
@@
 \hline  Hypothesis & strong %upward 
-			compatibility ~ & $\uparrow post^*$ effective 
+			compatibility ~ & $\uparrow \post^*$ effective 
 			& strong %upward 
-					compatibility \textit{+} $\uparrow post^*$ effective \\ \hline
+					compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline

### 2023-06-05 definitions.tex
sections
up

@@
 
-The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$, whether $s \to^* t$. It can be reformulated as asking whether 
-$s \in \pred^*(t)$ or asking whether $t \in \post^*(s)$. 
+The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$, whether $s \to^* t$.
+% It is equivalent to $s \in \pred^*(t)$ ?
+or $t \in \post^*(s)$ ?
 \alain{où est-ce utilisé ?}
@@
 	% \alain{
-Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
- \alain{utilisé ?}
+% Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
+% \alain{utilisé ?}
 %
@@
 We say that $\mathscr{S}$
-is {\em ideally effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
+is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
@@
 
-With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
+With the ideal-effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
 
 \begin{theorem}\cite{BFM-ic17}
-Coverability is decidable for any ideally effective ordered upward-compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
+Coverability is decidable for any ideal-effective ordered upward-compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
 \end{theorem}

### 2023-06-05 introduction.tex
up

@@
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-But we found another effective hypothesis {\sc State-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities.
-and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility.
+But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities.
+and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
 

### 2023-06-05 state-resilience.tex
up

@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideal effective WSTS with downward and upward compatibilities,
+{\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities,
 %\alain{a-t-on aussi la upward compatibility ?}
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility,
 $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.

### 2023-06-05 resilience-safe-up.tex
up

@@
 %
+
+Without the completion-post-effective hypothesis or the pred-basis hypothesis, coverability 
+becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}.
+
+Coverability reduce
+to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$.
+Indeed $s \to^* \uparrow t$ 
+reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.

### 2023-06-07 resilience.tex
up

@@
 In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
-%	If $\Safe = S \backslash \Bad$, then $\Safe$ is downward-closed iff $\Bad$ is upward-closed. 
-Let us formalize three resilience problems.
 
 \alain{faire le lien avec home space, et les formules dans Schnoebelen}
+
+Linked problems
+
+A set of configurations $H$ is an {\em home-space} for a set of configurations $X$ if every configuration reachable from $X$ can reach $H$.
+Resilience as introduced above essentially asks whether $\Safe$ is a home-space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
+The home-space problem is decidable for
+Petri nets when $X$ and $H$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
+To these author's knowledge, there are no general results for home-space in WSTS.
+
+\mathieu{vérifier au passage que c'est vrai qu'il y a pas de résultats home-space WSTS}
+
+Resilience and the home-space problem are also linked to the 
+model-checking of the ``from-all'' formulae $\forall s \in X~ \exists t \in Y~ s \to^* t$
+which has been shown decidable for Lossy Counter Machine
+when $X$ and $Y$ are semi-linear sets~\cite{DBLP:conf/rp/Schnoebelen10}.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for Lossy Counter Machines).
+
+% The Model-checking problem for LTL, CTL*, etc. is a natural generalization of the reachability problem. This problem is undecidable in general but there are simple formulae decidable
+
+Let us formalize three resilience problems.
 

### 2023-06-07 resilience-safe-up.tex
up

@@
 
-\alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are undecidable for WSTS with strong compatibility}
+% \alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem (RP), the bounded resilience problem (BRP) and the $k$-resilience problem (kRP) are undecidable for WSTS with strong compatibility}
 %
@@
 reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
+Hence, without the completion-post-effective hypothesis, 
+{\sc Resilience} is undecidable.

### 2023-06-07 application.tex
up

@@
 
-\mathieu{Tableau résumé de la situation pour les VASS}
+\begin{figure}
 
@@
 \end{center}
+	\caption{Decidability results for VASS.}
+\end{figure}
 
@@
 $ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience }.
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
 
@@
 
-%
-\iffalse
-%
-\subsection{Timed Automata}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
-
-\renewcommand{\A}{\mathcal{A}}
-\newcommand{\B}{\mathcal{B}}
-\renewcommand{\C}{\mathcal{C}}
-\newcommand{\Const}{\mathsf{Consts}}
-\newcommand{\Conf}{\mathsf{Conf}}
-\newcommand{\guards}{{\textsc{Guards}}}
-
-% \subsubsection{Guards, Clocks}
-
-A {\em guard} over a finite set of clocks $\Omega$ 
-is a comparison of the form
-$\omega \bowtie c$, where $ \omega \in \Omega$, $c \in \N$,
-and $\bowtie\in\{<,\leq,=,\geq,>\}$.
-%
-We denote by $\guards(\Omega)$ the {\em set of guards} over the set of 
-clocks $\Omega$.
-The {\em size} %$|g|$
- of a guard 
-$g=\omega \bowtie c$ is defined as %:
-$|g|=\log(c)$.
-A {\em clock valuation} is a function from $\Omega$ to $\N$;
-we write $\vec{0}$ to denote the clock valuation $\omega \mapsto 0$
-whenever the set $\Omega$ is clear from the context.
-For each clock valuation $v$ and each $t\in\N$ we denote
-by $v+t$ the clock valuation $\omega \mapsto v(\omega)+t$.
-%
-For each guard $g=\omega \bowtie c$ with $c\in\N$,
-we write $v\models g$ if $v(\omega)\bowtie c$.
-
-\iffalse
-We define an {\em empty guard} $g_\epsilon$ over a non-empty finite set of clocks
-$\Omega$ and to be of the form $\omega \geq 0$ for some 
-$\omega \in \Omega$. In particular, we
-defined $g_\epsilon$ such that for all $v \in \N^\Omega$ 
-we have
-$v \models g_\epsilon$, hence $g_\epsilon$ can be used as a guard that is always true. 
-\fi
-
-% \subsubsection{Timed automata}
-
-A timed automaton is a finite automaton extended with a finite set of clocks $\Omega$ that all progress at the same rate and that can individually be reset to zero.
-Moreover, every transition is labeled by a guard over
-$\Omega$  and by a set of clocks to be reset. \\
-
-\par\noindent\ignorespacesafterend
-Formally, a {\em timed automaton} ({\em TA} for short) is a tuple
-$\A=(Q,\Omega,R,q_{init},F)$, where
-\begin{samepage}
-\begin{itemize}
-	\item $Q$ is a non-empty finite {\em set of states}, 
-	\item $\Omega$ is a non-empty finite {\em set of clocks},
-	\item $R \subseteq Q\times\G(\Omega)\times \mathscr{P}(\Omega) \times Q$
-	is a finite {\em set of  rules},
-	\item $q_{init}\in Q$ is an {\em initial  state}, and 
-	\item $F\subseteq Q$ is a {\em set of final states}.
-\end{itemize}
-\end{samepage}
-
-\par\noindent\ignorespacesafterend
-We also refer to $\A$ as an $n$-TA if $|\Omega| = n$. 
-The {\em size} of $\A$ is defined as%:
-\[
-|\A| \ = \ |Q|+|\Omega|+|R|+\sum_{(q,g,U,q')\in R}|g|.
-\]
-Let 
-$\Const(\A) = \{ c\in\N \mid \exists(q,g,U,q')\in R, \ \exists \omega \in \Omega, \bowtie\in\{<,\leq,=,\geq,>\} : g = \omega \bowtie c \}$ denote the 
-set of constants that appear in the guards of the rules of $\A$.
-
-By $\Conf(\A)=Q\times\N^\Omega$ we denote the set of
-{\em configurations} of $\A$. 
-%We prefer however to denote a configuration in $\Const(\A)$ by $q(v)$ instead of $(q,v)$.\\
-We prefer however to abbreviate a configuration	%	 in $\Conf(\A)$	
-$(q,v)$ by $q(v)$.
-
-\begin{samepage}
-%\begin{definition}
-A TA $\A=(Q,\Omega,R,q_{init},F)$ induces the labeled transition system 
-$T_{\A} =  (\Conf(\A), \Lambda_{\A}, \rightarrow_{\A})$
-where $ \Lambda_{\A} = R \times \N $
-and 
-where $ \rightarrow_{\A}$ is defined such that, 
-for all $(\delta,t)\in R\times\N$ with  	$\delta = (q,g,U,q')\in R$,
-for all $q(v), q'(v') \in \Conf(\A)$,
-$q(v)\xrightarrow{\delta,t}_{\A} q'(v')$ if
-	$v+t\models g$, 
-	 $v'(u)=0$ for all $u \in U$ and $v'(\omega)=v(\omega)+t$ for all 
-	$\omega \in \Omega \setminus U$.
-%\end{definition}
-\end{samepage}
-
-A {\em run} from $q_0(v_0)$ to $q_n(v_n)$ in $\A$ is a path in the transition system $T_{\A}$, that is,
-a sequence 
-$\pi = q_0(v_0)\xrightarrow{\delta_1,t_1}_{\A}q_1(v_1)\cdots\xrightarrow{\delta_n,t_n}_{\A}q_n(v_n)$;
-it is called {\em reset-free} if for all $i \in \{1,\ldots,n\}$,
- $\delta_i = (g_i,\emptyset)$ for some guard $g_i$.
-
-We say $\pi$ is {\em accepting} if $q_0(v_0) = q_{init}(\vec{0})$ and $q_n \in F$. 
-\iffalse
-We say {\em reachability holds} for the TA $\A$ 
-if there exists an accepting run. %
-% if there is a run in $\A$ from $q_{init}(\vec{0})$   to some configuration $q(v)$ for some $q\in F$, and $v\in\N^\Omega$.
-%We refer to Figure~\ref{example pta} for an instance of a PTA for which reachability holds.
-\fi
-
-It is worth mentioning that there are further modes of time valuations and guards which exist in the literature, we refer
-to \cite{Andre19} for a recent overview. 
-%
-% \mh{Comment on difference between continuous and discrete time}
-Notably, we consider in this article only the case of timed automata over discrete time. It is worth mentioning that in
-the case of timed automata over continuous time (i.e. with clocks having values in $\R_{\geq 0}$),
-% However, for parametric timed automata with closed (i.e., non-strict) clock constraints and parameters restricted to ranging over integers, 1 standard digitisation techniques apply [HMP92, OW03], reducing the reachability problem over dense time to discrete (integer) time.
-techniques~\cite{HenzingerMP92,OuaknineW03} exist for reducing the reachability problem to discrete time in the case of closed (i.e. non-strict) clock constraints ranging over integers. \\
-
-\problemx{TA $k$-resilience problem}
-{A state $q$ of a TA $(Q, X, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
-{$\forall q' \in BAD \forall v,v' \in \N^X ~ (q(v) \rightarrow^* q'(v')) \implies \exists q'' \in SAFE \exists v'' \in \N^X ~ q'(v') \rightarrow^{\leq k} q''(v'')$ ?\newline}
-
-Analogously, we formulate the bounded resilience problem for WSTSs.
-
-\problemx{TA bounded resilience problem}
-{A state $q$ of a TA $(Q, X, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
-{$\exists k \geq 0 ~ \forall q' \in BAD \forall v,v' \in \N^X ~ (q(v) \rightarrow^* q'(v')) \implies \exists q'' \in SAFE \exists v'' \in \N^X ~ q'(v') \rightarrow^{\leq k} q''(v'')$ ?\newline}
-
-\textcolor{red}{I think there can be a discussion to be had here about how to quantify on the clock valuations}
-
-\textcolor{red}{Here one thing that could be interesting to try to formalize is: how to enforce that the time that passes is less than $k$, rather than the number of transitions.
-This is tricky to deal with I find but it should be more doable if for instance we use one counter automata, where the counter effect of the sequence can be quantified more explicitly I suppose ?
-But here you could also use a kinda special clock $x$ that is reset when you enter $BAD$ and is not reset between a state in $BAD$ and a state in $SAFE$, you could check that $x < k$.}
-
-\textcolor{red}{...
-I guess if you use $0/1$-TA then the problems become closer one to another ?
-Also of note is that $0/1$-TA induces transition systems with bounded branching, so I guess it may be interesting to investigate these first ?}
-
-A {\em $0/1$ timed automaton } ({\em $0/1$-TA} for short) is a tuple
-\[
-\B=(Q,X, \Delta_0, \Delta_1, q_{init}, F),
-\]
-\par\noindent\ignorespacesafterend
- where
-$\B_i=(Q,X, R_i, q_{init}, F)$ is a TA for all $i \in \{0,1\}$.
-For simplicity we define its {\em size}
-as $|\B|=|\B_0|+|\B_1|$.
-We analogously denote the constants of $\B$ 
-by $\Const(\B)$ and its configurations by  $\Conf(\B)$.
-
-\begin{samepage}
-%\begin{definition}
-A $0/1$ timed automaton $\B=(Q,X,R_0,R_1,q_{init},F)$ 
-induces the labeled transition system 
-$T_{\B} = (\Conf(\B), \lambda_{\B}, \rightarrow_{\B}) $
-where $ \lambda_{\B} = (R_0 \cup R_1) \times \{ 0,1\}$
-	and where $ \rightarrow_{\B}$
-	is defined such that
-	for all $q(z), q'(z') \in \Conf(\B)$, 
-	for all $(\delta,i) \in \lambda_{\B}$
-	with $\delta  = (q,g,U,q')\in R_i$
-	$q(v)\xrightarrow{\delta,i}_{\B} q'(v')$ if
-	$v+i \models g$, 
-	$v'(u)=0$ for all $u \in U$ and $v'(\omega)=v(\omega)+ i$ for all $\omega \in \Omega
-	\setminus U$. 
-%\end{definition}
-\end{samepage}
-
-As expected, we write $q(v)\xrightarrow{\delta,i}_{\B}q'(v')$ if 
-$q(v)\xrightarrow{\delta,i}_{\B}q'(v')$ for some 
-$i\in\{0,1\}$, and some $\delta \in R_i$.
-
-\subsection{One-Counter Automata}
-
-\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there}
-
-\problemx{OCA $k$-resilience problem}
-{A state $q$ of a OCA $(Q, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
-{$\forall q' \in BAD \forall n,n' \in \N ~ (q(n) \rightarrow^* q'(n')) \implies \exists q'' \in SAFE \exists n'' \in \N ~ q'(n') \rightarrow^{\leq k} q''(n'')$ ?\newline}
-
-\problemx{OCA bounded resilience problem}
-{A state $q$ of a OCA $(Q, \Delta)$, a set $SAFE \subseteq Q$, a set $BAD \subseteq Q$.}
-{$\exists k \geq 0 ~ \forall q' \in BAD \forall n,n' \in \N ~ (q(n) \rightarrow^* q'(n')) \implies \exists q'' \in SAFE \exists n'' \in \N ~ q'(n') \rightarrow^{\leq k} q''(n'')$ ?\newline}
-
-%
-\fi
-%
-
 \section{conclusion}

### 2023-06-07 resilience.tex
up

@@
 
-Linked problems
-
+% Linked problems
+{\bf Related problems.} 
 A set of configurations $H$ is an {\em home-space} for a set of configurations $X$ if every configuration reachable from $X$ can reach $H$.

### 2023-06-08 application.tex
rm vass state sketch

@@
 \end{proof}
-
-\mathieu{
-VASS donc la base de $\uparrow post^*(s_0)$ est calculable. 
-Intersection de $\Bad$ et $\uparrow \post^*(s_0)$ est upward-closed donc ça se fait
-					de décider si c'est  $\subseteq \pred^{\leq k}(\Safe)$
-	(re: $k$-résilience décidable voir Prop. 34)
-Problème: on pourrait avoir $\uparrow post^*(s_0) \neq  post^*(s_0)$.
-Si $ \Bad \cap \uparrow \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$ alors on a aussi
-$ \Bad \cap  \post^*(s_0) \subseteq  \pred^{\leq k}(\Safe)$.
-Si non, alors
-notre procédure décisionnelle pour vérifier
-ça
-donne un témoin.
-Le témoin, on peut vérifier si il est bien dans $\post^*(s_0)$
-(c'est juste la question de est-ce que on peut l'atteindre depuis $s_0$
-et l'accessibilité est décidable).
-Si il est effectivement dans $\post^*(s_0)$ alors c'est bon on a aussi un contre exemple
-qui prouve que
-$ \Bad \cap  \post^*(s_0) \nsubseteq  \pred^{\leq k}(\Safe)$.
-Si le témoin n'est pas dans $\post^*(s_0) \ldots$ alors
-$\ldots$
-j'imagine on peut reprendre la procédure décisionnelle sur 
-$ \Bad \cap \uparrow \post^*(s_0)$
-mais en excluant ce premier témoin,
-et ou bien on a que c'est bon si on enlève ce témoins alors on a bien 
-$\subseteq \pred^{\leq k}(\Safe)$
-ou bien ... on obtiens un nouveau témoin ... pour lui aussi on peut
-vérifier si il est effectivement dans $\post^*(s_0)$, et ainsi de suite.
-Si jamais la $k$-résilience n'est pas vérifiée, on vas bien finir par
-trouver un témoin que ça n'est pas vérifiée.
-Si jamais la $k$-résilience est vérifiée par contre ...
-$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
-doit être fini, on a dit.
-Est-ce qu'il y aurais un moyen de calculer ça plutôt ?
-Par exemple,
-on regarde
-$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$,
-comme on a le $\uparrow$ 
-devant $\uparrow \post^*(s_0)$
-alors
-deux choses se produisent,
-d'abords les "trous" entre les éléments de $ \post^*(s_0)$ sont comblés.
-Sauf que si un "trou" est un contre exemple et qu'on a qu'un nombre fini de
-trou alors "on voit bien" que c'est un trou en testant que
-ça appartiens ou non à $ \post^*(s_0)$ et "ça va" on peut s'en débarasser.
-L'autre chose qui se produit, c'est que on a les droites qui vont vers plus l'infini...
-Ça c'est plus difficile à gèrer parce que si
-il y a qu'un nombre fini d'éléments dans $ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
-et pas dans
-$ \Bad \cap  \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
-ça se fait,
-mais si il y a un nombre infini d'éléments dans
-$ \Bad \cap \uparrow \post^*(s_0)  \setminus \pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$
-potentiellement pas dans le plus petit,
-alors c'est plus compliqué de faire notre procédure décisionnelle.
-Il nous faudrait pouvoir décider si $\post^*(s_0)$ est fini ou non.
-Décider si $\post^*(s_0)$ est borné ou non.
-Est-ce que ça peut se réduire à un submarking problem ?
-Il faudrait prendre le complémentaire peut-être ...}
-
-\section{conclusion}
-
-We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...
-
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
-%
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
-
-%
-%
-%\subsection{Vector Addition System with States or PN}
-%
-%\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there} 
-%

### 2023-06-08 conclusion.tex
rm vass state sketch

\section{conclusion}

We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...

\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
%
\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}

%
%
%\subsection{Vector Addition System with States or PN}
%
%\textcolor{red}{Should be defined in a later 'application section' once we start writing any proof, for now I leave it there} 
%

### 2023-06-08 appendix.tex
up

@@
 
-\section{More Appendix thing if necessary}\label{appendix B}
+\section{More Appendix thing if necessary}\label{appendix brp VASS}
+
+\begin{theorem*}{\textbf{\ref{brp VASS}}.}
+{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
+.
+\end{theorem*}
+
+\begin{proof}
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+% $\Safe \subseteq Q \times \N^d$
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+% $q'(\omega)$
+$\uparrow q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+ $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
+% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+% \mathieu{Is it actually possible to check whether or not it is finite ?}
+% If it is finite then w
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If it is infinite then
+$k$-resilience does not hold 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
+
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards,
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+\end{proof}
+
+\section{More Appendix thing if necessary}\label{appendix W}

### 2023-06-08 application.tex
up

@@
 
-\begin{theorem}
+\begin{theorem}\label{brp VASS}
 {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
@@
 \begin{proof}
-Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
-Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
-% $\Safe \subseteq Q \times \N^d$
-$\Safe = S \setminus \Bad$
- downward-closed.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
-
-The set $\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider 
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
-
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-% $q'(\omega)$
-$\uparrow q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-Let us assume from now on 
-% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
- $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
-Because these subsets are upward-closed, we can compute a basis of
-% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
-% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
-$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
-
-We now consider the set 
-% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
-% \mathieu{Is it actually possible to check whether or not it is finite ?}
-% If it is finite then w
-We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
-If it is infinite then
-$k$-resilience does not hold 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
-$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
-
-In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
-and hence
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
-We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
-
-From $k_0$ onwards,
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is stationnary. 
-If the set is finite then we stop the procedure once we have checked for every element
-whether it is possible to reach $\Safe$ from it.
-If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
-hold
-for $k_m = \max(k_0, k_{\pi})$ with 
-$k_{\pi}$ the maximum of the length of the paths from 
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
-If it is infinite then
-for every $k \in \N$,
-$k$-resilience does not hold, 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
+\mathieu{Donner l'idée et se référencer à l'appendix~\ref{appendix brp VASS} pour la preuve complête.}
 \end{proof}

### 2023-06-08 resilience.tex
up
typo

@@
 Resilience and the home-space problem are also linked to the 
-model-checking of the ``from-all'' formulae $\forall s \in X~ \exists t \in Y~ s \to^* t$
+model-checking of the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
 which has been shown decidable for Lossy Counter Machine
@@
 %		
-We \alain{deduce} 
+We deduce 
 % \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
@@
 %	 and ideals are recursive 
+Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$.
 

### 2023-06-08 definitions.tex
up

@@
 %
-\iffalse
-\mathieu{ Maybe add this back ?
-A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
-$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
-${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
-ternary relation,
-denoted as the set of {\em labeled transitions}. }
-\fi
-%
-% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$.
 We write $s \rightarrow s'$ to denote $ (s,s') \in  {\rightarrow} $.
@@
 $\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. % \\
-\alain{supprimer ce qui est inutile}
-
-The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$, whether $s \to^* t$.
-% It is equivalent to $s \in \pred^*(t)$ ?
-or $t \in \post^*(s)$ ?
-\alain{où est-ce utilisé ?}
+The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two states $s, t \in S$, whether $s \to^* t$.
 
@@
 A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
-%\alain{utilisé ? We abbreviate $x \leq y \not\leq x$ by $x < y$.}
-% A partial ordering (a po) is an antisymmetric qo
-	% \alain{
-% Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
-% \alain{utilisé ?}
-%
-% and gives rise to a po between the equivalence classes ?
-% 	do we need po ?
 Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
@@
 A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$. 
+\alain{supprimer ce qui est inutile}
 

### 2023-06-08 definitions.tex
up

@@
 
-\noindent
+% \noindent
 A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
@@
 A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$.
-\alain{supprimer ce qui est inutile}
 
-\begin{definition}
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
-\end{definition}
 %
-% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
 %
@@
 %
-
-% Higman investigated ordered sets with the finite basis property.
-
-Wqo admits many other equivalent formulations. For instance: 
-$(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$)
-if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
-upward-closed subsets in $X$ eventually stabilizes, 
-%      i.e. there is a $k \in N$ such that $U_k = U_{k+1} = U_{k+2} = \cdots $ 
-if and only if any upward-closed subset $U \subseteq X$ has a
-finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
+Wqo admits many other equivalent formulations. 
+% For instance:  $(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$) if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of upward-closed subsets in $X$ eventually stabilizes, if and only if any upward-closed subset $U \subseteq X$ has a finite basis if and only if $(X,\leq)$ is well-founded (i.e.
+it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
 %
@@
 %
+Quasi-orderings that have no 
+%infinite antichains
+infinite subset of mutually incomparable elements (antichains)
+ enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
 
-\noindent
-Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward-closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
+In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement), and an upward-closed set $U$ is represented by its finite set of minimal elements.
+\\
 
-In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement).
-%		$D=\downarrow D$   $I_D=\{J_1, J_2,...,J_n\}$ 		satisfying $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ 
-and an upward-closed set $U$ is represented by its finite set of minimal elements. \\
-% $M_U$ of its  satisfying $U=\uparrow M_U$.
+\alain{supprimer ce qui est inutile}
 

### 2023-06-08 application.tex
up

@@
 
-\begin{theorem}{}
+\begin{theorem}\label{RP VASS}
 Resilience is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
@@
 		 & $\downarrow$~ $\uparrow$~ 
-
  \\ \hline
-   RP  
-   	& Decidable (Thm~\ref{down-up})  
-   		 & ??
-
+   RP  & Decidable (Thm~\ref{down-up})   & Decidable (Thm~\ref{RP VASS})
     \\ \hline
-   BRP  
-   &  Decidable (Thm~\ref{down-up}) 
-   		 & Decidable  ?
-
+   BRP   &  Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{brp VASS})
     \\ \hline
-      kRP  
-      & Decidable (Thm~\ref{down-up}) 
-      		& Decidable ?
-
+      kRP       & Decidable (Thm~\ref{down-up}) & Decidable (Thm~\ref{brp VASS})
        \\ \hline
-   sRP  
-   	& Decidable (Thm~\ref{post srp})
-   		 & ??
-
+   sRP    	& Decidable (Thm~\ref{post srp})	 & ??
     \\ \hline
-   BsRP  
-   &  Decidable (Thm~\ref{ref ozkan})
-   		 & ??
-
+   BsRP     &  Decidable (Thm~\ref{ref ozkan})	 & ?? 
     \\ \hline
-      ksRP   
-      & Decidable (Thm~\ref{ref ozkan})
-      		& ??
-
+      ksRP        & Decidable (Thm~\ref{ref ozkan}) & ?? 
        \\ \hline

### 2023-06-09 definitions.tex
up

@@
 
-We say that a WSTS $\mathscr{S}$ has the \emph{strong (upward) compatibility} when moreover for all 
+We say that a WSTS $\mathscr{S}$ has \emph{strong (upward) compatibility} when moreover for all 
 $s_1, t_1 , s_2 \in S$
@@
 Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
-
 We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
-
 For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
-When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
+When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
 Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
@@
 
-With the effective pred-basis hypothese, we obtain:
+With the effective pred-basis hypothesis, we obtain:
 
@@
 
-With the ideal-effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
+With the ideal-effective hypothesis, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
 

### 2023-06-09 resilience-safe-down.tex
sections
up

@@
 
-let us choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$.
+Let us choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$.
 %			at least one counter is unbounded. 
@@
 
-As an example, let us choose $\Safe=S-\Bad$ then since $\Bad=\uparrow \Bad$, we have: $\Safe=\downarrow \Safe$.
+As an example, let us choose $\Safe=S \setminus \Bad$. Then, since $\Bad=\uparrow \Bad$, we have: $\Safe=\downarrow \Safe$.
+
 %
@@
 it is not enough to conclude.
+
+\alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}

### 2023-06-09 resilience.tex
up

@@
 
-\mathieu{vérifier au passage que c'est vrai qu'il y a pas de résultats home-space WSTS}
+% \mathieu{vérifier au passage que c'est vrai qu'il y a pas de résultats home-space WSTS}
 
@@
 \begin{example}
-Let us consider a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
+Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
 then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g.
@@
 %
-\alain{il manque un résultat d'indecidabilité pour les 3 resilience pour les WSTS+...}
 %	 and ideals are recursive 
@@
 
-\iffalse
-
-\begin{proposition}\label{general}
-\textcolor{red}{
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Bad,\Safe)-resilient (CONJECTURE FOR NOW)
-}
-\end{proposition}
-
-\begin{proof}
-Resilient means that ...
-\end{proof}
-
-\fi
-
-% \mathieu{En fait de plus en plus j'ai l'impression que on peut se passer des neutres: même dans une situation où on aurait des configurations neutre j'aurais tendance à dire qu'on peut se ramener à un cas safe complémentaire de bad en duplicants les neutres en deux parties, une avec "on a vu un very bad récement il faut aller vers safe" (et qui dans le nouveau modèle seraient donc bad) et une "on a pas vu de very bad tout vas bien" (et qui dans le nouveau modèle seraient donc safe)
-% $=>$ Rédiger ça plus proprement}
 While considering a sequence of configurations, every prefix is always 

### 2023-06-09 resilience.tex
up

@@
 
-\alain{faire le lien avec home space, et les formules dans Schnoebelen}
-
-% Linked problems
 {\bf Related problems.} 
@@
   and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for Lossy Counter Machines).
-
-% The Model-checking problem for LTL, CTL*, etc. is a natural generalization of the reachability problem. This problem is undecidable in general but there are simple formulae decidable
 
@@
 
-\begin{remark}[Reformulation]\label{reformulation}
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Safe, \Bad)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-(\Safe, \Bad)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
-\end{remark}
+% \begin{remark}[Reformulation]\label{reformulation}
+% $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+% (\Safe, \Bad)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
+% $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
+% (\Safe, \Bad)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
+% \end{remark}
 

### 2023-06-09 resilience-safe-down.tex
up git push

@@
 \alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}
+
+\mathieu{
+Home-state indécidable pour les minsky machine
+(preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
+or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
+}

### 2023-06-09 resilience-safe-down.tex
up

@@
 }
+
+Without the [] hypothesis, the problem becomes undecidable. Indeed, zero-is-a-home-state problem
+(asking whether zero is a home state) is undecidable for Reset Petri-nets, where it is a particular case of resilience with $\Safe = \downarrow \Safe$, $\Bad = \uparrow \Bad$.
+This stems from the fact that zero-is-a-home-state problem is undecidable for Minsy machines [with more than one counter,].

### 2023-06-09 state-resilience.tex
up

@@
 \begin{proof}
-Consider Reset $d$-VASS
-with $\Safe$ % where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 
-	containing only the
-		configurations in $Q \times \{ \textbf{0} \}$
-	where $\textbf{0} = \min(\mathbb{N}^d)$, %$0$ 
-% \alain{tu n'as pas le doit d'écrire $0$ sans explication}
-and $\Bad = \mathbb{N}^d \setminus \Safe$.
-
-\mathieu{refaire la preuve}
+\mathieu{Refaire la preuve en utilisant les Minsky machines, i.e. si on pet résoudre la question dans les reset alors on peut la résoudre dans les Minsky machine ... et ça y est indécidable.}
 \end{proof}

### 2023-06-11 resilience-safe-up.tex
up

@@
 %
-%		cas étudié par les deux articles sur les graphes
-%
-Let us choose $\Bad$ to be the set of states from which there is no infinite run.
+
+We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
+It can be viewed as a generalizeation/extension of coverability, namely,
+it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
+
+% Let us choose $\Bad$ to be the set of states from which there is no infinite run.
 If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward-closed.
-\alain{ce qui précède est faux et pas convaincant: un vass qui seulement décroit ses compteurs}
-
-%Such a Turing machine, in combination with $M_\rightarrow$, allows computing
-%$\post(x)$ whenever the latter is finite.
+% \alain{ce qui précède est faux et pas convaincant: un vass qui seulement décroit ses compteurs}
 
@@
 
+As an
 example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e.
 from which it is always possible to fire a transition.
-This set of states is upward-closed and $\Safe=S-\Bad=\uparrow \Safe$.\alain{mal rédigé}
+This set of states is upward-closed and $\Safe=S \setminus \Bad=\uparrow \Safe$.\alain{mal rédigé}
 

### 2023-06-11 resilience.tex
up

@@
 \section{Resilience for WSTS}
+
+In this section we introduce the concept of resilience for well structured transition systems. We show that the resilience problems are decidable
+for  completion-post-effective $\omega^2$-WSTS with strong compatibility
+when [...]
+and for downward-compatible ideal-effective WSTS
+when [...]. 
+We show that resilience problems are however undecidable for WSTS with
+strong compatibility in general.
 

### 2023-06-11 reachability/article.tex
remains from pushdown parameter article added in

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\bibliographystyle{plainurl}% the mandatory bibstyle
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{mathtools}
\usepackage{todonotes}
\usepackage{microtype}

\usepackage{complexity}
\usepackage{amsmath}
\usepackage{dsfont}

\usepackage{stmaryrd}
\usepackage{MnSymbol}
\usepackage{graphicx}

\newcommand{\problemx}[3]{
	\vspace{0.2cm}
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf QUESTION:\ }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup
	\vspace{-0.2cm}
}

\newcounter{claimcounter}
\setcounter{claimcounter}{0}
\newtheorem{subclaim}{Subclaim}{}
\newtheorem*{theorem*}{Theorem}

\makeatletter

\usepackage{ textcomp }

\usepackage{stmaryrd}
\usepackage{MnSymbol}
\usepackage{graphicx}

\usepackage{wrapfig}

\newcommand\sg[1]{\todo[inline,size=\scriptsize]{#1 - \textbf{Stefan}}}
\newcommand\mh[1]{\todo[inline,size=\scriptsize]{#1 - \textbf{Mathieu}}}
\newcommand\sgchanged[1]{{\color{red}{'1}}}
\newcommand\mhchanged[1]{{\color{blue}{'1}}}

\renewcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\G}{\mathcal{G}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Time}{\mathbb{T}}
\newcommand{\Const}{\mathsf{Consts}}
\newcommand{\PTA}{\textsc{PTA}}
\newcommand{\MSO}{\textsc{MSO}}

\renewcommand{\poly}{\mathrm{poly}}

\newcommand{\semi}[1]{\xRightarrow{'1}}

\title{Parity games in pushdown systems with pebbles}

\titlerunning{Parity games in pushdown systems with pebbles}

% \author{Stefan G\"oller}{University of Kassel \and School of Electrical Engineering and Computer Science \and Kassel, Germany }{stefan.goeller@uni-kassel.de}{}{The author was supported by the Agence Nationale de la Recherche grant no.  ANR-17-CE40-0010.}
\author{Mathieu Hilaire}{Université Paris-Saclay\and
CNRS\and
ENS Paris-Saclay\and
Laboratoire Méthodes Formelles (LMF)\and
Gif-sur-Yvette, France
}{hilaire@lsv.fr}{}{This work was partly done while the author was supported by the 
Agence Nationale de la Recherche grant no.  ANR-17-CE40-0010.}

\authorrunning{M. Hilaire}

\Copyright{M. Hilaire}

%\ccsdesc[500]{Theory of computation~Timed and hybrid models}
%\ccsdesc{Theory of computation~Automata over infinite objects}
\ccsdesc[500]{Theory of computation~Automata extensions}

\keywords{Parity Games, Computational Complexity, Pushdown systems, Pebble Alternating Two Way Automata}

\category{}

\relatedversion{}
% \relatedversiondetails{Full Version}{}

\acknowledgements{The author would like to thank Benedikt Bollig and Stefan G\"oller for helpful discussions and feedback.}

\ArticleNo{70}
\begin{document}

\maketitle

\begin{abstract}
	We investigate the decidability and complexity of
	parity games over an extension of pushdown systems.
	More specifically, we consider pebble pushdown systems, in
	which pebbles can be used to remember some stack content already visited,
	and compared against.
	We determine decidability of the problem, and provide a nonelementary lower bound. 
\end{abstract}

\section{Introduction}

\subsubsection*{Background}

Stack-based automata are a popular formalism for modeling the sequential behavior of computer programs. 
Pushdown systems are probably the most studied models of stack-based automata.
They are
finite automata 
 extended with a stack that can be manipulated by pushing or popping stack symbols from some finite set, and can use the top of the stack to decide which transition to take next.
\textcolor{black}{To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
Pushdown systems can be used to model the behavior of recursive programs and,
as a result, a variety of 
 program analysis questions can be reduced to decision problems for games on pushdown systems.
They have applications for instance in inter-procedural control-flow analysis of recursive programs \cite{esparza1999automata, reps2005weighted}.

Two player games can be used in particular to
 represent the
interaction of a program with some environment, with the first player representing the program while the second player represents the environment. A winning condition then expresses a property required to hold however the environment behaves. Finding a winning strategy for the first player
thus provides some way 
 to ensure that the desired 
property, as expressed by the winning condition,
always holds \cite{arnold2003games}.

Games on the graphs of pushdown systems have been extensively studied.
It is known in particular from \cite{walukiewicz1996pushdown} that deciding the winner of a parity game on a graph of a pushdown system is an ${\sf EXPTIME}$-complete problem.
Several generalizations of the pushdown systems 
have been studied in depths, including
tree-pushdown systems \cite{guessarian1983pushdown}, ordered
multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and
strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.

We consider here a different approach to extending pushdown systems, by allowing them to test the stack content against precise values.
Pushdown systems which can test the stack content against regular-expressions are
equivalent to pushdown systems,
so pushdown systems which can test the stack content against specific words are too.
We consider instead checks against unspecified word values that can be instantiated one after the other,
providing the pushdown systems the ability to
store information on some stack contents for later comparisons.

Hence we propose in this paper to extend pushdown systems with a set of pebbles.
Compared to pebble word automata, instead of using pebbles
as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles
on their universe of stack contents. Dropping a pebble records the current stack content with an available pebble until it is lifted. Lifting a pebble then makes it available again.

A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
pebbles that it can drop on, and lift from 
words, using them as markers.
Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
\cite{globerman1996complexity, engelfriet1999tree}.
Automata with nested pebbles were also introduced
for tree-walking automata. 
It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}, the main trouble being that the tree-walking automata get lost rather easily.
Using pebbles is a remedy against getting lost along a tree, but 
the unrestricted use of pebbles leads to a class of tree languages much larger than the
regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
Thus, in both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline, and our model will follow this strict stack discipline too.

One motivation to the study of pebble pushdown systems is their ability to compare a stack content against some later stack content.
Similar type of storage for later comparison has already been used for instance
in register pushdown systems. Introduced in \cite{murawski2017reachability}
 as an extension of pushdown systems that 
 can keep data values in
 a set of registers and a stack,
register pushdown systems
enable
manipulation of
data values from an infinite domain,
and have applications for instance for
malware detection and XML schema
checking \cite{senda2021forward, senda2021ltl}.

Another closely related formalism is that of alternating tree automata with nested pebbles, as we 
consider in more details in Section 4. The approach of extending tree automata with pebbles was used in \cite{milo2000typechecking} to show that the XML typechecking problem is decidable, and in \cite{karhumaki2012jewels} to recognize first-order logic on trees. In both cases however, the input trees considered are finite, while 
simulation of plays in pebble pushdown graphs
requires the use of pebble alternating tree automata working on infinite trees
since the stack of a pebble pushdown system is in general unbounded.

The aim of the present paper is to study the decidability and complexity of
deciding
the winner of a parity game on a graph of a pushdown system with  
pebbles.

\subsubsection*{Contributions}

Our main result states that
deciding
the winner of a parity game on a graph of a pebble pushdown system
is decidable but nonelementary.
The contribution is two-fold.

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy.

	For the main contribution, we use pebble alternating two-way tree automata to solve pebble pushdown parity games.
Alternating two-way tree automata with parity winning conditions have been used in the past to solve parity games on pushdown graphs \cite{cachat2002two}. We show that pebble alternating two-way tree automata can similarily be used to solve parity games on pebble pushdown graphs.
We then prove  decidability 
by 
a reduction
 to the problem of model-checking $\MSO$ on a well chosen structure, whose decidability is
 guaranteed by
Muchnik's Theorem~\cite{Wal96}.

\subsubsection*{Overview}

In Section 2, we provide general notations and preliminary definitions. Section 3 will deal with the introduction of pebble pushdown games.
Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve pebble pushdown system parity games. We leave the proof of decidability itself  for Section 5.
Section 6 at last shows a nonelementary lower bound for the problem.

\section{Definitions}

\newcommand{\LCM}{\mathsf{LCM}}
\newcommand{\LOGSPACE}{\mathsf{LOGSPACE}}
\renewcommand{\MSO}{\mathsf{MSO}}
\newcommand{\SO}{\mathsf{SO}}

By $\Z $ we denote the {\em integers} and by $\N=\{0,1,\ldots\}$ we denote the {\em naturals}.
% For every $a,b\in\Z$ with $a\leq b$ we define $[a,b]=\{k\in\Z\mid a\leq k\leq b\}$.
% For every $n\geq 1$ we define $n\Z=\{n\cdot z\mid z\in\Z\}$.
% For every number $n\in\N$ we define $\log(n)=\min\{i+1\mid i\in\N, n\leq 2^i\}$, which is the smallest number of bits necessary to write down $n$ in binary.
For every finite alphabet $A$, $A^*$ is the set of finite
words with letters in  $A$, $A^\omega$ is the set of infinite words with letters in $A$, and
%by
 $A^\infty = A^\omega \cup A^*$. We denote the empty word in $A^*$ by $\epsilon$.
%For all $a\in A$ and all $w\in A^*$ let $|w|_a$ denote the number of occurrences of the letter $a$ in $w$.
%
% For every finite set $M\subset\N\setminus\{0\}$ let $\LCM(M)=\min\{n\geq 1\mid \forall m\in M\setminus\{0\}: m|n\}$ denote the least common multiple of the elements in $M$. 
% For any $j \in \N$ let $\LCM(j)=\LCM([1,j])$ denote the least common multiple of the numbers $\{1,\ldots,j\}$.

 For any two sets $X$ and $S$, let $X^S$ denote the set of all functions from $S$ to $X$.
For any set $S$ let 
%$\mathscr{P}(S) = \{ X \mid X \subseteq S \}$
$\mathcal{P}(S) = \{ X \mid X \subseteq S \}$
denote the power set of $S$.
%
A {\em partial function} $f$ from a set $X$ to a set $Y$ is a function defined on a subset $C$ of $X$ (possibly $X$ itself) with output values in $Y$. We extend the domain of a partial function to the whole set
by considering it as returning the bottom element $\bot$ when it is undefined.

A {\em graph} $G=(V,E)$ is a finite set of graph vertices $V$ with a set of graph edges $E \subseteq V \times V$. 
% Two graphs $G = (V,E)$ and $G'=(V',E')$ with sets of graph vertices $V=V'=V_n=\{1,2,...,n\}$ are said to be {\em isomorphic} if there is a permutation $\rho$ of $V_n$ such that $(u,v)$ is in the set of graph edges $E$ iff $(\rho(u),\rho(v))$ is in the set of graph edges $E'$.
%%%
%
For any set $V$, and a given infinite sequence $\pi = v_0 v_1 \ldots \in V^\omega$ of elements of $V$, we define the set 
$\text{\textit{Inf}}(\pi)$ of elements occurring infinitely often in $\pi$ as
$\text{\textit{Inf}}(\pi) = \{ v \in V \mid \forall i, \exists j > i, ~ v_j = v \}$.

\subsection{Logics}

We now brieﬂy review some standard deﬁnitions from mathematical logic.

\begin{samepage}
\begin{definition}
 A {\em vocabulary} $\tau$ is a set of relational symbols (denoted $P_1, P_2, \ldots $), each of which has a specified
 arity. A symbol $P \in \tau$ is called monadic if its arity is one, i.e., if it is used to
 denote sets.
%
A {\em $\tau$-structure} %(also called a {\em model})
%
$$ \mathfrak{A} = (A,	( P^{\mathfrak{A}} )_{P \in \tau})$$
%
consists of a set $A$ together with an interpretation of
 each $k$-ary relational symbol $P$ from $\tau$ as a $k$-ary relation on $A$; that is, a
set $P^{\mathfrak{A}} \subseteq A^k$.
% A structure $\mathfrak{A}$ is called finite if $A$ and $\sigma$ are finite sets. 
% The {\em universe} of a structure is typically denoted by a Roman letter corresponding to the name of the structure; that is, the universe of $\mathfrak{A}$ is the set $A$, the universe of $\mathfrak{B}$ is $B$, and so on. We shall also occasionally write $a \in \mathfrak{A}$ instead of $a \in A$.\newline
%
\end{definition}
\end{samepage}

%\noindent As an illustration, if $\tau$ consists of a single relational symbol $E$ of arity $2$, graphs are possible $\tau$-structures.

% Next, we deﬁne the  monadic second-order logic, or $\MSO$.

Next, we deﬁne monadic second-order formulas over a vocabulary $\tau$, or $\MSO[\tau]$ for short. %We define free variables, and the semantics of $\MSO(\tau)$ formulas.

\begin{samepage}
\begin{definition}\label{MSO}
We assume countably infinite sets of first-order variables {\em Var}
and of second-order variables {\em VAR}. First-order variables
will be typically denoted by $x, y, z, \ldots,$ with subscripts and superscripts,
whereas second-order variables will be typically denoted by $X, Y, Z, \ldots ,$
using subscripts and superscripts as well. We
inductively define %terms and 
formulas of the monadic second-order logic
over vocabulary $\tau$ %as follows: 
by the grammar:
$$ \phi :=  P(x_1 , \ldots , x_k) \mid x_1 = x_2 \mid X(x) 
		\mid \phi \vee \phi \mid \phi \wedge \phi \mid \neg \phi 
		\mid \exists x ~ \phi \mid  \forall x ~ \phi
		\mid \exists X ~ \phi \mid  \forall X ~ \phi		$$
where $P \in \tau$ is a $k$-ary relational symbol, $X \in \text{VAR}$, and $x, x_1, \ldots, x_k \in \text{Var}$.
\end{definition}
\end{samepage}

A formula that does not use existential ($\exists$) nor universal ($\forall$) quantiﬁers
is called {\em quantifier-free}.
%
% Given a set of formulas $S$, formulas constructed from formulas in $S$ using only the Boolean connectives $\vee$, $\wedge$, and $\neg$ are called {\em Boolean combinations} of formulas in $S$.
%
We shall use the standard shorthand $\phi  \to \psi $ for $ \neg \phi \vee \psi $ 
and $\phi \leftrightarrow \psi $ for
$(\phi  \to \psi ) \wedge (\psi  \to \phi )$.

If $\overrightarrow{x}$ is the tuple of all the free first-order variables of $\phi$,
and $\overrightarrow{X}$ is the tuple of all the free second-order variables of $\phi$ 
we write $\phi(\overrightarrow{x},\overrightarrow{X})$. A {\em sentence}
is a formula without free variables. We use the standard semantic of $\MSO$,
see Appendix~\ref{MSO appendix} for a more thorough induction over the formula.\\

Concerning the logic $\MSO$, we are interested in the following decision problem.

\problemx{$\MSO[\tau]$ model-checking}
{A $\tau$-structure $\mathfrak{A}$ and a $\MSO[\tau]$-sentence $\phi$.}
{Does $\mathfrak{A} \models \phi $ ?\newline}

\textcolor{black}{We define the problem in a general abstract manner, instead of focusing on particular representation for the input, since we will consider many structures of infinite size.}

{\em First-order logic over a vocabulary $\tau$} ($\FO[\tau]$ for short) is the fragment of $\MSO$ which is restricted to first-order quantifiers and variables: quantification is permitted only over individuals, and no second-order variables are used.

\subsection{Parity Games}

We first recall the definition of  two  player parity  games.
Two players, player $0$ and player $1$, move along an arena $\mathcal{G}=(V_0,V_1,E,\Omega)$ which is a labelled graph, composed of
two  disjoint  sets of vertices, $V_0$ and $V_1$ respectively associated with player $0$ and player $1$, 
a set of edges $E \subseteq V \times V$, where $V=V_0 \uplus V_1$,
% We require that the out-degree of each vertex is at least one. This ensures that every finite path in (V, E) can be prolonged.
and a mapping $\Omega:V \to \{0,1, \ldots, m\}$,
$m < \omega$, which
assigns a priority to each vertex. %Then the {\em initialized game} $(\mathcal{G}, v_I)$ is given with an initial vertex $v_I \in V$.

A {\em partial play} of $\mathcal{G}$ is a sequence of vertices $v_0 v_1 \ldots \in V^\infty$ connected by the transition relation, i.e. such that $(v_i, v_{i+1}) \in E$.

A {\em play} of $\mathcal{G}$ is a partial play that is {\em maximal} in the following sense: if it is finite, then the last vertex $v_k$ is such that for all $v' \in V$, $(v_k, v') \notin E$. Else, the play is an infinite sequence $ \pi  = v_0 v_1 \ldots \in  V^{\omega} $.

We consider min-parity games: in order to find out who wins a play, we need to take a
look at the positions which are reached infinitely often and compute their priorities using 
$\Omega$.
Player $0$ wins the play $\pi$  iff  $\min( \{ \Omega(v) \mid v \in \text{\textit{Inf}}( \pi) \})$ is even, else, player $1$ wins. The winner of a finite play is the player whose opponent is unable to move, i.e. the winner of a play such that the last vertex of the play is in $V_{0}$ (resp. $V_1$) is player $1$ (resp. player $0$).

A {\em strategy} $\sigma_i : V^* V_i \to V$ for player $i \in \{ 0,1\}$ is a function that, given a sequence of vertices $ w v$  with $w\in V^*$ and $v \in V_i$,
provides, if possible, a successor $v'$ such that $(v,v') \in E$.
A strategy is called {\em memoryless} if its output only depends on the final vertex of the sequence $v \in V_i$, i.e. if there exists a function $f_i : V_i \to V$ such
that for all $w\in V^*$, $v \in V_i$, we have $\sigma_i(wv) = f_i(v)$.

A partial play $\pi$ is {\em consistent} with a strategy $\sigma_i$ if sequences of vertices that end in $V_i$ along this play all have successors according to the strategy. Given a starting vertex $v_I$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play
starting with $v_I$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $v_I$. It is 
denoted by $\pi(v_I, \sigma_0, \sigma_1)$.

A strategy $\sigma_i$ for player $i$ from position $v_I \in V$ is a {\em winning strategy from $v_I$} if 
no matter which
strategy $\sigma_{1-i}$ the other player pursues,
% the resulting play  starting from $v_I$  
%  results in 
% player $i$ winning.
player $i$ wins the play $\pi(v_I, \sigma_0, \sigma_1)$.

The above definitions are essentially the same for finite and infinite arenas.
%
%
In this paper we are going to concern ourselves with parity games over potentially infinitely large arenas, \textcolor{black}{thus we define the following decision problem in a general abstract manner. In particular, we do not discuss representation of the input}.

\begin{samepage}
\problemx{Solving parity game}
{An arena $\mathcal{G}=(V_0,V_1,E,\Omega)$, an initial vertex $v_I \in V$.}
{Does player $0$ have a winning strategy from $v_I$ in the parity game $\mathcal{G}$?\newline}
\end{samepage}

One important property of parity games is that of {\em memoryless determinacy}: for a parity game and an initial vertex $v_I \in V$, one of the players has a memoryless winning strategy from $v_I$ \cite{zielonka1998infinite}.

\section{Pebble pushdown games}

In this section, we extend pushdown systems with a set of pebbles. 
Pebble pushdown systems will have the ability to lift or drop pebbles on their universe of stack contents according to a strict stack discipline. A \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) and a \textit{lift} pops the last dropped pebble. One can think of a pebble as a register that can store a stack content.

\begin{definition}
An $n$-pebble pushdown system ($n$-PPDS) is a tuple $\mathcal{Z} = (Q, \Gamma,  \Delta )$
where:
\begin{itemize}
\item $Q$ is a {\em finite set of control states},
\item $\Gamma$ is a {\em finite stack alphabet},
% \item $\$ \notin W$ is an {\em initial stack symbol},
\item  $\Delta   \subseteq  Q  \times (\Gamma \biguplus \{ \$ \})  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times Q  \times (\Gamma^* \cup \{ \! \! \text{\textit{lift}}, \text{\textit{drop}}\})$ is a {\em finite transition relation}, where ${\$ \notin \Gamma}$ is some bottom of stack symbol.
\end{itemize}
\end{definition}

A transition in an $n$-pebble pushdown system works like a transition in a pushdown system that additionally manages the $n$ pebbles.
The idea of a transtion $(q, a, i, S, q', \theta) \in \Delta$
is that, if the automaton $\mathcal{Z}$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\mathcal{Z}$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$\theta$.

A {\em pebble set} of $\mathcal{Z}$ is a set $P \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
{\em $P$-pebble assignment} is a function which maps each $j \in P$ to a word in $\$\Gamma^*$.
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, f )$, where 
$q$ is a state,
$w \in \$\Gamma^*$, and
$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
We call  $q$ the current state, $w$ the current stack, and
$f$ the current pebble assignment. 
We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
$f(j) = w_j$, for each
$j \leq i$.

The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. The transition graph of $\mathcal{Z}$ is $(V,E)$, where $V$ is the set of configurations. %and $E$ is such that
Moreover, 
for all $(q, a, i, S, q', \theta) \in \Delta$, with $a \in \Gamma$ (respectively $ \! a = \$ $),
 for all words $w \in \$ \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments

 \begin{samepage}
\noindent
 $f$ such that $f(j) = wa$ (respectively $f(j)= \$ $) for each
$j \in S$ 	and
 $f(j) \neq wa$ (resp. $f(j) \neq \$ $) for each $j \in \{ 1, \ldots , i \} \setminus S$,
%the following holds
we include in $E$ the following transitions:
\begin{itemize}
\item if $\theta \in \Gamma^*$, 
$((q,wa,f), (q',w\theta, f)) \in E $ \\
(respectively, $((q,\$,f), (q',\$\theta, f)) \in E $),
\item if $ \theta = \text{\textit{drop}}$, 
$((q,wa,f), (q',wa, f')) \in E $  \\
(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
where $f'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$f'(j) = f(j)$ for each
$j \leq i$, and $f'(i+1) = wa$ (respectively, $f'(i+1) = \$$), and
\item if $\theta = \text{\textit{lift}}$, and $ i \in S $%$i>0$
, 
$((q,wa,f), (q',wa, f')) \in E $ \\ 
(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
where $f'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$f'(j) = f(j)$ for each
$j < i$.
\end{itemize}
\end{samepage}

This defines a graph  $(V,E) $. 
%
To obtain
a parity game, it remains to define the sets $V_0$ and $V_1$ associating the vertices
to the two players, and the priorities of the configurations. One fixes a disjoint
union
 $Q = Q_0  \biguplus Q_1 $. Then 
 $V_0$ is the set of configurations with current state in $Q_0$ and 
 $V_1$ is the set of configurations with current state in $Q_1$.
The mapping  $\Omega$ is first
defined on $Q$, then  $\Omega (q,w,f) =  \Omega (q)$,  for all words $ w  \in \$ \Gamma^*$, all pebble assignments $f$, and 
$q  \in  Q$. So the player and the
priority only depend on the control states of $\mathcal{Z}$. 
This defines the {\em $n$-pebble pushdown game} $ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$.
The starting configuration is defined if we also fix an initial state 
$q_I  \in  Q$ as $v_I = (q_I, \$, f_I)$, where $f_I$ is the $\emptyset$-pebble assignment. \newline

The main problem we are interested in in this article is the following decision problem.

\problemx{Solving pebble pushdown parity game}
{A pebble pushdown system $\mathcal{Z}= (Q,\Gamma, \Delta )$, a disjoint
union $Q = Q_0  \biguplus Q_1 $, a priority mapping $\Omega$, an initial state $q_I \in Q$% and initial stack symbol $w_I \in W$
.}
{Does player $0$ have a winning strategy from $(q_I,\$, f_I)$ in the  parity game
$ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$ ?}
%$(\mathcal{G}_{\mathcal{Z}}, \Omega)$ ?\newline}

\section{Tree Automata}

In this section, we use pebble alternating two-way tree automata to solve pebble pushdown parity games. 
We introduce first alternating two-way tree automata, then pebble alternating two-way tree automata,
then explain the reduction. Decidability questions concerning pebble alternating two-way tree automata
are dealt with in Section~\ref{section decidability}.

\subsection{Alternating two-way tree automata}

Given a finite set $\Gamma$ of directions, a $\Gamma$-tree is a prefix closed set 
$T  \subseteq \Gamma^*$,
i.e., if $x.d  \in  T$, where 
$x  \in  \Gamma^*$ and $d  \in  \Gamma $, then $x  \in  T $. 
We will sometimes forget the “.” of the concatenation. 
The elements of $T$ are called {\em nodes}, the
 empty word  $\epsilon$
% word $\$ $
  is the {\em root} of $T $. For every $x.d  \in  T$, $d  \in  \Gamma$ the node $x$ is the
unique {\em parent} of $x.d$, and $x.d$ is a {\em $d$-child} of $x$. The {\em direction} of a node $x.d$
$( \neq   \epsilon )$ 
%$(\neq \$ )$
is $d$. 
The full infinite tree is 
$T = \Gamma^*$. A {\em path (branch)} of a tree $T$ is
a sequence  $\beta   \in  T^{\infty}$  such that 
$ \beta  = u_0 u_1 \ldots u_n$ or 
$ \beta  = u_0 u_1 u_2 \ldots$, where
$u_0 =  \epsilon$  and
 $\forall i < n,  \exists d  \in  \Gamma$, $u_{i+1} = u_i.d$. 
 A path can be finite or infinite.

For a finite set $X$, $\mathcal{B}^+ (X)$ is the set of positive Boolean formulas
over $X$ (i.e., Boolean formulas built from elements in $X$ using only $\wedge$ and $\vee$),
where we also allow the formulas $\text{\textit{true}}$ and $\text{\textit{false}}$. For a set $Y  \subseteq  X$ and a
formula $ \vartheta   \in  \mathcal{B}^+ (X)$, we say that $Y$ satisfies $ \vartheta$  iff assigning $\text{\textit{true}}$ to elements in
$Y$ and $\text{\textit{false}}$ to elements in $X \setminus Y$ makes  $\vartheta $ true.

The set $\text{\textit{types}}(\Gamma) = \{r \} \cup \Gamma$ 
describes the possible types of a node in the full infinite tree $\Gamma^*$. 
Here $r$ stands for the root, $\gamma \in \Gamma$ for a $\gamma$-child.

To navigate through the tree let $ext(\Gamma ) := \Gamma \biguplus \{ \epsilon , \uparrow \}$ be the extended set of directions: The symbol $\uparrow$ means “go to parent node”, 
and  $\epsilon$  means “stay on the present
node”. Formally we define $ \forall u  \in \$ \Gamma^* , d  \in  \Gamma, u.\epsilon  = u$ and $u.d. \! \uparrow= u$, while the node 
$ \epsilon . \! \uparrow $ is not defined.

\begin{definition}
An alternating two-way automaton over
 $\Gamma$-trees (ATWA) is a tuple
$\mathcal{A} = (Q,  \delta , q_I , 
\Omega)$ where
\begin{itemize}
\item $Q$ is a {\em finite set of states},
\item $\delta  : Q  \times  
%
\text{\textit{types}}(\Gamma)
%W \cup \$
  \to  \mathcal{B}^+ (Q \times ext(\Gamma))$ is a {\em transition function},
\item $q_I$ is an {\em initial control state}, and
\item $\Omega: Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
\end{itemize}
\end{definition}

The idea of a transition $ \delta (q, a) = ( q',\uparrow ) \wedge (q'',d)$ with $a,d\in\Gamma$ is the following: 
if the automaton
$\mathcal{A}$ is in state $q$ on the node $x.a$ of the 
tree $T$, it will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
another “copy” in state $q''$ to $x.a.d$. 
After that the two copies are
running independently. They may come again to the same node with two different
states.

More precisely a run of an alternating two-way automaton $\mathcal{A}$ over a tree
is
a tree $T_r$ in which every node is labelled by an
element of $Q \times \Gamma^*$. The latter tree is like the unfolding of the run, its structure
is different from the tree $\Gamma^*$, and its set of directions range rather on the naturals smaller than $|Q|*|ext(\Gamma)|$; a node in $T_r$ labelled by $(q,x)$ describes a “copy”
of the automaton that is in state $q$ and is situated at the node $x$ of $\Gamma^*$. 
Note
that many nodes of $T_r$ can correspond to the same node of $\Gamma^*$, because the
automaton can come back to a previously visited node. 
The label of a node
and its successors have to satisfy the transition function,
that is, for a node $y$ of $T_r$ labelled by $(q,x)$, and $\delta(q,t)= \vartheta$ where $t$ is the type of $x$, there is a (possibly empty) set 
 $Y = \{ (q_0,d_0),(q_1,d_1), \ldots (q_\ell, d_\ell) \}  \subseteq  Q \times ext(\Gamma)$, such that $Y$ satisfies  $\vartheta $, and for all
$(q_i,d_i)  \in  Y$, the node $y$ has a corresponding son $y.i$ labelled by $(q_i,x.d_i)$. Finally, the root of $T_r$ is labelled by $(q_I,\epsilon)$.

Remember that $x.d$ can be $x. \epsilon$  or $x. \! \uparrow$, and the latter is defined only if 
%$x \neq \$ $.
$x  \neq   \epsilon $.
So the run cannot go up from the root of the input tree. Note that it cannot use a
transition  $\delta (q, a) = \text{\textit{false}}$ since the formula $\text{\textit{false}}$ cannot be satisfied.
A run $T_r$ is accepting if all its infinite paths satisfy the parity acceptance 
condition
given by the priority function. 
% The finite paths of a run that end with a transition $ \vartheta $ equivalent to  $true$, which is viewed as successful termination. 
An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
%this path
the sequence $\gamma$ of corresponding labels
 is even, i.e.
%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
 $\min(\{ \Omega(q) \mid (q,x) \in \text{\textit{Inf}}(\gamma) \} $ is even. 
Such
a path in the run consists of following only one “copy” of the automaton.

% The generalization of twa with nested pebbles came recently, by two independent motivations: On the one hand, with the advancement of XML theory, finite state recognizers (with name n-pebble tree automata) were used in [21] to show that the XML typechecking problem is decidable. On the other hand, the concept of n-pebble tree-walking automata (n-ptwa) were defined in [9] to recognize first-orderlogic on trees.  Later, in [10]n-ptwa were extended with a more general pebble handling.

\subsection{Alternating two-way tree automata with pebbles}

Informally an $n$-pebble alternating two-way tree automaton works much like an alternating two-way tree automaton, but extended with a fixed set of pebbles, numbered from $1$ to $n$ that it can place on the tree. 
% At each time, pebbles $1,\ldots, i$ are placed on some nodes of the tree, for some $i \in \{1, 2, \ldots , n \}$, or no pebble are placed on the tree. 
To navigate through the tree the automaton can stay at the current node, move to its parent, to its 
 children, it can lift a pebble or place a pebble on the current node.
 Which of these transitions can be applied depends on the current state, the set
 of pebbles at the current node, 
and the type of the current node (root, $\gamma$-child for $\gamma \in \Gamma$). 
Operations $\text{\textit{lift}}$ and $\text{\textit{drop}}$ manage the
pebbles much like in pebble pushdown systems.
We indicate the possible kinds of moves of a pebble automaton by elements of the set 
$ext_2(\Gamma) = ext(\Gamma) \biguplus \{ \text{\textit{lift}}, \text{\textit{drop}}\}$.
% Clearly $\text{\textit{drop}}$ refers to dropping a pebble and $\text{\textit{lift}}$ to lifting a pebble. Note that the notation is  unambiguous by virtue of the stack discipline. 
% Finally recall that if $S$ is a set then $\mathcal{P}(S)$ denotes the powerset of $S$.

\begin{samepage}
\begin{definition}
For $n\geq 0$, an $n$-pebble alternating two-way automaton ($n$-PATWA) over  
$\Gamma$-trees is a tuple
$\mathcal{A} = (Q,   
\delta , q_I , 
\Omega)$ where
\begin{itemize}
\item $Q$ is a {\em finite set of states},
\item $\delta  : Q  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times 
\text{\textit{types}}(\Gamma)
  \to  \mathcal{B}^+ (Q \times ext_2(\Gamma))$ is a {\em transition function},
\item $q_I$ is an {\em initial control state}, and
\item $\Omega : Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
\end{itemize}
\end{definition}
\end{samepage}

The idea of a transition in a $n$-pebble alternating two-way tree automaton works like a transition in an alternating two-way tree automaton,
except the transition has additional conditions on the pebble assignment, and additional moves for taking care of the pebble.
Consider for instance 
that the automaton
 $\mathcal{A}$ is on the node $x.a$, in state $q$, pebbles $1, 2$ are on the tree, and exactly pebble $1$ is placed on the node $x.a$.
Then a transition
$\delta(q, 2, \{ 1\}, a) =  (q',\uparrow) \wedge (q'',\text{\textit{drop}})$
 will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
another “copy” in state $q''$ to $x.a$ with both pebble $1$ and pebble $3$ placed on the node. 
After that the two copies are
running independently.

%A {\em pebble set} of $\mathcal{A}$ is a set $P \subseteq \{ 1, \ldots, n\}$. 
For a tree structure $T$, a 
{\em $P$-pebble assignment} for a pebble set $P$ is a function which maps each $j \in P$ to a node in $T$.
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, x, f )$, where $x$ is a node, $q$ a state and
$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
We call $q$ the current state, $x$ the current node, and
$f$ the current pebble assignment. 
We also write $(q, x, x_{1} , \ldots , x_{i} )$ if 
$f(j) = x_j$, for each
$j \leq i$.

Then, formally, a run of an $n$-pebble alternating two-way automaton $\mathcal{A}$
over a tree is a tree $T_r$ in which every node is labelled by 
an $i$-configuration,
for some $i \in \{0, 1, \ldots, n\}$. A node in $T_r$ labelled by $(q, x, x_{1}, x_{2}, \ldots, x_{i})$ describes a “copy” of the automaton that is in state $q$, and is situated on the node $x$ of $\Gamma^*$, with the pebbles up to $i$ already dropped, with the $j$-th pebble for $j \leq i$ dropped on the node $x_{j}$.

The label of a
node and its successors have to satisfy the transition function. % However, recall there are restrictions on the drop-operation and the lift-operation : only
% the pebble with the number $i$ can be lifted and only the pebble with number $i - 1$ can be
% placed.
Formally, this works much the same way as for alternating two-way tree automata; 
that is, the root of $T_r$ is
labelled by $(q_I,\epsilon, f_I )$, and, for a node $y$ of $T_r$ labelled by an $i$-configuration $(q, x, f)$, and 
for
$\delta(q, i, S, t) = \vartheta$ where $t$ is
the type of $x$, 
and $f(j) = x$ for each $j \in S$, and $f(j) \neq x$ for each $j \in \{1,\ldots,i \} \setminus S$,
there is a (possibly empty) set 
$Y \subseteq Q \times ext_2(\Gamma )$, such that $Y$ satisfies $\vartheta$,
and for all $(q',d) \in Y$, the node $y$ has a corresponding son
labelled with %$(x.d, q', f')$, where
\begin{itemize}
\item if $d \in ext(\Gamma)$,
the $i$-configuration $(q ', x.d , f)$,
%
\item if $d = \text{\textit{drop}}$, the $(i+1)$-configuration  $ (q', x, f')$, where the domain of $f'$ is $\{ 1, \ldots, i, i+1\}$, $f'$ is equal to $f$ on the domain of $f$, and $f'(i+1) =x$, and
%
\item if $d = \text{\textit{lift}}$
and $f(i) = x$, i.e. if the pebble $i$ is on the current node%
, the $(i-1)$-configuration $(q', x, f')$, where $f'$ is the restriction of $f$ to $\{1, \ldots, i-1 \}$.
\end{itemize}

A run $T_r$ is {\em accepting} if all its infinite paths satisfy the parity acceptance 
condition 
given by the priority function $\Omega$.
% Again finite paths end with a transition with formula $ \vartheta $ equivalent to $\text{\textit{true}}$, which is viewed as successful termination. 
An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
%this path
the sequence $\gamma$ of corresponding labels
 is even, i.e.
%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
 $\min(\{ \Omega(q) \mid (q,x,f) \in \text{\textit{Inf}}(\gamma) \} $ is even.

By a pebble alternating tree-walking automaton (PATWA) we mean an $n$-PATWA for some $n$. 
%Note that a $n$-pebble automaton without pebbles is just an alternating tree-walking automaton.

\subsection{Reduction details}

Alternating two-way tree automata with parity winning conditions have already been used to solve parity games on pushdown graphs in \cite{cachat2002two}, with a technique first developed to solve the model-checking problem in \cite{kupferman2000automata}. The main
idea was to reduce a model-checking problem to an emptiness problem for the class of alternating two-way parity tree automata, and was then adapted to work on pushdown parity games.
%
%
This leads to the following reduction: given a 
parity game on pushdown graph $\mathcal{G}$ and a starting configuration $v_I$, we can
effectively construct an ATWA $\mathcal{A}$
such that
player $0$ has a winning strategy from $v_I$ in $\mathcal{G}$ iff $\mathcal{A}$ accepts the full
infinite tree $\Gamma^*$. The main idea of our reduction is that, when pushdown systems and ATWA are extended with a set of pebbles to manage, the technique used %is conserved.
still works.
We refer the reader to Appendix~\ref{appendix tree reduction} for more details on the proof.

\begin{theorem}~\label{theorem reduction PATWA}
For every $n$-PPDS $\mathcal{Z} = (Q_0 \biguplus Q_1, \Gamma,\Delta)$, 
priority mapping $\Omega$, and state $q_I \in Q_0 \biguplus Q_1$,
we can construct 
effectively
an $n$-PATWA $\mathcal{A}$ 
such that
player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the parity game 
$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$
iff
$\mathcal{A}$ accepts the full infinite tree $\Gamma^*$.
\end{theorem}

% Decidability of checking that the $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\Gamma^*$ is then dealt with later in the subsequent section.

\section{Decidability}\label{section decidability}

\section{Lower bound}

The aim of this section is to show a nonelementary lower bound for the 
problem of deciding whether player $0$ has a winning strategy from some starting configuration in
a pebble pushdown parity game.
% problem of deciding winning configuration for pebble pushdown parity games.

We reduce the satisfiability problem for first-order logic on words
to the problem of {\sc Solving pebble pushdown parity game}. 
We know from \cite{Sto74}
that satisfiability is nonelementary. 
%

% Here a function is in {\sf elementary} if it is  $O(h_k (n))$ for one of the $k$-fold exponential functions $h_k$ with $h_0 (n) = n$ and $h_{k+1} (n) = 2^{h_k (n)}$.

We define the tower function $T : \N \times \R \to \R$ by $T(0,r) =r$ and
$T(h+1,r)= 2^{T(h,r)}$ for all $h \in \N, r \in \R$. Thus $T(h,r)$ is a tower of $2$s of height $h$ with an $r$ sitting on top.
Observe that for all $n,h \in \N$ with $n\geq 1$, we have
$T(h, log^{(h)}(n)) = n$.
%
% 
Here a problem is in ${\sf ELEMENTARY}$ if it can be solved in time $O ( T(n, 0) )$.

For a finite alphabet $\Sigma$,  let $\tau (\Sigma )$ be the vocabulary consisting of a binary relational symbol $\leq$,  and a unary relational symbol $P_a$ for every $a \in \Sigma$. 
A word $w \in \Sigma^*$ can be seen as a $\tau (\Sigma )$-structure where the set of elements consists of a finite interval of $\N$. Elements are seen as positions in the word and for every element $i$, there exists precisely one $a \in \Sigma$ such that $i$ is in the interpretation of $P_a$. Lastly $\leq$ is interpreted as the linear order of $\N$.\\

It is well-known that if we are interested in the complexity of first-order or monadic
second-order model-checking on words, the alphabet can be assumed to be $\{0,1\}$ without loss of generality.
Thus the satisfiability problem is concerned with the following question.
\begin{samepage}
\problemx{$\FO$ satisfiability on words}
{A  
$\FO[\tau(\{0,1\})]$-sentence $\phi$.}
{Does there exists a word $w\in \{0, 1\}^*$, such that $w \models \phi$ ?\newline}
\end{samepage}

The result that is previously known and motivates the upcoming reduction consists in the following.

\begin{theorem}[\cite{Sto74}]\label{FOnonelementary}

%  Assume that $FPT \neq AW[*]$, and let $f$ be an elementary function and let $p$ be a polynomial.
%  There is no model-checking algorithm for first-order logic on the class of words 
%  whose running time is bounded by
%  $$ f(k) \cdot p(n)$$

% where $k$ denotes the size of the input sentence of the model-checking problem and $n$ the
% size of the input word.

The {\sc $\FO$ satisfiability on words} problem is not in {\sc ELEMENTARY}.
	% SAT(N,<) is not elementary-recursive
	% and in fact SAT(N,<) \not\in NSPACE(g(log_b(n),0)) for all b > 3.
\end{theorem}

\subsection{Reduction from FO SAT on words}

The following theorem states a polynomial time reduction from
{\sc $\FO$ satisfiability on words}
%We consider reduction 
towards the problem of %following problem.
{\sc Solving pebble pushdown parity game}.

\begin{samepage}
\begin{theorem}
{\sc FO satisfiability on words} is polynomial time reducible to {\sc Solving pebble pushdown parity game}.
\end{theorem}
\end{samepage}

\begin{proof}

\noindent
We start with
% a $\tau(\{0,1\})$-structure $w$ and
 a 
 $\FO[\tau(\{0,1\})]$%%
% $\FO$%
-formula $\phi$.
%
% Note that $w$ corresponds to a word $w \in \{0,1\}^*$ of size $n$.\\
%
% \noindent
We construct
a pebble pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
union $Q_\phi = Q_0  \biguplus Q_1 $, a priority mapping $\Omega_\phi$, and an initial state $q_I \in Q_\phi$
such that
player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the  parity game
$\mathcal{G}_{\mathcal{Z_\phi}, Q_0,Q_1, \Omega_\phi}$
if and only if
there exists a word $w \in \{0,1\}^*$ such that
$w  \models \phi $. \\

\noindent
We assume without loss of generality that $\phi$ is written in the prenex normal form
$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k ~ \psi$, where 
an even number $k$ of quantifiers alternate between existential and universal ones, and 
$\psi$ is quantifier-free and in disjunctive normal form.

Player $0$ starts by pushing a word $w$ onto the stack, then places the first pebble at the current position. The goal of this first pebble is to remember the word $\$w$, which corresponds to the 
word “guessed” by player $0$. This pebble is meant to enforce that the stack remain a prefix of $\$w$, with each stack content of the play corresponding to a position in the word. To enforce this, once pebble $1$ is placed, we allow in $\mathcal{Z}_\phi$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position on the word. A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to pebble $1$.

Then players alternate placing pebbles on prefixes of $\$w$. Universal variables correspond to pebbles placed by player $1$ and existential variables to pebbles placed by player $0$. 
Once players have chosen values for their respective variables, the pebble assignment $f$ is fixed. It is then time
for player $0$ to ensure that
$w$ satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.

Player $0$, being the existential player, choses one of the conjunctive clause in $\psi$, essentially claiming to be able to prove it. Then player $1$, the universal player, chooses one of the atomic formula in the clause to test. 
%The process can be seen in Figure~\ref{DNF}. 
Testing the atomic formula is the purpose of a gadget 
%$ \C_{\psi_{i,j}}$ built such that
in which
 player $0$ has a winning strategy 
 %in $ \C_{\psi_{i,j}}$
  if and only if 
$w$ satisfies %$\psi_{i,j}$
the atomic formula
 if the variables $x_1, \ldots, x_k$ are interpreted by 
$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.

The gadget depends on the type of the atomic formula. For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then check that the the top of the stack
$x$ is the right letter. For one checking that $x = x'$ where $x$ and $x'$ are both variables,
player $0$ goes to the position corresponding to $x$ and check against the presence of the pebble corresponding to $x'$ as well. If the formula is a negation, players exchange their roles.
\end{proof}

Then, the nonelementary lower bound follows from Theorem~\ref{FOnonelementary}.

\begin{theorem}

The problem of {\sc Solving pebble pushdown parity game} is not in {\sc ELEMENTARY}.

\end{theorem}

\section{Conclusion}

In this paper we have shown that
deciding
the winner of a parity game on a graph of a pebble pushdown system 
is nonelementary.

	For the nonelementary lower bound
we reduced the $\FO$ satisfiability problem on words \---- known to be nonelementary
from \cite{Sto74} \---- to the
problem of deciding whether player $0$ has a winning strategy from some starting configuration in 
a pebble pushdown parity game.

	For the decidability upper bound, we  used pebble alternating two-way tree automata to solve pebble pushdown parity games.
Alternating two-way tree automata (ATWA) with parity winning conditions have been used to solve parity games on pushdown graphs \cite{cachat2002two}. The technique was first developped to solve model-checking problem for pushdown graphs in \cite{kupferman2000automata}. 
We showed that, when pushdown graphs and ATWA are extended with a set of pebbles to manage, the technique
also works.
We then proved decidability 
% for checking that an $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\{0,1\}^*$ 
by %reducing it
a reduction
 to the problem of model-checking $\MSO$ on a well chosen structure. To ensure decidability of model-checking $\MSO$ on the chosen structure, we made use of 
Muchnik's Theorem: the structure is constructed by applying a successive number of times Muchnik's iteration to a structure $\mathfrak{A}_0$ for which model-checking $\MSO$ is decidable.
%  After providing this structure, we will show that  checking that an $n$-PATWA $\mathcal{A}$ accepts the full binary tree $\{0,1\}^*$ reduces itself to the problem of model-checking $\MSO$ on it.

\bibliography{bibliography/bib}

\appendix

\section{Reduction from PPDS to PATWA}\label{appendix tree reduction}
% \subsection{Construction}

We provide in this appendix more details on Theorem~\ref{theorem reduction PATWA}. Let us first recall the theorem.

\begin{theorem*}
For every $n$-PPDS $\mathcal{Z} = (Q_0 \biguplus Q_1, \Gamma,\Delta)$, 
a state $q_I \in Q = Q_0 \biguplus Q_1$,
and priority mapping $\Omega: Q \to \{1, \ldots, m\}$,
we can construct 
effectively
an $n$-PATWA $\mathcal{A}$ 
such that
player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the parity game 
$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$
iff
$\mathcal{A}$ accepts the full infinite tree $\Gamma^*$.
\end{theorem*}

We construct a pebble alternating two-way tree automaton $\mathcal{A}$ that determines if player
$0$ has a winning strategy from $(q_I, \$, f_I)$ in $\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$, i.e., wins every play, whatever player $1$ does. 
The
automaton $\mathcal{A}$ will simulate the transitions of the pebble pushdown system $\mathcal{Z}$ on the full
$\Gamma$-tree, guess nondeterministically the best moves of player $0$ and follow each
possible move of player $1$ using alternation.
This tree automaton can
go either up or down the tree: it goes down to simulate a rule that pushes
some new symbol, and it goes up to simulate a popping rule, and it drops or lifts a pebble to represent the dropping or lifting of a pebble. For each possible move, the control state has to
be updated in accordance with the pushdown transition rules. As we aim to
simulate a game, the tree automaton needs to be alternating: states
 associated to player $0$ states will have disjonctive booleans formulas
  while states associated to
player $1$ states will have conjonctive ones.
Finally, the priority mapping is inherited from $\Omega$.

As an example, the stack effect of a transition $(q, a, 0, \emptyset, q' , cb)  \in \Delta_{\mathcal{Z}} $ from a configuration $q(wa, \mu)$
of
the PPDS can be simulated by a $n$-PATWA over the full
$\Gamma$-tree from the node $ua$ by the following sequence of moves: 
$\uparrow, c, b$. Indeed
$wa \! \uparrow cb = wcb$. We have chosen suffix rewriting rather than prefix to conform with
the notation of the tree. 
The control states of $\mathcal{A}$ serve to store the control states of $\mathcal{Z}$.
In order to be able to simulate with many steps a transition of $\mathcal{Z}$, the control states of $\mathcal{A}$ are also used to remember the letters it has to write (see Figure~\ref{stack simulation} for an illustration).

\begin{center}
	\begin{figure}
		\hspace{2.1cm}
\includegraphics[width=0.66\textwidth]{figures/stack_simulation}
	\caption{Transition $(p, a, \emptyset, q, cb)$ of the RPDS
	%(long arrow)
	 simulated by the PATWA.}
	 \label{stack simulation}
	\end{figure}
\end{center}

Moving up or down the tree structure is enough in order to simulate the modifications of the stack
performed by transitions of $\mathcal{Z}$. Additionally, dropping or lifting pebble is enough to simulate the dropping or lifting of pebbles in the PPDS. As an example, we simulate
$(q, a, 0, \emptyset, q', \text{\textit{drop}})  \in   \Delta_{\mathcal{Z}} $ by the single move $\text{\textit{drop}}$, and
$(q, a, i, \{i\}, q', \text{\textit{lift}})  \in   \Delta_{\mathcal{Z}} $ by the single move 
$\text{\textit{lift}}$. Additionally, we simulate checks against pebble sets in $\mathcal{Z}$ by similar checks in the PATWA.

More formally now, the control states of $\mathcal{A}$ are tuples
from $( Q \times ext_2(\Gamma) )^*$  
composed of a control state of  $\mathcal{Z}$, and
a sequence of elements of $ ext_2(\Gamma)$.

The sequence of letters is used to remember the letters the automaton has to write. 
Each sequence of letters has at worst the length of a corresponding output of the transition relation of $\mathcal{Z}$, hence the sequences of letters of $\Gamma$ are more precisely elements of $\Gamma^{ \leq |\Delta_{\mathcal{Z}}|}$.

From a state $(q, \epsilon)$, for a transition
$(q, a, i, S, q', x) \in \Delta_{\mathcal{Z}}$,
the automaton's first move simulating the transition, be it the dropping of a pebble, the lifting of a pebble, or going up the tree, is accompanied by 
checks that the pebbles $1, \ldots, i$ are placed on the tree \---- or that no pebble is on the tree if $i=0$ \---- and checks against the presence of the pebbles of $S$.

From a state $(q , bx)$ ($bx \neq \epsilon$ because $b  \in  \Gamma$) the automaton goes down
in direction $b$, that is to say writes $b$, whatever it reads, and remembers the
(sub)word $x$ it still has to write and the state $q$ of the pushdown system.

If $q  \in  Q_1$ then $ (q,w)  \in  V_1$ for all $w \in \Gamma^{ \leq |\Delta_{\mathcal{Z}}|}$
and
$\mathcal{A}$
executes from $(q,\epsilon)$ all the possible moves of player $1$, to ensure that player $0$ can win after
each of these moves. But if $q  \in  Q_0$, $\mathcal{A}$ in $(q,\epsilon)$ chooses nondeterministically a move of
player $0$ and tries to make player $0$ win. \newline

The priority mapping of $\mathcal{A}$ is almost the same as the one of 
$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$: 
$\Omega_\A(q, x) =
\Omega (q)$. 
The initial state of $\mathcal{A}$ causes it to go deterministically to the starting
configuration of the game.

\begin{lemma}~\label{lemma reduction PATWA}
Player $0$ has a winning strategy from $(q_I, \$, f_I)$ in $\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$ iff $\mathcal{A}$ accepts the full
infinite tree $\Gamma^* $.

\end{lemma}

This concludes our construction for the proof of Theorem~\ref{theorem reduction PATWA}.

\end{document}

### 2023-06-11 reachability/bibliography/intro.tex
remains from pushdown parameter article added in

--- 2022-10-14 bibliography/succ.tex
@@
-\documentclass[times,envcountsame]{llncs}
-\usepackage{a4wide}
-\usepackage{times}
-\usepackage{amsmath,amssymb,eufrak,theorem}%mnsymbol}
-\usepackage{enumerate}
-\usepackage{complexity}
-\usepackage{gastex}
-
-\usepackage{xspace,tabularx,multicol}
-
-%\usepackage[thmmarks,standard,thref]{ntheorem}
-
-\iffalse
-
-\newtheorem{theorem}{{\bf Theorem}}[section]
-\newtheorem{corollary}[theorem]{{\bf Corollary}}
-\newtheorem{definition}[theorem]{{\bf Definition}}
-\newtheorem{example}[theorem]{{\bf Example}}
-\newtheorem{lemma}[theorem]{{\bf Lemma}}
-\newtheorem{proposition}[theorem]{{\bf Proposition}}
-\newtheorem{remark}[theorem]{{\bf Remark}}
-\newtheorem{fact}[theorem]{Fact}
-
-\fi
-
-\newtheorem{fact}[theorem]{Fact}
-
-\usepackage{stmaryrd,rotating,mathrsfs}
-
-%\usepackage[active]{srcltx}
-
-\iffalse
-\newcommand{\problemx}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf QUESTION: }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\newcommand{\problemy}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf OUTPUT: }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\fi
-
-\newcommand{\problemx}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf QUESTION:\ }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\newcommand{\problemy}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf OUTPUT:\ }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\renewcommand{\mod}{\text{mod }}
-\newcommand{\modulo}{\mod}
-
-%\renewcommand{\vec}{}
-
-\def\AC{{\mathrm{AC}}}
-\def\AG{{\mathsf{AG}}}
-\def\AND{{\mathrm{AND}}}
-\def\BIN{{\mathrm{BIN}}}
-\def\CRR{{\mathrm{CRR}}}
-\def\CTL{{\text{CTL}}}
-\def\div{{\mathrm{div}}}
-\def\dom{{\mathrm{dom}}}
-\def\EF{{\mathsf{EF}}}
-\def\AX{{\mathsf{AX}}}
-\def\EXPTIME{{\mathsf{EXPTIME}}}
-\def\inp{{\mathrm{in}}}
-\def\leaf{{\mathrm{leaf}}}
-\def\MAJ{{\mathrm{MAJ}}}
-\def\NC{{\mathrm{NC}}}
-\def\Omc{{\mathbb{O}}}
-\def\OR{{\mathrm{OR}}}
-\def\out{{\mathrm{out}}}
-\def\Pmc{{\mathcal{P}}}
-\def\PSPACE{{\mathsf{PSPACE}}}
-\def\TC{{\mathrm{TC}}}
-\def\U{{\mathsf{U}}}
-\def\WU{{\mathsf{WU}}}
-\def\X{{\mathsf{X}}}
-\def\EX{{\mathsf{EX}}}
-\def\E{{\mathsf{E}}}
-
-\newcommand{\prop}{\rho}
-
-\renewcommand{\L}{\mathsf{LOGSPACE}}
-
-%% \usepackage[thmmarks]{ntheorem}
-%% %%% Theorem environments
-%% \newtheorem{lemma}{Lemma}
-%% \newtheorem{theorem}{Theorem}
-%% \newtheorem{proposition}{Proposition}
-%% \newtheorem{corollary}{Corollary}
-
-%% \theoremstyle{definition}
-%% \theoremsymbol{\ensuremath{\lozenge}}
-%% \theorembodyfont{\normalfont}
-%% \newtheorem{definition}{Definition}
-
-%% \theoremstyle{nonumverplain}
-%% \theoremheaderfont{\normalfont\bf}
-%% \theorembodyfont{\normalfont}
-%% \theoremseparator{.}
-%% \theoremsymbol{\ensuremath{_\blacksquare}}
-%% \newtheorem{example}{Example}
-
-%% \theoremstyle{nonumberplain}
-%% \theoremheaderfont{\normalfont\itshape}
-%% \theorembodyfont{\normalfont}
-%% \theoremseparator{.}
-%% \theoremsymbol{\ensuremath{_\Box}}
-%% \newtheorem{proof}{Proof}
-
-% Symbols
-\newcommand{\naturals}{\ensuremath{\mathbf{N}}}
-\newcommand{\integers}{\ensuremath{\mathbf{Z}}}
-
-% Machines and their ingredients
-\newcommand{\automaton}{\ensuremath{A}}
-\newcommand{\clocations}{\ensuremath{Q}}
-\newcommand{\iclocation}{\ensuremath{q_{\mathit{in}}}}
-\newcommand{\fclocation}{\ensuremath{F}}
-\newcommand{\transitionrel}{\ensuremath{\Delta}}
-\newcommand{\transitionlabel}{\ensuremath{\lambda}}
-\newcommand{\clocationlabel}{\ensuremath{\xi}}
-\newcommand{\operations}{\ensuremath{Op}}
-\newcommand{\operation}{\ensuremath{op}}
-\newcommand{\add}[1]{\ensuremath{\mathsf{add}(#1)}}
-\newcommand{\zero}{\ensuremath{\mathsf{zero}}}
-\newcommand{\propositions}{\ensuremath{L}}
-\newcommand{\configuration}{\ensuremath{c}}
-\newcommand{\configurations}{\ensuremath{\mathcal{C}}}
-\newcommand{\run}{\ensuremath{\pi}}
-\newcommand{\trace}{\ensuremath{\tau}}
-\newcommand{\pocatuple}{\ensuremath{\langle \clocations,
-\iclocation, \fclocation, \transitionrel, \transitionlabel,
-\clocationlabel \rangle}}
-
-\newcommand{\circuit}{\ensuremath{\mathbb{C}}}
-
-\newcommand{\tm}{\ensuremath{\mathcal{M}}}
-
-%%% Names
-\newcommand{\oca}{\textsc{OCA} }
-\newcommand{\poca}{\textsc{POCA} }
-\newcommand{\Poca}{\mathbb{P}}
-\newcommand{\Soca}{\mathbb{S}}
-
-%%% Problems
-\newcommand{\stsat}{\textsc{Succinct 3-SAT} }
-\newcommand{\tsat}{\textsc{3-SAT} }
-\newcommand{\ltlpoca}{\textsc{LTL-POCA-MC} }
-
-%%% Complexity classes
-\newcommand{\nexptime}{\textsc{NExpTime} }
-\newcommand{\conexptime}{\textsc{co-NExpTime} }
-\newcommand{\pspace}{\textsc{PSpace} }
-
-%%% LTL
-\newcommand{\ltl}{\text{LTL}}
-\newcommand{\ctl}{\text{CTL}}
-%\newcommand{\until}{\ensuremath{\mathcal{~U~}}}
-\newcommand{\until}{\ensuremath{\U}}
-\newcommand{\eventually}{\ensuremath{\mathsf{F}}}
-\newcommand{\globally}{\ensuremath{\mathsf{G}}}
-\newcommand{\propalphabet}{\ensuremath{\Sigma}}
-\newcommand{\pastarred}{\ensuremath{(\mathcal{P}(\propalphabet))^{\omega}}}
-
-%%% Logic
-\newcommand{\true}{\ensuremath{\mathtt{tt}}}
-\newcommand{\false}{\ensuremath{\mathtt{ff}}}
-
-%%% Control commands
-\newcommand{\defemph}[1]{\emph{#1}}
-
-%\newtheorem{fact}{Fact}
-
-\newcommand{\LCM}{\text{LCM}}
-
-\begin{document}
-
-\title{Model Checking Succinct and Parametric One-Counter Automata}
-\author{Stefan G\"oller\inst{1} \and
-Christoph Haase\inst{2} \and
-Jo\"el Ouaknine\inst{2} \and
-James Worrell\inst{2}
-}
-
-\institute{
-Universit\"at
-Bremen, Institut f\"ur Informatik, Germany%\\
-%\email{goeller@informatik.uni-bremen.de}
-\and
-Oxford University Computing Laboratory, UK
-}
-
-\maketitle
-
-\begin{abstract}
-  We investigate the decidability and complexity of various model
-  checking problems over one-counter automata. More specifically, we
-  consider \emph{succinct} one-counter automata, in which additive
-  updates are encoded in binary, as well as \emph{parametric}
-  one-counter automata, in which additive updates may be given as
-  unspecified parameters. We fully determine the complexity of model
-  checking these automata against CTL, LTL, and modal $\mu$-calculus
-  specifications.
-\end{abstract}
-
-\section{Introduction}
-
 Counter automata, which comprise a finite-state controller together
@@
 parameters~\cite{BIL06-icalp,IJTW93-icalp}. We refer to such counter automata as
-\emph{succinct} and \emph{parametric} respectively, the former being
+\emph{succinct} and \emph{parametric} resp., the former being
 viewed as a subclass of the latter. Natural applications of such
-counter machines include the modeling of resource-bounded processes,
+counter machines include the modelling of resource-bounded processes,
 programs with lists, recursive or multi-threaded programs, and XML
 query evaluation; see, e.g.,
-\cite{ID06-tcs,BHIMV06,CR04}.
+\cite{BHIMV06,CR04,ID06-tcs}.
 
@@
 $\varphi$ is fixed) as well as the \emph{combined} complexity of this
-problem. Our main results are summarized in Table \ref{T results}.
-\begin{table}
+problem. Our main results are summarized in Table \ref{tab:results}.
+\begin{table}[t]
 \begin{center}
@@
 \end{tabular}
-\ \\[0.4cm]
-\caption{The complexity of CTL, the modal $\mu$-calculus, and LTL on SOCA and
-POCA.}{\label{T results}}
 \end{center}
+\label{tab:results}
+\caption{The complexity of CTL, the modal $\mu$-calculus, and LTL
+             on SOCA and POCA.}
 \end{table}
@@
 
-We note some interesting differences between our results and
-corresponding questions regarding finite automata. For the latter, the
-(combined) model checking problems for CTL, the $\mu$-calculus, and
-LTL are respectively known to be $\mathsf{P}$-complete, in
-$\mathsf{NP} \cap \coNP$, and $\PSPACE$-complete. Somewhat
-surprisingly, for SOCA and POCA, the complexity ordering is reversed
-and LTL becomes easier to model check than either CTL or the
-$\mu$-calculus.
+%% We note some interesting differences between our results and
+%% corresponding questions regarding finite automata. For the latter, the
+%% (combined) model checking problems for CTL, the $\mu$-calculus, and
+%% LTL are respectively known to be $\mathsf{P}$-complete, in
+%% $\mathsf{NP} \cap \coNP$, and $\PSPACE$-complete. Somewhat
+%% surprisingly, for SOCA and POCA, the complexity ordering is reversed
+%% and LTL becomes easier to model check than either CTL or the
+%% $\mu$-calculus.
 
@@
 an $\EXPSPACE$ lower bound. In a subsequent delicate stage of the
-proof, we show how to `split' the counter in order simultaneously to
+proof, we show how to partition the counter in order simultaneously to
 perform $\PSPACE$ computations in the counter and manipulate numbers
@@
 
-Our paper is organized as follows. In Section \ref{S Prelim} we
-introduce general notations.
-Section \ref{S CTL} deals with CTL and the modal $\mu$-calculus.
-LTL model checking is content of Section \ref{S LTL}.
-Finally, in Section \ref{S Concl} we give conclusions.
-
-\section{Preliminaries}{\label{S Prelim}}
-
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-\renewcommand{\O}{\mathbb{O}}
-%\renewcommand{\P}{\mathcal{P}}
-
-\newcommand{\trans}[1]{\stackrel{#1}{\longrightarrow}}
-\newcommand{\hook}[1]{\stackrel{#1}{\hookrightarrow}}
-\renewcommand{\rule}[2]{\stackrel{#2}{\stackrel{#1}{\longrightarrow}}}
-
-\renewcommand{\M}{\mathcal{M}}
-
-\newcommand{\bit}{\text{bit}}
-\renewcommand{\succ}{\text{succ}}
-
-\newcommand{\Reach}{\text{Reach}}
-\newcommand{\Moves}{\text{Moves}}
-\newcommand{\bin}{\text{bin}}
-
-By $\Z$ we denote the {\em integers} and by $\N=\{0,1,2,\ldots\}$ the
-denote the {\em naturals}.  For each $i,j\in\Z$ we define
-$[i,j]=\{k\in\Z\mid i\leq k\leq j\}$ and $[j]=[1,j]$.  For each
-$i,n\in\N$, let $\bit_i(n)$ denote the $i^{\text{th}}$ least
-significant bit of the binary representation of $n$. Hence
-$n=\sum_{i\in\N}2^i\cdot\bit_i(n)$.  By $\bin_m(n)=\bit_0(n)\cdots
-\bit_{m-1}(n)$ we denote the first $m$ least significant bits written
-from {\em left to right}. When $m$ is not important we just write
-$\bin(n)$.
-%For any non-empty subset $M\subseteq\Z^+$, define $\gcd(M)$ to be the largest
-%$k\geq 1$ that divides every element of $M$.
-%We define $\gcd(\emptyset)=0$.
-Let $p_i$ denote the $i^{\text{th}}$ prime number for each $i\geq 1$, i.e. $p_1=2$,
-$p_2=3$ and so on. We define $\log(n)=\min\{i\geq 1\mid 2^i> n\}$, in other
-words $\log(n)$ denotes the number of bits that are needed to represent $n$ in
-binary.
-All polynomials $p:\N\rightarrow\N$ that occur in this paper are assumed to satisfy
-$p(n)\geq n$ for each $n\in\N$.
-For each word $v=a_1\cdots a_n\in\Sigma^n$ over some finite alphabet $\Sigma$
-and each $i,j\in[n]$ define $v[i,j]=a_i\cdots a_j$ and
-$v(i)=v[i,i]$.
-
-\bigskip
-
-\renewcommand{\l}{\triangleright}
-\renewcommand{\r}{\triangleleft}
-\newcommand{\move}{\vdash}
-\noindent
-{\bf Turing machines and complexity theory:} In the following, we
-introduce $f(n)$ space-bounded deterministic Turing machines (DTMs) in
-a suitable way for proving lower bounds. These contain precisely one
-input tape and one working tape.  In our setting, the working alphabet is
- assumed to be $\{0,1,\l,\r\}$,
-where $\l$ is the {\em left marker}
-and $\r$ is the {\em right marker}.  The working tape of the initial
-configuration of such a DTM on an input $w\in\Sigma^n$ is
-assumed to be $\l0^{f(n)}\r$, whereas its input tape is $\l w\r$.
-Before we define DTMs, let $\Upsilon=\{-1,0,+1\}$ denote a set of
-directions.
-
-Formally, a {\em $f(n)$-space bounded deterministic Turing machine (DTM)} is a tuple
-$\M=(S,\Sigma,s_0,F,\mu)$, where $S$ is a finite set of {\em states},
-$\Sigma$ is a {\em finite input alphabet} with $\l,\r\not\in\Sigma$,
-$s_0\in S$ is an {\em initial state}, $F\subseteq S$ is a set of {\em final
-states}, and with $\Sigma_\M=\Sigma\uplus\{\l,\r\}$ we have that
-$$\mu\quad:\quad S\times\Sigma_\M\times\{0,1,\l,\r\}\ \rightarrow\
-S\times\Upsilon^2\times\{0,1,\l,\r\}$$
-is the {\em transition function}, where $\mu(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$ means
-that $\M$ is currently in state $s$, its input head reads $b_1$, its
-working tape reads $b_2$, and $\M$ changes to state $s'$, moves its
-input head in direction $\delta_1$,
-moves its working head in direction $\delta_2$, and
-writes the bit $b$.
-%We require
-%moreover that whenever $\M$'s input (resp. working) head reads $\l$, then $\M$
-%must move its input (resp. working) head to the right, i.e. $+1$, and whenever
-%$\M$'s input (resp.
-%working) head reads $\r$, %it must hold that
-%$\M$ moves its
-%input (resp. working) head to the left, i.e. $-1$. Formally,
-%whenever $\delta(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$, then
-%for each $i\in\{1,2\}$ we have that $b_i=\l$ implies $\delta_i=+1$ and $b_i=\r$
-%implies $\delta_i=-1$.
-Moreover, we require that $\M$ writes a marker precisely when it reads a marker,
-more formally: $b_2=m$ if and only if $b=m$ for each $m\in\{\l,\r\}$.
-As expected, a {\em configuration of $\M$} is a tuple $(s,\l w\r,\l w'\r,i,j)$,
-where $s\in S$ is the current state,
-$w\in\Sigma^n$ is an input,
-$w'\in\{0,1\}^{f(n)}$ is the content of the working tape of $\M$,
-$i\in[0,n+1]$ is the current input head position, and
-$j\in[0,f(n)+1]$ is the current working head position.
-%For configurations $C_1$ and $C_2$, we write $C_1\move_\M C_2$ whenever $\M$ can
-%move from $C_1$ to $C_2$ on one step.
-We define the {\em language $L(\M)$ of $\M$} to consist of all words
-$w\in\Sigma^*$ such that
-the {\em initial configuration}
-$(s_0,\l w\r,\l 0^{f(|w|)}\r,0,0)$ reaches a configuration
-$(s_0,\l w\r,\l w'\r,i,j)$ satisfying $s\in F$.
-
-By $\L$, $\PSPACE$, and $\EXPSPACE$ we denote the class of all
-problems that can be decided by DTM that
-is logarithmically, polynomially, exponentially space bounded,
-respectively. Recall that $\Sigma^0_1$ (resp. $\Pi^0_1$) is the class of
-all languages that are (resp. whose complements are) recursively enumerable.
-\newcommand{\Prop}{\mathcal{P}}
-\newcommand{\Op}{\text{Op}}
-%\newcommand{\add}{\mathsf{add}}
-\renewcommand{\A}{\mathbb{A}}
-%\newcommand{\EF}{\mathsf{EF}}
-%\newcommand{\CTL}{\mathsh{CTL}}
-\bigskip
-
-\noindent
-\textbf{Transition systems: }
-In the following, we fix a countable set of atomic propositions $\Prop$.
-A {\em transition system} is a tuple $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, where
-$S$ is a set of {\em states}, $S_\prop\subseteq S$ for each $\prop\in\Prop$ and
-$S_\prop$ is
-non-empty for finitely many $\prop\in\Prop$, and finally
-$\rightarrow \subseteq S\times S$ is a set of {\em transitions}.
-We prefer to use the infix notation $s_1\rightarrow s_2$ to abbreviate
-$(s_1,s_2)\in\rightarrow$.
-%A {\em finite path} of length $n$ is sequence of transitions
-%$\pi=s_0\rightarrow s_1\cdots \rightarrow s_n$, where $n\in\N $.
- An {\em infinite path} is an infinite sequence
-$\pi=s_0\rightarrow s_1\cdots$.
-% We denote
-%by $|\pi|$ the length of the path $\pi$, which we define to be $\omega$
-%if $\pi$ is infinite.
-For each infinite path
-$\pi= s_0\rightarrow s_1 \rightarrow \cdots$ and
-$i\in\N$, we denote by $\pi^i$ the path $s_i\rightarrow s_{i+1}\cdots$
-and by $\pi(i)$ the state $s_i$.
-Define the {\em trace of $\pi$} as $\trace(\pi):\N\rightarrow2^\Prop$ where
-$\trace(i)=\{\prop\in\Prop\mid s_i\in S_\prop\}$ for each $i\in\N$.
-\bigskip
-
-\noindent
-\textbf{Succinct and parametric one-counter automata: }
-A {\em succinct one-counter automaton (SOCA)} is a tuple
- $\Soca=(Q,\{Q_\prop\mid \prop\in\Prop\},E,\lambda)$, where $Q$ is a finite set
-of {\em control states}, $Q_\prop\subseteq Q$ for each $\prop\in\Prop$ and
-$Q_\rho$ is non-empty for finitely many $\prop\in\Prop$,
-$E\subseteq Q\times Q$ is a finite set of {\em transitions}, and
-$\lambda:E\rightarrow \Z\cup\{\zero\}$ labels the edges with decrements,
-increments, and zero tests.
-A {\em parametric one-counter automaton (POCA)} is a tuple
-$\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, where the first
-three components are same as for a SOCA,
-where $X$ is a finite set of {\em parameters over the natural numbers},
-and where $\lambda:E\rightarrow
-(\Z\cup\{\zero\}\cup\{\circ x\mid\circ\in\{+,-\}, x\in X\})$.
-For each assignment $\sigma:X\rightarrow\N$  the
-{\em induced SOCA} is defined as $\Poca^\sigma=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda')$
-where $\lambda'(e)=\circ\sigma(x)$ whenever $\lambda(e)=\circ x$
-for some $\circ\in\{+,-\}$ and $\lambda'(e)=\lambda(e)$ otherwise.
-If $X=\{x\}$ is a singleton, then we also write $\Poca(x)$ for $\Poca(X)$.
-The {\em size} of a POCA is defined as
-$|\Poca|=|Q|+|X|+|E|\cdot\max\{\log(|a|)\mid a\in\lambda(E)\cap\Z\}$.
-Hence, we represent each appearing
-integer in binary.
-The size of a SOCA is defined analogously.
-A SOCA $\Soca=(Q,\{Q_\prop\mid \prop\in\Prop\},E,\lambda)$ describes a transition system $T(\Soca)=(Q\times\N,\{Q_\prop\times\N\mid
-\prop\in\Prop\}, \rightarrow)$, where for each $q_1,q_2\in Q$ and each
-$n_1,n_2\in\N$ we have $q_1(n_1)\rightarrow q_2(n_2)$ if and only
-if either $\lambda(q_1,q_2)=n_2-n_1$, or
-($n_1=n_2=0$ and $\lambda(q_1,q_2)=\zero)$.
-
-\section{CTL Model Checking}{\label{S CTL}}
-
-In Section \ref{S CTL syn} we introduce the syntax and semantics of $\ctl$.
-Since the upper bounds for the modal $\mu$-calculus follow from
-known results and since the lower bounds follow from CTL, we do not introduce
-the modal $\mu$-calculus formally.
-In Section \ref{S Upper} we state an $\EXPSPACE$ upper bound for the combined
-complexity of $\ctl$ and the modal $\mu$-calculus on SOCA and a $\Pi^0_1$ upper bound
-on POCA. In Section \ref{S Serial} we recall basic notions from complexity theory
-such as serializability and results on Chinese remainder representation.
-In Section \ref{S CTL hardness} we prove the main result of this section, namely
-that the data complexity of $\ctl$ on SOCA is $\EXPSPACE$-hard.
-Finally, we prove a $\Pi^0_1$ lower bound for the data complexity
-of $\ctl$ on POCA in Section \ref{S CTL und}.
-
-\subsection{$\ctl$: Syntax and Semantics}{\label{S CTL syn}}
-\noindent
-Formulas $\varphi$ of $\ctl$ are given by the following
-grammar, where $\prop$ ranges over $\Prop$:
-
-$$
-\varphi\quad ::=\quad \prop\ \mid \neg\varphi\ \mid\ \varphi\wedge\varphi\ \mid\
-\EX\varphi\ \mid \E (\varphi\U\varphi)\ \mid\
-\E(\varphi\WU\varphi)
-$$
-
-\noindent
-We have the following abbreviations:
-$\varphi_1\vee\varphi_2=\neg(\neg\varphi_1\wedge\neg\varphi_2)$,
-$\varphi_1\rightarrow\varphi_2=\neg\varphi_1\vee\varphi_2$,
-$\true=\prop\vee\neg\prop$ for some atomic proposition $\prop\in\Prop$,
-and $\EF\varphi=\E(\true\U\varphi)$.
-The {\em size} of a $\ctl$ formula is inductively defined as follows:
-$|\prop|=1$ for each $\prop\in\Prop$, $|\neg\varphi|=|\EX\varphi|=|\varphi|+1$,
-$|\varphi_1\wedge\varphi_2|=|\E(\varphi_1\U\varphi_2)|=|\E(\varphi_1\WU\varphi_2)|=
-|\varphi_1|+|\varphi_2|+1$.
-Given a transition system $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, a state
-$s\in S$, and some $\CTL$ formula $\varphi$, define $(T,s)\models\varphi$ by
-induction on the structure of $\varphi$ as follows:
-
-\begin{eqnarray*}
-  (T,s)\models \prop &  \quad\iff\quad & s\in S_p\quad \text{ for each
-}\prop\in\Prop\\
-  (T,s)\models \varphi_1\wedge\varphi_2 &  \quad\iff\quad &
-(T,s)\models\varphi_1\text{ and }
-(T,s)\models \varphi_2 \\
-  (T,s)\models \neg\varphi &  \quad\iff\quad & (T,s)\not\models\varphi\\
-  (T,s)\models \EX\varphi &  \quad\iff\quad & (T,t)\models\varphi
-\text{ for some } t\in S \text{ with }s\rightarrow t\\
-  (T,s)\models \E(\varphi_1\U\varphi_2) &  \quad\iff\quad &
-\exists s_0,\cdots,s_n\in S, n\geq 0: s_0=s, (T,s_n)\models\varphi_2 \text{ such
-that}\\
-&& \forall i\in[0,n-1]: (T,s_i)\models\varphi_1\text{ and } s_i\rightarrow
-s_{i+1}\\
-  (T,s)\models \E(\varphi_1\WU\varphi_2) &  \quad\iff\quad &
-(T,s)\models\E(\varphi_1\U\varphi_2)\ \text{ or }\\
-&& \exists s_0,s_1,\ldots\in S:\forall i\geq 0: (T,s_{i})\models\varphi_1\text{ and }
-s_i\rightarrow s_{i+1}
-\end{eqnarray*}
-
-\noindent
-Let us define the {\em $\ctl$ model checking problem} on SOCA and POCA
-respectively.\\
-
-\problemx{$\ctl$ model checking on SOCA}
-{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
-$\varphi$.}
-{$(T(\Soca),q(n))\models\varphi$?}
-\
-
-\problemx{$\ctl$ model checking on POCA}
-{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
-$\varphi$.}
-{$(T(\Poca^\sigma),q(n))\models\varphi$ for every $\sigma:X\rightarrow\N$?}
-
-\subsection{Upper bounds}{\label{S Upper}}
-Let us first state upper bounds on model
-checking CTL and the modal $\mu$-calculus on SOCA.
- {\em One-counter automata (OCA)} are SOCA in which
- the numbers that occur in the transition labels are represented  in unary
- instead in binary.
-%In this section we state the upper bounds on model checking $\ctl$ on SOCA and
-%POCA. They follow from known results or are immediate.
-The following theorem gives an upper bound on model checking the
-{\em modal $\mu$-calculus} on OCA.
-Consult \cite{ArnNiw01} for more details on the modal $\mu$-calculus.
-
-\begin{theorem}[\cite{Serr06}]
-The combined complexity of the modal $\mu$-calculus on OCA is in $\PSPACE$.
-\end{theorem}
-
-\noindent
-Since every SOCA can be transformed into an exponentially larger OCA
- and since each $\ctl$ formula can be translated into an alternation-free
-$\mu$-calculus formula with a linear blowup, the following corollary is
-immediate.
-
-\begin{corollary}{\label{C CTL up}}
-The combined complexity of $\ctl$ and the modal $\mu$-calculus on SOCA is in $\EXPSPACE$.
-\end{corollary}
-
-\noindent
-The following upper bound for $\ctl$ on POCA is straightforward.
-
-\begin{corollary}
-The combined complexity of $\ctl$ and the modal $\mu$-calculus on POCA is in $\Pi^0_1$.
-\end{corollary}
-\begin{proof}
-Let $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$ be a POCA,
- $q\in Q$, and $n\in\N$, and $\varphi$ be
-a formula.
-From Corollary \ref{C CTL up} it follows that
-for each given $\sigma:X\rightarrow\N$ the question
-$
-(T(\Poca^\sigma),q(n))\models\varphi
-$
-is a decidable predicate.
-Hence, the question if
-$
-(T(\Poca^\sigma),q(n))\models\varphi
-$
-for each $\sigma:X\rightarrow\N$ is a $\Pi^0_1$-predicate since
-we can encode the set of all such assignments into the naturals
-via G\"odel encoding.
-\qed
-\end{proof}
-
-\subsection{Serializability and Chinese remainder representation}{\label{S Serial}}
-For a language $L\subseteq\Sigma^*$ let $\chi_L:\Sigma^*\rightarrow\{0,1\}$
-denote the {\em characteristic function of $L$}, i.e.
-$\chi_L(x)=1$ if $x\in L$ and $\chi_L(x)=0$ otherwise, for each $x\in\Sigma^*$.
-We define the {\em lexicographic
-order on $n$-bit strings} as $\preceq_n\subseteq\{0,1\}^n\times\{0,1\}^n$, where
-$x\preceq y$ if and only if
-$\bin_n(x)\leq\bin_n(y)$, e.g. $101\preceq_3 011$.
-
-\renewcommand{\C}{\mathcal{C}}
-\begin{definition}
-Let $\C$ be a complexity class.
-We say a language $L$ is {\em $\C$-serializable via some language
-$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$, and some language
-$U\in\C$ such that for all $x\in\{0,1\}^n$
-$$x\in L\qquad\!\!\iff\qquad\!\!
-\chi_U\left(x,0^{p(n)}\right)\quad \cdots\quad \chi_U\left(x,1^{p(n)}\right)\in
-R,$$ where with `$\cdots$'  we
-refer to $\preceq_{p(n)}$ in each of the second components.
-\end{definition}
-
-\noindent
-% it was shown that $\PSPACE$ is $\mathsf{P}$-serializable;
-%in \cite{HLSVW93} this result was sharpened to $\AC^0$-serializability,
-%see also \cite{Vol98}. It is not stated in \cite{HLSVW93,Vol98} but easy to
-%see from the proofs that \emph{logspace-uniform} $\AC^0$ suffices for serializing
-%$\PSPACE$, see the appendix for more details.
-The following theorem was proven in \cite{GoLo09}.
-
-\begin{theorem}[Theorem 22 in \cite{GoLo09}]{\label{T Serial}}
-For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
-logspace-uniformly $\mathsf{AC}^0$-serializable via $R$.
-\end{theorem}
-
-\noindent
-As an immediate consequence, we obtain the following corollary.
-
-\begin{corollary}{\label{C Serial}}
-For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
-$\L$-serializable via $R$.
-\end{corollary}
-
-\noindent
-We remark that our definition of serializability is adopted from \cite{GoLo09}
-and differs slightly from the standard
-notion as used in \cite{CaFu91,HLSVW93,Vol98}.
-We aim at lifting Corollary \ref{C Serial} via an appropriate notion of
-serializability to $\EXPSPACE$.
-
-\begin{definition}
-Let $\C$ be some complexity class.
-A language $L$ is {\em exponentially $\C$-serializable via some language
-$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$ and some language
-$U\in\C$ such that for all $x\in\{0,1\}^n$
-$$x\in L\qquad\!\!\iff\qquad\!\!
-\chi_U\left(x,0^{2^{p(n)}}\right)\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right)\in
-R,$$ where with `$\cdots$' we
-refer to $\preceq_{2^{p(n)}}$ in each of the second components.
-\end{definition}
-
-\noindent
-The following proposition is folklore and immediate.
-\begin{proposition}{\label{P Padding}}
-Let $L\subseteq\Sigma^*$ be in $\EXPSPACE$. Then there is some polynomial $q$
-such that the padded language
-$\{x\$^{2^{q(|x|)}}\mid x\in L\}\subseteq(\Sigma\uplus\{\$\})^*$ is in $\PSPACE$.
-\end{proposition}
-
-\begin{theorem}\label{T succ seri}
-For every language $L$ in $\EXPSPACE$ there is some regular language $R$ such
-that $L$ is exponentially $\L$-serializable via $R$.
-\end{theorem}
-\begin{proof}
-Let $L\subseteq\Sigma^*$ be some language in $\EXPSPACE$.
-Then by Proposition \ref{P Padding} for some polynomial $q$ the language
-$L'=\{x\$^{2^{q(|x|)}}\mid x\in L\}$ is in $\PSPACE$.
-Due to $\L$-serializability of $\PSPACE$ by Corollary \ref{C Serial}, there exists
-some polynomial $p'$, some regular language $R'$ and some $U'\in\L$
- such that for each $y\in(\Sigma\uplus\{\$\})^*$ we have
-\begin{eqnarray}
-y\in L'\quad\iff\quad \chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
-\chi_{U'}\left(y,1^{p'(|y|)}\right)\in R', \label{E Serial}
-\end{eqnarray}
-where `$\cdots$' refers to the lexicographic order $\preceq_{p'(|y|)}$.
-To prove the lemma we have to find some regular language $R$, some
-polynomial $p$ and some $U\in\L$
-such that for all $x\in\Sigma^*$ we have
-$$
-x\in L\quad \iff\quad \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \cdots\
-\chi_U\left(x,1^{2^{p(|x|)}}\right)\in R,
-$$
-where here `$\cdots$' refers to the lexicographic order
-$\preceq_{2^{p(|x|)}}$.
-For now, we choose $p$ sufficiently fast growing,
-let us postpone this to the end of the proof.
-Let us describe the language $U$ and consider inputs $(x,w)$
- with $x\in\Sigma^n$ and $w\in\{0,1\}^{2^{p(n)}}$;
-moreover let $y=x\$^{2^{q(|x|)}}$ be the padding of $x$.
-Our language $U$ consists of all such pairs $(x,w)$
-such that
-$w$ can be factorized as $w=bz0^l$ for some $b\in\{0,1\}$
-and some $l\geq 0$ such that
-$b=0$ or $(y,z)\in U'$.
-Roughly speaking, the goal of $U$ is to simulate $U'$
-on the corresponding padded word,
-however the length of the second component
-$w$ has to be a power of two, in other words the last $l$ bits of $w$
- contain redundant information.
-We will be able to filter out this redundant information via
-the bit $b$.
-Let us make this more precise.
-Let
-\begin{eqnarray*}
-\gamma' &\quad=\quad&
-\chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
-\chi_{U'}\left(y,1^{p'(|y|)}\right)\quad\text{and}\\
-\gamma &\quad=\quad& \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \ \cdots\ \
-\chi_U\left(x,1^{2^{p(|x|)}}\right).
-\end{eqnarray*}
-
-\noindent
-Then by definition of $U$, the first $2\cdot p'(|y|)$ bits of $\gamma$ are
-$1\gamma'(1)1\gamma'(2)\cdots 1\gamma'(p'(|y|))$, where
-the remaining bits of $\gamma$ are all $0$.
-Hence, when reading $\gamma$ in {\em pairs of bits},
-precisely when the first of the two bits
-is set to $1$ we read some relevant information.
-
-We have to give some regular language $R$ such that $\gamma'\in R'$ if and only
-if $\gamma\in R$.
-Recall that regular languages are closed under shuffle product $|\!|$ and
-homomorphisms.
-We define $R=\varphi(R'|\!|\{a\}^*)$, where
-$\varphi:\{0,1,a\}\rightarrow\{0,1\}^*$ is the following homomorphism:
-$\varphi(a)=00$, $\varphi(0)=10$, and $\varphi(1)=11$.
-
-As expected, we choose $p$ to be an arbitrary polynomial satisfying
-$2^{p(n)}\geq 1+p'(2^{q(n)}+n)$ for all $n\geq0$.
-Finally, let us show that $U$ is in $\L$.
-Let $(x,w)$ be an input to $U$ and let again $y=x\$^{2^{q(|x|)}}$ be the padding
-of $x$.
-Then it is straightforward to decide in logarithmic space if $w$ can be factorized as
-$w=bzv$ such that $b\in\{0,1\}$, $|z|=p'(2^{q(|x|)}+|x|)$ and
-$v\in\{0\}^*$.
-The only thing that might remain to be verified in logarithmic space is whether
-$(y,z)\in U'$, which in turn boils down to simulating some logarithmic space bounded TM
-for $U$ on input $(y,z)$.
-But this is possible since $|y|+|z|\leq 2\cdot(|x|+|w|)$ as
-$$
-|y|+|z|=|x|+2^{q(|x|)}+p'(2^{q(|x|)+|x|})\leq 2\cdot p'(2^{q(|x|)}+|x|)\leq
-2\cdot 2^{p(|x|)}\leq 2\cdot (|x|+|w|).
-$$
-\qed
-\end{proof}
-
-\renewcommand{\M}{\mathcal{M}}
-
-\noindent
-\textbf{Chinese remainder representation: }
- For every $m,M\in\N$ we denote by $\CRR_m(M)$ {\em the Chinese remainder
-representation} of $M$ as the Boolean tuple
-$(b_{i,c})_{i\in[m],0\leq c<p_i}$, where $b_{i,c}=1$ if $M\modulo p_i=c$
-and $b_{i,c}=0$ otherwise.
-The following theorem tells us that in logarithmic space we can
-compute the binary representation of a natural number that is given in Chinese
-remainder representation.
-It is a consequence of the result that division is in logspace-uniform
-$\mathsf{NC}^1$, proven in \cite{ChDaLi01}.
-
-\begin{theorem}[\cite{ChDaLi01} Theorem 3.3]{\label{T CRRBIN}}
-The following problem is in $\L$:
-
-INPUT: $\CRR_m(M),j\in[m]$, $b\in\{0,1\}$.
-
-QUESTION: $\bit_j(M\ \modulo 2^m)=b$?
-\end{theorem}
-
-\subsection{$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA}{\label{S CTL hardness}}
-
-In the rest of this section, we give the proof of $\EXPSPACE$-hardness
-of the data complexity of $\ctl$ on SOCA.  Let $L\subseteq\{0,1\}^*$ be an arbitrary
-language in $\EXPSPACE$. Then by Theorem \ref{T succ seri}, there is
-some regular language $R\subseteq\{0,1\}^*$ such that $L$ is
-exponentially $\L$-serializable via $R$. Hence there is some language $U\in\L$
-and some polynomial $p$ such that for all $x\in\{0,1\}^n$ we have
-
-\begin{eqnarray}{\label{E EXPSPACE char}}
-x\in L\qquad\iff\qquad \chi_{U}\left(x,0^{2^{p(n)}}\right)
-\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right) \in R,
-\end{eqnarray}
-where '$\cdots$' refers to the lexicographic order $\preceq_{2^{p(n)}}$
-on the bit strings on the right-hand side.
-
-\noindent
-For the rest of this section, let us fix an input $x_0\in\{0,1\}^n$.
-Let $N=p(n)$ and let $A=(Q,\{0,1\},q_0,\delta,F)$ be some
-deterministic finite automaton with $L(A)=R$.  Let us briefly recall
-what $A$ consists of: $Q$ is a finite set of {\em states}, $\{0,1\}$
-is the {\em input alphabet}, $q_0\in Q$ is the {\em initial state},
-$\delta:Q\times\{0,1\}\rightarrow Q$ is the {\em transition function},
-and $F\subseteq Q$ is the set of {\em final states}.  Let us describe
-equivalence (\ref{E EXPSPACE char}) differently: We have $x_0\in L$ if
-and only if the program in Figure \ref{F U} returns $\mathtt{true}$.
-
-\begin{figure}
-\normalsize
-
-\ \hspace{5cm} $q\in Q$; $q:=q_0$;\\[-0.4cm]
-
-\ \hspace{5cm} $d\in\N$; $d:=0$;
-
-\ \hspace{5cm} $b\in\{0,1\}$;\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{while }\  d\not=2^{2^N}$\ $\mathtt{ loop}$\\[-0.4cm]
-
-\ \hspace{6cm} $b:=\chi_U(x_0,\text{bin}_{2^N}(d))$;\\[-0.4cm]
-
-\ \hspace{6cm} $q:=\delta(q,b)$;\\[-0.4cm]
-
-\ \hspace{6cm} $d:=d+1$;\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{end loop}$\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{return}\ q\in F$;%\\[0.4cm]
-
-\caption{A program that returns $\mathtt{true}$ if and only if $x_0\in L$}{\label{F U}}
-\end{figure}
-
-\noindent
-Our goal is to mimic the execution of the program in Figure \ref{F U} by
-a fixed $\ctl$ formula and by a SOCA that depends on
-our input $x_0$.
-Before we start with the reduction, let us discuss the obstacles
-that arise:
-
-\begin{enumerate}[{\bf(A)}]
-\item We need some way of storing $d$ on the counter. Of course
-there are a lot of ways to do this, but since we want to access
-all bits of $d$ in the assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$, the most
-natural way is probably to represent $d$ in binary.
-However, for this $2^N$ bits are required.
- More problematically, we need to be able to check if $d$ is equal to
-$2^{2^N}$. This cannot be achieved by a transition in a SOCA
- that subtracts $2^{2^N}$, since the representation of this number
-requires exponentially many bits in $n$.
-\item As in \cite{GoLo10}, a solution to obstacle {\bf (A)} is to store $d$ in
-Chinese remainder representation with the first $2^N$ prime numbers.
-A polynomial number of bits (in $n$) suffice
-to represent each of the occurring prime numbers, but
-we need exponentially many of them.
-Thus, we cannot equip a polynomial size SOCA with transitions for
-each prime number, simply because there are too many of them.
-\item
-The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ implies that we need to simulate on
-the counter a logarithmically space bounded DTM for the language $U$
-on an exponentially
-large input (in $n$).
-Speaking in terms of the input size $n$, this means that we need to provide
-polynomially many bits on the counter that can be used to describe the
-working tape for this DTM. However, we need to provide
-some on-the-fly mechanism for reading the input.
-\end{enumerate}
-
-\iffalse
-
-\begin{enumerate}[(1)]
-\item For the representation of the variable $d$ we need $2^N+1$ bits
-  on our counter. More challenging, we need to check if $d$ is
-  equal to $2^{2^N}$. To check this, we cannot simply allow a
-  transition in our counter automaton that subtracts $2^{2^N}$,
-  since the representation of this number requires exponentially many
-  bits in $n$.
-\item As in \cite{GoLo10}, a solution to problem (1) is to store $d$
-  in Chinese remainder representation with the first $2^N+1$ primes.
-  Polynomially (in $n$) many bits suffice to represent each of the
-  occurring prime numbers, but we need exponentially many different
-  prime numbers, where each of the primes has up to exponentially many
-  residue classes that we cannot simply write down. So we cannot equip
-  the SOCA that we have to derive with transitions for each prime
-  number, simply since there are too many of them.
-
-\item The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ means that we need to simulate on
-our counter a logarithmically space bounded DTM for $U$ on an exponential
-sized input (in $n$).
-Speaking in terms of our input size $n$, this means that we need to provide
-polynomially many bits on our counter that can be used to describe the
-working tape for this DTM. However, we need to provide
-some on-the-fly mechanism for reading the input.
-\end{enumerate}
-
-\fi
-\noindent
-To achieve this, let us give a high-level description of how we proceed.
-In a first step, we carefully design a data structure on the counter
-and describe the intuition behind it.
-In a second step, we give five queries which we aim at implementing
-via fixed CTL formulas and by SOCA that can be computed from $x_0$
-in logarithmic space.
-
-\bigskip
-
-\noindent
-{\bf The data structure and how to access it: }
-Let $K=n+2^N+1$ denote the number of bits that are required to store an input
-for $U$ plus one.
-Let $\alpha=\log K$ denote the number of bits that we require for
-storing a pointer to an input for $U$ and let $\beta$
-be the number of bits that suffice for storing the
-$K^{\text{th}}$ prime. Hence $\alpha=O(N)$ and by
-the Prime Number Theorem, it follows that
-$\beta\in O(\log(K\log(K)))=O(N)$.
-The number $\alpha$ and such a sufficiently large number $\beta$ can be computed
- from $x_0$ in logarithmic space.
-
-Let us describe how we will interpret the counter in our reduction.
-Assume that the counter value is $v\in\N$.  Instead of
-treating $v$ as a natural number, we are interested only in the $l$
-least significant bits $V$ of the binary representation of $v$, where $l$
-is some number that is exponentially bounded in $n$; the precise value
-of $l$ will be made clear below.  Assume
-$V=\bit_0(v)\cdots\bit_{l-1}(v)$.  We read $V$ to be factorized into
-blocks of bits
-$$\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad V\quad=\quad I\ M\ C\ J\ X\ Y\ Z\ B,\qquad\qquad\qquad\qquad\qquad\qquad(\bigstar)$$ where
-\begin{itemize}
-\item $I\in\{0,1\}^{\alpha}$ represents a prime number index,
-\item $M\in\{0,1\}^\beta$ represents $I^{\text{th}}$
-prime number $p_I$,
-\item $C\in\{0,1\}^\beta$ represents some residue class modulo $M$,
-\item $J\in\{0,1\}^\alpha$ represents a pointer to some bit of $B$,
-\item
-each $X,Y$ and $Z$ consist of polynomially many bits (in $n$) and
-represent the working tape of three space bounded
-DTMs that we will comment on later in more detail, and
-\item $B\in\{0,1\}^{n+2^N+1}$ with $B=xB'$ for some
-$x\in\{0,1\}^n$ and some
-$B'\in\{0,1\}^{2^N+1}$,
-$B$ represents the current
-input for $U$, where
-$x$ is reserved to represent our input $x_0$
-and
-where $B'$ represents the counter $d$ from program in Figure \ref{F U} from above.
-The block $B'$ consists of $2^N+1$ bits since we want to be able to test
- if $d=2^{2^{N}}$.
-\end{itemize}
-
-\renewcommand{\Re}{\Rightarrow}
-\newcommand{\Li}{\Leftarrow}
-\newcommand{\re}{\rightarrow}
-\newcommand{\li}{\leftarrow}
-
-\noindent
-Let us introduce some more notation for addressing
-leftmost (starting to count from $0$) and rightmost bit
-positions in each of the above sequences of bits in $V$.
-For each such sequence $\Theta$ let $\Theta_{\Li}$ and $\Theta_{\Re}$ denote the
-respective positions of the leftmost and rightmost bits of $\Theta$ within the bit
-string $V$, e.g. $I_\Re=\alpha-1$ and $C_\Li=\alpha+\beta$.
-
-\medskip
-
-\noindent
-{\bf Important remark: } Throughout the rest of this section $v$ will
-denote an arbitrary natural number, moreover $I,M,C,J,X,Y,Z$, and $B$
-will implicitly be coupled with $v$.  Note that all of the bit strings
-have polynomial length in $n$ except for $B$.  Moreover, we identify each
-of the blocks with the natural number they represent.
-
-\medskip
-
-\noindent
-A very simple but important gadget that we need is to decide, for each
-$b\in\{0,1\}$ if the $i^{\text{th}}$ bit of $v$ is $b$.
-
-\begin{lemma}{\label{L bit}}
-For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\bit,b}$
-such that the following is computable in logarithmic space:
-
-INPUT: $i\in\N$ in unary.% and our input $x_0$.
-
-OUTPUT: A SOCA $\Soca_{\bit,i}$ and some control state $q_{\bit,i}$ such that
-$$
-(T(\Soca_{\bit,i}),q_{\bit,i}(v))\models\varphi_{\bit,b}\quad\Longleftrightarrow\quad
-\bit_i(v)=b.
-$$
-\end{lemma}
-\begin{proof}
-The SOCA $\Soca_{\bit,i}$ has the two atomic propositions $\tau_1$ and $\tau_2$ and is
-depicted below:
-
-\begin{center}
-\begin{picture}(85,20)(-20,-5)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-\put(13,-2){\small $\tau_1$}
-\put(41,-2){\small $\tau_2$}
-  \node(A)(20,0){$q_{\bit,i}$}
-  \node(B)(40,0){$\bullet$}
-\drawloop(A){$-2^{i+1}$}
-\drawedge(A,B){$-2^i$}
-\end{picture}
-\end{center}
-The simple way to check if $\bit_i(v)=1$ (resp. $\bit_i(v)=0$) is to
-repeatedly subtract $2^{i+1}$ from $v$ until no longer possible
-and then to check if $2^i$ can (resp. cannot) be subtracted.
-Hence we put
-\begin{eqnarray*}
-\varphi_{\bit,1}&\quad=\quad&
-\tau_1\wedge
-\EF(\tau_1\wedge\neg\EX \tau_1\wedge\EX
-\tau_2)\qquad\text{ and}\\
-\varphi_{\bit,0}&=&
-\tau_1\wedge
-\EF(\tau_1\wedge\neg\EX \tau_1\wedge\neg\EX
-\tau_2).
-\end{eqnarray*}
-\qed
-\end{proof}
-
-\noindent
-{\bf Queries that we need to implement: }
-We will implement the following five queries Q1 to Q5 by instances
-of the model checking problem, where each query builds on top
-of its preceding queries.
-
-\begin{enumerate}[(Q1)]
-\item \fbox{When assuming $C<M$, does $B\equiv C\ \mod M$ hold?}
-\item \fbox{Is $M$ the $I^{\text{th}}$ prime number, i.e. $M=p_I$?}
-%\item Let $\CRR_{2^\alpha}(B)=(b_{i,r})_{i\in[2^\alpha],0\leq r<p_i}$.
-%What is $b_{I,R}$? %resp. $b_{I,R}=1$ hold?
-\item \fbox{What is $\bit_J(B)$?}
-\item \fbox{Does $(B[1,n],B[n+1,n+2^N])\in U$ hold?}
-%\item Does $B'=2^{2^N}$ hold?
-\item \fbox{Does $x_0\in L$ hold?}
-\end{enumerate}
-
-\noindent
-We implement each of the five queries by providing fixed CTL formulas
-and SOCA that can be computed from $x_0$ in logarithmic space.
-$\EXPSPACE$-hardness of the data complexity of CTL on SOCA will hence
-follow from the implementation of Q5.
-First, let us give an implementation of Q1.
-
-\begin{lemma}{\label{L mod}} There exists some fixed $\ctl$ formula $\varphi_\mod$ such that we
-can compute from $x_0$ in logarithmic space some SOCA $\Soca_\mod$ and some
-control state $q_\mod$ such that
-$(T(\Soca_{\mod}),q_\mod(v))\models\varphi_\mod$ if and only if $B\equiv C\ \mod M$.
-\end{lemma}
-\begin{proof}
-The SOCA $\Soca_{\mod}$ contains the three atomic propositions $\prop_0,\prop_1$, and
-$\prop_2$ and is depicted in Figure \ref{F Smod}.
-The $\ctl$ formula $\varphi_\mod$ expresses that we
-traverse the upper
-sequence of diamonds and thereby repeatedly subtract $M$ from $B$.
-The number of diamonds both in the first and the second row equals $\beta$,
-the number of bits of $M$ and of $C$.
-In the upper row, one diamond corresponds to one bit of $M$. In case bit the rightmost bit of $M$
-(in other words bit $M_\Re$ of the counter) is $1$,
-which we can verify by a transition to the initial control state of
-the SOCA $\Soca_{\bit,M_\Re}$,
- we subtract $2^{B_\Li+\beta-1}$ from $B$, otherwise we do not modify the counter value.
-After that, we move on to the second diamond, which represents the second
-rightmost  bit of $M$, and so on.
-Hence, traversing a cycle that starts and ends
-in our initial control state $q_\mod $ will correspond to subtracting $M$ once.
- Then we traverse
-the lower sequence of diamonds and similarly subtract $C$, but this
-this time only once.
-Finally, after having traversed the lower sequence of diamonds and seeing the atomic
-proposition $\prop_2$, we check if $B=0$ by trying to subtract $2^{B_\Li}$.
-\begin{figure}
-\begin{center}
-\begin{picture}(100,120)(5,-41)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
-%\put(15,-5){$q(M_{\abs,\r})$}
-%\put(43,-5){$p_2$}
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)
-\put(10,25){\Large$\Soca_{\bit,M_\Re}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(58,20)(58,40)(35,40)
-\put(38,25){\Large$\Soca_{\bit,M_\Re-1}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,20)(108,20)(108,40)(85,40)
-\put(90,25){\Large$\Soca_{\bit,M_\Li}$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-  \node(1)(0,60){$q_{\mod}$}
-  \node(2O)(15,70){$\bullet$}
-  \node(2U)(15,50){$\bullet$}
-  \put(17,72){\small$\prop_1$}
-  \put(17,48){\small$\prop_0$}
-
-  \node(1A)(15,35){$q_{\bit,M_\Re}$}
-  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}
-
-  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}
-
-  \node(7A)(95,35){$q_{\bit,M_\Li}$}
-
-  \node(3)(30,60){$\bullet$}
-  \node(4O)(45,70){$\bullet$}
-
-  \put(47,72){\small$\prop_1$}
-  \put(47,48){\small$\prop_0$}
-  \node(4U)(45,50){$\bullet$}
-  \node(5)(60,60){$\bullet$}
-
-  \node(6)(80,60){$\bullet$}
-  \node(7O)(95,70){$\bullet$}
-  \put(97,72){\small$\prop_1$}
-  \put(97,48){\small$\prop_0$}
-
-  \node(7U)(95,50){$\bullet$}
-  \node(8)(110,60){$\bullet$}
-
-  \put(65,58){\huge$\cdots$}
-
-%\drawedge[curvedepth=-70](8,1){$0$}
-
-\drawedge(1,2O){$0$}
-\drawedge[ELside=r](1,2U){$0$}
-\drawedge[ELside=l,ELpos=60](2O,3){$-2^{B_\Li+\beta-1}$}
-\drawedge(2U,3){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](2O,1A){$0$}
-\drawedge(2U,1A){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](4O,4A){$0$}
-\drawedge(4U,4A){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](7O,7A){$0$}
-\drawedge(7U,7A){$0$}
-
-\drawedge(3,4O){$0$}
-\drawedge[ELside=r](3,4U){$0$}
-\drawedge[ELside=l,ELpos=70](4O,5){$-2^{B_\Li+\beta-2}$}
-\drawedge(4U,5){$0$}
-
-\drawedge(6,7O){$0$}
-\drawedge[ELside=r](6,7U){$0$}
-\drawedge[ELside=l,ELpos=60](7O,8){$-2^{B_\Li}$}
-\drawedge(7U,8){$0$}
-
-\drawcurve(110,60)(112,70)(55,80)(0,70)(0,62)
-
-%\drawedge[ELpos=70,curvedepth=1.5](8,9A){$0$}
-%\drawedge[ELside=r,ELpos=60](8,9B){$0$}
-  \put(60,81){$0$}
-%\drawedge[ELpos=50](8,9C){$0$}
-%\drawedge[ELside=r,ELpos=50](8,9D){$-2^{B_L}$}
-
-%%%%COPY
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,-40)(28,-40)(28,-20)(5,-20)
-\put(10,-35){\Large$\Soca_{\bit,C_\Re}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,-40)(58,-40)(58,-20)(35,-20)
-\put(38,-35){\Large$\Soca_{\bit,C_\Re-1}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,-40)(108,-40)(108,-20)(85,-20)
-\put(90,-35){\Large$\Soca_{\bit,C_\Li}$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-  \node(1C)(0,0){$\bullet$}
-  \drawedge[ELside=r](1,1C){$0$}
-  \node(2OC)(15,10){$\bullet$}
-  \node(2UC)(15,-10){$\bullet$}
-  \put(17,12){\small$\prop_1$}
-  \put(17,-12){\small$\prop_0$}
-
-  \node(1AC)(15,-25){$q_{\bit,C_\Re}$}
-  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}
-
-  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}
-
-  \node(7AC)(95,-25){$q_{\bit,C_\Li}$}
-
-  \node(3C)(30,0){$\bullet$}
-  \node(4OC)(45,10){$\bullet$}
-
-  \put(47,12){\small$\prop_1$}
-  \put(47,-12){\small$\prop_0$}
-  \node(4UC)(45,-10){$\bullet$}
-  \node(5C)(60,0){$\bullet$}
-
-  \node(6C)(80,0){$\bullet$}
-  \node(7OC)(95,10){$\bullet$}
-  \put(97,12){\small$\prop_1$}
-  \put(97,-12){\small$\prop_0$}
-
-  \node(7UC)(95,-10){$\bullet$}
-  \node(8C)(110,0){$\bullet$}
-
-  \node(9C)(110,-20){$\bullet$}
-
-\drawedge(8C,9C){$-2^{B_\Li}$}
-
-  \put(112,-2){$\prop_2$}
-
-  \put(65,-2){\huge$\cdots$}
-
-%\drawedge[curvedepth=-70](8,1){$0$}
-
-\drawedge(1C,2OC){$0$}
-\drawedge[ELside=r](1C,2UC){$0$}
-\drawedge[ELside=l,ELpos=60](2OC,3C){$-2^{B_\Li+\beta-1}$}
-\drawedge(2UC,3C){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](2OC,1AC){$0$}
-\drawedge(2UC,1AC){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](4OC,4AC){$0$}
-\drawedge(4UC,4AC){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](7OC,7AC){$0$}
-\drawedge(7UC,7AC){$0$}
-
-\drawedge(3C,4OC){$0$}
-\drawedge[ELside=r](3C,4UC){$0$}
-\drawedge[ELside=l,ELpos=70](4OC,5C){$-2^{B_\Li+\beta-2}$}
-\drawedge(4UC,5C){$0$}
-
-\drawedge(6C,7OC){$0$}
-\drawedge[ELside=r](6C,7UC){$0$}
-\drawedge[ELside=l,ELpos=60](7OC,8C){$-2^{B_\Li}$}
-\drawedge(7UC,8C){$0$}
-
-\end{picture}
-\caption{The SOCA $\Soca_\mod$}{\label{F Smod}
-for checking if $B\equiv C\ \mod M$.}
-\end{center}
-\end{figure}
-Finally, let us give the formula $\varphi_\mod$:
-$$
-\varphi_\mod\quad=\quad\E \left(\bigwedge_{i\in\{0,1\}} \prop_i
-\rightarrow \EX \varphi_{\bit,i} \right) \U (\prop_2\wedge\neg\EX\true)
-$$
-\qed
-\end{proof}
-
-\noindent
-We need the following
-proposition for implementing Q2. Its proof is straightforward.
-
-\begin{proposition}{\label{P prime}}
-The following problem is in $\PSPACE$:
-
-INPUT: $\bin(i)\$\bin(m)$
-
-QUESTION: $m=p_i$?
-\end{proposition}
-
-%\renewcommand{\prime}{\text{prime}}
-\newcommand{\prim}{\text{prime}}
-
-\begin{lemma}{\label{L prim}} There is some fixed $\ctl$ formula $\varphi_\prim$ such that
-from  $x_0$ we can compute in logarithmic space some SOCA
-$\Soca_\prim$ and some control state $q_\prim$ such that
-$(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$ %\quad\Leftrightarrow\quad
-if and only if $M=p_I$.
-\end{lemma}
-\begin{proof}
-By Proposition \ref{P prime} there is some $q(n)$ space-bounded DTM
-$\M=(S,\Sigma,s_0,F,\mu)$
-over the alphabet $\Sigma=\{0,1,\$\}$ that decides,
-given $\bin(i)\$\bin(m)$, whether $m=p_i$, where $q$ is some polynomial.
-The idea is to simulate $\M$ on input $z=I\$M$ using $X$ from $(\bigstar)$ as a working tape.
-For this, we define our block $X$ to consist of
-$l=q(|I|+|M|+1)=q(\alpha+\beta+1)$ many bits.
-Moreover, we assume w.l.o.g. that $\M$'s behaviour on input $I\$M$ is
-independent of the content of its initial working tape: This can be
-achieved by adding extra states to $\M$ that first write $0^l$ onto
-the working tape.
-Hence, the initial configuration of $\M$ on input $z$ is
-$(s_0,\l z\r,\l w \r,0,0)$ for some $w\in\{0,1\}^l$.
-Before we give our SOCA $\Soca_\prim$, we describe the computation of
-$\M$ on input $I\$M$ as a pseudo program that terminates if and only if $M=p_I$:\\
-
-\ \hspace{2cm} $s\in S$; $s:=s_0$;\quad  (current state of $\M$)\\[-0.4cm]
-
-\ \hspace{2cm} $i\in[0,\alpha+\beta+2]$; $i:=0$;\quad (current input head position)
-
-\ \hspace{2cm} $h\in[0,l+1]$; $h:=0$;\quad (current working head position)
-
-\ \hspace{2cm} $a\in\{0,1,\$,\l,\r\}$; $a:=\l$;\quad (current input symbol)
-
-\ \hspace{2cm} $b\in\{0,1,\l,\r\}$; $b:=\l$;\quad (current working tape
-symbol)\\[0.05cm]
-
-\ \hspace{2cm} $\mathtt{while }\  s\not\in F$\ $\mathtt{ loop}$\\[-0.4cm]
-
-\ \hspace{3cm} $a:=\begin{cases} \l& \text{ if }i=0\\
-\r &\text{ if }i=\alpha+\beta+2\\
-\$ &\text{ if }i=\alpha+1\\
-\bit_i(I) &\text{ if }i\in[1,\alpha]\\
-\bit_{i-\alpha-1}(M) &\text{ otherwise }
-\end{cases}$
-
-\ \hspace{3cm} $b:=\begin{cases} \l& \text{ if }h=0\\
-\r &\text{ if }h=l+1\\
-\bit_{h}(X) &\text{ otherwise }
-\end{cases}$
-
-%\\[-0.4cm]
-
-\ \hspace{3cm} Let $\mu(s,a,b)=(s',\delta_1,\delta_2,b')$.\\[-0.4cm]
-
-\ \hspace{3cm} $\mathtt{if}\ h\in[1,l]\ \ \mathtt{then}\  \ \bit_h(X):=b';\ \ \mathtt{fi}$\\[-0.4cm]
-
-\ \hspace{3cm} $s:=s '$;\\[-0.4cm]
-
-\ \hspace{3cm} $i:=i+\delta_1$;\\[-0.4cm]
-
-\ \hspace{3cm} $h:=h+\delta_2$;\\[-0.4cm]
-
-\ \hspace{2cm} $\mathtt{end loop}$\\[-0.2cm]
-
-\newcommand{\inpu}{\text{input}}
-\newcommand{\inz}{\text{input}_0}
-\newcommand{\ino}{\text{input}_1}
-\newcommand{\work}{\text{work}}
-\newcommand{\workz}{\text{work}_0}
-\newcommand{\worko}{\text{work}_1}
-
-\noindent
-Let us describe our SOCA $\Soca_\prim$.
-The control states $Q$ of $\Soca_\prim$ will contain
-$$
-S\times[0,\alpha+\beta+2]\times[0,l+2]\times
-\{0,1,\$,\l,\r\}\times\{0,1,\l,\r\}, \qquad\text{where}
-$$
-the first component corresponds to the variable $s$,
-the second component corresponds to the variable $i$,
-the third component corresponds to the variable $h$,
-the fourth component corresponds to the variable $a$, and
-finally the fifth component corresponds to the variable $b$
-of the above program.
-Moreover, our SOCA will contain
-the SOCA $\Soca_{\bit,I_\Li},\Soca_{\bit,I_\Li+1}\ldots,\Soca_{\bit,I_\Re}$ and
-$\Soca_{\bit,X_\Li},\Soca_{\bit,X_\Li+1}\ldots,\Soca_{\bit,X_\Re}$ in order to test if
-certain bits
-in $I$ and $X$ are set correctly.
-Hence $Q$ contains the control states
-$q_{\bit,I_\Li},q_{\bit,I_\Li+1}\ldots,q_{\bit,I_\Re}$ and
-$q_{\bit,X_\Li},q_{\bit,X_\Li+1}\ldots,q_{\bit,X_\Re}$ as well.
-
-\noindent
-We will provide the atomic propositions
-$\Prop=\{F,\inpu,\work,\inz,\ino,\workz,\worko\}$,
-where
-\begin{eqnarray*}
-Q_F&\ =\ &\{(s,i,h,a,b)\in Q\mid s\in F\},\\
-Q_\inpu&\ =\ &\{q_{\bit,k}\mid k\in[I_\Li,I_\Re]\},\\
-Q_\work&\ =\ &\{q_{\bit,k}\mid k\in[X_\Li,X_\Re]\},\\
-Q_{\inz}&\ =\ &\{(s,i,h,a,b)\in Q\mid a=0\},\\
-Q_{\ino}&\ =\ &\{(s,i,h,a,b)\in Q\mid a=1\},\\
-Q_{\workz}&\ =\ &\{(s,i,h,a,b)\in Q\mid b=0\},\\
-Q_{\worko}&\ =\ &\{(s,i,h,a,b)\in Q\mid b=1\}.\\
-\end{eqnarray*}
-We do not give all transitions of $\Soca_\prim$, but illustrate
-some cases by way of example.
-Let us give the outgoing transitions of the control state
-$(s,i,h,0,1)$, where $i\in[I_\Li,I_\Re]$ and $h\in[X_\Li,X_\Re]$,
-i.e. we currently scan the $i^{\text{th}}$ bit
-of $I$ and do not read a marker on our working tape.
-Moreover, let us assume $\mu(s,0,1)=(s',+1,-1,0)$, i.e. we change to state $s'$,
-we move the input head to the right, the working head to the left,
-and modify the current content on the working tape from $1$ to $0$.
-We realize the latter by subtracting $2^{X_\Li+h}$ from the counter,
- however we allow transitions
-to states $(s',i+1,h-1,a,b)$, {\em for each} $a,b\in\{0,1\}$, hence we guess
-the input tape symbol and the working tape symbol of the
-successor configuration.
-The $\ctl$ formula $\varphi_\prim$ will guarantee that our guessing was correct by
-accessing the control states $q_{\bit,k}$ where
-$k\in[I_\Li,I_\Re]\cup[M_\Li,M_\Re]$.
-
-\begin{center}
-\begin{picture}(85,73)(-20,4)
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-\put(-20,35){\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)}
-
-\put(50,35){\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)}
-
-\put(-12,68){\large$\Soca_{\bit,I_\Li+i}$}
-\node(I)(-3,60){$q_{\bit,I_\Li+i}$}
-  \put(-12,57){\tiny\bf$\inpu$}
-
-\put(58,68){\large$\Soca_{\bit,X_\Li+h}$}
-\node(X)(67,60){$q_{\bit,X_\Li+h}$}
-  \put(72,57){\tiny\bf$\work$}
-
-  \node(A)(30,40){$(s,i,h,0,1)$}
-  \put(39,40){\tiny\bf$\inz,\worko$}
-  \node(B1)(-35,10){$(s,i+1,h-1,0,0)$}
-  \put(-40,6){\tiny\bf$\inz,\workz$}
-
-  \put(0,6){\tiny\bf$\inz,\worko$}
-  \put(40,6){\tiny\bf$\ino,\workz$}
-  \put(80,6){\tiny\bf$\ino,\worko$}
-
-  \node(B2)(5,10){$(s,i+1,h-1,0,1)$}
-  \node(B3)(45,10){$(s,i+1,h-1,1,0)$}
-  \node(B4)(85,10){$(s,i+1,h-1,1,1)$}
-\drawedge[ELside=r](A,B1){$-2^{X_\Li+h}$}
-\drawedge[ELside=r,ELpos=70](A,B2){$-2^{X_\Li+h}$}
-\drawedge[ELside=r](A,B3){$-2^{X_\Li+h}$}
-\drawedge[ELside=l](A,B4){$-2^{X_\Li+h}$}
-
-\drawedge[ELside=l](A,I){$0$}
-\drawedge[ELside=l](A,X){$0$}
-
-\end{picture}
-\end{center}
-The other cases can be dealt with analogously.
-We put $q_\prim=(s_0,0,0,\l,\l)$.
-Our final formula is
-$$
-\varphi_\prim\quad=\quad\E
-\left(
-\bigwedge_{k\in\{0,1\}}
-\left(\text{input}_k\rightarrow\EX(\inpu\wedge\varphi_{\bit,k})\right)
-\wedge
-\left(\text{work}_k\rightarrow\EX(\work\wedge\varphi_{\bit,k})\right)
-\right)\U\ F.
-$$
- Then
-$(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$
-if and only if
-$I\$M\in L(\M)$ if and only if
-$M=p_I$.
-\qed
-\end{proof}
-
-\newcommand{\BIT}{\text{BIT}}
-
-\noindent
-Let us now give an implementation of query Q3.
-
-\begin{lemma}{\label{L BIT}}
-For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\BIT,b}$
-such that from $x_0$
-we can compute in logarithmic space some SOCA $\Soca_{\BIT}$ and some
-control state $q_\BIT$ such that
-$(T(\Soca_\BIT),q_\BIT(v))\models\varphi_{\BIT,b}$
-if and only if $\bit_J(B)=b$.
-\end{lemma}
-\renewcommand{\C}{\mathcal{C}}
-\renewcommand{\R}{\mathcal{R}}
-\begin{proof}[Sketch]
-By Lemma \ref{L prim}, there is a fixed $\ctl$
-formula $\varphi_\prim$ such that we can compute from $x_0$ a
-SOCA $\Soca_\prim$ that allows to check if $M=p_I$.
-Moreover, by Lemma \ref{L mod} there is a fixed $\ctl$
-formula $\varphi_\mod$ and a SOCA $\Soca_\mod$ that allows
-to check whether $B\equiv C \mod\ M$.
-In other words, we can deal with the Chinese remainder representation
-of $B$, but our goal is to access the the $J^{\text{th}}$ bit in the
-binary representation of $B$.
-So let us assume that $\R=\CRR_{K}(B)=(b_{i,c})_{1\leq i\leq K,0\leq c<p_i}$
-is the Chinese remainder representation of $B$.
-Note that we do not have $\R$ stored anywhere on the counter.
-However, the bit strings $I$ and $C$ serve as pointers to access the
-bit $b_{I,C}$ of $\R$.
-By Theorem \ref{T CRRBIN}, given $\R$ (on-the-fly by our pointers $I$ and $C$), our
-bit string $J$, and the bit $b$,
-we can decide if $\bit_J(B)=b$ in logarithmic space.
-So let $\M=(S,\Sigma,s_0,F,\mu)$ be some $k\cdot\log(m)$ space-bounded
-Turing machine machine for this, where $k\geq 1$ is some constant.
-Hence, in order to decide if
-$\bit_J(B)=b$ we need to simulate $\M$ on input $\langle \R, J, b\rangle$.
-For this we store the space that $\M$ requires on our reserved
-sequence of bits $Y$ from $(\bigstar)$, hence $Y$ consists of
-$l=k\cdot\log(|\R Jb|)=O(\beta^2+\alpha+1)=O(N^2)$ many bits.
-The definition of SOCA $\Soca_{\BIT,b}$ works analogously to the construction
-of $\Soca_{\prim}$ in the proof of Lemma \ref{L prim}
-i.e. we introduce in $\Soca_{\BIT,b}$ control states
-that remember the current bit of the input head, the current bit of
-the working head, and the position of the working head.
-The only difference is that a pointer to the input head cannot be
-completely stored
-in the control states.
-As mentioned above, for this we employ the bit blocks $I$ and $C$,
-and also the sequence $M$ for storing the $I^{\text{th}}$ prime number.
-In order to obtain bit $b_{I,C}$, we allow transitions to our SOCA $\Soca_\mod$
-(which we can compute in logarithmic
-space from $x_0$ by Lemma \ref{L mod}) and checking
-(i) via the fixed formula $\varphi_\prim$ whether $M=p_I$, and
-then (ii) via the fixed formula $\varphi_{\mod}$
-whether $B\equiv C\ \mod M$.
-Pointers to the remaining parts of the input for $\M$, namely $J$ and $b$,
-can directly be handled by the control states of $\Soca_{\BIT,b}$,
-in analogy to the proof of Lemma \ref{L prim}.
-By way of example we explain the behavior of $\Soca_{\BIT}$ when the input head
-of $\M$ currently scans bit $b_{I,C}$ of $\R$ and when getting
-to the successor configuration requires moving the input head to the left.
-Then we simply decrement $C$ by $1$ which corresponds to subtracting $2^{C_\Li}$
-from the counter. If, however, $C$ currently equals $0$, we need to
-decrement $I$ by $1$, overwrite $M$ with prime $p_{I-1}$, and finally
-overwrite $C$ with $p_{I-1}-1$. Decrementing $I$ by $1$ can simply achieved by
-subtracting $2^{I_\Li}$ from the counter. Overwriting $M$ with prime $p_{I-1}$
-can be done by repeatedly subtracting $1$ from $M$ (i.e. subtracting $2^{M_\Li}$
-from the counter) until $M$ equals $p_{I-1}$;
-checking if $M=p_{I-1}$ can be done via the fixed $\ctl$ formula $\varphi_{\prim}$
-and the SOCA $\Soca_{\prim}$ by Lemma \ref{L prim}.
-The other cases work analogously.
-\qed
-\end{proof}
-
-\noindent
-The following lemma implements query Q4.
-
-\begin{lemma}{\label{L U}}
-For each $b\in\{0,1\}$ there is some fixed $\ctl$ formula $\varphi_{U,b}$ such that
-from $x_0$
-we can compute in logarithmic space some SOCA $\Soca_U$ and some
-control state $q_U$ such that $(T(\Soca_U),q_U(v))\models\varphi_{U,b}$
-if and only if $b=\chi_U(B[1,n],B[n+1,n+2^N])$.
-\end{lemma}
-\begin{proof}[sketch]
-The proof is similar to the proof of Lemma \ref{L BIT}.
-Since $U$ is in $\L$, there is some $k\cdot\log(m)$-space bounded
-Turing machine that decides $U$, where $k\geq 1$ is some constant.
-In order to decide if $b=\chi_U(B[1,n],B[n+1,n+2^N])$ we
-need to simulate $\M$ on input $\langle B[1,n],B[n+1,n+2^N]\rangle$.
-We store the space that $\M$ requires on its working tape
-in the sequence of bits $Z$ from $(\bigstar)$, hence $Z$ consists of
-$k\cdot\log(n+2^N+1)=O(N)$ many bits.
-We use the bit sequence $J$ as a pointer for accessing bits of $B$.
-For reading the $J^{\text{th}}$ bit of $B$ we make use of the $\ctl$ formula
-$\varphi_{\BIT,b}$ and use the SOCA $\Soca_\BIT$ from Lemma \ref{L BIT}.
-The rest of the proof follows along the same lines as the proof of
-Lemma \ref{L prim}.
-\qed
-\end{proof}
-
-\noindent
-The following lemma implements query Q5 and concludes the
-$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA.
-
-\begin{lemma}
-There is some fixed $\ctl$ formula $\varphi_L$ such that from $x_0$
-we can compute in logarithmic space some SOCA $\Soca_L$ and some
-control state $q_L$ such that $(T(\Soca_L),q_L(0))\models\varphi_L$
-if and only if $x_0\in L$.
-\end{lemma}
-\begin{proof}
-First note that our $\ctl$ formula $\varphi_L$ will be evaluated in state $q_L(0)$.
-Recall that our bit sequence $B$ has length $n+2^N+1$ where $B$ is factorized
-as $B=xB'$
-for some $x\in\{0,1\}^n$ and some $B'\in\{0,1\}^{2^N+1}$.
-The SOCA $\Soca_L$ and the $\ctl$ formula $\varphi_L$ will mimic the execution
-of the program from Figure \ref{F U}.
-In the bit string $x$ we store the value $x_0$.
-The bit string $B'$ represents
-the variable $d$ of the program, hence we will initialize
-$B'$ with $0$.
-Note that incrementing $B'$ by $1$ corresponds to adding $2^{B_\Li+n}$ to the counter.
-Thus, checking when $d$ becomes $2^{2^N}$ for the first time boils down to checking
-when the $n+2^N+1^{\text{st}}$ bit of $B$ becomes $1$ for the first time.
- By Lemma
-\ref{L BIT} the $\ctl$ formula $\varphi_{\BIT,1}$ and the SOCA $\Soca_\BIT$ allow to test
-if the $J^{\text{th}}$ bit of $B$ equals $1$.
-Therefore we store in $J$ the number $n+2^N+1$.
-The following claim tells us that we can test whether we have
-initialized the counter correctly.
-Its proof is simple and therefore omitted.
-\newcommand{\init}{\text{init}}
-
-\medskip
-
-\noindent
-{\em Claim:} There is a fixed $\ctl$ formula $\varphi_\init$ such that from
-$x_0$ we can compute in logarithmic space some SOCA $\Soca_\init$ and
-some control state $q_\init$ such that
-$(T(\Soca_\init),q_\init(v))\models\varphi_\init$
-if and only if $J=n+2^N+1$ and $B=x_00^{2^N+1}$.\\
-
-\noindent
-Recall that $A=(Q,\{0,1\},q_0,\delta,F)$ is the deterministic finite automaton
-of the program in Figure \ref{F U} that needs to be simulated.
-Let us, before giving $\Soca_L$, define the auxiliary SOCA $\Soca_A$
-to be connected with the control state $q_U$ of $\Soca_U$ and with the control
-state $q_\BIT$ of $\Soca_\BIT$ along with the additional
-control states $S=Q\times\{0,1\}$ and
-the following transitions:
-$$
-(q,b)\xrightarrow{+2^{B_\Li+n}}(\delta(q,b),b'),\qquad
-(q,b)\xrightarrow{0}q_\BIT,\qquad\text{and}\qquad
-(q,b)\xrightarrow{0}q_U\qquad\text{for each }
-b,b'\in\{0,1\}.
-$$
-
-\noindent
-Moreover, $\Soca_A$ contains the atomic propositions
-$\{\bit_0,\bit_1,F\}$, where $S_{\bit_b}=Q\times\{b\}$ for each $b\in\{0,1\}$ and
-$S_F=F\times\{0,1\}$.
-Before giving $\varphi_L$, let us depict our SOCA $\Soca_L$ which has
-two additional labels $\rho$ and $\init$:
-
-\begin{center}
-\begin{picture}(85,53)(-40,-12)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-\put(-13,0){
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(58,20)(58,40)(35,40)
-\put(48.5,28){\Large$\Soca_{\init}$}
-}
-
-\put(-51,-35){
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(68,20)(68,40)(35,40)
-\put(48.5,25){\Large$\Soca_A$}
-}
-
-\put(0,23){\small $\prop$}
-\put(25,33){\small $\init$}
-  \node(A)(-30,20){$q_L$}
-  \node(B)(0,20){$\bullet$}
-  \node(C)(30,30){$q_\init$}
-  \node(D1)(-10,0){$(q_0,0)$}
-  \node(D2)(10,0){$(q_0,1)$}
-\drawloop(A){$+1$}
-\drawedge(A,B){$0$}
-\drawedge(B,C){$0$}
-\drawedge[ELside=r](B,D1){$0$}
-\drawedge(B,D2){$0$}
-\end{picture}
-\end{center}
-Let us define the auxiliary formula $\chi=\EX\varphi_\BIT$ that allows
-to test if $B'=2^{2^N}$.
-Our final formula is
-$$
-\varphi_L\quad=\quad\EF\left(\rho\wedge \EX(\init\wedge\varphi_\init)\wedge
-\E\left(\neg\chi\wedge\bigwedge_{b\in\{0,1\}}\bit_b\rightarrow
-\EX\varphi_{U,b}\right)
- \U\ (\chi\wedge F)
-\right)
-$$
-We have $x_0\in L$ if and only if
-the program from Figure \ref{F U} returns $\mathtt{true}$ if and only if
-$(T(\Soca_L),q_L(0))\models\varphi_L$.
-\qed
-\end{proof}
-
-\begin{theorem}
-The data and combined complexity of $\ctl$ and the modal $\mu$-calculus
- on SOCA is $\EXPSPACE$-complete.
-\label{thm:ctl-poca}
-\end{theorem}
-
-\iffalse
-\section{$\EF$ model checking is in $\PSPACE$}
-
-For a finite and non-empty set of integers $S\subseteq \Z\setminus\{0\}$ not
-containing $0$, we denote by {\em $\gcd(S)$} the greatest natural number that
-divides all numbers $\{|a|: a\in S\}$.
-A {\em linear equation} over $\Z$ is given by
-$$
-x\quad =\quad a_1\cdot x_1 + \cdots\ + a_k\cdot x_n,
-$$
-where $a_i\in\Z$ is a {\em coefficient} and $x_i$ is a {\em variable}
-ranging over $\N$ for each $i\in[n]$.
-We call $b\in\Z$ a {\em solution} to the above linear equation if there exists a
-mapping $\sigma:\{x_1,\ldots,x_n\}\rightarrow\N$ such that
-$b=\sum_{i\in[n]} a_i\cdot\sigma(x_i)$.
-An {\em arithmetic progression} with {\em offset} $a\in\Z$ and {\em period
-$b\in\Z$} is the set $a+b\N=\{a+b\cdot\N\mid n\in\N\}$.
-The following lemma has been used in \cite{AnthonyIPL,Chrobak,Shallit}.
-
-\begin{lemma}[\cite{lookup citation}]\label{L NT}
-Let $N\in\N$ and let
-\begin{eqnarray}
-x\quad=\quad a_1\cdot x_1 + \cdots\ + a_n\cdot x_n,\label{L Eq}
-\end{eqnarray}
-be a linear equation such that $1<a_1<\cdots a_n\leq N$ and where
-$b=\gcd(a_1,\ldots,a_n)$.
-Then, the set of solutions to (\ref{L Eq}) is some set
-$U\cup\ a+b\N$,
-where $U\subseteq[1,N^2]$,
-and
-$a=\min\{y>N^2\mid y\equiv 0\textsl{ mod }g\}$.
-\end{lemma}
-
-\begin{lemma}
-Let $N\in\N$ be some natural and let
-\begin{eqnarray}
-x\quad=\quad a_1\cdot x_1 + \cdots\ + a_n\cdot x_n,\label{L Eq neg}
-\end{eqnarray}
-be a linear equation
-where $-N\leq a_1<a_2<\cdots <a_n\leq N$ and where $g=\gcd(a_1,\ldots,a_n)$.
-Let $b=\gcd(a_1,\ldots,a_n)$ and let
-$a=\min\{y>N^2\mid y\equiv 0\textsl{ mod }b\}$.
-Then, the set of positive solutions to (\ref{L Eq neg}) is a set
-$U\cup\  a+ b\N$, where $U\subseteq[1,N^2]$ and the
-set of negative solutions to (\ref{L Eq neg}) is a set
-$V\cup -a-b\N$, where $V\subseteq[-N^2,-1]$.
-\end{lemma}
-\begin{proof}
-We only prove the lemma for the negative solutions.
-Let $c<0$ be an arbitrary negative solution. Thus,
-$c=d+a_1'\cdot m_1+\ldots +a_h'\cdot m_h$, where
-$d\in[-N,0]$ and
-$\{a_1',\ldots,a_h'\}\subseteq\{a_1,\ldots,a_n\}\cap\Z^-$ is a subset of
-the negative coefficients.
-Hence $c$ is the solution
-to some linear equation $x=b_1\cdot x_1+\ldots +b_l\cdot x_l$, where
-$\{b_1,\ldots,b_l\}\subseteq[-N,-1]$. By application of Lemma
-\ref{L Eq} and switching from positive integers to negative integers, the lemma
-follows.
-\qed
-\end{proof}
-
-\begin{lemma}
-  \label{lem:bounded-coefficients}
-  Let $N,m$ be natural numbers and let $a_1,\ldots, a_n$ be integers
-  such that $-N\le a_1 \le \ldots \le a_n\le N$, $m>N^3$ and
-  \begin{equation*}
-    m = a_1 \cdot x_1 + \cdots + a_n \cdot x_n
-  \end{equation*}
-  for some natural numbers $x_1, \ldots, x_n$. Then there exist
-  natural numbers $y_1, \ldots, y_n$ such that
-  \begin{equation*}
-    m = a_1 \cdot y_1 + \cdots + a_n \cdot y_n
-  \end{equation*}
-  and $y_i\le N^2$ for $1\le i<n$.
-\end{lemma}
-\begin{proof}
-  Let $a_\ell\in\{a_1,a_n\}$ be the $a_i$ with the largest
-  \emph{absolute} value. Define $x_i' =x_i \text{ mod } a_\ell$ for
-  $1\le i\le n, i\neq \ell$, and let $m'= \sum_{1\le i\le n, i\neq
-    \ell} a_i\cdot m_i$. Clearly, we have $m'<N^3$ and $m'\equiv m
-  \text{ mod } a_\ell$. Let $h=m-m'=a_\ell \cdot x_\ell'$ for some
-  integer $x_\ell'$. In case $a_\ell=a_n$, the lemma follows by
-  setting $y_i=x_i'$ for $1\le i\le n$. Otherwise, let $x_n^*= \lceil
-  \frac{h}{|a_1a_n|} \rceil$. The lemma follows by setting $y_i=x_i'$
-  for $1<i<n$, $y_1=\frac{x_n^*|a_1a_n|-h}{|a_1|}$ and $y_n=x_n'+
-  x_n^*|a_1|$.
-\end{proof}
-\begin{remark}
-  \label{rem:maximum-valley}
-  It follows from the previous lemma that $\sum_{a_i<0} a_i\cdot
-  y_i>-N^3$.
-\end{remark}
-
-\section{Periodicity analysis}
-
-\newcommand{\cycles}{\textrm{cycles}}
-
-Let us fix a SOCA $\O=(Q,\{Q_p\mid p\in\Prop\},\delta_0,\delta_{>0})$ for
- the rest of this section.
-
-In the following, we will treat $\O$ as a weighted directed graph $G(\O)$, where
-the weights are taken according to $\delta_{>0}$.
-More precisely, we define $G(\O)=(Q,\delta_{>0})$
-and write $q_1\stackrel{a}{\mapsto}q_2$ as an abbreviation for
-$(q_1,a,q_2)\in\delta_{0}$ and simply $q_1\mapsto q_2$ if
-$(q_1,a,q_2)\in\delta_{>0}$ for some $a\in\Z$.
-Note that every non-zero path in the transition system $T(\O)$ corresponds to a unique path in
-$G(\O)$.
-A {\em strongly connected component} of $G(\O)$ is an
-equivalence class of the equivalence relation $\mapsto^*$.
-For every finite path $\pi$ in $G(\O)$, let $w(\pi)$ denote the sum of all
-weights of edges of $\pi$.
-For every strongly connected component $C$ in $G(\O)$, let $\cycles(C)$ denote
-the set of chord-free cycles whose total weight does not summarize to $0$,
-more precisely
-the set of paths $\pi=q_0\stackrel{m_1}{\mapsto} q_2\cdots
-\stackrel{m_l}{\mapsto} q_l$, where $l\geq1$,
-$q_0=q_l$, such that $i\not= j$ implies $q_i\not=q_j$ for all
-$i,j\in[l]$ and moreover $\sum_{i\in[l]}m_i\not=0$.
-
-Let $M=\max\{|a|:\exists q_1,q_2\in Q. q_1\stackrel{a}{\mapsto}q_2\}$
-denote the maximal absolute value of a weight that appears in $G(\O)$.
-Moreover, let $g_C=\gcd(w(cycles(C)))$ denote the greatest common divisor
-of all non-zero weights of chord-free cycles in $C$.
-In the following define $$\Delta\ =\ \LCM\{g_C\mid C\text{ is a strongly
-connected component of } G(\O)\}.$$
-\noindent
-Note that $\Delta\in\exp(|\O|)$ since firstly $g_C\leq (k-1)\cdot M$ for each strongly
-connected component $C$ and secondly there are at most $|Q|$
-strongly connected components in $G(\O)$.
-Let $k=|Q|$.
-
-\begin{lemma}{\label{L Cycles}}
-Let $C$ be some strongly connected component of $G(\O)$, let $q\in Q$ be some
-control location, and let $n\geq k^2\cdot M$ be some natural.
-Then in $T(\O)$ there exists some non-zero path from $(q,n)$ of height $h$ with $h\equiv
-0\text{ mod }g_C$ and $|h|\leq k\cdot(k-1)\cdot M$  that visits each
-control location in $C$ at least once.
-\end{lemma}
-\begin{proof}
-Since $C$ is a strongly connected component and $n\geq k^2\cdot M$, there exists some
-non-zero path
-$$
-\pi\quad =\quad(q_1,n_1)\rightarrow(q_2,n_2)\ \cdots\ \rightarrow (q_l,n_l)
-$$
-in $T(\O)$ with $l\in[k(k-1)]$ that visits each control state in $C$ at least once
-(i.e. for all $p\in C$ there is some $i\in[l]$ such that $p=q_i$) such that
-moreover $q_1=q_l=q$ and $n_1=n$. Thus, it follows $|n_1-n_l|\leq k\cdot
-(k-1)\cdot M$.
-Observe that for every subpath $\pi[a,b]=(q_a,n_a)\cdots\rightarrow(q_b,n_b)$ of $\pi$ that is a chord-free
-cycle, we have $n_a\equiv n_b\text{ mod } g_C$ by definition of $g_C$.
-By successively cutting out subpaths of $\pi$ that correspond to chord-free
-cycles, it is apparent that $n_1\equiv n_l\text{ mod }g_C$.
-\qed
-\end{proof}
-
-Observe that Lemma \ref{L Cycles} implies that there is a non-zero path in
-$T(\O)$ that visits every
-control location in $C$ at least once from $(q,k^2\cdot M+\delta)$ with the
-desired properties with respect to the height.
-
-\begin{lemma}
-  Let $q\in Q$ be a state of some SCC $C$ and let $m,m'$ be natural
-  numbers such that $m-m'>k^2\cdot M+ k^3\cdot M^3$ and $m'>k\cdot
-  M$. Then $(q,m)\rightarrow^*(q,m')$ iff
-  $(q,m+\Delta)\rightarrow^*(q,m')$.
-\end{lemma}
-\begin{proof}
-  $(\Rightarrow)$ Let $q_1,\ldots,q_n\in C, 1\le n\le k$ be states on
-  loop-free cycles $\pi_{a_i}:q_i\rightarrow \ldots \rightarrow q_i$
-  with weights $a_i=w(\pi_{a_i})$ such that
-  $\text{gcd}(a_1,\ldots,a_n)=g_C$.  Let $\pi_i^1: q\rightarrow^* q_i$
-  and $\pi_i^2: q_i\rightarrow^* q$ be disjoint loop-free paths from
-  $q$ to $q_i$ respectively $q_i$ to $q$. Denote by
-  $\pi_i=\pi_i^1\cdot \pi_i^2$ a loop-free cycle from $q$ to itself
-  passing through $q_i, 1\le i\le n$.
-
-  Let $h=m-m'-\sum_{1\le i\le n}w(\pi_i)+\Delta$. Similarly to the
-  previous lemma, we find $g_C|m-m'$, $g_C|\sum_{1\le i\le n}w(\pi_i)$
-  and hence $g_C|h$. Moreover, $h>k^3\cdot M^3$ and thus we can derive
-  from Lemma \ref{lem:bounded-coefficients} that there are
-  $y_1,\ldots, y_n$ such that $\sum_{1\le i\le n}a_i\cdot y_i=h$ and
-  $y_i<k\cdot M$, $1\le i<n$.
-
-  Now define $\pi_i'=\pi_i^1\cdot (\pi_{a_i})^{y_i} \cdot \pi_i^2$ and
-  $\pi=\pi_1'\cdot\pi_2'\cdot\cdots\cdot\pi_n'$. Cleary,
-  $w(\pi)=m+\Delta-m'$. Moreover, $m'> k\cdot M$ ensures that is a
-  valid path. Hence $(q,m+\Delta)\rightarrow^*(q,m')$ is witnessed
-  by $\pi$.
-
-  $(\Leftarrow)$ The proof from the previous direction can directly
-  be adopted to show the statement.
-\end{proof}
-
-The following lemma is important for our periodicity analysis later.
-
-\begin{lemma}
-Let $q,q'\in Q$ and $n,n'\in\N$ such that
-$n-n'>4\cdot k^3\cdot M^2$ and
-$n'>k\cdot M$.
-If there is a path from $(q,n)$ to $(q',n')$ in $T(\O)$, then there
-is a path from $(q,n+\Delta)$ to $(q',n')$ in $T(\O)$.
-\end{lemma}
-\begin{proof}
-Let us fix some path
-$$
-\pi\quad=\quad(q_1,n_1)\rightarrow(q_2,n_2)\quad \cdots\quad \rightarrow (q_l,n_l),
-$$
-where $(q_1,n_1)=(q,n)$ and $(q_l,n_l)=(q',n')$.
-Since $n_1-n_l>4\cdot k^3\cdot M^2$ and since there are at most $k$ strongly connected components in $G(\O)$,  there
-exists a strongly connected component $C$ of $G(\O)$ and some subpath $\pi[s,t]$
-of $\pi$ such that the following holds
-\begin{enumerate}
-\item  $n_s-n_t>4\cdot k^2\cdot M^2$,
-\item $n_t\geq n'$, and
-\item $q_i\in C$ for each $i\in[s,t]$
-\end{enumerate}
-by the pigeonhole principle.
-Let $\delta=n_t-n_s<-4\cdot k^2\cdot M^2$ denote the height of the subpath $\pi[s,t]$.
-Let us treat the path $\pi[s,t]$ as a path in $G(\O)$. Now think of consecutively cutting out
-chord-free cycles in $\pi[s,t]$  until this is no longer possible, in a greedy
-manner from beginning to the end, say. Let us denote the resulting
-path (possibly empty) by $\rho$ and let  $h\in[-(k-1)\cdot M,(k-1)\cdot M]$ denote
-its weight.
-\begin{remark}{\label{R rho}}
-The path $\rho$ can be seen as a non-zero path in $T(\O)$ of length at most
-$k-1$ from $(q_s,n_s)$ to $(q_t,n_s+h)$.
-\end{remark}
-
-\noindent
-The following claim states that the height $\delta$ of $\pi[s,t]$ lies in some
-arithmetic progression of small offset and small period.
-
-\smallskip
-
-\noindent
-{\em Claim A.} We have $\delta\in (c+h)-g_C\N$,
-for some
-$c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ with $c\equiv 0\text{ mod }g_C$.
-
-\medskip
-
-\noindent
-Let us postpone the proof of Claim A and finish the proof of the lemma.
-It remains to show that there is some non-zero path
-$(q,n+\Delta)\rightarrow^*(q',n')$ in $T(\O)$.
-Before doing this, we claim that the following paths exist in $T(\O)$:
-\begin{enumerate}[(1)]
-\item $(q,n+\Delta)\rightarrow^*(q_s,n_s+\Delta)$.
-\item $(q_t,n_t)\rightarrow^*(q',n')$.
-\item $(q_s,n_t-h)\rightarrow^*(q_t,n_t)$.
-\end{enumerate}
-Point (1) is clear since $\pi[1,s]+\Delta$ witnesses this.
-Point (2) holds by factorization of $\pi$.
-For Point (3), observe that $\rho$ is a non-zero path from $(q_s,n_s)$ to $(q_t,n_s+h)$
-of height $h$ and of length at most $k-1$
-by Remark \ref{R rho}.
-Since $n_t\geq n'>(k-1)\cdot M\geq h$, there is a non-zero path
-from $(q_t,n_t-h)$ to $(q_t,n_t)$, namely
-$\rho+(-n_s+n_t-h)=\rho+(\delta-h)$.
- By combining Points (3) and (2) we obtain some non-zero path $(q_t,n-h)\rightarrow^*
-(q',n')$.
-Due to Point (1), it suffices to show that there is some
-non-zero path from $(q_s,n_s+\Delta)$ to $(q_s,n_t-h)$.
-In other words, we have to prove that in $G(\O)$ there is some path from
-$q_s$ to $q_s$ of weight
-\begin{eqnarray*}
-w\quad=\quad n_t-h-(n_s+\Delta)\quad=\quad n_t-n_s-h-\Delta\quad=\quad\delta-h-\Delta
-\end{eqnarray*}
-such that moreover the weight of any prefix of this path is strictly bigger
-than $-(n_s+\Delta)$.
-By Claim A we have $w=\delta-h-\Delta=(c+h)-r\cdot g_C-h-\Delta=c-r\cdot
-g_C-\Delta$ for
-some $r\in \N$ and some $c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ with
-$c\equiv 0\text{ mod } g_C$.
-%Since too $\Delta\equiv 0\text{ mod } g_c$, there exists some  $d\in\N$ such
-%that
-Since moreover $\Delta\equiv 0\text{ mod }g_C$, we have $w\equiv 0\text{
-mod }g_C$.
-Furthermore, we have $w<-3\cdot k^2\cdot M^2$.
-Since $n_s+\Delta>BLA>k^2\cdot M$, there is some non-zero path $\sigma$ of height
-$h_\sigma$ with $|h_\sigma|\leq k\cdot(k-1)\cdot M$ and $h_\sigma\equiv 0\text{
-mod } g_C$
-from $(q_s,n_s+\Delta)$ that visits each control location in $C$ at
-least one by Lemma \ref{L Cycles} .
-When taking the path $\sigma\circ\sigma$, the remaining weight is congruent $0$
-modulo $g_C$ and can be estimated
-as $$w-2\cdot h_\sigma\ <\ -3\cdot k^2\cdot M^2-2\cdot h_\sigma\ <
--3\cdot k^2\cdot M^2+2\cdot k\cdot (k-1)\cdot M<
--k^2\cdot M^2.
-$$
-Hence by Lemma \ref{L Eq neg} there exist $m_1,\ldots, m_n\in\N$ such that the
-remaining weight equals $\sum_{i=1}^n a_i\cdot m_i$.
-To finally obtain a path from $q_s$ to $q_s$ of weight $w$ in $G(\O)$, we
-modify the first traversal of $\sigma$ by choosing for each positive $a_i$ some
-chord-free cycle of weight $a_i$ through which we loop precisely $m_i$ many times, followed by
-the modification of the second traversal of $\sigma$, where we choose for each
-negative $a_i$ some chord-free cycle of weight $a_i$ through which we loop
-precisely $m_i$ many times.
-This concludes the proof of the lemma.
-
-\smallskip
-
-\noindent
-{\em Proof of Claim A.}\\
-\noindent
-Observe that $\delta=h+r$, where $r$ is some solution
-of the linear equation
-$$
-x\ =\ a_1\cdot x_1+\ \ldots\ +a_m\cdot x_m,
-$$
-where $\{a_1,\ldots, a_m\}\subseteq w(cycles(C))$ are weights of the chord-free cycles that
-were cut out from $\pi[s,t]$ in order to obtain $\rho$. %\subseteq[-k\cdot M,k\cdot M]$.
-Let $d=\gcd(a_1,\ldots,a_m)$ which is a multiple of $g_C$ by definition.
-Since we know that $\delta<-4\cdot k^2\cdot M^2$ and $h\in[-(k-1)\cdot M,(k-1)\cdot M]$, we have
-$$
-r=\delta-h\quad<\quad-4\cdot k^2\cdot M^2+(k-1)\cdot M\quad\leq\quad -3\cdot k^2\cdot M^2.
-$$
-Moreover, we have obviously $-\sqrt{3}\cdot k\cdot M<-k\cdot M\leq a_i\leq
-k\cdot M<\sqrt{3}\cdot k\cdot M$ for each $i\in[m]$.
-Now let us define $c=\max\{y<-3\cdot k^2\cdot M^2\mid y\equiv 0\text{ mod }
-d\}$.
-As $d\in[M]$, it follows $c\in[-3\cdot k^2\cdot M^2-M,-3\cdot k^2\cdot M^2]$ as
-required.
-By applying Lemma \ref{L Eq neg} we have that $r\in c-d\N$ and hence
-$\delta\in (c+h)-d\N$.
-Since $d$ is a multiple of $g_C$
-it follows $\delta\in (c+h)-g_C\N$.
-\qed
-\end{proof}
-
-\begin{theorem}
-Model checking EF over SOCA is $\PSPACE$-complete.
-\end{theorem}
-
-\fi
-
-\newcommand{\Two}{\mathbb{A}}
-\renewcommand{\add}{\mathsf{add}}
-
-\subsection{$\Pi^0_1$-Hardness of the Data Complexity of $\CTL$ on POCA}{\label{S CTL und}}
-In this section we show that there already
-exists a fixed $\CTL$ formula for which model checking of
-$\poca$ is $\Pi^0_1$-hard.
-We reduce from the complement of the emptiness problem for
-\emph{two-counter automata}, which is $\Sigma^0_1$-complete \cite{Min61}.
-Similar to a
-SOCA, a {\em two-counter automaton} is a tuple $\Two=(Q, E, \lambda)$,
-where $Q$ is a finite set of {\em control states},
-$E\subseteq Q\times Q$ is a set of {\em edges},
- however
-$\lambda: E \rightarrow \{\zero_1,\zero_2\}\cup \{\add_j(a)\mid j\in \{1,2\}, a\in
-\{-1,1\} \}$. The {\em configuration graph induced by $\Two$} is defined
-to be $G(\A)=(Q\times \N\times \N, \rightarrow)$, where
-$(p,i_1,i_2)\rightarrow (q,i_1+d_1,i_2+d_2)$ if and only if
-the following conditions are all satisfied:
-\begin{itemize}
-\item $(p,q)\in E$,
-%\item  $i_1,i_2,i_1+d_1,i_2+d_2\geq 0$,
-\item $\lambda(p,q)=\add_j(a)$ implies $d_j=a$ and $d_{3-j}=0$ for each
-$j\in\{1,2\}$, and
-\item $\lambda(p,q)=\zero_j$ implies $d_j=i_j=0$ and $d_{3-j}=0$ for each
-$j\in\{1,2\}$.
-\end{itemize}
-
-\noindent
-Let us define the following problem.
-\smallskip
-
-\problemx{Emptiness for two-counter automata}{Two-counter automaton $\Two=(Q,E,\lambda)$, control states
-$q_0,q_1\in Q$.}{Are there $m,n\in\N$ such that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ in
-$G(\Two)$?}
-
-\medskip
-
-\noindent
-The idea of our reduction is as follows: Given a two-counter automaton
-$\Two$, we construct a $\poca$ $\Poca(x)$ with one parameter in such a way
-that the two counters from $\A$ are encoded into the single counter
-from $\Poca(x)$ as follows: for the counter value $n$ of $\Poca(x)$,
-we have that $n\ \mod x$ encodes the value of the first and $n\ \div\ x$ encodes the
-value of the second counter of $\Two$.
-Hence, testing the first counter for zero corresponds to checking
-whether $n\equiv 0\ \mod x$ while testing the second counter for
-zero corresponds to checking  whether $n\geq x$.
-Incrementation (resp. decrementation) on the first counter can be
-mimicked by adding (resp. subtracting) $1$, whereas on counter two this
-corresponds to adding (resp. subtracting) $x$.
-Of course, we need to ensure that we do not overflow.
-For example, if $n\equiv -1\ \mod x$ and we would simulate to increment the first counter
-in the above manner, this
-would correspond to setting the first counter to zero and simultaneously
- incrementing the second counter. However, if the emptiness problem is
-solvable, then $x$ can be instantiated with a large enough
-value such that such an overflow does not occur.
-Our $\ctl$ formula will ensure that we do not overflow.
-\newcommand{\test}{\mathsf{test}}
-Before we give the reduction from the emptiness problem for two-counter
-automata, let us introduce some gadgets that we need in our reduction.
-\begin{lemma}{\label{L POCAs}}
-  There exists some fixed $\ctl$ formula $\varphi_\test$
-for which the following holds:
-\begin{enumerate}[(1)]
-\item
-There exists some fixed $\poca$ $\Poca_{\not\equiv 0}(x)$ with some control state
-$q_{\not\equiv 0}$ such that for each $\sigma:\{x\}\rightarrow\N$
- $$(\Poca_{\not\equiv 0}^\sigma,q_{\not\equiv0}(n)) \models
- \varphi_{\test} \quad\Longleftrightarrow\quad n\not
-  \equiv 0 \textsl{ mod } \sigma(x).$$
-\item There exists some fixed $\poca$ $\Poca_{\not\equiv-1}(x)$ with some control
-state $q_{\not\equiv -1}$ such that for each $\sigma:\{x\}\rightarrow\N$
- $$(\Poca_{\not\equiv -1}^\sigma,q_{\not\equiv-1}(n)) \models
- \varphi_\test \quad\Longleftrightarrow\quad n\not
-  \equiv -1 \textsl{ mod } \sigma(x).$$
-\item There exists some fixed $\poca$ $\Poca_{<x}(x)$ with some control
-state $q_{<x}$ such that for each $\sigma:\{x\}\rightarrow\N$
- $$(\Poca_{< x}^\sigma,q_{<x}(n)) \models
- \varphi_\test \quad\Longleftrightarrow\quad n< \sigma(x).$$
-
-\end{enumerate}
-\end{lemma}
-\begin{proof}
-We choose $\varphi_\test=\neg\EF\prop$, where $\prop$ is an atomic
-proposition.
-Below we depict $\Poca_{\not\equiv0}(x)$,
-$\Poca_{\not\equiv-1}(x)$, and $\Poca_{<x}(x)$, respectively.
-\begin{center}
-\begin{picture}(85,20)(-20,-3.5)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-\put(-65,0){
-\put(41,-2){\small $\prop$}
-  \node(A)(20,0){$q_{\not\equiv0}$}
-  \node(B)(40,0){$\bullet$}
-\drawloop(A){$-x$}
-\drawedge(A,B){$\zero$}
-}
-%%% AUTOMATON 2
-
-\put(-27,0){
-\put(61,-2){\small $\prop$}
-  \node(A)(20,0){$q_{\not\equiv-1}$}
-  \node(B)(40,0){$\bullet$}
-  \node(C)(60,0){$\bullet$}
-\drawloop(B){$-x$}
-\drawedge(A,B){$+1$}
-\drawedge(B,C){$\zero$}
-}
-
-%%% AUTOMATON 3
-
-\put(32,0){
-\put(61,-2){\small $\prop$}
-  \node(A)(20,0){$q_{< x}$}
-  \node(C)(40,0){$\bullet$}
-  \node(D)(60,0){$\bullet$}
-\drawloop(A){$-1$}
-\drawedge(A,C){$-x$}
-\drawedge(C,D){$\zero$}
-%\drawedge[ELside=r](C,D){$\zero$}
-}
-
-\end{picture}
-\end{center}
-Correctness of the lemma is immediate.
-\qed
-\end{proof}
-
-\newcommand{\up}{\mathsf{up}}
-
-\noindent
-Let us fix some instance of the emptiness problem for
- two-counter automata:
-Let $\langle\Two,q_0,q_1\rangle$ be an instance to the
-emptiness problem, where
-$\Two=(Q,E,\lambda)$ is some two-counter automaton.
-We will give a fixed $\ctl$ formula $\varphi$ such that from
-$\langle\Two,q_0,q_1\rangle$ we can construct some $\poca$ $\Poca(x)$ such that
-$(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ if and only if
-$(\Poca^\sigma,q_0(0))\models\varphi$ for some $\sigma:\{x\}\rightarrow\N$.
-
-The control states and transitions will of $\Poca(x)$ can be described by the
-following graph transformation rule that maps each pair of control states
-$(p,q)\in E$ of $\Two$ with $\lambda(p,q)=l$
- into corresponding control states in $\Poca(x)$, by
-possibly accessing the different $\poca$ of Lemma \ref{L POCAs}.
-\begin{center}
-\begin{picture}(85,30)(30,-10)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-\put(60,6){\Huge$\hookrightarrow$}
-
-%\put(41,-2){\small $\prop$}
-  \node(A)(20,10){$p$}
-  \node(B)(40,10){$q$}
-
-\drawedge(A,B){$l$}
-
-\put(109,12){\small$l$}
-
-  \node(C)(90,10){$p$}
-  \node(D)(110,10){$\bullet$}
-  \node(E)(130,10){$q$}
-  \node(F)(110,-10){$\test(l)$}
-
-%\drawloop(A){$-x$}
-\drawedge(C,D){$0$}
-\drawedge(D,E){$\up(l)$}
-\drawedge(D,F){$0$}
-\end{picture}
-\end{center}
-
-\noindent
-where
-$$
-\test(l)\quad =\quad \begin{cases} (\Poca_{\not\equiv 0}(x),q_{\not\equiv 0}) & \text{ if }\quad
-l\in\{ \add_1(-1),\zero_1\}\\
-(\Poca_{\not\equiv -1}(x),q_{\not\equiv -1})& \text{ if }\quad l=\add_1(+1)\\
-(\Poca_{<x}(x),q_{<x})& \text{ if }\quad l=\zero_2,
-\end{cases}
-$$
-
-\noindent
-and where
-$$
-\up(l)\quad =\quad \begin{cases}  a & \text{ if }\quad
-l=\add_1(a)\\
-\circ x& \text{ if }\quad l=\add_2(\circ 1)
-\text{ for some }\circ\in\{+,-\}\\
-0& \text{ otherwise.}
-\end{cases}
-$$
-
-\noindent
-Let us moreover ensure that every control state of the kind
-$q_{\tau}$ with $\tau\in\{\not\equiv 0,\not\equiv\! -1,\geq x\}$ has
-the label $\rho_\tau$.
-Before we give our final formula, let us introduce the following constraint
-formulas that guarantee that an overflow never occurs:
-\begin{eqnarray*}
-\psi_1&\quad=\quad& \add_1(+1)\longrightarrow \EX(\rho_{\not\equiv-1}
-\wedge\varphi_\test)\
-\\
-\psi_2&\quad=\quad&
-\add_1(-1)\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\varphi_\test)\\
-\psi_3&\quad=\quad&
-\zero_2\longrightarrow\EX(\rho_{<x}\wedge\varphi_{\test})\\
-\psi_4&\quad=\quad&
-\zero_1\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\neg\varphi_\test)
-\end{eqnarray*}
-
-\noindent
-Note that we do not need any constraint when changing the second counter.
-Let us introduce a label $f$ at control state $q_1$.
-Our final $\ctl$ formula $\varphi$ is
-$$
-\varphi\quad=\quad\E\left(\bigwedge_{i\in[4]}\psi_i\ \U\  f\right).
-$$
-\iffalse
-The correctness of the reduction follows from the following equivalences:
-\begin{itemize}
-\item $(T(\Poca),(q_0,0))\models \neg \varphi$
-\item for all $\sigma:\{x\}\rightarrow \N$ we have $(\Poca^\sigma(x),q_0(0))\not\models\varphi$
-\item for some $\sigma:\{x\}\rightarrow\N$ we have $(\Poca^\sigma(x),q_0(0))\models\varphi$
-\item for some $\exists\sigma:\{x\}\rightarrow\N$ there is some path
-$r_0(0)\rightarrow r_1(n_1)\cdots r_k(n_k)$ such that
-$r_0=q_0$, $r_k=q_1$, $r_k\in F$ and
-$(T(\Poca^\sigma(x)),r_i(n_i)
-
-%$(q_0,0)(q_1,n_1\mod\sigma(x),n_1\div\sigma(x))\cdots (q_k,n_k\mod\sigma(x),
-%n_k\div \sigma(x))$ in $T(\Poca^\sigma(x))$ such that
-
-\end{itemize}
-Then there is an assignment
-$\sigma$ such that $(T(\Poca^\sigma),(q_0,0))\models \varphi$. It is
-easily verified that the witnessing path $(q,0)(q_1,n_1)\ldots
-(q',n')$ in $T(\A'(\sigma))$ corresponds to the path
-$(q,0,0)(q_1,n_1\mod \sigma(x), n_1\div \sigma(x))\ldots (q',n'\mod
-\sigma(x), n'\div \sigma(x))$ in $T(\A)$. Conversely, if
-$(q,0,0)\rightarrow^* (q,n_1,n_2)$ in $T(\A)$, then the first counter
-of $\A$ does not exceed some value $u \in \naturals$. Consequently, by
-setting $\sigma(x):=u$, $(T(\A(\sigma)),(q,0))\models \varphi$.
-\fi
-We have that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ in $G(\A)$  if and only
-if
-$(T(\Poca^\sigma(x)),q_0(0))\models\varphi$ for some
-$\sigma:\{x\}\rightarrow\N$.
-We obtain the following theorem.
-
-\begin{theorem}
-  The data and combined complexity of $\CTL$ and the modal $\mu$-calculus
- on POCA is $\Pi^0_1$-complete.
-\end{theorem}
-
-\section{$\ltl$ Model Checking}{\label{S LTL}}
-
-%\subsection{Linear-time logic ($\ltl$)}
-Formulas of $\ltl$ are inductively defined according to the
-following grammar, where $\prop$ ranges over $\Prop$:
-\begin{equation*}
-  \varphi\quad ::=\quad \prop \mid \neg \varphi \mid \varphi \wedge \varphi \mid
-  \X \varphi \mid \varphi \until \varphi
-\end{equation*}
-The Boolean abbreviations and the formula $\true$ are defined in the same way as
-in CTL. The \emph{finally modality} $\eventually
-\varphi$ is an abbreviation for $\true \until \varphi$ and the
-\emph{globally modality} $\globally \varphi$ abbreviates $\neg
-\eventually \neg \varphi$.
-
-The semantics of LTL is given in terms of infinite paths in a
-transition system. Let $T=(S,\{S_\prop \mid \prop \in
-\Prop\},\rightarrow)$ be a transition system, let $\pi: s_0
-\rightarrow s_1 \rightarrow \cdots$ be an infinite path in $T$, and
-let $\varphi$ be an $\ltl$ formula, we define $(T,\pi)\models\varphi$
-by induction of the structure of $\varphi$, where $\prop$ ranges over
-$\Prop$:
-\begin{align*}
-  (T,\pi) \models  \prop &  \quad\iff\quad  \pi(0) \in S_\prop\\
-  (T,\pi) \models  \neg \varphi &\quad\iff\quad (T,\pi) \not \models \varphi\\
-  (T,\pi) \models \varphi_1 \wedge \varphi_2 &\quad\iff\quad (T,\pi)
-  \models \varphi_1 \text{ and } (T,\pi) \models \varphi_2\\
-  (T,\pi) \models \X \varphi & \quad\iff\quad (T,\pi^1) \models \varphi\\
-  (T,\pi) \models  \varphi_1 \until \varphi_2 & \quad\iff\quad
-  \exists j\ge 0: (T,\pi^j)\models \varphi_2
-  \text{ and } \forall 0\le i<j: (T,\pi^i)\models \varphi_1
-\end{align*}
-Let us now define the $\ltl$ model checking problem on SOCA
-and POCA.\\
-
-\problemx{$\ltl$ model checking on SOCA}
-{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary,
-  and an $\ltl$ formula $\varphi$.}
-{$(T(\Soca),\pi)\models\varphi$ for every infinite
-path $\pi$ with $\pi(0)=q(n)$?}
-\
-
-\problemx{$\ltl$ model checking on POCA}
-{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and an
-$\ltl$ formula $\varphi$.}
-{$(T(\Poca^\sigma),\pi)\models\varphi$ for every $\sigma:X\rightarrow\N$ and
-every infinite path $\pi$ with $\pi(0)=q(n)$?}
-
-\subsection{Upper bounds}
-
-A standard approach to $\ltl$ model checking is the automata-based
-approach, in which systems are modeled as non-deterministic B\"uchi
-automata (NBA). Given an NBA $\automaton$ and an $\ltl$ formula
-$\varphi$, the idea is to translate $\varphi$ into an NBA
-$\automaton_{\neg \varphi}$ such that $\automaton_{\neg \varphi}$
-accepts all words that violate $\varphi$.  Then, by checking for
-emptiness of the product automaton
-$\automaton'=\automaton\times\automaton_{\neg \varphi}$, it can be
-decided whether or not $\automaton$ generates traces that violate
-$\varphi$.  Emptiness can be decided by checking for recurrent
-reachability of a control state in the transition system induced by
-$\automaton'$, which can be performed non-deterministically in space logarithmically in the size of
-$\automaton'$. Vardi and Wolper showed in \cite{VW86} that the size of
-$\automaton_{\neg \varphi}$ is $2^{O(|\varphi|)}$, which yields that
-checking for emptiness of $\automaton'$ is in $\PSPACE$. The $\PSPACE$
-lower bound for finite state model checking was originally shown
-in \cite{LP85-popl} and hence carries over to SOCA.
-
-Constructing the product automaton of two NBA is a standard technique
-in model checking, see e.g.\ \cite{BK08} for a detailed treatment of
-this. Moreover, this construction can be adapted in a
-straight-forward way into the setting of POCA by equipping a POCA with
-a set of final states and defining emptiness with respect to B\"uchi
-acceptance condition in the standard way. For brevity, we do not give
-any further details. We then have that given a POCA $\Poca$, an $\ltl$
-formula $\varphi$, a state $q$ and $n\in\N$ we can construct a
-POCA $\Poca'$ as the product of $\Poca$ and $\automaton_{\neg\varphi}$
-such that for all assignments $\sigma$ we have that $\Poca'$ is empty
-with respect to B\"uchi acceptence condition if and only if
-$(T(\Poca^\sigma),q(n))\models \varphi$. Moreover, the size of
-$\Poca'$ is $2^{O(|\varphi|)}|\Poca|$.
-
-It has been shown in \cite{HKOW10} that checking emptiness with
-respect to B\"uchi acceptance condition is $\coNP$-complete for both
-SOCA and POCA, and in \cite{DG-jlc09} that it is $\NL$-complete for
-OCA. We use these results in order to establish upper bounds for our
-$\ltl$ model checking problems. Let us first consider the case of a
-fixed formula $\varphi$. In the terminology of the previous paragraph,
-$|\Poca'|=O(|\Poca|)$ and hence the data complexity of $\ltl$ on SOCA
-and POCA is in $\coNP$.  The matching $\coNP$ lower bound of the data
-complexity of LTL on SOCA follows from the $\NP$-hardness of the
-reachability problem for SOCA.
-%We obtain $\coNP$-hardness from the fact that
-%we can reduce deciding emptiness of some SOCA $\Soca$ with respect to
-%B\"uchi acceptance condition to $\ltl$ model checking of a slightly
-%altered SOCA $\Soca'$ against an easily constructed fixed $\ltl$
-%formula.
-
-\begin{proposition}
-The data complexity of $\ltl$ model checking on SOCA and POCA is
-  $\coNP$-complete.
-\end{proposition}
-As seen above, if both $\Poca$ and the formula $\varphi$ are part of
-the input then $|\Poca'|=2^{O(|\varphi)}\cdot|\Poca|$, and hence $\coNEXP$
-is an upper bound for the combined complexity of $\ltl$ model checking
-on both SOCA and POCA. However, we can improve this upper bound in the
-case of SOCA. Given a SOCA $\Soca$, let $m$ be the absolute value of
-the maximal increment or decrement performed on the transitions in
-$\Soca$ and let $\Soca'$ be the product automaton of $\Soca$ and
-$\automaton_{\neg \varphi}$. Let $\Soca''$ be obtained from $\Soca'$
-by replacing every transition labeled with $z$ with a sequence of
-fresh transitions and control locations of length $z$, where each
-transition is labeled with $+1$ respectively $-1$, depending on the
-sign of $z$. We conclude that
-$|\Soca''|=m\cdot|\Soca'|=m\cdot2^{O(|\varphi|)}\cdot|\Soca|$. Combining the result
-from \cite{DG-jlc09} together with the fact that $\ltl$ model checking
-of non-deterministic finite state automata is $\PSPACE$-hard
-\cite{LP85-popl}, we obtain the following proposition.
-
-\begin{proposition}
-The combined complexity of $\ltl$ on SOCA is $\PSPACE$-complete.
-\end{proposition}
-It remains to prove that the combined complexity of $\ltl$ model
-checking on POCA is $\coNEXP$-hard, which we will show in the next
-section.
-
-\subsection{Combined Complexity of $\ltl$ on POCA}
-
-We are now going to show $\coNEXP$-hardness of $\ltl$ model checking
-on POCA via a reduction from (the complement of) \textsf{Succinct
-  3-SAT}, which is a $\NEXP$-complete problem \cite{Papa94}. An input
-of \textsf{Succinct 3-SAT} is given by a Boolean circuit $\mathbb{C}$
-that encodes a Boolean formula $\psi_\circuit$ in 3-CNF,
-i.e.\ $\psi_\circuit=\bigwedge_{j\in [0,M]}(\ell^j_1\vee \ell^j_2\vee
-\ell^j_3)$ for some $M=O(2^{|\circuit|})$, and $\psi_\circuit$ is free
-over Boolean variables $y_1,\ldots,y_{N}$ for some
-$N=O(2^{|\circuit|})$. Let $j\in [0,M]$ be the index of a clause
-encoded in \emph{binary} and $k\in\{1,2,3\}$. On input $(j\cdot k)$,
-the output of $\circuit$ is $(i\cdot b)$, where $i\in [N]$ is the
-index of the Boolean variable that appears in literal $\ell_k^j$, and
-where $b=0$ when $\ell_k^j$ is negative and $b=1$ when $\ell_k^j$ is
-positive.
-\vspace*{0.25cm}
-\problemx{\textsf{Succinct 3-SAT}}
-{Boolean circuit $\circuit$}
-{Is $\psi_\circuit$ satisfiable?}
-\vspace*{0.25cm}
-
-%\medskip
-
-\noindent
-In order to establish $\coNEXP$-hardness for the combined complexity
-of $\ltl$ model checking, given an input $\circuit$ of
-$\textsf{Succinct 3-SAT}$, we construct a POCA $\Poca(x)$ and an
-$\ltl$ formula $\varphi$ such that $\psi_\circuit$ is satisfiable if
-and only if there is an assignment $\sigma$ such that
-$(T(\Poca^\sigma),q_{\mathsf{start}}(0))\models\varphi$ for some
-distinguished state $q_{\mathsf{start}}$ of $\Poca(x)$.
-%There are two key
-%ideas for the reduction.
-
-First, let us provide a suitable encoding of truth assignments by
-natural numbers. The encoding we use has also been employed for
-establishing lower bounds for model checking OCA~\cite{JKMZ04-ic}.
-Recall that $p_i$ denotes the $i^{\text{th}}$ prime number. Every
-natural number $x$ defines a truth assignment
-$\nu_x:\{y_1,\ldots,y_N\} \to \{ 0,1\}$ such that $\nu_x(y_i)=1$ iff
-$p_i$ divides $x$.  By the Prime Number Theorem, $p_N=O(N \log N)$ and
-hence $O(|\circuit|)$ bits are sufficient to represent $p_N$. Of
-course, since we need exponentially many prime numbers they cannot be
-hard-wired into $\Poca(x)$.
-
-%% Second, we employ $\ltl$ in order to perform arbitrary computations of
-%% space-bounded Turing machines on traces of runs of SOCA. However, we
-%% cannot make use of the approach we employed in $\ctl$. Intuitively
-%% speaking, $\ctl$ allows for performing tests on the counter value
-%% without destroying its value and this property has no correspondant in
-%% the $\ltl$ setting. We can however use atomic propositions to model
-%% computations along traces.
-
-\begin{figure}
-  \begin{center}
-    \begin{picture}(142,60)(0,0)
-      %   \put(-20,-28){\framebox(85,33){}}
-      %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
-      % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
-      % \node(A)(0,0){$q_0$}
-
-      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
-      \gasset{Nadjust=wh,Nadjustdist=1}
-      \gasset{curvedepth=0}
-
-      \put(0,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(6,7){$\Soca_{\mathsf{inc}}(j)$}
-      }
-
-      \put(35,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,8){$\Soca_{\circuit}(j\cdot 2)$}
-        \put(10,4){$=(i\cdot b)$ }
-      }
-      \put(35,45){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,8){$\Soca_{\circuit}(j\cdot 1)$}
-        \put(10,4){$=(i\cdot b)$ }
-      }
-      \put(35,5){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,8){$\Soca_{\circuit}(j\cdot 3)$}
-        \put(10,4){$=(i\cdot b)$ }
-      }
-
-      \put(70,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,6){$\Soca_{\mathsf{prime}}(i)=p_i$ }
-      }
-
-      \put(110,35){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,6){$\Soca_{\mathsf{divides}}(p_i,x)$ }
-        \put(8,11){$b=0?$}
-      }
-
-      \put(110,10){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,6){$\Soca_{\mathsf{not\_divides}}(p_i,x)$ }
-        \put(8,11){$b=1?$}
-      }
-
-      \node(A1)(12,29){}
-      \node(A2)(20,32){}
-
-      \node(B11)(40,8){}
-      \node(B12)(55,15){}
-
-      \node(B21)(40,28){}
-      \node(B22)(55,35){}
-
-      \node(B31)(40,48){}
-      \node(B32)(55,55){}
-
-      \node(C1)(73,36){}
-      \node(C2)(90,28){}
-
-      \node(D11)(115,39){}
-      \node(D12)(132,39){}
-
-      \node(D21)(113,13){}
-      \node(D22)(132,13){}
-
-      \node(E)(140,0){$\bullet$}
-      \node(F)(12,0){$q_{\mathsf{start}}$}
-
-      \drawedge(F,A1){$+x$}
-
-      \drawedge(A2,B11){$0$}
-      \drawedge(A2,B21){$0$}
-      \drawedge(A2,B31){$0$}
-
-      \drawedge(B12,C1){$0$}
-      \drawedge(B22,C1){$0$}
-      \drawedge(B32,C1){$0$}
-
-      \drawedge(C2,D11){$0$}
-      \drawedge(C2,D21){$0$}
-
-      \drawqbedge(D12,142,25,E){$0$}
-      \drawedge(D22,E){$0$}
-
-      \drawedge(E,F){$\mathsf{zero}$}
-    \end{picture}
-
-  \end{center}
-  \caption{High-level description of the automaton $\Poca(x)$ used for
-    the reduction.}
-  \label{fig:high-level}
-\end{figure}
-
-Let us now take a look at $\Poca(x)$, which is sketched in Figure
-\ref{fig:high-level}. It uses one parameter $x$ and employs several
-gadgets. The only gadgets manipulating the counter are
-$\Soca_{\mathsf{divides}}$ and $\Soca_{\mathsf{not\_divides}}$. The
-remaining gadgets are designed such that they communicate via
-designated propositional variables, and not as in Section \ref{S CTL
-  hardness} with the help of the counter. First, $\Poca(x)$ loads the
-value of the parameter $x$ on the counter.  Think of $x$ encoding a
-truth assignment of $\psi_\circuit$.  Next, $\Poca(x)$ traverses
-through $\Soca_{\mathsf{inc}}$, which initially chooses an arbitrary
-index $j$ identifying a clause of $\psi_\circuit$. Every time
-$\Soca_{\mathsf{inc}}$ is traversed afterwards, it increments $j$
-modulo $M+1$ and hereby moves on to the next clause.  Now $\Poca(x)$
-branches non-deterministically into a gadget $\Soca_\circuit$ in order
-to compute $(i\cdot b)$ from $\circuit$ on input $(j\cdot 1), (j\cdot
-2)$, resp.\ $(j\cdot 3)$. The index $i$ is then used as input to a
-gadget $\Soca_{\mathsf{prime}}$, which computes $p_i$. Then if $b=0$,
-it is checked in $\Soca_{\mathsf{not\_divides}}$ that $p_i$ does not
-divide the value of $x$, and likewise in $\Soca_{\mathsf{divides}}$
-that $p_i$ divides the value of $x$ if $b=1$. For checking the latter,
-the counter needs to be modified. After the checks have been finished,
-we restore the value $x$ on the counter and the process continues with
-clause $j+1~ \mod M+1$.
-
-It remains to show how the gadgets and the communication between them
-can be realized. The first observation is that the computations of
-$\Soca_{\mathsf{inc}}$, $\Soca_{\circuit}$ and
-$\Soca_{\mathsf{prime}}$ can be realized by space bounded DTM using no
-more than $p(|\circuit|)$ tape cells for their in- and output tape for
-some polynomial $p$ that is fixed once $\circuit$ is provided. Indeed,
-it is easily seen that incrementing modulo $M+1$, evaluating
-$\circuit$ and computing the $i^{\text{th}}$ prime number can be done
-by such a DTM. Thus, we now show how from a generic DTM $\M$, we can
-construct in logarithmic space a SOCA $\Soca_\M$ and some LTL formulas
-that mimic computations of $\M$ on traces of $\Soca_\M$.
-
-\begin{figure}
-  \begin{center}
-    \begin{picture}(140,100)(0,-52)
-      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
-      \gasset{Nadjust=wh,Nadjustdist=1}
-      \gasset{curvedepth=0}
-
-      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-      \node(START)(0,25){}
-      \node(A)(10,25){$\bullet$}
-      \node(B)(20,25){$q$}
-      \node(C1)(30,45){$\bullet$}
-      \node(C2)(30,35){$\bullet$}
-      \node(C3)(30,25){$\vdots$}
-      \node(C4)(30,15){$\bullet$}
-      \node(C5)(30,5){$\bullet$}
-      \node(D)(40,25){$\bullet$}
-%      \node(STATEOUT)(70,25){}
-
-%      \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-      \drawedge(START,A){}
-      \drawedge(A,B){}
-      \drawedge(B,C1){}
-      \drawedge[ELside=r](B,C2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(B,C4){}
-      \drawedge[ELside=r](B,C5){}
-      \drawedge(C1,D){}
-      \drawedge[ELside=r](C2,D){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(C4,D){}
-      \drawedge[ELside=r](C5,D){}
-%      \drawedge(D,STATEOUT){}
-
-      \put(9,21){$\zeta$}
-      \put(18,21){$\alpha$}
-      \put(29,47){$\mathsf{s}_0$}
-      \put(29,37){$\mathsf{s}_1$}
-      \put(27,12){$\mathsf{s}_{k-1}$}
-      \put(29,2){$\mathsf{s}_k$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-%%%%%%%%%%%%%%%%%% ihead
-
-%      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-%      \node(A)(5,25){$\bullet$}
-      \node(E)(50,25){$\bullet$}
-      \node(F1)(60,45){$\bullet$}
-      \node(F2)(60,35){$\bullet$}
-      \node(F3)(60,25){$\vdots$}
-      \node(F4)(60,15){$\bullet$}
-      \node(F5)(60,5){$\bullet$}
-      \node(G)(70,25){$\bullet$}
-%      \node(STATEOUT)(70,25){}
-
-      \drawedge(D,E){}
-
-%     \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-%      \drawedge(A,B){$0$}
-      \drawedge(E,F1){}
-      \drawedge[ELside=r](E,F2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(E,F4){}
-      \drawedge[ELside=r](E,F5){}
-      \drawedge(F1,G){}
-      \drawedge[ELside=r](F2,G){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(F4,G){}
-      \drawedge[ELside=r](F5,G){}
-%      \drawedge(D,STATEOUT){}
-
-%      \put(18,21){$\alpha$}
-      \put(58,47){$\mathsf{ih}_0$}
-      \put(58,37){$\mathsf{ih}_1$}
-      \put(57,12){$\mathsf{ih}_{m}$}
-      \put(56,2){$\mathsf{ih}_{m+1}$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-%%%%%%%%%%%%%%%%%% whead
-
-%      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-%      \node(A)(5,25){$\bullet$}
-      \node(E)(80,25){$\bullet$}
-      \drawedge(G,E){}
-      \node(F1)(90,45){$\bullet$}
-      \node(F2)(90,35){$\bullet$}
-      \node(F3)(90,25){$\vdots$}
-      \node(F4)(90,15){$\bullet$}
-      \node(F5)(90,5){$\bullet$}
-      \node(G)(100,25){$\bullet$}
-      \node(H)(100,0){$\bullet$}
-
-%     \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-%      \drawedge(A,B){$0$}
-      \drawedge(E,F1){}
-      \drawedge[ELside=r](E,F2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(E,F4){}
-      \drawedge[ELside=r](E,F5){}
-      \drawedge(F1,G){}
-      \drawedge[ELside=r](F2,G){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(F4,G){}
-      \drawedge[ELside=r](F5,G){}
-      \drawedge(G,H){}
-
-%      \put(18,21){$\alpha$}
-      \put(88,47){$\mathsf{wh}_0$}
-      \put(88,37){$\mathsf{wh}_1$}
-      \put(87,12){$\mathsf{wh}_{n}$}
-      \put(86,2){$\mathsf{wh}_{n+1}$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-      \node(I)(0,0){$\bullet$}
-      \drawedge(H,I){}
-
-%%%%%%%%%%%%%%%% wtape
-
-%      \node(WORKIN)(0,45){}
-      \node(A)(0,-15){$\bullet$}
-      \drawedge(I,A){}
-      \node(B1)(15,-25){$\bullet$}
-      \node(B2)(15,-5){$\bullet$}
-      \node(C)(30,-15){$\bullet$}
-      \node(D1)(45,-25){$\bullet$}
-      \node(D2)(45,-5){$\bullet$}
-      \node(E)(60,-15){$\bullet$}
-      \node(F)(70,-15){$\ldots$}
-      \node(G)(80,-15){$\bullet$}
-      \node(H1)( 95,-25){$\bullet$}
-      \node(H2)(95,-5){$\bullet$}
-      \node(I)(110,-15){$\bullet$}
-
-      \node(J)(110,-30){$\bullet$}
-      \node(K)(0,-30){$\bullet$}
-      \drawedge(I,J){}
-      \drawedge(J,K){}
-%      \node(J)(135,-15){$\bullet$}
-%      \node(DASH)(120,-25){}
-%      \node(WORKOUT)(145,-15){}
-
-%      \drawedge(WORKIN,A){}
-      \drawedge(A,B1){}
-      \drawedge(A,B2){}
-      \drawedge(B1,C){}
-      \drawedge(B2,C){}
-      \drawedge(C,D1){}
-      \drawedge(C,D2){}
-      \drawedge(D1,E){}
-      \drawedge(D2,E){}
-      \drawedge(G,H1){}
-      \drawedge(G,H2){}
-      \drawedge(H1,I){}
-      \drawedge(H2,I){}
-%      \drawedge(I,J){}
-%      \drawedge[dash={1.5}0](I,DASH){}
-
-%      \put(4,42){$\mathsf{w}_\l$}
-      \put(13,-9){$\mathsf{i}_{1,1}$}
-      \put(13,-29){$\mathsf{i}_{1,0}$}
-      \put(43,-9){$\mathsf{i}_{2,1}$}
-      \put(43,-29){$\mathsf{i}_{2,0}$}
-      \put(93,-9){$\mathsf{i}_{m,1}$}
-      \put(93,-29){$\mathsf{i}_{m,0}$}
-%      \put(120,42){$\mathsf{w}_\r$}
-%      \put(133,42){$\mathsf{yield}$}
-
-%%%%%%%%%%%%%%% wtape
-      \node(A)(0,-45){$\bullet$}
-
-      \drawedge(K,A){}
-
-      \node(B1)(15,-55){$\bullet$}
-      \node(B2)(15,-35){$\bullet$}
-      \node(C)(30,-45){$\bullet$}
-      \node(D1)(45,-55){$\bullet$}
-      \node(D2)(45,-35){$\bullet$}
-      \node(E)(60,-45){$\bullet$}
-      \node(F)(70,-45){$\ldots$}
-      \node(G)(80,-45){$\bullet$}
-      \node(H1)( 95,-55){$\bullet$}
-      \node(H2)(95,-35){$\bullet$}
-      \node(I)(110,-45){$\bullet$}
-      \node(J)(120,-45){$\bullet$}
-      \node(K)(120,50){$\bullet$}
-      \node(L)(20,50){$\bullet$}
-      \node(M)(20,25){}
-      \node(N)(130,-45){$\bullet$}
-      \node(P)(140,-45){}
-
-      \drawedge(I,J){}
-      \drawedge(J,K){}
-      \drawedge(K,L){}
-      \drawedge(L,M){}
-
-      \drawedge(J,N){}
-      \drawedge(N,P){}
-
-%      \node(DASH)(120,-25){}
-%      \node(WORKOUT)(145,-15){}
-
-%      \drawedge(WORKIN,A){}
-      \drawedge(A,B1){}
-      \drawedge(A,B2){}
-      \drawedge(B1,C){}
-      \drawedge(B2,C){}
-      \drawedge(C,D1){}
-      \drawedge(C,D2){}
-      \drawedge(D1,E){}
-      \drawedge(D2,E){}
-      \drawedge(G,H1){}
-      \drawedge(G,H2){}
-      \drawedge(H1,I){}
-      \drawedge(H2,I){}
-%      \drawedge(I,J){}
-%      \drawedge[dash={1.5}0](I,DASH){}
-
-%      \put(4,42){$\mathsf{w}_\l$}
-      \put(12,-39){$\mathsf{w}_{1,1}$}
-      \put(12,-58){$\mathsf{w}_{1,0}$}
-      \put(42,-39){$\mathsf{w}_{2,1}$}
-      \put(42,-58){$\mathsf{w}_{2,0}$}
-      \put(92,-39){$\mathsf{w}_{n,1}$}
-      \put(93,-58){$\mathsf{w}_{n,0}$}
-
-      \put(119,-49){$\gamma$}
-      \put(129,-49){$\beta$}
-
-%      \put(120,42){$\mathsf{w}_\r$}
-%      \put(133,42){$\mathsf{yield}$}
-
-    \end{picture}
-  \end{center}
-  \caption{SOCA $\Soca_\M$ for the simulation of space bounded Turing
-    machines.}
-  \label{fig:tm}
-\end{figure}
-
-Let $\M=(S, \Sigma, s_0, F, \mu)$ be a DTM with a fixed input tape
-with $m$ tape cells, and $n$ working tape cells, and let
-$S=\{s_0,\ldots,s_k\}$. Figure \ref{fig:tm} shows the SOCA $\Soca_\M$
-that we use for the simulation of $\M$. All transitions of $\Soca_\M$
-are implicitly labelled with $0$.  A simulation starts when $\Soca_\M$
-is entered at the location labeled with $\zeta$ and is finished when
-the location labeled with $\beta$ is reached.
-
-The sequence of propositions occurring on a trace starting from and
-ending in $q$ encodes a configuration of $\M$.  In detail,
-$\mathsf{s}_i$ indicates that $\M$ is in state $s_i$; $\mathsf{ih}_i$
-that the input tape head scans cell $i$; $\mathsf{wh}_i$ that the
-working tape head scans cell $i$; $\mathsf{i}_{i,b}$ that the
-$i^{\text{th}}$ bit of the input tape is set to $b$; and
-$\mathsf{w}_{i,b}$ that the $i^{\text{th}}$ bit of the working tape is
-set to $b$, where $i$ is in the respective range and $b\in \{0,1\}$.
-
-\newcommand{\state}{\mathsf{state}}
-\newcommand{\work}{\mathsf{work}}
-\newcommand{\iNp}{\mathsf{input}}
-\newcommand{\inh}{\mathsf{inhead}}
-\newcommand{\woh}{\mathsf{wohead}}
-\renewcommand{\succ}{\mathsf{succ}}
-
-Let us introduce some LTL formulas that allow for testing properties
-of the current configuration. Think of all of them as being evaluated
-in $q$. The formula $\state_i=\X \mathsf{s}_i$ for each $i\in[0,k]$
-expresses that the current state is $s_i$.  By the formula
-$\inh_{i}=\X\X\X\X \mathsf{ih}_i$ we express that the input head is at
-position $i$, where $i\in[m]$. Similarly, define the formulas
-$\woh_i$, $\work_{j,b}$, and $\iNp_{i,b}$ for expressing that the
-working head is at position $i$, that the $i^{\text{th}}$ bit of the
-input tape is $b$, and that the $j^{\text{th}}$ bit of the working
-tape is $b$, respectively, where $i\in[m]$,$j\in[n]$, and
-$b\in\{0,1\}$.
-
-The LTL formula below, assumed to be evaluated in $\alpha$, ensures
-that the transition function is correctly encoded into traces of
-$\Soca_\M$ for states $s\in S\setminus F$ whenever the input
-respectively working tape head does not scan a start ($\l$)
-respectively end marker ($\r$):
-
-\begin{multline*}
-\bigwedge_{s_l\in S\setminus F}
-\bigwedge_{i\in[m]}
-\bigwedge_{j\in[n]}
-\bigwedge_{b_1,b_2\in\{0,1\}}
-\state_l\wedge\inh_i\wedge\iNp_{i,b_1}\wedge\woh_j\wedge
-\work_{j,b_2} \longrightarrow\\
-\longrightarrow
-\left(
-\left(\X(\neg\alpha \wedge \neg \beta)
-\U(\alpha\wedge \succ(s_l,i,j,b_1,b_2)\right)
-\wedge
-\bigwedge_{j'\not= j}\bigwedge_{b\in\{0,1\}}
-\left(\work_{j',b}\leftrightarrow
-(\X\neg\alpha\U(\alpha\wedge\work_{j',b}))\right)
-\right).
-\end{multline*}
-Here, whenever $\mu(s_l,b_1,b_2)=(s_h,d_1,d_2,b)$, the formula
-\begin{align*}
-\succ(s_l,i,j,b_1,b_2)\quad =\quad\state_h\wedge\inh_{i+d_1}\wedge\woh_{j+d_2}\wedge
-\work_{j,b}
-\end{align*}
-guarantees that the correct bit is ``written'' to the working tape and
-that the state, the input head position, and the working tape position
-of the next configuration seen indeed match the successor
-configuration. A similar formula can be constructed for the case when
-one or both of the input or working heads point to a start
-respectively end marker. Once we have reached a final state $s_i\in
-F$, we require that $\Soca_\M$ is left which is expressed by the
-following formula when evaluated in $q$:
-\begin{align*}
-   \bigwedge_{s_i\in F}\state_i \longrightarrow (\neg \alpha \until
-   \beta).
-\end{align*}
-It is now easily seen that we can construct a formula $\Phi$ that is
-derived from a conjunction of the formulas from above such that the
-formula $\globally (\alpha \rightarrow \Phi)$ constraints paths
-through $\Soca_\M$ in such a way that their traces yield the encoding
-of a valid computation of $\M$.
-
-Let us now address towards ensuring that once we enter $\Soca_\M$ we
-initially traverse it in such a way that the trace corresponds to an
-initial configuration of $\M$. The formula
-\begin{align*}
-  \globally\left(\zeta \longrightarrow \X(\state_0 \wedge \inh_1 \wedge
-  \woh_1\wedge \bigwedge_{1\le j\le n}\mathsf{work}_{j,0} ) \right)
-\end{align*}
-makes sure that the heads of the input and working tape point to the
-first tape cell, that the working tape is filled with $0$s and that we
-are in the initial state. In case the input tape can be initialized
-with an arbitrary content, we are done. Otherwise, suppose that we
-want to transfer the first $j$ bits of the output of a TM $\M'$ from
-its corresponding SOCA $\Soca_{\M'}$ to the input of $\Soca_\M$.  For
-$b\in \{0,1\}$, let $\bar{b}=0$ if $b=1$ and $\bar{b}=1$ otherwise,
-and suppose that all atomic propositions are primed in $\Soca_{\M'}$.
-The formula
-\begin{align*}
-  \bigwedge_{1\le i\le j\atop{b\in \{0,1\}}}
-  \globally
-   ((\mathsf{w}'_{i,b} \wedge (\neg \alpha' \until \beta'))
-  \longrightarrow (\neg \mathsf{i}_{i,\bar{b}}\until \gamma)
-  %\right)
-\end{align*}
-guarantees that we traverse through the first $j$ bits of the
-component of $\Soca_\M$ representing the input tape of $\M$ in the
-same way as we traverse the first $j$ bits of the working tape
-component of $\M'$ in $\Soca_{\M'}$ when a computation has finished.
-Coming back to Figure \ref{fig:high-level}, we have thus seen how the
-SOCA $\Soca_{\mathsf{inc}}, \Soca_{\circuit}$ and
-$\Soca_{\mathsf{prime}}$ and the communication between them can be
-realized. The only major question left open is how we can perform a
-divisibility respectively non-divisibility test of the counter value
-with a prime number computed in $\Soca_{\mathsf{prime}}$. For this,
-let us consider the SOCA $\Soca_{\mathsf{divides}}$ from Figure
-\ref{fig:soca-divisibility}.
-
-\begin{figure}[t]
-  \begin{center}
-    \begin{picture}(140,30)(0,0)
-      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
-      \gasset{Nadjust=wh,Nadjustdist=1}
-      \gasset{curvedepth=0}
-
-      %% \node(WORKIN)(0,45){}
-      %% \node(A)(10,45){$\bullet$}
-      %% \node(B1)(25,55){$\bullet$}
-      %% \node(B2)(25,35){$\bullet$}
-      %% \node(C)(40,45){$\bullet$}
-      %% \node(D1)(55,55){$\bullet$}
-      %% \node(D2)(55,35){$\bullet$}
-      %% \node(E)(70,45){$\bullet$}
-      %% \node(F)(80,45){$\ldots$}
-      %% \node(G)(90,45){$\bullet$}
-      %% \node(H1)(105,55){$\bullet$}
-      %% \node(H2)(105,35){$\bullet$}
-      %% \node(I)(120,45){$\bullet$}
-      %% \node(J)(135,45){$\bullet$}
-      %% \node(DASH)(120,55){}
-      %% \node(WORKOUT)(145,45){}
-
-      %% \drawedge(WORKIN,A){}
-      %% \drawedge(A,B1){$0$}
-      %% \drawedge(A,B2){$0$}
-      %% \drawedge(B1,C){$0$}
-      %% \drawedge(B2,C){$0$}
-      %% \drawedge(C,D1){$0$}
-      %% \drawedge(C,D2){$0$}
-      %% \drawedge(D1,E){$0$}
-      %% \drawedge(D2,E){$0$}
-      %% \drawedge(G,H1){$0$}
-      %% \drawedge(G,H2){$0$}
-      %% \drawedge(H1,I){$0$}
-      %% \drawedge(H2,I){$0$}
-      %% \drawedge(I,J){$0$}
-      %% \drawedge[dash={1.5}0](I,DASH){}
-
-      %% \put(4,42){$\mathsf{work}_\l$}
-      %% \put(21,57){$\mathsf{work_{1,1}}$}
-      %% \put(21,32){$\mathsf{work_{1,0}}$}
-      %% \put(51,57){$\mathsf{work_{2,1}}$}
-      %% \put(51,32){$\mathsf{work_{2,0}}$}
-      %% \put(101,57){$\mathsf{work}_{N,1}$}
-      %% \put(101,32){$\mathsf{work}_{N,0}$}
-      %% \put(120,42){$\mathsf{work}_\r$}
-      %% \put(133,42){$\mathsf{yield}$}
-
-      \node(DIVIDESIN)(0,15){}
-      \node(A)(10,15){$\bullet$}
-      \node(B1)(25,25){$\bullet$}
-      \node(B2)(25,5){$\bullet$}
-      \node(C)(40,15){$\bullet$}
-      \node(D1)(55,25){$\bullet$}
-      \node(D2)(55,5){$\bullet$}
-      \node(E)(70,15){$\bullet$}
-      \node(F)(80,15){$\ldots$}
-      \node(G)(90,15){$\bullet$}
-      \node(H1)(105,25){$\bullet$}
-      \node(H2)(105,5){$\bullet$}
-      \node(I)(120,15){$\bullet$}
-      \node(J)(135,15){$\bullet$}
-      \node(DIVIDESOUT)(145,15){}
-
-      \drawedge(DIVIDESIN,A){}
-      \drawedge(A,B1){$0$}
-      \drawedge(A,B2){$0$}
-      \drawedge(B1,C){$-2^0$}
-      \drawedge(B2,C){$0$}
-      \drawedge(C,D1){$0$}
-      \drawedge(C,D2){$0$}
-      \drawedge(D1,E){$-2^1$}
-      \drawedge(D2,E){$0$}
-      \drawedge(G,H1){$0$}
-      \drawedge(G,H2){$0$}
-      \drawedge(H1,I){$-2^l$}
-      \drawedge(H2,I){$0$}
-      \drawedge(I,J){$\mathsf{\zero}$}
-      \drawbpedge[ELside=r](I,290,28,A,250,28){$0$}
-
-%      \put(4,12){$\mathsf{work}_\l$}
-      \put(22,27){$\prop_{0,1}$}
-      \put(22,2){$\prop_{0,0}$}
-      \put(52,27){$\prop_{1,1}$}
-      \put(52,2){$\prop_{1,0}$}
-      \put(102,27){$\prop_{l,1}$}
-      \put(102,2){$\prop_{l,0}$}
-%      \put(120,12){$\mathsf{work}_\r$}
-      \put(134,12){$\xi$}
-
-    \end{picture}
-  \end{center}
-  \caption{The SOCA $\Soca_{\mathsf{divides}}$  for testing the
-    counter for divisibility with some natural number of
-    bit length $l+1$.}
-  \label{fig:soca-divisibility}
-\end{figure}
-
-One cycle through $\Soca_{\mathsf{divides}}$ subtracts some natural
-number of bit length $l+1$ from the counter. However, in order to test
-for divisibility we need to make sure that we remain on the same path
-in every cycle. In the $\CTL$ setting, this problem was resolved by
-branching into the additional SOCA $\Soca_{\mathsf{bit}}$. In
-contrast, in $\ltl$ we cannot branch, but use the propositions
-$\prop_{j,b_j}, j\in [0,l], b\in \{0,1\}$ in order to stay on
-precisely one path in every cycle. Assuming that the number $p$ for
-which we want to test for divisibility with the current counter value
-is encoded as a sequence of propositions $\mathsf{w}_{j,b_j}$ of some
-SOCA $\Soca_\M$, the subsequent formula enforces that we always
-subtract $p$ in cycles of $\Soca_{\mathsf{divides}}$:
-\begin{align*}
-  \globally \left( \bigwedge_{0\le j\le l \atop{b_{j}\in \{0,1\}}}
-  ((\mathsf{w}_{j,b_j} \wedge (\neg \alpha \until \beta) ) \longrightarrow
-  ( \neg \prop_{j,\bar{b_j}} \until \xi)) \right).
-\end{align*}
-It is straight forward to derive a similar SOCA
-$\Soca_{\mathsf{not\_divides}}$ and an appropriate LTL formula for
-testing non-divisibility of the counter value with a previously
-computed prime number. Finally, we can also adopt these techniques in
-order to correctly handle the branching on $b$ performed in Figure
-\ref{fig:high-level}.
-
-In summary, by taking the disjoint union of all the gadgets from
-Figure \ref{fig:high-level}, their appendent $\ltl$ formula that we
-described in this section, connecting the gadgets correctly and taking
-the conjunction of the relevant $\ltl$ formulas, we can construct
-$\Poca(x)$ and an $\ltl$ formula $\varphi$ in logarithmic space such
-that there is an assignment $\sigma$ assigning a natural number to $x$
-such that $(T(\Poca^\sigma), q_{\mathsf{start}}(0))\not\models
-\varphi$ if and only if $\psi_\circuit$ given by an input $\circuit$
-of $\textsf{Succinct 3-SAT}$ is satisfiable.
-
-\begin{theorem}
-  The combined complexity of $\ltl$ model checking on POCA is
-  $\coNEXP$-complete.
-\end{theorem}
-
-\section{Conclusion}{\label{S Concl}}
-
-In this paper, we have settled the computational complexity of model
-checking CTL and LTL on SOCA and POCA with respect to data and
-combined complexity. Our proofs for providing lower bounds have
-introduced some involved concepts and techniques, which we believe may
-be of independent interest for providing lower bounds for decision
-problems in the verification of infinite state systems.
-
-An interesting aspect of future work could be the consideration of
-\emph{synthesis problems} for POCA. Given a POCA $\Poca(X)$ and an LTL
-formula $\varphi$, a natural question to pose is whether there
-\emph{exists an assignment} $\sigma$ such that $(T(\Poca^\sigma),
-\pi)\models \varphi$ for all infinite paths $\pi$ starting in some
-state of $T(\Poca^\sigma)$.  For CTL, such a problem is undecidable by
-Theorem \ref{thm:ctl-poca}, but we claim that it might be decidable
-for LTL. Moreover, the CTL fragment EF seems to be a good candidate of
-a branching-time logic for which model checking on POCA could be
-decidable, but this remains subject to further investigations.
-
-\bibliography{bib}
-\bibliographystyle{plain}
-
-%\thebibliography
-
-\iffalse
-
-\begin{thebibliography}{1}
-
-\bibitem{GoLo10}
-S.~G\"oller, M.~Lohrey.
-\newblock Branchning-time model checking of one-counter processes.
-\newblock In Proceedings of STACS 2010, to appear
-\newblock {\em Arxiv:0909.1102v1}
-
-\end{thebibliography}
-\fi
-
-\end{document}
+%For reasons of space, we have had to abbreviate or omit a number of
+%proofs; full details can however be found in the technical
+%report~\cite{techrep}.

### 2023-06-11 reachability/bibliography/socamc.tex
remains from pushdown parameter article added in

\documentclass{llncs}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{ifthen}

\begin{document}
\section{PSpace-hardness of CTL Model Checking of Succinct One-Counter
Automata}

\begin{figure}
%\begin{center}
%  \includegraphics[scale=0.4]{ctl-hardness.pdf}
%\end{center}
\caption{Gadget to check whether the $i$-th bit of the counter is
set (assuming no counter values larger than $2^{n+1}-1$).}
\label{fig:checker}
\end{figure}
Let $\mathcal{M}=\langle Q, \{0,1\}, \delta, q_0, F\rangle$ be a TM
which does not use more than $n$ tape cells during a computation,
where $n$ is polynomial in $|\mathcal{M}|$. Assume w.l.o.g.\ that
$\mathcal{M}$ does not make any right move when the head reads the
$n$-th tape cell and that it does not make any left-move when reading
the first tape cell. We define a succinct one-counter automaton
$\mathcal{A}=\langle P, \Delta \rangle$ and a CTL-formula $\varphi$
such that $\mathcal{M}$ reaches a final state if and only if
$\mathcal{A}\models \varphi$. The set of states $P$ of $\mathcal{A}$
contains elements $q_{i,b}$, where $q\in Q$, $1\le i\le n$ and $b\in
\{0,1\}$. Intuitively, when simulating a run of $\mathcal{M}$, we keep
the tape content in the counter and store the head position in the
states of $\mathcal{A}$. Thus, when $\mathcal{A}$ is in state
$q_{i,b}$, we ``mean'' that $\mathcal{M}$ is in state $q$, with head
in position $i$, reading $b$. Of course, we have to ensure that $b$ is
actually set respectively unset in the current counter value, but let
us postpone this problem for a moment. We define the transition
relation for $q_{i,b}$ as follows: First, suppose $b=0$ and consider
$\delta(q,0)=(q',b',d)$ ($q'$ is the next state, $b'$ is the content
to be written at the current tape cell and $d\in\{-1,1\}$ is the
direction the head moves). If $b'=0$, then $\Delta$ contains
$\{(q_{i,0},+0,q_{i+d,0}), (q_{i,0},+0,q_{i+d,1})\}$ and if $b'=1$ the
$\Delta$ contains $\{(q_{i,0},+2^i,q_{i+d,0}),
(q_{i,0},+2^i,q_{i+d,1})\}$. Intuitively, we set the $i$-th bit of the
counter accordingly, move the head and guess the content of the tape
cell we move to. The case $b=1$ is now straight-forward: If $b'=0$
then $\Delta$ contains $\{(q_{i,0},-2^i,q_{i+d,0}),
(q_{i,0},-2^i,q_{i+d,1})\}$ and if $b'=1$ the $\Delta$ contains
$\{(q_{i,0},+0,q_{i+d,0}), (q_{i,0},0,q_{i+d,1})\}$. Furthermore, each
state $q_{i,b}$ is labelled with a propositional variable $M$, which
indicates that those states come from $\mathcal{M}$. We now need a
gadget that can check for us that we guess the tape contents
correctly. But this is easy, consider the automaton shown in Figure
\ref{fig:checker}. It can reach the rightmost state only if the $i$-th
bit is set to one. Likewise, we can construct a similar automaton such
that the rightmost state can only be reached if the $i$-th bit of the
counter value is set to zero. Let $p_i^+$ respectively $p_i^-$ be the
initial states of the gadgets that can test whether the $i$-th bit of
the counter is set respectively not set. Then for each $q_{i,0}$ we
have a transition $(q_{i,0}, 0,p_i^-)\in \Delta$, and for each
$q_{i,1}$ we have $(q_{i,1}, 0, p_i^+)\in \Delta$. Finally, for each
$q\in F$ we have a transition $(q_{i,b}, +0, p_f)\Delta$ to a special
state $p_f\in P$ that is labelled with the proposition $F$.

Let us define $\varphi=\mathbf{E}((M\wedge \mathbf{EX}(\neg M \wedge
\mathbf{EX}^n \top)) \mathbf{U} F)$. It is now not difficult to check that
$\mathcal{A}\models \varphi$ if, and only if, $\mathcal{M}$ reaches a
final state.
\end{document}

### 2023-06-11 reachability/definitions.tex
remains from pushdown parameter article added in

\chapter{Definitions}\label{chapter definitions}

In this chapter, we introduce general notations and preliminary definitions.
First, we concern ourselves with basic mathematical objects and notations: 
numbers, words, (partial) functions.
Secondly, we introduce the general definition of transition systems.
Thirdly, we %[Parity games]
	define parity games.
Fourth, %we %[Logics] 
%		define the logics used in the thesis, the monadic-second order logic ($\MSO$) along with the first-order logic ($\FO$).
%Fifth, % [DTMs]
	we %define deterministic Turing Machines, and 
	define the complexity classes that will be of interest to us. 
%	We refer to \cite{Papa94,AroBar09} for further details on complexity. 
Lastly, % [Automata]
	we recall the definition of automata.

\iffalse
We assume the reader is familiar with 
Turing machines and standard complexity
classes such as $\LOGSPACE$, $\PSPACE$ and
$\EXPSPACE$. We refer to \cite{Papa94,AroBar09}
for further details on complexity.
\fi

%\subsection{Preliminaries}
\section{Preliminaries}

For any two sets $X$ and $S$, let $X^S$ denote the set of all functions from $S$ to $X$.
For any set $S$ let 
$\mathscr{P}(S) = \{ X \mid X \subseteq S \}$
denote the {\em power set} of $S$.
If a set $S$ is finite, we denote its {\em size} by $|S|$.

% \subsubsection{Numbers}
\subsection{Numbers}

By $\Z $ we denote the {\em integers} and by $\N=\{0,1,\ldots\}$ we denote the {\em non-negative integers}.
For every $a,b\in\Z$ with $a\leq b$ we define $[a,b]=\{k\in\Z\mid a\leq k\leq b\}$.
For every $n\geq 1$ we define $n\Z=\{n\cdot z\mid z\in\Z\}$.
For every number $n\in\N$ we define $\log(n)=\min\{i+1\mid i\in\N, n\leq 2^i\}$,
which is the 
%smallest 
number of bits necessary to write down $n$ in binary.
For every finite set $M\subset\N\setminus\{0\}$ let 
$\LCM(M)=\min\{n\geq 1\mid \forall m\in M: m | n\}$
denote the least common multiple of the elements in $M$. 
For any $j \in \N$ let 
$\LCM(j)=\LCM([1,j])$ denote the 
{\em least  common multiple} of the numbers $\{1,\ldots,j\}$.
For a real $r \in \R$, $\lfloor r \rfloor$ is the greater integer $z$ such that $z \leq r$.

% \subsubsection{Alphabets and words}
\subsection{Alphabets and words}

For every set $A$ we denote by $A^*$ the set of finite
sequences of elements of $A$, and we denote by $A^\omega$ the set of infinite sequences of elements of $A$. If the set $A$ is finite, $A$ will be called an {\em alphabet},
and we call the elements of $A^*$ {\em words} over $A$,
and the elements of $A^\omega$ {\em infinite words} over $A$. 
We denote
the union of $A^*$ and $A^\omega$ by
$A^\infty$.
We denote the {\em empty word} by $\varepsilon$.
For all $a\in A$ and all $w\in A^*$ let $|w|_a$ denote the number of occurrences
of the letter $a$ in $w$, while $|w|$ denotes the {\em length} of the word.
We denote by $A^n$ the set of words of length $n \in \N$ over %an alphabet 
$A$.

For two words $u, v \in A^*$, we denote by $u \cdot v$ (often abreviated $uv$) the
concatenation of the two sequences. We say $u$ is a {\em prefix} (resp. a {\em suffix}) of $v$ if there exists a word $w \in A^*$ such that $v = u \cdot w$ (resp. $v = wu$).

For a word $w = a_0 a_1 a_2  \ldots  a_{n-1}$ where $a_i \in A$ for all $i \in [0,n-1]$, we denote 
$a_i$ by $w[i]$ for $0 \leq i \leq n - 1$ and call it the {\em letter} at {\em position $i$}.
For each language $L\subseteq A^*$ let $\chi_L:A^*\rightarrow\{0,1\}$ denote its characteristic function defined as
$$
\chi_L(w)=\begin{cases}1&\text{if $w \in L,$}\\
	0 &\text{otherwise}.
\end{cases}
$$

%\subsubsection{Partial functions}
\subsection{Partial functions} \label{partial}

A {\em partial function} $f$ from a set $S$ to a set $X$ 
%is a function defined on a subset $C$ of $S$ (possibly $S$ itself) with output values in $X$, and 
that is defined on a subset $C \subseteq S$
is denoted by
$ f : S \rightharpoonup X $.
We call $C$ (resp.  $f(C)$) the {\em domain} of $f$ (resp. {\em image}) and write it 
$\text{Dom}(f)$ (resp. $\text{Im}(f)$).
If $C=S$ then %the function is defined on every element of $S$ and
$f$
 is called {\em total}.

\iffalse
Specifically, for a partial function $ f : S \rightharpoonup X $, and any $s \in S$, one has either
\begin{itemize}
\item   $ f(s)=x\in X$ (it is a single element in $X$), or
\item   $ f(s) $ is undefined%, in which case we consider $f$ as returning the bottom element $\bot$
.
\end{itemize}
\fi

For a partial function $ f : S \rightharpoonup X $, for notational purposes, we will consider some element $\bot_X \not\in X$ and 
$f$ can alternatively be associated with the function returning the 
bottom element $\bot_X$ when it is undefined. Thus we write $(X \biguplus \{ \bot_X \})^S$ for the set of all partial functions from $S$ to $X$, which we sometimes abbreviate as $(X \biguplus \{ \bot \})^S$% when $\bot_X \notin X$
.

For a partial function $ f : S \rightharpoonup X $, the {\em inverse image} of an element $x \in X$,
denoted by $f^{-1}(x)$, is the set
$\{ s \in S \mid f(s) = x \}.$

We say a function $f : \N \to \N$ is {\em monotonic non-decreasing} if $f$ for all $i,j \in \N$ 
 such that $i \leq j$, one also has $f(i) \leq f(j)$.
Let $f$ and $g$ denote total functions from $\N$ to $\N$. We say $f$ is 
{\em asymptotically bounded by $g$}
and write $f(n) \in O(g(n))$ if there exists a constant $M \in \N$ and
a natural number $n_o$ such that
$$ f(n) \leq M \cdot g(n)$$
holds for all $n \geq n_o$. We extend the notation to 
composition of functions i.e. if $f, g, h$ are all total functions from $\N$ to $\N$,
 $h$ is monotonic non-decreasing and $f(n) \in O(g(n))$, 
 then we sometimes write $h(f(n)) \in h(O(g(n)))$ for $h(f(n)) \in O(h(g(n)))$.

%\subsection{Transition systems}
\section{Transition systems}

A {\em labeled transition system} (LTS for short) is a tuple $T=(S, \Lambda, \rightarrow)$ where 
$S$ is a set of {\em configurations}, $ \Lambda$ is a set of {\em labels}, and 
${\rightarrow} \subseteq S\times \Lambda \times S$ is a 
ternary relation,
denoted as the set of {\em labeled transitions}. 
We
 prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as
%       $$p \ {\overset {a }{\rightarrow }} \ q$$
       $s  \xrightarrow{a}  s'$
to represent a transition from configuration $s$ to configuration $s'$ with label $a$. \\

\noindent
Labels can be used to represent the reading of an input, but also to represent an action performed during the transition or conditions that must hold in order to allow the use of the transition.

\iffalse
Labels can represent different things depending on the language of interest. Typical uses of labels include representing input expected, conditions that must be true to trigger the transition, or an action performed during the transition.
\fi

A {\em path} in a labeled transition system from a {\em source configuration} $s_0$
to a {\em target configuration} $s_n$ is a sequence 
$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$. 
We define the {\em concatenation} $ \pi_1 \pi_2$ of 
two paths $\pi_1$ and $\pi_2$ when the source configuration of $\pi_2$ is equal to the target configuration of $\pi_1$
as expected.
The {\em length} of 
$\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1} } s_n$
is defined as $|\pi|=n$. We say the path is {\em labeled} by $a_0 a_1 , \ldots a_{n-1}$.
For all $w \in \Lambda^*$, all $s,s' \in S$, we will write $s \xrightarrow{w } s'$ if there exists a path from $s$ to $s'$ labeled by $w$.

An {\em infinite path} is an infinite sequence
 $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots $.
For each infinite (resp. finite) path  $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ 
(resp. $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots \xrightarrow{a_{n-1}} s_n$)
and $i,j \in \N$ (resp. $i,j \in [0,n]$) with $i<j$ we denote
by $\pi[i,j]$ the path 
$s_i  \xrightarrow{a_i } s_{i+1}  \xrightarrow{a_{i+1} } \cdots  \xrightarrow{a_{j-1} } s_j$
 and by $\pi[i]$ the configuration $s_i$.
% 
% As expected, a {\em prefix} (resp. {\em suffix}) of a finite path $\pi$ is a path of the form $\pi[0,j]$ (resp. $\pi[i,n]$).  \\ 
As expected, a {\em prefix} of a finite or infinite path $\pi$ is a finite path of the form $\pi[0,j]$, and
a  {\em suffix} of a finite path $\pi$ is a path of the form $\pi[i,n]$.  \\
Given an infinite path $\pi = s_0 \xrightarrow{a_0 } s_1 \xrightarrow{a_1 } \cdots$ let
$\text{\textit{Inf}}(\pi) = \{ s \in S \mid \forall i ~ \exists j > i ~ s_j = s\}$.

\noindent
The set of {\em successors} of a configuration $s \in S$ is deﬁned as
$\{ s' \in S \mid \exists a \in \Lambda ~ s \xrightarrow{a} s'\}$.
A configuration without successors is called a {\em dead end}.  \\

A labeled transition system $(S, \Lambda, \rightarrow)$ is {\em deterministic} if for all configurations $s_1, s_2, s_3 \in S$ and all
$a \in \Lambda$,
 $ s_1  \xrightarrow{a} s_{2}$ and  
 $s_1  \xrightarrow{a} s_{3} $ implies $s_2 = s_3 $. \\

An {\em (unlabeled) transition system} is a pair $T = (S,\rightarrow )$ where $S$ is a set of 
{\em configurations} and  
$ {\rightarrow} \subseteq S \times S$ is a
binary relation 
on
the set of configurations, denoted as the set of {\em transitions}. 
We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from configuration $s$ to configuration $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline
Note 
 that an unlabeled transition system can be seen as a labeled transition system where the set of labels consists of only one element. 
Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
then defined as expected.

\iffalse
\textcolor{red}{
We are interested in the following decision problem.
%
\problemx{Reachability}
{A labeled transition system $T = (S, \Lambda, \rightarrow)$, two configurations $s,s' \in S$.}
{Is there a path in $T$ with source configuration $s$ and target configuration $s'$ ?\newline}
}
\fi

%\subsection{Games}
\section{Games}

% Games represent a model for interaction between decision makers. We study so-called turn-based games in extensive form with perfect information where the players have to make decisions again and again.
\iffalse
After its foundation by von Neumann and Morgenstern [vNM44] in the
1940s, game theory has quickly become a valuable tool for modelling the
interaction of several agents with individual and possibly conflicting objec-
tives. It has applications in many scientific fields such as economy, sociology,
biology, logic and computer science. Whereas the games used in the former
three areas are usually finite in the sense that any play of these games ends
after a finite number of steps (in fact often after only a single step), the
games used in logic and computer science are in general infinite.

@article{copeland1945john,
  title={John von Neumann and Oskar Morgenstern, theory of games and economic behavior},
  author={Copeland, Arthur H},
  journal={Bulletin of the American Mathematical Society},
  volume={51},
  number={7},
  pages={498--504},
  year={1945},
  publisher={American Mathematical Society}
}

More precisely, an infinite game is played for an infinite number of
rounds. In each round, one player chooses an action depending on the se-
quence of actions chosen in previous rounds. After infinitely many rounds,
an infinite sequence of actions has emerged, called the outcome of the game.
Each player receives a payoff determined by a payoff function mapping each
possible outcome to a real number in the interval [0, 1]. We will concen-
trate on games where the possible payoffs are just 0 and 1, in which case a
winning condition, given as an abstract set of possible outcomes, is used to
determine whether a player wins (payoff 0) or loses (payoff 1) an outcome.
In logic, games are used to evaluate a logical formula in a mathematical
structure and to compare mathematical structures with respect to their
(in)distinguishability by a certain logic. Games of the former kind are called
model checking games; games of the latter kind are called model comparison
games. Whereas model checking and model comparison games for plain
modal or first-order logic are finite games, infinite games arise as model
checking games for fixed point logics and as model comparison games for
infinitary logics.
In computer science, games are used for the verification and synthesis
of reactive systems, i.e. systems that interact with their environment. As
soon as the systems under consideration do not necessarily terminate, these
games become infinite. For example, the behaviour of an operating system
can be understood as an infinite game between the system and its users.

So far, the games used in logic and computer science have usually been
two-player zero-sum games, i.e. games with only two players where one
player wins if and only if the other player loses. For example, in a model
checking game, one player wants to show that the given formula holds in the
given structure whereas the other player wants to show that this is not the
case. Analogously, in a game used for verification, one player wants to show
that the given system is able to react against its environment in such a way
that the resulting execution path of the system fulfils the given specification
whereas the other player wants to show that this is not the case.

The research on infinite games in theoretical computer science is based on a
mixture of several motivations:
1. The beautiful classical theory of infinite games, as developed in descriptive
set theory, lacks an algorithmic content. Such an algorithmic orientation is
provided by the automata theoretic approach to infinite games, originating
in work of Church, B ̈uchi, McNaughton, and Rabin about fourty years ago.
2. Determinacy results for infinite games are closely related to complementation
results for logics and automata; thus, infinite games help to analyze logical
theories (the most prominent being the monadic theory S2S of two successor
functions, see e.g. [Th97]).
3. Games are a natural model of reactive computation, and infinite games are
thus a faithful representation of nonterminating reactive systems (for which
control problems can be solved in terms of providing winning strategies).
4. The model checking problem for logics like the μ-calculus can be formulated
as the question to determine the winner of an infinite game.
The purpose of this tutorial is to explain the core of the algorithmic (and
automata theoretic) theory of infinite games. In the present extended abstract,
only some key notions are explained, without technical details and proofs. In the
tutorial, more will be said about topics of current research as listed in Section 6
of this abstract.

@incollection{thomas1997languages,
  title={Languages, automata, and logic},
  author={Thomas, Wolfgang},
  booktitle={Handbook of formal languages},
  pages={389--455},
  year={1997},
  publisher={Springer}
}

\fi

A game is composed of an arena and a winning condition.
\iffalse
\mh{Two-player games oﬀer a very convenient framework to represent
interaction of a program with some (possibly hostile) environment. In this approach, the ﬁrst player represents the program while the second player simulates
the environment. A winning strategy expresses a property that must hold whatever the environment does, thus ﬁnding a winning strategy for the ﬁrst player
allows to synthesize a controller that restricts the program and ensures that the
property expressed by the winning condition always holds \cite{arnold2003games}.}
\fi
We will ﬁrst study
arenas and
 then introduce common winning conditions that will be of interest to us.

% \subsubsection{Arenas}
\subsection{Arenas}

\begin{samepage}
An {\em arena} is a tuple $A=(S_0, S_1,{\rightarrow})$  which is composed of
\begin{itemize}
	\item two  disjoint  sets of configurations, $S_0$ and $S_1$, %respectively associated with player $0$ and $1$, 
	whose disjoint union $S_0 \cup S_1$ we denote by $S$, and
	\item a relation of %labeled 
		transitions
	 ${\rightarrow} \subseteq S \times %\Lambda \times 
	 				S$. \\
\end{itemize}
\end{samepage}

\noindent
Note that $(S,{\rightarrow})$, where $S=S_0 \uplus S_1$, forms
a %labeled 
transition system, and, in particular, given a 
%labeled 
transition system,
one needs only to provide 
a partition of the set of configurations $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\
% We require that the out-degree of each configurations is at least one. This ensures that every finite path in (V, E) can be prolonged.

% \mh{I dont think I want to disallow dead-ends from the very beginning, means every time you define some arena you have to prove it has no dead-ends, seem tedious and unecessary.}

\noindent
The games we are interested in are played by two players, called {\em player $0$} and {\em player $1$}. We will often write player $i$ to denote a general player for $i \in \{0,1 \}$,
and we will call player $1-i$ its {\em opponent}.

% Observe that there is no restriction on the number of the successors of a configuration in an arena. Also, we don’t require that (V, E) is a bipartite graph with corresponding partition {V 0 , V 1 }.

\iffalse
\noindent
In the case where $\Lambda$ is a singleton $\{ a\}$, we will abuse notation
and denote the arena by
$(S_0, S_1,{\rightarrow},\Omega)$, and we will see $(S,{\rightarrow})$
as an unlabeled transition system.
\fi

% \subsubsection{Plays}
\subsection{Plays}

\noindent
%
%\noindent
A {\em play} in an arena $A=(S_0,S_1,\rightarrow)$ is a path in $(S,{\rightarrow})$ that is {\em maximal} in the following sense: it is either infinite or finite and if it is finite, then the target configuration is a dead end. 
A {\em partial play} in an arena $A=(S_0,S_1,\rightarrow)$ is a 
% sequence of configurations $s_{init} = s_0 s_1 \ldots \in S^\infty$ connected by the transition relation, i.e. such that $(v_i, v_{i+1}) \in T$.
prefix of a play in $A$.\\
%finite path in $(S,{\rightarrow})$.\\

\noindent
Essentially plays can be seen as this: the two players, player $0$ and player $1$, move along the labeled transition system, taking turns either inﬁnitely often or until a dead end is reached.

% \subsubsection{Strategies}
\subsection{Strategies}

% Before deﬁning formally what it means for a player to win a game, we need to introduce the notion of strategy.

A {\em strategy} for player $i \in \{ 0,1\}$ is a function $\sigma_i : S^* S_i \to S$ such
that $\sigma_i(ws)$ is a successor of $s$.
%
A strategy is called {\em memoryless} if its output only depends on the final configuration of the sequence $s \in S_i$, i.e. if $\sigma_i(ws) = \sigma_i(w's)$ for all $w,w' \in S^*$. Thus, a memoryless strategy for player $i$ can be written  as a function $\sigma_i: S_i \to S$.

A partial play $\pi = s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_n$ is {\em consistent} with a strategy $\sigma_i$ if sequences of configurations that end in $S_i$ along this play have all successors according to the strategy, i.e.
if for all $j\in[0,n-1]$, for all element $v_j \in V_i$ in the sequence $\pi$,
$v_{j+1} = \sigma_i(v_j)$.
 Given a configuration $s$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play starting with $s$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $s$. It is 
denoted by $\pi(s, \sigma_0, \sigma_1)$.\\

%\noindent
%Note that the above definitions are essentially the same for finite and infinite arenas.

% \subsubsection{Winning conditions}
\subsection{Winning conditions}

\newcommand{\win}{\textsc{Win}}

Once we have provided players with an arena, it remains to define  properly what the winning condition is in order to define a game. Given an arena $A = (S_0, S_1, \rightarrow)$, a {\em winning condition}
%$\win \subseteq S^\infty$
$\win$
is a subset  
of the set of maximal plays for $A$. A {\em game} is a pair $\mathcal{G}= (A, \win)$ which is composed of an arena $A$ and a winning condition $\win$.
Although a winning condition is in general an infinite object we are interested in winning conditions 
that can finitely be described.

\iffalse
We consider here several types of winning conditions. 
Essentially the 
 mapping $\Omega$ of an arena serves to colour the configurations of the labeled transition system.  Winning conditions then
 consist in acceptance conditions on colour sequences.
 \fi

A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $0$ if 
% no matter which
% strategy $\sigma_{1}$ the other player pursues,
% the play $\pi(s, \sigma_0, \sigma_1)$ is in $\win$.
every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
 from $s$} for player $1$ if 
% no matter which
% strategy $\sigma_{0}$ the other player pursues,
% the play $\pi(s, \sigma_0, \sigma_1)$ is not in $\win$.
every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.

\iffalse
This definition of winning strategies is independent from the winning condition itself, and will also apply to %B\"uchi games and min-
parity games. To differentiate, we mention the winning conditions of a winning strategy from $s_{init}$.
\fi

% \paragraph{Reachability Games}
\paragraph{Reachability Games}

We first consider reachability games: 
the winning condition $\win_F(S)$ is given by a set of final configurations
$F \subseteq S$, and $\win_F(S)$ is the set of maximal plays in $S^* F S^\infty$. 
We simply write
$\win_F$ in case the set $S$ is obvious from context.
To indicate that the winning condition of a game is a reachability winning condition, we will speak of {\em reachability games}.

With regards to reachability games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline

\begin{samepage}
\problemx{Reachability game}
{A reachability game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_F)$, an initial configuration $s \in S_0 \cup S_1$.}
{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
\end{samepage}

% \paragraph{Parity Games}
\paragraph{Parity Games}

Secondly, we consider min-parity games: 
the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
 $\Omega : S \to  [0, m]$.
An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
%this path
the sequence 
 is even, i.e.
 %$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even. 
% The winner of a finite play is the player whose opponent is unable to move, i.e. a play such that the last configuration of the play is in $S_{0}$ (resp. $S_1$) is player $1$ (resp. player $0$). 
A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
We simply write
$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
To indicate that the winning condition of a game is a parity winning condition, we will speak of
 {\em parity games}.

With regards to parity games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline

\begin{samepage}
\problemx{Parity game}
{A parity game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_\Omega)$, an initial configuration $s \in S_0 \cup S_1$.}
{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
\end{samepage}

%
\iffalse
Solving a parity game played on a finite graph means deciding, for a given starting position $v_{init}$, which of the two players has a winning strategy. It has been shown that this problem is in $\NP$ and 
co-$\NP$, more precisely $\UP$ and co-$\UP$. 
% Marcin Jurdziński (1998), "Deciding the winner in parity games is in UP∩ co-UP" (PDF), Information Processing Letters, Elsevier, 68 (3): 119–124, doi:10.1016/S0020-0190(98)00150-1
% @article{jurdzinski1998deciding,
%  title={Deciding the winner in parity games is in UP∩ co-UP},
%  author={Jurdzi{\'n}ski, Marcin},
%  journal={Information Processing Letters},
%  volume={68},
%  number={3},
%  pages={119--124},
%  year={1998},
%  publisher={Elsevier}
%}
\fi
%

% Muller condition: Inf() \in F

% Rabin condition :

% Streett Condition

% Rabin Chain condition

% max parity condition

% 1-winning  == reachability

\begin{center}
	\begin{figure}
			\hspace{3.81cm}
\includegraphics[width=0.4\textwidth]{figures/parity_game}
	\caption{An example of an arena. Circles  belong to player $0$, whereas diamonds belong to player $1$. Arrows represent transitions from a configuration to the next. The priority assigned by the priority mapping to each configuration is a number in $\{0,1,2,3\}$, next to the configuration it is assigned to. Note that we do not have a dead end in our example.}
	\label{example_parity_game}
	\end{figure}
\end{center}

\begin{example}
Let us consider the arena presented in  Figure~\ref{example_parity_game}. 
Configurations are partitioned into two sets: the circles, belonging to player $0$, and the diamonds,
belonging to player $1$. 
The priorities are $\{ 0,1,2,3\}$.
% The transitions and priority mappings may be derived from the picture.
% Note that we don't have a dead end in our example. 
Consider the game on the arena, where the winning condition is a parity condition.
Plays $\pi_1 = q_0 q_1 q_2 (q_4 q_7)^\omega$ and 
$\pi_2 = (q_0 q_1 q_2 q_4)^\omega$ are both winning for player $0$. Indeed, in $\pi_1$, 
the minimal priority encountered infinitely often is %that of $q_7$, that is, 
				$2$, 
whereas in $\pi_2$, it is %that of $q_2$, that is, 
			$0$. 
In both cases, it is an even number. On the other hand,
the plays $\pi_3 = (q_5 q_3 q_6)^\omega$ and $\pi_4=(q_5 q_6)^\omega$ are both winning for player $1$: the minimal priority encountered is odd, since it is $1$ for both.

One can take matters further: on this arena, player $0$ has a winning strategy for the parity game from $q_0$, $q_1$, $q_2$, $q_4$ and $q_7$, but not from $q_5$, $q_3$ nor $q_6$, where player $1$ has a winning strategy. 
\end{example}

% \mh{Say something abut positional determinacy}
One important property of parity games is that of {\em memoryless determinacy}: for a parity game
$\mathcal{G}= ((S_0,S_1,\rightarrow),\win_\Omega) $ and an initial configuration $s \in S_0 \cup S_1$, one of the players has a memoryless winning strategy from $s$ \cite{zielonka1998infinite}.

%\mh{
Of note is that several model checking problems %for pushdown automata 
can be expressed as decision problems for games: the most famous example is that the modal $\mu$-calculus model checking problem is polynomially equivalent to solving the parity game problem. 
%}

% \subsection{Complexity}
\section{Complexity}

We assume the reader is familiar with 
Turing machines% and standard complexity classes such as $\LOGSPACE$, $\PSPACE$ and $\EXPSPACE$. We refer to \cite{Papa94,AroBar09} for further details on complexity
.
%\iffalse
A proper definition of space bounded deterministic Turing machines is provided in 
Section~\ref{DTMs}, 
where a more thorough analysis of their behavior is required.
%\fi

We now provide an overview of time and space complexity classes. 
% The {\em running time} of a Turing machine is a function $f : \N \rightarrow \N$ such that for any input word $w$ of size $n$, the length of any computation on $w$ is at most $f(n)$. 
We say a Turing machine is
{\em $f(n)$-time bounded} if %its running time is $f$.
for any input word $w$ of size $n$, the length of any computation on $w$ is at most $f(n)$. 
We say a Turing machine is
{\em $f(n)$-space bounded} if $f(n)$ is the size of its working tape 
for any computation of the Turing machine on an input of length $n$.

% In the following, we assume the reader is familiar with other complexity classes such as
By $\P$, and $\EXP$ we denote the classes of all problems that can be decided by a deterministic Turing machine who is polynomially or exponentially time bounded, respectively and
by
 $\NP$, and $\NEXP$ we denote the classes of all problems that can be decided by a non-deterministic Turing machine that is polynomially or exponentially time bounded, respectively.

By ${\sf L}$, $\PSPACE$, and $\EXPSPACE$ we denote the classes of all problems that can be decided
by a deterministic Turing machine that is logarithmically, polynomially, exponentially space bounded, respectively. We do not explicitly define 
$\NPSPACE$ and $\NEXPSPACE$
since by Savitch’s theorem \cite{Sav70} they are equivalent to 
$\PSPACE$ and $\EXPSPACE$ respectively.
We do however define ${\sf NL}$ as the 
class of all problems that can be decided
by a nondeterministic Turing machine that is logarithmically space bounded.

% We abuse notations and say a language $L$ is in a complexity class if the problem of whether or not a word $w$ belongs to $L$ is itself in the complexity class.
%
%

% \def\powertower#1#2{#1\ifnum#2>1 ^{\powertower{#1}{\numexpr#2-1\relax}}\fi}
% $\powertower{2}{5}$

We define the tower function $T : \N \times \R \to \R$ by $T(0,r) =r$ and
$T(h+1,r)= 2^{T(h,r)}$ for all $h \in \N, r \in \R$. Thus $T(h,r)$ is a tower of $2$s of height $h$ with an $r$ sitting on top, i.e.

\iffalse
$$\text{height } h \begin{cases}&  2^{2^{2^{2^{2^{2^{2^
	 {{\hspace{0.13cm}}^{\cdot^{{\hspace{0.13cm}}^{\cdot^{
	 {\hspace{0.13cm}}^{\cdot^{{\hspace{0.13cm}}^{2^{2^{2^{2^r}}}}}}}}}}}}}}}}} 
	 = T(h,r)
	\end{cases} $$
\fi

$$	 \left.
    \begin{array}{ll}
T(h,r) = 2^{2^{2^{2^
	 {{\hspace{0.13cm}}^{\cdot^{{\hspace{0.13cm}}^{\cdot^{
	 {\hspace{0.13cm}}^{\cdot^{{\hspace{0.13cm}}^{2^r}}}}}}}}}}} 
    \end{array}
\right \} \text{height } h $$

% https://math-linux.com/latex-26/faq/latex-faq/article/horizontal-and-vertical-curly-braces-left-right-underbrace-and-overbrace

Observe that for all $n,h \in \N$ with $n\geq 1$, we have
$T(h, log^{(h)}(n)) = n$.
%
% 
Here a problem is in {\sf ELEMENTARY} if there exists $h \in \N$ such that
it can be solved in time $O ( T(h, 0) )$. By $h$-$\EXP$ we denote the class of all problems that can be decided by a deterministic Turing machine who is $T(h,f(n))$-time bounded for some polynomial function $f$.
% k-EXP is the union of DTIME(2^{n^i}_k ), for i ≥ 1.
The complexity class %$h$-$\NEXP$, 
$h$-$\EXPSPACE$ 
%and $h$-$\NEXPSPACE$ 
is defined analogously.

Finally, by $\PSPACE^\NEXP$ we denote the class of all problems that can be decided by a deterministic Turing machine that is polynomially space bounded, and that has access to results from an oracle  in $\NEXP$.

A true/false problem is {\em decidable} if there exists a 
time bounded Turing Machine 
% \textcolor{black}{for which all computations end in a finite number of steps and answer yes or no to the question asked by the problem.} 
that answers the question asked by the problem.
If there exists no such Turing Machine, the problem is {\em undecidable}. \\

\par\noindent\ignorespacesafterend
For a complexity class $C \in \{ \NP, \PSPACE, \NEXP, \PSPACE^\NEXP, \EXPSPACE, h-\EXP, h-\EXPSPACE\}$, we say a problem $P$ is {\em $C$-hard} when every problem in $C$
can be reduced in polynomial time to $P$. For a complexity class $C \in \{{\sf NL}, \P\}$
we 
say a problem $P$ is {\em $C$-hard} when every problem in $C$
can be reduced to $P$ using logarithmic space only.
When a problem $P$ is both $C$-hard and in $C$,
we call it {\em $C$-complete}.

 %
For more information on Turing machines in the broader definition, we refer the reader to  \cite{Papa94,AroBar09}, which provide further details on complexity theory as well. 
% Space bounded deterministic Turing machine will be introduced in more details in Section~\ref{DTMs} where a more thorough analysis of their behavior is required.

% \subsection{Automata}
\section{Automata}

\begin{samepage}
A {\em finite automaton}  is a tuple
$\A=(Q,\Sigma,R,q_{init},F)$, where
\begin{itemize}
	\item $Q$ is a finite {\em set of  states}, 
	\item $\Sigma$ is a {\em finite input alphabet},
	\item $R\subseteq Q\times \Sigma \times Q$
	is a %finite {\em set of  rules},
		{\em transition relation},
	\item $q_{init}\in Q$ is the {\em initial  state}, and 
	\item $F\subseteq Q$ is a {\em set of final  states}. \newline
\end{itemize}
\end{samepage}

\noindent
A finite automaton $\A=(Q,\Sigma,R,q_{init},F)$
 induces the finite labeled transition system $(Q, \Sigma, R)$.

The {\em size} of $\A$ is defined as $|\A| \ = \ |Q|+|\Sigma|+|R|$.

A {\em run} from $q_0$ to $q_n$ in $\A$ is a 
path in the labeled transition system $(Q, \Sigma, R)$ induced by $\A$,
and will be noted
$q_0 \xrightarrow{a_0} q_1 \xrightarrow{a_1} \cdots \xrightarrow{a_{n-1}} q_n$,
for $a_0, \ldots, a_{n-1} \in \Sigma$.
We sometimes use the abbreviation $q \rightarrow^* q'$
to denote a run of arbitrary length from $q$ to $q'$.
We say $\pi$ is {\em accepting} if
 $q_0 = q_{init}$
 and $q_n \in F$.
% We say {\em reachability holds} for $\A$ if there is an accepting path in $\A$.

We say a word $w \in \Sigma^*$ is accepted by $\A$ if 
$q_{init} \xrightarrow{w} q$ in $(Q, \Sigma, R)$
for some $q \in F$. % \mh{defined in the section about Transition systems}
%
The {\em language} of $\A$ is $L(\A)$ the
set of elements accepted by $\A$.
We say a language $L$ is {\em regular} if there exists a finite automaton $\A$
such that $L= L(\A)$.

\begin{center}
	\begin{figure}
			\hspace{3.81cm}
\includegraphics[width=0.4\textwidth]{figures/example_automata}
	\caption{A finite automaton that accepts $\{a,b\}^*ab\{a,b\}^* $. The automaton consists of three states, the input alphabet is $\{a,b\}$. The transitions are represented by arrows labeled with the corresponding input symbol. The initial state is $q_0$ and the set of final state is composed of the singular  state $q_f$.}
	\label{example_automata}
	\end{figure}
\end{center}

\begin{example}
	See Figure~\ref{example_automata} for an example of an automaton that accepts the language $\{a,b\}^*ab\{a,b\}^* $.
\end{example}

We are interested in the following decision problem.

\problemx{Automata reachability}
{A finite automaton $\A$.}
{Does 
%reachability hold for
there exists an accepting run in
 $\A$?\newline}

We refer the reader to \cite{Har78} for more details on
 %(deterministic) 
 finite automata and regular languages. \newpage

### 2023-06-11 reachability/pushdown.tex
remains from pushdown parameter article added in

\chapter{Parametric pushdown automata}\label{section ppda}

This chapter studies the computational complexity of reachability games and parity games in  parametric pushdown automata. Parametric pushdown automata provide a  formalism to
reason about 
recursive programs
making use of parametric constraints.
% Here, the stack can additionally be compared against parameters that can take unspecified word values. 
Recently, different variants of parametric pushdown automata have been introduced in the literature 
% [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
\cite{hague2011parameterised, esparza2016parameterized, fortin2017model} 
% In this work, following [21,20,15,27,12], we consider parametric pushdown systems
however as
parameterized asynchronous shared-memory systems.
These systems consist of a leader pushdown automaton
and arbitrarily many identical contributor pushdown automata, communicating via a shared memory
in the form of a register which can take finitely many values.
This variant have been shown to have applications
to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
\iffalse
21. V. Kahlon. Parameterization as abstraction: A tractable approach to the dataflow
analysis of concurrent programs. In LICS’08, pages 181–192. IEEE, 2008.
\fi

\iffalse
[2]
K. Apt and D. Kozen. Limits for automatic verification of finite-state concurrent systems. Information
Processing Letters (IPL), 1986.
\fi
% Whereas abstract interpretation is used for control and data abstractions, we propose the use of parameterization as an abstraction for concurrency. Leveraging abstract interpretation in conjunction with parameterization allows us to lift two desirable properties of sequential dataflow analysis to the concurrent domain, i.e., precision and scalability.
\iffalse
@inproceedings{kahlon2008parameterization,
  title={Parameterization as abstraction: A tractable approach to the dataflow analysis of concurrent programs},
  author={Kahlon, Vineet},
  booktitle={2008 23rd Annual IEEE Symposium on Logic in Computer Science},
  pages={181--192},
  year={2008},
  organization={IEEE}
}

See also:
S. A. German and P. A. Sistla. Reasoning about systems with many processes. J. ACM, 39(3):675-735, 1992.
\fi
\iffalse
for a more general architecture proposed by German and Sistla. Namely he
considered systems with a leader pushdown automaton and an arbitrary number
of copies of a contributor pushdown automaton; cf. Figure 1. Consecutively,
Esparza, Ganty and Majumdar [15] have proved that the problem is Pspace-
complete, thus surprisingly easy for such type of problem. La Torre, Muscholl,
and Walukiewicz [27] showed that the reachability problem remains decidable if
instead of pushdown automata one considers higher-order pushdown automata, or
any other automata model with some weak decidability properties. Reachability
was also shown to be decidable for parametric pushdown systems with dynamic
thread creation [32].

15. J. Esparza, P. Ganty, and R. Majumdar. Parameterized verification of asynchronous
shared-memory systems. J. ACM, 63(1):10:1–10:48, 2016. Journal version of
CAV’13.

27. S. La Torre, A. Muscholl, and I. Walukiewicz. Safety of parametrized asynchronous
shared-memory systems is almost always decidable. In CONCUR’15, volume 42 of
LIPIcs, pages 72–84. Schloss Dagstuhl - Leibniz-Zentrum für Informatik, 2015.

32. A. Muscholl, H. Seidl, and I. Walukiewicz. Reachability for dynamic parametric
processes. In VMCAI’17, volume 10145 of LNCS, pages 424–441. Springer, 2017.
\fi

We consider here a different approach to extending pushdown automata with parameters, by allowing them to test \textcolor{black}{equality of the stack content against parameters}. 
\iffalse
Pushdown systems which can test the
stack content against regular-expressions are equivalent to pushdown systems, so are pushdown
systems \textcolor{red}{which can test the stack content against specific words (not so relevant ? motivate more that parameters are used in comparisons rather than other things)}.
\fi
%
%
%
Perhaps a similar model consists in 
Pushdown automata 
%with regularity condition on the stack
with transitions that are conditioned by regular conditions on their stack content, which
can in particular
test the
stack content against
specific words. 
%
%
%
They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages),
and have
been used to establish that CTL$^*$ model checking remains decidable
when the formulas are allowed to include regular predicates on the stack content,
and to obtain 
model checking algorithms for LTL and CTL$^*$ model checking
for
pushdown automata \cite{finkel1997direct}.
Pushdown automata 
%with regularity condition on the stack
with transitions that are conditioned by regular conditions on their stack content
can be viewed as
a special case of stack automata as seen in \cite{hopcroft1969formal}.
Instead
of checking the stack content against regular conditions, or, in a more limited manner,
against specified word values,
we consider
checks against unspecified word values that can be instantiated using parameters.
This allows for instance to ensure equality between two stack contents appearing in two distincts configurations in a run.

% \mh{some more words about motivations}

% \textcolor{red}{One motivation to the study of parametric pushdown systems is their ability to check for equality of two stack contents that appear in a run}. 
Extensions of pushdown automata with storage for later comparisons
have been introduced for instance as register pushdown automata. 
Such automata possess registers in addition to their stack, and can keep data values in both.
Register pushdown automata have been shown to have applications for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}. 
In theory, since registers can be unfilled and later be filled again an unbounded number of times, the number of different data values a pushdown register automata can store in its registers in a run is unbounded.  
In \cite{murawski2017reachability} it was shown however that a register pushdown automaton can
only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers,
% Even though pushdown register automata have no bound on the number of elements of the alphabet that they can store at any instant, over the course of a run they can really “remember” at most 3r of them, 
i.e. for any run of a register pushdown automaton with $r$ registers there exists an equivalent run %, i.e.
 with the same initial and final conﬁgurations, but in which every configuration contains register assignments drawn from only $3r$ elements. This 
 %can 
 leads to
 the
  question
  of
   how useful the ability to unfill and later refill registers really is compared to a model that would only specify valuations once.

We provide formal definitions of parametric pushdown automata, parametric pushdown reachability games
and parametric pushdown parity games in Section~\ref{section def ppda}. We give an overview of our contribution in Section~\ref{contribution ppda}. The contribution consists in
proving parametric pushdown reachability games
and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
$n$ is fixed,
and providing a nonelementary lower bound for parametric pushdown reachability games in general.
We provide the proof of our result, which stretches along 
Section~\ref{lower bound PDA},
Section~\ref{pebbles},
% Section~\ref{ATWA and PATWA}, and
Section~\ref{section HPDA}.
In Section~\ref{discuss ppda} we close the chapter with a discussion about the methods used, with some directions for future work.

\iffalse
In this section we introduce parametric pushdown automata. 
% Being introduced by Bundala and Ouaknine in [10], we define parametric one-counter automata. These are automata that can manipulate a counter that can be incremented or decremented, parametrically or not, compared against constants or parameters, and with divisibility tests modulo constants. It is worth mentioning that the notion of parametric one-counter automata from [10] is slightly more expressive than ours, allowing more operations.
We investigate the decidability and complexity of 
% parity
several types of 
games over them. 
% More specifically, we consider pebble pushdown system, in which pebbles can be used to remember some stack content already visited, and compared against. 
We determine decidability of the problem, and provide a nonelementary lower bound.
\fi
%\subsection{Definitions}

\section{Definitions}\label{section def ppda}

\textcolor{black}{Parametric pushdown automata extend pushdown automata by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
A parametric pushdown automaton is then a finite automaton extended with a finite set of parameters $P$  and with a stack that can be manipulated by pushing or popping stack symbols and such that, moreover, the automaton can use the top of the stack, or check that the stack content corresponds to a parameter,  to decide which transition to take next.\\

\par\noindent\ignorespacesafterend
Formally, a {\em parametric pushdown automaton} ({\em PPDA} for short) 
is a tuple 
$\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
\begin{itemize}
\item $Q$ is a non-empty finite {\em set of  states},
\item $\Gamma$ is a non-empty finite {\em  stack alphabet},
\item $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$},
% \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Gamma^*$ is finite {\em  set of rules},
 \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules},
\item $q_{init}\in Q$ is an {\em initial  state}, 
\item $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
\item $F\subseteq Q$ is a {\em set of final  states}.
\end{itemize}

\par\noindent\ignorespacesafterend
The {\em size} of $\mathcal{Z}$ is defined as
$|\mathcal{Z}| = |Q| + |\Gamma| + |P| + |R| $.
We also refer to $\mathcal{Z}$ as an $n$-parametric pushdown automaton if $|P| = n$.
A {\em stack content} is a word from $ \Gamma^*$. 
As before we write the top
of the stack at the right of the word%(we are considering suffix rewriting as in Chapter 15)
. 
% A configuration is a  state and a stack content: $(q, w)$,
% which we rather write $q(w)$, where $w  \in  W^*$, $q  \in Q$. 
By $\Conf(\mathcal{Z})=Q\times  \Gamma^*$ we denote the set of
{\em configurations} of $\mathcal{Z}$. As usual we rather write $q(w)$ instead of $(q, w)$.
A {\em parameter valuation} is a function $\mu$ from $P$ to $\Gamma^*$.

\begin{samepage}
%\begin{definition}
A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
%labeled 
transition system  
$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
%\Lambda_{\mathcal{Z}}, 
\rightarrow_{\mathcal{Z},\mu})$ where
% $\Lambda_{\mathcal{Z}} = \Gamma \cup P$ and
for all $ q, q'  \in  Q$, 
for all $ w,w' \in \Gamma^* $, and 
for all $ a  \in  \Gamma$, 
\iffalse
\begin{eqnarray*}
q(wa) \xrightarrow{%a
}_{\mathcal{Z},\mu} q'(w') & \text{if} &  (q, a, q' , op)  \in  R, \\
q(\mu(p) ) \xrightarrow{%p
}_{\mathcal{Z},\mu} q'(\mu(p) %\textcolor{red}{\cdot} 
w) & \text{if} &  (q, p , q' , w)  \in  R.
\end{eqnarray*}
% \end{definition}
\fi
$q(wa) \xrightarrow{%a
}_{\mathcal{Z},\mu} q'(w')$
if there exists a rule in $R$
 of the form $(q,a,q', op)$ or $(q,p,q', op)$ with $\mu(p)=wa$,
such that
		% $\text{\textit{top}}(w) = a$
		% and 
% $w' = \theta(w)$.
either of the following holds
\begin{itemize}
\item $ op = \text{\textit{push}}^\gamma$ and $w' = wa\gamma$,

\item $ op = \text{\textit{pop}}$ and $w' = w$, or

\item $ op = \text{\textit{skip}}$ and $w' = wa$.
\end{itemize}
\end{samepage}

A {\em $\mu$-run} from $q_0(a_0)$ to $q_n(a_n)$ in $\mathcal{Z}$ is a corresponding
path in the transition system 
$T_{\mathcal{Z}}^\mu$ 
induced by $\mathcal{Z}$ and $\mu$.
%
%and will be noted
%$q_0(z_0) \xrightarrow{\lambda_0}_{\mathcal{Z},\mu} q_1(z_1) \xrightarrow{\lambda_1}_{\mathcal{Z},\mu} \cdots \xrightarrow{\lambda_{n-1}}_{\mathcal{Z},\mu} q_n(z_n)$.
%
As with PDA, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
and $q_n\in F$.
%
% We say {\em reachability holds} for $\mathcal{Z}$ if there is a $\mu$ such that there is a $\mu$-run in $\mathcal{Z}$ from $q_{init}(\gamma_{init})$   to some configuration $q(w)$ for some $q\in F$ and some $w \in \Gamma^*$.
% We refer to Figure~\ref{example_pda} for an instance of a PDA for which reachability holds.

In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
we prefer to write $q(w)\xrightarrow{%\lambda
}_{\mathcal{Z}, u}q'(w')$
to denote $q(w)\xrightarrow{%\lambda
}_{\mathcal{Z}, \mu}q'(w')$ and
will call the $\mu$-run an {\em $u$-run}. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\xrightarrow{%\lambda
}_{\mu}$ (resp. $\xrightarrow{%\lambda
}_{u}$)
instead of $\xrightarrow{%\lambda
}_{\mathcal{Z}, \mu}$ (resp. $\xrightarrow{%\lambda
}_{\mathcal{Z},u}$). \\

\paragraph{PPDA reachability}

Note that the reachability problem for parametric pushdown automata consists in
the following decision problem.

\problemx{$n$-PPDA reachability}
{An $n$-PPDA $\mathcal{Z}$.}
{Does 
%reachability hold for
an accepting $\mu$-run for some $\mu\in(\Gamma^*)^P$ exist in
 $\mathcal{Z}$?\newline}

% \subsubsection{Parametric pshdown games}
\paragraph{PPDA games}

We are moreover interested in games played on parametric pushdown automata's  
 transition systems.
Recall that given a 
%labeled 
transition system,
one needs only to provide %a mapping $\Omega$ and 
a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\

Thus, for an  $n$-parametric pushdown automaton 
$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we naturally partition the configurations of
%the transition system of 
 $\mathcal{Z}$
% $T^\mu_{\mathcal{Z}}$
into
$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
and
$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.

With these notations in mind one can define the arena
$$
A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
=
(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
parameter valuation $\mu: P\rightarrow \Gamma^*$.

Given additionally a partition of
$P$ into $P_0$ and $P_1$, 
we define 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
as
the arena that contains both the configurations and transitions of
the parameter valuation arena
$A_{P_0,P_1,\Gamma^*}$, as see on page~\pageref{parametric arena},
and that moreover contains the configurations and transitions of
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$ 
\---- albeit with configurations additionally including $\mu$ in the tuple to avoid confusion \----
for
all parameter valuations $\mu: P\rightarrow \Gamma^*$.
Moreover, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
we add a transition
from $\mu$ towards 
$q_{init}(\gamma_{init}, \mu)$ in $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$. \label{PPDA reachability game}

Given a 
priority function $\Omega: Q \to [0, m]$,
we consider the function
$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$
where we set
$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
for all
$w \in \Gamma^*$
and
for all $\mu \in (\Gamma^*)^{P}$,
and
$\widehat{\Omega}(\mu) = 0$
for all $\mu \in (\Gamma^* \cup \bot)^{P}$.

% $T_{\mathcal{Z}} = (\Conf(\mathcal{Z}), \rightarrow_{\mathcal{Z},\mu})$

\iffalse
\textcolor{red}{
Reachability %, B\"uchi
 and parity games on $\mathcal{G}_{(\mathcal{Z}, \mu,P_0, P_1, Q_0 , Q_1, \Omega)}$
are then clearly defined. They have the flaw however of not taking into account the
way the values assigned to the parameters is chosen. How values are assigned to
the parameters is the subject of the next Section.
}
\fi

We are interested in the following games and problems.

\problemx{Parametric pushdown reachability game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, and $ P = P_0 \biguplus P_1$.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{F \times \Gamma^* \times (\Gamma^*)^P})$ ?\newline}

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, and $ P = P_0 \biguplus P_1$, 
and a priority function $\Omega: Q \to [0, m]$%
%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}

\iffalse
\problemx{Parametric pushdown reachability game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, a disjoint
union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
.}
{Does player $0$ have a 
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}

\problemx{Parametric pushdown parity game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, a disjoint
union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
, 
a priority mapping $\Omega: Q \to \{0, \ldots m\}$%
%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
.}
{Does player $0$ have a 
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game, 
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
\fi

We write {\sc $n$-parametric reachability parity game}
and
{\sc $n$-parametric pushdown parity game} if 
 the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.

% \subsubsection{Contribution}
\subsection{Contribution}\label{contribution ppda}

The following theorems state our main results concerning {\sc Parametric pushdown parity game}
and {\sc Parametric pushdown reachability game}. The contribution is two-fold.

\begin{theorem}
{\sc $n$-parametric pushdown parity game} and {\sc $n$-parametric pushdown reachability game} are in $(n+1)$-$\EXP$.
\end{theorem}

\begin{theorem}
{\sc Parametric pushdown parity game} and {\sc Parametric pushdown reachability game} are not in $\ELEMENTARY$.
\end{theorem}

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.

	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown automata.
 We then show that higher-order pushdown automata parity games can be used to solve parity games on
 the transition systems of pebble pushdown automata,
using one additional stack level for each pebble.
 Since solving parity games on higher-order pushdown automata with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown automata with $n$ parameters.

% \subsubsection{Overview}
\subsection{Overview}

In Section~\ref{logics}, we provide preliminary definitions. 
Section~\ref{lower bound PDA} shows a nonelementary lower bound for the problem.
Section~\ref{pebbles} will deal with the introduction of pebble pushdown games.
% Section~\ref{ATWA and PATWA} is devoted to using pebble alternating tree automata with parity acceptance condition to solve pebble pushdown system parity games. 
% We leave the proof of decidability itself  for Section~\ref{section decidability}.
Section~\ref{section HPDA} is finally devoted to using higher-order pushdown automata parity games
to solve pebble pushdown automata parity games.

% \subsection{Logics}\label{logics}
\section{Logics}\label{logics}

We now brieﬂy review some standard deﬁnitions from mathematical logic.

\newcommand{\relational}{E}

\begin{samepage}
\begin{definition}
%
% A {\em vocabulary} $\sigma$ is a collection of {\em constant} symbols (denoted $c_1, \ldots , c_n, \ldots$), {\em relation}, or {\em predicate} symbols ($P_1, \ldots, P_n, \ldots $), and {\em function} symbols ($f_1, \ldots, f_n, \ldots$). Each relation and function symbol has an associated arity. \newline
 A {\em vocabulary} $\tau$ is a set of relational symbols (denoted $\relational_1, \ldots, \relational_n, \ldots $), each of which has a specified
 {\em arity} in $\N$. A symbol $\relational \in \tau$ is called monadic if its arity is one, i.e., if it is used to
 denote sets.
%
A {\em $\tau$-structure} (also called a {\em model})
%
$$ \mathfrak{A} = (A, %\{ c_i^{\mathfrak{A}} \}, 
			( \relational^{\mathfrak{A}} )_{\relational \in \tau}%, \{ f_i^{\mathfrak{A}} \}
			)$$
%
consists of a set $A$ together with an interpretation of
% \begin{itemize}
% \item each constant symbol $c_i$ from $\sigma$ as an element $ c_i^{\mathfrak{A}} \in A$,
% \item
 each $k$-ary relation symbol $\relational$ from $\tau$ as a $k$-ary relation on $A$; that is, a
set $\relational^{\mathfrak{A}} \subseteq A^k$.
% \item each $k$-ary function symbol $f_i$ from $\sigma$ as a function $f_i^{\mathfrak{A}}: A^k \to A$. \newline
%\end{itemize}
%
A structure $\mathfrak{A}$ is called finite if $A$ and $\tau$ are finite sets. The {\em universe} of
a structure is typically denoted by a Roman letter corresponding to the name
of the structure; that is, the universe of $\mathfrak{A}$ is the set $A$, the universe of $\mathfrak{B}$ is $B$, and so on. We shall also occasionally write $a \in \mathfrak{A}$ instead of $a \in A$.\newline
%
\end{definition}
\end{samepage}

For example, if $\tau$ consists of a single relational symbol ${ \rightarrow }$ of arity $2$, then possible structures for $\tau$ consist of transition systems. \newline

						\iffalse
For example, if $\sigma$ has constant symbols $0$, $1$, a binary relation symbol $<$,
and two binary function symbols $\cdot$ and $+$, then one possible structure for $\sigma$
is the real field $\mathfrak{R} = (\R, 0^{\mathfrak{R}}, 1^{\mathfrak{R}}, <^{\mathfrak{R}}, +^{\mathfrak{R}}, \cdot^{\mathfrak{R}})$, where $ 0^{\mathfrak{R}}, 1^{\mathfrak{R}}, <^{\mathfrak{R}}, +^{\mathfrak{R}}, \cdot^{\mathfrak{R}}$ have the expected meaning.
Typically, we shall omit the superscript with the name of the structure, using the same symbol for both a symbol in the vocabulary, and its interpretation in a structure. For example, we shall write
$ \mathfrak{R} = (\R, 0, 1, <, +, \cdot)$ for the real field.

A few notes on restrictions on vocabularies are in order. 
 Constants can be
 treated as functions of arity zero; however, we often need them separately, as
 in the finite case, we typically restrict vocabularies to {\em relational} ones: such vocabularies contain only relation symbols and constants. This is not a serious
restriction, as first-order logic defines, for each $k$-ary function $f$, its graph,
which is a $(k + 1)$-ary relation $\{( \overrightarrow{x}, f (\overrightarrow{x})) | \overrightarrow{x} \in A^k \}$. A vocabulary that consists exclusively of relation symbols (i.e., does not have constant and function symbols) is called {\em purely relational}.

Unless stated explicitly otherwise, we shall assume that:
\begin{itemize}
\item any vocabulary $\sigma$ is at most countable;
\item when we deal with finite structures, vocabularies $\tau$ are finite and relational.
\end{itemize}
						\fi

% Next, we deﬁne the  monadic second-order logic, or $\MSO$.

Next, we deﬁne first-order formulae over a vocabulary $\tau$, or $\FO[\tau]$ for short (or simply $\FO$ when the vocabulary is obvious from context). We define free variables, and
the semantics of $\FO[\tau]$ formulae.

\begin{definition}\label{FO}

We assume a countably infinite set of first-order variables {\em Var}.
% and of second-order variables {\em VAR}. 
First-order variables
will be typically denoted by $x, y, z, \ldots,$ with subscripts and superscripts.
% whereas second-order variables will be typically denoted by $X, Y, Z, \ldots ,$
% using subscripts and superscripts as well. 
We
inductively define %terms and 
formulae of the 
% monadic second-order 
first-order
logic
over vocabulary $\tau$ as follows:

\begin{samepage}
\begin{itemize}
% \item  Each variable $x$ is a term.
% \item  Each constant symbol $c$ is a term.
% \item  If $t_1 , \ldots , t_k$ are terms and $f$ is a $k$-ary function symbol, then  $f (t_1 , \ldots , t_k )$ is a term.
%
\item  If $x_1 , x_2$ are first-order variables, then $x_1 = x_2$ is an (atomic) formula.
\item  If $x_1 , \ldots , x_k$  are first-order variables and $\relational \in \tau$ is a $k$-ary relation symbol, then 
$\relational(x_1 , \ldots , x_k )$
is an (atomic) formula.
% \item If $x$ is a first-order variable and $X$ is a second-order variable, 
% then $X(x)$ is an (atomic) formula.
%
\item  If $ \phi_1 , \phi_2$ are formulae, then $\phi_1 \wedge \phi_2$ , $\phi_1 \vee \phi_2$, and 
$\neg \phi_1$ are formulae.
\item  If $\phi$  is a formula, and $x$ a first-order variable, then $\exists x ~ \phi$  and $\forall x ~ \phi$  are formulae.
% \item  If $\phi$  is a formula, and $X$ a second-order variable, then $\exists X ~ \phi$  and $\forall X ~ \phi$  are formulae.
\end{itemize}
\end{samepage}

\end{definition}

A formula that does not use existential ($\exists$) and universal ($\forall$) quantiﬁers
is called {\em quantifier-free}.
%
Given a set of formulae $S$, formulae constructed from formulae in $S$ using
only the Boolean connectives $\vee$, $\wedge$, and $\neg$ are called {\em Boolean combinations} of
formulae in $S$.
%
We shall use the standard shorthand $\phi  \to \psi $ for $ \neg \phi \vee \psi $ 
and $\phi \leftrightarrow \psi $ for
$(\phi  \to \psi ) \wedge (\psi  \to \phi )$. \\

The sets of free variables of a formula %or a term
 are defined as follows:

\newcommand{\free}{\text{\textit{free}}}

\begin{samepage}
\begin{itemize}
\item $\free(x_1 = x_2) = \{x_1, x_2 \}$,
\item $\free(\relational(x_1,\ldots, x_k)) = \{x_1, \ldots, x_k \}$,
% \item $\free(X(x)) = \{ x,X\}$,
% 
\item $\free(\neg \phi) = \free(\phi)$,
\item $\free(\phi \vee \psi)= \free(\phi) \cup \free(\psi)$,
\item $\free(\phi \wedge \psi)= \free(\phi) \cup  \free(\psi)$,
\item $\free(\exists x ~ \phi) = \free(\phi) \setminus \{ x \}$,
\item $\free(\forall x ~ \phi) = \free(\phi) \setminus \{ x \}$%,
% \item $\free(\exists X ~ \phi) = \free(\phi) \setminus \{ X \}$,
% \item $\free(\forall X ~ \phi) = \free(\phi) \setminus \{ X \}$%
.
\end{itemize}
\end{samepage}
If $\overrightarrow{x}$ is the tuple of all the free first-order variables of $\phi$,
% and $\overrightarrow{X}$ is the tuple of all the free second-order variables of $\phi$ 
we write $\phi(\overrightarrow{x}%,\overrightarrow{X}
)$. A {\em sentence}
is a formula without free variables.
%
%
For a finite set $\mathcal{V} \subseteq \text{Var} %\cup \text{VAR}
$, and a structure $\mathfrak{A}$,
we define a {\em $(\mathcal{V},\mathfrak{A})$ variable assignment} $\mu$ as a partial function from
$\text{Var} %\cup \text{VAR}
$ to $A$ whose domain is $\mathcal{V}$. For a first-order variable $x \in \text{Var}$, an element $a \in A$,
and a $(\mathcal{V},\mathfrak{A})$ variable assignment $\mu$ where $x \notin \mathcal{V}$,
 $\mu[x \mapsto a]$ denotes the 
 $((\mathcal{V} \cup \{x\}),\mathfrak{A})$ variable assignment $\mu'$ that maps
 $x$ to $a$ and, otherwise, coincides with $\mu$.
% For $X \in \text{VAR}$ and $B \subseteq A$, 
% $\mu[X \mapsto B]$ is defined similarly.

Given a $\tau$-structure $\mathfrak{A}$, we define inductively,
% for each term $t$ with free
% variables $(x_1 , \ldots , x_n )$ the value $t^{\mathfrak{A}} (\overrightarrow{a})$, where 
% $\overrightarrow{a} \in A^n$, and 
for each formula
$\phi$ with $\overrightarrow{x}%,\overrightarrow{X}
$ as free variables, the notion that 
$\mathfrak{A}$ satisfies $\phi$ 
for the
 $(\mathcal{V},\mathfrak{A})$ variable assignment $\mu$,
where $\overrightarrow{x}%,\overrightarrow{X}
 \subseteq \mathcal{V}$,
which we denote by
$\mathfrak{A} \models_\mu \phi$. 
If $\phi$ is a sentence, we just write 
$\mathfrak{A} \models \phi$.

\begin{samepage}
\begin{enumerate}
% \item   If $t$ is a constant symbol $c$, then the value of $t$ in $\mathfrak{A}$ is $c^{\mathfrak{A}}$.
%
% \item   If $t$ is a variable $x_i$, then the value of $t^{\mathfrak{A}} (\overrightarrow{a})$ is $a_i$.
%
% \item   If $t = f (t_1 , \ldots , t_k )$, then the value of $t^{\mathfrak{A}} (\overrightarrow{a})$ is 
% $f^{\mathfrak{A}} (t^{\mathfrak{A}}_1(\overrightarrow{a}), \ldots, t^{\mathfrak{A}}_k (\overrightarrow{a}))$.
%
\item   $\mathfrak{A} \models_\mu (x = y)$ if 
%$t^{\mathfrak{A}}_1(\overrightarrow{a}) = t^{\mathfrak{A}}_2(\overrightarrow{a})$
$ \mu(x) = \mu(y)$.
%
\item  $\mathfrak{A} \models_\mu \relational(x_1 , \ldots , x_k ) $ if
%$(t^{\mathfrak{A}}_1(\overrightarrow{a}), \ldots, t^{\mathfrak{A}}_k (\overrightarrow{a}))
$ (\mu(x_1), \ldots, \mu(a_k))
\in \relational^{\mathfrak{A}} $.
%
	%\item $\mathfrak{A} \models_\mu X(x)$ if
	%$\mu(x) \in \mu(X)$.
%
\item   $\mathfrak{A} \models_\mu \neg \phi$ if 
$\mathfrak{A} \models_\mu \phi $ does not hold.
%
\item   $\mathfrak{A} \models_\mu \phi_1 \wedge \phi_2$ if
$\mathfrak{A} \models_\mu \phi_1$ and 
$\mathfrak{A} \models_\mu \phi_2$.
%
\item   $\mathfrak{A} \models_\mu \phi_1 \vee \phi_2$ if
$\mathfrak{A} \models_\mu \phi_1$ or 
$\mathfrak{A} \models_\mu \phi_2$.
%
%
\item   % If $\psi (\overrightarrow{x},\overrightarrow{X} )$ is  $\exists y ~ \phi   (y, \overrightarrow{x} ,\overrightarrow{X})$, then 
$\mathfrak{A} \models_\mu \exists y ~ \phi   (y, \overrightarrow{x} %,\overrightarrow{X}
)$ iff
there exists
% a variable assignment $\mu'$ that differs from $\mu$ only on the value of $x$, unassigned in $\mu$,
$a \in A$
such that
$\mathfrak{A} \models_{\mu[y \mapsto a]} \phi$.
%
\item    % If $\psi (\overrightarrow{x},\overrightarrow{X})$ is  $\forall y ~ \phi   (y, \overrightarrow{x},\overrightarrow{X})$, then 
$\mathfrak{A} \models_\mu \forall y ~ \phi   (y, \overrightarrow{x}%,\overrightarrow{X}
)$ iff
for all
% variable assignment $\mu'$ that differs from $\mu$ only on the value of $x$, unassigned in $\mu$,
% such that
$a \in A$,
$\mathfrak{A} \models_{\mu[y \mapsto a]} \phi$.
%
	% \item    
	% $\mathfrak{A} \models_\mu \exists Y ~ \phi   ( \overrightarrow{x}, Y,\overrightarrow{X})$ iff
	% there exists
	% $B \subseteq A$
	% such that
	% $\mathfrak{A} \models_{\mu[Y \mapsto B]} \phi$.
%
	% \item   
	% $\mathfrak{A} \models_\mu \forall Y ~ \phi   ( \overrightarrow{x}, Y,\overrightarrow{X})$ iff
	% for all
	% $B \subseteq A$,
	% $\mathfrak{A} \models_{\mu[Y \mapsto B]} \phi$.
\end{enumerate}
\end{samepage}

Concerning the logic $\FO$, we are interested in the following decision problem.

\problemx{$\FO[\tau]$ model checking}
{A $\tau$-structure $\mathfrak{A}$ and a $\FO[\tau]$-sentence $\phi$.}
{Does $\mathfrak{A} \models \phi $ ?\newline}

% \subsection{A nonelementary lower bound}\label{lower bound PDA}
\section{A nonelementary lower bound}\label{lower bound PDA}

The aim of this section is to show a nonelementary lower bound for the 
problem of solving
parametric pushdown reachability games.
% problem of deciding winning configuration for pebble pushdown parity games.
It is trivial to show that
the problem of
solving parametric pushdown reachability game 
reduces itself to
the problem of
%solving parametric pushdown B\"uchi game
%and
%the problem of
solving parametric pushdown parity game.

We reduce the satisfiability problem for first-order logic on words
to solving {\sc Parametric pushdown reachability game}. 
We know from \cite{Sto74}
that satisfiability is nonelementary. 
% Here a function is elementary if it is  $O(h_k (n))$ for one of the $k$-fold exponential functions $h_k$ with $h_0 (n) = n$ and $h_{k+1} (n) = 2^{h_k (n)}$.

First we introduce in more detail what we call the satisfiability problem for first-order logic on the class of words.
Then we provide the reduction.

%\noindent

%\subsubsection{$\FO$ satisfiability on words}
\subsection{$\FO$ satisfiability on words}

In order to define $\FO$ satisfiability on words, we need to define precisely $\tau$-structures that correspond to words.

For a finite alphabet $\Sigma$,
 let $\tau (\Sigma )$ be the vocabulary consisting of a binary relation
symbol $\leq$, 
% a unary function symbol $S$, 
% two constant symbols $min$ and $max$, 
and a unary
relation symbol $\relational_a$ for every $a \in \Sigma$. 
A word structure over $\Sigma$ is a 
$\tau (\Sigma )$-structure $\mathfrak{W}$ with
the following properties:

	\begin{itemize}
	\item $W$ is a finite interval $[0,N]$ for some $ N \in \N$, 
	\item $<^\mathfrak{W}$ is the natural order of $\N$%,
	%	$min^\mathfrak{W}$ and 
	%	$max^\mathfrak{W}$ are 
	%	the minimum and maximum element of $\leq^\mathfrak{W}$, and 
	%	$S^\mathfrak{W}$ is the successor function associated with $\leq^\mathfrak{W}$, 
	%	where we let 
	%	$S^\mathfrak{W} (max^\mathfrak{W} ) = max^\mathfrak{W}$
,
	\item For every $i \in W$ there exists precisely one $a \in \Sigma$ such that $i \in 
	\relational_a^\mathfrak{W}$.
	\end{itemize}

We refer to elements $i \in W$ as the {\em positions} in the word (structure) and, for every position $i \in W$, to the unique $a$ such that $i \in \relational_a^\mathfrak{W}$ as the letter at $i$.

%It is obvious how to associate any word $w$ from the set $\Sigma^*$ of all words over $\Sigma$ with a word structure over $\Sigma$ and, conversely, how to associate any up to isomorphism unique word structure with a word in $\Sigma^*$. 
There is an obvious one to one correspondance between any word $w$ from the set $\Sigma^*$ of all words over $\Sigma$ and any word structure over $\Sigma$.
We identify words with the corresponding word structures
and write $w \in \Sigma^*$ to refer both to the word and the structure.

% The class of all words (or word structures) over any alphabet is denoted by $\mathbb{W}$. 
% The length of a word $\mathfrak{W}$ is denoted by $|\mathfrak{W}|$.

% definition of a subword

It is well-known that if we are interested in the complexity of first-order or monadic
second-order model checking or satisfiability on words, the alphabet can be assumed to be $\{0,1\}$ without loss of generality. \iffalse
This can be phrased as
follows:

\begin{claim} 
There is a linear time algorithm that, given an alphabet $\Sigma$, a sentence
$\phi \in \FO[\tau(\Sigma)]$ and a word structure $\mathfrak{W} \in \mathbb{W}$ over $\Sigma$, computes a sentence $\phi' \in \FO[ \tau (\{0, 1\}) ]$ and a word
$\mathfrak{W}' \in \{0, 1\}^*$ such that 
$|\phi'| \in O(|\phi|)$, $|\mathfrak{W}'| \in O(|\mathfrak{W}|)$, and 
$$ \mathfrak{W} \models \phi \Longleftrightarrow \mathfrak{W}'  \models \phi' .$$

\end{claim}

\fi
Thus when considering the satisfiability problem for first-order logic on the class of words, we can restrict ourselves to the alphabet $\{ 0,1\}^*$, and only consider the following problem.

\begin{samepage}
\problemx{$\FO$ satisfiability on words}
{A  
$\FO[\tau(\{0,1\})]$-sentence $\phi$.}
{Does there exist a word $w\in \{0, 1\}^*$, such that $w \models \phi$ ?\newline}
\end{samepage}

The result that is previously known and motivates the upcoming reduction consists in the following.

\begin{theorem}{\cite{Sto74}}\label{FOnonelementary}

%  Assume that $FPT \neq AW[*]$, and let $f$ be an elementary function and let $p$ be a polynomial.
%  There is no model checking algorithm for first-order logic on the class of words 
%  whose running time is bounded by
%  $$ f(k) \cdot p(n)$$

% where $k$ denotes the size of the input sentence of the model checking problem and $n$ the
% size of the input word.

The {\sc $\FO$ satisfiability on words} problem is not in {\sf ELEMENTARY}.
	% SAT(N,<) is not elementary-recursive
	% and in fact SAT(N,<) \not\in NSPACE(g(log_b(n),0)) for all b > 3.
\end{theorem}

% \subsubsection{Reduction from $\FO$ SAT on words}
\subsection{Reduction from $\FO$ SAT on words}

We consider reduction towards the problem of {\sc Parametric pushdown reachability game}.
The following theorem states a polynomial time reduction from
{\sc $\FO$ satisfiability on words}
%We consider reduction 
to %the problem of %following problem.
{\sc Parametric pushdown reachability game}.

\begin{samepage}
\begin{theorem}
{\sc $\FO$ satisfiability on words} is polynomial time reducible to {\sc Parametric pushdown reachability game}.
\end{theorem}
\end{samepage}

\begin{proof}

We start with
 a $\FO[\tau(\{0,1\})]$-formula $\phi$.
%
We construct
a  parametric pushdown automaton $\mathcal{Z}$,
 %a 
 disjoint
unions
 $Q = Q_0  \biguplus Q_1 $ and $P = P_0 \biguplus P_1$,
  and a priority mapping $\Omega$. 
We do so such that
player $0$ has a 
\iffalse
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$
\fi
		winning strategy from $\mu_\bot$ in the parity game
		$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ 
if and only if
there exists
a word $w \in \{0,1\}^*$ such that
%a word $\mathfrak{W} \in \{0,1\}^*$ such that
$w  \models \phi $, 
%$\mathfrak{W}  \models \phi $. \\
where 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
and
$\widehat{\Omega}$
correspond to the definitions on page~\pageref{PPDA reachability game}. \\

\noindent
We assume without loss of generality that $\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k ~ \psi$ is written in the prenex normal form%$\forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \psi$
, where 
%$Q_i$'s are
an even number $k$ of quantifiers alternate between existential and universal ones, and 
that $\psi$ is quantifier-free and in disjunctive normal form, i.e.
 $$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).$$

% \mh{All right, warning, $h$ could be different, need to take care of that.}

% \mh{assume that it starts with an universal, then it's alternating}

Our construction will include the following states and gadgets.

\begin{align*}
	&q_{init} &  & \text{the initial state} \\
%	&q_I' &  & \text{the \textit{post-}initial state} \\
%	& Q &	& \forall Q_i x_i \in \{Q_1 x_1, \ldots, Q_k x_k\}\\
	& \forall_i &	& \forall i \in \{1, 3, \ldots, k-1\}\\
	& \exists_i &	& \forall i \in \{2, 4 \ldots, k\}\\
%	& \C_{Q_i x_i},\text{ a gadget } & & \forall Q_i x_i \in \{Q_1 x_1, \ldots, Q_k x_k\}\\	
	& \C_{\forall x_i},\text{ a gadget } & &\forall i \in \{1, 3, \ldots, k-1\}\\	
	& \C_{\exists x_i},\text{ a gadget } & &\forall i \in \{2, 4 \ldots, k\}\\	
	& \vee & & \\
	& \wedge_i & & \forall i \in \{ 1, \ldots, l\} \\
	& \C_{\psi_{i,j}},  \text{ a gadget }&  & \forall i \in \{ 1, \ldots, l\}  \forall j \in \{ 1, \ldots, h_i\} \\
\end{align*}

Player $0$ starts by chosing a valuation $w$ for the first parameter (parameter $p_0$). The goal of this first parameter is to remember
the word $w$, which corresponds to the word ``guessed" by player $0$. 
This parameter is meant in part to enforce that 
the stack remains a prefix of $w$, with each stack content of the play corresponding to a position in the word.
 To enforce this, we allow in $\mathcal{Z}$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position in the word.
  A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to the value of parameter $p_0$. 
 The gadgets for these intermediary potential challenges will be excluded from the illustrations to preserve clarity of representation.
The other parameters are meant to correspond to the variables, with player $1$ parameters corresponding to universal variables and player $0$ parameters corresponding to existential variables.

\iffalse
Player $1$ then chose a valuation for the second parameter, then player $0$ chose
a valuation for the third, etc ... until for every  $i \in \{1, 3, \ldots, k-1\}$ 
player $1$ has chosen a valuation for $p_i$ and for every
$ i \in \{2, 4 \ldots, k\}$
player $0$ has chosen a valuation for $p_i$.
\fi
Once players have chosen values for their respective variables, the parameter assignment
$\mu$ is fixed. It is then time
for player $0$ to ensure that
%$\mathfrak{W}$ 
$w$
satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
$ |\mu(p_1)|-1, \ldots, |\mu(p_k)|-1$ respectively. This is the goal of the reachability game.

\begin{center}
	\begin{figure}
		\hspace{-0.25cm}
\includegraphics[width=1.01\textwidth]{figures/quantifiers}
	\caption{The begining of the automaton, and thus, of the game, essentially corresponds to a preprocessing steps dealing with the quantifiers. P%layer $0$ add $w$ to the stack, then
	%p
layers take turns %chosing particular
	proving the parameters chosen correspond to prefixes of $w$. Once every variable has been checked%assigned with the dropping of a corresponding pebble
	, players enters the second part of the game.}
		 \label{quantifiers}
	\end{figure}
\end{center}

The first step of the game is to check that the values assigned to the parameters correspond indeed to possible variables, that is, that the values of the parameters are prefixes of $w$. See Figure~\ref{quantifiers} for an illustration of what is happening.
%
% W.l.o.g, we assume $\psi$ to be in disjunctive normal form, that is, that 
%
% $$ \psi = \bigvee_{i \in \N} (\bigwedge_{j \in \N} \psi_{i,j} )$$
%
% where $\psi_{i,j}$ are atomic formula.

Recall $\psi$ is quantifier-free and in disjunctive normal form, i.e.
 $$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).$$

Player $0$, being the existential player, chooses one of the conjunctive clauses in $\psi$, essentially claiming to be able to prove it. Then player $1$, the universal player, chooses one of the atomic formulas in the clause to test. The process can be seen in Figure~\ref{DNF}. Testing the atomic formula is the purpose of a gadget $ \C_{\psi_{i,j}}$ built such that player $0$ has a winning strategy in $ \C_{\psi_{i,j}}$ if and only if 
$w$ satisfies $\psi_{i,j}$ if the variables $x_1, \ldots, x_k$ are interpreted by 
$ |\mu(p_1)|-1, \ldots, |\mu(p_k)|-1$ respectively.

\begin{center}
	\begin{figure}
		\hspace{2.4cm}
\includegraphics[width=0.6\textwidth]{figures/DNF}
	\caption{Player $0$ chooses which conjunctive clause to test and player $1$ chooses which of the atomic formula in the clause to test. State $\vee$ belongs to player $0$. For $i \in \{1, \ldots, l\}$, state $\wedge_i$ belongs to player $1$.}
		 \label{DNF}
	\end{figure}
\end{center}

Now we need only to describe the gadgets $ \C_{\psi_{i,j}}$. 
The gadget depends on the type of the atomic formula. For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then checks that the top of the stack
$x$ is the right letter. For one checking that $x = x'$ where $x$ and $x'$ are both variables,
player $0$ goes to the position corresponding to $x$ and checks against the parameter corresponding to $x'$ as well. If the formula is a negation, players exchange their roles. \\

\noindent
Thus we can in polynomial time in $|\phi|$ build a parametric pushdown automaton $\mathcal{Z}$ with $k+1$ parameters, 
 disjoint
unions $Q = Q_0  \biguplus Q_1 $ and $P = P_0 \biguplus P_1$,
and a priority mapping $\Omega$
such that
player $0$ has a 
\iffalse
strategy $\sigma_0$ for the parameter valuation game  such that
for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
player $0$ has a
winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ 
\fi
		winning strategy from $\mu_\bot$ in the parity game
		$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ 
if and only if
there exists
%a word $\mathfrak{W} \in \{0,1\}^*$ such that
a word $w \in \{0,1\}^*$ such that
%$\mathfrak{W}  \models \phi $.  \\
$w \models \phi$.
\end{proof}

Then, the nonelementary lower bound follows from Theorem~\ref{FOnonelementary}. 
% Indeed, if there was an elementary function $f$ and an algorithm deciding the  {\sc Parametric pushdown reachability game}  with running time bounded by $f(k)$, where $k$ denotes the size of the input parametric pushdown automaton, then the reduction would lead to a contradiction of the theorem.

\begin{theorem}

%The problem of 
{\sc Parametric pushdown reachability game} is not in {\sf ELEMENTARY}.

\end{theorem}

As a consequence, %the problem of {\sc Parametric pushdown parity game}
solving parametric pushdown parity games
%and the problem of {\sc Parametric pushdown B\"uchi game}
% are both
is 
 nonelementary too.

\begin{corollary}

%The problem of 
{\sc Parametric pushdown parity game} is not in {\sf ELEMENTARY}.

\end{corollary}

%\subsection{A decidability result from reduction to parity games on automata with pebbles}\label{pebbles}
\section{%A decidability result via r
Reduction to pebble pushdown automata parity games
}\label{pebbles}

% \mh{Motivates the use of pebbles. See pebbles on two-way automata, two-way tree automata, weighted automata, etc.}

% 21. Globerman, N., Harel, D.: Complexity results for two-way and multi-pebble automata and their logics. Theor. Comput. Sci. 169, 161–184 (1996)
%		same as [13] bellow
%
% Engelfriet, J., Hoogeboom, H.J.: Tree-walking pebble automata. In: Karhumäki, J., Maurer, H., Pa ̆un, G., Rozenberg, G. (Eds.) Jewels are Forever, Contributions to Theoretical Computer Science in Honor of Arto Salomaa, pp. 72–83. Springer, Berlin (1999) for tree-walking automata.
%
% @incollection{engelfriet1999tree,
%  title={Tree-walking pebble automata},
%  author={Engelfriet, Joost and Hoogeboom, Hendrik Jan},
%  booktitle={Jewels are forever},
%  pages={72--83},
%  year={1999},
%  publisher={Springer}
%}
% 
%
% Two-way pebble transducers for partial functions and their composition, Joost Engelfriet
%
% Tree-Walking Pebble automata, Joost Engelfriet and Hendrik Jan Hoogeboom 
A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
pebbles that it can drop on, and lift from 
words, using them as markers.
Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
\cite{globerman1996complexity, engelfriet1999tree}.
Automata with nested pebbles were also introduced
for tree-walking automata. 
It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}%, the main trouble being that tree-walking automata get lost rather easily
.
Using pebbles is a remedy against getting lost along a tree, but 
the unrestricted use of pebbles leads to a class of tree languages much larger than the
regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline. It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape, 
and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.

In this section, we extend these ideas to pushdown automata. Instead of using pebbles as markings on their input, pebble pushdown automata have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape, 
and a \textit{lift} pops the last dropped pebble 
while requiring that
% assuming
 it was placed on the current node. 
% Thus hierarchicality on the registers works slightly differently from hierarchicality on the pebbles, hierarchicality on the registers being a syntactic property, whereas hierarchicality on the pebbles is a semantical one.
One can think of a pebble as a register that can store a stack content for later comparisons.

We first define more formally our pebble pushdown automaton framework, and then provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.

%\subsubsection{Pebble pushdown automata}
\subsection{Pebble pushdown automata}

\newcommand{\ppda}{\mathcal{I}}
% \Denarius
% \vernal
% \mathscr{Z}
% \mathcal{I}
% \pluto
% \capricornus
% \Bumpeq
% \Pfund

An {\em $n$-pebble pushdown automaton} 
is a tuple 
$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
where
\begin{itemize}
\item $Q$ is a {\em finite set of  states},
\item $\Gamma$ is a {\em finite stack alphabet},
\item  $R  \subseteq  Q  \times \Gamma \times
		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
		 \times Q  \times (\Op(\Gamma) \cup \{\text{\textit{drop}}, \text{\textit{lift}}\})$ is a {\em finite set of rules},
		 where the fourth element $S$ of a rule $r$ is a subset of 
		$\mathcal{P}(\{1,\ldots,i\})$ where $i$ is the third element of $r$,
		and if the last element is
		$\text{\textit{lift}}$ then we additionally require $i \in S$,
\item $q_{init} \in Q$ is an {\em initial  state}, 
\item $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
\item $F \subseteq Q$ is a {\em set of final  states}.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% \mh{remind the reader what a partial function is}

\par\noindent\ignorespacesafterend 
Recall that for \iffalse a partial function $f$ from a set $S$ to a set $X$ 
%is a function defined on a subset $C$ of $S$ (possibly $S$ itself) with output values in $X$, and 
that is defined on a subset $C \subseteq S$
is denoted by
$ f : S \rightharpoonup X $.

For \fi a partial function $ f : S \rightharpoonup X $ (see page~\pageref{partial}), for notational purposes, we consider some element $\bot_X \not\in X$ and 
 associate $f$ with the function returning the 
bottom element $\bot_X$ when $f$ is undefined. Thus we write $(X \biguplus \{ \bot_X \})^S$ for the set of all partial functions from $S$ to $X$, which we here abbreviate as $(X \biguplus \{ \bot \})^S$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% An $i$-configuration of $\ppda$ is an element of
%$ Q\times W^* \times (W^*)^{\{1, \ldots i\}}$.
% By $\Conf_i(\ppda)$ we denote the set of $i$-configurations of $\ppda$.
By $\Conf(\ppda)=Q\times \Gamma^* \times (\Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$ we denote the set of
% By $\Conf(\ppda) = \bigcup_{0 \leq i \leq n} \Conf_i(\ppda)$ we denote the set of
{\em configurations} of $\ppda$. As expected, we rather write $q(w, \mu)$ instead of $(q, w, \mu)$.
An {\em$i$-configuration} for $i > 0$ of $\ppda$ is a configuration $q(z,\mu)$ where
$\text{Dom}(\mu) = \{1, \ldots, i\}$, while a $0$-configuration is a configuration 
 $q(z,\mu)$ where $\text{Dom}(\mu)=\emptyset$.

The idea of a transition $(q, a, i, S, q', m) \in R$
is that, if the automaton $\ppda$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
$\ppda$ goes to state 
$q'$ and makes modifications to the stack or the pebbles according to
$m$. 
% \mh{The set $S$ is used for testing the presence of certain pebbles.}
Note a pebble 
can be lifted only if the stack content which corresponds to the pebble
is the same as the current stack content. 
%$\Dzeta$ $\zeta$ $\Xi$
This is enforced by syntactically requiring
the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.

\iffalse
A {\em weak} $n$-pebble pushdown system is an $n$-pebble pushdown system $\ppda = (Q, \Gamma,  \Delta )$ in which $\text{\textit{lift}}$ operations also require the presence of the last pebble dropped, i.e.
for all $(q, a, i, S, q', \text{\textit{lift}}) \in \Delta$, we have $ i \in S $. 
%  is an additionnal restriction $f(i) = wa$ that needs to hold for the application of the $lift$ operation.
{\em Weak $n$-pebble pushdown games} are defined naturally as $n$-pebble pushdown games on 
weak $n$-pebble pushdown systems.

\fi

A {\em pebble set} of $\ppda$ is a set $U \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
{\em $U \!$-pebble assignment} is a function which maps each $j \in U$ to a word in $ \Gamma^*$.
The $\emptyset$-pebble assignment is denoted by $\mu_{init} :  \{ 1, \ldots, n\} \rightharpoonup  \Gamma^* $ and is the 
%partial function from  $\{ 1, \ldots, n\}$ to $ \Gamma^* $ that
%does not map any $j \in \{ 1, \ldots, n \}$ to an output.
totally undefined function.

\iffalse
For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, \mu )$, where $w \in \Gamma^*$, $q$ is a state and
$\mu$ a $\{ 1, \ldots , i \}$-pebble assignment. 
We call $w$ the current stack, $q$ the current state and
$\mu$ the current pebble assignment. 
We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
$\mu(j) = w_j$, for each
$j \leq i$.

The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. 
\fi

\begin{samepage}
%\begin{definition}
An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
 such that
for all $(q, a, i, S, q', m) \in R$, with $a \in \Gamma$,
 for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
 %(respectively $\mu(j)= {\gamma_{init}}  $) 
 for each
$j \in S$ 	and
 $\mu(j) \neq wa$ 
 %(resp. $\mu(j) \neq {\gamma_{init}} $) 
 for each $j \in \{ 1, \ldots , i \} \setminus S$,
the following holds
\begin{itemize}
\item if $m \in \Op(\Gamma)$, 
% $(q,wa,\mu) \rightarrow_{\ppda} (q',wm, \mu)$, 
either
\begin{itemize}
\item $ m = \text{\textit{push}}^\gamma$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa\gamma, \mu)$,

\item $ m = \text{\textit{pop}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',w, \mu)$, or

\item $ m = \text{\textit{skip}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu)$,
\end{itemize}
\item if $ m = \text{\textit{drop}}$, 
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$,  
where $\mu'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j \leq i$, and $\mu'(i+1) = wa$, and
\item if $m = \text{\textit{lift}}$, and $i \in S$, i.e. the last pebble dropped belong of the set of pebble we test the presence of,
$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$, 
where $\mu'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
$\mu'(j) = \mu(j)$, for each
$j < i$.
\end{itemize}
%\end{definition}
\end{samepage}

\par\noindent\ignorespacesafterend
We are interested in games over pebble pushdown automata, % namely, 
%reachability, B\"uchi, and parity games.
mainly, parity games.

Again, given
a 
%labeled
 transition system, one needs only to provide 
%a mapping $\Omega$ and 
a partition of the set of
configurations 
%$S$ into two sets $S_0$ and $S_1$
 to obtain an arena.
 %
Given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we partition the configurations of
$T_{\ppda}$
into
$\Conf_{\ppda,0}=Q_0\times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$
and
$\Conf_{\ppda,1}=Q_1\times  \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$.

With these notations in mind one can  define the arena
$$
A_{(\ppda, Q_0 , Q_1)}
=
(\Conf_{\ppda,0}, \Conf_{\ppda,1}, \rightarrow_{\ppda}).$$

As expected, given a 
priority function $\Omega: Q \to \{0, \ldots m\}$,
one naturally 
set the extension of $\Omega$ as
$\overline{\Omega}: \Conf(\ppda) \to \{0, \ldots m\}$
and
$\overline{\Omega}(q, w,\mu) = \Omega(q)$
for all
$w \in  \Gamma^*$
and
$\mu \in ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$. \\

Concerning pebble pushdown automata, we are interested in the following decision problem.

\problemx{$n$-pebble pushdown parity game}
{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?\newline}

% \subsubsection{Reduction from parameters to pebble pushdown automata}
\subsection{From parametric pushdown 
automata 
to pebble pushdown automata
}

We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.

\begin{theorem}\label{PPDA reduction}
%The problem of 
{\sc $n$-parametric pushdown parity game} 
is polynomial time reducible to
%the problem of 
{\sc $n$-pebble pushdown parity game}.
\end{theorem}

\begin{proof}[Sketch]

Let us fix some
$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
P%\{p_1, \ldots, p_n \}
, q_{init}, { \gamma_{init}} , F)$,
 disjoint
union $Q = Q_0  \biguplus Q_1$
and $P = P_0 \biguplus P_1$,
and
some priority function 
$\Omega : Q \times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^R \to [0, m ]$.

We construct
% polynomial time
  an 
 $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
a 
 disjoint
union $Q' = Q_0'  \biguplus Q_1'$
and a mapping $\Omega' : Q \times  \Gamma'^* \times ( \Gamma'^* \biguplus \{ \bot \})^{\{1, \ldots, n\}} \to \{1, \ldots, m \}$,
such that
 %:
%
player $0$ has a winning strategy from 
$(q'_{init},\gamma_{init}, \mu_{init})$
in
		\iffalse
		for the parity condition on
		$\mathcal{G}_{(\ppda, Q'_0, Q'_1, \Omega')}$.
		\fi
the parity game
$\mathcal{G'}= (A_{(\ppda, Q'_0, Q'_1)}, \win_{\overline{\Omega'}})$
if and only if
player $0$ has a 
		\iffalse
		strategy $\sigma_0$ for the parameter valuation game  such that
		for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
		player $0$ have a
		winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
		$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$. 
		\fi
winning strategy from $\mu_\bot$ in 
the parity game
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
where		 
$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
and
$\widehat{\Omega}$
correspond to the definitions on page~\pageref{PPDA reachability game}.

%
%
%
%

The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
(one for every 
%total 
parameter assignment)
 of the pushdown automaton with $n$ parameters by using pebbles to represent the parameters. 
The $n$-pebble pushdown automaton will first simulate the 
parameter valuation arena.
Players take turns chosing particular stack contents on which to place pebbles,
using a gadget similar as the one from Figure~\ref{quantifiers}. 
Once every parameter has been assigned with the dropping of a corresponding pebble, 
the pebble assignment $\mu$ is fixed.
%%
%
%
%
%
%
%

Further configurations in the $n$-pebble pushdown automaton then consist of a word representing the status of the stack, a state, and a fixed pebble assignment. For a fixed pebble assignment there is then a one for one correspondance between configurations of the $n$-parametric pushdown 
automaton 
%transition system with assignment $\mu$
and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
the set of states of the $n$-parametric pushdown automaton, and 
so are player $0$ states, player $1$ states and the priority mapping.
\end{proof}

\section{Reduction to higher-order pushdown automata}\label{section HPDA}

Higher-order pushdown automata (HPDA for short) were introduced as a 
generalization of pushdown automata \cite{aho1969nested,Gre70,maslov1976multilevel}. 
% Alfred V. Aho. Nested stack automata. J. ACM, 16(3):383–406, 1969.
% S. Greibach. Full AFL’s and nested iterated substitution. Inf. Control, 16(1):7–35, 1970.
% A.N. Maslov. Multilevel stack automata. Problemy Peredachi Informatsii, 12:55–62, 1976.
A stack of a pushdown automaton is seen as a {\em level $1$ stack}. 
A pushdown automaton of level $2$ (or $2$-HPDA) then works with a stack of level $1$ stacks. 
In addition to the ability to push
and to pop a symbol on the top-most level $1$ stack, an $2$-HPDA can copy or remove the entire topmost level $1$ stack. 
The definition generalizes to any $n \geq 2$, and $n$-HPDA are similarly defined for all level
$n$ as automata working with a stack of level $(n-1)$ stacks.
% HPDA have been extensively studied as language recognizers [Dam82,Eng83].

We recall the definition from \cite{cachat2007complexity} which itself is taken from \cite{knapik2002higher}. We then provide a
reduction from the problem of solving pebble pushdown parity games to the problem of solving
higher-order pushdown parity games.

\subsection{Higher-order pushdown automata}

A {\em level $1$ stack} (or {\em $1$-stack}) over an alphabet $\Gamma$ is 
simply a %word in $\Gamma^*$
	stack over $\Gamma$, i.e. a word  in $\Gamma^*$.
% an arbitrary sequence $w_0 w_1 \ldots w_n $ of elements of $\Gamma$ for some $n \in \N$. 
A  {\em level $n$ stack} (or {\em $n$-stack}) over an alphabet $\Gamma$, for $n \geq 2$, is a 
non-empty sequence
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
of
$(n-1)$-stacks over $\Gamma$,
for some $m \in \N$.
The set of $n$-stacks over $\Gamma$ is denoted by $\mathscr{S}_n(\Gamma)$,
or simply $\mathscr{S}_n$ in case the set $\Gamma$ is obvious from context.
The set of all stacks over $\Gamma$ is written $\mathscr{S}(\Gamma) = \bigcup_{n\in \N} \mathscr{S}_n(\Gamma)$.
We define $\epsilon_{1}$ as
$\epsilon \in \Gamma^*$ and we inductively define 
$\epsilon_{n} = \langle \epsilon_{n-1} \rangle$ in $\mathscr{S}_n$ for all $n > 1$.

A {\em higher-order stack operation} is a partial function from $\mathscr{S}(\Gamma)$ 
to $\mathscr{S}(\Gamma)$ which preserves the level of the input (i.e. the image of
an $n$-stack is an $n$-stack for all $n \in \N$). The {\em level } of an operation $op$
is the smallest $n \in \N$ such that $\text{Dom}(op) \cap \mathscr{S}_n \neq \emptyset$. 
The operations 
%we consider
additionally
 respect the hierarchicality of higher-order stacks, i.e. in
a level $n+1$ stack only the topmost level $n$ stack can be accessed. An operation
$op$ of level $n$, applied to a level $n+1$ stack 
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
of length $m \in \N$,
thus
returns the output
$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle op(s_m) \rangle$
if applicable.
The definition for  all levels of stacks greater than $n$ follows the same pattern.

% For all $m \in \N$, 
The following operations can be performed on a $1$-stacks of length $m \in \N$.
% $ w_0 w_1 \ldots w_m \in \Gamma^*$.
\begin{eqnarray*}
&\text{\textit{push}}_1^\gamma( w_0 w_1 \ldots w_m) &=\quad w_0 w_1 \ldots w_m \gamma   \text{ for all } \gamma \in  \Gamma, \\
&\text{\textit{pop}}_1(w_0 w_1 \ldots w_{m-1} w_m ) &=\quad w_0 w_1 \ldots w_{m-1}, \text{ if } m \geq 1\\
&\text{\textit{top}}( w_0 w_1 \ldots w_m ) &=\quad w_m.
\end{eqnarray*}
The operations added at level $n+1$ are the copy of the topmost $n$-stack
and the removal of the topmost $n$-stack.
More formally, if $ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$ is a stack of 
level $n > 1$, 
the following operations are possible.
%
\begin{eqnarray*}
&\text{\textit{push}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle \langle s_m \rangle , \\
&\text{\textit{push}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle, \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_j(s_m) \rangle,
\text{ if } 2 \leq j < n\\
&\text{\textit{push}}_1^\gamma( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{push}}_1^\gamma(s_m) \rangle,
\text{ for all } \gamma \in \Gamma,\\
&\text{\textit{pop}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
,\\
&\text{\textit{pop}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
\langle \text{\textit{pop}}_j(s_m) \rangle,
\text{ if } 1 \leq j < n\\
&\text{\textit{top}}( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad
\text{\textit{top}}(s_m).
\end{eqnarray*}
% We omit the stack alphabet $\Gamma$ from 
%The operation $pop_j$ is undefined on a stack whose top stack of level $j$ is empty.
The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\

Given a stack alphabet $\Gamma$ and $n \in \N$, 
we denote by  $\Op_n(\Gamma)$ the {\em base set of $n$-stack operations} as

\begin{itemize}
\item $\text{\textit{push}}_j$ for all $2 \leq j \leq n$,

\item $\text{\textit{push}}_1^\gamma$ for all $\gamma \in \Gamma$,

\item $\text{\textit{pop}}_j$ for all $1 \leq j \leq n$,

\item $\text{\textit{skip}}$, corresponding to the identity function of $\mathscr{S}(\Gamma)$.

\end{itemize}

% \begin{definition}
\begin{samepage}
\par\noindent\ignorespacesafterend
A {\em higher-order pushdown automata of level $n$} (or $n$-HPDA for short) 
is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
where 
\begin{itemize}
        \item $Q$ is a non-empty finite {\em set of  states},
        \item $\Gamma$ is a non-empty finite {\em  stack alphabet},
	\item $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules},
        \item $q_{init}$ is the {\em initial  state},
        \item $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
initial stack symbol}, and
	\item $F\subseteq Q$ is a {\em set of final  states}.
		\end{itemize}
\end{samepage}
% \end{definition}

\par\noindent\ignorespacesafterend
By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
we denote the set of {\em configurations} of an $n$-HPDA $\H$. 
%is a pair $(q, s)$ where $q \in Q$ and $s$ is an $n$-stack over $\Gamma$. 
As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.

An $n$-HPDA $\H=(Q, \Gamma, R, q_{init}, F)$
induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
where
for all $q,q'$ in $Q$,
for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
there exists some rule
%$(q,\gamma,q',\theta) \in R$
$(q,\gamma,q',op) \in R$
such that
$\text{\textit{top}}(s) = \gamma$
and
%$s' = \theta(s)$.
$s' = op(s)$.

\iffalse
For our constructions it would be simpler to assume that k-HPDS can work
also on stacks of lower levels, in particular on 1-stacks. Of course we can always
simulate a j-stack, for j < k with an k-stack but in the notation it requires some
additional parenthesis that make it less readable.

To define a game on the graph of a HPDA, we assign a player to each control
state, and we consider an initial configuration: a game structure on a HPDS H
is a tuple G = (H, P0 , P1 , s0 ), where P = P0 ⊎ P1 is a partition of the control
states of H, and s0 ∈ Sk . This extends naturally to a partition of the set of
configurations: with the notations of Section 2.1, V0 = P0 × Sk , V1 = P1 × Sk ,
and E is defined above.
\fi

Again we are interested in parity games.
As expected, given an 
 $n$-HPDA $\mathcal{H}$
 and a partition of
$Q$ into $Q_0$ and $Q_1$,
we partition the configurations of
$T_{\mathcal{H}}$
into
$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
and
$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
%
%
With these notations in mind one can define the arena
$$
A_{(\mathcal{H}, Q_0, Q_1)}
=
(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$$
\par\noindent\ignorespacesafterend
induced by an $n$-HPDA $\mathcal{H}$ and a partition of its set of states.

As expected, given a 
priority function $\Omega: Q \to [0, m]$,
we naturally extend the function
as follows,
by setting
$\Omega_{\mathscr{S}_n(\Gamma)}: \Conf(\mathcal{H}) \to [0, m]$
and
$\Omega_{\mathscr{S}_n(\Gamma)}(q, s) = \Omega(q)$
for all
$s \in \mathscr{S}_n(\Gamma)$.

Concerning higher-order pushdown automata, we are interested in the following decision problem. 
% \newpage

\begin{samepage}
\problemx{$n$-HPDA parity game}
{An $n$-HPDA $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
where $Q = Q_0  \biguplus Q_1 $, 
and
a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
{Does player $0$ have a winning strategy 
from $q_{init}(\text{\textit{push}}_1^{\gamma_{init}}(\epsilon_{n}))$ 
for the parity game 
$\mathcal{G}= (A_{(\H, Q_0 , Q_1)}, \win_{\Omega_{\mathscr{S}_n(\Gamma)}})$ ?\newline}
\end{samepage}

It was shown in \cite{Cach03} that {\sc $n$-HPDA parity game} can
be solved in $n$-$\EXP$. This also gives an $n$-$\EXP$ algorithm for the $\mu$-calculus
model checking over transitions systems induced by  $n$-HPDA.
In \cite{cachat2007complexity} the matching lower bound was showed, even in the case of reachability games,
hence showing $n$-$\EXP$-completeness of {\sc $n$-HPDA parity game}.

\begin{theorem}{\cite{ Cach03, cachat2007complexity}}
{\sc $n$-HPDA parity game} is $n$-$\EXP$-complete.
\end{theorem}

% \subsection{From pebble pushdown automata to higher-order pushdown automata}

In a $2$-HPDA, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
allows to ``copy'' the top level $1$ stack. The current word is hereby stored away and left untouched
until the next operation $\text{\textit{pop}}_2$,
while $\text{\textit{push}}_1$ and $\text{\textit{pop}}_1$ can be performed on the additional ``copy'' in the meantime. This behavior is similar to that of dropping and lifting a pebble. 
The main difference is that there is no operation to test that the ``copy'', after many updates, is again equal to the ``original'', i.e. there is no operation to syntactically test that the two topmost level $1$ stacks are identical.

% But cannot test

\iffalse
In an $n$-HPDA, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
allows to ``copy'' the top stack of level $j$ for all $2 \leq j \leq k$. 
\fi

In \cite{CaWoe03, Woeh05, carayol2006automates} however 
%the authors
Carayol and W\"ohrle
 introduced a variant of $n$-HPDA by extending the 
$\text{\textit{pop}}_j$ operations for $2 \leq j \leq n$ with a built-in equality test. 
For $2 \leq j \leq n$ 
the new operation
$\text{\textit{pop}}_j^=$ has the same effect as $\text{\textit{pop}}_j$, but
can only be applied if the two top level $j$ stacks coincide.
In \cite{carayol2006automates} it is seen as a symmetrical operation in comparison to
$\text{\textit{push}}_j$.\\

A {\em higher-order pushdown automaton with equality pop of level $n$} ($n$-HPDA$^=$ for short) 
is a higher-order pushdown automaton of level $n$ where in
% Instrn
$\Op_n(\Gamma)$
 the operation 
 %popk
 $\text{\textit{pop}}_j$  
  is replaced by 
%  pop=
%k 
$\text{\textit{push}}_j^=$
%for 1 ≤ k ≤ n. 
for $2 \leq j \leq n$.
We denote
this new set of operations by 
% Instr= n
$\Op_n^=(\Gamma)$. 
More formally the new operations $\text{\textit{push}}_j^=$ are defined as
\begin{eqnarray*}
&\text{\textit{pop}}_k^=( \langle s_0 \rangle \langle s_1 \rangle \ldots  \langle s_m \rangle \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle , \quad \text{ and }\\
&\text{\textit{pop}}_j^=( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle \langle s_1 \rangle \ldots 
\langle \text{\textit{pop}}_j^=(s_m) \rangle,
\text{ if } 2 \leq j < k. 
\end{eqnarray*}
%
%
%
%
\iffalse
Dans ce sous-paragraphe, nous présentons formellement les automates à pile
de niveau k sur le jeu d'opérations symétriques Opsk et sur le jeu d'opérations
classiques COpsk . Nous donnons à cette occasion un bref historique de la notion
d'automate à pile de piles. Enfin, nous établissons l'équivalence entre les deux
modèles d'automates à chaque niveau en tant qu'accepteurs de langages. Ce résultat a été obtenu dans [CW03] en collaboration ave Stefan Wöhrle. La preuve
complète apparait dans [Wöh05]. Indépendamment, ce résultat a été obtenu dans
[Fra05].
\fi
%
Transition systems induced by higher-order pushdown automata with equality pop of level $n$
are then defined as expected.  
Carayol and W\"ohrle proved \cite{Woeh05, carayol2006automates} that the two models,
namely
higher-order pushdown automata with equality pop of level $n$
and
higher-order pushdown automata of level $n$,
generate the same classes of transition systems.

\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA (resp. $n$-HPDA$^=$) then there
exists an $n$-HPDA$^=$ (resp. $n$-HPDA) $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}

\iffalse
\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA  then there
exists an $n$-HPDA$^=$  $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}
% Proposition 3.12

\begin{theorem}{\cite{Woeh05, carayol2006automates}}
If $\H$ is an $n$-HPDA$^=$ then there
exists an $n$-HPDA $\H'$ such that
$\H$ and $\H'$ %generate the same graph.
induce isomorphic transition systems.
\end{theorem}

% Proposition 3.18 If A is a higher-order pushdown system with equality pop
% of level n then there exists a higher-order pushdown system B of level n such
% that A and B generate the same graph.
\fi

From $n$-HPDA to $n$-HPDA$^=$ (Proposition 3.12 in \cite{Woeh05}) the proof relies on recreating a correct stack content to simulate higher-order 
$\text{\textit{pop}}$ operations by $\text{\textit{pop}}^=$, essentially ``guessing'' the correct stack content to be able to apply $\text{\textit{pop}}^=$. 
In the other direction (Proposition 3.18 in \cite{Woeh05}), 
the author
enrich the stack alphabet with new symbols stating which 
instructions have to be executed to recreate a
previous stack content. It is proven that such en encoding is possible since
there is 
for every stack $s$ of level $n$
% there is
 a unique shortest sequence of instructions which creates $s$ from
$\epsilon_n$. \\

Defined like {\sc $n$-HPDA parity game}, we write {\sc $n$-HPDA$^=$ parity game}  
in case the input is an $n$-HPDA$^=$ rather than an $n$-HPDA.

The algorithm from \cite{Cach03} actually provides an algorithmic solution to parity games on the graphs of the Caucal hierarchy. 
We skip a formal definition of a graph of level $n$ in the Caucal hierarchy, and refer the reader
to \cite{Cau02mfcs} for more details.
The $n$-$\EXP$ upper bound on $n$-HPDA parity games follows from the fact that every transition system induced by an $n$-HPDA $\H$ is a graph of the Caucal hierarchy \cite{Cach03, Woeh05},
whose vertices are almost in one-to-one correspondence with the configurations of $\H$.
In \cite{Woeh05} the converse direction is proven, i.e. that every graph of level $n$
of the Caucal hierarchy is generated by an $n$-HPDA.  
%
%
In \cite{carayol2006automates}
it is similarly shown that
every transition system induced by an $n$-HPDA$^=$ $\H$ is a graph of the Caucal hierarchy.
%
%
The algorithm from \cite{Cach03} hence lead to a solution for
$n$-HPDA$^=$ parity games as well.

\iffalse
\textcolor{red}{
Now then, because the class of $n$-HPDA$^=$ transitions systems
is the same as
the class of $n$-HPDA transitions systems, we have the following result.
}
\fi

\begin{theorem}\label{HPDA parity game}
{\sc $n$-HPDA$^=$ parity game} is in $n$-$\EXP$.
\end{theorem}

\iffalse
"To show that such an automaton exists we introduce the notion of a weak
popping higher-order pushdown automaton. A weak popping automaton is only
allowed to execute a pop instruction of level $j \geq 2$ if the two top level $j$ stacks
coincide. We skip a formal definition of a weak popping higher-order pushdown
automaton and just mention that even though this automaton model is equipped
with a built-in test on the equality of two stacks of the same level, it is equivalent
to the usual model. All proofs are given in the full version of this article [4]."

Pour pouvoir obtenir une notion pertinente d'ensemble rationnel de piles de
piles, il faut considérer un jeu d'opérations où la destruction inconditionnelle des
piles est remplacée par une version plus symétrique où la dernière pile de niveau ℓ
ne peut être détruite que si elle est égale à la précédente. Nous établissons dans le
paragraphe 4.1 l'équivalence (en temps qu'accepteurs de langages) les automates
à pile définis sur ces deux jeux d'opérations ( f. théorème 4.1.23). Nous montrons,
dans le paragraphe 4.6, le manque de pertinence de la notion de rationalité induite
par le jeu d'opérations classiques et par là même, nous établissons la nécessité de
considérer le jeu d'opérations symétriques.

L'opération de destruction symétrique a été introduite pour des raisons techniques dans 
[CW03] et a été utilisée de manière 
systématique dans [Car05] et
[Fra05]. 
Nous montrerons, dans le sous-paragraphe 4.1.5, que les automates à pile
de piles définis en utilisant les opérations destrk ou les opérations copyk acceptent
les mêmes langages.

Les automates à pile d'ordre supérieur ont été introduits dans les années 70.
Dans [Gre70℄, Greibach attribue l'idée de ces automates à Aho et Ullman. Ces
automates sont aussi définis dans [Mas76℄. Les automates sur COpsk diffèrent
légèrement des automates considérés par ces auteurs et de ceux onsidérés dans
[DG86, Eng91℄. La différence majeure réside dans la définition des piles de piles.
Les piles de niveau k sont des suites non vides de couples formés d'un symbole de
pile et d'une pile de niveau k − 1. Comme remarqué dans [KNU02℄, ces symboles
de pile supplémentaires peuvent être simulés dans le modèle des automates sur
COpsk .

Les langages a eptées par les automates à pile sur COpsk sont onnus sous
le nom de langages k -OI ou de langages indexés de niveau k . La première déno-
mination vient des travaux de Damm [Dam82]

Nous allons maintenant établir que les encodages des opérations simulent bien
les opérations sur les encodages des piles.

Lemme 4.1.21. Pour toute pile s ∈ Stacksk (Γ) et pour tout θ ∈ Opsk ,
[[ θ ]]k ([[ s ]]) =
{
[[ θ(s) ]]		si θ(s) est défini,	
non définie		sinon.

Proposition 4.1.22. Tout langage accepté par un automate à pile sur Opsk est
accepté par un automate à pile sur COpsk .

Démonstration. Soit A = (Γ,Σ,τ,Q,I,F,∆) un automate à pile sur Ops⋆k . Nous
définissons un automate B = (Γ,Σ,τ,Q,I,F,∆B ) où l'ensemble des transitions ∆B
est défini par:
∆B = {(p,x,η,q) | (p,x,θ,q) ∈ ∆ et η ∈ [[ θ ]]k }.
Par le lemme 4.1.21, il suit que pour tout w ∈ (Σ)∗ , p ∈ Q et s ∈ Stacksk (Γ):
w
(q0 , [ ]k ) −→ (p,s)
A
⇔
w
(q0 , [ ]k ) −→ (p,[[ s ]]).

Théorème 4.1.23. Les langages acceptés par les automates à pile sur Opsk et
sur COpsk coïncident.

\fi

\subsection{From pebble pushdown 
automata 
to higher-order pushdown 
automata
}

We use HPDA$^=$ parity games to simulate pebble pushdown automata
parity games.
A similar approach was used in \cite{carayol2006automates} to show that $(n + 2)$-level stack
automata could be used to simulate $n$-pebble alternating two-way word automata.

Let us start by discussing the simulation of the dropping and lifting of a pebble in the case of
a pebble pushdown automaton $\ppda$ with only one pebble.
As long as no pebble is dropped, a $2$-HPDA$^=$ $\H$ can simulate 
the behavior of a pebble pushdown automaton $\ppda$ with a $2$-stack containing a single $1$-stack
on which it performs the same operations $\ppda$ performs on its stack.
When the automaton $\ppda$ is in configuration $q(w)$ and drops a pebble, instead of making a pop or a push, we need
to %go to state $q'$ and word $w$ and 
	store the information that the pebble has been
dropped on $w$.
To simulate the next configuration in such a case, we use 
$\text{\textit{ push}}_2$
to store the information that the pebble has been dropped on the word $w$, leading to the new  $2$-stack
$\langle w \rangle \langle w \rangle$. 
Configurations afterwards have $2$-stacks of length two where the first component remains
$w$ until the pebble is lifted.
Lifting the pebble can be done only at the position the
pebble was dropped, by using $\text{\textit{pop}}^=_2$.
Checking that the node
corresponds to the one where the pebble has been dropped simply makes use of
the composition of $\text{\textit{pop}}^=_2$ and $\text{\textit{push}}_2$.
Checking the absence of the pebble can be done by challenging
the opponent to prove the presence of the pebble.

Now, let us detail the case when there is a second pebble, the construction
for each additional pebble being highly similar, where every additional pebble after
the first one would require the addition of another stack level.
Similarly as before, 
as long as no pebble is dropped, the stack of $\H$ 
is a $3$-stack containing a
single $2$-stack that contains a single $1$-stack.
To simulate configurations in the case one pebble have been dropped, the stack contains a single
 $2$-stack of the form $\langle w_1 \rangle \langle w \rangle$.
In the case the two pebbles have been dropped, 
the stack is of the form $ \langle \langle w_1 \rangle \langle w_2 \rangle \rangle
\langle \langle w_1 \rangle \langle w \rangle \rangle$.

Intuitively speaking, $w_1$ is, as before, the position where the first pebble is
placed, and $w_2$ is where the second pebble is placed. Dropping pebbles like this
again involves the cloning operation, only, each pebble operates at a different 
stack level: thus, dropping the first pebble will use $\text{\textit{push}}_2$
and dropping the second pebble will use $\text{\textit{push}}_3$. Knowing the number of pebbles dropped can be done by 
keeping track using the states of $\H$.
Lifting or
checking for the presence 
(or absence) 
of the pebbles functions on a similar
basis as before. Checking the presence of the first pebble uses 
operation $\text{\textit{pop}}^=_2$ followed by $\text{\textit{push}}_2$,
while checking the presence of the second pebble uses 
operation $\text{\textit{pop}}^=_3$ followed by $\text{\textit{push}}_3$.
Lifting the first pebble uses $\text{\textit{pop}}^=_2$, but
check first that the second pebble has already been lifted.
Lifting the second pebble simply uses $\text{\textit{pop}}^=_3$.

By expanding this reasoning inductively, we conclude that given $n \in N$, and
given an $n$-pebble pushdown automaton 
$\ppda = (Q, \Gamma, R, q_{init}, \gamma_{init}, F)$, where $Q= Q_0 \biguplus Q_1$,
one can compute an $(n+1)$-HPDA 
$\H = (Q', \Gamma, R', q_{init}', \gamma_{init}', F')$  where $Q'= Q_0' \biguplus Q_1'$, 
such 
that 
player $0$ has a winning strategy from $q'_{init}(\text{\textit{push}}_1^{\gamma_{init}'}(\epsilon_{n}))$ in $A_{(\H,Q'_0,Q'_1)}$ 
if
and only if
player $0$ has a winning strategy from $q(\gamma_{init},\mu_{init})$ in $A_{(\ppda,Q_0,Q_1)}$. 
Hence the following reduction.

\begin{theorem} 
{\sc $n$-pebble pushdown parity game} is polynomial time reducible
to {\sc $(n+1)$-HPDA$^=$ parity game}.
\end{theorem}

The reduction implies decidability of pebble pushdown automata parity game. Furthermore,
by Theorem~\ref{HPDA parity game}, it implies the following complexity result.

\begin{theorem}
{\sc $n$-pebble pushdown automata parity game} is in $(n+1)$-$\EXP$.
\end{theorem}

Finally, the following complexity result is due the above and Theorem~\ref{PPDA reduction}.

\begin{corollary}
{\sc $n$-parametric pushdown automata parity game} is in $(n+1)$-$\EXP$.
\end{corollary}

% \subsection{Discussion and open problems}
\section{Discussion and open problems}\label{discuss ppda}

In this section we have shown that
deciding
the winner of a parametric pushdown parity game or reachability game is nonelementary in general, but
decidable and in $(n+1)$-$\NEXP$ when the number $n$ of parameters is fixed.

	For the lower bound
we reduced the $\FO$ satisfiability problem on words \---- known to be nonelementary
from \cite{Sto74} \---- to the
problem of deciding whether player $0$ has a winning strategy for a parametric pushdown reachability game.

	For the decidability upper bound, we  used
	pebble pushdown parity games
	to solve 
	parametric pushdown parity games, and 
	 higher-order pushdown parity games to solve pebble pushdown parity games.
 Since solving parity games on higher-order pushdown automata with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown automata with $n$ parameters.

% \mh{question ofreducing PATWA to pebble pushdown is kinda easy, bot now other question: reducing pebble pushdown to parametric pushdown, doable or not ? since register have only 3r values they can 'remember' maybe it can be shown something along theses lines for pebble pushdown idk}

 Since a particular case of pushdown automata consists in one-counter automata, extensions of automata with a counter generally inherit known upper bounds on pushdown automata. 
This however is not the case in
the parametric extensions considered in this thesis, as parametric updates
% in PTOCA
 are not trivially handled by our PPDA model, i.e. a PPDA over a unary alphabet plus a bottom-of-stack symbol is not trivially a PTOCA nor a POCA. 
For solving parametric pushdown parity games, on the other hand, we introduced the notion of pebble pushdown parity games, and
pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
can be seen as a form of one-counter automata extended with pebbles. 
A pebble one-counter automata is then a one-counter automata that
can use a set of pebbles as markings on the set of non-negative integers, with a 
$\text{\textit{drop}}$ recording the current counter value, and a 
$\text{\textit{lift}}$ poping the last dropped
pebble while requiring that the current counter value 
corresponds to the one from the last dropped pebble.
%
%Hence, u
Unlike for parametric extensions, 
decidability of 
what should be called
pebble one-counter parity games follows from
decidability of pebble pushdown parity games. 
We believe it natural to study such a pebble one-counter automata model
on its own.
In particular we believe that the question of whether or not
solving pebble one-counter reachability games is nonelementary is worth investigating.
% It may very well be the case that the hierarchy collapses instead, at level e.g. 2 or 3.

Of note is that reachability games and parity games are 
not the only problems one can consider, as one can
also explore the complexity of reachability itself. 
In particular
one can 
% ask whether or not the parametric pushdown automata model described in this section  admits effective preservation of regularity.
study the precise complexity of the
{\sc $n$-PPDA reachability} problem.
Since reachability can be viewed as a reachability game, we know that
{\sc $n$-PPDA reachability} is decidable in $(n+1)$-$\EXP$, 
but the precise complexity of the problem remains unknown.
We are furthermore convinced that it is worthwhile to investigate further comparisons between parametric pushdown automata, pebble pushdown automata, 
%pebble alternating tree automata 
and register pushdown automata, especially since the latter have been shown to only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers.

% \newpage

### 2023-06-11 reachability/article.tex
up

@@
 \renewcommand{\R}{\mathbb{R}}
+\renewcommand{\H}{\mathcal{H}}
 \newcommand{\Z}{\mathbb{Z}}
@@
 \newcommand{\Const}{\mathsf{Consts}}
+\newcommand{\Conf}{\mathsf{Conf}}
 \newcommand{\PTA}{\textsc{PTA}}
@@
 \renewcommand{\poly}{\mathrm{poly}}
+
+\newcommand{\Op}{\mathsf{Op}}
+\newcommand{\win}{\textsc{Win}}
 
@@
 
+%\subsection{A decidability result from reduction to parity games on automata with pebbles}\label{pebbles}
+\section{%A decidability result via r
+Reduction to pebble pushdown automata parity games
+}\label{pebbles}
+
+% \mh{Motivates the use of pebbles. See pebbles on two-way automata, two-way tree automata, weighted automata, etc.}
+
+% 21. Globerman, N., Harel, D.: Complexity results for two-way and multi-pebble automata and their logics. Theor. Comput. Sci. 169, 161–184 (1996)
+%		same as [13] bellow
+%
+% Engelfriet, J., Hoogeboom, H.J.: Tree-walking pebble automata.
+In: Karhumäki, J., Maurer, H., Pa ̆un, G., Rozenberg, G.
+(Eds.) Jewels are Forever, Contributions to Theoretical Computer Science in Honor of Arto Salomaa, pp.
+72–83.
+Springer, Berlin (1999) for tree-walking automata.
+%
+% @incollection{engelfriet1999tree,
+%  title={Tree-walking pebble automata},
+%  author={Engelfriet, Joost and Hoogeboom, Hendrik Jan},
+%  booktitle={Jewels are forever},
+%  pages={72--83},
+%  year={1999},
+%  publisher={Springer}
+%}
+% 
+%
+% Two-way pebble transducers for partial functions and their composition, Joost Engelfriet
+%
+% Tree-Walking Pebble automata, Joost Engelfriet and Hendrik Jan Hoogeboom 
+A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
+pebbles that it can drop on, and lift from 
+words, using them as markers.
+Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
+\cite{globerman1996complexity, engelfriet1999tree}.
+Automata with nested pebbles were also introduced
+for tree-walking automata. 
+It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}%, the main trouble being that tree-walking automata get lost rather easily
+.
+Using pebbles is a remedy against getting lost along a tree, but 
+the unrestricted use of pebbles leads to a class of tree languages much larger than the
+regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
+Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline.
+It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
+and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.
+
+In this section, we extend these ideas to pushdown automata.
+Instead of using pebbles as markings on their input, pebble pushdown automata have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
+and a \textit{lift} pops the last dropped pebble 
+while requiring that
+% assuming
+ it was placed on the current node. 
+% Thus hierarchicality on the registers works slightly differently from hierarchicality on the pebbles, hierarchicality on the registers being a syntactic property, whereas hierarchicality on the pebbles is a semantical one.
+One can think of a pebble as a register that can store a stack content for later comparisons.
+
+We first define more formally our pebble pushdown automaton framework, and then provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
+
+%\subsubsection{Pebble pushdown automata}
+\subsection{Pebble pushdown automata}
+
+\newcommand{\ppda}{\mathcal{I}}
+% \Denarius
+% \vernal
+% \mathscr{Z}
+% \mathcal{I}
+% \pluto
+% \capricornus
+% \Bumpeq
+% \Pfund
+
+An {\em $n$-pebble pushdown automaton} 
+is a tuple 
+$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
+where
+\begin{itemize}
+\item $Q$ is a {\em finite set of  states},
+\item $\Gamma$ is a {\em finite stack alphabet},
+\item  $R  \subseteq  Q  \times \Gamma \times
+		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
+		 \times Q  \times (\Op(\Gamma) \cup \{\text{\textit{drop}}, \text{\textit{lift}}\})$ is a {\em finite set of rules},
+		 where the fourth element $S$ of a rule $r$ is a subset of 
+		$\mathcal{P}(\{1,\ldots,i\})$ where $i$ is the third element of $r$,
+		and if the last element is
+		$\text{\textit{lift}}$ then we additionally require $i \in S$,
+\item $q_{init} \in Q$ is an {\em initial  state}, 
+\item $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
+\item $F \subseteq Q$ is a {\em set of final  states}.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+% \mh{remind the reader what a partial function is}
+
+\par\noindent\ignorespacesafterend 
+Recall that for \iffalse a partial function $f$ from a set $S$ to a set $X$ 
+%is a function defined on a subset $C$ of $S$ (possibly $S$ itself) with output values in $X$, and 
+that is defined on a subset $C \subseteq S$
+is denoted by
+$ f : S \rightharpoonup X $.
+
+For \fi a partial function $ f : S \rightharpoonup X $ (see page~\pageref{partial}), for notational purposes, we consider some element $\bot_X \not\in X$ and 
+ associate $f$ with the function returning the 
+bottom element $\bot_X$ when $f$ is undefined.
+Thus we write $(X \biguplus \{ \bot_X \})^S$ for the set of all partial functions from $S$ to $X$, which we here abbreviate as $(X \biguplus \{ \bot \})^S$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% An $i$-configuration of $\ppda$ is an element of
+%$ Q\times W^* \times (W^*)^{\{1, \ldots i\}}$.
+% By $\Conf_i(\ppda)$ we denote the set of $i$-configurations of $\ppda$.
+By $\Conf(\ppda)=Q\times \Gamma^* \times (\Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$ we denote the set of
+% By $\Conf(\ppda) = \bigcup_{0 \leq i \leq n} \Conf_i(\ppda)$ we denote the set of
+{\em configurations} of $\ppda$. As expected, we rather write $q(w, \mu)$ instead of $(q, w, \mu)$.
+An {\em$i$-configuration} for $i > 0$ of $\ppda$ is a configuration $q(z,\mu)$ where
+$\text{Dom}(\mu) = \{1, \ldots, i\}$, while a $0$-configuration is a configuration 
+ $q(z,\mu)$ where $\text{Dom}(\mu)=\emptyset$.
+
+The idea of a transition $(q, a, i, S, q', m) \in R$
+is that, if the automaton $\ppda$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
+$\ppda$ goes to state 
+$q'$ and makes modifications to the stack or the pebbles according to
+$m$. 
+% \mh{The set $S$ is used for testing the presence of certain pebbles.}
+Note a pebble 
+can be lifted only if the stack content which corresponds to the pebble
+is the same as the current stack content. 
+%$\Dzeta$ $\zeta$ $\Xi$
+This is enforced by syntactically requiring
+the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.
+
+\iffalse
+A {\em weak} $n$-pebble pushdown system is an $n$-pebble pushdown system $\ppda = (Q, \Gamma,  \Delta )$ in which $\text{\textit{lift}}$ operations also require the presence of the last pebble dropped, i.e.
+for all $(q, a, i, S, q', \text{\textit{lift}}) \in \Delta$, we have $ i \in S $. 
+%  is an additionnal restriction $f(i) = wa$ that needs to hold for the application of the $lift$ operation.
+{\em Weak $n$-pebble pushdown games} are defined naturally as $n$-pebble pushdown games on 
+weak $n$-pebble pushdown systems.
+
+\fi
+
+A {\em pebble set} of $\ppda$ is a set $U \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
+{\em $U \!$-pebble assignment} is a function which maps each $j \in U$ to a word in $ \Gamma^*$.
+The $\emptyset$-pebble assignment is denoted by $\mu_{init} :  \{ 1, \ldots, n\} \rightharpoonup  \Gamma^* $ and is the 
+%partial function from  $\{ 1, \ldots, n\}$ to $ \Gamma^* $ that
+%does not map any $j \in \{ 1, \ldots, n \}$ to an output.
+totally undefined function.
+
+\iffalse
+For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, \mu )$, where $w \in \Gamma^*$, $q$ is a state and
+$\mu$ a $\{ 1, \ldots , i \}$-pebble assignment. 
+We call $w$ the current stack, $q$ the current state and
+$\mu$ the current pebble assignment. 
+We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
+$\mu(j) = w_j$, for each
+$j \leq i$.
+
+The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. 
+\fi
+
+\begin{samepage}
+%\begin{definition}
+An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
+a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
+ such that
+for all $(q, a, i, S, q', m) \in R$, with $a \in \Gamma$,
+ for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
+ %(respectively $\mu(j)= {\gamma_{init}}  $) 
+ for each
+$j \in S$ 	and
+ $\mu(j) \neq wa$ 
+ %(resp. $\mu(j) \neq {\gamma_{init}} $) 
+ for each $j \in \{ 1, \ldots , i \} \setminus S$,
+the following holds
+\begin{itemize}
+\item if $m \in \Op(\Gamma)$, 
+% $(q,wa,\mu) \rightarrow_{\ppda} (q',wm, \mu)$, 
+either
+\begin{itemize}
+\item $ m = \text{\textit{push}}^\gamma$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa\gamma, \mu)$,
+
+\item $ m = \text{\textit{pop}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',w, \mu)$, or
+
+\item $ m = \text{\textit{skip}}$ and $(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu)$,
+\end{itemize}
+\item if $ m = \text{\textit{drop}}$, 
+$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$,  
+where $\mu'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
+$\mu'(j) = \mu(j)$, for each
+$j \leq i$, and $\mu'(i+1) = wa$, and
+\item if $m = \text{\textit{lift}}$, and $i \in S$, i.e. the last pebble dropped belong of the set of pebble we test the presence of,
+$(q,wa,\mu) \rightarrow_{\ppda} (q',wa, \mu')$, 
+where $\mu'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
+$\mu'(j) = \mu(j)$, for each
+$j < i$.
+\end{itemize}
+%\end{definition}
+\end{samepage}
+
+\par\noindent\ignorespacesafterend
+We are interested in games over pebble pushdown automata, % namely, 
+%reachability, B\"uchi, and parity games.
+mainly, parity games.
+
+Again, given
+a 
+%labeled
+ transition system, one needs only to provide 
+%a mapping $\Omega$ and 
+a partition of the set of
+configurations 
+%$S$ into two sets $S_0$ and $S_1$
+ to obtain an arena.
+ %
+Given a partition of
+$Q$ into $Q_0$ and $Q_1$, 
+we partition the configurations of
+$T_{\ppda}$
+into
+$\Conf_{\ppda,0}=Q_0\times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$
+and
+$\Conf_{\ppda,1}=Q_1\times  \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$.
+
+With these notations in mind one can  define the arena
+$$
+A_{(\ppda, Q_0 , Q_1)}
+=
+(\Conf_{\ppda,0}, \Conf_{\ppda,1}, \rightarrow_{\ppda}).$$
+
+As expected, given a 
+priority function $\Omega: Q \to \{0, \ldots m\}$,
+one naturally 
+set the extension of $\Omega$ as
+$\overline{\Omega}: \Conf(\ppda) \to \{0, \ldots m\}$
+and
+$\overline{\Omega}(q, w,\mu) = \Omega(q)$
+for all
+$w \in  \Gamma^*$
+and
+$\mu \in ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$. \\
+
+Concerning pebble pushdown automata, we are interested in the following decision problem.
+
+\problemx{$n$-pebble pushdown parity game}
+{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
+where $Q = Q_0  \biguplus Q_1 $, 
+and
+a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
+{Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
+$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?\newline}
+
+% \subsubsection{Reduction from parameters to pebble pushdown automata}
+\subsection{From parametric pushdown 
+automata 
+to pebble pushdown automata
+}
+
+We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
+
+\begin{theorem}\label{PPDA reduction}
+%The problem of 
+{\sc $n$-parametric pushdown parity game} 
+is polynomial time reducible to
+%the problem of 
+{\sc $n$-pebble pushdown parity game}.
+\end{theorem}
+
+\begin{proof}[Sketch]
+
+Let us fix some
+$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
+P%\{p_1, \ldots, p_n \}
+, q_{init}, { \gamma_{init}} , F)$,
+ disjoint
+union $Q = Q_0  \biguplus Q_1$
+and $P = P_0 \biguplus P_1$,
+and
+some priority function 
+$\Omega : Q \times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^R \to [0, m ]$.
+
+We construct
+% polynomial time
+  an 
+ $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
+a 
+ disjoint
+union $Q' = Q_0'  \biguplus Q_1'$
+and a mapping $\Omega' : Q \times  \Gamma'^* \times ( \Gamma'^* \biguplus \{ \bot \})^{\{1, \ldots, n\}} \to \{1, \ldots, m \}$,
+such that
+ %:
+%
+player $0$ has a winning strategy from 
+$(q'_{init},\gamma_{init}, \mu_{init})$
+in
+		\iffalse
+		for the parity condition on
+		$\mathcal{G}_{(\ppda, Q'_0, Q'_1, \Omega')}$.
+		\fi
+the parity game
+$\mathcal{G'}= (A_{(\ppda, Q'_0, Q'_1)}, \win_{\overline{\Omega'}})$
+if and only if
+player $0$ has a 
+		\iffalse
+		strategy $\sigma_0$ for the parameter valuation game  such that
+		for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
+		player $0$ have a
+		winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
+		$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$. 
+		\fi
+winning strategy from $\mu_\bot$ in 
+the parity game
+$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
+where		 
+$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
+and
+$\widehat{\Omega}$
+correspond to the definitions on page~\pageref{PPDA reachability game}.
+
+%
+%
+%
+%
+
+The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
+(one for every 
+%total 
+parameter assignment)
+ of the pushdown automaton with $n$ parameters by using pebbles to represent the parameters. 
+The $n$-pebble pushdown automaton will first simulate the 
+parameter valuation arena.
+Players take turns chosing particular stack contents on which to place pebbles,
+using a gadget similar as the one from Figure~\ref{quantifiers}. 
+Once every parameter has been assigned with the dropping of a corresponding pebble, 
+the pebble assignment $\mu$ is fixed.
+%%
+%
+%
+%
+%
+%
+%
+
+Further configurations in the $n$-pebble pushdown automaton then consist of a word representing the status of the stack, a state, and a fixed pebble assignment.
+For a fixed pebble assignment there is then a one for one correspondance between configurations of the $n$-parametric pushdown
+automaton 
+%transition system with assignment $\mu$
+and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
+the set of states of the $n$-parametric pushdown automaton, and 
+so are player $0$ states, player $1$ states and the priority mapping.
+\end{proof}
+
+\section{Reduction to higher-order pushdown automata}\label{section HPDA}
+
+Higher-order pushdown automata (HPDA for short) were introduced as a 
+generalization of pushdown automata \cite{aho1969nested,Gre70,maslov1976multilevel}. 
+% Alfred V. Aho. Nested stack automata. J. ACM, 16(3):383–406, 1969.
+% S. Greibach. Full AFL’s and nested iterated substitution. Inf. Control, 16(1):7–35, 1970.
+% A.N. Maslov. Multilevel stack automata. Problemy Peredachi Informatsii, 12:55–62, 1976.
+A stack of a pushdown automaton is seen as a {\em level $1$ stack}. 
+A pushdown automaton of level $2$ (or $2$-HPDA) then works with a stack of level $1$ stacks. 
+In addition to the ability to push
+and to pop a symbol on the top-most level $1$ stack, an $2$-HPDA can copy or remove the entire topmost level $1$ stack. 
+The definition generalizes to any $n \geq 2$, and $n$-HPDA are similarly defined for all level
+$n$ as automata working with a stack of level $(n-1)$ stacks.
+% HPDA have been extensively studied as language recognizers [Dam82,Eng83].
+
+We recall the definition from \cite{cachat2007complexity} which itself is taken from \cite{knapik2002higher}. We then provide a
+reduction from the problem of solving pebble pushdown parity games to the problem of solving
+higher-order pushdown parity games.
+
+\subsection{Higher-order pushdown automata}
+
+A {\em level $1$ stack} (or {\em $1$-stack}) over an alphabet $\Gamma$ is 
+simply a %word in $\Gamma^*$
+	stack over $\Gamma$, i.e. a word  in $\Gamma^*$.
+% an arbitrary sequence $w_0 w_1 \ldots w_n $ of elements of $\Gamma$ for some $n \in \N$. 
+A  {\em level $n$ stack} (or {\em $n$-stack}) over an alphabet $\Gamma$, for $n \geq 2$, is a 
+non-empty sequence
+$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
+of
+$(n-1)$-stacks over $\Gamma$,
+for some $m \in \N$.
+The set of $n$-stacks over $\Gamma$ is denoted by $\mathscr{S}_n(\Gamma)$,
+or simply $\mathscr{S}_n$ in case the set $\Gamma$ is obvious from context.
+The set of all stacks over $\Gamma$ is written $\mathscr{S}(\Gamma) = \bigcup_{n\in \N} \mathscr{S}_n(\Gamma)$.
+We define $\epsilon_{1}$ as
+$\epsilon \in \Gamma^*$ and we inductively define 
+$\epsilon_{n} = \langle \epsilon_{n-1} \rangle$ in $\mathscr{S}_n$ for all $n > 1$.
+
+A {\em higher-order stack operation} is a partial function from $\mathscr{S}(\Gamma)$ 
+to $\mathscr{S}(\Gamma)$ which preserves the level of the input (i.e. the image of
+an $n$-stack is an $n$-stack for all $n \in \N$). The {\em level } of an operation $op$
+is the smallest $n \in \N$ such that $\text{Dom}(op) \cap \mathscr{S}_n \neq \emptyset$. 
+The operations 
+%we consider
+additionally
+ respect the hierarchicality of higher-order stacks, i.e. in
+a level $n+1$ stack only the topmost level $n$ stack can be accessed. An operation
+$op$ of level $n$, applied to a level $n+1$ stack 
+$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$
+of length $m \in \N$,
+thus
+returns the output
+$ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle op(s_m) \rangle$
+if applicable.
+The definition for  all levels of stacks greater than $n$ follows the same pattern.
+
+% For all $m \in \N$, 
+The following operations can be performed on a $1$-stacks of length $m \in \N$.
+% $ w_0 w_1 \ldots w_m \in \Gamma^*$.
+\begin{eqnarray*}
+&\text{\textit{push}}_1^\gamma( w_0 w_1 \ldots w_m) &=\quad w_0 w_1 \ldots w_m \gamma   \text{ for all } \gamma \in  \Gamma, \\
+&\text{\textit{pop}}_1(w_0 w_1 \ldots w_{m-1} w_m ) &=\quad w_0 w_1 \ldots w_{m-1}, \text{ if } m \geq 1\\
+&\text{\textit{top}}( w_0 w_1 \ldots w_m ) &=\quad w_m.
+\end{eqnarray*}
+The operations added at level $n+1$ are the copy of the topmost $n$-stack
+and the removal of the topmost $n$-stack.
+More formally, if $ \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle$ is a stack of 
+level $n > 1$, 
+the following operations are possible.
+%
+\begin{eqnarray*}
+&\text{\textit{push}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle \langle s_m \rangle , \\
+&\text{\textit{push}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle, \langle s_1 \rangle \ldots 
+\langle \text{\textit{push}}_j(s_m) \rangle,
+\text{ if } 2 \leq j < n\\
+&\text{\textit{push}}_1^\gamma( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &= \quad
+\langle s_0 \rangle \langle s_1 \rangle \ldots 
+\langle \text{\textit{push}}_1^\gamma(s_m) \rangle,
+\text{ for all } \gamma \in \Gamma,\\
+&\text{\textit{pop}}_n( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
+\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
+,\\
+&\text{\textit{pop}}_j( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle \langle s_m \rangle) &= \quad
+\langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_{m-1} \rangle
+\langle \text{\textit{pop}}_j(s_m) \rangle,
+\text{ if } 1 \leq j < n\\
+&\text{\textit{top}}( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad
+\text{\textit{top}}(s_m).
+\end{eqnarray*}
+% We omit the stack alphabet $\Gamma$ from 
+%The operation $pop_j$ is undefined on a stack whose top stack of level $j$ is empty.
+The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\
+
+Given a stack alphabet $\Gamma$ and $n \in \N$, 
+we denote by  $\Op_n(\Gamma)$ the {\em base set of $n$-stack operations} as
+
+\begin{itemize}
+\item $\text{\textit{push}}_j$ for all $2 \leq j \leq n$,
+
+\item $\text{\textit{push}}_1^\gamma$ for all $\gamma \in \Gamma$,
+
+\item $\text{\textit{pop}}_j$ for all $1 \leq j \leq n$,
+
+\item $\text{\textit{skip}}$, corresponding to the identity function of $\mathscr{S}(\Gamma)$.
+
+\end{itemize}
+
+% \begin{definition}
+\begin{samepage}
+\par\noindent\ignorespacesafterend
+A {\em higher-order pushdown automata of level $n$} (or $n$-HPDA for short) 
+is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
+where 
+\begin{itemize}
+        \item $Q$ is a non-empty finite {\em set of  states},
+        \item $\Gamma$ is a non-empty finite {\em  stack alphabet},
+	\item $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules},
+        \item $q_{init}$ is the {\em initial  state},
+        \item $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
+initial stack symbol}, and
+	\item $F\subseteq Q$ is a {\em set of final  states}.
+		\end{itemize}
+\end{samepage}
+% \end{definition}
+
+\par\noindent\ignorespacesafterend
+By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
+we denote the set of {\em configurations} of an $n$-HPDA $\H$. 
+%is a pair $(q, s)$ where $q \in Q$ and $s$ is an $n$-stack over $\Gamma$. 
+As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.
+
+An $n$-HPDA $\H=(Q, \Gamma, R, q_{init}, F)$
+induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
+where
+for all $q,q'$ in $Q$,
+for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
+$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
+there exists some rule
+%$(q,\gamma,q',\theta) \in R$
+$(q,\gamma,q',op) \in R$
+such that
+$\text{\textit{top}}(s) = \gamma$
+and
+%$s' = \theta(s)$.
+$s' = op(s)$.
+
+\iffalse
+For our constructions it would be simpler to assume that k-HPDS can work
+also on stacks of lower levels, in particular on 1-stacks. Of course we can always
+simulate a j-stack, for j < k with an k-stack but in the notation it requires some
+additional parenthesis that make it less readable.
+
+To define a game on the graph of a HPDA, we assign a player to each control
+state, and we consider an initial configuration: a game structure on a HPDS H
+is a tuple G = (H, P0 , P1 , s0 ), where P = P0 ⊎ P1 is a partition of the control
+states of H, and s0 ∈ Sk . This extends naturally to a partition of the set of
+configurations: with the notations of Section 2.1, V0 = P0 × Sk , V1 = P1 × Sk ,
+and E is defined above.
+\fi
+
+Again we are interested in parity games.
+As expected, given an 
+ $n$-HPDA $\mathcal{H}$
+ and a partition of
+$Q$ into $Q_0$ and $Q_1$,
+we partition the configurations of
+$T_{\mathcal{H}}$
+into
+$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
+and
+$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
+%
+%
+With these notations in mind one can define the arena
+$$
+A_{(\mathcal{H}, Q_0, Q_1)}
+=
+(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$$
+\par\noindent\ignorespacesafterend
+induced by an $n$-HPDA $\mathcal{H}$ and a partition of its set of states.
+
+As expected, given a 
+priority function $\Omega: Q \to [0, m]$,
+we naturally extend the function
+as follows,
+by setting
+$\Omega_{\mathscr{S}_n(\Gamma)}: \Conf(\mathcal{H}) \to [0, m]$
+and
+$\Omega_{\mathscr{S}_n(\Gamma)}(q, s) = \Omega(q)$
+for all
+$s \in \mathscr{S}_n(\Gamma)$.
+
+Concerning higher-order pushdown automata, we are interested in the following decision problem. 
+% \newpage
+
+\begin{samepage}
+\problemx{$n$-HPDA parity game}
+{An $n$-HPDA $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
+where $Q = Q_0  \biguplus Q_1 $, 
+and
+a priority mapping $\Omega: Q \to \{0, \ldots m\}$.}
+{Does player $0$ have a winning strategy 
+from $q_{init}(\text{\textit{push}}_1^{\gamma_{init}}(\epsilon_{n}))$ 
+for the parity game 
+$\mathcal{G}= (A_{(\H, Q_0 , Q_1)}, \win_{\Omega_{\mathscr{S}_n(\Gamma)}})$ ?\newline}
+\end{samepage}
+
+It was shown in \cite{Cach03} that {\sc $n$-HPDA parity game} can
+be solved in $n$-$\EXP$. This also gives an $n$-$\EXP$ algorithm for the $\mu$-calculus
+model checking over transitions systems induced by  $n$-HPDA.
+In \cite{cachat2007complexity} the matching lower bound was showed, even in the case of reachability games,
+hence showing $n$-$\EXP$-completeness of {\sc $n$-HPDA parity game}.
+
+\begin{theorem}{\cite{ Cach03, cachat2007complexity}}
+{\sc $n$-HPDA parity game} is $n$-$\EXP$-complete.
+\end{theorem}
+
+% \subsection{From pebble pushdown automata to higher-order pushdown automata}
+
+In a $2$-HPDA, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
+allows to ``copy'' the top level $1$ stack. The current word is hereby stored away and left untouched
+until the next operation $\text{\textit{pop}}_2$,
+while $\text{\textit{push}}_1$ and $\text{\textit{pop}}_1$ can be performed on the additional ``copy'' in the meantime. This behavior is similar to that of dropping and lifting a pebble. 
+The main difference is that there is no operation to test that the ``copy'', after many updates, is again equal to the ``original'', i.e.
+there is no operation to syntactically test that the two topmost level $1$ stacks are identical.
+
+% But cannot test
+
+\iffalse
+In an $n$-HPDA, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
+allows to ``copy'' the top stack of level $j$ for all $2 \leq j \leq k$. 
+\fi
+
+In \cite{CaWoe03, Woeh05, carayol2006automates} however 
+%the authors
+Carayol and W\"ohrle
+ introduced a variant of $n$-HPDA by extending the 
+$\text{\textit{pop}}_j$ operations for $2 \leq j \leq n$ with a built-in equality test. 
+For $2 \leq j \leq n$ 
+the new operation
+$\text{\textit{pop}}_j^=$ has the same effect as $\text{\textit{pop}}_j$, but
+can only be applied if the two top level $j$ stacks coincide.
+In \cite{carayol2006automates} it is seen as a symmetrical operation in comparison to
+$\text{\textit{push}}_j$.\\
+
+A {\em higher-order pushdown automaton with equality pop of level $n$} ($n$-HPDA$^=$ for short) 
+is a higher-order pushdown automaton of level $n$ where in
+% Instrn
+$\Op_n(\Gamma)$
+ the operation 
+ %popk
+ $\text{\textit{pop}}_j$  
+  is replaced by 
+%  pop=
+%k 
+$\text{\textit{push}}_j^=$
+%for 1 ≤ k ≤ n. 
+for $2 \leq j \leq n$.
+We denote
+this new set of operations by 
+% Instr= n
+$\Op_n^=(\Gamma)$. 
+More formally the new operations $\text{\textit{push}}_j^=$ are defined as
+\begin{eqnarray*}
+&\text{\textit{pop}}_k^=( \langle s_0 \rangle \langle s_1 \rangle \ldots  \langle s_m \rangle \langle s_m \rangle) &=\quad  \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle , \quad \text{ and }\\
+&\text{\textit{pop}}_j^=( \langle s_0 \rangle \langle s_1 \rangle \ldots \langle s_m \rangle) &=\quad \langle s_0 \rangle \langle s_1 \rangle \ldots 
+\langle \text{\textit{pop}}_j^=(s_m) \rangle,
+\text{ if } 2 \leq j < k. 
+\end{eqnarray*}
+%
+%
+%
+%
+\iffalse
+Dans ce sous-paragraphe, nous présentons formellement les automates à pile
+de niveau k sur le jeu d'opérations symétriques Opsk et sur le jeu d'opérations
+classiques COpsk . Nous donnons à cette occasion un bref historique de la notion
+d'automate à pile de piles. Enfin, nous établissons l'équivalence entre les deux
+modèles d'automates à chaque niveau en tant qu'accepteurs de langages. Ce résultat a été obtenu dans [CW03] en collaboration ave Stefan Wöhrle. La preuve
+complète apparait dans [Wöh05]. Indépendamment, ce résultat a été obtenu dans
+[Fra05].
+\fi
+%
+Transition systems induced by higher-order pushdown automata with equality pop of level $n$
+are then defined as expected.  
+Carayol and W\"ohrle proved \cite{Woeh05, carayol2006automates} that the two models,
+namely
+higher-order pushdown automata with equality pop of level $n$
+and
+higher-order pushdown automata of level $n$,
+generate the same classes of transition systems.
+
+\begin{theorem}{\cite{Woeh05, carayol2006automates}}
+If $\H$ is an $n$-HPDA (resp. $n$-HPDA$^=$) then there
+exists an $n$-HPDA$^=$ (resp. $n$-HPDA) $\H'$ such that
+$\H$ and $\H'$ %generate the same graph.
+induce isomorphic transition systems.
+\end{theorem}
+
+\iffalse
+\begin{theorem}{\cite{Woeh05, carayol2006automates}}
+If $\H$ is an $n$-HPDA  then there
+exists an $n$-HPDA$^=$  $\H'$ such that
+$\H$ and $\H'$ %generate the same graph.
+induce isomorphic transition systems.
+\end{theorem}
+% Proposition 3.12
+
+\begin{theorem}{\cite{Woeh05, carayol2006automates}}
+If $\H$ is an $n$-HPDA$^=$ then there
+exists an $n$-HPDA $\H'$ such that
+$\H$ and $\H'$ %generate the same graph.
+induce isomorphic transition systems.
+\end{theorem}
+
+% Proposition 3.18 If A is a higher-order pushdown system with equality pop
+% of level n then there exists a higher-order pushdown system B of level n such
+% that A and B generate the same graph.
+\fi
+
+From $n$-HPDA to $n$-HPDA$^=$ (Proposition 3.12 in \cite{Woeh05}) the proof relies on recreating a correct stack content to simulate higher-order 
+$\text{\textit{pop}}$ operations by $\text{\textit{pop}}^=$, essentially ``guessing'' the correct stack content to be able to apply $\text{\textit{pop}}^=$. 
+In the other direction (Proposition 3.18 in \cite{Woeh05}), 
+the author
+enrich the stack alphabet with new symbols stating which 
+instructions have to be executed to recreate a
+previous stack content. It is proven that such en encoding is possible since
+there is 
+for every stack $s$ of level $n$
+% there is
+ a unique shortest sequence of instructions which creates $s$ from
+$\epsilon_n$. \\
+
+Defined like {\sc $n$-HPDA parity game}, we write {\sc $n$-HPDA$^=$ parity game}  
+in case the input is an $n$-HPDA$^=$ rather than an $n$-HPDA.
+
+The algorithm from \cite{Cach03} actually provides an algorithmic solution to parity games on the graphs of the Caucal hierarchy. 
+We skip a formal definition of a graph of level $n$ in the Caucal hierarchy, and refer the reader
+to \cite{Cau02mfcs} for more details.
+The $n$-$\EXP$ upper bound on $n$-HPDA parity games follows from the fact that every transition system induced by an $n$-HPDA $\H$ is a graph of the Caucal hierarchy \cite{Cach03, Woeh05},
+whose vertices are almost in one-to-one correspondence with the configurations of $\H$.
+In \cite{Woeh05} the converse direction is proven, i.e. that every graph of level $n$
+of the Caucal hierarchy is generated by an $n$-HPDA.  
+%
+%
+In \cite{carayol2006automates}
+it is similarly shown that
+every transition system induced by an $n$-HPDA$^=$ $\H$ is a graph of the Caucal hierarchy.
+%
+%
+The algorithm from \cite{Cach03} hence lead to a solution for
+$n$-HPDA$^=$ parity games as well.
+
+\iffalse
+\textcolor{red}{
+Now then, because the class of $n$-HPDA$^=$ transitions systems
+is the same as
+the class of $n$-HPDA transitions systems, we have the following result.
+}
+\fi
+
+\begin{theorem}\label{HPDA parity game}
+{\sc $n$-HPDA$^=$ parity game} is in $n$-$\EXP$.
+\end{theorem}
+
+\iffalse
+"To show that such an automaton exists we introduce the notion of a weak
+popping higher-order pushdown automaton. A weak popping automaton is only
+allowed to execute a pop instruction of level $j \geq 2$ if the two top level $j$ stacks
+coincide. We skip a formal definition of a weak popping higher-order pushdown
+automaton and just mention that even though this automaton model is equipped
+with a built-in test on the equality of two stacks of the same level, it is equivalent
+to the usual model. All proofs are given in the full version of this article [4]."
+
+Pour pouvoir obtenir une notion pertinente d'ensemble rationnel de piles de
+piles, il faut considérer un jeu d'opérations où la destruction inconditionnelle des
+piles est remplacée par une version plus symétrique où la dernière pile de niveau ℓ
+ne peut être détruite que si elle est égale à la précédente. Nous établissons dans le
+paragraphe 4.1 l'équivalence (en temps qu'accepteurs de langages) les automates
+à pile définis sur ces deux jeux d'opérations ( f. théorème 4.1.23). Nous montrons,
+dans le paragraphe 4.6, le manque de pertinence de la notion de rationalité induite
+par le jeu d'opérations classiques et par là même, nous établissons la nécessité de
+considérer le jeu d'opérations symétriques.
+
+L'opération de destruction symétrique a été introduite pour des raisons techniques dans 
+[CW03] et a été utilisée de manière 
+systématique dans [Car05] et
+[Fra05]. 
+Nous montrerons, dans le sous-paragraphe 4.1.5, que les automates à pile
+de piles définis en utilisant les opérations destrk ou les opérations copyk acceptent
+les mêmes langages.
+
+Les automates à pile d'ordre supérieur ont été introduits dans les années 70.
+Dans [Gre70℄, Greibach attribue l'idée de ces automates à Aho et Ullman. Ces
+automates sont aussi définis dans [Mas76℄. Les automates sur COpsk diffèrent
+légèrement des automates considérés par ces auteurs et de ceux onsidérés dans
+[DG86, Eng91℄. La différence majeure réside dans la définition des piles de piles.
+Les piles de niveau k sont des suites non vides de couples formés d'un symbole de
+pile et d'une pile de niveau k − 1. Comme remarqué dans [KNU02℄, ces symboles
+de pile supplémentaires peuvent être simulés dans le modèle des automates sur
+COpsk .
+
+Les langages a eptées par les automates à pile sur COpsk sont onnus sous
+le nom de langages k -OI ou de langages indexés de niveau k . La première déno-
+mination vient des travaux de Damm [Dam82]
+
+Nous allons maintenant établir que les encodages des opérations simulent bien
+les opérations sur les encodages des piles.
+
+Lemme 4.1.21. Pour toute pile s ∈ Stacksk (Γ) et pour tout θ ∈ Opsk ,
+[[ θ ]]k ([[ s ]]) =
+{
+[[ θ(s) ]]		si θ(s) est défini,	
+non définie		sinon.
+
+Proposition 4.1.22. Tout langage accepté par un automate à pile sur Opsk est
+accepté par un automate à pile sur COpsk .
+
+Démonstration. Soit A = (Γ,Σ,τ,Q,I,F,∆) un automate à pile sur Ops⋆k . Nous
+définissons un automate B = (Γ,Σ,τ,Q,I,F,∆B ) où l'ensemble des transitions ∆B
+est défini par:
+∆B = {(p,x,η,q) | (p,x,θ,q) ∈ ∆ et η ∈ [[ θ ]]k }.
+Par le lemme 4.1.21, il suit que pour tout w ∈ (Σ)∗ , p ∈ Q et s ∈ Stacksk (Γ):
+w
+(q0 , [ ]k ) −→ (p,s)
+A
+⇔
+w
+(q0 , [ ]k ) −→ (p,[[ s ]]).
+
+Théorème 4.1.23. Les langages acceptés par les automates à pile sur Opsk et
+sur COpsk coïncident.
+
+\fi
+
+\subsection{From pebble pushdown 
+automata 
+to higher-order pushdown 
+automata
+}
+
+We use HPDA$^=$ parity games to simulate pebble pushdown automata
+parity games.
+A similar approach was used in \cite{carayol2006automates} to show that $(n + 2)$-level stack
+automata could be used to simulate $n$-pebble alternating two-way word automata.
+
+Let us start by discussing the simulation of the dropping and lifting of a pebble in the case of
+a pebble pushdown automaton $\ppda$ with only one pebble.
+As long as no pebble is dropped, a $2$-HPDA$^=$ $\H$ can simulate 
+the behavior of a pebble pushdown automaton $\ppda$ with a $2$-stack containing a single $1$-stack
+on which it performs the same operations $\ppda$ performs on its stack.
+When the automaton $\ppda$ is in configuration $q(w)$ and drops a pebble, instead of making a pop or a push, we need
+to %go to state $q'$ and word $w$ and 
+	store the information that the pebble has been
+dropped on $w$.
+To simulate the next configuration in such a case, we use 
+$\text{\textit{ push}}_2$
+to store the information that the pebble has been dropped on the word $w$, leading to the new  $2$-stack
+$\langle w \rangle \langle w \rangle$. 
+Configurations afterwards have $2$-stacks of length two where the first component remains
+$w$ until the pebble is lifted.
+Lifting the pebble can be done only at the position the
+pebble was dropped, by using $\text{\textit{pop}}^=_2$.
+Checking that the node
+corresponds to the one where the pebble has been dropped simply makes use of
+the composition of $\text{\textit{pop}}^=_2$ and $\text{\textit{push}}_2$.
+Checking the absence of the pebble can be done by challenging
+the opponent to prove the presence of the pebble.
+
+Now, let us detail the case when there is a second pebble, the construction
+for each additional pebble being highly similar, where every additional pebble after
+the first one would require the addition of another stack level.
+Similarly as before, 
+as long as no pebble is dropped, the stack of $\H$ 
+is a $3$-stack containing a
+single $2$-stack that contains a single $1$-stack.
+To simulate configurations in the case one pebble have been dropped, the stack contains a single
+ $2$-stack of the form $\langle w_1 \rangle \langle w \rangle$.
+In the case the two pebbles have been dropped, 
+the stack is of the form $ \langle \langle w_1 \rangle \langle w_2 \rangle \rangle
+\langle \langle w_1 \rangle \langle w \rangle \rangle$.
+
+Intuitively speaking, $w_1$ is, as before, the position where the first pebble is
+placed, and $w_2$ is where the second pebble is placed. Dropping pebbles like this
+again involves the cloning operation, only, each pebble operates at a different 
+stack level: thus, dropping the first pebble will use $\text{\textit{push}}_2$
+and dropping the second pebble will use $\text{\textit{push}}_3$. Knowing the number of pebbles dropped can be done by 
+keeping track using the states of $\H$.
+Lifting or
+checking for the presence 
+(or absence) 
+of the pebbles functions on a similar
+basis as before. Checking the presence of the first pebble uses 
+operation $\text{\textit{pop}}^=_2$ followed by $\text{\textit{push}}_2$,
+while checking the presence of the second pebble uses 
+operation $\text{\textit{pop}}^=_3$ followed by $\text{\textit{push}}_3$.
+Lifting the first pebble uses $\text{\textit{pop}}^=_2$, but
+check first that the second pebble has already been lifted.
+Lifting the second pebble simply uses $\text{\textit{pop}}^=_3$.
+
+By expanding this reasoning inductively, we conclude that given $n \in N$, and
+given an $n$-pebble pushdown automaton 
+$\ppda = (Q, \Gamma, R, q_{init}, \gamma_{init}, F)$, where $Q= Q_0 \biguplus Q_1$,
+one can compute an $(n+1)$-HPDA 
+$\H = (Q', \Gamma, R', q_{init}', \gamma_{init}', F')$  where $Q'= Q_0' \biguplus Q_1'$, 
+such 
+that 
+player $0$ has a winning strategy from $q'_{init}(\text{\textit{push}}_1^{\gamma_{init}'}(\epsilon_{n}))$ in $A_{(\H,Q'_0,Q'_1)}$ 
+if
+and only if
+player $0$ has a winning strategy from $q(\gamma_{init},\mu_{init})$ in $A_{(\ppda,Q_0,Q_1)}$. 
+Hence the following reduction.
+
+\begin{theorem} 
+{\sc $n$-pebble pushdown parity game} is polynomial time reducible
+to {\sc $(n+1)$-HPDA$^=$ parity game}.
+\end{theorem}
+
+The reduction implies decidability of pebble pushdown automata parity game. Furthermore,
+by Theorem~\ref{HPDA parity game}, it implies the following complexity result.
+
+\begin{theorem}
+{\sc $n$-pebble pushdown automata parity game} is in $(n+1)$-$\EXP$.
+\end{theorem}
+
+Finally, the following complexity result is due the above and Theorem~\ref{PPDA reduction}.
+
+\begin{corollary}
+{\sc $n$-parametric pushdown automata parity game} is in $(n+1)$-$\EXP$.
+\end{corollary}
+
 \end{document}

### 2023-06-11 reachability/article.tex
up

@@
 We consider here a different approach to extending pushdown systems, by allowing them to test the stack content against precise values.
-Pushdown systems which can test the stack content against regular-expressions are
-equivalent to pushdown systems,
-so pushdown systems which can test the stack content against specific words are too.
-We consider instead checks against unspecified word values that can be instantiated one after the other,
+% Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
+We consider 
+%instead
+here checks against unspecified word values that can be instantiated one after the other,
 providing the pushdown systems the ability to
@@
 $\text{\textit{Inf}}(\pi) = \{ v \in V \mid \forall i, \exists j > i, ~ v_j = v \}$.
+
+		\iffalse
 
@@
 
-\subsection{Parity Games}
-
-We first recall the definition of  two  player parity  games.
-Two players, player $0$ and player $1$, move along an arena $\mathcal{G}=(V_0,V_1,E,\Omega)$ which is a labelled graph, composed of
-two  disjoint  sets of vertices, $V_0$ and $V_1$ respectively associated with player $0$ and player $1$, 
-a set of edges $E \subseteq V \times V$, where $V=V_0 \uplus V_1$,
-% We require that the out-degree of each vertex is at least one. This ensures that every finite path in (V, E) can be prolonged.
-and a mapping $\Omega:V \to \{0,1, \ldots, m\}$,
-$m < \omega$, which
-assigns a priority to each vertex. %Then the {\em initialized game} $(\mathcal{G}, v_I)$ is given with an initial vertex $v_I \in V$.
-
-A {\em partial play} of $\mathcal{G}$ is a sequence of vertices $v_0 v_1 \ldots \in V^\infty$ connected by the transition relation, i.e. such that $(v_i, v_{i+1}) \in E$.
-
-A {\em play} of $\mathcal{G}$ is a partial play that is {\em maximal} in the following sense: if it is finite, then the last vertex $v_k$ is such that for all $v' \in V$, $(v_k, v') \notin E$.
-Else, the play is an infinite sequence $ \pi  = v_0 v_1 \ldots \in  V^{\omega} $.
-
-We consider min-parity games: in order to find out who wins a play, we need to take a
-look at the positions which are reached infinitely often and compute their priorities using 
-$\Omega$.
-Player $0$ wins the play $\pi$  iff  $\min( \{ \Omega(v) \mid v \in \text{\textit{Inf}}( \pi) \})$ is even, else, player $1$ wins.
-The winner of a finite play is the player whose opponent is unable to move, i.e.
-the winner of a play such that the last vertex of the play is in $V_{0}$ (resp.
-$V_1$) is player $1$ (resp.
-player $0$).
-
-A {\em strategy} $\sigma_i : V^* V_i \to V$ for player $i \in \{ 0,1\}$ is a function that, given a sequence of vertices $ w v$  with $w\in V^*$ and $v \in V_i$,
-provides, if possible, a successor $v'$ such that $(v,v') \in E$.
-A strategy is called {\em memoryless} if its output only depends on the final vertex of the sequence $v \in V_i$, i.e. if there exists a function $f_i : V_i \to V$ such
-that for all $w\in V^*$, $v \in V_i$, we have $\sigma_i(wv) = f_i(v)$.
-
-A partial play $\pi$ is {\em consistent} with a strategy $\sigma_i$ if sequences of vertices that end in $V_i$ along this play all have successors according to the strategy.
-Given a starting vertex $v_I$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play
-starting with $v_I$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $v_I$. It is 
-denoted by $\pi(v_I, \sigma_0, \sigma_1)$.
-
-A strategy $\sigma_i$ for player $i$ from position $v_I \in V$ is a {\em winning strategy from $v_I$} if 
-no matter which
-strategy $\sigma_{1-i}$ the other player pursues,
-% the resulting play  starting from $v_I$  
-%  results in 
-% player $i$ winning.
-player $i$ wins the play $\pi(v_I, \sigma_0, \sigma_1)$.
-
-The above definitions are essentially the same for finite and infinite arenas.
-%
-%
-In this paper we are going to concern ourselves with parity games over potentially infinitely large arenas, \textcolor{black}{thus we define the following decision problem in a general abstract manner.
-In particular, we do not discuss representation of the input}.
+		\fi
+
+\subsection{Games}
+
+% Games represent a model for interaction between decision makers.
+We study so-called turn-based games in extensive form with perfect information where the players have to make decisions again and again.
+\iffalse
+After its foundation by von Neumann and Morgenstern [vNM44] in the
+1940s, game theory has quickly become a valuable tool for modelling the
+interaction of several agents with individual and possibly conflicting objec-
+tives. It has applications in many scientific fields such as economy, sociology,
+biology, logic and computer science. Whereas the games used in the former
+three areas are usually finite in the sense that any play of these games ends
+after a finite number of steps (in fact often after only a single step), the
+games used in logic and computer science are in general infinite.
+
+@article{copeland1945john,
+  title={John von Neumann and Oskar Morgenstern, theory of games and economic behavior},
+  author={Copeland, Arthur H},
+  journal={Bulletin of the American Mathematical Society},
+  volume={51},
+  number={7},
+  pages={498--504},
+  year={1945},
+  publisher={American Mathematical Society}
+}
+
+More precisely, an infinite game is played for an infinite number of
+rounds. In each round, one player chooses an action depending on the se-
+quence of actions chosen in previous rounds. After infinitely many rounds,
+an infinite sequence of actions has emerged, called the outcome of the game.
+Each player receives a payoff determined by a payoff function mapping each
+possible outcome to a real number in the interval [0, 1]. We will concen-
+trate on games where the possible payoffs are just 0 and 1, in which case a
+winning condition, given as an abstract set of possible outcomes, is used to
+determine whether a player wins (payoff 0) or loses (payoff 1) an outcome.
+In logic, games are used to evaluate a logical formula in a mathematical
+structure and to compare mathematical structures with respect to their
+(in)distinguishability by a certain logic. Games of the former kind are called
+model checking games; games of the latter kind are called model comparison
+games. Whereas model checking and model comparison games for plain
+modal or first-order logic are finite games, infinite games arise as model
+checking games for fixed point logics and as model comparison games for
+infinitary logics.
+In computer science, games are used for the verification and synthesis
+of reactive systems, i.e. systems that interact with their environment. As
+soon as the systems under consideration do not necessarily terminate, these
+games become infinite. For example, the behaviour of an operating system
+can be understood as an infinite game between the system and its users.
+
+So far, the games used in logic and computer science have usually been
+two-player zero-sum games, i.e. games with only two players where one
+player wins if and only if the other player loses. For example, in a model
+checking game, one player wants to show that the given formula holds in the
+given structure whereas the other player wants to show that this is not the
+case. Analogously, in a game used for verification, one player wants to show
+that the given system is able to react against its environment in such a way
+that the resulting execution path of the system fulfils the given specification
+whereas the other player wants to show that this is not the case.
+
+The research on infinite games in theoretical computer science is based on a
+mixture of several motivations:
+1. The beautiful classical theory of infinite games, as developed in descriptive
+set theory, lacks an algorithmic content. Such an algorithmic orientation is
+provided by the automata theoretic approach to infinite games, originating
+in work of Church, B ̈uchi, McNaughton, and Rabin about fourty years ago.
+2. Determinacy results for infinite games are closely related to complementation
+results for logics and automata; thus, infinite games help to analyze logical
+theories (the most prominent being the monadic theory S2S of two successor
+functions, see e.g. [Th97]).
+3. Games are a natural model of reactive computation, and infinite games are
+thus a faithful representation of nonterminating reactive systems (for which
+control problems can be solved in terms of providing winning strategies).
+4. The model checking problem for logics like the μ-calculus can be formulated
+as the question to determine the winner of an infinite game.
+The purpose of this tutorial is to explain the core of the algorithmic (and
+automata theoretic) theory of infinite games. In the present extended abstract,
+only some key notions are explained, without technical details and proofs. In the
+tutorial, more will be said about topics of current research as listed in Section 6
+of this abstract.
+
+@incollection{thomas1997languages,
+  title={Languages, automata, and logic},
+  author={Thomas, Wolfgang},
+  booktitle={Handbook of formal languages},
+  pages={389--455},
+  year={1997},
+  publisher={Springer}
+}
+
+\fi
+
+A game is composed of an arena and a winning condition.
+\iffalse
+\mh{Two-player games oﬀer a very convenient framework to represent
+interaction of a program with some (possibly hostile) environment. In this approach, the ﬁrst player represents the program while the second player simulates
+the environment. A winning strategy expresses a property that must hold whatever the environment does, thus ﬁnding a winning strategy for the ﬁrst player
+allows to synthesize a controller that restricts the program and ensures that the
+property expressed by the winning condition always holds \cite{arnold2003games}.}
+\fi
+We will ﬁrst study
+arenas and
+ then introduce common winning conditions that will be of interest to us.
+
+ \subsubsection{Arenas}
 
 \begin{samepage}
-\problemx{Solving parity game}
-{An arena $\mathcal{G}=(V_0,V_1,E,\Omega)$, an initial vertex $v_I \in V$.}
-{Does player $0$ have a winning strategy from $v_I$ in the parity game $\mathcal{G}$?\newline}
+An {\em arena} is a tuple $A=(S_0, S_1,{\rightarrow})$  which is composed of
+\begin{itemize}
+	\item two  disjoint  sets of configurations, $S_0$ and $S_1$, %respectively associated with player $0$ and $1$, 
+	whose disjoint union $S_0 \cup S_1$ we denote by $S$, and
+	\item a relation of %labeled 
+		transitions
+	 ${\rightarrow} \subseteq S \times %\Lambda \times 
+	 				S$. \\
+\end{itemize}
 \end{samepage}
 
-One important property of parity games is that of {\em memoryless determinacy}: for a parity game and an initial vertex $v_I \in V$, one of the players has a memoryless winning strategy from $v_I$ \cite{zielonka1998infinite}.
+\noindent
+Note that $(S,{\rightarrow})$, where $S=S_0 \uplus S_1$, forms
+a %labeled 
+transition system, and, in particular, given a 
+%labeled 
+transition system,
+one needs only to provide 
+a partition of the set of configurations $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\
+% We require that the out-degree of each configurations is at least one. This ensures that every finite path in (V, E) can be prolonged.
+
+% \mh{I dont think I want to disallow dead-ends from the very beginning, means every time you define some arena you have to prove it has no dead-ends, seem tedious and unecessary.}
+
+\noindent
+The games we are interested in are played by two players, called {\em player $0$} and {\em player $1$}. We will often write player $i$ to denote a general player for $i \in \{0,1 \}$,
+and we will call player $1-i$ its {\em opponent}.
+
+% Observe that there is no restriction on the number of the successors of a configuration in an arena.
+Also, we don’t require that (V, E) is a bipartite graph with corresponding partition {V 0 , V 1 }.
+
+\iffalse
+\noindent
+In the case where $\Lambda$ is a singleton $\{ a\}$, we will abuse notation
+and denote the arena by
+$(S_0, S_1,{\rightarrow},\Omega)$, and we will see $(S,{\rightarrow})$
+as an unlabeled transition system.
+\fi
+
+ \subsubsection{Plays}
+
+\noindent
+%
+%\noindent
+A {\em play} in an arena $A=(S_0,S_1,\rightarrow)$ is a path in $(S,{\rightarrow})$ that is {\em maximal} in the following sense: it is either infinite or finite and if it is finite, then the target configuration is a dead end.
+A {\em partial play} in an arena $A=(S_0,S_1,\rightarrow)$ is a 
+% sequence of configurations $s_{init} = s_0 s_1 \ldots \in S^\infty$ connected by the transition relation, i.e. such that $(v_i, v_{i+1}) \in T$.
+prefix of a play in $A$.\\
+%finite path in $(S,{\rightarrow})$.\\
+
+\noindent
+Essentially plays can be seen as this: the two players, player $0$ and player $1$, move along the labeled transition system, taking turns either inﬁnitely often or until a dead end is reached.
+
+% \subsubsection{Strategies}
+\subsection{Strategies}
+
+% Before deﬁning formally what it means for a player to win a game, we need to introduce the notion of strategy.
+
+A {\em strategy} for player $i \in \{ 0,1\}$ is a function $\sigma_i : S^* S_i \to S$ such
+that $\sigma_i(ws)$ is a successor of $s$.
+%
+A strategy is called {\em memoryless} if its output only depends on the final configuration of the sequence $s \in S_i$, i.e.
+if $\sigma_i(ws) = \sigma_i(w's)$ for all $w,w' \in S^*$.
+Thus, a memoryless strategy for player $i$ can be written  as a function $\sigma_i: S_i \to S$.
+
+A partial play $\pi = s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_n$ is {\em consistent} with a strategy $\sigma_i$ if sequences of configurations that end in $S_i$ along this play have all successors according to the strategy, i.e.
+if for all $j\in[0,n-1]$, for all element $v_j \in V_i$ in the sequence $\pi$,
+$v_{j+1} = \sigma_i(v_j)$.
+ Given a configuration $s$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play starting with $s$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $s$.
+It is
+denoted by $\pi(s, \sigma_0, \sigma_1)$.\\
+
+%\noindent
+%Note that the above definitions are essentially the same for finite and infinite arenas.
+
+ \subsubsection{Winning conditions}
+
+Once we have provided players with an arena, it remains to define  properly what the winning condition is in order to define a game.
+Given an arena $A = (S_0, S_1, \rightarrow)$, a {\em winning condition}
+%$\win \subseteq S^\infty$
+$\win$
+is a subset  
+of the set of maximal plays for $A$. A {\em game} is a pair $\mathcal{G}= (A, \win)$ which is composed of an arena $A$ and a winning condition $\win$.
+Although a winning condition is in general an infinite object we are interested in winning conditions 
+that can finitely be described.
+
+\iffalse
+We consider here several types of winning conditions. 
+Essentially the 
+ mapping $\Omega$ of an arena serves to colour the configurations of the labeled transition system.  Winning conditions then
+ consist in acceptance conditions on colour sequences.
+ \fi
+
+A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
+ from $s$} for player $0$ if 
+% no matter which
+% strategy $\sigma_{1}$ the other player pursues,
+% the play $\pi(s, \sigma_0, \sigma_1)$ is in $\win$.
+every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
+On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
+ from $s$} for player $1$ if 
+% no matter which
+% strategy $\sigma_{0}$ the other player pursues,
+% the play $\pi(s, \sigma_0, \sigma_1)$ is not in $\win$.
+every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.
+
+\iffalse
+This definition of winning strategies is independent from the winning condition itself, and will also apply to %B\"uchi games and min-
+parity games. To differentiate, we mention the winning conditions of a winning strategy from $s_{init}$.
+\fi
+
+% \paragraph{Reachability Games}
+\paragraph{Reachability Games}
+
+We first consider reachability games: 
+the winning condition $\win_F(S)$ is given by a set of final configurations
+$F \subseteq S$, and $\win_F(S)$ is the set of maximal plays in $S^* F S^\infty$. 
+We simply write
+$\win_F$ in case the set $S$ is obvious from context.
+To indicate that the winning condition of a game is a reachability winning condition, we will speak of {\em reachability games}.
+
+With regards to reachability games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline
+
+\begin{samepage}
+\problemx{Reachability game}
+{A reachability game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_F)$, an initial configuration $s \in S_0 \cup S_1$.}
+{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
+\end{samepage}
+
+% \paragraph{Parity Games}
+\paragraph{Parity Games}
+
+Secondly, we consider min-parity games: 
+the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
+ $\Omega : S \to  [0, m]$.
+An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
+%this path
+the sequence 
+ is even, i.e.
+ %$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
+if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even. 
+% The winner of a finite play is the player whose opponent is unable to move, i.e. a play such that the last configuration of the play is in $S_{0}$ (resp. $S_1$) is player $1$ (resp. player $0$). 
+A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
+We simply write
+$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
+To indicate that the winning condition of a game is a parity winning condition, we will speak of
+ {\em parity games}.
+
+With regards to parity games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline
+
+\begin{samepage}
+\problemx{Parity game}
+{A parity game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_\Omega)$, an initial configuration $s \in S_0 \cup S_1$.}
+{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
+\end{samepage}
+
+%
+\iffalse
+Solving a parity game played on a finite graph means deciding, for a given starting position $v_{init}$, which of the two players has a winning strategy.
+It has been shown that this problem is in $\NP$ and
+co-$\NP$, more precisely $\UP$ and co-$\UP$. 
+% Marcin Jurdziński (1998), "Deciding the winner in parity games is in UP∩ co-UP" (PDF), Information Processing Letters, Elsevier, 68 (3): 119–124, doi:10.1016/S0020-0190(98)00150-1
+% @article{jurdzinski1998deciding,
+%  title={Deciding the winner in parity games is in UP∩ co-UP},
+%  author={Jurdzi{\'n}ski, Marcin},
+%  journal={Information Processing Letters},
+%  volume={68},
+%  number={3},
+%  pages={119--124},
+%  year={1998},
+%  publisher={Elsevier}
+%}
+\fi
+%
+
+% Muller condition: Inf() \in F
+
+% Rabin condition :
+
+% Streett Condition
+
+% Rabin Chain condition
+
+% max parity condition
+
+% 1-winning  == reachability
+
+\begin{center}
+	\begin{figure}
+			\hspace{3.81cm}
+\includegraphics[width=0.4\textwidth]{figures/parity_game}
+	\caption{An example of an arena.
+Circles  belong to player $0$, whereas diamonds belong to player $1$.
+Arrows represent transitions from a configuration to the next.
+The priority assigned by the priority mapping to each configuration is a number in $\{0,1,2,3\}$, next to the configuration it is assigned to.
+Note that we do not have a dead end in our example.}
+	\label{example_parity_game}
+	\end{figure}
+\end{center}
+
+\begin{example}
+Let us consider the arena presented in  Figure~\ref{example_parity_game}. 
+Configurations are partitioned into two sets: the circles, belonging to player $0$, and the diamonds,
+belonging to player $1$. 
+The priorities are $\{ 0,1,2,3\}$.
+% The transitions and priority mappings may be derived from the picture.
+% Note that we don't have a dead end in our example. 
+Consider the game on the arena, where the winning condition is a parity condition.
+Plays $\pi_1 = q_0 q_1 q_2 (q_4 q_7)^\omega$ and 
+$\pi_2 = (q_0 q_1 q_2 q_4)^\omega$ are both winning for player $0$. Indeed, in $\pi_1$, 
+the minimal priority encountered infinitely often is %that of $q_7$, that is, 
+				$2$, 
+whereas in $\pi_2$, it is %that of $q_2$, that is, 
+			$0$. 
+In both cases, it is an even number. On the other hand,
+the plays $\pi_3 = (q_5 q_3 q_6)^\omega$ and $\pi_4=(q_5 q_6)^\omega$ are both winning for player $1$: the minimal priority encountered is odd, since it is $1$ for both.
+
+One can take matters further: on this arena, player $0$ has a winning strategy for the parity game from $q_0$, $q_1$, $q_2$, $q_4$ and $q_7$, but not from $q_5$, $q_3$ nor $q_6$, where player $1$ has a winning strategy.
+\end{example}
+
+% \mh{Say something abut positional determinacy}
+One important property of parity games is that of {\em memoryless determinacy}: for a parity game
+$\mathcal{G}= ((S_0,S_1,\rightarrow),\win_\Omega) $ and an initial configuration $s \in S_0 \cup S_1$, one of the players has a memoryless winning strategy from $s$ \cite{zielonka1998infinite}.
+
+%\mh{
+Of note is that several model checking problems %for pushdown automata 
+can be expressed as decision problems for games: the most famous example is that the modal $\mu$-calculus model checking problem is polynomially equivalent to solving the parity game problem. 
+%}
 

### 2023-06-11 reachability/article.tex
up

@@
 % Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
-We consider 
-%instead
-here checks against unspecified word values that can be instantiated one after the other,
-providing the pushdown systems the ability to
-store information on some stack contents for later comparisons.
+Perhaps a similar model consists in 
+Pushdown automata 
+%with regularity condition on the stack
+with transitions that are conditioned by regular conditions on their stack content, which
+can in particular
+test the
+stack content against
+specific words. 
+They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages),
+and have
+been used to establish that CTL$^*$ model checking remains decidable
+when the formulas are allowed to include regular predicates on the stack content,
+and to obtain 
+model checking algorithms for LTL and CTL$^*$ model checking
+for
+pushdown automata \cite{finkel1997direct}.
+Pushdown automata 
+%with regularity condition on the stack
+with transitions that are conditioned by regular conditions on their stack content
+can be viewed as
+a special case of stack automata as seen in \cite{hopcroft1969formal}.
+Instead
+of checking the stack content against regular conditions, or, in a more limited manner,
+against specified word values,
+we consider
+checks against unspecified word values that can be instantiated using parameters.
+This allows for instance to ensure equality between two stack contents appearing in two distincts configurations in a run.
+
+Extensions of pushdown automata with storage for later comparisons
+have been introduced for instance as register pushdown automata. 
+Such automata possess registers in addition to their stack, and can keep data values in both.
+Register pushdown automata have been shown to have applications for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}. 
+In theory, since registers can be unfilled and later be filled again an unbounded number of times, the number of different data values a pushdown register automata can store in its registers in a run is unbounded.
+In \cite{murawski2017reachability} it was shown however that a register pushdown automaton can
+only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers,
+% Even though pushdown register automata have no bound on the number of elements of the alphabet that they can store at any instant, over the course of a run they can really “remember” at most 3r of them,
+i.e. for any run of a register pushdown automaton with $r$ registers there exists an equivalent run %, i.e.
+ with the same initial and final conﬁgurations, but in which every configuration contains register assignments drawn from only $3r$ elements. This 
+ %can 
+ leads to
+ the
+  question
+  of
+   how useful the ability to unfill and later refill registers really is compared to a model that would only specify valuations once.
 
@@
 pebbles.
+
+ \iffalse
+
+We provide formal definitions of parametric pushdown automata, parametric pushdown reachability games
+and parametric pushdown parity games in Section~\ref{section def ppda}. We give an overview of our contribution in Section~\ref{contribution ppda}. The contribution consists in
+proving parametric pushdown reachability games
+and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
+$n$ is fixed,
+and providing a nonelementary lower bound for parametric pushdown reachability games in general.
+We provide the proof of our result, which stretches along 
+Section~\ref{lower bound PDA},
+Section~\ref{pebbles},
+% Section~\ref{ATWA and PATWA}, and
+Section~\ref{section HPDA}.
+In Section~\ref{discuss ppda} we close the chapter with a discussion about the methods used, with some directions for future work.
+
+\fi
 

### 2023-06-11 reachability/article.tex
up

@@
 
-Our main result states that
-deciding
-the winner of a parity game on a graph of a pebble pushdown system
-is decidable but nonelementary.
-The contribution is two-fold.
+Our main result consists in
+proving parametric pushdown reachability games
+and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
+$n$ is fixed,
+and providing a nonelementary lower bound for parametric pushdown reachability games in general.
 
@@
 from \cite{Sto74}, to the
-problem of deciding whether a player has a winning strategy.
-
-	For the main contribution, we use pebble alternating two-way tree automata to solve pebble pushdown parity games.
-Alternating two-way tree automata with parity winning conditions have been used in the past to solve parity games on pushdown graphs \cite{cachat2002two}.
-We show that pebble alternating two-way tree automata can similarily be used to solve parity games on pebble pushdown graphs.
-We then prove  decidability 
-by 
-a reduction
- to the problem of model-checking $\MSO$ on a well chosen structure, whose decidability is
- guaranteed by
-Muchnik's Theorem~\cite{Wal96}.
+problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.
+
+	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown automata.
+ We then show that higher-order pushdown automata parity games can be used to solve parity games on
+ the transition systems of pebble pushdown automata,
+using one additional stack level for each pebble.
+ Since solving parity games on higher-order pushdown automata with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown automata with $n$ parameters.
 

### 2023-06-12 reachability/article.tex
up

@@
 
-We consider here a different approach to extending pushdown systems, by allowing them to test the stack content against precise values.
+ Parametric pushdown automata provide a  formalism to
+reason about 
+recursive programs
+making use of parametric constraints.
+% Here, the stack can additionally be compared against parameters that can take unspecified word values. 
+Recently, different variants of parametric pushdown automata have been introduced in the literature 
+% [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
+\cite{hague2011parameterised, esparza2016parameterized, fortin2017model} 
+% In this work, following [21,20,15,27,12], we consider parametric pushdown systems
+however as
+parameterized asynchronous shared-memory systems.
+These systems consist of a leader pushdown automaton
+and arbitrarily many identical contributor pushdown automata, communicating via a shared memory
+in the form of a register which can take finitely many values.
+This variant have been shown to have applications
+to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
+We consider here a different approach to extending pushdown automata with parameters, by allowing them to test equality of the stack content against parameters.
 % Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
 Perhaps a similar model consists in 
-Pushdown automata 
-%with regularity condition on the stack
+pushdown automata 
 with transitions that are conditioned by regular conditions on their stack content, which
@@
 checks against unspecified word values that can be instantiated using parameters.
-This allows for instance to ensure equality between two stack contents appearing in two distincts configurations in a run.
+This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.
 
@@
 
-Hence we propose in this paper to extend pushdown systems with a set of pebbles.
-Compared to pebble word automata, instead of using pebbles
-as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles
-on their universe of stack contents. Dropping a pebble records the current stack content with an available pebble until it is lifted. Lifting a pebble then makes it available again.
-
-A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
-pebbles that it can drop on, and lift from 
-words, using them as markers.
-Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
-\cite{globerman1996complexity, engelfriet1999tree}.
-Automata with nested pebbles were also introduced
-for tree-walking automata. 
-It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}, the main trouble being that the tree-walking automata get lost rather easily.
-Using pebbles is a remedy against getting lost along a tree, but 
-the unrestricted use of pebbles leads to a class of tree languages much larger than the
-regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
-Thus, in both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline, and our model will follow this strict stack discipline too.
-
-One motivation to the study of pebble pushdown systems is their ability to compare a stack content against some later stack content.
-Similar type of storage for later comparison has already been used for instance
-in register pushdown systems. Introduced in \cite{murawski2017reachability}
- as an extension of pushdown systems that 
- can keep data values in
- a set of registers and a stack,
-register pushdown systems
-enable
-manipulation of
-data values from an infinite domain,
-and have applications for instance for
-malware detection and XML schema
-checking \cite{senda2021forward, senda2021ltl}.
+Hence we propose in this paper to extend pushdown systems with a set of 
+parameters.
+%
+% Compared to pebble word automata, instead of using pebbles
+% as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles
+% on their universe of stack contents. Dropping a pebble records the current stack content with an available pebble until it is lifted. Lifting a pebble then makes it available again.
+
+% A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
+% pebbles that it can drop on, and lift from 
+% words, using them as markers.
+% Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
+% \cite{globerman1996complexity, engelfriet1999tree}.
+% Automata with nested pebbles were also introduced for tree-walking automata. 
+% It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}, the main trouble being that the tree-walking automata get lost rather easily.
+% Using pebbles is a remedy against getting lost along a tree, but  the unrestricted use of pebbles leads to a class of tree languages much larger than the regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
+% Thus, in both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline, and our model will follow this strict stack discipline too.
+%
+%
+% One motivation to the study of pebble pushdown systems is their ability to compare a stack content against some later stack content.
+% Similar type of storage for later comparison has already been used for instance in register pushdown systems.
+Introduced in \cite{murawski2017reachability}  as an extension of pushdown systems that can keep data values in a set of registers and a stack, register pushdown systems enable manipulation of data values from an infinite domain, and have applications for instance for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}.
 
@@
 In both cases however, the input trees considered are finite, while
-simulation of plays in pebble pushdown graphs
+simulation of plays in parametric pushdown graphs
 requires the use of pebble alternating tree automata working on infinite trees
-since the stack of a pebble pushdown system is in general unbounded.
+since the stack of a pushdown system is in general unbounded.
 
@@
 the winner of a parity game on a graph of a pushdown system with  
-pebbles.
+parameters.
 
@@
 In Section~\ref{discuss ppda} we close the chapter with a discussion about the methods used, with some directions for future work.
-
 \fi

### 2023-06-12 reachability/article.tex
up

@@
 
-Hence we propose in this paper to extend pushdown systems with a set of 
-parameters.
+% Hence we propose in this paper to extend pushdown systems with a set of  parameters.
 %
@@
 
-Another closely related formalism is that of alternating tree automata with nested pebbles, as we 
-consider in more details in Section 4.
+Another closely related formalism is that of alternating tree automata with nested pebbles%, as we consider in more details in Section 4
+.
 The approach of extending tree automata with pebbles was used in \cite{milo2000typechecking} to show that the XML typechecking problem is decidable, and in \cite{karhumaki2012jewels} to recognize first-order logic on trees.
@@
 
-In Section 2, we provide general notations and preliminary definitions. Section 3 will deal with the introduction of pebble pushdown games.
-Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve pebble pushdown system parity games. We leave the proof of decidability itself  for Section 5.
+In Section 2, we provide general notations and preliminary definitions. Section 3 will deal with the introduction of 
+parametric pushdown games.
+% Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve parametric pushdown system parity games.
+We leave the proof of decidability itself  for Section 5.
 Section 6 at last shows a nonelementary lower bound for the problem.

### 2023-06-12 reachability/article.tex
up

@@
 
-\title{Parity games in pushdown systems with pebbles}
-
-\titlerunning{Parity games in pushdown systems with pebbles}
+\title{Parity games in pushdown systems with parameters}
+
+\titlerunning{Parity games in pushdown systems with parameters}
 
@@
 	parity games over an extension of pushdown systems.
-	More specifically, we consider pebble pushdown systems, in
-	which pebbles can be used to remember some stack content already visited,
-	and compared against.
+	More specifically, we consider parameters pushdown systems, in
+	which parameters can be instantiated to some word over a stack alphabet,
+	and compared against the current stack content.
 	We determine decidability of the problem, and provide a nonelementary lower bound. 
@@
 
-In Section 2, we provide general notations and preliminary definitions. Section 3 will deal with the introduction of 
-parametric pushdown games.
+In Section 2, we provide general notations and preliminary definitions. 
+Section 3 will deal with the introduction of  parametric pushdown games.
 % Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve parametric pushdown system parity games.
 We leave the proof of decidability itself  for Section 5.
-Section 6 at last shows a nonelementary lower bound for the problem.
+Section 4 is Reduction to higher-order pushdown automata
+Section 5 at last shows a nonelementary lower bound for the problem.
 
@@
 % \paragraph{Reachability Games}
-\paragraph{Reachability Games}
+\subsection{Reachability Games}
 
@@
 % \paragraph{Parity Games}
-\paragraph{Parity Games}
+\subsection{Parity Games}
 
@@
 
-\section{Pebble pushdown games}
-
-In this section, we extend pushdown systems with a set of pebbles. 
-Pebble pushdown systems will have the ability to lift or drop pebbles on their universe of stack contents according to a strict stack discipline.
-A \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) and a \textit{lift} pops the last dropped pebble.
-One can think of a pebble as a register that can store a stack content.
-
-\begin{definition}
-An $n$-pebble pushdown system ($n$-PPDS) is a tuple $\mathcal{Z} = (Q, \Gamma,  \Delta )$
-where:
+\section{Parametric pushdown games}
+
+\textcolor{black}{Parametric pushdown automata extend pushdown automata by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
+A parametric pushdown automaton is then a finite automaton extended with a finite set of parameters $P$  and with a stack that can be manipulated by pushing or popping stack symbols and such that, moreover, the automaton can use the top of the stack, or check that the stack content corresponds to a parameter,  to decide which transition to take next.\\
+
+\par\noindent\ignorespacesafterend
+Formally, a {\em parametric pushdown automaton} ({\em PPDA} for short) 
+is a tuple 
+$\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
 \begin{itemize}
-\item $Q$ is a {\em finite set of control states},
-\item $\Gamma$ is a {\em finite stack alphabet},
-% \item $\$ \notin W$ is an {\em initial stack symbol},
-\item  $\Delta   \subseteq  Q  \times (\Gamma \biguplus \{ \$ \})  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times Q  \times (\Gamma^* \cup \{ \!
-\!
-\text{\textit{lift}}, \text{\textit{drop}}\})$ is a {\em finite transition relation}, where ${\$ \notin \Gamma}$ is some bottom of stack symbol.
+\item $Q$ is a non-empty finite {\em set of  states},
+\item $\Gamma$ is a non-empty finite {\em  stack alphabet},
+\item $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$},
+% \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Gamma^*$ is finite {\em  set of rules},
+ \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules},
+\item $q_{init}\in Q$ is an {\em initial  state}, 
+\item $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
+\item $F\subseteq Q$ is a {\em set of final  states}.
 \end{itemize}
-\end{definition}
-
-A transition in an $n$-pebble pushdown system works like a transition in a pushdown system that additionally manages the $n$ pebbles.
-The idea of a transtion $(q, a, i, S, q', \theta) \in \Delta$
-is that, if the automaton $\mathcal{Z}$ is in state $q$ with pebbles $1,\ldots, i$ dropped \---- or without pebble dropped if $i = 0$ \---- with top stack symbol $a$, and stack content which corresponds to the stack contents of the pebbles from $S$ and only these pebbles, then
-$\mathcal{Z}$ goes to state 
-$q'$ and makes modifications to the stack or the pebbles according to
-$\theta$.
-
-A {\em pebble set} of $\mathcal{Z}$ is a set $P \subseteq \{ 1, \ldots, n\}$. For a stack alphabet $\Gamma$, a 
-{\em $P$-pebble assignment} is a function which maps each $j \in P$ to a word in $\$\Gamma^*$.
-For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, f )$, where 
-$q$ is a state,
-$w \in \$\Gamma^*$, and
-$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
-We call  $q$ the current state, $w$ the current stack, and
-$f$ the current pebble assignment. 
-We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
-$f(j) = w_j$, for each
-$j \leq i$.
-
-The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. The transition graph of $\mathcal{Z}$ is $(V,E)$, where $V$ is the set of configurations. %and $E$ is such that
-Moreover, 
-for all $(q, a, i, S, q', \theta) \in \Delta$, with $a \in \Gamma$ (respectively $ \! a = \$ $),
- for all words $w \in \$ \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments
+
+\par\noindent\ignorespacesafterend
+The {\em size} of $\mathcal{Z}$ is defined as
+$|\mathcal{Z}| = |Q| + |\Gamma| + |P| + |R| $.
+We also refer to $\mathcal{Z}$ as an $n$-parametric pushdown automaton if $|P| = n$.
+A {\em stack content} is a word from $ \Gamma^*$. 
+As before we write the top
+of the stack at the right of the word%(we are considering suffix rewriting as in Chapter 15)
+. 
+% A configuration is a  state and a stack content: $(q, w)$,
+% which we rather write $q(w)$, where $w  \in  W^*$, $q  \in Q$. 
+By $\Conf(\mathcal{Z})=Q\times  \Gamma^*$ we denote the set of
+{\em configurations} of $\mathcal{Z}$. As usual we rather write $q(w)$ instead of $(q, w)$.
+A {\em parameter valuation} is a function $\mu$ from $P$ to $\Gamma^*$.
 
 \begin{samepage}
-\noindent
- $f$ such that $f(j) = wa$ (respectively $f(j)= \$ $) for each
-$j \in S$ 	and
- $f(j) \neq wa$ (resp. $f(j) \neq \$ $) for each $j \in \{ 1, \ldots , i \} \setminus S$,
-%the following holds
-we include in $E$ the following transitions:
+%\begin{definition}
+A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
+and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
+%labeled 
+transition system  
+$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
+%\Lambda_{\mathcal{Z}}, 
+\rightarrow_{\mathcal{Z},\mu})$ where
+% $\Lambda_{\mathcal{Z}} = \Gamma \cup P$ and
+for all $ q, q'  \in  Q$, 
+for all $ w,w' \in \Gamma^* $, and 
+for all $ a  \in  \Gamma$, 
+\iffalse
+\begin{eqnarray*}
+q(wa) \xrightarrow{%a
+}_{\mathcal{Z},\mu} q'(w') & \text{if} &  (q, a, q' , op)  \in  R, \\
+q(\mu(p) ) \xrightarrow{%p
+}_{\mathcal{Z},\mu} q'(\mu(p) %\textcolor{red}{\cdot} 
+w) & \text{if} &  (q, p , q' , w)  \in  R.
+\end{eqnarray*}
+% \end{definition}
+\fi
+$q(wa) \xrightarrow{%a
+}_{\mathcal{Z},\mu} q'(w')$
+if there exists a rule in $R$
+ of the form $(q,a,q', op)$ or $(q,p,q', op)$ with $\mu(p)=wa$,
+such that
+		% $\text{\textit{top}}(w) = a$
+		% and 
+% $w' = \theta(w)$.
+either of the following holds
 \begin{itemize}
-\item if $\theta \in \Gamma^*$, 
-$((q,wa,f), (q',w\theta, f)) \in E $ \\
-(respectively, $((q,\$,f), (q',\$\theta, f)) \in E $),
-\item if $ \theta = \text{\textit{drop}}$, 
-$((q,wa,f), (q',wa, f')) \in E $  \\
-(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
-where $f'$ is the $\{ 1, \ldots , i, i+1 \}$-pebble assignment such that
-$f'(j) = f(j)$ for each
-$j \leq i$, and $f'(i+1) = wa$ (respectively, $f'(i+1) = \$$), and
-\item if $\theta = \text{\textit{lift}}$, and $ i \in S $%$i>0$
-, 
-$((q,wa,f), (q',wa, f')) \in E $ \\ 
-(respectively, $((q,\$,f), (q',\$, f')) \in E $),\\
-where $f'$ is the $\{ 1, \ldots , i - 1 \}$-pebble assignment such that
-$f'(j) = f(j)$ for each
-$j < i$.
+\item $ op = \text{\textit{push}}^\gamma$ and $w' = wa\gamma$,
+
+\item $ op = \text{\textit{pop}}$ and $w' = w$, or
+
+\item $ op = \text{\textit{skip}}$ and $w' = wa$.
 \end{itemize}
@@
 
-This defines a graph  $(V,E) $. 
-%
-To obtain
-a parity game, it remains to define the sets $V_0$ and $V_1$ associating the vertices
-to the two players, and the priorities of the configurations. One fixes a disjoint
-union
- $Q = Q_0  \biguplus Q_1 $. Then 
- $V_0$ is the set of configurations with current state in $Q_0$ and 
- $V_1$ is the set of configurations with current state in $Q_1$.
-The mapping  $\Omega$ is first
-defined on $Q$, then  $\Omega (q,w,f) =  \Omega (q)$,  for all words $ w  \in \$ \Gamma^*$, all pebble assignments $f$, and 
-$q  \in  Q$. So the player and the
-priority only depend on the control states of $\mathcal{Z}$. 
-This defines the {\em $n$-pebble pushdown game} $ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$.
-The starting configuration is defined if we also fix an initial state 
-$q_I  \in  Q$ as $v_I = (q_I, \$, f_I)$, where $f_I$ is the $\emptyset$-pebble assignment. \newline
-
-The main problem we are interested in in this article is the following decision problem.
-
-\problemx{Solving pebble pushdown parity game}
-{A pebble pushdown system $\mathcal{Z}= (Q,\Gamma, \Delta )$, a disjoint
-union $Q = Q_0  \biguplus Q_1 $, a priority mapping $\Omega$, an initial state $q_I \in Q$% and initial stack symbol $w_I \in W$
+A {\em $\mu$-run} from $q_0(a_0)$ to $q_n(a_n)$ in $\mathcal{Z}$ is a corresponding
+path in the transition system 
+$T_{\mathcal{Z}}^\mu$ 
+induced by $\mathcal{Z}$ and $\mu$.
+%
+%and will be noted
+%$q_0(z_0) \xrightarrow{\lambda_0}_{\mathcal{Z},\mu} q_1(z_1) \xrightarrow{\lambda_1}_{\mathcal{Z},\mu} \cdots \xrightarrow{\lambda_{n-1}}_{\mathcal{Z},\mu} q_n(z_n)$.
+%
+As with PDA, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
+and $q_n\in F$.
+%
+% We say {\em reachability holds} for $\mathcal{Z}$ if there is a $\mu$ such that there is a $\mu$-run in $\mathcal{Z}$ from $q_{init}(\gamma_{init})$   to some configuration $q(w)$ for some $q\in F$ and some $w \in \Gamma^*$.
+% We refer to Figure~\ref{example_pda} for an instance of a PDA for which reachability holds.
+
+In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
+we prefer to write $q(w)\xrightarrow{%\lambda
+}_{\mathcal{Z}, u}q'(w')$
+to denote $q(w)\xrightarrow{%\lambda
+}_{\mathcal{Z}, \mu}q'(w')$ and
+will call the $\mu$-run an {\em $u$-run}. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\xrightarrow{%\lambda
+}_{\mu}$ (resp. $\xrightarrow{%\lambda
+}_{u}$)
+instead of $\xrightarrow{%\lambda
+}_{\mathcal{Z}, \mu}$ (resp. $\xrightarrow{%\lambda
+}_{\mathcal{Z},u}$). \\
+
+\subsection{PPDA reachability}
+
+Note that the reachability problem for parametric pushdown automata consists in
+the following decision problem.
+
+\problemx{$n$-PPDA reachability}
+{An $n$-PPDA $\mathcal{Z}$.}
+{Does 
+%reachability hold for
+an accepting $\mu$-run for some $\mu\in(\Gamma^*)^P$ exist in
+ $\mathcal{Z}$?\newline}
+
+% \subsubsection{Parametric pshdown games}
+\subsection{PPDA games}
+
+We are moreover interested in games played on parametric pushdown automata's  
+ transition systems.
+Recall that given a 
+%labeled 
+transition system,
+one needs only to provide %a mapping $\Omega$ and 
+a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\
+
+Thus, for an  $n$-parametric pushdown automaton 
+$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
+given a partition of
+$Q$ into $Q_0$ and $Q_1$, 
+we naturally partition the configurations of
+%the transition system of 
+ $\mathcal{Z}$
+% $T^\mu_{\mathcal{Z}}$
+into
+$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
+and
+$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.
+
+With these notations in mind one can define the arena
+$$
+A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
+=
+(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
+induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
+parameter valuation $\mu: P\rightarrow \Gamma^*$.
+
+Given additionally a partition of
+$P$ into $P_0$ and $P_1$, 
+we define 
+$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
+as
+the arena that contains both the configurations and transitions of
+the parameter valuation arena
+$A_{P_0,P_1,\Gamma^*}$, as see on page~\pageref{parametric arena},
+and that moreover contains the configurations and transitions of
+$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$ 
+\---- albeit with configurations additionally including $\mu$ in the tuple to avoid confusion \----
+for
+all parameter valuations $\mu: P\rightarrow \Gamma^*$.
+Moreover, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
+we add a transition
+from $\mu$ towards 
+$q_{init}(\gamma_{init}, \mu)$ in $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$. \label{PPDA reachability game}
+
+Given a 
+priority function $\Omega: Q \to [0, m]$,
+we consider the function
+$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$
+where we set
+$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
+for all
+$w \in \Gamma^*$
+and
+for all $\mu \in (\Gamma^*)^{P}$,
+and
+$\widehat{\Omega}(\mu) = 0$
+for all $\mu \in (\Gamma^* \cup \bot)^{P}$.
+
+% $T_{\mathcal{Z}} = (\Conf(\mathcal{Z}), \rightarrow_{\mathcal{Z},\mu})$
+
+\iffalse
+\textcolor{red}{
+Reachability %, B\"uchi
+ and parity games on $\mathcal{G}_{(\mathcal{Z}, \mu,P_0, P_1, Q_0 , Q_1, \Omega)}$
+are then clearly defined. They have the flaw however of not taking into account the
+way the values assigned to the parameters is chosen. How values are assigned to
+the parameters is the subject of the next Section.
+}
+\fi
+
+We are interested in the following games and problems.
+
+\problemx{Parametric pushdown reachability game}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
+, and $ P = P_0 \biguplus P_1$.}
+{Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
+$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{F \times \Gamma^* \times (\Gamma^*)^P})$ ?\newline}
+
+\problemx{Parametric pushdown parity game}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
+, and $ P = P_0 \biguplus P_1$, 
+and a priority function $\Omega: Q \to [0, m]$%
+%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
+%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
 .}
-{Does player $0$ have a winning strategy from $(q_I,\$, f_I)$ in the  parity game
-$ \mathcal{G}_{\mathcal{Z}, Q_0,Q_1,\Omega}$ ?}
-%$(\mathcal{G}_{\mathcal{Z}}, \Omega)$ ?\newline}
-
-\section{Tree Automata}
-
-In this section, we use pebble alternating two-way tree automata to solve pebble pushdown parity games. 
-We introduce first alternating two-way tree automata, then pebble alternating two-way tree automata,
-then explain the reduction. Decidability questions concerning pebble alternating two-way tree automata
-are dealt with in Section~\ref{section decidability}.
-
-\subsection{Alternating two-way tree automata}
-
-Given a finite set $\Gamma$ of directions, a $\Gamma$-tree is a prefix closed set 
-$T  \subseteq \Gamma^*$,
-i.e., if $x.d  \in  T$, where 
-$x  \in  \Gamma^*$ and $d  \in  \Gamma $, then $x  \in  T $. 
-We will sometimes forget the “.” of the concatenation. 
-The elements of $T$ are called {\em nodes}, the
- empty word  $\epsilon$
-% word $\$ $
-  is the {\em root} of $T $. For every $x.d  \in  T$, $d  \in  \Gamma$ the node $x$ is the
-unique {\em parent} of $x.d$, and $x.d$ is a {\em $d$-child} of $x$. The {\em direction} of a node $x.d$
-$( \neq   \epsilon )$ 
-%$(\neq \$ )$
-is $d$. 
-The full infinite tree is 
-$T = \Gamma^*$. A {\em path (branch)} of a tree $T$ is
-a sequence  $\beta   \in  T^{\infty}$  such that 
-$ \beta  = u_0 u_1 \ldots u_n$ or 
-$ \beta  = u_0 u_1 u_2 \ldots$, where
-$u_0 =  \epsilon$  and
- $\forall i < n,  \exists d  \in  \Gamma$, $u_{i+1} = u_i.d$. 
- A path can be finite or infinite.
-
-For a finite set $X$, $\mathcal{B}^+ (X)$ is the set of positive Boolean formulas
-over $X$ (i.e., Boolean formulas built from elements in $X$ using only $\wedge$ and $\vee$),
-where we also allow the formulas $\text{\textit{true}}$ and $\text{\textit{false}}$. For a set $Y  \subseteq  X$ and a
-formula $ \vartheta   \in  \mathcal{B}^+ (X)$, we say that $Y$ satisfies $ \vartheta$  iff assigning $\text{\textit{true}}$ to elements in
-$Y$ and $\text{\textit{false}}$ to elements in $X \setminus Y$ makes  $\vartheta $ true.
-
-The set $\text{\textit{types}}(\Gamma) = \{r \} \cup \Gamma$ 
-describes the possible types of a node in the full infinite tree $\Gamma^*$. 
-Here $r$ stands for the root, $\gamma \in \Gamma$ for a $\gamma$-child.
-
-To navigate through the tree let $ext(\Gamma ) := \Gamma \biguplus \{ \epsilon , \uparrow \}$ be the extended set of directions: The symbol $\uparrow$ means “go to parent node”, 
-and  $\epsilon$  means “stay on the present
-node”. Formally we define $ \forall u  \in \$ \Gamma^* , d  \in  \Gamma, u.\epsilon  = u$ and $u.d. \! \uparrow= u$, while the node 
-$ \epsilon . \! \uparrow $ is not defined.
-
-\begin{definition}
-An alternating two-way automaton over
- $\Gamma$-trees (ATWA) is a tuple
-$\mathcal{A} = (Q,  \delta , q_I , 
-\Omega)$ where
-\begin{itemize}
-\item $Q$ is a {\em finite set of states},
-\item $\delta  : Q  \times  
-%
-\text{\textit{types}}(\Gamma)
-%W \cup \$
-  \to  \mathcal{B}^+ (Q \times ext(\Gamma))$ is a {\em transition function},
-\item $q_I$ is an {\em initial control state}, and
-\item $\Omega: Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
-\end{itemize}
-\end{definition}
-
-The idea of a transition $ \delta (q, a) = ( q',\uparrow ) \wedge (q'',d)$ with $a,d\in\Gamma$ is the following: 
-if the automaton
-$\mathcal{A}$ is in state $q$ on the node $x.a$ of the 
-tree $T$, it will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
-another “copy” in state $q''$ to $x.a.d$. 
-After that the two copies are
-running independently. They may come again to the same node with two different
-states.
-
-More precisely a run of an alternating two-way automaton $\mathcal{A}$ over a tree
-is
-a tree $T_r$ in which every node is labelled by an
-element of $Q \times \Gamma^*$. The latter tree is like the unfolding of the run, its structure
-is different from the tree $\Gamma^*$, and its set of directions range rather on the naturals smaller than $|Q|*|ext(\Gamma)|$; a node in $T_r$ labelled by $(q,x)$ describes a “copy”
-of the automaton that is in state $q$ and is situated at the node $x$ of $\Gamma^*$. 
-Note
-that many nodes of $T_r$ can correspond to the same node of $\Gamma^*$, because the
-automaton can come back to a previously visited node. 
-The label of a node
-and its successors have to satisfy the transition function,
-that is, for a node $y$ of $T_r$ labelled by $(q,x)$, and $\delta(q,t)= \vartheta$ where $t$ is the type of $x$, there is a (possibly empty) set 
- $Y = \{ (q_0,d_0),(q_1,d_1), \ldots (q_\ell, d_\ell) \}  \subseteq  Q \times ext(\Gamma)$, such that $Y$ satisfies  $\vartheta $, and for all
-$(q_i,d_i)  \in  Y$, the node $y$ has a corresponding son $y.i$ labelled by $(q_i,x.d_i)$. Finally, the root of $T_r$ is labelled by $(q_I,\epsilon)$.
-
-Remember that $x.d$ can be $x. \epsilon$  or $x. \! \uparrow$, and the latter is defined only if 
-%$x \neq \$ $.
-$x  \neq   \epsilon $.
-So the run cannot go up from the root of the input tree. Note that it cannot use a
-transition  $\delta (q, a) = \text{\textit{false}}$ since the formula $\text{\textit{false}}$ cannot be satisfied.
-A run $T_r$ is accepting if all its infinite paths satisfy the parity acceptance 
-condition
-given by the priority function. 
-% The finite paths of a run that end with a transition $ \vartheta $ equivalent to  $true$, which is viewed as successful termination. 
-An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
-%this path
-the sequence $\gamma$ of corresponding labels
- is even, i.e.
-%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
- $\min(\{ \Omega(q) \mid (q,x) \in \text{\textit{Inf}}(\gamma) \} $ is even. 
-Such
-a path in the run consists of following only one “copy” of the automaton.
-
-% The generalization of twa with nested pebbles came recently, by two independent motivations: On the one hand, with the advancement of XML theory, finite state recognizers (with name n-pebble tree automata) were used in [21] to show that the XML typechecking problem is decidable.
-On the other hand, the concept of n-pebble tree-walking automata (n-ptwa) were defined in [9] to recognize first-orderlogic on trees.
-Later, in [10]n-ptwa were extended with a more general pebble handling.
-
-\subsection{Alternating two-way tree automata with pebbles}
-
-Informally an $n$-pebble alternating two-way tree automaton works much like an alternating two-way tree automaton, but extended with a fixed set of pebbles, numbered from $1$ to $n$ that it can place on the tree.
-% At each time, pebbles $1,\ldots, i$ are placed on some nodes of the tree, for some $i \in \{1, 2, \ldots , n \}$, or no pebble are placed on the tree. 
-To navigate through the tree the automaton can stay at the current node, move to its parent, to its 
- children, it can lift a pebble or place a pebble on the current node.
- Which of these transitions can be applied depends on the current state, the set
- of pebbles at the current node, 
-and the type of the current node (root, $\gamma$-child for $\gamma \in \Gamma$). 
-Operations $\text{\textit{lift}}$ and $\text{\textit{drop}}$ manage the
-pebbles much like in pebble pushdown systems.
-We indicate the possible kinds of moves of a pebble automaton by elements of the set 
-$ext_2(\Gamma) = ext(\Gamma) \biguplus \{ \text{\textit{lift}}, \text{\textit{drop}}\}$.
-% Clearly $\text{\textit{drop}}$ refers to dropping a pebble and $\text{\textit{lift}}$ to lifting a pebble. Note that the notation is  unambiguous by virtue of the stack discipline. 
-% Finally recall that if $S$ is a set then $\mathcal{P}(S)$ denotes the powerset of $S$.
-
-\begin{samepage}
-\begin{definition}
-For $n\geq 0$, an $n$-pebble alternating two-way automaton ($n$-PATWA) over  
-$\Gamma$-trees is a tuple
-$\mathcal{A} = (Q,   
-\delta , q_I , 
-\Omega)$ where
-\begin{itemize}
-\item $Q$ is a {\em finite set of states},
-\item $\delta  : Q  \times \{ 0, 1, \ldots, n\} \times \mathcal{P}(\{ 1, \ldots, n\}) \times 
-\text{\textit{types}}(\Gamma)
-  \to  \mathcal{B}^+ (Q \times ext_2(\Gamma))$ is a {\em transition function},
-\item $q_I$ is an {\em initial control state}, and
-\item $\Omega : Q \to \{ 0, 1, \ldots, m\}$ is a {\em priority function}.
-\end{itemize}
-\end{definition}
-\end{samepage}
-
-The idea of a transition in a $n$-pebble alternating two-way tree automaton works like a transition in an alternating two-way tree automaton,
-except the transition has additional conditions on the pebble assignment, and additional moves for taking care of the pebble.
-Consider for instance 
-that the automaton
- $\mathcal{A}$ is on the node $x.a$, in state $q$, pebbles $1, 2$ are on the tree, and exactly pebble $1$ is placed on the node $x.a$.
-Then a transition
-$\delta(q, 2, \{ 1\}, a) =  (q',\uparrow) \wedge (q'',\text{\textit{drop}})$
- will send a “copy” of $\mathcal{A}$  in state $q'$ to the node $x$ and
-another “copy” in state $q''$ to $x.a$ with both pebble $1$ and pebble $3$ placed on the node. 
-After that the two copies are
-running independently.
-
-%A {\em pebble set} of $\mathcal{A}$ is a set $P \subseteq \{ 1, \ldots, n\}$. 
-For a tree structure $T$, a 
-{\em $P$-pebble assignment} for a pebble set $P$ is a function which maps each $j \in P$ to a node in $T$.
-For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, x, f )$, where $x$ is a node, $q$ a state and
-$f$ a $\{ 1, \ldots , i \}$-pebble assignment \---- $f$ is the $\emptyset$-pebble assignment in case $i=0$. 
-We call $q$ the current state, $x$ the current node, and
-$f$ the current pebble assignment. 
-We also write $(q, x, x_{1} , \ldots , x_{i} )$ if 
-$f(j) = x_j$, for each
-$j \leq i$.
-
-Then, formally, a run of an $n$-pebble alternating two-way automaton $\mathcal{A}$
-over a tree is a tree $T_r$ in which every node is labelled by 
-an $i$-configuration,
-for some $i \in \{0, 1, \ldots, n\}$.
-A node in $T_r$ labelled by $(q, x, x_{1}, x_{2}, \ldots, x_{i})$ describes a “copy” of the automaton that is in state $q$, and is situated on the node $x$ of $\Gamma^*$, with the pebbles up to $i$ already dropped, with the $j$-th pebble for $j \leq i$ dropped on the node $x_{j}$.
-
-The label of a
-node and its successors have to satisfy the transition function. % However, recall there are restrictions on the drop-operation and the lift-operation : only
-% the pebble with the number $i$ can be lifted and only the pebble with number $i - 1$ can be
-% placed.
-Formally, this works much the same way as for alternating two-way tree automata; 
-that is, the root of $T_r$ is
-labelled by $(q_I,\epsilon, f_I )$, and, for a node $y$ of $T_r$ labelled by an $i$-configuration $(q, x, f)$, and 
-for
-$\delta(q, i, S, t) = \vartheta$ where $t$ is
-the type of $x$, 
-and $f(j) = x$ for each $j \in S$, and $f(j) \neq x$ for each $j \in \{1,\ldots,i \} \setminus S$,
-there is a (possibly empty) set 
-$Y \subseteq Q \times ext_2(\Gamma )$, such that $Y$ satisfies $\vartheta$,
-and for all $(q',d) \in Y$, the node $y$ has a corresponding son
-labelled with %$(x.d, q', f')$, where
-\begin{itemize}
-\item if $d \in ext(\Gamma)$,
-the $i$-configuration $(q ', x.d , f)$,
-%
-\item if $d = \text{\textit{drop}}$, the $(i+1)$-configuration  $ (q', x, f')$, where the domain of $f'$ is $\{ 1, \ldots, i, i+1\}$, $f'$ is equal to $f$ on the domain of $f$, and $f'(i+1) =x$, and
-%
-\item if $d = \text{\textit{lift}}$
-and $f(i) = x$, i.e. if the pebble $i$ is on the current node%
-, the $(i-1)$-configuration $(q', x, f')$, where $f'$ is the restriction of $f$ to $\{1, \ldots, i-1 \}$.
-\end{itemize}
-
-A run $T_r$ is {\em accepting} if all its infinite paths satisfy the parity acceptance 
-condition 
-given by the priority function $\Omega$.
-% Again finite paths end with a transition with formula $ \vartheta $ equivalent to $\text{\textit{true}}$, which is viewed as successful termination. 
-An infinite path $ \beta   \in  T_r^{\omega}$  satisfies the acceptance condition iff the smallest priority appearing infinitely often in
-%this path
-the sequence $\gamma$ of corresponding labels
- is even, i.e.
-%$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
- $\min(\{ \Omega(q) \mid (q,x,f) \in \text{\textit{Inf}}(\gamma) \} $ is even.
-
-By a pebble alternating tree-walking automaton (PATWA) we mean an $n$-PATWA for some $n$. 
-%Note that a $n$-pebble automaton without pebbles is just an alternating tree-walking automaton.
-
-\subsection{Reduction details}
-
-Alternating two-way tree automata with parity winning conditions have already been used to solve parity games on pushdown graphs in \cite{cachat2002two}, with a technique first developed to solve the model-checking problem in \cite{kupferman2000automata}.
-The main
-idea was to reduce a model-checking problem to an emptiness problem for the class of alternating two-way parity tree automata, and was then adapted to work on pushdown parity games.
-%
-%
-This leads to the following reduction: given a 
-parity game on pushdown graph $\mathcal{G}$ and a starting configuration $v_I$, we can
-effectively construct an ATWA $\mathcal{A}$
-such that
-player $0$ has a winning strategy from $v_I$ in $\mathcal{G}$ iff $\mathcal{A}$ accepts the full
-infinite tree $\Gamma^*$. The main idea of our reduction is that, when pushdown systems and ATWA are extended with a set of pebbles to manage, the technique used %is conserved.
-still works.
-We refer the reader to Appendix~\ref{appendix tree reduction} for more details on the proof.
-
-\begin{theorem}~\label{theorem reduction PATWA}
-For every $n$-PPDS $\mathcal{Z} = (Q_0 \biguplus Q_1, \Gamma,\Delta)$, 
-priority mapping $\Omega$, and state $q_I \in Q_0 \biguplus Q_1$,
-we can construct 
-effectively
-an $n$-PATWA $\mathcal{A}$ 
-such that
-player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the parity game 
-$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$
-iff
-$\mathcal{A}$ accepts the full infinite tree $\Gamma^*$.
-\end{theorem}
-
-% Decidability of checking that the $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\Gamma^*$ is then dealt with later in the subsequent section.
-
-\section{Decidability}\label{section decidability}
-
-\section{Lower bound}
-
-The aim of this section is to show a nonelementary lower bound for the 
-problem of deciding whether player $0$ has a winning strategy from some starting configuration in
-a pebble pushdown parity game.
-% problem of deciding winning configuration for pebble pushdown parity games.
-
-We reduce the satisfiability problem for first-order logic on words
-to the problem of {\sc Solving pebble pushdown parity game}. 
-We know from \cite{Sto74}
-that satisfiability is nonelementary. 
-%
-
-% Here a function is in {\sf elementary} if it is  $O(h_k (n))$ for one of the $k$-fold exponential functions $h_k$ with $h_0 (n) = n$ and $h_{k+1} (n) = 2^{h_k (n)}$.
-
-We define the tower function $T : \N \times \R \to \R$ by $T(0,r) =r$ and
-$T(h+1,r)= 2^{T(h,r)}$ for all $h \in \N, r \in \R$. Thus $T(h,r)$ is a tower of $2$s of height $h$ with an $r$ sitting on top.
-Observe that for all $n,h \in \N$ with $n\geq 1$, we have
-$T(h, log^{(h)}(n)) = n$.
-%
-% 
-Here a problem is in ${\sf ELEMENTARY}$ if it can be solved in time $O ( T(n, 0) )$.
-
-For a finite alphabet $\Sigma$,  let $\tau (\Sigma )$ be the vocabulary consisting of a binary relational symbol $\leq$,  and a unary relational symbol $P_a$ for every $a \in \Sigma$. 
-A word $w \in \Sigma^*$ can be seen as a $\tau (\Sigma )$-structure where the set of elements consists of a finite interval of $\N$.
-Elements are seen as positions in the word and for every element $i$, there exists precisely one $a \in \Sigma$ such that $i$ is in the interpretation of $P_a$.
-Lastly $\leq$ is interpreted as the linear order of $\N$.\\
-
-It is well-known that if we are interested in the complexity of first-order or monadic
-second-order model-checking on words, the alphabet can be assumed to be $\{0,1\}$ without loss of generality.
-Thus the satisfiability problem is concerned with the following question.
-\begin{samepage}
-\problemx{$\FO$ satisfiability on words}
-{A  
-$\FO[\tau(\{0,1\})]$-sentence $\phi$.}
-{Does there exists a word $w\in \{0, 1\}^*$, such that $w \models \phi$ ?\newline}
-\end{samepage}
-
-The result that is previously known and motivates the upcoming reduction consists in the following.
-
-\begin{theorem}[\cite{Sto74}]\label{FOnonelementary}
-
-%  Assume that $FPT \neq AW[*]$, and let $f$ be an elementary function and let $p$ be a polynomial.
-%  There is no model-checking algorithm for first-order logic on the class of words 
-%  whose running time is bounded by
-%  $$ f(k) \cdot p(n)$$
-
-% where $k$ denotes the size of the input sentence of the model-checking problem and $n$ the
-% size of the input word.
-
-The {\sc $\FO$ satisfiability on words} problem is not in {\sc ELEMENTARY}.
-	% SAT(N,<) is not elementary-recursive
-	% and in fact SAT(N,<) \not\in NSPACE(g(log_b(n),0)) for all b > 3.
-\end{theorem}
-
-\subsection{Reduction from FO SAT on words}
-
-The following theorem states a polynomial time reduction from
-{\sc $\FO$ satisfiability on words}
-%We consider reduction 
-towards the problem of %following problem.
-{\sc Solving pebble pushdown parity game}.
-
-\begin{samepage}
-\begin{theorem}
-{\sc FO satisfiability on words} is polynomial time reducible to {\sc Solving pebble pushdown parity game}.
-\end{theorem}
-\end{samepage}
-
-\begin{proof}
-
-\noindent
-We start with
-% a $\tau(\{0,1\})$-structure $w$ and
- a 
- $\FO[\tau(\{0,1\})]$%%
-% $\FO$%
--formula $\phi$.
-%
-% Note that $w$ corresponds to a word $w \in \{0,1\}^*$ of size $n$.\\
-%
-% \noindent
-We construct
-a pebble pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
-union $Q_\phi = Q_0  \biguplus Q_1 $, a priority mapping $\Omega_\phi$, and an initial state $q_I \in Q_\phi$
-such that
-player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the  parity game
-$\mathcal{G}_{\mathcal{Z_\phi}, Q_0,Q_1, \Omega_\phi}$
-if and only if
-there exists a word $w \in \{0,1\}^*$ such that
-$w  \models \phi $. \\
-
-\noindent
-We assume without loss of generality that $\phi$ is written in the prenex normal form
-$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k ~ \psi$, where 
-an even number $k$ of quantifiers alternate between existential and universal ones, and 
-$\psi$ is quantifier-free and in disjunctive normal form.
-
-Player $0$ starts by pushing a word $w$ onto the stack, then places the first pebble at the current position. The goal of this first pebble is to remember the word $\$w$, which corresponds to the 
-word “guessed” by player $0$.
-This pebble is meant to enforce that the stack remain a prefix of $\$w$, with each stack content of the play corresponding to a position in the word.
-To enforce this, once pebble $1$ is placed, we allow in $\mathcal{Z}_\phi$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position on the word.
-A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to pebble $1$.
-
-Then players alternate placing pebbles on prefixes of $\$w$. Universal variables correspond to pebbles placed by player $1$ and existential variables to pebbles placed by player $0$. 
-Once players have chosen values for their respective variables, the pebble assignment $f$ is fixed. It is then time
-for player $0$ to ensure that
-$w$ satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
-$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
-
-Player $0$, being the existential player, choses one of the conjunctive clause in $\psi$, essentially claiming to be able to prove it.
-Then player $1$, the universal player, chooses one of the atomic formula in the clause to test.
-%The process can be seen in Figure~\ref{DNF}. 
-Testing the atomic formula is the purpose of a gadget 
-%$ \C_{\psi_{i,j}}$ built such that
-in which
- player $0$ has a winning strategy 
- %in $ \C_{\psi_{i,j}}$
-  if and only if 
-$w$ satisfies %$\psi_{i,j}$
-the atomic formula
- if the variables $x_1, \ldots, x_k$ are interpreted by 
-$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
-
-The gadget depends on the type of the atomic formula.
-For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then check that the the top of the stack
-$x$ is the right letter. For one checking that $x = x'$ where $x$ and $x'$ are both variables,
-player $0$ goes to the position corresponding to $x$ and check against the presence of the pebble corresponding to $x'$ as well. If the formula is a negation, players exchange their roles.
-\end{proof}
-
-Then, the nonelementary lower bound follows from Theorem~\ref{FOnonelementary}.
-
-\begin{theorem}
-
-The problem of {\sc Solving pebble pushdown parity game} is not in {\sc ELEMENTARY}.
-
-\end{theorem}
-
-\section{Conclusion}
-
-In this paper we have shown that
-deciding
-the winner of a parity game on a graph of a pebble pushdown system 
-is nonelementary.
-
-	For the nonelementary lower bound
-we reduced the $\FO$ satisfiability problem on words \---- known to be nonelementary
-from \cite{Sto74} \---- to the
-problem of deciding whether player $0$ has a winning strategy from some starting configuration in 
-a pebble pushdown parity game.
-
-	For the decidability upper bound, we  used pebble alternating two-way tree automata to solve pebble pushdown parity games.
-Alternating two-way tree automata (ATWA) with parity winning conditions have been used to solve parity games on pushdown graphs \cite{cachat2002two}.
-The technique was first developped to solve model-checking problem for pushdown graphs in \cite{kupferman2000automata}.
-We showed that, when pushdown graphs and ATWA are extended with a set of pebbles to manage, the technique
-also works.
-We then proved decidability 
-% for checking that an $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\{0,1\}^*$ 
-by %reducing it
-a reduction
- to the problem of model-checking $\MSO$ on a well chosen structure. To ensure decidability of model-checking $\MSO$ on the chosen structure, we made use of 
-Muchnik's Theorem: the structure is constructed by applying a successive number of times Muchnik's iteration to a structure $\mathfrak{A}_0$ for which model-checking $\MSO$ is decidable.
-%  After providing this structure, we will show that  checking that an $n$-PATWA $\mathcal{A}$ accepts the full binary tree $\{0,1\}^*$ reduces itself to the problem of model-checking $\MSO$ on it.
-
-\bibliography{bibliography/bib}
-
-\appendix
-
-\section{Reduction from PPDS to PATWA}\label{appendix tree reduction}
-% \subsection{Construction}
-
-We provide in this appendix more details on Theorem~\ref{theorem reduction PATWA}. Let us first recall the theorem.
-
-\begin{theorem*}
-For every $n$-PPDS $\mathcal{Z} = (Q_0 \biguplus Q_1, \Gamma,\Delta)$, 
-a state $q_I \in Q = Q_0 \biguplus Q_1$,
-and priority mapping $\Omega: Q \to \{1, \ldots, m\}$,
-we can construct 
-effectively
-an $n$-PATWA $\mathcal{A}$ 
-such that
-player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the parity game 
-$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$
-iff
-$\mathcal{A}$ accepts the full infinite tree $\Gamma^*$.
-\end{theorem*}
-
-We construct a pebble alternating two-way tree automaton $\mathcal{A}$ that determines if player
-$0$ has a winning strategy from $(q_I, \$, f_I)$ in $\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$, i.e., wins every play, whatever player $1$ does. 
-The
-automaton $\mathcal{A}$ will simulate the transitions of the pebble pushdown system $\mathcal{Z}$ on the full
-$\Gamma$-tree, guess nondeterministically the best moves of player $0$ and follow each
-possible move of player $1$ using alternation.
-This tree automaton can
-go either up or down the tree: it goes down to simulate a rule that pushes
-some new symbol, and it goes up to simulate a popping rule, and it drops or lifts a pebble to represent the dropping or lifting of a pebble. For each possible move, the control state has to
-be updated in accordance with the pushdown transition rules. As we aim to
-simulate a game, the tree automaton needs to be alternating: states
- associated to player $0$ states will have disjonctive booleans formulas
-  while states associated to
-player $1$ states will have conjonctive ones.
-Finally, the priority mapping is inherited from $\Omega$.
-
-As an example, the stack effect of a transition $(q, a, 0, \emptyset, q' , cb)  \in \Delta_{\mathcal{Z}} $ from a configuration $q(wa, \mu)$
-of
-the PPDS can be simulated by a $n$-PATWA over the full
-$\Gamma$-tree from the node $ua$ by the following sequence of moves: 
-$\uparrow, c, b$. Indeed
-$wa \! \uparrow cb = wcb$. We have chosen suffix rewriting rather than prefix to conform with
-the notation of the tree. 
-The control states of $\mathcal{A}$ serve to store the control states of $\mathcal{Z}$.
-In order to be able to simulate with many steps a transition of $\mathcal{Z}$, the control states of $\mathcal{A}$ are also used to remember the letters it has to write (see Figure~\ref{stack simulation} for an illustration).
-
-\begin{center}
-	\begin{figure}
-		\hspace{2.1cm}
-\includegraphics[width=0.66\textwidth]{figures/stack_simulation}
-	\caption{Transition $(p, a, \emptyset, q, cb)$ of the RPDS
-	%(long arrow)
-	 simulated by the PATWA.}
-	 \label{stack simulation}
-	\end{figure}
-\end{center}
-
-Moving up or down the tree structure is enough in order to simulate the modifications of the stack
-performed by transitions of $\mathcal{Z}$. Additionally, dropping or lifting pebble is enough to simulate the dropping or lifting of pebbles in the PPDS. As an example, we simulate
-$(q, a, 0, \emptyset, q', \text{\textit{drop}})  \in   \Delta_{\mathcal{Z}} $ by the single move $\text{\textit{drop}}$, and
-$(q, a, i, \{i\}, q', \text{\textit{lift}})  \in   \Delta_{\mathcal{Z}} $ by the single move 
-$\text{\textit{lift}}$. Additionally, we simulate checks against pebble sets in $\mathcal{Z}$ by similar checks in the PATWA.
-
-More formally now, the control states of $\mathcal{A}$ are tuples
-from $( Q \times ext_2(\Gamma) )^*$  
-composed of a control state of  $\mathcal{Z}$, and
-a sequence of elements of $ ext_2(\Gamma)$.
-
-The sequence of letters is used to remember the letters the automaton has to write. 
-Each sequence of letters has at worst the length of a corresponding output of the transition relation of $\mathcal{Z}$, hence the sequences of letters of $\Gamma$ are more precisely elements of $\Gamma^{ \leq |\Delta_{\mathcal{Z}}|}$.
-
-From a state $(q, \epsilon)$, for a transition
-$(q, a, i, S, q', x) \in \Delta_{\mathcal{Z}}$,
-the automaton's first move simulating the transition, be it the dropping of a pebble, the lifting of a pebble, or going up the tree, is accompanied by 
-checks that the pebbles $1, \ldots, i$ are placed on the tree \---- or that no pebble is on the tree if $i=0$ \---- and checks against the presence of the pebbles of $S$.
-
-From a state $(q , bx)$ ($bx \neq \epsilon$ because $b  \in  \Gamma$) the automaton goes down
-in direction $b$, that is to say writes $b$, whatever it reads, and remembers the
-(sub)word $x$ it still has to write and the state $q$ of the pushdown system.
-
-If $q  \in  Q_1$ then $ (q,w)  \in  V_1$ for all $w \in \Gamma^{ \leq |\Delta_{\mathcal{Z}}|}$
+{Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
+$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}
+
+\iffalse
+\problemx{Parametric pushdown reachability game}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, a disjoint
+union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
+.}
+{Does player $0$ have a 
+strategy $\sigma_0$ for the parameter valuation game  such that
+for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
+player $0$ has a
+winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
+$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
+
+\problemx{Parametric pushdown parity game}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, a disjoint
+union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
+, 
+a priority mapping $\Omega: Q \to \{0, \ldots m\}$%
+%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
+%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
+.}
+{Does player $0$ have a 
+strategy $\sigma_0$ for the parameter valuation game  such that
+for all player $1$ strategies $\sigma_1$ for the parameter valuation game, 
+player $0$ has a
+winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
+$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
+\fi
+
+We write {\sc $n$-parametric reachability parity game}
 and
-$\mathcal{A}$
-executes from $(q,\epsilon)$ all the possible moves of player $1$, to ensure that player $0$ can win after
-each of these moves. But if $q  \in  Q_0$, $\mathcal{A}$ in $(q,\epsilon)$ chooses nondeterministically a move of
-player $0$ and tries to make player $0$ win. \newline
-
-The priority mapping of $\mathcal{A}$ is almost the same as the one of 
-$\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$: 
-$\Omega_\A(q, x) =
-\Omega (q)$. 
-The initial state of $\mathcal{A}$ causes it to go deterministically to the starting
-configuration of the game.
-
-\begin{lemma}~\label{lemma reduction PATWA}
-Player $0$ has a winning strategy from $(q_I, \$, f_I)$ in $\mathcal{G}_{\mathcal{Z},Q_0,Q_1,\Omega}$ iff $\mathcal{A}$ accepts the full
-infinite tree $\Gamma^* $.
-
-\end{lemma}
-
-This concludes our construction for the proof of Theorem~\ref{theorem reduction PATWA}.
+{\sc $n$-parametric pushdown parity game} if 
+ the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
 
@@
 
+\section{Lower bound}
+
+The aim of this section is to show a nonelementary lower bound for the 
+problem of deciding whether player $0$ has a winning strategy from some starting configuration in
+a pebble pushdown parity game.
+% problem of deciding winning configuration for pebble pushdown parity games.
+
+We reduce the satisfiability problem for first-order logic on words
+to the problem of {\sc Solving pebble pushdown parity game}. 
+We know from \cite{Sto74}
+that satisfiability is nonelementary. 
+%
+
+% Here a function is in {\sf elementary} if it is  $O(h_k (n))$ for one of the $k$-fold exponential functions $h_k$ with $h_0 (n) = n$ and $h_{k+1} (n) = 2^{h_k (n)}$.
+
+We define the tower function $T : \N \times \R \to \R$ by $T(0,r) =r$ and
+$T(h+1,r)= 2^{T(h,r)}$ for all $h \in \N, r \in \R$. Thus $T(h,r)$ is a tower of $2$s of height $h$ with an $r$ sitting on top.
+Observe that for all $n,h \in \N$ with $n\geq 1$, we have
+$T(h, log^{(h)}(n)) = n$.
+%
+% 
+Here a problem is in ${\sf ELEMENTARY}$ if it can be solved in time $O ( T(n, 0) )$.
+
+For a finite alphabet $\Sigma$,  let $\tau (\Sigma )$ be the vocabulary consisting of a binary relational symbol $\leq$,  and a unary relational symbol $P_a$ for every $a \in \Sigma$. 
+A word $w \in \Sigma^*$ can be seen as a $\tau (\Sigma )$-structure where the set of elements consists of a finite interval of $\N$.
+Elements are seen as positions in the word and for every element $i$, there exists precisely one $a \in \Sigma$ such that $i$ is in the interpretation of $P_a$.
+Lastly $\leq$ is interpreted as the linear order of $\N$.\\
+
+It is well-known that if we are interested in the complexity of first-order or monadic
+second-order model-checking on words, the alphabet can be assumed to be $\{0,1\}$ without loss of generality.
+Thus the satisfiability problem is concerned with the following question.
+\begin{samepage}
+\problemx{$\FO$ satisfiability on words}
+{A  
+$\FO[\tau(\{0,1\})]$-sentence $\phi$.}
+{Does there exists a word $w\in \{0, 1\}^*$, such that $w \models \phi$ ?\newline}
+\end{samepage}
+
+The result that is previously known and motivates the upcoming reduction consists in the following.
+
+\begin{theorem}[\cite{Sto74}]\label{FOnonelementary}
+
+%  Assume that $FPT \neq AW[*]$, and let $f$ be an elementary function and let $p$ be a polynomial.
+%  There is no model-checking algorithm for first-order logic on the class of words 
+%  whose running time is bounded by
+%  $$ f(k) \cdot p(n)$$
+
+% where $k$ denotes the size of the input sentence of the model-checking problem and $n$ the
+% size of the input word.
+
+The {\sc $\FO$ satisfiability on words} problem is not in {\sc ELEMENTARY}.
+	% SAT(N,<) is not elementary-recursive
+	% and in fact SAT(N,<) \not\in NSPACE(g(log_b(n),0)) for all b > 3.
+\end{theorem}
+
+\subsection{Reduction from FO SAT on words}
+
+The following theorem states a polynomial time reduction from
+{\sc $\FO$ satisfiability on words}
+%We consider reduction 
+towards the problem of %following problem.
+{\sc Solving pebble pushdown parity game}.
+
+\begin{samepage}
+\begin{theorem}
+{\sc FO satisfiability on words} is polynomial time reducible to {\sc Solving pebble pushdown parity game}.
+\end{theorem}
+\end{samepage}
+
+\begin{proof}
+
+\noindent
+We start with
+% a $\tau(\{0,1\})$-structure $w$ and
+ a 
+ $\FO[\tau(\{0,1\})]$%%
+% $\FO$%
+-formula $\phi$.
+%
+% Note that $w$ corresponds to a word $w \in \{0,1\}^*$ of size $n$.\\
+%
+% \noindent
+We construct
+a pebble pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
+union $Q_\phi = Q_0  \biguplus Q_1 $, a priority mapping $\Omega_\phi$, and an initial state $q_I \in Q_\phi$
+such that
+player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the  parity game
+$\mathcal{G}_{\mathcal{Z_\phi}, Q_0,Q_1, \Omega_\phi}$
+if and only if
+there exists a word $w \in \{0,1\}^*$ such that
+$w  \models \phi $. \\
+
+\noindent
+We assume without loss of generality that $\phi$ is written in the prenex normal form
+$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k ~ \psi$, where 
+an even number $k$ of quantifiers alternate between existential and universal ones, and 
+$\psi$ is quantifier-free and in disjunctive normal form.
+
+Player $0$ starts by pushing a word $w$ onto the stack, then places the first pebble at the current position. The goal of this first pebble is to remember the word $\$w$, which corresponds to the 
+word “guessed” by player $0$.
+This pebble is meant to enforce that the stack remain a prefix of $\$w$, with each stack content of the play corresponding to a position in the word.
+To enforce this, once pebble $1$ is placed, we allow in $\mathcal{Z}_\phi$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position on the word.
+A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to pebble $1$.
+
+Then players alternate placing pebbles on prefixes of $\$w$. Universal variables correspond to pebbles placed by player $1$ and existential variables to pebbles placed by player $0$. 
+Once players have chosen values for their respective variables, the pebble assignment $f$ is fixed. It is then time
+for player $0$ to ensure that
+$w$ satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
+$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
+
+Player $0$, being the existential player, choses one of the conjunctive clause in $\psi$, essentially claiming to be able to prove it.
+Then player $1$, the universal player, chooses one of the atomic formula in the clause to test.
+%The process can be seen in Figure~\ref{DNF}. 
+Testing the atomic formula is the purpose of a gadget 
+%$ \C_{\psi_{i,j}}$ built such that
+in which
+ player $0$ has a winning strategy 
+ %in $ \C_{\psi_{i,j}}$
+  if and only if 
+$w$ satisfies %$\psi_{i,j}$
+the atomic formula
+ if the variables $x_1, \ldots, x_k$ are interpreted by 
+$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
+
+The gadget depends on the type of the atomic formula.
+For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then check that the the top of the stack
+$x$ is the right letter. For one checking that $x = x'$ where $x$ and $x'$ are both variables,
+player $0$ goes to the position corresponding to $x$ and check against the presence of the pebble corresponding to $x'$ as well. If the formula is a negation, players exchange their roles.
+\end{proof}
+
+Then, the nonelementary lower bound follows from Theorem~\ref{FOnonelementary}.
+
+\begin{theorem}
+
+The problem of {\sc Solving pebble pushdown parity game} is not in {\sc ELEMENTARY}.
+
+\end{theorem}
+
+\section{Conclusion}
+
+In this paper we have shown that
+deciding
+the winner of a parity game on a graph of a pebble pushdown system 
+is nonelementary.
+
+	For the nonelementary lower bound
+we reduced the $\FO$ satisfiability problem on words \---- known to be nonelementary
+from \cite{Sto74} \---- to the
+problem of deciding whether player $0$ has a winning strategy from some starting configuration in 
+a pebble pushdown parity game.
+
+	For the decidability upper bound, we  used pebble alternating two-way tree automata to solve pebble pushdown parity games.
+Alternating two-way tree automata (ATWA) with parity winning conditions have been used to solve parity games on pushdown graphs \cite{cachat2002two}.
+The technique was first developped to solve model-checking problem for pushdown graphs in \cite{kupferman2000automata}.
+We showed that, when pushdown graphs and ATWA are extended with a set of pebbles to manage, the technique
+also works.
+We then proved decidability 
+% for checking that an $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\{0,1\}^*$ 
+by %reducing it
+a reduction
+ to the problem of model-checking $\MSO$ on a well chosen structure. To ensure decidability of model-checking $\MSO$ on the chosen structure, we made use of 
+Muchnik's Theorem: the structure is constructed by applying a successive number of times Muchnik's iteration to a structure $\mathfrak{A}_0$ for which model-checking $\MSO$ is decidable.
+%  After providing this structure, we will show that  checking that an $n$-PATWA $\mathcal{A}$ accepts the full binary tree $\{0,1\}^*$ reduces itself to the problem of model-checking $\MSO$ on it.
+
+\bibliography{bibliography/bib}
+
+\appendix
+
 \end{document}

### 2023-06-12 reachability/article.tex
up

@@
 
+		\iffalse
+
 \subsection{PPDA reachability}
@@
  $\mathcal{Z}$?\newline}
+
+		\fi
 
@@
 the parameter valuation arena
-$A_{P_0,P_1,\Gamma^*}$, as see on page~\pageref{parametric arena},
+$A_{P_0,P_1,\Gamma^*}$, as seen on page~\pageref{parametric arena},
 and that moreover contains the configurations and transitions of
@@
 \par\noindent\ignorespacesafterend 
-Recall that for \iffalse a partial function $f$ from a set $S$ to a set $X$ 
-%is a function defined on a subset $C$ of $S$ (possibly $S$ itself) with output values in $X$, and 
-that is defined on a subset $C \subseteq S$
-is denoted by
-$ f : S \rightharpoonup X $.
-
-For \fi a partial function $ f : S \rightharpoonup X $ (see page~\pageref{partial}), for notational purposes, we consider some element $\bot_X \not\in X$ and 
+Recall that for a partial function $ f : S \rightharpoonup X $% (see page~\pageref{partial})
+, for notational purposes, we consider some element $\bot_X \not\in X$ and 
  associate $f$ with the function returning the 

### 2023-06-12 reachability/parameters.tex
up

 \chapter{Parametric systems}\label{chapter addition of parameters}
% \chapter{Addition of parameters}

The infinite-state systems defined in the precedent chapter are widely used to model the behavior of computer programs. Essentially these approaches however address only concrete specifications.
For instance, one can ask, given a program, whether a certain set of states is reachable without some counter exceeding a threshold value $2048$, or one can ask whether a certain set of states is reachable after having more than $10$ time units elapse.
For certain types of computer programs however, such as embedded systems, the constraints depend on some environment, and concrete constraints are useful only when restricting oneself to a given concrete environment. 
In real life, it thus makes sense to rather ask, for instance, whether there exists a threshold value $t$ such that a certain set of states is reachable without some counter exceeding $t$, or to ask whether for all threshold $t$ a certain set of states is reachable after having more than $t$ time units elapse.
Hence the need to study under-specified systems, 
and parametric constraints. 
% For instance, one can wish to develop a model which allow to check that \textcolor{red}{an example: “message delivery within the time it takes to execute two assignment statements”}.

\iffalse
when studying the litterature on real time protocols, one sees that the desired timing properties for protocols are almost invariably parametric (cf [30,10,33]) because concrete timing constraints only make sense in the context of a given concrete environment

[10] H. Attiya, C. Dwork, N. Lynch and L. Stockmeyer. Bounds on the time to reach agreement in the presence of timing uncertainty. In Proc. 23 ACM STOC, 1991

[30] R. Strong, D. DOlev and F. Cristian. New latency bounds for atomic broadcast. In Proc. 11th IEEE RTSS, 1990.

[33] H. Weingberg and L. Zuck. Timed Ethernet: real-time formal specification of Ethernet. In Proc. 3r CONCUR, LNCS 630. Springer, 1992.
\fi
% Unfortunately, classical timed automata, where clocks can only be compared to constants, do not allow such a parametric reasoning.

In order to 
understand the behavior of such under-specified systems, Alur, Henzinger and Vardi have introduced parametric timed automata in their seminal paper \cite{AHV93-stoc}.
In addition to expressing concrete timing constraints, these can employ parametric constraints. Verification of the desired behavior of the system is then performed without concrete values. 
Their model laid the foundations for parametric reasoning about real time. 
% Parametric counter automata, defined in \cite{AHV93-stoc, annichini2000symbolic}, are similar to parametric timed automata. 
With similar considerations, it has also been natural to consider extensions of one-counter automata allowing updates that increase or decrease the counter by integer parameters \cite{IJTW93-icalp, BIL06-icalp}.

\iffalse
\cite{BIL06-icalp} Flat Parametric Counter Automata
\cite{IJTW93-icalp} New decidability results concerning two-way counter machines and applications. 
\fi
\iffalse
Work closely related to reachability in parametric one-counter automata is that of
Ibarra et al. [65], which shows decidability of reachability for a subset of the class of
deterministic parametric one-counter automata with sign tests. The decidability of
reachability over the whole class of such automata is stated as an open problem in
[65]. Note that although in this thesis parametric one-counter automata do not allow
negative counter values and sign tests, they allow non-determinism. Thus, the results
of this thesis are incomparable to those from [65].

[65] Oscar H. Ibarra, Tao Jiang, Nicholas Q. Trân, and Hui Wang. New decidability
results concerning two-way counter machines. SIAM Journal on Computing,
24(1):123–137, 1995.
\fi

In this thesis, we consider the problem of analysing parametric pushdown, one-counter and timed automata, i.e. automata extended with a stack, counter or clocks that can be compared with parameters.
These parameters can take unspecified values over infinite domains, which here consist in the set of words (for parametric pushdown automata) or the non-negative integers (for parametric one-counter automata and parametric timed automata). 
%
%
%
%
An {\em $\M$-parameter valuation} is then a function $\mu$ from a set $P$ of parameters to a 
% set $\Gamma^*$ of words
given set $\M$. We will omit the name of the set $\M$ in case it is obvious from context.
%
%
%
% or to the set $\N$ of non-negative integers, which can be seen as words over a unary alphabet.
Our interests lies in the parametric variants of 
%the problems of specification reachability games, parity games and emptyness. 
requirements such as safety or liveness properties: for instance, 
%to verify that the  model satisfies some safety or liveness property for at least one possible values of the parameters,
finding constraints on the parameters defining the set of all possible values for which the system satisfies a safety or liveness property, or verifying that the  
model satisfies some safety or liveness property for all possible values of the parameters.
% The latter problem can be solved, in the case of safety properties, as a parametric reachability problem.

\section{Parametric reachability}

The {\em parametric reachability problem} for an automaton with parameters (e.g. for a PTA) asks whether there exists a parameter valuation
$\mu$ such that 
a final configuration of the resulting concrete automaton becomes
reachable.
From here onwards, we omit ``parametric'' and
write simply ``reachability problem'' when it is clear from context that the input automaton has parameters.
Solving such a problem can be useful to verify safety
properties of automaton with parameters, i.e. to answer the question ``is a bad state reachable for some valuation
of the parameters''?

We are interested in how adding a set of parameters influences decidability and complexity properties. 
In particular, notice that the effects of adding a set of parameters depend on the model. Indeed, the reachability problem for parametric timed automata is
long known to be undecidable~\cite{AHV93-stoc}, whereas the reachability problem for timed automata is
$\PSPACE$-complete \cite{alur1994theory}. On the other hand,
reachability for succinct one-counter automata %without non-zero tests 
is $\NP$-complete \cite{HKOW-09concur}
and
reachability for parametric %succinct 
one-counter automata %without non-zero tests
 is both $\NP$-hard and in $\NEXP$ \cite{haase2012complexity}.

\iffalse
In the 1990’s timed automata have been introduced by Alur and Dill [?]. They extend
finite automata by clocks that can be compared against integer constants and provide a popular
formalism to reason about the behavior of real-time systems with desirable algorithmic properties;
For a more general means to specify the behavior of under-specified systems, such as embedded
systems, Alur, Henzinger and Vardi [?] have introduced parametric timed automata (PTA).

for many applications of counter machines, including
modelling computer programs, it is natural to consider more general types of up-
dates, such as adding integer constants to a counter [3, 5, 16] or adding integer
parameters [4, 12]. Parametric automata are used in various synthesis problems,
and to model open programs, whose behaviour depends on values input from
the environment [2]. In [20] parameters are also used to model resources (e.g.,
time, memory, dollars) consumed by transitions. The reachability problem for
parametric counter automata asks whether there exist values of the parameters
such that a given configuration is reachable from another given configuration.
\fi

%\subsubsection{Parameter valuation game}
\section{Parametric games}

As noted precedently, 
several model checking problems can be
expressed as decision problems for games.
For instance 
\textcolor{black}{
as mentioned 
in the case of pushdown automata, 
}
the modal $\mu$-calculus
model checking problem is polynomially equivalent to solving the parity game problem.
% \textcolor{red}{Since the model checking problem has a more varied range of applications than the reachability problem (mb have a source for that or smtg)},
S%ince model checking is typically more involved than simply solving the reachability problem, and s
ince we are interested in parametric model checking and specification,
we are naturally interested in parametric variants of reachability games and parity games.

% finding constraints on the parameters defining the set of all possible values for which the system satisfies a safety or liveness property, or verifying that the model satisfies some safety or liveness property for all possible values of the parameters.

One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a 
reachability game or a parity game
played on the arena induced by the
resulting concrete automaton
and some partition of its set of states. 
For determining such a set of parameter valuations, we incorporate the choice of the parameter valuation to the game. Each player then wants the parameter valuation to lead to an arena where it has a winning strategy.

In keeping with the spirit of alternation between players choosing successors in the arena, we consider here games where players
alternate choosing values for the parameters before alternating choosing successor configurations in the game on the induced arena.
For simplicity's sake, the values are assigned to the parameters one by one, alternating between values assigned by player $0$ and by player $1$.

A winning strategy for player $0$ allows to synthesize a controller that restricts the environment and ensures that {the property expressed by the winning condition always holds.}

\paragraph{Parameter valuation arenas}

We introduce here arenas over the set of partial $\M$-parameter valuations for a set of parameters
$P = P_0 \biguplus P_1$. The values are assigned to the parameters one by one, alternating between values assigned by player $0$ (corresponding to parameters in $P_0$) and by player $1$ (corresponding to parameters in $P_1$). Other options could have been considered: for instance, we could consider a fully existential approach 
(i.e. all parameters have their values assigned by player $0$), 
or a fully universal approach
(i.e. all parameters have their values assigned by player $1$). 
\iffalse
Note however that the alternating approach considered here encompasses both of the latter approaches: it suffices to include in the set $P$ parameters which do not appear in the rules of an automaton \---- or that appear only in rules that do not have any effect on whether or not a player has a winning strategy in reachability games or parity games played on resulting concrete automata \---- and to have a player assign values only to these ``unused'' parameters.  \\
\fi
Note however that the approach considered here encompasses both of the latter approaches:
the fully existential approach consists in the case where $P_1 = \emptyset$ and
the fully universal one, the case $P_0 = \emptyset$.

\iffalse
Given an ordering of the parameters $P = \{p_0, p_1, \ldots, p_k \}$ for some 
%\textcolor{red}{even}
 $k \in \N$, we partition the set of parameters into parameters with even numerotation, assigned to player $0$, and parameters with odd numerotation, assigned to player $1$.  
\fi

Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
%\textcolor{red}{even}
 $k \in \N$.
%
Then given a set $\M$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\M} = (S_0, S_1, \rightarrow_P)$ where
\begin{itemize}

\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,

\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_1$,

\iffalse
\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i+1}\}$ 
for some $i \in [0, \lceil k/2 \rceil - 1]$, plus the partial parameter valuation with domain $\emptyset$,
\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i}\}$
for some $i \in [0, \lfloor k/2 \rfloor]$, and
\fi

\item
$ \mu \rightarrow_P \mu'$,
if $\mu' \in (\M)^{\{p_0, \ldots, p_j\}}$ for some $j \in [0,k]$,
$\text{Dom}(\mu) = \text{Dom}(\mu') \setminus \{p_j\}$
and
 $\mu'(p) = \mu(p)$ for $p \in \text{Dom}(\mu)$. \label{parametric arena}
\end{itemize}

Note that the arena has dead ends, and that these are the configurations that correspond to parameter valuations.
Given a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1 $ for player $1$,
the last configuration of the resulting play $\pi(\mu_\bot,\sigma_0, \sigma_1)$
is called {\em the resulting parameter valuation}
 and is denoted by $\mu_{\sigma_0, \sigma_1}$, where 
 $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
 %We do not really consider winning conditions for parameter valuation games. Rather, once the resulting parameter valuation is produced, it serves to instantiate the parametric transition system for the next game.

Parametric valuation arenas are used as prefixes in parametric games. Once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
%
%
Indeed, where a concrete automaton induces a transition system, a parametric automaton induces one transition system for every possible parameter valuation.
We use $T^\mu_\A$ to denote the transition system induced by the concrete automaton corresponding to
a parametric automaton $\A$ with parameters $P$ taking values in $\M$ and a parameter valuation $\mu: P \to \M$.

\iffalse
Then for a parametric automaton, and given a partition of its states,

Thus, for an  $n$-parametric pushdown automaton 
$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
given a partition of
$Q$ into $Q_0$ and $Q_1$, 
we naturally partition the configurations of
%the transition system of 
 $\mathcal{Z}$
% $T^\mu_{\mathcal{Z}}$
into
$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
and
$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.

With these notations in mind one can define the arena
$$
A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
=
(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
parameter valuation $\mu: P\rightarrow \Gamma^*$.

\fi

\begin{center}
	\begin{figure}
		\hspace{2cm}
\includegraphics[width=0.7\textwidth]{figures/parameter_arena}
	\caption{An illustration of the arena for a parametric automaton $\A$ with 
%	two parameters, $p_0$ and $p_1$,	
	$P_0 = \{p_0\}$, $P_1 = \{p_1\}$ and
%	 that can take values in $[0,N]$ for some $N \in \N$. 
	$\M = [0,N]$ for some $N\in \N$.
	 Here $\mu_\bot$ is 
	%the partial parameter assignment with domain $\emptyset$,
	the totally undefined function and can be viewed rather as 
	the function that assigns $\bot$ to both parameters. For $i,j \in [0,N]$, $\mu_i$ is the partial parameter assignment with domain $\{p_0\}$ that maps $p_0$ to $i$ and $\mu_{i,j}$ is the parameter assignment that maps $p_0$ to $i$ and $p_1$ to $j$.
	 %Lastly $T^\mu_\A$ denotes the transition system induced by the concrete automaton corresponding to $\A$ with parameter valuation $\mu$.
	  As before, the circles belong to player $0$ and the diamonds to player $1$.} 
	\label{parametric_arena}
	\end{figure}
\end{center}

Given a parametric automaton $\A$ with a set of parameters $P = P_0 \biguplus P_1$ that can take 
values in the set $\M$,
a {\em parametric game} for $\A$ is a game played 
on the %following  arena $A_{\A}$.
%The 
arena $A_\A$
that include both 
% the configurations and transitions of
the parameter valuation arena
$A_{P_0, P_1,\M}$
and, % moreover contains 
for all parameter valuations $\mu: P\rightarrow \M$,
%the configurations and transitions of
the transition system %s 
% induced by the concrete automata corresponding to $\A$ with parameter valuation $\mu$
 $T_\A^\mu$
\---- albeit with configurations additionally indexed by $\mu$ %in the tuple
 to avoid confusion.
Moreover, for every 
dead end $\mu$
 in $A_{P_0,P_1,\M}$,
we add a transition
from the configuration $\mu \in \M^P$ to
$T_\A^\mu$.
See Figure~\ref{parametric_arena} for an illustration of such an arena.
Solving parametric games on an automaton with parameters (e.g. a POCA) then consists in answering whether there exists 
a winning strategy for player $0$
from the initial configuration of the parametric game.
Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.

\iffalse
Given a 
priority function $\Omega: Q \to [0, m]$,
we consider the function
$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$
where we set
$\widehat{\Omega}(q, w, \mu) = \Omega(q)$
for all				$w \in \Gamma^*$
and				for all $\mu \in (\Gamma^*)^{P}$,
and				$\widehat{\Omega}(\mu)=0$ for all $\mu \in (\Gamma^* \cup \bot)^{P}$.
\fi

### 2023-06-12 reachability/article.tex
up git p

@@
 
-\subsection{Games}
-
-% Games represent a model for interaction between decision makers.
-We study so-called turn-based games in extensive form with perfect information where the players have to make decisions again and again.
-\iffalse
-After its foundation by von Neumann and Morgenstern [vNM44] in the
-1940s, game theory has quickly become a valuable tool for modelling the
-interaction of several agents with individual and possibly conflicting objec-
-tives. It has applications in many scientific fields such as economy, sociology,
-biology, logic and computer science. Whereas the games used in the former
-three areas are usually finite in the sense that any play of these games ends
-after a finite number of steps (in fact often after only a single step), the
-games used in logic and computer science are in general infinite.
-
-@article{copeland1945john,
-  title={John von Neumann and Oskar Morgenstern, theory of games and economic behavior},
-  author={Copeland, Arthur H},
-  journal={Bulletin of the American Mathematical Society},
-  volume={51},
-  number={7},
-  pages={498--504},
-  year={1945},
-  publisher={American Mathematical Society}
-}
-
-More precisely, an infinite game is played for an infinite number of
-rounds. In each round, one player chooses an action depending on the se-
-quence of actions chosen in previous rounds. After infinitely many rounds,
-an infinite sequence of actions has emerged, called the outcome of the game.
-Each player receives a payoff determined by a payoff function mapping each
-possible outcome to a real number in the interval [0, 1]. We will concen-
-trate on games where the possible payoffs are just 0 and 1, in which case a
-winning condition, given as an abstract set of possible outcomes, is used to
-determine whether a player wins (payoff 0) or loses (payoff 1) an outcome.
-In logic, games are used to evaluate a logical formula in a mathematical
-structure and to compare mathematical structures with respect to their
-(in)distinguishability by a certain logic. Games of the former kind are called
-model checking games; games of the latter kind are called model comparison
-games. Whereas model checking and model comparison games for plain
-modal or first-order logic are finite games, infinite games arise as model
-checking games for fixed point logics and as model comparison games for
-infinitary logics.
-In computer science, games are used for the verification and synthesis
-of reactive systems, i.e. systems that interact with their environment. As
-soon as the systems under consideration do not necessarily terminate, these
-games become infinite. For example, the behaviour of an operating system
-can be understood as an infinite game between the system and its users.
-
-So far, the games used in logic and computer science have usually been
-two-player zero-sum games, i.e. games with only two players where one
-player wins if and only if the other player loses. For example, in a model
-checking game, one player wants to show that the given formula holds in the
-given structure whereas the other player wants to show that this is not the
-case. Analogously, in a game used for verification, one player wants to show
-that the given system is able to react against its environment in such a way
-that the resulting execution path of the system fulfils the given specification
-whereas the other player wants to show that this is not the case.
-
-The research on infinite games in theoretical computer science is based on a
-mixture of several motivations:
-1. The beautiful classical theory of infinite games, as developed in descriptive
-set theory, lacks an algorithmic content. Such an algorithmic orientation is
-provided by the automata theoretic approach to infinite games, originating
-in work of Church, B ̈uchi, McNaughton, and Rabin about fourty years ago.
-2. Determinacy results for infinite games are closely related to complementation
-results for logics and automata; thus, infinite games help to analyze logical
-theories (the most prominent being the monadic theory S2S of two successor
-functions, see e.g. [Th97]).
-3. Games are a natural model of reactive computation, and infinite games are
-thus a faithful representation of nonterminating reactive systems (for which
-control problems can be solved in terms of providing winning strategies).
-4. The model checking problem for logics like the μ-calculus can be formulated
-as the question to determine the winner of an infinite game.
-The purpose of this tutorial is to explain the core of the algorithmic (and
-automata theoretic) theory of infinite games. In the present extended abstract,
-only some key notions are explained, without technical details and proofs. In the
-tutorial, more will be said about topics of current research as listed in Section 6
-of this abstract.
-
-@incollection{thomas1997languages,
-  title={Languages, automata, and logic},
-  author={Thomas, Wolfgang},
-  booktitle={Handbook of formal languages},
-  pages={389--455},
-  year={1997},
-  publisher={Springer}
-}
-
-\fi
-
-A game is composed of an arena and a winning condition.
-\iffalse
-\mh{Two-player games oﬀer a very convenient framework to represent
-interaction of a program with some (possibly hostile) environment. In this approach, the ﬁrst player represents the program while the second player simulates
-the environment. A winning strategy expresses a property that must hold whatever the environment does, thus ﬁnding a winning strategy for the ﬁrst player
-allows to synthesize a controller that restricts the program and ensures that the
-property expressed by the winning condition always holds \cite{arnold2003games}.}
-\fi
+%ù \subsection{Games}
+
+A {\em game} is composed of an arena and a winning condition.
 We will ﬁrst study
@@
 
- \subsubsection{Arenas}
+% \subsubsection{Arenas}
 
@@
 	 ${\rightarrow} \subseteq S \times %\Lambda \times 
-	 				S$. \\
+	 				S$. 
 \end{itemize}
@@
 
- \subsubsection{Plays}
+%  \subsubsection{Plays}
 
 \noindent
-%
-%\noindent
 A {\em play} in an arena $A=(S_0,S_1,\rightarrow)$ is a path in $(S,{\rightarrow})$ that is {\em maximal} in the following sense: it is either infinite or finite and if it is finite, then the target configuration is a dead end.
@@
 
-% \subsubsection{Strategies}
-\subsection{Strategies}
-
-% Before deﬁning formally what it means for a player to win a game, we need to introduce the notion of strategy.
+% \subsection{Strategies}
 
@@
 
-%\noindent
-%Note that the above definitions are essentially the same for finite and infinite arenas.
-
- \subsubsection{Winning conditions}
+%  \subsubsection{Winning conditions}
 
@@
 
-% \paragraph{Reachability Games}
-\subsection{Reachability Games}
+% \subsection{Reachability Games}
 
@@
 
-With regards to reachability games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline
+With regards to reachability games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem.
 
@@
 
-% \paragraph{Parity Games}
-\subsection{Parity Games}
+% \subsection{Parity Games}
 
@@
 
-With regards to parity games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem. \newline
+With regards to parity games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem.
 
@@
 parameter valuation arena.
-Players take turns chosing particular stack contents on which to place pebbles,
-using a gadget similar as the one from Figure~\ref{quantifiers}. 
+Players take turns chosing particular stack contents on which to place pebbles. 
 Once every parameter has been assigned with the dropping of a corresponding pebble, 
@@
 Concerning higher-order pushdown automata, we are interested in the following decision problem.
-% \newpage
 

### 2023-06-12 reachability/article.tex
up

@@
 % sequence of configurations $s_{init} = s_0 s_1 \ldots \in S^\infty$ connected by the transition relation, i.e. such that $(v_i, v_{i+1}) \in T$.
-prefix of a play in $A$.\\
-%finite path in $(S,{\rightarrow})$.\\
-
-\noindent
+prefix of a play in $A$.
 Essentially plays can be seen as this: the two players, player $0$ and player $1$, move along the labeled transition system, taking turns either inﬁnitely often or until a dead end is reached.
@@
 that can finitely be described.
-
-\iffalse
-We consider here several types of winning conditions. 
-Essentially the 
- mapping $\Omega$ of an arena serves to colour the configurations of the labeled transition system.  Winning conditions then
- consist in acceptance conditions on colour sequences.
- \fi
 
@@
 every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.
-
-\iffalse
-This definition of winning strategies is independent from the winning condition itself, and will also apply to %B\"uchi games and min-
-parity games. To differentiate, we mention the winning conditions of a winning strategy from $s_{init}$.
-\fi
 
@@
 \end{samepage}
-
-%
-\iffalse
-Solving a parity game played on a finite graph means deciding, for a given starting position $v_{init}$, which of the two players has a winning strategy.
-It has been shown that this problem is in $\NP$ and
-co-$\NP$, more precisely $\UP$ and co-$\UP$. 
-% Marcin Jurdziński (1998), "Deciding the winner in parity games is in UP∩ co-UP" (PDF), Information Processing Letters, Elsevier, 68 (3): 119–124, doi:10.1016/S0020-0190(98)00150-1
-% @article{jurdzinski1998deciding,
-%  title={Deciding the winner in parity games is in UP∩ co-UP},
-%  author={Jurdzi{\'n}ski, Marcin},
-%  journal={Information Processing Letters},
-%  volume={68},
-%  number={3},
-%  pages={119--124},
-%  year={1998},
-%  publisher={Elsevier}
-%}
-\fi
-%
-
-% Muller condition: Inf() \in F
-
-% Rabin condition :
-
-% Streett Condition
-
-% Rabin Chain condition
-
-% max parity condition
-
-% 1-winning  == reachability
 
@@
 
-% \mh{Say something abut positional determinacy}
 One important property of parity games is that of {\em memoryless determinacy}: for a parity game
@@
 Formally, a {\em parametric pushdown automaton} ({\em PPDA} for short) 
-is a tuple 
+is a tuple \\
 $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
@@
 
+As noted precedently, 
+several model checking problems can be
+expressed as decision problems for games.
+For instance 
+\textcolor{black}{
+as mentioned 
+in the case of pushdown automata, 
+}
+the modal $\mu$-calculus
+model checking problem is polynomially equivalent to solving the parity game problem.
+% \textcolor{red}{Since the model checking problem has a more varied range of applications than the reachability problem (mb have a source for that or smtg)},
+S%ince model checking is typically more involved than simply solving the reachability problem, and s
+ince we are interested in parametric model checking and specification,
+we are naturally interested in parametric variants of reachability games and parity games.
+
+% finding constraints on the parameters defining the set of all possible values for which the system satisfies a safety or liveness property, or verifying that the model satisfies some safety or liveness property for all possible values of the parameters.
+
+One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a 
+reachability game or a parity game
+played on the arena induced by the
+resulting concrete automaton
+and some partition of its set of states. 
+For determining such a set of parameter valuations, we incorporate the choice of the parameter valuation to the game.
+Each player then wants the parameter valuation to lead to an arena where it has a winning strategy.
+
+In keeping with the spirit of alternation between players choosing successors in the arena, we consider here games where players
+alternate choosing values for the parameters before alternating choosing successor configurations in the game on the induced arena.
+For simplicity's sake, the values are assigned to the parameters one by one, alternating between values assigned by player $0$ and by player $1$.
+
+A winning strategy for player $0$ allows to synthesize a controller that restricts the environment and ensures that {the property expressed by the winning condition always holds.}
+
+\paragraph{Parameter valuation arenas}
+
+We introduce here arenas over the set of partial $\M$-parameter valuations for a set of parameters
+$P = P_0 \biguplus P_1$.
+The values are assigned to the parameters one by one, alternating between values assigned by player $0$ (corresponding to parameters in $P_0$) and by player $1$ (corresponding to parameters in $P_1$).
+Other options could have been considered: for instance, we could consider a fully existential approach
+(i.e. all parameters have their values assigned by player $0$), 
+or a fully universal approach
+(i.e. all parameters have their values assigned by player $1$). 
+\iffalse
+Note however that the alternating approach considered here encompasses both of the latter approaches: it suffices to include in the set $P$ parameters which do not appear in the rules of an automaton \---- or that appear only in rules that do not have any effect on whether or not a player has a winning strategy in reachability games or parity games played on resulting concrete automata \---- and to have a player assign values only to these ``unused'' parameters.
+\\
+\fi
+Note however that the approach considered here encompasses both of the latter approaches:
+the fully existential approach consists in the case where $P_1 = \emptyset$ and
+the fully universal one, the case $P_0 = \emptyset$.
+
+\iffalse
+Given an ordering of the parameters $P = \{p_0, p_1, \ldots, p_k \}$ for some 
+%\textcolor{red}{even}
+ $k \in \N$, we partition the set of parameters into parameters with even numerotation, assigned to player $0$, and parameters with odd numerotation, assigned to player $1$.  
+\fi
+
+Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
+%\textcolor{red}{even}
+ $k \in \N$.
+%
+Then given a set $\M$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\M} = (S_0, S_1, \rightarrow_P)$ where
+\begin{itemize}
+
+\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,
+
+\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_1$,
+
+\iffalse
+\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i+1}\}$ 
+for some $i \in [0, \lceil k/2 \rceil - 1]$, plus the partial parameter valuation with domain $\emptyset$,
+\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i}\}$
+for some $i \in [0, \lfloor k/2 \rfloor]$, and
+\fi
+
+\item
+$ \mu \rightarrow_P \mu'$,
+if $\mu' \in (\M)^{\{p_0, \ldots, p_j\}}$ for some $j \in [0,k]$,
+$\text{Dom}(\mu) = \text{Dom}(\mu') \setminus \{p_j\}$
+and
+ $\mu'(p) = \mu(p)$ for $p \in \text{Dom}(\mu)$. \label{parametric arena}
+\end{itemize}
+
+Note that the arena has dead ends, and that these are the configurations that correspond to parameter valuations.
+Given a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1 $ for player $1$,
+the last configuration of the resulting play $\pi(\mu_\bot,\sigma_0, \sigma_1)$
+is called {\em the resulting parameter valuation}
+ and is denoted by $\mu_{\sigma_0, \sigma_1}$, where 
+ $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
+ %We do not really consider winning conditions for parameter valuation games.
+Rather, once the resulting parameter valuation is produced, it serves to instantiate the parametric transition system for the next game.
+
+Parametric valuation arenas are used as prefixes in parametric games.
+Once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
+%
+%
+Indeed, where a concrete automaton induces a transition system, a parametric automaton induces one transition system for every possible parameter valuation.
+We use $T^\mu_\A$ to denote the transition system induced by the concrete automaton corresponding to
+a parametric automaton $\A$ with parameters $P$ taking values in $\M$ and a parameter valuation $\mu: P \to \M$.
+
+\iffalse
+Then for a parametric automaton, and given a partition of its states,
+
+Thus, for an  $n$-parametric pushdown automaton 
+$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
+given a partition of
+$Q$ into $Q_0$ and $Q_1$, 
+we naturally partition the configurations of
+%the transition system of 
+ $\mathcal{Z}$
+% $T^\mu_{\mathcal{Z}}$
+into
+$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
+and
+$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.
+
+With these notations in mind one can define the arena
+$$
+A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
+=
+(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
+induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
+parameter valuation $\mu: P\rightarrow \Gamma^*$.
+
+\fi
+
+\begin{center}
+	\begin{figure}
+		\hspace{2cm}
+\includegraphics[width=0.7\textwidth]{figures/parameter_arena}
+	\caption{An illustration of the arena for a parametric automaton $\A$ with 
+%	two parameters, $p_0$ and $p_1$,	
+	$P_0 = \{p_0\}$, $P_1 = \{p_1\}$ and
+%	 that can take values in $[0,N]$ for some $N \in \N$. 
+	$\M = [0,N]$ for some $N\in \N$.
+	 Here $\mu_\bot$ is 
+	%the partial parameter assignment with domain $\emptyset$,
+	the totally undefined function and can be viewed rather as 
+	the function that assigns $\bot$ to both parameters.
+For $i,j \in [0,N]$, $\mu_i$ is the partial parameter assignment with domain $\{p_0\}$ that maps $p_0$ to $i$ and $\mu_{i,j}$ is the parameter assignment that maps $p_0$ to $i$ and $p_1$ to $j$.
+	 %Lastly $T^\mu_\A$ denotes the transition system induced by the concrete automaton corresponding to $\A$ with parameter valuation $\mu$.
+	  As before, the circles belong to player $0$ and the diamonds to player $1$.} 
+	\label{parametric_arena}
+	\end{figure}
+\end{center}
+
+Given a parametric automaton $\A$ with a set of parameters $P = P_0 \biguplus P_1$ that can take 
+values in the set $\M$,
+a {\em parametric game} for $\A$ is a game played 
+on the %following  arena $A_{\A}$.
+%The 
+arena $A_\A$
+that include both 
+% the configurations and transitions of
+the parameter valuation arena
+$A_{P_0, P_1,\M}$
+and, % moreover contains 
+for all parameter valuations $\mu: P\rightarrow \M$,
+%the configurations and transitions of
+the transition system %s 
+% induced by the concrete automata corresponding to $\A$ with parameter valuation $\mu$
+ $T_\A^\mu$
+\---- albeit with configurations additionally indexed by $\mu$ %in the tuple
+ to avoid confusion.
+Moreover, for every 
+dead end $\mu$
+ in $A_{P_0,P_1,\M}$,
+we add a transition
+from the configuration $\mu \in \M^P$ to
+$T_\A^\mu$.
+See Figure~\ref{parametric_arena} for an illustration of such an arena.
+Solving parametric games on an automaton with parameters (e.g. a POCA) then consists in answering whether there exists 
+a winning strategy for player $0$
+from the initial configuration of the parametric game.
+Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.
+
 We are moreover interested in games played on parametric pushdown automata's  

### 2023-06-12 reachability/article.tex
up

@@
 \par\noindent\ignorespacesafterend
-Formally, a {\em parametric pushdown automaton} ({\em PPDA} for short) 
-is a tuple \\
-$\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
-\begin{itemize}
-\item $Q$ is a non-empty finite {\em set of  states},
-\item $\Gamma$ is a non-empty finite {\em  stack alphabet},
-\item $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$},
+A {\em parametric pushdown automaton} ({\em PPDA} for short) 
+is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
+ $Q$ is a non-empty finite {\em set of  states};
+ $\Gamma$ is a non-empty finite {\em  stack alphabet};
+ $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$};
 % \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Gamma^*$ is finite {\em  set of rules},
- \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules},
-\item $q_{init}\in Q$ is an {\em initial  state}, 
-\item $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
-\item $F\subseteq Q$ is a {\em set of final  states}.
-\end{itemize}
-
-\par\noindent\ignorespacesafterend
+  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules};
+ $q_{init}\in Q$ is an {\em initial  state};
+ $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
+ $F\subseteq Q$ is a {\em set of final  states}.
+
 The {\em size} of $\mathcal{Z}$ is defined as
@@
 $\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
-where
-\begin{itemize}
-\item $Q$ is a {\em finite set of  states},
-\item $\Gamma$ is a {\em finite stack alphabet},
-\item  $R  \subseteq  Q  \times \Gamma \times
+where $Q$ is a {\em finite set of  states}; $\Gamma$ is a {\em finite stack alphabet};
+  $R  \subseteq  Q  \times \Gamma \times
 		 \{ 0, 1, \ldots n\} \times \mathcal{P}(\{1,\ldots,n\})
@@
 		and if the last element is
-		$\text{\textit{lift}}$ then we additionally require $i \in S$,
-\item $q_{init} \in Q$ is an {\em initial  state}, 
-\item $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}, and
-\item $F \subseteq Q$ is a {\em set of final  states}.
-\end{itemize}
+		$\text{\textit{lift}}$ then we additionally require $i \in S$;
+ $q_{init} \in Q$ is an {\em initial  state};
+ $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
+ $F \subseteq Q$ is a {\em set of final  states}.
 

### 2023-06-12 reachability/article.tex
up

@@
 problem of deciding whether player $0$ has a winning strategy from some starting configuration in
-a pebble pushdown parity game.
+a parametric pushdown parity game.
 % problem of deciding winning configuration for pebble pushdown parity games.
@@
 We reduce the satisfiability problem for first-order logic on words
-to the problem of {\sc Solving pebble pushdown parity game}. 
+to the problem of {\sc Parametric pushdown parity game}. 
 We know from \cite{Sto74}
@@
 towards the problem of %following problem.
-{\sc Solving pebble pushdown parity game}.
+{\sc Parametric pushdown parity game}.
 
@@
 \begin{theorem}
-{\sc FO satisfiability on words} is polynomial time reducible to {\sc Solving pebble pushdown parity game}.
+{\sc FO satisfiability on words} is polynomial time reducible to {\sc Parametric pushdown parity game}.
 \end{theorem}
@@
 We construct
-a pebble pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
+a parametric pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
 union $Q_\phi = Q_0  \biguplus Q_1 $, a priority mapping $\Omega_\phi$, and an initial state $q_I \in Q_\phi$
@@
 
-Player $0$ starts by pushing a word $w$ onto the stack, then places the first pebble at the current position. The goal of this first pebble is to remember the word $\$w$, which corresponds to the 
-word “guessed” by player $0$.
-This pebble is meant to enforce that the stack remain a prefix of $\$w$, with each stack content of the play corresponding to a position in the word.
-To enforce this, once pebble $1$ is placed, we allow in $\mathcal{Z}_\phi$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position on the word.
-A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to pebble $1$.
-
-Then players alternate placing pebbles on prefixes of $\$w$. Universal variables correspond to pebbles placed by player $1$ and existential variables to pebbles placed by player $0$. 
-Once players have chosen values for their respective variables, the pebble assignment $f$ is fixed. It is then time
+Player $0$ starts by chosing a valuation $w$ for the first parameter (parameter $p_0$). The goal of this first parameter is to remember
+the word $w$, which corresponds to the word ``guessed" by player $0$. 
+This parameter is meant in part to enforce that 
+the stack remains a prefix of $w$, with each stack content of the play corresponding to a position in the word.
+ To enforce this, we allow in $\mathcal{Z}$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position in the word.
+  A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to the value of parameter $p_0$. 
+ The gadgets for these intermediary potential challenges will be excluded from the illustrations to preserve clarity of representation.
+The other parameters are meant to correspond to the variables, with player $1$ parameters corresponding to universal variables and player $0$ parameters corresponding to existential variables.
+
+Once players have chosen values for their respective variables, the parameter assignment
+$\mu$ is fixed. It is then time
 for player $0$ to ensure that
-$w$ satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
-$ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
+%$\mathfrak{W}$ 
+$w$
+satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
+$ |\mu(p_1)|-1, \ldots, |\mu(p_k)|-1$ respectively. This is the goal of the reachability game.
+
+The first step of the game is to check that the values assigned to the parameters correspond indeed to possible variables, that is, that the values of the parameters are prefixes of $w$.
+
+Recall $\psi$ is quantifier-free and in disjunctive normal form, i.e.
+ $$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).$$
 
@@
 
+Now we need only to describe the gadgets $ \C_{\psi_{i,j}}$. 
 The gadget depends on the type of the atomic formula.
-For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then check that the the top of the stack
+For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then checks that the top of the stack
 $x$ is the right letter. For one checking that $x = x'$ where $x$ and $x'$ are both variables,
-player $0$ goes to the position corresponding to $x$ and check against the presence of the pebble corresponding to $x'$ as well. If the formula is a negation, players exchange their roles.
+player $0$ goes to the position corresponding to $x$ and checks against the parameter corresponding to $x'$ as well. If the formula is a negation, players exchange their roles. 
 \end{proof}
@@
 \begin{theorem}
-
-The problem of {\sc Solving pebble pushdown parity game} is not in {\sc ELEMENTARY}.
-
+{\sc Parametric pushdown parity game} is not in {\sc ELEMENTARY}.
 \end{theorem}

### 2023-06-12 reachability/article.tex
up

@@
 
-As noted precedently, 
-several model checking problems can be
+Several model checking problems can be
 expressed as decision problems for games.
 For instance 
-\textcolor{black}{
-as mentioned 
 in the case of pushdown automata, 
-}
 the modal $\mu$-calculus
@@
 % \textcolor{red}{Since the model checking problem has a more varied range of applications than the reachability problem (mb have a source for that or smtg)},
-S%ince model checking is typically more involved than simply solving the reachability problem, and s
-ince we are interested in parametric model checking and specification,
+Since we are interested in parametric model checking and specification,
 we are naturally interested in parametric variants of reachability games and parity games.
-
-% finding constraints on the parameters defining the set of all possible values for which the system satisfies a safety or liveness property, or verifying that the model satisfies some safety or liveness property for all possible values of the parameters.
 
@@
 
-\paragraph{Parameter valuation arenas}
-
-We introduce here arenas over the set of partial $\M$-parameter valuations for a set of parameters
+% \subsection{Parameter valuation arenas}
+
+We introduce here arenas over the set of partial parameter valuations for a set of parameters
 $P = P_0 \biguplus P_1$.
@@
 (i.e. all parameters have their values assigned by player $1$). 
-\iffalse
-Note however that the alternating approach considered here encompasses both of the latter approaches: it suffices to include in the set $P$ parameters which do not appear in the rules of an automaton \---- or that appear only in rules that do not have any effect on whether or not a player has a winning strategy in reachability games or parity games played on resulting concrete automata \---- and to have a player assign values only to these ``unused'' parameters.
-\\
-\fi
 Note however that the approach considered here encompasses both of the latter approaches:
@@
 
-\iffalse
-Given an ordering of the parameters $P = \{p_0, p_1, \ldots, p_k \}$ for some 
-%\textcolor{red}{even}
- $k \in \N$, we partition the set of parameters into parameters with even numerotation, assigned to player $0$, and parameters with odd numerotation, assigned to player $1$.  
-\fi
-
 Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
@@
 %
-Then given a set $\M$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\M} = (S_0, S_1, \rightarrow_P)$ where
+Then given a alphabet $\Gamma$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\Gamma} = (S_0, S_1, \rightarrow_P)$ where
 \begin{itemize}
 
-\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,
-
-\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_1$,
-
-\iffalse
-\item $S_0$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i+1}\}$ 
-for some $i \in [0, \lceil k/2 \rceil - 1]$, plus the partial parameter valuation with domain $\emptyset$,
-\item $S_1$ is the set of all partial $\M$-parameter valuations with domain $\{p_0, p_1, \ldots, p_{2i}\}$
-for some $i \in [0, \lfloor k/2 \rfloor]$, and
-\fi
+\item $S_0$ is the set of all partial parameter valuations with images in $\Gamma^*$ and domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,
+
+\item $S_1$ is the set of all partial $\M$-parameter valuations  with images in $\Gamma^*$ and  domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_1$,
 
@@
 $ \mu \rightarrow_P \mu'$,
-if $\mu' \in (\M)^{\{p_0, \ldots, p_j\}}$ for some $j \in [0,k]$,
+if $\mu' \in (\Gamma^*)^{\{p_0, \ldots, p_j\}}$ for some $j \in [0,k]$,
 $\text{Dom}(\mu) = \text{Dom}(\mu') \setminus \{p_j\}$
@@
 Rather, once the resulting parameter valuation is produced, it serves to instantiate the parametric transition system for the next game.
-
-Parametric valuation arenas are used as prefixes in parametric games.
+Parametric valuation arenas can then be used as prefixes in parametric games.
 Once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
@@
 We use $T^\mu_\A$ to denote the transition system induced by the concrete automaton corresponding to
-a parametric automaton $\A$ with parameters $P$ taking values in $\M$ and a parameter valuation $\mu: P \to \M$.
+a parametric automaton $\A$ with parameters $P$ taking values in $\Gamma^*$ and a parameter valuation $\mu: P \to \Gamma^*$.
 
 \iffalse
-Then for a parametric automaton, and given a partition of its states,
-
-Thus, for an  $n$-parametric pushdown automaton 
-$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
-given a partition of
-$Q$ into $Q_0$ and $Q_1$, 
-we naturally partition the configurations of
-%the transition system of 
- $\mathcal{Z}$
-% $T^\mu_{\mathcal{Z}}$
-into
-$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
-and
-$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$.
-
-With these notations in mind one can define the arena
-$$
-A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
-=
-(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
-induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
-parameter valuation $\mu: P\rightarrow \Gamma^*$.
-
-\fi
-
 \begin{center}
@@
 \end{center}
+\fi
 
 Given a parametric automaton $\A$ with a set of parameters $P = P_0 \biguplus P_1$ that can take 
-values in the set $\M$,
+values in the set $\Gamma^*$,
 a {\em parametric game} for $\A$ is a game played 
@@
 the parameter valuation arena
-$A_{P_0, P_1,\M}$
+$A_{P_0, P_1,\Gamma}$
 and, % moreover contains 
-for all parameter valuations $\mu: P\rightarrow \M$,
+for all parameter valuations $\mu: P\rightarrow \Gamma^*$,
 %the configurations and transitions of
@@
 dead end $\mu$
- in $A_{P_0,P_1,\M}$,
+ in $A_{P_0,P_1,\Gamma}$,
 we add a transition
-from the configuration $\mu \in \M^P$ to
+from the configuration $\mu \in (\Gamma^*)^P$ to
 $T_\A^\mu$.
-See Figure~\ref{parametric_arena} for an illustration of such an arena.
-Solving parametric games on an automaton with parameters (e.g. a POCA) then consists in answering whether there exists 
+% See Figure~\ref{parametric_arena} for an illustration of such an arena.
+Solving parametric games on an automaton with parameters then consists in answering whether there exists 
 a winning strategy for player $0$
@@
 Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$. 
-
-We are moreover interested in games played on parametric pushdown automata's  
- transition systems.
 Recall that given a 
@@
 one needs only to provide %a mapping $\Omega$ and 
-a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\
+a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena.
 
@@
 With these notations in mind one can define the arena
-$$
+\[
 A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
 =
-(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$$
+(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})\]
 induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
@@
 
-Given additionally a partition of
+From a partition of
 $P$ into $P_0$ and $P_1$, 
-we define 
+we thus define 
 $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
@@
 the parameter valuation arena
-$A_{P_0,P_1,\Gamma^*}$, as seen on page~\pageref{parametric arena},
+$A_{P_0,P_1,\Gamma}$,
 and that moreover contains the configurations and transitions of
@@
 for all $\mu \in (\Gamma^* \cup \bot)^{P}$.
-
-% $T_{\mathcal{Z}} = (\Conf(\mathcal{Z}), \rightarrow_{\mathcal{Z},\mu})$
-
-\iffalse
-\textcolor{red}{
-Reachability %, B\"uchi
- and parity games on $\mathcal{G}_{(\mathcal{Z}, \mu,P_0, P_1, Q_0 , Q_1, \Omega)}$
-are then clearly defined. They have the flaw however of not taking into account the
-way the values assigned to the parameters is chosen. How values are assigned to
-the parameters is the subject of the next Section.
-}
-\fi
 

### 2023-06-12 reachability/article.tex
up

@@
 
+\iffalse
 Given a parametric automaton $\A$ with a set of parameters $P = P_0 \biguplus P_1$ that can take 
@@
 a {\em parametric game} for $\A$ is a game played 
-on the %following  arena $A_{\A}$.
-%The 
+on the 
 arena $A_\A$
 that include both 
-% the configurations and transitions of
 the parameter valuation arena
 $A_{P_0, P_1,\Gamma}$
-and, % moreover contains 
+and, %   
 for all parameter valuations $\mu: P\rightarrow \Gamma^*$,
-%the configurations and transitions of
 the transition system %s 
-% induced by the concrete automata corresponding to $\A$ with parameter valuation $\mu$
  $T_\A^\mu$
@@
 % See Figure~\ref{parametric_arena} for an illustration of such an arena.
-Solving parametric games on an automaton with parameters then consists in answering whether there exists 
-a winning strategy for player $0$
-from the initial configuration of the parametric game.
-Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$. 
+\fi
+
 Recall that given a 
@@
 a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. 
-
 Thus, for an  $n$-parametric pushdown automaton 
@@
 
-We are interested in the following games and problems.
-
 \problemx{Parametric pushdown reachability game}
@@
 
-\iffalse
-\problemx{Parametric pushdown reachability game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, a disjoint
-union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
-.}
-{Does player $0$ have a 
-strategy $\sigma_0$ for the parameter valuation game  such that
-for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
-player $0$ has a
-winning strategy from $q_{init}(\gamma_{init})$ for the reachability condition on
-$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
-
-\problemx{Parametric pushdown parity game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, a disjoint
-union $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
-, 
-a priority mapping $\Omega: Q \to \{0, \ldots m\}$%
-%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
-%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
-.}
-{Does player $0$ have a 
-strategy $\sigma_0$ for the parameter valuation game  such that
-for all player $1$ strategies $\sigma_1$ for the parameter valuation game, 
-player $0$ has a
-winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
-$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$ ?\newline}
-\fi
-
 We write {\sc $n$-parametric reachability parity game}
@@
  the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
+Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.
+We are interested in the following games and problems.
 

### 2023-06-13 abstract.tex
up

We investigate and expand upon the notion of resilience as defined in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state. We introduce a more general definition of resilience and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
and the $k$-(state-)resilience problem (kRP). The (general) resilience problem 
% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$. 
We remark that it is an instance of the home space problem defined in the framework of VASS.
State resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.

In a well-structured framework we show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility. Moreover, state-resilience,
bounded-state-resilience and
 $k$-state-resilience
are undecidable for strongly upward-compatible WSTS with effective pred-basis.

On the positive side, we find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$, and moreover the resilience problem is decidable for ideal-effective WSTS with 
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$
and
the additional hypothesis that
for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.

We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

Finally, we study the resilience for VASS, where most problems are shown decidable.

### 2023-06-13 introduction.tex
up

@@
 The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$.
-We remark that it is an instance of the home space problem defined in the framework of VASS in \cite{DBLP:journals/corr/abs-2207-02697}.
+We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697}.
 State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is ...
@@
 
-\item We generalize the main theorem of \cite{} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
+\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
  WSTS with effective 
@@
 
-\item We study the resilience for VASS and most of problems are shown decidable.
+\item We study the resilience for VASS where most of the problems are shown decidable.
 \end{itemize}

### 2023-06-13 abstract.tex
up

@@
-We investigate and expand upon the notion of resilience as defined in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
+We investigate and expand upon the notion of resilience as defined by Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
 We introduce a more general definition of resilience and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)

### 2023-06-13 abstract.tex
up

@@
 and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
-$\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
+$\uparrow$ $\post^*$ basis hypothesis leads to undecidability, but we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
 

### 2023-06-13 introduction.tex
up

@@
 Think for instance of a supply chain, or an airport’s air trafic control.
-In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack [pushed it towards some kind of ‘error state’] and, if it can, whether it can perform the return in a satisfactory timeframe.
+In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack
+pushed it towards some kind of ‘error state’ and, if it can, whether it can perform the return in a satisfactory timeframe.
 In the exemple of air trafic control, an accident i.e.
-for instance a delay in the arrival of a plance can lead to conflicts of resolution (i.e.
+for instance a delay in the arrival of a plane can lead to conflicts of resolution (i.e.
 landing) which lead to prompting planes to remain on standby.
 Since their fuel quantity is limited it is hence pretty important to be able to know that trafic will return to its normal pace quickly enough.
-This is the resilience problem.
-Resilience is the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states.
+We formalize this as the resilience problem.
+Resilience is then the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states. \\
 
@@
 
-{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
-Notification to authors: September 16, 2022}
+% {\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm) Notification to authors: September 16, 2022}
 
-trop d'hypotheses trop fortes sur les WSTS
+% trop d'hypotheses trop fortes sur les WSTS
 
+\noindent
 {\bf Our contributions}
-
 \begin{itemize}
-
 \item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
@@
 We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697}.
-State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is ...
+State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.
 
@@
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities.
+Nevertheless, we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities.
 and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
@@
 In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-All these papers consider particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed.
-
+All these papers are concerned particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed,
+even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
 \begin{center}
 	\begin{figure}
-			\hspace{0.75cm}
-\includegraphics[width=0.85\textwidth]{FigureB}
+			\hspace{0.74cm}
+\includegraphics[width=0.87\textwidth]{FigureB}
 	\caption{A channel system with three automata and four channels.
@@
 Consider the channel system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time. 
-
 The automata on the left and right represents the aircrafts attempting to land, while the middle one represent the control tower.

### 2023-06-13 resilience-safe-down.tex
up

--- 2023-06-09 resilience-safe-down.tex
@@
 
-Let us choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$.
-%			at least one counter is unbounded. 
-This set is upward-closed by (upward) compatibility in VASS.
+Let us now consider the case $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.
+It is of interest to note this case can be linked to the problem of mutual exclusion: indeed, one can choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$, and $\Safe$ to be the downward-closed complement of $\Bad$.
+In this case, our contribution consist in the following:
 
-As an example, let us choose $\Safe=S \setminus \Bad$. Then, since $\Bad=\uparrow \Bad$, we have: $\Safe=\downarrow \Safe$.
-
-%
-% \alain{definir  ideal-effective WSTS}
 %
@@
 
-\alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}
+% \alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}
+
+% \mathieu{
+% Home-state indécidable pour les minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
+% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.}
+
+Without the downward-compatible hypothesis, the problem becomes undecidable. Indeed, it is 
+undecidable for Reset 
+%Petri-nets
+VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
+This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.

### 2023-06-13 resilience-safe-up.tex
up

@@
 We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
-It can be viewed as a generalizeation/extension of coverability, namely,
+It can be viewed as a generalizeation
+ of coverability, namely,
 it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
@@
 
-As an
+\iffalse
+As a
 example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e.
@@
 This set of states is upward-closed and $\Safe=S \setminus \Bad=\uparrow \Safe$.\alain{mal rédigé}
+\fi
 
@@
 Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
-
 %	If $k <  \min(K,n)$, then we check every path of length smaller
@@
 becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}. 
-
 Coverability reduce
-to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$.
-Indeed $s \to^* \uparrow t$ 
+to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
+as $s \to^* \uparrow t$ 
 reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
 Hence, without the completion-post-effective hypothesis, 
-{\sc Resilience} is undecidable.
+{\sc Resilience} itself is undecidable.

### 2023-06-13 resilience.tex
up

--- 2023-06-09 resilience.tex
@@
 \section{Resilience for WSTS}
+
+In this section we introduce the concept of resilience, bounded-resilience and $k$-resilience for well structured transition systems. We show that the resilience problems are decidable
+for  completion-post-effective $\omega^2$-WSTS with strong compatibility
+in the case $\Bad$ is downward-closed and $\Safe $ is upward-closed,
+and that resilience is decidable for downward-compatible ideal-effective WSTS
+when $\Safe$ is downward-closed and $\Bad $ is upward-closed. 
+We show that the three resilience problems are however undecidable for WSTS with
+strong compatibility in general.
 
@@
 Surprinsingly, we have not found decidability results about the resilience problems for WSTS.
-
 %		
@@
 decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
-
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 
-{\bf Synthesis of the main results}\label{synthesis}
+{\bf Synthesis of the main decidability results}\label{synthesis}
 

### 2023-06-13 appendix.tex
up

@@
 \end{proof}
+
+\section{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+
+Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
+% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
 

### 2023-06-13 state-resilience.tex
up

@@
 %{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-
 %		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
-
 The three previous problems become:
@@
 Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
-
 As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
@@
 Reset VASS extend the basic VASS model with special “reset
-transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.\\
-
-% More Formally {\em reset ... }
-
+transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
 \begin{definition}
@@
 $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
-
-\begin{itemize}
-
+\begin{samepage}\begin{itemize}
 \item  $t = (q,add(\textbf{z}),q') \in T$
@@
 and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-
 \item $t = (q,reset(\gamma),q') \in T$ 
@@
 $\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize}
+\end{itemize} \end{samepage}
 
-\iffalse
-\problemx{$d$-VASS zero-reachability}
-{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
-{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\} 
-\fi
-
-It is well known that Reset $d$-VASS are WSTS [ref ? relire WSTS everywhere]. 
+It is well known that Reset $d$-VASS are WSTS \cite{dufourd1998reset}. 
 Moreover, we are interested in the 
@@
 \end{proof}
-
-\iffalse
-\alain{definir downward reachability problem.....
-downward-closed problem given a state $s$ of a WSTS
-% in the regarded class 
-with strong compatibility 
-and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. }
-\fi
 

### 2023-06-13 state-resilience.tex
up

@@
 
-It is well known that Reset $d$-VASS are WSTS \cite{dufourd1998reset}. 
+It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
 Moreover, we are interested in the 
 %following
- undecidable \cite{araki1976PN} decision problem%:
-\iffalse
-\alain{ce qui suit est inutile}
-
-\begin{samepage}
-\problemx{Zero-reachability}
-{A Reset $d$-VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
-{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
-\end{samepage}
-%
-\alain{ce qui précède est inutile}
-\fi
-~ of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
 Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
@@
 A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
-%\alain{mélange de s et (p,u)}
  $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
@@
 To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
-% $pre(t)=\{ u \in S \mid \exists u' \in S$ in a labeled WSTS such that $u \xrightarrow{t} u'$\} 
 $pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
@@
 as the following formula
-% \[ ~ \forall s \in S, ~ s_0 \rightarrow^* s \implies \exists s' \in U_t, ~ s \rightarrow^{*} s'\] 
 \[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
@@
 The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS, whose proof can be seen in Appendix~\ref{appendix}.
-
 \begin{proposition}\label{liveness reset}
@@
 $t$-liveness} is undecidable.
-
 Since {\sc %$Reset $d$-VASS 
@@
 $t$-liveness} is undecidable.
-
 % \begin{proposition}
 % \end{proof}
-
 % \begin{proposition}
@@
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
-
 \[ \post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} \cap \Bad \subseteq \pred^*(\Safe)\]
-% \alain{$\equiv $  ???}
-
 Since $B_{\uparrow \post^*(s)}$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
@@
 Results synthesis in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
-
 \begin{center}
@@
 
-Unfortunately the {\sc State-resilience} problem is undecidable for WSTS.
+Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by Reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
 
@@
 
-\begin{proof}
-\mathieu{Refaire la preuve en utilisant les Minsky machines, i.e. si on pet résoudre la question dans les reset alors on peut la résoudre dans les Minsky machine ... et ça y est indécidable.}
-\end{proof}
+% \begin{proof}
+% \mathieu{Refaire la preuve en utilisant les Minsky machines, i.e. si on pet résoudre la question dans les reset alors on peut la résoudre dans les Minsky machine ... et ça y est indécidable.}
+% \end{proof}
 

### 2023-06-13 application.tex
up

@@
 The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
-\alain{vérifier que la plus grande existe}
+%\alain{vérifier que la plus grande existe}
 Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
@@
 \begin{theorem}{}
-Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
+Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear.
 \end{theorem}
@@
 \begin{theorem}{}
-Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear sets.
+Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
 \end{theorem}
@@
 % Again let us without loss of generality hence let us consider only the case when $\Safe$ is semilinear and $\Bad = S \setminus \Safe$ is semilinear too.
-\mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
+% \mathieu{Expliciter davantage la preuve ? $\pred^*(\Safe)$ semilinéaire car définissable en Presburger ? Puis inclusion de semilinéaires décidables.}
 \end{proof}
@@
 
+\iffalse
 \begin{figure}
-
 \begin{center}
@@
 \end{figure}
+\fi
 
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 \end{proposition}
@@
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
-
 The set $\Bad$ admits a finite basis $B_\Bad$.
@@
 and thus, are in $\Bad$, by upward-closedness.
-
 Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
@@
 
-\iffalse
-\mathieu{
-resilience never hold in the case of VAS. 
-What happens when you add states ? \\
-So you consider for instance $q_0(u_k)$ instead of just $u_k$. \\
-Everything reachable from $q_0(u_k)$ is still of the form $q(u)$ with u greater than $v_{\Bad}$
-(we'd have $v_{\Bad}$ sum of the coordinate of the elements from the basis of $\Bad$ when not considering states).
-So now, let us fix some state $q$.
-Either there is an element of the basis of $\Bad$ with state $q$.
-Then $q(u)$ is necessarily in $\Bad$ (If this hold for all $q \in Q$, then again $k$-resilience does not hold).
-Either there is none.
-If there is none, then $q(\omega)$ is in the complement of $\Bad$, i.e. $\Safe$.
-Then, of course there are probably other elements in $\Safe$ but from
-the perspective of $\Bad$ and the extremely high elements $q_i(u_k)$  these ‘smaller’ elements are all but unreachable. 
-And the question becomes rather `can we reach $q(\omega)$ from every element of $\Bad$ save from maybe the few smaller ones ?’
-Since $q(\omega)$ here is upward-closed this is decidable i.e. it is asking coverability questions. 
-It’s not 100\% the question «asking, for every element in the basis of $\Bad$, whether or not $q(0)$ is coverable from it», since you could have, for smaller elements in $\Bad$, other elements of $\Safe$ you can reach, but it seem close.
-This would lead to a decision scheme.
-Need to write this more formally/properly.}
-\fi
+This all changes when you add control states.
+For a VASS with control states in $Q$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\uparrow q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
+ $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+This lead to a decision procedure, written in detail in Appendix~\ref{appendix brp VASS}, which lead to the following decidability result:
 
@@
 
-\begin{proof}
-\mathieu{Donner l'idée et se référencer à l'appendix~\ref{appendix brp VASS} pour la preuve complête.}
-\end{proof}
+% \begin{proof}
+% \mathieu{Donner l'idée et se référencer à l' pour la preuve complête.}
+% \end{proof}

### 2023-06-14 conclusion.tex
up

@@
 
-We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...
+We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
+% The problem [] is []. 
+%In [] the problem was introduced. We expand on the notion. 
+We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general. We also exhibited classes of WSTS with decidable resilience, namely, 
+completion-post-effective $\omega^2$-WSTS with strong compatibility (in the case $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$) and 
+downward-compatible
+ideal-effective WSTS
+(in the case $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$).
+We additionnally generalized decidability results from~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
 
-\alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
-%
-\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
+Several questions still remain.
+% We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...
+For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done. 
+One other angle of attack not studied here is synthesis of sets $\Bad$ and/or $\Safe$ for resilience,
+for instance, given a set $\Safe$, find the maximal upward-closed (resp. downward-closed) subset 
+$\Bad$ so that a system is $(\Safe,\Bad)$-resilient. 
+% \alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
+
+One could also extend upon the classes of set $\Bad$ and $\Safe$ considered.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward closed subsets.
+Finally, while we mention VASS, there are other computational models such as pushdown automata, one-counter automata or timed automata, which can be viewed as WSTS~\cite{DBLP:journals/tcs/FinkelS01}, in which a more detailed analysis of the resilience problems could have useful applications.
+
+% \alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}
 

### 2023-06-14 resilience.tex
up
new bibliographic reference

@@
 {\bf Related problems.} 
-A set of configurations $H$ is an {\em home-space} for a set of configurations $X$ if every configuration reachable from $X$ can reach $H$.
-Resilience as introduced above essentially asks whether $\Safe$ is a home-space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
-The home-space problem is decidable for
+A set of configurations $H$ is an {\em home space} for a set of configurations $X$ if every configuration reachable from $X$ can reach $H$.
+Resilience as introduced above essentially asks whether $\Safe$ is a home space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
+The home space problem is decidable for
 Petri nets when $X$ and $H$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
-To these author's knowledge, there are no general results for home-space in WSTS.
+To these author's knowledge, there are no general results for home space in WSTS.
 
-% \mathieu{vérifier au passage que c'est vrai qu'il y a pas de résultats home-space WSTS}
-
-Resilience and the home-space problem are also linked to the 
+Resilience and the home space problem are also linked to the 
 model-checking of the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
@@
 {\bf Synthesis of the main decidability results}\label{synthesis}
-
-% \mathieu{En rouge quand il s'agit de conjectures}
-
-% \alain{simplifier le tableau en enlevant les deux dernières colonnes}
-
 \begin{center}

### 2023-06-14 introduction.tex
new bibliographic reference
up

@@
 The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$.
-We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697}.
+We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
 State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.
@@
 \includegraphics[width=0.87\textwidth]{FigureB}
-	\caption{A channel system with three automata and four channels.
-Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+	\caption{A lossy VASS with three subsystems and four counters.
+Leftmost and rightmost systems $A$ and $B$ represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
 					\label{air control}
@@
 
-Consider the channel system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time. 
-The automata on the left and right represents the aircrafts attempting to land, while the middle one represent the control tower.
-The control tower can send messages to the aircrafts using channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using respectively channel $c_3$ and $c_4$.
-Remark channels $c_3$ and $c_4$ have a unary alphabet
-in the exemple and therefore can be seen as counters.
+\begin{example}\label{Example}
+Consider the vector addition system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time. 
+The automata $A$ on the left and $B$ on the right represent the aircrafts attempting to land, while the middle one $M$ represents the control tower.
+The control tower can communicate information to the aircrafts using
+% channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using 
+the first two shared counters, and these in turn can send messages to the control tower
+using the last two shared counters. 
+% respectively channel $c_3$ and $c_4$. 
+% Remark channels $c_3$ and $c_4$ have a unary alphabet in the exemple and therefore can be seen as counters.
 
 Safety requires that only one aircraft at a time attempt to land. The role of the control tower is to ensure this;
-there are two possible choices: aircraft A waits for aircraft B to land
-or vice-versa. To make its choice known to the aircrafts, the control tower will send them messages using the channels. 
-We assume that the channel are
-lossy \-- i.e. an arbitrary number of messages may be lost from the channels \--  but that they are still such that out of ten messages, at least one is not lost.
-In this example, the safe, desirable state would be one where the aircrafts have both landed. 
-Resilience here expresses that the aircrafts state will both land, while explicit resilience or $k$-resilience asks whether they will do so in at most $k$ steps. 
-Explicit resilience is of interest since in general we want to ensure the safe state will
+there are two possible choices: aircraft $A$ waits for aircraft $B$ to land
+or vice-versa. To make its choice known to the aircrafts, the control tower will use the counters. 
+We assume that the VASS is
+lossy \-- i.e. [...] \--  but that they are still such that [out of ten messages, at least one is not lost].
+In this example, the safe, desirable state would be one where the aircrafts have both landed, i.e.
+$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(v) \mid v \in \N^4\}$, and $\Bad$ is the complement of $\Safe$. 
+% We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.
+Resilience here expresses that the aircrafts will both land, while explicit resilience or $k$-resilience asks whether they will do so in at most $k$ steps. 
+Explicit resilience is of interest since in general we want to ensure the safe space will
 be reached in
-the least amount of step possible. 
+the least amount of steps possible. 
 In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time.
-With one out of ten messages guarranteed to not be lost, $40$-resilience hold in the example.
+With [one out of ten messages guarranteed to not be lost], $40$-resilience hold in the example.
+\end{example}
 

### 2023-06-14 definitions.tex
up

@@
 %   In this section, we introduce general notations and preliminary definitions.
-
-%The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).
-
-% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}
 
@@
 
+\mathieu{Mentionner le qo utilisé dans l'example~\ref{Example}}
+
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
@@
 %
-\iffalse
-%
-\begin{lemma}
-(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
-\end{lemma}
-
-\begin{proof}
-Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
-be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
-hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
-increasing subsequence.
-\end{proof}
-%
-\fi
 %
@@
 %
-\iffalse
-%
-
-\begin{lemma}{(Higman [40])} 
-If $\leq$ is a wqo then any upward-closed $I$ has a
-finite basis.
-\end{lemma}
-
-% \textcolor{red}{Expliquer ce que c'est une base d'abords.}
-
-\begin{proof}
-The set of minimal elements of $I$ is a basis because $\leq$ is well-founded. It
-only contains a finite number of non-equivalent elements otherwise they would make
-an infinite sequence contradicting the wqo assumption.
-\end{proof}
-
-\begin{lemma}  \label{upward-closed stablizes}
-If $\leq$ is a wqo then any infinite increasing sequence $I_0 \subseteq I_1 \subseteq I_2 \subseteq \cdots$ of
-upward-closed sets eventually stabilizes, i.e. there is a $k \in N$ such that 
-$I_k = I_{k+1} = I_{k+2} = \cdots $.
-\end{lemma}
-
-\begin{proof}
-Assue we have a counter-example.
-We extract an infinite subsequence where
-inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$.
-Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$.
-The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
-contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
-closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
-\end{proof}
-%
-\fi
 %
@@
 
-\alain{supprimer ce qui est inutile}
-
 \noindent
@@
 equipped with a wqo ${\leq} \subseteq S \times S$ such that   
-%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
 the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
@@
 
+\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}
+
 Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
-
-% \textcolor{red}{Define 'has effective pred-basis'.
-Maybe it should be included in WSTS definition, maybe it can be another def.
-I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}
-\iffalse
-\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
-If $\mathscr{S}$ is an WSTS and $U \subseteq S$ is an upward-closed set of states, then $\pred^*(U )$ is upward-
-closed.
-\end{proposition}
-%
-Proof.
-Assume $s \in \pred^* (U )$.
-Then $s \rightarrow^* t$ for some $t \in U $.
-If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$.
-Then $t' \in U$ and $s' \in \pred^*(U )$.
-\fi
 
@@
 
-%\alain{étrange ces deux propositions, l'une suffirait}
-%%%%%%%%%%%%%%%%%
-\iffalse
-%%%%%%%%%%%%%
-Proof.
-Assume $s \in \pred (I )$.
-Then $s \rightarrow t$ for some $t \in I $.
-If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$.
-Then $t' \in I$ and $s' \in \pred(I )$.
-%%%%%%%%%%%
-\fi
-%%%%%%%%%%
+\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}
 
@@
 We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
-\\
-
 With the effective pred-basis hypothesis, we obtain:

### 2023-06-14 resilience.tex
up

@@
 
+\iffalse
 \problemx{resilience problem (RP)}
@@
 {$\Bad \longrightarrow^{*} \Safe$ ?\newline}
-%
-%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
 
-%We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
-%
+\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
 
@@
 
+\fi
+
+\problemx{resilience problems}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
+{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
+({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
+({\sc $k$-resilience problem (kRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
+ $k$-resilient ?\newline}
+
 \begin{example}
@@
 
-% \begin{remark}[Reformulation]\label{reformulation}
-% $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-% (\Safe, \Bad)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
-% $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
-% (\Safe, \Bad)-$k$-resilient iff $\Bad \subseteq \pred^{\leq k}(\Safe)$.
-% \end{remark}
-
 These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
@@
 In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
+\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
 
 Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
-%		
-We deduce 
-% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
-% \mathieu{changer ça}
-decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
+We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
@@
 %	 and ideals are recursive 
-Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$.
+Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
+and in the general case when $\Safe = \downarrow \Safe$.
 
@@
 Appendix~\ref{case up up} for results.
+
+\mathieu{Beaucoup plus dévellopper le paragraphe ci-dessus et expliciter qu'à partir de maintenant, on considère $\Bad$ complémentaire de $\Safe$ par défaut.}
 

### 2023-06-14 resilience.tex
réuni les trois problèmes dans un même environement mais le k m'embête

@@
 
-\problemx{resilience problems}
+\problemx{Resilience problems}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
@@
 ({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
-({\sc $k$-resilience problem (kRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
+({\sc bounded resilience problem (BRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
  $k$-resilient ?\newline}

### 2023-06-14 state-resilience.tex
réuni les trois problèmes dans un même environement mais le k m'embête

@@
 
+\iffalse
+
 \problemx{State-resilience problem (SRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-% 		an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
-% 		{$ ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe, ~ s' \rightarrow^{*} s''$ ?\newline}
 {$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
 {$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient. ?\newline}
 
-%\textcolor{blue}{
-%justification des definitions: si on enlève la condition  an upward-closed set $U$ (extended resilience $U$ qq) reachability reduces to resilience (avec $U=\{x\}$ donc indecidable pour les modèles à reachability indécidable.
-%}
+\fi
+
+\problemx{State resilience problems}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
+{ ({\sc state-resilience problem (SRP)}) $\post^*(s) \longrightarrow^{*} \Safe$ ?\newline  
+({\sc $k$-state-resilience problem (kSRP)})		$\post^*(s) \longrightarrow^{\leq k} \Safe$ ?\newline
+({\sc bounded-state-resilience problem (BSRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
+ $k$-resilient ?\newline}
 

### 2023-06-14 definitions.tex
up

@@
 
-\mathieu{Mentionner le qo utilisé dans l'example~\ref{Example}}
-
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
@@
 %
+%
+% \mathieu{Mentionner le qo utilisé dans l'example~\ref{Example}}
+% As an illustration, Example~\ref{Example} uses a quasi order based on the  canonical order of $\N^4$
+As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
 %
@@
 
-\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}
-
 Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
+The Lossy VASS given in Example~\ref{Example}, for instance, is a WSTS.
+% \mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}
 
@@
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-
-\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}
+%
+% \mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}
+As an illustration, the Lossy VASS given in Example~\ref{Example} is post-effective,
+ideal-effective, and
+has effective pred-basis.
 

### 2023-06-14 resilience-safe-down.tex
up

@@
 %%%%%%%%
-\subsection{Case: $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.}
+\subsection{Case: $\Safe=\downarrow \Safe$.}
 %
@@
 
-Without the downward-compatible hypothesis, the problem becomes undecidable. Indeed, it is 
+Without the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed, the problem becomes undecidable. Indeed, it is 
 undecidable for Reset 
@@
 This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.
+See Appendix~\ref{HS-Minsk} for a more detailed construction.
+
+\mathieu{Remarquer que les résultats sont aussi valide si on n'a pas safe et bad complémentaires mais juste $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$}

### 2023-06-14 resilience.tex
up

@@
 Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
-then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g.
-$q(-k-1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$.
+then resilience hold but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g. $q(k+1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$. 
 \end{example}
@@
 In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
-\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
+%\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
+Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed and downward-closed.
 
@@
 In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
-%
-%	 and ideals are recursive 
 Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
@@
 \end{center}
-
-%We first assume that the safety property is given by an upward-closed set and the bad condition by a decidable downward-closed set. 
-% \textcolor{red}{Seems like a reasonable assumption to me.}
-
-%From these considerations, we formulate instances of the abstract resilience problems for well-
-%structured transition systems.

### 2023-06-14 definitions.tex
up

@@
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
-The Lossy VASS given in Example~\ref{Example}, for instance, is a WSTS.
-% \mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}
+The lossy VASS given in Example~\ref{Example}, for instance, is a WSTS.
+% \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est un WSTS}
 
@@
 %
-% \mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}
-As an illustration, the Lossy VASS given in Example~\ref{Example} is post-effective,
+% \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est effectif}
+As an illustration, the lossy VASS given in Example~\ref{Example} is post-effective,
 ideal-effective, and

### 2023-06-14 introduction.tex
up

@@
 We assume that the VASS is
-lossy \-- i.e. [...] \--  but that they are still such that [out of ten messages, at least one is not lost].
+lossy \-- i.e. 
+ the contents of a counter can spontaneously get lower at any time
+ \--  but that they are still such that out of ten incrementations of a counter, at least one is not rendered void.
 In this example, the safe, desirable state would be one where the aircrafts have both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(v) \mid v \in \N^4\}$, and $\Bad$ is the complement of $\Safe$. 
+$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4\}$, and $\Bad$ is the complement of $\Safe$. 
 % We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.
@@
 In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time.
-With [one out of ten messages guarranteed to not be lost], $40$-resilience hold in the example.
+With one out of ten increment of a counter guarranteed to not be lost, $40$-resilience hold in the example.
 \end{example}

### 2023-06-14 resilience.tex
up

@@
 model-checking of the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-which has been shown decidable for Lossy Counter Machine
+which has been shown decidable for Lossy Counter Machine (LCM)
 when $X$ and $Y$ are semi-linear sets~\cite{DBLP:conf/rp/Schnoebelen10}.
@@
 ``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
-  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for Lossy Counter Machines).
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
 
@@
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed and downward-closed.
+Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
 
 Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
-We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
-In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
+% We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
+In particular, the existing proofs of resilience use the decidability of reachability~\cite{DBLP:journals/corr/abs-2207-02697}; but we wish to decide resilience for models with undecidable reachability.
 Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$

### 2023-06-14 resilience-safe-down.tex
up

@@
 
-\mathieu{Remarquer que les résultats sont aussi valide si on n'a pas safe et bad complémentaires mais juste $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$}
+Remark we did not make use of the hypothesis $\Bad$ complement of $\Safe$, simply 
+$\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complement of each others.

### 2023-06-14 resilience.tex
up

@@
 
-While considering a sequence of configurations, every prefix is always 
-either in the case that the latest visited $\Bad$ configuration has already been followed by a $\Safe$ configuration
+Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, it is useful to note that in a sequence of states of $S$, every prefix is always 
+either in the case that the latest visited $\Bad$ state has already been followed by a 
+$\Safe$ state
 or in the case that it still has to. 
-Thus, in context, there are no true ``neutral'' configurations: either a neutral configuration is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ configuration, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ configuration and has, accordingly, to reach a $\Safe$ one.
-Hence we will consider not only $\Bad$ and $\Safe$ disjoint but
+Thus, \emph{in context}, there are no true ``neutral'' state: either a neutral state is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ state, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ state and has, accordingly, to reach a $\Safe$ one.
+Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but
 % w.l.o.g.
 	also
-$\Bad$ complement of $\Safe$,
+that $\Bad$ is the complement of $\Safe$,
 in which case
@@
 Appendix~\ref{case up up} for results.
-
-\mathieu{Beaucoup plus dévellopper le paragraphe ci-dessus et expliciter qu'à partir de maintenant, on considère $\Bad$ complémentaire de $\Safe$ par défaut.}
 

### 2023-06-14 state-resilience.tex
up

@@
 However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
-%
-%			However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
-%
-%\problemx{(I,J)-$k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
-%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%
-%\problemx{general $k$-resilience problem for WSTS}
-%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
-%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-%		Again, we first consider the case where the recovery time is bounded by a $k \in \N$.
+Since we still consider $\Bad$ and $\Safe$ complement of each others, this is akin to asking whether from $\post^*(s_0)$ there must exist a path to $\Safe$. 
 The three previous problems become:
@@
 As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
-$\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$ 
+$\Safe = \uparrow \Safe$ (and $\Bad = \downarrow \Bad$) 
 and where 
-$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+$\Safe = \downarrow \Safe$ (and $\Bad = \uparrow \Bad$).
 
-% \mathieu{Ici on s'intéresse seulement aux cas $\uparrow \downarrow$ et $\downarrow \uparrow$.}
-
-\subsection{Case: $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$}
+\subsection{Case: $\Safe = \uparrow \Safe$% and $\Bad = \downarrow \Bad$
+}
 
@@
  reducible to {\sc State-resilience}.
-
-% \noindent
-% {\bf Reset VASS}
 
@@
 $pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
-
-\iffalse
-\alain{ou alors $t=(u,u')$ such that $u \rightarrow u'$ mais alors c'est bizarre de nommer par la même lettre $t$ des couples différents $(u,u'), (v,v'),...$ dans un WSTS...ou alors on appelle transition un couple $(u,u')$ such that $u \rightarrow u'$ mais plus de lettre $t$...ou on définit un WSTS par un nombre fini de (shéma de ) transition $t_i$ qui sont compatibles par compatibilité mais ça contraint à un nombre fini (comme dans les VASS) ou alors on ne parle ici que des VASS et pas des WSTS mais ça suffira puisque state resilience pour VASS implique pour WSTS BREF il faut réfléchir à ce qu'est une transition...}
-
-\mathieu{En train de penser que le mieux serait de définir la $t$-liveness pas pour les WSTS mais pour les (Reset) VASS uniquement et ensuite dire SRP indécidable dans WSTS car indécidable pour les Reset VASS.}
-
-\fi
 
@@
 VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
-\iffalse
-
-\begin{proposition}\cite{DBLP:journals/corr/abs-2108-00889}
-Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.
-% \alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
-\end{proposition}
-
-\fi
 It is well-known that 
@@
 \end{proof}
-
-%il existe des modèles où on peut calculer $\uparrow \post^*$: inserted channel systems: on sait calculer $\post^*$ qui est rationnel donc on sait calculer $\uparrow \post^*$.
-
-%
-%si on garde strong mais on enlève effective basis of $\post^*$:
-%
-%\begin{proposition}(à prouver)
-%{\sc $k$-resilience}, hence {\sc bounded resilience}, is undecidable for effective WSTS with strong compatibility.
-%\end{proposition}
-%
-%\begin{proof}
-%pas de effective basis of $\post^*$. leur algorithme n'est plus un algo. ex reset PN ?  \textcolor{red}{pas certain ! à faire Mathieu}
-%\end{proof}
 
@@
 
-\subsection{Case: $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$}
+\subsection{Case: $\Safe = \downarrow \Safe$% and $\Bad = \uparrow \Bad$
+}
 

### 2023-06-14 introduction.tex
section 1

@@
 even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
+A set $X$ is an \emph{home space} for a system $(S,s_0)$ \cite{DBLP:conf/ac/MemmiV86} if from all reachable states from $s_0$, it is always possible to reach $X$; recently Memmi extended this definition as follows: A set $X$ is an \emph{$A$-home space} for a system $S$ \cite{memmi2023invariants} if from all states in $A$, it is always possible to reach $X$.
+
 \begin{center}

### 2023-06-15 introduction.tex
up

@@
 	\caption{A lossy VASS with three subsystems and four counters.
-Leftmost and rightmost systems $A$ and $B$ represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
 					\label{air control}
@@
 \begin{example}\label{Example}
-Consider the vector addition system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time. 
+Consider the vector addition system in Figure~\ref{air control}. It models a scenario in which two aircrafts want to land in an airport at the same time. 
 The automata $A$ on the left and $B$ on the right represent the aircrafts attempting to land, while the middle one $M$ represents the control tower.
@@
 
-Safety requires that only one aircraft at a time attempt to land. The role of the control tower is to ensure this;
+Safety requires that only one aircraft at a time attempts to land. The role of the control tower is to ensure this;
 there are two possible choices: aircraft $A$ waits for aircraft $B$ to land

### 2023-06-15 state-resilience.tex
up

@@
 The three previous problems become:
-
-\iffalse
 
@@
 
-\fi
-
+\iffalse
 \problemx{State resilience problems}
@@
  $k$-resilient ?\newline}
+\fi
 
 Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
-As in the Section~3, we study decidability results for different pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed. We will consider here the cases where 
-$\Safe = \uparrow \Safe$ (and $\Bad = \downarrow \Bad$) 
-and where 
-$\Safe = \downarrow \Safe$ (and $\Bad = \uparrow \Bad$).
+As in the Section~3, we study decidability results for $\Safe$ downward-closed and upward-closed.
 
@@
 \begin{proposition}\label{reductions}
-% In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
 {\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
@@
 $U_t=\uparrow pre(t)$ 
-% \alain{$\pred(t)$ n'est pas défini et $\pred$ est une mauvaise notation pour ça: plutôt: $pre(t)=\{ u \in S \mid \exists u' \in S$  such that $u \xrightarrow{t} u'$\} }
 is the upward closure of the preconditions to fire transition $t$.  
 The problem reduces itself to {\sc State-resilience}
-where $\Safe = U_t$ and $\Bad = Q \times \N^d \setminus \Safe$.
+where $\Safe = U_t$.
 \end{proof}
 
-% To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. 
 The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS, whose proof can be seen in Appendix~\ref{appendix}.
@@
 when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$.
+$\Safe=\uparrow \Safe$.
 \end{theorem}
@@
  {\sc $k$-state-resilience}. 
-%
-%
-\iffalse
-% \begin{corollary}\label{bsrp up down}
-Bounded-state-resilience and 
-$k$ state-resilience are undecidable for strongly compatible WSTS with effective pred-basis
-when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$.
-% \end{corollary}
-\fi
-%
-%\begin{proof} 
-%  From the two reductions above. 
 \end{proof}
@@
 when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$.
+$\Safe=\uparrow \Safe$.
 \end{corollary}
@@
 Then, because {\sc $n_0$-state-resilience} is decidable, 
-checking $\uparrow \post^*(s) \cap \Bad \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+checking $\uparrow \post^*(s) %\cap \Bad
+ \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
 and {\sc State-resilience} is decidable.
@@
 Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
-\iffalse
-Then, one would be able to decide whether an element $m \in \min(\uparrow \post^*)$ is reachable.
-%in counter $i$ 
-by examining if there is %some vector $v$ in the basis such that $v(i)=0$
-$0$ in the basis%
-. But reachability of $0$  %in a counter $i$ 
-is undecidable for Reset VASS. 
-% sont WSTS with strong compatibility. et l'accessibilité est undecidable.
-%	$\uparrow \post^*$ n'est pas calculable pour les LCS non plus car $\uparrow \post^*= \uparrow \downarrow \post^*$ n'est pas calculable ???
-\fi
 Then, {\sc State-resilience} would be decidable for Reset VASS, which leads to a contradiction.
@@
 {\sc State-resilience} is decidable for 
-%(not necessarily strong)
  WSTS with effective 
@@
 %\alain{notation fluctuante}
- be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$
-and $\Bad$ a downward-closed set given by its finite set of (maximal) ideals.
+ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
-\[ \post^*(s) \cap \Bad \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} \cap \Bad \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and $\Bad$ is decidable, we can directly compute $ B_s \cap \Bad$.
-% For every element of $ B_p \cap J$, checking that it is in $\pred^*(I)$ is the same as checking coverability of I, and thus decidable.
-We can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, and hence check that $B_{\uparrow \post^*(s)} \cap \Bad \setminus \pred^*(\Safe) = \emptyset$. 
+\[ \post^*(s) %\cap \Bad
+ \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} %\cap \Bad 
+ \subseteq \pred^*(\Safe)\]
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} %\cap \Bad
+ \setminus \pred^*(\Safe) = \emptyset$. 
 \end{proof}
-
-% In the proof of Thm. 1, it was crucial that we have strong compatibility. This approach does not work
-% for WSTSs in general. We loose precision when we only demand compatibility. Thus, we conjecture
-% that both resilience problems are undecidable for WSTSs in general, but this question remains still open.
-%
-% Well, maybe their version of resilience but resilience with * seems fine.
-%
 
@@
 which is incorrect.
-% An example of this is ...
 
@@
 
-Results synthesis in the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$:
+Results synthesis in the case $\Safe = \uparrow \Safe$:
 \begin{center}
@@
 
-Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by Reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
+Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by Reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
 
@@
 when
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
+$\Safe=\downarrow \Safe$.
 \end{theorem}
 
-% \begin{proof}
-% \mathieu{Refaire la preuve en utilisant les Minsky machines, i.e. si on pet résoudre la question dans les reset alors on peut la résoudre dans les Minsky machine ... et ça y est indécidable.}
-% \end{proof}
-
-Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
-is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
+Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$
+is symmetrical to the case $\Safe = \uparrow \Safe$.
 %
@@
 Lemma~\ref{Lemma intersection 2} can be used to show that
-if $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$,
+if $\Safe = \downarrow \Safe$,
 then
-$\post^*(s) \cap \Bad \subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) \cap \Bad \subseteq \pred^*(\Safe)$.
+$\post^*(s) %\cap \Bad 
+\subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) 
+%\cap \Bad 
+\subseteq \pred^*(\Safe)$.
 
@@
 {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities,
-%\alain{a-t-on aussi la upward compatibility ?}
-%\mathieu{oui}
-$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+$\Safe = \downarrow \Safe$.
 \end{theorem}
@@
 \begin{proof}
-In order to decide whether $\post^{\leq n}(s) \cap \Bad \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
+In order to decide whether $\post^{\leq n}(s) %\cap \Bad
+ \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
 one looking for a resilience certificate and one looking for a non-resilience certificate.
@@
 Notice that, by 
-%standard \alain{upward ? downward ?}
 upward compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
@@
 Procedure 1 stops when it finds an invariant $D$ such that
-$D  \cap \Bad \subseteq \pred^*(\Safe)$. 
+$D  %\cap \Bad
+ \subseteq \pred^*(\Safe)$. 
 Indeed
-$D  \cap \Bad \subseteq  \pred^*(\Safe)$ implies
-$\downarrow \post^*(s) \cap \Bad \subseteq  \pred^*(\Safe)$
+$D % \cap \Bad 
+\subseteq  \pred^*(\Safe)$ implies
+$\downarrow \post^*(s) % \cap \Bad
+ \subseteq  \pred^*(\Safe)$
 since $ \downarrow \post^*(s)  \subseteq D$.
@@
 The second procedure iteratively computes
-$\post^{\leq n}(s) \cap \Bad$
+$\post^{\leq n}(s) %\cap \Bad
+$
 until it finds an element
@@
 \end{proof}
-
-% \mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of $\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}
 
@@
 {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility,
-$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+$\Safe = \downarrow \Safe$.
 \end{corollary}

### 2023-06-15 state-resilience.tex
up

@@
 However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
-Since we still consider $\Bad$ and $\Safe$ complement of each others, this is akin to asking whether from $\post^*(s_0)$ there must exist a path to $\Safe$. 
+Since we still consider $\Bad$ and $\Safe$ complement of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
 The three previous problems become:
@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
+% {$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
+{$\post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
 
@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-%
+% { $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+{ $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
 
@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient. ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient ?\newline}
 

### 2023-06-15 state-resilience.tex
up

@@
 \subseteq S$.}
-% {$\exists s' \in D ~ s \to^* s'$? \newline}\alain{ecrire $s  \to^* D$}
 {$s  \to^* D$? \newline}
@@
 , a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-%i 			.e. given a state $s$ of a WSTS
-% in the regarded class 
-%with strong compatibility 
-%and a decidable downward-closed set $D$, it can be decided whether $\exists s' \in D ~ s \to^* s'$. 
 \end{proposition}
 
-% \alain{rappeler l'idée de la preuve}
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
@@
 
-% \mathieu{Quelque chose de plus à dire sur la charactérisation?}
-
 More concretely,
@@
 It is well-known that 
-% Petri nets are WSTS with strong  compatibility (Thm.~6.1 in \cite{DBLP:journals/tcs/FinkelS01}). 
-% Since 
 VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
@@
 There exist classes of WSTSs with strong 
-%(upward)
- compatibility for which there don't exist an algorithm computing a basis of $\uparrow \post^*$.
+ compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$.
 \end{proposition}
@@
 
-Results synthesis in the case $\Safe = \uparrow \Safe$:
+{\bf Results synthesis in the case $\Safe = \uparrow \Safe$}
 \begin{center}

### 2023-06-15 application.tex
up

@@
 
-\begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
-\end{proposition}
-
 \begin{theorem}{}
@@
 \begin{proof}
-When asking about {\sc Resilience}, in context, there are no true ``neutral'' configurations.
+\iffalse When asking about {\sc Resilience}, in context, there are no true ``neutral'' configurations.
 Indeed, from a VASS $V$, $\Safe$ and $\Bad$ semilinear, one can build VASS $V'$, $\Safe'$ semilinear, and $\Bad'$ complement of $\Safe'$ such that {\sc Resilience} holds for
@@
 
-Without loss of generality hence let us consider only the case when $\Safe$ is semilinear
+Without loss of generality hence let us consider only \fi
+We consider the case when $\Safe$ is semilinear
 and $\Bad = S \setminus \Safe$ is semilinear too.
@@
 
-Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logics.
+Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logic.
 Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.

### 2023-06-15 appendix.tex
up

@@
 
-\section{Proof of Theorem~\ref{liveness reset}}\label{appendix}
+Reset VASS extend the basic VASS model with special “reset
+transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
+\begin{definition}
+A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
+% $\mathds{Z}^d$
+labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
+%$T \subseteq Q \times \mathds{Z}^d \times Q$
+$T \subseteq Q \times Op \times Q$
+ will be referred to as the {\em control-transitions} of $V$,
+where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
+\end{definition}
 
-Recall Theorem~\ref{liveness reset}.
+Again $Q \times \N^d$
+ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
+\begin{samepage}\begin{itemize}
+\item  $t = (q,add(\textbf{z}),q') \in T$
+% then for all $\textbf{u} \in \N^d$ such that  
+% $\textbf{u}+\textbf{z} \geq 0$
+% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
+\item $t = (q,reset(\gamma),q') \in T$ 
+% then for all $\textbf{u} \in \N^d$ 
+% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+\end{itemize} \end{samepage}
+
+It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
+Moreover, we are interested in the 
+%following
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
+Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
+which we then reduce to {\sc State-resilience}.
+A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
+live. This leads to the following problem.
+
+\problemx{$t$-liveness}
+{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{Is $t$ live in $s_0$ ? \\}
+
+To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
+$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+
+\begin{proposition}\label{reductions}
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness 
+in a 
+% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
+Reset $d$-VASS $(Q,T)$ 
+with initial configuration $s_0$
+as the following formula
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
+where
+$U_t=\uparrow pre(t)$ 
+is the upward closure of the preconditions to fire transition $t$.  
+The problem reduces itself to {\sc State-resilience}
+where $\Safe = U_t$.
+\end{proof}
+
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS%, whose proof can be seen in Appendix~\ref{appendix}%
+.
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
+\end{proposition}
+
+% \section{Proof of Theorem~\ref{liveness reset}}\label{appendix}
+
+% Recall Theorem~\ref{liveness reset}.
 
@@
 
-\begin{theorem*}[Adjusting Theorem~5.5 from \cite{peterson1981petri}]
-The  Reset Petri Net zero reachability problem can be reduced to the liveness problem.
-\end{theorem*}
+%\begin{proposition}\label{liveness reset}
+% The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
+%\end{proposition}
 
@@
 Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero. 
+\end{proof}
 
-\end{proof}
+Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
+% the following:
+%
+%
+{\sc %$ Reset $d$-VASS 
+$t$-liveness} is undecidable.
 

### 2023-06-15 state-resilience.tex
up

@@
 
-Reset VASS extend the basic VASS model with special “reset
-transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
-\begin{definition}
-A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
-% $\mathds{Z}^d$
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
-%$T \subseteq Q \times \mathds{Z}^d \times Q$
-$T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control-transitions} of $V$,
-where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
-
-Again $Q \times \N^d$
- denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
-$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
-\begin{samepage}\begin{itemize}
-\item  $t = (q,add(\textbf{z}),q') \in T$
-% then for all $\textbf{u} \in \N^d$ such that  
-% $\textbf{u}+\textbf{z} \geq 0$
-% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
-and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-\item $t = (q,reset(\gamma),q') \in T$ 
-% then for all $\textbf{u} \in \N^d$ 
-% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
-and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize} \end{samepage}
-
-It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
-Moreover, we are interested in the 
-%following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
-Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-%
-We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
-which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
-live. This leads to the following problem.
-
-\problemx{$t$-liveness}
-{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
-{Is $t$ live in $s_0$ ? \\}
-
-To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
-$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
-
-\begin{proposition}\label{reductions}
-{\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
-\end{proposition}
-
-\begin{proof}
-We reformulate $t$-liveness 
-in a 
-% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
-Reset $d$-VASS $(Q,T)$ 
-with initial configuration $s_0$
-as the following formula
-\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
-where
-$U_t=\uparrow pre(t)$ 
-is the upward closure of the preconditions to fire transition $t$.  
-The problem reduces itself to {\sc State-resilience}
-where $\Safe = U_t$.
-\end{proof}
-
-The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS, whose proof can be seen in Appendix~\ref{appendix}.
-\begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
-\end{proposition}
-
-Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
-% the following:
-%
-%
-{\sc %$ Reset $d$-VASS 
-$t$-liveness} is undecidable.
 Since {\sc %$Reset $d$-VASS 

### 2023-06-16 appendix.tex
up

@@
 
-The well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
-Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed.
-%		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
+We relax for now the hypothesis that $\Bad$ is the complement of $\Safe$, thus it is possible for $\Safe$ and $\Bad$ to share closure properties, i.e. for them to be both downward-closed.
 
@@
 It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}.
-Moreover, we are interested in the 
+
+\section{Reset VASS and state-resilience}\label{appendix}
+
+We are now interested in the 
 %following
@@
 
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
+Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
+%This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
 \section{More Appendix thing if necessary}\label{appendix brp VASS}

### 2023-06-16 resilience-safe-down.tex
up

@@
 Let us now consider the case $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.
-It is of interest to note this case can be linked to the problem of mutual exclusion: indeed, one can choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$, and $\Safe$ to be the downward-closed complement of $\Bad$.
+It is of interest to note this case can be linked to the problem of mutual exclusion.
+Indeed the well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
+%  indeed, one can choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$, and $\Safe$ to be the downward-closed complement of $\Bad$.
+
  In this case, our contribution consist in the following:

### 2023-06-16 resilience.tex
up

@@
 
-In this section we introduce the concept of resilience, bounded-resilience and $k$-resilience for well structured transition systems. We show that the resilience problems are decidable
+In this section we introduce the concepts of resilience, bounded-resilience and $k$-resilience for well structured transition systems. We show that the resilience problems are decidable
 for  completion-post-effective $\omega^2$-WSTS with strong compatibility
-in the case $\Bad$ is downward-closed and $\Safe $ is upward-closed,
+in the case of an upward-closed safe set,
 and that resilience is decidable for downward-compatible ideal-effective WSTS
-when $\Safe$ is downward-closed and $\Bad $ is upward-closed. 
+when the set of safe spaces is downward-closed. 
 We show that the three resilience problems are however undecidable for WSTS with
@@
 In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
+
+Let us formalize three resilience problems.
+
+\problemx{resilience problem (RP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
+
+% \alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
+
+\problemx{$k$-resilience problem (kRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
+{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
+
+\problemx{bounded resilience problem (BRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
+{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
+ $k$-resilient ?\newline}
+
+\iffalse
+
+\problemx{Resilience problems}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
+{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
+({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
+({\sc bounded resilience problem (BRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
+ $k$-resilient ?\newline}
+
+\fi
+
+\iffalse
+\begin{example}
+Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
+then resilience hold but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g. $q(k+1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$. 
+\end{example}
+\fi
+
+These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
+So we restrict our framework to the class of infinite-state WSTS.
+Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
+%\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
+Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
+
+Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
+% We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
+In particular, the existing proofs of resilience use the decidability of reachability;
+\mathieu{quelles preuves ? pour quelle notion de résilience ?} but we wish to decide resilience for models with undecidable reachability.
+Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
+and in the general case when $\Safe = \downarrow \Safe$.
+
+Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, it is useful to note that in a sequence of states of $S$, every prefix is always 
+either in the case that the latest visited $\Bad$ state has already been followed by a 
+$\Safe$ state
+or in the case that it still has to. 
+Thus, \emph{in context}, there is no true ``neutral'' state: either a neutral state is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ state, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ state and has, accordingly, to reach a $\Safe$ one.
+Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but
+% w.l.o.g.
+	also
+that $\Bad$ is the complement of $\Safe$,
+in which case
+if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
+Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. are both downward-closed or both upward-closed,
+have also been studied, 
+see 
+Appendix~\ref{case down down}
+and
+Appendix~\ref{case up up} for results.
 
@@
   and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
-
-Let us formalize three resilience problems.
-
-\iffalse
-\problemx{resilience problem (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
-
-\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}
-
-\problemx{$k$-resilience problem (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
-
-\problemx{bounded resilience problem (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
-%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
- $k$-resilient ?\newline}
-
-\fi
-
-\problemx{Resilience problems}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
-{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
-({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
-({\sc bounded resilience problem (BRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
- $k$-resilient ?\newline}
-
-\begin{example}
-Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
-then resilience hold but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g. $q(k+1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$. 
-\end{example}
-
-These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
-So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
-%\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
-
-Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
-% We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
-In particular, the existing proofs of resilience use the decidability of reachability~\cite{DBLP:journals/corr/abs-2207-02697}; but we wish to decide resilience for models with undecidable reachability.
-Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
-and in the general case when $\Safe = \downarrow \Safe$.
-
-Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, it is useful to note that in a sequence of states of $S$, every prefix is always 
-either in the case that the latest visited $\Bad$ state has already been followed by a 
-$\Safe$ state
-or in the case that it still has to. 
-Thus, \emph{in context}, there are no true ``neutral'' state: either a neutral state is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ state, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ state and has, accordingly, to reach a $\Safe$ one.
-Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but
-% w.l.o.g.
-	also
-that $\Bad$ is the complement of $\Safe$,
-in which case
-if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
-Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. are both downward-closed or both upward-closed,
-have also been studied, 
-see 
-Appendix~\ref{case down down}
-and
-Appendix~\ref{case up up} for results.
 

### 2023-06-16 resilience.tex
up

@@
 Resilience and the home space problem are also linked to the 
-model-checking of the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-which has been shown decidable for Lossy Counter Machine (LCM)
-when $X$ and $Y$ are semi-linear sets~\cite{DBLP:conf/rp/Schnoebelen10}.
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+% which has been shown
+is decidable for Lossy Counter Machine (LCM)
+when $X$ and $Y$ are semi-linear sets.
 Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 

### 2023-06-16 resilience-safe-up.tex
up

@@
-\subsection{Case: $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$.}
+\subsection{Case: $\Safe=\uparrow \Safe$.}
 
@@
 It can be viewed as a generalizeation
- of coverability, namely,
+ of coverability, as
 it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
@@
 
-Without the completion-post-effective hypothesis or the pred-basis hypothesis, coverability 
+Without the completion-post-effective hypothesis nor the pred-basis hypothesis, coverability 
 becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}. 
@@
 {\sc Resilience} itself is undecidable.
+
+Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.

### 2023-06-16 resilience.tex
up

@@
 {\bf Related problems.} 
-A set of configurations $H$ is an {\em home space} for a set of configurations $X$ if every configuration reachable from $X$ can reach $H$.
+A set of states $H$ is an {\em home space} for a set of states $X$ if every state reachable from $X$ can reach $H$.
+A {\em home space} is a state whose corresponding singleton set is a home space.
 Resilience as introduced above essentially asks whether $\Safe$ is a home space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.

### 2023-06-16 state-resilience.tex
up

@@
 However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
-Since we still consider $\Bad$ and $\Safe$ complement of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
+Since we still consider $\Bad$ and $\Safe$ complements of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
 The three previous problems become:
@@
 
-Unfortunately, {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
+We start with the case $\Safe = \uparrow \Safe$.
+%
+Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
 This stems from the fact that it is undecidable in the particular case of Reset VASS,
 where  $t$-liveness is both undecidable and 
- reducible to {\sc State-resilience}.
-
-Since {\sc %$Reset $d$-VASS 
-$t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
-Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
-This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+ reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
 
@@
 \begin{proof}
-{\sc State-resilience} itself is undecidable from Proposition~\ref{reductions} since 
-{\sc %Reset $d$-VASS 
-$t$-liveness} is undecidable.
-% \begin{proposition}
-% \end{proof}
-% \begin{proposition}
-In WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
+{\sc State-resilience} itself is undecidable % from Proposition~\ref{reductions} 
+since 
+{\sc $t$-liveness} is undecidable in Reset VASS
+and reducible to {\sc State-resilience}.
+Additionally, in WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
 reducible to {\sc $k$-state-resilience}:
-% \end{proposition}
-% 
-% \begin{proof}
 since $\Safe=\uparrow \Safe$ and
@@
 $n_0$, then iteratively check whether $k$-state-resilience 
-% \alain{$k$-state resilience ou $k$-state-resilience ?}
 hold for $k$ from $0$ to $n_0$.  
-% \end{proof}
 Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
-% \end{proposition}
-%
-% \begin{proof}
 since 
 $\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
-Hence the %equivalence, and
- undecidability of {\sc Bounded-state-resilience}
+Hence the undecidability of {\sc Bounded-state-resilience}
 and 
@@
 On the positive side, let us recall a result about {\sc Bounded-state-resilience} decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
-
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}

### 2023-06-16 appendix.tex
up

@@
 
-Reset VASS extend the basic VASS model with special “reset
+Reset-VASS extend the basic VASS model with special “reset
 transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
 \begin{definition}
-A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
+A {\em reset-VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
 % $\mathds{Z}^d$
@@
 
-\section{Reset VASS and state-resilience}\label{appendix}
+\section{Reset-VASS and state-resilience}\label{appendix}
 
@@
 
-The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS%, whose proof can be seen in Appendix~\ref{appendix}%
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
 .
 \begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
+The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
 \end{proposition}
@@
 %\begin{proposition}\label{liveness reset}
-% The  zero reachability problem can be reduced to the $t$-liveness problem in Reset VASS.
+% The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
 %\end{proposition}

### 2023-06-16 state-resilience.tex
up

@@
 Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
-This stems from the fact that it is undecidable in the particular case of Reset VASS,
+This stems from the fact that it is undecidable in the particular case of reset-VASS,
 where  $t$-liveness is both undecidable and 
@@
 since 
-{\sc $t$-liveness} is undecidable in Reset VASS
+{\sc $t$-liveness} is undecidable in reset-VASS
 and reducible to {\sc State-resilience}.
@@
 VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework. Moreover, we may show:
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  \iffalse Moreover, we may show:
 
@@
 \begin{proof}
-Let us show that Reset VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for Reset VASS. 
-Then, {\sc State-resilience} would be decidable for Reset VASS, which leads to a contradiction.
-\end{proof}
+Let us show that reset-VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
+Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for reset-VASS. 
+Then, {\sc State-resilience} would be decidable for reset-VASS, which leads to a contradiction.
+\end{proof} \fi
+Moreover there exist classes of WSTSs with strong 
+ compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$, such as reset-VASS.
 
-Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience }.
-Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889}:
+Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
 
@@
 
+We now consider the case $\Safe = \downarrow \Safe$.
+%
 Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by Reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
 

### 2023-06-16 application.tex
up

@@
 \section{Resilience for VASS}
+
+In this section we continue to study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+$\Safe$ is upward-closed. Lacking downward compatibility or a more relaxed hypothesis that
+for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
+VASS do not inherit the decidability results for WSTS in the case $\Safe$ is downward-closed.
+In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.
+We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
 
@@
 
+The above results concern only unbounded resilience. We consider now 
+{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
+the path from $\Bad$ to $\Safe$
+when $\Safe$ is downward-closed. Unfortunately, we have the following:
+
 \begin{proposition}

### 2023-06-16 resilience-safe-down.tex
up
section 3

@@
 
-Let us now consider the case $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.
+Let us now consider the case $\Safe=\downarrow \Safe$ hence $\Bad=\uparrow \Bad$.
 It is of interest to note this case can be linked to the problem of mutual exclusion.
-Indeed the well-known mutual exclusion property can be modelized, in a $d$-VASS with $k$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the number of processes that are allowed to be simultaneously in the critical section.
+Indeed the well-known mutual exclusion property can be modelized, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
 Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
@@
 
- In this case, our contribution consist in the following:
+%	 In the case $\Safe=\downarrow \Safe$, our contribution consist in the following:
 

### 2023-06-16 resilience-safe-down.tex
up

--- 2023-06-16 resilience-safe-down.tex
@@
 
- In this case, our contribution consist in the following:
+ In the case $\Safe=\downarrow \Safe$, our contribution consist in the following:
 
@@
 %Petri-nets
-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
+VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
 This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.

### 2023-06-16 appendix.tex
up

@@
-\section{Case: $\Bad=\downarrow \Bad$ and $\Safe=\downarrow \Safe$}\label{case down down}
+\section{Resilience when $\Safe$ and $\Bad$ have the same closure properties}
 
-We relax for now the hypothesis that $\Bad$ is the complement of $\Safe$, thus it is possible for $\Safe$ and $\Bad$ to share closure properties, i.e. for them to be both downward-closed.
+We relax for now the hypothesis that $\Bad$ is the complement of $\Safe$, thus it is possible for $\Safe$ and $\Bad$ to share closure properties, i.e.
+for them to be both downward-closed or upward-closed, and we explore both possiblities.
+
+\subsection{Case: $\Bad=\downarrow \Bad$ and $\Safe=\downarrow \Safe$}\label{case down down}
 
 % In the case \Safe and \Bad are both downward-closed, they can both be finite.
+We start with the case $\Safe = \downarrow \Safe$ and $\Bad = \downarrow \Bad$.
+%
+Unfortunately, in this case the resilience problem is undecidable.
 
 \begin{theorem}\label{down-down}
-The resilience problem is undecidable for  effective WSTS with  strong  compatibility such that
+The {\sc Resilience problem} is undecidable for  effective WSTS with  strong  compatibility such that
 %	with more than one minimal element, 
@@
 
-\section{Case: $\Bad=\uparrow \Bad$ and $\Safe=\uparrow \Safe$.}\label{case up up}
+\subsection{Case: $\Bad=\uparrow \Bad$ and $\Safe=\uparrow \Safe$.}\label{case up up}
 
@@
 
-Since $\Bad \cap \Safe = \emptyset$, the subcase $\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction since two upward-closed sets have not an empty intersection.
-We then consider the case where $\Bad$ is as upward-closed
+We now consider the case where $\Safe = \uparrow \Safe$ and $\Bad$ is not the complement of 
+$\Safe$ and we'd want $\Bad$ to have the same closure properties as $\Safe$.
+We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the subcase 
+$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction, since two upward-closed sets have not an empty intersection.
+We thus consider instead the case where $\Bad$ is as upward-closed
 as it can be without intersecting with $\Safe$,
@@
 such that $\Bad = U \setminus \Safe$.
-
-\iffalse
-\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
-\mathieu{Pas convaincu que ça serve, je vais enlever}
-% hop
-\fi
-
-example: ?
 

### 2023-06-16 resilience-safe-down.tex
up

--- 2023-06-16 resilience-safe-down.tex
@@
 for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed, the problem becomes undecidable. Indeed, it is 
-undecidable for Reset 
-%Petri-nets
-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
+undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
 This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.

### 2023-06-16 appendix.tex
up

@@
 
-\section{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+\section{Reductions to reset-VASS}
 
+Most of our undecidability results make use of reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
+
+Reachability is undecidable for reset-VASS because, with
+
+\subsection{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+
+As mentionned in Section~$3$, {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$. This is due to 
 Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
@@
 
-\section{Reset-VASS and state-resilience}\label{appendix}
+\subsection{Reset-VASS and state-resilience}\label{appendix}
 

### 2023-06-16 appendix.tex
up

--- 2023-06-16 appendix.tex
@@
 
-\section{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+\section{Reductions to reset-VASS}
 
+Most of our undecidability results make use of reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
+This has consequences for {\sc Resilience}.
+For instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element.
+Even then, though, {\sc Resilience} (resp.
+{\sc State-resilience}) is not strictly reachability, and we will need to delve more into the details to obtain undecidability results.
+
+This section is dedicated to studying reset-VASS for showing undecidability of {\sc Resilience} and {\sc State-resilience} for WSTS in general.
+
+\subsection{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+
+As mentionned in Section~$3$, {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$. 
 Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).

### 2023-06-16 appendix.tex
up

@@
 
-As mentionned in Section~$3$, {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$. 
-Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
+We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
+Let us provide a more detailed undecidability proof here.
+
+First, let us recall the home-state problem: 
+whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
+
+Home-state is undecidable for Minsky machine with $3+$ counters.
+
+Because $2$-counter Minsky machine termination is undecidable.
+
+$2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ which 
+third counter starts at least at $1$, and simulates $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
+
+Then the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+
+% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
 % or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
 
+Thus home-state/resilience is undecidable for Minsky machines.
+
+Executions of Minsky machines can be simulated by reset-VASS. 
+% Let us recall more formally what a reset-VASS is.
 Reset-VASS extend the basic VASS model with special “reset
@@
 A {\em reset-VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
-% $\mathds{Z}^d$
 labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
-%$T \subseteq Q \times \mathds{Z}^d \times Q$
 $T \subseteq Q \times Op \times Q$
@@
 It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
+Reset-VASS can simulate executions of a Minsky machine.
+
+Thus home-state/resilience is undecidable for reset-VASS and hence for WSTS as well.
 

### 2023-06-16 state-resilience.tex
up

@@
 \problemx{State-resilience problem (SRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
 % {$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
@@
 \problemx{$k$-state-resilience problem (kSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
 % { $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
@@
 \problemx{bounded-state-resilience problem (BSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
 {$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient ?\newline}
@@
 
-We start with the case $\Safe = \uparrow \Safe$.
+We start with the case $\Safe = \uparrow \Safe$ hence $\Bad = \downarrow \Bad$.
 %
@@
 
-We now consider the case $\Safe = \downarrow \Safe$.
+We now consider the case $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
 %

### 2023-06-16 application.tex
section 3

@@
-\section{Computing the maximal resilient subsystem}
-
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
-
-\begin{definition}
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
-\end{definition}
-
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
-The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
-%\alain{vérifier que la plus grande existe}
-Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
-Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
-Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
-%
-For example, VASS allows such guards.
-  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
-
-\begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
-\end{theorem}
-
-\begin{proof}
-Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
-%
-%
-Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
 \end{proof}
@@
 %
+
+\section{Resilience for VASS}
+
 Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
@@
 Let us remark that the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$, the maximal $X$-resilient VASS of $V$ is also a computable VASS $V_X$ and that $V_X$ is independant of the initial state $v_0$.
-
-\section{Resilience for VASS}
 

### 2023-06-16 resilience.tex
section 3

@@
 \end{center}
+
+\subsection{Computing the maximal resilient subsystem}
+
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+
+\begin{definition}
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
+\end{definition}
+
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
+The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
+%\alain{vérifier que la plus grande existe}
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
+Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
+%
+For example, VASS allows such guards.
+  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
+\end{theorem}
+
+\begin{proof}
+Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+%
+%
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.

### 2023-06-16 resilience.tex
up
section 3

@@
 \begin{definition}
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient (resp. $(s_0,X)$-resilient) if $S  \xrightarrow{*} X$ (resp. $\post^*(s_0) \xrightarrow{*} X$) is satisfied.  
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient if $S  \xrightarrow{*} X$ is satisfied.  
 \end{definition}
@@
 Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
+
+It is also possible to define $(s_0,X)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} X$.

### 2023-06-16 definitions.tex
up

--- 2023-06-14 definitions.tex
@@
 %
+As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
+%
 Quasi-orderings that have no 
@@
  enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
-
+%
+%
 In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement), and an upward-closed set $U$ is represented by its finite set of minimal elements.
@@
 Let us now recall the (most general) definition of well-structured transition systems.
-
 \begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
@@
 
-\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est un WSTS}
-
 Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
+The lossy VASS given in Example~\ref{Example} ({(\bf Flight Controller System)}), for instance, is a WSTS.
+ \mathieu{Faire la preuve une fois que j'ai l'equation}
 
@@
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-
-\mathieu{Mentionner que le Lossy VASS de l'example~\ref{Example} est effectif}
+%
+% \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est effectif}
+As an illustration, the lossy VASS given in Example~\ref{Example} ({(\bf Flight Controller System)}) is post-effective,
+ideal-effective, and
+has effective pred-basis.
 

### 2023-06-16 resilience-safe-down.tex
up

@@
 $\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$
 and
@@
 The resilience problem is decidable for ideal-effective downward-compatible WSTS with 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$.
+$\Safe=\downarrow \Safe$.
 \end{corollary}

### 2023-06-17 reachability/article.tex
section

@@
 
-Stack-based automata are a popular formalism for modeling the sequential behavior of computer programs. 
-Pushdown systems are probably the most studied models of stack-based automata.
-They are
+{\bf Context}
+Stack-based automata \textcolor{red}{pourquoi ne pas commencer directement avec lesPushdown systems ?}are a popular formalism for modeling the sequential behavior of computer programs. 
+Pushdown systems \textcolor{red}{or stack-automata ?}are probably the most studied models of stack-based automata.
+\textcolor{red}{inutile: They are
 finite automata 
- extended with a stack that can be manipulated by pushing or popping stack symbols from some finite set, and can use the top of the stack to decide which transition to take next.
-\textcolor{black}{To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
-Pushdown systems can be used to model the behavior of recursive programs and,
+ extended with a stack that can be manipulated by pushing or popping stack symbols from some finite set, and can use the top of the stack to decide which transition to take next.}
+\textcolor{blue}{ok je comprends que tu signales la différence mais c'est trop détaillé pour cet endroit du texte: To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
+
+{\bf Games on pushdown systems}Pushdown systems can be used to model the behavior of recursive programs and,
 as a result, a variety of 
@@
 
+{\bf Parametric pushdown automata or systems ?}
  Parametric pushdown automata provide a  formalism to
@@
 
+{\bf Extensions of  pushdown automata/systems ?}
 Extensions of pushdown automata with storage for later comparisons

### 2023-06-17 reachability/article.tex
section

@@
 They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages),
-and have
+and have \textcolor{red}{ phrases trop longues}
 been used to establish that CTL$^*$ model checking remains decidable
@@
 
-\subsubsection*{Contributions}
+\subsubsection*{Our contributions}
 

### 2023-06-19 definitions.tex
up

@@
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
-The lossy VASS given in Example~\ref{Example} ({(\bf Flight Controller System)}), for instance, is a WSTS.
+The lossy VASS given in Example~\ref{Example} ({\bf Flight Controller System}), for instance, is a WSTS.
  \mathieu{Faire la preuve une fois que j'ai l'equation}
@@
 % \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est effectif}
-As an illustration, the lossy VASS given in Example~\ref{Example} ({(\bf Flight Controller System)}) is post-effective,
+As an illustration, the lossy VASS given in Example~\ref{Example} ({\bf Flight Controller System}) is post-effective,
 ideal-effective, and

### 2023-06-19 definitions.tex
up

@@
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
-The lossy VASS given in Example~\ref{Example} ({\bf Flight Controller System}), for instance, is a WSTS.
+The lossy VASS given in the {\bf Flight Controller System} example, for instance, is a WSTS.
  \mathieu{Faire la preuve une fois que j'ai l'equation}
@@
 % \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est effectif}
-As an illustration, the lossy VASS given in Example~\ref{Example} ({\bf Flight Controller System}) is post-effective,
+As an illustration, the lossy VASS given in the {\bf Flight Controller System} example is post-effective,
 ideal-effective, and

### 2023-06-19 introduction.tex
named the ex
up

@@
 	\begin{figure}
-			\hspace{0.74cm}
-\includegraphics[width=0.87\textwidth]{FigureB}
+			\hspace{0.75cm}
+\includegraphics[width=0.85\textwidth]{FigureB}
 	\caption{A lossy VASS with three subsystems and four counters.
@@
 \begin{example}\label{Example}
+{(\bf Flight Controller System).}
 Consider the vector addition system in Figure~\ref{air control}. It models a scenario in which two aircrafts want to land in an airport at the same time. 
@@
 We assume that the VASS is
-lossy \-- i.e. 
- the contents of a counter can spontaneously get lower at any time
- \--  but that they are still such that out of ten incrementations of a counter, at least one is not rendered void.
+lossy, i.e. 
+ the contents of a counter can spontaneously get lower at any time.
+ Lossiness is represented by using transitions decreasing any counter by one from every state. 
+We ask however that the system is not too lossy.
+We add the condition that for every prefix of an execution, the number of ``lossiness transition'' for a counter is less than nine tenth of the number of transitions incrementing said counter.
 In this example, the safe, desirable state would be one where the aircrafts have both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4\}$, and $\Bad$ is the complement of $\Safe$. 
+$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$, and $\Bad$ is the complement of $\Safe$. 
 % We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.

### 2023-06-20 reachability/article.tex
up

@@
 
-\keywords{Parity Games, Computational Complexity, Pushdown systems, Pebble Alternating Two Way Automata}
+\keywords{Parity Games, Computational Complexity, Pushdown systems}
 
@@
 	and compared against the current stack content.
-	We determine decidability of the problem, and provide a nonelementary lower bound. 
+	We determine the problem
+	belongs to $(n+1)$-$\EXP$ in case the number of parameters
+$n$ is fixed, and provide a nonelementary lower bound for the general case. 
 \end{abstract}

### 2023-06-20 reachability/article.tex
up

@@
 
-{\bf Context}
-Stack-based automata \textcolor{red}{pourquoi ne pas commencer directement avec lesPushdown systems ?}are a popular formalism for modeling the sequential behavior of computer programs. 
-Pushdown systems \textcolor{red}{or stack-automata ?}are probably the most studied models of stack-based automata.
-\textcolor{red}{inutile: They are
-finite automata 
- extended with a stack that can be manipulated by pushing or popping stack symbols from some finite set, and can use the top of the stack to decide which transition to take next.}
-\textcolor{blue}{ok je comprends que tu signales la différence mais c'est trop détaillé pour cet endroit du texte: To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
+% {\bf Context}
 
@@
 strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
+\textcolor{blue}{ok je comprends que tu signales la différence mais c'est trop détaillé pour cet endroit du texte: To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
 

### 2023-06-20 reachability/article.tex
up

@@
 
-{\bf Games on pushdown systems}Pushdown systems can be used to model the behavior of recursive programs and,
+{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs and,
 as a result, a variety of 
@@
 specific words. 
-They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages),
-and have \textcolor{red}{ phrases trop longues}
+They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
+They have \textcolor{red}{ phrases trop longues}
 been used to establish that CTL$^*$ model checking remains decidable

### 2023-06-20 reachability/article.tex
up

@@
 ANR-17-CE40-0010.}
-\author{Mathieu Hilaire}{Université Paris-Saclay\and
-CNRS\and
-ENS Paris-Saclay\and
-Laboratoire Méthodes Formelles (LMF)\and
+\author{Mathieu Hilaire}{Université Paris-Saclay,
+CNRS,
+ENS Paris-Saclay,
+LMF,
 Gif-sur-Yvette, France
@@
 
-\acknowledgements{The author would like to thank Benedikt Bollig and Stefan G\"oller for helpful discussions and feedback.}
+\acknowledgements{The author would like to thank Benedikt Bollig, Stefan G\"oller and Alain Finkel for helpful discussions and feedback.}
 

### 2023-06-20 reachability/article.tex
up

@@
 {\bf Parametric pushdown automata or systems ?}
- Parametric pushdown automata provide a  formalism to
+ Parametric pushdown systems provide a  formalism to
 reason about 
@@
 % Here, the stack can additionally be compared against parameters that can take unspecified word values. 
-Recently, different variants of parametric pushdown automata have been introduced in the literature 
+Recently, different variants of parametric pushdown systems have been introduced in the literature 
 % [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
@@
 These systems consist of a leader pushdown automaton
-and arbitrarily many identical contributor pushdown automata, communicating via a shared memory
+and arbitrarily many identical contributor pushdown systems, communicating via a shared memory
 in the form of a register which can take finitely many values.
@@
 to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
-We consider here a different approach to extending pushdown automata with parameters, by allowing them to test equality of the stack content against parameters.
+We consider here a different approach to extending pushdown systems with parameters, by allowing them to test equality of the stack content against parameters.
 % Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
 Perhaps a similar model consists in 
-pushdown automata 
+pushdown systems 
 with transitions that are conditioned by regular conditions on their stack content, which
@@
 for
-pushdown automata \cite{finkel1997direct}.
-Pushdown automata 
+pushdown systems \cite{finkel1997direct}.
+Pushdown systems 
 %with regularity condition on the stack
@@
 {\bf Extensions of  pushdown automata/systems ?}
-Extensions of pushdown automata with storage for later comparisons
-have been introduced for instance as register pushdown automata. 
+Extensions of pushdown systems with storage for later comparisons
+have been introduced for instance as register pushdown systems. 
 Such automata possess registers in addition to their stack, and can keep data values in both.
-Register pushdown automata have been shown to have applications for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}. 
-In theory, since registers can be unfilled and later be filled again an unbounded number of times, the number of different data values a pushdown register automata can store in its registers in a run is unbounded.
+Register pushdown systems have been shown to have applications for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}. 
+In theory, since registers can be unfilled and later be filled again an unbounded number of times, the number of different data values a pushdown register systems can store in its registers in a run is unbounded.
 In \cite{murawski2017reachability} it was shown however that a register pushdown automaton can
 only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers,
-% Even though pushdown register automata have no bound on the number of elements of the alphabet that they can store at any instant, over the course of a run they can really “remember” at most 3r of them,
+% Even though pushdown register systems have no bound on the number of elements of the alphabet that they can store at any instant, over the course of a run they can really “remember” at most 3r of them,
 i.e. for any run of a register pushdown automaton with $r$ registers there exists an equivalent run %, i.e.
@@
 
-We provide formal definitions of parametric pushdown automata, parametric pushdown reachability games
+We provide formal definitions of parametric pushdown systems, parametric pushdown reachability games
 and parametric pushdown parity games in Section~\ref{section def ppda}. We give an overview of our contribution in Section~\ref{contribution ppda}. The contribution consists in
@@
 
-	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown automata.
- We then show that higher-order pushdown automata parity games can be used to solve parity games on
- the transition systems of pebble pushdown automata,
+	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown systems.
+ We then show that higher-order pushdown systems parity games can be used to solve parity games on
+ the transition systems of pebble pushdown systems,
 using one additional stack level for each pebble.
- Since solving parity games on higher-order pushdown automata with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown automata with $n$ parameters.
+ Since solving parity games on higher-order pushdown systems with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown systems with $n$ parameters.
 
@@
 Section 3 will deal with the introduction of  parametric pushdown games.
-% Section 4 is devoted to using pebble alternating tree automata with parity acceptance condition to solve parametric pushdown system parity games.
+% Section 4 is devoted to using pebble alternating tree systems with parity acceptance condition to solve parametric pushdown system parity games.
 We leave the proof of decidability itself  for Section 5.
-Section 4 is Reduction to higher-order pushdown automata
+Section 4 is Reduction to higher-order pushdown systems
 Section 5 at last shows a nonelementary lower bound for the problem.
@@
 %\mh{
-Of note is that several model checking problems %for pushdown automata 
+Of note is that several model checking problems %for pushdown systems 
 can be expressed as decision problems for games: the most famous example is that the modal $\mu$-calculus model checking problem is polynomially equivalent to solving the parity game problem. 
@@
 
-\textcolor{black}{Parametric pushdown automata extend pushdown automata by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
+\textcolor{black}{Parametric pushdown systems extend pushdown systems by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
 A parametric pushdown automaton is then a finite automaton extended with a finite set of parameters $P$  and with a stack that can be manipulated by pushing or popping stack symbols and such that, moreover, the automaton can use the top of the stack, or check that the stack content corresponds to a parameter,  to decide which transition to take next.\\
@@
 \par\noindent\ignorespacesafterend
-A {\em parametric pushdown automaton} ({\em PPDA} for short) 
+A {\em parametric pushdown system} ({\em PPDS} for short) 
 is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
@@
 %
-As with PDA, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
+As with pushdown systems, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
 and $q_n\in F$.
@@
 
-\subsection{PPDA reachability}
-
-Note that the reachability problem for parametric pushdown automata consists in
+\subsection{PPDS reachability}
+
+Note that the reachability problem for parametric pushdown systems consists in
 the following decision problem.
 
-\problemx{$n$-PPDA reachability}
-{An $n$-PPDA $\mathcal{Z}$.}
+\problemx{$n$-PPDS reachability}
+{An $n$-PPDS $\mathcal{Z}$.}
 {Does 
@@
 % \subsubsection{Parametric pshdown games}
-\subsection{PPDA games}
+\subsection{PPDS games}
 
@@
 For instance 
-in the case of pushdown automata, 
+in the case of pushdown systems, 
 the modal $\mu$-calculus
@@
 (\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})\]
-induced by a PPDA $\mathcal{Z}$, a partition of its set of states, and a
+induced by a PPDS $\mathcal{Z}$, a partition of its set of states, and a
 parameter valuation $\mu: P\rightarrow \Gamma^*$.
@@
 \section{%A decidability result via r
-Reduction to pebble pushdown automata parity games
+Reduction to pebble pushdown systems parity games
 }\label{pebbles}
 
-% \mh{Motivates the use of pebbles. See pebbles on two-way automata, two-way tree automata, weighted automata, etc.}
-
-% 21. Globerman, N., Harel, D.: Complexity results for two-way and multi-pebble automata and their logics. Theor. Comput. Sci. 169, 161–184 (1996)
-%		same as [13] bellow
-%
-% Engelfriet, J., Hoogeboom, H.J.: Tree-walking pebble automata.
-In: Karhumäki, J., Maurer, H., Pa ̆un, G., Rozenberg, G.
-(Eds.) Jewels are Forever, Contributions to Theoretical Computer Science in Honor of Arto Salomaa, pp.
-72–83.
-Springer, Berlin (1999) for tree-walking automata.
-%
-% @incollection{engelfriet1999tree,
-%  title={Tree-walking pebble automata},
-%  author={Engelfriet, Joost and Hoogeboom, Hendrik Jan},
-%  booktitle={Jewels are forever},
-%  pages={72--83},
-%  year={1999},
-%  publisher={Springer}
-%}
-% 
-%
-% Two-way pebble transducers for partial functions and their composition, Joost Engelfriet
-%
-% Tree-Walking Pebble automata, Joost Engelfriet and Hendrik Jan Hoogeboom 
 A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
@@
 
-In this section, we extend these ideas to pushdown automata.
-Instead of using pebbles as markings on their input, pebble pushdown automata have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
+In this section, we extend these ideas to pushdown systems.
+Instead of using pebbles as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
 and a \textit{lift} pops the last dropped pebble 
@@
 %\subsubsection{Pebble pushdown automata}
-\subsection{Pebble pushdown automata}
+\subsection{Pebble pushdown systems}
 
@@
 \par\noindent\ignorespacesafterend
-We are interested in games over pebble pushdown automata, % namely, 
+We are interested in games over pebble pushdown systems, % namely, 
 %reachability, B\"uchi, and parity games.
@@
 
-Concerning pebble pushdown automata, we are interested in the following decision problem.
+Concerning pebble pushdown systems, we are interested in the following decision problem.
 
@@
 
-% \subsubsection{Reduction from parameters to pebble pushdown automata}
+% \subsubsection{Reduction from parameters to pebble pushdown systems}
 \subsection{From parametric pushdown 
-automata 
-to pebble pushdown automata
+systems 
+to pebble pushdown systems
 }
@@
 
-\section{Reduction to higher-order pushdown automata}\label{section HPDA}
-
-Higher-order pushdown automata (HPDA for short) were introduced as a 
-generalization of pushdown automata \cite{aho1969nested,Gre70,maslov1976multilevel}. 
-% Alfred V. Aho. Nested stack automata. J. ACM, 16(3):383–406, 1969.
+\section{Reduction to higher-order pushdown systems}\label{section HPDA}
+
+Higher-order pushdown systems (HPDS for short) were introduced as a 
+generalization of pushdown systems \cite{aho1969nested,Gre70,maslov1976multilevel}. 
+% Alfred V. Aho. Nested stack systems. J. ACM, 16(3):383–406, 1969.
 % S. Greibach. Full AFL’s and nested iterated substitution. Inf. Control, 16(1):7–35, 1970.
-% A.N. Maslov. Multilevel stack automata. Problemy Peredachi Informatsii, 12:55–62, 1976.
+% A.N. Maslov. Multilevel stack systems. Problemy Peredachi Informatsii, 12:55–62, 1976.
 A stack of a pushdown automaton is seen as a {\em level $1$ stack}. 
-A pushdown automaton of level $2$ (or $2$-HPDA) then works with a stack of level $1$ stacks. 
+A pushdown automaton of level $2$ (or $2$-HPDS) then works with a stack of level $1$ stacks. 
 In addition to the ability to push
-and to pop a symbol on the top-most level $1$ stack, an $2$-HPDA can copy or remove the entire topmost level $1$ stack. 
-The definition generalizes to any $n \geq 2$, and $n$-HPDA are similarly defined for all level
-$n$ as automata working with a stack of level $(n-1)$ stacks.
-% HPDA have been extensively studied as language recognizers [Dam82,Eng83].
+and to pop a symbol on the top-most level $1$ stack, an $2$-HPDS can copy or remove the entire topmost level $1$ stack. 
+The definition generalizes to any $n \geq 2$, and $n$-HPDS are similarly defined for all level
+$n$ as systems working with a stack of level $(n-1)$ stacks.
+% HPDS have been extensively studied as language recognizers [Dam82,Eng83].
 
@@
 
-\subsection{Higher-order pushdown automata}
+\subsection{Higher-order pushdown systems}
 
@@
 \par\noindent\ignorespacesafterend
-A {\em higher-order pushdown automata of level $n$} (or $n$-HPDA for short) 
+A {\em higher-order pushdown systems of level $n$} (or $n$-HPDS for short) 
 is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
@@
 By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
-we denote the set of {\em configurations} of an $n$-HPDA $\H$. 
+we denote the set of {\em configurations} of an $n$-HPDS $\H$. 
 %is a pair $(q, s)$ where $q \in Q$ and $s$ is an $n$-stack over $\Gamma$. 
@@
 
-An $n$-HPDA $\H=(Q, \Gamma, R, q_{init}, F)$
+An $n$-HPDS $\H=(Q, \Gamma, R, q_{init}, F)$
 induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
@@
 As expected, given an 
- $n$-HPDA $\mathcal{H}$
+ $n$-HPDS $\mathcal{H}$
  and a partition of
@@
 \par\noindent\ignorespacesafterend
-induced by an $n$-HPDA $\mathcal{H}$ and a partition of its set of states.
+induced by an $n$-HPDS $\mathcal{H}$ and a partition of its set of states.
 
@@
 
-Concerning higher-order pushdown automata, we are interested in the following decision problem.
+Concerning higher-order pushdown systems, we are interested in the following decision problem.
 
 \begin{samepage}
-\problemx{$n$-HPDA parity game}
-{An $n$-HPDA $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
+\problemx{$n$-HPDS parity game}
+{An $n$-HPDS $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
 where $Q = Q_0  \biguplus Q_1 $, 
@@
 
-It was shown in \cite{Cach03} that {\sc $n$-HPDA parity game} can
+It was shown in \cite{Cach03} that {\sc $n$-HPDS parity game} can
 be solved in $n$-$\EXP$. This also gives an $n$-$\EXP$ algorithm for the $\mu$-calculus
-model checking over transitions systems induced by  $n$-HPDA.
+model checking over transitions systems induced by  $n$-HPDS.
 In \cite{cachat2007complexity} the matching lower bound was showed, even in the case of reachability games,
-hence showing $n$-$\EXP$-completeness of {\sc $n$-HPDA parity game}.
+hence showing $n$-$\EXP$-completeness of {\sc $n$-HPDS parity game}.
 
 \begin{theorem}{\cite{ Cach03, cachat2007complexity}}
-{\sc $n$-HPDA parity game} is $n$-$\EXP$-complete.
+{\sc $n$-HPDS parity game} is $n$-$\EXP$-complete.
 \end{theorem}
 
-% \subsection{From pebble pushdown automata to higher-order pushdown automata}
-
-In a $2$-HPDA, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
+% \subsection{From pebble pushdown systems to higher-order pushdown systems}
+
+In a $2$-HPDS, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
 allows to ``copy'' the top level $1$ stack. The current word is hereby stored away and left untouched
@@
 \iffalse
-In an $n$-HPDA, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
+In an $n$-HPDS, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
 allows to ``copy'' the top stack of level $j$ for all $2 \leq j \leq k$. 
@@
 Carayol and W\"ohrle
- introduced a variant of $n$-HPDA by extending the 
+ introduced a variant of $n$-HPDS by extending the 
 $\text{\textit{pop}}_j$ operations for $2 \leq j \leq n$ with a built-in equality test. 
@@
 
-A {\em higher-order pushdown automaton with equality pop of level $n$} ($n$-HPDA$^=$ for short) 
+A {\em higher-order pushdown automaton with equality pop of level $n$} ($n$-HPDS$^=$ for short) 
 is a higher-order pushdown automaton of level $n$ where in
@@
 %
-Transition systems induced by higher-order pushdown automata with equality pop of level $n$
+Transition systems induced by higher-order pushdown systems with equality pop of level $n$
 are then defined as expected.  
@@
 namely
-higher-order pushdown automata with equality pop of level $n$
+higher-order pushdown systems with equality pop of level $n$
 and
-higher-order pushdown automata of level $n$,
+higher-order pushdown systems of level $n$,
 generate the same classes of transition systems.
@@
 \begin{theorem}{\cite{Woeh05, carayol2006automates}}
-If $\H$ is an $n$-HPDA (resp. $n$-HPDA$^=$) then there
-exists an $n$-HPDA$^=$ (resp. $n$-HPDA) $\H'$ such that
+If $\H$ is an $n$-HPDS (resp. $n$-HPDS$^=$) then there
+exists an $n$-HPDS$^=$ (resp. $n$-HPDS) $\H'$ such that
 $\H$ and $\H'$ %generate the same graph.
@@
 \begin{theorem}{\cite{Woeh05, carayol2006automates}}
-If $\H$ is an $n$-HPDA  then there
-exists an $n$-HPDA$^=$  $\H'$ such that
+If $\H$ is an $n$-HPDS  then there
+exists an $n$-HPDS$^=$  $\H'$ such that
 $\H$ and $\H'$ %generate the same graph.
@@
 \begin{theorem}{\cite{Woeh05, carayol2006automates}}
-If $\H$ is an $n$-HPDA$^=$ then there
-exists an $n$-HPDA $\H'$ such that
+If $\H$ is an $n$-HPDS$^=$ then there
+exists an $n$-HPDS $\H'$ such that
 $\H$ and $\H'$ %generate the same graph.
@@
 
-From $n$-HPDA to $n$-HPDA$^=$ (Proposition 3.12 in \cite{Woeh05}) the proof relies on recreating a correct stack content to simulate higher-order 
+From $n$-HPDS to $n$-HPDS$^=$ (Proposition 3.12 in \cite{Woeh05}) the proof relies on recreating a correct stack content to simulate higher-order 
 $\text{\textit{pop}}$ operations by $\text{\textit{pop}}^=$, essentially ``guessing'' the correct stack content to be able to apply $\text{\textit{pop}}^=$. 
@@
 
-Defined like {\sc $n$-HPDA parity game}, we write {\sc $n$-HPDA$^=$ parity game}  
-in case the input is an $n$-HPDA$^=$ rather than an $n$-HPDA.
+Defined like {\sc $n$-HPDS parity game}, we write {\sc $n$-HPDS$^=$ parity game}  
+in case the input is an $n$-HPDS$^=$ rather than an $n$-HPDS.
 
@@
 to \cite{Cau02mfcs} for more details.
-The $n$-$\EXP$ upper bound on $n$-HPDA parity games follows from the fact that every transition system induced by an $n$-HPDA $\H$ is a graph of the Caucal hierarchy \cite{Cach03, Woeh05},
+The $n$-$\EXP$ upper bound on $n$-HPDS parity games follows from the fact that every transition system induced by an $n$-HPDS $\H$ is a graph of the Caucal hierarchy \cite{Cach03, Woeh05},
 whose vertices are almost in one-to-one correspondence with the configurations of $\H$.
 In \cite{Woeh05} the converse direction is proven, i.e. that every graph of level $n$
-of the Caucal hierarchy is generated by an $n$-HPDA.  
+of the Caucal hierarchy is generated by an $n$-HPDS.  
 %
@@
 it is similarly shown that
-every transition system induced by an $n$-HPDA$^=$ $\H$ is a graph of the Caucal hierarchy.
+every transition system induced by an $n$-HPDS$^=$ $\H$ is a graph of the Caucal hierarchy.
 %
@@
 The algorithm from \cite{Cach03} hence lead to a solution for
-$n$-HPDA$^=$ parity games as well.
+$n$-HPDS$^=$ parity games as well.
 
@@
 \textcolor{red}{
-Now then, because the class of $n$-HPDA$^=$ transitions systems
+Now then, because the class of $n$-HPDS$^=$ transitions systems
 is the same as
-the class of $n$-HPDA transitions systems, we have the following result.
+the class of $n$-HPDS transitions systems, we have the following result.
 }
@@
 \begin{theorem}\label{HPDA parity game}
-{\sc $n$-HPDA$^=$ parity game} is in $n$-$\EXP$.
+{\sc $n$-HPDS$^=$ parity game} is in $n$-$\EXP$.
 \end{theorem}
@@
 \subsection{From pebble pushdown 
-automata 
+systems 
 to higher-order pushdown 
-automata
+systems
 }
 
-We use HPDA$^=$ parity games to simulate pebble pushdown automata
+We use HPDS$^=$ parity games to simulate pebble pushdown systems
 parity games.
 A similar approach was used in \cite{carayol2006automates} to show that $(n + 2)$-level stack
-automata could be used to simulate $n$-pebble alternating two-way word automata.
+systems could be used to simulate $n$-pebble alternating two-way word systems.
 
@@
 a pebble pushdown automaton $\ppda$ with only one pebble.
-As long as no pebble is dropped, a $2$-HPDA$^=$ $\H$ can simulate 
+As long as no pebble is dropped, a $2$-HPDS$^=$ $\H$ can simulate 
 the behavior of a pebble pushdown automaton $\ppda$ with a $2$-stack containing a single $1$-stack
@@
 $\ppda = (Q, \Gamma, R, q_{init}, \gamma_{init}, F)$, where $Q= Q_0 \biguplus Q_1$,
-one can compute an $(n+1)$-HPDA 
+one can compute an $(n+1)$-HPDS 
 $\H = (Q', \Gamma, R', q_{init}', \gamma_{init}', F')$  where $Q'= Q_0' \biguplus Q_1'$, 
@@
 {\sc $n$-pebble pushdown parity game} is polynomial time reducible
-to {\sc $(n+1)$-HPDA$^=$ parity game}.
+to {\sc $(n+1)$-HPDS$^=$ parity game}.
 \end{theorem}
 
-The reduction implies decidability of pebble pushdown automata parity game. Furthermore,
+The reduction implies decidability of pebble pushdown systems parity game. Furthermore,
 by Theorem~\ref{HPDA parity game}, it implies the following complexity result.
@@
 \begin{theorem}
-{\sc $n$-pebble pushdown automata parity game} is in $(n+1)$-$\EXP$.
+{\sc $n$-pebble pushdown systems parity game} is in $(n+1)$-$\EXP$.
 \end{theorem}
@@
 \begin{corollary}
-{\sc $n$-parametric pushdown automata parity game} is in $(n+1)$-$\EXP$.
+{\sc $n$-parametric pushdown systems parity game} is in $(n+1)$-$\EXP$.
 \end{corollary}
@@
 
-	For the decidability upper bound, we  used pebble alternating two-way tree automata to solve pebble pushdown parity games.
-Alternating two-way tree automata (ATWA) with parity winning conditions have been used to solve parity games on pushdown graphs \cite{cachat2002two}.
-The technique was first developped to solve model-checking problem for pushdown graphs in \cite{kupferman2000automata}.
-We showed that, when pushdown graphs and ATWA are extended with a set of pebbles to manage, the technique
-also works.
-We then proved decidability 
-% for checking that an $n$-PATWA $\mathcal{A}$ accepts the full infinite tree $\{0,1\}^*$ 
-by %reducing it
-a reduction
- to the problem of model-checking $\MSO$ on a well chosen structure. To ensure decidability of model-checking $\MSO$ on the chosen structure, we made use of 
-Muchnik's Theorem: the structure is constructed by applying a successive number of times Muchnik's iteration to a structure $\mathfrak{A}_0$ for which model-checking $\MSO$ is decidable.
-%  After providing this structure, we will show that  checking that an $n$-PATWA $\mathcal{A}$ accepts the full binary tree $\{0,1\}^*$ reduces itself to the problem of model-checking $\MSO$ on it.
+	For the decidability upper bound, we
 

### 2023-06-20 reachability/article.tex
upg
up

@@
 
-{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs and,
-as a result, a variety of 
+{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs.
+As a result, a variety of 
  program analysis questions can be reduced to decision problems for games on pushdown systems.
@@
 
-{\bf Parametric pushdown automata or systems ?}
+{\bf Parametric pushdown systems}
  Parametric pushdown systems provide a  formalism to
@@
 
-\subsubsection*{Overview}
-
+{\bf Overview. }
 In Section 2, we provide general notations and preliminary definitions. 
@@
 It is
-denoted by $\pi(s, \sigma_0, \sigma_1)$.\\
+denoted by $\pi(s, \sigma_0, \sigma_1)$.
 
@@
 
-	For the decidability upper bound, we
+	For the upper bound, we replaced parameters by pebbles acting as registers, leading to a more general model, and then used a reducton to higher-order pushdown systems parity games to solve parity games.
+Since solving parity games on higher-order pushdown systems with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown systems with $n$ parameters.
 

### 2023-06-20 reachability/article.tex
up

@@
 
-{\bf Extensions of  pushdown automata/systems ?}
+{\bf Related models.}
 Extensions of pushdown systems with storage for later comparisons

### 2023-06-21 abstract.tex
section abstract

--- 2023-06-13 abstract.tex
@@
-We investigate and expand upon the notion of resilience as defined in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
-We introduce a more general definition of resilience and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
-and the $k$-(state-)resilience problem (kRP). The (general) resilience problem 
-% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
-asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$. 
-We remark that it is an instance of the home space problem defined in the framework of VASS.
-State resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.
-
-In a well-structured framework we show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility. Moreover, state-resilience,
-bounded-state-resilience and
- $k$-state-resilience
-are undecidable for strongly upward-compatible WSTS with effective pred-basis.
-
-On the positive side, we find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$, and moreover the resilience problem is decidable for ideal-effective WSTS with
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$
-and
-the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
-
-We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
- WSTS with effective 
-$\uparrow$ $\post^*$ basis
-when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
-$\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
-
-Finally, we study the resilience for VASS, where most problems are shown decidable.
+\noindent
+Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
+We consider six types of resilience (one of them is the home-state property) defined by the two subsets Bad and Safe and by the existence of a bound on the length of runs starting from Bad and reaching Safe.
+We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
+We first show that most of all resilience problems are undecidable for WSTS with strong compatibility and with effective pred-basis.
+Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility.
+Moreover, some resilience properties are decidable for two other classes of WSTS : (1) for WSTS with effective
+$\uparrow$ $\post^*$ basis and (2) for ideal-effective WSTS with downward and upward compatibilities. Finally, we study the resilience for VASS where most of the problems are shown decidable.

### 2023-06-21 introduction.tex
section abstract

@@
 {\bf Context.} 
-Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain).
+Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges}).
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
@@
 
+State resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.
+
+We investigate and expand upon the notion of resilience as defined by Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, which asks whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
+We introduce a more general definition of resilience and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
+and the $k$-(state-)resilience problem (kRP). The (general) resilience problem 
+% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
+asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$.
+
+In a well-structured framework we show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility. Moreover, state-resilience,
+bounded-state-resilience and
+ $k$-state-resilience
+are undecidable for strongly upward-compatible WSTS with effective pred-basis.
+
+On the positive side, we find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
+and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$, and moreover the resilience problem is decidable for ideal-effective WSTS with
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$
+and
+the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+
+We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
+ WSTS with effective 
+$\uparrow$ $\post^*$ basis
+when
+$\Safe=\uparrow \Safe$
+and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
+$\uparrow$ $\post^*$ basis hypothesis leads to undecidability, but we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
+
+Finally, we study the resilience for VASS, where most problems are shown decidable.
 %

### 2023-06-21 abstract.tex
section abstract

@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-state property) defined by the two subsets Bad and Safe and by the existence of a bound on the length of runs starting from Bad and reaching Safe.
+We consider six types of resilience (one of them is the home-space property) defined by Bad is an upward/downward closed set or it is a subset of the reachability set from an initial state, Safe is upward-closed ($\Safe=\uparrow \Safe$) or Safe is downward-closed ($\Safe=\downarrow \Safe$) (Bad is the complementary of Safe) and by the existence (or not) of a bound $k$ on the length of runs starting from Bad and reaching Safe.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
-We first show that most of all resilience problems are undecidable for WSTS with strong compatibility and with effective pred-basis.
-Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility.
-Moreover, some resilience properties are decidable for two other classes of WSTS : (1) for WSTS with effective
-$\uparrow$ $\post^*$ basis and (2) for ideal-effective WSTS with downward and upward compatibilities. Finally, we study the resilience for VASS where most of the problems are shown decidable.
+We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis).
+Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$.
+Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
+$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$.
+Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for extensions and variations of VASS; most of the resilience properties are shown decidable.

### 2023-06-21 resilience-safe-up.tex
section 3
section abstract

@@
 
-We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
+We start with the case $\Safe=\uparrow \Safe$ hence $\Bad=\downarrow \Bad$. 
 It can be viewed as a generalizeation
@@
 
-Without the completion-post-effective hypothesis nor the pred-basis hypothesis, coverability 
-becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}. 
-Coverability reduce
+Let us remark that coverability reduce
 to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
@@
 reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
+%
+Let us recall that the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}, hence the strongly increasing well-structured nets are not completion-post-effective hypothesis nor have the pred-basis hypothesis.
+%
+Without the completion-post-effective hypothesis nor the pred-basis hypothesis \alain{où est-ce prouvé ?}, coverability 
+becomes undecidable. For instance
+
 Hence, without the completion-post-effective hypothesis, 
 {\sc Resilience} itself is undecidable.
-
+\alain{c'est confus}
 Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using

### 2023-06-21 resilience.tex
up
section abstract

@@
 
+\mathieu{Rappeler l'example  les Safe Bad considérés et dire ce que ça veut dire en termes de chemins etc.}
+
+Recall the {\bf Flight Controller System} example, the safe state corresponds to the aircrafts having both landed, i.e.
+$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4\}$,
+and $\Bad$ is the complement of $\Safe$. 
+% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
+%%  Then the resilience problem asks
+% wheher
+
 \iffalse
@@
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in Example~\ref{Example}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
+Remark that in Example~\ref{Example} ({(\bf Flight Controller System)}), $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
 
@@
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad %& $\uparrow$~ $\uparrow$~ 
-		& $\uparrow$~ $\downarrow$~ 
-		 & $\downarrow$~ $\uparrow$~ 
+\hline   \ %& $\uparrow$~ $\uparrow$~ 
+		& $\Safe = {\uparrow \Safe}$~ 
+		 & $\Safe = {\downarrow \Safe}$~ 
 		 %& $\downarrow$~ $\downarrow$~
@@
 Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
-
+%		\end{proof}
+\alain{impossible de mettre end{proof} ???...et la preuve ne se finit pas clairement}
+%
 It is also possible to define $(s_0,X)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} X$.

### 2023-06-21 abstract.tex
section abstract

@@
 $\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$.
-Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for extensions and variations of VASS; most of the resilience properties are shown decidable.
+Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for variations of VASS (lossy counter machines, integer VASS and continuous VASS); most of the resilience properties are shown decidable.

### 2023-06-21 introduction.tex
removed the 'not too lossy hypothesis', thinking it doesn't add enough to warrant inclusion

@@
  Lossiness is represented by using transitions decreasing any counter by one from every state. 
-We ask however that the system is not too lossy.
+% We ask however that the system is not too lossy.
 We add the condition that for every prefix of an execution, the number of ``lossiness transition'' for a counter is less than nine tenth of the number of transitions incrementing said counter.
@@
 In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time.
-With one out of ten increment of a counter guarranteed to not be lost, $40$-resilience hold in the example.
+With one out of ten increment of a counter guarranteed to not be lost, $8$-resilience hold in the example.
 \end{example}

### 2023-06-21 introduction.tex
section abstract

--- 2023-06-19 introduction.tex
@@
 pushed it towards some kind of ‘error state’ and, if it can, whether it can perform the return in a satisfactory timeframe. 
+%%%%%%%%
 In the exemple of air trafic control, an accident i.e.
@@
 
-%
-%{\bf MFCS 2023
-%Abstract submission deadline:    	 April 24th (AoE) 
-%Paper submission deadline:    	April 28th (AoE)} \\
-%%
-%{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
-%Submission Deadline 	May 2, 2023}\\
+State-resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.
+
+{\bf State of the art.} 
+In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience.
+Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+All these papers are concerned particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed,
+even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
+
+Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, considered two decidability questions: xxxx whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
+xxxxxx
+
+% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
+asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$.
+
+\alain{state-resilience is an instance of the home space A set $X$ is an \emph{home space} for a system $(S,s_0)$ \cite{DBLP:conf/ac/MemmiV86} if from all reachable states from $s_0$, it is always possible to reach $X$; resilience is an instance of recently Memmi extended this definition as follows: A set $X$ is an \emph{$A$-home space} for a system $S$ \cite{memmi2023invariants} if from all states in $A$, it is always possible to reach $X$.}
 
@@
 \begin{itemize}
-\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
+\item We investigate and expand upon the notion of resilience as defined by Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}
+We introduce a (more) general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
 and the $k$-(state-)resilience problem (kRP).
-The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$.
-We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
+The (general) resilience problem is the following: let $Bad,Safe \subseteq S$ be two subsets of states, S is $(Bad,Safe)$-resilient if $\post^*(Bad)	\subseteq \pred^*(Safe)$.
+We remark that it is an instance of the home- space problem first defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
 State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.
 
-\item Surprinsingly, the general undecidability statements were not known neither proved. We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
-
-%\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
-
-\item We find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and two sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
-
-\item The resilience problem is decidable for ideal-effective WSTS with 
-$\Safe=\downarrow \Safe$
-and $\Bad=\uparrow \Bad$
-and
-the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+\item Undecidable results: Surprinsingly, the general undecidability statements were not known neither proved.
+We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
+Moreover, state-resilience,
+bounded-state-resilience and
+ $k$-state-resilience
+are undecidable for strongly upward-compatible WSTS with effective pred-basis.
 
@@
 
-\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
+%\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
+
+\item We find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
+and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
+%		two sets $\Bad = \downarrow \Bad$
+
+\item The resilience problem is decidable for ideal-effective WSTS with 
+$\Safe=\downarrow \Safe$
+and $\Bad=\uparrow \Bad$
+and
+the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+%
+%%%
+
+\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is still decidable for 
  WSTS with effective 
@@
 and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
+%
 
@@
 \end{itemize}
-
-{\bf Related works.} 
-In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience.
-Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-All these papers are concerned particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed,
-even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
-A set $X$ is an \emph{home space} for a system $(S,s_0)$ \cite{DBLP:conf/ac/MemmiV86} if from all reachable states from $s_0$, it is always possible to reach $X$; recently Memmi extended this definition as follows: A set $X$ is an \emph{$A$-home space} for a system $S$ \cite{memmi2023invariants} if from all states in $A$, it is always possible to reach $X$.
 

### 2023-06-21 resilience.tex
solved endproof
up

@@
 
-\mathieu{Rappeler l'example  les Safe Bad considérés et dire ce que ça veut dire en termes de chemins etc.}
-
-Recall the {\bf Flight Controller System} example, the safe state corresponds to the aircrafts having both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,5}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4\}$,
+Recall the {\bf Flight Controller System} example, where the safe state corresponds to the aircrafts having both landed, i.e.
+$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
 and $\Bad$ is the complement of $\Safe$. 
@@
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in Example~\ref{Example} ({(\bf Flight Controller System)}), $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
+Remark that in the {\bf Flight Controller System} example, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
 
@@
 Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
-%		\end{proof}
+		\end{proof}
 \alain{impossible de mettre end{proof} ???...et la preuve ne se finit pas clairement}

### 2023-06-21 resilience.tex
up

@@
 % $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
-%%  Then the resilience problem asks
 % wheher
+The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.
 
@@
 
-Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, it is useful to note that in a sequence of states of $S$, every prefix is always 
+% Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, 
+It is useful to note that in a sequence of states% of $S$
+, every prefix is always 
 either in the case that the latest visited $\Bad$ state has already been followed by a 

### 2023-06-21 resilience-safe-up.tex
up

@@
 %
-Without the completion-post-effective hypothesis nor the pred-basis hypothesis \alain{où est-ce prouvé ?}, coverability 
-becomes undecidable. For instance
+Moreover, this implies from the reduction that resilience for strongly increasing well-structured nets is  undecidable. Hence resilience for WSTS in general is undecidable.
 
-Hence, without the completion-post-effective hypothesis, 
-{\sc Resilience} itself is undecidable.
-\alain{c'est confus}
 Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using

### 2023-06-21 introduction.tex
section abstract

@@
 
+{\bf Resiliences.}  We now define the resilience and the state-resilience : the (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $Bad,Safe \subseteq S$, is stated as follows: $\mathscr{S}$ is (Bad,Safe)-resilient if $\post^*(Bad) \subseteq \pred^*(Safe)$; similarly, $\mathscr{S}$ is (Bad,Safe)-state-resilient if $\post^*(s_0) \cap Bad \subseteq \pred^*(Safe)$.
+Resilience implies state-resilience but a system can be state-resilient but not resilient.
+
+and we define three decidability questions for each of type of resilience:
+ The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
+and the $k$-(state-)resilience problem (kRP).
+
 {\bf State of the art.}
-In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience.
+In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} many interesting intuitions (without detailed proofs) about resilience in the framework of process algebra.
+They built the composition of the process and its adversary as a transition system and they gave conditions that insure that the composed transition system is an effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
+In this framework and under these hypotheses, resilience reduces to coverability which is decidable on WSTS.
+Since the definition of resilience is only given informally in the introduction as "a process is resilient to an adversary if its observable behaviour in the presence of adversarial action is equivalent to its behaviour in the absence of the adversary", it still remains unclear what is formally resilience.
+%%%%%%%%%
+
+in 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and in 2022, Okzan \cite{DBLP:conf/gg/Ozkan22} studied well-structured graph transformation systems and they considered two state-resilience problems : the k-State-Resilience Problem and the Bounded-State-Resilience Problem.
+They studied these two problems for WSTS with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
+In these papers, $\Bad$ is downward-closed and $\Safe$ is upward-closed
+(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases).
+%%%%%%%%%
+
+ In 1986, Memmi and Vautherin introduced the notion of home-space \cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
+%			from all reachable states from $s_0$, it is always possible to reach $H$ 
+$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$, we say that $H$ is an home-space for $X$ if $\post^*(X) \subseteq \pred^*(H)$.
+In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
+In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X,H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
+%
+
 Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-All these papers are concerned particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed,
-even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
+
+No relation is established in previous papers with the HSP.
+
+State-resilience is an instance of the home-space property.
 
@@
 asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$.
-
-\alain{state-resilience is an instance of the home space A set $X$ is an \emph{home space} for a system $(S,s_0)$ \cite{DBLP:conf/ac/MemmiV86} if from all reachable states from $s_0$, it is always possible to reach $X$; resilience is an instance of recently Memmi extended this definition as follows: A set $X$ is an \emph{$A$-home space} for a system $S$ \cite{memmi2023invariants} if from all states in $A$, it is always possible to reach $X$.}
 
@@
 \item We investigate and expand upon the notion of resilience as defined by Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}
-We introduce a (more) general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
-and the $k$-(state-)resilience problem (kRP).
-The (general) resilience problem is the following: let $Bad,Safe \subseteq S$ be two subsets of states, S is $(Bad,Safe)$-resilient if $\post^*(Bad)	\subseteq \pred^*(Safe)$.
-We remark that it is an instance of the home- space problem first defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
+We introduce a (more) general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience:We remark that it is an instance of the home- space problem first defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
 State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.

### 2023-06-21 reachability/article.tex
up

@@
 and we will call player $1-i$ its {\em opponent}. 
-
-% Observe that there is no restriction on the number of the successors of a configuration in an arena.
-Also, we don’t require that (V, E) is a bipartite graph with corresponding partition {V 0 , V 1 }.
-
-\iffalse
-\noindent
-In the case where $\Lambda$ is a singleton $\{ a\}$, we will abuse notation
-and denote the arena by
-$(S_0, S_1,{\rightarrow},\Omega)$, and we will see $(S,{\rightarrow})$
-as an unlabeled transition system.
-\fi
-
-%  \subsubsection{Plays}
-
-\noindent
 A {\em play} in an arena $A=(S_0,S_1,\rightarrow)$ is a path in $(S,{\rightarrow})$ that is {\em maximal} in the following sense: it is either infinite or finite and if it is finite, then the target configuration is a dead end.
@@
 that can finitely be described.
-
 A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
@@
 
-\begin{samepage}
 \problemx{Reachability game}
 {A reachability game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_F)$, an initial configuration $s \in S_0 \cup S_1$.}
-{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
-\end{samepage}
-
-% \subsection{Parity Games}
+{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$? \\}
 

### 2023-06-21 introduction.tex
section intro

@@
 
-{\bf Resiliences.}  We now define the resilience and the state-resilience : the (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $Bad,Safe \subseteq S$, is stated as follows: $\mathscr{S}$ is (Bad,Safe)-resilient if $\post^*(Bad) \subseteq \pred^*(Safe)$; similarly, $\mathscr{S}$ is (Bad,Safe)-state-resilient if $\post^*(s_0) \cap Bad \subseteq \pred^*(Safe)$.
+{\bf Home-spaces}
+ In 1986, Memmi and Vautherin introduced the notion of home-space \cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
+%			from all reachable states from $s_0$, it is always possible to reach $H$ 
+$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$, we say that $H$ is an home-space for $X$ if $\post^*(X) \subseteq \pred^*(H)$.
+In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
+In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
+%%%%%
+
+{\bf Resiliences.}  
+%		We now give the definition of two kinds of resilience: the (general) resilience and the state-resilience. 
+The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $Bad,Safe \subseteq S$ can be stated as an home-space problem : $\mathscr{S}$ is (Bad,Safe)-resilient if $\post^*(Bad) \subseteq \pred^*(Safe)$;
+similarly, $\mathscr{S}$ is (Bad,Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap Bad \subseteq \pred^*(Safe)$.
 Resilience implies state-resilience but a system can be state-resilient but not resilient.
-
-and we define three decidability questions for each of type of resilience:
- The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
-and the $k$-(state-)resilience problem (kRP).
+We will study three decidability questions for each of type of resilience.
+The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is (Bad,Safe)-resilient (for two given subsets of states $Bad,Safe \subseteq S$).
+The $k$-resilience problem (kRP) is to decide whether $\post^*(Bad) \subseteq \pred^{\leq k}(Safe)$ (for a given $k$) and 
+the bounded resilience problem (BRP) is to decide whether there exists an $k$ such that $\mathscr{S}$ is (Bad,Safe)-$k$-resilient. We adapt these three definitions to state-resilience.
 
@@
 
-in 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and in 2022, Okzan \cite{DBLP:conf/gg/Ozkan22} studied well-structured graph transformation systems and they considered two state-resilience problems : the k-State-Resilience Problem and the Bounded-State-Resilience Problem.
-They studied these two problems for WSTS with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-In these papers, $\Bad$ is downward-closed and $\Safe$ is upward-closed
-(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases).
+in 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and in 2022, Okzan \cite{DBLP:conf/gg/Ozkan22} proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
+$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, hence $\Bad$ is downward-closed
+(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases). The paper of Prasad and Zuck \cite{DBLP:journals/corr/PrasadZ16} is not quoted by Ozkan and Würdemann.
+
+Both Prasad \& Zuck and Ozkan \& Würdemann did not make the relation between resilience and the home-space property.
 %%%%%%%%%
-
- In 1986, Memmi and Vautherin introduced the notion of home-space \cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
-%			from all reachable states from $s_0$, it is always possible to reach $H$ 
-$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$, we say that $H$ is an home-space for $X$ if $\post^*(X) \subseteq \pred^*(H)$.
-In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
-In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X,H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
 %
-
-Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$.
-
-No relation is established in previous papers with the HSP.
-
-State-resilience is an instance of the home-space property.
-
-Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}, considered two decidability questions: xxxx whether a system can reach a safe state in a bounded number of steps whenever it reached a bad state.
-xxxxxx
-
-% is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. 
-asks whether a set of safe elements $\Safe$ is reachable for all elements in a set of bad elements $\Bad$.
+Remark that the reflexive downward compatibility of a WSTS implies that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to prove the announced result by Prasad and Zuck.
 

### 2023-06-21 introduction.tex
section intro

@@
 %%%%%%%%
-In the exemple of air trafic control, an accident i.e.
+\alain{fait double emploi avec l'exemple si on le développe ce que je ne suis plus sûr...:  In the exemple of air trafic control, an accident i.e.
 for instance a delay in the arrival of a plane can lead to conflicts of resolution (i.e.
@@
 
-State-resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.
+State-resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.}
 
@@
 
-in 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and in 2022, Okzan \cite{DBLP:conf/gg/Ozkan22} proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
+In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan in 2022, \cite{DBLP:conf/gg/Ozkan22} proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
 $\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, hence $\Bad$ is downward-closed
@@
 %
-Remark that the reflexive downward compatibility of a WSTS implies that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to prove the announced result by Prasad and Zuck.
+Remark also that the reflexive downward compatibility of an effective WSTS implies that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to prove the announced result by Prasad and Zuck.
 

### 2023-06-21 reachability/article.tex
up

@@
 
-% {\bf Context}
-
-{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs.
+{\bf Context} Pushdown systems can be used to model the behavior of recursive programs.
 As a result, a variety of 
@@
 making use of parametric constraints. 
-% Here, the stack can additionally be compared against parameters that can take unspecified word values. 
 Recently, different variants of parametric pushdown systems have been introduced in the literature 
-% [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
 \cite{hague2011parameterised, esparza2016parameterized, fortin2017model} 
-% In this work, following [21,20,15,27,12], we consider parametric pushdown systems
 however as
@@
 We consider here a different approach to extending pushdown systems with parameters, by allowing them to test equality of the stack content against parameters.
-% Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
 Perhaps a similar model consists in 
@@
 Pushdown systems 
-%with regularity condition on the stack
 with transitions that are conditioned by regular conditions on their stack content
@@
 
-% Hence we propose in this paper to extend pushdown systems with a set of  parameters.
-%
-% Compared to pebble word automata, instead of using pebbles
-% as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles
-% on their universe of stack contents. Dropping a pebble records the current stack content with an available pebble until it is lifted. Lifting a pebble then makes it available again.
-
-% A pebble automaton is a two-way finite state automaton that uses a fixed, finite number of
-% pebbles that it can drop on, and lift from 
-% words, using them as markers.
-% Pebble automata recognize regular languages only, provided the life times of the pebbles, i.e. the times between dropping a pebble and lifting it again, are properly nested
-% \cite{globerman1996complexity, engelfriet1999tree}.
-% Automata with nested pebbles were also introduced for tree-walking automata. 
-% It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}, the main trouble being that the tree-walking automata get lost rather easily.
-% Using pebbles is a remedy against getting lost along a tree, but  the unrestricted use of pebbles leads to a class of tree languages much larger than the regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
-% Thus, in both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline, and our model will follow this strict stack discipline too.
-%
-%
-% One motivation to the study of pebble pushdown systems is their ability to compare a stack content against some later stack content.
-% Similar type of storage for later comparison has already been used for instance in register pushdown systems.
-Introduced in \cite{murawski2017reachability}  as an extension of pushdown systems that can keep data values in a set of registers and a stack, register pushdown systems enable manipulation of data values from an infinite domain, and have applications for instance for malware detection and XML schema checking \cite{senda2021forward, senda2021ltl}.
-
-Another closely related formalism is that of alternating tree automata with nested pebbles%, as we consider in more details in Section 4
-.
+Another closely related formalism is that of alternating tree automata with nested pebbles.
 The approach of extending tree automata with pebbles was used in \cite{milo2000typechecking} to show that the XML typechecking problem is decidable, and in \cite{karhumaki2012jewels} to recognize first-order logic on trees.
@@
 parameters.
-
- \iffalse
-
-We provide formal definitions of parametric pushdown systems, parametric pushdown reachability games
-and parametric pushdown parity games in Section~\ref{section def ppda}. We give an overview of our contribution in Section~\ref{contribution ppda}. The contribution consists in
-proving parametric pushdown reachability games
-and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
-$n$ is fixed,
-and providing a nonelementary lower bound for parametric pushdown reachability games in general.
-We provide the proof of our result, which stretches along 
-Section~\ref{lower bound PDA},
-Section~\ref{pebbles},
-% Section~\ref{ATWA and PATWA}, and
-Section~\ref{section HPDA}.
-In Section~\ref{discuss ppda} we close the chapter with a discussion about the methods used, with some directions for future work.
-\fi
 
@@
 By $\Z $ we denote the {\em integers} and by $\N=\{0,1,\ldots\}$ we denote the {\em naturals}.
-% For every $a,b\in\Z$ with $a\leq b$ we define $[a,b]=\{k\in\Z\mid a\leq k\leq b\}$.
-% For every $n\geq 1$ we define $n\Z=\{n\cdot z\mid z\in\Z\}$.
-% For every number $n\in\N$ we define $\log(n)=\min\{i+1\mid i\in\N, n\leq 2^i\}$, which is the smallest number of bits necessary to write down $n$ in binary.
 For every finite alphabet $A$, $A^*$ is the set of finite
 words with letters in  $A$, $A^\omega$ is the set of infinite words with letters in $A$, and
-%by
  $A^\infty = A^\omega \cup A^*$. We denote the empty word in $A^*$ by $\epsilon$.
-%For all $a\in A$ and all $w\in A^*$ let $|w|_a$ denote the number of occurrences of the letter $a$ in $w$.
-%
-% For every finite set $M\subset\N\setminus\{0\}$ let $\LCM(M)=\min\{n\geq 1\mid \forall m\in M\setminus\{0\}: m|n\}$ denote the least common multiple of the elements in $M$. 
-% For any $j \in \N$ let $\LCM(j)=\LCM([1,j])$ denote the least common multiple of the numbers $\{1,\ldots,j\}$.
 
@@
 For any set $S$ let 
-%$\mathscr{P}(S) = \{ X \mid X \subseteq S \}$
 $\mathcal{P}(S) = \{ X \mid X \subseteq S \}$
@@
 A {\em graph} $G=(V,E)$ is a finite set of graph vertices $V$ with a set of graph edges $E \subseteq V \times V$. 
-% Two graphs $G = (V,E)$ and $G'=(V',E')$ with sets of graph vertices $V=V'=V_n=\{1,2,...,n\}$ are said to be {\em isomorphic} if there is a permutation $\rho$ of $V_n$ such that $(u,v)$ is in the set of graph edges $E$ iff $(\rho(u),\rho(v))$ is in the set of graph edges $E'$.
-%%%
-%
 For any set $V$, and a given infinite sequence $\pi = v_0 v_1 \ldots \in V^\omega$ of elements of $V$, we define the set 
@@
 
-%ù \subsection{Games}
-
 A {\em game} is composed of an arena and a winning condition.
@@
 
-% \subsubsection{Arenas}
-
 \begin{samepage}
@@
 \begin{itemize}
-	\item two  disjoint  sets of configurations, $S_0$ and $S_1$, %respectively associated with player $0$ and $1$, 
+	\item two  disjoint  sets of configurations, $S_0$ and $S_1$, 
 	whose disjoint union $S_0 \cup S_1$ we denote by $S$, and
-	\item a relation of %labeled 
+	\item a relation of  
 		transitions
-	 ${\rightarrow} \subseteq S \times %\Lambda \times 
-	 				S$. 
+	 ${\rightarrow} \subseteq S \times 	S$. 
 \end{itemize}

### 2023-06-21 reachability/article.tex
up

@@
 
-\subsubsection*{Background}
-
-{\bf Context} Pushdown systems can be used to model the behavior of recursive programs.
+{\bf Background.} Pushdown systems can be used to model the behavior of recursive programs.
 As a result, a variety of 
@@
 strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
-\textcolor{blue}{ok je comprends que tu signales la différence mais c'est trop détaillé pour cet endroit du texte: To underline the fact that we are concerned with the graph of configurations and not the language recognized, we use the name pushdown system rather than pushdown automaton.}
 
@@
 
-\section{Definitions}
+\section{Preliminaries}
 
@@
 
-\begin{samepage}
-An {\em arena} is a tuple $A=(S_0, S_1,{\rightarrow})$  which is composed of
-\begin{itemize}
-	\item two  disjoint  sets of configurations, $S_0$ and $S_1$, 
+An {\em arena} is a tuple $A=(S_0, S_1,{\rightarrow})$  which is composed of two  disjoint  sets of configurations, $S_0$ and $S_1$, 
 	whose disjoint union $S_0 \cup S_1$ we denote by $S$, and
-	\item a relation of  
+ a relation of  
 		transitions
 	 ${\rightarrow} \subseteq S \times 	S$. 
-\end{itemize}
-\end{samepage}
-
-\noindent
 Note that $(S,{\rightarrow})$, where $S=S_0 \uplus S_1$, forms
@@
 one needs only to provide 
-a partition of the set of configurations $S$ into two sets $S_0$ and $S_1$ to obtain an arena. \\
-% We require that the out-degree of each configurations is at least one. This ensures that every finite path in (V, E) can be prolonged.
-
-% \mh{I dont think I want to disallow dead-ends from the very beginning, means every time you define some arena you have to prove it has no dead-ends, seem tedious and unecessary.}
-
-\noindent
+a partition of the set of configurations $S$ into two sets $S_0$ and $S_1$ to obtain an arena.
+
 The games we are interested in are played by two players, called {\em player $0$} and {\em player $1$}. We will often write player $i$ to denote a general player for $i \in \{0,1 \}$,
@@
 	\caption{An example of an arena.
-Circles  belong to player $0$, whereas diamonds belong to player $1$.
+Circles  belong to player~$0$, whereas diamonds belong to player~$1$.
 Arrows represent transitions from a configuration to the next.
 The priority assigned by the priority mapping to each configuration is a number in $\{0,1,2,3\}$, next to the configuration it is assigned to.
-Note that we do not have a dead end in our example.}
+Note that we do not have any dead end in our example.}
 	\label{example_parity_game}
@@
 Let us consider the arena presented in  Figure~\ref{example_parity_game}. 
-Configurations are partitioned into two sets: the circles, belonging to player $0$, and the diamonds,
+Configurations are partitioned into two sets: circles belonging to player $0$ and diamonds
 belonging to player $1$. 
@@
 % The transitions and priority mappings may be derived from the picture.
-% Note that we don't have a dead end in our example. 
-Consider the game on the arena, where the winning condition is a parity condition.
 Plays $\pi_1 = q_0 q_1 q_2 (q_4 q_7)^\omega$ and 
@@
 the plays $\pi_3 = (q_5 q_3 q_6)^\omega$ and $\pi_4=(q_5 q_6)^\omega$ are both winning for player $1$: the minimal priority encountered is odd, since it is $1$ for both.
-
 One can take matters further: on this arena, player $0$ has a winning strategy for the parity game from $q_0$, $q_1$, $q_2$, $q_4$ and $q_7$, but not from $q_5$, $q_3$ nor $q_6$, where player $1$ has a winning strategy.
@@
 
-%\mh{
-Of note is that several model checking problems %for pushdown systems 
-can be expressed as decision problems for games: the most famous example is that the modal $\mu$-calculus model checking problem is polynomially equivalent to solving the parity game problem. 
-%}
-
 \section{Parametric pushdown games}
 
-\textcolor{black}{Parametric pushdown systems extend pushdown systems by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.}
+Parametric pushdown systems extend pushdown systems by allowing the stack to be compared against parameters that can be assigned values which are words over the stack alphabet.
 A parametric pushdown automaton is then a finite automaton extended with a finite set of parameters $P$  and with a stack that can be manipulated by pushing or popping stack symbols and such that, moreover, the automaton can use the top of the stack, or check that the stack content corresponds to a parameter,  to decide which transition to take next.\\
@@
 }_{\mathcal{Z}, \mu}$ (resp. $\xrightarrow{%\lambda
-}_{\mathcal{Z},u}$). \\
+}_{\mathcal{Z},u}$). 
+Since we are concerned with the graph of configurations and not the language recognized, we omit the input alphabet from the model and will use the name pushdown system rather than pushdown automaton.
 
@@
 the modal $\mu$-calculus
-model checking problem is polynomially equivalent to solving the parity game problem.
-% \textcolor{red}{Since the model checking problem has a more varied range of applications than the reachability problem (mb have a source for that or smtg)},
-Since we are interested in parametric model checking and specification,
-we are naturally interested in parametric variants of reachability games and parity games.
+model checking problem is polynomially equivalent to solving the parity game problem~\cite{emerson1993model}.
+We are interested in this article in parametric variants of reachability games and parity games.
 
@@
  $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
- %We do not really consider winning conditions for parameter valuation games.
-Rather, once the resulting parameter valuation is produced, it serves to instantiate the parametric transition system for the next game.
-Parametric valuation arenas can then be used as prefixes in parametric games.
-Once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
+We then use parametric valuation arenas as prefixes in parametric games: once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
 %
@@
 	 %Lastly $T^\mu_\A$ denotes the transition system induced by the concrete automaton corresponding to $\A$ with parameter valuation $\mu$.
-	  As before, the circles belong to player $0$ and the diamonds to player $1$.} 
+	  As before, the circles belong to player $0$ and the diamonds to player~$1$.} 
 	\label{parametric_arena}
@@
 
-Recall that given a 
-%labeled 
-transition system,
-one needs only to provide %a mapping $\Omega$ and 
-a partition of the set of configuration $S$ into two sets $S_0$ and $S_1$ to obtain an arena. 
-Thus, for an  $n$-parametric pushdown automaton 
+More formally, for an  $n$-parametric pushdown automaton 
 $\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
@@
 we naturally partition the configurations of
-%the transition system of 
  $\mathcal{Z}$
-% $T^\mu_{\mathcal{Z}}$
 into
@@
 $\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$. 
-
-With these notations in mind one can define the arena
-\[
+With these notations in mind we define the arena
+$
 A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
 =
-(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})\]
-induced by a PPDS $\mathcal{Z}$, a partition of its set of states, and a
+(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$
+induced by $\mathcal{Z}$, the partition $Q = Q_0 \biguplus Q_1$, and a
 parameter valuation $\mu: P\rightarrow \Gamma^*$.
-
 From a partition of
@@
 
-Given a 
-priority function $\Omega: Q \to [0, m]$,
-we consider the function
-$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$
+We extend any 
+priority function $\Omega: Q \to [0, m]$
+into
+$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$,
 where we set
@@
 Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.
-We are interested in the following games and problems.
-
-%\subsection{A decidability result from reduction to parity games on automata with pebbles}\label{pebbles}
-\section{%A decidability result via r
-Reduction to pebble pushdown systems parity games
+
+\section{Reduction to pebble pushdown systems parity games
 }\label{pebbles}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% \mh{remind the reader what a partial function is}
-
-\par\noindent\ignorespacesafterend 
 Recall that for a partial function $ f : S \rightharpoonup X $% (see page~\pageref{partial})
@@
 $m$. 
-% \mh{The set $S$ is used for testing the presence of certain pebbles.}
 Note a pebble 
@@
 is the same as the current stack content. 
-%$\Dzeta$ $\zeta$ $\Xi$
 This is enforced by syntactically requiring
 the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.
-
-\iffalse
-A {\em weak} $n$-pebble pushdown system is an $n$-pebble pushdown system $\ppda = (Q, \Gamma,  \Delta )$ in which $\text{\textit{lift}}$ operations also require the presence of the last pebble dropped, i.e.
-for all $(q, a, i, S, q', \text{\textit{lift}}) \in \Delta$, we have $ i \in S $. 
-%  is an additionnal restriction $f(i) = wa$ that needs to hold for the application of the $lift$ operation.
-{\em Weak $n$-pebble pushdown games} are defined naturally as $n$-pebble pushdown games on 
-weak $n$-pebble pushdown systems.
-
-\fi
 
@@
 totally undefined function.
-
-\iffalse
-For $0 \leq i \leq n$, an {\em $i$-configuration} is a tuple $(q, w, \mu )$, where $w \in \Gamma^*$, $q$ is a state and
-$\mu$ a $\{ 1, \ldots , i \}$-pebble assignment. 
-We call $w$ the current stack, $q$ the current state and
-$\mu$ the current pebble assignment. 
-We also write $(q,w, w_{1} , \ldots , w_{i} )$ if 
-$\mu(j) = w_j$, for each
-$j \leq i$.
-
-The set of configurations is the union of all $i$-configurations for $0 \leq i \leq n$. 
-\fi
 
@@
 $\Conf_{\ppda,1}=Q_1\times  \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^{\{1, \ldots n\}}$. 
-
-With these notations in mind one can  define the arena
-$$
+We  define the arena
+$
 A_{(\ppda, Q_0 , Q_1)}
 =
-(\Conf_{\ppda,0}, \Conf_{\ppda,1}, \rightarrow_{\ppda}).$$
-
+(\Conf_{\ppda,0}, \Conf_{\ppda,1}, \rightarrow_{\ppda}).$
 As expected, given a 
@@
 {Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
-$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?\newline}
-
-% \subsubsection{Reduction from parameters to pebble pushdown systems}
+$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?}
+
 \subsection{From parametric pushdown 
@@
 \begin{theorem}\label{PPDA reduction}
-%The problem of 
 {\sc $n$-parametric pushdown parity game} 
 is polynomial time reducible to
-%the problem of 
 {\sc $n$-pebble pushdown parity game}.
@@
 $n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
-P%\{p_1, \ldots, p_n \}
-, q_{init}, { \gamma_{init}} , F)$,
+P, q_{init}, { \gamma_{init}} , F)$,
  disjoint
@@
 We construct
-% polynomial time
   an 
@@
 such that
- %:
-%
 player $0$ has a winning strategy from 
@@
 
-%
-%
-%
-%
-
 The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
 (one for every 
-%total 
 parameter assignment)
@@
 the pebble assignment $\mu$ is fixed.
-%%
-%
-%
-%
-%
-%
-%
 
@@
 automaton 
-%transition system with assignment $\mu$
 and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
@@
 We start with
-% a $\tau(\{0,1\})$-structure $w$ and
  a 
- $\FO[\tau(\{0,1\})]$%%
-% $\FO$%
+ $\FO[\tau(\{0,1\})]$
 -formula $\phi$.
-%
-% Note that $w$ corresponds to a word $w \in \{0,1\}^*$ of size $n$.\\
-%
-% \noindent
 We construct

### 2023-06-22 appendix.tex
up

@@
 
-First, let us recall the home-state problem: 
-whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
+First, let us recall the home-state problem
+
+\problemx{Home-state}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
+{$S \longrightarrow^{*} \{s\}$ ?\newline}
+
+% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
 
 Home-state is undecidable for Minsky machine with $3+$ counters.
-
-Because $2$-counter Minsky machine termination is undecidable.
-
-$2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ which 
-third counter starts at least at $1$, and simulates $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
-
-Then the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
+From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
+such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
 
@@
 
-Thus home-state/resilience is undecidable for Minsky machines.
+Thus {\sc Home-state} is undecidable for Minsky machines. Hence {\sc Resilience} is undecidable.
 

### 2023-06-22 appendix.tex
up

@@
 We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
-
 % Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
 % or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
+Thus {\sc Home-state} is undecidable for Minsky machines, and hence {\sc Resilience} is undecidable.
 
-Thus {\sc Home-state} is undecidable for Minsky machines. Hence {\sc Resilience} is undecidable.
-
-Executions of Minsky machines can be simulated by reset-VASS. 
+Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
 % Let us recall more formally what a reset-VASS is.
@@
 It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
-Reset-VASS can simulate executions of a Minsky machine.
-
-Thus home-state/resilience is undecidable for reset-VASS and hence for WSTS as well.
+Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
 

### 2023-06-22 introduction.tex
up

@@
 {\bf Context.} 
-Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges}).
+Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety~\cite{challenges}.
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
@@
 {\bf Home-spaces}
- In 1986, Memmi and Vautherin introduced the notion of home-space \cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
+ In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
 %			from all reachable states from $s_0$, it is always possible to reach $H$ 
@@
 %		We now give the definition of two kinds of resilience: the (general) resilience and the state-resilience. 
-The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $Bad,Safe \subseteq S$ can be stated as an home-space problem : $\mathscr{S}$ is (Bad,Safe)-resilient if $\post^*(Bad) \subseteq \pred^*(Safe)$;
-similarly, $\mathscr{S}$ is (Bad,Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap Bad \subseteq \pred^*(Safe)$.
+The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $\Bad,\Safe \subseteq S$ can be stated as an home-space problem : $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if $\post^*(\Bad) \subseteq \pred^*(\Safe)$;
+similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
 Resilience implies state-resilience but a system can be state-resilient but not resilient.
 We will study three decidability questions for each of type of resilience.
-The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is (Bad,Safe)-resilient (for two given subsets of states $Bad,Safe \subseteq S$).
-The $k$-resilience problem (kRP) is to decide whether $\post^*(Bad) \subseteq \pred^{\leq k}(Safe)$ (for a given $k$) and 
-the bounded resilience problem (BRP) is to decide whether there exists an $k$ such that $\mathscr{S}$ is (Bad,Safe)-$k$-resilient. We adapt these three definitions to state-resilience.
+The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is ($\Bad,\Safe$)-resilient (for two given subsets of states $\Bad,\Safe \subseteq S$).
+The $k$-resilience problem (kRP) is to decide whether $\post^*(\Bad) \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
+the bounded resilience problem (BRP) is to decide whether there exists an $k$ such that $\mathscr{S}$ is ($\Bad,\Safe$)-$k$-resilient. We adapt these three definitions to state-resilience.
 
@@
 the least amount of steps possible. 
-In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time.
-With one out of ten increment of a counter guarranteed to not be lost, $8$-resilience hold in the example.
+In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time. Fortunately, $8$-resilience hold in the example.
 \end{example}

### 2023-06-22 resilience.tex
up

@@
 
-% Given a set of states $S$, with two subsets $\Bad$ and $\Safe$, 
 It is useful to note that in a sequence of states% of $S$
@@
 either in the case that the latest visited $\Bad$ state has already been followed by a 
-$\Safe$ state
-or in the case that it still has to. 
-Thus, \emph{in context}, there is no true ``neutral'' state: either a neutral state is contextually ``$\Safe$'' because it follows a visit in a $\Safe$ state, either it is contextually ``$\Bad$'' because it follows a visit in a $\Bad$ state and has, accordingly, to reach a $\Safe$ one.
+$\Safe$ state \-- and thus the latest state is contextually ``$\Safe$'' \--
+or in the case that it still has to \-- and thus the latest state is contextually ``$\Bad$''. 
+Thus, \emph{in context}, there is no true ``neutral'' state.
 Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but

### 2023-06-22 resilience.tex
up

@@
 {\bf Related problems.} 
-A set of states $H$ is an {\em home space} for a set of states $X$ if every state reachable from $X$ can reach $H$.
-A {\em home space} is a state whose corresponding singleton set is a home space.
-Resilience as introduced above essentially asks whether $\Safe$ is a home space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
-The home space problem is decidable for
+A set of states $H$ is an {\em home-space} for a set of states $X$ if every state reachable from $X$ can reach $H$.
+A {\em home-state} is a state whose corresponding singleton set is a home-space.
+Resilience as introduced above essentially asks whether $\Safe$ is a home-space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
+The home-space problem is decidable for
 Petri nets when $X$ and $H$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
-To these author's knowledge, there are no general results for home space in WSTS.
+To these author's knowledge, there are no general results for home-space in WSTS.
 
-Resilience and the home space problem are also linked to the 
+Resilience and the home-space problem are also linked to the 
 model-checking of basic reachability and safety formulae. 

### 2023-06-22 resilience.tex
Removed home-space section from resilience since it is now introduced and discussed in the introduction

@@
 {\bf Related problems.} 
-A set of states $H$ is an {\em home-space} for a set of states $X$ if every state reachable from $X$ can reach $H$.
-A {\em home-state} is a state whose corresponding singleton set is a home-space.
-Resilience as introduced above essentially asks whether $\Safe$ is a home-space for $\Bad$, whereas $k$-resilience and bounded-resilience focus more on the length of the paths from $\Bad$ to $\Safe$.
-The home-space problem is decidable for
-Petri nets when $X$ and $H$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
-To these author's knowledge, there are no general results for home-space in WSTS.
-
 Resilience and the home-space problem are also linked to the 

### 2023-06-23 abstract.tex
section intro

@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by Bad is an upward/downward closed set or it is a subset of the reachability set from an initial state, Safe is upward-closed ($\Safe=\uparrow \Safe$) or Safe is downward-closed ($\Safe=\downarrow \Safe$) (Bad is the complementary of Safe) and by the existence (or not) of a bound $k$ on the length of runs starting from Bad and reaching Safe.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set Safe  ($\Safe=\uparrow \Safe$) or downward-closed set ($\Safe=\downarrow \Safe$) (Bad is generally the complementary of Safe) and by the existence of a bound $k$ on the length of runs starting from Bad and reaching Safe.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.

### 2023-06-23 conclusion.tex
up
section intro

@@
 One could also extend upon the classes of set $\Bad$ and $\Safe$ considered.
-As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward closed subsets.
-Finally, while we mention VASS, there are other computational models such as pushdown automata, one-counter automata or timed automata, which can be viewed as WSTS~\cite{DBLP:journals/tcs/FinkelS01}, in which a more detailed analysis of the resilience problems could have useful applications.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets.
+Finally, while we mention VASS, but a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
 

### 2023-06-23 introduction.tex
section intro

@@
 %%%%%%%%
-\alain{fait double emploi avec l'exemple si on le développe ce que je ne suis plus sûr...:  In the exemple of air trafic control, an accident i.e.
+%\alain{fait double emploi avec l'exemple si on le développe ce que je ne suis plus sûr...:  In the exemple of air trafic control, an accident i.e.
 for instance a delay in the arrival of a plane can lead to conflicts of resolution (i.e.
@@
 We formalize this as the resilience problem.
-Resilience is then the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states. \\
-
-State-resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.}
+%Resilience is then the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states. \\
+%
+%State-resilience only consider bad elements that are reachable from an initial state, while bounded-resilience and $k$-resilience ask whether safe is reachable in a bounded number of step and $k$ or less steps respectively.}
 
@@
 similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
-Resilience implies state-resilience but a system can be state-resilient but not resilient.
 We will study three decidability questions for each of type of resilience.
@@
 In this framework and under these hypotheses, resilience reduces to coverability which is decidable on WSTS.
-Since the definition of resilience is only given informally in the introduction as "a process is resilient to an adversary if its observable behaviour in the presence of adversarial action is equivalent to its behaviour in the absence of the adversary", it still remains unclear what is formally resilience.
+%Since the definition of resilience is only given informally in the introduction as "a process is resilient to an adversary if its observable behaviour in the presence of adversarial action is equivalent to its behaviour in the absence of the adversary", it still remains unclear what is formally resilience.
 %%%%%%%%%
@@
 $\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, hence $\Bad$ is downward-closed
-(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases). The paper of Prasad and Zuck \cite{DBLP:journals/corr/PrasadZ16} is not quoted by Ozkan and Würdemann.
-
+%		(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases). 
+%		The paper of Prasad and Zuck \cite{DBLP:journals/corr/PrasadZ16} is not quoted by Ozkan and Würdemann.
+Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
 Both Prasad \& Zuck and Ozkan \& Würdemann did not make the relation between resilience and the home-space property.
@@
 %
-Remark also that the reflexive downward compatibility of an effective WSTS implies that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to prove the announced result by Prasad and Zuck.
 
@@
 % trop d'hypotheses trop fortes sur les WSTS
-
+\alain{harmonier les noms des pb}
 \noindent
@@
 \begin{itemize}
-\item We investigate and expand upon the notion of resilience as defined by Okan Özkan and Nick Würdemann~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}
-We introduce a (more) general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience:We remark that it is an instance of the home- space problem first defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697, memmi2023invariants}.
-State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.
+%\item We investigate and expand upon the notion of state-resilience ~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we study three decidability questions for each of type of resilience.
+We remark that it is an instance of the home-space problem~\cite{DBLP:journals/corr/abs-2207-02697,memmi2023invariants}.
 
-\item Undecidable results: Surprinsingly, the general undecidability statements were not known neither proved.
+\item Surprinsingly, the general undecidability statements were not known neither proved.
 We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
-Moreover, state-resilience,
-bounded-state-resilience and
- $k$-state-resilience
-are undecidable for strongly upward-compatible WSTS with effective pred-basis.
-
-\item Unfortunately, {\sc State-resilience},
+%Moreover, state-resilience,
+%bounded-state-resilience and
+% $k$-state-resilience
+%are undecidable for WSTS with strong compatibility having an effective pred-basis. 
+Moreover, {\sc State-resilience},
 {\sc Bounded-state-resilience} and
@@
 when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$. We made a reduction of zero-reachability in reset-VASS to state-resilience in reset-VASS that are WSTS with strong upward-compatibility and with effective pred-basis.
+$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to state-resilience in reset-VASS.
 
@@
 
-\item We find classes of WSTS that have a \emph{decidable} resilience. The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
+\item The three resilience problems are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
 %		two sets $\Bad = \downarrow \Bad$
@@
 
-\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is still decidable for 
- WSTS with effective 
-$\uparrow$ $\post^*$ basis
-when
-$\Safe=\uparrow \Safe$
-and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
+\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis.
+%: {\sc State-resilience} is still decidable for 
+% WSTS with effective 
+%$\uparrow$ $\post^*$ basis
+%when
+%$\Safe=\uparrow \Safe$
+%and $\Bad=\downarrow \Bad$. 
+We show that removing the effective 
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-Nevertheless, we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities.
-and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.
+We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
 %
 
-\item We study the resilience for VASS where most of the problems are shown decidable.
+\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
 \end{itemize}
+
+\alain{sur le dessin, enlever les vecteurs (0000) et juste dire qu'un arc non étiqueté ne modifie pas les compteurs.}
 
@@
 	\caption{A lossy VASS with three subsystems and four counters.
-Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ \alain{C pour controlleur, faire le dessin plus petit}represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
 					\label{air control}
@@
 Resilience here expresses that the aircrafts will both land, while explicit resilience or $k$-resilience asks whether they will do so in at most $k$ steps. 
-Explicit resilience is of interest since in general we want to ensure the safe space will
+Explicit resilience is of interest since in general we want to ensure the safe space will \alain{on n'utilise nulle part le terme explicit resilience....}
 be reached in

### 2023-06-23 state-resilience.tex
up
section intro

@@
 However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
+In the {\bf Flight Controller System} example for instance, $(q_{A,1}, q_{M,4}, q_{B,0})(\textbf{0})$
+belongs to $\Bad$, but is not reachable from $(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{0})$,
+hence it may seem unnecessary to check for the existence of a path to $\Safe$ from this element.
 Since we still consider $\Bad$ and $\Safe$ complements of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
@@
 \problemx{State-resilience problem (SRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
 % {$\Bad \cap \post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
@@
 \problemx{$k$-state-resilience problem (kSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
 % { $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
@@
 \problemx{bounded-state-resilience problem (BSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and two sets $\Safe \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
 {$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient ?\newline}

### 2023-06-23 introduction.tex
up

@@
 
-\alain{sur le dessin, enlever les vecteurs (0000) et juste dire qu'un arc non étiqueté ne modifie pas les compteurs.}
-
 \begin{center}
 	\begin{figure}
-			\hspace{0.75cm}
-\includegraphics[width=0.85\textwidth]{FigureB}
+			\hspace{2cm}
+\includegraphics[width=0.65\textwidth]{FigureB}
 	\caption{A lossy VASS with three subsystems and four counters.
-Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $M$ \alain{C pour controlleur, faire le dessin plus petit}represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
+Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $C$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.
+Arrows represent control-transitions.
+Unlabelled arrows by default do not modify counter values.}
 					\label{air control}
@@
 
-%\alain{il faut mieux expliquer le modèle, le rôle des 4 canaux, des messages, remarquer que 2 canaux sont des compteurs,....}
-
 % contrôle de trafic aérien, perturbation dans le planning → Bad (par exemple, trop d’avions qui cherchent à atterir en même temps et qui doivent être mis en stand-by au dessus de l’aéroport?) → Safe (trafic fluide à nouveau).

### 2023-06-23 resilience.tex
section intro

@@
 
-In this section we introduce the concepts of resilience, bounded-resilience and $k$-resilience for well structured transition systems. We show that the resilience problems are decidable
+\alain{pourquoi redire ça, c'est dans l'intro : In this section we introduce the concepts of resilience, bounded-resilience and $k$-resilience for well structured transition systems.
+We show that the resilience problems are decidable
 for  completion-post-effective $\omega^2$-WSTS with strong compatibility
-in the case of an upward-closed safe set,
+with $\Safe=\uparrow \Safe$,
 and that resilience is decidable for downward-compatible ideal-effective WSTS
-when the set of safe spaces is downward-closed. 
+when $\Safe=\downarrow \Safe$. 
 We show that the three resilience problems are however undecidable for WSTS with
-strong compatibility in general.
+strong compatibility in general.}
 
@@
 
-Recall the {\bf Flight Controller System} example, where the safe state corresponds to the aircrafts having both landed, i.e.
+In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
 $\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
@@
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in the {\bf Flight Controller System} example, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
+Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
 
@@
 		\end{proof}
-\alain{impossible de mettre end{proof} ???...et la preuve ne se finit pas clairement}
 %

### 2023-06-23 state-resilience.tex
section intro

@@
 
-Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$.
-However, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
-In the {\bf Flight Controller System} example for instance, $(q_{A,1}, q_{M,4}, q_{B,0})(\textbf{0})$
+Resilience is a strong property that implies that from every element of $\Bad$ there must exist a path to $\Safe$.
+However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
+In the {\bf Flight Controller System}, the state $(q_{A,1}, q_{M,4}, q_{B,0})(\textbf{0})$
 belongs to $\Bad$, but is not reachable from $(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{0})$,

### 2023-06-27 reachability/article.tex
up

@@
 {\bf Background.} Pushdown systems can be used to model the behavior of recursive programs.
-As a result, a variety of 
+They have applications for instance in inter-procedural control-flow analysis of recursive programs \cite{esparza1999automata, reps2005weighted}.
+More generally, a variety of 
  program analysis questions can be reduced to decision problems for games on pushdown systems.
-They have applications for instance in inter-procedural control-flow analysis of recursive programs \cite{esparza1999automata, reps2005weighted}.
-
-Two player games can be used in particular to
+
+In particular, two player games can be used to
  represent the
@@
 always holds \cite{arnold2003games}.
-
 Games on the graphs of pushdown systems have been extensively studied.
-It is known in particular from \cite{walukiewicz1996pushdown} that deciding the winner of a parity game on a graph of a pushdown system is an ${\sf EXPTIME}$-complete problem.
-Several generalizations of the pushdown systems 
-have been studied in depths, including
-tree-pushdown systems \cite{guessarian1983pushdown}, ordered
-multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and
-strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
-
-{\bf Parametric pushdown systems}
+It is known from \cite{walukiewicz1996pushdown} that deciding the winner of a parity game on a graph of a pushdown system is an ${\sf EXPTIME}$-complete problem.
+
+{\bf Parametric constraints.}
  Parametric pushdown systems provide a  formalism to
@@
 parameterized asynchronous shared-memory systems.
-These systems consist of a leader pushdown automaton
+These consist of a leader pushdown system
 and arbitrarily many identical contributor pushdown systems, communicating via a shared memory
 in the form of a register which can take finitely many values.
-This variant have been shown to have applications
-to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
-We consider here a different approach to extending pushdown systems with parameters, by allowing them to test equality of the stack content against parameters.
+They have applications
+in the dataflow analysis of concurrent programs~\cite{kahlon2008parameterization}. 
+We consider in this article a different approach to extending pushdown systems with parameters, by allowing equality test of the stack against parameters.
 Perhaps a similar model consists in 
@@
 They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
-They have \textcolor{red}{ phrases trop longues}
-been used to establish that CTL$^*$ model checking remains decidable
-when the formulas are allowed to include regular predicates on the stack content,
-and to obtain 
+They have 
+been used to  obtain 
 model checking algorithms for LTL and CTL$^*$ model checking
 for
-pushdown systems \cite{finkel1997direct}.
-Pushdown systems 
-with transitions that are conditioned by regular conditions on their stack content
-can be viewed as
-a special case of stack automata as seen in \cite{hopcroft1969formal}.
+pushdown systems~\cite{finkel1997direct}.
 Instead
-of checking the stack content against regular conditions, or, in a more limited manner,
+of checking the stack content against regular conditions, or
 against specified word values,
 we consider
-checks against unspecified word values that can be instantiated using parameters.
+checks against unspecified word values. These will be instantiated using parameters,
+and compared against using parametric guards.
 This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.
@@
 {\bf Related models.}
+Several generalizations of the pushdown systems 
+have been studied in depths, including
+ ordered multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and
+strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
 Extensions of pushdown systems with storage for later comparisons
@@
 only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers,
-% Even though pushdown register systems have no bound on the number of elements of the alphabet that they can store at any instant, over the course of a run they can really “remember” at most 3r of them,
-i.e. for any run of a register pushdown automaton with $r$ registers there exists an equivalent run %, i.e.
+i.e. for any run of a register pushdown automaton with $r$ registers there exists an equivalent run 
  with the same initial and final conﬁgurations, but in which every configuration contains register assignments drawn from only $3r$ elements. This  
- %can 
  leads to
@@
 the winner of a parity game on a graph of a pushdown system with  
-parameters.
+parametric guards.
 
@@
 $\text{\textit{Inf}}(\pi) = \{ v \in V \mid \forall i, \exists j > i, ~ v_j = v \}$.
-
-		\iffalse
-
-\subsection{Logics}
-
-We now brieﬂy review some standard deﬁnitions from mathematical logic.
-
-\begin{samepage}
-\begin{definition}
- A {\em vocabulary} $\tau$ is a set of relational symbols (denoted $P_1, P_2, \ldots $), each of which has a specified
- arity. A symbol $P \in \tau$ is called monadic if its arity is one, i.e., if it is used to
- denote sets.
-%
-A {\em $\tau$-structure} %(also called a {\em model})
-%
-$$ \mathfrak{A} = (A,	( P^{\mathfrak{A}} )_{P \in \tau})$$
-%
-consists of a set $A$ together with an interpretation of
- each $k$-ary relational symbol $P$ from $\tau$ as a $k$-ary relation on $A$; that is, a
-set $P^{\mathfrak{A}} \subseteq A^k$.
-% A structure $\mathfrak{A}$ is called finite if $A$ and $\sigma$ are finite sets. 
-% The {\em universe} of a structure is typically denoted by a Roman letter corresponding to the name of the structure; that is, the universe of $\mathfrak{A}$ is the set $A$, the universe of $\mathfrak{B}$ is $B$, and so on.
-We shall also occasionally write $a \in \mathfrak{A}$ instead of $a \in A$.\newline
-%
-\end{definition}
-\end{samepage}
-
-%\noindent As an illustration, if $\tau$ consists of a single relational symbol $E$ of arity $2$, graphs are possible $\tau$-structures.
-
-% Next, we deﬁne the  monadic second-order logic, or $\MSO$.
-
-Next, we deﬁne monadic second-order formulas over a vocabulary $\tau$, or $\MSO[\tau]$ for short. %We define free variables, and the semantics of $\MSO(\tau)$ formulas.
-
-\begin{samepage}
-\begin{definition}\label{MSO}
-We assume countably infinite sets of first-order variables {\em Var}
-and of second-order variables {\em VAR}. First-order variables
-will be typically denoted by $x, y, z, \ldots,$ with subscripts and superscripts,
-whereas second-order variables will be typically denoted by $X, Y, Z, \ldots ,$
-using subscripts and superscripts as well. We
-inductively define %terms and 
-formulas of the monadic second-order logic
-over vocabulary $\tau$ %as follows: 
-by the grammar:
-$$ \phi :=  P(x_1 , \ldots , x_k) \mid x_1 = x_2 \mid X(x) 
-		\mid \phi \vee \phi \mid \phi \wedge \phi \mid \neg \phi 
-		\mid \exists x ~ \phi \mid  \forall x ~ \phi
-		\mid \exists X ~ \phi \mid  \forall X ~ \phi		$$
-where $P \in \tau$ is a $k$-ary relational symbol, $X \in \text{VAR}$, and $x, x_1, \ldots, x_k \in \text{Var}$.
-\end{definition}
-\end{samepage}
-
-A formula that does not use existential ($\exists$) nor universal ($\forall$) quantiﬁers
-is called {\em quantifier-free}.
-%
-% Given a set of formulas $S$, formulas constructed from formulas in $S$ using only the Boolean connectives $\vee$, $\wedge$, and $\neg$ are called {\em Boolean combinations} of formulas in $S$.
-%
-We shall use the standard shorthand $\phi  \to \psi $ for $ \neg \phi \vee \psi $ 
-and $\phi \leftrightarrow \psi $ for
-$(\phi  \to \psi ) \wedge (\psi  \to \phi )$.
-
-If $\overrightarrow{x}$ is the tuple of all the free first-order variables of $\phi$,
-and $\overrightarrow{X}$ is the tuple of all the free second-order variables of $\phi$ 
-we write $\phi(\overrightarrow{x},\overrightarrow{X})$. A {\em sentence}
-is a formula without free variables. We use the standard semantic of $\MSO$,
-see Appendix~\ref{MSO appendix} for a more thorough induction over the formula.\\
-
-Concerning the logic $\MSO$, we are interested in the following decision problem.
-
-\problemx{$\MSO[\tau]$ model-checking}
-{A $\tau$-structure $\mathfrak{A}$ and a $\MSO[\tau]$-sentence $\phi$.}
-{Does $\mathfrak{A} \models \phi $ ?\newline}
-
-\textcolor{black}{We define the problem in a general abstract manner, instead of focusing on particular representation for the input, since we will consider many structures of infinite size.}
-
-{\em First-order logic over a vocabulary $\tau$} ($\FO[\tau]$ for short) is the fragment of $\MSO$ which is restricted to first-order quantifiers and variables: quantification is permitted only over individuals, and no second-order variables are used.
-
-		\fi
 
@@
 Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
-%\textcolor{red}{even}
  $k \in \N$.
@@
 
-%\subsubsection{Pebble pushdown automata}
 \subsection{Pebble pushdown systems}
@@
 \newcommand{\ppda}{\mathcal{I}}
-% \Denarius
-% \vernal
-% \mathscr{Z}
-% \mathcal{I}
-% \pluto
-% \capricornus
-% \Bumpeq
-% \Pfund
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% An $i$-configuration of $\ppda$ is an element of
-%$ Q\times W^* \times (W^*)^{\{1, \ldots i\}}$.
-% By $\Conf_i(\ppda)$ we denote the set of $i$-configurations of $\ppda$.
 By $\Conf(\ppda)=Q\times \Gamma^* \times (\Gamma^* \biguplus \{ \bot \})^{\{1, \ldots, n\}}$ we denote the set of
-% By $\Conf(\ppda) = \bigcup_{0 \leq i \leq n} \Conf_i(\ppda)$ we denote the set of
 {\em configurations} of $\ppda$. As expected, we rather write $q(w, \mu)$ instead of $(q, w, \mu)$.
@@
 The $\emptyset$-pebble assignment is denoted by $\mu_{init} :  \{ 1, \ldots, n\} \rightharpoonup  \Gamma^* $ and is the 
-%partial function from  $\{ 1, \ldots, n\}$ to $ \Gamma^* $ that
-%does not map any $j \in \{ 1, \ldots, n \}$ to an output.
 totally undefined function.
@@
 \begin{samepage}
-%\begin{definition}
 An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
@@
  for all words $w \in  \Gamma^*$, and for all $\{ 1, \ldots , i \}$-pebble assignments $\mu$ such that $\mu(j) = wa$ 
- %(respectively $\mu(j)= {\gamma_{init}}  $) 
  for each
@@
  $\mu(j) \neq wa$ 
- %(resp. $\mu(j) \neq {\gamma_{init}} $) 
  for each $j \in \{ 1, \ldots , i \} \setminus S$,
@@
 \item if $m \in \Op(\Gamma)$, 
-% $(q,wa,\mu) \rightarrow_{\ppda} (q',wm, \mu)$, 
 either
@@
 \end{itemize}
-%\end{definition}
 \end{samepage}
@@
 \par\noindent\ignorespacesafterend
-We are interested in games over pebble pushdown systems, % namely, 
-%reachability, B\"uchi, and parity games.
+We are interested in games over pebble pushdown systems, 
 mainly, parity games.
@@
 a 
-%labeled
  transition system, one needs only to provide 
-%a mapping $\Omega$ and 
 a partition of the set of
 configurations 
-%$S$ into two sets $S_0$ and $S_1$
  to obtain an arena.
@@
 generalization of pushdown systems \cite{aho1969nested,Gre70,maslov1976multilevel}. 
-% Alfred V. Aho. Nested stack systems. J. ACM, 16(3):383–406, 1969.
-% S. Greibach. Full AFL’s and nested iterated substitution. Inf. Control, 16(1):7–35, 1970.
-% A.N. Maslov. Multilevel stack systems. Problemy Peredachi Informatsii, 12:55–62, 1976.
 A stack of a pushdown automaton is seen as a {\em level $1$ stack}. 
@@
 	stack over $\Gamma$, i.e. a word  in $\Gamma^*$.
-% an arbitrary sequence $w_0 w_1 \ldots w_n $ of elements of $\Gamma$ for some $n \in \N$. 
 A  {\em level $n$ stack} (or {\em $n$-stack}) over an alphabet $\Gamma$, for $n \geq 2$, is a 
@@
 The operations 
-%we consider
 additionally
@@
 
-% For all $m \in \N$, 
 The following operations can be performed on a $1$-stacks of length $m \in \N$.
-% $ w_0 w_1 \ldots w_m \in \Gamma^*$.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-% We omit the stack alphabet $\Gamma$ from 
-%The operation $pop_j$ is undefined on a stack whose top stack of level $j$ is empty.
 The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\
@@
 
-% \begin{definition}
 \begin{samepage}
@@
 \end{samepage}
-% \end{definition}
 
@@
 we denote the set of {\em configurations} of an $n$-HPDS $\H$. 
-%is a pair $(q, s)$ where $q \in Q$ and $s$ is an $n$-stack over $\Gamma$. 
 As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.
@@
 and
-%$s' = \theta(s)$.
 $s' = op(s)$.
-
-\iffalse
-For our constructions it would be simpler to assume that k-HPDS can work
-also on stacks of lower levels, in particular on 1-stacks. Of course we can always
-simulate a j-stack, for j < k with an k-stack but in the notation it requires some
-additional parenthesis that make it less readable.
-
-To define a game on the graph of a HPDA, we assign a player to each control
-state, and we consider an initial configuration: a game structure on a HPDS H
-is a tuple G = (H, P0 , P1 , s0 ), where P = P0 ⊎ P1 is a partition of the control
-states of H, and s0 ∈ Sk . This extends naturally to a partition of the set of
-configurations: with the notations of Section 2.1, V0 = P0 × Sk , V1 = P1 × Sk ,
-and E is defined above.
-\fi
 
@@
 
-% \subsection{From pebble pushdown systems to higher-order pushdown systems}
-
-In a $2$-HPDS, the operation $\text{\textit{push}}_2$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
+In a $2$-HPDS, the operation $\text{\textit{push}}_2$ 
 allows to ``copy'' the top level $1$ stack. The current word is hereby stored away and left untouched
@@
 
-% But cannot test
-
-\iffalse
-In an $n$-HPDS, the operation $\text{\textit{push}}_j$ % allowig to copy a part of the stack, are responsible for the fact the hierarchy of HPDS is strict.
-allows to ``copy'' the top stack of level $j$ for all $2 \leq j \leq k$. 
-\fi
-
 In \cite{CaWoe03, Woeh05, carayol2006automates} however 
-%the authors
 Carayol and W\"ohrle
@@
 %
-\iffalse
-Dans ce sous-paragraphe, nous présentons formellement les automates à pile
-de niveau k sur le jeu d'opérations symétriques Opsk et sur le jeu d'opérations
-classiques COpsk . Nous donnons à cette occasion un bref historique de la notion
-d'automate à pile de piles. Enfin, nous établissons l'équivalence entre les deux
-modèles d'automates à chaque niveau en tant qu'accepteurs de langages. Ce résultat a été obtenu dans [CW03] en collaboration ave Stefan Wöhrle. La preuve
-complète apparait dans [Wöh05]. Indépendamment, ce résultat a été obtenu dans
-[Fra05].
-\fi
 %
@@
 \end{theorem}
-
-\iffalse
-\begin{theorem}{\cite{Woeh05, carayol2006automates}}
-If $\H$ is an $n$-HPDS  then there
-exists an $n$-HPDS$^=$  $\H'$ such that
-$\H$ and $\H'$ %generate the same graph.
-induce isomorphic transition systems.
-\end{theorem}
-% Proposition 3.12
-
-\begin{theorem}{\cite{Woeh05, carayol2006automates}}
-If $\H$ is an $n$-HPDS$^=$ then there
-exists an $n$-HPDS $\H'$ such that
-$\H$ and $\H'$ %generate the same graph.
-induce isomorphic transition systems.
-\end{theorem}
-
-% Proposition 3.18 If A is a higher-order pushdown system with equality pop
-% of level n then there exists a higher-order pushdown system B of level n such
-% that A and B generate the same graph.
-\fi
 
@@
 for every stack $s$ of level $n$
-% there is
  a unique shortest sequence of instructions which creates $s$ from
@@
 
-\iffalse
-\textcolor{red}{
-Now then, because the class of $n$-HPDS$^=$ transitions systems
-is the same as
-the class of $n$-HPDS transitions systems, we have the following result.
-}
-\fi
-
 \begin{theorem}\label{HPDA parity game}
@@
 \end{theorem}
-
-\iffalse
-"To show that such an automaton exists we introduce the notion of a weak
-popping higher-order pushdown automaton. A weak popping automaton is only
-allowed to execute a pop instruction of level $j \geq 2$ if the two top level $j$ stacks
-coincide. We skip a formal definition of a weak popping higher-order pushdown
-automaton and just mention that even though this automaton model is equipped
-with a built-in test on the equality of two stacks of the same level, it is equivalent
-to the usual model. All proofs are given in the full version of this article [4]."
-
-Pour pouvoir obtenir une notion pertinente d'ensemble rationnel de piles de
-piles, il faut considérer un jeu d'opérations où la destruction inconditionnelle des
-piles est remplacée par une version plus symétrique où la dernière pile de niveau ℓ
-ne peut être détruite que si elle est égale à la précédente. Nous établissons dans le
-paragraphe 4.1 l'équivalence (en temps qu'accepteurs de langages) les automates
-à pile définis sur ces deux jeux d'opérations ( f. théorème 4.1.23). Nous montrons,
-dans le paragraphe 4.6, le manque de pertinence de la notion de rationalité induite
-par le jeu d'opérations classiques et par là même, nous établissons la nécessité de
-considérer le jeu d'opérations symétriques.
-
-L'opération de destruction symétrique a été introduite pour des raisons techniques dans 
-[CW03] et a été utilisée de manière 
-systématique dans [Car05] et
-[Fra05]. 
-Nous montrerons, dans le sous-paragraphe 4.1.5, que les automates à pile
-de piles définis en utilisant les opérations destrk ou les opérations copyk acceptent
-les mêmes langages.
-
-Les automates à pile d'ordre supérieur ont été introduits dans les années 70.
-Dans [Gre70℄, Greibach attribue l'idée de ces automates à Aho et Ullman. Ces
-automates sont aussi définis dans [Mas76℄. Les automates sur COpsk diffèrent
-légèrement des automates considérés par ces auteurs et de ceux onsidérés dans
-[DG86, Eng91℄. La différence majeure réside dans la définition des piles de piles.
-Les piles de niveau k sont des suites non vides de couples formés d'un symbole de
-pile et d'une pile de niveau k − 1. Comme remarqué dans [KNU02℄, ces symboles
-de pile supplémentaires peuvent être simulés dans le modèle des automates sur
-COpsk .
-
-Les langages a eptées par les automates à pile sur COpsk sont onnus sous
-le nom de langages k -OI ou de langages indexés de niveau k . La première déno-
-mination vient des travaux de Damm [Dam82]
-
-Nous allons maintenant établir que les encodages des opérations simulent bien
-les opérations sur les encodages des piles.
-
-Lemme 4.1.21. Pour toute pile s ∈ Stacksk (Γ) et pour tout θ ∈ Opsk ,
-[[ θ ]]k ([[ s ]]) =
-{
-[[ θ(s) ]]		si θ(s) est défini,	
-non définie		sinon.
-
-Proposition 4.1.22. Tout langage accepté par un automate à pile sur Opsk est
-accepté par un automate à pile sur COpsk .
-
-Démonstration. Soit A = (Γ,Σ,τ,Q,I,F,∆) un automate à pile sur Ops⋆k . Nous
-définissons un automate B = (Γ,Σ,τ,Q,I,F,∆B ) où l'ensemble des transitions ∆B
-est défini par:
-∆B = {(p,x,η,q) | (p,x,θ,q) ∈ ∆ et η ∈ [[ θ ]]k }.
-Par le lemme 4.1.21, il suit que pour tout w ∈ (Σ)∗ , p ∈ Q et s ∈ Stacksk (Γ):
-w
-(q0 , [ ]k ) −→ (p,s)
-A
-⇔
-w
-(q0 , [ ]k ) −→ (p,[[ s ]]).
-
-Théorème 4.1.23. Les langages acceptés par les automates à pile sur Opsk et
-sur COpsk coïncident.
-
-\fi
 
@@
 When the automaton $\ppda$ is in configuration $q(w)$ and drops a pebble, instead of making a pop or a push, we need
-to %go to state $q'$ and word $w$ and 
+to
 	store the information that the pebble has been
@@
 a parametric pushdown parity game.
-% problem of deciding winning configuration for pebble pushdown parity games.
 
@@
 %
-
-% Here a function is in {\sf elementary} if it is  $O(h_k (n))$ for one of the $k$-fold exponential functions $h_k$ with $h_0 (n) = n$ and $h_{k+1} (n) = 2^{h_k (n)}$.
 
@@
 \begin{theorem}[\cite{Sto74}]\label{FOnonelementary}
-
-%  Assume that $FPT \neq AW[*]$, and let $f$ be an elementary function and let $p$ be a polynomial.
-%  There is no model-checking algorithm for first-order logic on the class of words 
-%  whose running time is bounded by
-%  $$ f(k) \cdot p(n)$$
-
-% where $k$ denotes the size of the input sentence of the model-checking problem and $n$ the
-% size of the input word.
-
 The {\sc $\FO$ satisfiability on words} problem is not in {\sc ELEMENTARY}.
-	% SAT(N,<) is not elementary-recursive
-	% and in fact SAT(N,<) \not\in NSPACE(g(log_b(n),0)) for all b > 3.
 \end{theorem}
@@
 {\sc $\FO$ satisfiability on words}
-%We consider reduction 
-towards the problem of %following problem.
+towards the problem of 
 {\sc Parametric pushdown parity game}.
@@
 for player $0$ to ensure that
-%$\mathfrak{W}$ 
 $w$
@@
 Recall $\psi$ is quantifier-free and in disjunctive normal form, i.e.
- $$\phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).$$
+ \[ \phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).\]
 
@@
 Then player $1$, the universal player, chooses one of the atomic formula in the clause to test.
-%The process can be seen in Figure~\ref{DNF}. 
 Testing the atomic formula is the purpose of a gadget 
-%$ \C_{\psi_{i,j}}$ built such that
 in which
  player $0$ has a winning strategy 
- %in $ \C_{\psi_{i,j}}$
   if and only if 
-$w$ satisfies %$\psi_{i,j}$
+$w$ satisfies 
 the atomic formula
@@
 
+For solving parametric pushdown parity games, on the other hand, we introduced the notion of pebble pushdown parity games, and
+pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
+can be seen as a form of one-counter automata extended with pebbles. 
+A pebble one-counter automata is then a one-counter automata that
+can use a set of pebbles as markings on the set of non-negative integers, with a 
+$\text{\textit{drop}}$ recording the current counter value, and a 
+$\text{\textit{lift}}$ poping the last dropped
+pebble while requiring that the current counter value 
+corresponds to the one from the last dropped pebble.
+%
+%Hence, u
+Unlike for parametric extensions, 
+decidability of 
+what should be called
+pebble one-counter parity games follows from
+decidability of pebble pushdown parity games. 
+We believe it natural to study such a pebble one-counter automata model
+on its own.
+In particular we believe that the question of whether or not
+solving pebble one-counter reachability games is nonelementary is worth investigating.
+% It may very well be the case that the hierarchy collapses instead, at level e.g. 2 or 3.
+
+Of note is that reachability games and parity games are 
+not the only problems one can consider, as one can
+also explore the complexity of reachability itself. 
+In particular
+one can 
+% ask whether or not the parametric pushdown automata model described in this section  admits effective preservation of regularity.
+study the precise complexity of the
+{\sc $n$-PPDA reachability} problem.
+Since reachability can be viewed as a reachability game, we know that
+{\sc $n$-PPDA reachability} is decidable in $(n+1)$-$\EXP$, 
+but the precise complexity of the problem remains unknown.
+We are furthermore convinced that it is worthwhile to investigate further comparisons between parametric pushdown automata, pebble pushdown automata, 
+%pebble alternating tree automata 
+and register pushdown automata, especially since the latter have been shown to only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers.
+
 \bibliography{bibliography/bib}

### 2023-06-28 resilience.tex
section

@@
 
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$.
+We now briefly investigate the question to control a system (non-resilient for a set $Safe$) such that its controlled behavior is resilient for $Safe$.
+%
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{Safe}$.
 
@@
 
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(X)$-resilient.
-The \emph{maximal $X$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_X=(S,\rightarrow_X,\leq)$ where $\rightarrow_X$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_X$ is $X$-resilient.
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(Safe)$-resilient.
+The \emph{maximal $Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{Safe}=(S,\rightarrow_{Safe},\leq)$ where $\rightarrow_{Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{Safe}$ is $Safe$-resilient.
 %\alain{vérifier que la plus grande existe}
@@
 %
-For example, VASS allows such guards.
+For example, VASS and most of concrete WSTS allow such guards.
   %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
@@
 \begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $X$, the maximal resilient subsystem is a computable WSTS.
+Given an effective WSTS with effective pred-basis and an upward-closed set $Safe$, the maximal resilient subsystem is a computable WSTS.
 \end{theorem}
@@
 \begin{proof}
-Since $X \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(X)=\{x_1,x_2,...,x_p\}$. 
+Since $Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(Safe)=\{x_1,x_2,...,x_p\}$. 
 %
 %
-Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(X)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
 		\end{proof}
 %
-It is also possible to define $(s_0,X)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} X$.
+It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.

### 2023-06-29 introduction.tex
section intro
up

@@
 % trop d'hypotheses trop fortes sur les WSTS
-\alain{harmonier les noms des pb}
 \noindent
@@
 % We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.
-Resilience here expresses that the aircrafts will both land, while explicit resilience or $k$-resilience asks whether they will do so in at most $k$ steps. 
-Explicit resilience is of interest since in general we want to ensure the safe space will \alain{on n'utilise nulle part le terme explicit resilience....}
+Resilience here expresses that the aircrafts will both land, while $k$-resilience asks whether they will do so in at most $k$ steps. 
+$k$-resilience is of interest since in general we want to ensure the safe space will 
 be reached in

### 2023-06-29 resilience.tex
up

@@
 
-\alain{pourquoi redire ça, c'est dans l'intro : In this section we introduce the concepts of resilience, bounded-resilience and $k$-resilience for well structured transition systems.
-We show that the resilience problems are decidable
-for  completion-post-effective $\omega^2$-WSTS with strong compatibility
-with $\Safe=\uparrow \Safe$,
-and that resilience is decidable for downward-compatible ideal-effective WSTS
-when $\Safe=\downarrow \Safe$. 
-We show that the three resilience problems are however undecidable for WSTS with
-strong compatibility in general.}
+% \alain{pourquoi redire ça, c'est dans l'intro : In this section we introduce the concepts of resilience, bounded-resilience and $k$-resilience for well structured transition systems.
+We show that the resilience problems are decidable for  completion-post-effective $\omega^2$-WSTS with strong compatibility with $\Safe=\uparrow \Safe$, and that resilience is decidable for downward-compatible ideal-effective WSTS when $\Safe=\downarrow \Safe$.
+We show that the three resilience problems are however undecidable for WSTS with strong compatibility in general.}
 

### 2023-06-30 reachability/article.tex
up

@@
  to ensure that the desired 
-property, as expressed by the winning condition,
+property
 always holds \cite{arnold2003games}.
@@
 {\bf Parametric constraints.}
- Parametric pushdown systems provide a  formalism to
-reason about 
-recursive programs
-making use of parametric constraints. 
+% Parametric pushdown systems provide a  formalism to reason about  recursive programs making use of parametric constraints. 
+For certain types of computer programs, such as embedded systems, under-specified constraints are necessary rather than only considering concrete specifications.
+Unspecified values represented by parameters have been used
+for instance in parametric timed automata~\cite{AHV93-stoc} and parametric one-counter automata~\cite{haase2012complexity, lechner2015complexity}.
+Verification of the desired properties of the system is then performed without concrete values.
+ We consider in this article a similar approach, only applied to pushdown systems.
+Rather than comparing a counter against a parametric number, we allow comparison of a stack content against a parametric word.
+  Perhaps a similar model consists in  pushdown systems with transitions that are conditioned by regular conditions on their stack content, which can in particular test the stack content against specific words.
+They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
+They have  been used to  obtain  model checking algorithms for LTL and CTL$^*$ model checking
+for pushdown systems~\cite{finkel1997direct}.
+Instead of checking the stack content against regular conditions, or against specified word values, we consider checks against unspecified word values.
+These will be instantiated using parameters, and compared against using parametric guards.
+This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.
+
+{\bf Related models.}
 Recently, different variants of parametric pushdown systems have been introduced in the literature 
@@
 in the dataflow analysis of concurrent programs~\cite{kahlon2008parameterization}. 
-We consider in this article a different approach to extending pushdown systems with parameters, by allowing equality test of the stack against parameters.
-Perhaps a similar model consists in 
-pushdown systems 
-with transitions that are conditioned by regular conditions on their stack content, which
-can in particular
-test the
-stack content against
-specific words. 
-They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
-They have 
-been used to  obtain 
-model checking algorithms for LTL and CTL$^*$ model checking
-for
-pushdown systems~\cite{finkel1997direct}.
-Instead
-of checking the stack content against regular conditions, or
-against specified word values,
-we consider
-checks against unspecified word values. These will be instantiated using parameters,
-and compared against using parametric guards.
-This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.
-
-{\bf Related models.}
-Several generalizations of the pushdown systems 
-have been studied in depths, including
- ordered multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and
-strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
+We consider in this article a different approach to extending pushdown systems with parameters, by making them appear directly in equality tests against the stack.
+
+% Several generalizations of the pushdown systems  have been studied in depths, including ordered multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
 Extensions of pushdown systems with storage for later comparisons
@@
 
-	For the nonelementary lower bound,
-we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
-from \cite{Sto74}, to the
-problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.
-
 	For the upper bound, we start by replacing parameters by pebbles acting as registers, leading to a more general model, pebble pushdown systems.
@@
 
-{\bf Overview. }
-In Section 2, we provide general notations and preliminary definitions. 
-Section 3 will deal with the introduction of  parametric pushdown games.
-% Section 4 is devoted to using pebble alternating tree systems with parity acceptance condition to solve parametric pushdown system parity games.
-We leave the proof of decidability itself  for Section 5.
-Section 4 is Reduction to higher-order pushdown systems
-Section 5 at last shows a nonelementary lower bound for the problem.
+	For the nonelementary lower bound,
+we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
+from \cite{Sto74}, to the
+problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.
+
+% {\bf Overview. }
+% In Section 2, we provide general notations and preliminary definitions. 
+% Section 3 will deal with the introduction of  parametric pushdown games.
+% Section 4 is Reduction to higher-order pushdown systems
+% Section 5 at last shows a nonelementary lower bound for the problem.
 

### 2023-06-30 reachability/article.tex
up

@@
 % Parametric pushdown systems provide a  formalism to reason about  recursive programs making use of parametric constraints. 
-For certain types of computer programs, such as embedded systems, under-specified constraints are necessary rather than only considering concrete specifications.
+For certain types of computer programs, such as embedded systems, it can
+be necessary to consider under-specified constraints rather than concrete specifications only.
 Unspecified values represented by parameters have been used
@@
 	\begin{figure}
-			\hspace{3.81cm}
-\includegraphics[width=0.4\textwidth]{figures/parity_game}
+			\hspace{3.95cm}
+\includegraphics[width=0.35\textwidth]{figures/parity_game}
 	\caption{An example of an arena.
@@
 Let us consider the arena presented in  Figure~\ref{example_parity_game}. 
-Configurations are partitioned into two sets: circles belonging to player $0$ and diamonds
+Configurations are partitioned into circles belonging to player $0$ and diamonds
 belonging to player $1$. 
 The priorities are $\{ 0,1,2,3\}$.
-% The transitions and priority mappings may be derived from the picture.
 Plays $\pi_1 = q_0 q_1 q_2 (q_4 q_7)^\omega$ and 

### 2023-07-04 reachability/article.tex
Too long, 15 pages have to shorten to 12

@@
 \usepackage{wrapfig}
-
-\newcommand\sg[1]{\todo[inline,size=\scriptsize]{#1 - \textbf{Stefan}}}
-\newcommand\mh[1]{\todo[inline,size=\scriptsize]{#1 - \textbf{Mathieu}}}
-\newcommand\sgchanged[1]{{\color{red}{'1}}}
-\newcommand\mhchanged[1]{{\color{blue}{'1}}}
 
@@
 
-% \author{Stefan G\"oller}{University of Kassel \and School of Electrical Engineering and Computer Science \and Kassel, Germany }{stefan.goeller@uni-kassel.de}{}{The author was supported by the Agence Nationale de la Recherche grant no.
-ANR-17-CE40-0010.}
 \author{Mathieu Hilaire}{Université Paris-Saclay,
@@
 
-\acknowledgements{The author would like to thank Benedikt Bollig, Stefan G\"oller and Alain Finkel for helpful discussions and feedback.}
+\acknowledgements{The author would like to thank Benedikt Bollig, Stefan G\"oller and Alain Finkel for helpful discussions and feedback.
+The author would also like to thank the anonymous reviewers from FSTTCS 2022 for their helpful suggestions and remarks.}
 
@@
 	parity games over an extension of pushdown systems.
-	More specifically, we consider parameters pushdown systems, in
+	More specifically, we consider parametric pushdown systems, in
 	which parameters can be instantiated to some word over a stack alphabet,
@@
 thus provides some way 
- to ensure that the desired 
-property
-always holds \cite{arnold2003games}.
-Games on the graphs of pushdown systems have been extensively studied.
+ to ensure the desired 
+property~\cite{arnold2003games}.
 It is known from \cite{walukiewicz1996pushdown} that deciding the winner of a parity game on a graph of a pushdown system is an ${\sf EXPTIME}$-complete problem.
@@
 {\bf Parametric constraints.}
-% Parametric pushdown systems provide a  formalism to reason about  recursive programs making use of parametric constraints. 
 For certain types of computer programs, such as embedded systems, it can
@@
 Unspecified values represented by parameters have been used
-for instance in parametric timed automata~\cite{AHV93-stoc} and parametric one-counter automata~\cite{haase2012complexity, lechner2015complexity}.
-Verification of the desired properties of the system is then performed without concrete values.
+for instance in parametric timed automata~\cite{AHV93-stoc} and parametric one-counter automata~\cite{haase2012complexity, lechner2015complexity} in which
+verification of the desired properties of the system is then performed without concrete values.
  We consider in this article a similar approach, only applied to pushdown systems.
@@
 They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
-They have  been used to  obtain  model checking algorithms for LTL and CTL$^*$ model checking
+They have  been used to  obtain  algorithms for LTL and CTL$^*$ model checking
 for pushdown systems~\cite{finkel1997direct}.
-Instead of checking the stack content against regular conditions, or against specified word values, we consider checks against unspecified word values.
-These will be instantiated using parameters, and compared against using parametric guards.
+Instead of checking the stack content against regular conditions, or against specified word values, we consider checks against unspecified word values, instantiated using parameters, and compared against using parametric guards.
 This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.
@@
 in the dataflow analysis of concurrent programs~\cite{kahlon2008parameterization}. 
-We consider in this article a different approach to extending pushdown systems with parameters, by making them appear directly in equality tests against the stack.
-
-% Several generalizations of the pushdown systems  have been studied in depths, including ordered multi-pushdown systems \cite{breveglieri1996multi, atig2012model}, annotated higher-order pushdown systems \cite{maslov1976multilevel, broadbent2012saturation}, and strongly normed multi-pushdown systems \cite{czerwinski2012reachability}.
+This article's approach to extending pushdown systems with parameters is different in that it makes them appear directly in equality tests against the stack.
+
 Extensions of pushdown systems with storage for later comparisons
@@
 
-% {\bf Overview. }
-% In Section 2, we provide general notations and preliminary definitions. 
-% Section 3 will deal with the introduction of  parametric pushdown games.
-% Section 4 is Reduction to higher-order pushdown systems
-% Section 5 at last shows a nonelementary lower bound for the problem.
-
 \section{Preliminaries}
@@
 Note that $(S,{\rightarrow})$, where $S=S_0 \uplus S_1$, forms
-a %labeled 
+a   
 transition system, and, in particular, given a  
-%labeled 
 transition system,
@@
 
-% \subsection{Strategies}
-
 A {\em strategy} for player $i \in \{ 0,1\}$ is a function $\sigma_i : S^* S_i \to S$ such
@@
 
-%  \subsubsection{Winning conditions}
-
 Once we have provided players with an arena, it remains to define  properly what the winning condition is in order to define a game.
 Given an arena $A = (S_0, S_1, \rightarrow)$, a {\em winning condition}
-%$\win \subseteq S^\infty$
 $\win$
@@
  from $s$} for player $0$ if 
-% no matter which
-% strategy $\sigma_{1}$ the other player pursues,
-% the play $\pi(s, \sigma_0, \sigma_1)$ is in $\win$.
 every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
@@
  from $s$} for player $1$ if 
-% no matter which
-% strategy $\sigma_{0}$ the other player pursues,
-% the play $\pi(s, \sigma_0, \sigma_1)$ is not in $\win$.
 every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.
-
-% \subsection{Reachability Games}
 
@@
 An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
-%this path
 the sequence 
  is even, i.e.
- %$\min ~ \text{\textit{Inf}}( \Omega (r( \beta ))) $
 if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even.  
-% The winner of a finite play is the player whose opponent is unable to move, i.e. a play such that the last configuration of the play is in $S_{0}$ (resp. $S_1$) is player $1$ (resp. player $0$). 
 A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
@@
 $\pi_2 = (q_0 q_1 q_2 q_4)^\omega$ are both winning for player $0$. Indeed, in $\pi_1$, 
-the minimal priority encountered infinitely often is %that of $q_7$, that is, 
+the minimal priority encountered infinitely often is   
 				$2$, 
-whereas in $\pi_2$, it is %that of $q_2$, that is, 
+whereas in $\pi_2$, it is   
 			$0$. 
@@
  $P$ is a finite {\em   set of parameters}, \textcolor{black}{with $\Gamma \cap P = \emptyset$};
-% \item  $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Gamma^*$ is finite {\em  set of rules},
   $R   \subseteq  Q  \times (\Gamma \biguplus P)  \times Q  \times \Op(\Gamma)$ is finite {\em  set of rules};
@@
 As before we write the top
-of the stack at the right of the word%(we are considering suffix rewriting as in Chapter 15)
-. 
-% A configuration is a  state and a stack content: $(q, w)$,
-% which we rather write $q(w)$, where $w  \in  W^*$, $q  \in Q$. 
+of the stack at the right of the word. 
 By $\Conf(\mathcal{Z})=Q\times  \Gamma^*$ we denote the set of
@@
 \begin{samepage}
-%\begin{definition}
 A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
 and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
-%labeled 
 transition system  
 $T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
-%\Lambda_{\mathcal{Z}}, 
 \rightarrow_{\mathcal{Z},\mu})$ where
-% $\Lambda_{\mathcal{Z}} = \Gamma \cup P$ and
 for all $ q, q'  \in  Q$, 
@@
 for all $ a  \in  \Gamma$, 
-\iffalse
-\begin{eqnarray*}
-q(wa) \xrightarrow{%a
-}_{\mathcal{Z},\mu} q'(w') & \text{if} &  (q, a, q' , op)  \in  R, \\
-q(\mu(p) ) \xrightarrow{%p
-}_{\mathcal{Z},\mu} q'(\mu(p) %\textcolor{red}{\cdot} 
-w) & \text{if} &  (q, p , q' , w)  \in  R.
-\end{eqnarray*}
-% \end{definition}
-\fi
-$q(wa) \xrightarrow{%a
-}_{\mathcal{Z},\mu} q'(w')$
+$q(wa) \rightarrow_{\mathcal{Z},\mu} q'(w')$
 if there exists a rule in $R$
@@
 such that
-		% $\text{\textit{top}}(w) = a$
-		% and 
-% $w' = \theta(w)$.
 either of the following holds
@@
 
-A {\em $\mu$-run} from $q_0(a_0)$ to $q_n(a_n)$ in $\mathcal{Z}$ is a corresponding
-path in the transition system 
-$T_{\mathcal{Z}}^\mu$ 
-induced by $\mathcal{Z}$ and $\mu$.
-%
-%and will be noted
-%$q_0(z_0) \xrightarrow{\lambda_0}_{\mathcal{Z},\mu} q_1(z_1) \xrightarrow{\lambda_1}_{\mathcal{Z},\mu} \cdots \xrightarrow{\lambda_{n-1}}_{\mathcal{Z},\mu} q_n(z_n)$.
-%
-As with pushdown systems, we say $\pi$ is {\em accepting} if $q_0 = q_{init}$, $a_0=\gamma_{init}$,
-and $q_n\in F$.
-%
-% We say {\em reachability holds} for $\mathcal{Z}$ if there is a $\mu$ such that there is a $\mu$-run in $\mathcal{Z}$ from $q_{init}(\gamma_{init})$   to some configuration $q(w)$ for some $q\in F$ and some $w \in \Gamma^*$.
-% We refer to Figure~\ref{example_pda} for an instance of a PDA for which reachability holds.
-
 In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
-we prefer to write $q(w)\xrightarrow{%\lambda
-}_{\mathcal{Z}, u}q'(w')$
-to denote $q(w)\xrightarrow{%\lambda
-}_{\mathcal{Z}, \mu}q'(w')$ and
-will call the $\mu$-run an {\em $u$-run}. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\xrightarrow{%\lambda
-}_{\mu}$ (resp. $\xrightarrow{%\lambda
-}_{u}$)
-instead of $\xrightarrow{%\lambda
-}_{\mathcal{Z}, \mu}$ (resp. $\xrightarrow{%\lambda
-}_{\mathcal{Z},u}$). 
+we prefer to write $q(w)\rightarrow_{\mathcal{Z}, u}q'(w')$
+to denote $q(w)\rightarrow_{\mathcal{Z}, \mu}q'(w')$. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\rightarrow_{\mu}$ (resp. $\rightarrow_{u}$)
+instead of $\rightarrow_{\mathcal{Z}, \mu}$ (resp. $\rightarrow_{\mathcal{Z},u}$). 
 Since we are concerned with the graph of configurations and not the language recognized, we omit the input alphabet from the model and will use the name pushdown system rather than pushdown automaton.
 
-		\iffalse
-
-\subsection{PPDS reachability}
-
-Note that the reachability problem for parametric pushdown systems consists in
-the following decision problem.
-
-\problemx{$n$-PPDS reachability}
-{An $n$-PPDS $\mathcal{Z}$.}
-{Does 
-%reachability hold for
-an accepting $\mu$-run for some $\mu\in(\Gamma^*)^P$ exist in
- $\mathcal{Z}$?\newline}
-
-		\fi
-
-% \subsubsection{Parametric pshdown games}
 \subsection{PPDS games}
 
-Several model checking problems can be
-expressed as decision problems for games.
-For instance 
-in the case of pushdown systems, 
-the modal $\mu$-calculus
-model checking problem is polynomially equivalent to solving the parity game problem~\cite{emerson1993model}.
 We are interested in this article in parametric variants of reachability games and parity games.
-
 One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a 
@@
 A winning strategy for player $0$ allows to synthesize a controller that restricts the environment and ensures that {the property expressed by the winning condition always holds.}
-
-% \subsection{Parameter valuation arenas}
 
@@
 a parametric automaton $\A$ with parameters $P$ taking values in $\Gamma^*$ and a parameter valuation $\mu: P \to \Gamma^*$.
-
-\iffalse
-\begin{center}
-	\begin{figure}
-		\hspace{2cm}
-\includegraphics[width=0.7\textwidth]{figures/parameter_arena}
-	\caption{An illustration of the arena for a parametric automaton $\A$ with 
-%	two parameters, $p_0$ and $p_1$,	
-	$P_0 = \{p_0\}$, $P_1 = \{p_1\}$ and
-%	 that can take values in $[0,N]$ for some $N \in \N$. 
-	$\M = [0,N]$ for some $N\in \N$.
-	 Here $\mu_\bot$ is 
-	%the partial parameter assignment with domain $\emptyset$,
-	the totally undefined function and can be viewed rather as 
-	the function that assigns $\bot$ to both parameters.
-For $i,j \in [0,N]$, $\mu_i$ is the partial parameter assignment with domain $\{p_0\}$ that maps $p_0$ to $i$ and $\mu_{i,j}$ is the parameter assignment that maps $p_0$ to $i$ and $p_1$ to $j$.
-	 %Lastly $T^\mu_\A$ denotes the transition system induced by the concrete automaton corresponding to $\A$ with parameter valuation $\mu$.
-	  As before, the circles belong to player $0$ and the diamonds to player~$1$.} 
-	\label{parametric_arena}
-	\end{figure}
-\end{center}
-\fi
-
-\iffalse
-Given a parametric automaton $\A$ with a set of parameters $P = P_0 \biguplus P_1$ that can take 
-values in the set $\Gamma^*$,
-a {\em parametric game} for $\A$ is a game played 
-on the 
-arena $A_\A$
-that include both 
-the parameter valuation arena
-$A_{P_0, P_1,\Gamma}$
-and, %   
-for all parameter valuations $\mu: P\rightarrow \Gamma^*$,
-the transition system %s 
- $T_\A^\mu$
-\---- albeit with configurations additionally indexed by $\mu$ %in the tuple
- to avoid confusion.
-Moreover, for every 
-dead end $\mu$
- in $A_{P_0,P_1,\Gamma}$,
-we add a transition
-from the configuration $\mu \in (\Gamma^*)^P$ to
-$T_\A^\mu$.
-% See Figure~\ref{parametric_arena} for an illustration of such an arena.
-\fi
 
@@
 \problemx{Parametric pushdown reachability game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
-, and $ P = P_0 \biguplus P_1$.}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$.}
 {Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
@@
 \problemx{Parametric pushdown parity game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $%which naturally lead to a partition of $T_{\mathcal{Z}}$
-, and $ P = P_0 \biguplus P_1$, 
-and a priority function $\Omega: Q \to [0, m]$%
-%which naturally lead to a mapping on $\Conf(\mathcal{Z})$
-%, and an initial state $q_{init} \in Q$% and initial stack symbol $w_{init} \in W$
-.}
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
+and a priority function $\Omega: Q \to [0, m]$.}
 {Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
@@
 for tree-walking automata. 
-It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}%, the main trouble being that tree-walking automata get lost rather easily
-.
+It is known that tree-walking automata do not recognize all regular tree languages \cite{bojanczyk2008tree}.
 Using pebbles is a remedy against getting lost along a tree, but 
@@
 Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline.
-It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
+It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available)
 and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.
@@
 In this section, we extend these ideas to pushdown systems.
-Instead of using pebbles as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available) % and returns to the beginning of the tape,
+Instead of using pebbles as markings on their input, pebble pushdown systems have the ability to lift or drop pebbles on their universe of stack contents; a \textit{drop} simply records the current stack content with a fresh pebble (such a pebble should be available)
 and a \textit{lift} pops the last dropped pebble 
 while requiring that
-% assuming
  it was placed on the current node. 
-% Thus hierarchicality on the registers works slightly differently from hierarchicality on the pebbles, hierarchicality on the registers being a syntactic property, whereas hierarchicality on the pebbles is a semantical one.
 One can think of a pebble as a register that can store a stack content for later comparisons.
@@
 
-Recall that for a partial function $ f : S \rightharpoonup X $% (see page~\pageref{partial})
-, for notational purposes, we consider some element $\bot_X \not\in X$ and 
+Recall that for a partial function $ f : S \rightharpoonup X $, for notational purposes, we consider some element $\bot_X \not\in X$ and 
  associate $f$ with the function returning the 
@@
 in
-		\iffalse
-		for the parity condition on
-		$\mathcal{G}_{(\ppda, Q'_0, Q'_1, \Omega')}$.
-		\fi
 the parity game
@@
 player $0$ has a 
-		\iffalse
-		strategy $\sigma_0$ for the parameter valuation game  such that
-		for all player $1$ strategies $\sigma_1$ for the parameter valuation game,
-		player $0$ have a
-		winning strategy from $q_{init}(\gamma_{init})$ for the parity condition on
-		$\mathcal{G}_{(\mathcal{Z}, \mu_{\sigma_0, \sigma_1}, P_0, P_1, Q_0 , Q_1, \Omega)}$. 
-		\fi
 winning strategy from $\mu_\bot$ in 
@@
 $n$ as systems working with a stack of level $(n-1)$ stacks.
-% HPDS have been extensively studied as language recognizers [Dam82,Eng83].
 
@@
 A {\em level $1$ stack} (or {\em $1$-stack}) over an alphabet $\Gamma$ is 
-simply a %word in $\Gamma^*$
+simply a 
 	stack over $\Gamma$, i.e. a word  in $\Gamma^*$.
@@
 there exists some rule
-%$(q,\gamma,q',\theta) \in R$
 $(q,\gamma,q',op) \in R$
@@
 is a higher-order pushdown automaton of level $n$ where in
-% Instrn
 $\Op_n(\Gamma)$
  the operation 
- %popk
  $\text{\textit{pop}}_j$  
   is replaced by 
-%  pop=
-%k 
 $\text{\textit{push}}_j^=$
-%for 1 ≤ k ≤ n. 
 for $2 \leq j \leq n$.
@@
 this new set of operations by 
-% Instr= n
 $\Op_n^=(\Gamma)$. 
@@
 %
-%
 Transition systems induced by higher-order pushdown systems with equality pop of level $n$
@@
 exists an $n$-HPDS$^=$ (resp. $n$-HPDS) $\H'$ such that
-$\H$ and $\H'$ %generate the same graph.
+$\H$ and $\H'$ 
 induce isomorphic transition systems.
@@
 %
-%Hence, u
 Unlike for parametric extensions, 
@@
 solving pebble one-counter reachability games is nonelementary is worth investigating.
-% It may very well be the case that the hierarchy collapses instead, at level e.g. 2 or 3.
 
@@
 one can 
-% ask whether or not the parametric pushdown automata model described in this section  admits effective preservation of regularity.
 study the precise complexity of the
@@
 We are furthermore convinced that it is worthwhile to investigate further comparisons between parametric pushdown automata, pebble pushdown automata, 
-%pebble alternating tree automata 
 and register pushdown automata, especially since the latter have been shown to only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers.

### 2023-07-04 reachability/article.tex
Too long, 15 pages have to shorten to 12

@@
 
-We first consider reachability games: 
+We first consider {\em reachability games}: 
 the winning condition $\win_F(S)$ is given by a set of final configurations
 $F \subseteq S$, and $\win_F(S)$ is the set of maximal plays in $S^* F S^\infty$. 
-We simply write
-$\win_F$ in case the set $S$ is obvious from context.
-To indicate that the winning condition of a game is a reachability winning condition, we will speak of {\em reachability games}.
-
-With regards to reachability games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem.
-
-\problemx{Reachability game}
-{A reachability game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_F)$, an initial configuration $s \in S_0 \cup S_1$.}
-{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$? \\}
-
-Secondly, we consider min-parity games: 
+Secondly, we consider {\em parity games}: 
 the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
@@
 We simply write
-$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
-To indicate that the winning condition of a game is a parity winning condition, we will speak of
- {\em parity games}.
-
-With regards to parity games, we are going to concern ourselves with games over potentially infinitely large arenas, and we are interested in the following decision problem.
-
-\begin{samepage}
-\problemx{Parity game}
-{A parity game $\mathcal{G}= ((S_0, S_1,{\rightarrow}),\win_\Omega)$, an initial configuration $s \in S_0 \cup S_1$.}
-{Does player $0$ have a winning strategy from $s$ in $\mathcal{G}$?}
-\end{samepage}
+$\win_F$  and $\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
 
@@
 For simplicity's sake, the values are assigned to the parameters one by one, alternating between values assigned by player $0$ and by player $1$.
-
-A winning strategy for player $0$ allows to synthesize a controller that restricts the environment and ensures that {the property expressed by the winning condition always holds.}
 
@@
 
-\subsection{From parametric pushdown 
-systems 
-to pebble pushdown systems
-}
-
-We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
-
-\begin{theorem}\label{PPDA reduction}
+Note
 {\sc $n$-parametric pushdown parity game} 
@@
 {\sc $n$-pebble pushdown parity game}.
-\end{theorem}
-
-\begin{proof}[Sketch]
-
-Let us fix some
-$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
-P, q_{init}, { \gamma_{init}} , F)$,
- disjoint
-union $Q = Q_0  \biguplus Q_1$
-and $P = P_0 \biguplus P_1$,
-and
-some priority function 
-$\Omega : Q \times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^R \to [0, m ]$.
-
-We construct
-  an 
- $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
-a 
- disjoint
-union $Q' = Q_0'  \biguplus Q_1'$
-and a mapping $\Omega' : Q \times  \Gamma'^* \times ( \Gamma'^* \biguplus \{ \bot \})^{\{1, \ldots, n\}} \to \{1, \ldots, m \}$,
-such that
-player $0$ has a winning strategy from 
-$(q'_{init},\gamma_{init}, \mu_{init})$
-in
-the parity game
-$\mathcal{G'}= (A_{(\ppda, Q'_0, Q'_1)}, \win_{\overline{\Omega'}})$
-if and only if
-player $0$ has a 
-winning strategy from $\mu_\bot$ in 
-the parity game
-$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
-where		 
-$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
-and
-$\widehat{\Omega}$
-correspond to the definitions on page~\pageref{PPDA reachability game}.
-
-The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
-(one for every 
-parameter assignment)
- of the pushdown automaton with $n$ parameters by using pebbles to represent the parameters. 
-The $n$-pebble pushdown automaton will first simulate the 
-parameter valuation arena.
-Players take turns chosing particular stack contents on which to place pebbles. 
-Once every parameter has been assigned with the dropping of a corresponding pebble, 
-the pebble assignment $\mu$ is fixed.
-
-Further configurations in the $n$-pebble pushdown automaton then consist of a word representing the status of the stack, a state, and a fixed pebble assignment.
-For a fixed pebble assignment there is then a one for one correspondance between configurations of the $n$-parametric pushdown
-automaton 
-and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
-the set of states of the $n$-parametric pushdown automaton, and 
-so are player $0$ states, player $1$ states and the priority mapping.
-\end{proof}
+See appendix for more details.
 
@@
 
-For a finite alphabet $\Sigma$,  let $\tau (\Sigma )$ be the vocabulary consisting of a binary relational symbol $\leq$,  and a unary relational symbol $P_a$ for every $a \in \Sigma$. 
-A word $w \in \Sigma^*$ can be seen as a $\tau (\Sigma )$-structure where the set of elements consists of a finite interval of $\N$.
-Elements are seen as positions in the word and for every element $i$, there exists precisely one $a \in \Sigma$ such that $i$ is in the interpretation of $P_a$.
-Lastly $\leq$ is interpreted as the linear order of $\N$.\\
-
-It is well-known that if we are interested in the complexity of first-order or monadic
-second-order model-checking on words, the alphabet can be assumed to be $\{0,1\}$ without loss of generality.
-Thus the satisfiability problem is concerned with the following question.
+The satisfiability problem is concerned with the following question.
 \begin{samepage}
@@
 {A  
-$\FO[\tau(\{0,1\})]$-sentence $\phi$.}
+$\FO$-sentence $\phi$.}
 {Does there exists a word $w\in \{0, 1\}^*$, such that $w \models \phi$ ?\newline}
@@
  a 
- $\FO[\tau(\{0,1\})]$
--formula $\phi$.
-We construct
-a parametric pushdown system $\mathcal{Z}_\phi = (Q_\phi,W_\phi, \Delta_\phi )$, with a disjoint
-union $Q_\phi = Q_0  \biguplus Q_1 $, a priority mapping $\Omega_\phi$, and an initial state $q_I \in Q_\phi$
-such that
-player $0$ has a winning strategy from $(q_I, \$, f_I)$ in the  parity game
-$\mathcal{G}_{\mathcal{Z_\phi}, Q_0,Q_1, \Omega_\phi}$
-if and only if
-there exists a word $w \in \{0,1\}^*$ such that
-$w  \models \phi $. \\
-
-\noindent
+ $\FO$-formula $\phi$.
 We assume without loss of generality that $\phi$ is written in the prenex normal form
@@
 $\psi$ is quantifier-free and in disjunctive normal form.
+We construct
+a parametric parity game
+in which
+player $0$ has a winning strategy
+if and only if
+there exists a word $w \in \{0,1\}^*$ such that
+$w  \models \phi $.
 
@@
   A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to the value of parameter $p_0$. 
- The gadgets for these intermediary potential challenges will be excluded from the illustrations to preserve clarity of representation.
 The other parameters are meant to correspond to the variables, with player $1$ parameters corresponding to universal variables and player $0$ parameters corresponding to existential variables.
@@
 satisfies $\psi$ if the variables $x_1, \ldots, x_k$ are interpreted by 
-$ |\mu(p_1)|-1, \ldots, |\mu(p_k)|-1$ respectively. This is the goal of the reachability game.
-
-The first step of the game is to check that the values assigned to the parameters correspond indeed to possible variables, that is, that the values of the parameters are prefixes of $w$.
-
+$ |\mu(p_1)|-1, \ldots, |\mu(p_k)|-1$ respectively. 
 Recall $\psi$ is quantifier-free and in disjunctive normal form, i.e.
- \[ \phi = \forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).\]
-
+ $ \psi = %\forall x_1 \exists x_2 \forall x_3 \ldots \exists x_k 
+ \bigvee_{i=1}^l(\bigwedge_{j=1}^{h_i} \psi_{i,j}(x_1, \ldots, x_k)).$
 Player $0$, being the existential player, choses one of the conjunctive clause in $\psi$, essentially claiming to be able to prove it.
@@
 $ |f(1)|-1, \ldots, |f(k)|-1$ respectively.
-
-Now we need only to describe the gadgets $ \C_{\psi_{i,j}}$. 
-The gadget depends on the type of the atomic formula.
+The gadget will depend on the type of the atomic formula.
 For a formula checking the letter at the position of a variable $x$, player $0$ goes to the position corresponding to $x$, then checks that the top of the stack
@@
 
+\subsection{From parametric pushdown 
+systems 
+to pebble pushdown systems
+}
+
+We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
+
+\begin{theorem}\label{PPDA reduction}
+{\sc $n$-parametric pushdown parity game} 
+is polynomial time reducible to
+{\sc $n$-pebble pushdown parity game}.
+\end{theorem}
+
+\begin{proof}[Sketch]
+
+Let us fix some
+$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
+P, q_{init}, { \gamma_{init}} , F)$,
+ disjoint
+union $Q = Q_0  \biguplus Q_1$
+and $P = P_0 \biguplus P_1$,
+and
+some priority function 
+$\Omega : Q \times \Gamma^* \times ( \Gamma^* \biguplus \{ \bot \})^R \to [0, m ]$.
+
+We construct
+  an 
+ $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
+a 
+ disjoint
+union $Q' = Q_0'  \biguplus Q_1'$
+and a mapping $\Omega' : Q \times  \Gamma'^* \times ( \Gamma'^* \biguplus \{ \bot \})^{\{1, \ldots, n\}} \to \{1, \ldots, m \}$,
+such that
+player $0$ has a winning strategy from 
+$(q'_{init},\gamma_{init}, \mu_{init})$
+in
+the parity game
+$\mathcal{G'}= (A_{(\ppda, Q'_0, Q'_1)}, \win_{\overline{\Omega'}})$
+if and only if
+player $0$ has a 
+winning strategy from $\mu_\bot$ in 
+the parity game
+$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
+where		 
+$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
+and
+$\widehat{\Omega}$
+correspond to the definitions on page~\pageref{PPDA reachability game}.
+
+The transition graph of the $n$-pebble pushdown automaton $\ppda$ will simulate the possible transitions graphs 
+(one for every 
+parameter assignment)
+ of the pushdown automaton with $n$ parameters by using pebbles to represent the parameters. 
+The $n$-pebble pushdown automaton will first simulate the 
+parameter valuation arena.
+Players take turns chosing particular stack contents on which to place pebbles. 
+Once every parameter has been assigned with the dropping of a corresponding pebble, 
+the pebble assignment $\mu$ is fixed.
+
+Further configurations in the $n$-pebble pushdown automaton then consist of a word representing the status of the stack, a state, and a fixed pebble assignment.
+For a fixed pebble assignment there is then a one for one correspondance between configurations of the $n$-parametric pushdown
+automaton 
+and these of the $n$-pebble pushdown automaton. The set of states of the $n$-pebble pushdown automaton apart from the initial gadget is the same as 
+the set of states of the $n$-parametric pushdown automaton, and 
+so are player $0$ states, player $1$ states and the priority mapping.
+\end{proof}
+
 \end{document}

### 2023-07-04 reachability/article.tex
Too long, 13 pages and something, have to shorten to 12

@@
 Our main result consists in
-proving parametric pushdown reachability games
-and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
+proving parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
 $n$ is fixed,
-and providing a nonelementary lower bound for parametric pushdown reachability games in general.
+and providing a nonelementary lower bound for parametric pushdown parity games in general.
 
@@
 from \cite{Sto74}, to the
-problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.
+problem of deciding whether a player has a winning strategy for parametric pushdown parity games.
 
@@
 
-We first consider {\em reachability games}: 
-the winning condition $\win_F(S)$ is given by a set of final configurations
-$F \subseteq S$, and $\win_F(S)$ is the set of maximal plays in $S^* F S^\infty$. 
-Secondly, we consider {\em parity games}: 
+We consider {\em parity games}: 
 the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
@@
 We simply write
-$\win_F$  and $\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
+$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
 
@@
 
-In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
-we prefer to write $q(w)\rightarrow_{\mathcal{Z}, u}q'(w')$
-to denote $q(w)\rightarrow_{\mathcal{Z}, \mu}q'(w')$. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\rightarrow_{\mu}$ (resp. $\rightarrow_{u}$)
-instead of $\rightarrow_{\mathcal{Z}, \mu}$ (resp. $\rightarrow_{\mathcal{Z},u}$). 
-Since we are concerned with the graph of configurations and not the language recognized, we omit the input alphabet from the model and will use the name pushdown system rather than pushdown automaton.
-
-\subsection{PPDS games}
-
-We are interested in this article in parametric variants of reachability games and parity games.
-One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a 
-reachability game or a parity game
+\textbf{PPDS games.}
+We are interested in this article in parametric variants of parity games.
+One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a parity game
 played on the arena induced by the
@@
 
-\problemx{Parametric pushdown reachability game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$.}
-{Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
-$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{F \times \Gamma^* \times (\Gamma^*)^P})$ ?\newline}
-
 \problemx{Parametric pushdown parity game}
@@
 
-We write {\sc $n$-parametric reachability parity game}
-and
+We write 
 {\sc $n$-parametric pushdown parity game} if 
  the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
-Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.
 
@@
 
-Note
+\begin{theorem}\label{PPDA reduction}
 {\sc $n$-parametric pushdown parity game} 
@@
 {\sc $n$-pebble pushdown parity game}.
-See appendix for more details.
+\end{theorem}
 
@@
 In particular we believe that the question of whether or not
-solving pebble one-counter reachability games is nonelementary is worth investigating.
-
-Of note is that reachability games and parity games are 
+solving pebble one-counter parity games is nonelementary is worth investigating.
+
+Of note is that parity games are 
 not the only problems one can consider, as one can
-also explore the complexity of reachability itself. 
+also explore the complexity of reachability. 
 In particular
@@
 {\sc $n$-PPDA reachability} problem.
-Since reachability can be viewed as a reachability game, we know that
+Since reachability can be viewed as a reachability game, which can be simulated using a parity game, we know that
 {\sc $n$-PPDA reachability} is decidable in $(n+1)$-$\EXP$, 

### 2023-07-04 reachability/article.tex
removed example since ppl know what a parity game looks like in general it's not like i'm inventing a huge new concept or anything

@@
 $\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.
-
-\begin{center}
-	\begin{figure}
-			\hspace{3.95cm}
-\includegraphics[width=0.35\textwidth]{figures/parity_game}
-	\caption{An example of an arena.
-Circles  belong to player~$0$, whereas diamonds belong to player~$1$.
-Arrows represent transitions from a configuration to the next.
-The priority assigned by the priority mapping to each configuration is a number in $\{0,1,2,3\}$, next to the configuration it is assigned to.
-Note that we do not have any dead end in our example.}
-	\label{example_parity_game}
-	\end{figure}
-\end{center}
-
-\begin{example}
-Let us consider the arena presented in  Figure~\ref{example_parity_game}. 
-Configurations are partitioned into circles belonging to player $0$ and diamonds
-belonging to player $1$. 
-The priorities are $\{ 0,1,2,3\}$.
-Plays $\pi_1 = q_0 q_1 q_2 (q_4 q_7)^\omega$ and 
-$\pi_2 = (q_0 q_1 q_2 q_4)^\omega$ are both winning for player $0$. Indeed, in $\pi_1$, 
-the minimal priority encountered infinitely often is   
-				$2$, 
-whereas in $\pi_2$, it is   
-			$0$. 
-In both cases, it is an even number. On the other hand,
-the plays $\pi_3 = (q_5 q_3 q_6)^\omega$ and $\pi_4=(q_5 q_6)^\omega$ are both winning for player $1$: the minimal priority encountered is odd, since it is $1$ for both.
-One can take matters further: on this arena, player $0$ has a winning strategy for the parity game from $q_0$, $q_1$, $q_2$, $q_4$ and $q_7$, but not from $q_5$, $q_3$ nor $q_6$, where player $1$ has a winning strategy.
-\end{example}
 

### 2023-07-04 reachability/article.tex
Satisfy the 12 pages limit now

@@
 of the set of maximal plays for $A$. A {\em game} is a pair $\mathcal{G}= (A, \win)$ which is composed of an arena $A$ and a winning condition $\win$.
-Although a winning condition is in general an infinite object we are interested in winning conditions 
-that can finitely be described.
-A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
- from $s$} for player $0$ if 
-every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
-On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
- from $s$} for player $1$ if 
-every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.
-
-We consider {\em parity games}: 
+Although a winning condition is in general an infinite object we are interested in parity games winning conditions, 
+which can finitely be described. In 
+{\em parity games},
 the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
@@
 
+A strategy $\sigma_0$ for player $0$ from position $s \in S$ is a {\em winning strategy
+ from $s$} for player $0$ if 
+every maximal play starting from $s$ and consistent with $\sigma_0$ is in $\win$.
+On the other hand, a strategy $\sigma_1$ for player $1$ from position $s \in S$ is a {\em winning strategy
+ from $s$} for player $1$ if 
+every maximal play starting from $s$ and consistent with $\sigma_1$ is not in $\win$.
+
 One important property of parity games is that of {\em memoryless determinacy}: for a parity game
@@
 A {\em parametric pushdown system} ({\em PPDS} for short) 
-is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}, F),$ where
+is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}),$ where
  $Q$ is a non-empty finite {\em set of  states};
@@
  $q_{init}\in Q$ is an {\em initial  state};
- $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
- $F\subseteq Q$ is a {\em set of final  states}.
+ $ \gamma_{init} \in \Gamma$ is an {\em initial stack symbol}.
 
@@
 \begin{samepage}
-A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init},F)$ 
+A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init})$ 
 and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
@@
 
-We introduce here arenas over the set of partial parameter valuations for a set of parameters
-$P = P_0 \biguplus P_1$.
+For an  $n$-parametric pushdown automaton 
+$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init})$,
+and a partition
+$P = P_0 \biguplus P_1$,
+we introduce first arenas over the set of partial parameter valuations.
 The values are assigned to the parameters one by one, alternating between values assigned by player $0$ (corresponding to parameters in $P_0$) and by player $1$ (corresponding to parameters in $P_1$).
@@
 Assume an ordering of the parameters $P_0 \biguplus P_1 = \{p_0, p_1, \ldots, p_k \}$ for some 
- $k \in \N$.
-%
-Then given a alphabet $\Gamma$, a {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\Gamma} = (S_0, S_1, \rightarrow_P)$ where
+ $k \in \N$, the {\em parameter valuation arena} is a tuple $A_{P_0,P_1,\Gamma} = (S_0, S_1, \rightarrow_P)$ where
 \begin{itemize}
@@
 
-Note that the arena has dead ends, and that these are the configurations that correspond to parameter valuations.
+The arena has dead ends which correspond to parameter valuations.
 Given a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1 $ for player $1$,
@@
  $\mu_\bot$ is the partial parameter valuation with domain $\emptyset$. 
-We then use parametric valuation arenas as prefixes in parametric games: once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
-%
-%
-Indeed, where a concrete automaton induces a transition system, a parametric automaton induces one transition system for every possible parameter valuation.
-We use $T^\mu_\A$ to denote the transition system induced by the concrete automaton corresponding to
-a parametric automaton $\A$ with parameters $P$ taking values in $\Gamma^*$ and a parameter valuation $\mu: P \to \Gamma^*$.
-
-More formally, for an  $n$-parametric pushdown automaton 
-$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init},F)$,
-given a partition of
-$Q$ into $Q_0$ and $Q_1$, 
-we naturally partition the configurations of
- $\mathcal{Z}$
-into
-$\Conf_{\mathcal{Z},0}=Q_0\times \Gamma^*$
-and
-$\Conf_{\mathcal{Z},1}=Q_1\times \Gamma^*$. 
-With these notations in mind we define the arena
+The parametric valuation arena is a prefix of the full parametric games: once the resulting parameter valuation is produced by the two players, it serves to instantiate the transition system for continuing the game.
+%
+%
+
+We then obtain the full parametric game's arena $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$ by, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
+ adding a transition
+from $\mu$ towards the initial configuration
+$(q_{init}(\gamma_{init}), \mu)$ 
+of
+the arena
 $
@@
 =
-(\Conf_{\mathcal{Z},0}, \Conf_{\mathcal{Z},1}, \rightarrow_{\mathcal{Z},\mu})$
-induced by $\mathcal{Z}$, the partition $Q = Q_0 \biguplus Q_1$, and a
-parameter valuation $\mu: P\rightarrow \Gamma^*$.
-From a partition of
-$P$ into $P_0$ and $P_1$, 
-we thus define 
-$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
-as
-the arena that contains both the configurations and transitions of
-the parameter valuation arena
-$A_{P_0,P_1,\Gamma}$,
-and that moreover contains the configurations and transitions of
-$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$ 
-\---- albeit with configurations additionally including $\mu$ in the tuple to avoid confusion \----
-for
-all parameter valuations $\mu: P\rightarrow \Gamma^*$.
-Moreover, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
-we add a transition
-from $\mu$ towards 
-$q_{init}(\gamma_{init}, \mu)$ in $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1, \mu)}$. \label{PPDA reachability game}
+(Q_0\times \Gamma^*, Q_1\times \Gamma^*, \rightarrow_{\mathcal{Z},\mu})$
+with configurations additionally including $\mu$ in the tuple.
+
+ \label{PPDA reachability game}
 
@@
 $\widehat{\Omega}(\mu) = 0$
-for all $\mu \in (\Gamma^* \cup \bot)^{P}$.
+for all $\mu \in (\Gamma^* \cup \bot)^{P}$. We consider the following decision problem.
 
 \problemx{Parametric pushdown parity game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
+{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} )$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
 and a priority function $\Omega: Q \to [0, m]$.}
@@
 regular tree languages, in fact to all tree languages in $\NSPACE(log ~ n)$.
-Thus, in  both pebble word automata and pebble tree-walking automata, the placement of the pebble follows a strict stack discipline.
-It is traditional hence to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available)
-and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.
+It is traditional since pebbles follows a strict stack discipline to represent syntactically the dropping and lifting of pebbles by operations \textit{lift} and \textit{drop}; a \textit{drop} simply records the current position with a fresh pebble (such a pebble should be available)  and a \textit{lift} pops the last dropped pebble if the current position corresponds to the one recorded by it.
 
@@
 
-We first define more formally our pebble pushdown automaton framework, and then provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
-
-\subsection{Pebble pushdown systems}
+\paragraph*{Pebble pushdown systems}
 
@@
 is a tuple 
-$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init}, F)$
+$\ppda = (Q, \Gamma,  R, q_{init}, \gamma_{init})$
 where $Q$ is a {\em finite set of  states}; $\Gamma$ is a {\em finite stack alphabet};
@@
  $q_{init} \in Q$ is an {\em initial  state};
- $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}; and
- $F \subseteq Q$ is a {\em set of final  states}.
+ $\gamma_{init} \in \Gamma$ is an {\em initial stack symbol}.
 
@@
 This is enforced by syntactically requiring
-the last pebble dropped $i$ is in the set $S$ used for testing the presence of certain pebbles.
+the last pebble dropped $i$ is in the set $S$ used for testing the presence of  pebbles.
 
@@
 
-\begin{samepage}
-An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$ induces 
+An $n$-pebble pushdown automaton $\ppda =  (Q, \Gamma,  R, q_{init}, {\gamma_{init}} )$ induces 
 a transition system  $T_{\ppda} = (\Conf(\ppda), \rightarrow_{\ppda})$ 
@@
 the following holds
+	%\begin{samepage}
 \begin{itemize}
@@
 \end{itemize}
-\end{samepage}
-
-\par\noindent\ignorespacesafterend
-We are interested in games over pebble pushdown systems, 
-mainly, parity games.
+	%\end{samepage}
 
@@
 \problemx{$n$-pebble pushdown parity game}
-{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
+{An $n$-pebble pushdown automaton $\ppda= (Q, \Gamma,  R, q_{init}, {\gamma_{init}})$, 
 where $Q = Q_0  \biguplus Q_1 $, 
@@
 {Does player $0$ have a winning strategy from $q_{init}(\gamma_{init}, \mu_{init})$ for the parity game 
-$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ?}
+$\mathcal{G}= (A_{(\ppda, Q_0 , Q_1)}, \win_{\overline{\Omega}})$ ? \\}
+
+Since pebbles can store stack contents, they can be used to simulate parameters. From this we deduce the following theorem. See Appendix~\ref{appendix A} for more details on the construction.
 
@@
 
-\subsection{Higher-order pushdown systems}
+\paragraph*{Higher-order pushdown systems}
 
@@
 \end{eqnarray*}
-The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.  \\
+The operations $pop_1$ and $top$ are undefined on a stack whose top $1$-stack is empty.
 
@@
 A {\em higher-order pushdown systems of level $n$} (or $n$-HPDS for short) 
-is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init}, F)$,
+is a tuple $\H=(Q, \Gamma, R, q_{init}, \gamma_{init})$,
 where 
-\begin{itemize}
-        \item $Q$ is a non-empty finite {\em set of  states},
-        \item $\Gamma$ is a non-empty finite {\em  stack alphabet},
-	\item $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules},
-        \item $q_{init}$ is the {\em initial  state},
-        \item $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
-initial stack symbol}, and
-	\item $F\subseteq Q$ is a {\em set of final  states}.
-		\end{itemize}
+ $Q$ is a non-empty finite {\em set of  states} ;
+         $\Gamma$ is a non-empty finite {\em  stack alphabet};
+	 $R\subseteq Q\times \Gamma \times Q \times \Op_n(\Gamma)$ is a finite {\em set of  rules};
+         $q_{init}$ is the {\em initial  state} ;
+         $\gamma_{init} \in \Gamma$ is the {\em %bottom-of-stack
+initial stack symbol}.
 \end{samepage}
 
-\par\noindent\ignorespacesafterend
 By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
@@
 As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.
-
-An $n$-HPDS $\H=(Q, \Gamma, R, q_{init}, F)$
+An $n$-HPDS $\H=(Q, \Gamma, R, q_{init})$
 induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
@@
 With these notations in mind one can define the arena
-$$
+$
 A_{(\mathcal{H}, Q_0, Q_1)}
 =
-(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$$
-\par\noindent\ignorespacesafterend
+(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$
 induced by an $n$-HPDS $\mathcal{H}$ and a partition of its set of states.
@@
 \problemx{$n$-HPDS parity game}
-{An $n$-HPDS $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}} , F)$, 
+{An $n$-HPDS $\H= (Q, \Gamma,  R, q_{init}, {\gamma_{init}})$, 
 where $Q = Q_0  \biguplus Q_1 $, 
@@
 given an $n$-pebble pushdown automaton 
-$\ppda = (Q, \Gamma, R, q_{init}, \gamma_{init}, F)$, where $Q= Q_0 \biguplus Q_1$,
+$\ppda = (Q, \Gamma, R, q_{init}, \gamma_{init})$, where $Q= Q_0 \biguplus Q_1$,
 one can compute an $(n+1)$-HPDS 
-$\H = (Q', \Gamma, R', q_{init}', \gamma_{init}', F')$  where $Q'= Q_0' \biguplus Q_1'$, 
+$\H = (Q', \Gamma, R', q_{init}', \gamma_{init}')$  where $Q'= Q_0' \biguplus Q_1'$, 
 such 
@@
 The reduction implies decidability of pebble pushdown systems parity game. Furthermore,
-by Theorem~\ref{HPDA parity game}, it implies the following complexity result.
+by Theorem~\ref{HPDA parity game} and Theorem~\ref{PPDA reduction}, it implies the following complexity result.
 
 \begin{theorem}
-{\sc $n$-pebble pushdown systems parity game} is in $(n+1)$-$\EXP$.
+{\sc $n$-pebble pushdown systems parity game} and {\sc $n$-parametric pushdown systems parity game} are in $(n+1)$-$\EXP$.
 \end{theorem}
-
-Finally, the following complexity result is due the above and Theorem~\ref{PPDA reduction}.
-
-\begin{corollary}
-{\sc $n$-parametric pushdown systems parity game} is in $(n+1)$-$\EXP$.
-\end{corollary}
 
@@
 deciding
-the winner of a parity game on a graph of a pebble pushdown system 
-is nonelementary.
-
-	For the nonelementary lower bound
-we reduced the $\FO$ satisfiability problem on words \---- known to be nonelementary
+the winner of a parametric parity game
+is nonelementary, reducing the $\FO$ satisfiability problem on words \---- known to be nonelementary
 from \cite{Sto74} \---- to the
 problem of deciding whether player $0$ has a winning strategy from some starting configuration in 
-a pebble pushdown parity game.
+a parametric pushdown parity game.
 
@@
 
-For solving parametric pushdown parity games, on the other hand, we introduced the notion of pebble pushdown parity games, and
-pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
+For solving parametric pushdown parity games we introduced the notion of pebble pushdown parity games.
+Pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
 can be seen as a form of one-counter automata extended with pebbles. 
@@
 
-\subsection{From parametric pushdown 
+\setcounter{theorem}{0}
+
+\section{From parametric pushdown 
 systems 
 to pebble pushdown systems
-}
-
-We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games.
+}\label{appendix A}
+
+We now provide a reduction from the problem of solving parametric pushdown parity games to the problem of solving pebble pushdown parity games. Recall the theorem from Section~\ref{pebbles}.
 
@@
 $n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
-P, q_{init}, { \gamma_{init}} , F)$,
+P, q_{init}, { \gamma_{init}})$,
  disjoint
@@
   an 
- $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}} , F')$,
+ $n$-pebble pushdown automaton $\ppda =(Q', \Gamma,  R', q_{init}', {\gamma_{init}})$,
 a 

### 2023-07-04 introduction.tex
section

@@
 %			from all reachable states from $s_0$, it is always possible to reach $H$ 
-$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$, we say that $H$ is an home-space for $X$ if $\post^*(X) \subseteq \pred^*(H)$.
+$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
 In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
@@
 {\bf State of the art.}
-In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} many interesting intuitions (without detailed proofs) about resilience in the framework of process algebra.
+In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} interesting results (without detailed proofs) about resilience in the framework of process algebra.
 They built the composition of the process and its adversary as a transition system and they gave conditions that insure that the composed transition system is an effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
@@
 
-In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan in 2022, \cite{DBLP:conf/gg/Ozkan22} proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
-$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, hence $\Bad$ is downward-closed
+In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
+$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, (hence $\Bad$ is downward-closed).
 %		(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases). 
 %		The paper of Prasad and Zuck \cite{DBLP:journals/corr/PrasadZ16} is not quoted by Ozkan and Würdemann.
+
 Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
-Both Prasad \& Zuck and Ozkan \& Würdemann did not make the relation between resilience and the home-space property.
+However, neither Prasad \& Zuck nor Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
 %%%%%%%%%
@@
 
-\item Surprinsingly, the general undecidability statements were not known neither proved.
+\item Surprinsingly, the general undecidability statements about resilience were not known neither proved.
 We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
@@
 %and $\Bad=\downarrow \Bad$. 
-We show that removing the effective 
+We also show that removing the effective 
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
@@
 	\caption{A lossy VASS with three subsystems and four counters.
-Leftmost and rightmost systems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $C$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.
+Leftmost and rightmost subsystems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $C$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.
 Arrows represent control-transitions.

### 2023-07-04 resilience.tex
section

@@
 
-Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
-% We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
-In particular, the existing proofs of resilience use the decidability of reachability;
-\mathieu{quelles preuves ? pour quelle notion de résilience ?} but we wish to decide resilience for models with undecidable reachability.
-Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
-and in the general case when $\Safe = \downarrow \Safe$.
+%Surprinsingly, we have not found decidability results about the resilience problems for WSTS.	
+%% We deduce decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
+%In particular, the existing proofs of resilience use the decidability of reachability;
+%\mathieu{quelles preuves ? pour quelle notion de résilience ?} but we wish to decide resilience for models with undecidable reachability.
+%Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$
+%and in the general case when $\Safe = \downarrow \Safe$.
 

### 2023-07-05 introduction.tex
section

@@
 {\bf State of the art.}
-In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} interesting results (without detailed proofs) about resilience in the framework of process algebra.
+In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} interesting definitions and results (without detailed proofs) about resilience in the framework of process algebra.
 They built the composition of the process and its adversary as a transition system and they gave conditions that insure that the composed transition system is an effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
@@
  the contents of a counter can spontaneously get lower at any time.
- Lossiness is represented by using transitions decreasing any counter by one from every state. 
+\alain{contradictoire, à enlever:  Lossiness is represented by using transitions decreasing any counter by one from every state.}
 % We ask however that the system is not too lossy.
 We add the condition that for every prefix of an execution, the number of ``lossiness transition'' for a counter is less than nine tenth of the number of transitions incrementing said counter.
-In this example, the safe, desirable state would be one where the aircrafts have both landed, i.e.
+In this example, the safe states would be ones where the aircrafts have both landed, i.e.
 $\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$, and $\Bad$ is the complement of $\Safe$. 
 % We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.
-Resilience here expresses that the aircrafts will both land, while $k$-resilience asks whether they will do so in at most $k$ steps. 
-$k$-resilience is of interest since in general we want to ensure the safe space will 
-be reached in
-the least amount of steps possible. 
-In this exemple, for instance, aircrafts cannot simply wait in the air an infinite amount of time. Fortunately, $8$-resilience hold in the example.
+Resilience here expresses that the aircrafts will both land, while $k$-resilience is the property insuring that they will do so in at most $k$ steps. 
+%$k$-resilience is of interest since in general we want to ensure the safe space will 
+%be reached in
+%the least amount of steps possible. 
+In this exemple, aircrafts cannot simply wait in the air an unbounded amount of time: here, $8$-resilience hold.\alain{expliquer}
 \end{example}

### 2023-07-05 reachability/article.tex
reachability

@@
 \renewcommand{\R}{\mathbb{R}}
-\renewcommand{\H}{\mathcal{H}}
+\renewcommand{\H}{\mathcal{C}}
 \newcommand{\Z}{\mathbb{Z}}
@@
 A {\em parametric pushdown system} ({\em PPDS} for short) 
-is a tuple $\mathcal{Z} = (Q, \Gamma, P, R, q_{init},\gamma_{init}),$ where
+is a tuple $\mathcal{A} = (Q, \Gamma, P, R, q_{init},\gamma_{init}),$ where
  $Q$ is a non-empty finite {\em set of  states};
@@
 
-The {\em size} of $\mathcal{Z}$ is defined as
-$|\mathcal{Z}| = |Q| + |\Gamma| + |P| + |R| $.
-We also refer to $\mathcal{Z}$ as an $n$-parametric pushdown automaton if $|P| = n$.
+The {\em size} of $\mathcal{A}$ is defined as
+$|\mathcal{A}| = |Q| + |\Gamma| + |P| + |R| $.
+We also refer to $\mathcal{A}$ as an $n$-parametric pushdown automaton if $|P| = n$.
 A {\em stack content} is a word from $ \Gamma^*$. 
@@
 of the stack at the right of the word. 
-By $\Conf(\mathcal{Z})=Q\times  \Gamma^*$ we denote the set of
-{\em configurations} of $\mathcal{Z}$. As usual we rather write $q(w)$ instead of $(q, w)$.
+By $\Conf(\mathcal{A})=Q\times  \Gamma^*$ we denote the set of
+{\em configurations} of $\mathcal{A}$. As usual we rather write $q(w)$ instead of $(q, w)$.
 A {\em parameter valuation} is a function $\mu$ from $P$ to $\Gamma^*$.
@@
 \begin{samepage}
-A parametric pushdown automaton  $\mathcal{Z}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init})$ 
+A parametric pushdown automaton  $\mathcal{A}= (Q, \Gamma, P,u R, q_{init}, \gamma_{init})$ 
 and a parameter valuation $\mu: P\rightarrow \Gamma^*$ induce the 
 transition system  
-$T_{\mathcal{Z}}^{\mu} = (\Conf(\mathcal{Z}), 
-\rightarrow_{\mathcal{Z},\mu})$ where
+$T_{\mathcal{A}}^{\mu} = (\Conf(\mathcal{A}), 
+\rightarrow_{\mathcal{A},\mu})$ where
 for all $ q, q'  \in  Q$, 
@@
 for all $ a  \in  \Gamma$, 
-$q(wa) \rightarrow_{\mathcal{Z},\mu} q'(w')$
+$q(wa) \rightarrow_{\mathcal{A},\mu} q'(w')$
 if there exists a rule in $R$
@@
 For an  $n$-parametric pushdown automaton 
-$\mathcal{Z}= (Q, \Gamma, P, R, q_{init}, \gamma_{init})$,
+$\mathcal{A}= (Q, \Gamma, P, R, q_{init}, \gamma_{init})$,
 and a partition
@@
 
-\item $S_0$ is the set of all partial parameter valuations with images in $\Gamma^*$ and domain $\{p_0, p_1, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,
+\item $S_0$ is the set of all partial parameter valuations with images in $\Gamma^*$ and domain $\{p_0, \ldots, p_{j}\} \setminus \{p_j\}$ where $\p_j \in P_0$,
 
@@
 
-We then obtain the full parametric game's arena $A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$ by, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
+We then obtain the full parametric game's arena $A_{(\mathcal{A}, P_0, P_1, Q_0 , Q_1)}$ by, for every configuration $\mu$ in $A_{P_0,P_1,\Gamma^*}$,
  adding a transition
@@
 $
-A_{(\mathcal{Z}, Q_0, Q_1, \mu)}
+A_{(\mathcal{A}, Q_0, Q_1, \mu)}
 =
-(Q_0\times \Gamma^*, Q_1\times \Gamma^*, \rightarrow_{\mathcal{Z},\mu})$
+(Q_0\times \Gamma^*, Q_1\times \Gamma^*, \rightarrow_{\mathcal{A},\mu})$
 with configurations additionally including $\mu$ in the tuple.
@@
 into
-$\widehat{\Omega}: \Conf(\mathcal{Z}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$,
+$\widehat{\Omega}: \Conf(\mathcal{A}) \cup (\Gamma^* \cup \bot)^{[0,n]} \to [0, m]$,
 where we set
@@
 \problemx{Parametric pushdown parity game}
-{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} )$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
+{A parametric pushdown automaton $\mathcal{A}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init}, \gamma_{init} )$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$, 
 and a priority function $\Omega: Q \to [0, m]$.}
 {Does player $0$ have a winning strategy from $\mu_\bot$ in the parity game \\
-$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}
+$\mathcal{G} =(A_{(\mathcal{A}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$ ?\newline}
 
@@
 {\sc $n$-parametric pushdown parity game} if 
- the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
+ the number of parameters $n=|P|$ of $\mathcal{A}$ is fixed by the problem.
 
@@
 
-\newcommand{\ppda}{\mathcal{I}}
+\newcommand{\ppda}{\mathcal{B}}
 
@@
 \begin{samepage}
-\par\noindent\ignorespacesafterend
+%\par\noindent\ignorespacesafterend
 A {\em higher-order pushdown systems of level $n$} (or $n$-HPDS for short) 
@@
 
-By $\Conf(\mathcal{H}) = Q \times \mathscr{S}_n(\Gamma)$
+By $\Conf(\H) = Q \times \mathscr{S}_n(\Gamma)$
 we denote the set of {\em configurations} of an $n$-HPDS $\H$. 
-As usual we abbreviate $(q,s) \in \Conf(\mathcal{H})$ as $q(s)$.
+As usual we abbreviate $(q,s) \in \Conf(\H)$ as $q(s)$.
 An $n$-HPDS $\H=(Q, \Gamma, R, q_{init})$
-induces the transition system $\T_\H = (Conf(\mathcal{H}), \rightarrow_{\mathcal{H}})$
+induces the transition system $\T_\H = (Conf(\H), \rightarrow_{\H})$
 where
@@
 for all $s,s'$ in $\mathscr{S}_n(\Gamma)$,
-$q(s) \rightarrow_{\mathcal{H}} q'(s')$ if
+$q(s) \rightarrow_{\H} q'(s')$ if
 there exists some rule
@@
 As expected, given an 
- $n$-HPDS $\mathcal{H}$
+ $n$-HPDS $\H$
  and a partition of
@@
 we partition the configurations of
-$T_{\mathcal{H}}$
+$T_{\H}$
 into
-$\Conf_{\mathcal{H},0}=Q_0\times \mathscr{S}_n(\Gamma)$
-and
-$\Conf_{\mathcal{H},1}=Q_1\times\mathscr{S}_n(\Gamma)$.
+$\Conf_{\H,0}=Q_0\times \mathscr{S}_n(\Gamma)$
+and
+$\Conf_{\H,1}=Q_1\times\mathscr{S}_n(\Gamma)$.
 %
@@
 $
-A_{(\mathcal{H}, Q_0, Q_1)}
+A_{(\H, Q_0, Q_1)}
 =
-(\Conf_{\mathcal{H},0}, \Conf_{\mathcal{H},1}, \rightarrow_{\mathcal{H}})$
-induced by an $n$-HPDS $\mathcal{H}$ and a partition of its set of states.
+(\Conf_{\H,0}, \Conf_{\H,1}, \rightarrow_{\H})$
+induced by an $n$-HPDS $\H$ and a partition of its set of states.
 
@@
 by setting
-$\Omega_{\mathscr{S}_n(\Gamma)}: \Conf(\mathcal{H}) \to [0, m]$
+$\Omega_{\mathscr{S}_n(\Gamma)}: \Conf(\H) \to [0, m]$
 and
@@
 In \cite{carayol2006automates} it is seen as a symmetrical operation in comparison to
-$\text{\textit{push}}_j$.\\
+$\text{\textit{push}}_j$.
 
@@
  a unique shortest sequence of instructions which creates $s$ from
-$\epsilon_n$. \\
+$\epsilon_n$.
 
@@
 in case the input is an $n$-HPDS$^=$ rather than an $n$-HPDS.
-
 The algorithm from \cite{Cach03} actually provides an algorithmic solution to parity games on the graphs of the Caucal hierarchy. 
@@
 it is similarly shown that
-every transition system induced by an $n$-HPDS$^=$ $\H$ is a graph of the Caucal hierarchy.
+every transition system induced by an $n$-HPDS$^=$ $\H$ is a graph of level $n$ of the Caucal hierarchy.
 %
@@
 the stack remains a prefix of $w$, with each stack content of the play corresponding to a position in the word.
- To enforce this, we allow in $\mathcal{Z}$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position in the word.
+ To enforce this, we allow in $\mathcal{A}$, after any decision, for a player to challenge the assumption that the last movement of the other player led to a position in the word.
   A player thus challenged only has the ability to add things onto the stack, and win if and only if able to find a way back to the value of parameter $p_0$. 
@@
 Let us fix some
-$n$-parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, 
+$n$-parametric pushdown automaton $\mathcal{A}= (Q, \Gamma, 
 P, q_{init}, { \gamma_{init}})$,
@@
 the parity game
-$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
+$\mathcal{G} =(A_{(\mathcal{A}, P_0, P_1, Q_0 , Q_1)}, \win_{\widehat{\Omega}})$,
 where		 
-$A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}$
+$A_{(\mathcal{A}, P_0, P_1, Q_0 , Q_1)}$
 and

### 2023-07-05 conclusion.tex

@@
 
-We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
+We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
 % The problem [] is []. 
@@
 (in the case $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$).
-We additionnally generalized decidability results from~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
+%We additionnally generalized decidability results from~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
 
@@
 % We will analyse in detail the complexities of the different resilience. resilience for other computable models like stack automata,...
-For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done. 
+For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
 One other angle of attack not studied here is synthesis of sets $\Bad$ and/or $\Safe$ for resilience,
@@
 One could also extend upon the classes of set $\Bad$ and $\Safe$ considered.
-As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets \cite{DBLP:journals/fmsd/BertrandS13}.
 Finally, while we mention VASS, but a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.

### 2023-08-04 resilience-safe-down.tex
section

@@
 
-Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is downward compatible if
+Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is \emph{downward compatible} if
 for all $s_1, s_2, t_1 \in S$ with
@@
 $\Safe=\downarrow \Safe$.
+\alain{ou
+{\sc Resilience} is decidable for ideal-effective downward-compatible WSTS with 
+$\Safe=\downarrow \Safe$.}
 \end{corollary}

### 2023-08-07 introduction.tex

@@
 \item Surprinsingly, the general undecidability statements about resilience were not known neither proved.
-We show that resilience and state-resilience problems are both undecidable for WSTS with strong compatibility.
+We show that resilience and {\sc State-resilience} problems are both undecidable for WSTS with strong compatibility.
+\alain{harminiser les notations des pbs {\sc State-resilience} ou  State-resilience ?}
 %Moreover, state-resilience,
@@
 when
-$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to state-resilience in reset-VASS.
+$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {\sc State-resilience} in reset-VASS.
 

### 2023-08-07 resilience.tex

@@
 
-We now briefly investigate the question to control a system (non-resilient for a set $Safe$) such that its controlled behavior is resilient for $Safe$.
+We now briefly investigate the question to control a system such that its controlled behavior may always reach $Safe$.
 %
 Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{Safe}$. 
+%
+%		\begin{definition}
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient if $S  \xrightarrow{*} Safe$ is satisfied.  
+%		\end{definition}
+%
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(Safe)$-resilient.
 
-\begin{definition}
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $X$-resilient if $S  \xrightarrow{*} X$ is satisfied.  
-\end{definition}
-
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(Safe)$-resilient.
 The \emph{maximal $Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{Safe}=(S,\rightarrow_{Safe},\leq)$ where $\rightarrow_{Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{Safe}$ is $Safe$-resilient.
@@
 %
-It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
+%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.

### 2023-08-08 introduction.tex
up

@@
 The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $\Bad,\Safe \subseteq S$ can be stated as an home-space problem : $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if $\post^*(\Bad) \subseteq \pred^*(\Safe)$;
+\mathieu{Pas la définition exacte,  $\Bad \subseteq \pred^*(\Safe)$ n'est pas la même chose que $\post^*(\Bad) \subseteq \pred^*(\Safe)$ sauf cas particulier comme typiquement Bad et Safe complémentaire.
+Changer la formulation.}
 similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
@@
 \alain{harminiser les notations des pbs {\sc State-resilience} ou  State-resilience ?}
+\mathieu{je pense dans l'intro on peut conserver State-resilience et par contre une fois le probleme de {\sc State-resilience} proprement défini c'est cette notation qu'il faudrait utiliser}
 %Moreover, state-resilience,

### 2023-08-08 introduction.tex
up

@@
 We assume that the VASS is
-lossy, i.e. 
- the contents of a counter can spontaneously get lower at any time.
-\alain{contradictoire, à enlever:  Lossiness is represented by using transitions decreasing any counter by one from every state.}
+lossy;  Lossiness will be represented by using transitions decreasing any counter by one from every state.
 % We ask however that the system is not too lossy.

### 2023-08-08 resilience-safe-down.tex
up

@@
 \begin{corollary}
-The resilience problem is decidable for ideal-effective downward-compatible WSTS with 
+{\sc Resilience} is decidable for ideal-effective downward-compatible WSTS with 
 $\Safe=\downarrow \Safe$.
-\alain{ou
-{\sc Resilience} is decidable for ideal-effective downward-compatible WSTS with 
-$\Safe=\downarrow \Safe$.}
 \end{corollary}

### 2023-08-08 resilience-safe-down.tex
Removing unecessary abreviation precisions

@@
 \begin{theorem}\label{up-down}
-The resilience problem is decidable for ideal-effective WSTS with 
+{\sc Resilience}  is decidable for ideal-effective WSTS with 
 $\Safe=\downarrow \Safe$

### 2023-08-08 resilience-safe-up.tex
Removing unecessary abreviation precisions

@@
  sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
-The resilience problem (RP), the bounded resilience problem (BRP)
-and the $k$-resilience problem (kRP) are decidable.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
 \end{theorem}
@@
 The %uniform 
-resilience problem (RP) can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
 Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
@@
 Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we deduce that both the 
-  resilience problem (RP) and the bounded resilience problem (BRP) are decidable.
+  resilience problem and the bounded resilience problem are decidable.
 
-\iffalse
-	\end{proof}
-
-	\begin{theorem}\label{k-down-up}
-	Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS 
-	with strong compatibility and the predbasis hypothesis, and two finite sets: $\Bad$ 
-	and $\Safe$.
-	The  $k$-resilience problem (kRP) is decidable 
-	%	for $k \geq \min(K,n)$ where $n$ satisfies $ \uparrow \pred^n(\Safe)=  \uparrow 	\pred^*(\Safe)$. 
-	\textcolor{red}{CONJECTURE for now}
-	 \end{theorem}
-
-	\begin{proof}
-
-We begin to compute $K$ and $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.
-If $K=\infty$ then $\mathscr{S}$ is not resilient for Safe and Bad.
-Now, if $k \geq min(K,n)$, we conclude that is $k$-resilient; 
-%	if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, 
-%	and then kRP is decidable.
-\fi
-
-Let us now show that the $k$-resilience problem (kRP), with $k \in \mathbb{N}$, is also decidable.
+Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
 Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
@@
 \end{proof}
-
-% \alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem (RP), the bounded resilience problem (BRP) and the $k$-resilience problem (kRP) are undecidable for WSTS with strong compatibility}
-%
 

### 2023-08-08 resilience.tex
Removing unecessary abreviation precisions

@@
 
-\problemx{resilience problem (RP)}
+\problemx{Resilience}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
@@
 
-\problemx{$k$-resilience problem (kRP)}
+\problemx{$k$-resilience problem}
 {A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
@@
 
-\problemx{bounded resilience problem (BRP)}
+\problemx{Bounded resilience}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}

### 2023-08-08 state-resilience.tex
Removing unecessary abreviation precisions

@@
 
-\problemx{State-resilience problem (SRP)}
+\problemx{State-resilience}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
@@
 
-\problemx{$k$-state-resilience problem (kSRP)}
+\problemx{$k$-state-resilience}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
@@
 
-\problemx{bounded-state-resilience problem (BSRP)}
+\problemx{bounded-state-resilience}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}

### 2023-08-08 resilience.tex
up

@@
 
-\problemx{Resilience}
+\problemx{Resilience (RP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
@@
 
-\problemx{$k$-resilience problem}
+\problemx{$k$-resilience (kRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
@@
 
-\problemx{Bounded resilience}
+\problemx{Bounded resilience (BRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}

### 2023-08-08 state-resilience.tex
up

@@
 
-\problemx{State-resilience}
+\problemx{State-resilience (SRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
@@
 
-\problemx{$k$-state-resilience}
+\problemx{$k$-state-resilience (kSRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
@@
 
-\problemx{bounded-state-resilience}
+\problemx{bounded-state-resilience (BSRP)}
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}

### 2023-08-08 application.tex
solved endproof
section

@@
-\end{proof}
 %		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
@@
 
-\section{Resilience for VASS}
+\section{Resilience for VASS and variations}
+
+In this section we continue to study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+$\Safe$ is upward-closed. Lacking downward compatibility or a more relaxed hypothesis that
+for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
+VASS do not inherit the decidability results for WSTS in the case $\Safe$ is downward-closed.
+In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.
+We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
 
@@
 Let us remark that the basis of the four upward-closed sets $X$ are computable in $2^{2^n}$, the maximal $X$-resilient VASS of $V$ is also a computable VASS $V_X$ and that $V_X$ is independant of the initial state $v_0$.
-
-In this section we continue to study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
-$\Safe$ is upward-closed. Lacking downward compatibility or a more relaxed hypothesis that
-for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
-VASS do not inherit the decidability results for WSTS in the case $\Safe$ is downward-closed.
-In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.
-We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
 
@@
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system \alain{pas défini}when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 \end{proposition}
@@
 
-This all changes when you add control states.
-For a VASS with control states in $Q$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+This all changes when one considers VASS (having control states).
+For a VASS $V = (Q,T)$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
 then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.

### 2023-08-08 introduction.tex
up

@@
 The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is ($\Bad,\Safe$)-resilient (for two given subsets of states $\Bad,\Safe \subseteq S$).
-The $k$-resilience problem (kRP) is to decide whether $\post^*(\Bad) \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
-the bounded resilience problem (BRP) is to decide whether there exists an $k$ such that $\mathscr{S}$ is ($\Bad,\Safe$)-$k$-resilient. We adapt these three definitions to state-resilience.
+The $k$-resilience problem is to decide whether $\post^*(\Bad) \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
+the bounded resilience problem is to decide whether there exists an $k$ such that $\mathscr{S}$ is ($\Bad,\Safe$)-$k$-resilient. We adapt these three definitions to state-resilience.
 

### 2023-08-14 appendix.tex
up

@@
 
-We relax for now the hypothesis that $\Bad$ is the complement of $\Safe$, thus it is possible for $\Safe$ and $\Bad$ to share closure properties, i.e.
-for them to be both downward-closed or upward-closed, and we explore both possiblities.
+We relax now the hypothesis that $\Bad$ is the complement of $\Safe$.
+In this appendix it is thus possible for $\Safe$ and $\Bad$ to share closure properties, i.e.
+for them to be both downward-closed or upward-closed.
+We will now explore both possiblities.
 
@@
 \begin{theorem}\label{down-down}
-The {\sc Resilience problem} is undecidable for  effective WSTS with  strong  compatibility such that
+{\sc Resilience} is undecidable for  effective WSTS with  strong  compatibility such that
 %	with more than one minimal element, 
@@
 \begin{theorem}\label{up-up}
-The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
+{\sc Resilience} is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
 and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
@@
 \begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
+In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the {\sc Bounded resilience} is equivalent to {\sc Resilience}.
 \end{proposition}
@@
 \begin{corollary}\label{B-up-up}
-The bounded resilience problem is decidable for WSTS with effective pred-basis,
+{\sc Bounded resilience}  is decidable for WSTS with effective pred-basis,
 strong compatibility
@@
 \begin{theorem}\label{k-up-up}
-The % three %uniform 
-$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
+{\sc $k$-resilience}  is decidable for WSTS with effective pred-basis, strong %upward-
 compatibility, $\Safe=\uparrow \Safe$
@@
 
-Most of our undecidability results make use of reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
-This has consequences for {\sc Resilience}.
-For instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element.
-Even then, though, {\sc Resilience} (resp.
-{\sc State-resilience}) is not strictly reachability, and we will need to delve more into the details to obtain undecidability results.
+Most of our undecidability results rely on reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
+This has consequences for {\sc Resilience}; for instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element.
+Even then, {\sc Resilience} (resp.
+{\sc State-resilience}) is not strictly speaking the same as reachability, and we will need to delve more into the details to obtain undecidability results.
 
@@
 
-\subsection{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}
+\subsection{Home-state is undecidable for Minsky machine, reset-VASS and WSTS}\label{HS-Minsk}
 
-We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
+We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 Let us provide a more detailed undecidability proof here.
@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
-{$S \longrightarrow^{*} \{s\}$ ?\newline}
+{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}
 
 % whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
+% post ∗ (S ) ⊆ pred ∗ (s_0)
 
@@
 % or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
-Thus {\sc Home-state} is undecidable for Minsky machines, and hence {\sc Resilience} is undecidable.
+Thus {\sc Home-state} is undecidable for Minsky machines.
+Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}.
+Hence {\sc Resilience} is undecidable for Minsky machines.
 
@@
 %following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
-Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset VASS, which consists in, given a
+Reset VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
 deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
@@
 \end{proof}
-
-\section{More Appendix thing if necessary}\label{appendix W}

### 2023-08-14 appendix.tex
up

@@
 $\Safe$ and we'd want $\Bad$ to have the same closure properties as $\Safe$.
-We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the subcase 
-$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ leads to a contradiction, since two upward-closed sets have not an empty intersection.
-We thus consider instead the case where $\Bad$ is as upward-closed
+We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the conditions 
+$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ would lead to a contradiction, since two upward-closed sets have not an empty intersection.
+We consider instead the case where $\Bad$ is as upward-closed
 as it can be without intersecting with $\Safe$,
@@
 such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
-We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
+We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
 % Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
@@
 \begin{definition}
-A {\em reset-VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
+A {\em reset-VASS} in dimension $d$ %(reset-VASS for short)
+ is a finite 
 labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
@@
 
-It is well known that Reset $d$-VASS are WSTS~\cite{dufourd1998reset}. 
+It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
 Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
@@
 %following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset VASS, which consists in, given a
-Reset VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
 deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
 %
-We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
 which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
+A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
 live. This leads to the following problem.
@@
 \problemx{$t$-liveness}
-{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
 {Is $t$ live in $s_0$ ? \\}
@@
 \begin{proposition}\label{reductions}
-{\sc $t$-liveness} is reducible to {\sc State-resilience} in Reset $d$-VASS.
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
 \end{proposition}
@@
 % WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
-Reset $d$-VASS $(Q,T)$ 
+reset-VASS $(Q,T)$ 
 with initial configuration $s_0$
@@
 
-\mathieu{À refaire avec des Reset VASS}
-
 %\begin{proposition}\label{liveness reset}
@@
 \begin{proof}
-If we wish to determine if $0 \in R(C_1, \mu_1)$ in any
-reset Petri net $C_1 = (P_1, T_1, I_1, O_1, R_1 )$ (where $R_1(t) =$ places that are reset when t fires off), with initial marking $\mu_1$, we construct a
-reset petri net $C_2 = (P_2, T_2, I_2, O_2, R_2 )$ with initial marking
-$\mu_2$ which is live if and only if the zero marking is not reachable from $\mu_1$.
+
+We wish to determine if $q(\textbf{0})$ is reachable in a reset-VASS $V_1$
+with
+initial vector $\textbf{u}$ and initial state $q$.
+To do so, we construct a reset-VASS $V_2$
+with initial vector $\textbf{v}$ and initial state $q'$
+which is live if and only if 
+$\textbf{0}$
+is reachable from $\textbf{u}$.
+
+The new reset-VASS is constructed by adding two new places
+and
+$|V_1|+2$ new transitions.
+
+We first modify all transitions to include $r_1$ both as an input and an output
+
+\mathieu{Ok non pas comme ça}
 
@@
 
-Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
+Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
 % the following:
@@
 %
-{\sc %$ Reset $d$-VASS 
+{\sc %$ reset-VASS 
 $t$-liveness} is undecidable.
 
-Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for Reset $d$-VASS, which are WSTS with strong compatibility.
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
 Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.

### 2023-08-14 appendix.tex
up

@@
 
-The new reset-VASS is constructed by adding two new places
+The new reset-VASS is constructed by adding four new vector coordinate
 and
-$|V_1|+2$ new transitions.
+$|V_1|+4$ new transitions.
 
-We first modify all transitions to include $r_1$ both as an input and an output
+We first modify all transitions to include a $-1$ on the first additional vector coordinate and a $+1$ on the second additional vector coordinate.
+We also add a transition with only $+1$ on the first additional vector coordinate and $-1$ on the second one.
 
-\mathieu{Ok non pas comme ça}
+% The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$ and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
 
-The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$
-and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
+% We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
 
-We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $r_1$.
+The initial vector $\textbf{v}$ is constructed from $\textbf{u}$ by having at least $1$ on the first two additional vector coordinate.
 
-The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
+% The initial marking $\mu_2$ will include a token in $r_1$.
 
-Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$
-and "freezing" the marking of $P_1$.
+These two vector coordinate serve to mark that the run is ongoing.
+As long as they are both non-zero, the adjusted transitions of $V_1$ are live and can be used normally.
+Thus any vector which is reachable in $V_1$ is also reachable [in $V_2$].
 
-Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
+% The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
+% Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
 
-The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
+We add an additional transition $s$ which reset the first two additional vector coordinate.
+% Transition $s$ is defined to have $r_1$ as an input and a null output.
+This allows to disable the transitions of $V_1$ and to "freeze" the vector of $V_1$.
+% This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$ and "freezing" the marking of $P_1$.
 
-Now we need to see if that marking is zero.
+% Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
 
-We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
+These first two additional vector coordinate together with this additional transition allow
+to reach aby reachable vector and then to freeze.
 
-If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if
-$s$ fires in the marking zero.
+% The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
 
-If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
+Now we need to see if that vector is zero.
+% Now we need to see if that marking is zero.
 
-In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
+We introduce two new additional vector coordinate and $d$ new transitions which have
+$-1$ on the $d$-th vector coordinate and $+1$ on this penultimate coordinate.
+
+% We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
+
+If $\textbf{0}$ is reachable, then these $d$ new transitions are not live. Everything is dead if $s$ fires in $\textbf{0}$.
+
+% If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if $s$ fires in the marking zero.
+
+If we can always have some vector coordinate be not zero, then we can always fire some transition to add $+1$ to the penultimate additional coordinate.
+% If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
+In this case we can add to the first additional coordinate again and assure that all transitions are live.
+% In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
+We must be sure that $V_2$ is live even if $V_1$ is {\em not} live.
+% We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
+% This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
 Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
+This is accomplished by a transition $s'$ which arbitrarily increments every other vector coordinate 
+while decrementing the penultimate one and, additionally, a transition which 
+substract from the last additional vector coordinate and add to the penultimate one.
 
-Now if $0$ is reachable
-in $R(C_1, \mu_1)$ 
-then the net $C_2$ can also reach this marking
-in the place of $P_1$ by executing the same sequence of transition firings.
+Now if $\textbf{0}$ is reachable from $\textbf{u}$ in $V_1$, then $V_2$ can also reach a similar
+vector $(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transition.
+Then $s$ can freeze the $V_1$ subset,
+% Now if $0$ is reachable in $R(C_1, \mu_1)$  then the net $C_2$ can also reach this marking in the place of $P_1$ by executing the same sequence of transition firings.
 Then $s$ can fire, freezing the $C_1$ subset.
-Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $0$ is reachable then $C_2$ is not live.
+ending up in $0^{d+4}$, and $V_2$ is not live.
+% Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
+Thus if $\textbf{0}$ is reachable then $V_2$ is not live.
 
-Conversey, if $C_2$ is not live, then a marking $\mu$ must be reachable in which 
-$\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token
-(in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.)
-% (Rappel: tant que il y a un token quelque part, on peut mettre un token en r_2
-% et r_2 commence sans token)
-If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero. 
+Conversely, if $V_2$ is not live, then a vector is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.
+% Conversely, if $C_2$ is not live, then a marking $\mu$ must be reachable in which  $\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token (in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.).
+If it is not possible to have (), then the transitions are not live, and the
+$d$-dimensional subset of the current vector must be zero.
+% If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
+% Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero. 
+Thus if $V_2$ is not live then 
+$\textbf{0}$ is reachable.
 \end{proof}

### 2023-08-14 appendix.tex
up

@@
 
+\mathieu{Thinking about using reset-VAS to simplify notations.}
+
 We wish to determine if $q(\textbf{0})$ is reachable in a reset-VASS $V_1$
@@
 
-\section{More Appendix thing if necessary}\label{appendix brp VASS}
+\section{Resilience for VASS}\label{appendix brp VASS}
 
-\begin{theorem*}{\textbf{\ref{brp VASS}}.}
+As mentionned in Section~\ref{section VASS}
+{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable.
+This stems from the fact that there are simple conditions in VASS that ensure neither {\sc $k$-resilience }  nor {\sc Bounded resilience} hold.
+Restricting oneself to the case where these conditions are not met simplifies the problem.
+
+\begin{theorem*}{{\em \textbf{\ref{brp VASS}}}.}
 {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$

### 2023-08-14 appendix.tex
up

@@
 
-As mentionned in Section~\ref{section VASS}
+As mentionned in Section~\ref{section VASS},
 {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable.
@@
 Restricting oneself to the case where these conditions are not met simplifies the problem.
+Let us provide a more detailed decidability proof here.
 

### 2023-08-14 application.tex
up

@@
 
-\section{Resilience for VASS and variations}
+\section{Resilience for VASS and variations}\label{section VASS}
 
@@
 \begin{theorem}{}
-Resilience is decidable for VASS when $\Safe$ and $\Bad$ are semilinear sets.
+{\sc Resilience} is decidable for VASS when $\Safe$ and $\Bad$ are semilinear sets.
 \end{theorem}
@@
 
-Resilience is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logic.
+{\sc Resilience} is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logic.
 Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
@@
 \begin{theorem}{}
-Resilience is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear.
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear.
 \end{theorem}
@@
 \begin{theorem}{}
-Resilience is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
+{\sc Resilience} is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
 \end{theorem}
@@
 \begin{theorem}\label{RP VASS}
-Resilience is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
+{\sc Resilience} is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
 the rationals with addition and order.

### 2023-08-15 appendix.tex
up

--- 2023-08-14 appendix.tex
@@
 %This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
-
-\section{More Appendix thing if necessary}\label{appendix brp VASS}
-
-\begin{theorem*}{\textbf{\ref{brp VASS}}.}
-{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
-.
-\end{theorem*}
-
-\begin{proof}
-Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
-Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
-% $\Safe \subseteq Q \times \N^d$
-$\Safe = S \setminus \Bad$
- downward-closed.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
-
-The set $\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider 
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
-
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-% $q'(\omega)$
-$\uparrow q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-Let us assume from now on 
-% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
- $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
-Because these subsets are upward-closed, we can compute a basis of
-% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
-% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
-$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
-
-We now consider the set 
-% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
-% \mathieu{Is it actually possible to check whether or not it is finite ?}
-% If it is finite then w
-We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
-If it is infinite then
-$k$-resilience does not hold 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
-$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
-
-In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
-and hence
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
-We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
-
-From $k_0$ onwards,
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is stationnary. 
-If the set is finite then we stop the procedure once we have checked for every element
-whether it is possible to reach $\Safe$ from it.
-If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
-hold
-for $k_m = \max(k_0, k_{\pi})$ with 
-$k_{\pi}$ the maximum of the length of the paths from 
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
-If it is infinite then
-for every $k \in \N$,
-$k$-resilience does not hold, 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-\end{proof}

### 2023-08-15 application.tex
up

@@
 
-\begin{theorem}\label{brp VASS}
+\begin{theorem*}{{\em \textbf{\ref{brp VASS}}}.}
 {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
 .
-\end{theorem}
+\end{theorem*}
 
-% \begin{proof}
-% \mathieu{Donner l'idée et se référencer à l' pour la preuve complête.}
-% \end{proof}
+\begin{proof}
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+% $\Safe \subseteq Q \times \N^d$
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+% $q'(\omega)$
+$\uparrow q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+% $q_{j_1}(\omega), q_{j_2}(\omega), \ldots q_{j_\ell}(\omega)$ are
+ $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+% $\pred^{\leq k}(q_{j_1}(\omega))$, $\pred^{\leq k}(q_{j_2}(\omega))$, \ldots
+% $\pred^{\leq k}(q_{j_\ell}(\omega))$.
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+% $\Bad \setminus (\pred^{\leq k}(q_{j_1}(\omega)) \cup \ldots \pred^{\leq k}(q_{j_\ell}(\omega)) )$.
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+% \mathieu{Is it actually possible to check whether or not it is finite ?}
+% If it is finite then w
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If it is infinite then
+$k$-resilience does not hold 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
+
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards,
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+\end{proof}

### 2023-08-15 appendix.tex
up

@@
 \end{proof}
-
-\section{Reductions to reset-VASS}
-
-Most of our undecidability results rely on reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
-This has consequences for {\sc Resilience}; for instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element.
-Even then, {\sc Resilience} (resp.
-{\sc State-resilience}) is not strictly speaking the same as reachability, and we will need to delve more into the details to obtain undecidability results.
-
-This section is dedicated to studying reset-VASS for showing undecidability of {\sc Resilience} and {\sc State-resilience} for WSTS in general.
-
-\subsection{Home-state is undecidable for Minsky machine, reset-VASS and WSTS}\label{HS-Minsk}
-
-We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
-Let us provide a more detailed undecidability proof here.
-
-First, let us recall the home-state problem
-
-\problemx{Home-state}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
-{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}
-
-% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
-% post ∗ (S ) ⊆ pred ∗ (s_0)
-
-Home-state is undecidable for Minsky machine with $3+$ counters.
-This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
-From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
-such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
-We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
-% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
-% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
-Thus {\sc Home-state} is undecidable for Minsky machines.
-Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}.
-Hence {\sc Resilience} is undecidable for Minsky machines.
-
-Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
-% Let us recall more formally what a reset-VASS is.
-Reset-VASS extend the basic VASS model with special “reset
-transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
-\begin{definition}
-A {\em reset-VASS} in dimension $d$ %(reset-VASS for short)
- is a finite 
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
-$T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control-transitions} of $V$,
-where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
-
-Again $Q \times \N^d$
- denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
-$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
-\begin{samepage}\begin{itemize}
-\item  $t = (q,add(\textbf{z}),q') \in T$
-% then for all $\textbf{u} \in \N^d$ such that  
-% $\textbf{u}+\textbf{z} \geq 0$
-% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
-and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-\item $t = (q,reset(\gamma),q') \in T$ 
-% then for all $\textbf{u} \in \N^d$ 
-% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
-and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize} \end{samepage}
-
-It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
-
-\subsection{Reset-VASS and state-resilience}\label{appendix}
-
-We are now interested in the 
-%following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-%
-We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
-which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
-live. This leads to the following problem.
-
-\problemx{$t$-liveness}
-{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
-{Is $t$ live in $s_0$ ? \\}
-
-To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
-$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
-
-\begin{proposition}\label{reductions}
-{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
-\end{proposition}
-
-\begin{proof}
-We reformulate $t$-liveness 
-in a 
-% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
-reset-VASS $(Q,T)$ 
-with initial configuration $s_0$
-as the following formula
-\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
-where
-$U_t=\uparrow pre(t)$ 
-is the upward closure of the preconditions to fire transition $t$.  
-The problem reduces itself to {\sc State-resilience}
-where $\Safe = U_t$.
-\end{proof}
-
-The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
-.
-\begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
-\end{proposition}
-
-% \section{Proof of Theorem~\ref{liveness reset}}\label{appendix}
-
-% Recall Theorem~\ref{liveness reset}.
-
-%\begin{proposition}\label{liveness reset}
-% The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
-%\end{proposition}
-
-Let us describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri}.
-
-\begin{proof}
-
-\mathieu{Thinking about using reset-VAS to simplify notations.}
-
-We wish to determine if $q(\textbf{0})$ is reachable in a reset-VASS $V_1$
-with
-initial vector $\textbf{u}$ and initial state $q$.
-To do so, we construct a reset-VASS $V_2$
-with initial vector $\textbf{v}$ and initial state $q'$
-which is live if and only if 
-$\textbf{0}$
-is reachable from $\textbf{u}$.
-
-The new reset-VASS is constructed by adding four new vector coordinate
-and
-$|V_1|+4$ new transitions.
-
-We first modify all transitions to include a $-1$ on the first additional vector coordinate and a $+1$ on the second additional vector coordinate.
-We also add a transition with only $+1$ on the first additional vector coordinate and $-1$ on the second one.
-
-% The reset PN $C_2$ is constructed from $C_1$ by the addition of two places $r_1$ and $r_2$ and $|P_1| +2$ transitions $s$, $\{ s_p | p \in P_1 \}$ and $s'$.
-
-% We first modify all transitions of $T_1$ to include $r_1$ as both an input and an output.
-
-The initial vector $\textbf{v}$ is constructed from $\textbf{u}$ by having at least $1$ on the first two additional vector coordinate.
-
-% The initial marking $\mu_2$ will include a token in $r_1$.
-
-These two vector coordinate serve to mark that the run is ongoing.
-As long as they are both non-zero, the adjusted transitions of $V_1$ are live and can be used normally.
-Thus any vector which is reachable in $V_1$ is also reachable [in $V_2$].
-
-% The place $r_1$ is a 'run' place; as long as the token remains in $r_1$ the transitions of $T_1$ can fire normally.
-% Thus any marking which is reachable in the places of $P_1$ is also reachable in $C_2$.
-
-We add an additional transition $s$ which reset the first two additional vector coordinate.
-% Transition $s$ is defined to have $r_1$ as an input and a null output.
-This allows to disable the transitions of $V_1$ and to "freeze" the vector of $V_1$.
-% This allows the token in $r_1$ to be removed, disabling all transitions in $T_1$ and "freezing" the marking of $P_1$.
-
-% Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
-These first two additional vector coordinate together with this additional transition allow
-to reach aby reachable vector and then to freeze.
-
-% The place $r_1$ and transition $s$ allow the net $C_1$ to reach any reachable marking and then for $s$ to fire and freeze the net at that marking.
-
-Now we need to see if that vector is zero.
-% Now we need to see if that marking is zero.
-
-We introduce two new additional vector coordinate and $d$ new transitions which have
-$-1$ on the $d$-th vector coordinate and $+1$ on this penultimate coordinate.
-
-% We introduce a new place $r_2$ and new transitions $s_p$ which have $p$ as input and $r_2$ as output.
-
-If $\textbf{0}$ is reachable, then these $d$ new transitions are not live. Everything is dead if $s$ fires in $\textbf{0}$.
-
-% If $p$ can ever become zero, $s_p$ is not live. The entire net is dead if $s$ fires in the marking zero.
-
-If we can always have some vector coordinate be not zero, then we can always fire some transition to add $+1$ to the penultimate additional coordinate.
-% If we can always have some $p_i$ not be zero, then we can always fire some $s_{p_i}$, putting a token in $r_2$.
-In this case we can add to the first additional coordinate again and assure that all transitions are live.
-% In this case we must put a token back in $r_1$ and assure that all transitions in $C_2$ are live.
-We must be sure that $V_2$ is live even if $V_1$ is {\em not} live.
-% We must be sure that $C_2$ is live even if $C_1$ is {\em not} live.
-% This is accomplished by a transition $s'$ which  "floods" the net $C_2$ with tokens, assuring that every transition is live if a token is ever put in $r_2$.
-Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-This is accomplished by a transition $s'$ which arbitrarily increments every other vector coordinate 
-while decrementing the penultimate one and, additionally, a transition which 
-substract from the last additional vector coordinate and add to the penultimate one.
-
-Now if $\textbf{0}$ is reachable from $\textbf{u}$ in $V_1$, then $V_2$ can also reach a similar
-vector $(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transition.
-Then $s$ can freeze the $V_1$ subset,
-% Now if $0$ is reachable in $R(C_1, \mu_1)$  then the net $C_2$ can also reach this marking in the place of $P_1$ by executing the same sequence of transition firings.
-Then $s$ can fire, freezing the $C_1$ subset.
-ending up in $0^{d+4}$, and $V_2$ is not live.
-% Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $\textbf{0}$ is reachable then $V_2$ is not live.
-
-Conversely, if $V_2$ is not live, then a vector is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.
-% Conversely, if $C_2$ is not live, then a marking $\mu$ must be reachable in which  $\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token (in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.).
-If it is not possible to have (), then the transitions are not live, and the
-$d$-dimensional subset of the current vector must be zero.
-% If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
-% Thus if $C_2$ is not live then a marking is reachable in which the marking of each $p$ in $P_1$ is zero. 
-Thus if $V_2$ is not live then 
-$\textbf{0}$ is reachable.
-\end{proof}
-
-Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
-% the following:
-%
-%
-{\sc %$ reset-VASS 
-$t$-liveness} is undecidable.
-
-Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
-Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
-%This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

### 2023-08-15 resilience-safe-down.tex
up

@@
 Without the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed, the problem becomes undecidable. Indeed, it is 
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed, the problem becomes undecidable.
+
+ Indeed, it is 
 undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
 This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.
-See Appendix~\ref{HS-Minsk} for a more detailed construction.
+% See Appendix~\ref{HS-Minsk} for a more detailed construction.
+
+	% 	From Appendix
+	%
+% \subsection{Home-state is undecidable for Minsky machine, reset-VASS and WSTS}\label{HS-Minsk}
+	%
+% We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+Let us provide a more detailed undecidability proof here.
+
+First, let us recall the home-state problem
+
+\problemx{Home-state}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
+{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}
+
+% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
+% post ∗ (S ) ⊆ pred ∗ (s_0)
+
+Home-state is undecidable for Minsky machine with $3+$ counters.
+This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
+From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
+such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
+% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
+% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
+Thus {\sc Home-state} is undecidable for Minsky machines.
+Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}.
+Hence {\sc Resilience} is undecidable for Minsky machines.
+
+Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
+% Let us recall more formally what a reset-VASS is.
+Reset-VASS extend the basic VASS model with special “reset
+transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
+\begin{definition}
+A {\em reset-VASS} in dimension $d$ %(reset-VASS for short)
+ is a finite 
+labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
+$T \subseteq Q \times Op \times Q$
+ will be referred to as the {\em control-transitions} of $V$,
+where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
+\end{definition}
+
+Again $Q \times \N^d$
+ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
+\begin{samepage}\begin{itemize}
+\item  $t = (q,add(\textbf{z}),q') \in T$
+% then for all $\textbf{u} \in \N^d$ such that  
+% $\textbf{u}+\textbf{z} \geq 0$
+% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
+\item $t = (q,reset(\gamma),q') \in T$ 
+% then for all $\textbf{u} \in \N^d$ 
+% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+\end{itemize} \end{samepage}
+
+It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
+Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
 

### 2023-08-15 state-resilience.tex
up

@@
  reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+We are now interested in the 
+%following
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
+which we then reduce to {\sc State-resilience}.
+A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
+live. This leads to the following problem.
+
+\problemx{$t$-liveness}
+{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{Is $t$ live in $s_0$ ? \\}
+
+To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
+$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+
+\begin{proposition}\label{reductions}
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness 
+in a 
+% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
+reset-VASS $(Q,T)$ 
+with initial configuration $s_0$
+as the following formula
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
+where
+$U_t=\uparrow pre(t)$ 
+is the upward closure of the preconditions to fire transition $t$.  
+The problem reduces itself to {\sc State-resilience}
+where $\Safe = U_t$.
+\end{proof}
+
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
+.
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
+\end{proposition}
+
+Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
+% the following:
+%
+%
+{\sc %$ reset-VASS 
+$t$-liveness} is undecidable.
+
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
+Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
+%This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
 
@@
 Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in Appendix~\ref{HS-Minsk}.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in % Appendix~\ref{HS-Minsk}.
 

### 2023-08-16 application.tex
VCMAI

@@
 %
-\iffalse \mathieu{Defining it like this makes it a LTS rather than a unlabeled TS - maybe talk about how we can 'forget' the labels to obtain an unlabeled TS ?}
-\alain{les VASS ne sont pas vraiment étiquetés, sauf si on veut le faire, car on peut dire que les VASS sont définis à partir d'un nombre fini de control-transitions  $t = (p, \textbf{z}, q)$ générant une infinité de transtions notées $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ puisque pour tout s, $p(\textbf{u+s}) \xrightarrow{t} q(\textbf{v+s})$} \fi
 When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
@@
 Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
-This lead to a decision procedure, written in detail in Appendix~\ref{appendix brp VASS}, which lead to the following decidability result:
+This lead to a decision procedure%, written in detail in Appendix~\ref{appendix brp VASS},
+ which lead to the following decidability result:
 
-\begin{theorem*}{{\em \textbf{\ref{brp VASS}}}.}
+\begin{theorem}
 {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
 .
-\end{theorem*}
+\end{theorem}
 

### 2023-08-16 introduction.tex
up

@@
 %		We now give the definition of two kinds of resilience: the (general) resilience and the state-resilience. 
-The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $\Bad,\Safe \subseteq S$ can be stated as an home-space problem : $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if $\post^*(\Bad) \subseteq \pred^*(\Safe)$;
-\mathieu{Pas la définition exacte,  $\Bad \subseteq \pred^*(\Safe)$ n'est pas la même chose que $\post^*(\Bad) \subseteq \pred^*(\Safe)$ sauf cas particulier comme typiquement Bad et Safe complémentaire.
-Changer la formulation.}
-similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if $\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
+The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $\Bad,\Safe \subseteq S$ consists of the following: $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if
+$\Bad \subseteq \pred^*(\Safe)$. In the case 
+it 
+can be stated as an home-space problem : $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if $\post^*(\Bad) \subseteq \pred^*(\Safe)$.
+Similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if  
+$\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
 We will study three decidability questions for each of type of resilience.
@@
 \item Surprinsingly, the general undecidability statements about resilience were not known neither proved.
-We show that resilience and {\sc State-resilience} problems are both undecidable for WSTS with strong compatibility.
-\alain{harminiser les notations des pbs {\sc State-resilience} ou  State-resilience ?}
-\mathieu{je pense dans l'intro on peut conserver State-resilience et par contre une fois le probleme de {\sc State-resilience} proprement défini c'est cette notation qu'il faudrait utiliser}
+We show that resilience and { State-resilience} problems are both undecidable for WSTS with strong compatibility.
+% \alain{harminiser les notations des pbs { State-resilience} ou  State-resilience ?}
+% \mathieu{je pense dans l'intro on peut conserver State-resilience et par contre une fois le probleme de {\ State-resilience} proprement défini c'est cette notation qu'il faudrait utiliser}
 %Moreover, state-resilience,
@@
 %are undecidable for WSTS with strong compatibility having an effective pred-basis. 
-Moreover, {\sc State-resilience},
-{\sc Bounded-state-resilience} and
-{\sc $k$-state-resilience}
+Moreover, { State-resilience},
+{ Bounded-state-resilience} and
+{ $k$-state-resilience}
 are undecidable for strongly upward-compatible WSTS with effective pred-basis
 when
-$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {\sc State-resilience} in reset-VASS.
+$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {\ State-resilience} in reset-VASS.
 
@@
 \item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis.
-%: {\sc State-resilience} is still decidable for 
+%: {\ State-resilience} is still decidable for 
 % WSTS with effective 
@@
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
+We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({ $k$-state-resilience} and { bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
 %

### 2023-08-17 state-resilience.tex
up

@@
 
-We are now interested in the 
+The
 %following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS consists in, given a
 reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
@@
 
-To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
+To better discuss $t$-liveness let us define more formally the upward-closed set of configurations such that $t$ is enabled
 $pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
@@
 \[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
-where
-$U_t=\uparrow pre(t)$ 
-is the upward closure of the preconditions to fire transition $t$.  
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
 The problem reduces itself to {\sc State-resilience}
-where $\Safe = U_t$.
+where $\Safe = pre(t)$.
 \end{proof}

### 2023-08-17 introduction.tex
section

@@
 \includegraphics[width=0.65\textwidth]{FigureB}
-	\caption{A lossy VASS with three subsystems and four counters.
+	\caption{A lossy (distributed) VASS with three subsystems and four counters.
 Leftmost and rightmost subsystems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $C$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.
 Arrows represent control-transitions.
-Unlabelled arrows by default do not modify counter values.}
+Unlabelled arrows do not modify counter values.}
 					\label{air control}
@@
 {(\bf Flight Controller System).}
-Consider the vector addition system in Figure~\ref{air control}. It models a scenario in which two aircrafts want to land in an airport at the same time. 
-The automata $A$ on the left and $B$ on the right represent the aircrafts attempting to land, while the middle one $M$ represents the control tower.
+Consider the (distributed) VASS in Figure~\ref{air control}. It models a scenario in which two aircrafts want to land in an airport at the same time. 
+The VASSs $A$ on the left and $B$ on the right represent the aircrafts attempting to land, while the middle one $M$ represents the control tower.
 The control tower can communicate information to the aircrafts using
@@
 We assume that the VASS is
-lossy;  Lossiness will be represented by using transitions decreasing any counter by one from every state.
+lossy;  \alain{à quoi sert cette phrase ?Lossiness will be represented by using transitions decreasing any counter by one from every state.}
 % We ask however that the system is not too lossy.
 We add the condition that for every prefix of an execution, the number of ``lossiness transition'' for a counter is less than nine tenth of the number of transitions incrementing said counter.
-In this example, the safe states would be ones where the aircrafts have both landed, i.e.
+The safe states would be ones where the aircrafts have both landed, i.e.
 $\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$, and $\Bad$ is the complement of $\Safe$. 
@@
 %the least amount of steps possible. 
-In this exemple, aircrafts cannot simply wait in the air an unbounded amount of time: here, $8$-resilience hold.\alain{expliquer}
+In this exemple, aircrafts cannot simply wait in the air an unbounded amount of time: here, $8$-resilience hold but $8$-resilience don't.\alain{expliquer}
 \end{example}

### 2023-08-17 resilience.tex
section

@@
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
- $k$-resilient ?\newline}
+{$\exists k \geq 0$ such that {$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}}
+
+%$\mathscr{S}$ is %uniformely
+% $k$-resilient ?\newline}
 
@@
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
+Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
 In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.

### 2023-08-17 application.tex
up

@@
 When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
 
@@
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system \alain{pas défini}when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 \end{proposition}

### 2023-08-17 appendix.tex
up

--- 2023-08-15 appendix.tex
@@
 \[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
-where
-$U_t=\uparrow pre(t)$ 
-is the upward closure of the preconditions to fire transition $t$.  
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
 The problem reduces itself to {\sc State-resilience}
-where $\Safe = U_t$.
+where $\Safe = pre(t)$.
 \end{proof}

### 2023-08-17 state-resilience.tex
Qed

--- 2023-08-08 state-resilience.tex
@@
 and 
- {\sc $k$-state-resilience}. 
+ {\sc $k$-state-resilience}. \qed
 \end{proof}
@@
  \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-and {\sc State-resilience} is decidable.
+and {\sc State-resilience} is decidable. \qed
 \end{proof}
@@
 $({\uparrow  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$.
+$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$. \qed
 \end{proof}
@@
 Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} %\cap \Bad
- \setminus \pred^*(\Safe) = \emptyset$. 
+ \setminus \pred^*(\Safe) = \emptyset$.  \qed
 \end{proof}
@@
 In the other direction, since $A \subseteq {\downarrow A}$, the inclusion $({\downarrow  A}) \cap U \subseteq D$ implies $A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
+\qed
 \end{proof}
@@
 until it finds an element
-not in $\pred^*(\Safe)$.
+not in $\pred^*(\Safe)$. \qed
 \end{proof}

### 2023-08-17 application.tex
Qed
section

--- 2023-08-16 application.tex
@@
 \begin{proof}
-The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^d$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^{2d}$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
 % For Z-VAS we observe that we are asking for inclusion between linear sets. 
@@
 The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is also in EXPSPACE. 
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). 
 %			\alain{and may be in coNP-complete. Recall that reachability is P-complete for continuous VASS}
@@
 when $\Safe$ is downward-closed. Unfortunately, we have the following:
-
+\alain{incohérence des notations: VASS ou vector addition system with states and plus bas seulement vector addition system ??? VAS ???}
 \begin{proposition}

### 2023-08-17 application.tex
section

@@
 $\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
-{\sc Resilience} is decidable.
+{\sc Resilience} is decidable. \qed
 \end{proof}
@@
 Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
+\qed
 \end{proof}
@@
 Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
+\qed
 \end{proof}
@@
 $\Safe$ in $k$ or less steps.
+\qed
 \end{proof}

### 2023-08-17 state-resilience.tex
section

@@
 {A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
-{$\exists k \geq 0$ such that $\mathscr{S}$ is $k$-state-resilient ?\newline}
+{$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+%			$\mathscr{S}$ is $k$-state-resilient ?\newline}
 

### 2023-08-19 introduction.tex
Removed \Bad from the intro

@@
 {\bf Resiliences.}  
-%		We now give the definition of two kinds of resilience: the (general) resilience and the state-resilience. 
-The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and two subsets of states $\Bad,\Safe \subseteq S$ consists of the following: $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if
-$\Bad \subseteq \pred^*(\Safe)$. In the case 
-it 
-can be stated as an home-space problem : $\mathscr{S}$ is ($\Bad,\Safe$)-resilient if $\post^*(\Bad) \subseteq \pred^*(\Safe)$.
-Similarly, $\mathscr{S}$ is (\Bad,\Safe)-state-resilient for an initial state $s_0$ if  
-$\post^*(s_0) \cap \Bad \subseteq \pred^*(\Safe)$.
+The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and a subset of states $\Safe \subseteq S$ consists of the following: $\mathscr{S}$ is {\em $\Safe$-resilient} if
+$S \subseteq \pred^*(\Safe)$. %In the case 
+% it
+It 
+can be stated as an home-space problem : $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
+Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
+$\post^*(s_0) \subseteq \pred^*(\Safe)$.
 We will study three decidability questions for each of type of resilience.
-The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is ($\Bad,\Safe$)-resilient (for two given subsets of states $\Bad,\Safe \subseteq S$).
-The $k$-resilience problem is to decide whether $\post^*(\Bad) \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
-the bounded resilience problem is to decide whether there exists an $k$ such that $\mathscr{S}$ is ($\Bad,\Safe$)-$k$-resilient. We adapt these three definitions to state-resilience.
+The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient (for a given subset of states $\Safe \subseteq S$).
+The $k$-resilience problem is to decide whether $S \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
+the bounded resilience problem is to decide whether there exists an $k$ such that $\mathscr{S}$ is $\Safe$-$k$-resilient. We adapt these three definitions to state-resilience.
 
@@
 
-%\item safe clos haut, bad clos par le bas: resilience = reachability d'un clos par le bas (anti-coverability)
-
 \item The three resilience problems are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
-%		two sets $\Bad = \downarrow \Bad$
 
@@
 %$\Safe=\uparrow \Safe$
-%and $\Bad=\downarrow \Bad$. 
 We also show that removing the effective 
@@
 The safe states would be ones where the aircrafts have both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$, and $\Bad$ is the complement of $\Safe$. 
-% We remark that $\Safe$ and $\Bad$ are here both upward and downward closed.
+$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$%, and $\Bad$ is the complement of $\Safe$. 
+.
 Resilience here expresses that the aircrafts will both land, while $k$-resilience is the property insuring that they will do so in at most $k$ steps. 

### 2023-08-22 introduction.tex
up

@@
 The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient (for a given subset of states $\Safe \subseteq S$).
+The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
+whether $\Bad \subseteq \pred^*(\Safe)$.
 The $k$-resilience problem is to decide whether $S \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
@@
 
-\begin{center}
 	\begin{figure}
 			\hspace{2cm}
-\includegraphics[width=0.65\textwidth]{FigureB}
-	\caption{A lossy (distributed) VASS with three subsystems and four counters.
-Leftmost and rightmost subsystems $A$ and $B$ represent aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one $C$ represents the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.
-Arrows represent control-transitions.
-Unlabelled arrows do not modify counter values.}
-					\label{air control}
+\includegraphics[width=0.65\textwidth]{FigureC}
+	\caption{A reset-VASS.}
+					\label{r-V}
 	\end{figure}
@@
 \begin{example}\label{Example}
-{(\bf Flight Controller System).}
-Consider the (distributed) VASS in Figure~\ref{air control}. It models a scenario in which two aircrafts want to land in an airport at the same time. 
-The VASSs $A$ on the left and $B$ on the right represent the aircrafts attempting to land, while the middle one $M$ represents the control tower.
-The control tower can communicate information to the aircrafts using
-% channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using 
-the first two shared counters, and these in turn can send messages to the control tower
-using the last two shared counters. 
-% respectively channel $c_3$ and $c_4$. 
-% Remark channels $c_3$ and $c_4$ have a unary alphabet in the exemple and therefore can be seen as counters.
-
-Safety requires that only one aircraft at a time attempts to land. The role of the control tower is to ensure this;
-there are two possible choices: aircraft $A$ waits for aircraft $B$ to land
-or vice-versa. To make its choice known to the aircrafts, the control tower will use the counters. 
-We assume that the VASS is
-lossy;  \alain{à quoi sert cette phrase ?Lossiness will be represented by using transitions decreasing any counter by one from every state.}
-% We ask however that the system is not too lossy.
-We add the condition that for every prefix of an execution, the number of ``lossiness transition'' for a counter is less than nine tenth of the number of transitions incrementing said counter.
-The safe states would be ones where the aircrafts have both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$%, and $\Bad$ is the complement of $\Safe$. 
-.
-Resilience here expresses that the aircrafts will both land, while $k$-resilience is the property insuring that they will do so in at most $k$ steps. 
-%$k$-resilience is of interest since in general we want to ensure the safe space will 
-%be reached in
-%the least amount of steps possible. 
-In this exemple, aircrafts cannot simply wait in the air an unbounded amount of time: here, $8$-resilience hold but $8$-resilience don't.\alain{expliquer}
-\end{example}
-
-% contrôle de trafic aérien, perturbation dans le planning → Bad (par exemple, trop d’avions qui cherchent à atterir en même temps et qui doivent être mis en stand-by au dessus de l’aéroport?) → Safe (trafic fluide à nouveau).
-K-résilience important car les avions peuvent pas rester indéfiniment en stand-by (re : le carburant est un réactif limitant)
+{(\bf Reset-VASS example).}

### 2023-08-22 introduction.tex
up

@@
 
+ \begin{center}
 	\begin{figure}
-			\hspace{2cm}
-\includegraphics[width=0.65\textwidth]{FigureC}
-	\caption{A reset-VASS.}
+	%		\hspace{3cm}
+\includegraphics[width=0.45\textwidth]{FigureC}
+\hspace{0.5cm}
+\includegraphics[width=0.45\textwidth]{FigureD}
+	\caption{Two reset-VASS with three states and one counter.}
 					\label{r-V}
@@
 {(\bf Reset-VASS example).}
+Consider the leftmost reset-VASS from Figure~\ref{r-V}.
+Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
+Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V$.
+However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of lenght less than $n$ towards $q_1(0)$.
+Consider now the rightmost reset-VASS, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$.
+In this reset-VASS { Bounded resilience} hold, since $10$-resilience hold.
+Indeed, from any configuration with value of the counter bigger than $10$, 
+\end{example}

### 2023-08-22 introduction.tex
up

@@
 	%		\hspace{3cm}
-\includegraphics[width=0.45\textwidth]{FigureC}
-\hspace{0.5cm}
-\includegraphics[width=0.45\textwidth]{FigureD}
+\includegraphics[width=0.4\textwidth]{FigureC}
+\hspace{0.6cm}
+\includegraphics[width=0.4\textwidth]{FigureD}
 	\caption{Two reset-VASS with three states and one counter.}
@@
 Consider now the rightmost reset-VASS, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$.
-In this reset-VASS { Bounded resilience} hold, since $10$-resilience hold.
-Indeed, from any configuration with value of the counter bigger than $10$, 
+In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold.
+Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the reset transition.
+From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less.
+From
+$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps.
 \end{example}

### 2023-08-22 resilience.tex
Qed
up

@@
 
-In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$ and $\Safe ,  \Bad \neq \emptyset$.
+In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
 %	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
@@
 
+\iffalse
 In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
@@
 The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.
+\fi
 
@@
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
-Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
+% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.
 
@@
 Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
+\qed
 		\end{proof}

### 2023-08-22 application.tex
up

@@
 
+\iffalse
 Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
@@
 A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
+\fi
 

### 2023-08-22 definitions.tex
up

@@
 Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
-The lossy VASS given in the {\bf Flight Controller System} example, for instance, is a WSTS.
- \mathieu{Faire la preuve une fois que j'ai l'equation}
+% The lossy VASS given in the {\bf Flight Controller System} example, for instance, is a WSTS.
 
@@
 %
-% \mathieu{Mentionner que le lossy VASS de l'example~\ref{Example} est effectif}
-As an illustration, the lossy VASS given in the {\bf Flight Controller System} example is post-effective,
-ideal-effective, and
-has effective pred-basis.
+% As an illustration, the lossy VASS given in the {\bf Flight Controller System} example is post-effective, ideal-effective, and has effective pred-basis.
 
@@
 \end{theorem}
+
+Let us recall finally that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
+$||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
+%
+Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
+For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+%
+When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.

### 2023-08-22 introduction.tex
up

@@
 In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
-$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$, (hence $\Bad$ is downward-closed).
+$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$ and $\Bad$ downward-closed.
 %		(even though \cite{DBLP:conf/gg/Ozkan22} also discuss, albeit briefly, the other possible cases). 

### 2023-08-22 appendix.tex
up

@@
-\section{Resilience when $\Safe$ and $\Bad$ have the same closure properties}
-
-We relax now the hypothesis that $\Bad$ is the complement of $\Safe$.
-In this appendix it is thus possible for $\Safe$ and $\Bad$ to share closure properties, i.e.
-for them to be both downward-closed or upward-closed.
-We will now explore both possiblities.
-
-\subsection{Case: $\Bad=\downarrow \Bad$ and $\Safe=\downarrow \Safe$}\label{case down down}
-
-% In the case \Safe and \Bad are both downward-closed, they can both be finite.
-We start with the case $\Safe = \downarrow \Safe$ and $\Bad = \downarrow \Bad$.
-%
-Unfortunately, in this case the resilience problem is undecidable.
-
-\begin{theorem}\label{down-down}
-{\sc Resilience} is undecidable for  effective WSTS with  strong  compatibility such that
-%	with more than one minimal element, 
-$\Safe=\downarrow \Safe$
-and $\Bad=\downarrow \Bad$.
-\end{theorem}
-
-\begin{proof}
-If the set $S$ of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ has an unique minimal element $m$, then $m$ belongs to both $\Safe$ and $\Bad$ which contradicts the assumption $\Safe \cap \Bad= \emptyset$.
-So let us consider the case of a set $S$ with at least two minimal elements $m_1$ and $m_2$.
-The problem of whether $m_2$ is reachable from $m_1$ reduces itself to the resilience problem by considering $\Safe=\downarrow m_2 = \{ m_2\}$ and $\Bad=\downarrow m_1 = \{ m_1\}$.
-By undecidability of the reachability problem for effective WSTS with strong compatibility we conclude.
-\end{proof}
-
-\subsection{Case: $\Bad=\uparrow \Bad$ and $\Safe=\uparrow \Safe$.}\label{case up up}
-
-% \alain{2 clos par le haut ont une intersection non vide....}
-
-We now consider the case where $\Safe = \uparrow \Safe$ and $\Bad$ is not the complement of 
-$\Safe$ and we'd want $\Bad$ to have the same closure properties as $\Safe$.
-We still however maintain the hypothesis that $\Bad \cap \Safe = \emptyset$, thus the conditions 
-$\Safe = \uparrow \Safe$ and $\Bad = \uparrow \Bad$ would lead to a contradiction, since two upward-closed sets have not an empty intersection.
-We consider instead the case where $\Bad$ is as upward-closed
-as it can be without intersecting with $\Safe$,
-i.e. 
-such that there exists an upward-closed set $U$
-such that $\Bad = U \setminus \Safe$.
-
-\begin{theorem}\label{up-up}
-{\sc Resilience} is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
-and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS,  $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$ and $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$.
-Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
-\begin{proposition}
-In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the {\sc Bounded resilience} is equivalent to {\sc Resilience}.
-\end{proposition}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
-compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
-% \alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
-and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
-Hence the equivalence.
-\end{proof}
-
-\begin{corollary}\label{B-up-up}
-{\sc Bounded resilience}  is decidable for WSTS with effective pred-basis,
-strong compatibility
- $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
-\end{corollary}
-
-\begin{theorem}\label{k-up-up}
-{\sc $k$-resilience}  is decidable for WSTS with effective pred-basis, strong %upward-
-compatibility, $\Safe=\uparrow \Safe$
-and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
-%	and \Bad is upward-closed or downward-closed.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
-Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong (upward) compatibility,  $\pred^{\leq n}(\Safe)=\uparrow \pred^{\leq n}(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$.
-% Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm.
-Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$.
-% Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
-Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
-% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
-hence the resilience problem is decidable.
-\end{proof}
-
 \section{Reductions to reset-VASS}

### 2023-08-22 resilience.tex
up

@@
 %	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
-The \emph{resilience problem} (resp.
-the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe,\Bad)$ is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp.
-a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
-We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp.
-$\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.
-$\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
-In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
+The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe)$ is to decide whether from \emph{any} state in %$\Bad$
+$S$, \emph{there exists} a path (resp. a path of length smaller than or equal to $k$) that reaches a state in $\Safe$. We use the notation 
+% $\Bad \longrightarrow^{*} \Safe$ (resp. $\Bad \longrightarrow^{\leq k} \Safe$)
+$S \longrightarrow^{*} \Safe$ (resp. $S \longrightarrow^{\leq k} \Safe$)
+ for 
+ %$\forall x \in \Bad, \exists y \in \Safe$ 
+$\forall x \in S, \exists y \in \Safe$ 
+ such that $x \longrightarrow^{*} y$ 
+ %(resp.  $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
+  (resp.  $\forall x \in S, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
+  In our framework, $\Safe \subseteq S$  is possibly infinite but  must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
 
@@
 \problemx{Resilience (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
+{$S \longrightarrow^{*} \Safe$ ?\newline}
 
@@
 \problemx{$k$-resilience (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
-{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a sets $\Safe \subseteq S$.}
+{$S \longrightarrow^{\leq k} \Safe$ ?\newline}
 
 \problemx{Bounded resilience (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
 %{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
-{$\exists k \geq 0$ such that {$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}}
+{$\exists k \geq 0$ such that {$S \longrightarrow^{\leq k} \Safe$ ?\newline}}
 
@@
 
-\iffalse
-
-\problemx{Resilience problems}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
-{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
-({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
-({\sc bounded resilience problem (BRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
- $k$-resilient ?\newline}
-
-\fi
-
-\iffalse
-In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
-$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
-and $\Bad$ is the complement of $\Safe$. 
-% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
-% wheher
-The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.
-\fi
-
-\iffalse
-\begin{example}
-Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
-then resilience hold but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g. $q(k+1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$. 
-\end{example}
-\fi
-
 These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
 So we restrict our framework to the class of infinite-state WSTS.
-Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$.
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed.
+Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
+.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.
 %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.}
@@
 %and in the general case when $\Safe = \downarrow \Safe$.
-
-It is useful to note that in a sequence of states% of $S$
-, every prefix is always 
-either in the case that the latest visited $\Bad$ state has already been followed by a 
-$\Safe$ state \-- and thus the latest state is contextually ``$\Safe$'' \--
-or in the case that it still has to \-- and thus the latest state is contextually ``$\Bad$''. 
-Thus, \emph{in context}, there is no true ``neutral'' state.
-Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but
-% w.l.o.g.
-	also
-that $\Bad$ is the complement of $\Safe$,
-in which case
-if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
-Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. are both downward-closed or both upward-closed,
-have also been studied, 
-see 
-Appendix~\ref{case down down}
-and
-Appendix~\ref{case up up} for results.
 

### 2023-08-22 resilience-safe-up.tex
up

@@
 
-We start with the case $\Safe=\uparrow \Safe$ hence $\Bad=\downarrow \Bad$. 
+We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$. 
 It can be viewed as a generalizeation
@@
 it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
-
-% Let us choose $\Bad$ to be the set of states from which there is no infinite run.
-If we are in an upward compatible ordered transition system (like VASS, lossy channel systems,...), then $\Bad$ is downward-closed.
-% \alain{ce qui précède est faux et pas convaincant: un vass qui seulement décroit ses compteurs}
-
-%%%%
-
-\iffalse
-As a
-example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e.
-from which it is always possible to fire a transition.
-This set of states is upward-closed and $\Safe=S \setminus \Bad=\uparrow \Safe$.\alain{mal rédigé}
-\fi
-
-%  Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
-%  Alain Finkel & Philippe Schnoebelen (2001): Well-structured transition systems everywhere! Theor. Comput. Sci. 256(1-2), pp. 63–92, doi:10.1016/S0304-3975(00)00102-X.
-
-%Transfering the abstract resilience problems into this framework,
-%it is therefore reasonable to demand that both propositions, \Safe and \Bad, are given by 
-%upward-closed or downward-closed sets.
-
-%
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two 
-%finite \alain{non, confusion entre ensemble et base}
- sets $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 \begin{proof}
-Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$.
+Let  $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
 The %uniform 
@@
 
-Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
+Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$. \qed
 %	If $k <  \min(K,n)$, then we check every path of length smaller
@@
 Let us remark that coverability reduce
-to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
+to resilience when  $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$,
 as $s \to^* \uparrow t$ 
-reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
+reduces to $\Safe$-resilience for $\Safe = \uparrow t$ and $\Bad = \downarrow s$.
 %

### 2023-08-22 resilience-safe-up.tex
up

@@
 
+Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
 Let us remark that coverability reduce
@@
 reduces to $\Safe$-resilience for $\Safe = \uparrow t$ and $\Bad = \downarrow s$.
-%
+\mathieu{No longer the case here since $\Bad$ needs be the complement of $\Safe$.}
 Let us recall that the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}, hence the strongly increasing well-structured nets are not completion-post-effective hypothesis nor have the pred-basis hypothesis.
@@
 Moreover, this implies from the reduction that resilience for strongly increasing well-structured nets is  undecidable. Hence resilience for WSTS in general is undecidable.
-
-Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.

### 2023-08-22 resilience-safe-down.tex
Qed
section
up

@@
 $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
-the resilience problem is decidable.
+the resilience problem is decidable. \qed
 \end{proof}
@@
 % \alain{qui appartient à qui ?}
-belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$.
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$. \qed
 \end{proof}
@@
 
-Without the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed, the problem becomes undecidable.
-
+Without the hypothesis that
+{\em for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}, the problem is undecidable.
  Indeed, it is 
@@
 Home-state is undecidable for Minsky machine with $3+$ counters.
-This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
+This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61,Min67}.
 From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 

### 2023-08-22 resilience-safe-down.tex
up

@@
 
-Without the hypothesis that
-{\em for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}, the problem is undecidable.
+In the case
+of a ideal-effective WSTS 
+where the hypothesis that
+{for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}
+is not met, the problem is undecidable.
  Indeed, it is 
-undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$.
+undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$.
 This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.

### 2023-08-22 resilience.tex
up

@@
 Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
-.
+. Since $\Safe \subseteq \pred^*(\Safe)$, one only need to decide whether 
+the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.
 In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.
@@
 
-We now briefly investigate the question to control a system such that its controlled behavior may always reach $Safe$.
+We now briefly investigate the question to control a system such that its controlled behavior may always reach $\Safe$.
 %
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{Safe}$. 
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $\Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{\Safe}$. 
 %
 %		\begin{definition}
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient if $S  \xrightarrow{*} Safe$ is satisfied.  
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient if $S  \xrightarrow{*} \Safe$ is satisfied.  
 %		\end{definition}
 %
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(Safe)$-resilient.
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(\Safe)$-resilient.
 
-The \emph{maximal $Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{Safe}=(S,\rightarrow_{Safe},\leq)$ where $\rightarrow_{Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{Safe}$ is $Safe$-resilient.
+The \emph{maximal $\Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{\Safe}=(S,\rightarrow_{\Safe},\leq)$ where $\rightarrow_{\Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{\Safe}$ is $\Safe$-resilient.
 %\alain{vérifier que la plus grande existe}
@@
 \begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $Safe$, the maximal resilient subsystem is a computable WSTS.
+Given an effective WSTS with effective pred-basis and an upward-closed set $\Safe$, the maximal resilient subsystem is a computable WSTS.
 \end{theorem}
@@
 \begin{proof}
-Since $Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(Safe)=\{x_1,x_2,...,x_p\}$. 
+Since $\Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(\Safe)=\{x_1,x_2,...,x_p\}$.
 %
 %
-Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
-\qed
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(\Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
 		\end{proof}

### 2023-08-22 state-resilience.tex
up

@@
 
-Resilience is a strong property that implies that from every element of $\Bad$ there must exist a path to $\Safe$.
-However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from $\Bad \cap \post^*(s_0)$, there must exist a path to $\Safe$.
-In the {\bf Flight Controller System}, the state $(q_{A,1}, q_{M,4}, q_{B,0})(\textbf{0})$
-belongs to $\Bad$, but is not reachable from $(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{0})$,
-hence it may seem unnecessary to check for the existence of a path to $\Safe$ from this element.
-Since we still consider $\Bad$ and $\Safe$ complements of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
+Resilience is a strong property that implies that from every element %of $\Bad$
+ there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from 
+ $
+ % \Bad \cap 
+ \post^*(s_0)$, there must exist a path to $\Safe$. 
+%In the {\bf Flight Controller System}, the state $(q_{A,1}, q_{M,4}, q_{B,0})(\textbf{0})$ belongs to $\Bad$, but is not reachable from $(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{0})$, hence it may seem unnecessary to check for the existence of a path to $\Safe$ from this element.
+%
+% Since we still consider $\Bad$ and $\Safe$ complements of each others, this is akin to asking whether from $\post^*(s_0)$ there always exists a path to $\Safe$. 
 The three previous problems become:
@@
 
-We start with the case $\Safe = \uparrow \Safe$ hence $\Bad = \downarrow \Bad$.
+We start with the case $\Safe = \uparrow \Safe$.
 %
@@
 In such a case the algorithm from
-\cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold,
+\cite{DBLP:conf/gg/Ozkan22} would deduce that {\sc $1$-state-resilience} does not hold,
 which is incorrect.
@@
 
-We now consider the case $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
+We now consider the case $\Safe = \downarrow \Safe$.
 %
 Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in % Appendix~\ref{HS-Minsk}.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+% Appendix~\ref{HS-Minsk}. 
+Section~\ref{safe-down}.
 

### 2023-08-22 application.tex
up

@@
-%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
-%
-%
-%\noindent
-%{\bf Vector addition system with states}
 %
@@
 \begin{theorem}{}
-{\sc Resilience} is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear.
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
 \end{theorem}
@@
 when $\Safe$ is downward-closed. Unfortunately, we have the following:
-\alain{incohérence des notations: VASS ou vector addition system with states and plus bas seulement vector addition system ??? VAS ???}
+% \alain{incohérence des notations: VASS ou vector addition system with states and plus bas seulement vector addition system ??? VAS ???}
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 \end{proposition}
@@
  $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
-consider a given vector addition system $V$.
+consider a given VAS $V$.
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
@@
  $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
-Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for vector addition system when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
 This lead to a decision procedure%, written in detail in Appendix~\ref{appendix brp VASS},

### 2023-08-22 application.tex
up

@@
 \begin{theorem}{}
-{\sc Resilience} is decidable for VASS when $\Safe$ and $\Bad$ are semilinear sets.
+{\sc Resilience} is decidable for VASS when $\Safe$ 
+%and $\Bad$ are semilinear sets.
+is a semilinear set.
 \end{theorem}
@@
 \begin{theorem}{}
-{\sc Resilience} is decidable for lossy counter machines when $\Safe$ and $\Bad$ are semilinear sets.
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$ 
+%and $\Bad$ are semilinear sets.
+is a semilinear set.
 \end{theorem}
@@
 \begin{theorem}\label{RP VASS}
-{\sc Resilience} is decidable for continuous VASS when $\Safe$ and $\Bad$ are definable in the existential theory of
-the rationals with addition and order.
+{\sc Resilience} is decidable for continuous VASS when $\Safe$ 
+%and $\Bad$ are
+is
+%s
+definable in the existential theory of the rationals with addition and order.
 \end{theorem}
@@
 \end{proof}
-
-\iffalse
-\begin{figure}
-\begin{center}
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \Safe~\Bad 
-		& $\uparrow$~ $\downarrow$~ 
-		 & $\downarrow$~ $\uparrow$~ 
- \\ \hline
-   RP  & Decidable (Thm~\ref{down-up})   & Decidable (Thm~\ref{RP VASS})
-    \\ \hline
-   BRP   &  Decidable (Thm~\ref{down-up})  & Decidable (Thm~\ref{brp VASS})
-    \\ \hline
-      kRP       & Decidable (Thm~\ref{down-up}) & Decidable (Thm~\ref{brp VASS})
-       \\ \hline
-   sRP    	& Decidable (Thm~\ref{post srp})	 & ??
-    \\ \hline
-   BsRP     &  Decidable (Thm~\ref{ref ozkan})	 & ?? 
-    \\ \hline
-      ksRP        & Decidable (Thm~\ref{ref ozkan}) & ?? 
-       \\ \hline
-
-\end{tabular}
-\end{center}
-	\caption{Decidability results for VASS.}
-\end{figure}
-\fi
 

### 2023-08-22 application.tex
up

@@
 
-\begin{theorem}{}
+\begin{theorem}\label{SL VASS}
 {\sc Resilience} is decidable for VASS when $\Safe$ 
@@
 
-\begin{theorem}{}
+\begin{theorem}\label{SL Lossy}
 {\sc Resilience} is decidable for lossy counter machines when $\Safe$ 
@@
 
-\begin{theorem}{}
+\begin{theorem}\label{SL int-VASS}
 {\sc Resilience} is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
@@
 \end{proof}
+
+Remark that
+Theorem~\ref{SL Lossy} and
+Theorem~\ref{SL int-VASS}
+(resp. Theorem~\ref{RP VASS})
+did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
+semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
+the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
 

### 2023-08-22 application.tex
up

@@
 \begin{theorem}\label{SL int-VASS}
-{\sc Resilience} is decidable for integer VASS when $\Safe$ and $\Bad$ are semilinear.
+{\sc Resilience} is decidable for integer VASS when $\Safe$ %and $\Bad$ are
+is a
+semilinear set.
 \end{theorem}
@@
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$% and $\Bad = \uparrow \Bad$
+.
 \end{proposition}
@@
 \begin{theorem}
-{\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
+{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+$\Safe = \downarrow \Safe$%
+%$\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
 .

### 2023-08-22 conclusion.tex
section
up

@@
-\section{conclusion}
+\section{Conclusion}
 
@@
 We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general. We also exhibited classes of WSTS with decidable resilience, namely, 
-completion-post-effective $\omega^2$-WSTS with strong compatibility (in the case $\Bad = \downarrow \Bad$ and $\Safe = \uparrow \Safe$) and 
+completion-post-effective $\omega^2$-WSTS with strong compatibility (in the case 
+%$\Bad = \downarrow \Bad$ and
+ $\Safe = \uparrow \Safe$) and 
 downward-compatible
 ideal-effective WSTS
-(in the case $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$).
+(in the case %$\Bad = \uparrow \Bad$ and
+ $\Safe = \downarrow \Safe$).
 %We additionnally generalized decidability results from~\cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
@@
 For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
-One other angle of attack not studied here is synthesis of sets $\Bad$ and/or $\Safe$ for resilience,
-for instance, given a set $\Safe$, find the maximal upward-closed (resp. downward-closed) subset 
-$\Bad$ so that a system is $(\Safe,\Bad)$-resilient. 
+One other angle of attack not studied here is synthesis of sets %$\Bad$ and/or 
+$\Safe$ for resilience,
+for instance, %given a set $\Safe$,
+ find the maximal upward-closed (resp. downward-closed) subset 
+% $\Bad$ so that a system is $(\Safe,\Bad)$-resilient.
+$\Safe$ so that a system is $\Safe$-resilient. 
 % \alain{trouver les \Bad~ et \Safe~ maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ?}
-
-One could also extend upon the classes of set $\Bad$ and $\Safe$ considered.
+Another question could be to analyse the resilience in the framework of a controller and its environment. 
+One could also extend upon the classes of set $\Safe$ %and $\Safe$
+ considered.
 As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets \cite{DBLP:journals/fmsd/BertrandS13}.

### 2023-08-22 resilience-safe-down.tex
up

@@
 %%%%%%%%
-\subsection{Case: $\Safe=\downarrow \Safe$.}
+\subsection{Case: $\Safe=\downarrow \Safe$.}\label{safe-down}
 %
@@
 
-Remark we did not make use of the hypothesis $\Bad$ complement of $\Safe$, simply 
+Remark we did not make use of the
+% hypothesis
+property $\Bad$ complement of $\Safe$, simply 
 $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 

### 2023-08-22 resilience-safe-up.tex
up

@@
-\subsection{Case: $\Safe=\uparrow \Safe$.}
+\subsection{Case: $\Safe=\uparrow \Safe$.}\label{safe-up}
 
@@
 
-Remark we did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply using
+Remark we did not make use of the 
+% hypothesis
+property that $\Bad$ is the complement of $\Safe$, simply using
 $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 

### 2023-08-22 definitions.tex
up

@@
-\section{Well-structured transition systems}\label{section definitions}
+\section{Well-structured transition systems and VASS}\label{section definitions}
 
@@
 
-Let us recall finally that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+Let us recall the  definition of vector addition system with states. 
+ \begin{definition} 
+A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+ \end{definition} 
 % The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e.
@@
 When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.
+
 A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.

### 2023-08-22 resilience-safe-down.tex
ip

@@
  is a finite 
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
+labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and 
 $T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control-transitions} of $V$,
+is the set of {\em control-transitions},
 where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 

### 2023-08-22 resilience-safe-down.tex
up

@@
  is a finite 
-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and 
+labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, 
 $T \subseteq Q \times Op \times Q$
-is the set of {\em control-transitions},
-where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+is the set of {\em control-transitions}, and $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
 		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.

### 2023-08-22 appendix.tex
up

@@
-\section{Reductions to reset-VASS}
-
-Most of our undecidability results rely on reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}.
-This has consequences for {\sc Resilience}; for instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element.
-Even then, {\sc Resilience} (resp.
-{\sc State-resilience}) is not strictly speaking the same as reachability, and we will need to delve more into the details to obtain undecidability results.
-
-This section is dedicated to studying reset-VASS for showing undecidability of {\sc Resilience} and {\sc State-resilience} for WSTS in general.
-
-\subsection{Home-state is undecidable for Minsky machine, reset-VASS and WSTS}\label{HS-Minsk}
-
-We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
-Let us provide a more detailed undecidability proof here.
-
-First, let us recall the home-state problem
-
-\problemx{Home-state}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
-{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}
-
-% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
-% post ∗ (S ) ⊆ pred ∗ (s_0)
-
-Home-state is undecidable for Minsky machine with $3+$ counters.
-This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
-From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
-such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
-We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
-% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
-% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
-Thus {\sc Home-state} is undecidable for Minsky machines.
-Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}.
-Hence {\sc Resilience} is undecidable for Minsky machines.
-
-Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
-% Let us recall more formally what a reset-VASS is.
-Reset-VASS extend the basic VASS model with special “reset
-transitions” that resets (set to $0$) some coordinates in the vector. Let us recall their definition here.
-\begin{definition}
-A {\em reset-VASS} in dimension $d$ %(reset-VASS for short)
- is a finite 
-labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
-$T \subseteq Q \times Op \times Q$
- will be referred to as the {\em control-transitions} of $V$,
-where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
-
-Again $Q \times \N^d$
- denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
-$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
-\begin{samepage}\begin{itemize}
-\item  $t = (q,add(\textbf{z}),q') \in T$
-% then for all $\textbf{u} \in \N^d$ such that  
-% $\textbf{u}+\textbf{z} \geq 0$
-% $q(\textbf{u}) \xrightarrow{\textbf{z}} q'(\textbf{u}+\textbf{z})$,
-and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-\item $t = (q,reset(\gamma),q') \in T$ 
-% then for all $\textbf{u} \in \N^d$ 
-% $q(\textbf{u}) \xrightarrow{z} q'(\textbf{u}')$,  where 
-and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize} \end{samepage}
-
-It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
-
-\subsection{Reset-VASS and state-resilience}\label{appendix}
+\section{Reset-VASS and state-resilience}\label{appendix}
 

### 2023-08-22 resilience.tex
up

@@
    &  Decidable (Thm~\ref{down-up}) 
-   		 & ?? 
+   		 & ?
    	%	 & ??
@@
       & Decidable (Thm~\ref{down-up}) 
-      		& ?? 
+      		& ? 
       	%	& ??
@@
 
+%%%%%%%
+\iffalse
+%%%%%%%%%%%%
 \subsection{Computing the maximal resilient subsystem}
@@
 %It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
+
+%%%%%%%
+\fi
+%%%%%%%%%

### 2023-08-22 state-resilience.tex
up
section

@@
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
-   BSRP & Undecidable (Thm~\ref{srp up down}) &  ??  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
-      kSRP & Undecidable (Thm~\ref{srp up down}) & ?? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+   BSRP & Undecidable (Thm~\ref{srp up down}) &  ?  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+      kSRP & Undecidable (Thm~\ref{srp up down}) & ? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
  \end{tabular}
@@
 \end{corollary}
+
+%\subsection{Computing the maximal resilient subsystem}

### 2023-08-22 resilience-safe-up.tex
up

@@
 Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
-Let us now define $K_{\mathscr{S}}(\Safe,\Bad)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
 %
@@
 %	(if all $k_{i,j}$ are finite) else $K=\infty$.
-We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe,\Bad)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe,\Bad)$-resilient with $K_{\mathscr{S}}(\Safe,\Bad)$ finite.
+We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
 
@@
 
-Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$. \qed
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$. \qed
 %	If $k <  \min(K,n)$, then we check every path of length smaller

### 2023-08-23 application.tex
section
up

@@
 A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
-\fi
+%%%%%%%%
 
@@
 
+%%%%%%%%%%
+\fi
+%%%%%%%%%%%
+
 \begin{theorem}\label{SL VASS}
@@
 $V'$, $\Safe'$ and $\Bad'$.
-For the construction, $V'$ will have two copies of the control states of $V$; $Q_{A}$ and $Q_{B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp.
+For the construction, $V'$ will have two copies of the control-states of $V$; $Q_{A}$ and $Q_{B}$, and will have two other VASS running in parallel checking whether or not the current configuration belong to $\Safe$ resp.
 $\Bad$.
@@
 
-This all changes when one considers VASS (having control states).
+This all changes when one considers VASS (having control-states).
 For a VASS $V = (Q,T)$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----

### 2023-08-23 resilience-safe-down.tex
up

@@
 \problemx{Home-state}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
-{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and two states $s_0, s \in S$.}
+{$\post^*(s_0) \subseteq \pred^*(\{s\})$ ?\newline}
 
@@
 
-Home-state is undecidable for Minsky machine with $3+$ counters.
+Home-state is undecidable for Minsky machine with at least $3$ counters.
 This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61,Min67}.
@@
 such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
-We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
+We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
 % Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).

### 2023-08-23 resilience-safe-up.tex
up

@@
 
+The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
+Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of %strongly 
+increasing recursive functions from \cite{FMP-ic04} defined as
+
+$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+
+	\noindent	
+	where $TM_j$ is the $j$-th Turing machine (in a classical enumeration)
+	which moreover begins
+by writing the integer $j$ on its tape,	
+	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+	The
+	transition system 
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+	is a WSTS
+	and
+	has the property that
+	it is
+	$\uparrow (1,1)$-resilient
+	iff $TM_j$ halts on input $j$,
+	since $(1,1)$ is coverable from $(0,0)$
+	iff $TM_j$ halts on input $j$. 
+	Hence there is no Turing machine which correctly determines 
+	resilience and halts whenever its input is a
+	WSTS.
+\iffalse
+
 Let us remark that coverability reduce
@@
 %
-Moreover, this implies from the reduction that resilience for strongly increasing well-structured nets is  undecidable. Hence resilience for WSTS in general is undecidable.
+Moreover, this implies from the reduction that resilience for strongly increasing well-structured nets is  undecidable. 
+\fi
+Hence resilience for WSTS in general is undecidable.

### 2023-08-23 introduction.tex
up

@@
 %			from all reachable states from $s_0$, it is always possible to reach $H$ 
-$\post^*(s_0) \subseteq \pred^*(H)$; it could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
+$\post^*(s_0) \subseteq \pred^*(H)$. If the set $H$ contains a single element, this element is an {\em home-state}.
+It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
 In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
@@
 	\begin{figure}
-	%		\hspace{3cm}
-\includegraphics[width=0.4\textwidth]{FigureC}
-\hspace{0.6cm}
-\includegraphics[width=0.4\textwidth]{FigureD}
-	\caption{Two reset-VASS with three states and one counter.}
+			\hspace{0.8cm}
+%\includegraphics[width=0.35\textwidth]{FigureC}
+%			\hspace{0.8cm}
+%\includegraphics[width=0.35\textwidth]{FigureD}
+\includegraphics[width=0.75\textwidth]{FigureCD}
+	\caption{Two reset-VASS with three control-states and one counter.}
 					\label{r-V}
@@
 {(\bf Reset-VASS example).}
-Consider the leftmost reset-VASS from Figure~\ref{r-V}.
+Consider the %leftmost
+ reset-VASS $V_1$ from Figure~\ref{r-V}.
 Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
-Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V$.
+Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$.
 However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of lenght less than $n$ towards $q_1(0)$.
-Consider now the rightmost reset-VASS, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$.
-In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold.
+Consider now the %rightmost
+ reset-VASS $V_2$, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$ compared to $V_1$. In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold. 
 Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the reset transition.
@@
 \end{example}
+
+Hence we can see how adding transitions can 
+turn a system that is not bounded resilient into one that is.

### 2023-08-23 resilience-safe-down.tex
up

@@
 
-First, let us recall the home-state problem
+First, let us recall the home-space problem
 
-\problemx{Home-state}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and two states $s_0, s \in S$.}
-{$\post^*(s_0) \subseteq \pred^*(\{s\})$ ?\newline}
+\problemx{Home-space}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ a states $s_0 \in S$ and a subset $H \subseteq S$.}
+{$\post^*(s_0) \subseteq \pred^*(H)$ ?\newline}
 
@@
 
-Home-state is undecidable for Minsky machine with at least $3$ counters.
+Home-space is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
 This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61,Min67}.

### 2023-08-23 state-resilience.tex
up

@@
 
+Remark that {\sc State-resilience} is {\sc Home-space} with input set $\Safe$.
 \iffalse
@@
 \fi
-
 Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
@@
  reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+We now recall the
+%following
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
+which we then reduce to {\sc State-resilience}.
+A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
+live. This leads to the following problem.
+
+\problemx{$t$-liveness}
+{A reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{Is $t$ live in $s_0$ ? \\}
+
+To better discuss $t$-liveness let us define more formally the upward-closed set of configurations such that $t$ is enabled
+$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+
+\begin{proposition}\label{reductions}
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness 
+in a 
+% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
+reset-VASS $(Q,T)$ 
+with initial configuration $s_0$
+as the following formula
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
+The problem reduces itself to {\sc State-resilience}
+where $\Safe = pre(t)$.
+\end{proof}
+
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS, whose proof can be seen in Appendix~\ref{appendix}%
+.
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
+\end{proposition}
 

### 2023-08-23 appendix.tex
up

@@
 \section{Reset-VASS and state-resilience}\label{appendix}
-
-We are now interested in the 
-%following
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-%
-We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
-which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
-live. This leads to the following problem.
-
-\problemx{$t$-liveness}
-{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
-{Is $t$ live in $s_0$ ? \\}
-
-To better discuss $t$-liveness let us define more formally the upward-closed set of configurations such that $t$ is enabled
-$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
-
-\begin{proposition}\label{reductions}
-{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
-\end{proposition}
-
-\begin{proof}
-We reformulate $t$-liveness 
-in a 
-% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
-reset-VASS $(Q,T)$ 
-with initial configuration $s_0$
-as the following formula
-\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
-The problem reduces itself to {\sc State-resilience}
-where $\Safe = pre(t)$.
-\end{proof}
-
-The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
-.
-\begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
-\end{proposition}
 
@@
 
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
+\end{proposition}
+
+%\begin{theorem*}{{\em \textbf{\ref{brp VASS}}}.}
+% {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
+%.
+%\end{theorem*}
+
 Let us describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri}.
@@
 \begin{proof}
-
-\mathieu{Thinking about using reset-VAS to simplify notations.}
 
@@
 \end{proof}
-
-Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
-% the following:
-%
-%
-{\sc %$ reset-VASS 
-$t$-liveness} is undecidable.
-
-Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
-Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
-%This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

### 2023-08-23 state-resilience.tex
up

@@
 \begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
+The  zero reachability problem can be reduced to {\sc $t$-liveness}.
 \end{proposition}
+
+Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
+% the following:
+%
+%
+{\sc %$ reset-VASS 
+$t$-liveness} is undecidable.
+
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
+Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
+%This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
 

### 2023-08-23 resilience-safe-down.tex
up

@@
 % or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
-Thus {\sc Home-state} is undecidable for Minsky machines.
-Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}.
+Thus {\sc Home-space} is undecidable for Minsky machines.
+Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-space} is a particular instance of {\sc Resilience}.
 Hence {\sc Resilience} is undecidable for Minsky machines.

### 2023-08-23 appendix.tex
up

@@
 
-\begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
-\end{proposition}
+Let us recall the undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+We describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri} reducing the problem to {\sc $t$-liveness}. \\
 
-%\begin{theorem*}{{\em \textbf{\ref{brp VASS}}}.}
-% {\sc $k$-resilience }  and {\sc Bounded resilience} in VASS when $\Bad = \uparrow \Bad$ and $\Safe = S \setminus \Bad$ are decidable%, and moreover {\sc $k$-resilience } is in $\EXPSPACE$
-%.
-%\end{theorem*}
-
-Let us describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri}.
+\noindent
+\textbf{Proposition~\ref{liveness reset}}
+%\label{liveness reset}
+{\em The  zero reachability problem can be reduced to {\sc $t$-liveness}.}
 
@@
 with
-initial vector $\textbf{u}$ and initial state $q$.
+initial vector $\textbf{u}$ and initial state $q_0$.
 To do so, we construct a reset-VASS $V_2$
-with initial vector $\textbf{v}$ and initial state $q'$
+with initial vector $\textbf{v}$ and initial state $q'_0$
 which is live if and only if 
-$\textbf{0}$
-is reachable from $\textbf{u}$.
+$q(\textbf{0})$
+is reachable from $q_0(\textbf{u})$.
 
@@
 and
-$|V_1|+4$ new transitions.
+$|V_1|+4$ new transitions for every control-state.
 
@@
 As long as they are both non-zero, the adjusted transitions of $V_1$ are live and can be used normally.
-Thus any vector which is reachable in $V_1$ is also reachable [in $V_2$].
+Thus any vector which is reachable in $V_1$ is also reachable% [in $V_2$]
+.
 
@@
 
-We add an additional transition $s$ which reset the first two additional vector coordinate.
+We add an additional transition $s$ which reset the first two additional vector coordinate from control-state $q$.
 % Transition $s$ is defined to have $r_1$ as an input and a null output.
@@
 These first two additional vector coordinate together with this additional transition allow
-to reach aby reachable vector and then to freeze.
+to reach any reachable vector and control-state $q$ and then to freeze the configuration.
 
@@
 
-If $\textbf{0}$ is reachable, then these $d$ new transitions are not live. Everything is dead if $s$ fires in $\textbf{0}$.
+If $q(\textbf{0})$ is reachable, then these $d$ new transitions are not live. Everything is dead if $s$ fires in $q(\textbf{0})$.
 
@@
 
-Now if $\textbf{0}$ is reachable from $\textbf{u}$ in $V_1$, then $V_2$ can also reach a similar
-vector $(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transition.
+Now if $q(\textbf{0})$ is reachable from $q_0(\textbf{u})$ in $V_1$, then $V_2$ can also reach a similar
+vector $q(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transition.
 Then $s$ can freeze the $V_1$ subset,
@@
 Then $s$ can fire, freezing the $C_1$ subset.
-ending up in $0^{d+4}$, and $V_2$ is not live.
+ending up in $q(0^{d+4})$, and $V_2$ is not live.
 % Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $\textbf{0}$ is reachable then $V_2$ is not live.
+Thus if $q(\textbf{0})$ is reachable then $V_2$ is not live.
 
@@
 % Conversely, if $C_2$ is not live, then a marking $\mu$ must be reachable in which  $\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token (in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.).
-If it is not possible to have (), then the transitions are not live, and the
+If it is not possible to have a strictly positive integer in one of the last two coordinates, then the transitions are not live, and the
 $d$-dimensional subset of the current vector must be zero.
@@
 Thus if $V_2$ is not live then 
-$\textbf{0}$ is reachable.
+$q(\textbf{0})$ is reachable.
 \end{proof}

### 2023-08-23 resilience-safe-down.tex
up

@@
 
+\iffalse
 	% 	From Appendix
@@
 % post ∗ (S ) ⊆ pred ∗ (s_0)
+\fi
 
@@
 We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
+Remark that in the construction, from any input with at least $1$ on its third counter, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
 % Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
 % or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
-Thus {\sc Home-space} is undecidable for Minsky machines.
-Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-space} is a particular instance of {\sc Resilience}.
+Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in a $3$-counter Minsky machine
+from any input with at least $1$ on its third counter is undecidable. 
 Hence {\sc Resilience} is undecidable for Minsky machines.

### 2023-08-23 introduction.tex
up
section

@@
 %			from all reachable states from $s_0$, it is always possible to reach $H$ 
-$\post^*(s_0) \subseteq \pred^*(H)$. If the set $H$ contains a single element, this element is an {\em home-state}.
+$\post^*(s_0) \subseteq \pred^*(H)$. If the home-space contains a single element, this element is an {\em home-state}.
 It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
@@
  reset-VASS $V_1$ from Figure~\ref{r-V}.
-Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
+Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, \alain{quelle est la logique des majuscules ?} { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
 Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$.
-However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of lenght less than $n$ towards $q_1(0)$.
+However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$.
 Consider now the %rightmost
- reset-VASS $V_2$, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$ compared to $V_1$. In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold. 
+ reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold. 
 Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the reset transition.

### 2023-08-23 introduction.tex
section

@@
 Consider now the %rightmost
- reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold. 
-Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the reset transition.
+ reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, { Bounded resilience} hold, since $6$-resilience hold. 
+Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.
 From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less.

### 2023-08-23 resilience-safe-up.tex
section

@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$. 
-It can be viewed as a generalizeation
+Resilience can be viewed as a generalizeation
  of coverability, as
-it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
+it asks whether for \emph{every} element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
 

### 2023-08-23 resilience-safe-down.tex
section

@@
 
-Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is \emph{downward compatible} if
+Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is \emph{downward compatible} \cite{DBLP:journals/tcs/FinkelS01} if
 for all $s_1, s_2, t_1 \in S$ with

### 2023-08-23 introduction.tex
up

@@
  reset-VASS $V_1$ from Figure~\ref{r-V}.
-Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, \alain{quelle est la logique des majuscules ?} { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
-Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$.
-However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$.
+Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {resilience}, 
+%\alain{quelle est la logique des majuscules ?} 
+{$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
+Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$.
+However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$.
 Consider now the %rightmost
- reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, { Bounded resilience} hold, since $6$-resilience hold. 
+ reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
 Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.

### 2023-08-23 resilience-safe-down.tex
section

@@
 From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
-such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
 We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.

### 2023-08-23 appendix.tex
section

@@
 with
-initial vector $\textbf{u}$ and initial state $q_0$.
+initial vector $\textbf{u}$ and initial \alain{control-}state $q_0$.\alain{incohérent: control-state !, notation $q_0(\textbf{u})$  à utiliser}
 To do so, we construct a reset-VASS $V_2$
-with initial vector $\textbf{v}$ and initial state $q'_0$
+with initial vector $\textbf{v}$ and initial state $q'_0$ \alain{idem}
 which is live if and only if 
@@
 The initial vector $\textbf{v}$ is constructed from $\textbf{u}$ by having at least $1$ on the first two additional vector coordinate.
+\alain{un dessin aiderait...}
 
@@
 % Note that all transitions of $T_1$ are in conflict and, by construction if not by definition, no more than one transition can fire at a time.
-
+\alain{donne des noms, numéros,...}
 These first two additional vector coordinate together with this additional transition allow
@@
 % Now we need to see if that marking is zero.
-
+\alain{structurer la preuve pour la rendre lisible, nommer les objets, faire des paragraphes,...}
 We introduce two new additional vector coordinate and $d$ new transitions which have
@@
 Now if $q(\textbf{0})$ is reachable from $q_0(\textbf{u})$ in $V_1$, then $V_2$ can also reach a similar
-vector $q(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transition.
+vector $q(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transitions.
 Then $s$ can freeze the $V_1$ subset,

### 2023-08-23 appendix.tex
section

@@
 deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-We describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri} reducing the problem to {\sc $t$-liveness}. \\
+We describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness}. \\
 
@@
 
+{\bf construction of $V_2$}
 The new reset-VASS is constructed by adding four new vector coordinate
@@
 We introduce two new additional vector coordinate and $d$ new transitions which have
-$-1$ on the $d$-th vector coordinate and $+1$ on this penultimate coordinate.
+$-1$ on the $d$-th vector coordinate and $+1$ on this penultimate coordinate. \\
 
@@
 
-If $q(\textbf{0})$ is reachable, then these $d$ new transitions are not live. Everything is dead if $s$ fires in $q(\textbf{0})$.
+Let us now prove that $V_2$ is live if and only if 
+$q(\textbf{0})$
+is reachable from $q_0(\textbf{u})$.
+
+{\bf Suppose that $q(\textbf{0})$ is reachable from $q_0(\textbf{u})$} in V1 or V2 ????\\
+Then these \alain{vague} $d$ new transitions are not live. Everything \alain{vague} is dead if $s$ fires in $q(\textbf{0})$.
 
@@
 
-Conversely, if $V_2$ is not live, then a vector is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.
+{\bf Suppose that $V_2$ is not live and prove that 
+$q(\textbf{0})$ is reachable in V12 ???}\\
+
+if $V_2$ is not live, then a vector \alain{vague} is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.
 % Conversely, if $C_2$ is not live, then a marking $\mu$ must be reachable in which  $\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token (in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.).
-If it is not possible to have a strictly positive integer in one of the last two coordinates, then the transitions are not live, and the
+If it is not possible to have a strictly positive integer in one of the last two coordinates, then the transitions \alain{vague} are not live, and the
 $d$-dimensional subset of the current vector must be zero.

### 2023-08-24 appendix.tex
section

@@
 
-We add an additional transition $s$ which reset the first two additional vector coordinate from control-state $q$.
+We add an additional transition $s$ \alain{s est d'habitude un état}
+which reset the first two additional vector coordinate from control-state $q$.
 % Transition $s$ is defined to have $r_1$ as an input and a null output.
@@
 $q(\textbf{0})$
-is reachable from $q_0(\textbf{u})$.
+is reachable from $q_0(\textbf{u})$ in $V_1$ ?????.
 
@@
 Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-This is accomplished by a transition $s'$ which arbitrarily increments every other vector coordinate 
+This is accomplished by a transition $s'$  \alain{s' est d'habitude un état}
+which arbitrarily increments every other vector coordinate 
 while decrementing the penultimate one and, additionally, a transition which 
@@
 
-Now if $q(\textbf{0})$ is reachable from $q_0(\textbf{u})$ in $V_1$, then $V_2$ can also reach a similar
-vector $q(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence of transitions.
-Then $s$ can freeze the $V_1$ subset,
+Now if $q(\textbf{0})$ is reachable from $q_0(\textbf{u})$ in $V_1$ by a sequence $\sigma$, then $V_2$ can also reach a similar
+vector $q(0,0,0,\ldots, 0, 1,0,0,0)$ by executing the same sequence $\sigma$.
+Then $s$ can freeze the $V_1$ subset, \alain{quel subset ?}
 % Now if $0$ is reachable in $R(C_1, \mu_1)$  then the net $C_2$ can also reach this marking in the place of $P_1$ by executing the same sequence of transition firings.
@@
 % Since $mu(p_i) = 0$ for all $p_i \in P_1$, no transition $s_{p_i}$ with $p_i \in P_1$ can fire, and $C_2$ is dead.
-Thus if $q(\textbf{0})$ is reachable then $V_2$ is not live.
+Thus if $q(\textbf{0})$ is reachable in ???? then $V_2$ is not live.
 
 {\bf Suppose that $V_2$ is not live and prove that 
-$q(\textbf{0})$ is reachable in V12 ???}\\
+$q(\textbf{0})$ is reachable in ???}\\
 
-if $V_2$ is not live, then a vector \alain{vague} is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.
+if $V_2$ is not live, then a vector \alain{vague} is reachable in which the last two coordinate are equal to $0$ and there is no reachable vector in which either is equal to at least one.\alain{pas compréhensible facilement}
 % Conversely, if $C_2$ is not live, then a marking $\mu$ must be reachable in which  $\mu(r_2) = 0$ and there is no reachable state in which $r_2$ has a token (in particular, since we do not allow token removal from $r_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $r_2$.).
 If it is not possible to have a strictly positive integer in one of the last two coordinates, then the transitions \alain{vague} are not live, and the
-$d$-dimensional subset of the current vector must be zero.
+$d$-dimensional subset \alain{subset ou vecteur ?}
+of the current vector must be zero.
 % If $r_2$ has no token and cannot get any, then the transitions $s_p$ are not live, and the markings of $p$ for $p \in P_1$ must be zero. 
@@
 Thus if $V_2$ is not live then 
-$q(\textbf{0})$ is reachable.
+$q(\textbf{0})$ is reachable in ???.
 \end{proof}

### 2023-08-28 appendix.tex
Acting on the comments on the appendix

@@
 with
-initial vector $\textbf{u}$ and initial \alain{control-}state $q_0$.\alain{incohérent: control-state !, notation $q_0(\textbf{u})$  à utiliser}
+initial configuration $q_0(\textbf{u})$.
 To do so, we construct a reset-VASS $V_2$
-with initial vector $\textbf{v}$ and initial state $q'_0$ \alain{idem}
+and an initial configuration $q'_0(\textbf{v})$ 
 which is live if and only if 
@@
 
-We add an additional transition $s$ \alain{s est d'habitude un état}
+We add an additional transition $t$ %\alain{s est d'habitude un état}
 which reset the first two additional vector coordinate from control-state $q$.
@@
 Transition $s'$ has $r_2$ as its input and every place of $C_2$ as output
-This is accomplished by a transition $s'$  \alain{s' est d'habitude un état}
+This is accomplished by a transition $t'$  %\alain{s' est d'habitude un état}
 which arbitrarily increments every other vector coordinate 

### 2023-08-29 appendix.tex
section

@@
 $q(\textbf{0})$
-is reachable from $q_0(\textbf{u})$.
+is reachable from $q_0(\textbf{u})$.\alain{in V2 ? préciser tt le temps dans quel vass}
 

### 2023-08-29 appendixPN.tex
Been writing an alternate version of the appendix dealing with reset PN rather than reset VASS. Wanted to try to see if reset VASS could have a similar level of clarity and compare boths but right now I am of the opinion that the Petri net version is much clearer and that VASS formalism is not really well ajusted to the proof methodology

\section{Reset-VASS and state-resilience}\label{appendix}

% \section{Proof of Theorem~\ref{liveness reset}}\label{appendix}

% Recall Theorem~\ref{liveness reset}.

%\begin{proposition}\label{liveness reset}
% The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
%\end{proposition}

Let us recall the undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
We describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness}. \\

\noindent
\textbf{Proposition~\ref{liveness reset}}
%\label{liveness reset}
{\em The  zero reachability problem can be reduced to {\sc $t$-liveness}.}

\begin{proof}

For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS. We recall that Petri nets and VASS are equivalents models~%\cite{qui?}
[??].

A reset Petri net $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ... r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.

If we wish to determine if a marking in which every place is empty is reachable for a reset Petri net
 $N_1 = (\Pi_1, \Sigma_1, E_1, R_1, \mu_1)$ % with initial marking $\mu_1$, 
 we construct a reset Petri net
 $N_2 = (\Pi_2, \Sigma_2, E_2, R_2, \mu_2)$ % with initial marking $\mu_2$
  which is live if and only if the {\em zero marking}, which assigns $0$ to every place of $\Pi_1$, is not reachable from $\mu_1$ in $N_1$. \\

\begin{samepage}
{\bf Construction of $N_2$} \\
\indent
The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_1$ and $p_2$ and $|\Pi_1| +2$ transitions $t_{p_1}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_2}$.
We first modify all transitions of $N_1$ to include $p_1$ as both an input and an output.
The initial marking $\mu_2$ will include a token in $p_1$ and no token in $p_2$. 
This new place $p_1$ serve to mark that the run is ongoing. As long as it contains a token, the adjusted transitions of $N_1$ are live and can be used normally. Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
We add an additional transition $t_{p_1}$ %\alain{s est d'habitude un état}
which has $p_1$ as an input and a null output.
This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $\Pi_1$ in $N_2$.
 The place $p_1$ and transition $t_{p_1}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_1}$ to fire and freeze the net at that marking.
We introduce a new place $p_2$ and new transitions $t_p$, for all $p \in \Pi_1$, which have $p$ as input and $p_2$ as output.
Lastly, we add a transition $t_{p_2}$ with $p_2$ as its output and every place of $\Pi_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
put in $p_2$.
\end{samepage}

 \begin{center}
	\begin{figure}
\includegraphics[width=0.75\textwidth]{FigurePN}
\caption{Construction of $N_2$ from $N_1$.}
	\end{figure}
\end{center}

Let us now prove that $t_2$ is live if and only if 
the zero marking
is not reachable from $\mu_1$ in $N_1$. \\

{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live}

Indeed the marking with zero in every place of $\Pi_1$ and in $p_2$ is reachable in $N_2$, by executing the same sequence of transition firings. Then $t_{p_1}$ can fire, leading to the marking which assigns zero to every place of $\Pi_2$. From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_1}$, and, finally, nor is $t_{p_2}$. Thus $t_{p_2}$ is not live. \\

{\bf Suppose that $t_{p_2}$ is not live, then the zero marking is reachable from $\mu_1$ in $N_1$}

Indeed, if $t_{p_2}$ is not live, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
This means that no transition $t_p$ is live in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
Thus, every place of $\Pi_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
it can be reached without using the transitions $t_p$ or $t_{p_2}$. Since $t_{p_1}$ do not modify the places inherited from $\Pi_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $\Pi_2$ inherited from $\Pi_1$ is devoid of token using only the transitions inherited from $N_1$. Thus the zero marking is reachable from $\mu_1$ in $N_1$.
\end{proof}

### 2023-08-30 appendixPN.tex
section

@@
 
-For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS. We recall that Petri nets and VASS are equivalents models~%\cite{qui?}
-[??].
+For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS.
+It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
 

### 2023-08-30 appendixPN.tex
section

@@
 
+Let us recall the {\em zero-reachability} problem in (reset-)VASS.
+
+\noindent
+input: a $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$ \\
+question:  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?
+
+The zero-reachability problem is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
+%%%%%
+\iffalse
+%%%%
 Let us recall the undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
@@
 deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-We describe therein how to adjust Theorem~5.5 from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness}. \\
+%%
+\fi
+%%%%
+
+For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS.
+It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
+The same equivalence hold between reset-VASS and reset Petri nets.
+
+We describe therein how to adjust Theorem~5.5 \alain{que dit le theorem 5.5 ?}from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets. \\
+
+A \emph{reset Petri net} $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ...
+r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
+{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
 
@@
 %\label{liveness reset}
-{\em The  zero reachability problem can be reduced to {\sc $t$-liveness}.}
+{\em In reset-Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
 
 \begin{proof}
-
-For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS.
-It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
-
-A reset Petri net $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ...
-r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
-{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
 
@@
 \end{proof}
+
+\alain{numérotation à faire}
+\noindent
+\textbf{Corollary~\ref{liveness reset VASS}}
+%\label{liveness reset}
+{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}

### 2023-08-30 appendixPN.tex
section

@@
 
-For the sake of clarity, the construction detailled here concerns itself with reset Petri nets rather than reset-VASS.
+For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets rather than reset-VASS.
+
+A \emph{reset Petri net} $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ...
+r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
+{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
+
 It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
@@
 
-A \emph{reset Petri net} $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ...
-r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
-{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
-
 \noindent
@@
 %\label{liveness reset}
-{\em In reset-Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
+{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
 

### 2023-08-30 appendixPN.tex
section

@@
 {\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
-
+\alain{pourquoi ces notations étranges ? note N=(P,T,...) comme pour les Petri normaux et pas r les places, et pas sigma les transitions,...}
 It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.

### 2023-08-30 appendixPN.tex
section
up

@@
 
-For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets rather than reset-VASS.
+For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS.
 
@@
 
-We describe therein how to adjust Theorem~5.5 \alain{que dit le theorem 5.5 ?}from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets. \\
+We describe therein how to adjust Theorem~5.5 \alain{que dit le theorem 5.5 ?}\mathieu{reachability se réduit à liveness pour les Petri nets}from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
+\\
 

### 2023-08-30 appendixPN.tex
up g

--- 2023-08-30 appendixPN.tex
@@
 
-For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets rather than reset-VASS.
+For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS.
 
-A \emph{reset Petri net} $N = (\Pi, \Sigma, E, R, \mu)$ consists of a finite set of {\em places} $\Pi = \{r_1, r_2, ...
-r_{|\Pi|}\}$, a finite set of {\em transitions} $\Sigma = \{t_1, t_2, ..., t_{|\Sigma|} \}$, a finite set of {\em arcs} $E \subseteq \Pi \times \Sigma \cup \Sigma \times \Pi$, a finite set of
-{\em reset-arcs} $R \subseteq \Sigma \times \Pi$, and an {\em initial marking} $\mu : \Pi \to \N$.
-
+A \emph{reset Petri net} $N = (P, T, E, R, \mu)$ consists of a finite set of {\em places} $P = \{p_1, p_2, ...
+p_{|P|}\}$, a finite set of {\em transitions} $T = \{t_1, t_2, ..., t_{|T|} \}$, a finite set of {\em arcs} $E \subseteq (P \times T) \cup (T \times P)$, a finite set of
+{\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
+% \alain{pourquoi ces notations étranges ? note N=(P,T,...) comme pour les Petri normaux et pas r les places, et pas sigma les transitions,...}
 It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
@@
 If we wish to determine if a marking in which every place is empty is reachable for a reset Petri net
- $N_1 = (\Pi_1, \Sigma_1, E_1, R_1, \mu_1)$ % with initial marking $\mu_1$, 
+ $N_1 = (P_1, T_1, E_1, R_1, \mu_1)$ % with initial marking $\mu_1$, 
  we construct a reset Petri net
- $N_2 = (\Pi_2, \Sigma_2, E_2, R_2, \mu_2)$ % with initial marking $\mu_2$
-  which is live if and only if the {\em zero marking}, which assigns $0$ to every place of $\Pi_1$, is not reachable from $\mu_1$ in $N_1$. \\
+ $N_2 = (P_2, T_2, E_2, R_2, \mu_2)$ % with initial marking $\mu_2$
+  which is live if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
 
@@
 \indent
-The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_1$ and $p_2$ and $|\Pi_1| +2$ transitions $t_{p_1}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_2}$.
+The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_1$ and $p_2$ and $|P_1| +2$ transitions $t_{p_1}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_2}$.
 We first modify all transitions of $N_1$ to include $p_1$ as both an input and an output.
@@
 which has $p_1$ as an input and a null output.
-This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $\Pi_1$ in $N_2$.
+This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $P_1$ in $N_2$.
  The place $p_1$ and transition $t_{p_1}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_1}$ to fire and freeze the net at that marking.
-We introduce a new place $p_2$ and new transitions $t_p$, for all $p \in \Pi_1$, which have $p$ as input and $p_2$ as output.
-Lastly, we add a transition $t_{p_2}$ with $p_2$ as its output and every place of $\Pi_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
+We introduce a new place $p_2$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_2$ as output.
+Lastly, we add a transition $t_{p_2}$ with $p_2$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
 put in $p_2$.
@@
 
-Let us now prove that $t_2$ is live if and only if 
+Let us now prove that $t_{p_2}$ is live if and only if 
 the zero marking
@@
 
-Indeed the marking with zero in every place of $\Pi_1$ and in $p_2$ is reachable in $N_2$, by executing the same sequence of transition firings.
-Then $t_{p_1}$ can fire, leading to the marking which assigns zero to every place of $\Pi_2$.
+Indeed the marking with zero in every place of $P_1$ and in $p_2$ is reachable in $N_2$, by executing the same sequence of transition firings.
+Then $t_{p_1}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
 From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_1}$, and, finally, nor is $t_{p_2}$.
@@
 This means that no transition $t_p$ is live in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
-Thus, every place of $\Pi_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
+Thus, every place of $P_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
 it can be reached without using the transitions $t_p$ or $t_{p_2}$.
-Since $t_{p_1}$ do not modify the places inherited from $\Pi_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $\Pi_2$ inherited from $\Pi_1$ is devoid of token using only the transitions inherited from $N_1$.
+Since $t_{p_1}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
 Thus the zero marking is reachable from $\mu_1$ in $N_1$.

### 2023-08-30 appendixPN.tex
up

@@
 \indent
-The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_1$ and $p_2$ and $|P_1| +2$ transitions $t_{p_1}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_2}$.
-We first modify all transitions of $N_1$ to include $p_1$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $p_1$ and no token in $p_2$. 
-This new place $p_1$ serve to mark that the run is ongoing.
+The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_a$ and $p_b$ and $|P_1| +2$ transitions $t_{p_a}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_b}$.
+We first modify all transitions of $N_1$ to include $p_a$ as both an input and an output.
+The initial marking $\mu_2$ will include a token in $p_a$ and no token in $p_b$. 
+This new place $p_a$ serve to mark that the run is ongoing.
 As long as it contains a token, the adjusted transitions of $N_1$ are live and can be used normally.
 Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
-We add an additional transition $t_{p_1}$ %\alain{s est d'habitude un état}
-which has $p_1$ as an input and a null output.
+We add an additional transition $t_{p_a}$ %\alain{s est d'habitude un état}
+which has $p_a$ as an input and a null output.
 This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $P_1$ in $N_2$.
- The place $p_1$ and transition $t_{p_1}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_1}$ to fire and freeze the net at that marking.
-We introduce a new place $p_2$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_2$ as output.
-Lastly, we add a transition $t_{p_2}$ with $p_2$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
+ The place $p_a$ and transition $t_{p_a}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_a}$ to fire and freeze the net at that marking.
+We introduce a new place $p_b$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_b$ as output.
+Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
 put in $p_2$.
@@
 
-Indeed the marking with zero in every place of $P_1$ and in $p_2$ is reachable in $N_2$, by executing the same sequence of transition firings.
-Then $t_{p_1}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
-From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_1}$, and, finally, nor is $t_{p_2}$.
+Indeed the marking with zero in every place of $P_1$ and in $p_b$ is reachable in $N_2$, by executing the same sequence of transition firings.
+Then $t_{p_a}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
+From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
 Thus $t_{p_2}$ is not live.
@@
 it can be reached without using the transitions $t_p$ or $t_{p_2}$.
-Since $t_{p_1}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
+Since $t_{p_a}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
 Thus the zero marking is reachable from $\mu_1$ in $N_1$.

### 2023-08-30 appendixPN.tex
up

@@
 
-For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS.
+For the sake of clarity \alain{bof, une autre raison?}\mathieu{for better visualisation? better representation?}, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS.
 

### 2023-08-30 appendixPN.tex
section

--- 2023-08-30 appendixPN.tex
@@
 input: a $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$ \\
-question:  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?
+question:  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\
 
@@
 %%%%
-
-For the sake of clarity \alain{bof, une autre raison?}, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS.
-
+%
+For a better representation, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS. 
 A \emph{reset Petri net} $N = (P, T, E, R, \mu)$ consists of a finite set of {\em places} $P = \{p_1, p_2, ...
@@
 
-We describe therein how to adjust Theorem~5.5 \alain{que dit le theorem 5.5 ?}\mathieu{reachability se réduit à liveness pour les Petri nets}from \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
+We describe therein how to "adjust" Theorem~5.5 in  \cite{peterson1981petri} that shows that reachability can be reduced to liveness for Petri nets \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
 \\

### 2023-08-30 appendixPN.tex
up
section

@@
 
-\noindent
-input: a $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$ \\
-question:  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\
+\problemx{zero-reachability}
+{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
 
@@
 
-We describe therein how to "adjust" Theorem~5.5 in  \cite{peterson1981petri} that shows that reachability can be reduced to liveness for Petri nets \cite{peterson1981petri} reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
+We describe therein how to "adjust" Theorem~5.5 in  \cite{peterson1981petri}, that shows that reachability can be reduced to liveness for Petri nets \cite{peterson1981petri}, reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
 \\
@@
 
-Let us now prove that $t_{p_2}$ is live if and only if 
+Let us now prove that $t_{p_2}$ is live in $N_2$ if and only if 
 the zero marking
@@
 
-{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live}
+{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live} in $N_2$.
 
@@
 From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
-Thus $t_{p_2}$ is not live.
+Thus $t_{p_2}$ is not live \alain{dire tt le temps dans quel PN in $N_2$}.
 \\
 
-{\bf Suppose that $t_{p_2}$ is not live, then the zero marking is reachable from $\mu_1$ in $N_1$}
+{\bf Suppose that $t_{p_2}$ is not live in $N_2$, then the zero marking is reachable from $\mu_1$ in $N_1$}
 
-Indeed, if $t_{p_2}$ is not live, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
+Indeed, if $t_{p_2}$ is not live \alain{in $N_2$}, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
 This means that no transition $t_p$ is live in $\mu$ since any transition $t_p$ can place a token in $p_2$. 

### 2023-08-30 appendixPN.tex
section

@@
 {A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
-{  $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
 
@@
 
-We describe therein how to "adjust" Theorem~5.5 in  \cite{peterson1981petri}, that shows that reachability can be reduced to liveness for Petri nets \cite{peterson1981petri}, reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
-\\
+We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
+
+%that shows that reachability can be reduced to liveness for Petri nets \cite{peterson1981petri}, 
+%reducing the {\em zero-reachability} problem to {\sc $t$-liveness} in reset Petri nets.
 

### 2023-08-31 appendixPN.tex
section
up

@@
 {\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
-% \alain{pourquoi ces notations étranges ? note N=(P,T,...) comme pour les Petri normaux et pas r les places, et pas sigma les transitions,...}
+\alain{comment sont représentés, sur la figure 2, les arcs resets ?}
 It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
@@
  $N_2 = (P_2, T_2, E_2, R_2, \mu_2)$ % with initial marking $\mu_2$
-  which is live if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
+  which is live in $N_2$ if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
 
@@
 Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
-We add an additional transition $t_{p_a}$ %\alain{s est d'habitude un état}
+We add an additional transition $t_{p_a}$ 
 which has $p_a$ as an input and a null output.
@@
 We introduce a new place $p_b$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_b$ as output.
-Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live if a token is ever
+Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live in $N_2$ if a token is ever
 put in $p_2$.
@@
 
-{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live} in $N_2$.
+{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live in $N_2$}.
 
@@
 From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
-Thus $t_{p_2}$ is not live \alain{dire tt le temps dans quel PN in $N_2$}.
+Thus $t_{p_2}$ is not live in $N_2$.
 \\
@@
 
-Indeed, if $t_{p_2}$ is not live \alain{in $N_2$}, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
-This means that no transition $t_p$ is live in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
+Indeed, if $t_{p_2}$ is not live in $N_2$, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
+This means that no transition $t_p$ is live in $N_2$ in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
 Thus, every place of $P_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 

### 2023-08-31 appendixPN.tex
up

@@
 \section{Reset-VASS and state-resilience}\label{appendix}
+
+\newcommand{\arrowcircle}[1][]{%
+  \begin{tikzpicture}[#1]
+    \draw[->] (0,0ex) -- (2em,0ex);
+    \draw (1em,0ex) circle (0.7ex);
+  \end{tikzpicture}%
+}
 
@@
 {\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
-\alain{comment sont représentés, sur la figure 2, les arcs resets ?}
 It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
@@
 \includegraphics[width=0.75\textwidth]{FigurePN}
-\caption{Construction of $N_2$ from $N_1$.}
+\caption{Construction of $N_2$ from $N_1$.
+Places are represented using circles, transitions are represented using lines.
+Arcs are represented by $\rightarrow$ and reset-arcs by $\arrowcircle$.
+Note that since the construction of $N_2$ from $N_1$ do not add new reset-arcs, the only reset-arcs in $N_2$ are the ones inherited from $N_1$.}
 	\end{figure}

### 2023-08-31 final.tex
up

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%

\bibliographystyle{splncs04}% the mandatory bibstyle
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{mathtools}
\usepackage{todonotes}
\usepackage{microtype}

\usepackage{complexity}
\usepackage{amsmath}

\usepackage{stmaryrd}
\usepackage{dsfont}

\usepackage{mathrsfs}
\usepackage{mathalpha}
\usepackage{amsmath}
\usepackage{amsfonts}

\usepackage{ textcomp }

\usepackage{stmaryrd}
\usepackage{wrapfig}

%

\newcommand{\problemx}[3]{
	\vspace{0.2cm}
\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
\begingroup\clubpenalty10000\widowpenalty10000
\setbox0\hbox{\bf INPUT:\ }\setbox1\hbox{\bf QUESTION:\ }
\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
\vskip-\parskip\noindent
\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
\vskip-\parskip\noindent
\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
\endgroup
	\vspace{-0.2cm}
}

\newcounter{claimcounter}
\setcounter{claimcounter}{0}
\newtheorem{subclaim}{Subclaim}{}
\newtheorem{fact}{Fact}{}

\makeatletter

\renewcommand{\poly}{\mathrm{poly}}

\newcommand{\alain}[1]{\todo[inline,color=red!20]{{\bf AF:} #1}}
\newcommand{\mathieu}[1]{\todo[inline,color=blue!20]{{\bf MH:} #1}}

\newcommand{\pred}{\textsf{pred}}
\newcommand{\post}{\textsf{post}}

\newcommand{\Bad}{\textsf{Bad}}
\newcommand{\Safe}{\textsf{Safe}}

\begin{document}
%
\title{Resilience and Home-Space for WSTS \thanks{This work was partly done while the author was supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}} 
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alain Finkel\inst{1,2,3} \and Mathieu Hilaire\inst{1,3}}
% This research has been supported by ANR programme BraVAS (ANR-17-CE40-0028).
%
\authorrunning{A. Finkel, M. Hilaire.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Université Paris-Saclay, CNRS, ENS Paris-Saclay, LMF, Gif-sur-Yvette, France %\email{hilaire@lsv.fr}
\and
Institut Universitaire de France \and Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028)
}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
\noindent
Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
We consider six types of resilience (one of them is the home-space property) defined by an upward closed set Safe  ($\Safe=\uparrow \Safe$) or downward-closed set ($\Safe=\downarrow \Safe$) (Bad is generally the complementary of Safe) and by the existence of a bound $k$ on the length of minimal runs starting from Bad and reaching Safe. We study the decidability of each type of resilience for two kinds of models: WSTS and VASS. We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis). Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$. Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective 
$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$. Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for variations of VASS (lossy counter machines, integer VASS and continuous VASS); most of the resilience properties are shown decidable.
\end{abstract}

\keywords{Verification, Resilience, Home-Space, Well-structured transition systems, Vector addition system with states}

\newcommand{\LCM}{\mathsf{LCM}}
\newcommand{\LOGSPACE}{\mathsf{LOGSPACE}}
\newcommand{\MSO}{\mathsf{MSO}}
\newcommand{\SO}{\mathsf{SO}}

 \newcommand{\N}{\mathds{N}}

\section{Introduction}\label{section introduction}

{\bf Context.} 
Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety~\cite{challenges}. Some systems are subjects at frequent intervals to accidents, attacks or changes. Think for instance of a supply chain, or an airport’s air trafic control. In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack
pushed it towards some kind of ‘error state’ and, if it can, whether it can perform the return in a satisfactory timeframe.

{\bf Home-spaces}
 In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if 
$\post^*(s_0) \subseteq \pred^*(H)$. If the home-space contains a single element, this element is an {\em home-state}.
It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$. In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report). In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
%%%%%

{\bf Resiliences.}  
The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and a subset of states $\Safe \subseteq S$ consists of the following: $\mathscr{S}$ is {\em $\Safe$-resilient} if
$S \subseteq \pred^*(\Safe)$. It 
can be stated as an home-space problem : $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
$\post^*(s_0) \subseteq \pred^*(\Safe)$.
We will study three decidability questions for each of type of resilience.
The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient (for a given subset of states $\Safe \subseteq S$).
The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
whether $\Bad \subseteq \pred^*(\Safe)$.
The $k$-resilience problem is to decide whether $S \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
the bounded resilience problem is to decide whether there exists an $k$ such that $\mathscr{S}$ is $\Safe$-$k$-resilient. We adapt these three definitions to state-resilience.

{\bf State of the art.}
In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} interesting definitions and results (without detailed proofs) about resilience in the framework of process algebra. They built the composition of the process and its adversary as a transition system and they gave conditions that insure that the composed transition system is an effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions). In this framework and under these hypotheses, resilience reduces to coverability which is decidable on WSTS. 
%%%%%%%%%

In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective 
$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$ and $\Bad$ downward-closed.

Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
However, neither Prasad \& Zuck nor Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
%%%%%%%%%
%

\noindent
{\bf Our contributions}
\begin{itemize}
\item Surprinsingly, the general undecidability statements about resilience were not known neither proved. We show that resilience and { State-resilience} problems are both undecidable for WSTS with strong compatibility. 
Moreover, {State-resilience},
{Bounded-state-resilience} and
{$k$-state-resilience}
are undecidable for strongly upward-compatible WSTS with effective pred-basis
when
$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {\ State-resilience} in reset-VASS.

\item The three resilience problems are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.

\item The resilience problem is decidable for ideal-effective WSTS with 
$\Safe=\downarrow \Safe$
and
the additional hypothesis that
for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
%
%%%

\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis.
We also show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({ $k$-state-resilience} and { bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
%

\item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
\end{itemize}

 \begin{center}
	\begin{figure}
			\hspace{0.8cm}
\includegraphics[width=0.75\textwidth]{FigureCD}
	\caption{Two reset-VASS with three control-states and one counter.}
					\label{r-V}
	\end{figure}
\end{center}

\begin{example}\label{Example}
{(\bf Reset-VASS example).}
Consider the 
 reset-VASS $V_1$ from Figure~\ref{r-V}.
Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {resilience}, 
{$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$. Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$. However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$. Consider now the 
 reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition. From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less. From 
$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps. 
\end{example}

Hence we can see how adding transitions can 
turn a system that is not bounded resilient into one that is.

\section{Well-structured transition systems and VASS}\label{section definitions}

\noindent
 A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
 {\em states} and  
 $ {\rightarrow} \subseteq S \times S$ is a
 binary relation 
 on
 the set of states, denoted as the set of {\em transitions}. 
%
We write $s \rightarrow s'$ to denote $ (s,s') \in  {\rightarrow} $.
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$; we denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
 $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
 $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
By iterating $\pred$ and $\post$ we obtain  
$\post^n(s) = \{ s' \in S \mid  ~ s \xrightarrow{}^n s'\}$
and
$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
However, we are generally more interested in
$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$, $\post^*(s)= \bigcup_{1 \leq i} \post^i(s)$
and
$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. 
The {\em reachability problem} asks, given a transition system $\mathscr{S} = (S, \to)$, two states $s, t \in S$, whether $s \to^* t$.

%

A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
To any subset $A \subseteq X$, we may associate
its {\em upward-closure},
 $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
 and its 
 {\em downward-closure},
 $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
as $\uparrow x$ (resp. $\downarrow x$).
%
A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$.

 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
%
%
%
Wqo admits many other equivalent formulations. 
%
%
As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
%
Quasi-orderings that have no 
infinite subset of mutually incomparable elements (antichains)
 enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
%
%
In what follows, a downward-closed set $D$ is represented by its finite set of ideals (or by the minimal elements of its upward-closed complement), and an upward-closed set $U$ is represented by its finite set of minimal elements. \\

\noindent
Let us now recall the (most general) definition of well-structured transition systems.
\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

We say that a WSTS $\mathscr{S}$ has \emph{strong (upward) compatibility} when moreover for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow t_2$.

Several families of formal models of processes give rise to WSTSs in a natural way, e.g. Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.

\noindent
{\bf On effectivity}
Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.
We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x$, with $x \in X_i $. 
We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
%
We say that $\mathscr{S}$
is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
%

Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. 
With the effective pred-basis hypothesis, we obtain:

\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$. Hence coverability is decidable.
\end{theorem}

With the ideal-effective hypothesis, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.

\begin{theorem}\cite{BFM-ic17}
Coverability is decidable for any ideal-effective ordered upward-compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
\end{theorem}

Let us recall the  definition of vector addition system with states. 
 \begin{definition} 
A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}. 
 \end{definition} 
%
Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with a $d$-VASS $V$.
For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
%
When in the context of a $d$-VASS, we denote $0^d$ by $\textbf{0}$.

A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence it can be only defined by $T$.

\input{resilience}

\input{state-resilience}

\input{application}

 	\input{conclusion}

\bibliography{bibliography/bib}

\appendix

\input{appendixPN}

\end{document}

### 2023-08-31 final.tex
up

@@
 
-\input{resilience}
+\section{Resilience for WSTS}
+
+In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
+The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe)$ is to decide whether from \emph{any} state in 
+$S$, \emph{there exists} a path (resp.
+a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
+Resilience is then the Home-Space problem (defined in the introduction) for the set $\Safe$.
+We use the notation
+$S \longrightarrow^{*} \Safe$ (resp. $S \longrightarrow^{\leq k} \Safe$)
+ for 
+$\forall x \in S, \exists y \in \Safe$ 
+ such that $x \longrightarrow^{*} y$ 
+  (resp.  $\forall x \in S, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
+  In our framework, $\Safe \subseteq S$  is possibly infinite but  must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
+
+Let us formalize three resilience problems.
+
+\problemx{Resilience (RP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
+{$S \longrightarrow^{*} \Safe$ ?\newline}
+
+\problemx{$k$-resilience (kRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a sets $\Safe \subseteq S$.}
+{$S \longrightarrow^{\leq k} \Safe$ ?\newline}
+
+\problemx{Bounded resilience (BRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
+{$\exists k \geq 0$ such that {$S \longrightarrow^{\leq k} \Safe$ ?\newline}}
+
+These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
+So we restrict our framework to the class of infinite-state WSTS.
+Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
+. Since $\Safe \subseteq \pred^*(\Safe)$, one only need to decide whether 
+the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.
+
+{\bf Related problems.} 
+Resilience and the home-space problem are also linked to the 
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+is decidable for Lossy Counter Machine (LCM)
+when $X$ and $Y$ are semi-linear sets.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
+
+\input{resilience-safe-up}
+
+\input{resilience-safe-down}
+
+{\bf Synthesis of the main decidability results}\label{synthesis}
+\begin{center}
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \ %& $\uparrow$~ $\uparrow$~ 
+		& $\Safe = {\uparrow \Safe}$~ 
+		 & $\Safe = {\downarrow \Safe}$~ 
+		 %& $\downarrow$~ $\downarrow$~
+ \\ \hline
+   RP %& Decidable (Thm~\ref{up-up}) 
+   	& Decidable (Thm~\ref{down-up})  
+   		 & Decidable (Thm~\ref{up-down}) 
+   	%	 & Undecidable (Thm~\ref{down-down})
+    \\ \hline
+   BRP %& Decidable (Corollary~\ref{B-up-up}) 
+   &  Decidable (Thm~\ref{down-up}) 
+   		 & ?
+   	%	 & ??
+    \\ \hline
+      kRP %& Decidable (Thm~\ref{k-up-up}) 
+      & Decidable (Thm~\ref{down-up}) 
+      		& ? 
+      	%	& ??
+       \\ \hline
+ \end{tabular}
+\end{center}
+
+%%%%%%%
+\iffalse
+%%%%%%%%%%%%
+\subsection{Computing the maximal resilient subsystem}
+
+We now briefly investigate the question to control a system such that its controlled behavior may always reach $\Safe$.
+%
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $\Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{\Safe}$. 
+%
+%		\begin{definition}
+We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient if $S  \xrightarrow{*} \Safe$ is satisfied.  
+%		\end{definition}
+%
+We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(\Safe)$-resilient.
+
+The \emph{maximal $\Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{\Safe}=(S,\rightarrow_{\Safe},\leq)$ where $\rightarrow_{\Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{\Safe}$ is $\Safe$-resilient.
+%\alain{vérifier que la plus grande existe}
+Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
+Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
+Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
+%
+For example, VASS and most of concrete WSTS allow such guards.
+  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
+
+\begin{theorem}{}
+Given an effective WSTS with effective pred-basis and an upward-closed set $\Safe$, the maximal resilient subsystem is a computable WSTS.
+\end{theorem}
+
+\begin{proof}
+Since $\Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(\Safe)=\{x_1,x_2,...,x_p\}$.
+%
+%
+Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(\Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
+		\end{proof}
+%
+%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
+
+%%%%%%%
+\fi
+%%%%%%%%%
 

### 2023-08-31 final.tex
up

@@
 
-\input{resilience-safe-up}
+\subsection{Case: $\Safe=\uparrow \Safe$.}\label{safe-up}
 
-\input{resilience-safe-down}
+We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$. 
+Resilience can be viewed as a generalizeation
+ of coverability, as
+it asks whether for \emph{every} element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
+
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
+$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
+A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
+it is post-effective, and there exists a Turing Machine $M_\downarrow$ that computes, on input $s \in S$, 
+ $\downarrow s$, and some Turing Machine $M_{\uparrow^C}$ that computes, on input
+ $\{s_1, s_2, \ldots, s_m\} \in S^m$, 
+ the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$.
+Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).
+
+Let us recall two other results in \cite{BFM-ic17}.
+Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+%
+
+%
+\begin{theorem}\label{down-up}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+\begin{proof}
+Let  $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+
+Let us prove that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+%
+Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
+From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $\downarrow b_j$ is covered from $\downarrow x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
+%	for every $y \in J$, 
+there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
+
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow b_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
+%
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
+Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+%
+
+%
+We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
+
+This implies that resilience and bounded resilience are equivalent to coverability.
+Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we deduce that both the 
+  resilience problem and the bounded resilience problem are decidable.
+
+Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
+If $\downarrow b_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow b_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$. \qed
+\end{proof}
+
+Remark that the previous proof don't make use of the 
+property that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
+Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
+increasing recursive functions from \cite{FMP-ic04} defined as
+
+$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+
+	\noindent	
+	where $TM_j$ is the $j$-th Turing machine (in a classical enumeration)
+	which moreover begins
+by writing the integer $j$ on its tape,	
+	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+	The
+	transition system 
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+	is a WSTS
+	and
+	has the property that
+	it is
+	$\uparrow (1,1)$-resilient
+	iff $TM_j$ halts on input $j$,
+	since $(1,1)$ is coverable from $(0,0)$
+	iff $TM_j$ halts on input $j$. 
+	Hence there is no Turing machine which correctly determines 
+	resilience and halts whenever its input is a
+	WSTS.
+Hence resilience for WSTS in general is undecidable.
+
+\subsection{Case: $\Safe=\downarrow \Safe$.}\label{safe-down}
+
+Let us now consider the case $\Safe=\downarrow \Safe$ hence $\Bad=\uparrow \Bad$.
+It is of interest to note this case can be linked to the problem of mutual exclusion.
+Indeed the well-known mutual exclusion property can be modelized, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
+
+%
+\begin{theorem}\label{up-down}
+{\sc Resilience}  is decidable for ideal-effective WSTS with 
+$\Safe=\downarrow \Safe$
+and
+the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+\end{theorem}
+
+\begin{proof}
+By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
+The resilience problem can be reformulated as 
+$\Bad \subseteq  \pred^*(\Safe)$.
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of upward- or downward-closed 
+ subsets.
+Hence we can compute the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both upward-closed.
+Since
+$\Bad \subseteq \pred^*(\Safe)$
+can be reformulated as
+$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
+the resilience problem is decidable. \qed
+\end{proof}
+
+Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is \emph{downward compatible} \cite{DBLP:journals/tcs/FinkelS01} if
+for all $s_1, s_2, t_1 \in S$ with
+$s_2 \leq s_1$ and $s_1 \to t_1$
+there
+exists $t_2 \in S$ with
+$t_2 \leq t_1$ and $s_2 \to^* t_2$.
+
+\begin{corollary}
+{\sc Resilience} is decidable for ideal-effective downward-compatible WSTS with 
+$\Safe=\downarrow \Safe$.
+\end{corollary}
+
+\begin{proof}
+
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \downarrow \pred^*(D)$.
+By downward closure, there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$.
+By definition of $\pred^*(D)$ then there exists 
+$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+
+By downward compatibility $a_0 \to a_1$
+implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
+$x \to^* a'_1$.
+More generally $a_i \to a_{i+1}$ and
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
+$a'_i \to^* a'_{i+1}$,
+and, by induction,
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+with $d' \leq d$.
+Since
+$d'$ 
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$. \qed
+\end{proof}
+
+In the case
+of a ideal-effective WSTS 
+where
+the additional hypothesis that
+for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed
+is not met,
+the above construction
+can provide a proof
+of non-resilience
+i.e. when
+$\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) \neq \emptyset$
+then
+$\Bad \not\subseteq \downarrow\pred^*(\Safe)$
+and hence
+$\Bad \not\subseteq \pred^*(\Safe)$.
+When $\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) = \emptyset$
+however
+it is not enough to conclude.
+
+In the case
+of a ideal-effective WSTS 
+where the hypothesis that
+{for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}
+is not met, the problem is undecidable.
+ Indeed, it is 
+undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$.
+This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.
+
+Home-space is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
+This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61,Min67}.
+From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
+We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
+Remark that in the construction, from any input with at least $1$ on its third counter, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
+Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in a $3$-counter Minsky machine
+from any input with at least $1$ on its third counter is undecidable. 
+Hence {\sc Resilience} is undecidable for Minsky machines.
+
+Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
+Reset-VASS extend the basic VASS model with special “reset
+transitions” that set to $0$ some coordinates in the vector. Let us recall their definition here.
+\begin{definition}
+A {\em reset-VASS} in dimension $d$ 
+ is a finite 
+labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, 
+$T \subseteq Q \times Op \times Q$
+is the set of {\em control-transitions}, and $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
+\end{definition}
+
+Again $Q \times \N^d$
+ denotes the set of configurations of $V$.
+For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
+\begin{samepage}\begin{itemize}
+\item  $t = (q,add(\textbf{z}),q') \in T$
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
+\item $t = (q,reset(\gamma),q') \in T$ 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+\end{itemize} \end{samepage}
+
+It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
+Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
+
+Remark we did not make use of the
+property $\Bad$ complement of $\Safe$, simply 
+$\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complement of each others.
 
@@
 \begin{tabular}{ | l | c | c | c | r |}
-\hline   \ %& $\uparrow$~ $\uparrow$~ 
+\hline   \ 
 		& $\Safe = {\uparrow \Safe}$~ 
 		 & $\Safe = {\downarrow \Safe}$~ 
-		 %& $\downarrow$~ $\downarrow$~
  \\ \hline
-   RP %& Decidable (Thm~\ref{up-up}) 
+   RP 
    	& Decidable (Thm~\ref{down-up})  
    		 & Decidable (Thm~\ref{up-down}) 
-   	%	 & Undecidable (Thm~\ref{down-down})
     \\ \hline
-   BRP %& Decidable (Corollary~\ref{B-up-up}) 
+   BRP 
    &  Decidable (Thm~\ref{down-up}) 
    		 & ?
-   	%	 & ??
     \\ \hline
-      kRP %& Decidable (Thm~\ref{k-up-up}) 
+      kRP 
       & Decidable (Thm~\ref{down-up}) 
       		& ? 
-      	%	& ??
        \\ \hline
@@
 \end{center}
-
-%%%%%%%
-\iffalse
-%%%%%%%%%%%%
-\subsection{Computing the maximal resilient subsystem}
-
-We now briefly investigate the question to control a system such that its controlled behavior may always reach $\Safe$.
-%
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be WSTS and $\Safe \subseteq S$ be an upward-closed set given by its finite minimal basis $B_{\Safe}$. 
-%
-%		\begin{definition}
-We say that $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient if $S  \xrightarrow{*} \Safe$ is satisfied.  
-%		\end{definition}
-%
-We may verify that  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\Safe$-resilient iff  $\mathscr{S}=(S, \rightarrow, \leq)$ is $\pred^*(\Safe)$-resilient.
-
-The \emph{maximal $\Safe$-resilient subsystem} of $\mathscr{S}=(S,\rightarrow,\leq)$ is defined as $\mathscr{S}_{\Safe}=(S,\rightarrow_{\Safe},\leq)$ where $\rightarrow_{\Safe}$ is the largest restriction of $\rightarrow$ that insures that $\mathscr{S}_{\Safe}$ is $\Safe$-resilient.
-%\alain{vérifier que la plus grande existe}
-Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$.
-Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$.
-Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq)$ are still WSTS.
-%
-For example, VASS and most of concrete WSTS allow such guards.
-  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
-
-\begin{theorem}{}
-Given an effective WSTS with effective pred-basis and an upward-closed set $\Safe$, the maximal resilient subsystem is a computable WSTS.
-\end{theorem}
-
-\begin{proof}
-Since $\Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(\Safe)=\{x_1,x_2,...,x_p\}$.
-%
-%
-Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(\Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
-		\end{proof}
-%
-%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
-
-%%%%%%%
-\fi
-%%%%%%%%%
 

### 2023-08-31 final.tex
up

@@
 
-In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Okzan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
+In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
 $\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$ and $\Bad$ downward-closed.
@@
 
-\input{state-resilience}
+\section{State-resilience}
+
+Resilience is a strong property that implies that from every element 
+ there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from 
+ $ \post^*(s_0)$, there must exist a path to $\Safe$. 
+The three previous problems become:
+
+\problemx{State-resilience (SRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
+{$\post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
+
+\problemx{$k$-state-resilience (kSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
+{ $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+\problemx{bounded-state-resilience (BSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
+{$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+Remark that {\sc State-resilience} is {\sc Home-space} with input set $\Safe$.
+Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
+As in the Section~3, we study decidability results for $\Safe$ downward-closed and upward-closed.
+
+\subsection{Case: $\Safe = \uparrow \Safe$}
+
+We start with the case $\Safe = \uparrow \Safe$.
+Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
+This stems from the fact that it is undecidable in the particular case of reset-VASS,
+where  $t$-liveness is both undecidable and 
+ reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+We now recall the
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
+which we then reduce to {\sc State-resilience}.
+A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
+ $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
+live. This leads to the following problem.
+
+\problemx{$t$-liveness}
+{A reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{Is $t$ live in $s_0$ ? \\}
+
+To better discuss $t$-liveness let us define more formally the upward-closed set of configurations such that $t$ is enabled
+$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+
+\begin{proposition}\label{reductions}
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness 
+in a 
+reset-VASS $(Q,T)$ 
+with initial configuration $s_0$
+as the following formula
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
+The problem reduces itself to {\sc State-resilience}
+where $\Safe = pre(t)$.
+\end{proof}
+
+The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS, whose proof can be seen in Appendix~\ref{appendix}%
+.
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to {\sc $t$-liveness}.
+\end{proposition}
+
+Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
+{\sc $t$-liveness} is undecidable.
+
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
+Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
+
+\begin{theorem}\label{srp up down}
+{\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for strongly compatible WSTS with effective pred-basis
+when
+$\Safe=\uparrow \Safe$.
+\end{theorem}
+
+\begin{proof}
+{\sc State-resilience} itself is undecidable 
+since 
+{\sc $t$-liveness} is undecidable in reset-VASS
+and reducible to {\sc State-resilience}.
+Additionally, in WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
+reducible to {\sc $k$-state-resilience}:
+since $\Safe=\uparrow \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
+compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+We compute 
+$n_0$, then iteratively check whether $k$-state-resilience 
+hold for $k$ from $0$ to $n_0$.  
+Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+since 
+$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
+Hence the undecidability of {\sc Bounded-state-resilience}
+and 
+ {\sc $k$-state-resilience}. \qed
+\end{proof}
+
+On the positive side, let us recall a result about {\sc Bounded-state-resilience} decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+when
+$\Safe=\uparrow \Safe$.
+\end{theorem}
+
+We may immediately generalyse this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
+
+\begin{corollary}\label{postcomputable}
+{\sc State-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+when
+$\Safe=\uparrow \Safe$.
+\end{corollary}
+
+\begin{proof}
+Since $\mathscr{S}$ is a WSTS there exists $n_0 \in \N$ such that
+$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute this $n_0$ by iteratively computing 
+$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
+$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
+returning $n$ if that is the case.
+Then, because {\sc $n_0$-state-resilience} is decidable, 
+checking $\uparrow \post^*(s) 
+ \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+and {\sc State-resilience} is decidable. \qed
+\end{proof}
+
+The proof of Theorem~\ref{ref ozkan} rely on the computability of $\uparrow \post^*(s)$ and on the following lemma.
+
+\begin{lemma}\label{Lemma intersection}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
+Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
+Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\uparrow A}$, the inclusion
+$({\uparrow  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$. \qed
+\end{proof}
+
+The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
+
+\problemx{Downward-reachability problem}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
+\subseteq S$.}
+{$s  \to^* D$? \newline}
+
+%\"Ozkan
+\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
+For finite-branching WSTS, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+\end{proposition}
+
+The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
+$B_{\uparrow \post^*(s)} \cap D = \emptyset$, 
+where $B_{\uparrow \post^*(s)}$ is a basis of $\uparrow \post^*(s)$,
+ that is equivalent to $\post^*(s)\cap D = \emptyset$ by
+Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
+$U_n = \uparrow \post^{\leq n}(s)$ until it becomes stationnary. 
+Decidability of downward-reachability leads to the decidability of the following stop condition:
+asking whether $S \setminus U_n$ is reachable from $s$.
+
+More concretely,
+VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
+It is well-known that 
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
+Hence {\sc State-resilience} is decidable for  
+VASS.
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  \iffalse Moreover, we may show:
+
+\begin{proposition}
+There exist classes of WSTSs with strong 
+ compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$.
+\end{proposition}
+
+\begin{proof}
+Let us show that reset-VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
+Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for reset-VASS. 
+Then, {\sc State-resilience} would be decidable for reset-VASS, which leads to a contradiction.
+\end{proof} \fi
+Moreover there exist classes of WSTSs with strong 
+ compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$, such as reset-VASS.
+
+Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
+
+\begin{theorem}\label{post srp}
+{\sc State-resilience} is decidable for 
+ WSTS with effective 
+$\uparrow$ $\post^*$ basis
+when
+$\Safe=\uparrow \Safe$. 
+\end{theorem}
+
+\begin{proof}
+Let $B_{\uparrow \post^*(s)}$ 
+%\alain{notation fluctuante}
+ be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$.
+By applying Lemma~\ref{Lemma intersection} twice, we obtain
+\[ \post^*(s) %\cap \Bad
+ \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} %\cap \Bad 
+ \subseteq \pred^*(\Safe)\]
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} %\cap \Bad
+ \setminus \pred^*(\Safe) = \emptyset$.  \qed
+\end{proof}
+
+However when removing strong compatibility, precision is lost.
+Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
+ $\uparrow \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
+despite having 
+$\post^* (s) \cap S \subseteq \pred( \Safe)$.
+In such a case the algorithm from
+\cite{DBLP:conf/gg/Ozkan22} would deduce that {\sc $1$-state-resilience} does not hold,
+which is incorrect.
+
+Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
+{\sc Bounded-state-resilience}.
+
+{\bf Results synthesis in the case $\Safe = \uparrow \Safe$}
+\begin{center}
+\begin{tabular}{ | l | c | c | c | c |}
+\hline  Hypothesis & strong %upward 
+			compatibility ~ & $\uparrow \post^*$ effective 
+			& strong %upward 
+					compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
+   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
+   BSRP & Undecidable (Thm~\ref{srp up down}) &  ?  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+      kSRP & Undecidable (Thm~\ref{srp up down}) & ? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
+ \end{tabular}
+\end{center}
+
+\subsection{Case: $\Safe = \downarrow \Safe$% and $\Bad = \uparrow \Bad$
+}
+
+We now consider the case $\Safe = \downarrow \Safe$.
+%
+Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+% Appendix~\ref{HS-Minsk}. 
+Section~\ref{safe-down}.
+
+\begin{theorem}\label{srp down up}
+{\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
+when
+$\Safe=\downarrow \Safe$.
+\end{theorem}
+
+Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$
+is symmetrical to the case $\Safe = \uparrow \Safe$.
+%
+For instance one can write the following lemma:
+
+\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap U \subseteq D$  iff $ (\downarrow  A) \cap U \subseteq D$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap U \subseteq D$.
+Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$.
+Let us show that $({\downarrow A}) \cap U \subseteq {\downarrow (A \cap U)}$.
+Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$.
+Since $x \in U$ and $U$ is upward-closed, we also have $a \in U$.
+Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
+In the other direction, since $A \subseteq {\downarrow A}$, the inclusion $({\downarrow  A}) \cap U \subseteq D$ implies $A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
+\qed
+\end{proof}
+
+In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
+then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
+Lemma~\ref{Lemma intersection 2} can be used to show that
+if $\Safe = \downarrow \Safe$,
+then
+$\post^*(s) %\cap \Bad 
+\subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) 
+%\cap \Bad 
+\subseteq \pred^*(\Safe)$.
+
+\begin{theorem}\label{downward srp}
+{\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities,
+$\Safe = \downarrow \Safe$.
+\end{theorem}
+
+\begin{proof}
+In order to decide whether $\post^{\leq n}(s) %\cap \Bad
+ \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
+one looking for a resilience certificate and one looking for a non-resilience certificate.
+Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward-closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
+Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
+Notice that, by 
+upward compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
+
+Procedure 1 stops when it finds an invariant $D$ such that
+$D  %\cap \Bad
+ \subseteq \pred^*(\Safe)$. 
+Indeed
+$D % \cap \Bad 
+\subseteq  \pred^*(\Safe)$ implies
+$\downarrow \post^*(s) % \cap \Bad
+ \subseteq  \pred^*(\Safe)$
+since $ \downarrow \post^*(s)  \subseteq D$.
+
+The second procedure iteratively computes
+$\post^{\leq n}(s) %\cap \Bad
+$
+until it finds an element
+not in $\pred^*(\Safe)$. \qed
+\end{proof}
+
+Strong downward compatibility implies furthermore the decidability
+of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.
+
+\begin{corollary}\label{downward brp}
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility,
+$\Safe = \downarrow \Safe$.
+\end{corollary}
+
+%\subsection{Computing the maximal resilient subsystem}
 

### 2023-08-31 final.tex
up

@@
 Let $B_{\uparrow \post^*(s)}$ 
-%\alain{notation fluctuante}
  be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
-\[ \post^*(s) %\cap \Bad
- \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)} %\cap \Bad 
+\[ \post^*(s) 
+ \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)}  
  \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} %\cap \Bad
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} 
  \setminus \pred^*(\Safe) = \emptyset$.  \qed
@@
 \begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong %upward 
-			compatibility ~ & $\uparrow \post^*$ effective 
-			& strong %upward 
-					compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
+\hline  Hypothesis & strong compatibility ~ & $\uparrow \post^*$ effective 
+			& strong compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
    SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
@@
 
-\subsection{Case: $\Safe = \downarrow \Safe$% and $\Bad = \uparrow \Bad$
-}
+\subsection{Case: $\Safe = \downarrow \Safe$}
 
@@
 As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-% Appendix~\ref{HS-Minsk}. 
 Section~\ref{safe-down}.
@@
 then
-$\post^*(s) %\cap \Bad 
+$\post^*(s)  
 \subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) 
-%\cap \Bad 
 \subseteq \pred^*(\Safe)$.
@@
 \begin{proof}
-In order to decide whether $\post^{\leq n}(s) %\cap \Bad
+In order to decide whether $\post^{\leq n}(s) 
  \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
@@
 Procedure 1 stops when it finds an invariant $D$ such that
-$D  %\cap \Bad
- \subseteq \pred^*(\Safe)$. 
+$D   \subseteq \pred^*(\Safe)$. 
 Indeed
-$D % \cap \Bad 
-\subseteq  \pred^*(\Safe)$ implies
-$\downarrow \post^*(s) % \cap \Bad
- \subseteq  \pred^*(\Safe)$
+$D \subseteq  \pred^*(\Safe)$ implies
+$\downarrow \post^*(s)  \subseteq  \pred^*(\Safe)$
 since $ \downarrow \post^*(s)  \subseteq D$.
@@
 The second procedure iteratively computes
-$\post^{\leq n}(s) %\cap \Bad
-$
+$\post^{\leq n}(s) $
 until it finds an element
@@
 
-%\subsection{Computing the maximal resilient subsystem}
-
 \input{application}

### 2023-08-31 final.tex
up

@@
 VASS.
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  \iffalse Moreover, we may show:
-
-\begin{proposition}
-There exist classes of WSTSs with strong 
- compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$.
-\end{proposition}
-
-\begin{proof}
-Let us show that reset-VASS, that are effective WSTSs with strong compatibility, don't enjoy the property that $\uparrow \post^*$ is computable.
-Suppose that one are able to compute a finite basis of $\uparrow \post^*$ for reset-VASS. 
-Then, {\sc State-resilience} would be decidable for reset-VASS, which leads to a contradiction.
-\end{proof} \fi
+However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  
 Moreover there exist classes of WSTSs with strong 
@@
 
-\input{application}
-
- 	\input{conclusion}
+\section{Resilience for VASS and variations}\label{section VASS}
+
+In this section we continue to study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+$\Safe$ is upward-closed. Lacking downward compatibility or a more relaxed hypothesis that
+for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
+VASS do not inherit the decidability results for WSTS in the case $\Safe$ is downward-closed.
+In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.
+We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
+
+\begin{theorem}\label{SL VASS}
+{\sc Resilience} is decidable for VASS when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We consider the case when $\Safe$ is semilinear
+and $\Bad = S \setminus \Safe$ is semilinear too.
+{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
+We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
+Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
+\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).\]
+Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
+$\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
+{\sc Resilience} is decidable. \qed
+\end{proof}
+
+{\sc Resilience} is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logic.
+Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
+
+\begin{theorem}\label{SL Lossy}
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
+Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
+\qed
+\end{proof}
+
+Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.
+
+\begin{theorem}\label{SL int-VASS}
+{\sc Resilience} is decidable for integer VASS when $\Safe$ 
+is a
+semilinear set.
+\end{theorem}
+
+\begin{proof}
+The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^{2d}$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\end{proof}
+
+Continuous VASS \cite{DBLP:journals/tocl/BlondinFHH17} are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+
+\begin{theorem}\label{RP VASS}
+{\sc Resilience} is decidable for continuous VASS when $\Safe$  is
+definable in the existential theory of the rationals with addition and order.
+\end{theorem}
+
+\begin{proof}
+The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). \qed
+\end{proof}
+
+Remark that
+Theorem~\ref{SL Lossy} and
+Theorem~\ref{SL int-VASS}
+(resp. Theorem~\ref{RP VASS})
+did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
+semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
+the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+The above results concern only unbounded resilience. We consider now 
+{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
+the path from $\Bad$ to $\Safe$
+when $\Safe$ is downward-closed. Unfortunately, we have the following:
+\begin{proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$.
+\end{proposition}
+
+\begin{proof}
+Consider a given $k \in \N$,
+ $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
+consider a given VAS $V$.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
+The set $\Bad$ admits a finite basis $B_\Bad$.
+Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+
+All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
+and thus, are in $\Bad$, by upward-closedness.
+Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
+Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
+\qed
+\end{proof}
+
+This all changes when one considers VASS (having control-states).
+For a VASS $V = (Q,T)$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\uparrow q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
+ $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+This lead to a decision procedure
+ which lead to the following decidability result:
+\begin{theorem}
+{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+$\Safe = \downarrow \Safe$.
+\end{theorem}
+
+\begin{proof}
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider 
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\uparrow q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+ $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If it is infinite then
+$k$-resilience does not hold 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
+
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards,
+$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+\qed
+\end{proof}
+
+\section{Conclusion and perspectives}
+
+We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
+We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general. We also exhibited classes of WSTS with decidable resilience, namely, 
+completion-post-effective $\omega^2$-WSTS with strong compatibility (in the case 
+ $\Safe = \uparrow \Safe$) and 
+downward-compatible
+ideal-effective WSTS
+(in the case 
+ $\Safe = \downarrow \Safe$).
+
+Several questions still remain.
+For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
+One other angle of attack not studied here is synthesis of sets 
+$\Safe$ for resilience,
+for instance, 
+ find the maximal upward-closed (resp. downward-closed) subset 
+$\Safe$ so that a system is $\Safe$-resilient. 
+Another question could be to analyse the resilience in the framework of a controller and its environment. 
+One could also extend upon the classes of set $\Safe$ 
+ considered.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets \cite{DBLP:journals/fmsd/BertrandS13}.
+Finally, while we mention VASS, but a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
 

### 2023-08-31 final.tex
up

@@
 
-\input{appendixPN}
+\section{Reset-VASS and state-resilience}\label{appendix}
+
+\newcommand{\arrowcircle}[1][]{%
+  \begin{tikzpicture}[#1]
+    \draw[->] (0,0ex) -- (2em,0ex);
+    \draw (1em,0ex) circle (0.7ex);
+  \end{tikzpicture}%
+}
+
+Let us recall the {\em zero-reachability} problem in (reset-)VASS.
+
+\problemx{zero-reachability}
+{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+
+The zero-reachability problem is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
+For a better representation, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS. 
+A \emph{reset Petri net} $N = (P, T, E, R, \mu)$ consists of a finite set of {\em places} $P = \{p_1, p_2, ...
+p_{|P|}\}$, a finite set of {\em transitions} $T = \{t_1, t_2, ..., t_{|T|} \}$, a finite set of {\em arcs} $E \subseteq (P \times T) \cup (T \times P)$, a finite set of
+{\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
+It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
+The same equivalence hold between reset-VASS and reset Petri nets.
+
+We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
+
+\noindent
+\textbf{Proposition~\ref{liveness reset}}
+{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
+
+\begin{proof}
+
+If we wish to determine if a marking in which every place is empty is reachable for a reset Petri net
+ $N_1 = (P_1, T_1, E_1, R_1, \mu_1)$ 
+ we construct a reset Petri net
+ $N_2 = (P_2, T_2, E_2, R_2, \mu_2)$ 
+  which is live in $N_2$ if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
+
+\begin{samepage}
+{\bf Construction of $N_2$} \\
+\indent
+The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_a$ and $p_b$ and $|P_1| +2$ transitions $t_{p_a}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_b}$.
+We first modify all transitions of $N_1$ to include $p_a$ as both an input and an output.
+The initial marking $\mu_2$ will include a token in $p_a$ and no token in $p_b$. 
+This new place $p_a$ serve to mark that the run is ongoing.
+As long as it contains a token, the adjusted transitions of $N_1$ are live and can be used normally.
+Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
+We add an additional transition $t_{p_a}$ 
+which has $p_a$ as an input and a null output.
+This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $P_1$ in $N_2$.
+ The place $p_a$ and transition $t_{p_a}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_a}$ to fire and freeze the net at that marking.
+We introduce a new place $p_b$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_b$ as output.
+Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live in $N_2$ if a token is ever
+put in $p_2$.
+\end{samepage}
+
+ \begin{center}
+	\begin{figure}
+\includegraphics[width=0.75\textwidth]{FigurePN}
+\caption{Construction of $N_2$ from $N_1$.
+Places are represented using circles, transitions are represented using lines.
+Arcs are represented by $\rightarrow$ and reset-arcs by $\arrowcircle$.
+Note that since the construction of $N_2$ from $N_1$ do not add new reset-arcs, the only reset-arcs in $N_2$ are the ones inherited from $N_1$.}
+	\end{figure}
+\end{center}
+
+Let us now prove that $t_{p_2}$ is live in $N_2$ if and only if 
+the zero marking
+is not reachable from $\mu_1$ in $N_1$. \\
+
+{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live in $N_2$}.
+
+Indeed the marking with zero in every place of $P_1$ and in $p_b$ is reachable in $N_2$, by executing the same sequence of transition firings.
+Then $t_{p_a}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
+From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
+Thus $t_{p_2}$ is not live in $N_2$.
+\\
+
+{\bf Suppose that $t_{p_2}$ is not live in $N_2$, then the zero marking is reachable from $\mu_1$ in $N_1$}
+
+Indeed, if $t_{p_2}$ is not live in $N_2$, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
+This means that no transition $t_p$ is live in $N_2$ in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
+Thus, every place of $P_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
+it can be reached without using the transitions $t_p$ or $t_{p_2}$.
+Since $t_{p_a}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
+Thus the zero marking is reachable from $\mu_1$ in $N_1$.
+\end{proof}
+
+\alain{numérotation à faire}
+\noindent
+\textbf{Corollary~\ref{liveness reset VASS}}
+{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
 

### 2023-08-31 appendixPN.tex
section

@@
 \includegraphics[width=0.75\textwidth]{FigurePN}
-\caption{Construction of $N_2$ from $N_1$.
-Places are represented using circles, transitions are represented using lines.
-Arcs are represented by $\rightarrow$ and reset-arcs by $\arrowcircle$.
-Note that since the construction of $N_2$ from $N_1$ do not add new reset-arcs, the only reset-arcs in $N_2$ are the ones inherited from $N_1$.}
+\caption{Construction of $N_2$ from $N_1$.}
+%			As usual, places are represented by circles, transitions by lines. Arcs are represented by $\rightarrow$.
+%			and reset-arcs by $\arrowcircle$. 
+%		Note that $N_2$ do not contain new reset-arcs that were not in $N_1$.}
+%		the only reset-arcs in $N_2$ are the ones inherited from $N_1$.}
 	\end{figure}
@@
 is not reachable from $\mu_1$ in $N_1$. \\
+
+\alain{dans le dessin Fig 2, écrire reset Petri net $N_1$ et pas Petri net}
 

### 2023-08-31 final.tex
up

@@
 
-Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility.
+Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong upward-compatibility.
 Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
@@
 \includegraphics[width=0.75\textwidth]{FigurePN}
-\caption{Construction of $N_2$ from $N_1$.
-Places are represented using circles, transitions are represented using lines.
-Arcs are represented by $\rightarrow$ and reset-arcs by $\arrowcircle$.
-Note that since the construction of $N_2$ from $N_1$ do not add new reset-arcs, the only reset-arcs in $N_2$ are the ones inherited from $N_1$.}
+\caption{Construction of $N_2$ from $N_1$.}
 	\end{figure}

### 2023-08-31 final.tex
up
Added the " in front of Ozkan

@@
 
-In 2021, Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
+In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
 $\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$ and $\Bad$ downward-closed.
@@
 Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
-However, neither Prasad \& Zuck nor Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
+However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
 %%%%%%%%%
@@
 The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
-Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
+\"Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
 
@@
 
-\bibliography{bibliography/bib}
+\bibliography{bib}
 
@@
 
-\alain{numérotation à faire}
 \noindent
-\textbf{Corollary~\ref{liveness reset VASS}}
+\textbf{Corollary}
 {\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}

### 2023-08-31 final.tex
section

--- 2023-08-31 final.tex
@@
 We now recall the
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
 reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
@@
 \problemx{$t$-liveness}
-{A reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, a transition $t \in T$, an initial configuration $s_0 \in Q \times \N^d$}
 {Is $t$ live in $s_0$ ? \\}
 
-To better discuss $t$-liveness let us define more formally the upward-closed set of configurations such that $t$ is enabled
-$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+To better discuss \alain{flou}$t$-liveness, let us define more formally the upward-closed set of configurations such that $t$ is enabled:
+$pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
 

### 2023-08-31 final.tex
section

@@
  A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
- {\em states} and  
+ {\em states (or configurations)} and  
  $ {\rightarrow} \subseteq S \times S$ is a
@@
 
-Let us recall the  definition of vector addition system with states. 
+Let us recall the  definition of vector addition system with (control-)states. 
  \begin{definition} 
@@
 which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
- $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
+A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a configuration
+ $p(\textbf{u})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \to^t $ ($t$ is enabled in $p(\textbf{u})$). We say that the whole reset-VASS is live if all its control-transitions are
 live. This leads to the following problem.
@@
 
-To better discuss \alain{flou}$t$-liveness, let us define more formally the upward-closed set of configurations such that $t$ is enabled:
+To better discuss \alain{flou}$t$-liveness, let us define the upward-closed set of configurations \alain{configuration ou state, tjs le même mot stp}such that $t$ is enabled:
 $pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.

### 2023-08-31 final.tex
section

@@
 A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a configuration
- $p(\textbf{u})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \to^t $ ($t$ is enabled in $p(\textbf{u})$). We say that the whole reset-VASS is live if all its control-transitions are
+ $p(\textbf{u})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} $ ($t$ is enabled in $p(\textbf{u})$). We say that the whole reset-VASS is live if all its control-transitions are
 live. This leads to the following problem.
@@
 
-To better discuss \alain{flou}$t$-liveness, let us define the upward-closed set of configurations \alain{configuration ou state, tjs le même mot stp}such that $t$ is enabled:
-$pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.
+%		 better discuss \alain{flou}$t$-liveness, 
+Let us define the set of configurations such that $t$ is enabled:
+$pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $pre(t)$ is upward-closed.
 

### 2023-08-31 final.tex
section

@@
 ~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
-The problem reduces itself to {\sc State-resilience}
+The previous formula reduces itself to {\sc State-resilience}
 where $\Safe = pre(t)$.
@@
 
-The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS, whose proof can be seen in Appendix~\ref{appendix}%
+The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets, whose proof can be seen in Appendix~\ref{appendix}%
 .

### 2023-08-31 final.tex
section

@@
 
-On the positive side, let us recall a result about {\sc Bounded-state-resilience} decidability (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
@@
 
-We may immediately generalyse this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
+We may immediately generalize this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
 

### 2023-08-31 final.tex
section

@@
 \begin{proof}
-Since $\mathscr{S}$ is a WSTS there exists $n_0 \in \N$ such that
-$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute this $n_0$ by iteratively computing 
+Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
+$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
 $\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
@@
 %\"Ozkan
+\alain{pourquoi ne pas juste écrire Proposition 1 \cite{DBLP:conf/gg/Ozkan22}et enlever "notre" proposition 3 ?}
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
@@
 Let $B_{\uparrow \post^*(s)}$ 
- be a basis of $\uparrow \post^*(s)$, $B_\Safe$ a basis of $\Safe$.
+ be a basis of $\uparrow \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
@@
 
-However when removing strong compatibility, precision is lost.
+However when removing strong compatibility, some precision is lost.
 Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
@@
 $\post^* (s) \cap S \subseteq \pred( \Safe)$.
-In such a case the algorithm from
+In such a case, the algorithm in
 \cite{DBLP:conf/gg/Ozkan22} would deduce that {\sc $1$-state-resilience} does not hold,

### 2023-08-31 final.tex
section

@@
 %\"Ozkan
-\alain{pourquoi ne pas juste écrire Proposition 1 \cite{DBLP:conf/gg/Ozkan22}et enlever "notre" proposition 3 ?}
+%\alain{pourquoi ne pas juste écrire Proposition 1 \cite{DBLP:conf/gg/Ozkan22}et enlever "notre" proposition 3 ?}
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
@@
 \textbf{Proposition~\ref{liveness reset}}
-{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
+{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\
+
+We deduce that:\\
+
+\noindent
+\textbf{Corollary}
+{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
 
@@
 
-\noindent
-\textbf{Corollary}
-{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
-
 \end{document}

### 2023-08-31 final.tex
up

@@
 The set $\Bad$ admits a finite basis $B_\Bad$.
-Consider $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider 
+Consider the vector $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider the vector
 $\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
@@
 q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider 
+Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider the vector
 $\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.

### 2023-08-31 final.tex
up
section

@@
 {$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
-Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V_1$.
+Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any state of $V_1$.
 However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$.
@@
  reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
-Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.
+Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.
 From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less.
@@
  A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
- {\em states (or configurations)} and  
+ {\em states} and  
  $ {\rightarrow} \subseteq S \times S$ is a
@@
 We say that $\mathscr{S}$
-is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
+is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
@@
 %
-Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with a $d$-VASS $V$.
-For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+Subsetquently, $Q \times \N^d$ is the set of states of the transition system associated with a $d$-VASS $V$.
+For all states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
 %
@@
 Again $Q \times \N^d$
- denotes the set of configurations of $V$.
-For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
+ denotes the set of states of $V$.
+For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
 $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
@@
 which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a configuration
+A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a state
  $p(\textbf{u})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} $ ($t$ is enabled in $p(\textbf{u})$). We say that the whole reset-VASS is live if all its control-transitions are
@@
 \problemx{$t$-liveness}
-{A reset-VASS $V=(Q,T)$ of dimension $d$, a transition $t \in T$, an initial configuration $s_0 \in Q \times \N^d$}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, a transition $t \in T$, an initial state $s_0 \in Q \times \N^d$}
 {Is $t$ live in $s_0$ ? \\}
@@
 %		 better discuss \alain{flou}$t$-liveness, 
-Let us define the set of configurations such that $t$ is enabled:
+Let us define the set of states such that $t$ is enabled:
 $pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $pre(t)$ is upward-closed.
@@
 reset-VASS $(Q,T)$ 
-with initial configuration $s_0$
+with initial state $s_0$
 as the following formula
@@
 \begin{center}
+\resizebox{\textwidth}{!}{
 \begin{tabular}{ | l | c | c | c | c |}
@@
  \end{tabular}
+ }
 \end{center}
@@
 
-All configurations reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
+All states reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
 and thus, are in $\Bad$, by upward-closedness.
@@
 $\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all configurations reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
 of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
@@
 
-We deduce that:\\
+We may then deduce that:\\
 

### 2023-08-31 final.tex
up

@@
  \begin{definition} 
-A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+A {\em vector addition system with (control-)states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
  \end{definition} 

### 2023-08-31 final.tex
up

@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set Safe  ($\Safe=\uparrow \Safe$) or downward-closed set ($\Safe=\downarrow \Safe$) (Bad is generally the complementary of Safe) and by the existence of a bound $k$ on the length of minimal runs starting from Bad and reaching Safe.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\downarrow \Safe$) ($\Bad$ is generally the complementary of $\Safe$) and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
@@
 $\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$.
-Finally, we study the resilience for VASS with semi-linear subsets Bad and Safe and for variations of VASS (lossy counter machines, integer VASS and continuous VASS); most of the resilience properties are shown decidable.
+Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines, integer VASS and continuous VASS); most of the resilience properties are shown decidable.
 \end{abstract}
@@
 
-%		 better discuss \alain{flou}$t$-liveness, 
 Let us define the set of states such that $t$ is enabled:
@@
 
-%\"Ozkan
-%\alain{pourquoi ne pas juste écrire Proposition 1 \cite{DBLP:conf/gg/Ozkan22}et enlever "notre" proposition 3 ?}
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}

### 2023-08-31 final.tex
up

@@
 %
-\title{Resilience and Home-Space for WSTS \thanks{This work was partly done while the author was supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}} 
+\title{Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}} 
 %

### 2023-08-31 final.tex
up

@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\downarrow \Safe$) ($\Bad$ is generally the complementary of $\Safe$) and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\downarrow \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.

### 2023-08-31 final.tex
up

@@
 $\post^*(s_0) \subseteq \pred^*(\Safe)$.
-We will study three decidability questions for each of type of resilience.
+We will study three decidability questions for each type of set $\Safe$.
 The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient (for a given subset of states $\Safe \subseteq S$).
@@
 \item Surprinsingly, the general undecidability statements about resilience were not known neither proved.
-We show that resilience and { State-resilience} problems are both undecidable for WSTS with strong compatibility.
-Moreover, {State-resilience},
-{Bounded-state-resilience} and
+We show that resilience and {state-resilience} problems are both undecidable for WSTS with strong compatibility.
+Moreover, {state-resilience},
+{bounded-state-resilience} and
 {$k$-state-resilience}
@@
 when
-$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {\ State-resilience} in reset-VASS.
+$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
 
@@
 $\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({ $k$-state-resilience} and { bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
+We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
 %

### 2023-08-31 final.tex
up

@@
 
-{\bf Home-spaces}
+{\bf Home-spaces.}
  In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
@@
 Consider now the
- reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
+ reset-VASS $V_2$, which is $V_1$ plus a $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
 Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.
@@
 \noindent
-{\bf On effectivity}
+{\bf On effectivity.}
 Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms

### 2023-08-31 final.tex
up

@@
 {\bf On effectivity.}
-Let a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be a WSTS. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
 ($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
@@
 a path of length smaller than or equal to $k$) that reaches a state in $\Safe$.
-Resilience is then the Home-Space problem (defined in the introduction) for the set $\Safe$.
+Resilience is then akin to the Home-Space problem (defined in the introduction) for the set $\Safe$.
 We use the notation
@@
 
-Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
 The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.

### 2023-08-31 final.tex
up

@@
 Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo.
+The completion is always finitely branching but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
 $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
 A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-it is post-effective, and there exists a Turing Machine $M_\downarrow$ that computes, on input $s \in S$, 
- $\downarrow s$, and some Turing Machine $M_{\uparrow^C}$ that computes, on input
+it is post-effective, and there exists an algorithm $M_\downarrow$ that computes, on input $s \in S$, 
+ $\downarrow s$, and some algorithm $M_{\uparrow^C}$ that computes, on input
  $\{s_1, s_2, \ldots, s_m\} \in S^m$, 

### 2023-08-31 final.tex
up

@@
 
-Remark that the previous proof don't make use of the 
+Remark that the above proof doesn't make use of the 
 property that $\Bad$ is the complement of $\Safe$, simply using

### 2023-08-31 final.tex
up

@@
 	\noindent	
-	where $TM_j$ is the $j$-th Turing machine (in a classical enumeration)
+	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
 	which moreover begins
@@
 	$\uparrow (1,1)$-resilient
-	iff $TM_j$ halts on input $j$,
+	iff TM$_j$ halts on input $j$,
 	since $(1,1)$ is coverable from $(0,0)$
-	iff $TM_j$ halts on input $j$. 
+	iff TM$_j$ halts on input $j$. 
 	Hence there is no Turing machine which correctly determines 

### 2023-08-31 final.tex
up

@@
 undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$.
-This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.
+This stems from the fact that it is undecidable for Minsky machines with more than one counter whether zero is a home-state.
 
 Home-space is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
-This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61,Min67}.
+This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
 From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 

### 2023-08-31 final.tex
up

@@
 {for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}
-is not met, the problem is undecidable.
- Indeed, it is 
-undecidable for reset-VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state for $\N^d \setminus \textbf{0}$, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$.
-This stems from the fact that it is undecidable for Minsky machines with more than one counter whether zero is a home-state.
-
-Home-space is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
+is not met, {\sc Resilience} is undecidable.
+This stems from the fact that it is undecidable for Minsky machines with more than one counter,
+which can be simulated by WSTS.
+
+Indeed, {\sc Home-space} is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
 This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.

### 2023-08-31 final.tex
up

@@
 
-Indeed, {\sc Home-space} is undecidable for Minsky machine with at least $3$ counters, since it is undecidable in the particular case where $H$ is a singleton.
+Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
 This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
@@
 It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since Reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
 

### 2023-08-31 final.tex
up

@@
 $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complement of each others.
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
 
@@
 %
-Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$.
+Unfortunately the {\sc State-resilience} problem is undecidable in this case.
 As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in

### 2023-08-31 final.tex
up

@@
 Unfortunately the {\sc State-resilience} problem is undecidable in this case.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+As for resilience in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
 Section~\ref{safe-down}.

### 2023-08-31 final.tex
up

--- 2023-08-31 final.tex
@@
 %
-Unfortunately the {\sc State-resilience} problem is undecidable in this case.
-As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+Unfortunately {\sc State-resilience} is undecidable in this case.
+As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
 Section~\ref{safe-down}.

### 2023-08-31 final.tex
up

@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility,
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.

### 2023-08-31 final.tex
up

@@
 
-In this section we continue to study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
-$\Safe$ is upward-closed. Lacking downward compatibility or a more relaxed hypothesis that
+In this section we study VASS rather than WSTS in general. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+$\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
 for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 

### 2023-08-31 final.tex
up

@@
 for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
-VASS do not inherit the decidability results for WSTS in the case $\Safe$ is downward-closed.
+VASS do not inherit the decidability results for WSTS in the case $\Safe = \downarrow \Safe$.
 In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.

### 2023-08-31 final.tex
up

@@
 
-In this section we study VASS rather than WSTS in general. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+In this section we study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
 $\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
@@
 
-{\sc Resilience} is also decidable for other class of counter machines for which the reachability relation can be expressed in a decidable logic.
+{\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
 Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.

### 2023-08-31 final.tex
up

@@
 For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
-One other angle of attack not studied here is synthesis of sets 
+One other angle of attack not studied here is the synthesis of sets 
 $\Safe$ for resilience,
 for instance, 
- find the maximal upward-closed (resp. downward-closed) subset 
+ finding the maximal upward-closed (resp. downward-closed) subset 
 $\Safe$ so that a system is $\Safe$-resilient. 
@@
  considered.
-As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets \cite{DBLP:journals/fmsd/BertrandS13}.
-Finally, while we mention VASS, but a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets~\cite{DBLP:journals/fmsd/BertrandS13}.
+Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
 

### 2023-08-31 final.tex
section
up

@@
 %
-\author{Alain Finkel\inst{1,2,3} \and Mathieu Hilaire\inst{1,3}}
+\author{Alain Finkel\inst{1,2} \and Mathieu Hilaire\inst{1}}
 % This research has been supported by ANR programme BraVAS (ANR-17-CE40-0028).
@@
 \and
-Institut Universitaire de France \and Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028)
-}
+Institut Universitaire de France }
+
+%\and Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028)
+%}
 %
@@
 Consider now the
- reset-VASS $V_2$, which is $V_1$ plus a $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
+ reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
 Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.

### 2023-08-31 final.tex
up
section

@@
 
-%\and Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028)
-%}
 %
@@
 exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
-on input $x$, with $x \in X_i $. 
+on input $x \in X$. 
 We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting

### 2023-08-31 final.tex
section

@@
 \problemx{$k$-resilience (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a sets $\Safe \subseteq S$.}
+{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a set $\Safe \subseteq S$.}
 {$S \longrightarrow^{\leq k} \Safe$ ?\newline}
@@
 A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-it is post-effective, and there exists an algorithm $M_\downarrow$ that computes, on input $s \in S$, 
- $\downarrow s$, and some algorithm $M_{\uparrow^C}$ that computes, on input
- $\{s_1, s_2, \ldots, s_m\} \in S^m$, 
- the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$.
+it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+ $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+ $\{s_1, s_2, \ldots, s_m\} \in S^m$.\alain{que signifie $S^m$ ? pourquoi pas $S$ ?}
 Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).

### 2023-09-02 final.tex
up

@@
  $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
- $\{s_1, s_2, \ldots, s_m\} \in S^m$.\alain{que signifie $S^m$ ? pourquoi pas $S$ ?}
+ $s_1, s_2, \ldots, s_m \in S$.
 Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).

### 2023-09-02 final.tex
up

@@
 In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
-$\uparrow$ $\post^*$ basis) and with $\Safe=\uparrow \Safe$ and $\Bad$ downward-closed.
+$\uparrow$ $\post^*$ basis) in the case $\Safe=\uparrow \Safe$ and $\Bad = \downarrow \Bad$.
 

### 2023-09-02 final.tex
up

@@
 infinite subset of mutually incomparable elements (antichains)
- enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.
+ enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$.
 %

### 2023-10-12 VCMAI revised version/final.tex
section

--- 2023-09-02 final.tex
@@
 \\
+Let us recall a result of Jancar \cite{DBLP:journals/ipl/Jancar99} that proved: that a wqo $(X,\leq)$ is an $\omega^2-wqo$ iff $(Ideals(X),\subseteq)$ is a wqo.
 
@@
 We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
+When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\uparrow U_k)_k$ instead.
 With the effective pred-basis hypothesis, we obtain:

### 2023-10-12 VMCAI-revised-version/final.tex
up

--- 2023-10-12 VCMAI revised version/final.tex
@@
 \\
-Let us recall a result of Jancar \cite{DBLP:journals/ipl/Jancar99} that proved: that a wqo $(X,\leq)$ is an $\omega^2-wqo$ iff $(Ideals(X),\subseteq)$ is a wqo.
 
@@
 
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
-Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
+Several families of formal models of processes give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
+Petri nets when inclusion between markings is used as the well-ordering (WSTS with strong compatibility) and lossy channel systems with the subword ordering (WSTS, compatibility is not strong).
+%%%%
+
+Let us recall a result of Jancar \cite{DBLP:journals/ipl/Jancar99} that we will use as a simple definition of an $\omega^2$-wqo: a wqo $(X,\leq)$ is an $\omega^2$-wqo iff $(Ideals(X),\subseteq)$ is a wqo.
+There exists wqo that are not $\omega^2$-wqo (see the use of the Rado ordering in \cite{DBLP:journals/ipl/Jancar99}).
+
+\begin{definition} \cite{DBLP:journals/corr/abs-1208-4549}
+A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
+is an $\omega^2$-WSTS if $(S, \leq)$ is an $\omega^2$-wqo.
+\end{definition}
+All naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, Petri nets, VASS, reset/transfer VASS, affine VASS with $d$ counters and $Q$ a finite set of control-states are $\omega^2$-WSTS for the usual ordering ($=\times \leq^d$) on $S=Q \times \N^d$.
+Similarly, lossy channel systems with $d$ channels and $Q$ a finite set of control-states are $\omega^2$-WSTS
+for the ordering $=\times \sqsubseteq^d$ (where $\sqsubseteq$ is the subword ordering on $\Sigma^*$) on $S= Q \times (\Sigma^*)^d$.\\
+%%%%%%
 

### 2023-10-13 version-0-VMCAI/abstract.tex
section

--- 2023-06-23 abstract.tex
@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set Safe  ($\Safe=\uparrow \Safe$) or downward-closed set ($\Safe=\downarrow \Safe$) (Bad is generally the complementary of Safe) and by the existence of a bound $k$ on the length of runs starting from Bad and reaching Safe.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set Safe  ($\Safe=\uparrow \Safe$) or downward-closed set ($\Safe=\downarrow \Safe$) (Bad is generally the complementary of Safe) and by the existence of a bound $k$ on the length of minimal runs starting from Bad and reaching Safe.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.

### 2023-10-13 version-0-VMCAI/bibliography/icalp.tex
section

--- 2022-10-14 bibliography/succ.tex
@@
 \documentclass[times,envcountsame]{llncs}
-\usepackage{a4wide}
+
 \usepackage{times}
@@
 \usepackage{gastex}
+\usepackage{url}
 
@@
 
-%\usepackage[thmmarks,standard,thref]{ntheorem}
-
-\iffalse
-
-\newtheorem{theorem}{{\bf Theorem}}[section]
-\newtheorem{corollary}[theorem]{{\bf Corollary}}
-\newtheorem{definition}[theorem]{{\bf Definition}}
-\newtheorem{example}[theorem]{{\bf Example}}
-\newtheorem{lemma}[theorem]{{\bf Lemma}}
-\newtheorem{proposition}[theorem]{{\bf Proposition}}
-\newtheorem{remark}[theorem]{{\bf Remark}}
+\usepackage{wrapfig}
+%\usepackage{floatfig}
+
 \newtheorem{fact}[theorem]{Fact}
 
-\fi
-
-\newtheorem{fact}[theorem]{Fact}
-
 \usepackage{stmaryrd,rotating,mathrsfs}
 
-%\usepackage[active]{srcltx}
-
-\iffalse
-\newcommand{\problemx}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf QUESTION: }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\newcommand{\problemy}[3]{
-\par\noindent\underline{\sc#1}\par\nobreak\vskip.2\baselineskip
-\begingroup\clubpenalty10000\widowpenalty10000
-\setbox0\hbox{\bf INPUT: }\setbox1\hbox{\bf OUTPUT: }
-\dimen0=\wd0\ifnum\wd1>\dimen0\dimen0=\wd1\fi
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box0\hfil}\hangindent\dimen0\hangafter1\ignorespaces#2\par
-\vskip-\parskip\noindent
-\hbox to\dimen0{\box1\hfil}\hangindent\dimen0\hangafter1\ignorespaces#3\par
-\endgroup}
-
-\fi
+\newcommand{\prim}{\text{prime}}
 
@@
 
-\def\AC{{\mathrm{AC}}}
+\def\AC{{\mathsf{AC}}}
 \def\AG{{\mathsf{AG}}}
@@
 
-\title{Model Checking Succinct and Parametric One-Counter Automata}
+\title{Model Checking Succinct and Parametric\\ One-Counter Automata}
 \author{Stefan G\"oller\inst{1} \and
@@
 \section{Introduction}
-
-Counter automata, which comprise a finite-state controller together
-with a number of counter variables, are a fundamental and widely-studied
-computational model. One of the earliest results about counter
-automata, which appeared in a seminal paper of Minsky's five decades
-ago, is the fact that two counters suffice to achieve Turing
-completeness~\cite{Min61}.
-
-Following Minsky's work, much research has been directed towards
-studying restricted classes of counter automata and related
-formalisms. Among others, we note the use of restrictions to a single
-counter, on the kinds of allowable tests on the counters, on the
-underlying topology of the finite controller (such as
-flatness~\cite{CJ98-cav,LS08-atva}), and on the types of computations
-considered (such as reversal-boundedness~\cite{ID06-tcs}). Counter
-automata are also closely related to Petri nets and pushdown automata.
-
-In Minsky's original formulation, counters were represented as integer
-variables that could be incremented, decremented, or tested for
-equality with zero by the finite-state controller. More recently,
-driven by complexity-theoretic considerations on the one hand, and
-potential applications on the other, researchers have investigated
-additional primitive operations on counters, such as additive updates
-encoded in binary~\cite{BHIMV06,LS08-atva} or even
-in \emph{parametric} form, i.e., whose precise values depend on
-parameters~\cite{BIL06-icalp,IJTW93-icalp}. We refer to such counter automata as
-\emph{succinct} and \emph{parametric} respectively, the former being
-viewed as a subclass of the latter. Natural applications of such
-counter machines include the modeling of resource-bounded processes,
-programs with lists, recursive or multi-threaded programs, and XML
-query evaluation; see, e.g.,
-\cite{ID06-tcs,BHIMV06,CR04}.
-
-In most cases, investigations have centered around the decidability
-and complexity of the \emph{reachability} problem, i.e., whether a
-given control state can be reached starting from the initial
-configuration of the counter automaton. Various instances of the
-reachability problem for succinct and parametric counter automata are
-examined, for example, in~\cite{DG-jlc09,HKOW10,IJTW93-icalp}.
-
-The aim of the present paper is to study the decidability and
-complexity of \emph{model checking} for succinct and parametric
-one-counter automata. In view of Minsky's result, we restrict our
-attention to \emph{succinct one-counter automata (SOCA)} and
-\emph{parametric one-counter automata (POCA)}. On the specification
-side, we focus on the three most prominent formalisms in the
-literature, namely the temporal logics CTL and LTL, as well as the
-modal $\mu$-calculus. For a counter automaton $\mathbb{A}$ and a
-specification $\varphi$, we therefore consider the question of
-deciding whether $\mathbb{A} \models \varphi$,
-in case of POCA for all values of the parameters,
-%\footnote{In the case of
-%parametric automata, one asks whether the automaton satisfies the
-%formula \emph{for all} legal values of the parameters.}
-and investigate both the \emph{data} complexity (in which the formula
-$\varphi$ is fixed) as well as the \emph{combined} complexity of this
-problem. Our main results are summarized in Table \ref{T results}.
-\begin{table}
-\begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}m{3cm}
-  >{\centering\arraybackslash}m{2cm} |
-  >{\centering\arraybackslash}m{3.3cm} |
-> {\centering\arraybackslash}m{3.3cm} |}\\
-  &  &{\bf SOCA}&{\bf POCA}\\ \hline
-{\raisebox{-2ex}[0ex]{{\bf CTL, $\mu$-calculus}}}&   {\bf data} &   & \\ \cline{2-2}
-{\raisebox{-2ex}[-5ex]{{$\phantom{\text{CTL}}$}}}&  {\bf combined}&
-{\raisebox{2ex}[0ex]{$\EXPSPACE$-complete}} &
-{\raisebox{2ex}[0ex]{ $\Pi^0_1$-complete}}\\ \cline{1-4}
-{\raisebox{-2ex}[0ex]{{\bf LTL}}}&  {\bf data} &
-\multicolumn{2}{c|} {\raisebox{0ex}[0ex]{$\coNP$-complete}}
-\\ \cline{2-4}
-{\raisebox{-2ex}[-5ex]{{$\phantom{\text{CTL}}$}}}&  {\bf combined}&
-$\PSPACE$-complete &
-$\coNEXP$-complete\\ \cline{1-4}
-\end{tabular}
-\ \\[0.4cm]
-\caption{The complexity of CTL, the modal $\mu$-calculus, and LTL on SOCA and
-POCA.}{\label{T results}}
-\end{center}
-\end{table}
-
-One of the motivations for our work was the recent discovery that
-reachability is decidable and in fact $\NP$-complete for both SOCA and
-POCA~\cite{HKOW10}. We were also influenced by the work of Demri and
-Gascon on model checking extensions of LTL over non-succinct,
-non-parametric one-counter automata~\cite{DG-jlc09}, as well as the
-recent result of G\"oller and Lohrey establishing that model checking
-CTL on such counter automata is $\PSPACE$-complete \cite{GoLo10}.
-
-We note some interesting differences between our results and
-corresponding questions regarding finite automata. For the latter, the
-(combined) model checking problems for CTL, the $\mu$-calculus, and
-LTL are respectively known to be $\mathsf{P}$-complete, in
-$\mathsf{NP} \cap \coNP$, and $\PSPACE$-complete. Somewhat
-surprisingly, for SOCA and POCA, the complexity ordering is reversed
-and LTL becomes easier to model check than either CTL or the
-$\mu$-calculus.
-
-On a technical level, the most intricate result is the
-$\EXPSPACE$-hardness of CTL model checking for SOCA, which requires
-several steps. We first show that $\mathsf{EXPSPACE}$ is `exponentially
-$\mathsf{LOGSPACE}$-serializable', adapting the known proof that
-$\PSPACE$ is $\mathsf{LOGSPACE}$-serializable. Unfortunately, and in
-contrast to~\cite{GoLo10}, this does not immediately provide
-an $\EXPSPACE$ lower bound. In a subsequent delicate stage of the
-proof, we show how to `split' the counter in order simultaneously to
-perform $\PSPACE$ computations in the counter and manipulate numbers
-of exponential size in a SOCA of polynomial size.
-
-Our paper is organized as follows. In Section \ref{S Prelim} we
-introduce general notations.
-Section \ref{S CTL} deals with CTL and the modal $\mu$-calculus.
-LTL model checking is content of Section \ref{S LTL}.
-Finally, in Section \ref{S Concl} we give conclusions.
-
-\section{Preliminaries}{\label{S Prelim}}
+\input{intro.tex}
+
+For reasons of space, we have had to abbreviate or omit a number of
+proofs; full details can however be found in the technical
+report~\cite{techrep}.
+
+\section{Preliminaries}
 
@@
 \newcommand{\bin}{\text{bin}}
-
-By $\Z$ we denote the {\em integers} and by $\N=\{0,1,2,\ldots\}$ the
-denote the {\em naturals}.  For each $i,j\in\Z$ we define
-$[i,j]=\{k\in\Z\mid i\leq k\leq j\}$ and $[j]=[1,j]$.  For each
-$i,n\in\N$, let $\bit_i(n)$ denote the $i^{\text{th}}$ least
-significant bit of the binary representation of $n$. Hence
-$n=\sum_{i\in\N}2^i\cdot\bit_i(n)$.  By $\bin_m(n)=\bit_0(n)\cdots
-\bit_{m-1}(n)$ we denote the first $m$ least significant bits written
-from {\em left to right}. When $m$ is not important we just write
-$\bin(n)$.
-%For any non-empty subset $M\subseteq\Z^+$, define $\gcd(M)$ to be the largest
-%$k\geq 1$ that divides every element of $M$.
-%We define $\gcd(\emptyset)=0$.
-Let $p_i$ denote the $i^{\text{th}}$ prime number for each $i\geq 1$, i.e. $p_1=2$,
-$p_2=3$ and so on. We define $\log(n)=\min\{i\geq 1\mid 2^i> n\}$, in other
-words $\log(n)$ denotes the number of bits that are needed to represent $n$ in
-binary.
-All polynomials $p:\N\rightarrow\N$ that occur in this paper are assumed to satisfy
-$p(n)\geq n$ for each $n\in\N$.
-For each word $v=a_1\cdots a_n\in\Sigma^n$ over some finite alphabet $\Sigma$
-and each $i,j\in[n]$ define $v[i,j]=a_i\cdots a_j$ and
-$v(i)=v[i,i]$.
-
-\bigskip
-
 \renewcommand{\l}{\triangleright}
@@
 \newcommand{\move}{\vdash}
-\noindent
-{\bf Turing machines and complexity theory:} In the following, we
-introduce $f(n)$ space-bounded deterministic Turing machines (DTMs) in
-a suitable way for proving lower bounds. These contain precisely one
-input tape and one working tape.  In our setting, the working alphabet is
- assumed to be $\{0,1,\l,\r\}$,
-where $\l$ is the {\em left marker}
-and $\r$ is the {\em right marker}.  The working tape of the initial
-configuration of such a DTM on an input $w\in\Sigma^n$ is
-assumed to be $\l0^{f(n)}\r$, whereas its input tape is $\l w\r$.
-Before we define DTMs, let $\Upsilon=\{-1,0,+1\}$ denote a set of
-directions.
-
-Formally, a {\em $f(n)$-space bounded deterministic Turing machine (DTM)} is a tuple
-$\M=(S,\Sigma,s_0,F,\mu)$, where $S$ is a finite set of {\em states},
-$\Sigma$ is a {\em finite input alphabet} with $\l,\r\not\in\Sigma$,
-$s_0\in S$ is an {\em initial state}, $F\subseteq S$ is a set of {\em final
-states}, and with $\Sigma_\M=\Sigma\uplus\{\l,\r\}$ we have that
-$$\mu\quad:\quad S\times\Sigma_\M\times\{0,1,\l,\r\}\ \rightarrow\
-S\times\Upsilon^2\times\{0,1,\l,\r\}$$
-is the {\em transition function}, where $\mu(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$ means
-that $\M$ is currently in state $s$, its input head reads $b_1$, its
-working tape reads $b_2$, and $\M$ changes to state $s'$, moves its
-input head in direction $\delta_1$,
-moves its working head in direction $\delta_2$, and
-writes the bit $b$.
-%We require
-%moreover that whenever $\M$'s input (resp. working) head reads $\l$, then $\M$
-%must move its input (resp. working) head to the right, i.e. $+1$, and whenever
-%$\M$'s input (resp.
-%working) head reads $\r$, %it must hold that
-%$\M$ moves its
-%input (resp. working) head to the left, i.e. $-1$. Formally,
-%whenever $\delta(s,b_1,b_2)=(s',\delta_1,\delta_2,b)$, then
-%for each $i\in\{1,2\}$ we have that $b_i=\l$ implies $\delta_i=+1$ and $b_i=\r$
-%implies $\delta_i=-1$.
-Moreover, we require that $\M$ writes a marker precisely when it reads a marker,
-more formally: $b_2=m$ if and only if $b=m$ for each $m\in\{\l,\r\}$.
-As expected, a {\em configuration of $\M$} is a tuple $(s,\l w\r,\l w'\r,i,j)$,
-where $s\in S$ is the current state,
-$w\in\Sigma^n$ is an input,
-$w'\in\{0,1\}^{f(n)}$ is the content of the working tape of $\M$,
-$i\in[0,n+1]$ is the current input head position, and
-$j\in[0,f(n)+1]$ is the current working head position.
-%For configurations $C_1$ and $C_2$, we write $C_1\move_\M C_2$ whenever $\M$ can
-%move from $C_1$ to $C_2$ on one step.
-We define the {\em language $L(\M)$ of $\M$} to consist of all words
-$w\in\Sigma^*$ such that
-the {\em initial configuration}
-$(s_0,\l w\r,\l 0^{f(|w|)}\r,0,0)$ reaches a configuration
-$(s_0,\l w\r,\l w'\r,i,j)$ satisfying $s\in F$.
-
-By $\L$, $\PSPACE$, and $\EXPSPACE$ we denote the class of all
-problems that can be decided by DTM that
-is logarithmically, polynomially, exponentially space bounded,
-respectively. Recall that $\Sigma^0_1$ (resp. $\Pi^0_1$) is the class of
-all languages that are (resp. whose complements are) recursively enumerable.
+
+%By $\L$, $\PSPACE$, and $\EXPSPACE$
+%we denote the class of all problems that can be decided by some deterministic
+%Turing machine that
+%logarithmically, polynomially, and exponentially space bounded, respectively.
 \newcommand{\Prop}{\mathcal{P}}
@@
 \renewcommand{\A}{\mathbb{A}}
-%\newcommand{\EF}{\mathsf{EF}}
-%\newcommand{\CTL}{\mathsh{CTL}}
-\bigskip
-
-\noindent
-\textbf{Transition systems: }
-In the following, we fix a countable set of atomic propositions $\Prop$.
+
+By $\Z$ we denote the {\em integers} and by $\N=\{0,1,2,\ldots\}$ the {\em naturals}.
+For each $i,j\in\Z$ we define $[i,j]=\{k\in\Z\mid i\leq k\leq j\}$ and $[j]=[1,j]$.
+For each $i,n\in\N$, let $\bit_i(n)$ denote the $i^{\text{th}}$ least significant bit of
+the binary representation of $n$. Hence $n=\sum_{i\in\N}2^i\cdot\bit_i(n)$.
+By $\bin_m(n)=\bit_0(n)\cdots \bit_{m-1}(n)$ we denote the first $m$ least significant
+bits written from {\em left to right}.
+%When $m$ is not important we just write $\bin(n)$.
+Let $p_i$ denote the $i^{\text{th}}$ prime number for each $i\geq 1$.
+We define $\log(n)=\min\{i\ge 1 \mid 2^i> n\}$, i.e.
+$\log(n)$ denotes the number of bits that are needed to represent $n$ in
+binary.
+%All polynomials $p:\N\rightarrow\N$ that occur in this paper, are assumed to satisfy
+%$p(n)\geq n$ for each $n\in\N$.
+For each word $v=a_1\cdots a_n\in\Sigma^n$ over some finite alphabet $\Sigma$
+and each $i,j\in[n]$ define $v[i,j]=a_i\cdots a_j$ and
+$v(i)=v[i,i]$.
+For the rest of the paper, we fix a countable set of {\em atomic propositions} $\Prop$.
 A {\em transition system} is a tuple $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, where
@@
 $\rightarrow \subseteq S\times S$ is a set of {\em transitions}.
-We prefer to use the infix notation $s_1\rightarrow s_2$ to abbreviate
+We prefer to use the infix notation $s_1\rightarrow s_2$ instead of
 $(s_1,s_2)\in\rightarrow$.
-%A {\em finite path} of length $n$ is sequence of transitions
-%$\pi=s_0\rightarrow s_1\cdots \rightarrow s_n$, where $n\in\N $.
  An {\em infinite path} is an infinite sequence
-$\pi=s_0\rightarrow s_1\cdots$.
+$\pi=s_0\rightarrow s_1\rightarrow \cdots$.
 % We denote
@@
 For each infinite path
-$\pi= s_0\rightarrow s_1 \rightarrow \cdots$ and
-$i\in\N$, we denote by $\pi^i$ the path $s_i\rightarrow s_{i+1}\cdots$
+$\pi= s_0\rightarrow s_1 \rightarrow \cdots$ and each
+$i\in\N$, we denote by $\pi^i$ the suffix $s_i\rightarrow
+s_{i+1}\rightarrow\cdots$
 and by $\pi(i)$ the state $s_i$.
-Define the {\em trace of $\pi$} as $\trace(\pi):\N\rightarrow2^\Prop$ where
-$\trace(i)=\{\prop\in\Prop\mid s_i\in S_\prop\}$ for each $i\in\N$.
-\bigskip
-
-\noindent
-\textbf{Succinct and parametric one-counter automata: }
-A {\em succinct one-counter automaton (SOCA)} is a tuple
+% Comment: Formal definition not needed
+%Define the {\em trace of $\pi$} as $\trace(\pi):\N\rightarrow2^\Prop$ where
+%$\trace(i)=\{\prop\in\Prop\mid s_i\in S_\prop\}$  for each $i\in\N$.
+%\noindent
+%\textbf{Succinct and parametric one-counter automata: }
+A {\em SOCA} is a tuple
  $\Soca=(Q,\{Q_\prop\mid \prop\in\Prop\},E,\lambda)$, where $Q$ is a finite set
@@
 $E\subseteq Q\times Q$ is a finite set of {\em transitions}, and
-$\lambda:E\rightarrow \Z\cup\{\zero\}$ labels the edges with decrements,
-increments, and zero tests.
-A {\em parametric one-counter automaton (POCA)} is a tuple
+$\lambda:E\rightarrow \Z\cup\{\zero\}$.
+We call a SOCA $\Soca$ with $\lambda:E\rightarrow
+\{-1,0,1\}\cup\{\zero\}$ a {\em unary one-counter automaton (OCA)}.
+A {\em POCA} is a tuple
 $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, where the first
 three components are same as for a SOCA,
-where $X$ is a finite set of {\em parameters over the natural numbers},
-and where $\lambda:E\rightarrow
-(\Z\cup\{\zero\}\cup\{\circ x\mid\circ\in\{+,-\}, x\in X\})$.
+$X$ is a finite set of {\em parameters over the naturals},
+and $\lambda:E\rightarrow
+(\Z\cup\{\zero\}\cup\{-x, +x\mid x\in X\})$.
 For each assignment $\sigma:X\rightarrow\N$  the
-{\em induced SOCA} is defined as $\Poca^\sigma=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda')$
-where $\lambda'(e)=\circ\sigma(x)$ whenever $\lambda(e)=\circ x$
-for some $\circ\in\{+,-\}$ and $\lambda'(e)=\lambda(e)$ otherwise.
-If $X=\{x\}$ is a singleton, then we also write $\Poca(x)$ for $\Poca(X)$.
+induced SOCA is defined as $\Poca^\sigma=(Q,E,\lambda')$
+where $\lambda'(e)=\sigma(x)$ if $\lambda(e)=x$,
+$\lambda'(e)=-\sigma(x)$ if $\lambda(e)=-x$, and
+$\lambda'(e)=\lambda(e)$ otherwise.
+%% $\lambda'(e)=\circ\sigma(x)$ whenever $\lambda(e)=\circ x$
+%% for some $\circ\in\{+,-\}$, $x\in X$, and $\lambda'(e)=\lambda(e)$ otherwise.
+If $X=\{x\}$ we also write $\Poca(x)$ instead of $\Poca(X)$.
 The {\em size} of a POCA is defined as
@@
 \prop\in\Prop\}, \rightarrow)$, where for each $q_1,q_2\in Q$ and each
-$n_1,n_2\in\N$ we have $q_1(n_1)\rightarrow q_2(n_2)$ if and only
-if either $\lambda(q_1,q_2)=n_2-n_1$, or
-($n_1=n_2=0$ and $\lambda(q_1,q_2)=\zero)$.
-
-\section{CTL Model Checking}{\label{S CTL}}
-
-In Section \ref{S CTL syn} we introduce the syntax and semantics of $\ctl$.
-Since the upper bounds for the modal $\mu$-calculus follow from
-known results and since the lower bounds follow from CTL, we do not introduce
-the modal $\mu$-calculus formally.
-In Section \ref{S Upper} we state an $\EXPSPACE$ upper bound for the combined
-complexity of $\ctl$ and the modal $\mu$-calculus on SOCA and a $\Pi^0_1$ upper bound
-on POCA. In Section \ref{S Serial} we recall basic notions from complexity theory
-such as serializability and results on Chinese remainder representation.
-In Section \ref{S CTL hardness} we prove the main result of this section, namely
-that the data complexity of $\ctl$ on SOCA is $\EXPSPACE$-hard.
-Finally, we prove a $\Pi^0_1$ lower bound for the data complexity
-of $\ctl$ on POCA in Section \ref{S CTL und}.
-
-\subsection{$\ctl$: Syntax and Semantics}{\label{S CTL syn}}
-\noindent
-Formulas $\varphi$ of $\ctl$ are given by the following
+$n_1,n_2\in\N$ we have $q_1(n_1)\rightarrow q_2(n_2)$ iff
+either $\lambda(q_1,q_2)=n_2-n_1$, or both
+$n_1=n_2=0$ and $\lambda(q_1,q_2)=\zero$.
+
+\section{$\CTL$ Model Checking}
+Formulas $\varphi$ of CTL are given by the following
 grammar, where $\prop$ ranges over $\Prop$:
-
-$$
+\begin{equation*}
 \varphi\quad ::=\quad \prop\ \mid \neg\varphi\ \mid\ \varphi\wedge\varphi\ \mid\
@@
 \E(\varphi\WU\varphi)
-$$
+\end{equation*}
+\iffalse
 
@@
 $\varphi_1\rightarrow\varphi_2=\neg\varphi_1\vee\varphi_2$,
-$\true=\prop\vee\neg\prop$ for some atomic proposition $\prop\in\Prop$,
-and $\EF\varphi=\E(\true\U\varphi)$.
-The {\em size} of a $\ctl$ formula is inductively defined as follows:
-$|\prop|=1$ for each $\prop\in\Prop$, $|\neg\varphi|=|\EX\varphi|=|\varphi|+1$,
-$|\varphi_1\wedge\varphi_2|=|\E(\varphi_1\U\varphi_2)|=|\E(\varphi_1\WU\varphi_2)|=
-|\varphi_1|+|\varphi_2|+1$.
+$\true=\prop\vee\neg\prop$ for some atomic proposition $\prop\in\Prop$, $\false=\neg\true$,
+$\EF\varphi=\E(\true\U\varphi)$, and $\AX\varphi=\neg\EX\neg\varphi$.
+%The {\em size} of a CTL formula is inductively defined as follows:
+%$|\prop|=1$ for each $\prop\in\Prop$, $|\neg\varphi|=|\EX\varphi|=|\varphi|+1$,
+%$|\varphi_1\wedge\varphi_2|=|\E\varphi_1\U\varphi_2|=|\E\varphi_1\WU\varphi_2|=
+%|\varphi_1|+|\varphi_2|+1$.
+\fi
 Given a transition system $T=(S,\{S_\prop\mid \prop\in\Prop\},\rightarrow)$, a state
@@
 induction on the structure of $\varphi$ as follows:
-
 \begin{eqnarray*}
-  (T,s)\models \prop &  \quad\iff\quad & s\in S_p\quad \text{ for each
+  (T,s)\models \prop &  \quad\iff\quad & s\in S_\prop\quad \text{ for each
 }\prop\in\Prop\\
@@
   (T,s)\models \E(\varphi_1\U\varphi_2) &  \quad\iff\quad &
-\exists s_0,\cdots,s_n\in S, n\geq 0: s_0=s, (T,s_n)\models\varphi_2 \text{ such
-that}\\
+\exists s_0,\ldots,s_n\in S, n\geq 0: s_0=s, (T,s_n)\models\varphi_2, \\
 && \forall i\in[0,n-1]: (T,s_i)\models\varphi_1\text{ and } s_i\rightarrow
@@
   (T,s)\models \E(\varphi_1\WU\varphi_2) &  \quad\iff\quad &
-(T,s)\models\E(\varphi_1\U\varphi_2)\ \text{ or }\\
-&& \exists s_0,s_1,\ldots\in S:\forall i\geq 0: (T,s_{i})\models\varphi_1\text{ and }
+(T,s)\models\E(\varphi_1\U\varphi_2)\ \text{ or } \exists s_0,s_1,\ldots\in
+S.\forall i\geq 0:\\
+&& s=s_0,(T,s_{i})\models\varphi_1\text{ and }
 s_i\rightarrow s_{i+1}
@@
 \noindent
-Let us define the {\em $\ctl$ model checking problem} on SOCA and POCA
-respectively.\\
-
-\problemx{$\ctl$ model checking on SOCA}
-{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
+Subsequently we use the standard abbreviations for disjunction, and
+implication.  Moreover, we define $\true=\prop\vee\neg\prop$ for some
+$\prop\in\Prop$ and $\EF\varphi=\E\ (\true\ \U\varphi)$.  Let us
+define the {\em CTL model checking problem} on SOCA and POCA
+resp.\\[-0.3cm]
+
+\problemx{CTL model checking on SOCA}
+{SOCA $\Soca=(Q,E,\lambda)$, $q\in Q$, $n\in\N$ in binary, a CTL formula
 $\varphi$.}
-{$(T(\Soca),q(n))\models\varphi$?}
-\
-
-\problemx{$\ctl$ model checking on POCA}
-{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and a $\ctl$ formula
+{Does $(T(\Soca),q(n))\models\varphi$?}
+\ \\[-0.4cm]
+
+\problemx{CTL model checking on POCA}
+{POCA $\Poca(X)=(Q,E,\lambda)$, $q\in Q$, $n\in\N$ in binary,  a CTL formula
 $\varphi$.}
-{$(T(\Poca^\sigma),q(n))\models\varphi$ for every $\sigma:X\rightarrow\N$?}
-
-\subsection{Upper bounds}{\label{S Upper}}
-Let us first state upper bounds on model
-checking CTL and the modal $\mu$-calculus on SOCA.
- {\em One-counter automata (OCA)} are SOCA in which
- the numbers that occur in the transition labels are represented  in unary
- instead in binary.
-%In this section we state the upper bounds on model checking $\ctl$ on SOCA and
-%POCA. They follow from known results or are immediate.
-The following theorem gives an upper bound on model checking the
-{\em modal $\mu$-calculus} on OCA.
-Consult \cite{ArnNiw01} for more details on the modal $\mu$-calculus.
-
-\begin{theorem}[\cite{Serr06}]
-The combined complexity of the modal $\mu$-calculus on OCA is in $\PSPACE$.
-\end{theorem}
-
-\noindent
-Since every SOCA can be transformed into an exponentially larger OCA
- and since each $\ctl$ formula can be translated into an alternation-free
-$\mu$-calculus formula with a linear blowup, the following corollary is
-immediate.
-
-\begin{corollary}{\label{C CTL up}}
-The combined complexity of $\ctl$ and the modal $\mu$-calculus on SOCA is in $\EXPSPACE$.
-\end{corollary}
-
-\noindent
-The following upper bound for $\ctl$ on POCA is straightforward.
-
-\begin{corollary}
-The combined complexity of $\ctl$ and the modal $\mu$-calculus on POCA is in $\Pi^0_1$.
-\end{corollary}
-\begin{proof}
-Let $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$ be a POCA,
- $q\in Q$, and $n\in\N$, and $\varphi$ be
-a formula.
-From Corollary \ref{C CTL up} it follows that
-for each given $\sigma:X\rightarrow\N$ the question
-$
-(T(\Poca^\sigma),q(n))\models\varphi
-$
-is a decidable predicate.
-Hence, the question if
-$
-(T(\Poca^\sigma),q(n))\models\varphi
-$
-for each $\sigma:X\rightarrow\N$ is a $\Pi^0_1$-predicate since
-we can encode the set of all such assignments into the naturals
-via G\"odel encoding.
-\qed
-\end{proof}
-
-\subsection{Serializability and Chinese remainder representation}{\label{S Serial}}
-For a language $L\subseteq\Sigma^*$ let $\chi_L:\Sigma^*\rightarrow\{0,1\}$
-denote the {\em characteristic function of $L$}, i.e.
-$\chi_L(x)=1$ if $x\in L$ and $\chi_L(x)=0$ otherwise, for each $x\in\Sigma^*$.
+{Does $(T(\Poca^\sigma),q(n))\models\varphi$ for every $\sigma:X\rightarrow\N$?}
+
+\subsection{Upper Bounds}{\label{S Upper}}
+\noindent
+%\textbf{Upper bounds:} First, let us state the upper bounds on model
+%checking CTL on SOCA and
+% Comment: Moved to preliminaries
+%% {\em One-counter automata (OCA)} are SOCA in which
+%% the numbers that occur in the transition labels are represented  in unary
+%% instead in binary.
+%% The following theorem gives an upper bound on model checking the
+%% {\em modal $\mu$-calculus} on OCA.
+Due to space restrictions we do not formally introduce the modal
+$\mu$-calculus and refer to \cite{ArnNiw01} for more details instead.
+In \cite{Serr06} Serre showed that the combined complexity of the
+modal $\mu$-calculus on OCA is in $\PSPACE$.  Since every SOCA can be
+transformed into an OCA of exponential size, and since each CTL
+formula can be translated into an alternation-free $\mu$-calculus
+formula with a linear blowup, the following proposition is immediate
+by adjusting the resulting $\mu$-calculus formula appropriately.
+Moreover, this immediately implies the containment in the arithmetic
+hierarchy of the combined complexity of the modal $\mu$-calculus and
+CTL on POCA.
+
+%a $\Pi^0_1$ upper bound for the combined complexity of
+%the modal $\mu$-calculus and CTL on POCA.
+\begin{proposition}{\label{P CTL up}}
+For the modal $\mu$-calculus and CTL the combined complexity on SOCA is in
+$\EXPSPACE$, whereas it is in $\Pi^0_1$ on POCA.
+\end{proposition}
+
+\subsection{Hardness of the Data Complexity of CTL on SOCA}{\label{S CTL hardness}}
+Before we prove $\EXPSPACE$-hardness of the data complexity of CTL on SOCA, we
+introduce some notions and results from complexity theory.
+Given a language $L\subseteq\Sigma^*$, let $\chi_L:\Sigma^*\rightarrow\{0,1\}$
+denote the {\em characteristic function of $L$}.
 We define the {\em lexicographic
-order on $n$-bit strings} as $\preceq_n\subseteq\{0,1\}^n\times\{0,1\}^n$, where
-$x\preceq y$ if and only if
-$\bin_n(x)\leq\bin_n(y)$, e.g. $101\preceq_3 011$.
-
+order on $n$-bit strings}
+% Comment: obsolete
+%as $\preceq_n\subseteq\{0,1\}^n\times\{0,1\}^n$, where
+by $x\preceq_n y$ if and only if
+$\bin_n(x)\leq\bin_n(y)$, e.g. $011 \preceq_3 101$.
 \renewcommand{\C}{\mathcal{C}}
-\begin{definition}
-Let $\C$ be a complexity class.
-We say a language $L$ is {\em $\C$-serializable via some language
-$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$, and some language
-$U\in\C$ such that for all $x\in\{0,1\}^n$
+We say a language $L$ is {\em exponentially $\C$-serializable via some language
+$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$ and some language
+$U\in\C$ s.t.\ for all $x\in\{0,1\}^n$\\[-0.2cm]
 $$x\in L\qquad\!\!\iff\qquad\!\!
-\chi_U\left(x,0^{p(n)}\right)\quad \cdots\quad \chi_U\left(x,1^{p(n)}\right)\in
-R,$$ where with `$\cdots$'  we
-refer to $\preceq_{p(n)}$ in each of the second components.
-\end{definition}
-
-\noindent
-% it was shown that $\PSPACE$ is $\mathsf{P}$-serializable;
-%in \cite{HLSVW93} this result was sharpened to $\AC^0$-serializability,
-%see also \cite{Vol98}. It is not stated in \cite{HLSVW93,Vol98} but easy to
-%see from the proofs that \emph{logspace-uniform} $\AC^0$ suffices for serializing
-%$\PSPACE$, see the appendix for more details.
-The following theorem was proven in \cite{GoLo09}.
-
-\begin{theorem}[Theorem 22 in \cite{GoLo09}]{\label{T Serial}}
-For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
-logspace-uniformly $\mathsf{AC}^0$-serializable via $R$.
-\end{theorem}
-
-\noindent
-As an immediate consequence, we obtain the following corollary.
-
-\begin{corollary}{\label{C Serial}}
-For every $L$ in $\PSPACE$ there is some regular language $R$ such that $L$ is
-$\L$-serializable via $R$.
-\end{corollary}
-
-\noindent
-We remark that our definition of serializability is adopted from \cite{GoLo09}
-and differs slightly from the standard
-notion as used in \cite{CaFu91,HLSVW93,Vol98}.
-We aim at lifting Corollary \ref{C Serial} via an appropriate notion of
-serializability to $\EXPSPACE$.
-
-\begin{definition}
-Let $\C$ be some complexity class.
-A language $L$ is {\em exponentially $\C$-serializable via some language
-$R\subseteq\{0,1\}^*$} if there is some polynomial $p(n)$ and some language
-$U\in\C$ such that for all $x\in\{0,1\}^n$
-$$x\in L\qquad\!\!\iff\qquad\!\!
-\chi_U\left(x,0^{2^{p(n)}}\right)\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right)\in
-R,$$ where with `$\cdots$' we
-refer to $\preceq_{2^{p(n)}}$ in each of the second components.
-\end{definition}
-
-\noindent
-The following proposition is folklore and immediate.
-\begin{proposition}{\label{P Padding}}
-Let $L\subseteq\Sigma^*$ be in $\EXPSPACE$. Then there is some polynomial $q$
-such that the padded language
-$\{x\$^{2^{q(|x|)}}\mid x\in L\}\subseteq(\Sigma\uplus\{\$\})^*$ is in $\PSPACE$.
-\end{proposition}
+\chi_U\left(x \cdot 0^{2^{p(n)}}\right)\quad \cdots\quad \chi_U\left(x
+\cdot 1^{2^{p(n)}}\right)\in
+R,\\[-0.1cm]
+$$ where the bit strings on the right-hand side of the
+concatenation symbol are enumerated in
+lexicographic order.
+%% by `$\cdots$' we
+%% refer to $\preceq_{2^{p(n)}}$ in each of the second components.
+This definition is a padded variant of the serializability notion used
+in \cite{GoLo09}, which in turn is a variant of the serializability notion
+from \cite{CaFu91,HLSVW93,Vol98}.
+Some subtle technical padding arguments are required
+to lift $\AC^0$-serializability of $\PSPACE$, proven in
+Theorem 22 in \cite{GoLo09}, to exponential
+$\L$-serializability of $\EXPSPACE$.%, for proving the following theorem.
 
@@
 \end{theorem}
-\begin{proof}
-Let $L\subseteq\Sigma^*$ be some language in $\EXPSPACE$.
-Then by Proposition \ref{P Padding} for some polynomial $q$ the language
-$L'=\{x\$^{2^{q(|x|)}}\mid x\in L\}$ is in $\PSPACE$.
-Due to $\L$-serializability of $\PSPACE$ by Corollary \ref{C Serial}, there exists
-some polynomial $p'$, some regular language $R'$ and some $U'\in\L$
- such that for each $y\in(\Sigma\uplus\{\$\})^*$ we have
-\begin{eqnarray}
-y\in L'\quad\iff\quad \chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
-\chi_{U'}\left(y,1^{p'(|y|)}\right)\in R', \label{E Serial}
-\end{eqnarray}
-where `$\cdots$' refers to the lexicographic order $\preceq_{p'(|y|)}$.
-To prove the lemma we have to find some regular language $R$, some
-polynomial $p$ and some $U\in\L$
-such that for all $x\in\Sigma^*$ we have
-$$
-x\in L\quad \iff\quad \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \cdots\
-\chi_U\left(x,1^{2^{p(|x|)}}\right)\in R,
-$$
-where here `$\cdots$' refers to the lexicographic order
-$\preceq_{2^{p(|x|)}}$.
-For now, we choose $p$ sufficiently fast growing,
-let us postpone this to the end of the proof.
-Let us describe the language $U$ and consider inputs $(x,w)$
- with $x\in\Sigma^n$ and $w\in\{0,1\}^{2^{p(n)}}$;
-moreover let $y=x\$^{2^{q(|x|)}}$ be the padding of $x$.
-Our language $U$ consists of all such pairs $(x,w)$
-such that
-$w$ can be factorized as $w=bz0^l$ for some $b\in\{0,1\}$
-and some $l\geq 0$ such that
-$b=0$ or $(y,z)\in U'$.
-Roughly speaking, the goal of $U$ is to simulate $U'$
-on the corresponding padded word,
-however the length of the second component
-$w$ has to be a power of two, in other words the last $l$ bits of $w$
- contain redundant information.
-We will be able to filter out this redundant information via
-the bit $b$.
-Let us make this more precise.
-Let
-\begin{eqnarray*}
-\gamma' &\quad=\quad&
-\chi_{U'}\left(y,0^{p'(|y|)}\right)\ \ \cdots\ \
-\chi_{U'}\left(y,1^{p'(|y|)}\right)\quad\text{and}\\
-\gamma &\quad=\quad& \chi_U\left(x,0^{2^{p(|x|)}}\right)\ \ \cdots\ \
-\chi_U\left(x,1^{2^{p(|x|)}}\right).
-\end{eqnarray*}
-
-\noindent
-Then by definition of $U$, the first $2\cdot p'(|y|)$ bits of $\gamma$ are
-$1\gamma'(1)1\gamma'(2)\cdots 1\gamma'(p'(|y|))$, where
-the remaining bits of $\gamma$ are all $0$.
-Hence, when reading $\gamma$ in {\em pairs of bits},
-precisely when the first of the two bits
-is set to $1$ we read some relevant information.
-
-We have to give some regular language $R$ such that $\gamma'\in R'$ if and only
-if $\gamma\in R$.
-Recall that regular languages are closed under shuffle product $|\!|$ and
-homomorphisms.
-We define $R=\varphi(R'|\!|\{a\}^*)$, where
-$\varphi:\{0,1,a\}\rightarrow\{0,1\}^*$ is the following homomorphism:
-$\varphi(a)=00$, $\varphi(0)=10$, and $\varphi(1)=11$.
-
-As expected, we choose $p$ to be an arbitrary polynomial satisfying
-$2^{p(n)}\geq 1+p'(2^{q(n)}+n)$ for all $n\geq0$.
-Finally, let us show that $U$ is in $\L$.
-Let $(x,w)$ be an input to $U$ and let again $y=x\$^{2^{q(|x|)}}$ be the padding
-of $x$.
-Then it is straightforward to decide in logarithmic space if $w$ can be factorized as
-$w=bzv$ such that $b\in\{0,1\}$, $|z|=p'(2^{q(|x|)}+|x|)$ and
-$v\in\{0\}^*$.
-The only thing that might remain to be verified in logarithmic space is whether
-$(y,z)\in U'$, which in turn boils down to simulating some logarithmic space bounded TM
-for $U$ on input $(y,z)$.
-But this is possible since $|y|+|z|\leq 2\cdot(|x|+|w|)$ as
-$$
-|y|+|z|=|x|+2^{q(|x|)}+p'(2^{q(|x|)+|x|})\leq 2\cdot p'(2^{q(|x|)}+|x|)\leq
-2\cdot 2^{p(|x|)}\leq 2\cdot (|x|+|w|).
-$$
-\qed
-\end{proof}
-
 \renewcommand{\M}{\mathcal{M}}
-
-\noindent
-\textbf{Chinese remainder representation: }
- For every $m,M\in\N$ we denote by $\CRR_m(M)$ {\em the Chinese remainder
+%\noindent
+%\textbf{Chinese remainder representation: }
+A further concept we use is the Chinese remainder
+representation of a natural number. For every $m,M\in\N$ we denote by $\CRR_m(M)$ {\em the Chinese remainder
 representation} of $M$ as the Boolean tuple
@@
 The following theorem tells us that in logarithmic space we can
-compute the binary representation of a natural number that is given in Chinese
+compute the binary representation of a natural number from its Chinese
 remainder representation.
-It is a consequence of the result that division is in logspace-uniform
-$\mathsf{NC}^1$, proven in \cite{ChDaLi01}.
+This result is a consequence of \cite{ChDaLi01}, where it is shown that division is in logspace-uniform
+$\mathsf{NC}^1$.
 
@@
 
-QUESTION: $\bit_j(M\ \modulo 2^m)=b$?
+QUESTION: Is $\bit_j(M\ \modulo 2^m)=b$?
 \end{theorem}
-
-\subsection{$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA}{\label{S CTL hardness}}
-
-In the rest of this section, we give the proof of $\EXPSPACE$-hardness
-of the data complexity of $\ctl$ on SOCA.  Let $L\subseteq\{0,1\}^*$ be an arbitrary
-language in $\EXPSPACE$. Then by Theorem \ref{T succ seri}, there is
-some regular language $R\subseteq\{0,1\}^*$ such that $L$ is
-exponentially $\L$-serializable via $R$. Hence there is some language $U\in\L$
-and some polynomial $p$ such that for all $x\in\{0,1\}^n$ we have
-
+In the rest of this section, we sketch the proof of $\EXPSPACE$-hardness of
+the data complexity of CTL on SOCA.
+Let $L\subseteq\{0,1\}^*$ be an arbitrary language in $\EXPSPACE$.
+Then by Theorem \ref{T succ seri}, there is some regular
+language $R\subseteq\{0,1\}^*$ s.t.\ $L$ is exponentially
+$\L$-serializable via $R$. Hence there is some
+language $U\in\L$ s.t.\ for all $x\in\{0,1\}^n$ we have\\[-0.2cm]
 \begin{eqnarray}{\label{E EXPSPACE char}}
-x\in L\qquad\iff\qquad \chi_{U}\left(x,0^{2^{p(n)}}\right)
-\quad \cdots\quad \chi_U\left(x,1^{2^{p(n)}}\right) \in R,
+x\in L\qquad\iff\qquad \chi_{U}\left(x \cdot 0^{2^{p(n)}}\right)
+\quad \cdots\quad \chi_U\left(x \cdot 1^{2^{p(n)}}\right) \in R,
 \end{eqnarray}
-where '$\cdots$' refers to the lexicographic order $\preceq_{2^{p(n)}}$
-on the bit strings on the right-hand side.
-
-\noindent
-For the rest of this section, let us fix an input $x_0\in\{0,1\}^n$.
-Let $N=p(n)$ and let $A=(Q,\{0,1\},q_0,\delta,F)$ be some
-deterministic finite automaton with $L(A)=R$.  Let us briefly recall
-what $A$ consists of: $Q$ is a finite set of {\em states}, $\{0,1\}$
-is the {\em input alphabet}, $q_0\in Q$ is the {\em initial state},
-$\delta:Q\times\{0,1\}\rightarrow Q$ is the {\em transition function},
-and $F\subseteq Q$ is the set of {\em final states}.  Let us describe
-equivalence (\ref{E EXPSPACE char}) differently: We have $x_0\in L$ if
-and only if the program in Figure \ref{F U} returns $\mathtt{true}$.
-
-\begin{figure}
+\noindent
+where the bit strings on the right-hand side of the concatenation
+symbol are enumerated in lexicographic order.  For the rest of this
+section, let us fix an input $x_0\in\{0,1\}^n$.  Let $N=p(n)$ and
+$A=(Q,\{0,1\},q_0,\delta,F)$ be some deterministic finite automaton
+with $L(A)=R$.  Let us describe equivalence (\ref{E EXPSPACE char})
+differently: We have $x_0\in L$ iff the program in Fig.\ \ref{F prog}
+returns $\mathtt{true}$.  We are going to mimic the execution of the
+program by a fixed CTL formula and a SOCA that can be computed from
+$x_0$ in logarithmic space.
+\begin{wrapfigure}{r}{0.3\textwidth}
+%  \begin{center}
 \normalsize
-
-\ \hspace{5cm} $q\in Q$; $q:=q_0$;\\[-0.4cm]
-
-\ \hspace{5cm} $d\in\N$; $d:=0$;
-
-\ \hspace{5cm} $b\in\{0,1\}$;\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{while }\  d\not=2^{2^N}$\ $\mathtt{ loop}$\\[-0.4cm]
-
-\ \hspace{6cm} $b:=\chi_U(x_0,\text{bin}_{2^N}(d))$;\\[-0.4cm]
-
-\ \hspace{6cm} $q:=\delta(q,b)$;\\[-0.4cm]
-
-\ \hspace{6cm} $d:=d+1$;\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{end loop}$\\[-0.4cm]
-
-\ \hspace{5cm} $\mathtt{return}\ q\in F$;%\\[0.4cm]
-
-\caption{A program that returns $\mathtt{true}$ if and only if $x_0\in L$}{\label{F U}}
-\end{figure}
-
-\noindent
-Our goal is to mimic the execution of the program in Figure \ref{F U} by
-a fixed $\ctl$ formula and by a SOCA that depends on
-our input $x_0$.
+\vspace{-0.5cm}
+ $q\in Q$; $q:=q_0$;\\[-0.4cm]
+
+ $d\in\N$; $d:=0$;
+
+$b\in\{0,1\}$;\\[-0.4cm]
+
+ $\mathtt{while }\  d\not=2^{2^N}$\ $\mathtt{ loop}$\\[-0.4cm]
+
+$\ \ \ \ b:=\chi_U(x_0\cdot\text{bin}_{2^N}(d))$;\\[-0.4cm]
+
+ $\ \ \ \ q:=\delta(q,b)$;\\[-0.4cm]
+
+$\ \ \ \ d:=d+1$;\\[-0.4cm]
+
+$\mathtt{end loop}$\\[-0.4cm]
+
+ $\mathtt{return}\ q\in F$;%\\[0.4cm]
+
+ % \end{center}
+\caption{A program that returns $\mathtt{true}$ iff $x_0\in L$.}{\label{F
+U}}{\label{F prog}}
+\vspace{-0.3cm}
+\end{wrapfigure}
 Before we start with the reduction, let us discuss the obstacles
@@
 
-\begin{enumerate}[{\bf(A)}]
-\item We need some way of storing $d$ on the counter. Of course
+\textbf{(A)} We need some way of storing $d$ on the counter. Of course
 there are a lot of ways to do this, but since we want to access
-all bits of $d$ in the assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$, the most
+all bits of $d$ in the assignment $b:=\chi_U(x_0\cdot \bin_{2^N}(d))$, the most
 natural way is probably to represent $d$ in binary.
@@
 requires exponentially many bits in $n$.
-\item As in \cite{GoLo10}, a solution to obstacle {\bf (A)} is to store $d$ in
+\textbf{(B)} As in \cite{GoLo10}, a solution to obstacle (A) is to store $d$ in
 Chinese remainder representation with the first $2^N$ prime numbers.
@@
 each prime number, simply because there are too many of them.
-\item
-The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ implies that we need to simulate on
+\textbf{(C)}
+The assignment $b:=\chi_U(x_0\cdot \bin_{2^N}(d))$ implies that we need to simulate on
 the counter a logarithmically space bounded DTM for the language $U$
@@
 some on-the-fly mechanism for reading the input.
-\end{enumerate}
-
-\iffalse
-
-\begin{enumerate}[(1)]
-\item For the representation of the variable $d$ we need $2^N+1$ bits
-  on our counter. More challenging, we need to check if $d$ is
-  equal to $2^{2^N}$. To check this, we cannot simply allow a
-  transition in our counter automaton that subtracts $2^{2^N}$,
-  since the representation of this number requires exponentially many
-  bits in $n$.
-\item As in \cite{GoLo10}, a solution to problem (1) is to store $d$
-  in Chinese remainder representation with the first $2^N+1$ primes.
-  Polynomially (in $n$) many bits suffice to represent each of the
-  occurring prime numbers, but we need exponentially many different
-  prime numbers, where each of the primes has up to exponentially many
-  residue classes that we cannot simply write down. So we cannot equip
-  the SOCA that we have to derive with transitions for each prime
-  number, simply since there are too many of them.
-
-\item The assignment $b:=\chi_U(x_0,\bin_{2^N}(d))$ means that we need to simulate on
-our counter a logarithmically space bounded DTM for $U$ on an exponential
-sized input (in $n$).
-Speaking in terms of our input size $n$, this means that we need to provide
-polynomially many bits on our counter that can be used to describe the
-working tape for this DTM. However, we need to provide
-some on-the-fly mechanism for reading the input.
-\end{enumerate}
-
-\fi
-\noindent
-To achieve this, let us give a high-level description of how we proceed.
-In a first step, we carefully design a data structure on the counter
-and describe the intuition behind it.
-In a second step, we give five queries which we aim at implementing
-via fixed CTL formulas and by SOCA that can be computed from $x_0$
+
+\noindent
+Let us give a high-level description of the $\EXPSPACE$-hardness proof.
+In the first step, we carefully design a data structure on the counter
+and explain the intuition behind it.
+In the second step, we list five queries which we aim at implementing via
+fixed CTL formulas and by SOCA that can be computed from the input $x_0$
 in logarithmic space.
 
-\bigskip
+\smallskip
 
@@
  Let $K=n+2^N+1$ denote the number of bits that are required to store an input
-for $U$ plus one.
+for $U$.
 Let $\alpha=\log K$ denote the number of bits that we require for
@@
 the Prime Number Theorem, it follows that
-$\beta\in O(\log(K\log(K)))=O(N)$.
+$\beta= O(\log(K\log(K)))=O(N)$.
 The number $\alpha$ and such a sufficiently large number $\beta$ can be computed
@@
 
-Let us describe how we will interpret the counter in our reduction.
-Assume that the counter value is $v\in\N$.  Instead of
-treating $v$ as a natural number, we are interested only in the $l$
-least significant bits $V$ of the binary representation of $v$, where $l$
-is some number that is exponentially bounded in $n$; the precise value
-of $l$ will be made clear below.  Assume
-$V=\bit_0(v)\cdots\bit_{l-1}(v)$.  We read $V$ to be factorized into
-blocks of bits
-$$\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad V\quad=\quad I\ M\ C\ J\ X\ Y\ Z\ B,\qquad\qquad\qquad\qquad\qquad\qquad(\bigstar)$$ where
-\begin{itemize}
-\item $I\in\{0,1\}^{\alpha}$ represents a prime number index,
-\item $M\in\{0,1\}^\beta$ represents $I^{\text{th}}$
-prime number $p_I$,
-\item $C\in\{0,1\}^\beta$ represents some residue class modulo $M$,
-\item $J\in\{0,1\}^\alpha$ represents a pointer to some bit of $B$,
-\item
-each $X,Y$ and $Z$ consist of polynomially many bits (in $n$) and
-represent the working tape of three space bounded
-DTMs that we will comment on later in more detail, and
-\item $B\in\{0,1\}^{n+2^N+1}$ with $B=xB'$ for some
-$x\in\{0,1\}^n$ and some
-$B'\in\{0,1\}^{2^N+1}$,
-$B$ represents the current
-input for $U$, where
-$x$ is reserved to represent our input $x_0$
-and
-where $B'$ represents the counter $d$ from program in Figure \ref{F U} from above.
-The block $B'$ consists of $2^N+1$ bits since we want to be able to test
- if $d=2^{2^{N}}$.
-\end{itemize}
-
-\renewcommand{\Re}{\Rightarrow}
-\newcommand{\Li}{\Leftarrow}
-\newcommand{\re}{\rightarrow}
-\newcommand{\li}{\leftarrow}
-
-\noindent
-Let us introduce some more notation for addressing
-leftmost (starting to count from $0$) and rightmost bit
-positions in each of the above sequences of bits in $V$.
-For each such sequence $\Theta$ let $\Theta_{\Li}$ and $\Theta_{\Re}$ denote the
-respective positions of the leftmost and rightmost bits of $\Theta$ within the bit
-string $V$, e.g. $I_\Re=\alpha-1$ and $C_\Li=\alpha+\beta$.
-
-\medskip
-
-\noindent
-{\bf Important remark: } Throughout the rest of this section $v$ will
-denote an arbitrary natural number, moreover $I,M,C,J,X,Y,Z$, and $B$
-will implicitly be coupled with $v$.  Note that all of the bit strings
-have polynomial length in $n$ except for $B$.  Moreover, we identify each
-of the blocks with the natural number they represent.
-
-\medskip
-
-\noindent
-A very simple but important gadget that we need is to decide, for each
-$b\in\{0,1\}$ if the $i^{\text{th}}$ bit of $v$ is $b$.
-
-\begin{lemma}{\label{L bit}}
-For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\bit,b}$
-such that the following is computable in logarithmic space:
-
-INPUT: $i\in\N$ in unary.% and our input $x_0$.
-
-OUTPUT: A SOCA $\Soca_{\bit,i}$ and some control state $q_{\bit,i}$ such that
-$$
-(T(\Soca_{\bit,i}),q_{\bit,i}(v))\models\varphi_{\bit,b}\quad\Longleftrightarrow\quad
-\bit_i(v)=b.
-$$
-\end{lemma}
-\begin{proof}
-The SOCA $\Soca_{\bit,i}$ has the two atomic propositions $\tau_1$ and $\tau_2$ and is
-depicted below:
-
+\begin{wrapfigure}{r}{0.4\textwidth}
 \begin{center}
-\begin{picture}(85,20)(-20,-5)
-%   \put(-20,-28){\framebox(85,33){}}
-  %\gasset{Nadjust=w,Nadjustdist=2,Nh=6,Nmr=1}
- % \gasset{Nadjust=w,Nadjustdist=2,Nh=6}%$,Nmr=1}
- % \node(A)(0,0){$q_0$}
+\begin{picture}(85,13)(5,-10)
 
@@
 \drawedge(A,B){$-2^i$}
+
+\put(4,-10){
+\scriptsize
+$%\tau_1\wedge
+\varphi_{\bit,1}=\tau_1\wedge\EF(\tau_1\wedge\neg\EX \tau_1\wedge\EX
+\tau_2)$
+}
+\put(4,-14){
+\scriptsize
+$%\tau_1\wedge
+\varphi_{\bit,0}=\tau_1\wedge\EF(\tau_1\wedge\neg\EX \tau_1\wedge\neg\EX
+\tau_2)$
+}
+
 \end{picture}
 \end{center}
-The simple way to check if $\bit_i(v)=1$ (resp. $\bit_i(v)=0$) is to
-repeatedly subtract $2^{i+1}$ from $v$ until no longer possible
-and then to check if $2^i$ can (resp. cannot) be subtracted.
-Hence we put
-\begin{eqnarray*}
-\varphi_{\bit,1}&\quad=\quad&
-\tau_1\wedge
-\EF(\tau_1\wedge\neg\EX \tau_1\wedge\EX
-\tau_2)\qquad\text{ and}\\
-\varphi_{\bit,0}&=&
-\tau_1\wedge
-\EF(\tau_1\wedge\neg\EX \tau_1\wedge\neg\EX
-\tau_2).
-\end{eqnarray*}
-\qed
-\end{proof}
-
-\noindent
-{\bf Queries that we need to implement: }
-We will implement the following five queries Q1 to Q5 by instances
-of the model checking problem, where each query builds on top
-of its preceding queries.
-
+
+\caption{SOCA $\Soca_{\bit,i}$ and CTL formulas
+$\varphi_{\bit,b}$ for checking if $\bit_i(v)=b$.}  {\label{F Bit}}
+\end{wrapfigure}
+Let us describe the data structure on the counter in our reduction.
+Assume that the counter value is $v\in\N$.
+We are interested only in the $l$ least
+significant bits of the binary representation of $v$, where $l$ is some number
+that is exponentially bounded in $n$; the
+value of $l$ will be made clear below.
+Assume $V=\bit_0(v)\cdots\bit_{l-1}(v)$.
+We imagine $V$ to be factorized into blocks of bits\\[-0.4cm]
+ $$V\ \ =\ \ I\ M\ C\ J\ X\ Y\ Z\ B\quad\qquad(*)$$
+where
+$I\in\{0,1\}^{\alpha}$ is a prime number index;
+$M\in\{0,1\}^\beta$ is intended to represent the $I^{\text{th}}$
+prime number $p_I$;
+$C\in\{0,1\}^\beta$ is some residue class
+modulo $M$;
+ $J\in\{0,1\}^\alpha$ represents a pointer to some bit of $B$;
+$X,Y$ and $Z$ consist of polynomially many bits (in $n$) and are
+ intended to represent the working tape of three space-bounded
+DTMs that we will comment on later in more detail; and
+$B\in\{0,1\}^{n+2^N}$ with $B=x_0B'$ for some
+$B'\in\{0,1\}^{2^N}$.
+Our intention is that $B$ represents the current input for $U$, and
+in particular
+$B'$ represents the counting variable $d$ from Fig. \ref{F prog}.
+%The block $B'$ consists of $2^N+1$ bits since we want to be able to test
+% if $d=2^{2^{N}}$.
+\noindent
+Throughout the rest of this section, $v$
+will denote an arbitrary natural number.
+Moreover
+$I,M,C,J,X,Y,Z$ and $B$ will implicitly be coupled with $v$ via the
+factorization $(*)$.
+Note that all of the bit strings have polynomial length in $n$ except for
+$B$.
+Subsequently, we identify each of the blocks with the natural number they represent. A simple but powerful gadget, which will subsequently be used
+to check for each $b\in\{0,1\}$ if the $i^{\text{th}}$ bit
+of the counter is $b$, is
+shown in Fig.\ \ref{F Bit}. We have that
+$q_{\bit,i}(v)$ satisfies $\varphi_{\bit,b}$ iff $\bit_i(v)=b$,
+for each $b\in\{0,1\}$.
+
+%
+\renewcommand{\Re}{\Rightarrow}
+\newcommand{\Li}{\Leftarrow}
+\newcommand{\re}{\rightarrow}
+\newcommand{\li}{\leftarrow}
+%
+%\noindent
+\iffalse
+Let us introduce some more notation for addressing
+leftmost (starting to count from $0$) and rightmost bit
+positions in each of the above sequences from $(*)$.
+For each such sequence $\Theta$ let $\Theta_\Li$ denote
+the {\em absolute position of the leftmost bit} of $\Theta$ in $V$
+and let $\Theta_\re$ denote
+the {\em position of the rightmost bit of $\Theta$ relative
+to $\Theta_\Li$}, i.e. $\Theta_\re=|\Theta|-1$.
+For example $C_\Li=\alpha+\beta$ and $J_\re=\alpha-1$.
+\fi
+%
+
+\smallskip
+\noindent
+{\bf Queries that we need to implement: } Next, we list five queries that we aim
+at answering by instances of the model checking problem.
+Each query is based on its preceding queries.
 \begin{enumerate}[(Q1)]
-\item \fbox{When assuming $C<M$, does $B\equiv C\ \mod M$ hold?}
-\item \fbox{Is $M$ the $I^{\text{th}}$ prime number, i.e. $M=p_I$?}
-%\item Let $\CRR_{2^\alpha}(B)=(b_{i,r})_{i\in[2^\alpha],0\leq r<p_i}$.
-%What is $b_{I,R}$? %resp. $b_{I,R}=1$ hold?
-\item \fbox{What is $\bit_J(B)$?}
-\item \fbox{Does $(B[1,n],B[n+1,n+2^N])\in U$ hold?}
-%\item Does $B'=2^{2^N}$ hold?
-\item \fbox{Does $x_0\in L$ hold?}
+\item  Assuming $C<M$, does $C\equiv B\ \mod M$ hold?
+\item  Is $M$ the $I^{\text{th}}$ prime number, i.e. $M=p_I$?
+\item  What is $\bit_J(B)$?
+\item  Does $(B[1,n] \cdot B[n+1,n+2^N])\in U$ hold?
+\item  Does $x_0\in L$ hold?
 \end{enumerate}
-
-\noindent
-We implement each of the five queries by providing fixed CTL formulas
-and SOCA that can be computed from $x_0$ in logarithmic space.
-$\EXPSPACE$-hardness of the data complexity of CTL on SOCA will hence
-follow from the implementation of Q5.
-First, let us give an implementation of Q1.
-
-\begin{lemma}{\label{L mod}} There exists some fixed $\ctl$ formula $\varphi_\mod$ such that we
-can compute from $x_0$ in logarithmic space some SOCA $\Soca_\mod$ and some
-control state $q_\mod$ such that
-$(T(\Soca_{\mod}),q_\mod(v))\models\varphi_\mod$ if and only if $B\equiv C\ \mod M$.
+\noindent
+$\EXPSPACE$-hardness of data complexity of CTL on SOCA will hence follow from the
+implementation of query Q5.
+%Before we give the implementation of Q1,
+Let $\gamma=O(N)$ denote the absolute
+value of the leftmost bit position of $B$ in $V$.
+Hence, when the $\gamma^{\text{th}}$ bit of $V$ is set to $1$ and we subtract
+$2^\gamma$ from the counter, then the leftmost bit of $B$ is set to $0$.
+Similarly, let $\mu$ denote the leftmost bit position of $M$ in $V$.
+Q1 can now be realized as follows.
+
+\begin{lemma}{\label{L mod}} There is a  CTL formula $\varphi_\mod$ s.t.
+we can compute from $x_0$ in logarithmic space a SOCA $\Soca_\mod$ and a
+control state $q_\mod$ s.t.
+$(T(\Soca_{\mod}),q_\mod(v))\models\varphi_\mod$ iff $C\equiv B\ \mod M$.
 \end{lemma}
 \begin{proof}
-The SOCA $\Soca_{\mod}$ contains the three atomic propositions $\prop_0,\prop_1$, and
-$\prop_2$ and is depicted in Figure \ref{F Smod}.
-The $\ctl$ formula $\varphi_\mod$ expresses that we
-traverse the upper
-sequence of diamonds and thereby repeatedly subtract $M$ from $B$.
-The number of diamonds both in the first and the second row equals $\beta$,
-the number of bits of $M$ and of $C$.
-In the upper row, one diamond corresponds to one bit of $M$. In case bit the rightmost bit of $M$
-(in other words bit $M_\Re$ of the counter) is $1$,
+For brevity, we illustrate the special case $C=0$, i.e.
+$(T(\Soca_\mod),q_\mod(v))\models\varphi_\mod$ iff
+$B\equiv 0\ \mod M$.
+The SOCA $\Soca_{\mod}$ contains four atomic propositions $\prop_0,\prop_1,y,z$
+ and is depicted below. The $\CTL$ formula $\varphi_\mod$ expresses that we
+traverse the sequence of diamonds and thereby repeatedly subtract $M$ from $B$.
+The number of diamonds equals $\beta$, the number of bits of $M$.
+One diamond corresponds to one bit of $M$. In case bit $\beta$ of $M$ is $1$,
 which we can verify by a transition to the initial control state of
-the SOCA $\Soca_{\bit,M_\Re}$,
- we subtract $2^{B_\Li+\beta-1}$ from $B$, otherwise we do not modify the counter value.
-After that, we move on to the second diamond, which represents the second
-rightmost  bit of $M$, and so on.
-Hence, traversing a cycle that starts and ends
-in our initial control state $q_\mod $ will correspond to subtracting $M$ once.
- Then we traverse
-the lower sequence of diamonds and similarly subtract $C$, but this
-this time only once.
-Finally, after having traversed the lower sequence of diamonds and seeing the atomic
-proposition $\prop_2$, we check if $B=0$ by trying to subtract $2^{B_\Li}$.
-\begin{figure}
+the SOCA $\Soca_{\bit,\mu+\beta}$ (see Fig. \ref{F Bit}),
+ we subtract $2^{\gamma+\beta}$ from $B$, otherwise we do not modify the counter value.
+This process is repeated until we reach the last diamond in which we consider
+the first bit of $M$.
+Finally,  the transition from $q_{\mod}$ to the control state satisfying $z$
+serves for checking if $B=0$ by trying to subtract $2^\gamma$.
+%\begin{figure}
 \begin{center}
-\begin{picture}(100,120)(5,-41)
+\begin{picture}(100,44)(5,33)
 %   \put(-20,-28){\framebox(85,33){}}
@@
 %\put(43,-5){$p_2$}
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,20)(28,20)(28,40)(5,40)
-\put(10,25){\Large$\Soca_{\bit,M_\Re}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,20)(58,20)(58,40)(35,40)
-\put(38,25){\Large$\Soca_{\bit,M_\Re-1}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,20)(108,20)(108,40)(85,40)
-\put(90,25){\Large$\Soca_{\bit,M_\Li}$}
+\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,33)(28,33)(28,47)(5,47)
+\put(10,35){\large$\Soca_{\bit,\mu+\beta}$}
+
+\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,33)(58,33)(58,47)(35,47)
+\put(38,35){\large$\Soca_{\bit,\mu+\beta-1}$}
+
+\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,33)(108,33)(108,47)(85,47)
+\put(90,35){\large$\Soca_{\bit,\mu}$}
 
@@
   \node(1)(0,60){$q_{\mod}$}
-  \node(2O)(15,70){$\bullet$}
-  \node(2U)(15,50){$\bullet$}
-  \put(17,72){\small$\prop_1$}
-  \put(17,48){\small$\prop_0$}
-
-  \node(1A)(15,35){$q_{\bit,M_\Re}$}
-  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}
-
-  \node(4A)(45,35){$q_{\bit,M_\Re-1}$}
-
-  \node(7A)(95,35){$q_{\bit,M_\Li}$}
+  \put(-6,62){$y$}
+  \node(z)(0,40){$\bullet$}
+  \put(-3,40){$z$}
+\drawedge[ELside=r](1,z){$-2^{\gamma}$}
+  \node(2O)(15,65){$\bullet$}
+  \node(2U)(15,55){$\bullet$}
+  \put(17,67){\small$\prop_1$}
+  \put(17,53){\small$\prop_0$}
+
+  \node(1A)(15,42){$q_{\bit,\mu+\beta}$}
+  \node(4A)(45,42){$q_{\bit,\mu+\beta-1}$}
+
+  \node(4A)(45,42){$q_{\bit,\mu+\beta-1}$}
+
+  \node(7A)(95,42){$q_{\bit,\mu}$}
 
   \node(3)(30,60){$\bullet$}
-  \node(4O)(45,70){$\bullet$}
-
-  \put(47,72){\small$\prop_1$}
-  \put(47,48){\small$\prop_0$}
-  \node(4U)(45,50){$\bullet$}
+  \node(4O)(45,65){$\bullet$}
+
+  \put(47,67){\small$\prop_1$}
+  \put(47,53){\small$\prop_0$}
+  \node(4U)(45,55){$\bullet$}
   \node(5)(60,60){$\bullet$}
@@
   \node(6)(80,60){$\bullet$}
-  \node(7O)(95,70){$\bullet$}
-  \put(97,72){\small$\prop_1$}
-  \put(97,48){\small$\prop_0$}
-
-  \node(7U)(95,50){$\bullet$}
+  \node(7O)(95,65){$\bullet$}
+  \put(97,67){\small$\prop_1$}
+  \put(97,53){\small$\prop_0$}
+
+  \node(7U)(95,55){$\bullet$}
   \node(8)(110,60){$\bullet$}
@@
 \drawedge[ELside=r](1,2U){$0$}
-\drawedge[ELside=l,ELpos=60](2O,3){$-2^{B_\Li+\beta-1}$}
+\drawedge[ELside=l,ELpos=60](2O,3){$-2^{\gamma+\beta}$}
 \drawedge(2U,3){$0$}
@@
 \drawedge[ELside=r](3,4U){$0$}
-\drawedge[ELside=l,ELpos=70](4O,5){$-2^{B_\Li+\beta-2}$}
+\drawedge[ELside=l,ELpos=70](4O,5){$-2^{\gamma+\beta-1}$}
 \drawedge(4U,5){$0$}
@@
 \drawedge[ELside=r](6,7U){$0$}
-\drawedge[ELside=l,ELpos=60](7O,8){$-2^{B_\Li}$}
+\drawedge[ELside=l,ELpos=60](7O,8){$-2^{\gamma}$}
 \drawedge(7U,8){$0$}
 
-\drawcurve(110,60)(112,70)(55,80)(0,70)(0,62)
+\drawcurve(110,60)(112,62)(55,75)(0,63)(0,62)
 
@@
 %\drawedge[ELside=r,ELpos=60](8,9B){$0$}
-  \put(60,81){$0$}
+  \put(60,76){$0$}
 %\drawedge[ELpos=50](8,9C){$0$}
@@
 
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](5,-40)(28,-40)(28,-20)(5,-20)
-\put(10,-35){\Large$\Soca_{\bit,C_\Re}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](35,-40)(58,-40)(58,-20)(35,-20)
-\put(38,-35){\Large$\Soca_{\bit,C_\Re-1}$}
-
-\drawpolygon[Nframe=n,Nfill=y,fillgray=.9](85,-40)(108,-40)(108,-20)(85,-20)
-\put(90,-35){\Large$\Soca_{\bit,C_\Li}$}
-
-\gasset{Nframe=n,loopdiam=9,ELdist=.7}
-\gasset{Nadjust=wh,Nadjustdist=1}
-\gasset{curvedepth=0}
-
-  \node(1C)(0,0){$\bullet$}
-  \drawedge[ELside=r](1,1C){$0$}
-  \node(2OC)(15,10){$\bullet$}
-  \node(2UC)(15,-10){$\bullet$}
-  \put(17,12){\small$\prop_1$}
-  \put(17,-12){\small$\prop_0$}
-
-  \node(1AC)(15,-25){$q_{\bit,C_\Re}$}
-  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}
-
-  \node(4AC)(45,-25){$q_{\bit,C_\Re-1}$}
-
-  \node(7AC)(95,-25){$q_{\bit,C_\Li}$}
-
-  \node(3C)(30,0){$\bullet$}
-  \node(4OC)(45,10){$\bullet$}
-
-  \put(47,12){\small$\prop_1$}
-  \put(47,-12){\small$\prop_0$}
-  \node(4UC)(45,-10){$\bullet$}
-  \node(5C)(60,0){$\bullet$}
-
-  \node(6C)(80,0){$\bullet$}
-  \node(7OC)(95,10){$\bullet$}
-  \put(97,12){\small$\prop_1$}
-  \put(97,-12){\small$\prop_0$}
-
-  \node(7UC)(95,-10){$\bullet$}
-  \node(8C)(110,0){$\bullet$}
-
-  \node(9C)(110,-20){$\bullet$}
-
-\drawedge(8C,9C){$-2^{B_\Li}$}
-
-  \put(112,-2){$\prop_2$}
-
-  \put(65,-2){\huge$\cdots$}
-
-%\drawedge[curvedepth=-70](8,1){$0$}
-
-\drawedge(1C,2OC){$0$}
-\drawedge[ELside=r](1C,2UC){$0$}
-\drawedge[ELside=l,ELpos=60](2OC,3C){$-2^{B_\Li+\beta-1}$}
-\drawedge(2UC,3C){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](2OC,1AC){$0$}
-\drawedge(2UC,1AC){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](4OC,4AC){$0$}
-\drawedge(4UC,4AC){$0$}
-
-\drawedge[curvedepth=-6,ELpos=70,ELside=r](7OC,7AC){$0$}
-\drawedge(7UC,7AC){$0$}
-
-\drawedge(3C,4OC){$0$}
-\drawedge[ELside=r](3C,4UC){$0$}
-\drawedge[ELside=l,ELpos=70](4OC,5C){$-2^{B_\Li+\beta-2}$}
-\drawedge(4UC,5C){$0$}
-
-\drawedge(6C,7OC){$0$}
-\drawedge[ELside=r](6C,7UC){$0$}
-\drawedge[ELside=l,ELpos=60](7OC,8C){$-2^{B_\Li}$}
-\drawedge(7UC,8C){$0$}
-
 \end{picture}
-\caption{The SOCA $\Soca_\mod$}{\label{F Smod}
-for checking if $B\equiv C\ \mod M$.}
+%\caption{The diamond SOCA $\Soca_\mod$}{\label{F Smod}}
 \end{center}
-\end{figure}
-Finally, let us give the formula $\varphi_\mod$:
-$$
-\varphi_\mod\quad=\quad\E \left(\bigwedge_{i\in\{0,1\}} \prop_i
-\rightarrow \EX \varphi_{\bit,i} \right) \U (\prop_2\wedge\neg\EX\true)
-$$
+%\end{figure}
+
+\noindent
+We put
+$
+\varphi_\mod=\E \left(\bigwedge_{b\in\{0,1\}} \prop_b
+\rightarrow \EX \varphi_{\bit,b} \right) \U (y\wedge\neg\EX z)
+$.
 \qed
@@
 
-\noindent
-We need the following
-proposition for implementing Q2. Its proof is straightforward.
-
-\begin{proposition}{\label{P prime}}
-The following problem is in $\PSPACE$:
-
-INPUT: $\bin(i)\$\bin(m)$
-
-QUESTION: $m=p_i$?
-\end{proposition}
-
-%\renewcommand{\prime}{\text{prime}}
-\newcommand{\prim}{\text{prime}}
-
-\begin{lemma}{\label{L prim}} There is some fixed $\ctl$ formula $\varphi_\prim$ such that
-from  $x_0$ we can compute in logarithmic space some SOCA
-$\Soca_\prim$ and some control state $q_\prim$ such that
+\newcommand{\BIT}{\text{BIT}}
+\renewcommand{\R}{\mathcal{R}}
+
+\noindent
+Let us give some informal ideas on how to implement the queries
+Q2 to Q5. We strongly recommend
+the reader to consult the technical report \cite{techrep}
+to understand the technical subtleties.
+
+For implementing Q2 we simulate with a SOCA $\Soca_\prim$ some
+polynomially space-bounded DTM
+that decides, on the input $\langle I,M\rangle$, whether
+$M=p_I$. We use the bit string $X$ from $(*)$ for storing the working tape
+of this DTM on the counter.
+The current input and working tape symbol and the position of the
+input and working tape in the counter can directly be hard-wired
+into the control states of $\Soca_\prim$.
+We can construct a fixed CTL formula that
+simulates the computation of this DTM.
+
+Implementing Q3, i.e. deciding the $J^{\text{th}}$ bit of
+$B$, is more involved.
+Recall that $B$ consists of $n+2^N$ bits and $J$ consists of
+$\alpha=O(N)$ bits.
+Hence checking if $\bit_J(B)=1$ cannot be done in a similar fashion as in
+Fig. \ref{F Bit}, since $J$ is too big.
+The solution is the following: By making use of $\Soca_\prim$, one can initialize
+ $M$ with $p_I$ and after that
+decide if $C\equiv B\ \mod p_I$
+by making use of $\Soca_{\mod}$ and $\varphi_\mod$
+from Lemma \ref{L mod}.
+Hence, we can access bits of the Chinese remainder representation of $B$.
+Let us assume that $\R=\CRR_{K}(B)=(b_{i,c})_{1\leq i\leq K,0\leq c<p_i}$
+is the Chinese remainder representation of $B$.
+Observe that $|\R|$ is exponential in $n$ and
+$\R$ is not stored anywhere on the counter.
+However we
+can use the bit strings $I$ and $C$ as pointers to access
+the bit $b_{I,C}$ of $\R$.
+By Theorem \ref{T CRRBIN}, given $\R$ (in our case on-the-fly by the pointers
+$I$ and $C$), the bit string $J$ and $b\in\{0,1\}$,
+we can decide if $\bit_J(B)=b$
+by simulating a logarithmically space-bounded DTM
+on the input $\langle\R,J,b\rangle$ of exponential size.
+In the block $Y$ of $(*)$ we reserve the space that this DTM requires.
+Q4 can be implemented similarly as Q3 by simulating
+a logarithmically space-bounded DTM that decides $U$ on input
+$B[1,n]\cdot B[n+1,n+2^N]$ of exponential size.
+We use $Z$ for simulating the working tape and the bit sequence $J$ as a
+pointer to access the bits of $B$.
+
+For implementing Q5 we
+simulate the program from in Fig.\ \ref{F prog}.
+Recall that our bit sequence $B$ is of length $n+2^N$.
+We initialize the first $n$ bits of $B$ with $x_0$.
+The remaining bit sequence $B'$ stores the variable $d$ of the program, initialized with $0$
+and being repeatedly incremented by adding $2^{\gamma+n}$.
+%Note that incrementing $B'$ by $1$ corresponds to adding $2^{B_\Li+n}$ to the one-counter.
+Thus, checking when $d$ becomes $2^{2^N}$ for the first time boils down to checking
+when $B'$ overflows for the first time.
+This can be checked by initializing $J$ appropriately and
+being able to access the $J^{\text{th}}$ bit via query Q3.
+The states of the automaton $A$ can directly be handled by the control states of the
+SOCA. To obtain $\chi_U(x_0,\bin_{2^N}(d))$, we invoke the query Q4 and store
+this bit in the control state of the SOCA. This concludes our $\EXPSPACE$-hardness proof.
+
+\begin{theorem}
+The data complexity and the combined complexity of CTL
+and the modal $\mu$-calculus on SOCA is $\EXPSPACE$-complete.
+\end{theorem}
+
+\iffalse
+\begin{lemma}{\label{L prim}} There is some fixed $\CTL$ formula $\varphi_\prim$ s.t.\
+from our input $x_0$ we can compute in logarithmic space some SOCA
+$\Soca_\prim$ and some control state $q_\prim$ s.t.\
 $(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$ %\quad\Leftrightarrow\quad
-if and only if $M=p_I$.
+iff $M=p_I$.
 \end{lemma}
 \begin{proof}
-By Proposition \ref{P prime} there is some $q(n)$ space-bounded DTM
+By Proposition \ref{P prime} there is some $q(n)$ space-bounded Turing machine
 $\M=(S,\Sigma,s_0,F,\mu)$
@@
 given $\bin(i)\$\bin(m)$, whether $m=p_i$, where $q$ is some polynomial.
-The idea is to simulate $\M$ on input $z=I\$M$ using $X$ from $(\bigstar)$ as a working tape.
-For this, we define our block $X$ to consist of
+The idea is to simulate $\M$ on input $z=I\$M$ and using $X$ as a working tape.
+For this, we define our block $X$ to consist of precisely
 $l=q(|I|+|M|+1)=q(\alpha+\beta+1)$ many bits.
-Moreover, we assume w.l.o.g. that $\M$'s behaviour on input $I\$M$ is
-independent of the content of its initial working tape: This can be
+Moreover, we assume w.l.o.g.\ that $\M$'s behaviour on input $I\$M$ is
+independent on the content of its initial working tape: This can be
 achieved by adding extra states to $\M$ that first write $0^l$ onto
@@
 $(s_0,\l z\r,\l w \r,0,0)$ for some $w\in\{0,1\}^l$.
+\iffalse
 Before we give our SOCA $\Soca_\prim$, we describe the computation of
-$\M$ on input $I\$M$ as a pseudo program that terminates if and only if $M=p_I$:\\
+$\M$ on input $I\$M$ as a pseudo program that terminates iff $M=p_I$:\\
 
@@
 
-\ \hspace{3cm} $\mathtt{if}\ h\in[1,l]\ \ \mathtt{then}\  \ \bit_h(X):=b';\ \ \mathtt{fi}$\\[-0.4cm]
-
 \ \hspace{3cm} $s:=s '$;\\[-0.4cm]
 
+\ \hspace{3cm} $\bit_h(X):=b'$;\\[-0.4cm]
+
 \ \hspace{3cm} $i:=i+\delta_1$;\\[-0.4cm]
@@
 
+\fi
 \newcommand{\inpu}{\text{input}}
@@
 Moreover, our SOCA will contain
-the SOCA $\Soca_{\bit,I_\Li},\Soca_{\bit,I_\Li+1}\ldots,\Soca_{\bit,I_\Re}$ and
-$\Soca_{\bit,X_\Li},\Soca_{\bit,X_\Li+1}\ldots,\Soca_{\bit,X_\Re}$ in order to test if
+the SOCA $\Soca_{\bit,I_\Li},\ldots,\Soca_{\bit,I_\Re}$ and
+$\Soca_{\bit,X_\Li},\ldots,\Soca_{\bit,X_\Re}$ in order to test if
 certain bits
@@
 Hence $Q$ contains the control states
-$q_{\bit,I_\Li},q_{\bit,I_\Li+1}\ldots,q_{\bit,I_\Re}$ and
-$q_{\bit,X_\Li},q_{\bit,X_\Li+1}\ldots,q_{\bit,X_\Re}$ as well.
+$q_{\bit,I_\Li},\ldots,q_{\bit,I_\Re}$ and
+$q_{\bit,X_\Li},\ldots,q_{\bit,X_\Re}$ as well.
 
@@
 We do not give all transitions of $\Soca_\prim$, but illustrate
-some cases by way of example.
+them at an example.
 Let us give the outgoing transitions of the control state
-$(s,i,h,0,1)$, where $i\in[I_\Li,I_\Re]$ and $h\in[X_\Li,X_\Re]$,
-i.e. we currently scan the $i^{\text{th}}$ bit
-of $I$ and do not read a marker on our working tape.
+$(s,i,h,0,1)$, where $i\in[I_\Li,I_\Re]$ and $h\in[X_\Li,X_\Re]$, i.e. we currently scan
+a bit of $I$ and do not read a marker on our working tape.
 Moreover, let us assume $\mu(s,0,1)=(s',+1,-1,0)$, i.e. we change to state $s'$,
@@
 and modify the current content on the working tape from $1$ to $0$.
-We realize the latter by subtracting $2^{X_\Li+h}$ from the counter,
- however we allow transitions
-to states $(s',i+1,h-1,a,b)$, {\em for each} $a,b\in\{0,1\}$, hence we guess
+We realize the latter by subtracting $2^{X_\Li+h}$, however we allow transitions
+to states $(s',i+1,h-1,a,b)$, for each $a,b\in\{0,1\}$, hence we guess
 the input tape symbol and the working tape symbol of the
 successor configuration.
-The $\ctl$ formula $\varphi_\prim$ will guarantee that our guessing was correct by
+Our CTL formula $\varphi_\prim$ will guarantee that our guessing was correct by
 accessing the control states $q_{\bit,k}$ where
@@
 The other cases can be dealt with analogously.
-We put $q_\prim=(s_0,0,0,\l,\l)$.
 Our final formula is
@@
 \left(\text{work}_k\rightarrow\EX(\work\wedge\varphi_{\bit,k})\right)
-\right)\U\ F.
-$$
- Then
+\right)\U F.
+$$
+We put $q_\prim=(s_0,0,0,\l,\l)$. Then
 $(T(\Soca_\prim),q_\prim(v))\models\varphi_\prim$
-if and only if
-$I\$M\in L(\M)$ if and only if
+iff
+$I\$M\in L(\M)$ iff
 $M=p_I$.
@@
 \noindent
-Let us now give an implementation of query Q3.
+Let us give an answer to question (iii).
 
 \begin{lemma}{\label{L BIT}}
-For each bit $b\in\{0,1\}$ there exists a fixed $\ctl$ formula $\varphi_{\BIT,b}$
-such that from $x_0$
-we can compute in logarithmic space some SOCA $\Soca_{\BIT}$ and some
-control state $q_\BIT$ such that
-$(T(\Soca_\BIT),q_\BIT(v))\models\varphi_{\BIT,b}$
-if and only if $\bit_J(B)=b$.
+For each bit $b\in\{0,1\}$ there exists a fixed CTL formula
+$\varphi_{\BIT,b}$ s.t.\ from our input $x_0$ we can compute in
+logarithmic space some SOCA $\Soca_{\BIT}$ and some control state
+$q_\BIT$ s.t. $(T(\Soca_\BIT),q_\BIT(v))\models\varphi_{\BIT,b}$ iff
+$\bit_J(B)=b$.
 \end{lemma}
@@
 \begin{proof}[Sketch]
-By Lemma \ref{L prim}, there is a fixed $\ctl$
-formula $\varphi_\prim$ such that we can compute from $x_0$ a
+By Lemma \ref{L prim}, there is a fixed CTL
+formula $\varphi_\prim$ s.t.\ we can compute from $x_0$ a
 SOCA $\Soca_\prim$ that allows to check if $M=p_I$.
-Moreover, by Lemma \ref{L mod} there is a fixed $\ctl$
+Analogously, by Lemma \ref{L mod} there is a fixed CTL
 formula $\varphi_\mod$ and a SOCA $\Soca_\mod$ that allows
-to check whether $B\equiv C \mod\ M$.
+to check whether $C=B\ \mod M$.
 In other words, we can deal with the Chinese remainder representation
-of $B$, but our goal is to access the the $J^{\text{th}}$ bit in the
-binary representation of $B$.
+of $B$, but we aim at accessing the $J^{\text{th}}$ bit of $B$.
 So let us assume that $\R=\CRR_{K}(B)=(b_{i,c})_{1\leq i\leq K,0\leq c<p_i}$
 is the Chinese remainder representation of $B$.
-Note that we do not have $\R$ stored anywhere on the counter.
-However, the bit strings $I$ and $C$ serve as pointers to access the
-bit $b_{I,C}$ of $\R$.
-By Theorem \ref{T CRRBIN}, given $\R$ (on-the-fly by our pointers $I$ and $C$), our
+%The idea is that the bit strings $I$ and $C$ serve as pointers for accessing
+%bits of $\R$.
+%To prove the lemma, in a nutshell, we translate our input from Chinese remainder
+%representation to binary representation.
+By Corollary \ref{C CRRBIN}, given $\R$, our
 bit string $J$, and the bit $b$,
@@
 Turing machine machine for this, where $k\geq 1$ is some constant.
-Hence, in order to decide if
+Hence, in order to decide
 $\bit_J(B)=b$ we need to simulate $\M$ on input $\langle \R, J, b\rangle$.
 For this we store the space that $\M$ requires on our reserved
-sequence of bits $Y$ from $(\bigstar)$, hence $Y$ consists of
+sequence of bits $Y$, hence $Y$ consists of
 $l=k\cdot\log(|\R Jb|)=O(\beta^2+\alpha+1)=O(N^2)$ many bits.
-The definition of SOCA $\Soca_{\BIT,b}$ works analogously to the construction
-of $\Soca_{\prim}$ in the proof of Lemma \ref{L prim}
+The crucial difficulty is that we do not have $\R$ stored anywhere on our
+counter, since $\R$ consists of has exponentially many bits in $n$.
+The solution to this problem is to access bit strings $I$ and $C$
+and to treat them as pointers to access bits of $\R$.
+The definition of SOCA $\Soca_{\BIT,b}$ works analogously as the construction
+of $\Soca_{\prim}$ as in the proof of Lemma \ref{L prim}
 i.e. we introduce in $\Soca_{\BIT,b}$ control states
@@
 the working head, and the position of the working head.
-The only difference is that a pointer to the input head cannot be
-completely stored
+The only difference is that a pointer to the input head cannot completely stored
 in the control states.
@@
 (i) via the fixed formula $\varphi_\prim$ whether $M=p_I$, and
-then (ii) via the fixed formula $\varphi_{\mod}$
-whether $B\equiv C\ \mod M$.
+then (ii) via the fixed formula $\varphi_{\mod,1}$ (resp. $\varphi_{\mod,0}$)
+ whether $B\ \mod M=C$ (resp. $B\ \mod M\not=C$).
 Pointers to the remaining parts of the input for $\M$, namely $J$ and $b$,
@@
 in analogy to the proof of Lemma \ref{L prim}.
-By way of example we explain the behavior of $\Soca_{\BIT}$ when the input head
-of $\M$ currently scans bit $b_{I,C}$ of $\R$ and when getting
+Let us exemplarily explain the behavior of $\Soca_{\BIT}$ when the input head
+of $\M$ currently scans bit $b_{I,C}$ of $\R$ and let us assume that getting
 to the successor configuration requires moving the input head to the left.
 Then we simply decrement $C$ by $1$ which corresponds to subtracting $2^{C_\Li}$
-from the counter. If, however, $C$ currently equals $0$, we need to
+from our counter. If, however, $C$ currently equals $0$, we need to
 decrement $I$ by $1$, overwrite $M$ with prime $p_{I-1}$, and finally
 overwrite $C$ with $p_{I-1}-1$. Decrementing $I$ by $1$ can simply achieved by
-subtracting $2^{I_\Li}$ from the counter. Overwriting $M$ with prime $p_{I-1}$
+subtracting $2^{I_\Li}$ from our counter. Overwriting $M$ with prime $p_{I-1}$
 can be done by repeatedly subtracting $1$ from $M$ (i.e. subtracting $2^{M_\Li}$
 from the counter) until $M$ equals $p_{I-1}$;
-checking if $M=p_{I-1}$ can be done via the fixed $\ctl$ formula $\varphi_{\prim}$
+checking if $M=p_{I-1}$ can be done via the fixed CTL formula $\varphi_{\prim}$
 and the SOCA $\Soca_{\prim}$ by Lemma \ref{L prim}.
@@
 \noindent
-The following lemma implements query Q4.
+The following lemma answers question (iv).
 
 \begin{lemma}{\label{L U}}
-For each $b\in\{0,1\}$ there is some fixed $\ctl$ formula $\varphi_{U,b}$ such that
-from $x_0$
+For each $b\in\{0,1\}$ there is some fixed CTL formula $\varphi_{U,b}$ s.t.\
+from our input $x_0$
 we can compute in logarithmic space some SOCA $\Soca_U$ and some
-control state $q_U$ such that $(T(\Soca_U),q_U(v))\models\varphi_{U,b}$
-if and only if $b=\chi_U(B[1,n],B[n+1,n+2^N])$.
+control state $q_U$ s.t.\ $(T(\Soca_U),q_U)\models\varphi_{U,b}$
+iff $b=\chi_U(B[1,n],B[n+1,n+2^N])$.
 \end{lemma}
-\begin{proof}[sketch]
-The proof is similar to the proof of Lemma \ref{L BIT}.
-Since $U$ is in $\L$, there is some $k\cdot\log(m)$-space bounded
-Turing machine that decides $U$, where $k\geq 1$ is some constant.
-In order to decide if $b=\chi_U(B[1,n],B[n+1,n+2^N])$ we
-need to simulate $\M$ on input $\langle B[1,n],B[n+1,n+2^N]\rangle$.
-We store the space that $\M$ requires on its working tape
-in the sequence of bits $Z$ from $(\bigstar)$, hence $Z$ consists of
-$k\cdot\log(n+2^N+1)=O(N)$ many bits.
-We use the bit sequence $J$ as a pointer for accessing bits of $B$.
-For reading the $J^{\text{th}}$ bit of $B$ we make use of the $\ctl$ formula
-$\varphi_{\BIT,b}$ and use the SOCA $\Soca_\BIT$ from Lemma \ref{L BIT}.
-The rest of the proof follows along the same lines as the proof of
-Lemma \ref{L prim}.
-\qed
-\end{proof}
-
-\noindent
-The following lemma implements query Q5 and concludes the
-$\EXPSPACE$-hardness of the data complexity of $\ctl$ on SOCA.
+
+\noindent
+The following Lemma answers question (v) and hence the proof of
+$\EXPSPACE$-hardness of the data complexity of CTL on SOCA.
 
 \begin{lemma}
-There is some fixed $\ctl$ formula $\varphi_L$ such that from $x_0$
+There is some fixed CTL formula $\varphi_L$ s.t.\ from our input $x_0$
 we can compute in logarithmic space some SOCA $\Soca_L$ and some
-control state $q_L$ such that $(T(\Soca_L),q_L(0))\models\varphi_L$
-if and only if $x_0\in L$.
+control state $q_L$ s.t.\ $(T(\Soca_L),q_L)\models\varphi_L$
+iff $x_0\in L$.
 \end{lemma}
 \begin{proof}
-First note that our $\ctl$ formula $\varphi_L$ will be evaluated in state $q_L(0)$.
-Recall that our bit sequence $B$ has length $n+2^N+1$ where $B$ is factorized
+First note that our CTL formula $\varphi_L$ will be evaluated in state $q_L(0)$.
+Recall that our bit sequence $B$ had length $n+2^N+1$ where $B$ is factorized
 as $B=xB'$
 for some $x\in\{0,1\}^n$ and some $B'\in\{0,1\}^{2^N+1}$.
-The SOCA $\Soca_L$ and the $\ctl$ formula $\varphi_L$ will mimic the execution
-of the program from Figure \ref{F U}.
-In the bit string $x$ we store the value $x_0$.
-The bit string $B'$ represents
+The SOCA $\Soca_L$ and the CTL formula $\varphi_L$ will mimic the execution
+of the program from Fig.\ \ref{F U}.
+In the bit string $x$ we want to store the value $x_0$.
+The bit string $B'$ is meant to represent
 the variable $d$ of the program, hence we will initialize
@@
  By Lemma
-\ref{L BIT} the $\ctl$ formula $\varphi_{\BIT,1}$ and the SOCA $\Soca_\BIT$ allow to test
+\ref{L BIT} the CTL formula $\varphi_{\BIT,1}$ and the SOCA $\Soca_\BIT$ allow to test
 if the $J^{\text{th}}$ bit of $B$ equals $1$.
@@
 The following claim tells us that we can test whether we have
-initialized the counter correctly.
+initialized our counter correctly.
 Its proof is simple and therefore omitted.
@@
 \noindent
-{\em Claim:} There is a fixed $\ctl$ formula $\varphi_\init$ such that from
+{\em Claim:} There is a fixed CTL formula $\varphi_\init$ s.t.\ from
 $x_0$ we can compute in logarithmic space some SOCA $\Soca_\init$ and
-some control state $q_\init$ such that
+some control state $q_\init$ s.t.\
 $(T(\Soca_\init),q_\init(v))\models\varphi_\init$
-if and only if $J=n+2^N+1$ and $B=x_00^{2^N+1}$.\\
+iff $J=n+2^N+1$ and $B=x_00^{2^N+1}$.\\
 
@@
 Recall that $A=(Q,\{0,1\},q_0,\delta,F)$ is the deterministic finite automaton
-of the program in Figure \ref{F U} that needs to be simulated.
-Let us, before giving $\Soca_L$, define the auxiliary SOCA $\Soca_A$
+that needs to be simulated.
+Let us, in a first step, define the auxiliary SOCA $\Soca_A$
 to be connected with the control state $q_U$ of $\Soca_U$ and with the control
@@
 Moreover, $\Soca_A$ contains the atomic propositions
-$\{\bit_0,\bit_1,F\}$, where $S_{\bit_b}=Q\times\{b\}$ for each $b\in\{0,1\}$ and
+$\{\bit_0,\bit_1,F\}$, where with $S_{\bit_b}=Q\times\{b\}$ for each $b\in\{0,1\}$ and
 $S_F=F\times\{0,1\}$.
@@
 \end{center}
-Let us define the auxiliary formula $\chi=\EX\varphi_\BIT$ that allows
-to test if $B'=2^{2^N}$.
+Let us define the auxiliary formula $\chi=\EX\varphi_\BIT$.
 Our final formula is
 $$
-\varphi_L\quad=\quad\EF\left(\rho\wedge \EX(\init\wedge\varphi_\init)\wedge
+\varphi_L=\EF\left(\rho\wedge \EX(\init\wedge\varphi_\init)\wedge
 \E\left(\neg\chi\wedge\bigwedge_{b\in\{0,1\}}\bit_b\rightarrow
 \EX\varphi_{U,b}\right)
- \U\ (\chi\wedge F)
+ \U (\chi\wedge F)
 \right)
 $$
-We have $x_0\in L$ if and only if
-the program from Figure \ref{F U} returns $\mathtt{true}$ if and only if
+We have $x_0\in L$ iff
 $(T(\Soca_L),q_L(0))\models\varphi_L$.
@@
 \begin{theorem}
-The data and combined complexity of $\ctl$ and the modal $\mu$-calculus
- on SOCA is $\EXPSPACE$-complete.
-\label{thm:ctl-poca}
+Data complexity of CTL on SOCA is $\EXPSPACE$-hard.
 \end{theorem}
@@
 We call $b\in\Z$ a {\em solution} to the above linear equation if there exists a
-mapping $\sigma:\{x_1,\ldots,x_n\}\rightarrow\N$ such that
+mapping $\sigma:\{x_1,\ldots,x_n\}\rightarrow\N$ s.t.\
 $b=\sum_{i\in[n]} a_i\cdot\sigma(x_i)$.
@@
 \end{eqnarray}
-be a linear equation such that $1<a_1<\cdots a_n\leq N$ and where
+be a linear equation s.t.\ $1<a_1<\cdots a_n\leq N$ and where
 $b=\gcd(a_1,\ldots,a_n)$.
@@
 
+\fi
 \newcommand{\Two}{\mathbb{A}}
@@
 
-\subsection{$\Pi^0_1$-Hardness of the Data Complexity of $\CTL$ on POCA}{\label{S CTL und}}
-In this section we show that there already
-exists a fixed $\CTL$ formula for which model checking of
-$\poca$ is $\Pi^0_1$-hard.
-We reduce from the complement of the emptiness problem for
-\emph{two-counter automata}, which is $\Sigma^0_1$-complete \cite{Min61}.
-Similar to a
-SOCA, a {\em two-counter automaton} is a tuple $\Two=(Q, E, \lambda)$,
-where $Q$ is a finite set of {\em control states},
-$E\subseteq Q\times Q$ is a set of {\em edges},
- however
-$\lambda: E \rightarrow \{\zero_1,\zero_2\}\cup \{\add_j(a)\mid j\in \{1,2\}, a\in
-\{-1,1\} \}$. The {\em configuration graph induced by $\Two$} is defined
-to be $G(\A)=(Q\times \N\times \N, \rightarrow)$, where
-$(p,i_1,i_2)\rightarrow (q,i_1+d_1,i_2+d_2)$ if and only if
-the following conditions are all satisfied:
-\begin{itemize}
-\item $(p,q)\in E$,
-%\item  $i_1,i_2,i_1+d_1,i_2+d_2\geq 0$,
-\item $\lambda(p,q)=\add_j(a)$ implies $d_j=a$ and $d_{3-j}=0$ for each
-$j\in\{1,2\}$, and
-\item $\lambda(p,q)=\zero_j$ implies $d_j=i_j=0$ and $d_{3-j}=0$ for each
-$j\in\{1,2\}$.
-\end{itemize}
-
-\noindent
-Let us define the following problem.
-\smallskip
-
-\problemx{Emptiness for two-counter automata}{Two-counter automaton $\Two=(Q,E,\lambda)$, control states
-$q_0,q_1\in Q$.}{Are there $m,n\in\N$ such that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ in
-$G(\Two)$?}
-
-\medskip
-
-\noindent
+\subsection{Hardness of the data complexity of $\CTL$ on POCA}{\label{S CTL und}}
+We now show that there exists a fixed $\CTL$ formula for which model
+checking of $\poca$ is $\Pi^0_1$-hard by a reduction from the emptiness
+problem for \emph{two-counter automata}, which is
+$\Pi^0_1$-complete~\cite{Min61}. Similar to a SOCA, a two-counter
+automaton $\A$ consists of a finite set of control states and transitions
+between them. However, each transition of $\Two$ acts on two counters,
+which it can in- and decrement and test for $\zero$.
+
+%% is a tuple $\Two=(Q, E, \lambda)$ with a finite set of
+%% contro states $Q$, an edge relation $E\subseteq Q\times Q$, however
+%% $\lambda: E \rightarrow \{\zero_1,\zero_2\}\cup \{\add_j(a)\mid j\in
+%% \{1,2\}, a\in \{-1,1\} \}$.
+
+%% The {\em directed graph induced by $\Two$}
+%% is defined to be $G(\A)=(Q\times \N\times \N, \rightarrow)$, where
+%% $(p,i_1,i_2)\rightarrow (q,i_1+d_1,i_2+d_2)$ iff $(p,q)\in E$,
+%% $\lambda(p,q)=\add_j(a)$ implies $d_j=a$ and $d_{3-j}=0$, and
+%% $\lambda(p,q)=\zero_j$ implies $d_j=i_j=0$ and $d_{3-j}=0$.  The {\em
+%%   reachability problem on two-counter automata} asks, given a
+%% two-counter automaton $\Two=(Q,E,\lambda)$, and control states
+%% $q_0,q_1\in Q$, whether there are $m,n\in\N$ such that
+%% $(q_0,0,0)\rightarrow^*(q_1,m,n)$ in $G(\Two)$?
+
+%\noindent
+
 The idea of our reduction is as follows: Given a two-counter automaton
-$\Two$, we construct a $\poca$ $\Poca(x)$ with one parameter in such a way
-that the two counters from $\A$ are encoded into the single counter
-from $\Poca(x)$ as follows: for the counter value $n$ of $\Poca(x)$,
-we have that $n\ \mod x$ encodes the value of the first and $n\ \div\ x$ encodes the
-value of the second counter of $\Two$.
-Hence, testing the first counter for zero corresponds to checking
-whether $n\equiv 0\ \mod x$ while testing the second counter for
-zero corresponds to checking  whether $n\geq x$.
-Incrementation (resp. decrementation) on the first counter can be
-mimicked by adding (resp. subtracting) $1$, whereas on counter two this
-corresponds to adding (resp. subtracting) $x$.
-Of course, we need to ensure that we do not overflow.
-For example, if $n\equiv -1\ \mod x$ and we would simulate to increment the first counter
-in the above manner, this
-would correspond to setting the first counter to zero and simultaneously
- incrementing the second counter. However, if the emptiness problem is
-solvable, then $x$ can be instantiated with a large enough
-value such that such an overflow does not occur.
-Our $\ctl$ formula will ensure that we do not overflow.
+$\Two$, we construct a $\poca$ $\Poca(x)$ with one parameter in such a
+way that the two counters from $\A$ are encoded into the single
+counter from $\Poca(x)$. Given a counter value $n$ of $\Poca(x)$,
+$n\ \mod x$ encodes the value of the first, and $n\ \div\ x$ encodes
+the value of the second counter of $\Two$.  Hence, testing whether the
+first equals $0$ corresponds to checking whether $n\equiv 0\ \mod
+x$, while testing whether the second counter equals $0$ corresponds to
+checking whether $n\geq x$.  Incrementing (resp.\ decrementing) the
+first counter of $\A$ can be mimicked by adding (resp.\ subtracting) $1$,
+whereas on counter two this corresponds to adding (resp. subtracting)
+$x$. Of course, we need $\ctl$ formulas to ensure that we do not
+overflow when simulating an increment of the first counter of $\A$.
+For instance, if $n\equiv -1\ \mod x$ and we want to
+simulate an increment of the first counter of $\Two$ in that way, we
+would actually set the first counter to $0$ and simultaneously
+increment the second counter. However, if $\Two$ is not empty, then
+$x$ can be instantiated with a large enough value such that such an
+overflow does not occur. Conversely, if $\Two$ is empty then there is
+no such instantiation.  \iffalse
+
 \newcommand{\test}{\mathsf{test}}
-Before we give the reduction from the emptiness problem for two-counter
+Before we give the reduction from the reachability problem in two-counter
 automata, let us introduce some gadgets that we need in our reduction.
 \begin{lemma}{\label{L POCAs}}
-  There exists some fixed $\ctl$ formula $\varphi_\test$
+  There exists some fixed CTL formula $\varphi_\test$
 for which the following holds:
@@
 There exists some fixed $\poca$ $\Poca_{\not\equiv 0}(x)$ with some control state
-$q_{\not\equiv 0}$ such that for each $\sigma:\{x\}\rightarrow\N$
+$q_{\not\equiv 0}$ such that
  $$(\Poca_{\not\equiv 0}^\sigma,q_{\not\equiv0}(n)) \models
- \varphi_{\test} \quad\Longleftrightarrow\quad n\not
+ \varphi_{\test} \quad\Leftrightarrow\quad n\not
   \equiv 0 \textsl{ mod } \sigma(x).$$
 \item There exists some fixed $\poca$ $\Poca_{\not\equiv-1}(x)$ with some control
-state $q_{\not\equiv -1}$ such that for each $\sigma:\{x\}\rightarrow\N$
+state $q_{\not\equiv -1}$ such that
  $$(\Poca_{\not\equiv -1}^\sigma,q_{\not\equiv-1}(n)) \models
- \varphi_\test \quad\Longleftrightarrow\quad n\not
+ \varphi_\test \quad\Leftrightarrow\quad n\not
   \equiv -1 \textsl{ mod } \sigma(x).$$
-\item There exists some fixed $\poca$ $\Poca_{<x}(x)$ with some control
-state $q_{<x}$ such that for each $\sigma:\{x\}\rightarrow\N$
- $$(\Poca_{< x}^\sigma,q_{<x}(n)) \models
- \varphi_\test \quad\Longleftrightarrow\quad n< \sigma(x).$$
+\item There exists some fixed $\poca$ $\Poca_{\geq x}(x)$ with some control
+state $q_{\geq x}$ such that
+ $$(\Poca_{\geq x}^\sigma,q_{\geq x}(n)) \models
+ \varphi_\test \quad\Leftrightarrow\quad n\geq \sigma(x).$$
 
@@
 proposition.
-Below we depict $\Poca_{\not\equiv0}(x)$,
-$\Poca_{\not\equiv-1}(x)$, and $\Poca_{<x}(x)$, respectively.
+The following figure depicts $\Poca_{\not\equiv0}(x)$,
+$\Poca_{\not\equiv-1}(x)$, and $\Poca_{\geq x}(x)$, resp.
 \begin{center}
-\begin{picture}(85,20)(-20,-3.5)
+\begin{picture}(85,30)(-20,-12)
 %   \put(-20,-28){\framebox(85,33){}}
@@
 
-\put(-65,0){
+\put(-60,0){
 \put(41,-2){\small $\prop$}
@@
 
-\put(-27,0){
+\put(-15,0){
 \put(61,-2){\small $\prop$}
@@
 
-\put(32,0){
-\put(61,-2){\small $\prop$}
-  \node(A)(20,0){$q_{< x}$}
-  \node(C)(40,0){$\bullet$}
-  \node(D)(60,0){$\bullet$}
-\drawloop(A){$-1$}
-\drawedge(A,C){$-x$}
-\drawedge(C,D){$\zero$}
-%\drawedge[ELside=r](C,D){$\zero$}
+\put(50,7){
+\put(41,-2){\small $\prop$}
+  \node(A)(20,0){$q_{\geq x}$}
+  \node(B)(20,-20){$\bullet$}
+  \node(C)(40,-20){$\bullet$}
+  \node(D)(40,0){$\bullet$}
+\drawloop(A){$+1$}
+\drawedge(A,B){$+1$}
+\drawedge(B,C){$-x$}
+\drawedge[ELside=r](C,D){$\zero$}
 }
@@
 \noindent
-Let us fix some instance of the emptiness problem for
+Let us fix some instance of the reachability problem for
  two-counter automata:
-Let $\langle\Two,q_0,q_1\rangle$ be an instance to the
-emptiness problem, where
+Let $\langle\Two,q_0,q_1\rangle$ be an input to the
+reachability problem, where
 $\Two=(Q,E,\lambda)$ is some two-counter automaton.
-We will give a fixed $\ctl$ formula $\varphi$ such that from
+We will give a fixed CTL formula $\varphi$ such that from
 $\langle\Two,q_0,q_1\rangle$ we can construct some $\poca$ $\Poca(x)$ such that
-$(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ if and only if
+$(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ iff
 $(\Poca^\sigma,q_0(0))\models\varphi$ for some $\sigma:\{x\}\rightarrow\N$.
@@
 (\Poca_{\not\equiv -1}(x),q_{\not\equiv -1})& \text{ if }\quad l=\add_1(+1)\\
-(\Poca_{<x}(x),q_{<x})& \text{ if }\quad l=\zero_2,
+(\Poca_{\geq x}(x),q_{\geq x})& \text{ if }\quad l=\zero_2,
 \end{cases}
@@
 \text{ for some }\circ\in\{+,-\}\\
-0& \text{ otherwise.}
+0& \text{ otherwise}
 \end{cases}
@@
 Let us moreover ensure that every control state of the kind
-$q_{\tau}$ with $\tau\in\{\not\equiv 0,\not\equiv\! -1,\geq x\}$ has
+$q_{\tau}$ with $\tau\in\{\not\equiv 0,\not\equiv -1,\geq x\}$ has
 the label $\rho_\tau$.
@@
 formulas that guarantee that an overflow never occurs:
+
 \begin{eqnarray*}
@@
 \psi_3&\quad=\quad&
-\zero_2\longrightarrow\EX(\rho_{<x}\wedge\varphi_{\test})\\
+\zero_1\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\neg\varphi_\test)\\
 \psi_4&\quad=\quad&
-\zero_1\longrightarrow\EX(\rho_{\not\equiv 0}\wedge\neg\varphi_\test)
+\zero_2\longrightarrow\EX(\rho_{\geq x}\wedge\neg\varphi_{\test})\\
 \end{eqnarray*}
@@
 \noindent
-Note that we do not need any constraint when changing the second counter.
 Let us introduce a label $f$ at control state $q_1$.
-Our final $\ctl$ formula $\varphi$ is
+Our final CTL formula $\varphi$ is
 $$
@@
 \fi
-We have that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ in $G(\A)$  if and only
-if
+We have that $(q_0,0,0)\rightarrow^*(q_1,m,n)$ for some $m,n\in\N$ in $G(\A)$
+iff
 $(T(\Poca^\sigma(x)),q_0(0))\models\varphi$ for some
@@
 We obtain the following theorem.
-
+\fi
 \begin{theorem}
-  The data and combined complexity of $\CTL$ and the modal $\mu$-calculus
- on POCA is $\Pi^0_1$-complete.
+  The data and combined complexity of $\CTL$ on POCA is
+  $\Pi^0_1$-complete.
+  \label{thm:ctl-poca}
 \end{theorem}
 
-\section{$\ltl$ Model Checking}{\label{S LTL}}
-
-%\subsection{Linear-time logic ($\ltl$)}
-Formulas of $\ltl$ are inductively defined according to the
-following grammar, where $\prop$ ranges over $\Prop$:
+\section{$\ltl$ Model Checking}
+Formulas of $\ltl$ are given by the following grammar, where $\prop$
+ranges over $\Prop$:
 \begin{equation*}
-  \varphi\quad ::=\quad \prop \mid \neg \varphi \mid \varphi \wedge \varphi \mid
+  \varphi \quad ::= \quad \prop \mid \neg \varphi \mid \varphi
+  \wedge \varphi \mid
   \X \varphi \mid \varphi \until \varphi
 \end{equation*}
-The Boolean abbreviations and the formula $\true$ are defined in the same way as
-in CTL. The \emph{finally modality} $\eventually
-\varphi$ is an abbreviation for $\true \until \varphi$ and the
-\emph{globally modality} $\globally \varphi$ abbreviates $\neg
-\eventually \neg \varphi$.
-
-The semantics of LTL is given in terms of infinite paths in a
+%% The Boolean abbreviations and $\true$ and $\false$ are defined in the
+%% same way as in $\CTL$. The \emph{eventually modality} $\eventually
+%% \varphi$ is an abbreviation for $\true \until \varphi$ and the
+%% \emph{globally modality} $\globally \varphi$ abbreviates $\neg
+%% \eventually \neg \varphi$.
+
+The semantics of $\ltl$ is given in terms of infinite paths in a
 transition system. Let $T=(S,\{S_\prop \mid \prop \in
-\Prop\},\rightarrow)$ be a transition system, let $\pi: s_0
-\rightarrow s_1 \rightarrow \cdots$ be an infinite path in $T$, and
-let $\varphi$ be an $\ltl$ formula, we define $(T,\pi)\models\varphi$
-by induction of the structure of $\varphi$, where $\prop$ ranges over
-$\Prop$:
+\Prop\},\rightarrow)$ be a transition system, $\pi= s_0 \rightarrow
+s_1 \rightarrow \cdots$ an infinite path in $T$ and $\varphi$ an
+$\ltl$ formula, we define $(T,\pi)\models \varphi$ by induction on the
+structure $\varphi$.
 \begin{align*}
-  (T,\pi) \models  \prop &  \quad\iff\quad  \pi(0) \in S_\prop\\
-  (T,\pi) \models  \neg \varphi &\quad\iff\quad (T,\pi) \not \models \varphi\\
-  (T,\pi) \models \varphi_1 \wedge \varphi_2 &\quad\iff\quad (T,\pi)
-  \models \varphi_1 \text{ and } (T,\pi) \models \varphi_2\\
-  (T,\pi) \models \X \varphi & \quad\iff\quad (T,\pi^1) \models \varphi\\
-  (T,\pi) \models  \varphi_1 \until \varphi_2 & \quad\iff\quad
-  \exists j\ge 0: (T,\pi^j)\models \varphi_2
-  \text{ and } \forall 0\le i<j: (T,\pi^i)\models \varphi_1
+  (T,\pi) \models  \prop &  \iff  \pi(0) \in S_\prop, \prop \in \Prop &
+  (T,\pi) \models  \neg \varphi &\iff \pi \not \models \varphi\\
+  (T,\pi) \models \varphi_1 \wedge \varphi_2 &\iff
+  \forall i\in \{1,2\}:(T,\pi) \models \varphi_i &
+  (T,\pi) \models \X \varphi & \iff (T,\pi^1) \models \varphi
 \end{align*}
-Let us now define the $\ltl$ model checking problem on SOCA
-and POCA.\\
+\vspace*{-1.0cm}
+\begin{align*}
+  (T,\pi) \models  \varphi_1 \until \varphi_2  \iff
+  \exists j\ge 0: (T,\pi^j) \models \varphi_2
+  \text{ and } \forall 0\le i<j: (T,\pi^i) \models \varphi_1
+\end{align*}
+%% Let us now define the $\ltl$ model checking problem on SOCA
+%% respectively POCA.\\
 
 \problemx{$\ltl$ model checking on SOCA}
-{SOCA $\Soca=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary,
-  and an $\ltl$ formula $\varphi$.}
-{$(T(\Soca),\pi)\models\varphi$ for every infinite
-path $\pi$ with $\pi(0)=q(n)$?}
+{SOCA $\Soca=(Q,E,\lambda)$, $q\in Q$, $n\in\N$ in binary,
+  an $\ltl$ formula $\varphi$.}
+{Does $(T(\Soca),\pi)\models\varphi$ for all infinite
+paths $\pi$ with $\pi(0)=q(n)$?}
 \
@@
 \problemx{$\ltl$ model checking on POCA}
-{POCA $\Poca(X)=(Q,\{Q_\prop\mid\prop\in\Prop\},E,\lambda)$, $q\in Q$, $n\in\N$ in binary, and an
+{POCA $\Poca(X)=(Q,E,\lambda)$, $q\in Q$, $n\in\N$ in binary, an
 $\ltl$ formula $\varphi$.}
-{$(T(\Poca^\sigma),\pi)\models\varphi$ for every $\sigma:X\rightarrow\N$ and
-every infinite path $\pi$ with $\pi(0)=q(n)$?}
-
-\subsection{Upper bounds}
+{Does $(T(\Poca^\sigma),\pi)\models\varphi$ for all $\sigma:X\rightarrow\N$
+and for all infinite paths $\pi$ with $\pi(0)=q(n)$?}
+
+\subsection{Upper Bounds}
 
 A standard approach to $\ltl$ model checking is the automata-based
-approach, in which systems are modeled as non-deterministic B\"uchi
-automata (NBA). Given an NBA $\automaton$ and an $\ltl$ formula
-$\varphi$, the idea is to translate $\varphi$ into an NBA
-$\automaton_{\neg \varphi}$ such that $\automaton_{\neg \varphi}$
-accepts all words that violate $\varphi$.  Then, by checking for
-emptiness of the product automaton
-$\automaton'=\automaton\times\automaton_{\neg \varphi}$, it can be
-decided whether or not $\automaton$ generates traces that violate
-$\varphi$.  Emptiness can be decided by checking for recurrent
-reachability of a control state in the transition system induced by
-$\automaton'$, which can be performed non-deterministically in space logarithmically in the size of
-$\automaton'$. Vardi and Wolper showed in \cite{VW86} that the size of
-$\automaton_{\neg \varphi}$ is $2^{O(|\varphi|)}$, which yields that
-checking for emptiness of $\automaton'$ is in $\PSPACE$. The $\PSPACE$
-lower bound for finite state model checking was originally shown
-in \cite{LP85-popl} and hence carries over to SOCA.
-
-Constructing the product automaton of two NBA is a standard technique
-in model checking, see e.g.\ \cite{BK08} for a detailed treatment of
-this. Moreover, this construction can be adapted in a
-straight-forward way into the setting of POCA by equipping a POCA with
-a set of final states and defining emptiness with respect to B\"uchi
-acceptance condition in the standard way. For brevity, we do not give
-any further details. We then have that given a POCA $\Poca$, an $\ltl$
-formula $\varphi$, a state $q$ and $n\in\N$ we can construct a
-POCA $\Poca'$ as the product of $\Poca$ and $\automaton_{\neg\varphi}$
-such that for all assignments $\sigma$ we have that $\Poca'$ is empty
-with respect to B\"uchi acceptence condition if and only if
-$(T(\Poca^\sigma),q(n))\models \varphi$. Moreover, the size of
-$\Poca'$ is $2^{O(|\varphi|)}|\Poca|$.
-
-It has been shown in \cite{HKOW10} that checking emptiness with
-respect to B\"uchi acceptance condition is $\coNP$-complete for both
-SOCA and POCA, and in \cite{DG-jlc09} that it is $\NL$-complete for
-OCA. We use these results in order to establish upper bounds for our
-$\ltl$ model checking problems. Let us first consider the case of a
-fixed formula $\varphi$. In the terminology of the previous paragraph,
-$|\Poca'|=O(|\Poca|)$ and hence the data complexity of $\ltl$ on SOCA
-and POCA is in $\coNP$.  The matching $\coNP$ lower bound of the data
-complexity of LTL on SOCA follows from the $\NP$-hardness of the
-reachability problem for SOCA.
-%We obtain $\coNP$-hardness from the fact that
-%we can reduce deciding emptiness of some SOCA $\Soca$ with respect to
-%B\"uchi acceptance condition to $\ltl$ model checking of a slightly
-%altered SOCA $\Soca'$ against an easily constructed fixed $\ltl$
-%formula.
+approach, in which behaviours of \emph{systems} are modelled as
+non-deterministic B\"uchi automata (NBA). Given an NBA $\automaton$
+modelling a system and an $\ltl$ formula $\varphi$, the idea is to
+translate $\varphi$ into an NBA $\automaton_{\neg \varphi}$ of size
+$2^{O(|\varphi|)}$ such that the language of $\automaton\times
+\automaton_{\neg \varphi}$ is empty iff $\varphi$ holds on all
+infinite traces of $\automaton$. The concept of B\"uchi automata can
+easily be adopted to the setting of counter automata. Then a
+\emph{B\"uchi-SOCA} is not empty if there is an infinite path on which
+some designated control states occurs infinitely often. The latter
+boils down to just checking for recurrent reachability. Moreover, the
+B\"uchi-SOCA obtained from the product of a SOCA and an NBA can be
+defined and constructed in a straightforward way. We omit details for brevity.
+
+It was shown in \cite{HKOW10} that checking emptiness is
+$\coNP$-complete for both B\"uchi-SOCA and B\"uchi-POCA, and in
+\cite{DG-jlc09} that it is $\NL$-complete for B\"uchi-OCA. We use
+these results for establishing upper bounds for the $\ltl$ model
+checking problems.
+
+For every fixed $\ltl$ formula $\varphi$, and every POCA $\Poca$, the
+size of $\Poca\times \automaton_{\neg \varphi}$ is $O(|\Poca|)$, hence
+the data complexity of $\ltl$ on SOCA and POCA is in $\coNP$. Hardness
+for $\coNP$ follows from $\NP$-hardness of reachability using a fixed
+formula $\true \until \rho$ for some $\rho\in \Prop$.
+
+If both $\Poca$ and $\varphi$ are part of the input then $|\Poca\times
+\automaton_{\neg \varphi}|=|\Poca|\cdot 2^{O(|\varphi|)}$, and hence
+\cite{HKOW10} gives a $\coNEXP$ upper bound for the combined
+complexity of $\ltl$ model checking on both SOCA and POCA. This upper
+bound can however be improved for SOCA. Given a SOCA $\Soca$, let $m$
+be the absolute value of the maximum increment or decrement on the
+transitions in $\Soca$.  Let $\Soca'$ be the B\"uchi-SOCA obtained
+from the product $\Soca\times\automaton_{\neg \varphi}$ by replacing
+every transition labeled with $z$ with a sequence of fresh transitions
+and control states of length $z$, where, depending on the sign of
+$z$, each transition is labeled with $+1$ resp.\ $-1$. We have
+$|\Soca'|= m\cdot|\Soca|\cdot 2^{O(|\varphi|)}$, and hence the $\NL$
+upper bound for emptiness of B\"uchi-OCA from \cite{DG-jlc09} yields a
+$\PSPACE$ upper bound.
 
@@
   The data complexity of $\ltl$ model checking on SOCA and POCA is
-  $\coNP$-complete.
+  $\coNP$-complete, the combined complexity of $\ltl$ model
+  checking on SOCA is $\PSPACE$-complete, and the combined complexity
+  of $\ltl$ model checking on POCA is in $\coNEXP$.
 \end{proposition}
-As seen above, if both $\Poca$ and the formula $\varphi$ are part of
-the input then $|\Poca'|=2^{O(|\varphi)}\cdot|\Poca|$, and hence $\coNEXP$
-is an upper bound for the combined complexity of $\ltl$ model checking
-on both SOCA and POCA. However, we can improve this upper bound in the
-case of SOCA. Given a SOCA $\Soca$, let $m$ be the absolute value of
-the maximal increment or decrement performed on the transitions in
-$\Soca$ and let $\Soca'$ be the product automaton of $\Soca$ and
-$\automaton_{\neg \varphi}$. Let $\Soca''$ be obtained from $\Soca'$
-by replacing every transition labeled with $z$ with a sequence of
-fresh transitions and control locations of length $z$, where each
-transition is labeled with $+1$ respectively $-1$, depending on the
-sign of $z$. We conclude that
-$|\Soca''|=m\cdot|\Soca'|=m\cdot2^{O(|\varphi|)}\cdot|\Soca|$. Combining the result
-from \cite{DG-jlc09} together with the fact that $\ltl$ model checking
-of non-deterministic finite state automata is $\PSPACE$-hard
-\cite{LP85-popl}, we obtain the following proposition.
-
-\begin{proposition}
-The combined complexity of $\ltl$ on SOCA is $\PSPACE$-complete.
-\end{proposition}
-It remains to prove that the combined complexity of $\ltl$ model
-checking on POCA is $\coNEXP$-hard, which we will show in the next
-section.
-
-\subsection{Combined Complexity of $\ltl$ on POCA}
-
-We are now going to show $\coNEXP$-hardness of $\ltl$ model checking
-on POCA via a reduction from (the complement of) \textsf{Succinct
-  3-SAT}, which is a $\NEXP$-complete problem \cite{Papa94}. An input
-of \textsf{Succinct 3-SAT} is given by a Boolean circuit $\mathbb{C}$
-that encodes a Boolean formula $\psi_\circuit$ in 3-CNF,
-i.e.\ $\psi_\circuit=\bigwedge_{j\in [0,M]}(\ell^j_1\vee \ell^j_2\vee
-\ell^j_3)$ for some $M=O(2^{|\circuit|})$, and $\psi_\circuit$ is free
-over Boolean variables $y_1,\ldots,y_{N}$ for some
-$N=O(2^{|\circuit|})$. Let $j\in [0,M]$ be the index of a clause
-encoded in \emph{binary} and $k\in\{1,2,3\}$. On input $(j\cdot k)$,
-the output of $\circuit$ is $(i\cdot b)$, where $i\in [N]$ is the
-index of the Boolean variable that appears in literal $\ell_k^j$, and
-where $b=0$ when $\ell_k^j$ is negative and $b=1$ when $\ell_k^j$ is
-positive.
-\vspace*{0.25cm}
-\problemx{\textsf{Succinct 3-SAT}}
-{Boolean circuit $\circuit$}
-{Is $\psi_\circuit$ satisfiable?}
-\vspace*{0.25cm}
-
-%\medskip
-
-\noindent
-In order to establish $\coNEXP$-hardness for the combined complexity
-of $\ltl$ model checking, given an input $\circuit$ of
-$\textsf{Succinct 3-SAT}$, we construct a POCA $\Poca(x)$ and an
-$\ltl$ formula $\varphi$ such that $\psi_\circuit$ is satisfiable if
-and only if there is an assignment $\sigma$ such that
-$(T(\Poca^\sigma),q_{\mathsf{start}}(0))\models\varphi$ for some
-distinguished state $q_{\mathsf{start}}$ of $\Poca(x)$.
-%There are two key
-%ideas for the reduction.
-
-First, let us provide a suitable encoding of truth assignments by
-natural numbers. The encoding we use has also been employed for
-establishing lower bounds for model checking OCA~\cite{JKMZ04-ic}.
-Recall that $p_i$ denotes the $i^{\text{th}}$ prime number. Every
-natural number $x$ defines a truth assignment
-$\nu_x:\{y_1,\ldots,y_N\} \to \{ 0,1\}$ such that $\nu_x(y_i)=1$ iff
-$p_i$ divides $x$.  By the Prime Number Theorem, $p_N=O(N \log N)$ and
-hence $O(|\circuit|)$ bits are sufficient to represent $p_N$. Of
-course, since we need exponentially many prime numbers they cannot be
-hard-wired into $\Poca(x)$.
-
-%% Second, we employ $\ltl$ in order to perform arbitrary computations of
-%% space-bounded Turing machines on traces of runs of SOCA. However, we
-%% cannot make use of the approach we employed in $\ctl$. Intuitively
-%% speaking, $\ctl$ allows for performing tests on the counter value
-%% without destroying its value and this property has no correspondant in
-%% the $\ltl$ setting. We can however use atomic propositions to model
-%% computations along traces.
+
+\subsection{Hardness of the Combined Complexity of $\ltl$ on POCA}
+
+We are now going to sketch a proof of $\coNEXP$-hardness of $\ltl$
+model checking on POCA via a reduction from the complement of
+the $\NEXP$-complete \textsf{Succinct 3-SAT} problem \cite{Papa94}.
+An input of \textsf{Succinct 3-SAT} is given by a
+Boolean circuit $\mathbb{C}$ that encodes a Boolean formula $\psi$ in
+3-CNF, i.e.\ $\psi=\bigwedge_{0\le j< M}(\ell^j_1\vee \ell^j_2\vee
+\ell^j_3)$. Let $j\in [M]$ be the index of a clause encoded in
+\emph{binary} and $k\in\{1,2,3\}$. Assume that $\psi$ uses
+$N$ different variables $y_1,\ldots,y_N$.
+ On input $(j\cdot k)$, the output
+of $\circuit$ is $(i\cdot b)$, where $i\in [N]$
+ is the index of the Boolean variable that appears in
+literal $\ell_k^j$, and where $b=0$ when $\ell_k^j$ is negative
+and $b=1$ when $\ell_k^j$ is positive.
+ \textsf{Succinct 3-SAT} is to decide whether $\psi$ is satisfiable. Fig.\
+\ref{fig:high-level} depicts on a high-level the POCA $\Poca(x)$ derived from
+$\circuit$ that we are using in our reduction.
+
+As a first step, let us provide a suitable encoding of truth assignments
+by natural numbers. The encoding we use has also been employed for establishing
+lower bounds for model checking OCA~\cite{JKMZ04-ic}.
+%%%%%% HIER KOENNTE MAN ZITIEREN
+Recall that $p_i$ denotes the $i^{\text{th}}$ prime
+number.
+ Every natural number $n$ defines a truth assignment
+ $\nu_n:\{y_1,\ldots,y_N\} \to \{ 0,1\}$ such that $\nu_n(y_i)=1$
+iff $p_i$ divides $n$.
+By the Prime
+Number Theorem, $p_N=O(N \log N)$ and hence $O(|\circuit|)$ bits are
+sufficient to represent $p_N$. Of course, since we need exponentially
+many prime numbers they cannot be hard-wired into $\Poca(x)$.
 
@@
   \begin{center}
-    \begin{picture}(142,60)(0,0)
+    \begin{picture}(120,52)(0,0)
       %   \put(-20,-28){\framebox(85,33){}}
@@
       \put(0,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(6,7){$\Soca_{\mathsf{inc}}(j)$}
+        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(18,0)(18,15)(0,15)
+        \put(3,7){$\Soca_{\mathsf{inc}}(j)$}
       }
 
-      \put(35,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
+      \put(30,25){
+        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(18,0)(18,15)(0,15)
         \put(3,8){$\Soca_{\circuit}(j\cdot 2)$}
-        \put(10,4){$=(i\cdot b)$ }
+        \put(4,4){$=(i\cdot b)$ }
       }
-      \put(35,45){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
+      \put(30,45){
+        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(18,0)(18,12)(0,12)
         \put(3,8){$\Soca_{\circuit}(j\cdot 1)$}
-        \put(10,4){$=(i\cdot b)$ }
+        \put(4,4){$=(i\cdot b)$ }
       }
-      \put(35,5){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
+      \put(30,6){
+        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(18,0)(18,12)(0,12)
         \put(3,8){$\Soca_{\circuit}(j\cdot 3)$}
-        \put(10,4){$=(i\cdot b)$ }
+        \put(4,4){$=(i\cdot b)$ }
       }
 
-      \put(70,25){
-        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
-        \put(3,6){$\Soca_{\mathsf{prime}}(i)=p_i$ }
+      \put(62,28){
+        \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(18,0)(18,12)(0,12)
+        \put(3,8){$\Soca_{\mathsf{prime}}(i)$}
+        \put(5,4){$=p_i$}
       }
 
-      \put(110,35){
+      \put(87,35){
         \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
@@
 
-      \put(110,10){
+      \put(87,10){
         \drawpolygon[Nframe=n,Nfill=y,fillgray=.9](0,0)(25,0)(25,15)(0,15)
@@
 
-      \node(A1)(12,29){}
-      \node(A2)(20,32){}
-
-      \node(B11)(40,8){}
-      \node(B12)(55,15){}
-
-      \node(B21)(40,28){}
-      \node(B22)(55,35){}
-
-      \node(B31)(40,48){}
-      \node(B32)(55,55){}
-
-      \node(C1)(73,36){}
-      \node(C2)(90,28){}
-
-      \node(D11)(115,39){}
-      \node(D12)(132,39){}
-
-      \node(D21)(113,13){}
-      \node(D22)(132,13){}
-
-      \node(E)(140,0){$\bullet$}
-      \node(F)(12,0){$q_{\mathsf{start}}$}
+      \node(A1)(5,29){}
+      \node(A2)(13,32){}
+
+      \node(B11)(33,16){}
+      \node(B12)(44,16){}
+
+      \node(B21)(33,27){}
+      \node(B22)(44,27){}
+
+      \node(B31)(33,48){}
+      \node(B32)(44,48){}
+
+      \node(C1)(65,36){}
+      \node(C2)(77,36){}
+
+      \node(D11)(90,39){}
+      \node(D12)(109,39){}
+
+      \node(D21)(90,13){}
+      \node(D22)(109,13){}
+
+      \node(E)(115,0){$\bullet$}
+      \node(F)(5,0){$q_{\mathsf{start}}$}
 
@@
 
-      \drawqbedge(D12,142,25,E){$0$}
+      \drawqbedge(D12,115,25,E){$0$}
       \drawedge(D22,E){$0$}
@@
   \end{center}
-  \caption{High-level description of the automaton $\Poca(x)$ used for
-    the reduction.}
+  \caption{High-level description of the POCA $\Poca(x)$ used for the
+    reduction from $\textsf{Succinct 3-SAT}$.}
   \label{fig:high-level}
@@
 
-Let us now take a look at $\Poca(x)$, which is sketched in Figure
-\ref{fig:high-level}. It uses one parameter $x$ and employs several
-gadgets. The only gadgets manipulating the counter are
-$\Soca_{\mathsf{divides}}$ and $\Soca_{\mathsf{not\_divides}}$. The
-remaining gadgets are designed such that they communicate via
+Let us now take a look at $\Poca(x)$. It uses one parameter $x$ and
+employs several gadgets. Only the gadgets $\Soca_{\mathsf{divides}}$
+and $\Soca_{\mathsf{not\_divides}}$ manipulate the counter.
+All gadgets are designed so that they communicate via
 designated propositional variables, and not as in Section \ref{S CTL
-  hardness} with the help of the counter. First, $\Poca(x)$ loads the
+  hardness} with the help of the counter. Starting
+in $q_{\mathsf{start}}$, $\Poca(x)$ first loads the
 value of the parameter $x$ on the counter.  Think of $x$ encoding a
-truth assignment of $\psi_\circuit$.  Next, $\Poca(x)$ traverses
-through $\Soca_{\mathsf{inc}}$, which initially chooses an arbitrary
-index $j$ identifying a clause of $\psi_\circuit$. Every time
-$\Soca_{\mathsf{inc}}$ is traversed afterwards, it increments $j$
-modulo $M+1$ and hereby moves on to the next clause.  Now $\Poca(x)$
-branches non-deterministically into a gadget $\Soca_\circuit$ in order
-to compute $(i\cdot b)$ from $\circuit$ on input $(j\cdot 1), (j\cdot
-2)$, resp.\ $(j\cdot 3)$. The index $i$ is then used as input to a
-gadget $\Soca_{\mathsf{prime}}$, which computes $p_i$. Then if $b=0$,
-it is checked in $\Soca_{\mathsf{not\_divides}}$ that $p_i$ does not
-divide the value of $x$, and likewise in $\Soca_{\mathsf{divides}}$
-that $p_i$ divides the value of $x$ if $b=1$. For checking the latter,
-the counter needs to be modified. After the checks have been finished,
-we restore the value $x$ on the counter and the process continues with
-clause $j+1~ \mod M+1$.
-
-It remains to show how the gadgets and the communication between them
-can be realized. The first observation is that the computations of
-$\Soca_{\mathsf{inc}}$, $\Soca_{\circuit}$ and
-$\Soca_{\mathsf{prime}}$ can be realized by space bounded DTM using no
-more than $p(|\circuit|)$ tape cells for their in- and output tape for
-some polynomial $p$ that is fixed once $\circuit$ is provided. Indeed,
-it is easily seen that incrementing modulo $M+1$, evaluating
-$\circuit$ and computing the $i^{\text{th}}$ prime number can be done
-by such a DTM. Thus, we now show how from a generic DTM $\M$, we can
-construct in logarithmic space a SOCA $\Soca_\M$ and some LTL formulas
-that mimic computations of $\M$ on traces of $\Soca_\M$.
-
-\begin{figure}
-  \begin{center}
-    \begin{picture}(140,100)(0,-52)
-      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
-      \gasset{Nadjust=wh,Nadjustdist=1}
-      \gasset{curvedepth=0}
-
-      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-      \node(START)(0,25){}
-      \node(A)(10,25){$\bullet$}
-      \node(B)(20,25){$q$}
-      \node(C1)(30,45){$\bullet$}
-      \node(C2)(30,35){$\bullet$}
-      \node(C3)(30,25){$\vdots$}
-      \node(C4)(30,15){$\bullet$}
-      \node(C5)(30,5){$\bullet$}
-      \node(D)(40,25){$\bullet$}
-%      \node(STATEOUT)(70,25){}
-
-%      \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-      \drawedge(START,A){}
-      \drawedge(A,B){}
-      \drawedge(B,C1){}
-      \drawedge[ELside=r](B,C2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(B,C4){}
-      \drawedge[ELside=r](B,C5){}
-      \drawedge(C1,D){}
-      \drawedge[ELside=r](C2,D){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(C4,D){}
-      \drawedge[ELside=r](C5,D){}
-%      \drawedge(D,STATEOUT){}
-
-      \put(9,21){$\zeta$}
-      \put(18,21){$\alpha$}
-      \put(29,47){$\mathsf{s}_0$}
-      \put(29,37){$\mathsf{s}_1$}
-      \put(27,12){$\mathsf{s}_{k-1}$}
-      \put(29,2){$\mathsf{s}_k$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-%%%%%%%%%%%%%%%%%% ihead
-
-%      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-%      \node(A)(5,25){$\bullet$}
-      \node(E)(50,25){$\bullet$}
-      \node(F1)(60,45){$\bullet$}
-      \node(F2)(60,35){$\bullet$}
-      \node(F3)(60,25){$\vdots$}
-      \node(F4)(60,15){$\bullet$}
-      \node(F5)(60,5){$\bullet$}
-      \node(G)(70,25){$\bullet$}
-%      \node(STATEOUT)(70,25){}
-
-      \drawedge(D,E){}
-
-%     \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-%      \drawedge(A,B){$0$}
-      \drawedge(E,F1){}
-      \drawedge[ELside=r](E,F2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(E,F4){}
-      \drawedge[ELside=r](E,F5){}
-      \drawedge(F1,G){}
-      \drawedge[ELside=r](F2,G){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(F4,G){}
-      \drawedge[ELside=r](F5,G){}
-%      \drawedge(D,STATEOUT){}
-
-%      \put(18,21){$\alpha$}
-      \put(58,47){$\mathsf{ih}_0$}
-      \put(58,37){$\mathsf{ih}_1$}
-      \put(57,12){$\mathsf{ih}_{m}$}
-      \put(56,2){$\mathsf{ih}_{m+1}$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-%%%%%%%%%%%%%%%%%% whead
-
-%      \node(DASH)(20,40){}
-%      \node(STATEIN)(-5,25){}
-%      \node(A)(5,25){$\bullet$}
-      \node(E)(80,25){$\bullet$}
-      \drawedge(G,E){}
-      \node(F1)(90,45){$\bullet$}
-      \node(F2)(90,35){$\bullet$}
-      \node(F3)(90,25){$\vdots$}
-      \node(F4)(90,15){$\bullet$}
-      \node(F5)(90,5){$\bullet$}
-      \node(G)(100,25){$\bullet$}
-      \node(H)(100,0){$\bullet$}
-
-%     \drawedge[dash={1.5}0](DASH,B){}
-%      \drawedge(STATEIN,A){}
-%      \drawedge(A,B){$0$}
-      \drawedge(E,F1){}
-      \drawedge[ELside=r](E,F2){}
-%     \drawedge(B,C3){$0$}
-      \drawedge(E,F4){}
-      \drawedge[ELside=r](E,F5){}
-      \drawedge(F1,G){}
-      \drawedge[ELside=r](F2,G){}
-%     \drawedge(C3,D){$0$}
-      \drawedge(F4,G){}
-      \drawedge[ELside=r](F5,G){}
-      \drawedge(G,H){}
-
-%      \put(18,21){$\alpha$}
-      \put(88,47){$\mathsf{wh}_0$}
-      \put(88,37){$\mathsf{wh}_1$}
-      \put(87,12){$\mathsf{wh}_{n}$}
-      \put(86,2){$\mathsf{wh}_{n+1}$}
-%      \put(60,21){$\mathsf{s}_\r$}
-
-      \node(I)(0,0){$\bullet$}
-      \drawedge(H,I){}
-
-%%%%%%%%%%%%%%%% wtape
-
-%      \node(WORKIN)(0,45){}
-      \node(A)(0,-15){$\bullet$}
-      \drawedge(I,A){}
-      \node(B1)(15,-25){$\bullet$}
-      \node(B2)(15,-5){$\bullet$}
-      \node(C)(30,-15){$\bullet$}
-      \node(D1)(45,-25){$\bullet$}
-      \node(D2)(45,-5){$\bullet$}
-      \node(E)(60,-15){$\bullet$}
-      \node(F)(70,-15){$\ldots$}
-      \node(G)(80,-15){$\bullet$}
-      \node(H1)( 95,-25){$\bullet$}
-      \node(H2)(95,-5){$\bullet$}
-      \node(I)(110,-15){$\bullet$}
-
-      \node(J)(110,-30){$\bullet$}
-      \node(K)(0,-30){$\bullet$}
-      \drawedge(I,J){}
-      \drawedge(J,K){}
-%      \node(J)(135,-15){$\bullet$}
-%      \node(DASH)(120,-25){}
-%      \node(WORKOUT)(145,-15){}
-
-%      \drawedge(WORKIN,A){}
-      \drawedge(A,B1){}
-      \drawedge(A,B2){}
-      \drawedge(B1,C){}
-      \drawedge(B2,C){}
-      \drawedge(C,D1){}
-      \drawedge(C,D2){}
-      \drawedge(D1,E){}
-      \drawedge(D2,E){}
-      \drawedge(G,H1){}
-      \drawedge(G,H2){}
-      \drawedge(H1,I){}
-      \drawedge(H2,I){}
-%      \drawedge(I,J){}
-%      \drawedge[dash={1.5}0](I,DASH){}
-
-%      \put(4,42){$\mathsf{w}_\l$}
-      \put(13,-9){$\mathsf{i}_{1,1}$}
-      \put(13,-29){$\mathsf{i}_{1,0}$}
-      \put(43,-9){$\mathsf{i}_{2,1}$}
-      \put(43,-29){$\mathsf{i}_{2,0}$}
-      \put(93,-9){$\mathsf{i}_{m,1}$}
-      \put(93,-29){$\mathsf{i}_{m,0}$}
-%      \put(120,42){$\mathsf{w}_\r$}
-%      \put(133,42){$\mathsf{yield}$}
-
-%%%%%%%%%%%%%%% wtape
-      \node(A)(0,-45){$\bullet$}
-
-      \drawedge(K,A){}
-
-      \node(B1)(15,-55){$\bullet$}
-      \node(B2)(15,-35){$\bullet$}
-      \node(C)(30,-45){$\bullet$}
-      \node(D1)(45,-55){$\bullet$}
-      \node(D2)(45,-35){$\bullet$}
-      \node(E)(60,-45){$\bullet$}
-      \node(F)(70,-45){$\ldots$}
-      \node(G)(80,-45){$\bullet$}
-      \node(H1)( 95,-55){$\bullet$}
-      \node(H2)(95,-35){$\bullet$}
-      \node(I)(110,-45){$\bullet$}
-      \node(J)(120,-45){$\bullet$}
-      \node(K)(120,50){$\bullet$}
-      \node(L)(20,50){$\bullet$}
-      \node(M)(20,25){}
-      \node(N)(130,-45){$\bullet$}
-      \node(P)(140,-45){}
-
-      \drawedge(I,J){}
-      \drawedge(J,K){}
-      \drawedge(K,L){}
-      \drawedge(L,M){}
-
-      \drawedge(J,N){}
-      \drawedge(N,P){}
-
-%      \node(DASH)(120,-25){}
-%      \node(WORKOUT)(145,-15){}
-
-%      \drawedge(WORKIN,A){}
-      \drawedge(A,B1){}
-      \drawedge(A,B2){}
-      \drawedge(B1,C){}
-      \drawedge(B2,C){}
-      \drawedge(C,D1){}
-      \drawedge(C,D2){}
-      \drawedge(D1,E){}
-      \drawedge(D2,E){}
-      \drawedge(G,H1){}
-      \drawedge(G,H2){}
-      \drawedge(H1,I){}
-      \drawedge(H2,I){}
-%      \drawedge(I,J){}
-%      \drawedge[dash={1.5}0](I,DASH){}
-
-%      \put(4,42){$\mathsf{w}_\l$}
-      \put(12,-39){$\mathsf{w}_{1,1}$}
-      \put(12,-58){$\mathsf{w}_{1,0}$}
-      \put(42,-39){$\mathsf{w}_{2,1}$}
-      \put(42,-58){$\mathsf{w}_{2,0}$}
-      \put(92,-39){$\mathsf{w}_{n,1}$}
-      \put(93,-58){$\mathsf{w}_{n,0}$}
-
-      \put(119,-49){$\gamma$}
-      \put(129,-49){$\beta$}
-
-%      \put(120,42){$\mathsf{w}_\r$}
-%      \put(133,42){$\mathsf{yield}$}
-
-    \end{picture}
-  \end{center}
-  \caption{SOCA $\Soca_\M$ for the simulation of space bounded Turing
-    machines.}
-  \label{fig:tm}
-\end{figure}
-
-Let $\M=(S, \Sigma, s_0, F, \mu)$ be a DTM with a fixed input tape
-with $m$ tape cells, and $n$ working tape cells, and let
-$S=\{s_0,\ldots,s_k\}$. Figure \ref{fig:tm} shows the SOCA $\Soca_\M$
-that we use for the simulation of $\M$. All transitions of $\Soca_\M$
-are implicitly labelled with $0$.  A simulation starts when $\Soca_\M$
-is entered at the location labeled with $\zeta$ and is finished when
-the location labeled with $\beta$ is reached.
-
-The sequence of propositions occurring on a trace starting from and
-ending in $q$ encodes a configuration of $\M$.  In detail,
-$\mathsf{s}_i$ indicates that $\M$ is in state $s_i$; $\mathsf{ih}_i$
-that the input tape head scans cell $i$; $\mathsf{wh}_i$ that the
-working tape head scans cell $i$; $\mathsf{i}_{i,b}$ that the
-$i^{\text{th}}$ bit of the input tape is set to $b$; and
-$\mathsf{w}_{i,b}$ that the $i^{\text{th}}$ bit of the working tape is
-set to $b$, where $i$ is in the respective range and $b\in \{0,1\}$.
-
-\newcommand{\state}{\mathsf{state}}
-\newcommand{\work}{\mathsf{work}}
-\newcommand{\iNp}{\mathsf{input}}
-\newcommand{\inh}{\mathsf{inhead}}
-\newcommand{\woh}{\mathsf{wohead}}
-\renewcommand{\succ}{\mathsf{succ}}
-
-Let us introduce some LTL formulas that allow for testing properties
-of the current configuration. Think of all of them as being evaluated
-in $q$. The formula $\state_i=\X \mathsf{s}_i$ for each $i\in[0,k]$
-expresses that the current state is $s_i$.  By the formula
-$\inh_{i}=\X\X\X\X \mathsf{ih}_i$ we express that the input head is at
-position $i$, where $i\in[m]$. Similarly, define the formulas
-$\woh_i$, $\work_{j,b}$, and $\iNp_{i,b}$ for expressing that the
-working head is at position $i$, that the $i^{\text{th}}$ bit of the
-input tape is $b$, and that the $j^{\text{th}}$ bit of the working
-tape is $b$, respectively, where $i\in[m]$,$j\in[n]$, and
-$b\in\{0,1\}$.
-
-The LTL formula below, assumed to be evaluated in $\alpha$, ensures
-that the transition function is correctly encoded into traces of
-$\Soca_\M$ for states $s\in S\setminus F$ whenever the input
-respectively working tape head does not scan a start ($\l$)
-respectively end marker ($\r$):
-
-\begin{multline*}
-\bigwedge_{s_l\in S\setminus F}
-\bigwedge_{i\in[m]}
-\bigwedge_{j\in[n]}
-\bigwedge_{b_1,b_2\in\{0,1\}}
-\state_l\wedge\inh_i\wedge\iNp_{i,b_1}\wedge\woh_j\wedge
-\work_{j,b_2} \longrightarrow\\
-\longrightarrow
-\left(
-\left(\X(\neg\alpha \wedge \neg \beta)
-\U(\alpha\wedge \succ(s_l,i,j,b_1,b_2)\right)
-\wedge
-\bigwedge_{j'\not= j}\bigwedge_{b\in\{0,1\}}
-\left(\work_{j',b}\leftrightarrow
-(\X\neg\alpha\U(\alpha\wedge\work_{j',b}))\right)
-\right).
-\end{multline*}
-Here, whenever $\mu(s_l,b_1,b_2)=(s_h,d_1,d_2,b)$, the formula
-\begin{align*}
-\succ(s_l,i,j,b_1,b_2)\quad =\quad\state_h\wedge\inh_{i+d_1}\wedge\woh_{j+d_2}\wedge
-\work_{j,b}
-\end{align*}
-guarantees that the correct bit is ``written'' to the working tape and
-that the state, the input head position, and the working tape position
-of the next configuration seen indeed match the successor
-configuration. A similar formula can be constructed for the case when
-one or both of the input or working heads point to a start
-respectively end marker. Once we have reached a final state $s_i\in
-F$, we require that $\Soca_\M$ is left which is expressed by the
-following formula when evaluated in $q$:
-\begin{align*}
-   \bigwedge_{s_i\in F}\state_i \longrightarrow (\neg \alpha \until
-   \beta).
-\end{align*}
-It is now easily seen that we can construct a formula $\Phi$ that is
-derived from a conjunction of the formulas from above such that the
-formula $\globally (\alpha \rightarrow \Phi)$ constraints paths
-through $\Soca_\M$ in such a way that their traces yield the encoding
-of a valid computation of $\M$.
-
-Let us now address towards ensuring that once we enter $\Soca_\M$ we
-initially traverse it in such a way that the trace corresponds to an
-initial configuration of $\M$. The formula
-\begin{align*}
-  \globally\left(\zeta \longrightarrow \X(\state_0 \wedge \inh_1 \wedge
-  \woh_1\wedge \bigwedge_{1\le j\le n}\mathsf{work}_{j,0} ) \right)
-\end{align*}
-makes sure that the heads of the input and working tape point to the
-first tape cell, that the working tape is filled with $0$s and that we
-are in the initial state. In case the input tape can be initialized
-with an arbitrary content, we are done. Otherwise, suppose that we
-want to transfer the first $j$ bits of the output of a TM $\M'$ from
-its corresponding SOCA $\Soca_{\M'}$ to the input of $\Soca_\M$.  For
-$b\in \{0,1\}$, let $\bar{b}=0$ if $b=1$ and $\bar{b}=1$ otherwise,
-and suppose that all atomic propositions are primed in $\Soca_{\M'}$.
-The formula
-\begin{align*}
-  \bigwedge_{1\le i\le j\atop{b\in \{0,1\}}}
-  \globally
-   ((\mathsf{w}'_{i,b} \wedge (\neg \alpha' \until \beta'))
-  \longrightarrow (\neg \mathsf{i}_{i,\bar{b}}\until \gamma)
-  %\right)
-\end{align*}
-guarantees that we traverse through the first $j$ bits of the
-component of $\Soca_\M$ representing the input tape of $\M$ in the
-same way as we traverse the first $j$ bits of the working tape
-component of $\M'$ in $\Soca_{\M'}$ when a computation has finished.
-Coming back to Figure \ref{fig:high-level}, we have thus seen how the
-SOCA $\Soca_{\mathsf{inc}}, \Soca_{\circuit}$ and
-$\Soca_{\mathsf{prime}}$ and the communication between them can be
-realized. The only major question left open is how we can perform a
-divisibility respectively non-divisibility test of the counter value
-with a prime number computed in $\Soca_{\mathsf{prime}}$. For this,
-let us consider the SOCA $\Soca_{\mathsf{divides}}$ from Figure
-\ref{fig:soca-divisibility}.
-
-\begin{figure}[t]
-  \begin{center}
-    \begin{picture}(140,30)(0,0)
-      \gasset{Nframe=n,loopdiam=9,ELdist=.7}
-      \gasset{Nadjust=wh,Nadjustdist=1}
-      \gasset{curvedepth=0}
-
-      %% \node(WORKIN)(0,45){}
-      %% \node(A)(10,45){$\bullet$}
-      %% \node(B1)(25,55){$\bullet$}
-      %% \node(B2)(25,35){$\bullet$}
-      %% \node(C)(40,45){$\bullet$}
-      %% \node(D1)(55,55){$\bullet$}
-      %% \node(D2)(55,35){$\bullet$}
-      %% \node(E)(70,45){$\bullet$}
-      %% \node(F)(80,45){$\ldots$}
-      %% \node(G)(90,45){$\bullet$}
-      %% \node(H1)(105,55){$\bullet$}
-      %% \node(H2)(105,35){$\bullet$}
-      %% \node(I)(120,45){$\bullet$}
-      %% \node(J)(135,45){$\bullet$}
-      %% \node(DASH)(120,55){}
-      %% \node(WORKOUT)(145,45){}
-
-      %% \drawedge(WORKIN,A){}
-      %% \drawedge(A,B1){$0$}
-      %% \drawedge(A,B2){$0$}
-      %% \drawedge(B1,C){$0$}
-      %% \drawedge(B2,C){$0$}
-      %% \drawedge(C,D1){$0$}
-      %% \drawedge(C,D2){$0$}
-      %% \drawedge(D1,E){$0$}
-      %% \drawedge(D2,E){$0$}
-      %% \drawedge(G,H1){$0$}
-      %% \drawedge(G,H2){$0$}
-      %% \drawedge(H1,I){$0$}
-      %% \drawedge(H2,I){$0$}
-      %% \drawedge(I,J){$0$}
-      %% \drawedge[dash={1.5}0](I,DASH){}
-
-      %% \put(4,42){$\mathsf{work}_\l$}
-      %% \put(21,57){$\mathsf{work_{1,1}}$}
-      %% \put(21,32){$\mathsf{work_{1,0}}$}
-      %% \put(51,57){$\mathsf{work_{2,1}}$}
-      %% \put(51,32){$\mathsf{work_{2,0}}$}
-      %% \put(101,57){$\mathsf{work}_{N,1}$}
-      %% \put(101,32){$\mathsf{work}_{N,0}$}
-      %% \put(120,42){$\mathsf{work}_\r$}
-      %% \put(133,42){$\mathsf{yield}$}
-
-      \node(DIVIDESIN)(0,15){}
-      \node(A)(10,15){$\bullet$}
-      \node(B1)(25,25){$\bullet$}
-      \node(B2)(25,5){$\bullet$}
-      \node(C)(40,15){$\bullet$}
-      \node(D1)(55,25){$\bullet$}
-      \node(D2)(55,5){$\bullet$}
-      \node(E)(70,15){$\bullet$}
-      \node(F)(80,15){$\ldots$}
-      \node(G)(90,15){$\bullet$}
-      \node(H1)(105,25){$\bullet$}
-      \node(H2)(105,5){$\bullet$}
-      \node(I)(120,15){$\bullet$}
-      \node(J)(135,15){$\bullet$}
-      \node(DIVIDESOUT)(145,15){}
-
-      \drawedge(DIVIDESIN,A){}
-      \drawedge(A,B1){$0$}
-      \drawedge(A,B2){$0$}
-      \drawedge(B1,C){$-2^0$}
-      \drawedge(B2,C){$0$}
-      \drawedge(C,D1){$0$}
-      \drawedge(C,D2){$0$}
-      \drawedge(D1,E){$-2^1$}
-      \drawedge(D2,E){$0$}
-      \drawedge(G,H1){$0$}
-      \drawedge(G,H2){$0$}
-      \drawedge(H1,I){$-2^l$}
-      \drawedge(H2,I){$0$}
-      \drawedge(I,J){$\mathsf{\zero}$}
-      \drawbpedge[ELside=r](I,290,28,A,250,28){$0$}
-
-%      \put(4,12){$\mathsf{work}_\l$}
-      \put(22,27){$\prop_{0,1}$}
-      \put(22,2){$\prop_{0,0}$}
-      \put(52,27){$\prop_{1,1}$}
-      \put(52,2){$\prop_{1,0}$}
-      \put(102,27){$\prop_{l,1}$}
-      \put(102,2){$\prop_{l,0}$}
-%      \put(120,12){$\mathsf{work}_\r$}
-      \put(134,12){$\xi$}
-
-    \end{picture}
-  \end{center}
-  \caption{The SOCA $\Soca_{\mathsf{divides}}$  for testing the
-    counter for divisibility with some natural number of
-    bit length $l+1$.}
-  \label{fig:soca-divisibility}
-\end{figure}
-
-One cycle through $\Soca_{\mathsf{divides}}$ subtracts some natural
-number of bit length $l+1$ from the counter. However, in order to test
-for divisibility we need to make sure that we remain on the same path
-in every cycle. In the $\CTL$ setting, this problem was resolved by
-branching into the additional SOCA $\Soca_{\mathsf{bit}}$. In
-contrast, in $\ltl$ we cannot branch, but use the propositions
-$\prop_{j,b_j}, j\in [0,l], b\in \{0,1\}$ in order to stay on
-precisely one path in every cycle. Assuming that the number $p$ for
-which we want to test for divisibility with the current counter value
-is encoded as a sequence of propositions $\mathsf{w}_{j,b_j}$ of some
-SOCA $\Soca_\M$, the subsequent formula enforces that we always
-subtract $p$ in cycles of $\Soca_{\mathsf{divides}}$:
-\begin{align*}
-  \globally \left( \bigwedge_{0\le j\le l \atop{b_{j}\in \{0,1\}}}
-  ((\mathsf{w}_{j,b_j} \wedge (\neg \alpha \until \beta) ) \longrightarrow
-  ( \neg \prop_{j,\bar{b_j}} \until \xi)) \right).
-\end{align*}
-It is straight forward to derive a similar SOCA
-$\Soca_{\mathsf{not\_divides}}$ and an appropriate LTL formula for
-testing non-divisibility of the counter value with a previously
-computed prime number. Finally, we can also adopt these techniques in
-order to correctly handle the branching on $b$ performed in Figure
-\ref{fig:high-level}.
-
-In summary, by taking the disjoint union of all the gadgets from
-Figure \ref{fig:high-level}, their appendent $\ltl$ formula that we
-described in this section, connecting the gadgets correctly and taking
-the conjunction of the relevant $\ltl$ formulas, we can construct
-$\Poca(x)$ and an $\ltl$ formula $\varphi$ in logarithmic space such
-that there is an assignment $\sigma$ assigning a natural number to $x$
-such that $(T(\Poca^\sigma), q_{\mathsf{start}}(0))\not\models
-\varphi$ if and only if $\psi_\circuit$ given by an input $\circuit$
-of $\textsf{Succinct 3-SAT}$ is satisfiable.
+truth assignment of $\psi$.  Next, $\Poca(x)$ traverses through
+$\Soca_{\mathsf{inc}}$, which initially chooses an arbitrary index $j$
+identifying a clause of $\psi$. Every time $\Soca_{\mathsf{inc}}$ is
+traversed afterwards, it increments $j$ modulo $N$ and hereby moves on
+to the next clause.  Now $\Poca(x)$ branches non-deterministically
+into a gadget $\Soca_\circuit$ in order to compute $(i\cdot b)$ from
+$\circuit$ on input $(j\cdot 1), (j\cdot 2)$, resp.\ $(j\cdot 3)$. The
+index $i$ is then used as input to a gadget $\Soca_{\mathsf{prime}}$,
+which computes $p_i$. Then if $b=0$, it is checked that $p_i$ does not
+divide $x$, and likewise that $p_i$ divides $x$ if $b=1$. Those
+checks need to modify the counter. After they
+have finished, we restore the value $x$ on the counter and the
+process continues with clause $j+1~ \mod N$. We can construct an
+$\ltl$ formula $\varphi$ that ensures that all gadgets work and
+communicate correctly, and prove that $\psi$ is satisfiable iff there
+is an assignment $\sigma$ and an infinite path $\pi=
+q_{\mathsf{start}}(0)\rightarrow \cdots$ such that $(T(\Poca^\sigma),
+\pi) \not \models \varphi$. The gadgets $\Soca_{\mathsf{inc}}$,
+$\Soca_{\mathsf{circuit}}$ and $\Soca_{\mathsf{inc}}$ can be realized
+by simulating space-bounded Turing machines with SOCA and some
+appropriate $\ltl$ formulas. Here it is important that our LTL formula
+$\varphi$ is not fixed.  Divisibility resp. non-divisibility is
+checked similar as in the $\CTL$ case, cf.\ Lemma \ref{L mod}. We
+refer the reader to the technical report for further details
+\cite{techrep}.
 
@@
 
-\section{Conclusion}{\label{S Concl}}
-
+\section{Conclusion}
 In this paper, we have settled the computational complexity of model
-checking CTL and LTL on SOCA and POCA with respect to data and
-combined complexity. Our proofs for providing lower bounds have
-introduced some involved concepts and techniques, which we believe may
-be of independent interest for providing lower bounds for decision
-problems in the verification of infinite state systems.
-
-An interesting aspect of future work could be the consideration of
-\emph{synthesis problems} for POCA. Given a POCA $\Poca(X)$ and an LTL
-formula $\varphi$, a natural question to pose is whether there
-\emph{exists an assignment} $\sigma$ such that $(T(\Poca^\sigma),
-\pi)\models \varphi$ for all infinite paths $\pi$ starting in some
-state of $T(\Poca^\sigma)$.  For CTL, such a problem is undecidable by
-Theorem \ref{thm:ctl-poca}, but we claim that it might be decidable
-for LTL. Moreover, the CTL fragment EF seems to be a good candidate of
-a branching-time logic for which model checking on POCA could be
-decidable, but this remains subject to further investigations.
-
-\bibliography{bib}
-\bibliographystyle{plain}
+checking CTL, the modal $\mu$-calculus and LTL on SOCA and POCA with
+respect to data and combined complexity. Our proofs for providing
+lower bounds have introduced some non-trivial concepts and techniques,
+which we believe may be of independent interest for providing lower
+bounds for decision problems in the verification of infinite state
+systems.
+
+An interesting aspect of future work could be to consider
+\emph{synthesis problems} for POCA. Given a POCA $\Poca(X)$ and a
+formula $\varphi$, a natural question to ask is whether there
+\emph{exists an assignment} $\sigma$ such that $(T(\Poca^\sigma), \pi)
+\models \varphi$ on all infinite paths $\pi$ starting in some state of
+$T(\Poca^\sigma)$. For CTL resp. the modal $\mu$-calculus, such
+a problem is undecidable by Theorem \ref{thm:ctl-poca}. However for
+LTL it seems conceivable that this problem can be translated into a
+sentence of a decidable fragment of Presburger arithmetic with
+divisibility, similar to those studied in \cite{BozgaI05}.
+
+\footnotesize
+
+%\bibliography{bib}
+%\bibliographystyle{plain}
+
+\def\cprime{$'$}
+\begin{thebibliography}{10}
+
+\bibitem{ArnNiw01}
+A.~Arnold and D.~Niwi{\'n}ski.
+\newblock {\em Rudiments of $\mu$-calculus}, volume 146 of {\em Studies in
+  Logic and the Foundations of Mathematics}.
+\newblock {N}orth-{H}olland, 2001.
+
+\bibitem{BHIMV06}
+A.~Bouajjani, M.~Bozga, P.~Habermehl, R.~Iosif, P.~Moro, and T.~Vojnar.
+\newblock Programs with lists are counter automata.
+\newblock In {\em Proc. of CAV}, volume 4144 of {\em LNCS}, pages 517--531. Springer, 2006.
+
+\bibitem{BIL06-icalp}
+M.~Bozga, R.~Iosif, and Y.~Lakhnech.
+\newblock Flat parametric counter automata.
+\newblock In {\em Proc. of ICALP}, volume 4052 of {\em LNCS}, pages 577--588. Springer, 2006.
+
+\bibitem{BozgaI05}
+Marius Bozga and Radu Iosif.
+\newblock On decidability within the arithmetic of addition and divisibility.
+\newblock In {\em Proc. of FOSSACS}, volume 3441 of {\em LNCS}, pages 425--439. Springer,
+  2005.
+
+\bibitem{CaFu91}
+Jin-Yi Cai and Merrick Furst.
+\newblock P{SPACE} survives constant-width bottlenecks.
+\newblock {\em International Journal of Foundations of Computer Science}, 2(1):67--76, 1991.
+
+\bibitem{CR04}
+Cristiana Chitic and Daniela Rosu.
+\newblock On validation of xml streams using finite state machines.
+\newblock In {\em Proc. of WebDB}, pages 85--90. ACM, 2004.
+
+\bibitem{ChDaLi01}
+Andrew Chiu, George Davida, and Bruce Litow.
+\newblock Division in logspace-uniform {$\rm NC\sp 1$}.
+\newblock {\em Theoretical Informatics and Applications. Informatique
+  Th\'eorique et Applications}, 35(3):259--275, 2001.
+
+\bibitem{CJ98-cav}
+H.~Comon and Y.~Jurski.
+\newblock Multiple counters automata, safety analysis and presburger
+  arithmetic.
+\newblock In {\em Proc. of {CAV}}, volume 1427 of {\em LNCS}, pages 268--279. Springer, 1998.
+
+\bibitem{DG-jlc09}
+St{\'e}phane Demri and R{\'e}gis Gascon.
+\newblock The effects of bounding syntactic resources on {P}resburger {LTL}.
+\newblock {\em Journal of Logic and Computation}, 19(6):1541--1575, December 2009.
+
+\bibitem{techrep}
+Stefan G{\"o}ller, Christoph Haase, Jo{\"e}l Ouaknine, and James Worrell.
+\newblock Model checking succinct and parametric one-counter automata.
+\newblock Technical report, University of Bremen, 2010.
+\newblock Available via\\
+  \url{http://www.informatik.uni-bremen.de/tdki/research/papers/succ.pdf}.
+
+\bibitem{GoLo09}
+Stefan G{\"o}ller and Markus Lohrey.
+\newblock Branching-time model checking of one-counter processes.
+\newblock Technical report, arXiv.org, 2009.
+\newblock {\tt http://arxiv.org/abs/0909.1102}.
+
+\bibitem{GoLo10}
+Stefan G{\"o}ller and Markus Lohrey.
+\newblock Branching-time model checking of one-counter processes.
+\newblock In {\em Proc. of STACS}, pages 405--416. IFIB Schloss Dagstuhl, 2010.
+
+\bibitem{HKOW10}
+Christoph Haase, Stephan Kreutzer, Joel Ouaknine, and James Worrell.
+\newblock Reachability in parametric one-counter automata.
+\newblock 2010.
+\newblock Submitted. Available via:\\
+  \url{http://www.comlab.ox.ac.uk/files/2833/iandc.pdf}.
+
+\bibitem{HLSVW93}
+Ulrich Hertrampf, Clemens Lautemann, Thomas Schwentick, Heribert Vollmer, and
+  Klaus~W. Wagner.
+\newblock On the power of polynomial time bit-reductions.
+\newblock In {\em Proc. of CoCo}, pages 200--207. IEEE Computer Society Press, 1993.
+
+\bibitem{IJTW93-icalp}
+O.~H. Ibarra, T.~Jiang, N.~Tr{\^a}n, and H.~Wang.
+\newblock New decidability results concerning two-way counter machines and
+  applications.
+\newblock In {\em Proc. of ICALP}, volume 700 of {\em LNCS}, pages 313--324. Springer, 1993.
+
+\bibitem{ID06-tcs}
+Oscar~H. Ibarra and Zhe Dang.
+\newblock On the solvability of a class of diophantine equations and
+  applications.
+\newblock {\em Theor. Comput. Sci.}, 352(1):342--346, 2006.
+
+\bibitem{JKMZ04-ic}
+P.~Jan\v{c}ar, A.~Ku\v{c}era, F.~Moller, and Z.~Sawa.
+\newblock {DP} lower bounds for equivalence-checking and model-checking of
+  one-counter automata.
+\newblock {\em Information Computation}, 188(1):1--19, 2004.
+
+\bibitem{LS08-atva}
+J.~Leroux and G.~Sutre.
+\newblock Flat counter automata almost everywhere!
+\newblock In {\em Proc. of {ATVA}'05}, volume 3707 of {\em LNCS}, pages 489--503. Springer, 2005.
+
+\bibitem{Min61}
+Marvin~L. Minsky.
+\newblock Recursive unsolvability of {P}ost's problem of ``tag'' and other
+  topics in theory of {T}uring machines.
+\newblock {\em Annals of Mathematics. Second Series}, 74:437--455, 1961.
+
+\bibitem{Papa94}
+C.~H. Papadimitriou.
+\newblock {\em Computational Complexity}.
+\newblock Addison Wesley, 1994.
+
+\bibitem{Serr06}
+Olivier Serre.
+\newblock Parity games played on transition graphs of one-counter processes.
+\newblock In  {\em Proc. of FOSSACS}, volume 3921 of {\em LNCS}, pages 337--351. Springer, 2006.
+
+\bibitem{Vol98}
+Heribert Vollmer.
+\newblock A generalized quantifier concept in computational complexity theory.
+\newblock Technical report, arXiv.org, 1998.
+\newblock \url{http://arxiv.org/abs/cs.CC/9809115}.
+
+\end{thebibliography}
 

### 2023-10-13 version-0-VMCAI/bibliography/ltl-hardness.tex
section

\documentclass{llncs}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{ifthen}

%% \usepackage[thmmarks]{ntheorem}
%% %%% Theorem environments
%% \newtheorem{lemma}{Lemma}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{proposition}{Proposition}
%% \newtheorem{corollary}{Corollary}

%% \theoremstyle{definition}
%% \theoremsymbol{\ensuremath{\lozenge}}
%% \theorembodyfont{\normalfont}
%% \newtheorem{definition}{Definition}

%% \theoremstyle{nonumverplain}
%% \theoremheaderfont{\normalfont\bf}
%% \theorembodyfont{\normalfont}
%% \theoremseparator{.}
%% \theoremsymbol{\ensuremath{_\blacksquare}}
%% \newtheorem{example}{Example}

%% \theoremstyle{nonumberplain}
%% \theoremheaderfont{\normalfont\itshape}
%% \theorembodyfont{\normalfont}
%% \theoremseparator{.}
%% \theoremsymbol{\ensuremath{_\Box}}
%% \newtheorem{proof}{Proof}

% Symbols
\newcommand{\naturals}{\ensuremath{\mathbf{N}}}
\newcommand{\integers}{\ensuremath{\mathbf{Z}}}

% Machines and their ingredients
\newcommand{\automaton}{\ensuremath{\mathcal{A}}}
\newcommand{\clocations}{\ensuremath{Q}}
\newcommand{\iclocation}{\ensuremath{q_{\mathit{in}}}}
\newcommand{\fclocation}{\ensuremath{F}}
\newcommand{\transitionrel}{\ensuremath{\Delta}}
\newcommand{\transitionlabel}{\ensuremath{\lambda}}
\newcommand{\clocationlabel}{\ensuremath{\xi}}
\newcommand{\operations}{\ensuremath{Op}}
\newcommand{\operation}{\ensuremath{op}}
\newcommand{\add}[1]{\ensuremath{\mathit{add}(#1)}}
\newcommand{\zero}{\ensuremath{\mathit{zero}}}
\newcommand{\propositions}{\ensuremath{L}}
\newcommand{\configuration}{\ensuremath{c}}
\newcommand{\configurations}{\ensuremath{\mathcal{C}}}
\newcommand{\run}{\ensuremath{\pi}}
\newcommand{\trace}{\ensuremath{\tau}}
\newcommand{\pocatuple}{\ensuremath{\langle \clocations,
\iclocation, \fclocation, \transitionrel, \transitionlabel,
\clocationlabel \rangle}}

\newcommand{\circuit}{\ensuremath{\mathcal{C}}}

\newcommand{\tm}{\ensuremath{\mathcal{M}}}

%%% Names
\newcommand{\oca}{\textsc{OCA} }
\newcommand{\poca}{\textsc{POCA} }

%%% Problems
\newcommand{\stsat}{\textsc{Succinct 3-SAT} }
\newcommand{\tsat}{\textsc{3-SAT} }
\newcommand{\ltlpoca}{\textsc{LTL-POCA-MC} }

%%% Complexity classes
\newcommand{\nexptime}{\textsc{NExpTime} }
\newcommand{\conexptime}{\textsc{co-NExpTime} }
\newcommand{\pspace}{\textsc{PSpace} }

%%% LTL
\newcommand{\ltl}{\textsc{LTL} }
\newcommand{\until}{\ensuremath{\mathcal{~U~}}}
\newcommand{\propalphabet}{\ensuremath{\Sigma}}
\newcommand{\pastarred}{\ensuremath{(\mathcal{P}(\propalphabet))^{\omega}}}

%%% Logic
\newcommand{\true}{\ensuremath{\mathit{true}}}
\newcommand{\false}{\ensuremath{\mathit{false}}}

%%% Control commands
\newcommand{\defemph}[1]{\emph{#1}}

\begin{document}
\section{LTL Model Checking}
\begin{figure}
  \begin{center}
    \includegraphics{graphics/divisibility-1.pdf}
  \end{center}
  \begin{center}
    \includegraphics{graphics/divisibility-2.pdf}
  \end{center}
  \caption{Gadgets for testing the counter for divisibility and
    non-divisibility with $p$.}
  \label{fig:divisibility-gadgets}
\end{figure}
\begin{figure}
  \begin{center}
    \includegraphics{graphics/divisibility-0.pdf}
  \end{center}
  \caption{Gadget for adding a number between $0$ and $-2^{k+1}+1$ to
    the counter.}
  \label{fig:number-diamond}
\end{figure}

Let $\propalphabet$ be a set of \defemph{propsions}. In the following,
capital arabic letters such as $A$ will identify elements from
$\propalphabet$.  The set of \ltl formula is inductively defined
according to the following grammar:
\begin{equation*}
  \varphi ::= A \mid \neg \varphi \mid \varphi_1 \wedge \varphi_2 \mid
  \bigcirc \varphi \mid \varphi_1 \until \varphi_2
\end{equation*}
We define the Boolean abbreviations in a standard way, i.e.,
$\varphi_1 \vee \varphi_2 := \neg (\neg \varphi_1 \wedge \neg
\varphi_2)$, $\varphi_1\rightarrow \varphi_2:=\neg \varphi_1 \vee
\varphi_2$ and moreover $\top:=A\vee \neg A$ for some $A\in
\propalphabet$. The temporal abbreviations $\Diamond$
(\defemph{eventually}) and $\Box$ (\defemph{globally}) are defined as
$\Diamond \varphi:= \top \until \varphi$ and $\Box \varphi := \neg
\Diamond \neg \varphi$.  Let $\trace \in \pastarred$ be a trace, the
containment of $\trace$ in the trace language of $\varphi$ ($\trace
\models \varphi$) is defined by induction on the structure of
$\varphi$:
\begin{eqnarray*}
  \trace \models  A &  \iff & A\in\trace(0)\\
  \trace \models  \neg \varphi &\iff & \trace \not \models \varphi\\
  \trace \models \varphi_1 \wedge \varphi_2 &\iff& \trace
  \models \varphi_1 \text{ and } \trace \models \varphi_2\\
  \trace \models \bigcirc \varphi & \iff & \tau^1 \models \varphi\\
  \trace \models  \varphi_1 \until \varphi_2 & \iff &
  \text{there is } j\in\naturals \text{ s.t. } \tau^j\models \varphi_2
  \text{ and for all } i<j: \tau^j\models \varphi_1
\end{eqnarray*}
Thus, the language of a formula $\varphi$ is $L(\varphi):=\{ \tau \in
\pastarred : \tau\models \varphi\}$.  Given an \oca $\automaton$, we
write $\automaton\models \varphi$ iff $L(\automaton)\subseteq
L(\varphi)$. We can now define the LTL model checking problem for
parametric one-counter automata.
\begin{definition}[$\ltlpoca$]
  Let $\automaton$ be a $\poca$ and $\varphi$ an LTL formula. We write
  $\automaton \models \varphi$ iff for all assignments $\sigma:
  \rho(\automaton)\to \integers$ we have $\automaton(\sigma) \models
  \varphi$.  The \defemph{LTL model checking of parametric one-counter
    automata ($\ltlpoca$)} is to decide $\mathcal{A}\models \varphi$.
\end{definition}
A general automata theoretic approach to check a system given by a
finite state automaton $\automaton$ against an LTL specification
$\varphi$ was introduced by Vardi and Wolper in
\cite{VW86-lics}. Their idea is to translate the negation of
$\varphi$, which captures all bad behaviours, into a B\"uchi automaton
$\automaton_{\neg \varphi}$ and then to check for language emptiness
of the product automaton $\mathcal{B}=\automaton\times
\automaton_{\neg \varphi}$. Moreover, they shown that the size of the
automaton $\automaton_{\neg \varphi}$ is in
$\mathcal{O}(2^{|\varphi|})$, hence $|\mathcal{B}|\in \mathcal{O}
(|\automaton|\times 2^{|\varphi|})$.

We can employ this technique in a straight-forward way in our
setting. Given a \poca $\automaton$ and an LTL formula $\varphi$, in
order to check that $\automaton \not \models \varphi$, we need to find
an assignment $\sigma$ to $\rho(\automaton)$ such that
$\automaton(\sigma)\not\models \varphi$. Consequently, it is
sufficient to check whether $L(\automaton \times \automaton_{\neg
\varphi}) \neq \emptyset$, for which we have established the
complexity in the previous section.
\begin{proposition}
  \ltlpoca is in $\conexptime$.
\end{proposition}
Subsequently, we show that this upper bound cannot be improved.
Our prototype problem for showing $\conexptime$-hardness is $\stsat$
\cite{P94-book}. As in classical $\tsat$, the problem posed by an
instance of $\stsat$ is to find a satisfying assigment of a
propositional formula $\chi$ in 3-DNF. The difference however is that
$\chi$ is represented by a Boolean circuit. This representation allows
for an exponentially succinct representation of $\chi$, which renders
\stsat $\nexptime$-complete \cite{P94-book}.
\begin{definition}[\stsat]
  An instance of \defemph{$\stsat$} is a Boolean circuit $\circuit$
  with input $y;\ell$, where $y\in \{0, \ldots, 2^m-1\}$ is a number
  encoded in binary that identifies a clause for some arbitrary but
  fixed $m\in\naturals$, and $1\le \ell\le 3$ accesses the respective
  literal of the clause. The output of $\circuit$ is $z;b$, where
  $1\le z\le t$ is a number encoded in binary identifying a variable
  $X_z$ for some fixed $1\le t\le 3\cdot 2^m$, and $b\in\{0,1\}$
  indicates whether or not $X_z$ is negated. The propositional formula
  encoded by $\circuit$ is
  \begin{equation*}
    \chi(X_1,\ldots,X_t) = \bigwedge_{0\le y<2^m}\left(
    \bigvee_{\circuit(y,\ell)=z;0\atop 1\le \ell\le 3} \neg X_z
    \vee \bigvee_{C(y,\ell)=z;1\atop 1\le \ell \le 3} X_z \right).
  \end{equation*}
  \stsat is to decide whether there exists a valuation $v: \{X_1,
  \ldots X_{t}\}\to \{\true, \false\}$ such that $\chi$ evaluates to
  true under $v$.
\end{definition}
In order to establish $\conexptime$-hardness of $\ltlpoca$, given an
instance $\circuit$ of \stsat we construct a \poca $\automaton$ with
only one parameter and an LTL formula $\varphi$ such that the formula
$\chi$ induced by $\circuit$ is satisfiable iff $\automaton
\not\models \neg\varphi$. There are two key ideas for the reduction:
First, we choose a suitable encoding of arbirtray valuations in the
naturals so that we can use the parameter to guess and store
them. Second, we employ the expressive power of LTL in order to
perform arbitrary computations of \pspace Turing machines on traces of
runs of $\automaton$.

Regarding the encoding, let $X_1,\ldots, X_t$ be the propositional
variables ocurring in $\chi$, $p_1, \ldots, p_t$ the first $t$ prime
numbers and $a\in \naturals$. We define a valuation $v:\{X_1,\ldots,
X_t\}\to \{\true,\false\}$ such that $a\equiv 0 \mod p_i$ iff
$v(X_i)=\false, 1\le i\le t$. Thus, any number $a$ corresponds to a
valuation. Conversely, the Chinese Remainder Theorem guarantees that
for any valuation a suitable $a$ exists. Note that the Prime Number
Theorem ensures that the binary representation of the $p_i$ is
polynomial in $t$.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.9]{graphics/structure-0.pdf}
  \end{center}
  \caption{High-level description of the automaton $\automaton$ used for
  the reduction.}
  \label{fig:high-level}
\end{figure}

Let us now take a high-level look at the $\poca$ $\automaton$ that is
used for the reduction and sketched in Figure \ref{fig:high-level}.
$\automaton$ uses one parameter $a$ and employs several gadgets
(depicted as rectangles) that communicate via designated propsitional
variables. First, $\automaton$ loads the value of the parameter $a$
into the counter. It then passes through a gadget that initially
chooses a number $c$ identifying a clause of $\chi$ encoded as a
sequence of propositions. This gadget is constructed such that it
outputs the increment of $c$ modulo $2^m$ each time it is passed
thereafter. Now $\automaton$ branches non-deterministically into a
gadget $\automaton_\circuit$ in order to choose one literal of the
respective clause, i.e., evaluates the circuit $\circuit$ for input
$c;1, c;2$, respectively $c;3$. The identifier $z$ of the variable
$X_z$ of the respective literal, again encoded as a sequence of
propositions, is then used as input to a gadget $\automaton_p$ which
computes the $z$-th prime number $p_z$. Having computed $p_z$, if
$b=0$, i.e., $X_z$ is negative in $c$, it is checked that
$p_z|a$. Symmetrically, if $X_z$ is positive it is checked that
$p_z\not|a$. After the checks have been finished, it is ensured that
the counter has value $a$ and the process continues with clause $c+1
\mod 2^m$. Given an LTL formula $\varphi$ that ensures that all
gadgets work and communicate properly, it is clear that $\chi$ is
satisfiable iff $\automaton \not \models \neg \varphi$.

So it remains to show how the gadgets can be realized. Observe that
the computations performed by the gadgets $\automaton_\circuit$ and
$\automaton_p$ can be realized by $\pspace$ Turing machines:
evaluating a Boolean circuit clearly is in $\pspace$, as well is
computing the $i$-th prime number for $1\le i\le t$. Therefore, we
subsequently show how a generic $\pspace$ Turing machine can be
encoded as an \oca together with a family of LTL formulas that ensure
proper behaviour and communication.

\begin{figure}
\begin{center}
\includegraphics{graphics/tm-oca-0.pdf}
\end{center}
\caption{The \poca $\automaton_\tm$ used to generically perform
  computations of arbitrary \pspace Turing machines.}
\label{fig:tm}
\end{figure}
% assume TM goes into infinite loop if it does not accept
Let $\tm=\langle S, \{0, 1\}, s_0, \delta, F \rangle$ be a
deterministic Turing machine which does not use more than $n$ tape
cells during a computation, where $n$ is polynomial in
$|\tm|$. Following standard methods for showing $\pspace$-hardness of
LTL satisfiability, in order to encode computations of $\tm$ we
introduce propositions $T_{i,b}^\epsilon$ and $T^s_{i,b}$, $1\le i\le
n, b\in\{0,1\}, s\in S$. When simulating a computation,
$T_{i,b}^\epsilon$ indicates that the content of the $i$-th tape cell
is $b$ and $T_{i,b}^s$ additionally signals that the head currently
reads the $i$-th tape cell in state $s\in S$.  For technical
convenience, in the following $T_{n+1,b}^s$ abbreviates $\top$ for any
$s\in S\cup \{\epsilon\}, b\in\{0,1\}$.

Consider now the \poca $\automaton_\tm$ shown in Figure
\ref{fig:tm}. For each $1\le i\le n$ it contains a column of locations
such that each $T_{i,b}^\epsilon$ and $T_{i,b}^s$ occurs as a label of
exactly one location.  When simulating a computation of $\tm$, as long
as we have not reached an accepting state, we loop through the upper
component of $\automaton_\tm$. Once the simulated computation reaches
an accepting state, the final tape content is encoded into a trace of
a run in the lower component of $\automaton_\tm$, i.e., we pass
through a location labelled with $A_{i,b}$ iff the $i$-th tape cell
contains $b$ when the computation has finished for $1\le i\le n, b\in
\{0,1\}$.  The propositional variable $\$$ is used as a special marker
the allows for detecting when the upper component of $\automaton_\tm$
is traversed for the first time. We now give LTL formulas that
restrict the traces of runs of $\automaton_\tm$ to correspond to
computations of $\tm$.

At first, we need to make sure that the path we initially take in the
upper component corresponds to an initial configuration of $\tm$. As
previously said, we want to be able to use the output of one Turing
machine as the input to another.  Assume the output is encoded by
propositions $A_{1,b_1}',\ldots, A_{n,b_n}'$, the following formula
ensures that they are correctly reused as input to $\automaton_\tm$:
\begin{equation*}
  \Box \left( \bigwedge_{b\in \{0,1\}}(A_{1,b}' \rightarrow (( \neg
  T_{1,\bar{b}}^{s_i}\wedge \bigwedge_{s\in (S\setminus\{s_i\}\cup
    \{\epsilon\},\atop b'\in\{0,1\}} \neg T_{1,b'}^s) \until \$)
  \wedge \bigwedge_{2\le i\le n,\atop b\in\{0,1\}} A'_{i,b}
  \rightarrow ((\neg T_{i,\bar{b}}^\epsilon \wedge\bigwedge_{s\in
    S,\atop b'\in \{0,1\}} T_{i,b'}^s)\until \$)) \right)
\end{equation*}
%% by propositions Suppose the
%% initial tape content is given by $b_1,\ldots, b_n \in \{0,1\}$, this
%% can be achieved by the following formula:
%% \begin{equation}
%%   \left(\bigwedge_{s\in (S\setminus \{s_i\} \cup \{\epsilon\})\atop b\in\{0,1\}}
%%   \neg T_{1,b}^s\wedge \neg T_{1,\bar{b_1}}^{s_i}
%%   \wedge
%%   \bigwedge_{2\le i\le n} \bigwedge_{s\in S\atop
%%     b\in\{0,1\}}\neg T_{i,b}^s \wedge \neg
%%   T_{i,\bar{b_i}}^\epsilon\right)
%%   \until \$
%% \end{equation}
The reason for writing the above formula in this form is that it
allows us to transfer inputs from outputs of Turing machines even if
they are not directly connected to each other.

%todo: Add reason why we write formula like that
We can now give LTL formulas that constraint runs of $\automaton_\tm$
such that the propositions seen in the corresponding trace
successively encode valid configurations of $\tm$ with respect to
$\delta$.  First, whenever we see three consecutive propositions
indicating that the head is currently not reading any tape cell
corresponding to them, we require that the content of the second does
not change in the next encoded configuration. A special case is the
first tape cell, which needs to be treated seperately. The formula
enforcing this constraint is
\begin{equation*}
   \Box\Bigg(
     \bigwedge_{b_1,b_2\in \{0,1\}} \Big(
  (T_{1,b_1}^\epsilon \wedge \bigcirc^2 T_{2,b_2}^\epsilon) \rightarrow
  \bigcirc^{2n+1} (T_{1,b_1}^\epsilon \vee A_{1,b_1})
  \Big)\wedge
\end{equation*}
\begin{equation*}
  \wedge\bigwedge_{1\le i\le n-1\atop b_0,b_1,b_2\in \{0,1\}}
  \Big( (T_{i,b_0}^\epsilon \wedge \bigcirc^{2}T_{i+1,b_1}^\epsilon \wedge
  \bigcirc^{4}T_{i+2,b_2}^\epsilon) \rightarrow
  \bigcirc^{2(n+1)+1}(T_{i+1,b_1}^\epsilon\vee A_{i+1,b_1}) \Big)\Bigg)
\end{equation*}
%% \begin{equation}
%% \rightarrow
%%   (\bigcirc^{2n}(T_{i,b_0}\vee A_{i,b_0}) \wedge
%%    \wedge
%%   \bigcirc^{2(n+2)}(T_{i+2,b_2}\vee A_{i+2,b_2})\Big)\Bigg)
%%   \Box(\bigwedge_{1\le i\le n-1\atop b_0,b_1,b_2\in \{0,1\}}
%%   ( (T_{i,b_0} \wedge \bigcirc^{2}T_{i+1,b_1} \wedge
%%   \bigcirc^{4}T_{i+2,b_2}) \rightarrow\\ \rightarrow
%%   (\bigcirc^{2n}(T_{i,b_0}\vee A_{i,b_0}) \wedge
%%   \bigcirc^{2(n+1)}(T_{i+1,b_1}\vee A_{i+1,b_1}) \wedge
%%   \bigcirc^{2(n+2)}(T_{i+2,b_2}\vee A_{i+2,b_2})) \wedge\\
%% \end{equation}
Second, whenever we see a proposition indicating that in the
corresponding configuration the head is reading the respective
tape cell, we need to ensure that all the neighbour tape
cells are updated according to $\delta$:
\begin{equation*}
  \Box\Bigg(\bigwedge_{1\le i\le n-1\atop b_0,b_1,b_2\in \{0,1\}}
  \bigwedge_{s\in S} \Big( (T_{i,b_0}^\epsilon \wedge
  \bigcirc^{2}T_{i+1,b_1}^s \wedge \bigcirc^{4}T_{i+2,b_2}^\epsilon)
  \rightarrow
\end{equation*}
\begin{equation*}
  \rightarrow \left\{
    \begin{array}{ll}
      \bigcirc^{2n+1} T_{i,b_0}^{s'} \wedge \bigcirc^{2(n+1)+1}
      T_{i+1,b}^\epsilon \wedge \bigcirc^{2(n+2)+1} T_{i+2,b_2}^\epsilon))& \text{ if }
      \delta(s,b_1)=(s',b,-1), s'\not\in F\\
      \bigcirc^{2n+1} T_{i,b_1}^\epsilon \wedge \bigcirc^{2(n+1)+1}
      T_{i+1,b}^\epsilon \wedge \bigcirc^{2(n+2)+1} T_{i+2,b_2}^{s'}))& \text{ if }
      \delta(s,b_1)=(s',b,+1),s'\not\in F\\
      \bigcirc^{2n+1} A_{i,b_1} \wedge \bigcirc^{2(n+1)+1}
      A_{i+1,b} \wedge \bigcirc^{2(n+2)+1} A_{i+2,b_2}))& \text{ if }
      \delta(s,b_1)=(s',b,d), s'\in F
    \end{array}
    \right.
\end{equation*}
%% \begin{equation*}
%%     \wedge
%%     (T_{0,b_0}^q \wedge \bigcirc^{2} T_{1,b_1}) \rightarrow
%%     \left \{
%%     \begin{array}{ll}
%%       \bigcirc^{2n} T_{0,b} \wedge \bigcirc^{2(n+1)} T_{1,b_1}))
%%       & \text{ if } \delta(q,b_1)=(q',b,R) \text{ and } q'\neq q_f\\
%%       \bigcirc^{2n} A_{0,b} \wedge \bigcirc^{2(n+1)} A_{1,b_1}))
%%       & \text{ if } \delta(q,b_1)=(q',b,R) \text{ and } q'= q_f\\
%%     \end{array}
%%     \right.
%% \end{equation*}
Here, $A_{n+1,b}$ also abbreviates $\top$ for $b\in\{0,1\}$. In order
to improve readability, we have omitted the special case when the head
reads the first tape cell. The above formula can however
straight-forwardly be adopted to take this into account.

Coming back to Figure \ref{fig:high-level}, it now is easy to see how
we can implement a gadget that chooses a clause $c$ and subsequently
increments $c$ modulo $2^m$ each time it is passed. It is just a
$\pspace$ Turing machine that uses its own output as input and
increments it modulo $2^m$. A valid input for the first time it is
used can be provided by a suitable LTL formula.

We contiune with showing how the counter value can be tested for
{(non\-)} divisibility with a prime number that has previously been
computed. Suppose the prime number is encoded in propositions
$A_{1,b_1},\ldots,A_{k+1,b_{k+1}}$. As previously mentioned, Figure
\ref{fig:divisibility-gadgets} provides gadgets that can test the
counter for (non-)divisibility with some parametric value $p$.
Observe that both gadgets ensure that the counter value is $0$ when
they are left. We can obtain a generic version of those gadgets by
replacing every edge that is labelled with $\mathit{add}(-p)$ with the
gadget from Figure \ref{fig:number-diamond}. We just have constraint
the runs such that each time we pass through the gadget from Figure
\ref{fig:number-diamond} we take a path whose sum adds up to the
number encoded by the $A_{i,b_i}$. This is ensured by the following
formula:
\begin{equation*}
  \Box \left( \bigwedge_{1\le i\le k+1,\atop b\in\{0,1\}} A_{i,b}
  \rightarrow (\neg N_{i-1,\bar{b}}\until \$) \right)
\end{equation*}

Finally, it is not difficult to see that the techniques we used in the
previous paragraphs can also be used to correctly handle the branching
on $b$ performed in Figure \ref{fig:high-level}.

In summary, we can take the disjoint union of all the gadgets from
Figure \ref{fig:high-level} and their appendant LTL formula that we
described in this section, connect the gadgets correctly and take the
conjuction of the LTL formulas in order to obtain an automaton
$\automaton$ and a formula $\varphi$ such that $\automaton \not
\models \neg \varphi$ iff $\chi$ is satisfiable.

\begin{theorem}
  \ltlpoca is $\conexptime$-complete.
\end{theorem}

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2023-10-13 version-0-VMCAI/definitions.tex
section

--- 2023-08-22 definitions.tex
@@
 for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
-Let $X,Y \subseteq S$ and $k \in \mathbb{N}$.
-We denote $X \longrightarrow^{*} Y$ (resp.
+Let $X,Y \subseteq S$ and $k \in \mathbb{N}$; we denote $X \longrightarrow^{*} Y$ (resp.
 $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp.
@@
 %
-Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with $V$.
-For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+Subsetquently, $Q \times \N^d$ is the set of configurations of the transition system associated with a $d$-VASS $V$.
+For all configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
 %
@@
 
-A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence it can be only defined by $T$.

### 2023-10-13 VMCAI-revised-version/final.tex
section

--- 2023-10-12 VCMAI revised version/final.tex
@@
 
-Several families of formal models of processes give rise to WSTSs in a natural way, e.g.
-Petri nets when inclusion between markings is used as the well-ordering and lossy channel systems with the subword ordering.
+Several families of formal models of processes give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
+compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets and data Petri nets), reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
+Similarly, lossy channel systems with $d$ channels and $Q$ a finite set of control-states are WSTS (with a non-strong compatibility)
+for the ordering $=\times \sqsubseteq^d$ (where $\sqsubseteq$ is the subword ordering on $\Sigma^*$) on $S= Q \times (\Sigma^*)^d$.\\
+%%%%
+
+By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that possesses an excellent characteristic that allows to obtain more algorithms:
+
+\begin{definition}
+A wqo $(X, \leq)$ is an \emph{$\omega^2$-wqo} if $(Ideals(X), \subseteq)$ forms a wqo.
+\end{definition}
+
+%Using Proposition 4.3 in \cite{DBLP:journals/corr/abs-1208-4549}, we recall a simple definition of a strict subclass of wqos that has a very nice property: a wqo $(X,\leq)$ is an $\omega^2$-wqo iff $(Ideals(X),\subseteq)$ is a wqo.
+
+Although there exists wqo that are not $\omega^2$-wqo (see the Rado ordering in \cite{DBLP:journals/ipl/Jancar99}), all naturally occurring wqos are $\omega^2$-wqo, perhaps to the notable exception of ﬁnite graphs quasi-ordered by the graph minor relation, which are wqo but not known to be $\omega^2$-wqo.
+The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is a an $\omega^2$-wqo (Proposition 4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
+Now we are able to define $\omega^2$-WSTS:
+
+\begin{definition} \cite{DBLP:journals/corr/abs-1208-4549}
+A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
+is an $\omega^2$-WSTS if $(S, \leq)$ is an $\omega^2$-wqo.
+\end{definition}
+Since almost of usual wqo are $\omega^2$-wqo, all naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, reset/transfer VASS are $\omega^2$-WSTS.\\
+%%%%%%
 

### 2023-10-13 VMCAI-revised-version/final.tex
up

@@
 \end{lemma}
-
-\begin{proof}
-Let us suppose that $A \cap U \subseteq D$.
-Then ${\downarrow (A \cap U)} \subseteq {\downarrow D} = D$.
-Let us show that $({\downarrow A}) \cap U \subseteq {\downarrow (A \cap U)}$.
-Let $x \in ({\downarrow A}) \cap U$, then there exists $a \in A$ such that $x \leq a$.
-Since $x \in U$ and $U$ is upward-closed, we also have $a \in U$.
-Hence $a \in A \cap U$ and then $x \in { \downarrow (A \cap U)}$.
-In the other direction, since $A \subseteq {\downarrow A}$, the inclusion $({\downarrow  A}) \cap U \subseteq D$ implies $A \cap U \subseteq ({\uparrow  A}) \cap U \subseteq D$.
-\qed
-\end{proof}
 

### 2023-10-13 VMCAI-revised-version/final.tex
Theorem 10 and theorem 12 are only one thm now

@@
 
+Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.
+
 \begin{theorem}\label{SL VASS}
-{\sc Resilience} is decidable for VASS when $\Safe$  is a semilinear set.
+{\sc Resilience} is decidable for VASS, and integer VASS, when $\Safe$  is a semilinear set.
 \end{theorem}
@@
 \begin{proof}
-We consider the case when $\Safe$ is semilinear
+We consider in VASS the case of when $\Safe$ is semilinear
 and $\Bad = S \setminus \Safe$ is semilinear too.
@@
 $\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
-{\sc Resilience} is decidable. \qed
+{\sc Resilience} is decidable. 
+The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^{2d}$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\qed
 \end{proof}
@@
 \qed
-\end{proof}
-
-Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.
-
-\begin{theorem}\label{SL int-VASS}
-{\sc Resilience} is decidable for integer VASS when $\Safe$ 
-is a
-semilinear set.
-\end{theorem}
-
-\begin{proof}
-The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^{2d}$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
 \end{proof}

### 2023-10-13 VMCAI-revised-version/final.tex
section

@@
 
+There exists another type of WSTS which enjoy the downward compatibility.
+
+\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
+A {\em downward compatible well-structured transition system} (dc-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
+is a transition system $(S, \rightarrow)$
+equipped with a wqo ${\leq} \subseteq S \times S$ such that   
+the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, there exists 
+	$t_2 \in S$ with 
+	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.
+\end{definition}
+
+Let us reformulate the downward compatibility property.
+
+\begin{lemma}
+For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible; (2) for all downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
+\end{lemma}
+
+To simplify the notations, we write without ambiguity  $\pred_{\mathscr{S}}^*(D)$ by  $\pred^*(D)$.
+\begin{proof}
+Let us prove $1 \implies 2$. 
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \downarrow \pred^*(D)$.
+By downward closure, there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$.
+By definition of $\pred^*(D)$ then there exists 
+$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+%
+By downward compatibility $a_0 \to a_1$
+implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
+$x \to^* a'_1$.
+More generally $a_i \to a_{i+1}$ and
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
+$a'_i \to^* a'_{i+1}$,
+and, by induction,
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+with $d' \leq d$.
+Since
+$d'$ 
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\downarrow\pred^*(D) \subseteq \pred^*(D)$.
+Since $\pred^*(D) \subseteq \downarrow \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\downarrow\pred^*(D)$ is downward-closed.
+
+Let us prove $2 \implies 1$.
+Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Consider $\pred^*(\downarrow t_1)$; since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\downarrow t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\downarrow t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\downarrow t_1)$.
+This means that there exists $t_2 \leq t_1$ such that
+$s_2 \rightarrow^{*} t_2$. Hence, we proved that $\mathscr{S}$ is downward compatible.
+\qed
+
+\end{proof}
+
 \noindent
@@
 
-Let us recall that a system $\mathscr{S}=(S,\rightarrow, \leq)$ is \emph{downward compatible} \cite{DBLP:journals/tcs/FinkelS01} if
-for all $s_1, s_2, t_1 \in S$ with
-$s_2 \leq s_1$ and $s_1 \to t_1$
-there
-exists $t_2 \in S$ with
-$t_2 \leq t_1$ and $s_2 \to^* t_2$.
-
 \begin{corollary}
@@
 
-Let $D$ be a downward-closed subset of $S$
-and let $x \in \downarrow \pred^*(D)$.
-By downward closure, there exists
-$y \in \pred^*(D)$ 
-such that $x \leq y$.
-By definition of $\pred^*(D)$ then there exists 
-$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
-$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
-
-By downward compatibility $a_0 \to a_1$
-implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
-$x \to^* a'_1$.
-More generally $a_i \to a_{i+1}$ and
-$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
-$a'_i \to^* a'_{i+1}$,
-and, by induction,
- $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
-with $d' \leq d$.
-Since
-$d'$ 
-belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$. \qed
+%
+%Let $D$ be a downward-closed subset of $S$
+%and let $x \in \downarrow \pred^*(D)$.
+%By downward closure, there exists
+%$y \in \pred^*(D)$ 
+%such that $x \leq y$.
+%By definition of $\pred^*(D)$ then there exists 
+%$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+%$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+%
+%By downward compatibility $a_0 \to a_1$
+%implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
+%$x \to^* a'_1$.
+%More generally $a_i \to a_{i+1}$ and
+%$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
+%$a'_i \to^* a'_{i+1}$,
+%and, by induction,
+% $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+%with $d' \leq d$.
+%Since
+%$d'$ 
+%belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$. \qed
 \end{proof}

### 2023-10-13 VMCAI-revised-version/final.tex
section

@@
 The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
- $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
+ $\post_{\mathscr{S}}(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
 The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
- $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
+ $\pred_{\mathscr{S}}(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
+To simplify the notations, we write without ambiguity $\post_{\mathscr{S}}(D)$ by $\post(D)$ and  $\pred_{\mathscr{S}}(D)$ by  $\pred(D)$.
 By iterating $\pred$ and $\post$ we obtain  
@@
 
-To simplify the notations, we write without ambiguity  $\pred_{\mathscr{S}}^*(D)$ by  $\pred^*(D)$.
 \begin{proof}

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 Let us prove $2 \implies 1$.
-Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Consider $\pred^*(\downarrow t_1)$; since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\downarrow t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\downarrow t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\downarrow t_1)$.
+Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\downarrow t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\downarrow t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\downarrow t_1)$.
 This means that there exists $t_2 \leq t_1$ such that

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 
-\begin{lemma}
+\begin{lemma}\label{downward compatible}
 For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible; (2) for all downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for ideal-effective WSTS with 
-$\Safe=\downarrow \Safe$
-and
-the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+{\sc Resilience}  is decidable for ideal-effective downward compatible WSTS with 
+$\Safe=\downarrow \Safe$.
 \end{theorem}
@@
 \begin{proof}
-By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
+By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
 The resilience problem can be reformulated as 
@@
 
-\begin{corollary}
-{\sc Resilience} is decidable for ideal-effective downward-compatible WSTS with 
-$\Safe=\downarrow \Safe$.
-\end{corollary}
-
-\begin{proof}
-
-%
-%Let $D$ be a downward-closed subset of $S$
-%and let $x \in \downarrow \pred^*(D)$.
-%By downward closure, there exists
-%$y \in \pred^*(D)$ 
-%such that $x \leq y$.
-%By definition of $\pred^*(D)$ then there exists 
-%$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
-%$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
-%
-%By downward compatibility $a_0 \to a_1$
-%implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
-%$x \to^* a'_1$.
-%More generally $a_i \to a_{i+1}$ and
-%$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
-%$a'_i \to^* a'_{i+1}$,
-%and, by induction,
-% $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
-%with $d' \leq d$.
-%Since
-%$d'$ 
-%belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$. \qed
-\end{proof}
-
 In the case
-of a ideal-effective WSTS 
-where
-the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed
-is not met,
+of a ideal-effective WSTS, not necesarly downward compatible,
 the above construction
@@
 In the case
-of a ideal-effective WSTS 
-where the hypothesis that
-{for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed}
-is not met, {\sc Resilience} is undecidable.
+of a ideal-effective WSTS, {\sc Resilience} is undecidable.
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 
-With the ideal-effective hypothesis, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.
-
-\begin{theorem}\cite{BFM-ic17}
-Coverability is decidable for any ideal-effective ordered upward-compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
+Let us recall that a Well Behaved Transition Systems (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ contains no infinite antichains (but it can be not well founded).
+The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
+With the ideal-effective hypothesis, we obtain the decidability of coverability for WBTS.
+
+\begin{theorem}\cite{DBLP:journals/lmcs/BlondinFM17}
+Coverability is decidable for any ideal-effective WBTS.
 \end{theorem}

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 %
-Wqo admits many other equivalent formulations. 
+Wqo admits many other equivalent formulations (see for example \cite{schmitz:cel-00727025}). 
 %
@@
 
-Let us recall that a Well Behaved Transition Systems (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ contains no infinite antichains (but it can be not well founded).
+Let us recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
 The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 %
-Wqo admits many other equivalent formulations (see for example \cite{schmitz:cel-00727025}). 
+Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff $(X,\leq)$ is well founded (i.e. there is no infinite strictly decreasing sequence
+of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements). See other equivalences in \cite{schmitz:cel-00727025}. 
 %
@@
 Quasi-orderings that have no 
-infinite subset of mutually incomparable elements (antichains)
+infinite antichains
  enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$.

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 %
-Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff $(X,\leq)$ is well founded (i.e. there is no infinite strictly decreasing sequence
-of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements). See other equivalences in \cite{schmitz:cel-00727025}. 
+Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$ iff $(X,\leq)$ is well founded (i.e.
+there is no infinite strictly decreasing sequence
+of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements of $X$). See other equivalences in \cite{schmitz:cel-00727025}. 
 %

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
-%
-%
-%
-Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$ iff $(X,\leq)$ is well founded (i.e.
-there is no infinite strictly decreasing sequence
+This property is equivalent to : every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$.
+%
+%
+%
+Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff $(X,\leq)$ is well founded (i.e. there is no infinite strictly decreasing sequence
 of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements of $X$). See other equivalences in \cite{schmitz:cel-00727025}. 

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+The completion is always finitely branching but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
+$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS.
+A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
+it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+ $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+ $s_1, s_2, \ldots, s_m \in S$.
+Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).\\
+
 Let us recall the  definition of vector addition system with (control-)states. 
@@
 
-In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
+In a (not necessarly ordered) transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
 The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe)$ is to decide whether from \emph{any} state in 
@@
 Resilience is then akin to the Home-Space problem (defined in the introduction) for the set $\Safe$.
+Resilience can also be viewed as a generalizeation
+ of coverability, as
+it asks whether for \emph{every} element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
 We use the notation 
@@
 the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.
-
+In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.  \\
+
+Surprinsingly, the general undecidability statement about resilience was not known neither proved. 
+We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility.
+
+\begin{theorem}\label{indec WSTS}
+Resilience for $\omega^2$-WSTS with strong compatibility and for upward closed set $\Safe$ is undecidable.
+\end{theorem}
+
+\begin{proof}
+Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
+increasing recursive functions from \cite{FMP-ic04} defined as
+
+$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+
+	\noindent	
+	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
+	which moreover begins
+by writing the integer $j$ on its tape,	
+	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+	The
+	transition system 
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+	is a $\omega^2$-WSTS with strong compatibility
+	and
+	has the property that
+	it is
+	$\uparrow (1,1)$-resilient
+	iff TM$_j$ halts on input $j$,
+	since $(1,1)$ is coverable from $(0,0)$
+	iff TM$_j$ halts on input $j$. 
+	Hence there is no Turing machine which correctly determines 
+	resilience and halts whenever its input is a
+	WSTS.
+Hence resilience for $\omega^2$-WSTS with strong compatibility is undecidable.\qed
+\end{proof}
+
+\alain{ce qui suit est à mettre ailleurs ?}
 {\bf Related problems.} 
@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Resilience can be viewed as a generalizeation
- of coverability, as
-it asks whether for \emph{every} element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.
-
-Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set).
-A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
- $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
- $s_1, s_2, \ldots, s_m \in S$.
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).
-
-Let us recall two other results in \cite{BFM-ic17}.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-%
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we consider the completion-post-effective hypothesis to catch decidability.
 
@@
 \begin{proof}
+Let us first recall two results in \cite{BFM-ic17}.
+Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+%
+
 Let  $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
@@
 
-The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
-Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
-increasing recursive functions from \cite{FMP-ic04} defined as
-
-$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
-		(n,n+k) &\text{ otherwise,} \end{dcases*} $
-
-	\noindent	
-	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
-	which moreover begins
-by writing the integer $j$ on its tape,	
-	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
-	The
-	transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-	is a WSTS
-	and
-	has the property that
-	it is
-	$\uparrow (1,1)$-resilient
-	iff TM$_j$ halts on input $j$,
-	since $(1,1)$ is coverable from $(0,0)$
-	iff TM$_j$ halts on input $j$. 
-	Hence there is no Turing machine which correctly determines 
-	resilience and halts whenever its input is a
-	WSTS.
-Hence resilience for WSTS in general is undecidable.
+%				The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
+%%%
+%%%
+%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
+%increasing recursive functions from \cite{FMP-ic04} defined as
+%
+%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+%		
+%	\noindent	
+%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
+%	which moreover begins
+%by writing the integer $j$ on its tape,	
+%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+%	The
+%	transition system 
+%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+%	is a WSTS
+%	and
+%	has the property that
+%	it is
+%	$\uparrow (1,1)$-resilient
+%	iff TM$_j$ halts on input $j$,
+%	since $(1,1)$ is coverable from $(0,0)$
+%	iff TM$_j$ halts on input $j$. 
+%	Hence there is no Turing machine which correctly determines 
+%	resilience and halts whenever its input is a
+%	WSTS.
+%Hence resilience for WSTS in general is undecidable.
 
@@
 It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.\alain{déjà dit avant}
 

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 
-A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence it can be only defined by $T$.
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence it can be only defined by $T$.\\
+
+VASS can extended with resets.
+
+\begin{definition}
+A {\em reset-VASS} in dimension $d$ 
+ is a finite 
+labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, 
+$T \subseteq Q \times Op \times Q$
+is the set of {\em control-transitions}, and $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
+		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
+\end{definition}
+
+Again $Q \times \N^d$
+ denotes the set of states of $V$.
+For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
+$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
+\begin{samepage}\begin{itemize}
+\item  $t = (q,add(\textbf{z}),q') \in T$
+and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
+\item $t = (q,reset(\gamma),q') \in T$ 
+and
+$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+\end{itemize} \end{samepage}
 
@@
 
+\alain{ce qui suit est à mettre ici ?}
+{\bf Related problems.} 
+Resilience and the home-space problem are also linked to the 
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+is decidable for Lossy Counter Machine (LCM)
+when $X$ and $Y$ are semi-linear sets.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
+
 These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
@@
 \begin{theorem}\label{indec WSTS}
-Resilience for $\omega^2$-WSTS with strong compatibility and for upward closed set $\Safe$ is undecidable.
+Resilience for post-effective $\omega^2$-WSTS with strong compatibility and for upward closed set $\Safe$ is undecidable.
 \end{theorem}
@@
 	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-	is a $\omega^2$-WSTS with strong compatibility
+	is a post-effective $\omega^2$-WSTS with strong compatibility
 	and
@@
 
-\alain{ce qui suit est à mettre ailleurs ?}
-{\bf Related problems.} 
-Resilience and the home-space problem are also linked to the 
-model-checking of basic reachability and safety formulae. 
-In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-is decidable for Lossy Counter Machine (LCM)
-when $X$ and $Y$ are semi-linear sets.
-Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
- t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
-whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
-``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
-  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
-
 \subsection{Case: $\Safe=\uparrow \Safe$.}\label{safe-up}
@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we consider the completion-post-effective hypothesis to catch decidability.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the completion-post-effective hypothesis) we demonstrate that the three resilience problems are decidable.
 
@@
 \end{theorem}
-
+\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ?}
 \begin{proof}
@@
 however
-it is not enough to conclude.
+it is not enough to conclude.\\
 
@@
 of a ideal-effective WSTS, {\sc Resilience} is undecidable.
+
+\begin{theorem}\label{indec WSTS with dcs}
+Resilience for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+\end{theorem}
+
+\begin{proof}
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,
 which can be simulated by WSTS.
-
 Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
+\alain{pas facile à lire, à réécrire et à repenser}
 This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
@@
 transitions” that set to $0$ some coordinates in the vector. Let us recall their definition here.
-\begin{definition}
-A {\em reset-VASS} in dimension $d$ 
- is a finite 
-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, 
-$T \subseteq Q \times Op \times Q$
-is the set of {\em control-transitions}, and $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
-		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
-
-Again $Q \times \N^d$
- denotes the set of states of $V$.
-For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
-$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when 
-\begin{samepage}\begin{itemize}
-\item  $t = (q,add(\textbf{z}),q') \in T$
-and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-\item $t = (q,reset(\gamma),q') \in T$ 
-and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
-\end{itemize} \end{samepage}
 
 It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.\alain{déjà dit avant}
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective}\qed
+\end{proof}
 
@@
 Remark that {\sc State-resilience} is {\sc Home-space} with input set $\Safe$.
-Since these problems are undecidable for general infinite-state transition systems, we restrict our study to WSTS.
+Since these problems are undecidable for general infinite-state transition systems, we still restrict our study to WSTS.
 As in the Section~3, we study decidability results for $\Safe$ downward-closed and upward-closed.

### 2023-10-14 VMCAI-revised-version/final.tex
section

@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the completion-post-effective hypothesis) we demonstrate that the three resilience problems are decidable.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the completion-post-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
 
@@
 In the case
-of a ideal-effective WSTS, {\sc Resilience} is undecidable.
+of a ideal-effective (upward compatible) WSTS, {\sc Resilience} is undecidable.
 
@@
 \begin{proof}
+\alain{pas facile à lire, à réécrire et à repenser}
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,
@@
 Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-\alain{pas facile à lire, à réécrire et à repenser}
 This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.

### 2023-10-15 VMCAI-revised-version/final.tex
section

@@
 %
-As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.
-%
+As an example, $(\N^d, \leq)$, the set of vectors of $d$ natural numbers (where $d$ is finite) with component-wise order is a wqo.\\
 Quasi-orderings that have no 
 infinite antichains
- enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$.
+ enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$ See for example \cite{BFM-ic17}.
 %
@@
 
+We introduce five types of effectivity that will be necessary to decide some resilence problems for differents models of WSTS.\\
 \noindent
 {\bf On effectivity.}
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be a WSTS. We say that $\mathscr{S}$ is {\em effective} if there exists a pair of algorithms
+Let $\mathscr{S}=(S, \rightarrow, \leq)$ be a WSTS. We say that $\mathscr{S}$ is 
+\begin{itemize}
+\item {\em effective} if there exists a pair of algorithms
 ($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
@@
 $M_\leq$ the ordering relation “$\leq$”.
-We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
+\item {\em post-effective} if $\mathscr{S}$ is effective, and if there
 exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
 on input $x \in X$. 
-We say that $\mathscr{S}$ has {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+\item {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
 %
-We say that $\mathscr{S}$
-is {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
+\item {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-%
-
+\item {\em completion-post-effective} if
+$\mathscr{S}$ is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+ $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+ $s_1, s_2, \ldots, s_m \in S$.
+\end{itemize}
+\alain{le rappel ci-dessous est-il entièrement utile ?}
 Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
@@
 $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS.
-A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
- $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
- $s_1, s_2, \ldots, s_m \in S$.
+%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
+%it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+% $s_1, s_2, \ldots, s_m \in S$.
 Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).\\

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 \end{theorem}
-
+\alain{peut-on ajouter des hypothèses d'effectivity ?
+peut-on montrer que l'exemple n'est pas pred effectif ?
+l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais 3 ?)?
+on voit que l'exemple n'est pas completion-post-effective vu le theoreme qui suit}
 \begin{proof}
@@
 \begin{proof}
-\alain{pas facile à lire, à réécrire et à repenser}
+\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" . preuve pas facile à lire, à réécrire et à repenser}
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 
-Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}. 
+Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
 Reset-VASS extend the basic VASS model with special “reset
-transitions” that set to $0$ some coordinates in the vector. Let us recall their definition here.
-
-It is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for WSTS in general as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective}\qed
+transitions” that set to $0$ some coordinates in the vector. 
+%			Let us recall their definition here.
+
+%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective}\qed
 \end{proof}
@@
 $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.\\
 

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+The completion is always \emph{finitely branching} but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
+It is proved in  \cite{BFM-ic17} that $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS.
+\\
+
 There exists another type of WSTS which enjoy the downward compatibility.
@@
 \qed
-
-\end{proof}
-
-We introduce five types of effectivity that will be necessary to decide some resilence problems for differents models of WSTS.\\
+\end{proof}
+
+%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
+%it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+% $s_1, s_2, \ldots, s_m \in S$.
+
+We introduce five types of effectiveness \cite{BFM-ic17} that will be essential for solving various resilience problems in different models of WSTS.
 \noindent
-{\bf On effectivity.}
+%			{\bf On effectivity.}
 Let $\mathscr{S}=(S, \rightarrow, \leq)$ be a WSTS. We say that $\mathscr{S}$ is 
@@
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-\item {\em completion-post-effective} if
-$\mathscr{S}$ is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+\item {\em completion-post-effective} if $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
  $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
@@
 
-Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
-The completion is always finitely branching but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
-$\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS.
-%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-%it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
-% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
-% $s_1, s_2, \ldots, s_m \in S$.
-Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS (we don't need the pred-basis hypothesis).\\
+Coverability is  decidable for a class of WSTS without the pred-basis hypothesis.
+
+\begin{theorem} (Theorem 44 in \cite{BFM-ic17})
+Coverability is  decidable for completion-post-effective WSTS.
+\end{theorem}
 

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 %			{\bf On effectivity.}
-Let $\mathscr{S}=(S, \rightarrow, \leq)$ be a WSTS. We say that $\mathscr{S}$ is 
+We say that a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
 \begin{itemize}
@@
 %
-\item {\em ideal-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
+\item {\em ideally-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
 to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-\item {\em completion-post-effective} if $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
- $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
- $s_1, s_2, \ldots, s_m \in S$.
+%
+\item {\em completion-post-effective} if $\mathscr{S}$ is ideally-effective and its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective.
+%
+%and there exist an algorithm $M_\downarrow$ that computes 
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+% $s_1, s_2, \ldots, s_m \in S$.
+%
 \end{itemize}
-\alain{le rappel ci-dessous est-il entièrement utile ?}
+\alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
 Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
@@
 
-Coverability is  decidable for a class of WSTS without the pred-basis hypothesis.
-
-\begin{theorem} (Theorem 44 in \cite{BFM-ic17})
+Coverability is  decidable for a class of WSTS without the effective pred-basis hypothesis.
+
+\begin{theorem} \cite{BFM-ic17}
 Coverability is  decidable for completion-post-effective WSTS.
@@
 \end{proof}
+ Let us remark that the transition system 
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not ideally effective neither completion-post-effective.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 This implies that resilience and bounded resilience are equivalent to coverability.
-Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we deduce that both the 
+Since coverability is decidable for completion-post-effective WSTS \cite{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.
@@
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective}\qed
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis....}\qed
 \end{proof}

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 \end{theorem}
-\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ?}
+\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
 \begin{proof}
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities,
+{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities,
 $\Safe = \downarrow \Safe$.
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 $M_\leq$ the ordering relation “$\leq$”.
+%
 \item {\em post-effective} if $\mathscr{S}$ is effective, and if there
@@
 on input $x \in X$.
+%
+\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
+to the encoding of the ideal $\downarrow s$ is computable and inclusion of ideals is decidable.
+%
+\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
+%
 \item {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
-%
-\item {\em ideally-effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a state $s$
-to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
-%
-\item {\em completion-post-effective} if $\mathscr{S}$ is ideally-effective and its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective.
 %
@@
 \end{itemize}
+%Most of usual WSTS enjoy most of all effectivities.
+
 \alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
@@
 The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
-With the ideal-effective hypothesis, we obtain the decidability of coverability for WBTS.
+With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS.
 
 \begin{theorem}\cite{DBLP:journals/lmcs/BlondinFM17}
-Coverability is decidable for any ideal-effective WBTS.
+Coverability is decidable for any ideally-effective WBTS.
 \end{theorem}
@@
 \begin{theorem} \cite{BFM-ic17}
-Coverability is  decidable for completion-post-effective WSTS.
+Coverability is  decidable for completion-post-effective and ideally-effective WSTS.
 \end{theorem}
@@
 \begin{theorem}\label{indec WSTS}
-Resilience for post-effective $\omega^2$-WSTS with strong compatibility and for upward closed set $\Safe$ is undecidable.
+Resilience for post-effective and ideally effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
 \end{theorem}
@@
 peut-on montrer que l'exemple n'est pas pred effectif ?
-l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais 3 ?)?
+l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais pas 3 car ça permettrait de décider l'arrêt ?)?
 on voit que l'exemple n'est pas completion-post-effective vu le theoreme qui suit}
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not ideally effective neither completion-post-effective.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not completion-post-effective.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective and ideally-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 This implies that resilience and bounded resilience are equivalent to coverability.
-Since coverability is decidable for completion-post-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+Since coverability is decidable for completion-post-effective and ideally-effective WSTS \cite{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.
@@
 \begin{proof}
-\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" . preuve pas facile à lire, à réécrire et à repenser}
+\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?...... preuve pas claire, à réécrire}
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,
@@
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis....}\qed
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis, completion et ideally effective....}\qed
 \end{proof}
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities,
+{\sc State-resilience} is decidable for ideally-effective and completion-post-effective WSTS with downward and upward compatibilities,
 $\Safe = \downarrow \Safe$.
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective and completion-post-effective WSTS with strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.
@@
 
-In this section we study VASS. Since they are completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+In this section we study VASS. Since they are ideally-effective and completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
 $\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
@@
 We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general. We also exhibited classes of WSTS with decidable resilience, namely, 
-completion-post-effective $\omega^2$-WSTS with strong compatibility (in the case 
+completion-post-effective and ideally-effective WSTS with strong compatibility (in the case 
  $\Safe = \uparrow \Safe$) and 
 downward-compatible
-ideal-effective WSTS
+ideally-effective WSTS
 (in the case 

### 2023-10-15 VMCAI-revised-version/final.tex
effectivity

@@
 \end{itemize}
-%Most of usual WSTS enjoy most of all effectivities.
+Most of usual WSTS are post-effective and ideally-effective.
 
@@
 \begin{theorem}\label{indec WSTS with dcs}
-Resilience for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
-\end{theorem}
-
-\begin{proof}
-\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?...... preuve pas claire, à réécrire}
+{\sc Resilience} for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+\end{theorem}
+\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?}
+
+\begin{proof}
+\alain{preuve pas claire, à réécrire}
 This stems from the fact that it is undecidable for Minsky machines with more than one counter,
@@
 Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}.
-Reset-VASS extend the basic VASS model with special “reset
-transitions” that set to $0$ some coordinates in the vector. 
+
+%Reset-VASS extend the basic VASS model with special “reset
+%transitions” that set to $0$ some coordinates in the vector. 
 %			Let us recall their definition here.

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 \item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
-to the encoding of the ideal $\downarrow s$ is computable and inclusion of ideals is decidable.
-%
-\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
+to the encoding of the ideal $\downarrow s$ is computable, inclusion of ideals is decidable and $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
+%
+%\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
 %
@@
 \begin{theorem} \cite{BFM-ic17}
-Coverability is  decidable for completion-post-effective and ideally-effective WSTS.
+Coverability is decidable for ideally-effective WSTS.
 \end{theorem}
@@
 
-Surprinsingly, the general undecidability statement about resilience was not known neither proved. 
+Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
 We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility.

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{theorem}\label{indec WSTS}
-Resilience for post-effective and ideally effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
+Resilience for post-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
 \end{theorem}
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not completion-post-effective.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not ideally-effective.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective and ideally-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be an ideally-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 This implies that resilience and bounded resilience are equivalent to coverability.
-Since coverability is decidable for completion-post-effective and ideally-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+Since coverability is decidable for ideally-effective WSTS \cite{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.
@@
 In the case
-of a ideal-effective (upward compatible) WSTS, {\sc Resilience} is undecidable.
+of an ideal-effective (upward compatible) WSTS, {\sc Resilience} is undecidable.
 
@@
 \end{theorem}
-\alain{est-ce sur qu'on peut rajouter l'hypothese "ideal-effective" ? oui. et completion-post-effective ?}
 
@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For finite-branching WSTS, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For effective finite-branching WSTS, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideally-effective and completion-post-effective WSTS with downward and upward compatibilities,
+{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities,
 $\Safe = \downarrow \Safe$.
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective and completion-post-effective WSTS with strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.
@@
 We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
-We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general. We also exhibited classes of WSTS with decidable resilience, namely, 
-completion-post-effective and ideally-effective WSTS with strong compatibility (in the case 
+We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general.
+We also exhibited classes of WSTS with decidable resilience, namely ideally-effective WSTS with strong compatibility (in the case
  $\Safe = \uparrow \Safe$) and 

### 2023-10-16 VMCAI-revised-version/final.tex
up

@@
 Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong upward-compatibility.
-Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility.
 
@@
 and $\Bad = S \setminus \Safe$ is semilinear too.
+We want to use the fact that
+it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
+$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
 {\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
@@
 Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
-\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad iff \quad \Bad \subseteq \pred^*(\Safe).\]
-Since it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
-$\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}, 
-{\sc Resilience} is decidable. 
-The reachability relation of integer $d$-VASS is definable by a sentence in Presburger in $\mathbb{Z}^{2d}$ hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
+and hence, {\sc Resilience} is decidable. 
+The reachability relation of integer $d$-VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
 \qed

### 2023-10-16 VMCAI-revised-version/final.tex
up

@@
 We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have checked found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If the set is finite then we stop the procedure once we have found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
 If it is infinite then
-$k$-resilience does not hold 
-and we eventually find a witness that $k$-resilience does not hold.
+$k$-resilience does not hold, 
+hence we eventually find a witness that $k$-resilience does not hold.
 Indeed, 

### 2023-10-16 VMCAI-revised-version/final.tex
up
effectivity

@@
 
-In the case
-of an ideal-effective (upward compatible) WSTS, {\sc Resilience} is undecidable.
+In the case of (upward compatible) WSTS with ideally-effectivity, {\sc Resilience} is undecidable.
 
 \begin{theorem}\label{indec WSTS with dcs}
-{\sc Resilience} for ideal-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+{\sc Resilience} for ideally-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
 \end{theorem}
@@
 Remark that
-Theorem~\ref{SL Lossy} and
-Theorem~\ref{SL int-VASS}
+Theorem~\ref{SL Lossy} 
 (resp. Theorem~\ref{RP VASS})

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
-With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS.
+With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS (without the effective pred-basis hypothesis).
 
@@
 \end{theorem}
-
-Coverability is  decidable for a class of WSTS without the effective pred-basis hypothesis.
-
-\begin{theorem} \cite{BFM-ic17}
-Coverability is decidable for ideally-effective WSTS.
-\end{theorem}
+%
+%Coverability is  decidable for a class of WSTS without the effective pred-basis hypothesis.
+%
+%\begin{theorem} \cite{BFM-ic17}
+%Coverability is decidable for ideally-effective WSTS.
+%\end{theorem}
 

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 
-\begin{theorem}\label{indec WSTS}
+\begin{proposition}\label{indec WSTS}
 Resilience for post-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
-\end{theorem}
+\end{proposition}
 \alain{peut-on ajouter des hypothèses d'effectivity ?
@@
 
-\begin{theorem}\label{indec WSTS with dcs}
+\begin{proposition}\label{indec WSTS with dcs}
 {\sc Resilience} for ideally-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
-\end{theorem}
+\end{proposition}
 

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 
-We may immediately generalize this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
-
-\begin{corollary}\label{postcomputable}
+We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
+
+\begin{proposition}\label{postcomputable}
 {\sc State-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
@@
 $\Safe=\uparrow \Safe$.
-\end{corollary}
+\end{proposition}
 
@@
 VASS.
+
 However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  
@@
 
-Keeping the $\uparrow \post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+Keeping the $\uparrow \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
 Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
@@
 
-\begin{theorem}\label{srp down up}
+\begin{proposition}\label{srp down up}
 {\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
@@
 $\Safe=\downarrow \Safe$.
-\end{theorem}
+\end{proposition}
 

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 
-\item The three resilience problems are decidable for completion-post-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
-
-\item The resilience problem is decidable for ideal-effective WSTS with 
+\item The three resilience problems are decidable for ideally-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
+
+\item The resilience problem is decidable for ideally-effective WSTS with 
 $\Safe=\downarrow \Safe$
@@
 We also show that removing the effective 
-$\uparrow$ $\post^*$ basis hypothesis leads to undecidability.
-We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility).
+$\uparrow$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, most WSTS don't enjoy this property.
+We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideally-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility).
 %
@@
 %
-\item {\em post-effective} if $\mathscr{S}$ is effective, and if there
+\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
 exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
@@
 %
-\item {\em effective pred-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+\item {\em effective $\pred$-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
 any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
@@
 %
+\item {\em effective $\post^*$-basis} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if there exists an algorithm accepting
+any state $s \in S$ and returning a finite basis of $\uparrow \post^*(s)$.
 \end{itemize}
-Most of usual WSTS are post-effective and ideally-effective.
+
+%Most of usual WSTS are post-effective and ideally-effective.
 
@@
 \begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
-{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for WSTS $S$ with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for effective $\post^*$-basis WSTS with strong compatibility
 when
@@
 \begin{proposition}\label{postcomputable}
-{\sc State-resilience} is decidable for WSTS with strong compatibility and such that $\uparrow \post^*(s)$ is computable for $s \in S$
+{\sc State-resilience} is decidable for effective $\post^*$-basis WSTS with strong compatibility
 when
@@
 \begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for 
- WSTS with effective 
-$\uparrow$ $\post^*$ basis
+{\sc State-resilience} is decidable for effective $\post^*$-basis 
+ WSTS
 when
@@
 
-Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
+Thus in case of effective $\post^*$-basis WSTS (when the compatibility is not strong), we don't know the decidability status of {\sc $k$-state-resilience} and 
 {\sc Bounded-state-resilience}.

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
-Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
+Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $x \leq y$ and $x \in U$ then $y \in U $.
 A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
@@
 its {\em upward-closure},
- $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
+ $\uparrow A = \{x \in X \mid \exists a \in A ~ a \leq y\}$
  and its 
@@
 \begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
-A {\em downward compatible well-structured transition system} (dc-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
+A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
 is a transition system $(S, \rightarrow)$
@@
 \end{theorem}
+
+Downward-WSTS enjoy a powerfull property.
+
+\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01}
+Effective finitely branching downward-WSTS are effective $\post^*$-basis.
+\end{theorem}
 %
@@
 \end{itemize} \end{samepage}
+
+\alain{ici un dessin comme celui que j'ai envoyé sur skype...à rajouter les downward WSTS}
 
@@
 
-The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis. What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis.
+What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
+\alain{One first answer is the class of effective downward-WSTS which are precisely effective $\post^*$-basis}.
+Moreover,
 \"Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for ideal-effective downward compatible WSTS with 
+{\sc Resilience}  is decidable for ideal-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
@@
 
-\begin{theorem}\label{srp up down}
+\begin{proposition}\label{srp up down}
 {\sc State-resilience},
@@
 $\Safe=\uparrow \Safe$.
-\end{theorem}
+\end{proposition}
 
@@
 What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
-\alain{One first answer is the class of effective downward-WSTS which are precisely effective $\post^*$-basis}.
+\alain{One first answer is the class of effective finitely branching downward-WSTS which are precisely effective $\post^*$-basis}.
 Moreover,

### 2023-10-16 VMCAI-revised-version/final.tex
up

@@
 
-{\bf Synthesis of the main decidability results}\label{synthesis}
-\begin{center}
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \ 
-		& $\Safe = {\uparrow \Safe}$~ 
-		 & $\Safe = {\downarrow \Safe}$~ 
- \\ \hline
-   RP 
-   	& Decidable (Thm~\ref{down-up})  
-   		 & Decidable (Thm~\ref{up-down}) 
-    \\ \hline
-   BRP 
-   &  Decidable (Thm~\ref{down-up}) 
-   		 & ?
-    \\ \hline
-      kRP 
-      & Decidable (Thm~\ref{down-up}) 
-      		& ? 
-       \\ \hline
- \end{tabular}
-\end{center}
-
 \section{State-resilience}
@@
 {\sc Bounded-state-resilience}.
-
-{\bf Results synthesis in the case $\Safe = \uparrow \Safe$}
-\begin{center}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{ | l | c | c | c | c |}
-\hline  Hypothesis & strong compatibility ~ & $\uparrow \post^*$ effective 
-			& strong compatibility \textit{+} $\uparrow \post^*$ effective \\ \hline
-   SRP & Undecidable (Thm~\ref{srp up down}) & Decidable (Thm~\ref{post srp})  & Decidable (Thm~\ref{postcomputable})\\ \hline
-   BSRP & Undecidable (Thm~\ref{srp up down}) &  ?  & Decidable (Thm~\ref{ref ozkan}) \\ \hline
-      kSRP & Undecidable (Thm~\ref{srp up down}) & ? & Decidable (Thm~\ref{ref ozkan}) \\ \hline
- \end{tabular}
- }
-\end{center}
 

### 2023-10-16 VMCAI-revised-version/final.tex
up
effectivity

@@
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
-This property is equivalent to : every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$.
+This property is equivalent to the \emph{finite decomposition} property : every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$.
 %
@@
 
-There exists another type of WSTS which enjoy the downward compatibility.
+Another type of WSTS exists that enjoys downward compatibility.
 
@@
 \end{definition}
+
+There are fewer downward-WSTS than WSTS, but we can still mention FIFO automata with insertion errors and Basic Process Algebra (BPA).
 
@@
 \noindent
-\textbf{Proposition~\ref{liveness reset}}
+\textbf{Proposition~\ref{liveness reset}.}
 {\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\

### 2023-10-16 VMCAI-revised-version/final.tex
up

--- 2023-10-16 VMCAI-revised-version/final.tex
@@
 When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\uparrow U_k)_k$ instead.
-With the effective pred-basis hypothesis, we obtain:
-
-\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
+With the effective pred-basis hypothesis, we obtain:\\
+
+\noindent
+\textbf{Theorem.}
+{\em ~\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
  A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
 Hence coverability is decidable.
-\end{theorem}
+}\\
 
@@
 The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
-With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS (without the effective pred-basis hypothesis).
-
-\begin{theorem}\cite{DBLP:journals/lmcs/BlondinFM17}
+With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS (without the effective pred-basis hypothesis).\\
+
+\noindent
+\textbf{Theorem.}
+{\em ~\cite{DBLP:journals/lmcs/BlondinFM17}
 Coverability is decidable for any ideally-effective WBTS.
-\end{theorem}
+}\\
 

### 2023-10-16 VMCAI-revised-version/final.tex
up

--- 2023-10-16 VMCAI-revised-version/final.tex
@@
 
-\begin{theorem}\cite{DBLP:journals/lmcs/BlondinFM17}
+\begin{theorem}(Blondin~\cite{DBLP:journals/lmcs/BlondinFM17})
 Coverability is decidable for any ideally-effective WBTS.
@@
 On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
-\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}\label{ref ozkan}
+\begin{theorem}\label{ref ozkan}
+(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889})
 {\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for effective $\post^*$-basis WSTS with strong compatibility

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 Let us now recall the (most general) definition of well-structured transition systems.
-\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
+\begin{definition}(Finkel \cite{DBLP:journals/iandc/Finkel90})
 A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 
-\begin{definition} \cite{DBLP:journals/corr/abs-1208-4549}
+\begin{definition}(Finkel and Goubault-Larrecq \cite{DBLP:journals/corr/abs-1208-4549}
 A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 
-\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
+\begin{definition}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
 A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 
-\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
+\begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
  A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
@@
 
-\begin{theorem}(Blondin~\cite{DBLP:journals/lmcs/BlondinFM17})
+\begin{theorem}(Blondin and al. ~\cite{DBLP:journals/lmcs/BlondinFM17})
 Coverability is decidable for any ideally-effective WBTS.
@@
 
-\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01}
+\begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
 Effective finitely branching downward-WSTS are effective $\post^*$-basis.
@@
 \begin{theorem}\label{ref ozkan}
-(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889})
+(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
 {\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for effective $\post^*$-basis WSTS with strong compatibility

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 
-\begin{definition}(Finkel and Goubault-Larrecq \cite{DBLP:journals/corr/abs-1208-4549}
+\begin{definition}(Finkel and Goubault-Larrecq \cite{DBLP:journals/corr/abs-1208-4549})
 A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 \begin{lemma}\label{downward compatible}
-For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible; (2) for all downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
+For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
 \end{lemma}
@@
 
-We introduce five types of effectiveness \cite{BFM-ic17} that will be essential for solving various resilience problems in different models of WSTS.
+We recall and introduce five types of \emph{effectiveness} that will be essential for solving various resilience problems in different models of WSTS.
 \noindent
@@
 
-Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$. \qed
-\end{proof}
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.
+\end{proof} \qed
 
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for ideally-effective $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+{\sc Resilience} for ideally-effective and effective pred-basis $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
 \end{proposition}
@@
 \alain{preuve pas claire, à réécrire}
-This stems from the fact that it is undecidable for Minsky machines with more than one counter,
-which can be simulated by WSTS.
-Indeed, {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
+\alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
+%which can be simulated by WSTS.
+First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
+%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
 From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter.
-We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then lower the first two counters until they reach $0$, then, and only then, finally lower the third counter until it reaches $0$.
-Remark that in the construction, from any input with at least $1$ on its third counter, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
-Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in a $3$-counter Minsky machine
-from any input with at least $1$ on its third counter is undecidable. 
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input in $\uparrow(0,0,1)$ 
+%			with at least $1$ on its third counter. 
+We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
+Remark that in the construction, from any input in $\uparrow(0,0,1)$, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
+Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61,Min67}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
-
-Executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}.
-
+\alain{écrire qqchose ici}
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
+\alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
 %Reset-VASS extend the basic VASS model with special “reset
@@
 %			Let us recall their definition here.
-
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}. 
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for $\omega^2$-WSTS as well.\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis, completion et ideally effective....}\qed
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, $\post^*$-basis $\omega^2$-WSTS with strong compatibility as well.\qed
+%\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
 \end{proof}
@@
 
-The computability of $\uparrow \post^*(s)$ however seems a strong hypothesis.
-What are the WSTS for which $\uparrow \post^*(s)$ is computable for $s \in S$ ?
-\alain{One first answer is the class of effective finitely branching downward-WSTS which are precisely effective $\post^*$-basis}.
-Moreover,
 \"Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
@@
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
+{\sc State-resilience} is undecidable for ideally-effective and effective pred-basis $\omega^2$-WSTS with strong compatibility
 when

### 2023-10-16 VMCAI-revised-version/final.tex
effectivity

@@
 A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. 
-Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $x \leq y$ and $x \in U$ then $y \in U $.
+Given a quasi-ordering $(X,\leq)$, an {\em upward-closed set} is any set $U \subseteq X$ such that if $x \leq y$ and $x \in U$ then $y \in U $.
 A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
@@
 its {\em upward-closure},
- $\uparrow A = \{x \in X \mid \exists a \in A ~ a \leq y\}$
+ $\uparrow A = \{x \in X \mid \exists a \in A ~ a \leq x\}$
  and its 
  {\em downward-closure},
- $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
+ $\downarrow A = \{x \in X \mid \exists a \in A ~ x \leq a\}$. 
 We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
@@
 %
-A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$.
+A {\em basis} of an upward-closed set $U$ is a set $U_b$ such that $U = \uparrow U_b$; similarly, a {\em basis} of a downward-closed set $D$ is a set $D_b$ such that $D = \downarrow D_b$.
 
@@
 Wqo admits many other equivalent formulations like : a qo $(X,\leq)$ is a wqo iff $(X,\leq)$ is well founded (i.e. there is no infinite strictly decreasing sequence
-of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements of $X$). See other equivalences in \cite{schmitz:cel-00727025}. 
+of elements of $X$) and $(X,\leq)$ contains no infinite antichains (an antichain is a subset of mutually incomparable elements of $X$).
+See other equivalences in \cite{DBLP:phd/hal/Halfon18,schmitz:cel-00727025}.
 %
@@
 infinite antichains
- enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = \downarrow (J_1 \cup J_2 \cup \ldots \cup J_n)$ See for example \cite{BFM-ic17}.
+ enjoy a similar \emph{finite decomposition} property than wqo: every downward-closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2, \ldots, J_n$ such that $D = J_1 \cup J_2 \cup \ldots \cup J_n$ See for example \cite{BFM-ic17}.
 %
@@
 
-Several families of formal models of processes give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
+Several families of formal models of processes  \cite{DBLP:journals/tcs/FinkelS01} give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
 compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets and data Petri nets), reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
@@
 When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\uparrow U_k)_k$ instead.
+
+Recall the \emph{coverability problem} for ordered transition systems.
+
+\problemx{Coverability problem}
+{An ordered transition system $\mathscr{S}=(S,\rightarrow,\longrightarrow)$ and two states $s_0,s \in S$.}
+{ $s_0 \in \pred^*(\uparrow s )$ ? \newline}
+
+%An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
 With the effective pred-basis hypothesis, we obtain:
@@
 Subsetquently, $Q \times \N^d$ is the set of states of the transition system associated with a $d$-VASS $V$.
-For all states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$
+For all states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and for every control-transition $t = (p, \textbf{z}, q)$, we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$.
 %
@@
 
-A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence it can be only defined by $T$.\\
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence one only needs $T$.\\
 
@@
   (resp.  $\forall x \in S, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
-  In our framework, $\Safe \subseteq S$  is possibly infinite but  must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.
+  In our framework, $\Safe \subseteq S$  is possibly infinite but  must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqos and semilinear sets in $\mathbb{N}^d$ have finite representations.
 
@@
 Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
-. Since $\Safe \subseteq \pred^*(\Safe)$, one only need to decide whether 
+. Since $\Safe \subseteq \pred^*(\Safe)$, one only needs to decide whether 
 the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.

### 2023-10-17 VMCAI-revised-version/final.tex
Retiré la notion d'invariant et donné juste les détails de la procédure et une source quand à la faisabilité de l'énumération

@@
 \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+\mathieu{non, sinon on aurait décidabilité pour la state résilience}
 %This stems from the fact that it is undecidable for Minsky machines with more than one counter,
@@
 Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61,Min67}. 
+from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
@@
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
+The simulation operates by building a reset-VASS such that
+ an execution from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS simulates
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ a state of the form $(s',0,0,0)$, and
+ correspondingly,
+  if there is a sequence from 
+  $(s_0,0,0,0)$ to 
+  $(s_f,0,0,0)$ in the Minsky machine
+  there is a sequence in the reset-VASS which simulates
+  the sequence and ends
+  in a state of the form $(s,0, \ldots O)$
+  (from Theorem~$5$ in \cite{araki1976PN}).
+Hence $\downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+resilience in the corresponding reset-VASS.
 Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, $\post^*$-basis $\omega^2$-WSTS with strong compatibility as well.\qed
@@
 one looking for a resilience certificate and one looking for a non-resilience certificate.
-Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward-closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
-Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
+
+Procedure 1 enumerates every downward closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
+The computability of the enumeration comes from the ideally-effective hypothesis, as seen in
+Lemma~4.3. from \cite{DBLP:journals/lmcs/BlondinFM17}.
+The procedure then check for every downward closed subset $D_i$ whether $\downarrow \post^*(D_i) \subseteq D_i$. 
+Every subset $D_i$ such that $\downarrow \post^*(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
 Notice that, by 
-upward compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.
-
-Procedure 1 stops when it finds an invariant $D$ such that
+upward compatibility, $\downarrow \post^*(s)$ is such a subset and may eventually be found.
+
+Procedure 1 stops when it finds a subset $D$ such that
+$\downarrow \post^*(D) \subseteq D$,
+$s \in D$ and
 $D   \subseteq \pred^*(\Safe)$. 

### 2023-10-17 VMCAI-revised-version/final.tex
up

@@
 The simulation operates by building a reset-VASS such that
- an execution from 
+there exists
+ a sequence from 
 an initial state %$(s,x_1 \ldots x_n)$
 to
-a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS simulates
+a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS if and only if
+there exists
 a sequence
@@
  to
- a state of the form $(s',0,0,0)$, and
- correspondingly,
-  if there is a sequence from 
-  $(s_0,0,0,0)$ to 
-  $(s_f,0,0,0)$ in the Minsky machine
-  there is a sequence in the reset-VASS which simulates
-  the sequence and ends
-  in a state of the form $(s,0, \ldots O)$
+ a state of the form $(s',0,0,0)$
   (from Theorem~$5$ in \cite{araki1976PN}).

### 2023-10-17 VMCAI-revised-version/final.tex
up

@@
 $\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$.
-Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines, integer VASS and continuous VASS); most of the resilience properties are shown decidable.
+Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS); most of the resilience properties are shown decidable.
 \end{abstract}
@@
 
-\alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
-Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
-We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
-For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
+% \alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
+% Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+% We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
+% For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
 When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
@@
 \problemx{Coverability problem}
-{An ordered transition system $\mathscr{S}=(S,\rightarrow,\longrightarrow)$ and two states $s_0,s \in S$.}
+{An ordered transition system $\mathscr{S}=(S,\rightarrow, \leq)$ and two states $s_0,s \in S$.}
 { $s_0 \in \pred^*(\uparrow s )$ ? \newline}
@@
 Let us recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
-The class of WBTS is strictly larger than WSTS: for example, integer VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
+The class of WBTS is strictly larger than WSTS: for example, $\mathbb{Z}-$VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
 With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS (without the effective pred-basis hypothesis).
@@
 
-\alain{ce qui suit est à mettre ici ?}
-{\bf Related problems.} 
-Resilience and the home-space problem are also linked to the 
-model-checking of basic reachability and safety formulae. 
-In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-is decidable for Lossy Counter Machine (LCM)
-when $X$ and $Y$ are semi-linear sets.
-Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
- t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
-whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
-``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
-  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
-
 These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems.
@@
 \end{proposition}
-\alain{peut-on ajouter des hypothèses d'effectivity ?
+% \alain{peut-on ajouter des hypothèses d'effectivity ?
 peut-on montrer que l'exemple n'est pas pred effectif ?
@@
 \alain{preuve pas claire, à réécrire}
-\alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
-\mathieu{non, sinon on aurait décidabilité pour la state résilience}
+% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
 %This stems from the fact that it is undecidable for Minsky machines with more than one counter,
@@
 %	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
-From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $(0,0,0)$ from any input in $\uparrow(0,0,1)$ 
+From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \uparrow (0,0,1)$. 
 %			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches a control-state indicative of termination, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $\uparrow(0,0,1)$, it is not possible to reach $(1,0,0)$, $(0,1,0)$ or $(1,1,0)$.
-Based on this construction, the problem of deciding whether the downward closed set $\downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-from any input in $\uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
+Remark that in the construction, from any input in $Q_{M'} \times \uparrow(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
+Based on this construction, the problem of deciding whether the downward closed set $Q_{M'} \times \downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+from any input in $Q_{M'} \times \uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
@@
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
-\alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
+% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
 %Reset-VASS extend the basic VASS model with special “reset
@@
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS such that
+The simulation operates by building a reset-VASS in which
 there exists
@@
 to
-a state of the form $ (s, 0,  \ldots 0)$  in the reset-VASS if and only if
+a state of the form $ q(0,  \ldots 0)$ if and only if
 there exists
@@
  to
- a state of the form $(s',0,0,0)$
+ a state of the form $q_f(0,0,0)$
   (from Theorem~$5$ in \cite{araki1976PN}).
-Hence $\downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
-resilience in the corresponding reset-VASS.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, $\post^*$-basis $\omega^2$-WSTS with strong compatibility as well.\qed
+Hence $Q_{M'} \times \downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+resilience hold in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, 
+$\omega^2$-WSTS with strong compatibility as well.\qed
 %\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities,
-$\Safe = \downarrow \Safe$.
+{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities, effective pred-basis,
+and $\Safe = \downarrow \Safe$.
 \end{theorem}
@@
 \begin{proof}
-In order to decide whether $\post^{\leq n}(s) 
+In order to decide whether %$\post^{\leq n}(s)
+$\post^{ *}(s)  
  \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
@@
 until it finds an element
-not in $\pred^*(\Safe)$. \qed
+not in $\pred^*(\Safe)$.
+
+If resilience hold, then procedure 1 terminates since it eventually finds
+ $D = \downarrow \post^*(s)$ 
+ such that  $D \subseteq  \pred^*(\Safe)$.
+ If resilience does not hold, then procedure 2 terminates since
+there exists a witness that resilience does not hold which can eventually be found. \qed
 \end{proof}
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with effective pred-basis, strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.
@@
 
-Integer VASS or $\mathbb{Z}-$VASS \cite{DBLP:conf/rp/HaaseH14} are VASS that are allowed to take values from the integers.
+%Integer VASS or 
+$\mathbb{Z}^d-$VASS (resp. $\mathbb{Z}-$VASS)~\cite{DBLP:conf/rp/HaaseH14} are $d$-VASS (resp. VASS) that are allowed to take values from the integers.
 
 \begin{theorem}\label{SL VASS}
-{\sc Resilience} is decidable for VASS, and integer VASS, when $\Safe$  is a semilinear set.
+{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
 \end{theorem}
@@
 and hence, {\sc Resilience} is decidable. 
-The reachability relation of integer $d$-VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+The reachability relation of 
+%integer $d$-VASS
+ $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
 \qed
@@
 {\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
-Recall that lossy counter machines \cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
+Recall that lossy counter machines (LCM)~\cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
+
+Resilience and the home-space problem are also linked to the 
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+is decidable for lossy counter machine
+when $X$ and $Y$ are semi-linear sets.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
 

### 2023-10-17 VMCAI-revised-version/final.tex
effectivity

@@
 
-We recall and introduce five types of \emph{effectiveness} that will be essential for solving various resilience problems in different models of WSTS.
-\noindent
-%			{\bf On effectivity.}
-We say that a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
+To obtain decidability results, we must introduce some notions of \emph{effectiveness}.
+We use a mixture of the notions defined both by Blondin and al.
+in \cite{DBLP:journals/lmcs/BlondinFM17}) and by Halfon in \cite{DBLP:phd/hal/Halfon18}.
+%			 that will be essential for solving various resilience problems in different models of WSTS. 
+First, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ are \emph{effective}, i.e. that they satisfy the five following properties:
 \begin{itemize}
-\item {\em effective} if there exists a pair of algorithms
-($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
-$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
-$M_\leq$ the ordering relation “$\leq$”.
+\item there is a computational representation for $S$ for which membership in $S$ is decidable,
+\item there exists an algorithm that decides whether $s \rightarrow t$ is true or not, for any states $s,t \in S$,
+%
+%			We also suppose that all considered wqos $(S,\leq)$ satisfy the three following conditions : 
+\item $\leq$ is decidable,
+\item there is a computational representation for $Ideals(S)$ for which membership in $Ideals(S)$ is decidable, and
+\item  inclusion of ideals is decidable. 
+\end{itemize}
+%
+Blondin and al.
+showed (Lemma 4.3 in \cite{DBLP:journals/lmcs/BlondinFM17}) that under the previous hypotheses, one are able to enumerate \emph{downward closed sets} (by their finite decomposition in ideals), to decide inclusion between downward closed sets and to decide if a state belongs to a given finite set of ideals.
+But the five properties are not sufficient to compute the complementary and the intersection of downward closed sets.
+%
+Halfon introduced \emph{ideally effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of intersections of two filters and intersections of ideals.
+%	 (hence of downward closed sets).
+From now, we suppose that wqo are ideally effective and this the case of most known wqo \cite{DBLP:phd/hal/Halfon18}.
+%
+To build algorithms, we also need some effective hypotheses on the computations of some sets of predecessors and successors. More precisely, we say that a WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
+\begin{itemize}
+%\item {\em effective} if there exists a pair of algorithms
+%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
+%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
+%$M_\leq$ the ordering relation “$\leq$”.
 %
 \item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
-exists an algorithm that computes $|\post(x)| \in \N \cup \{\infty\}$ 
-on input $x \in X$.
-%
-\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
-to the encoding of the ideal $\downarrow s$ is computable, inclusion of ideals is decidable and $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
-%
+exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
+on input $s \in S$.
+%
+%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
+%to the encoding of the ideal $\downarrow s$ is computable.
+%
+\item {\em $\post$-ideal-effective} if there exists an algorithm accepting
+any ideal $I \in Ideals(S)$ and returning $\downarrow \post(I)$, expressed as a finite union of (maximal) ideals.
 %\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
 %
-\item {\em effective $\pred$-basis} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
-any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
+\item {\em $\pred$-basis effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+any state $s \in S$ and returning a finite basis of $\uparrow \pred(\uparrow s)$.
 %
@@
 %
-\item {\em effective $\post^*$-basis} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if there exists an algorithm accepting
+\item {\em $\post^*$-basis effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if there exists an algorithm accepting
 any state $s \in S$ and returning a finite basis of $\uparrow \post^*(s)$.

### 2023-10-17 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{definition}(Finkel \cite{DBLP:journals/iandc/Finkel90})
-A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
+A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
 is a transition system $(S, \rightarrow)$
@@
 %%%%
+
+But there is a more general class of ordered transition systems than WSTS for which coverability is still decidable: recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
+The class of WBTS is strictly larger than WSTS: for example, $\mathbb{Z}^d-$VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
+\\
 
@@
 %			 that will be essential for solving various resilience problems in different models of WSTS. 
-First, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ are \emph{effective}, i.e. that they satisfy the five following properties:
+First, to simplify, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ are \emph{effective}, i.e.
+that they satisfy the five following properties:
 \begin{itemize}
@@
 %
-Halfon introduced \emph{ideally effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of intersections of two filters and intersections of ideals.
+Halfon introduced \emph{ideally effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
 %	 (hence of downward closed sets).
-From now, we suppose that wqo are ideally effective and this the case of most known wqo \cite{DBLP:phd/hal/Halfon18}.
-%
-To build algorithms, we also need some effective hypotheses on the computations of some sets of predecessors and successors. More precisely, we say that a WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
+
+%
+Starting now, we assume that all wqo are ideally effective, which is the case for most well-known wqo \cite{DBLP:phd/hal/Halfon18}.
+In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors.
+More precisely, we say that a WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ is
 \begin{itemize}
@@
 %
-\item {\em $\post$-ideal-effective} if there exists an algorithm accepting
+\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
 any ideal $I \in Ideals(S)$ and returning $\downarrow \post(I)$, expressed as a finite union of (maximal) ideals.
@@
 %
-\item {\em $\pred$-basis effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if there exists an algorithm accepting
+\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
 any state $s \in S$ and returning a finite basis of $\uparrow \pred(\uparrow s)$.
@@
 %
-\item {\em $\post^*$-basis effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if there exists an algorithm accepting
+\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
 any state $s \in S$ and returning a finite basis of $\uparrow \post^*(s)$.
@@
 %An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
-With the effective pred-basis hypothesis, we obtain:
+With the effective pred-basis effective hypothesis, we obtain:
 
@@
 
-Let us recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
-The class of WBTS is strictly larger than WSTS: for example, $\mathbb{Z}-$VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
-With the ideally-effective hypothesis, we obtain the decidability of coverability for WBTS (without the effective pred-basis hypothesis).
+With the post-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis effective hypothesis).
 
 \begin{theorem}(Blondin and al. ~\cite{DBLP:journals/lmcs/BlondinFM17})
-Coverability is decidable for any ideally-effective WBTS.
+Coverability is decidable for any post-ideal-effective WBTS.
 \end{theorem}
@@
 \begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
-Effective finitely branching downward-WSTS are effective $\post^*$-basis.
+Effective finitely branching downward-WSTS are $\post^*$-basis-effective.
 \end{theorem}
@@
 \end{itemize} \end{samepage}
+
+As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo and reset-VASS are $\post$-effective, $\post$-ideal-effective and $\pred$-basis-effective but they are not $\post^*$-basis-effective (VASS are $\post^*$-basis-effective).
 
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not ideally-effective.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not post-ideal effective.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be an ideally-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a post-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 \end{theorem}
+
 \alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
+
 \begin{proof}
@@
 This implies that resilience and bounded resilience are equivalent to coverability.
-Since coverability is decidable for ideally-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+Since coverability is decidable for post-effective  WSTS \cite{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.
@@
 
+\alain{maintenant on a toutes les possibilités de manipuler les ideaux, on doit pouvoir enlever strong compatibility}
+\begin{theorem}\label{xxx}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a pred-basis-effective WSTS with strong compatibility ???? and a set $\Safe = \uparrow \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+\begin{proof}
+on calcule $pre^*(Safe)$ qui est clos par le haut et on teste si $Bad = \downarrow Bad$  est inclus dans $pre^*(Safe)$, ce qui revient à tester si $Bad \cap (S-pre^*(Safe))=\emptyset$ ce qui revient à calculer l'intersection de 2 clos par le bas ce qui est possible (ideally effective wqo).
+
+\end{proof} \qed
 %				The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for ideal-effective downward-WSTS with 
+{\sc Resilience}  is decidable for post-ideal-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
@@
 $\Bad \subseteq  \pred^*(\Safe)$.
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of upward- or downward-closed 
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is post-ideal-effective, we can compute intersections of upward- or downward-closed 
  subsets.
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for ideally-effective and effective pred-basis $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+{\sc Resilience} for pred-basis-effective ???? $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
 \end{proposition}
@@
 resilience hold in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  ideally-effective, effective pred-basis, 
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  pred-basis-effective, 
 $\omega^2$-WSTS with strong compatibility as well.\qed
@@
 More concretely,
-VASS for instance are $\uparrow \post^*$-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
+VASS for instance are $ \post^*$-basis-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
 It is well-known that 
@@
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for ideally-effective and effective pred-basis $\omega^2$-WSTS with strong compatibility
+{\sc State-resilience} is undecidable for post-ideal-effective ???? and effective pred-basis $\omega^2$-WSTS with strong compatibility
 when
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for ideally-effective WSTS with downward and upward compatibilities, effective pred-basis,
+{\sc State-resilience} is decidable for post-ideal-effective WSTS with downward and upward compatibilities, effective pred-basis,
 and $\Safe = \downarrow \Safe$.
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideally-effective WSTS with effective pred-basis, strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for post-ideal-effective ???? WSTS with effective pred-basis, strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.

### 2023-10-18 VMCAI-revised-version/final.tex
effectivity

@@
 turn a system that is not bounded resilient into one that is. 
+\alain{à quoi sert cette remarque ?}
 
@@
 
+An (upward) compatible transition relation $ \rightarrow$ is \emph{reflexive} when the sequence $s_2 \rightarrow^{*} t_2$ is not empty: formally,
+for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+	$t_2 \in S$ with 
+	$t_1 \leq t_2$ and $s_2 \rightarrow^{+} t_2$.
+%
 We say that a WSTS $\mathscr{S}$ has \emph{strong (upward) compatibility} when moreover for all 
@@
 Several families of formal models of processes  \cite{DBLP:journals/tcs/FinkelS01} give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
-compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets and data Petri nets), reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
+compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets) data Petri nets, reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
 Similarly, lossy channel systems with $d$ channels and $Q$ a finite set of control-states are WSTS (with a non-strong compatibility)
@@
 
-But there is a more general class of ordered transition systems than WSTS for which coverability is still decidable: recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
+But there is a more general class of (upward) compatible ordered transition systems than WSTS for which coverability is still decidable: recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
 The class of WBTS is strictly larger than WSTS: for example, $\mathbb{Z}^d-$VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.
@@
 
-By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that possesses an excellent characteristic that allows to obtain more algorithms:
+By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that possesses a characteristic that allows to obtain more algorithms:
 
@@
 
-Although there exists wqo that are not $\omega^2$-wqo (see the Rado ordering in \cite{DBLP:journals/ipl/Jancar99}), all naturally occurring wqos are $\omega^2$-wqo, perhaps to the notable exception of ﬁnite graphs quasi-ordered by the graph minor relation, which are wqo but not known to be $\omega^2$-wqo.
-The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is a an $\omega^2$-wqo (Proposition 4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
+Although there exists wqo that are not $\omega^2$-wqo (see the Rado ordering in \cite{DBLP:journals/ipl/Jancar99}), all naturally occurring wqos are $\omega^2$-wqo, perhaps to the notable exception of ﬁnite graphs well-quasi-ordered by the graph minor relation.
+%			which are wqo but not known to be $\omega^2$-wqo.
+The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is an $\omega^2$-wqo (Proposition 4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
 Now we are able to define $\omega^2$-WSTS:
@@
 %			 that will be essential for solving various resilience problems in different models of WSTS. 
-First, to simplify, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ are \emph{effective}, i.e.
-that they satisfy the five following properties:
-\begin{itemize}
+First, to simplify, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five following properties:
+\begin{enumerate}
+\item there exists an algorithm that decides whether $s \rightarrow t$ is true or not, for any states $s,t \in S$,
+%
 \item there is a computational representation for $S$ for which membership in $S$ is decidable,
-\item there exists an algorithm that decides whether $s \rightarrow t$ is true or not, for any states $s,t \in S$,
 %
@@
 \item  inclusion of ideals is decidable. 
-\end{itemize}
+\end{enumerate}
 %
@@
 Halfon introduced \emph{ideally effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
+Most well-known wqo are ideally effective .
+\cite{DBLP:phd/hal/Halfon18}.
 %	 (hence of downward closed sets).
@@
 %
-Starting now, we assume that all wqo are ideally effective, which is the case for most well-known wqo \cite{DBLP:phd/hal/Halfon18}.
+Starting now, we assume that all considered WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five previous properties and that the wqo $\leq$ is ideally effective.
+We refer to such WBTS as \emph{effective}.
+
 In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors.
@@
 %An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
-With the effective pred-basis effective hypothesis, we obtain:
+With the pred-basis-effective hypothesis, we obtain:
 
 \begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
- A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+ A finite basis of $ \pred^*(U)$ is computable for any pred-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
 Hence coverability is decidable.
@@
 
-With the post-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis effective hypothesis).
+With the post-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis-effective hypothesis).
 
@@
 \begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
-Effective finitely branching downward-WSTS are $\post^*$-basis-effective.
+Post-effective finitely branching downward-WSTS with reflexive compatibility are $\post^*$-basis-effective.
 \end{theorem}
@@
 
-VASS can extended with resets.
+VASS can be extended with resets.
 
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not post-ideal effective.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not post-ideal-effective.
 
@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the completion-post-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the post-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a post-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a post-ideal-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 
-\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
+%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
 
@@
 
-This implies that resilience and bounded resilience are equivalent to coverability.
+This implies that resilience and bounded resilience can be reduced to coverability.
 Since coverability is decidable for post-effective WSTS \cite{BFM-ic17}, we deduce that both the 
@@
 
-\alain{maintenant on a toutes les possibilités de manipuler les ideaux, on doit pouvoir enlever strong compatibility}
+\alain{maintenant on doit pouvoir enlever strong compatibility}
 \begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a pred-basis-effective WSTS with strong compatibility ???? and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a pred-basis-effective WSTS and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 \begin{proof}
-on calcule $pre^*(Safe)$ qui est clos par le haut et on teste si $Bad = \downarrow Bad$  est inclus dans $pre^*(Safe)$, ce qui revient à tester si $Bad \cap (S-pre^*(Safe))=\emptyset$ ce qui revient à calculer l'intersection de 2 clos par le bas ce qui est possible (ideally effective wqo).
-
+The resilience problem can be reformulated as 
+$\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is pred-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
+Since the wqo is ideally effective, we can compute the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable.
+%%%%%%%
+\alain{faire les deux autres résilience}
 \end{proof} \qed
@@
 The resilience problem can be reformulated as 
-$\Bad \subseteq  \pred^*(\Safe)$.
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is post-ideal-effective, we can compute intersections of upward- or downward-closed 
+$\Bad \subseteq  \pred^*(\Safe)$ and also as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
+Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward (and downward-closed) 
  subsets.
@@
 $S \setminus \pred^*(\Safe)$,
-which are both upward-closed.
-Since
-$\Bad \subseteq \pred^*(\Safe)$
-can be reformulated as
-$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
+which are both upward-closed. Hence
 the resilience problem is decidable. \qed
@@
 In the case
-of a ideal-effective WSTS, not necesarly downward compatible,
+of a post-ideal-effective WSTS, not necesarly downward compatible,
 the above construction
@@
 
-In the case of (upward compatible) WSTS with ideally-effectivity, {\sc Resilience} is undecidable.
+In the case of (upward compatible) WSTS with pred-basis-effectivity, {\sc Resilience} is undecidable.
 
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for pred-basis-effective ???? $\omega^2$-WSTS with strong compatibility and for downward closed set $\Safe$ is undecidable.
+{\sc Resilience} for pred-basis-effective ???? $\omega^2$-WSTS with strong compatibility and with 
+$\Safe=\downarrow \Safe$ is undecidable.
 \end{proposition}
@@
 (\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
-{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for effective $\post^*$-basis WSTS with strong compatibility
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\post^*$-basis-effective WSTS with strong compatibility
 when
@@
 \begin{proposition}\label{postcomputable}
-{\sc State-resilience} is decidable for effective $\post^*$-basis WSTS with strong compatibility
+{\sc State-resilience} is decidable for $\post^*$-basis-effective WSTS with strong compatibility
 when
@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For effective finite-branching WSTS, a basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \uparrow \post^{\leq n}(s)$ until it becomes stationnary. 
+$U_n = \uparrow \post^{\leq n}(s)$ \alain{il manque l'hypothèse post-effective}until it becomes stationnary. 
 Decidability of downward-reachability leads to the decidability of the following stop condition:
@@
 
-More concretely,
-VASS for instance are $ \post^*$-basis-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
+For instance,
+VASS are $ \post^*$-basis-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
 It is well-known that 
@@
 \begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for effective $\post^*$-basis 
- WSTS
+{\sc State-resilience} is decidable for $\post^*$-basis-effective WSTS
 when
@@
  \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and we can compute a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} 
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute \alain{comment calcule-t-on une telle base sans l'hypothèse pred-basis-effective qui n'est pas dans les hypothèses du théorème ?
+Est-elle chez Ozkan ?} a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)}
  \setminus \pred^*(\Safe) = \emptyset$.  \qed
@@
 
-Thus in case of effective $\post^*$-basis WSTS (when the compatibility is not strong), we don't know the decidability status of {\sc $k$-state-resilience} and 
+Thus in case of $\post^*$-basis-effective WSTS (when the compatibility is not strong), we don't know the decidability status of {\sc $k$-state-resilience} and 
 {\sc Bounded-state-resilience}.
@@
 As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-Section~\ref{safe-down}.
+Section~\ref{safe-down}.\alain{trop rapide}
 
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for post-ideal-effective ???? and effective pred-basis $\omega^2$-WSTS with strong compatibility
+{\sc State-resilience} is undecidable for post-ideal-effective ???? and pred-basis-effective $\omega^2$-WSTS with strong compatibility
 when
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for post-ideal-effective WSTS with downward and upward compatibilities, effective pred-basis,
+{\sc State-resilience} is decidable for post-effective, post-ideal-effective and effective pred-basis WSTS with downward and upward compatibilities
 and $\Safe = \downarrow \Safe$.
@@
 Procedure 1 enumerates every downward closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
-The computability of the enumeration comes from the ideally-effective hypothesis, as seen in
-Lemma~4.3. from \cite{DBLP:journals/lmcs/BlondinFM17}.
+The computability of the enumeration comes from the hypothesis that $\leq$ is ideally-effective (Lemma~4.3.
+from \cite{DBLP:journals/lmcs/BlondinFM17}).
 The procedure then check for every downward closed subset $D_i$ whether $\downarrow \post^*(D_i) \subseteq D_i$. 
+\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
+L'hypothèse "post-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
+Il manquait la post-effective dans les hypothèses}
 Every subset $D_i$ such that $\downarrow \post^*(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
@@
 The second procedure iteratively computes
-$\post^{\leq n}(s) $
+$\post^{\leq n}(s) $ \alain{ici on utilise post-effective que j'ai rajoutée}
 until it finds an element
@@
 of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.
+\alain{trop rapide}
 
@@
 We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general.
-We also exhibited classes of WSTS with decidable resilience, namely ideally-effective WSTS with strong compatibility (in the case
+We also exhibited classes of WSTS with decidable resilience, namely post-ideal-effective WSTS with strong compatibility (in the case
  $\Safe = \uparrow \Safe$) and 
 downward-compatible
-ideally-effective WSTS
+effective WSTS \alain{vérifier les effectivités}
 (in the case 

### 2023-10-18 VMCAI-revised-version/final.tex
up

@@
 
-Hence we can see how adding transitions can 
-turn a system that is not bounded resilient into one that is. 
-\alain{à quoi sert cette remarque ?}
+% Hence we can see how adding transitions can 
+% turn a system that is not bounded resilient into one that is. 
+% \alain{à quoi sert cette remarque ?}
 
@@
 \alain{ici un dessin comme celui que j'ai envoyé sur skype...à rajouter les downward WSTS}
+
+ \begin{center}
+	\begin{figure}
+	\hspace{1.5cm}
+\includegraphics[width=0.70\textwidth]{WSTS_taxonomy}
+	\caption{A taxonomy of WSTS and variants.}
+	\end{figure}
+\end{center}
 
@@
 
-Continuous VASS \cite{DBLP:journals/tocl/BlondinFHH17} are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+Continuous VASS~\cite{DBLP:journals/tocl/BlondinFHH17}, or $\mathds{Q}^d$-VASS are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
 

### 2023-10-18 VMCAI-revised-version/final.tex
up

@@
 
-Continuous VASS~\cite{DBLP:journals/tocl/BlondinFHH17}, or $\mathds{Q}^d$-VASS are a relaxation of classical discrete VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical discrete $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
 

### 2023-10-18 VMCAI-revised-version/final.tex
post-effective replaced with the macro i.e. $\posthimBHseffective

@@
 We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis).
-Then we prove the decidability of resilience for completion-post-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$.
+Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$.
 Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
@@
 {$k$-state-resilience}
-are undecidable for strongly upward-compatible WSTS with effective pred-basis
+are undecidable for strongly upward-compatible $\pred$-basis-effective WSTS
 when
@@
 
-%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-post-effective} if
-%it is post-effective, and there exists an algorithm $M_\downarrow$ that computes 
+%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-$\post$-effective} if
+%it is $\post$-effective, and there exists an algorithm $M_\downarrow$ that computes 
 % $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
@@
 any ideal $I \in Ideals(S)$ and returning $\downarrow \post(I)$, expressed as a finite union of (maximal) ideals.
-%\item {\em completion-post-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is post-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
+%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
 %
@@
 
-%Most of usual WSTS are post-effective and ideally-effective.
+%Most of usual WSTS are $\post$-effective and ideally-effective.
 
@@
 
-With the post-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis-effective hypothesis).
+With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis-effective hypothesis).
 
 \begin{theorem}(Blondin and al. ~\cite{DBLP:journals/lmcs/BlondinFM17})
-Coverability is decidable for any post-ideal-effective WBTS.
+Coverability is decidable for any $\post$-ideal-effective WBTS.
 \end{theorem}
@@
 \begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
-Post-effective finitely branching downward-WSTS with reflexive compatibility are $\post^*$-basis-effective.
+$\post$-effective finitely branching downward-WSTS with reflexive compatibility are $\post^*$-basis-effective.
 \end{theorem}
@@
 \begin{proposition}\label{indec WSTS}
-Resilience for post-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
+Resilience for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
 \end{proposition}
@@
 l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais pas 3 car ça permettrait de décider l'arrêt ?)?
-on voit que l'exemple n'est pas completion-post-effective vu le theoreme qui suit}
+on voit que l'exemple n'est pas completion-$\post$-effective vu le theoreme qui suit}
 \begin{proof}
@@
 	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-	is a post-effective $\omega^2$-WSTS with strong compatibility
+	is a $\post$-effective $\omega^2$-WSTS with strong compatibility
 	and
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not post-ideal-effective.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective.
 
@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the post-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a post-ideal-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 This implies that resilience and bounded resilience can be reduced to coverability.
-Since coverability is decidable for post-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.
@@
 \end{proof} \qed
-%				The completion-post-effective hypothesis is needed to decide {\sc Resilience}.
+%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
 %%%
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for post-ideal-effective downward-WSTS with 
+{\sc Resilience}  is decidable for $\post$-ideal-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
@@
 In the case
-of a post-ideal-effective WSTS, not necesarly downward compatible,
+of a $\post$-ideal-effective WSTS, not necesarly downward compatible,
 the above construction
@@
 $\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et post-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
 \end{proof}
@@
 {\sc $k$-state-resilience}
-are undecidable for strongly compatible WSTS with effective pred-basis
+are undecidable for strongly compatible $\pred$-basis-effective WSTS
 when
@@
 On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+%
 \begin{theorem}\label{ref ozkan}
 (\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
-{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\post^*$-basis-effective WSTS with strong compatibility
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
 when
@@
 \begin{proposition}\label{postcomputable}
-{\sc State-resilience} is decidable for $\post^*$-basis-effective WSTS with strong compatibility
+{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
 when
@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For $\pred$-basis-effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \uparrow \post^{\leq n}(s)$ \alain{il manque l'hypothèse post-effective}until it becomes stationnary. 
+$U_n = \uparrow \post^{\leq n}(s)$ \alain{il manque l'hypothèse $\post$-effective}until it becomes stationnary. 
 Decidability of downward-reachability leads to the decidability of the following stop condition:
@@
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for post-ideal-effective ???? and pred-basis-effective $\omega^2$-WSTS with strong compatibility
+{\sc State-resilience} is undecidable for $\post$-ideal-effective ???? and pred-basis-effective $\omega^2$-WSTS with strong compatibility
 when
@@
 %
-For instance one can write the following lemma:
+% For instance one can write the following lemma:
+%  Page 14: "For instance one can write the following lemma" => Rewrite this: we will need the following lemma for this and that.
+We will need the following lemma in order to obtain symmetrical results.
 
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for post-effective, post-ideal-effective and effective pred-basis WSTS with downward and upward compatibilities
+{\sc State-resilience} is decidable for $\post$-effective, $\post$-ideal-effective and effective pred-basis WSTS with downward and upward compatibilities
 and $\Safe = \downarrow \Safe$.
@@
 \alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
-L'hypothèse "post-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
-Il manquait la post-effective dans les hypothèses}
+L'hypothèse "$\post$-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
+Il manquait la $\post$-effective dans les hypothèses}
 Every subset $D_i$ such that $\downarrow \post^*(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
@@
 The second procedure iteratively computes
-$\post^{\leq n}(s) $ \alain{ici on utilise post-effective que j'ai rajoutée}
+$\post^{\leq n}(s) $ \alain{ici on utilise $\post$-effective que j'ai rajoutée}
 until it finds an element
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for post-ideal-effective ???? WSTS with effective pred-basis, strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\pred$-basis-effective, $\post$-ideal-effective ???? WSTS, strong downward compatibility and upward compatibility when
 $\Safe = \downarrow \Safe$.
@@
 
-In this section we study VASS. Since they are ideally-effective and completion-post-effective WSTS, they inherit the decidability results for WSTS in the case 
+In this section we study VASS. Since they are ideally-effective and completion-$\post$-effective WSTS, they inherit the decidability results for WSTS in the case 
 $\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
@@
 We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general.
-We also exhibited classes of WSTS with decidable resilience, namely post-ideal-effective WSTS with strong compatibility (in the case
+We also exhibited classes of WSTS with decidable resilience, namely $\post$-ideal-effective WSTS with strong compatibility (in the case
  $\Safe = \uparrow \Safe$) and 

### 2023-10-18 VMCAI-revised-version/final.tex
pred-basis-effective replaced with the macro i.e. $\predhimBHsbasis-effective for coherency

@@
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
-We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective pred-basis).
+We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective $\pred$-basis).
 Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$.
@@
 %An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
-With the pred-basis-effective hypothesis, we obtain:
+With the $\pred$-basis-effective hypothesis, we obtain:
 
 \begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
- A finite basis of $ \pred^*(U)$ is computable for any pred-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+ A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
 Hence coverability is decidable.
@@
 
-With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the pred-basis-effective hypothesis).
+With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the $\pred$-basis-effective hypothesis).
 
@@
 %
-%Coverability is  decidable for a class of WSTS without the effective pred-basis hypothesis.
+%Coverability is  decidable for a class of WSTS without the effective $\pred$-basis hypothesis.
 %
@@
 \begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a pred-basis-effective WSTS and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 $\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is pred-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
+The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
 Since the wqo is ideally effective, we can compute the intersection of
@@
 
-In the case of (upward compatible) WSTS with pred-basis-effectivity, {\sc Resilience} is undecidable.
+In the case of (upward compatible) WSTS with $\pred$-basis-effectivity, {\sc Resilience} is undecidable.
 
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for pred-basis-effective ???? $\omega^2$-WSTS with strong compatibility and with 
+{\sc Resilience} for $\pred$-basis-effective ???? $\omega^2$-WSTS with strong compatibility and with 
 $\Safe=\downarrow \Safe$ is undecidable.
@@
 resilience hold in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  pred-basis-effective, 
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
 $\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective pred-basis et ideally effective....}
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally effective....}
 \end{proof}
@@
 and reducible to {\sc State-resilience}.
-Additionally, in WSTS with strong compatibility and effective pred-basis, {\sc Bounded-state-resilience} is
+Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
 reducible to {\sc $k$-state-resilience}:
@@
 hold for $k$ from $0$ to $n_0$.  
-Furthermore, in WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
 since 
@@
  \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and we can compute \alain{comment calcule-t-on une telle base sans l'hypothèse pred-basis-effective qui n'est pas dans les hypothèses du théorème ?
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute \alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ?
 Est-elle chez Ozkan ?} a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)}
@@
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for $\post$-ideal-effective ???? and pred-basis-effective $\omega^2$-WSTS with strong compatibility
+{\sc State-resilience} is undecidable for $\post$-ideal-effective ???? and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
 when
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for $\post$-effective, $\post$-ideal-effective and effective pred-basis WSTS with downward and upward compatibilities
+{\sc State-resilience} is decidable for $\post$-effective, $\post$-ideal-effective and effective $\pred$-basis WSTS with downward and upward compatibilities
 and $\Safe = \downarrow \Safe$.

### 2023-10-18 VMCAI-revised-version/final.tex
ideally effective vs ideally-effective, kept only the one with '-'

@@
 %
-Halfon introduced \emph{ideally effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
-Most well-known wqo are ideally effective .
+Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
+Most well-known wqo are ideally-effective .
 \cite{DBLP:phd/hal/Halfon18}.
@@
 %
-Starting now, we assume that all considered WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five previous properties and that the wqo $\leq$ is ideally effective.
+Starting now, we assume that all considered WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five previous properties and that the wqo $\leq$ is ideally-effective.
 We refer to such WBTS as \emph{effective}.
@@
 The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
-Since the wqo is ideally effective, we can compute the intersection of
+Since the wqo is ideally-effective, we can compute the intersection of
 $\Bad$
@@
 $\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally effective....}
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
 \end{proof}
@@
 \begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for $\post^*$-basis-effective WSTS
+{\sc State-resilience} is decidable for $\pred$-effective, $\post^*$-basis-effective WSTS
 when

### 2023-10-18 VMCAI-revised-version/final.tex
up

@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For $\pred$-basis-effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \uparrow \post^{\leq n}(s)$ \alain{il manque l'hypothèse $\post$-effective}until it becomes stationnary. 
+$U_n = \uparrow \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
+until it becomes stationnary. 
 Decidability of downward-reachability leads to the decidability of the following stop condition:
@@
 \begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for $\pred$-effective, $\post^*$-basis-effective WSTS
+{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
 when

### 2023-10-18 VMCAI-revised-version/final.tex
up

@@
  \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and we can compute \alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ?
-Est-elle chez Ozkan ?} a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)}
+Since $B_{\uparrow \post^*(s)}$ is finite and we can compute 
+%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
+a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} 
  \setminus \pred^*(\Safe) = \emptyset$.  \qed

### 2023-10-18 VMCAI-revised-version/final.tex
Addressed the - Proof of Theorem 3: Why are the elements in the basis of Safe denoted with a b while there is also the set of bad states?- comment

@@
 
-Let  $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
-The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$.
+Let  $\{s_1,s_2,...,s_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
 Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
 
-Let us prove that $b_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow b_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
-%
-Suppose that $b_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq b_j$.
-From Proposition 30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow b_j$ in $\hat{\mathscr{S}}$, hence $\downarrow b_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow b_j \subseteq J$ then 
+Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow s_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+%
+Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
+From Proposition~30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow s_j$ in $\hat{\mathscr{S}}$, hence $\downarrow s_j$ is covered from $\downarrow x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow s_j \subseteq J$ then 
 %	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq b_j$ such that $x \xrightarrow{k} y'  \geq b_j$ in $\mathscr{S}$ and then $b_j$ is coverable from $x$ in $\mathscr{S}$.
-
-Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow b_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
-%
-Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
+there exists $x \in I$ and $y' \geq s_j$ such that $x \xrightarrow{k} y'  \geq s_j$ in $\mathscr{S}$ and then $s_j$ is coverable from $x$ in $\mathscr{S}$.
+
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
+%
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow s_j$ is not coverable from $J_i$.
 Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
@@
 Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
-Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow b_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow b_j$ is not coverable from $J_i$.
-If $\downarrow b_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow b_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow s_j$ is not coverable from $J_i$.
+If $\downarrow s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
 

### 2023-10-18 VMCAI-revised-version/final.tex
What to do with pred-basis-effectivity in case Safe is downward closed?

@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for $\post$-effective, $\post$-ideal-effective and effective $\pred$-basis WSTS with downward and upward compatibilities
+{\sc State-resilience} is decidable for $\post$-effective, $\pred$-basis-effective, $\post$-ideal-effective  WSTS with downward and upward compatibilities
 and $\Safe = \downarrow \Safe$.
@@
 from \cite{DBLP:journals/lmcs/BlondinFM17}).
-The procedure then check for every downward closed subset $D_i$ whether $\downarrow \post^*(D_i) \subseteq D_i$. 
+The procedure then checks for every downward closed subset $D_i$ whether $\downarrow \post^*(D_i) \subseteq D_i$. 
 \alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
@@
 $D   \subseteq \pred^*(\Safe)$. 
+\mathieu{ici $\pred^*(\Safe)$ est downward-closed.
+du coup pas sur que la $\pred$-basis-effectivité telle qu'elle est actuellement nous permette de trancher ...
+à moins que les hypothèses d'effectivté sur les idéaux nous permettent de guarantir qu'on puisse calculer le $\pred$ d'un idéal ?}
 Indeed
@@
 \alain{trop rapide}
+Indeed, with strong downward compatibility, 
+$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is,
+and becomes possible to represent \mathieu{... and compute ?} and compare
+$\pred^{\leq k}(\Safe)$ 
+to other downward closed subsets.
 

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-18 VMCAI-revised-version/final.tex
@@
 any state $s \in S$ and returning a finite basis of $\uparrow \post^*(s)$.
+%
+\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+any ideal $I \in Ideals(S)$ and returning a finite basis of $\downarrow \pred^*(I)$.
 \end{itemize}
@@
 As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo and reset-VASS are $\post$-effective, $\post$-ideal-effective and $\pred$-basis-effective but they are not $\post^*$-basis-effective (VASS are $\post^*$-basis-effective).
-
-\alain{ici un dessin comme celui que j'ai envoyé sur skype...à rajouter les downward WSTS}
 
@@
 
-\alain{maintenant on doit pouvoir enlever strong compatibility}
 \begin{theorem}\label{xxx}
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for $\post$-ideal-effective downward-WSTS with 
+{\sc Resilience}  is decidable for $\pred^*$-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
@@
 \begin{proof}
-By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
-The resilience problem can be reformulated as 
-$\Bad \subseteq  \pred^*(\Safe)$ and also as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
-Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward (and downward-closed) 
+Let $\mathscr{S}$ be a $\pred^*$-effective downward-WSTS with 
+$\Safe=\downarrow \Safe$.
+By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-effective.
+The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
+Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
  subsets.
-Hence we can compute the intersection of
+Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
 $\Bad$
@@
 $S \setminus \pred^*(\Safe)$,
-which are both upward-closed. Hence
+which are both upward-closed. We may decide whether this intersection is empty, hence
 the resilience problem is decidable. \qed
@@
 In the case
-of a $\post$-ideal-effective WSTS, not necesarly downward compatible,
+of a $\pred^*$-effective WSTS, not necesarly downward compatible,
 the above construction
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for $\pred$-basis-effective ???? $\omega^2$-WSTS with strong compatibility and with 
+{\sc Resilience} for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
 $\Safe=\downarrow \Safe$ is undecidable.
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for $\post$-effective, $\post$-ideal-effective and effective $\pred$-basis WSTS with downward and upward compatibilities
+{\sc State-resilience} is decidable for $\post$-effective, $\pred$-basis-effective, $\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
 and $\Safe = \downarrow \Safe$.
@@
 \begin{proof}
+Since $\mathscr{S}$ is $\pred^*$-basis-effective, $\Safe = \downarrow \Safe$ and $\pred^*(\Safe)=\downarrow \pred^*(\Safe)$, we may compute the finite decomposition in ideals of $\pred^*(\Safe)$.
+
 In order to decide whether %$\post^{\leq n}(s)
@@
 from \cite{DBLP:journals/lmcs/BlondinFM17}).
-The procedure then check for every downward closed subset $D_i$ whether $\downarrow \post^*(D_i) \subseteq D_i$. 
-\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
+The procedure then checks for every downward-closed subset $D_i$ whether $\downarrow \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
+%\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
 L'hypothèse "$\post$-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
 Il manquait la $\post$-effective dans les hypothèses}
-Every subset $D_i$ such that $\downarrow \post^*(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
+Every subset $D_i$ such that $\downarrow \post(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
 Notice that, by 
@@
 
-Procedure 1 stops when it finds a subset $D$ such that
-$\downarrow \post^*(D) \subseteq D$,
+Procedure 1 stops when it finds a downward-closed subset $D$ such that
+$\downarrow \post(D) \subseteq D$,
 $s \in D$ and
 $D   \subseteq \pred^*(\Safe)$. 
+%\mathieu{ici $\pred^*(\Safe)$ est downward-closed.
+du coup pas sur que la $\pred$-basis-effectivité telle qu'elle est actuellement nous permette de trancher ...
+à moins que les hypothèses d'effectivté sur les idéaux nous permettent de guarantir qu'on puisse calculer le $\pred$ d'un idéal ?}
+
 Indeed
@@
 The second procedure iteratively computes
-$\post^{\leq n}(s) $ \alain{ici on utilise $\post$-effective que j'ai rajoutée}
+$\post^{\leq n}(s) $ (this is effective because $\mathscr{S}$ is $\post$-effective)
+%			\alain{ici on utilise $\post$-effective que j'ai rajoutée}
 until it finds an element
@@
 
-We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
 We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general.
@@
  $\Safe = \downarrow \Safe$).
+
+We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
 

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 \keywords{Verification, Resilience, Home-Space, Well-structured transition systems, Vector addition system with states}
+
+\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée}
 
@@
 
-In this section we study VASS. Since they are ideally-effective and completion-$\post$-effective WSTS, they inherit the decidability results for WSTS in the case 
+In this section we study VASS.
+Since they enjoy many properties of effectivity ($\leq$ is ideally-effective, VASS are $\post$-effective, $\post$-ideal-effective, $\pred$-basis-effective, $\post^*$-basis-effective), they inherit the decidability results for WSTS in the case
 $\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
@@
 \begin{proof}
-We consider in VASS the case of when $\Safe$ is semilinear
+We consider the case where $\Safe$ is semilinear
 and $\Bad = S \setminus \Safe$ is semilinear too.
-We want to use the fact that
+We want to use the fact that, for VASS,
 it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
@@
 
-$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical discrete $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
 
@@
 the path from $\Bad$ to $\Safe$
-when $\Safe$ is downward-closed. Unfortunately, we have the following:
+when $\Safe$ is downward-closed. Surprisingly, we have the following result for VAS (not VASS):
 \begin{proposition}

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility. 
-
+%
+\alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
 \begin{proposition}\label{indec WSTS}
@@
 We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider $\omega^2$-WSTS with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
 
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WSTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 \begin{proof}
-Let us first recall two results in \cite{BFM-ic17}.
+Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
 Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for $\pred^*$-effective downward-WSTS with 
+{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
@@
 \begin{proof}
-Let $\mathscr{S}$ be a $\pred^*$-effective downward-WSTS with 
+Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
 $\Safe=\downarrow \Safe$.
-By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-effective.
+By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
 The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
@@
 In the case
-of a $\pred^*$-effective WSTS, not necesarly downward compatible,
+of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
 the above construction

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{lemma}\label{downward compatible}
-For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
+For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
 \end{lemma}
@@
 such that $x \leq y$.
-By definition of $\pred^*(D)$ then there exists 
-$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+By definition of $\pred^*(D)$, there exists 
+$d \in D$ such that: (1) either $y=d$ and then $x \in D \subseteq \pred^*(D)$ or (2) there exist $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
 $y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
@@
 It is of interest to note this case can be linked to the problem of mutual exclusion.
-Indeed the well-known mutual exclusion property can be modelized, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
+Indeed the well-known mutual exclusion property can be modeled, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
 Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
@@
 
-We complemented previous results by providing some undecidability proofs for resilience and state-resilience in general.
-We also exhibited classes of WSTS with decidable resilience, namely $\post$-ideal-effective WSTS with strong compatibility (in the case
- $\Safe = \uparrow \Safe$) and 
-downward-compatible
-effective WSTS \alain{vérifier les effectivités}
-(in the case 
- $\Safe = \downarrow \Safe$).
-
-We have extended the state-resilience problem introduced in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}.
+We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
+We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
+ $\Safe = \uparrow \Safe$ and in the case 
+ $\Safe = \downarrow \Safe$.
 

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 Blondin and al.
-showed (Lemma 4.3 in \cite{DBLP:journals/lmcs/BlondinFM17}) that under the previous hypotheses, one are able to enumerate \emph{downward closed sets} (by their finite decomposition in ideals), to decide inclusion between downward closed sets and to decide if a state belongs to a given finite set of ideals.
-But the five properties are not sufficient to compute the complementary and the intersection of downward closed sets.
+showed (Lemma 4.3 in \cite{DBLP:journals/lmcs/BlondinFM17}) that under the previous hypotheses, one are able to enumerate \emph{downward-closed sets} (by their finite decomposition in ideals), to decide inclusion between downward-closed sets and to decide if a state belongs to a given finite set of ideals.
+But the five properties are not sufficient to compute the complementary and the intersection of downward-closed sets.
 %
@@
 \cite{DBLP:phd/hal/Halfon18}.
-%	 (hence of downward closed sets).
+%	 (hence of downward-closed sets).
 
@@
 Again $Q \times \N^d$
- denotes the set of states of $V$.
+ is the set of states of $V$.
 For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
@@
 and $\textbf{u}+\textbf{z} = \textbf{v} \geq 0$,
-\item $t = (q,reset(\gamma),q') \in T$ 
+\item $t = (q,reset(i),q') \in T$ 
 and
-$\textbf{v}[\gamma] = 0$, and $\textbf{v}[\gamma'] = \textbf{u}[\gamma']$ for all $\gamma' \in \{1,\ldots, d\} \setminus \gamma$.
+$\textbf{v}[i] = 0$, and $\textbf{v}[i'] = \textbf{u}[i']$ for all $i' \in \{1,\ldots, d\} \setminus i$.
 \end{itemize} \end{samepage}
@@
 Remark that in the construction, from any input in $Q_{M'} \times \uparrow(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward closed set $Q_{M'} \times \downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
 from any input in $Q_{M'} \times \uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
@@
 which we then reduce to {\sc State-resilience}.
-A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a state
- $p(\textbf{u})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} $ ($t$ is enabled in $p(\textbf{u})$). We say that the whole reset-VASS is live if all its control-transitions are
+A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists two states
+ $p(\textbf{u}),p'(\textbf{u'})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} p'(\textbf{u'}) $. 
+%		($t$ is enabled in $p(\textbf{u})$). 
+We say that the whole reset-VASS is live if all its control-transitions are
 live. This leads to the following problem.
@@
 
-Procedure 1 enumerates every downward closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
+Procedure 1 enumerates every downward-closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
 The computability of the enumeration comes from the hypothesis that $\leq$ is ideally-effective (Lemma~4.3.
@@
 $\pred^{\leq k}(\Safe)$ 
-to other downward closed subsets.
+to other downward-closed subsets.
 

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 
+\newcommand{\pre}{\textsf{pre}}
+
 \newcommand{\pred}{\textsf{pred}}
@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\downarrow \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\mathop{\downarrow} \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
@@
 Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
-$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\downarrow \Safe$.
+$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
 Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS); most of the resilience properties are shown decidable.
@@
 
-\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée}
+\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée- peutx-tu changer pour review 3: $A = \mathop{\downarrow} B$ instead of $A = \mathop{\downarrow} B$: this dangling arrow is typographically horrific!}
 
@@
 It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
-In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report).
+In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report \cite{de1989decidability}).
 In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
@@
 In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
-$\uparrow$ $\post^*$ basis) in the case $\Safe=\uparrow \Safe$ and $\Bad = \downarrow \Bad$.
+$\uparrow$ $\post^*$ basis) in the case $\Safe=\uparrow \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
 
@@
 \item The resilience problem is decidable for ideally-effective WSTS with 
-$\Safe=\downarrow \Safe$
+$\Safe=\mathop{\downarrow} \Safe$
 and
@@
  {\em downward-closure},
- $\downarrow A = \{x \in X \mid \exists a \in A ~ x \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
-as $\uparrow x$ (resp. $\downarrow x$).
-%
-A {\em basis} of an upward-closed set $U$ is a set $U_b$ such that $U = \uparrow U_b$; similarly, a {\em basis} of a downward-closed set $D$ is a set $D_b$ such that $D = \downarrow D_b$.
+ $\mathop{\downarrow} A = \{x \in X \mid \exists a \in A ~ x \leq a\}$. 
+We abbreviate $\uparrow \{x\}$ (resp. $\mathop{\downarrow} \{x\}$)
+as $\uparrow x$ (resp. $\mathop{\downarrow} x$).
+%
+A {\em basis} of an upward-closed set $U$ is a set $U_b$ such that $U = \uparrow U_b$; similarly, a {\em basis} of a downward-closed set $D$ is a set $D_b$ such that $D = \mathop{\downarrow} D_b$.
 
@@
 
-Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$.
+Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where $Ideals(S)$ is the set of ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\mathop{\downarrow} \post_{\mathscr{S}}(I)$.
 The completion is always \emph{finitely branching} but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo.
@@
 Let $D$ be a downward-closed subset of $S$
-and let $x \in \downarrow \pred^*(D)$.
+and let $x \in \mathop{\downarrow} \pred^*(D)$.
 By downward closure, there exists
@@
 $d'$ 
-belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\downarrow\pred^*(D) \subseteq \pred^*(D)$.
-Since $\pred^*(D) \subseteq \downarrow \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\downarrow\pred^*(D)$ is downward-closed.
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\mathop{\downarrow}\pred^*(D) \subseteq \pred^*(D)$.
+Since $\pred^*(D) \subseteq \mathop{\downarrow} \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\mathop{\downarrow}\pred^*(D)$ is downward-closed.
 
 Let us prove $2 \implies 1$.
-Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\downarrow t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\downarrow t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\downarrow t_1)$.
+Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\mathop{\downarrow} t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\mathop{\downarrow} t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\mathop{\downarrow} t_1)$.
 This means that there exists $t_2 \leq t_1$ such that
@@
 %
-Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\downarrow s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
+Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\mathop{\downarrow} s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
 Most well-known wqo are ideally-effective .
@@
 \item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any ideal $I \in Ideals(S)$ and returning $\downarrow \post(I)$, expressed as a finite union of (maximal) ideals.
+any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
 %\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
@@
 \item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any ideal $I \in Ideals(S)$ and returning a finite basis of $\downarrow \pred^*(I)$.
+any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
 \end{itemize}
@@
 
-We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\downarrow \Bad$.
+We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
 Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
@@
 Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \downarrow x$, there exists an ideal $J \supseteq \downarrow y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
 Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
@@
 
-Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\downarrow s_j$ is coverable (for inclusion) from $\downarrow x$ in $\hat{\mathscr{S}}$.
+Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}}$.
 %
 Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
-From Proposition~30, there exist an ideal $J$ and a run $\downarrow x \xrightarrow{k} J$ where $J \supseteq \downarrow y \supseteq \downarrow s_j$ in $\hat{\mathscr{S}}$, hence $\downarrow s_j$ is covered from $\downarrow x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\downarrow s_j \subseteq J$ then 
+From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
 %	for every $y \in J$, 
@@
 
-Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\downarrow s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
-%
-Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\downarrow s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow s_j$ is not coverable from $J_i$.
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\mathop{\downarrow} s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
+%
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
 Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
@@
 Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
-Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\downarrow s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\downarrow s_j$ is not coverable from $J_i$.
-If $\downarrow s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\downarrow s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
+If $\mathop{\downarrow} s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\mathop{\downarrow} s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
 
@@
 property that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$, thus 
+$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\uparrow \Safe$, thus 
 the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
@@
 
-\subsection{Case: $\Safe=\downarrow \Safe$.}\label{safe-down}
-
-Let us now consider the case $\Safe=\downarrow \Safe$ hence $\Bad=\uparrow \Bad$.
+\subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
+
+Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\uparrow \Bad$.
 It is of interest to note this case can be linked to the problem of mutual exclusion.
@@
 {\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\downarrow \Safe$.
+$\Safe=\mathop{\downarrow} \Safe$.
 \end{theorem}
@@
 Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\downarrow \Safe$.
+$\Safe=\mathop{\downarrow} \Safe$.
 By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
@@
 i.e. when
-$\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) \neq \emptyset$
+$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
 then
-$\Bad \not\subseteq \downarrow\pred^*(\Safe)$
+$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
 and hence
 $\Bad \not\subseteq \pred^*(\Safe)$.
-When $\Bad \cap (S \setminus \downarrow\pred^*(\Safe)) = \emptyset$
+When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
 however
@@
 {\sc Resilience} for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
-$\Safe=\downarrow \Safe$ is undecidable.
+$\Safe=\mathop{\downarrow} \Safe$ is undecidable.
 \end{proposition}
@@
 Remark that in the construction, from any input in $Q_{M'} \times \uparrow(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \downarrow (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
 from any input in $Q_{M'} \times \uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
@@
   (from Theorem~$5$ in \cite{araki1976PN}).
-Hence $Q_{M'} \times \downarrow (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
 resilience hold in the reset-VASS simulating $M'$.
@@
 property $\Bad$ complement of $\Safe$, simply 
-$\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$, thus 
+$\Bad=\uparrow \Bad$ and $\Safe=\mathop{\downarrow} \Safe$, thus 
 the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.\\
@@
 
-Let us define the set of states such that $t$ is enabled:
-$pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $pre(t)$ is upward-closed.
+Let us define the set of states, $\pre(t)$, from which a  control-transition $t$ is enabled:
+$\pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v}) \in Q \times \N^d$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $\pre(t)$ is upward-closed.
 
@@
 \[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in \pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
 The previous formula reduces itself to {\sc State-resilience}
-where $\Safe = pre(t)$.
+where $\Safe = \pre(t)$.
 \end{proof}
@@
 
-Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
-{\sc $t$-liveness} is undecidable.
-
-Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong upward-compatibility.
+We may now prove the undecidability of the three state-resilience problems.
 
@@
 \begin{proof}
-{\sc State-resilience} itself is undecidable 
-since 
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
+{\sc $t$-liveness} is undecidable.
+Because
 {\sc $t$-liveness} is undecidable in reset-VASS
-and reducible to {\sc State-resilience}.
+and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
 Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
@@
 
-\subsection{Case: $\Safe = \downarrow \Safe$}
-
-We now consider the case $\Safe = \downarrow \Safe$.
+\subsection{Case: $\Safe = \mathop{\downarrow} \Safe$}
+
+We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
 %
 Unfortunately {\sc State-resilience} is undecidable in this case.
-As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
 Section~\ref{safe-down}.\alain{trop rapide}
@@
 when
-$\Safe=\downarrow \Safe$.
+$\Safe=\mathop{\downarrow} \Safe$.
 \end{proposition}
 
-Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$
+Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \mathop{\downarrow} \Safe$
 is symmetrical to the case $\Safe = \uparrow \Safe$.
@@
 Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap U \subseteq D$  iff $ (\downarrow  A) \cap U \subseteq D$.
+Then $A \cap U \subseteq D$  iff $ (\mathop{\downarrow}  A) \cap U \subseteq D$.
 \end{lemma}
@@
 Lemma~\ref{Lemma intersection 2} can be used to show that
-if $\Safe = \downarrow \Safe$,
+if $\Safe = \mathop{\downarrow} \Safe$,
 then
 $\post^*(s)  
-\subseteq \pred^*(\Safe)$  iff $ (\downarrow  \post^*(s)) 
+\subseteq \pred^*(\Safe)$  iff $ (\mathop{\downarrow}  \post^*(s)) 
 \subseteq \pred^*(\Safe)$.
@@
 {\sc State-resilience} is decidable for $\post$-effective, $\pred$-basis-effective, $\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
-and $\Safe = \downarrow \Safe$.
+and $\Safe = \mathop{\downarrow} \Safe$.
 \end{theorem}
@@
 \begin{proof}
-Since $\mathscr{S}$ is $\pred^*$-basis-effective, $\Safe = \downarrow \Safe$ and $\pred^*(\Safe)=\downarrow \pred^*(\Safe)$, we may compute the finite decomposition in ideals of $\pred^*(\Safe)$.
+Since $\mathscr{S}$ is $\pred^*$-basis-effective, $\Safe = \mathop{\downarrow} \Safe$ and $\pred^*(\Safe)=\mathop{\downarrow} \pred^*(\Safe)$, we may compute the finite decomposition in ideals of $\pred^*(\Safe)$.
 
@@
 from \cite{DBLP:journals/lmcs/BlondinFM17}).
-The procedure then checks for every downward-closed subset $D_i$ whether $\downarrow \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
+The procedure then checks for every downward-closed subset $D_i$ whether $\mathop{\downarrow} \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
 %\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
@@
 Il manquait la $\post$-effective dans les hypothèses}
-Every subset $D_i$ such that $\downarrow \post(D_i) \subseteq D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
+Every subset $D_i$ such that $\mathop{\downarrow} \post(D_i) \subseteq D_i$ is an “over-approximation” of $\mathop{\downarrow} \post^*(s)$ if it contains $s$.
 Notice that, by 
-upward compatibility, $\downarrow \post^*(s)$ is such a subset and may eventually be found.
+upward compatibility, $\mathop{\downarrow} \post^*(s)$ is such a subset and may eventually be found.
 
 Procedure 1 stops when it finds a downward-closed subset $D$ such that
-$\downarrow \post(D) \subseteq D$,
+$\mathop{\downarrow} \post(D) \subseteq D$,
 $s \in D$ and
@@
 $D \subseteq  \pred^*(\Safe)$ implies
-$\downarrow \post^*(s)  \subseteq  \pred^*(\Safe)$
-since $ \downarrow \post^*(s)  \subseteq D$.
+$\mathop{\downarrow} \post^*(s)  \subseteq  \pred^*(\Safe)$
+since $ \mathop{\downarrow} \post^*(s)  \subseteq D$.
 
@@
 If resilience hold, then procedure 1 terminates since it eventually finds
- $D = \downarrow \post^*(s)$ 
+ $D = \mathop{\downarrow} \post^*(s)$ 
  such that  $D \subseteq  \pred^*(\Safe)$.
@@
 {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\pred$-basis-effective, $\post$-ideal-effective ???? WSTS, strong downward compatibility and upward compatibility when
-$\Safe = \downarrow \Safe$.
+$\Safe = \mathop{\downarrow} \Safe$.
 \end{corollary}
@@
 for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
-VASS do not inherit the decidability results for WSTS in the case $\Safe = \downarrow \Safe$.
-In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed.
+VASS do not inherit the decidability results for WSTS in the case $\Safe = \mathop{\downarrow} \Safe$.
+In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed and $\Bad = S \setminus \Safe$.
 We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
@@
 \begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$.
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
 \end{proposition}
@@
 Consider a given $k \in \N$,
- $\Bad \subseteq \N^d$ upward-closed and $\Safe \subseteq \N^d$ downward-closed, and
+ $\emptyset \neq \Bad \subseteq \N^d$ is upward-closed and $\Safe \subseteq \N^d$ is downward-closed, and
 consider a given VAS $V$.
 Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
-The set $\Bad$ admits a finite basis $B_\Bad$.
+The set $\Bad \neq \emptyset$ admits a finite basis $B_\Bad$.
 Consider the vector $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider the vector
@@
  $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
-Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \uparrow \Bad$.
 This lead to a decision procedure
@@
 {\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
-$\Safe = \downarrow \Safe$.
+$\Safe = \mathop{\downarrow} \Safe$.
 \end{theorem}
@@
  $\Safe = \uparrow \Safe$ and in the case 
- $\Safe = \downarrow \Safe$.
+ $\Safe = \mathop{\downarrow} \Safe$.
 
@@
 Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
+
+\alain{remercier les reviewers !}
 

### 2023-10-19 VMCAI-revised-version/final.tex
mathop{uparrow}

@@
 Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\uparrow \Safe$) or a downward-closed set ($\Safe=\mathop{\downarrow} \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
+We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\mathop{\uparrow} \Safe$) or a downward-closed set ($\Safe=\mathop{\downarrow} \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
 We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
 We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective $\pred$-basis).
-Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \uparrow \Safe$.
+Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \mathop{\uparrow} \Safe$.
 Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
-$\uparrow$ $\post^*$ basis and $\Safe=\uparrow \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
+$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
 Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS); most of the resilience properties are shown decidable.
@@
 
-In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$ (called effective
-$\uparrow$ $\post^*$ basis) in the case $\Safe=\uparrow \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
-
-Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
+In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$ (called effective
+$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
+
+Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
 However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
@@
 when
-$\Safe=\uparrow \Safe$. We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-
-\item The three resilience problems are decidable for ideally-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \uparrow \Safe$.
+$\Safe=\mathop{\uparrow} \Safe$. We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
+
+\item The three resilience problems are decidable for ideally-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \mathop{\uparrow} \Safe$.
 
@@
 We also show that removing the effective 
-$\uparrow$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, most WSTS don't enjoy this property.
+$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, most WSTS don't enjoy this property.
 We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideally-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility).
@@
 its {\em upward-closure},
- $\uparrow A = \{x \in X \mid \exists a \in A ~ a \leq x\}$
+ $\mathop{\uparrow} A = \{x \in X \mid \exists a \in A ~ a \leq x\}$
  and its 
@@
  $\mathop{\downarrow} A = \{x \in X \mid \exists a \in A ~ x \leq a\}$. 
-We abbreviate $\uparrow \{x\}$ (resp. $\mathop{\downarrow} \{x\}$)
-as $\uparrow x$ (resp. $\mathop{\downarrow} x$).
-%
-A {\em basis} of an upward-closed set $U$ is a set $U_b$ such that $U = \uparrow U_b$; similarly, a {\em basis} of a downward-closed set $D$ is a set $D_b$ such that $D = \mathop{\downarrow} D_b$.
+We abbreviate $\mathop{\uparrow} \{x\}$ (resp. $\mathop{\downarrow} \{x\}$)
+as $\mathop{\uparrow} x$ (resp. $\mathop{\downarrow} x$).
+%
+A {\em basis} of an upward-closed set $U$ is a set $U_b$ such that $U = \mathop{\uparrow} U_b$; similarly, a {\em basis} of a downward-closed set $D$ is a set $D_b$ such that $D = \mathop{\downarrow} D_b$.
 
  A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
-This property is equivalent to the \emph{finite decomposition} property : every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\uparrow B$.
+This property is equivalent to the \emph{finite decomposition} property : every upward-closed set $\emptyset \neq U \subseteq X$ admits a finite basis $B \subseteq X$ such that $U=\mathop{\uparrow} B$.
 %
@@
 %it is $\post$-effective, and there exists an algorithm $M_\downarrow$ that computes 
-% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
 % $s_1, s_2, \ldots, s_m \in S$.
@@
 %
-Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\mathop{\downarrow} s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\uparrow s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
+Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\mathop{\downarrow} s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\mathop{\uparrow} s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
 Most well-known wqo are ideally-effective .
@@
 \item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any state $s \in S$ and returning a finite basis of $\uparrow \pred(\uparrow s)$.
+any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
 %
 %and there exist an algorithm $M_\downarrow$ that computes 
-% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\uparrow^C}$ that computes the ideal decomposition of $S \setminus \uparrow \{ s_1, s_2, \ldots, s_m\}$, on input
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
 % $s_1, s_2, \ldots, s_m \in S$.
@@
 \item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any state $s \in S$ and returning a finite basis of $\uparrow \post^*(s)$.
+any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
 %
@@
 % \alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
-% Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.
+% Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\mathop{\uparrow} s )$ is computable for every $s \in S$.
 % We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
@@
 We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
-When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\uparrow U_k)_k$ instead.
+When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\mathop{\uparrow} U_k)_k$ instead.
 
@@
 {An ordered transition system $\mathscr{S}=(S,\rightarrow, \leq)$ and two states $s_0,s \in S$.}
-{ $s_0 \in \pred^*(\uparrow s )$ ? \newline}
+{ $s_0 \in \pred^*(\mathop{\uparrow} s )$ ? \newline}
 
@@
 	it is
-	$\uparrow (1,1)$-resilient
+	$\mathop{\uparrow} (1,1)$-resilient
 	iff TM$_j$ halts on input $j$,
@@
 
-\subsection{Case: $\Safe=\uparrow \Safe$.}\label{safe-up}
-
-We start with the case $\Safe=\uparrow \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
+\subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
+
+We start with the case $\Safe=\mathop{\uparrow} \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
 Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
@@
 \begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 property that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\uparrow \Safe$, thus 
+$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
 the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
@@
 \begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \uparrow \Safe$.
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
 {\sc Resilience}, {\sc Bounded resilience} 
@@
 %	it is
-%	$\uparrow (1,1)$-resilient
+%	$\mathop{\uparrow} (1,1)$-resilient
 %	iff TM$_j$ halts on input $j$,
@@
 
-Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\uparrow \Bad$.
+Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\mathop{\uparrow} \Bad$.
 It is of interest to note this case can be linked to the problem of mutual exclusion.
@@
 From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \uparrow (0,0,1)$. 
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
 %			with at least $1$ on its third counter. 
 We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \uparrow(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
 Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
 \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \uparrow(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
@@
 property $\Bad$ complement of $\Safe$, simply 
-$\Bad=\uparrow \Bad$ and $\Safe=\mathop{\downarrow} \Safe$, thus 
+$\Bad=\mathop{\uparrow} \Bad$ and $\Safe=\mathop{\downarrow} \Safe$, thus 
 the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.\\
@@
 
-\subsection{Case: $\Safe = \uparrow \Safe$}
-
-We start with the case $\Safe = \uparrow \Safe$.
+\subsection{Case: $\Safe = \mathop{\uparrow} \Safe$}
+
+We start with the case $\Safe = \mathop{\uparrow} \Safe$.
 Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
@@
 when
-$\Safe=\uparrow \Safe$.
+$\Safe=\mathop{\uparrow} \Safe$.
 \end{proposition}
@@
 reducible to {\sc $k$-state-resilience}:
-since $\Safe=\uparrow \Safe$ and
+since $\Safe=\mathop{\uparrow} \Safe$ and
 $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
-compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
 and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
 We compute 
@@
 hold for $k$ from $0$ to $n_0$.  
-Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\uparrow \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
 since 
-$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = {\uparrow \pred^*(\Safe)} = \pred^*(\Safe)$.
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
 Hence the undecidability of {\sc Bounded-state-resilience}
@@
 when
-$\Safe=\uparrow \Safe$.
+$\Safe=\mathop{\uparrow} \Safe$.
 \end{theorem}
@@
 Then, because {\sc $n_0$-state-resilience} is decidable, 
-checking $\uparrow \post^*(s) 
+checking $\mathop{\uparrow} \post^*(s) 
  \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
@@
 
-The proof of Theorem~\ref{ref ozkan} rely on the computability of $\uparrow \post^*(s)$ and on the following lemma.
+The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
 
@@
 Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\uparrow  A) \cap D \subseteq U$.
+Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
 \end{lemma}
@@
 \begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\uparrow (A \cap D)} \subseteq {\uparrow U} = U$.
-Let us show that $({\uparrow A}) \cap D \subseteq {\uparrow (A \cap D)}$.
-Let $x \in ({\uparrow A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
+Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
+Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
 Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \uparrow (A \cap D)}$.
+Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
 In the other direction,
-since $A \subseteq {\uparrow A}$, the inclusion
-$({\uparrow  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\uparrow  A}) \cap D \subseteq U$. \qed
+since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
+$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
 \end{proof}
@@
 \begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\uparrow \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
-$B_{\uparrow \post^*(s)} \cap D = \emptyset$, 
-where $B_{\uparrow \post^*(s)}$ is a basis of $\uparrow \post^*(s)$,
+$B_{\mathop{\uparrow} \post^*(s)} \cap D = \emptyset$, 
+where $B_{\mathop{\uparrow} \post^*(s)}$ is a basis of $\mathop{\uparrow} \post^*(s)$,
  that is equivalent to $\post^*(s)\cap D = \emptyset$ by
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \uparrow \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
+$U_n = \mathop{\uparrow} \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
 until it becomes stationnary. 
@@
 It is well-known that 
-VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\uparrow \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\mathop{\uparrow} \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
 Hence {\sc State-resilience} is decidable for  
@@
 
-However, the hypothesis that $\uparrow \post^*$ is computable cannot be tested in the general WSTS framework.  
+However, the hypothesis that $\mathop{\uparrow} \post^*$ is computable cannot be tested in the general WSTS framework.  
 Moreover there exist classes of WSTSs with strong 
- compatibility for which there doesn't exist an algorithm computing a basis of $\uparrow \post^*$, such as reset-VASS.
-
-Keeping the $\uparrow \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+ compatibility for which there doesn't exist an algorithm computing a basis of $\mathop{\uparrow} \post^*$, such as reset-VASS.
+
+Keeping the $\mathop{\uparrow} \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
 Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
@@
 when
-$\Safe=\uparrow \Safe$. 
+$\Safe=\mathop{\uparrow} \Safe$. 
 \end{theorem}
@@
 \begin{proof}
-Let $B_{\uparrow \post^*(s)}$ 
- be a basis of $\uparrow \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
+Let $B_{\mathop{\uparrow} \post^*(s)}$ 
+ be a basis of $\mathop{\uparrow} \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
 By applying Lemma~\ref{Lemma intersection} twice, we obtain
 \[ \post^*(s) 
- \subseteq \pred^*(\Safe) \text{ iff } B_{\uparrow \post^*(s)}  
+ \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
  \subseteq \pred^*(\Safe)\]
-Since $B_{\uparrow \post^*(s)}$ is finite and we can compute 
+Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite and we can compute 
 %\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
-a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\uparrow \post^*(s)} 
+a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\mathop{\uparrow} \post^*(s)} 
  \setminus \pred^*(\Safe) = \emptyset$.  \qed
@@
 However when removing strong compatibility, some precision is lost.
-Since $\pred(\uparrow \Safe)$ is not necessarily upward-closed, it is possible to have 
- $\uparrow \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
+Since $\pred(\mathop{\uparrow} \Safe)$ is not necessarily upward-closed, it is possible to have 
+ $\mathop{\uparrow} \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
 despite having 
@@
 Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \mathop{\downarrow} \Safe$
-is symmetrical to the case $\Safe = \uparrow \Safe$.
+is symmetrical to the case $\Safe = \mathop{\uparrow} \Safe$.
 %
@@
 Since they enjoy many properties of effectivity ($\leq$ is ideally-effective, VASS are $\post$-effective, $\post$-ideal-effective, $\pred$-basis-effective, $\post^*$-basis-effective), they inherit the decidability results for WSTS in the case
-$\Safe = \uparrow \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
+$\Safe = \mathop{\uparrow} \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
 for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
@@
 This all changes when one considers VASS (having control-states).
-For a VASS $V = (Q,T)$, a set $\Bad = \uparrow \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
 then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
 If there is none, then
-$\uparrow q(\textbf{0})$
+$\mathop{\uparrow} q(\textbf{0})$
  is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
  $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
-Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \uparrow \Bad$.
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
 This lead to a decision procedure
@@
 If there is none, then
-$\uparrow q(\textbf{0})$
+$\mathop{\uparrow} q(\textbf{0})$
  is in the complement of $\Bad$, i.e. $\Safe$.
@@
 Let us assume from now on 
- $\uparrow q_{j_1}(\textbf{0}), \uparrow q_{j_2}(\textbf{0}), \ldots \uparrow q_{j_\ell}(\textbf{0})$ are
+ $\mathop{\uparrow} q_{j_1}(\textbf{0}), \mathop{\uparrow} q_{j_2}(\textbf{0}), \ldots \mathop{\uparrow} q_{j_\ell}(\textbf{0})$ are
 all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
 Because these subsets are upward-closed, we can compute a basis of
-$\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\uparrow q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))$.
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\mathop{\uparrow} q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
 
 We now consider the set 
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
 We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
@@
 if
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
 is
@@
 with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
 and hence from which it is not possible at all to reach
 $\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
 $ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
@@
 and hence
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$.
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
 We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
@@
 From $k_0$ onwards,
-$\Bad \setminus (\pred^{\leq k}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\uparrow q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
 is stationnary. 
@@
 $k_{\pi}$ the maximum of the length of the paths from 
-$\Bad \setminus (\pred^{\leq k_0}(\uparrow q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\uparrow q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
 If it is infinite then
@@
 if
-$\Bad \setminus (\pred^{*}(\uparrow q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\uparrow q_{j_\ell}(\textbf{0})) )$
+$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
 is
@@
 with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\uparrow q_{j_i}(\textbf{0}))$,
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
 and hence from which it is not possible at all to reach
@@
 We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
- $\Safe = \uparrow \Safe$ and in the case 
+ $\Safe = \mathop{\uparrow} \Safe$ and in the case 
  $\Safe = \mathop{\downarrow} \Safe$.

### 2023-10-19 VMCAI-revised-version/final.tex
effectivity

@@
 
-\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée- peutx-tu changer pour review 3: $A = \mathop{\downarrow} B$ instead of $A = \mathop{\downarrow} B$: this dangling arrow is typographically horrific!}
+%\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée- peutx-tu changer pour review 3: $A = \mathop{\downarrow} B$ instead of $A = \mathop{\downarrow} B$: this dangling arrow is typographically horrific!}
 
@@
 
-By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that possesses a characteristic that allows to obtain more algorithms:
+By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that allow to construct coverability trees:
 
@@
 
-Again $Q \times \N^d$
- is the set of states of $V$.
+%Again $Q \times \N^d$
+% is the set of states of $V$.
 For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
@@
 . Since $\Safe \subseteq \pred^*(\Safe)$, one only needs to decide whether 
-the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.
-In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.  \\
-
+the complement of $\Safe$ is in $\pred^*(\Safe)$. From now on, we use $\Bad$ to denote the complement of $\Safe$.\\
+%In   	 \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.  \\
+%
 Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
@@
 %%%%%%%
-\alain{faire les deux autres résilience}
+\alain{prouver, si possible, les deux autres résilience}
 \end{proof} \qed
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
-$\Safe=\mathop{\downarrow} \Safe$ is undecidable.
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+$\Safe=\mathop{\downarrow} \Safe$.
 \end{proposition}
@@
 \alain{écrire qqchose ici}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN}~\cite{dufourd1998reset}. 
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
 % \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
@@
 to
-a state of the form $ q(0,  \ldots 0)$ if and only if
+a state $ q(0,  \ldots 0)$ if and only if
 there exists
@@
  to
- a state of the form $q_f(0,0,0)$
+ the state $q_f(0,0,0)$
   (from Theorem~$5$ in \cite{araki1976PN}).

### 2023-10-20 VMCAI-revised-version/final.tex
u

--- 2023-10-19 VMCAI-revised-version/final.tex
@@
 
+\mathieu{Reviewer 1: we can consider that the example at the bottom of page 8, even if it is a WSTS, is a very particular case very different from VASS, lossy counter, Petri nets, communicating automata, etc.
+and wonder if classes containing such  systems are interesting per se.
+}
 % Hence we can see how adding transitions can 
@@
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for $\post$-ideal-effective ???? and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
+{\sc State-resilience} is undecidable for 
+%$\post$-ideal-effective ???? and
+ $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
 when

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-19 VMCAI-revised-version/final.tex
@@
 \begin{proposition}\label{indec WSTS}
-Resilience for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
+{\sc Resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
 \end{proposition}

### 2023-10-20 VMCAI-revised-version/final.tex
Rearranged a bit the t-liveness section, déjà on commence par parler de la t-liveness et on fait moins de back and forth entre zero-reachability et la t-liveness, comme ça quand on parle de la zero reachability c'est déjà davantage évident pq la t-liveness c'est très proche de la resilience. et aussi du coup quand j'introduit la zero-reachability j'ai rajouté une phrase pour parler du fait que c'est pas le même type de quantificateur ce qui rends la réduction nontriviale

@@
 
-We now recall the
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-%
-We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
-which we then reduce to {\sc State-resilience}.
 A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists two states
@@
 
+It now remains to argue why {\sc $t$-liveness} is undecidable.
+We  recall the
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+Remark that the quantification over the path is existential rather than universal i.e. 
+{\em zero-reachability}
+does not asks that zero remains reachable from an initial state $s_0$ at all times but rather, that there exists at least one way to reach it.
+Despite this it is possible to reduce 
+the {\em zero-reachability problem} 
+to
+{\sc $t$-liveness}.
 The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets, whose proof can be seen in Appendix~\ref{appendix}%

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
 In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors.
-More precisely, we say that a WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ is
-\begin{itemize}
+More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is
+\begin{enumerate}
 %\item {\em effective} if there exists a pair of algorithms
@@
 any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
-\end{itemize}
+\end{enumerate}
+
+Counter machines are effective but don't enjoy any other effectivities among the list of five.
+Reset VASS (hence VASS) and (front) lossy fifo automata are $\post$-effective, $\post$-ideal-effective, and $\pred$-basis-effective.
+There exist $\post$-effective WSTS that are not $\post$-ideal-effective (Proposition 35 in \cite{BFM-ic17}); there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition 36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition 45 in \cite{BFM-ic17}).
+Reset VASS are not $\post^*$-basis-effective since $\post^*$-basis-effectiveness allows to compute the finite set of the set of minimal reachable states, hence it would allow to decide the zero-reachability problem that is undecidable for reset-VASS.
+However, VASS are $\post^*$-basis-effective \cite{DBLP:journals/corr/abs-2108-00889} and LCS (with $d$ fifo channels) are also $\post^*$-basis-effective because we have $(q,w_1,w_2,...,w_d) \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)  \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)$ is reachable from $(q_0,u_1,u_2,...,u_d)$ that is decidable in LCS.
+%
+Reset-VASS are not $\pred^*$-basis-effective because computing a finite basis of $\mathop{\downarrow} \pred^*(q,0,...,0)$ would allow to decide whether $(q,0,...,0)$ is reachable that is undecidable for reset-VASS.
+
+VASS are not $\pred^*$-basis-effective ????
 
@@
 
-As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo and reset-VASS are $\post$-effective, $\post$-ideal-effective and $\pred$-basis-effective but they are not $\post^*$-basis-effective (VASS are $\post^*$-basis-effective).
+As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo.
 

### 2023-10-20 VMCAI-revised-version/final.tex

@@
 Reset-VASS are not $\pred^*$-basis-effective because computing a finite basis of $\mathop{\downarrow} \pred^*(q,0,...,0)$ would allow to decide whether $(q,0,...,0)$ is reachable that is undecidable for reset-VASS.
-
-VASS are not $\pred^*$-basis-effective ????
+By using the decidability of reachability and Theorem 3.11 in \cite{DBLP:journals/acta/ValkJ85}, we may prove that VASS are $\pred^*$-basis-effective.
+A more complete study of these properties will be done in the long version of this paper.
 

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@
 \begin{proof}
+\mathieu{preuve résumé / concise à écrire}
+\iffalse
 \alain{preuve pas claire, à réécrire}
@@
 %\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
+\fi
 \end{proof}
@@
 \begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for $\post$-effective, $\pred$-basis-effective, $\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
+{\sc State-resilience} is decidable for $\post$-effective, 
+%$\pred$-basis-effective, 
+$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
 and $\Safe = \mathop{\downarrow} \Safe$.
@@
 
-Strong downward compatibility implies furthermore the decidability
-of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.
-\alain{trop rapide}
-Indeed, with strong downward compatibility, 
-$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is,
-and becomes possible to represent \mathieu{... and compute ?} and compare
-$\pred^{\leq k}(\Safe)$ 
-to other downward-closed subsets.
+In the case of strong downward compatibility,
+$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is.
+Hence, 
+assuming a $\pred^*$-basis-effective variation
+consisting in an algorithm
+accepting any ideal $I$ and returning a finite basis of
+ $\mathop{\downarrow} \pred(I)$, it becomes possible to perform the same procedures as above except with comparisons against $\pred^{\leq k}(\Safe)$ rather than $\pred^*(\Safe)$, and hence {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable in this case.
 
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\pred$-basis-effective, $\post$-ideal-effective ???? WSTS, strong downward compatibility and upward compatibility when
-$\Safe = \mathop{\downarrow} \Safe$.
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
+$\Safe = \mathop{\downarrow} \Safe$
+and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
+ $\mathop{\downarrow} \pred(I)$.
 \end{corollary}
+
+\mathieu{Donner des exemples dans cette classe.}
 
@@
 
+\section{Reset-VASS and Minsky-machines}\label{minsk}
+
+\begin{proposition}\label{indec WSTS with dcs}
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{proposition}
+
+\begin{proof}
+\alain{preuve pas claire, à réécrire}
+% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
+%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
+%which can be simulated by WSTS.
+First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
+%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
+From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
+%			with at least $1$ on its third counter. 
+We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
+Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+Hence {\sc Resilience} is undecidable for Minsky machines.
+\alain{écrire qqchose ici}
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
+% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
+%Reset-VASS extend the basic VASS model with special “reset
+%transitions” that set to $0$ some coordinates in the vector. 
+%			Let us recall their definition here.
+%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
+The simulation operates by building a reset-VASS in which
+there exists
+ a sequence from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state $ q(0,  \ldots 0)$ if and only if
+there exists
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ the state $q_f(0,0,0)$
+  (from Theorem~$5$ in \cite{araki1976PN}).
+Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+resilience hold in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.\qed
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
+\end{proof}
+
 \section{Reset-VASS and state-resilience}\label{appendix}

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@
 
-\mathieu{Reviewer 1: we can consider that the example at the bottom of page 8, even if it is a WSTS, is a very particular case very different from VASS, lossy counter, Petri nets, communicating automata, etc.
-and wonder if classes containing such  systems are interesting per se.
-}
 % Hence we can see how adding transitions can 
@@
 
-By applying Proposition 4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that allow to construct coverability trees:
+By applying Proposition~4.3 from \cite{DBLP:journals/corr/abs-1208-4549}, we can reintroduce a straightforward definition of a specific proper subset of wqos that allow to construct coverability trees:
 
@@
 
-%Using Proposition 4.3 in \cite{DBLP:journals/corr/abs-1208-4549}, we recall a simple definition of a strict subclass of wqos that has a very nice property: a wqo $(X,\leq)$ is an $\omega^2$-wqo iff $(Ideals(X),\subseteq)$ is a wqo.
+%Using Proposition~4.3 in \cite{DBLP:journals/corr/abs-1208-4549}, we recall a simple definition of a strict subclass of wqos that has a very nice property: a wqo $(X,\leq)$ is an $\omega^2$-wqo iff $(Ideals(X),\subseteq)$ is a wqo.
 
@@
 %			which are wqo but not known to be $\omega^2$-wqo.
-The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is an $\omega^2$-wqo (Proposition 4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
+The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is an $\omega^2$-wqo (Proposition~4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
 Now we are able to define $\omega^2$-WSTS:
@@
 Reset VASS (hence VASS) and (front) lossy fifo automata are $\post$-effective, $\post$-ideal-effective, and $\pred$-basis-effective.
-There exist $\post$-effective WSTS that are not $\post$-ideal-effective (Proposition 35 in \cite{BFM-ic17}); there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition 36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition 45 in \cite{BFM-ic17}).
+There exist $\post$-effective WSTS that are not $\post$-ideal-effective (Proposition~35 in \cite{BFM-ic17}); there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition~36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition~45 in \cite{BFM-ic17}).
 Reset VASS are not $\post^*$-basis-effective since $\post^*$-basis-effectiveness allows to compute the finite set of the set of minimal reachable states, hence it would allow to decide the zero-reachability problem that is undecidable for reset-VASS.
@@
 	\begin{figure}
-	\hspace{1.5cm}
-\includegraphics[width=0.70\textwidth]{WSTS_taxonomy}
+	\hspace{2.3cm}
+\includegraphics[width=0.60\textwidth]{WSTS_taxonomy}
 	\caption{A taxonomy of WSTS and variants.}
@@
 %In   	 \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed.  \\
-%
+
 Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
@@
 
+\mathieu{Reviewer 1: the example at the bottom of page 8, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
+
 \subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
@@
 
-Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.
-\end{proof} \qed
-
-Remark that the above proof doesn't make use of the 
-property that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.\qed
+\end{proof}
 
@@
 $S \setminus \pred^*(\Safe)$,
-which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable.
+which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable. \qed\end{proof} 
 %%%%%%%
 \alain{prouver, si possible, les deux autres résilience}
-\end{proof} \qed
+
+\begin{corollary}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{corollary}
+
+Remark that the above proofs do not make use of the 
+property that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
 %				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity
up

@@
 %
-\alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
+% \alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
 \begin{proposition}\label{indec WSTS}
@@
 \end{proposition}
-% \alain{peut-on ajouter des hypothèses d'effectivity ?
+ \alain{peut-on ajouter des hypothèses d'effectivity ?
 peut-on montrer que l'exemple n'est pas pred effectif ?
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective.
-
-\mathieu{Reviewer 1: the example at the bottom of page 8, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective
+	nor $\pred$-effective.
+
+\mathieu{Reviewer 1: the example at the bottom of \textbf{ page 8} (now 10), even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
+
+For strongly compatible WSTS,
+$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$
+when $\Safe = \mathop[\uparrow}(\Safe)$. 
+$\mathop{\uparrow} \pred^{\leq k}(\Safe)$
+$k_min$ such that 
+$\pred^{\leq k_{min}}(\Safe) = \pred^{*}(\Safe) $
 

### 2023-10-20 VMCAI-revised-version/final.tex
corrigé une typi
effectivity

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
 \noindent
-Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state then it can also leave this bad state and reach a safe state.
-We consider six types of resilience (one of them is the home-space property) defined by an upward closed set $\Safe$  ($\Safe=\mathop{\uparrow} \Safe$) or a downward-closed set ($\Safe=\mathop{\downarrow} \Safe$), and by the existence of a bound $k$ on the length of minimal runs starting from $\Bad$ and reaching $\Safe$, where $\Bad$ is generally the complementary of $\Safe$.
-We study the decidability of each type of resilience for two kinds of models: WSTS and VASS.
-We first show that most of all resilience problems are undecidable for WSTS with strong compatibility (and with effective $\pred$-basis).
-Then we prove the decidability of resilience for completion-$\post$-effective $\omega^2$-WSTS with strong compatibility (almost all known WSTS are in this class) and $\Safe = \mathop{\uparrow} \Safe$.
+Resilience of unperfect systems is a key property for improving safety by insuring that if a system could go into a bad state in $\Bad$ then it can also leave this bad state and reach a safe state in $\Safe$.
+We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$).
+We study the decidability of each type of resilience for many models: Well Behaved Transition Systems (WBTS), WSTS, VASS and other variations of counter machines.
+%
+We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) for downward-closed sets $\Safe$.
+
+We then show that resilience is decidable for WBTS with the computability of the finite basis of the downward-closure of the set of (immediate) successors of a downward-closed set ($\post$-basis-effective) and for WSTS with the (well-known) $\pred$-basis-effective hypothesis.
+%
+%		most of all resilience problems are decidable for WSTS with the good effective hypothesis. R
+%
+Most of the resilience properties are shown decidable for other classes like downward-WSTS, VASS, lossy counter machines, integer VASS and continuous VASS.
+
+%Then we prove the decidability of resilience for WBTS and WSTS with strong compatibility, adapted effective hypothesis and for upward-closed sets $\Safe$.
 Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
-$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
-Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS); most of the resilience properties are shown decidable.
+%$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
+%Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS)
 \end{abstract}
@@
 \mathieu{Donner des exemples dans cette classe.}
+\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
 

### 2023-10-20 VMCAI-revised-version/final.tex
AF: prouver, si possible, les deux autres résilience

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
 
-For strongly compatible WSTS,
-$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$
-when $\Safe = \mathop[\uparrow}(\Safe)$. 
-$\mathop{\uparrow} \pred^{\leq k}(\Safe)$
-$k_min$ such that 
-$\pred^{\leq k_{min}}(\Safe) = \pred^{*}(\Safe) $
-
 \subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
@@
 \alain{prouver, si possible, les deux autres résilience}
+Strong compatibility implies furthermore the decidability
+of {\sc $k$-resilience} and {\sc bounded-resilience}.
 
@@
 \end{corollary}
+
+\begin{proof}
+For strongly compatible WSTS,
+$\pred^{k}(\Safe) = \mathop{\uparrow} \pred^{k}(\Safe)$,
+and hence
+$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$,
+when $\Safe = \mathop{\uparrow} \Safe$. 
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\mathop{\uparrow} \pred^{\leq k}(\Safe) = \pred^{\leq k}(\Safe)$.
+Like above, we can test if the intersection of $\Bad$
+and
+$S \setminus \pred^{\leq k}(\Safe)$ is empty or not, hence {\sc $k$-resilience} is decidable.
+To decide {\sc Bounded resilience}, we check {\sc $k$-resilience}
+starting with $k=0$ until we find some $k_0$ such that either {\sc $k_0$-resilience} holds,
+either $\pred^{\leq k_0}(\Safe) = \pred^{\leq k_0+1}(\Safe)$, whichever comes first.
+The convergence of $(\pred^{\leq k}(\Safe) )_{k\in\N}$ guarantees the latter eventually happens.
+\end{proof}
 

### 2023-10-20 VMCAI-revised-version/final.tex
transmission des modifications sur l'abstract

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
  Let us remark that the transition system 
-	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective.
-
-\mathieu{Reviewer 1: the example at the bottom of page 8, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
-
-\subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
-
-We start with the case $\Safe=\mathop{\uparrow} \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
-
-%
-\begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
-{\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{theorem}
-
-%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
-
-\begin{proof}
-Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-%
-
-Let  $\{s_1,s_2,...,s_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
-The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
-Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}}$.
-%
-Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
-From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq s_j$ such that $x \xrightarrow{k} y'  \geq s_j$ in $\mathscr{S}$ and then $s_j$ is coverable from $x$ in $\mathscr{S}$.
-
-Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\mathop{\downarrow} s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
-%
-Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
-Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
-%
-
-%
-We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
-
-This implies that resilience and bounded resilience can be reduced to coverability.
-Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
-  resilience problem and the bounded resilience problem are decidable.
-
-Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
-Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
-If $\mathop{\downarrow} s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\mathop{\downarrow} s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
-
-Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.\qed
-\end{proof}
-
-\begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
-{\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{theorem}
-
-\begin{proof}
-The resilience problem can be reformulated as 
-$\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
-Since the wqo is ideally-effective, we can compute the intersection of
-$\Bad$
-and
-$S \setminus \pred^*(\Safe)$,
-which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable. \qed\end{proof} 
-%%%%%%%
-\alain{prouver, si possible, les deux autres résilience}
-
-\begin{corollary}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{corollary}
-
-Remark that the above proofs do not make use of the 
-property that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
-
-%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
-%%%
-%%%
-%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
-%increasing recursive functions from \cite{FMP-ic04} defined as
-%
-%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
-%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
-%		
-%	\noindent	
-%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
-%	which moreover begins
-%by writing the integer $j$ on its tape,	
-%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
-%	The
-%	transition system 
-%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-%	is a WSTS
-%	and
-%	has the property that
-%	it is
-%	$\mathop{\uparrow} (1,1)$-resilient
-%	iff TM$_j$ halts on input $j$,
-%	since $(1,1)$ is coverable from $(0,0)$
-%	iff TM$_j$ halts on input $j$. 
-%	Hence there is no Turing machine which correctly determines 
-%	resilience and halts whenever its input is a
-%	WSTS.
-%Hence resilience for WSTS in general is undecidable.
-
-\subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
-
-Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\mathop{\uparrow} \Bad$.
-It is of interest to note this case can be linked to the problem of mutual exclusion.
-Indeed the well-known mutual exclusion property can be modeled, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
-Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
-
-%
-\begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\mathop{\downarrow} \Safe$.
-\end{theorem}
-
-\begin{proof}
-Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\mathop{\downarrow} \Safe$.
-By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
-The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
-Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
- subsets.
-Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
-$\Bad$
-and
-$S \setminus \pred^*(\Safe)$,
-which are both upward-closed. We may decide whether this intersection is empty, hence
-the resilience problem is decidable. \qed
-\end{proof}
-
-In the case
-of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
-the above construction
-can provide a proof
-of non-resilience
-i.e. when
-$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
-then
-$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
-and hence
-$\Bad \not\subseteq \pred^*(\Safe)$.
-When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
-however
-it is not enough to conclude.\\
-
-In the case of (upward compatible) WSTS with $\pred$-basis-effectivity, {\sc Resilience} is undecidable.
+	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective
+	nor $\pred$-effective.
+
+\mathieu{Reviewer 1: the example at the bottom of \textbf{ page 8} (now 10), even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
+
+In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable.
 
@@
 
+\subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
+
+We start with the case $\Safe=\mathop{\uparrow} \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
+
+%
+\begin{theorem}\label{down-up}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
+
+\begin{proof}
+Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
+Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+%
+
+Let  $\{s_1,s_2,...,s_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
+Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+
+Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}}$.
+%
+Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
+From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
+%	for every $y \in J$, 
+there exists $x \in I$ and $y' \geq s_j$ such that $x \xrightarrow{k} y'  \geq s_j$ in $\mathscr{S}$ and then $s_j$ is coverable from $x$ in $\mathscr{S}$.
+
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\mathop{\downarrow} s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
+%
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
+Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+%
+
+%
+We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
+
+This implies that resilience and bounded resilience can be reduced to coverability.
+Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+  resilience problem and the bounded resilience problem are decidable.
+
+Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
+If $\mathop{\downarrow} s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\mathop{\downarrow} s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.\qed
+\end{proof}
+
+\begin{theorem}\label{xxx}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+\begin{proof}
+The resilience problem can be reformulated as 
+$\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
+Since the wqo is ideally-effective, we can compute the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable. \qed\end{proof} 
+%%%%%%%
+%%\alain{prouver, si possible, les deux autres résilience}
+Strong compatibility implies furthermore the decidability
+of {\sc $k$-resilience} and {\sc bounded-resilience}.
+
+\begin{corollary}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{corollary}
+
+\begin{proof}
+For strongly compatible WSTS,
+$\pred^{k}(\Safe) = \mathop{\uparrow} \pred^{k}(\Safe)$,
+and hence
+$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$,
+when $\Safe = \mathop{\uparrow} \Safe$. 
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\mathop{\uparrow} \pred^{\leq k}(\Safe) = \pred^{\leq k}(\Safe)$.
+Like above, we can test if the intersection of $\Bad$
+and
+$S \setminus \pred^{\leq k}(\Safe)$ is empty or not, hence {\sc $k$-resilience} is decidable.
+To decide {\sc Bounded resilience}, we check {\sc $k$-resilience}
+starting with $k=0$ until we find some $k_0$ such that either {\sc $k_0$-resilience} holds,
+either $\pred^{\leq k_0}(\Safe) = \pred^{\leq k_0+1}(\Safe)$, whichever comes first.
+The convergence of $(\pred^{\leq k}(\Safe) )_{k\in\N}$ guarantees the latter eventually happens.
+\end{proof}
+
+Remark that the above proofs do not make use of the 
+property that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
+%%%
+%%%
+%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
+%increasing recursive functions from \cite{FMP-ic04} defined as
+%
+%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+%		
+%	\noindent	
+%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
+%	which moreover begins
+%by writing the integer $j$ on its tape,	
+%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+%	The
+%	transition system 
+%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+%	is a WSTS
+%	and
+%	has the property that
+%	it is
+%	$\mathop{\uparrow} (1,1)$-resilient
+%	iff TM$_j$ halts on input $j$,
+%	since $(1,1)$ is coverable from $(0,0)$
+%	iff TM$_j$ halts on input $j$. 
+%	Hence there is no Turing machine which correctly determines 
+%	resilience and halts whenever its input is a
+%	WSTS.
+%Hence resilience for WSTS in general is undecidable.
+
+\subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
+
+Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\mathop{\uparrow} \Bad$.
+It is of interest to note this case can be linked to the problem of mutual exclusion.
+Indeed the well-known mutual exclusion property can be modeled, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
+
+%
+\begin{theorem}\label{up-down}
+{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{theorem}
+
+\begin{proof}
+Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
+$\Safe=\mathop{\downarrow} \Safe$.
+By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
+The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
+Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
+ subsets.
+Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both upward-closed. We may decide whether this intersection is empty, hence
+the resilience problem is decidable. \qed
+\end{proof}
+
+In the case
+of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
+the above construction
+can provide a proof
+of non-resilience
+i.e. when
+$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
+then
+$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
+and hence
+$\Bad \not\subseteq \pred^*(\Safe)$.
+When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
+however
+it is not enough to conclude.\\
+
 Remark we did not make use of the

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@
 \begin{proof}
+First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
+$\Safe = Q_{M} \times \mathop{\downarrow} (1,1,0)$.
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} such that a sequence from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state $ q(0,  \ldots 0)$ if and only if
+there exists
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ the state $q_f(0,0,0)$
+Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+resilience hold in the reset-VASS simulating $M'$.
+Hence {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.
 \mathieu{preuve résumé / concise à écrire}
@@
 Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed
+$\omega^2$-WSTS with strong compatibility as well.
 %\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
 \fi
-\end{proof}
+\qed \end{proof}
 

### 2023-10-20 VMCAI-revised-version/final.tex
up

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
 \begin{proof}
-\mathieu{preuve résumé / concise à écrire}
+First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
+$\Safe = Q_{M} \times \mathop{\downarrow} (1,1,0)$,
+since it is possible to guarantee that state $q_f(0,0,0)$ is the only element of
+$Q_{M} \times \mathop{\downarrow} (1,1,0)$
+potentially reachable from 
+$Q_{M} \times \mathop{\uparrow} (0,0,1)$.
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
+% from  a state $p({\bf u})$
+% to
+reaching
+ a state 
+%
+$ q({\bf 0})$ corresponds to one that reaches 
+$q_f(0,0,0)$
+ in the Minsky machine and vice versa.
+ By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
+ we obtain that 
+ {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.
+%\mathieu{preuve résumé / concise à écrire}
 \iffalse
+\alain{preuve pas claire, à réécrire}
+% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
+%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
+%which can be simulated by WSTS.
+First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
+%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
+From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
+%			with at least $1$ on its third counter. 
+We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
+Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+Hence {\sc Resilience} is undecidable for Minsky machines.
+\alain{écrire qqchose ici}
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
+% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
+%Reset-VASS extend the basic VASS model with special “reset
+%transitions” that set to $0$ some coordinates in the vector. 
+%			Let us recall their definition here.
+%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
+The simulation operates by building a reset-VASS in which
+there exists
+ a sequence from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state $ q(0,  \ldots 0)$ if and only if
+there exists
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ the state $q_f(0,0,0)$
+  (from Theorem~$5$ in \cite{araki1976PN}).
+Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
+resilience hold in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
+\fi
+\qed \end{proof}
+
+\subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
+
+We start with the case $\Safe=\mathop{\uparrow} \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
+Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
+
+%
+\begin{theorem}\label{down-up}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
+
+\begin{proof}
+Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
+Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+%
+
+Let  $\{s_1,s_2,...,s_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
+Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+
+Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}}$.
+%
+Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
+From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
+Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
+%	for every $y \in J$, 
+there exists $x \in I$ and $y' \geq s_j$ such that $x \xrightarrow{k} y'  \geq s_j$ in $\mathscr{S}$ and then $s_j$ is coverable from $x$ in $\mathscr{S}$.
+
+Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\mathop{\downarrow} s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
+%
+Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
+Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
+%
+
+%
+We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
+
+This implies that resilience and bounded resilience can be reduced to coverability.
+Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+  resilience problem and the bounded resilience problem are decidable.
+
+Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
+Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
+If $\mathop{\downarrow} s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\mathop{\downarrow} s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
+
+Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.\qed
+\end{proof}
+
+\begin{theorem}\label{xxx}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+\begin{proof}
+The resilience problem can be reformulated as 
+$\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
+Since the wqo is ideally-effective, we can compute the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable. \qed\end{proof} 
+%%%%%%%
+%%\alain{prouver, si possible, les deux autres résilience}
+Strong compatibility implies furthermore the decidability
+of {\sc $k$-resilience} and {\sc bounded-resilience}.
+
+\begin{corollary}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{corollary}
+
+\begin{proof}
+For strongly compatible WSTS,
+$\pred^{k}(\Safe) = \mathop{\uparrow} \pred^{k}(\Safe)$,
+and hence
+$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$,
+when $\Safe = \mathop{\uparrow} \Safe$. 
+Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\mathop{\uparrow} \pred^{\leq k}(\Safe) = \pred^{\leq k}(\Safe)$.
+Like above, we can test if the intersection of $\Bad$
+and
+$S \setminus \pred^{\leq k}(\Safe)$ is empty or not, hence {\sc $k$-resilience} is decidable.
+To decide {\sc Bounded resilience}, we check {\sc $k$-resilience}
+starting with $k=0$ until we find some $k_0$ such that either {\sc $k_0$-resilience} holds,
+either $\pred^{\leq k_0}(\Safe) = \pred^{\leq k_0+1}(\Safe)$, whichever comes first.
+The convergence of $(\pred^{\leq k}(\Safe) )_{k\in\N}$ guarantees the latter eventually happens.
+\end{proof}
+
+Remark that the above proofs do not make use of the 
+property that $\Bad$ is the complement of $\Safe$, simply using
+$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
+%%%
+%%%
+%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
+%increasing recursive functions from \cite{FMP-ic04} defined as
+%
+%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
+%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
+%		
+%	\noindent	
+%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
+%	which moreover begins
+%by writing the integer $j$ on its tape,	
+%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
+%	The
+%	transition system 
+%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
+%	is a WSTS
+%	and
+%	has the property that
+%	it is
+%	$\mathop{\uparrow} (1,1)$-resilient
+%	iff TM$_j$ halts on input $j$,
+%	since $(1,1)$ is coverable from $(0,0)$
+%	iff TM$_j$ halts on input $j$. 
+%	Hence there is no Turing machine which correctly determines 
+%	resilience and halts whenever its input is a
+%	WSTS.
+%Hence resilience for WSTS in general is undecidable.
+
+\subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
+
+Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\mathop{\uparrow} \Bad$.
+It is of interest to note this case can be linked to the problem of mutual exclusion.
+Indeed the well-known mutual exclusion property can be modeled, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
+Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
+
+%
+\begin{theorem}\label{up-down}
+{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{theorem}
+
+\begin{proof}
+Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
+$\Safe=\mathop{\downarrow} \Safe$.
+By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
+The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
+%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
+Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
+ subsets.
+Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
+$\Bad$
+and
+$S \setminus \pred^*(\Safe)$,
+which are both upward-closed. We may decide whether this intersection is empty, hence
+the resilience problem is decidable. \qed
+\end{proof}
+
+In the case
+of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
+the above construction
+can provide a proof
+of non-resilience
+i.e. when
+$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
+then
+$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
+and hence
+$\Bad \not\subseteq \pred^*(\Safe)$.
+When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
+however
+it is not enough to conclude.\\
+
+Remark we did not make use of the
+property $\Bad$ complement of $\Safe$, simply 
+$\Bad=\mathop{\uparrow} \Bad$ and $\Safe=\mathop{\downarrow} \Safe$, thus 
+the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.\\
+
+\section{State-resilience}
+
+Resilience is a strong property that implies that from every element 
+ there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from 
+ $ \post^*(s_0)$, there must exist a path to $\Safe$. 
+The three previous problems become:
+
+\problemx{State-resilience (SRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
+{$\post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
+
+\problemx{$k$-state-resilience (kSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
+{ $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+\problemx{bounded-state-resilience (BSRP)}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
+{$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+Remark that {\sc State-resilience} is {\sc Home-space} with input set $\Safe$.
+Since these problems are undecidable for general infinite-state transition systems, we still restrict our study to WSTS.
+As in the Section~3, we study decidability results for $\Safe$ downward-closed and upward-closed.
+
+\subsection{Case: $\Safe = \mathop{\uparrow} \Safe$}
+
+We start with the case $\Safe = \mathop{\uparrow} \Safe$.
+Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
+This stems from the fact that it is undecidable in the particular case of reset-VASS,
+where  $t$-liveness is both undecidable and 
+ reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
+
+A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists two states
+ $p(\textbf{u}),p'(\textbf{u'})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} p'(\textbf{u'}) $. 
+%		($t$ is enabled in $p(\textbf{u})$). 
+We say that the whole reset-VASS is live if all its control-transitions are
+live. This leads to the following problem.
+
+\problemx{$t$-liveness}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, a transition $t \in T$, an initial state $s_0 \in Q \times \N^d$}
+{Is $t$ live in $s_0$ ? \\}
+
+Let us define the set of states, $\pre(t)$, from which a  control-transition $t$ is enabled:
+$\pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v}) \in Q \times \N^d$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $\pre(t)$ is upward-closed.
+
+\begin{proposition}\label{reductions}
+{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
+\end{proposition}
+
+\begin{proof}
+We reformulate $t$-liveness 
+in a 
+reset-VASS $(Q,T)$ 
+with initial state $s_0$
+as the following formula
+\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
+~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in \pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
+The previous formula reduces itself to {\sc State-resilience}
+where $\Safe = \pre(t)$.
+\end{proof}
+
+It now remains to argue why {\sc $t$-liveness} is undecidable.
+We  recall the
+ undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
+reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
+deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+%
+Remark that the quantification over the path is existential rather than universal i.e. 
+{\em zero-reachability}
+does not asks that zero remains reachable from an initial state $s_0$ at all times but rather, that there exists at least one way to reach it.
+Despite this it is possible to reduce 
+the {\em zero-reachability problem} 
+to
+{\sc $t$-liveness}.
+The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets, whose proof can be seen in Appendix~\ref{appendix}%
+.
+\begin{proposition}\label{liveness reset}
+The  zero reachability problem can be reduced to {\sc $t$-liveness}.
+\end{proposition}
+
+We may now prove the undecidability of the three state-resilience problems.
+
+\begin{proposition}\label{srp up down}
+{\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for strongly compatible $\pred$-basis-effective WSTS
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+\end{proposition}
+
+\begin{proof}
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
+{\sc $t$-liveness} is undecidable.
+Because
+{\sc $t$-liveness} is undecidable in reset-VASS
+and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
+Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
+reducible to {\sc $k$-state-resilience}:
+since $\Safe=\mathop{\uparrow} \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
+compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
+We compute 
+$n_0$, then iteratively check whether $k$-state-resilience 
+hold for $k$ from $0$ to $n_0$.  
+Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+since 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
+Hence the undecidability of {\sc Bounded-state-resilience}
+and 
+ {\sc $k$-state-resilience}. \qed
+\end{proof}
+
+On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
+%
+\begin{theorem}\label{ref ozkan}
+(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+\end{theorem}
+
+We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
+
+\begin{proposition}\label{postcomputable}
+{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+\end{proposition}
+
+\begin{proof}
+Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
+$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
+$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
+$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
+returning $n$ if that is the case.
+Then, because {\sc $n_0$-state-resilience} is decidable, 
+checking $\mathop{\uparrow} \post^*(s) 
+ \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+and {\sc State-resilience} is decidable. \qed
+\end{proof}
+
+The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
+
+\begin{lemma}\label{Lemma intersection}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
+\end{lemma}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
+Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
+Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
+$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
+\end{proof}
+
+\"Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
+
+\problemx{Downward-reachability problem}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
+\subseteq S$.}
+{$s  \to^* D$? \newline}
+
+\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
+For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+\end{proposition}
+
+The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
+$B_{\mathop{\uparrow} \post^*(s)} \cap D = \emptyset$, 
+where $B_{\mathop{\uparrow} \post^*(s)}$ is a basis of $\mathop{\uparrow} \post^*(s)$,
+ that is equivalent to $\post^*(s)\cap D = \emptyset$ by
+Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
+$U_n = \mathop{\uparrow} \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
+until it becomes stationnary. 
+Decidability of downward-reachability leads to the decidability of the following stop condition:
+asking whether $S \setminus U_n$ is reachable from $s$.
+
+For instance,
+VASS are $ \post^*$-basis-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
+It is well-known that 
+VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\mathop{\uparrow} \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
+Hence {\sc State-resilience} is decidable for  
+VASS.
+
+However, the hypothesis that $\mathop{\uparrow} \post^*$ is computable cannot be tested in the general WSTS framework.  
+Moreover there exist classes of WSTSs with strong 
+ compatibility for which there doesn't exist an algorithm computing a basis of $\mathop{\uparrow} \post^*$, such as reset-VASS.
+
+Keeping the $\mathop{\uparrow} \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
+
+\begin{theorem}\label{post srp}
+{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
+when
+$\Safe=\mathop{\uparrow} \Safe$. 
+\end{theorem}
+
+\begin{proof}
+Let $B_{\mathop{\uparrow} \post^*(s)}$ 
+ be a basis of $\mathop{\uparrow} \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
+By applying Lemma~\ref{Lemma intersection} twice, we obtain
+\[ \post^*(s) 
+ \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
+ \subseteq \pred^*(\Safe)\]
+Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite and we can compute 
+%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
+a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\mathop{\uparrow} \post^*(s)} 
+ \setminus \pred^*(\Safe) = \emptyset$.  \qed
+\end{proof}
+
+However when removing strong compatibility, some precision is lost.
+Since $\pred(\mathop{\uparrow} \Safe)$ is not necessarily upward-closed, it is possible to have 
+ $\mathop{\uparrow} \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
+despite having 
+$\post^* (s) \cap S \subseteq \pred( \Safe)$.
+In such a case, the algorithm in
+\cite{DBLP:conf/gg/Ozkan22} would deduce that {\sc $1$-state-resilience} does not hold,
+which is incorrect.
+
+Thus in case of $\post^*$-basis-effective WSTS (when the compatibility is not strong), we don't know the decidability status of {\sc $k$-state-resilience} and 
+{\sc Bounded-state-resilience}.
+
+\subsection{Case: $\Safe = \mathop{\downarrow} \Safe$}
+
+We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
+%
+Unfortunately {\sc State-resilience} is undecidable in this case.
+As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+Section~\ref{safe-down}.\alain{trop rapide}
+
+\begin{proposition}\label{srp down up}
+{\sc State-resilience} is undecidable for 
+%$\post$-ideal-effective ???? and
+ $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
+when
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{proposition}
+
+Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \mathop{\downarrow} \Safe$
+is symmetrical to the case $\Safe = \mathop{\uparrow} \Safe$.
+%
+% For instance one can write the following lemma:
+%  Page 14: "For instance one can write the following lemma" => Rewrite this: we will need the following lemma for this and that.
+We will need the following lemma in order to obtain symmetrical results.
+
+\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
+Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap U \subseteq D$  iff $ (\mathop{\downarrow}  A) \cap U \subseteq D$.
+\end{lemma}
+
+In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
+then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
+Lemma~\ref{Lemma intersection 2} can be used to show that
+if $\Safe = \mathop{\downarrow} \Safe$,
+then
+$\post^*(s)  
+\subseteq \pred^*(\Safe)$  iff $ (\mathop{\downarrow}  \post^*(s)) 
+\subseteq \pred^*(\Safe)$.
+
+\begin{theorem}\label{downward srp}
+{\sc State-resilience} is decidable for $\post$-effective, 
+%$\pred$-basis-effective, 
+$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
+and $\Safe = \mathop{\downarrow} \Safe$.
+\end{theorem}
+
+\begin{proof}
+Since $\mathscr{S}$ is $\pred^*$-basis-effective, $\Safe = \mathop{\downarrow} \Safe$ and $\pred^*(\Safe)=\mathop{\downarrow} \pred^*(\Safe)$, we may compute the finite decomposition in ideals of $\pred^*(\Safe)$.
+
+In order to decide whether %$\post^{\leq n}(s)
+$\post^{ *}(s)  
+ \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
+one looking for a resilience certificate and one looking for a non-resilience certificate.
+
+Procedure 1 enumerates every downward-closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
+The computability of the enumeration comes from the hypothesis that $\leq$ is ideally-effective (Lemma~4.3.
+from \cite{DBLP:journals/lmcs/BlondinFM17}).
+The procedure then checks for every downward-closed subset $D_i$ whether $\mathop{\downarrow} \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
+%\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
+L'hypothèse "$\post$-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
+Il manquait la $\post$-effective dans les hypothèses}
+Every subset $D_i$ such that $\mathop{\downarrow} \post(D_i) \subseteq D_i$ is an “over-approximation” of $\mathop{\downarrow} \post^*(s)$ if it contains $s$.
+Notice that, by 
+upward compatibility, $\mathop{\downarrow} \post^*(s)$ is such a subset and may eventually be found.
+
+Procedure 1 stops when it finds a downward-closed subset $D$ such that
+$\mathop{\downarrow} \post(D) \subseteq D$,
+$s \in D$ and
+$D   \subseteq \pred^*(\Safe)$. 
+%\mathieu{ici $\pred^*(\Safe)$ est downward-closed.
+du coup pas sur que la $\pred$-basis-effectivité telle qu'elle est actuellement nous permette de trancher ...
+à moins que les hypothèses d'effectivté sur les idéaux nous permettent de guarantir qu'on puisse calculer le $\pred$ d'un idéal ?}
+
+Indeed
+$D \subseteq  \pred^*(\Safe)$ implies
+$\mathop{\downarrow} \post^*(s)  \subseteq  \pred^*(\Safe)$
+since $ \mathop{\downarrow} \post^*(s)  \subseteq D$.
+
+The second procedure iteratively computes
+$\post^{\leq n}(s) $ (this is effective because $\mathscr{S}$ is $\post$-effective)
+%			\alain{ici on utilise $\post$-effective que j'ai rajoutée}
+until it finds an element
+not in $\pred^*(\Safe)$.
+
+If resilience hold, then procedure 1 terminates since it eventually finds
+ $D = \mathop{\downarrow} \post^*(s)$ 
+ such that  $D \subseteq  \pred^*(\Safe)$.
+ If resilience does not hold, then procedure 2 terminates since
+there exists a witness that resilience does not hold which can eventually be found. \qed
+\end{proof}
+
+In the case of strong downward compatibility,
+$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is.
+Hence, 
+assuming a $\pred^*$-basis-effective variation
+consisting in an algorithm
+accepting any ideal $I$ and returning a finite basis of
+ $\mathop{\downarrow} \pred(I)$, it becomes possible to perform the same procedures as above except with comparisons against $\pred^{\leq k}(\Safe)$ rather than $\pred^*(\Safe)$, and hence {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable in this case.
+
+\begin{corollary}\label{downward brp}
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
+$\Safe = \mathop{\downarrow} \Safe$
+and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
+ $\mathop{\downarrow} \pred(I)$.
+\end{corollary}
+
+\mathieu{Donner des exemples dans cette classe.}
+\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
+
+\section{Resilience for VASS and variations}\label{section VASS}
+
+In this section we study VASS.
+Since they enjoy many properties of effectivity ($\leq$ is ideally-effective, VASS are $\post$-effective, $\post$-ideal-effective, $\pred$-basis-effective, $\post^*$-basis-effective), they inherit the decidability results for WSTS in the case
+$\Safe = \mathop{\uparrow} \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
+for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
+VASS do not inherit the decidability results for WSTS in the case $\Safe = \mathop{\downarrow} \Safe$.
+In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed and $\Bad = S \setminus \Safe$.
+We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
+
+%Integer VASS or 
+$\mathbb{Z}^d-$VASS (resp. $\mathbb{Z}-$VASS)~\cite{DBLP:conf/rp/HaaseH14} are $d$-VASS (resp. VASS) that are allowed to take values from the integers.
+
+\begin{theorem}\label{SL VASS}
+{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We consider the case where $\Safe$ is semilinear
+and $\Bad = S \setminus \Safe$ is semilinear too.
+We want to use the fact that, for VASS,
+it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
+$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
+{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
+We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
+Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
+\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
+and hence, {\sc Resilience} is decidable. 
+The reachability relation of 
+%integer $d$-VASS
+ $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\qed
+\end{proof}
+
+{\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
+Recall that lossy counter machines (LCM)~\cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
+
+Resilience and the home-space problem are also linked to the 
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+is decidable for lossy counter machine
+when $X$ and $Y$ are semi-linear sets.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
+
+\begin{theorem}\label{SL Lossy}
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
+Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
+\qed
+\end{proof}
+
+$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+
+\begin{theorem}\label{RP VASS}
+{\sc Resilience} is decidable for continuous VASS when $\Safe$  is
+definable in the existential theory of the rationals with addition and order.
+\end{theorem}
+
+\begin{proof}
+The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). \qed
+\end{proof}
+
+Remark that
+Theorem~\ref{SL Lossy} 
+(resp. Theorem~\ref{RP VASS})
+did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
+semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
+the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+The above results concern only unbounded resilience. We consider now 
+{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
+the path from $\Bad$ to $\Safe$
+when $\Safe$ is downward-closed. Surprisingly, we have the following result for VAS (not VASS):
+\begin{proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
+\end{proposition}
+
+\begin{proof}
+Consider a given $k \in \N$,
+ $\emptyset \neq \Bad \subseteq \N^d$ is upward-closed and $\Safe \subseteq \N^d$ is downward-closed, and
+consider a given VAS $V$.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
+The set $\Bad \neq \emptyset$ admits a finite basis $B_\Bad$.
+Consider the vector $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider the vector
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+
+All states reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
+and thus, are in $\Bad$, by upward-closedness.
+Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
+Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
+\qed
+\end{proof}
+
+This all changes when one considers VASS (having control-states).
+For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\mathop{\uparrow} q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
+ $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
+Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
+This lead to a decision procedure
+ which lead to the following decidability result:
+\begin{theorem}
+{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+$\Safe = \mathop{\downarrow} \Safe$.
+\end{theorem}
+
+\begin{proof}
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider the vector
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\mathop{\uparrow} q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+ $\mathop{\uparrow} q_{j_1}(\textbf{0}), \mathop{\uparrow} q_{j_2}(\textbf{0}), \ldots \mathop{\uparrow} q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\mathop{\uparrow} q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If it is infinite then
+$k$-resilience does not hold, 
+hence we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
+
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards,
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+\qed
+\end{proof}
+
+\section{Conclusion and perspectives}
+
+We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
+We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
+ $\Safe = \mathop{\uparrow} \Safe$ and in the case 
+ $\Safe = \mathop{\downarrow} \Safe$.
+
+Several questions still remain.
+For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
+One other angle of attack not studied here is the synthesis of sets 
+$\Safe$ for resilience,
+for instance, 
+ finding the maximal upward-closed (resp. downward-closed) subset 
+$\Safe$ so that a system is $\Safe$-resilient. 
+Another question could be to analyse the resilience in the framework of a controller and its environment. 
+One could also extend upon the classes of set $\Safe$ 
+ considered.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets~\cite{DBLP:journals/fmsd/BertrandS13}.
+Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
+
+\alain{remercier les reviewers !}
+
+\bibliography{bib}
+
+\appendix
+
+\section{Reset-VASS and Minsky-machines}\label{minsk}
+
+\begin{proposition}\label{indec WSTS with dcs}
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{proposition}
+
+\begin{proof}
 \alain{preuve pas claire, à réécrire}
@@
 %\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
-\fi
-\end{proof}
-
-\subsection{Case: $\Safe=\mathop{\uparrow} \Safe$.}\label{safe-up}
-
-We start with the case $\Safe=\mathop{\uparrow} \Safe$, hence $\Bad=\mathop{\downarrow} \Bad$.
-Since Resilience is undecidable for  $\omega^2$-WSTS with strong compatibility (Theorem \ref{indec WSTS}), we still consider WSTS, and even WBTS, with strong compatibility but by strengthening the assumptions of effectiveness (we now consider the $\post$-ideal-effective hypothesis) and we demonstrate that the three resilience problems are now decidable.
-
-%
-\begin{theorem}\label{down-up}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
-{\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{theorem}
-
-%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
-
-\begin{proof}
-Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
-It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
-%
-
-Let  $\{s_1,s_2,...,s_m\}$ be the (unique) minimal basis of $\Safe$ and  $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$.
-The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
-Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
-
-Let us prove that $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}}$.
-%
-Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
-From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
-Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
-%	for every $y \in J$, 
-there exists $x \in I$ and $y' \geq s_j$ such that $x \xrightarrow{k} y'  \geq s_j$ in $\mathscr{S}$ and then $s_j$ is coverable from $x$ in $\mathscr{S}$.
-
-Hence we obtain: $\mathscr{S}$ is resilient iff for all $i=1,..,n$ and $j= 1,..m$, $\mathop{\downarrow} s_j$ is coverable from ideal $J_i$ in $\hat{\mathscr{S}}$.
-%
-Let us denote by $k_{i,j}$ the length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ and let $k_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
-Let us now define $K_{\mathscr{S}}(\Safe)=\max(k_{i,j} \mid i=1,..,n$ and $j= 1,..m$).
-%
-
-%
-We now have $\mathscr{S}$ is resilient iff $K_{\mathscr{S}}(\Safe)$ is finite iff $\mathscr{S}$ is $K_{\mathscr{S}}(\Safe)$-resilient with $K_{\mathscr{S}}(\Safe)$ finite.
-
-This implies that resilience and bounded resilience can be reduced to coverability.
-Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
-  resilience problem and the bounded resilience problem are decidable.
-
-Let us now show that the $k$-resilience problem, with $k \in \mathbb{N}$, is also decidable.
-Let us denote by $k'_{i,j}$ the \emph{minimal} length of a covering sequence that covers $\mathop{\downarrow} s_j$ from $J_i$ in $\hat{\mathscr{S}}$ if it exists and let $k'_{i,j}\stackrel{\text{def}}{=}\infty$ if $\mathop{\downarrow} s_j$ is not coverable from $J_i$.
-If $\mathop{\downarrow} s_j$ is coverable from $J_i$, we first compute an $k_{i,j}$, and then we compute $k'_{i,j}$ by iteratively checking whether there exists a sequence of length $0,1,...,k_{i,j}-1$ that covers $\mathop{\downarrow} s_j$ from $J_i$ until we find the minimal one which is necessarly smaller (or equal to) than $k_{i,j}$.
-
-Let us now define $K'_{\mathscr{S}}(\Safe)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe)$.\qed
-\end{proof}
-
-\begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
-{\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{theorem}
-
-\begin{proof}
-The resilience problem can be reformulated as 
-$\Bad \subseteq  \pred^*(\Safe)$ that is equivalent to $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-The set $\pred^*(\Safe)$ is upward-closed, since $\Safe$ is upward-closed. Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\pred^*(\Safe)$.
-Since the wqo is ideally-effective, we can compute the intersection of
-$\Bad$
-and
-$S \setminus \pred^*(\Safe)$,
-which are both downward-closed, and test if this intersection is empty or not. Hence, the resilience problem is decidable. \qed\end{proof} 
-%%%%%%%
-%%\alain{prouver, si possible, les deux autres résilience}
-Strong compatibility implies furthermore the decidability
-of {\sc $k$-resilience} and {\sc bounded-resilience}.
-
-\begin{corollary}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
-\end{corollary}
-
-\begin{proof}
-For strongly compatible WSTS,
-$\pred^{k}(\Safe) = \mathop{\uparrow} \pred^{k}(\Safe)$,
-and hence
-$\pred^{\leq k}(\Safe) = \mathop{\uparrow} \pred^{\leq k}(\Safe)$,
-when $\Safe = \mathop{\uparrow} \Safe$. 
-Since $\mathscr{S}=(S,\rightarrow, \leq)$ is $\pred$-basis-effective, we can compute a basis of $\mathop{\uparrow} \pred^{\leq k}(\Safe) = \pred^{\leq k}(\Safe)$.
-Like above, we can test if the intersection of $\Bad$
-and
-$S \setminus \pred^{\leq k}(\Safe)$ is empty or not, hence {\sc $k$-resilience} is decidable.
-To decide {\sc Bounded resilience}, we check {\sc $k$-resilience}
-starting with $k=0$ until we find some $k_0$ such that either {\sc $k_0$-resilience} holds,
-either $\pred^{\leq k_0}(\Safe) = \pred^{\leq k_0+1}(\Safe)$, whichever comes first.
-The convergence of $(\pred^{\leq k}(\Safe) )_{k\in\N}$ guarantees the latter eventually happens.
-\end{proof}
-
-Remark that the above proofs do not make use of the 
-property that $\Bad$ is the complement of $\Safe$, simply using
-$\Bad=\mathop{\downarrow} \Bad$ and $\Safe=\mathop{\uparrow} \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
-
-%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
-%%%
-%%%
-%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
-%increasing recursive functions from \cite{FMP-ic04} defined as
-%
-%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
-%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
-%		
-%	\noindent	
-%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
-%	which moreover begins
-%by writing the integer $j$ on its tape,	
-%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
-%	The
-%	transition system 
-%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-%	is a WSTS
-%	and
-%	has the property that
-%	it is
-%	$\mathop{\uparrow} (1,1)$-resilient
-%	iff TM$_j$ halts on input $j$,
-%	since $(1,1)$ is coverable from $(0,0)$
-%	iff TM$_j$ halts on input $j$. 
-%	Hence there is no Turing machine which correctly determines 
-%	resilience and halts whenever its input is a
-%	WSTS.
-%Hence resilience for WSTS in general is undecidable.
-
-\subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
-
-Let us now consider the case $\Safe=\mathop{\downarrow} \Safe$ hence $\Bad=\mathop{\uparrow} \Bad$.
-It is of interest to note this case can be linked to the problem of mutual exclusion.
-Indeed the well-known mutual exclusion property can be modeled, in a $d$-VASS with $d$ counters, by the property that a special counter $c_{mutex}$ must be bounded by $k \geq 1$ which counts the (maximal) number of processes that are allowed to be simultaneously in the critical section.
-Then, the set $\Safe =  \{c_{mutex} \leq k\} \times \mathbb{N}^{d-1}$ is downward-closed		and $\Bad =\{c_{mutex} \geq k+1\} \times  \mathbb{N}^{d-1} $ is the upward-closed complementary of $\Safe$.
-
-%
-\begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\mathop{\downarrow} \Safe$.
-\end{theorem}
-
-\begin{proof}
-Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
-$\Safe=\mathop{\downarrow} \Safe$.
-By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
-The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
-Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
- subsets.
-Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
-$\Bad$
-and
-$S \setminus \pred^*(\Safe)$,
-which are both upward-closed. We may decide whether this intersection is empty, hence
-the resilience problem is decidable. \qed
-\end{proof}
-
-In the case
-of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
-the above construction
-can provide a proof
-of non-resilience
-i.e. when
-$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
-then
-$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
-and hence
-$\Bad \not\subseteq \pred^*(\Safe)$.
-When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
-however
-it is not enough to conclude.\\
-
-Remark we did not make use of the
-property $\Bad$ complement of $\Safe$, simply 
-$\Bad=\mathop{\uparrow} \Bad$ and $\Safe=\mathop{\downarrow} \Safe$, thus 
-the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.\\
-
-\section{State-resilience}
-
-Resilience is a strong property that implies that from every element 
- there must exist a path to $\Safe$. However, when one considers a system with an initial state $s_0$, it could be sufficient to only ask that from 
- $ \post^*(s_0)$, there must exist a path to $\Safe$. 
-The three previous problems become:
-
-\problemx{State-resilience (SRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
-{$\post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
-
-\problemx{$k$-state-resilience (kSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
-{ $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-
-\problemx{bounded-state-resilience (BSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
-{$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-
-Remark that {\sc State-resilience} is {\sc Home-space} with input set $\Safe$.
-Since these problems are undecidable for general infinite-state transition systems, we still restrict our study to WSTS.
-As in the Section~3, we study decidability results for $\Safe$ downward-closed and upward-closed.
-
-\subsection{Case: $\Safe = \mathop{\uparrow} \Safe$}
-
-We start with the case $\Safe = \mathop{\uparrow} \Safe$.
-Unfortunately, in this case {\sc State-resilience} is undecidable for (general) WSTS even with strong upward-compatibility.
-This stems from the fact that it is undecidable in the particular case of reset-VASS,
-where  $t$-liveness is both undecidable and 
- reducible to {\sc State-resilience}. This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.
-
-A control-transition $t$ of a reset-VASS is {\em live} in a state $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists two states
- $p(\textbf{u}),p'(\textbf{u'})$ such that $q(\textbf{v}) \to^* p(\textbf{u}) \xrightarrow{t} p'(\textbf{u'}) $. 
-%		($t$ is enabled in $p(\textbf{u})$). 
-We say that the whole reset-VASS is live if all its control-transitions are
-live. This leads to the following problem.
-
-\problemx{$t$-liveness}
-{A reset-VASS $V=(Q,T)$ of dimension $d$, a transition $t \in T$, an initial state $s_0 \in Q \times \N^d$}
-{Is $t$ live in $s_0$ ? \\}
-
-Let us define the set of states, $\pre(t)$, from which a  control-transition $t$ is enabled:
-$\pre(t)=\{ p(\textbf{u}) \in Q \times \N^d \mid \exists q(\textbf{v}) \in Q \times \N^d$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$. Remark that $\pre(t)$ is upward-closed.
-
-\begin{proposition}\label{reductions}
-{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
-\end{proposition}
-
-\begin{proof}
-We reformulate $t$-liveness 
-in a 
-reset-VASS $(Q,T)$ 
-with initial state $s_0$
-as the following formula
-\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
-~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in \pre(t), ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]  
-The previous formula reduces itself to {\sc State-resilience}
-where $\Safe = \pre(t)$.
-\end{proof}
-
-It now remains to argue why {\sc $t$-liveness} is undecidable.
-We  recall the
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
-%
-Remark that the quantification over the path is existential rather than universal i.e. 
-{\em zero-reachability}
-does not asks that zero remains reachable from an initial state $s_0$ at all times but rather, that there exists at least one way to reach it.
-Despite this it is possible to reduce 
-the {\em zero-reachability problem} 
-to
-{\sc $t$-liveness}.
-The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets, whose proof can be seen in Appendix~\ref{appendix}%
-.
-\begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to {\sc $t$-liveness}.
-\end{proposition}
-
-We may now prove the undecidability of the three state-resilience problems.
-
-\begin{proposition}\label{srp up down}
-{\sc State-resilience},
-{\sc Bounded-state-resilience} and
-{\sc $k$-state-resilience}
-are undecidable for strongly compatible $\pred$-basis-effective WSTS
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-\end{proposition}
-
-\begin{proof}
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
-{\sc $t$-liveness} is undecidable.
-Because
-{\sc $t$-liveness} is undecidable in reset-VASS
-and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
-Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
-reducible to {\sc $k$-state-resilience}:
-since $\Safe=\mathop{\uparrow} \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
-compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
-We compute 
-$n_0$, then iteratively check whether $k$-state-resilience 
-hold for $k$ from $0$ to $n_0$.  
-Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
-since 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
-Hence the undecidability of {\sc Bounded-state-resilience}
-and 
- {\sc $k$-state-resilience}. \qed
-\end{proof}
-
-On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
-%
-\begin{theorem}\label{ref ozkan}
-(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
-{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-\end{theorem}
-
-We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
-
-\begin{proposition}\label{postcomputable}
-{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-\end{proposition}
-
-\begin{proof}
-Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
-$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
-$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
-$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
-returning $n$ if that is the case.
-Then, because {\sc $n_0$-state-resilience} is decidable, 
-checking $\mathop{\uparrow} \post^*(s) 
- \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-and {\sc State-resilience} is decidable. \qed
-\end{proof}
-
-The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
-
-\begin{lemma}\label{Lemma intersection}
-Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
-\end{lemma}
-
-\begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
-Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
-Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
-$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
-\end{proof}
-
-\"Ozkan \cite{DBLP:conf/gg/Ozkan22} argues that it is precisely the WSTS for which the following problem is decidable.
-
-\problemx{Downward-reachability problem}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and a downward-closed set $D
-\subseteq S$.}
-{$s  \to^* D$? \newline}
-
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-\end{proposition}
-
-The idea behind the proof is the following. For deciding whether a downward-closed set $D$ is reachable from $s$, one checks whether
-$B_{\mathop{\uparrow} \post^*(s)} \cap D = \emptyset$, 
-where $B_{\mathop{\uparrow} \post^*(s)}$ is a basis of $\mathop{\uparrow} \post^*(s)$,
- that is equivalent to $\post^*(s)\cap D = \emptyset$ by
-Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \mathop{\uparrow} \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
-until it becomes stationnary. 
-Decidability of downward-reachability leads to the decidability of the following stop condition:
-asking whether $S \setminus U_n$ is reachable from $s$.
-
-For instance,
-VASS are $ \post^*$-basis-effective WSTS \cite{DBLP:journals/corr/abs-2108-00889}. 
-It is well-known that 
-VASS are WSTS with strong compatibility and since there is an algorithm that computes a finite basis of  $\mathop{\uparrow} \post^*(s)$, \cite{DBLP:conf/gg/Ozkan22} deduced that {\sc Bounded-state-resilience} is decidable for VASS.
-Hence {\sc State-resilience} is decidable for  
-VASS.
-
-However, the hypothesis that $\mathop{\uparrow} \post^*$ is computable cannot be tested in the general WSTS framework.  
-Moreover there exist classes of WSTSs with strong 
- compatibility for which there doesn't exist an algorithm computing a basis of $\mathop{\uparrow} \post^*$, such as reset-VASS.
-
-Keeping the $\mathop{\uparrow} \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
-Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
-
-\begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
-when
-$\Safe=\mathop{\uparrow} \Safe$. 
-\end{theorem}
-
-\begin{proof}
-Let $B_{\mathop{\uparrow} \post^*(s)}$ 
- be a basis of $\mathop{\uparrow} \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
-By applying Lemma~\ref{Lemma intersection} twice, we obtain
-\[ \post^*(s) 
- \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
- \subseteq \pred^*(\Safe)\]
-Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite and we can compute 
-%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
-a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\mathop{\uparrow} \post^*(s)} 
- \setminus \pred^*(\Safe) = \emptyset$.  \qed
-\end{proof}
-
-However when removing strong compatibility, some precision is lost.
-Since $\pred(\mathop{\uparrow} \Safe)$ is not necessarily upward-closed, it is possible to have 
- $\mathop{\uparrow} \post^* (s) \cap S \not\subseteq \pred(\Safe)$,
-despite having 
-$\post^* (s) \cap S \subseteq \pred( \Safe)$.
-In such a case, the algorithm in
-\cite{DBLP:conf/gg/Ozkan22} would deduce that {\sc $1$-state-resilience} does not hold,
-which is incorrect.
-
-Thus in case of $\post^*$-basis-effective WSTS (when the compatibility is not strong), we don't know the decidability status of {\sc $k$-state-resilience} and 
-{\sc Bounded-state-resilience}.
-
-\subsection{Case: $\Safe = \mathop{\downarrow} \Safe$}
-
-We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
-%
-Unfortunately {\sc State-resilience} is undecidable in this case.
-As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-Section~\ref{safe-down}.\alain{trop rapide}
-
-\begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for 
-%$\post$-ideal-effective ???? and
- $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
-when
-$\Safe=\mathop{\downarrow} \Safe$.
-\end{proposition}
-
-Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \mathop{\downarrow} \Safe$
-is symmetrical to the case $\Safe = \mathop{\uparrow} \Safe$.
-%
-% For instance one can write the following lemma:
-%  Page 14: "For instance one can write the following lemma" => Rewrite this: we will need the following lemma for this and that.
-We will need the following lemma in order to obtain symmetrical results.
-
-\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
-Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap U \subseteq D$  iff $ (\mathop{\downarrow}  A) \cap U \subseteq D$.
-\end{lemma}
-
-In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
-then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
-Lemma~\ref{Lemma intersection 2} can be used to show that
-if $\Safe = \mathop{\downarrow} \Safe$,
-then
-$\post^*(s)  
-\subseteq \pred^*(\Safe)$  iff $ (\mathop{\downarrow}  \post^*(s)) 
-\subseteq \pred^*(\Safe)$.
-
-\begin{theorem}\label{downward srp}
-{\sc State-resilience} is decidable for $\post$-effective, 
-%$\pred$-basis-effective, 
-$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
-and $\Safe = \mathop{\downarrow} \Safe$.
-\end{theorem}
-
-\begin{proof}
-Since $\mathscr{S}$ is $\pred^*$-basis-effective, $\Safe = \mathop{\downarrow} \Safe$ and $\pred^*(\Safe)=\mathop{\downarrow} \pred^*(\Safe)$, we may compute the finite decomposition in ideals of $\pred^*(\Safe)$.
-
-In order to decide whether %$\post^{\leq n}(s)
-$\post^{ *}(s)  
- \subseteq \pred^*(\Safe)$, we execute two procedures in parallel,
-one looking for a resilience certificate and one looking for a non-resilience certificate.
-
-Procedure 1 enumerates every downward-closed subsets in some fixed order $D_1$ , $D_2$ , $\ldots$ by their ideal decomposition.
-The computability of the enumeration comes from the hypothesis that $\leq$ is ideally-effective (Lemma~4.3.
-from \cite{DBLP:journals/lmcs/BlondinFM17}).
-The procedure then checks for every downward-closed subset $D_i$ whether $\mathop{\downarrow} \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
-%\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
-L'hypothèse "$\post$-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
-Il manquait la $\post$-effective dans les hypothèses}
-Every subset $D_i$ such that $\mathop{\downarrow} \post(D_i) \subseteq D_i$ is an “over-approximation” of $\mathop{\downarrow} \post^*(s)$ if it contains $s$.
-Notice that, by 
-upward compatibility, $\mathop{\downarrow} \post^*(s)$ is such a subset and may eventually be found.
-
-Procedure 1 stops when it finds a downward-closed subset $D$ such that
-$\mathop{\downarrow} \post(D) \subseteq D$,
-$s \in D$ and
-$D   \subseteq \pred^*(\Safe)$. 
-%\mathieu{ici $\pred^*(\Safe)$ est downward-closed.
-du coup pas sur que la $\pred$-basis-effectivité telle qu'elle est actuellement nous permette de trancher ...
-à moins que les hypothèses d'effectivté sur les idéaux nous permettent de guarantir qu'on puisse calculer le $\pred$ d'un idéal ?}
-
-Indeed
-$D \subseteq  \pred^*(\Safe)$ implies
-$\mathop{\downarrow} \post^*(s)  \subseteq  \pred^*(\Safe)$
-since $ \mathop{\downarrow} \post^*(s)  \subseteq D$.
-
-The second procedure iteratively computes
-$\post^{\leq n}(s) $ (this is effective because $\mathscr{S}$ is $\post$-effective)
-%			\alain{ici on utilise $\post$-effective que j'ai rajoutée}
-until it finds an element
-not in $\pred^*(\Safe)$.
-
-If resilience hold, then procedure 1 terminates since it eventually finds
- $D = \mathop{\downarrow} \post^*(s)$ 
- such that  $D \subseteq  \pred^*(\Safe)$.
- If resilience does not hold, then procedure 2 terminates since
-there exists a witness that resilience does not hold which can eventually be found. \qed
-\end{proof}
-
-In the case of strong downward compatibility,
-$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is.
-Hence, 
-assuming a $\pred^*$-basis-effective variation
-consisting in an algorithm
-accepting any ideal $I$ and returning a finite basis of
- $\mathop{\downarrow} \pred(I)$, it becomes possible to perform the same procedures as above except with comparisons against $\pred^{\leq k}(\Safe)$ rather than $\pred^*(\Safe)$, and hence {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable in this case.
-
-\begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
-$\Safe = \mathop{\downarrow} \Safe$
-and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
- $\mathop{\downarrow} \pred(I)$.
-\end{corollary}
-
-\mathieu{Donner des exemples dans cette classe.}
-\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
-
-\section{Resilience for VASS and variations}\label{section VASS}
-
-In this section we study VASS.
-Since they enjoy many properties of effectivity ($\leq$ is ideally-effective, VASS are $\post$-effective, $\post$-ideal-effective, $\pred$-basis-effective, $\post^*$-basis-effective), they inherit the decidability results for WSTS in the case
-$\Safe = \mathop{\uparrow} \Safe$. Lacking downward compatibility or a more relaxed hypothesis that
-for all downward-closed set $D$, the set $\pred^*(D)$ is downward-closed, 
-VASS do not inherit the decidability results for WSTS in the case $\Safe = \mathop{\downarrow} \Safe$.
-In this section, we work to re-establish decidability results for VASS when $\Safe$ is downward-closed and $\Bad = S \setminus \Safe$.
-We also extend decidability to resilience for semilinear sets rather than simply upward and downward-closed ones.
-
-%Integer VASS or 
-$\mathbb{Z}^d-$VASS (resp. $\mathbb{Z}-$VASS)~\cite{DBLP:conf/rp/HaaseH14} are $d$-VASS (resp. VASS) that are allowed to take values from the integers.
-
-\begin{theorem}\label{SL VASS}
-{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
-\end{theorem}
-
-\begin{proof}
-We consider the case where $\Safe$ is semilinear
-and $\Bad = S \setminus \Safe$ is semilinear too.
-We want to use the fact that, for VASS,
-it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
-$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
-{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
-We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
-Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
-\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
-and hence, {\sc Resilience} is decidable. 
-The reachability relation of 
-%integer $d$-VASS
- $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
-\qed
-\end{proof}
-
-{\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
-Recall that lossy counter machines (LCM)~\cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
-
-Resilience and the home-space problem are also linked to the 
-model-checking of basic reachability and safety formulae. 
-In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-is decidable for lossy counter machine
-when $X$ and $Y$ are semi-linear sets.
-Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
- t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
-whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
-``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
-  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
-
-\begin{theorem}\label{SL Lossy}
-{\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
-\end{theorem}
-
-\begin{proof}
-We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
-Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
-\qed
-\end{proof}
-
-$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
-
-\begin{theorem}\label{RP VASS}
-{\sc Resilience} is decidable for continuous VASS when $\Safe$  is
-definable in the existential theory of the rationals with addition and order.
-\end{theorem}
-
-\begin{proof}
-The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). \qed
-\end{proof}
-
-Remark that
-Theorem~\ref{SL Lossy} 
-(resp. Theorem~\ref{RP VASS})
-did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
-semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
-the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
-
-The above results concern only unbounded resilience. We consider now 
-{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
-the path from $\Bad$ to $\Safe$
-when $\Safe$ is downward-closed. Surprisingly, we have the following result for VAS (not VASS):
-\begin{proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
-\end{proposition}
-
-\begin{proof}
-Consider a given $k \in \N$,
- $\emptyset \neq \Bad \subseteq \N^d$ is upward-closed and $\Safe \subseteq \N^d$ is downward-closed, and
-consider a given VAS $V$.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition of $V$.
-The set $\Bad \neq \emptyset$ admits a finite basis $B_\Bad$.
-Consider the vector $\textbf{v}_{\Bad}$ obtained by summing all members of the basis of $\Bad$ and then consider the vector
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-
-All states reachable from $\textbf{u}_k$ in $k$ or less steps are above $ \textbf{v}_{\Bad} $
-and thus, are in $\Bad$, by upward-closedness.
-Hence  $\Safe$ is not reachable from $\textbf{u}_k$ in $k$ or less steps  and {$k$-resilience} does not hold.
-Since the reasoning hold for all $k \in \N$, {\sc Bounded-resilience} does not hold either.
-\qed
-\end{proof}
-
-This all changes when one considers VASS (having control-states).
-For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-$\mathop{\uparrow} q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
- $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
-Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
-This lead to a decision procedure
- which lead to the following decidability result:
-\begin{theorem}
-{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
-$\Safe = \mathop{\downarrow} \Safe$.
-\end{theorem}
-
-\begin{proof}
-Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
-Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
-$\Safe = S \setminus \Bad$
- downward-closed.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
-
-The set $\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider the vector
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
-
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-$\mathop{\uparrow} q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-Let us assume from now on 
- $\mathop{\uparrow} q_{j_1}(\textbf{0}), \mathop{\uparrow} q_{j_2}(\textbf{0}), \ldots \mathop{\uparrow} q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
-Because these subsets are upward-closed, we can compute a basis of
-$\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\mathop{\uparrow} q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
-
-We now consider the set 
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
-We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
-If it is infinite then
-$k$-resilience does not hold, 
-hence we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
-$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
-
-In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
-and hence
-$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
-We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
-
-From $k_0$ onwards,
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is stationnary. 
-If the set is finite then we stop the procedure once we have checked for every element
-whether it is possible to reach $\Safe$ from it.
-If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
-hold
-for $k_m = \max(k_0, k_{\pi})$ with 
-$k_{\pi}$ the maximum of the length of the paths from 
-$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
-If it is infinite then
-for every $k \in \N$,
-$k$-resilience does not hold, 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-\qed
-\end{proof}
-
-\section{Conclusion and perspectives}
-
-We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
-We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
- $\Safe = \mathop{\uparrow} \Safe$ and in the case 
- $\Safe = \mathop{\downarrow} \Safe$.
-
-Several questions still remain.
-For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
-One other angle of attack not studied here is the synthesis of sets 
-$\Safe$ for resilience,
-for instance, 
- finding the maximal upward-closed (resp. downward-closed) subset 
-$\Safe$ so that a system is $\Safe$-resilient. 
-Another question could be to analyse the resilience in the framework of a controller and its environment. 
-One could also extend upon the classes of set $\Safe$ 
- considered.
-As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets~\cite{DBLP:journals/fmsd/BertrandS13}.
-Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
-
-\alain{remercier les reviewers !}
-
-\bibliography{bib}
-
-\appendix
-
-\section{Reset-VASS and Minsky-machines}\label{minsk}
-
-\begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
-$\Safe=\mathop{\downarrow} \Safe$.
-\end{proposition}
-
-\begin{proof}
-\alain{preuve pas claire, à réécrire}
-% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
-% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
-%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
-%which can be simulated by WSTS.
-First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
-From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
-%			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
-Hence {\sc Resilience} is undecidable for Minsky machines.
-\alain{écrire qqchose ici}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
-% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
-%Reset-VASS extend the basic VASS model with special “reset
-%transitions” that set to $0$ some coordinates in the vector. 
-%			Let us recall their definition here.
-%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS in which
-there exists
- a sequence from 
-an initial state %$(s,x_1 \ldots x_n)$
-to
-a state $ q(0,  \ldots 0)$ if and only if
-there exists
-a sequence
- in the Minsky machine
- from a corresponding initial state
- to
- the state $q_f(0,0,0)$
-  (from Theorem~$5$ in \cite{araki1976PN}).
-Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
-resilience hold in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
 \end{proof}

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@
 
-\mathieu{Reviewer 1: the example at the bottom of \textbf{ page 8} (now 10), even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se. }
-
-In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable.
+\mathieu{Reviewer 1: the example at the bottom of \textbf{ page~\pageref{indec WSTS}}, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se.
+}
+
+In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
 
@@
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
-{\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
+ {\sc Resilience} is decidable.
 \end{theorem}
@@
 \begin{corollary}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$, {\sc Bounded resilience} 
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.  {\sc Bounded resilience} 
 and {\sc $k$-resilience} are decidable.
@@
 The convergence of $(\pred^{\leq k}(\Safe) )_{k\in\N}$ guarantees the latter eventually happens.
-\end{proof}
+\qed\end{proof}
 

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@
 First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
-$\Safe = Q_{M} \times \mathop{\downarrow} (1,1,0)$,
-since it is possible to guarantee that state $q_f(0,0,0)$ is the only element of
-$Q_{M} \times \mathop{\downarrow} (1,1,0)$
+$\Safe = \mathop{\downarrow} \Safe$,
+since it is possible to guarantee that state $q_f(0,0,0)$ is the only element 
+% of $Q_{M} \times \mathop{\downarrow} (1,1,0)$
 potentially reachable from 
-$Q_{M} \times \mathop{\uparrow} (0,0,1)$.
+$\Bad = Q_{M} \times \mathop{\uparrow} (0,0,1)$ that is not already inside $\Bad$.
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
@@
  in the Minsky machine and vice versa.
+ Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
  By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
  we obtain that 
- {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\pred$-basis-effective, 
+ {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
 $\omega^2$-WSTS with strong compatibility as well.
@@
 Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
+% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
 from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
-\alain{écrire qqchose ici}
+
+% \alain{écrire qqchose ici}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
@@
 {\sc $k$-state-resilience}
-are undecidable for strongly compatible $\pred$-basis-effective WSTS
+are undecidable for strongly compatible,
+$\post$-effective, $\pred$-basis-effective WSTS
 when
@@
 Unfortunately {\sc State-resilience} is undecidable in this case.
-As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-Section~\ref{safe-down}.\alain{trop rapide}
+As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines.
+Indeed, {\sc State-resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
+$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
+ As seen in 
+Proposition~\ref{indec WSTS with dcs},
+ the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS
+by sequences reaching a $\Safe'$ state from a $\Bad'$ one,
+ and vice versa, hence the following result. 
+ \alain{trop rapide}
+\mathieu{tjr trop rapide ?}
 
@@
 %$\post$-ideal-effective ???? and
+$\post$-effective,
  $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
 $\Safe=\mathop{\downarrow} \Safe$.
@@
 We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
+Based on this construction, the problem of deciding whether the downward-closed set 
+$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
+% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.
-\alain{écrire qqchose ici}
+
+% \alain{écrire qqchose ici}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 

### 2023-10-20 VMCAI-revised-version/final.tex

--- 2023-10-20 VMCAI-revised-version/final.tex
@@
  in the Minsky machine and vice versa.
+ Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
  By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
@@
 We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
+Based on this construction, the problem of deciding whether the downward-closed set 
+$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
 \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
 Hence {\sc Resilience} is undecidable for Minsky machines.

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity

@@
 	$S_j = (\N^2, \{f_j, g\}, \leq)$ in the previous proof is not $\post$-ideal-effective
-	nor $\pred$-effective.
+	nor $\pred$-basis-effective.
 

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity

@@
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
-Think for instance of a supply chain, or an airport’s air trafic control.
+%				Think for instance of a supply chain, or an airport’s air trafic control. 
 In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack
@@
 
-{\bf Home-spaces.}
- In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for a system $\mathscr{S} = (S,\rightarrow )$ with an initial state $s_0$ : a subset $H \subseteq S$ is an \emph{home-space}  if
-$\post^*(s_0) \subseteq \pred^*(H)$. If the home-space contains a single element, this element is an {\em home-state}.
-It could be easily generalized for two subsets $X,H$ and we say that $H$ is an \emph{home-space for $X$} if $\post^*(X) \subseteq \pred^*(H)$.
-In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (only published as an internal report \cite{de1989decidability}).
+%%%%%
+\noindent
+{\bf Resiliences.}  
+%
+%			If the home-space contains a single element, this element is an {\em home-state}.
+Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$ (i.e., $\post^*(X) \subseteq \pred^*(H)$).
+%
+A transition system is {\em resilient for $(\Bad,\Safe)$} if $(\Bad,\Safe)$ satisfies the home-space property.
+A transition system is {\em resilient for $\Safe$} if it is resilient for $(\Bad,\Safe)$ with $\Bad=S \setminus \Safe$.
+It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ with $\Bad=\{s_0\}$.
+%
+%			(resp. $\post^*(s_0)$)
+%                  It  is state-resilient if
+%			from every state, there is a run that goes into $\Safe$.
+%
+%			$\mathscr{S}$ is {\em resilient for $\Safe$} if from every state, there is a run that goes into  $\Safe$.
+%			$S \subseteq \pred^*(\Safe)$. 
+%			this is the $(S,\Safe)$ home-space problem.
+%			($\post^*(S) \subseteq \pred^*(\Safe)$).
+%
+% 		for $S,\Safe$: $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
+%		Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
+%		$\post^*(s_0) \subseteq \pred^*(\Safe)$.
+%		We will study three decidability questions for each type of set $\Safe$.
+%		The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient 			(for a given subset of states $\Safe \subseteq S$).
+%		The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
+%		whether $\Bad \subseteq \pred^*(\Safe)$.
+The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ (i.e., $S \subseteq \pred^{\leq k}(\Safe)$) and
+the {\em bounded resilience} problem is to decide whether there exists an $k$ such that the system is $k$-resilient.
+
+\noindent
+{\bf State of the art.}
+%
+%			{\bf Home-spaces.}
+
+In 2016, Prasad and Zuck introduced in \cite{DBLP:journals/corr/PrasadZ16} interesting definitions and results (without detailed proofs) about resilience in the framework of process algebra.
+They show that resilience is decidable for effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
+
+%		reduces to coverability which is decidable on WSTS.
+%		They gave conditions that insure that a transition system is an effective WSTS with both upward and reflexive 		downward compatibilities (and some other technical conditions). 
+%              In this framework and under these hypotheses, 
+%			resilience reduces to coverability which is decidable on WSTS. 
+%%%%%%%%%
+
+In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$ (i.e., $\mathop{\uparrow} \post^*(s)$) when $\Safe$ is upward-closed.
+%
+%			(called effective 
+%			$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
+%%%%%%%%%
+
+In 1986, Memmi and Vautherin introduced the (restricted) notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for $X$ a singleton.
+In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}).
 In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
-%%%%%
-
-{\bf Resiliences.}  
-The (general) resilience property for a transition system $\mathscr{S} = (S,\rightarrow )$ and a subset of states $\Safe \subseteq S$ consists of the following: $\mathscr{S}$ is {\em $\Safe$-resilient} if
-$S \subseteq \pred^*(\Safe)$. It 
-can be stated as an home-space problem : $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
-Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
-$\post^*(s_0) \subseteq \pred^*(\Safe)$.
-We will study three decidability questions for each type of set $\Safe$.
-The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient (for a given subset of states $\Safe \subseteq S$).
-The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
-whether $\Bad \subseteq \pred^*(\Safe)$.
-The $k$-resilience problem is to decide whether $S \subseteq \pred^{\leq k}(\Safe)$ (for a given $k$) and 
-the bounded resilience problem is to decide whether there exists an $k$ such that $\mathscr{S}$ is $\Safe$-$k$-resilient. We adapt these three definitions to state-resilience.
-
-{\bf State of the art.}
-In 2016, Prasad and Zuck introduced in  \cite{DBLP:journals/corr/PrasadZ16} interesting definitions and results (without detailed proofs) about resilience in the framework of process algebra.
-They built the composition of the process and its adversary as a transition system and they gave conditions that insure that the composed transition system is an effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
-In this framework and under these hypotheses, resilience reduces to coverability which is decidable on WSTS.
-%%%%%%%%%
-
-In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$ (called effective
-$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
-
-Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
-However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
-%%%%%%%%%
-%
 
@@
 \item Surprinsingly, the general undecidability statements about resilience were not known neither proved.
-We show that resilience and {state-resilience} problems are both undecidable for WSTS with strong compatibility.
+We show that resilience and {state-resilience} (i.e.
+$X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility.
 Moreover, {state-resilience},
@@
 $\Safe=\mathop{\uparrow} \Safe$. We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-
-\item The three resilience problems are decidable for ideally-effective $\omega^2$-WSTS with strong compatibility and $\Safe = \mathop{\uparrow} \Safe$.
-
-\item The resilience problem is decidable for ideally-effective WSTS with 
+%
+\item We clarify the different effectiness conditions that allow decidability of resilience.
+%
+\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility .
+%
+\item The resilience problem is decidable for post-ideal-effective downward-WSTS with 
 $\Safe=\mathop{\downarrow} \Safe$
-and
-the additional hypothesis that
-for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
+%and
+%the additional hypothesis that
+%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
 %
@@
 
-\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the  strong compatibility hypothesis.
-We also show that removing the effective 
+\item We remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
+We show that removing the effective
 $\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, most WSTS don't enjoy this property.
-We extend and prove the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for ideally-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for ideally-effective WSTS with strong downward compatibility).
-%
-
+We extend the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
+%
+\item Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
+However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
+%
 \item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
 \end{itemize}
+
+\noindent
+{\bf Plan of the paper.}
 

### 2023-10-22 VMCAI-revised-version/final.tex
effectivity

@@
 We show that resilience and {state-resilience} (i.e.
-$X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility.
+$X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed.
 Moreover, {state-resilience},
@@
 {$k$-state-resilience}
-are undecidable for strongly upward-compatible $\pred$-basis-effective WSTS
-when
-$\Safe=\mathop{\uparrow} \Safe$. We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-%
-\item We clarify the different effectiness conditions that allow decidability of resilience.
-%
-\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility .
-%
-\item The resilience problem is decidable for post-ideal-effective downward-WSTS with 
-$\Safe=\mathop{\downarrow} \Safe$
+are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
+$\Safe$ is upward-closed. 
+%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
+\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility.
+Moreover, the resilience problem is decidable for post-ideal-effective downward-WSTS with
+$\Safe$ upward-closed
 %and
@@
 %%%
-
+%
+\item We clarify the different effectiness conditions that allow decidability of the six resilience.
+%
 \item We remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
@@
 
+\alain{énoncer pour les reset-VASS c'est plus fort}
 \begin{proposition}\label{srp up down}

### 2023-10-22 VMCAI-revised-version/final.tex
effectivity

@@
 %
-\item We remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
+\item We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
 We show that removing the effective
-$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, most WSTS don't enjoy this property.
-We extend the main result in  \cite{DBLP:journals/corr/PrasadZ16} : the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
-%
-\item Remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
+$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.
+We still extend  \cite{DBLP:journals/corr/PrasadZ16} as follows: the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
+%
+\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
 However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.

### 2023-10-22 VMCAI-revised-version/final.tex
effectivity

@@
 \end{proposition}
- \alain{peut-on ajouter des hypothèses d'effectivity ?
+% \alain{peut-on ajouter des hypothèses d'effectivity ?
 peut-on montrer que l'exemple n'est pas pred effectif ?
@@
 
-\alain{énoncer pour les reset-VASS c'est plus fort}
+%\alain{énoncer pour les reset-VASS c'est plus fort}
 \begin{proposition}\label{srp up down}
@@
 {\sc $k$-state-resilience}
-are undecidable for strongly compatible,
+are undecidable for reset-VASS, hence for strongly compatible,
 $\post$-effective, $\pred$-basis-effective WSTS
@@
  As seen in 
-Proposition~\ref{indec WSTS with dcs},
+Proposition~\ref{indec WSTS with dcs},\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
  the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS
@@
  and vice versa, hence the following result. 
- \alain{trop rapide}
-\mathieu{tjr trop rapide ?}
+% \alain{trop rapide}
+%\mathieu{tjr trop rapide ?}
+%\alain{ça va}
 
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for 
+{\sc State-resilience} is undecidable for reset-VASS, hence also for
 %$\post$-ideal-effective ???? and

### 2023-10-23 VMCAI-revised-version/final.tex
effectivity

@@
 
-\begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
+\begin{theorem}(Finkel and Schnoebelen, [Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}], Abdulla and al., [Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}])
  A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.

### 2023-10-23 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-22 VMCAI-revised-version/final.tex
@@
 \keywords{Verification, Resilience, Home-Space, Well-structured transition systems, Vector addition system with states}
+
+\alain{j'ai enlevé les noms lorsqu'on cite un resultat}
 
@@
 Let us now recall the (most general) definition of well-structured transition systems.
-\begin{definition}(Finkel \cite{DBLP:journals/iandc/Finkel90})
+\begin{definition}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
 A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 Another type of WSTS exists that enjoys downward compatibility.
-
-\begin{definition}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
+%
+%		Finkel and Schnoebelen, 
+%
+\begin{definition}[Definition 5.1, \cite{DBLP:journals/tcs/FinkelS01}]
 A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 With the $\pred$-basis-effective hypothesis, we obtain:
-
-\begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
+%
+%			(Finkel and Schnoebelen, Abdulla and al., 
+%
+\begin{theorem}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
  A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
@@
 With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the $\pred$-basis-effective hypothesis).
-
-\begin{theorem}(Blondin and al. ~\cite{DBLP:journals/lmcs/BlondinFM17})
+%
+%		(Blondin and al., 
+%
+\begin{theorem}[Corollaire 4.4, ~\cite{DBLP:journals/lmcs/BlondinFM17}]
 Coverability is decidable for any $\post$-ideal-effective WBTS.
@@
 Downward-WSTS enjoy a powerfull property.
-
-\begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
-$\post$-effective finitely branching downward-WSTS with reflexive compatibility are $\post^*$-basis-effective.
+%
+%		Finkel and Schnoebelen, 
+%
+\begin{theorem}[Proposition 5.4, \cite{DBLP:journals/tcs/FinkelS01}]
+Finitely branching downward-WSTS with reflexive compatibility and $\post$-effective are $\post^*$-basis-effective.
 \end{theorem}
@@
 %
-\begin{theorem}\label{ref ozkan}
-(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
+%		(\"Ozkan,   \"Ozkan and W{\"{u}}rdemann,  
+%
+\alain{je ne comprends pas pourquoi les citations ne sortent pas comme les autres}
+\begin{theorem}\label{ref ozkan}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]
 {\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
@@
 
-\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
+\begin{proposition}[Proposition 1, \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
 For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.

### 2023-10-23 VMCAI-revised-version/final.tex
effectivity
appendix minsky

--- 2023-10-23 VMCAI-revised-version/final.tex
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
-$\Safe=\mathop{\downarrow} \Safe$.
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathopp{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
 \end{proposition}
@@
 First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
+	This is due to the undecidability of $2$-counter Minsky machine termination.
 From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
@@
 from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
-Hence {\sc Resilience} is undecidable for Minsky machines.
+Note that by chosing $s_0 \in Q_{M'} \times \mathop{\uparrow} (0,0,1)$,
+deciding whether the downward-closed set 
+$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable
+from every element of $\post^*(s_0) \cup Q_{M'} \times \mathop{\uparrow} (0,0,1)$ 
+is again equivalent to asking whether $M$ terminate on every input.
+Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
 
@@
 to
-a state $ q(0,  \ldots 0)$ if and only if
+a state $ q({\bf 0})$ if and only if
 there exists
@@
   (from Theorem~$5$ in \cite{araki1976PN}).
-Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
-resilience hold in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
+  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
+Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
+{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
+and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
 $\omega^2$-WSTS with strong compatibility as well.\qed

### 2023-10-23 VMCAI-revised-version/final.tex
appendix minsky

@@
 
-\begin{proposition}\label{indec WSTS with dcs}
+We now delve into more details into the undecidability results 
+upon which rely
+Theorem~\ref{indec WSTS with dcs}
+and
+Theorem~\ref{srp down up},
+namely, that
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathopp{\downarrow}~\Safe$
+are undecidable for Minsky machines
+and 
+for reset-VASS.
+
+The problems are undecidable for a Minsky machine with at least $3$ counters when
+$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
+We then consider a reset-VASS simulating the Minsky machine.
+We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way.
+% The simulation however is 
+Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0,0)$ in the Minsky machine, 
+this weak simulation will be enough to obtain undecidability of both problems.
+
+\begin{proposition}\label{both minsk indec}
 {\sc Resilience} and {\sc State-resilience} when $\Safe = \mathopp{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.

### 2023-10-23 VMCAI-revised-version/final.tex
appendix minsky

@@
 namely, that
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathopp{\downarrow}~\Safe$
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
 are undecidable for Minsky machines
@@
 \begin{proposition}\label{both minsk indec}
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathopp{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
 \end{proposition}

### 2023-10-23 VMCAI-revised-version/final.tex
appendix minsky

@@
 upon which rely
-Theorem~\ref{indec WSTS with dcs}
+Proposition~\ref{indec WSTS with dcs}
 and
-Theorem~\ref{srp down up},
+Proposition~\ref{srp down up},
 namely, that

### 2023-10-23 VMCAI-revised-version/final.tex
appendix minsky

@@
 \mathieu{Reviewer 1: the example at the bottom of \textbf{ page~\pageref{indec WSTS}}, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se.
-}
+=> citer un autre résultat où les increasing recursive functions ont été utilisées ?}
 
@@
 Proposition~\ref{indec WSTS with dcs},\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
+\mathieu{Elle apparaît dans le corps du texte aussi mais j'avais utilisé deux fois le même label ce qui est corrigé maintenant.}
  the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS

### 2023-10-23 VMCAI-revised-version/final.tex

@@
 \keywords{Verification, Resilience, Home-Space, Well-structured transition systems, Vector addition system with states}
+
+\alain{j'ai ajouté les numéros et j'ai enlevé les noms dans les {theorem, proposition, definition} qu'on cite}
 
@@
 With the $\pred$-basis-effective hypothesis, we obtain:
-
-\begin{theorem}(Finkel and Schnoebelen, [Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}], Abdulla and al., [Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}])
+%
+%			(Finkel and Schnoebelen, Abdulla and al., 
+%
+\begin{theorem}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
  A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.

### 2023-10-23 VMCAI-revised-version/final.tex
appendix minsky

@@
  {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.
-%\mathieu{preuve résumé / concise à écrire}
-\iffalse
-\alain{preuve pas claire, à réécrire}
-% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
-% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
-%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
-%which can be simulated by WSTS.
-First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
-From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
-%			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
-Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
-% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
-Hence {\sc Resilience} is undecidable for Minsky machines.
-
-% \alain{écrire qqchose ici}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
-% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
-%Reset-VASS extend the basic VASS model with special “reset
-%transitions” that set to $0$ some coordinates in the vector. 
-%			Let us recall their definition here.
-%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS in which
-there exists
- a sequence from 
-an initial state %$(s,x_1 \ldots x_n)$
-to
-a state $ q(0,  \ldots 0)$ if and only if
-there exists
-a sequence
- in the Minsky machine
- from a corresponding initial state
- to
- the state $q_f(0,0,0)$
-  (from Theorem~$5$ in \cite{araki1976PN}).
-Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
-resilience hold in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
-\fi
-\qed \end{proof}
+$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
 

### 2023-10-23 VMCAI-revised-version/final.tex
effectivity

@@
 
-\alain{j'ai ajouté les numéros et j'ai enlevé les noms dans les {theorem, proposition, definition} qu'on cite}
+%\alain{j'ai ajouté les numéros et j'ai enlevé les noms dans les {theorem, proposition, definition} qu'on cite}
 
@@
 => citer un autre résultat où les increasing recursive functions ont été utilisées ?}
+\alain{ma réponse: termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.}
 
@@
  As seen in 
-Proposition~\ref{indec WSTS with dcs},\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
-\mathieu{Elle apparaît dans le corps du texte aussi mais j'avais utilisé deux fois le même label ce qui est corrigé maintenant.}
+Proposition~\ref{indec WSTS with dcs},
+%\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
+%\mathieu{Elle apparaît dans le corps du texte aussi mais j'avais utilisé deux fois le même label ce qui est corrigé maintenant.}
  the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS

### 2023-10-24 VMCAI-revised-version/final.tex
Adding table of results

@@
 \usepackage{amsfonts}
+\usepackage{multirow}
 
@@
 % \alain{à quoi sert cette remarque ?}
+
+\begin{center}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \ 
+		& $\Safe = {\uparrow \Safe}$~ 
+		 & $\Safe = {\downarrow \Safe}$~ 
+ \\ \hline
+   \multirow{3}{*}{RP} 
+  	& WBTS (post-ideal-effective) Thm~4  
+   		 & Downard-WSTS (pred*-basis-effective) Thm~6 \\ 
+~	& WBTS (pred-basis-effective) Thm~5 
+		& VASS \\\cline{2-3}
+~	& WSTS (post-effective) Prop~1 & WSTS (pred-basis-effective) Prop~2    \\ \hline
+   \multirow{3}{*}{BRP} 
+  	& WBTS (post-ideal-effective) Thm~4  
+   		 & ? \\ 
+~	& WBTS (pred-basis-effective) Thm~5 
+		& (VASS ici aussi) \\\cline{2-3}
+~	& ? & ?    \\ \hline
+   \multirow{3}{*}{kRP} 
+ 	& WBTS (post-ideal-effective) Thm~4  
+   		 & ? \\ 
+~	& WBTS (pred-basis-effective) Thm~5 
+		& (VASS ici aussi) \\\cline{2-3}
+~	& ? & ?    \\ \hline
+
+   \multirow{2}{*}{sRP} 
+  	& WBTS (post-ideal-effective) Thm~8 
+   		 & WSTS (downard + upward) Thm~9 \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~5 & WSTS (pred-basis + ??) Prop~8    \\ \hline
+   \multirow{2}{*}{BsRP} 
+  	& WBTS (pred + post*-ideal-effective) Thm~7 [18,10]  
+   		 & WSTS (d+up) Cor~1 \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~5 & ?    \\ \hline
+   \multirow{2}{*}{ksRP} 
+  	& WBTS (pred + post*-ideal-effective) Thm~7 [18,10]  
+   		 & WSTS (d+up) Cor~1 \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~5 & ?    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
 

### 2023-10-24 VMCAI-revised-version/final.tex
Adding table of results

--- 2023-10-23 VMCAI-revised-version/final.tex
@@
 % \alain{à quoi sert cette remarque ?}
+
+\begin{center}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \ 
+		& $\Safe = {\uparrow \Safe}$~ 
+		 & $\Safe = {\downarrow \Safe}$~ 
+ \\ \hline
+   \multirow{3}{*}{RP} 
+  	& WBTS (post-ideal-effective) Thm~\ref{down-up}
+   		 & Downard-WSTS (pred*-basis-effective) Thm~\ref{up-down} \\ 
+~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+		& VASS \\\cline{2-3}
+~	& WSTS (post-effective) Prop~\ref{indec WSTS} & WSTS (pred-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
+   \multirow{3}{*}{BRP} 
+  	& WBTS (post-ideal-effective) Thm~\ref{down-up}
+   		 & ? \\ 
+~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+		& (VASS ici aussi) \\\cline{2-3}
+~	& ? & ?    \\ \hline
+   \multirow{3}{*}{kRP} 
+ 	& WBTS (post-ideal-effective) Thm~\ref{down-up} 
+   		 & ? \\ 
+~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+		& (VASS ici aussi) \\\cline{2-3}
+~	& ? & ?    \\ \hline
+
+   \multirow{2}{*}{sRP} 
+  	& WBTS (post-ideal-effective) Thm~\ref{post srp}
+   		 & WSTS (downard + upward) Thm~9 \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~\ref{srp up down} & WSTS (pred-basis + ??) Prop~\ref{srp up down}    \\ \hline
+   \multirow{2}{*}{BsRP} 
+  	& WBTS (pred + post*-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+   		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~\ref{srp up down} & ?    \\ \hline
+   \multirow{2}{*}{ksRP} 
+  	& WBTS (pred + post*-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+   		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
+~	& WSTS (pred-...) Prop~\ref{srp up down} & ?    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
 

### 2023-10-24 VMCAI-revised-version/final.tex
Adding references and such to the figure (gonna need another pass/check-up if we introduce new references/definitions unfortunately
changing problems environments to regroup the three pb

@@
 	\begin{figure}
-	\hspace{2.3cm}
-\includegraphics[width=0.60\textwidth]{WSTS_taxonomy}
+	\hspace{1.2cm}
+\includegraphics[width=0.80\textwidth]{WSTS_taxonomy_large}
 	\caption{A taxonomy of WSTS and variants.}
@@
 
+\iffalse
 \problemx{Resilience (RP)}
@@
 {$\exists k \geq 0$ such that {$S \longrightarrow^{\leq k} \Safe$ ?\newline}}
+\fi
+
+\problemx{Resilience problems}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$.}
+{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
+({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
+({\sc bounded resilience problem (BRP)})	$\exists k \geq 0 ~ S \longrightarrow^{\leq k} \Safe$ ?\newline}
 
@@
 
+\iffalse
+
 \problemx{State-resilience (SRP)}
@@
 {$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
+
+\fi
+
+\problemx{State resilience problems}
+{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, and a set $\Safe \subseteq S$.}
+{ ({\sc state-resilience problem (SRP)}) $\post^*(s) \longrightarrow^{*} \Safe$ ?\newline  
+({\sc $k$-state-resilience problem (kSRP)})		$\post^*(s) \longrightarrow^{\leq k} \Safe$ ?\newline
+({\sc bounded-state-resilience problem (BSRP)})	$\exists k \geq 0 $ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$?\newline}
 

### 2023-10-24 VMCAI-revised-version/final.tex
Starting to rearrange the proofs towards the appendix

@@
 
-\begin{proof}
+\iffalse\begin{proof}
 Let us prove $1 \implies 2$. 
@@
 \qed
-\end{proof}
+\end{proof}\fi
 
@@
 
+\section{Lemma technical proofs}
+
+Let us now prove more formally some of the lemmas and propositions ...
+
+\mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
+
+\noindent
+\textbf{Lemma~\ref{downward compatible}.}
+{\em For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.}
+
+\begin{proof}
+Let us prove $1 \implies 2$. 
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \mathop{\downarrow} \pred^*(D)$.
+By downward closure, there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$.
+By definition of $\pred^*(D)$, there exists 
+$d \in D$ such that: (1) either $y=d$ and then $x \in D \subseteq \pred^*(D)$ or (2) there exist $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+%
+By downward compatibility $a_0 \to a_1$
+implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
+$x \to^* a'_1$.
+More generally $a_i \to a_{i+1}$ and
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
+$a'_i \to^* a'_{i+1}$,
+and, by induction,
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+with $d' \leq d$.
+Since
+$d'$ 
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\mathop{\downarrow}\pred^*(D) \subseteq \pred^*(D)$.
+Since $\pred^*(D) \subseteq \mathop{\downarrow} \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\mathop{\downarrow}\pred^*(D)$ is downward-closed.
+
+Let us prove $2 \implies 1$.
+Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\mathop{\downarrow} t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\mathop{\downarrow} t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\mathop{\downarrow} t_1)$.
+This means that there exists $t_2 \leq t_1$ such that
+$s_2 \rightarrow^{*} t_2$. Hence, we proved that $\mathscr{S}$ is downward compatible.
+\qed
+\end{proof}
+
 \section{Reset-VASS and Minsky-machines}\label{minsk}
@@
 
+\mathieu{mettre ici le détail de la  prop5}
+
+\section{Thm 9}
+
+\mathieu{mettre ici le détail du Theo 9}
+
+\section{More details on VASS}
+
+\mathieu{mettre ici le détail du Theo 13}
+
 \end{document}

### 2023-10-24 VMCAI-revised-version/final.tex
up
retiré preuve prop 5

@@
 
+\iffalse
  \begin{center}
@@
 \end{example}
-
+%%%%%
+
+%%%%%
 % Hence we can see how adding transitions can 
@@
 % \alain{à quoi sert cette remarque ?}
+\fi
 
@@
 
-We may now prove the undecidability of the three state-resilience problems.
+% We may now prove the undecidability of the three state-resilience problems.
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction implies 
+{\sc $t$-liveness} is undecidable.
+Because
+{\sc $t$-liveness} is undecidable in reset-VASS
+and reducible to {\sc State-resilience}, we 
+deduce the following.
 
@@
 
-\begin{proof}
+\iffalse \begin{proof}
 Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
@@
  {\sc $k$-state-resilience}. \qed
-\end{proof}
+\end{proof} \fi
 
@@
 
-\begin{proof}
+\iffalse \begin{proof}
 Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
@@
 $A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
-\end{proof}
+\end{proof} \fi
 
@@
 
+Additionnally ...
+
+\noindent
+\textbf{Lemma~\ref{Lemma intersection}.}
+{\em Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
+}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
+Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
+Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
+$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
+\end{proof}
+
 \section{Reset-VASS and Minsky-machines}\label{minsk}
@@
 
+\noindent
+\textbf{Proposition~\ref{srp up down}.}
+{\em {\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for reset-VASS, hence for strongly compatible,
+$\post$-effective, $\pred$-basis-effective WSTS
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+}
+
+\begin{proof}
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
+{\sc $t$-liveness} is undecidable.
+Because
+{\sc $t$-liveness} is undecidable in reset-VASS
+and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
+Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
+reducible to {\sc $k$-state-resilience}:
+since $\Safe=\mathop{\uparrow} \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
+compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
+We compute 
+$n_0$, then iteratively check whether $k$-state-resilience 
+hold for $k$ from $0$ to $n_0$.  
+Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+since 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
+Hence the undecidability of {\sc Bounded-state-resilience}
+and 
+ {\sc $k$-state-resilience}. \qed
+\end{proof}
+
 \section{Thm 9}

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
 % We may now prove the undecidability of the three state-resilience problems.
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction implies 
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the reduction implies 
 {\sc $t$-liveness} is undecidable.
-Because
-{\sc $t$-liveness} is undecidable in reset-VASS
-and reducible to {\sc State-resilience}, we 
+By reduction to {\sc State-resilience}, and straightforward reduction to the other two state resilience problems, we 
 deduce the following.

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
    \multirow{3}{*}{RP} 
-  	& WBTS (post-ideal-effective) Thm~\ref{down-up}
-   		 & Downard-WSTS (pred*-basis-effective) Thm~\ref{up-down} \\ 
-~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
+   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down} \\ 
+~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
 		& VASS \\\cline{2-3}
-~	& WSTS (post-effective) Prop~\ref{indec WSTS} & WSTS (pred-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
+~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
    \multirow{3}{*}{BRP} 
-  	& WBTS (post-ideal-effective) Thm~\ref{down-up}
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
    		 & ? \\ 
-~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
 		& (VASS ici aussi) \\\cline{2-3}
@@
    \multirow{3}{*}{kRP} 
- 	& WBTS (post-ideal-effective) Thm~\ref{down-up} 
+ 	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
    		 & ? \\ 
-~	& WBTS (pred-basis-effective) Thm~\ref{xxx}
+~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
 		& (VASS ici aussi) \\\cline{2-3}
@@
    \multirow{2}{*}{sRP} 
-  	& WBTS (post-ideal-effective) Thm~\ref{post srp}
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{post srp}
    		 & WSTS (downard + upward) Thm~9 \\ \cline{2-3}
-~	& WSTS (pred-...) Prop~\ref{srp up down} & WSTS (pred-basis + ??) Prop~\ref{srp up down}    \\ \hline
+~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & WSTS ($\pred$-basis + ??) Prop~\ref{srp up down}    \\ \hline
    \multirow{2}{*}{BsRP} 
-  	& WBTS (pred + post*-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+  	& WBTS ($\pred$ + $\post^*$-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
    		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS (pred-...) Prop~\ref{srp up down} & ?    \\ \hline
+~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & ?    \\ \hline
    \multirow{2}{*}{ksRP} 
-  	& WBTS (pred + post*-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+  	& WBTS ($\pred$ + $\post^*$-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
    		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS (pred-...) Prop~\ref{srp up down} & ?    \\ \hline
+~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & ?    \\ \hline
 
@@
 We use a mixture of the notions defined both by Blondin and al.
-in \cite{DBLP:journals/lmcs/BlondinFM17}) and by Halfon in \cite{DBLP:phd/hal/Halfon18}.
+in \cite{DBLP:journals/lmcs/BlondinFM17} and by Halfon in \cite{DBLP:phd/hal/Halfon18}.
 %			 that will be essential for solving various resilience problems in different models of WSTS. 

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
    \multirow{2}{*}{BsRP} 
-  	& WBTS ($\pred$ + $\post^*$-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+  	& WBTS ($\pred$ + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
    		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
@@
    \multirow{2}{*}{ksRP} 
-  	& WBTS ($\pred$ + $\post^*$-ideal-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+  	& WBTS ($\pred$ + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
    		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
 The previous formula reduces itself to {\sc State-resilience}
-where $\Safe = \pre(t)$.
+where $\Safe = \pre(t)$.\qed
 \end{proof}

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

@@
    		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down} \\ 
-~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
 		& VASS \\\cline{2-3}
-~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
+~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
    \multirow{3}{*}{BRP} 
   	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
-   		 & ? \\ 
-~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		& (VASS ici aussi) \\\cline{2-3}
+   		 & VASS \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		&   \\\cline{2-3}
 ~	& ? & ?    \\ \hline
@@
  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
-   		 & ? \\ 
-~	& WBTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		& (VASS ici aussi) \\\cline{2-3}
+   		 & VASS \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		& \\\cline{2-3}
 ~	& ? & ?    \\ \hline
@@
    \multirow{2}{*}{sRP} 
-  	& WBTS ($\post$-ideal-effective) Thm~\ref{post srp}
-   		 & WSTS (downard + upward) Thm~9 \\ \cline{2-3}
-~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & WSTS ($\pred$-basis + ??) Prop~\ref{srp up down}    \\ \hline
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} NON
+   		 & WSTS (downward + upward) Thm~9 \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Reset-VASS, WSTS ($\pred$-basis-effective+...) Prop~\ref{srp up down}    \\ \hline
    \multirow{2}{*}{BsRP} 
-  	& WBTS ($\pred$ + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
-   		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & ?    \\ \hline
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & ?    \\ \hline
    \multirow{2}{*}{ksRP} 
-  	& WBTS ($\pred$ + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
-   		 & WSTS (d+up) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS ($\pred$-...) Prop~\ref{srp up down} & ?    \\ \hline
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & ?    \\ \hline
 

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

@@
    		 & WSTS (downward + upward) Thm~9 \\ \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Reset-VASS, WSTS ($\pred$-basis-effective+...) Prop~\ref{srp up down}    \\ \hline
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}    \\ \hline
    \multirow{2}{*}{BsRP} 
@@
 \alain{je ne comprends pas pourquoi les citations ne sortent pas comme les autres}
-\begin{theorem}\label{ref ozkan}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]
+\begin{theorem}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}
 {\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity
up

@@
 		&   \\\cline{2-3}
-~	& ? & ?    \\ \hline
+~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}  & ?    \\ \hline
    \multirow{3}{*}{kRP} 
@@
 \begin{proposition}\label{indec WSTS}
-{\sc Resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ is undecidable.
+{\sc Resilience} and {\sc Bounded resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ are undecidable.
 \end{proposition}
@@
 	WSTS.
-Hence resilience for $\omega^2$-WSTS with strong compatibility is undecidable.\qed
+Hence {\sc Resilience} for $\omega^2$-WSTS with strong compatibility is undecidable. Furthermore, strong compatibility implies the existence of $k\in\N$ such that
+$\pred^*(\Safe) = \pred^{\leq k}(\Safe)$, hence {\sc Bounded-resilience} for $\omega^2$-WSTS with strong compatibility is undecidable as well.\qed
 \end{proof}
@@
 	nor $\pred$-basis-effective.
-
-\mathieu{Reviewer 1: the example at the bottom of \textbf{ page~\pageref{indec WSTS}}, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se.
+  Termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.
+
+% \mathieu{Reviewer 1: the example at the bottom of \textbf{ page~\pageref{indec WSTS}}, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se.
 => citer un autre résultat où les increasing recursive functions ont été utilisées ?}
-\alain{ma réponse: termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.}
+% \alain{ma réponse: termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.}
 
@@
 It now remains to argue why {\sc $t$-liveness} is undecidable.
-We  recall the
- undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS of dimension $d$, which consists in, given a
-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
-deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
+Recall the
+ undecidable~\cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS% of dimension $d$%, which consists in, given a reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$, deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$%
+ .
+
+\problemx{zero-reachability}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, and $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+
 %
@@
 %
-\alain{je ne comprends pas pourquoi les citations ne sortent pas comme les autres}
 \begin{theorem}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
 Several families of formal models of processes  \cite{DBLP:journals/tcs/FinkelS01} give rise to WSTSs in a natural way with disfferent compatibilities, e.g.
-compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets) data Petri nets, reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
+compatible counter machines like VASS with $d$ counters and $Q$ a finite set of control-states (and equivalently Petri nets) data Petri nets\nocite{DBLP:journals/fuin/LazicNORW08}, reset/transfer VASS are WSTS with strong compatibility for the usual ordering $=\times \leq^d$ on the set of states $S=Q \times \N^d$.
 Similarly, lossy channel systems with $d$ channels and $Q$ a finite set of control-states are WSTS (with a non-strong compatibility)

### 2023-10-24 VMCAI-revised-version/final.tex
up

@@
 
-\mathieu{Donner des exemples dans cette classe.}
-\alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
+% \mathieu{Donner des exemples dans cette classe.}
+% \alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
+
+$\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
+Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96, DBLP:journals/iandc/CeceFI96}.
 
@@
  which lead to the following decidability result:
-\begin{theorem}
+\begin{theorem}\label{vass down}
 {\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
@@
 \end{theorem}
+
+\begin{proof}
+\iffalse
+Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
+Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
+$\Safe = S \setminus \Bad$
+ downward-closed.
+Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
+
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider the vector
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\mathop{\uparrow} q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+Let us assume from now on 
+ $\mathop{\uparrow} q_{j_1}(\textbf{0}), \mathop{\uparrow} q_{j_2}(\textbf{0}), \ldots \mathop{\uparrow} q_{j_\ell}(\textbf{0})$ are
+all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
+Because these subsets are upward-closed, we can compute a basis of
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\mathop{\uparrow} q_{j_2}(\textbf{0}))$, \ldots
+$\pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+
+We now consider the set 
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
+We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
+If the set is finite then we stop the procedure once we have found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
+If it is infinite then
+$k$-resilience does not hold, 
+hence we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps.
+This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
+$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
+thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
+
+In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
+and hence
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
+We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
+
+From $k_0$ onwards,
+$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
+\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is stationnary. 
+If the set is finite then we stop the procedure once we have checked for every element
+whether it is possible to reach $\Safe$ from it.
+If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
+hold
+for $k_m = \max(k_0, k_{\pi})$ with 
+$k_{\pi}$ the maximum of the length of the paths from 
+$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
+If it is infinite then
+for every $k \in \N$,
+$k$-resilience does not hold, 
+and we eventually find a witness that $k$-resilience does not hold.
+Indeed, 
+if
+$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
+is
+infinite
+then
+there exists an element of the form $q(\textbf{u} )$
+with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
+that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
+and hence from which it is not possible at all to reach
+$\Safe$ in $k$ or less steps. 
+\fi
+\mathieu{Write small text and refer to appendix~\ref{appendix vass}}
+\qed
+\end{proof}
+
+\section{Conclusion and perspectives}
+
+We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
+We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
+ $\Safe = \mathop{\uparrow} \Safe$ and in the case 
+ $\Safe = \mathop{\downarrow} \Safe$.
+
+Several questions still remain.
+For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
+One other angle of attack not studied here is the synthesis of sets 
+$\Safe$ for resilience,
+for instance, 
+ finding the maximal upward-closed (resp. downward-closed) subset 
+$\Safe$ so that a system is $\Safe$-resilient. 
+Another question could be to analyse the resilience in the framework of a controller and its environment. 
+One could also extend upon the classes of set $\Safe$ 
+ considered.
+As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets~\cite{DBLP:journals/fmsd/BertrandS13}.
+Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
+
+\alain{remercier les reviewers !}
+
+\bibliography{bib}
+
+\appendix
+
+\section{Lemma technical proofs}
+
+Let us now prove more formally some of the lemmas and propositions ...
+
+\mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
+
+\noindent
+\textbf{Lemma~\ref{downward compatible}.}
+{\em For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.}
+
+\begin{proof}
+Let us prove $1 \implies 2$. 
+Let $D$ be a downward-closed subset of $S$
+and let $x \in \mathop{\downarrow} \pred^*(D)$.
+By downward closure, there exists
+$y \in \pred^*(D)$ 
+such that $x \leq y$.
+By definition of $\pred^*(D)$, there exists 
+$d \in D$ such that: (1) either $y=d$ and then $x \in D \subseteq \pred^*(D)$ or (2) there exist $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
+$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
+%
+By downward compatibility $a_0 \to a_1$
+implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
+$x \to^* a'_1$.
+More generally $a_i \to a_{i+1}$ and
+$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
+$a'_i \to^* a'_{i+1}$,
+and, by induction,
+ $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
+with $d' \leq d$.
+Since
+$d'$ 
+belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\mathop{\downarrow}\pred^*(D) \subseteq \pred^*(D)$.
+Since $\pred^*(D) \subseteq \mathop{\downarrow} \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\mathop{\downarrow}\pred^*(D)$ is downward-closed.
+
+Let us prove $2 \implies 1$.
+Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\mathop{\downarrow} t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\mathop{\downarrow} t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\mathop{\downarrow} t_1)$.
+This means that there exists $t_2 \leq t_1$ such that
+$s_2 \rightarrow^{*} t_2$. Hence, we proved that $\mathscr{S}$ is downward compatible.
+\qed
+\end{proof}
+
+Additionnally ...
+
+\noindent
+\textbf{Lemma~\ref{Lemma intersection}.}
+{\em Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
+Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
+}
+
+\begin{proof}
+Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
+Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
+Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
+Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
+Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
+In the other direction,
+since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
+$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
+$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
+\end{proof}
+
+\section{Reset-VASS and Minsky-machines}\label{minsk}
+
+We now delve into more details into the undecidability results 
+upon which rely
+Proposition~\ref{indec WSTS with dcs}
+and
+Proposition~\ref{srp down up},
+namely, that
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
+are undecidable for Minsky machines
+and 
+for reset-VASS.
+
+The problems are undecidable for a Minsky machine with at least $3$ counters when
+$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
+We then consider a reset-VASS simulating the Minsky machine.
+We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way.
+% The simulation however is 
+Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0,0)$ in the Minsky machine, 
+this weak simulation will be enough to obtain undecidability of both problems.
+
+\begin{proposition}\label{both minsk indec}
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
+\end{proposition}
+
+\begin{proof}
+\alain{preuve pas claire, à réécrire}
+% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
+% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
+%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
+%which can be simulated by WSTS.
+First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
+	This is due to the undecidability of $2$-counter Minsky machine termination.
+From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
+such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
+%			with at least $1$ on its third counter. 
+We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
+Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
+Based on this construction, the problem of deciding whether the downward-closed set 
+$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
+% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
+from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
+Note that by chosing $s_0 \in Q_{M'} \times \mathop{\uparrow} (0,0,1)$,
+deciding whether the downward-closed set 
+$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable
+from every element of $\post^*(s_0) \cup Q_{M'} \times \mathop{\uparrow} (0,0,1)$ 
+is again equivalent to asking whether $M$ terminate on every input.
+Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
+
+% \alain{écrire qqchose ici}
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
+% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
+%Reset-VASS extend the basic VASS model with special “reset
+%transitions” that set to $0$ some coordinates in the vector. 
+%			Let us recall their definition here.
+%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
+The simulation operates by building a reset-VASS in which
+there exists
+ a sequence from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state $ q({\bf 0})$ if and only if
+there exists
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ the state $q_f(0,0,0)$
+  (from Theorem~$5$ in \cite{araki1976PN}).
+  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
+Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
+{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
+and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.\qed
+%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
+\end{proof}
+
+\section{Reset-VASS and state-resilience}\label{appendix}
+
+\newcommand{\arrowcircle}[1][]{%
+  \begin{tikzpicture}[#1]
+    \draw[->] (0,0ex) -- (2em,0ex);
+    \draw (1em,0ex) circle (0.7ex);
+  \end{tikzpicture}%
+}
+
+Let us recall the {\em zero-reachability} problem in (reset-)VASS.
+
+\problemx{zero-reachability}
+{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+
+The zero-reachability problem is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
+For a better representation, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS. 
+A \emph{reset Petri net} $N = (P, T, E, R, \mu)$ consists of a finite set of {\em places} $P = \{p_1, p_2, ...
+p_{|P|}\}$, a finite set of {\em transitions} $T = \{t_1, t_2, ..., t_{|T|} \}$, a finite set of {\em arcs} $E \subseteq (P \times T) \cup (T \times P)$, a finite set of
+{\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
+It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
+The same equivalence hold between reset-VASS and reset Petri nets.
+
+We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
+
+\noindent
+\textbf{Proposition~\ref{liveness reset}.}
+{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\
+
+We may then deduce that:\\
+
+\noindent
+\textbf{Corollary}
+{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
+
+\begin{proof}
+
+If we wish to determine if a marking in which every place is empty is reachable for a reset Petri net
+ $N_1 = (P_1, T_1, E_1, R_1, \mu_1)$ 
+ we construct a reset Petri net
+ $N_2 = (P_2, T_2, E_2, R_2, \mu_2)$ 
+  which is live in $N_2$ if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
+
+\begin{samepage}
+{\bf Construction of $N_2$} \\
+\indent
+The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_a$ and $p_b$ and $|P_1| +2$ transitions $t_{p_a}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_b}$.
+We first modify all transitions of $N_1$ to include $p_a$ as both an input and an output.
+The initial marking $\mu_2$ will include a token in $p_a$ and no token in $p_b$. 
+This new place $p_a$ serve to mark that the run is ongoing.
+As long as it contains a token, the adjusted transitions of $N_1$ are live and can be used normally.
+Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
+We add an additional transition $t_{p_a}$ 
+which has $p_a$ as an input and a null output.
+This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $P_1$ in $N_2$.
+ The place $p_a$ and transition $t_{p_a}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_a}$ to fire and freeze the net at that marking.
+We introduce a new place $p_b$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_b$ as output.
+Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live in $N_2$ if a token is ever
+put in $p_2$.
+\end{samepage}
+
+ \begin{center}
+	\begin{figure}
+\includegraphics[width=0.75\textwidth]{FigurePN}
+\caption{Construction of $N_2$ from $N_1$.}
+	\end{figure}
+\end{center}
+
+Let us now prove that $t_{p_2}$ is live in $N_2$ if and only if 
+the zero marking
+is not reachable from $\mu_1$ in $N_1$. \\
+
+{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live in $N_2$}.
+
+Indeed the marking with zero in every place of $P_1$ and in $p_b$ is reachable in $N_2$, by executing the same sequence of transition firings.
+Then $t_{p_a}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
+From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
+Thus $t_{p_2}$ is not live in $N_2$.
+\\
+
+{\bf Suppose that $t_{p_2}$ is not live in $N_2$, then the zero marking is reachable from $\mu_1$ in $N_1$}
+
+Indeed, if $t_{p_2}$ is not live in $N_2$, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
+This means that no transition $t_p$ is live in $N_2$ in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
+Thus, every place of $P_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
+it can be reached without using the transitions $t_p$ or $t_{p_2}$.
+Since $t_{p_a}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
+Thus the zero marking is reachable from $\mu_1$ in $N_1$.
+\end{proof}
+
+\mathieu{mettre ici le détail de la  prop5}
+
+\noindent
+\textbf{Proposition~\ref{srp up down}.}
+{\em {\sc State-resilience},
+{\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}
+are undecidable for reset-VASS, hence for strongly compatible,
+$\post$-effective, $\pred$-basis-effective WSTS
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+}
+
+\begin{proof}
+Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
+{\sc $t$-liveness} is undecidable.
+Because
+{\sc $t$-liveness} is undecidable in reset-VASS
+and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
+Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
+reducible to {\sc $k$-state-resilience}:
+since $\Safe=\mathop{\uparrow} \Safe$ and
+$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
+compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
+and there exists $n_0 \in \N$ such that 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
+We compute 
+$n_0$, then iteratively check whether $k$-state-resilience 
+hold for $k$ from $0$ to $n_0$.  
+Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
+since 
+$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
+Hence the undecidability of {\sc Bounded-state-resilience}
+and 
+ {\sc $k$-state-resilience}. \qed
+\end{proof}
+
+\section{More details on VASS}\label{appendix vass}
+
+\mathieu{mettre ici le détail du Theo 13}
+
+\noindent
+\textbf{Theorem~\ref{vass down}.}
+{\em {\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+$\Safe = \mathop{\downarrow} \Safe$.
+}
 
@@
 
-\section{Conclusion and perspectives}
-
-We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
-We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
- $\Safe = \mathop{\uparrow} \Safe$ and in the case 
- $\Safe = \mathop{\downarrow} \Safe$.
-
-Several questions still remain.
-For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models. 
-One other angle of attack not studied here is the synthesis of sets 
-$\Safe$ for resilience,
-for instance, 
- finding the maximal upward-closed (resp. downward-closed) subset 
-$\Safe$ so that a system is $\Safe$-resilient. 
-Another question could be to analyse the resilience in the framework of a controller and its environment. 
-One could also extend upon the classes of set $\Safe$ 
- considered.
-As with semilinear sets for VASS, one could study resilience for sets defined in a boolean logic on upward and downward-closed subsets~\cite{DBLP:journals/fmsd/BertrandS13}.
-Finally, while we mention VASS, a more detailed analysis of the resilience problems could be also done for other computational models such as pushdown automata, one-counter automata or timed automata.
-
-\alain{remercier les reviewers !}
-
-\bibliography{bib}
-
-\appendix
-
-\section{Lemma technical proofs}
-
-Let us now prove more formally some of the lemmas and propositions ...
-
-\mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
-
-\noindent
-\textbf{Lemma~\ref{downward compatible}.}
-{\em For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.}
-
-\begin{proof}
-Let us prove $1 \implies 2$. 
-Let $D$ be a downward-closed subset of $S$
-and let $x \in \mathop{\downarrow} \pred^*(D)$.
-By downward closure, there exists
-$y \in \pred^*(D)$ 
-such that $x \leq y$.
-By definition of $\pred^*(D)$, there exists 
-$d \in D$ such that: (1) either $y=d$ and then $x \in D \subseteq \pred^*(D)$ or (2) there exist $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
-$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
-%
-By downward compatibility $a_0 \to a_1$
-implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
-$x \to^* a'_1$.
-More generally $a_i \to a_{i+1}$ and
-$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
-$a'_i \to^* a'_{i+1}$,
-and, by induction,
- $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
-with $d' \leq d$.
-Since
-$d'$ 
-belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\mathop{\downarrow}\pred^*(D) \subseteq \pred^*(D)$.
-Since $\pred^*(D) \subseteq \mathop{\downarrow} \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\mathop{\downarrow}\pred^*(D)$ is downward-closed.
-
-Let us prove $2 \implies 1$.
-Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\mathop{\downarrow} t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\mathop{\downarrow} t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\mathop{\downarrow} t_1)$.
-This means that there exists $t_2 \leq t_1$ such that
-$s_2 \rightarrow^{*} t_2$. Hence, we proved that $\mathscr{S}$ is downward compatible.
-\qed
-\end{proof}
-
-Additionnally ...
-
-\noindent
-\textbf{Lemma~\ref{Lemma intersection}.}
-{\em Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
-Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
-}
-
-\begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
-Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
-Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
-$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
-\end{proof}
-
-\section{Reset-VASS and Minsky-machines}\label{minsk}
-
-We now delve into more details into the undecidability results 
-upon which rely
-Proposition~\ref{indec WSTS with dcs}
-and
-Proposition~\ref{srp down up},
-namely, that
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
-are undecidable for Minsky machines
-and 
-for reset-VASS.
-
-The problems are undecidable for a Minsky machine with at least $3$ counters when
-$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
-We then consider a reset-VASS simulating the Minsky machine.
-We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way.
-% The simulation however is 
-Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0,0)$ in the Minsky machine, 
-this weak simulation will be enough to obtain undecidability of both problems.
-
-\begin{proposition}\label{both minsk indec}
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
-\end{proposition}
-
-\begin{proof}
-\alain{preuve pas claire, à réécrire}
-% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
-% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
-%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
-%which can be simulated by WSTS.
-First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-	This is due to the undecidability of $2$-counter Minsky machine termination.
-From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
-%			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
-Based on this construction, the problem of deciding whether the downward-closed set 
-$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
-% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
-Note that by chosing $s_0 \in Q_{M'} \times \mathop{\uparrow} (0,0,1)$,
-deciding whether the downward-closed set 
-$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable
-from every element of $\post^*(s_0) \cup Q_{M'} \times \mathop{\uparrow} (0,0,1)$ 
-is again equivalent to asking whether $M$ terminate on every input.
-Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
-
-% \alain{écrire qqchose ici}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
-% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
-%Reset-VASS extend the basic VASS model with special “reset
-%transitions” that set to $0$ some coordinates in the vector. 
-%			Let us recall their definition here.
-%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS in which
-there exists
- a sequence from 
-an initial state %$(s,x_1 \ldots x_n)$
-to
-a state $ q({\bf 0})$ if and only if
-there exists
-a sequence
- in the Minsky machine
- from a corresponding initial state
- to
- the state $q_f(0,0,0)$
-  (from Theorem~$5$ in \cite{araki1976PN}).
-  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
-Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
-{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
-and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
-\end{proof}
-
-\section{Reset-VASS and state-resilience}\label{appendix}
-
-\newcommand{\arrowcircle}[1][]{%
-  \begin{tikzpicture}[#1]
-    \draw[->] (0,0ex) -- (2em,0ex);
-    \draw (1em,0ex) circle (0.7ex);
-  \end{tikzpicture}%
-}
-
-Let us recall the {\em zero-reachability} problem in (reset-)VASS.
-
-\problemx{zero-reachability}
-{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
-{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
-
-The zero-reachability problem is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
-For a better representation, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS. 
-A \emph{reset Petri net} $N = (P, T, E, R, \mu)$ consists of a finite set of {\em places} $P = \{p_1, p_2, ...
-p_{|P|}\}$, a finite set of {\em transitions} $T = \{t_1, t_2, ..., t_{|T|} \}$, a finite set of {\em arcs} $E \subseteq (P \times T) \cup (T \times P)$, a finite set of
-{\em reset-arcs} $R \subseteq T \times P$, and an {\em initial marking} $\mu : P \to \N$.
-It is well-known that Petri nets and VASS are equivalents models with regards of the decidability of reachability problems: an $d$-VASS can be encoded into a Petri net with $d+2$ places and conversely, a Petri net with $p$ places can be encoded into a $p$-VASS~\cite{DBLP:journals/siglog/Schmitz16}.
-The same equivalence hold between reset-VASS and reset Petri nets.
-
-We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
-
-\noindent
-\textbf{Proposition~\ref{liveness reset}.}
-{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\
-
-We may then deduce that:\\
-
-\noindent
-\textbf{Corollary}
-{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
-
-\begin{proof}
-
-If we wish to determine if a marking in which every place is empty is reachable for a reset Petri net
- $N_1 = (P_1, T_1, E_1, R_1, \mu_1)$ 
- we construct a reset Petri net
- $N_2 = (P_2, T_2, E_2, R_2, \mu_2)$ 
-  which is live in $N_2$ if and only if the {\em zero marking}, which assigns $0$ to every place of $P_1$, is not reachable from $\mu_1$ in $N_1$. \\
-
-\begin{samepage}
-{\bf Construction of $N_2$} \\
-\indent
-The reset Petri net $N_2$ is constructed from $N_1$ by the addition of two places $p_a$ and $p_b$ and $|P_1| +2$ transitions $t_{p_a}$, $\{ t_p \mid p \in P_1 \}$ and $t_{p_b}$.
-We first modify all transitions of $N_1$ to include $p_a$ as both an input and an output.
-The initial marking $\mu_2$ will include a token in $p_a$ and no token in $p_b$. 
-This new place $p_a$ serve to mark that the run is ongoing.
-As long as it contains a token, the adjusted transitions of $N_1$ are live and can be used normally.
-Thus any marking which is reachable in the places of $N_1$ is also reachable in the corresponding places of $N_2$.
-We add an additional transition $t_{p_a}$ 
-which has $p_a$ as an input and a null output.
-This allows to disable the transitions of $N_1$ and to "freeze" the marking of the places of $P_1$ in $N_2$.
- The place $p_a$ and transition $t_{p_a}$ allow the net $N_2$ to reach any reachable marking in $N_1$ and then for $t_{p_a}$ to fire and freeze the net at that marking.
-We introduce a new place $p_b$ and new transitions $t_p$, for all $p \in P_1$, which have $p$ as input and $p_b$ as output.
-Lastly, we add a transition $t_{p_b}$ with $p_b$ as its output and every place of $P_2$ as output, which ``floods'' the net with tokens, assuring that every transition is live in $N_2$ if a token is ever
-put in $p_2$.
-\end{samepage}
-
- \begin{center}
-	\begin{figure}
-\includegraphics[width=0.75\textwidth]{FigurePN}
-\caption{Construction of $N_2$ from $N_1$.}
-	\end{figure}
-\end{center}
-
-Let us now prove that $t_{p_2}$ is live in $N_2$ if and only if 
-the zero marking
-is not reachable from $\mu_1$ in $N_1$. \\
-
-{\bf Suppose that the zero marking is reachable from $\mu_1$ in $N_1$, then $t_{p_2}$ is not live in $N_2$}.
-
-Indeed the marking with zero in every place of $P_1$ and in $p_b$ is reachable in $N_2$, by executing the same sequence of transition firings.
-Then $t_{p_a}$ can fire, leading to the marking which assigns zero to every place of $P_2$.
-From this marking the transitions $t_p$ are not live and neither are the transitions inherited from $N_1$ nor $t_{p_a}$, and, finally, nor is $t_{p_2}$.
-Thus $t_{p_2}$ is not live in $N_2$.
-\\
-
-{\bf Suppose that $t_{p_2}$ is not live in $N_2$, then the zero marking is reachable from $\mu_1$ in $N_1$}
-
-Indeed, if $t_{p_2}$ is not live in $N_2$, then a marking $\mu$ must be reachable in which $\mu(p_2) = 0$ and there is no reachable state in which $p_2$ has a token (in particular, since we do not allow token removal from $p_2$, the marking $\mu$ must be reached in a sequence of transitions that do not place any token in $p_2$).
-This means that no transition $t_p$ is live in $N_2$ in $\mu$ since any transition $t_p$ can place a token in $p_2$. 
-Thus, every place of $P_2$ inherited from $N_1$ must be devoid of token. Moreover, since the marking $\mu$ must be reached  in a sequence of transitions that do not place any token in $p_2$, 
-it can be reached without using the transitions $t_p$ or $t_{p_2}$.
-Since $t_{p_a}$ do not modify the places inherited from $P_1$ nor does it enable new transitions, the reachability of $\mu$ implies the reachability of a marking where every place of $P_2$ inherited from $P_1$ is devoid of token using only the transitions inherited from $N_1$.
-Thus the zero marking is reachable from $\mu_1$ in $N_1$.
-\end{proof}
-
-\mathieu{mettre ici le détail de la  prop5}
-
-\noindent
-\textbf{Proposition~\ref{srp up down}.}
-{\em {\sc State-resilience},
-{\sc Bounded-state-resilience} and
-{\sc $k$-state-resilience}
-are undecidable for reset-VASS, hence for strongly compatible,
-$\post$-effective, $\pred$-basis-effective WSTS
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-}
-
-\begin{proof}
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
-{\sc $t$-liveness} is undecidable.
-Because
-{\sc $t$-liveness} is undecidable in reset-VASS
-and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
-Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
-reducible to {\sc $k$-state-resilience}:
-since $\Safe=\mathop{\uparrow} \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
-compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
-We compute 
-$n_0$, then iteratively check whether $k$-state-resilience 
-hold for $k$ from $0$ to $n_0$.  
-Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
-since 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
-Hence the undecidability of {\sc Bounded-state-resilience}
-and 
- {\sc $k$-state-resilience}. \qed
-\end{proof}
-
-\section{Thm 9}
-
-\mathieu{mettre ici le détail du Theo 9}
-
-\section{More details on VASS}
-
-\mathieu{mettre ici le détail du Theo 13}
-
 \end{document}

### 2023-10-24 VMCAI-revised-version/final.tex
up
effectivity

@@
 \end{center}
+
+	\vspace{-1.2cm}
 
@@
 $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
-Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96, DBLP:journals/iandc/CeceFI96}.
+Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96,DBLP:journals/iandc/CeceFI96}.
 
@@
 
-\alain{remercier les reviewers !}
+\alain{remercier les reviewers !} We thank the reviewers of the VMCAI 2024 conference for their numerous and relevant comments and improvement suggestions.
 

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

@@
 
-\alain{remercier les reviewers !} We thank the reviewers of the VMCAI 2024 conference for their numerous and relevant comments and improvement suggestions.
+\alain{remercier les reviewers !} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.
 
@@
 %which can be simulated by WSTS.
-First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
-	This is due to the undecidability of $2$-counter Minsky machine termination.
-From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
+First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
+%	This is due to the undecidability of termination for $2$-counter Minsky machines.
+From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$
+such that $M$ terminates on all initial states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
 %			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
-Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$.
-Based on this construction, the problem of deciding whether the downward-closed set 
-$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable in the $3$-counter Minsky machine $M'$
+We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
+%
+Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
+%
+Based on this construction, the downward-closed set 
+$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable (in $M'$)
 % \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable~\cite{Min61}. 
-Note that by chosing $s_0 \in Q_{M'} \times \mathop{\uparrow} (0,0,1)$,
+from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates on all initial states; this last problem is undecidable~\cite{Min61}. 
+Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
 deciding whether the downward-closed set 
-$Q_{M'} \times \N^3 \setminus Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is reachable
-from every element of $\post^*(s_0) \cup Q_{M'} \times \mathop{\uparrow} (0,0,1)$ 
-is again equivalent to asking whether $M$ terminate on every input.
+$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable
+from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
+is again equivalent to asking whether $M$ terminates on every initial state.
 Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.

### 2023-10-24 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{proof}
-\alain{preuve pas claire, à réécrire}
+\alain{j'ai commencé à  réécrire mais je n'ai pas fini car il y a des choses non expliquées que je ne suis pas sûr de comprendre...}
 % \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
@@
 First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
+\alain{pourquoi prendre 3 compteurs et pas 2 ? On ne comprend pas}
 %	This is due to the undecidability of termination for $2$-counter Minsky machines.
-From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$
-such that $M$ terminates on all initial states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
+From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$.
+We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
+
+We observe that $M$ terminates from all (initial) states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
+\alain{pourquoi  ? On ne comprend pas}
 %			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
 %
 Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
+\alain{pourquoi   $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ ?}
+\alain{pourquoi  $Q' \times \mathop{\uparrow}(0,0,1)$ et pas $Q' \times \mathop{\uparrow}(0,0,0)$? On ne comprend pas}
+%
+\alain{l'ensemble $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ est redondant car il est égal à $Q' \times \{(m,n,0) \mid m,n \in \N\}=Q' \times \N^2 \times \{0\}$}
+%
+\alain{les ensembles $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ et $Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ sonts les mêmes et égaux à $Q' \times \N^2 \times \{0\}$.
+Pourquoi avoir deux notations différentes ??
+ça complique inutilement la lecture.
+Et prendre la troisième, la plus simple $Q' \times \N^2 \times \{0\}$ qui n'apparait pas...}
+%
 %
@@
 % \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates on all initial states; this last problem is undecidable~\cite{Min61}. 
+from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates from all (initial) states;
+%
+\alain{a déjà été dit qq lignes avant}
+%
+ this last problem is undecidable~\cite{Min61}. 
 Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
@@
 from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
-is again equivalent to asking whether $M$ terminates on every initial state.
+%
+\alain{n'a-t-on pas $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1) = Q' \times \mathop{\uparrow} (0,0,1)$ ?
+a-t-on $Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1)) = Q' \times  (0,0,0)$ ?
+si oui pourquoi faire compliqué ?}
+%
+is again equivalent to asking 
+%
+\alain{asking: non}
+%
+whether $M$ terminates on every initial state.
+%
+\alain{je trouve ça pas clair}
+%
 Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
@@
 a state $ q({\bf 0})$ if and only if
+%
+\alain{attention notations incohérentes pour (q,0,0,0) et $ q({\bf 0})$}
+%
 there exists
@@
   Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
+%
+\alain{désolé, pas clair}
+%
 Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 

### 2023-10-25 VMCAI-revised-version/final.tex
up

--- 2023-10-24 VMCAI-revised-version/final.tex
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
-$\Safe=\mathop{\downarrow} \Safe$.
+{\sc Resilience} is undecidable for reset-VASS when $\Safe=\mathop{\downarrow} \Safe$, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
 \end{proposition}
@@
 \begin{proof}
-First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
-$\Safe = \mathop{\downarrow} \Safe$,
-since it is possible to guarantee that state $q_f(0,0,0)$ is the only element 
-% of $Q_{M} \times \mathop{\downarrow} (1,1,0)$
-potentially reachable from 
-$\Bad = Q_{M} \times \mathop{\uparrow} (0,0,1)$ that is not already inside $\Bad$.
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
-% from  a state $p({\bf u})$
-% to
-reaching
- a state 
-%
-$ q({\bf 0})$ corresponds to one that reaches 
-$q_f(0,0,0)$
- in the Minsky machine and vice versa.
- Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
- By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
- we obtain that 
- {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
+\nocite{araki1976PN,dufourd1998reset}
+{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
+deciding structural termination of reset-VASS, which is undecidable.
+\qed \end{proof}
+\mathieu{GOTO}
 
@@
 %
-Unfortunately {\sc State-resilience} is undecidable in this case.
-As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the corresponding problem in Minsky machines.
-Indeed, {\sc State-resilience} is undecidable for a Minsky machine $M$ with at least $3$ counters when
-$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
- As seen in 
-Proposition~\ref{indec WSTS with dcs},
-%\alain{la Proposition~\ref{indec WSTS with dcs} n'apparait que dans les annéexes...}
-%\mathieu{Elle apparaît dans le corps du texte aussi mais j'avais utilisé deux fois le même label ce qui est corrigé maintenant.}
- the sequences of $M$ reaching a $\Safe$ state from a $\Bad$ one can be simulated in a reset-VASS
-by sequences reaching a $\Safe'$ state from a $\Bad'$ one,
- and vice versa, hence the following result. 
-% \alain{trop rapide}
-%\mathieu{tjr trop rapide ?}
-%\alain{ça va}
+Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the 
+structural termination of reset-VASS.
+\mathieu{GOTO}
 
 \begin{proposition}\label{srp down up}
-{\sc State-resilience} is undecidable for reset-VASS, hence also for
-%$\post$-ideal-effective ???? and
+{\sc State-resilience} is undecidable for reset-VASS when
+$\Safe=\mathop{\downarrow} \Safe$, hence also for
 $\post$-effective,
- $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility
-when
-$\Safe=\mathop{\downarrow} \Safe$.
+ $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
  \end{proposition}
@@
 \begin{proof}
-Consider a given $k \in \N$,
- $\emptyset \neq \Bad \subseteq \N^d$ is upward-closed and $\Safe \subseteq \N^d$ is downward-closed, and
+Consider a given $k \in \N$, $\Safe \subseteq \N^d$ is downward-closed
+ such that its upward-closed complement $\Bad \subseteq \N^d$ is nonempty, and
 consider a given VAS $V$. 
@@
 
-\section{Reset-VASS and Minsky-machines}\label{minsk}
-
-We now delve into more details into the undecidability results 
-upon which rely
-Proposition~\ref{indec WSTS with dcs}
-and
-Proposition~\ref{srp down up},
-namely, that
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
-are undecidable for Minsky machines
-and 
-for reset-VASS.
-
-The problems are undecidable for a Minsky machine with at least $3$ counters when
-$\Safe$ is downward-closed, since it is possible to guarantee that state $q_f(0,0,0)$ is the only $\Safe$ element potentially reachable from $\Bad$.
-We then consider a reset-VASS simulating the Minsky machine.
-We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way.
-% The simulation however is 
-Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0,0)$ in the Minsky machine, 
-this weak simulation will be enough to obtain undecidability of both problems.
-
-\begin{proposition}\label{both minsk indec}
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
-\end{proposition}
-
-\begin{proof}
-\alain{preuve pas claire, à réécrire}
-% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
-% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
-%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
-%which can be simulated by WSTS.
-First, let us show that {\sc Resilience} is undecidable for $3$-counter machines.
-%	This is due to the undecidability of termination for $2$-counter Minsky machines.
-From an $2$-counter machine $M=(Q, q_f,...)$ where $Q$ is the set of control-states and $q_f \in Q$ is the final control-state, one constructs an associated $3$-counter machine $M'=(Q', q'_f,...)$ with a final control-state $q_f' \in Q'$
-such that $M$ terminates on all initial states iff $q_f'(0,0,0)$ is reachable (in $M'$) from any state in $Q' \times \mathop{\uparrow} (0,0,1)$. 
-%			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then $M'$ decreases its first two counters until they reach $0$, then, and only then, finally $M'$ decreases the third counter until it reaches $0$.
-%
-Remark that $Q' \times \{ (m,0,0), (0,n,0),(m,n,0) \mid m,n \in \N\}$ is not reachable from any state in $Q' \times \mathop{\uparrow}(0,0,1)$.
-%
-Based on this construction, the downward-closed set 
-$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable (in $M'$)
-% \alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
-from any state in $Q' \times \mathop{\uparrow}(0,0,1)$ iff $M$ terminates on all initial states; this last problem is undecidable~\cite{Min61}. 
-Note that by chosing $s_0 \in Q' \times \mathop{\uparrow} (0,0,1)$,
-deciding whether the downward-closed set 
-$Q' \times (\N^3 \setminus \mathop{\uparrow}(0,0,1))$ is reachable
-from every state in $\post^*(s_0) \cup Q' \times \mathop{\uparrow} (0,0,1)$ 
-is again equivalent to asking whether $M$ terminates on every initial state.
-Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
-
-% \alain{écrire qqchose ici}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
-% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
-%Reset-VASS extend the basic VASS model with special “reset
-%transitions” that set to $0$ some coordinates in the vector. 
-%			Let us recall their definition here.
-%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS in which
-there exists
- a sequence from 
-an initial state %$(s,x_1 \ldots x_n)$
-to
-a state $ q({\bf 0})$ if and only if
-there exists
-a sequence
- in the Minsky machine
- from a corresponding initial state
- to
- the state $q_f(0,0,0)$
-  (from Theorem~$5$ in \cite{araki1976PN}).
-  Moreover, the reset-VASS is able to emulate zero tests on the counters, hence it is able to guarrantee that $ q({\bf 0})$ is its only $\Safe$ state reachable from a $\Bad$ state as well.
-Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
-{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
-and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed
-%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
-\end{proof}
-
 \section{Reset-VASS and state-resilience}\label{appendix}

### 2023-10-25 VMCAI-revised-version/final.tex
up

@@
 
-%Integer VASS or 
-$\mathbb{Z}^d-$VASS (resp. $\mathbb{Z}-$VASS)~\cite{DBLP:conf/rp/HaaseH14} are $d$-VASS (resp. VASS) that are allowed to take values from the integers.
-
-\begin{theorem}\label{SL VASS}
-{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
-\end{theorem}
-
-\begin{proof}
-We consider the case where $\Safe$ is semilinear
-and $\Bad = S \setminus \Safe$ is semilinear too.
-We want to use the fact that, for VASS,
-it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
-$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
-{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
-We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
-Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
-\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
-and hence, {\sc Resilience} is decidable. 
-The reachability relation of 
-%integer $d$-VASS
- $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
-\qed
-\end{proof}
-
-{\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
-Recall that lossy counter machines (LCM)~\cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
-
-Resilience and the home-space problem are also linked to the 
-model-checking of basic reachability and safety formulae. 
-In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
-is decidable for lossy counter machine
-when $X$ and $Y$ are semi-linear sets.
-Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
- t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
-whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
-``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
-  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
-
-\begin{theorem}\label{SL Lossy}
-{\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
-\end{theorem}
-
-\begin{proof}
-We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
-Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
-\qed
-\end{proof}
-
-$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
-
-\begin{theorem}\label{RP VASS}
-{\sc Resilience} is decidable for continuous VASS when $\Safe$  is
-definable in the existential theory of the rationals with addition and order.
-\end{theorem}
-
-\begin{proof}
-The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
-the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). \qed
-\end{proof}
-
-Remark that
-Theorem~\ref{SL Lossy} 
-(resp. Theorem~\ref{RP VASS})
-did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
-semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
-the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
-
-The above results concern only unbounded resilience. We consider now 
-{\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of
-the path from $\Bad$ to $\Safe$
-when $\Safe$ is downward-closed. Surprisingly, we have the following result for VAS (not VASS):
+Surprisingly, when $\Safe$ is downward-closed, we have the following result for VAS (not VASS):
 \begin{proposition}
@@
 
+Let us now extend decidability of {\sc Resilience} for semilinear sets and VASS variants.
+%
+%Integer VASS or 
+$\mathbb{Z}^d-$VASS (resp. $\mathbb{Z}-$VASS)~\cite{DBLP:conf/rp/HaaseH14} are $d$-VASS (resp. VASS) that are allowed to take values from the integers.
+
+\begin{theorem}\label{SL VASS}
+{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We consider the case where $\Safe$ is semilinear
+and $\Bad = S \setminus \Safe$ is semilinear too.
+We want to use the fact that, for VASS,
+it is decidable whether $\post^*(X) \subseteq \pred^*(Y)$ when 
+$X$ and $Y$ are both semilinear sets~\cite{DBLP:journals/corr/abs-2207-02697}.
+{\sc Resilience} asks whether $\Bad \subseteq \pred^*(\Safe)$.
+We have $\post^*(\Bad) \setminus \Safe = \Bad$ by extension of $S \setminus \Safe = \Bad$.
+Thus $\post^*(\Bad) = \Bad \cup (\Safe \cap \post^*(\Bad))$. Since $(\Safe \cap \post^*(\Bad)) \subseteq \pred^*(\Safe)$, we have
+\[\post^*(\Bad) \subseteq \pred^*(\Safe) \quad \text{iff} \quad \Bad \subseteq \pred^*(\Safe).\]
+and hence, {\sc Resilience} is decidable. 
+The reachability relation of 
+%integer $d$-VASS
+ $\mathbb{Z}^d-$VASS is definable by a well-formed formula with no free variables in the first-order theory of the integers with addition and orders (Presburger arithmetic) in $\mathbb{Z}^{2d}$, hence	$\Bad \subseteq \pred^*(\Safe)$ is decidable when $\Safe$ and $\Bad$ are semilinear sets.
+\qed
+\end{proof}
+
+{\sc Resilience} is also decidable for other classes of counter machines for which the reachability relation can be expressed in a decidable logic.
+Recall that lossy counter machines (LCM)~\cite{DBLP:conf/rp/Schnoebelen10} are counter machines that may loose tokens in each control-state.
+
+Resilience and the home-space problem are also linked to the 
+model-checking of basic reachability and safety formulae. 
+In particular~\cite{DBLP:conf/rp/Schnoebelen10} shows that the ``from-all'' formula $\forall s \in X~ \exists t \in Y~ s \to^* t$
+is decidable for lossy counter machine
+when $X$ and $Y$ are semi-linear sets.
+Other decidable formulae include ``one-to-one'' ($\exists s \in  X ~ \exists t \in  Y ~ s \to^* 
+ t$), and ``all-to-same'' ($\exists t \in  Y ~ \forall s \in  X ~ s \to^*  t$),
+whereas ``one-to-all'' ($\exists s \in  X ~ \forall t \in  Y ~ s \to^*  t$), 
+``all-to-all'' ($\forall s \in  X ~ \forall t \in  Y ~ s \to^*  t$)
+  and ``to-all'' ($\forall t \in  Y  ~ \exists s \in  X ~ s \to^*  t$) are undecidable (again, for LCM).
+
+\begin{theorem}\label{SL Lossy}
+{\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
+\end{theorem}
+
+\begin{proof}
+We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
+Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
+\qed
+\end{proof}
+
+$\mathds{Q}^d$-VASS~\cite{DBLP:journals/tocl/BlondinFHH17} (or continuous VASS) are a relaxation of classical (discrete) $d$-VASS in which transitions can be fired a fractional number of times, and consequently counters may contain a fractional number of tokens.
+
+\begin{theorem}\label{RP VASS}
+{\sc Resilience} is decidable for continuous VASS when $\Safe$  is
+definable in the existential theory of the rationals with addition and order.
+\end{theorem}
+
+\begin{proof}
+The reachability relation of continuous VASS is definable by a sentence of linear size in the existential theory of
+the rationals with addition and order whose complexity is EXPSPACE \cite{DBLP:journals/tocl/BlondinFHH17}. Hence, $\Bad \subseteq \pred^*(\Safe)$ is decidable (and also in EXPSPACE). \qed
+\end{proof}
+
+Remark that
+Theorem~\ref{SL Lossy} 
+(resp. Theorem~\ref{RP VASS})
+did not make use of the hypothesis that $\Bad$ is the complement of $\Safe$, simply relying on the
+semilinearity of the set (resp. its definability in the existential theory of the rationals with addition and order). Thus 
+the mentionned theorems still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+
+% The above results concern only unbounded resilience.
+We consider now  {\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of the path from $\Bad$ to $\Safe$ when $\Safe$ is downward-closed.
+
 \section{Conclusion and perspectives}

### 2023-10-25 VMCAI-revised-version/final.tex
up

@@
 In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
+
+The structural termination problem consists in deciding if the termination problem holds for  for every initial state.
+Since termination is w.l.o.g.
+reachability of $q_f({\bf 0})$ for some control state $q_f$, we can formulate the structural termination problem as
+
+\problemx{Structural termination}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, a control state $q_f \in Q$.}
+{ $Q \times \longrightarrow^{*} q_f({\bf 0})$ ? \\}
 

### 2023-10-25 VMCAI-revised-version/final.tex
up

@@
 {A reset-VASS $V=(Q,T)$ of dimension $d$, a control state $q_f \in Q$.}
-{ $Q \times \longrightarrow^{*} q_f({\bf 0})$ ? \\}
+{ $Q \times \N^d \longrightarrow^{*} q_f({\bf 0})$ ? \\}
 

### 2023-10-25 VMCAI-revised-version/final.tex
up

--- 2023-10-25 VMCAI-revised-version/final.tex
@@
 
-In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
+In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
+{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
+deciding structural termination of reset-VASS, which is undecidable.
+
+Indeed, the structural termination problem consists in deciding if %the termination problem holds for 
+a system terminates on every initial state. 
+Since termination is w.l.o.g.
+reachability of $q_f({\bf 0})$ \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g.
+??} for some control state $q_f$, we can formulate the structural termination problem as
+
+\problemx{Structural termination}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, a control state $q_f \in Q$.}
+{ $Q \times \N^d \longrightarrow^{*} q_f({\bf 0})$ ? \\}
+
+The undecidability of which leads to the following proposition.
+
+\mathieu{GOTO}
 
@@
 
-\begin{proof}
 \nocite{araki1976PN,dufourd1998reset}
+\iffalse
+\begin{proof}
 {\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
@@
 \qed \end{proof}
-\mathieu{GOTO}
+\fi
 

### 2023-10-25 VMCAI-revised-version/final.tex
up

@@
 a system terminates on every initial state.
-Since termination is w.l.o.g.
-reachability of $q_f({\bf 0})$ \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g.
-??} for some control state $q_f$, we can formulate the structural termination problem as
 
 \problemx{Structural termination}
-{A reset-VASS $V=(Q,T)$ of dimension $d$, a control state $q_f \in Q$.}
-{ $Q \times \N^d \longrightarrow^{*} q_f({\bf 0})$ ? \\}
+{A reset-VASS $V=(Q,T)$ of dimension $d$.}
+{ %$Q \times \N^d \longrightarrow^{*} q_f({\bf 0})$ ? 
+Do all runs for every initial state terminate ?\\}
 
@@
 
-\mathieu{GOTO}
+% \mathieu{GOTO}
 
@@
 
+\begin{proof}
 \nocite{araki1976PN,dufourd1998reset}
-\iffalse
-\begin{proof}
 {\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
-deciding structural termination of reset-VASS, which is undecidable.
-\qed \end{proof}
-\fi
+deciding structural termination of reset-VASS,
+since termination is w.l.o.g. reachability of $q_f({\bf 0})$ 
+ \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g. ??} 
+for some control state $q_f$. \qed \end{proof}
 
@@
 % 
-Remark that the quantification over the path is existential rather than universal i.e. 
-{\em zero-reachability}
-does not asks that zero remains reachable from an initial state $s_0$ at all times but rather, that there exists at least one way to reach it.
+{\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
+{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q(\textbf{0})$,
+which could be unreachable from other reachable states. 
+%
 Despite this it is possible to reduce 
@@
 structural termination of reset-VASS.
-\mathieu{GOTO}
+%\mathieu{GOTO}
 

### 2023-10-25 VMCAI-revised-version/final.tex
effectivity

@@
 %			If the home-space contains a single element, this element is an {\em home-state}.
-Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$ (i.e., $\post^*(X) \subseteq \pred^*(H)$).
+Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$
+%				(i.e., $\post^*(X) \subseteq \pred^*(H)$).
 %
 A transition system is {\em resilient for $(\Bad,\Safe)$} if $(\Bad,\Safe)$ satisfies the home-space property.
-A transition system is {\em resilient for $\Safe$} if it is resilient for $(\Bad,\Safe)$ with $\Bad=S \setminus \Safe$.
-It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ with $\Bad=\{s_0\}$.
+A transition system is {\em resilient for $\Safe$} if it is resilient for $(\Bad,\Safe)$ with $\Bad$ is the complementary to $ \Safe$.
+It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ where $\Bad$ contains an unique state.
 %
@@
 %		whether $\Bad \subseteq \pred^*(\Safe)$.
-The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ (i.e., $S \subseteq \pred^{\leq k}(\Safe)$) and
+The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ 
+%
+%		(i.e., $S \subseteq \pred^{\leq k}(\Safe)$)
+%
+ and 
 the {\em bounded resilience} problem is to decide whether there exists an $k$ such that the system is $k$-resilient.
@@
 
-In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$ (i.e., $\mathop{\uparrow} \post^*(s)$) when $\Safe$ is upward-closed.
+In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$
+%
+%			(i.e., $\mathop{\uparrow} \post^*(s)$) 
+%
+when $\Safe$ is upward-closed.
 %
@@
 %
-\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of $\mathop{\uparrow} \post^*(s)$ for all state $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
+\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
+%
+%			$\mathop{\uparrow} \post^*(s)$ 
+%
+for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
 However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.

### 2023-10-25 VMCAI-revised-version/final.tex
effectivity

@@
 
-In 1986, Memmi and Vautherin introduced the (restricted) notion of home-space~\cite{DBLP:conf/ac/MemmiV86} for $X$ a singleton.
+In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} in Petri nets for $X$ a singleton.
 In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}).
@@
 \noindent
-{\bf Our contributions}
+{\bf Our main contributions}
 \begin{itemize}
@@
 We show that resilience and {state-resilience} (i.e.
-$X=\{s_0\}$) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed.
-Moreover, {state-resilience},
-{bounded-state-resilience} and
-{$k$-state-resilience}
-are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
-$\Safe$ is upward-closed. 
+$X$ a singleton) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed.
+%Moreover, {state-resilience},
+%{bounded-state-resilience} and
+%{$k$-state-resilience}
+%are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
+%$\Safe$ is upward-closed. 
 %			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-\item When $\Safe$ is upward-closed, the three resilience problems are decidable for both post-ideal-effective WBTS with strong (upward) compatibility and for effective pred-basis WSTS with strong (upward) compatibility.
-Moreover, the resilience problem is decidable for post-ideal-effective downward-WSTS with
-$\Safe$ upward-closed
-%and
+\item The three resilience problems are decidable for post-ideal-effective Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
+%
+\item  Resilience is also decidable, when 
+$\Safe$ is upward-closed, for effective pred-basis WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
+%
+%    and
 %the additional hypothesis that
@@
 %
+%				and for post-ideal-effective downward-WSTS ????
 %%%
 %
-\item We clarify the different effectiness conditions that allow decidability of the six resilience.
-%
-\item We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
+\item We clarify the different effectiness conditions on WBTS and WSTS that allow the decidability of the six resilience.
+%
+\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} ; we also show that relaxing some hypothesis leads to undecidability.
+%
+%			à utiliser dans le texte ???
+%
+%We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
 We show that removing the effective
-$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.
-We still extend  \cite{DBLP:journals/corr/PrasadZ16} as follows: the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities ({$k$-state-resilience} and {bounded-state-resilience} are decidable for post-ideal-effective WSTS with strong downward compatibility).
-%
-\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
-%
-%			$\mathop{\uparrow} \post^*(s)$ 
-%
-for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
-However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
+%$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.
+
+\item We show that the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities (Theorem \ref{downward srp} extends Theorem 1 in \cite{DBLP:journals/corr/PrasadZ16}).
+The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WSTS with strong downward compatibility.
+%
+%			à utiliser dans le texte ???
+%
+%\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
+%%
+%%			$\mathop{\uparrow} \post^*(s)$ 
+%%
+%for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
+%However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
 %
@@
 \end{itemize}
+
+Most of our results about the different models and different types of resilience are in Figure YYYY.
 
@@
 
+\alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
+
 \section{Well-structured transition systems and VASS}\label{section definitions}
@@
 
+\alain{il manque des refs dans la fig 1 qui sont déjà dans la biblio: WBTS, omega2,on peut mettre les front lossy avec les LCM et LCS}
+%
 \section{Resilience for WSTS}
@@
  \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g. ??} 
+\alain{qu'entends-tu par deterministic ici ?}
 for some control state $q_f$. \qed \end{proof}
@@
 
-\alain{remercier les reviewers !} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.
+\section*{Acknowledgements} We express our thanks to the reviewers of the VMCAI 2024 Conference for their numerous and relevant comments and improvement suggestions.
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 This all changes when one considers VASS (having control-states).
-For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
 then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
@@
 
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with state $q$ \----
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
 then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
@@
 \fi
+The set $\Bad$ admits a finite basis 
+$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
+q_{i_m}(\textbf{v}_{i_m}) \}$.
+Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
+and then consider the vector
+$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
+For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
+of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
+
+For all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
+then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
+If there is none, then
+$\mathop{\uparrow} q(\textbf{0})$
+ is in the complement of $\Bad$, i.e. $\Safe$.
+
+From then on, we can split $\Safe$ into two subsets: an upward-closed 
+union of sets $\mathop{\uparrow} q(\textbf{0})$ and a remaining downward-closed subset.
+The key now is that elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps.
 \mathieu{Write small text and refer to appendix~\ref{appendix vass}}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 
-From then on, we can split $\Safe$ into two subsets: an upward-closed 
-union of sets $\mathop{\uparrow} q(\textbf{0})$ and a remaining downward-closed subset.
-The key now is that elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps.
+From then on, we partition $\Safe$ into two subsets: an upward-closed 
+union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
+The key now is that
+there is only a finite number of elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} \leq \textbf{u}_k$ for all of which we will check reachability of $\Safe$ in $k$ or less steps, while 
+ elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps.
 \mathieu{Write small text and refer to appendix~\ref{appendix vass}}

### 2023-10-26 VMCAI-revised-version/final.tex
up
corrigé une erreur dans le tableau

@@
 \begin{center}
-\resizebox{\textwidth}{!}{
+ \resizebox{1.2\textwidth}{!}{
 \begin{tabular}{ | l | c | c | c | r |}
@@
 		&   \\\cline{2-3}
-~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}  & ?    \\ \hline
+~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS}  & ?    \\ \hline
    \multirow{3}{*}{kRP} 

### 2023-10-26 VMCAI-revised-version/final.tex
Added new tentative table for results

@@
 \alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
+
+\begin{center}
+ \resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \
+		& WBTS model (with $\Safe = {\uparrow \Safe}$)~ 
+		 & Decidability ~ 
+ \\ \hline
+   \multirow{3}{*}{RP} 
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
+   		 &    \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		&  \\\cline{2-3}
+~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & Undecidable    \\ \hline
+   \multirow{3}{*}{BRP} 
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
+   		 & \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		&   \\\cline{2-3}
+~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS}  & Undecidable    \\ \hline
+   \multirow{3}{*}{kRP} 
+ 	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
+   		 & \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		& \\\cline{2-3}
+~	& ? & Undecidable   \\ \hline
+
+   \multirow{2}{*}{sRP} 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} NON
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable   \\ \hline
+   \multirow{2}{*}{BsRP} 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
+   \multirow{2}{*}{ksRP} 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
+
+\begin{center}
+ \resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \ 
+		 & WBTS model (with $\Safe = {\downarrow \Safe}$)~ 
+		 & Decidability
+ \\ \hline
+   \multirow{3}{*}{RP} 
+   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
+   		 	& \multirow{2}{*}{Decidable}  \\
+
+		& VASS & \\\cline{2-3}
+		& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs} 
+		& Undecidable   \\ \hline
+   \multirow{2}{*}{BRP} 
+   		 & VASS 
+   		 & Decidable   \\\cline{2-3}
+~	&  ? & Undecidable    \\ \hline
+   \multirow{2}{*}{kRP} 
+~	& VASS
+		& Decidable \\\cline{2-3}
+~	& ? & Undecidable    \\ \hline
+
+   \multirow{2}{*}{sRP} 
+   		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
+~	& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
+   \multirow{2}{*}{BsRP} 
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
+~	&  ? & Undecidable   \\ \hline
+   \multirow{2}{*}{ksRP} 
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
+~	 & ? & Undecidable    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

--- 2023-10-26 VMCAI-revised-version/final.tex
@@
 \begin{center}
- \resizebox{1.2\textwidth}{!}{
+ \resizebox{\textwidth}{!}{
 \begin{tabular}{ | l | c | c | c | r |}
@@
 \alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
+
+\begin{center}
+ \resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \
+		& {\bf WBTS model (with $\Safe = \mathop{\uparrow} \Safe$) }~ 
+		 & ~ {\bf Decidability}  ~ 
+ \\ \hline
+   \multirow{3}{*}{ {\bf RP }} 
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
+   		 &    \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		&  \\\cline{2-3}
+~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & Undecidable    \\ \hline
+   \multirow{3}{*}{{\bf BRP }} 
+  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
+   		 & \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		&   \\\cline{2-3}
+~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS}  & Undecidable    \\ \hline
+   \multirow{3}{*}{ {\bf kRP } } 
+ 	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
+   		 & \multirow{2}{*}{Decidable}  \\ 
+~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
+		& \\\cline{2-3}
+~	& ? & Undecidable   \\ \hline
+
+   \multirow{2}{*}{ {\bf sRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} NON
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable   \\ \hline
+   \multirow{2}{*}{ {\bf BsRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
+   \multirow{2}{*}{ {\bf ksRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
+   		 & Decidable \\ \cline{2-3}
+~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
+
+\begin{center}
+ \resizebox{\textwidth}{!}{
+\begin{tabular}{ | l | c | c | c | r |}
+\hline   \
+		& {\bf WBTS model (with $\Safe = \mathop{\downarrow} \Safe$)} ~ 
+		 & ~ {\bf Decidability}  ~ 
+ \\ \hline
+   \multirow{3}{*}{ {\bf RP }} 
+   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
+   		 	& \multirow{2}{*}{Decidable}  \\
+
+		& VASS & \\\cline{2-3}
+		& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs} 
+		& Undecidable   \\ \hline
+   \multirow{2}{*}{{\bf BRP }} 
+   		 & VASS 
+   		 & Decidable   \\\cline{2-3}
+~	&  ? & Undecidable    \\ \hline
+   \multirow{2}{*}{{\bf kRP }} 
+~	& VASS
+		& Decidable \\\cline{2-3}
+~	& ? & Undecidable    \\ \hline
+
+   \multirow{2}{*}{{\bf sRP }} 
+   		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
+~	& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
+   \multirow{2}{*}{{\bf BsRP }} 
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
+~	&  ? & Undecidable   \\ \hline
+   \multirow{2}{*}{{\bf ksRP }} 
+   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
+~	 & ? & Undecidable    \\ \hline
+
+ \end{tabular}
+  }
+\end{center}
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
  elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps.
-\mathieu{Write small text and refer to appendix~\ref{appendix vass}}
-\qed
+Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed, in particular the set of
+ elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$
+ and the set of elements from which the upward-closed subset of $\Safe$ is reachable
+ are both upward-closed sets a basis of which we can construct, and hence can be checked for inclusion. We refer to appendix~\ref{appendix vass} for more details.\qed
 \end{proof}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 is stationnary. 
+\mathieu{
+We check the finiteness of the set by comparing the basis of $\Bad$ and
+$(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+}
 If the set is finite then we stop the procedure once we have checked for every element

### 2023-10-26 VMCAI-revised-version/final.tex
up

--- 2023-10-26 VMCAI-revised-version/final.tex
@@
 is stationnary. 
+\mathieu{
+We check the finiteness of the set by comparing the finite basis of $\Bad$ and
+$(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+}
 If the set is finite then we stop the procedure once we have checked for every element

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
    \multirow{2}{*}{sRP} 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} NON
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} 
    		 & WSTS (downward + upward) Thm~9 \\ \cline{2-3}
@@
    \multirow{2}{*}{ {\bf sRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} NON
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} 
    		 & Decidable \\ \cline{2-3}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 
+\iffalse
 \begin{center}
@@
 \end{center}
+\fi
 
@@
 		& \\\cline{2-3}
-~	& ? & Undecidable   \\ \hline
-
-   \multirow{2}{*}{ {\bf sRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} 
-   		 & Decidable \\ \cline{2-3}
+~	&  & Undecidable   \\ \hline
+
+   \multirow{3}{*}{ {\bf sRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective)  
+   		 & \multirow{2}{*}{Decidable} \\
+   	& Thm~\ref{post srp} &	 \\ \cline{2-3}
 ~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable   \\ \hline
-   \multirow{2}{*}{ {\bf BsRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
-   		 & Decidable \\ \cline{2-3}
+   \multirow{3}{*}{ {\bf BsRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) 
+   		 & \multirow{2}{*}{Decidable} \\
+   	& Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+   	& \\ 	  \cline{2-3}
 ~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
-   \multirow{2}{*}{ {\bf ksRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
-   		 & Decidable \\ \cline{2-3}
+   \multirow{3}{*}{ {\bf ksRP } } 
+  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective)
+   		 & \multirow{2}{*}{Decidable} \\
+   	&  Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  		& \\	  \cline{2-3}
 ~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
@@
 		& VASS & \\\cline{2-3}
-		& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs} 
+		& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}~ 
 		& Undecidable   \\ \hline
@@
    		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
-~	& Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
+~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down} ~ & Undecidable  \\ \hline
    \multirow{2}{*}{{\bf BsRP }} 
@@
 
-Let us now prove more formally some of the lemmas and propositions ...
-
+We have omitted proofs for Lemma~\ref{downward compatible} and Lemma~\ref{Lemma intersection}
+up until now
 \mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 		& VASS & \\\cline{2-3}
-		& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}~ 
+		& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs} 
 		& Undecidable   \\ \hline
@@
 
-   \multirow{2}{*}{{\bf sRP }} 
+   \multirow{2}{*}{{\bf sRP ~ ~ }} 
    		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
-~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down} ~ & Undecidable  \\ \hline
+~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
    \multirow{2}{*}{{\bf BsRP }} 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 		& \\\cline{2-3}
-~	&  & Undecidable   \\ \hline
+~	& ? & Undecidable   \\ \hline
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 We have omitted proofs for Lemma~\ref{downward compatible} and Lemma~\ref{Lemma intersection}
-up until now
-\mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
+up until now since these are fairly straightforward technical proofs.
+% \mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
+We provide now both. \\
 
@@
 
-Additionnally ...
+% Additionnally ...
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 {\sc $t$-liveness} is undecidable.
-By reduction to {\sc State-resilience}, and straightforward reduction to the other two state resilience problems, we 
-deduce the following.
+We deduce undecidability by reduction to {\sc State-resilience}, and straightforward reductions to the other two state resilience problems, as detailed in Appendix~\ref{appendix}.
 
@@
 
-\mathieu{mettre ici le détail de la  prop5}
+Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies {\sc $t$-liveness} is undecidable as well.
+This leads to the undecidability of {\sc State resilience}, {\sc Bounded-state-resilience} and
+{\sc $k$-state-resilience}. 
+The first follows straight from Proposition~\ref{reductions},
+while the other two
+follow from straightforward reductions
+we nonetheless
+develop in more details here.
 
@@
 
-\mathieu{mettre ici le détail du Theo 13}
+% \mathieu{mettre ici le détail du Theo 13}
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 $\mathop{\uparrow} q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$. One can then exhibit upward-closed subsets of 
- $\Safe$ and compute basis for the sets of elements from which they are reachable in at most $k$ steps. 
+ is in the complement of $\Bad$, i.e. $\Safe$.
+ From then on, we partition $\Safe$ into two subsets: an upward-closed 
+union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
+The key now is that
+there is only a finite number of elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} \leq \textbf{u}_k$ for all of which we will check reachability of $\Safe$ in $k$ or less steps, while 
+ elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps. 
+Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed, in particular we can compute the basis for the sets of elements from which the upward-closed subset of $\Safe$ is reachable in at most $k$ steps.
 Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
-This lead to a decision procedure
+This lead to a decision procedure detailed in appendix~\ref{appendix vass},
  which lead to the following decidability result:
@@
 \end{theorem}
-
-\begin{proof}
-\iffalse
-Let $V$ be a given $d$-VASS and consider a given $k \in \N$.
-Consider $\Bad \subseteq Q \times \N^d$ upward-closed and 
-$\Safe = S \setminus \Bad$
- downward-closed.
-Let us call $c_{\max}$ the maximal absolute value of a constant appearing in a coordinate of a transition.
-
-The set $\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider the vector
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
-
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-$\mathop{\uparrow} q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-Let us assume from now on 
- $\mathop{\uparrow} q_{j_1}(\textbf{0}), \mathop{\uparrow} q_{j_2}(\textbf{0}), \ldots \mathop{\uparrow} q_{j_\ell}(\textbf{0})$ are
-all subsets of $\Safe$ and that their union contain all upward-closed subsets of $\Safe$.
-Because these subsets are upward-closed, we can compute a basis of
-$\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))$, $\pred^{\leq k}(\mathop{\uparrow} q_{j_2}(\textbf{0}))$, \ldots
-$\pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
-
-We now consider the set 
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
-We inductively check for any element in the set whether or not it is possible to reach from it the set $\Safe$ in $k$ or less steps.
-If the set is finite then we stop the procedure once we have found a witness that $k$-resilience does not hold or once we have checked for every element, whichever comes first.
-If it is infinite then
-$k$-resilience does not hold, 
-hence we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps.
-This also means that, even in the case where $\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite, we only need to check reachability of $\Safe$ in $k$ or less steps for the elements with vectorial component at most
-$ \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$. Binding the search space
-thus leads to an $\EXPSPACE$ upper bound for {\sc $k$-resilience}.
-
-In order to deal with {\sc Bounded resilience} now, remark that there exists some $k_0 \in \N$ for which
-and hence
-$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$.
-We start the decision procedure by checking {\sc $k$-resilience} from $0$ until $k_0$.
-
-From $k_0$ onwards,
-$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
-\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is stationnary. 
-If the set is finite then we stop the procedure once we have checked for every element
-whether it is possible to reach $\Safe$ from it.
-If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
-hold
-for $k_m = \max(k_0, k_{\pi})$ with 
-$k_{\pi}$ the maximum of the length of the paths from 
-$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
-If it is infinite then
-for every $k \in \N$,
-$k$-resilience does not hold, 
-and we eventually find a witness that $k$-resilience does not hold.
-Indeed, 
-if
-$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
-is
-infinite
-then
-there exists an element of the form $q(\textbf{u} )$
-with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
-that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
-and hence from which it is not possible at all to reach
-$\Safe$ in $k$ or less steps. 
-\fi
-The set $\Bad$ admits a finite basis 
-$B_\Bad = \{ q_{i_1}(\textbf{v}_{i_1}), q_{i_2}(\textbf{v}_{i_2}), \ldots,
-q_{i_m}(\textbf{v}_{i_m}) \}$.
-Consider the vector $\textbf{v}_{\Bad} = \Sigma_{1 \leq j \leq m} \textbf{v}_{i_j}$,
-and then consider the vector
-$\textbf{u}_k = \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, c_{\max}, \ldots, c_{\max})$.
-For all $p \in Q$, all states reachable from $p(\textbf{u}_k)$ in $k$ or less steps are 
-of the form $q(\textbf{v})$ with $ \textbf{v} > \textbf{v}_{\Bad}$.
-
-For all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
-then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none.
-If there is none, then
-$\mathop{\uparrow} q(\textbf{0})$
- is in the complement of $\Bad$, i.e. $\Safe$.
-
-From then on, we partition $\Safe$ into two subsets: an upward-closed 
-union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
-The key now is that
-there is only a finite number of elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} \leq \textbf{u}_k$ for all of which we will check reachability of $\Safe$ in $k$ or less steps, while 
- elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps.
-Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed, in particular the set of
- elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$
- and the set of elements from which the upward-closed subset of $\Safe$ is reachable
- are both upward-closed sets a basis of which we can construct, and hence can be checked for inclusion. We refer to appendix~\ref{appendix vass} for more details.\qed
-\end{proof}
 
@@
 is stationnary. 
-\mathieu{
+%\mathieu{
 We check the finiteness of the set by comparing the finite basis of $\Bad$ and
 $(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
-}
+%}
 If the set is finite then we stop the procedure once we have checked for every element

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
  is in the complement of $\Bad$, i.e. $\Safe$.
- From then on, we partition $\Safe$ into two subsets: an upward-closed 
+Based on this, we partition $\Safe$ into two subsets: an upward-closed 
 union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
@@
  elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps. 
-Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed, in particular we can compute the basis for the sets of elements from which the upward-closed subset of $\Safe$ is reachable in at most $k$ steps.
+Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed.
+Indeed we can compute the basis for the sets of elements from which our upward-closed subset of $\Safe$ is reachable in at most $k$ steps.
 Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
-This lead to a decision procedure detailed in appendix~\ref{appendix vass},
+This leads to a decision procedure detailed in appendix~\ref{appendix vass},
  which lead to the following decidability result:

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 
-A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence one only needs $T$.\\
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton; hence one only needs $T$.
+
+ \begin{center}
+	\begin{figure}
+	\hspace{1.2cm}
+\includegraphics[width=0.80\textwidth]{WSTS_taxonomy_large}
+	\caption{A taxonomy of WSTS and variants.}
+	\end{figure}
+\end{center}
+
+	\vspace{-0.8cm}
+
+% \alain{il manque des refs dans la fig 1 qui sont déjà dans la biblio: WBTS, omega2,on peut mettre les front lossy avec les LCM et LCS}
+%
 
@@
 
- \begin{center}
-	\begin{figure}
-	\hspace{1.2cm}
-\includegraphics[width=0.80\textwidth]{WSTS_taxonomy_large}
-	\caption{A taxonomy of WSTS and variants.}
-	\end{figure}
-\end{center}
-
-	\vspace{-1.2cm}
-
-\alain{il manque des refs dans la fig 1 qui sont déjà dans la biblio: WBTS, omega2,on peut mettre les front lossy avec les LCM et LCS}
-%
 \section{Resilience for WSTS}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 {\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
-deciding structural termination of reset-VASS, which is undecidable.
-
-Indeed, the structural termination problem consists in deciding if %the termination problem holds for 
-a system terminates on every initial state.
-
-\problemx{Structural termination}
+deciding structural %termination of reset-VASS, 
+			existential halting of reset-VASS, 
+which is undecidable.
+
+Indeed, the structural existential halting problem consists in deciding if 
+a system %terminates
+have a terminating (i.e. finite) run
+ from every initial state.
+
+% \problemx{Structural universal halting}
+% {A reset-VASS $V=(Q,T)$ of dimension $d$.}
+% { Do all runs for every initial state terminate ?\\}
+
+\problemx{Structural existential halting}
 {A reset-VASS $V=(Q,T)$ of dimension $d$.}
-{ %$Q \times \N^d \longrightarrow^{*} q_f({\bf 0})$ ? 
-Do all runs for every initial state terminate ?\\}
+{ For every initial state, does there exists a run that terminates ?\\}
 
@@
 \nocite{araki1976PN,dufourd1998reset}
-{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to
-deciding structural termination of reset-VASS,
-since termination is w.l.o.g. reachability of $q_f({\bf 0})$ 
- \mathieu{est-ce qu'on a pas besoin d'une hypothèse d'être deterministic ici pour que ce soit vraiment w.l.o.g. ??} 
-\alain{qu'entends-tu par deterministic ici ?}
-for some control state $q_f$. \qed \end{proof}
+{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to {\sc Structural existential halting} of reset-VASS,
+since a run terminates w.l.o.g. if it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 \noindent
-{\bf Our main contributions}
+{\bf Our main contributions.}
 \begin{itemize}
@@
 
-Most of our results about the different models and different types of resilience are in Figure YYYY.
-
-\noindent
-{\bf Plan of the paper.}
+Most of our results about the different models and different types of resilience are in the following tables. The different notions of effectivity are detailed on page~\pageref{on effectivity}.
 
@@
 
-\alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
-
 \begin{center}
@@
 
+% \alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
+
+\noindent
+{\bf Plan of the paper.}
+
+% Section 2 introduit les notions, Section 3 et 4 les résultats de résilience et state résilience pour les WBTS, Section 5 sur les VASS et variantes
+Section~2 introduces the necessary definitions of WSTS and variations as well as the different notions of effectivity.
+Section~3 and 4 are concerned with the general framework of WSTS and contain our results on the three resilience problems
+and the three state-resilience problems respectively. Lastly, Section~5 concerns itself with the resilience problems for VASS and variations of VASS.
+
 \section{Well-structured transition systems and VASS}\label{section definitions}
@@
 More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is
-\begin{enumerate}
+\begin{enumerate}\label{on effectivity}
 %\item {\em effective} if there exists a pair of algorithms
@@
 
-	\vspace{-0.8cm}
+	\vspace{-0.9cm}
 
@@
 $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
-Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96,DBLP:journals/iandc/CeceFI96}.
+Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96%,DBLP:journals/iandc/CeceFI96
+%
+%% GOTO bibliographic reference when doing the bibtex final.aux did not work
+}.
 

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 
-The undecidability of which leads to the following proposition.
+{\sc Structural existential halting} is undecidable which leads to the following proposition.
 
@@
 {\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to {\sc Structural existential halting} of reset-VASS,
-since a run terminates w.l.o.g. if it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}
+since a run terminates w.l.o.g. iff it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}
 
@@
 Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the 
-structural termination of reset-VASS.
+{\sc Structural existential halting} of reset-VASS.
 %\mathieu{GOTO}
@@
 
-\section{Lemma technical proofs}
+\section{Proofs of technical lemmas}
 
@@
 
-\section{More details on VASS}\label{appendix vass}
+\section{Decision procedure for VASS when $\Safe = \mathop{\downarrow} \Safe$}\label{appendix vass}
 
@@
 %\mathieu{
-We check the finiteness of the set by comparing the finite basis of $\Bad$ and
+We check the finiteness of the set by comparing the elements from the finite basis of $\Bad$ and
 $(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$.
-This leads to a decision procedure detailed in appendix~\ref{appendix vass},
+We check the finiteness of the set by comparing the elements from the finite basis of both upward-closed sets. This leads to a decision procedure detailed in appendix~\ref{appendix vass},
  which lead to the following decidability result:
@@
 $(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+%
+As an example, in dimension $2$, for an upward-closed set $U$, $\mathop{\uparrow} q(n,m) \setminus U$ is finite if and only if $U$ contains an element of the form $q(l,m')$ with $m' \geq m$ and $l \leq n$
+and one of the form $q(n',l)$ with $n' \geq n$ and $l \leq m$. Indeed, if it does, then
+elements of $\mathop{\uparrow} q(n,m)$ with coordinates sufficiently large are greater than these and thus belong to $U$.
+If not, then the infinite set $\{q(n,m') \mid m' \geq m \}$ or the infinite set
+$\{q(n',m) \mid n' \geq n \}$   belong to $\mathop{\uparrow} q(n,m) \setminus U$ which is hence not finite.
 %}

### 2023-10-26 VMCAI-revised-version/final.tex
up

--- 2023-10-26 VMCAI-revised-version/final.tex
@@
 We check the finiteness of the set by comparing the elements from the finite basis of $\Bad$ and
-$(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+$(\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
+%
+As an example, in dimension $2$, for an upward-closed set $U$, $(\mathop{\uparrow} q(n,m)) \setminus U$ is finite if and only if $U$ contains an element of the form $q(n',m')$ with $m' \geq m$ and $n' \leq n$,
+and one of the form $q(n'',m'')$ with $n'' \geq n$ and $m'' \leq m$. Indeed, if it does, then
+elements of $\mathop{\uparrow} q(n,m)$ with coordinates sufficiently large are greater than these and thus belong to $U$ by upward-closedness.
+If not, then either the infinite set $\{q(n,m') \mid m' \geq m \}$ or the infinite set
+$\{q(n',m) \mid n' \geq n \}$   belong to $\mathop{\uparrow} q(n,m) \setminus U$ which is hence not finite.
 %}
-If the set is finite then we stop the procedure once we have checked for every element
+
+If the set $\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$ is finite then we stop the procedure once we have checked for every element
 whether it is possible to reach $\Safe$ from it.
@@
 $\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
-If it is infinite then
+If the set is infinite then
 for every $k \in \N$,

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
 
+\section{Structural halting in reset-VASS}
+
+We claim {\sc Structual existential halting} in reset-VASS is undecidable.  
+This is based on the construction providng the undecidability of structural termination in Lossy counter machines and in reset-VASS~\cite{dufourd1998reset,mayr1998lossy}.
+Strictly speaking undecidability of
+{\sc Structual existential halting} is not considered in
+\cite{mayr1998lossy},
+but 
+the encoding of Minsky machines can be reused
+since 
+ {\sc Structual existential halting}
+ is equivalent to
+ structural termination in the case of deterministic Minsky machines and thus,
+ is undecidable for Minsky machines in general.
+
 \section{Decision procedure for VASS when $\Safe = \mathop{\downarrow} \Safe$}\label{appendix vass}

### 2023-10-26 VMCAI-revised-version/final.tex
up

@@
  is equivalent to
- structural termination in the case of deterministic Minsky machines and thus,
- is undecidable for Minsky machines in general.
+ structural termination in the case of deterministic Minsky machines and is thus
+ undecidable for Minsky machines in general.
 

### 2023-10-27 VMCAI-revised-version/final.tex
effectivity

@@
 
-But there is a more general class of (upward) compatible ordered transition systems than WSTS for which coverability is still decidable: recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ that contains no infinite antichains (but it can be not well founded).
+But there is a more general class of (upward) compatible ordered transition systems than WSTS for which coverability is still decidable: recall that a Well Behaved Transition System (WBTS) \cite{DBLP:journals/lmcs/BlondinFM17} is an upward compatible ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $(S,\leq)$ contains no infinite antichains (but it can be not well founded).
 The class of WBTS is strictly larger than WSTS: for example, $\mathbb{Z}^d-$VASS under the lexicographical ordering are WBTS but not WSTS \cite{DBLP:journals/lmcs/BlondinFM17}.

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{lemma}\label{downward compatible}
-For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.
+For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
 \end{lemma}

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

@@
    \multirow{3}{*}{ {\bf RP }} 
-   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
+   		 & Downard-WBTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
    		 	& \multirow{2}{*}{Decidable}  \\
@@
    \multirow{2}{*}{{\bf sRP ~ ~ }} 
-   		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
+   		 & WBTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
 ~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
    \multirow{2}{*}{{\bf BsRP }} 
-   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
+   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
 ~	&  ? & Undecidable   \\ \hline
    \multirow{2}{*}{{\bf ksRP }} 
-   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
+   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
 ~	 & ? & Undecidable    \\ \hline
@@
 %
-\begin{definition}[Definition 5.1, \cite{DBLP:journals/tcs/FinkelS01}]
-A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
-is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ such that   
+\begin{definition}
+A {\em downward (compatible) WBTS (short, downward-WBTS)} $\mathscr{S}=(S, \rightarrow, \leq)$
+is an ordered transition system $(S, \rightarrow, \leq)$ such that $\leq$ is without infinite antichains and
 the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
@@
 	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.
+
+A  {\em downward (compatible) WSTS (short, downward-WSTS)} \cite[Definition 5.1]{DBLP:journals/tcs/FinkelS01} is a downward-WBTS where $\leq$ is a wqo.
 \end{definition}
@@
 \begin{theorem}\label{up-down}
-{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
+{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WBTS with 
 $\Safe=\mathop{\downarrow} \Safe$.
@@
 \begin{proof}
-Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
+Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WBTS with 
 $\Safe=\mathop{\downarrow} \Safe$.
@@
 In the case
-of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
+of a $\pred^*$-basis-effective WBTS, not necesarly downward compatible,
 the above construction
@@
 are undecidable for reset-VASS, hence for strongly compatible,
-$\post$-effective, $\pred$-basis-effective WSTS
+$\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS
 when
@@
 
-In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
+In the case of a WBTS with \emph{downward} compatibility, not necessarily strong,
 then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
@@
 %$\pred$-basis-effective, 
-$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
+$\post$-ideal-effective and $\pred^*$-basis-effective WBTS with downward and upward compatibilities
 and $\Safe = \mathop{\downarrow} \Safe$.
@@
 \begin{corollary}\label{downward brp}
-{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
+{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WBTS, strong downward compatibility and upward compatibility when
 $\Safe = \mathop{\downarrow} \Safe$

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

@@
 We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$).
-We study the decidability of each type of resilience for many models: Well Behaved Transition Systems (WBTS), WSTS, VASS and other variations of counter machines.
-%
-We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) for downward-closed sets $\Safe$.
+We study the decidability of each type of resilience for many infinite-state transitiuon systems: Well Behaved Transition Systems (WBTS), Well Structured Transition Systems (WSTS), VASS and other variations of counter machines.
+%
+We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) when $\Safe$ are downward-closed sets.
 
@@
 \begin{proof}
+\alain{où utilise-t-on la strong compatibility ? et si on ne l'utilise pas, c'est le theo 8...}
 Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
@@
 
+\alain{plus fort que prop 6}
 \begin{theorem}\label{post srp}
@@
 \begin{proof}
-Let $B_{\mathop{\uparrow} \post^*(s)}$ 
- be a basis of $\mathop{\uparrow} \post^*(s)$, and $B_\Safe$ a basis of $\Safe$.
-By applying Lemma~\ref{Lemma intersection} twice, we obtain
+Let $\mathscr{S}$ be a $\pred$-basis-effective, $\post^*$-basis-effective WSTS.
+%
+Since $\mathscr{S}$ is $\post^*$-basis-effective, let $B_{\mathop{\uparrow} \post^*(s)}$ 
+ be a finite basis of $\mathop{\uparrow} \post^*(s)$, hence $\mathop{\uparrow} \post^*(s) = \mathop{\uparrow} B_{\mathop{\uparrow} \post^*(s)}$.
+%
+Let $B_\Safe$ be a finite basis of $\Safe$, hence $\Safe = .\mathop{\uparrow} B_\Safe$.
+%
+Since $\mathscr{S}$ is $\pred$-basis-effective, we may compute a finite basis $B_{\pred^*(\Safe)}$ of $\pred^*(\Safe)$ from $B_\Safe$.
+%
+\alain{je pense qu'il y a une erreur, on n'utilise pas le Lemme 2}
+%
+By applying Lemma~\ref{Lemma intersection} twice,
+we obtain that 
 \[ \post^*(s) 
@@
  \subseteq \pred^*(\Safe)\]
-Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite and we can compute 
+%
+%
+\alain{je pense qu'il y a une erreur, je propose plutôt}
+%
+Recall that $\mathscr{S}$ is state-resilient iff  $\post^*(s) 
+ \subseteq \pred^*(\Safe)$ iff 
+\[ \mathop{\uparrow} B_{\mathop{\uparrow} \post^*(s)} 
+ \subseteq \mathop{\uparrow}  B_{\pred^*(\Safe)} \text{ iff } \forall b \in B_{\mathop{\uparrow} \post^*(s)}  \exists b' \in B_{\pred^*(\Safe)} \text{ such that } b' \leq b \]
+%
+Since both $B_{\mathop{\uparrow} \post^*(s)}$ and $B_{\pred^*(\Safe)}$ are finite, {\sc State-resilience} is decidable.
+%
+\alain{ce qui suit est à enlever je l'ai commenté (à vérifier)}
+%
+%			Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite
 %\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
-a basis of $\pred^*(\Safe)$ from $B_\Safe$, we can check that $B_{\mathop{\uparrow} \post^*(s)} 
- \setminus \pred^*(\Safe) = \emptyset$.  \qed
+%			we can check that $B_{\mathop{\uparrow} \post^*(s)} 
+%			 \setminus \pred^*(\Safe) = \emptyset$.  \qed
 \end{proof}

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

@@
 \end{theorem}
-
-We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
-
-\begin{proposition}\label{postcomputable}
-{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-\end{proposition}
-
-\begin{proof}
-\alain{où utilise-t-on la strong compatibility ? et si on ne l'utilise pas, c'est le theo 8...}
-Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
-$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
-$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
-$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
-returning $n$ if that is the case.
-Then, because {\sc $n_0$-state-resilience} is decidable, 
-checking $\mathop{\uparrow} \post^*(s) 
- \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-and {\sc State-resilience} is decidable. \qed
-\end{proof}
-
+%
+%
+%\alain{ce qui suit est à supprimer car subsumé par Theo 8}
+%%
+%We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
+%
+%\begin{proposition}\label{postcomputable}
+%{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
+%when
+%$\Safe=\mathop{\uparrow} \Safe$.
+%\end{proposition}
+%
+%\begin{proof}
+%\alain{où utilise-t-on la strong compatibility ? et si on ne l'utilise pas, c'est le theo 8...}
+%Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
+%$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
+%$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
+%$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
+%returning $n$ if that is the case.
+%Then, because {\sc $n_0$-state-resilience} is decidable, 
+%checking $\mathop{\uparrow} \post^*(s) 
+% \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
+%and {\sc State-resilience} is decidable. \qed
+%\end{proof}
+%%
+%\alain{ce qui précède est à supprimer car subsumé par Theo 8}
+%
 The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
@@
 
-\alain{plus fort que prop 6}
+%\alain{Theo 8 est plus fort que prop 6}
 \begin{theorem}\label{post srp}
@@
 %
-\alain{je pense qu'il y a une erreur, on n'utilise pas le Lemme 2}
 %
@@
 \[ \post^*(s) 
- \subseteq \pred^*(\Safe) \text{ iff } B_{\mathop{\uparrow} \post^*(s)}  
+ \subseteq \pred^*(\Safe) \text{ iff } \mathop{\uparrow} \post^*(s)
  \subseteq \pred^*(\Safe)\]
@@
 %
-\alain{je pense qu'il y a une erreur, je propose plutôt}
-%
-Recall that $\mathscr{S}$ is state-resilient iff  $\post^*(s) 
- \subseteq \pred^*(\Safe)$ iff 
+%
+%Recall that $\mathscr{S}$ is state-resilient iff  $\post^*(s) 
+% \subseteq \pred^*(\Safe)$ iff 
+Now the last inclusion is equivalent to:
 \[ \mathop{\uparrow} B_{\mathop{\uparrow} \post^*(s)} 
@@
 %
-Since both $B_{\mathop{\uparrow} \post^*(s)}$ and $B_{\pred^*(\Safe)}$ are finite, {\sc State-resilience} is decidable.
-%
-\alain{ce qui suit est à enlever je l'ai commenté (à vérifier)}
+Since both $B_{\mathop{\uparrow} \post^*(s)}$ and $B_{\pred^*(\Safe)}$ are finite, {\sc State-resilience} is decidable.\qed
+%
 %

### 2023-10-28 VMCAI-revised-version/final.tex
up

@@
 %			which are wqo but not known to be $\omega^2$-wqo.
-The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is an $\omega^2$-wqo (Proposition~4.5 in \cite{DBLP:journals/corr/abs-1208-4549}).
+The class of $\omega^2$-wqo is robust because every datatype in the following list - natural number, finite set, ﬁnite product, ﬁnite sum, finite disjoint sum, ﬁnite words, ﬁnite multisets and ﬁnite trees - is an $\omega^2$-wqo \cite[Proposition~4.5]{DBLP:journals/corr/abs-1208-4549}.
 Now we are able to define $\omega^2$-WSTS:
@@
 %
-Blondin and al.
-showed (Lemma 4.3 in \cite{DBLP:journals/lmcs/BlondinFM17}) that under the previous hypotheses, one are able to enumerate \emph{downward-closed sets} (by their finite decomposition in ideals), to decide inclusion between downward-closed sets and to decide if a state belongs to a given finite set of ideals.
+Blondin and al. showed \cite[Lemma 4.3]{DBLP:journals/lmcs/BlondinFM17} that under the
+%Blondin and al. showed (Lemma 4.3 in \cite{DBLP:journals/lmcs/BlondinFM17}) that under the 
+previous hypotheses, one are able to enumerate \emph{downward-closed sets} (by their finite decomposition in ideals), to decide inclusion between downward-closed sets and to decide if a state belongs to a given finite set of ideals.
 But the five properties are not sufficient to compute the complementary and the intersection of downward-closed sets.
@@
 Reset VASS (hence VASS) and (front) lossy fifo automata are $\post$-effective, $\post$-ideal-effective, and $\pred$-basis-effective.
-There exist $\post$-effective WSTS that are not $\post$-ideal-effective (Proposition~35 in \cite{BFM-ic17}); there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition~36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition~45 in \cite{BFM-ic17}).
+There exist $\post$-effective WSTS that are not $\post$-ideal-effective \cite[Proposition~35]{BFM-ic17}; there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition~36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition~45 in \cite{BFM-ic17}).
 Reset VASS are not $\post^*$-basis-effective since $\post^*$-basis-effectiveness allows to compute the finite set of the set of minimal reachable states, hence it would allow to decide the zero-reachability problem that is undecidable for reset-VASS.
@@
 the above results still hold in the more general case where $\Bad$ and $\Safe$ are not complements of each others.
+Remark furthermore that
+	reset-VASS are $\pred$-basis-effective
+	hence satisfy the conditions from
+	Theorem~\ref{xxx}
+	which thus implies decidability
+	of {\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} in reset-VASS in the case $\Safe = \mathop{\uparrow} \Safe$.
 
@@
 {\sc $t$-liveness}.
-The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets, whose proof can be seen in Appendix~\ref{appendix}%
+% The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets,
+The following Proposition is a variation  to reset-VASS 
+of~\cite[Theorem~5.5]{peterson1981petri} originally stated for Petri nets,
+ whose proof can be seen in Appendix~\ref{appendix}%
 .
@@
 \begin{proof}
-We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
+% We deduce from Theorem 3.6 in \cite{DBLP:conf/rp/Schnoebelen10} 
+We deduce from \cite[Theorem 3.6]{DBLP:conf/rp/Schnoebelen10} 
+that $\pred^*(\Safe)$ is a computable semilinear set if $\Safe$ is semilinear.
 Hence since the inclusion between two semilinear sets is decidable, we deduce that $\Bad \subseteq \pred^*(\Safe)$ is decidable if $\Bad$ is semilinear.
@@
 
-We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
+We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by 
+% "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
+"adjusting" a similar result for Petri nets~\cite[Theorem~5.5]{peterson1981petri}. \\
 

### 2023-10-29 VMCAI-revised-version/final.tex
effectivity

@@
 %			If the home-space contains a single element, this element is an {\em home-state}.
-Given a transition system (where $S$ is the set of states) and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$
+Given a transition system whose set of states is $S$ and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$
 %				(i.e., $\post^*(X) \subseteq \pred^*(H)$).

### 2023-10-29 VMCAI-revised-version/final.tex
effectivity

@@
 %			If the home-space contains a single element, this element is an {\em home-state}.
-Given a transition system whose set of states is $S$ and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$
+Given a transition system whose set of states is $S$ and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$.
+%
+In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} in Petri nets for $X$ a singleton.
+In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}).
+In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
+%
 %				(i.e., $\post^*(X) \subseteq \pred^*(H)$).
-%
+
 A transition system is {\em resilient for $(\Bad,\Safe)$} if $(\Bad,\Safe)$ satisfies the home-space property.
@@
 
-In 2016, Prasad and Zuck introduced in \cite{DBLP:journals/corr/PrasadZ16} interesting definitions and results (without detailed proofs) about resilience in the framework of process algebra.
-They show that resilience is decidable for effective WSTS with both upward and reflexive downward compatibilities (and some other technical conditions).
-
+In 2016, Prasad and Zuck introduced in \cite{DBLP:journals/corr/PrasadZ16} intuitions, definitions (without any connection to the concept of the home-space) and results (without detailed proofs) about resilience in the framework of process algebra.
+They show that resilience is decidable for effective Well Structured Transition Systems (WSTS) with both upward and reflexive downward compatibilities and some other technical conditions (that are defined in Section \ref{section definitions}).
+%
 %		reduces to coverability which is decidable on WSTS.
@@
 %%%%%%%%%
-
-In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$
+%
+In 2021, \"Ozkan and Würdemann  \cite{DBLP:journals/corr/abs-2108-00889} and \"Ozkan \cite{DBLP:conf/gg/Ozkan22}, in 2022, proved the decidability of the bounded state-resilience problem and the $k$-state-resilience problem for WSTS  with strong compatibility (see Section \ref{section definitions}) and with the supplementary (strong) hypothesis that there exists an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$
 %
@@
 %%%%%%%%%
-
-In 1986, Memmi and Vautherin introduced the notion of home-space~\cite{DBLP:conf/ac/MemmiV86} in Petri nets for $X$ a singleton.
-In 1989, de Frutos Escrig and Johnen proved that the home-space problem (for $X$ a singleton and $H$ a finite union of linear sets with the same period) was decidable for VASS (\cite{de1989decidability}).
-In 2023, Jancar and Leroux proved the decidability of the (complete) semilinear home-space problem ($X$ and $H$ are both semilinear)  in VASS \cite{DBLP:journals/corr/abs-2207-02697}.
 
@@
 %			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-\item The three resilience problems are decidable for post-ideal-effective Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
+\item The three resilience problems are decidable for post-ideal-effective (Definition \ref{eff}) Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
 %
 \item  Resilience is also decidable, when 
-$\Safe$ is upward-closed, for effective pred-basis WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
+$\Safe$ is upward-closed, for effective pred-basis (Definition \ref{eff}) WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
 %
@@
 %
-\item We clarify the different effectiness conditions on WBTS and WSTS that allow the decidability of the six resilience.
-%
-\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} ; we also show that relaxing some hypothesis leads to undecidability.
+\item We clarify the different effectiness conditions on WBTS and WSTS (Definition \ref{eff}) that allow the decidability of the six resilience.
+%
+\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability.
 %
@@
 
-Most of our results about the different models and different types of resilience are in the following tables. The different notions of effectivity are detailed on page~\pageref{on effectivity}.
+Most of our results about the different models and different types of resilience are in the following tables. The different notions of effectivity are given in Definition \ref{eff}.
+%				detailed on page~\pageref{on effectivity}.
 
@@
 
-In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors.
-More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is
+In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors. More precisely:
+%
+\begin{definition}\label{eff}
+%
+We say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is
+%
 \begin{enumerate}\label{on effectivity}
@@
 \end{enumerate}
-
+%
+\end{definition}
+%
 Counter machines are effective but don't enjoy any other effectivities among the list of five.

### 2023-10-29 VMCAI-revised-version/final.tex
effectivity

@@
 %			 that will be essential for solving various resilience problems in different models of WSTS. 
-First, to simplify, we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five following properties:
+First, to simplify and w.l.o.g., we suppose that all classes of considered ordered transition systems $\mathscr{S}=(S, \rightarrow, \leq)$ satisfy the five following properties:
 \begin{enumerate}
@@
 Halfon introduced \emph{ideally-effective wqo} as wqo that essentially allow to compute representations of (principal) ideals $\mathop{\downarrow} s$, of the complementary of an ideal, of the complementary of filters (a filter is a set $\mathop{\uparrow} s$ for $s \in S$) and to compute representations of finite intersections of filters and finite intersections of ideals.
-Most well-known wqo are ideally-effective .
-\cite{DBLP:phd/hal/Halfon18}.
+Most well-known wqo are ideally-effective \cite{DBLP:phd/hal/Halfon18}.
 %	 (hence of downward-closed sets).
@@
 Reset VASS (hence VASS) and (front) lossy fifo automata are $\post$-effective, $\post$-ideal-effective, and $\pred$-basis-effective.
-There exist $\post$-effective WSTS that are not $\post$-ideal-effective \cite[Proposition~35]{BFM-ic17}; there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition~36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition~45 in \cite{BFM-ic17}).
+There exist $\post$-effective WSTS that are not $\post$-ideal-effective \cite[Proposition~35]{BFM-ic17}; there exist $\post$-ideal-effective WSTS that are not $\post$-effective \cite[Proposition~36]{BFM-ic17}.There also exist WSTS that are not $\pred$-basis-effective \cite[Proposition~45]{BFM-ic17}.
 Reset VASS are not $\post^*$-basis-effective since $\post^*$-basis-effectiveness allows to compute the finite set of the set of minimal reachable states, hence it would allow to decide the zero-reachability problem that is undecidable for reset-VASS.
-However, VASS are $\post^*$-basis-effective \cite{DBLP:journals/corr/abs-2108-00889} and LCS (with $d$ fifo channels) are also $\post^*$-basis-effective because we have $(q,w_1,w_2,...,w_d) \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)  \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)$ is reachable from $(q_0,u_1,u_2,...,u_d)$ that is decidable in LCS.
+However, VASS are $\post^*$-basis-effective \cite[Proposition 2]{DBLP:conf/gg/Ozkan22} and LCS (with $d$ fifo channels) are also $\post^*$-basis-effective because we have $(q,w_1,w_2,...,w_d) \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)  \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)$ is reachable from $(q_0,u_1,u_2,...,u_d)$ that is decidable in LCS.
 %
@@
 
-Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven.
-We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility. 
+Surprisingly, the general undecidability statement regarding resilience had neither been known nor proven (it is simply mentioned as a future work in the conclusion of \cite{DBLP:journals/corr/abs-2108-00889}).
+We show that the resilience problem is undecidable for $\omega^2$-WSTS with strong compatibility and natural effectiveness hypothesis. 
 %

### 2023-10-29 VMCAI-revised-version/final.tex
effectivity

@@
    		 & WBTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
-~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
+~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp down up}  & Undecidable  \\ \hline
    \multirow{2}{*}{{\bf BsRP }} 
@@
 	nor $\pred$-basis-effective.
-  Termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.
+  Termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectiveness hypothesis on recursive functions.
 
@@
 
-In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
+In the case of $\pred$-basis-effective (upward) compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
 {\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
@@
 a system %terminates
-have a terminating (i.e. finite) run
+have a \alain{un ou tous les run ?}terminating (i.e. finite) run
  from every initial state.
@@
 {A reset-VASS $V=(Q,T)$ of dimension $d$.}
-{ For every initial state, does there exists a run that terminates ?\\}
+{ For every initial state, does there exist a run that terminates ?\\}
 
@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS when $\Safe=\mathop{\downarrow} \Safe$, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
+{\sc Resilience} is undecidable for reset-VASS when $\Safe=\mathop{\downarrow} \Safe$, hence it is also undecidable for $\post$-effective and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
 \end{proposition}

### 2023-10-30 VMCAI-revised-version/final.tex
up
effectivity

@@
 %			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
-\item The three resilience problems are decidable for post-ideal-effective (Definition \ref{eff}) Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up})
-%
-\item  Resilience is also decidable, when 
+%
+\item In the subsequent three figures, we present the majority of our results pertaining to various models and resilience categories.
+The figures depict the boundaries of decidability within the context of well-known VASS and reset-VASS models.
+%
+\item In particular, the three resilience problems are decidable for post-ideal-effective (Definition \ref{eff}) Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up}).
+Resilience is also decidable, when
 $\Safe$ is upward-closed, for effective pred-basis (Definition \ref{eff}) WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
+%
 %
@@
 %
-\item We generalize  the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability.
+\item We generalize with Theorem \ref{post srp} the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability (Proposition \ref{srp up down}).
 %
@@
 
-\item We show that the three state-resilience problems are decidable for post-ideal-effective WSTS with downward and upward compatibilities (Theorem \ref{downward srp} extends Theorem 1 in \cite{DBLP:journals/corr/PrasadZ16}).
-The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WSTS with strong downward compatibility.
+\item We show that the three state-resilience problems are decidable for post-ideal-effective WBTS with downward and upward compatibilities and $\Safe$ downward-closed (Theorem \ref{downward srp} extends \cite[Theorem 1]{DBLP:journals/corr/PrasadZ16}).
+The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WBTS with strong downward compatibility.
 %
@@
 
-Most of our results about the different models and different types of resilience are in the following tables. The different notions of effectivity are given in Definition \ref{eff}.
 %				detailed on page~\pageref{on effectivity}.
@@
 \end{center}
+\fi
+
+ \begin{center}
+ 	\begin{figure}
+ 	\hspace{-2.cm}
+\includegraphics[width=1.40\textwidth]{All_results}
+\caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
+\iffalse
+ \begin{center}
+	\begin{figure}
+%	\hspace{1.2cm}
+\includegraphics[width=0.70\textwidth]{Safe_up_R}
+	\end{figure}
+\end{center}
+
+ \begin{center}
+	\begin{figure}
+\includegraphics[width=0.70\textwidth]{Safe_up_State_R}
+	\end{figure}
+\end{center}
+
+ \begin{center}
+	\begin{figure}
+\includegraphics[width=0.70\textwidth]{Safe_down_R}
+	\end{figure}
+\end{center}
+\fi
 

### 2023-10-30 VMCAI-revised-version/final.tex
effectivity

@@
 We complemented previous results (decidability of two state-resilience problems for a restricted class of WSTS in~\cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) by providing some undecidability proofs for resilience and state-resilience in general.
-We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience, in the case
- $\Safe = \mathop{\uparrow} \Safe$ and in the case 
- $\Safe = \mathop{\downarrow} \Safe$.
+We exhibited classes of WBTS, WSTS, VASS and extensions of VASS with decidable resilience.
+%in the case  $\Safe = \mathop{\uparrow} \Safe$ and in the case  $\Safe = \mathop{\downarrow} \Safe$.
 
@@
 For instance, we have been concerned with decidability only, and a detailed complexity analysis of the different resilience problems still remains to be done for concrete models.
-One other angle of attack not studied here is the synthesis of sets 
-$\Safe$ for resilience,
-for instance, 
- finding the maximal upward-closed (resp. downward-closed) subset 
-$\Safe$ so that a system is $\Safe$-resilient. 
+%
+%One other angle of attack not studied here is the synthesis of sets 
+%$\Safe$ for resilience,
+%for instance, 
+% finding the maximal upward-closed (resp. downward-closed) subset 
+%$\Safe$ so that a system is $\Safe$-resilient. 
+%
 Another question could be to analyse the resilience in the framework of a controller and its environment. 

### 2023-10-30 VMCAI-revised-version/final.tex
effectivity

@@
 We consider six types of resilience (one of them is the home-space property) defined by an upward-closed set or a downward-closed set $\Safe$, and by the existence of a bound on the length of minimal runs starting from a set $\Bad$ and reaching $\Safe$ ($\Bad$ is generally the complementary of $\Safe$).
-We study the decidability of each type of resilience for many infinite-state transitiuon systems: Well Behaved Transition Systems (WBTS), Well Structured Transition Systems (WSTS), VASS and other variations of counter machines.
-%
-We first show that all resilience problems are undecidable for effective WSTS with strong compatibility and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) when $\Safe$ are downward-closed sets.
-
-We then show that resilience is decidable for WBTS with the computability of the finite basis of the downward-closure of the set of (immediate) successors of a downward-closed set ($\post$-basis-effective) and for WSTS with the (well-known) $\pred$-basis-effective hypothesis.
+
+%We study the decidability of each type of resilience for many infinite-state transitiuon systems: Well Behaved Transition Systems (WBTS), Well Structured Transition Systems (WSTS), VASS and other variations of counter machines.
+%
+We first show that all resilience problems are undecidable for effective Well Structured Transition Systems (WSTS) with strong compatibility.
+%
+%and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) when $\Safe$ are downward-closed sets.
+%
+We then show that resilience is decidable for Well Behaved Transition Systems (WBTS) and for WSTS with adapted effectiveness hypotheses.
+%
+%with the computability of the finite basis of the downward-closure of the set of (immediate) successors of a downward-closed set ($\post$-basis-effective) 
+%
+%
+%			with the (well-known) $\pred$-basis-effective hypothesis.
 %
@@
 %
-Most of the resilience properties are shown decidable for other classes like downward-WSTS, VASS, lossy counter machines, integer VASS and continuous VASS.
+Most of the resilience properties are shown decidable for other classes like WSTS with the downward compatibility, VASS, lossy counter machines, reset-VASS, integer VASS and continuous VASS.
 

### 2023-10-30 VMCAI-revised-version/final.tex
up

@@
 a system %terminates
-have a \alain{un ou tous les run ?}terminating (i.e. finite) run
+have %a \alain{un ou tous les run ?}
+at least one terminating (i.e. finite) run
  from every initial state.

### 2023-10-30 VMCAI-revised-version/final.tex
up

@@
  	\hspace{-2.cm}
-\includegraphics[width=1.40\textwidth]{All_results}
-\caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
-Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+\includegraphics[width=1.30\textwidth]{All_results}
+\caption{ Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.
+Detailed notions of effectivity can be found on page~\pageref{on effectivity}.
+The different models are furthermore introduced in Section~2 and Section~3.}
 	\end{figure}
 \end{center}
+
+\vspace{-0.5cm}
 
@@
 \end{definition}
-Since almost of usual wqo are $\omega^2$-wqo, all naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, reset/transfer VASS are $\omega^2$-WSTS.\\
+Since almost all usual wqo are $\omega^2$-wqo, all naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, reset/transfer VASS are $\omega^2$-WSTS.\\
 %%%%%%

### 2023-10-30 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-30 VMCAI-revised-version/final.tex
@@
 Reset-VASS are not $\pred^*$-basis-effective because computing a finite basis of $\mathop{\downarrow} \pred^*(q,0,...,0)$ would allow to decide whether $(q,0,...,0)$ is reachable that is undecidable for reset-VASS.
-By using the decidability of reachability and Theorem 3.11 in \cite{DBLP:journals/acta/ValkJ85}, we may prove that VASS are $\pred^*$-basis-effective.
+By using the decidability of reachability and \cite[Theorem 3.11]{DBLP:journals/acta/ValkJ85}, we may prove that VASS are $\pred^*$-basis-effective.
 A more complete study of these properties will be done in the long version of this paper.
@@
 \begin{theorem}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
- A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
+ A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis.
 Hence coverability is decidable.

### 2023-10-30 VMCAI-revised-version/final.tex
up

--- 2023-10-30 VMCAI-revised-version/final.tex
@@
 Detailed notions of effectivity can be found on page~\pageref{on effectivity}.
-The different models are furthermore introduced in Section~2 and Section~3.}
+The different models are furthermore detailled in Section~2 and Section~3.}
 	\end{figure}

### 2023-10-30 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-30 VMCAI-revised-version/final.tex
@@
 
-In the case of $\pred$-basis-effective (upward) compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
-{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
-deciding structural %termination of reset-VASS, 
-			existential halting of reset-VASS, 
-which is undecidable.
-
-Indeed, the structural existential halting problem consists in deciding if 
-a system %terminates
-have a \alain{un ou tous les run ?}terminating (i.e. finite) run
- from every initial state.
+%In the case of $\pred$-basis-effective (upward) compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
+%{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
+%deciding structural %termination of reset-VASS, 
+%			existential halting of reset-VASS, 
+%which is undecidable.
+
+We now prove that {\sc Resilience} is undecidable for reset-VASS, hence it will be also undecidable for (upward) compatible WSTS when $\Safe = \mathop{\downarrow} \Safe$.
+%
+%Indeed, the structural existential halting problem consists in deciding if 
+%a system %terminates
+%have a terminating (i.e. finite) run
+% from every initial state.
 
@@
 
-{\sc Structural existential halting} is undecidable which leads to the following proposition.
+{\sc Structural existential halting} is undecidable since termination is undecidable for deterministic counter machines.
+
+%			which leads to the following proposition.
 

### 2023-10-31 VMCAI-revised-version/final.tex

@@
 
-We now prove that {\sc Resilience} is undecidable for reset-VASS, hence it will be also undecidable for (upward) compatible WSTS when $\Safe = \mathop{\downarrow} \Safe$.
+We now prove that {\sc Resilience} is undecidable for reset-VASS when $\Safe = \mathop{\downarrow} \Safe$. We will use the undecidability of the structural existential halting problem.
+%
+%			hence it will be also undecidable for (upward) compatible WSTS when $\Safe = \mathop{\downarrow} \Safe$.
 %
@@
 \problemx{Structural existential halting}
-{A reset-VASS $V=(Q,T)$ of dimension $d$.}
-{ For every initial state, does there exist a run that terminates ?\\}
-
-{\sc Structural existential halting} is undecidable since termination is undecidable for deterministic counter machines.
-
+%			{A reset-VASS $V=(Q,T)$ of dimension $d$.}
+{A transition system $\mathscr{S}=(S,\rightarrow)$}
+{ For every initial state $s_0 \in S$, does there exist a run from $s_0$ that terminates ?\\}
+
+{\sc Structural existential halting} is undecidable for reset-VASS since termination is undecidable for deterministic counter machines. 
+\alain{REFERENCE !!!!}
 %			which leads to the following proposition.
@@
 Let us first recall two results in \cite{BFM-ic17} that are stated for WSTS but are also true for WBTS since the proofs rely only on compatibility and not on the property of wqo.
-Proposition 30 establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
+\cite[Proposition 30]{BFM-ic17} establishes a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ and the runs of its completion $\hat{\mathscr{S}}$.
 It states that if $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
-Proposition 29 establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+\cite[Proposition 29]{BFM-ic17} establishes that if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
 Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
@@
 Suppose that $s_j$ is coverable from $x$ then there exists a run $x \xrightarrow{k} y \geq s_j$.
-From Proposition~30, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
+From \cite[Proposition 30]{BFM-ic17}, there exist an ideal $J$ and a run $\mathop{\downarrow} x \xrightarrow{k} J$ where $J \supseteq \mathop{\downarrow} y \supseteq \mathop{\downarrow} s_j$ in $\hat{\mathscr{S}}$, hence $\mathop{\downarrow} s_j$ is covered from $\mathop{\downarrow} x$.
 Conversely, if $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ with $\mathop{\downarrow} s_j \subseteq J$ then 
@@
 This implies that resilience and bounded resilience can be reduced to coverability.
-Since coverability is decidable for $\post$-effective WSTS \cite{BFM-ic17}, we deduce that both the 
+Since coverability is decidable for $\post$-ideal-effective WBTS \cite[Theorem 44]{BFM-ic17}, we deduce that both the 
   resilience problem and the bounded resilience problem are decidable.

### 2023-10-31 VMCAI-revised-version/final.tex
Re-basculé sur Minsky resilience pour l'indec des reset-VASS

@@
 
-We now prove that {\sc Resilience} is undecidable for reset-VASS when $\Safe = \mathop{\downarrow} \Safe$. We will use the undecidability of the structural existential halting problem.
+We now prove that {\sc Resilience} is undecidable for reset-VASS when $\Safe = \mathop{\downarrow} \Safe$.
+
+\mathieu{J'ai remis une vielle version que j'avais écrit avant de basculer sur le structural existential halting, je pense guarder ça en main text et détailler davantage dans l'appendix dit moi ce que tu en pense}
+In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
+
+\begin{proposition}\label{indec WSTS with dcs}
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+$\Safe=\mathop{\downarrow} \Safe$.
+\end{proposition}
+
+\begin{proof}
+First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $2$ counters when
+$\Safe = \mathop{\downarrow} \Safe$,
+since it is possible to have $\Safe = \{q_f(0,0)\}$. 
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
+% from  a state $p({\bf u})$
+% to
+reaching
+ a state 
+%
+$ q({\bf 0})$ corresponds to one that reaches 
+$q_f(0,0)$
+ in the Minsky machine and vice versa.
+  By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
+ we obtain that 
+ {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
+
+	\iffalse
+
+We will use the undecidability of the structural existential halting problem.
 %
@@
 since a run terminates w.l.o.g. iff it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}
+
+	\fi
 

### 2023-10-31 VMCAI-revised-version/final.tex
effectivity

@@
 \begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
+{\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\post$-effective and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
 $\Safe=\mathop{\downarrow} \Safe$.

### 2023-10-31 VMCAI-revised-version/final.tex
Appendix minsk back again

--- 2023-10-31 VMCAI-revised-version/final.tex
@@
 
+\section{Reset-VASS and Minsky-machines}\label{minsk}
+
+We now delve into more details into the undecidability results 
+upon which rely
+Proposition~\ref{indec WSTS with dcs}
+and
+Proposition~\ref{srp down up},
+namely, that
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
+are undecidable for Minsky machines
+and 
+for reset-VASS.
+
+The problems are undecidable for a Minsky machine with at least $2$ counters when
+$\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
+% The simulation however is 
+Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0)$ in the Minsky machine, 
+this weak simulation will be enough to obtain undecidability of both problems.
+
+\begin{proposition}\label{both minsk indec}
+{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
+\end{proposition}
+
+\begin{proof}
+First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
+	This is due to the undecidability of $2$-counter Minsky machine termination.
+From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
+such that there exists a terminating run in $M$ for all inputs iff machine $M'$ can reach $q_f'(0,0)$ from any input. 
+%			with at least $1$ on its third counter. 
+We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the two counters until they reach $0$, then, and only then, go in $q_f'$.
+Based on this construction, the problem of deciding whether the downward-closed set 
+$\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
+from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable~\cite{Min61}. 
+Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
+deciding whether the downward-closed set 
+$ \{q_f'(0,0) \}$ is reachable
+from every element of $\post^*(s_0)$ 
+is again equivalent to asking whether $M$ terminate on every input.
+Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
+
+% \alain{écrire qqchose ici}
+Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
+% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
+%Reset-VASS extend the basic VASS model with special “reset
+%transitions” that set to $0$ some coordinates in the vector. 
+%			Let us recall their definition here.
+%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
+The simulation operates by building a reset-VASS in which
+there exists
+ a sequence from 
+an initial state %$(s,x_1 \ldots x_n)$
+to
+a state $ q({\bf 0})$ if and only if
+there exists
+a sequence
+ in the Minsky machine
+ from a corresponding initial state
+ to
+ the state $q_f(0,0,0)$
+  (from Theorem~$5$ in \cite{araki1976PN}).
+Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
+{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
+Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
+and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
+$\omega^2$-WSTS with strong compatibility as well.\qed
+\end{proof}
+
+\iffalse
 \section{Structural halting in reset-VASS}
@@
  undecidable for Minsky machines in general.
+\fi
 

### 2023-10-31 VMCAI-revised-version/final.tex
effectivity

--- 2023-10-31 VMCAI-revised-version/final.tex
@@
 since it is possible to have $\Safe = \{q_f(0,0)\}$. 
+\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ?}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 

### 2023-10-31 VMCAI-revised-version/final.tex
effectivity

@@
 since it is possible to have $\Safe = \{q_f(0,0)\}$. 
-\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ?}
+\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 

### 2023-10-31 VMCAI-revised-version/final.tex

--- 2023-10-31 VMCAI-revised-version/final.tex
@@
 since it is possible to have $\Safe = \{q_f(0,0)\}$. 
+\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 

### 2023-10-31 VMCAI-revised-version/final.tex
up

@@
 \begin{proof}
-First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
-	This is due to the undecidability of $2$-counter Minsky machine termination.
+First, note that deciding whether from all inputs, all runs are terminating
+ is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well~\cite{Min61}. Hence 
+  deciding whether from all inputs, there exists at least one terminating run
+ is undecidable for Minsky machines with at least $2$ counters, since,
+ for deterministic Minsky machines, the two problems are one and the same.
+
+Now, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
+%	This is due to the undecidability of $2$-counter Minsky machine structural termination.
 From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
@@
 $\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
-from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable~\cite{Min61}. 
+from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable. 
 Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
@@
 
-% \alain{écrire qqchose ici}
 Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
@@
 %      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS in which
+The simulation operates by building a reset-VASS 
+which mimick the behaviors of the Minsky machines' counters by using two mirror counters 
+for each of them,
+and in which
 there exists
@@
 an initial state %$(s,x_1 \ldots x_n)$
+	with equal mirror counters
 to
@@
   (from Theorem~$5$ in \cite{araki1976PN}).
+  []
 Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 

### 2023-10-31 VMCAI-revised-version/final.tex
up

@@
 %
-Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the 
-{\sc Structural existential halting} of reset-VASS.
+Unfortunately {\sc State-resilience} is undecidable in this case. 
 %\mathieu{GOTO}
+Unfortunately {\sc State-resilience} is undecidable in this case.
+As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
+Section~\ref{safe-down} and devellopped more in Appendix~\ref{minsk}.
 
@@
   (from Theorem~$5$ in \cite{araki1976PN}).
-  []
+  [... from initial states with unequal mirror counters ... -> go back to a state with equal mirror counters -> then ...]
 Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 

### 2023-10-31 VMCAI-revised-version/final.tex
up

@@
 As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-Section~\ref{safe-down} and devellopped more in Appendix~\ref{minsk}.
+Section~\ref{safe-down} and developped more in Appendix~\ref{minsk}.
 

### 2023-10-31 VMCAI-revised-version/final.tex
up

@@
 
-We now prove that {\sc Resilience} is undecidable for reset-VASS when $\Safe = \mathop{\downarrow} \Safe$.
-
-\mathieu{J'ai remis une vielle version que j'avais écrit avant de basculer sur le structural existential halting, je pense guarder ça en main text et détailler davantage dans l'appendix dit moi ce que tu en pense}
-In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
+We now prove that {\sc Resilience} is undecidable for reset-VASS when 
+$\Safe = \mathop{\downarrow} \Safe$.
+
+% \mathieu{J'ai remis une vielle version que j'avais écrit avant de basculer sur le structural existential halting, je pense guarder ça en main text et détailler davantage dans l'appendix dit moi ce que tu en pense}
+% In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
 
@@
 since it is possible to have $\Safe = \{q_f(0,0)\}$. 
+This stems from the fact that it is undecidable whether for a Minsky machine $M$ there exists a terminating run for every possible initial state.
 \alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
@@
  in the Minsky machine and vice versa.
-  By combining both constructions, which can be seen in more details in Appendix~\ref{minsk},
+ The construction rely on using pairs of counters to represent counters of the Minsky machine,
+where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
+otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
+Using the construction, which can be seen in more details in Appendix~\ref{minsk},
  we obtain that 

### 2023-10-31 VMCAI-revised-version/final.tex
up

@@
 for each of them,
-and in which
+and where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation,
+otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
+The construction is such that
 there exists

### 2023-11-02 VMCAI-revised-version/final.tex
up

@@
 \problemx{zero-reachability}
-{A reset-VASS $V=(Q,T)$ of dimension $d$, and $p(\textbf{u}) \in Q \times \N^d$.} 
-{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, $q_f \in Q$, $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $%\exists q \in Q, ~ 
+p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}
 
@@
 {\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
-{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q(\textbf{0})$,
+{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
 which could be unreachable from other reachable states. 
@@
 \begin{proposition}\label{liveness reset}
-The  zero reachability problem can be reduced to {\sc $t$-liveness}.
+The  zero-reachability problem can be reduced to {\sc $t$-liveness}.
 \end{proposition}
@@
 %\mathieu{GOTO}
-Unfortunately {\sc State-resilience} is undecidable in this case.
-As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in
-Section~\ref{safe-down} and developped more in Appendix~\ref{minsk}.
+This stems from undeciability of zero-reachability in reset-VASS.
 
@@
  \end{proposition}
+
+\begin{proof}
+Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
+we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
+$q_0({\bf 0})$
+by a transition reseting every counter and
+changing the control-state to $q_0$.
+Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
+iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $N'$,
+which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
+and $s_0 = q_0({\bf 0})$.
+\end{proof}
 
@@
 \problemx{zero-reachability}
-{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
-{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}
+{A $d$-reset-VASS $V=(Q,T)$, $q_f \in Q$ and $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $%\exists q \in Q, ~ 
+p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}
 
@@
 \textbf{Proposition~\ref{liveness reset}.}
-{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\
+{\em In reset Petri nets, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}\\
 
@@
 \textbf{Corollary}
-{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}
+{\em In reset-VASS, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}
 
@@
 First, note that deciding whether from all inputs, all runs are terminating
- is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well~\cite{Min61}. Hence 
+ is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well. Hence 
   deciding whether from all inputs, there exists at least one terminating run
@@
 
-\iffalse
-\section{Structural halting in reset-VASS}
-
-We claim {\sc Structual existential halting} in reset-VASS is undecidable.  
-This is based on the construction providng the undecidability of structural termination in Lossy counter machines and in reset-VASS~\cite{dufourd1998reset,mayr1998lossy}.
-Strictly speaking undecidability of
-{\sc Structual existential halting} is not considered in
-\cite{mayr1998lossy},
-but 
-the encoding of Minsky machines can be reused
-since 
- {\sc Structual existential halting}
- is equivalent to
- structural termination in the case of deterministic Minsky machines and is thus
- undecidable for Minsky machines in general.
-\fi
-
 \section{Decision procedure for VASS when $\Safe = \mathop{\downarrow} \Safe$}\label{appendix vass}

### 2023-11-02 VMCAI-revised-version/final.tex
new simpler proof for undecidability

@@
 \begin{proof}
+Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
+we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
+$q_0({\bf 0})$
+by a transition reseting every counter and
+changing the control-state to $q_0$.
+Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
+iff $q_f({\bf 0})$ is reachable from every element in $V'$,
+which corresponds to {\sc Resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$.
+\end{proof}
+
+	\iffalse
+\begin{proof}
 First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $2$ counters when
@@
 $\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
+	\fi
 
@@
 Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
-iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $N'$,
+iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $V'$,
 which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
@@
 
+	\iffalse
 \section{Reset-VASS and Minsky-machines}\label{minsk}
@@
 upon which rely
-Proposition~\ref{indec WSTS with dcs}
-and
-Proposition~\ref{srp down up},
+Proposition~\ref{indec WSTS with dcs},
+% and
+% Proposition~\ref{srp down up},
 namely, that
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$
-are undecidable for Minsky machines
+{\sc Resilience} %and {\sc State-resilience} 
+when $\Safe = \mathop{\downarrow}~\Safe$
+%are
+is
+%
+ undecidable for Minsky machines
 and 
@@
 
-The problems are undecidable for a Minsky machine with at least $2$ counters when
+The 
+% problems are
+problem is undecidable for a Minsky machine with at least $2$ counters when
 $\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
@@
   [... from initial states with unequal mirror counters ... -> go back to a state with equal mirror counters -> then ...]
+  [ initial states with unequal mirror counters should be allowed to have some kind of backtracking to them]
 Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
@@
 \end{proof}
+
+	\fi
 

### 2023-11-02 VMCAI-revised-version/final.tex
New proof only for prop~2 since the other prop use same proof

@@
 
-%In the case of $\pred$-basis-effective (upward) compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.  This stems from the fact
-%{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some control-state $q_f$ is equivalent to
-%deciding structural %termination of reset-VASS, 
-%			existential halting of reset-VASS, 
-%which is undecidable.
-
 We now prove that {\sc Resilience} is undecidable for reset-VASS when 
-$\Safe = \mathop{\downarrow} \Safe$.
-
-% \mathieu{J'ai remis une vielle version que j'avais écrit avant de basculer sur le structural existential halting, je pense guarder ça en main text et détailler davantage dans l'appendix dit moi ce que tu en pense}
-% In the case of $\pred$-basis-effective upward compatible WSTS, {\sc Resilience} is undecidable when $\Safe = \mathop{\downarrow} \Safe$.
+$\Safe = \mathop{\downarrow} \Safe$ by a reduction from the
+ undecidable~\cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS% of dimension $d$%, which consists in, given a reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$, deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$%
+ .
+
+\problemx{zero-reachability}
+{A reset-VASS $V=(Q,T)$ of dimension $d$, $q_f \in Q$, $p(\textbf{u}) \in Q \times \N^d$.} 
+{  $%\exists q \in Q, ~ 
+p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}
 
@@
 \begin{proof}
-Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
-we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
+Let $V$ be a reset-VASS, $q_0({\bf 0})$ an initial state of $V$,
+and $q_f({\bf 0})$ a terminal state of $V$.
+We build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
 $q_0({\bf 0})$
 by a transition reseting every counter and
-changing the control-state to $q_0$.
-Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
-iff $q_f({\bf 0})$ is reachable from every element in $V'$,
+changing the control-state to $q_0$. Remark that the set of states 
+reachable from $q_0({\bf 0})$ is the same in $V'$ and in $V$.
+Then $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$
+iff $q_f({\bf 0})$ is reachable from every state in $V'$,
 which corresponds to {\sc Resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$.
+Indeed, if $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$,
+then, since $q_0({\bf 0})$ is reachable from every state in $V'$,
+$q_f({\bf 0})$ is reachable from every state in $V'$.
+In the other direction, if $q_f({\bf 0})$ is reachable from every state in $V'$,
+then in particular $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V'$,
+hence $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V'$.
 \end{proof}
-
-	\iffalse
-\begin{proof}
-First, note that {\sc Resilience} is undecidable for a Minsky machine $M$ with at least $2$ counters when
-$\Safe = \mathop{\downarrow} \Safe$,
-since it is possible to have $\Safe = \{q_f(0,0)\}$. 
-This stems from the fact that it is undecidable whether for a Minsky machine $M$ there exists a terminating run for every possible initial state.
-\alain{et alors ? d'où sais-tu que le pb est indécidable pour safe=q,0 ? Bien sûr il l'est mais pas immédiatement puisque ce n'est pas le reachability}
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset} in such a way that any sequence 
-% from  a state $p({\bf u})$
-% to
-reaching
- a state 
-%
-$ q({\bf 0})$ corresponds to one that reaches 
-$q_f(0,0)$
- in the Minsky machine and vice versa.
- The construction rely on using pairs of counters to represent counters of the Minsky machine,
-where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation, 
-otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
-Using the construction, which can be seen in more details in Appendix~\ref{minsk},
- we obtain that 
- {\sc Resilience} is undecidable for reset-VASS when $\Safe = \downarrow \Safe$, and hence for  $\post$-effective, $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed \end{proof}
-	\fi
-
-	\iffalse
-
-We will use the undecidability of the structural existential halting problem.
-%
-%			hence it will be also undecidable for (upward) compatible WSTS when $\Safe = \mathop{\downarrow} \Safe$.
-%
-%Indeed, the structural existential halting problem consists in deciding if 
-%a system %terminates
-%have a terminating (i.e. finite) run
-% from every initial state.
-
-% \problemx{Structural universal halting}
-% {A reset-VASS $V=(Q,T)$ of dimension $d$.}
-% { Do all runs for every initial state terminate ?\\}
-
-\problemx{Structural existential halting}
-%			{A reset-VASS $V=(Q,T)$ of dimension $d$.}
-{A transition system $\mathscr{S}=(S,\rightarrow)$}
-{ For every initial state $s_0 \in S$, does there exist a run from $s_0$ that terminates ?\\}
-
-{\sc Structural existential halting} is undecidable for reset-VASS since termination is undecidable for deterministic counter machines. 
-\alain{REFERENCE !!!!}
-%			which leads to the following proposition.
-
-% \mathieu{GOTO}
-
-\begin{proposition}\label{indec WSTS with dcs}
-{\sc Resilience} is undecidable for reset-VASS when $\Safe=\mathop{\downarrow} \Safe$, hence it is also undecidable for $\post$-effective and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
-\end{proposition}
-
-\begin{proof}
-\nocite{araki1976PN,dufourd1998reset}
-{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to {\sc Structural existential halting} of reset-VASS,
-since a run terminates w.l.o.g. iff it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}
-
-	\fi
 
@@
 
+\noindent
 It now remains to argue why {\sc $t$-liveness} is undecidable.
-Recall the
- undecidable~\cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS% of dimension $d$%, which consists in, given a reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$, deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$%
- .
+Recall {\sc zero-reachability} in reset-VASS
+is undecidable~\cite{araki1976PN}.
+\iffalse
 
@@
 % 
+\fi
 {\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
-{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
+{\sc zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
 which could be unreachable from other reachable states. 
@@
 Despite this it is possible to reduce 
-the {\em zero-reachability problem} 
+the {\sc zero-reachability} 
 to
@@
 \begin{proposition}\label{liveness reset}
-The  zero-reachability problem can be reduced to {\sc $t$-liveness}.
+{\sc zero-reachability} can be reduced to {\sc $t$-liveness}.
 \end{proposition}
@@
 % We may now prove the undecidability of the three state-resilience problems.
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the reduction implies 
+Since {\sc zero-reachability} for reset-VASS is undecidable \cite{dufourd1998reset}, the reduction implies 
 {\sc $t$-liveness} is undecidable.
@@
 \end{proposition}
-
-\iffalse \begin{proof}
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
-{\sc $t$-liveness} is undecidable.
-Because
-{\sc $t$-liveness} is undecidable in reset-VASS
-and reducible to {\sc State-resilience}, we also deduce that {\sc State-resilience} is undecidable for reset-VASS, which are WSTS with strong compatibility.
-Additionally, in WSTS with strong compatibility and effective $\pred$-basis, {\sc Bounded-state-resilience} is
-reducible to {\sc $k$-state-resilience}:
-since $\Safe=\mathop{\uparrow} \Safe$ and
-$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong 
-compatibility, then $\pred^{\leq n}(\Safe)= \mathop{\uparrow}~\pred^{\leq n}(\Safe)$ for all $n \in \N$,
-and there exists $n_0 \in \N$ such that 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^*(\Safe) = \pred^*(\Safe)$.
-We compute 
-$n_0$, then iteratively check whether $k$-state-resilience 
-hold for $k$ from $0$ to $n_0$.  
-Furthermore, in WSTS with strong compatibility and effective $\pred$-basis,  $\Safe=\mathop{\uparrow} \Safe$, {\sc Bounded-state-resilience} is equivalent to {\sc State-resilience},
-since 
-$\pred^{\leq n_0}(\Safe) = \mathop{\uparrow} \pred^{\leq n_0}(\Safe) = {\mathop{\uparrow} \pred^*(\Safe)} = \pred^*(\Safe)$.
-Hence the undecidability of {\sc Bounded-state-resilience}
-and 
- {\sc $k$-state-resilience}. \qed
-\end{proof} \fi
 
@@
 \begin{proposition}[Proposition 1, \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
-For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
+For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of 
+$\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
 \end{proposition}
@@
 %\mathieu{GOTO}
-This stems from undeciability of zero-reachability in reset-VASS.
+This stems from undeciability of {\sc zero-reachability} in reset-VASS,
+as seen in the reduction proof of Proposition~\ref{indec WSTS with dcs}
 
@@
 
-Let us recall the {\em zero-reachability} problem in (reset-)VASS.
+Let us recall the {\sc zero-reachability} problem in (reset-)VASS.
 
@@
 
-The zero-reachability problem is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
+{\sc zero-reachability} is decidable in VASS and it is undecidable in reset-VASS \cite{araki1976PN}.
 For a better representation, we will use reset Petri nets \cite{dufourd1998reset} rather than reset-VASS. 
@@
 
-We will show that the {\em zero-reachability} problem can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by 
+We will show that  {\sc zero-reachability}  can be reduced to {\sc $t$-liveness} in reset Petri nets; this is done by 
 % "adjusting" a similar result for Petri nets (Theorem~5.5 in  \cite{peterson1981petri}).\\
@@
 \textbf{Proposition~\ref{liveness reset}.}
-{\em In reset Petri nets, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}\\
+{\em In reset Petri nets, {\sc zero-reachability} can be reduced to {\sc $t$-liveness}.}\\
 
@@
 \textbf{Corollary}
-{\em In reset-VASS, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}
+{\em In reset-VASS, {\sc zero-reachability} can be reduced to {\sc $t$-liveness}.}
 
@@
 
-Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies {\sc $t$-liveness} is undecidable as well.
+Since {\sc zero-reachability} for reset-VASS is undecidable, the reduction implies {\sc $t$-liveness} is undecidable as well.
 This leads to the undecidability of {\sc State resilience}, {\sc Bounded-state-resilience} and
@@
 \begin{proof}
-Since the zero-reachability problem for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
+Since {\sc zero-reachability} for reset-VASS is undecidable \cite{dufourd1998reset}, the previous reduction (Proposition \ref{liveness reset}) implies 
 {\sc $t$-liveness} is undecidable.

### 2023-11-02 VMCAI-revised-version/final.tex
up

@@
 iff $q_f({\bf 0})$ is reachable from every state in $V'$,
-which corresponds to {\sc Resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$.
+which corresponds to {\sc Resilience} in $V'$ for $\Safe =  \{ q_f({\bf 0}) \}$.
 Indeed, if $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$,
@@
 then in particular $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V'$,
-hence $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V'$.
+hence $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$.
 \end{proof}
@@
 This stems from undeciability of {\sc zero-reachability} in reset-VASS,
-as seen in the reduction proof of Proposition~\ref{indec WSTS with dcs}
+as seen in the reduction proof of Proposition~\ref{indec WSTS with dcs}.
 

### 2023-11-02 VMCAI-revised-version/final.tex
up

@@
 \end{center}
+\newpage
 
@@
 \fi
-{\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
+\iffalse {\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
 {\sc zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
@@
 Despite this it is possible to reduce 
-the {\sc zero-reachability} 
+ {\sc zero-reachability} 
 to
 {\sc $t$-liveness}.
+\fi
 % The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets,
@@
 
-Let us recall the {\sc zero-reachability} problem in (reset-)VASS.
+Let us recall the {\em zero-reachability} problem in (reset-)VASS.
 

### 2023-11-02 VMCAI-revised-version/final.tex
up

@@
 \end{proposition}
+
+Recall that on the other end {\sc Resilience} is decidable for reset-VASS when $\Safe = \mathop{\uparrow} \Safe$ by Theorem~\ref{xxx}.
+This suggests that the undecidability of {\sc State-resilience} 
+comes more from the fact $\post^*(s_0)$ is not constructible in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.  
+\mathieu{À relire encore}
 

### 2023-11-02 VMCAI-revised-version/final.tex
up

--- 2023-11-02 VMCAI-revised-version/final.tex
@@
 \end{proposition}
+
+Recall that on the other end {\sc Resilience} is decidable for reset-VASS when $\Safe = \mathop{\uparrow} \Safe$ by Theorem~\ref{xxx}.
+This suggests that the undecidability of {\sc State-resilience} 
+comes more from the fact $\post^*(s_0)$ is not constructible~\cite{mayr1998lossy} in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.  
+\mathieu{À relire encore}
 

### 2023-11-02 VMCAI-revised-version/final.tex
up

--- 2023-11-02 VMCAI-revised-version/final.tex
@@
 \end{proposition}
+
+Recall that on the other end {\sc Resilience} is decidable for reset-VASS when $\Safe = \mathop{\uparrow} \Safe$ by Theorem~\ref{xxx}.
+This suggests that the undecidability of {\sc State-resilience} 
+comes more from the fact $\post^*(s_0)$ is not constructible~\cite{dufourd1998reset} in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.
+\nocite{mayr1998lossy}
+\mathieu{À relire encore}
 

### 2023-11-02 VMCAI-revised-version/final.tex
up

--- 2023-11-02 VMCAI-revised-version/final.tex
@@
 This suggests that the undecidability of {\sc State-resilience} 
-comes more from the fact $\post^*(s_0)$ is not constructible in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.  
+comes more from the fact $\post^*(s_0)$ is not constructible~\cite{DBLP:conf/icalp/DufordJS99} in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.
+% \nocite{mayr1998lossy}
 \mathieu{À relire encore}

### 2023-11-02 VMCAI-revised-version/final.tex
up

--- 2023-11-02 VMCAI-revised-version/final.tex
@@
 
-Recall that on the other end {\sc Resilience} is decidable for reset-VASS when $\Safe = \mathop{\uparrow} \Safe$ by Theorem~\ref{xxx}.
+Recall that on the other hand {\sc Resilience} is decidable for reset-VASS when $\Safe = \mathop{\uparrow} \Safe$ by Theorem~\ref{xxx}.
 This suggests that the undecidability of {\sc State-resilience} 
-comes more from the fact $\post^*(s_0)$ is not constructible~\cite{dufourd1998reset} in reset-VASS rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.
+comes more from the fact $\mathop{\uparrow} \post^*(s_0)$ is not constructible in reset-VASS \-- a consequence of the undecidability of reachability \-- rather than from the difficulties inherent in searching for paths from $\post^*(s_0)$ to $\Safe$.
 \nocite{mayr1998lossy}
-\mathieu{À relire encore}
 

### 2023-11-02 VMCAI-revised-version/final.tex
ANR grant

@@
 \textbf{Lemma~\ref{downward compatible}.}
-{\em For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred_{\mathscr{S}}^*(D)$ is downward-closed.}
+{\em For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.}
 

### 2023-11-03 VMCAI-revised-version/final.tex

@@
 then in particular $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V'$,
-hence $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$.
+hence $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $V$.\qed
 \end{proof}

### 2023-11-03 VMCAI-revised-version/final.tex
retiré les iffalse et les \comments

@@
 
-%We study the decidability of each type of resilience for many infinite-state transitiuon systems: Well Behaved Transition Systems (WBTS), Well Structured Transition Systems (WSTS), VASS and other variations of counter machines.
-%
 We first show that all resilience problems are undecidable for effective Well Structured Transition Systems (WSTS) with strong compatibility.
-%
-%and for upward-closed sets $\Safe$; resilience is also undecidable for reset-VASS (that are WSTS with effective $\pred$-basis) when $\Safe$ are downward-closed sets.
-%
 We then show that resilience is decidable for Well Behaved Transition Systems (WBTS) and for WSTS with adapted effectiveness hypotheses.
-%
-%with the computability of the finite basis of the downward-closure of the set of (immediate) successors of a downward-closed set ($\post$-basis-effective) 
-%
-%
-%			with the (well-known) $\pred$-basis-effective hypothesis.
-%
-%		most of all resilience problems are decidable for WSTS with the good effective hypothesis. R
-%
 Most of the resilience properties are shown decidable for other classes like WSTS with the downward compatibility, VASS, lossy counter machines, reset-VASS, integer VASS and continuous VASS.
-
-%Then we prove the decidability of resilience for WBTS and WSTS with strong compatibility, adapted effective hypothesis and for upward-closed sets $\Safe$.
-Moreover, some resilience properties are decidable for three other classes of WSTS : (1) WSTS with effective
-%$\mathop{\uparrow}$ $\post^*$ basis and $\Safe=\mathop{\uparrow} \Safe$ ; (2) ideal-effective WSTS with downward and upward compatibilities ; and (3) ideal-effective downward-compatible WSTS with $\Safe=\mathop{\downarrow} \Safe$.
-%Finally, we study the resilience for VASS with semi-linear subsets $\Bad$ and $\Safe$ and for variations of VASS (lossy counter machines,  $\mathbb{Z}-$VASS and continuous VASS)
 \end{abstract}
@@
 \keywords{Verification, Resilience, Home-Space, Well-structured transition systems, Vector addition system with states}
-
-%\alain{j'ai ajouté les numéros et j'ai enlevé les noms dans les {theorem, proposition, definition} qu'on cite}
-
-%\alain{ok pour la notation $q(v)$, je vois qu'elle est déjà utilisée- peutx-tu changer pour review 3: $A = \mathop{\downarrow} B$ instead of $A = \mathop{\downarrow} B$: this dangling arrow is typographically horrific!}
 
@@
 Some systems are subjects at frequent intervals to accidents, attacks or changes.
-%				Think for instance of a supply chain, or an airport’s air trafic control. 
 In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack
@@
 {\bf Resiliences.}  
-%
-%			If the home-space contains a single element, this element is an {\em home-state}.
 Given a transition system whose set of states is $S$ and let $X,H \subseteq S$, we say that $(X,H)$ satisfies the \emph{home-space} property ~\cite{DBLP:conf/ac/MemmiV86} if the reachability set from $X$ is included in the set of predecessors of $H$.
@@
 %
-%				(i.e., $\post^*(X) \subseteq \pred^*(H)$).
 
@@
 It is {\em state-resilient} if it is resilient for $(\Bad,\Safe)$ where $\Bad$ contains an unique state.
-%
-%			(resp. $\post^*(s_0)$)
-%                  It  is state-resilient if
-%			from every state, there is a run that goes into $\Safe$.
-%
-%			$\mathscr{S}$ is {\em resilient for $\Safe$} if from every state, there is a run that goes into  $\Safe$.
-%			$S \subseteq \pred^*(\Safe)$. 
-%			this is the $(S,\Safe)$ home-space problem.
-%			($\post^*(S) \subseteq \pred^*(\Safe)$).
-%
-% 		for $S,\Safe$: $\mathscr{S}$ is $\Safe$-resilient if $\post^*(S) \subseteq \pred^*(\Safe)$.
-%		Similarly, $\mathscr{S}$ is $\Safe$-state-resilient for an initial state $s_0$ if  
-%		$\post^*(s_0) \subseteq \pred^*(\Safe)$.
-%		We will study three decidability questions for each type of set $\Safe$.
-%		The resilience problem is to decide whether a transition system $\mathscr{S} = (S,\rightarrow )$ is $\Safe$-resilient 			(for a given subset of states $\Safe \subseteq S$).
-%		The resilience problem could be easily generalized for two subsets $\Bad$ and $\Safe$ by asking
-%		whether $\Bad \subseteq \pred^*(\Safe)$.
 The {\em $k$-resilience} problem, for $k\geq0$, is to decide whether from any state is it always possible to reach $\Safe$ with a run of length smaller than $k$ 
-%
-%		(i.e., $S \subseteq \pred^{\leq k}(\Safe)$)
-%
  and 
@@
 {\bf State of the art.}
-%
-%			{\bf Home-spaces.}
 
@@
 %
-%		reduces to coverability which is decidable on WSTS.
-%		They gave conditions that insure that a transition system is an effective WSTS with both upward and reflexive 		downward compatibilities (and some other technical conditions). 
-%              In this framework and under these hypotheses, 
-%			resilience reduces to coverability which is decidable on WSTS. 
-%%%%%%%%%
 %
@@
 %
-%			(i.e., $\mathop{\uparrow} \post^*(s)$) 
-%
 when $\Safe$ is upward-closed.
 %
-%			(called effective 
-%			$\mathop{\uparrow}$ $\post^*$ basis) in the case $\Safe=\mathop{\uparrow} \Safe$ and $\Bad = \mathop{\downarrow} \Bad$.
-%%%%%%%%%
 
@@
 $X$ a singleton) problems are both undecidable for WSTS with strong compatibility and for $\Safe$ upward-closed or downward-closed.
-%Moreover, {state-resilience},
-%{bounded-state-resilience} and
-%{$k$-state-resilience}
-%are undecidable for reset-VASS (that are strongly compatible $\pred$-basis-effective WSTS) when
-%$\Safe$ is upward-closed. 
-%			We made a reduction of zero-reachability in reset-VASS to {state-resilience} in reset-VASS.
+
 %
@@
 %
-%
-%    and
-%the additional hypothesis that
-%for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
-%
-%				and for post-ideal-effective downward-WSTS ????
-%%%
-%
+
 \item We clarify the different effectiness conditions on WBTS and WSTS (Definition \ref{eff}) that allow the decidability of the six resilience.
@@
 \item We generalize with Theorem \ref{post srp} the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability (Proposition \ref{srp up down}).
-%
-%			à utiliser dans le texte ???
-%
-%We first remark that the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} can be generalized by relaxing the  strong compatibility hypothesis.
-We show that removing the effective
-%$\mathop{\uparrow}$ $\post^*$-basis hypothesis leads to undecidability; however, this hypothesis seems very strong since, many WSTS (like reset-VASS) don't enjoy this property.
 
@@
 The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WBTS with strong downward compatibility.
-%
-%			à utiliser dans le texte ???
-%
-%\item We remark that the reflexive downward compatibility (of an effective WSTS) hypothesis in  \cite{DBLP:journals/corr/PrasadZ16} implies the existence of an algorithm that computes a finite basis of the upward-closure of the reachability set from $s$,
-%%
-%%			$\mathop{\uparrow} \post^*(s)$ 
-%%
-%for all states $s$; this provides a way to use \cite{DBLP:journals/corr/abs-2108-00889} for proving the announced result by Prasad and Zuck.
-%However, neither Prasad \& Zuck nor \"Ozkan \& Würdemann established the strong relation between resilience and the home-space property.
-%
+
 \item We study the resilience problems for VASS and variations of VASS where most of the resilience problems are shown decidable.
 \end{itemize}
-
-%				detailed on page~\pageref{on effectivity}.
-
-\iffalse
- \begin{center}
-	\begin{figure}
-			\hspace{0.8cm}
-\includegraphics[width=0.75\textwidth]{FigureCD}
-	\caption{Two reset-VASS with three control-states and one counter.}
-					\label{r-V}
-	\end{figure}
-\end{center}
-
-\begin{example}\label{Example}
-{(\bf Reset-VASS example).}
-Consider the 
- reset-VASS $V_1$ from Figure~\ref{r-V}.
-Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {resilience}, 
-{$k$-resilience} and {bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$.
-Consider now $\Safe = \{q_1(0)\} $; then {resilience} hold: it is possible to reach $q_1(0)$ from any state of $V_1$.
-However, {bounded resilience} does not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of length less than $n$ towards $q_1(0)$.
-Consider now the
- reset-VASS $V_2$, which is $V_1$ plus an $+1$ transition from $q_1$ back to $q_0$. In $V_2$, {bounded resilience} hold, since $6$-resilience hold. 
-Indeed, from any state with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the (unique) reset transition.
-From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less.
-From
-$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps. 
-\end{example}
-%%%%%
-
-%%%%%
-% Hence we can see how adding transitions can 
-% turn a system that is not bounded resilient into one that is. 
-% \alain{à quoi sert cette remarque ?}
-\fi
-
-\iffalse
-\begin{center}
- \resizebox{\textwidth}{!}{
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \ 
-		& $\Safe = {\uparrow \Safe}$~ 
-		 & $\Safe = {\downarrow \Safe}$~ 
- \\ \hline
-   \multirow{3}{*}{RP} 
-  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
-   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down} \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		& VASS \\\cline{2-3}
-~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS} & Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs}    \\ \hline
-   \multirow{3}{*}{BRP} 
-  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
-   		 & VASS \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		&   \\\cline{2-3}
-~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS}  & ?    \\ \hline
-   \multirow{3}{*}{kRP} 
- 	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
-   		 & VASS \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		& \\\cline{2-3}
-~	& ? & ?    \\ \hline
-
-   \multirow{2}{*}{sRP} 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{post srp} 
-   		 & WSTS (downward + upward) Thm~9 \\ \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}    \\ \hline
-   \multirow{2}{*}{BsRP} 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
-   		 & WSTS (downward + upward) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & ?    \\ \hline
-   \multirow{2}{*}{ksRP} 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  
-   		 & WSTS (downward + upward) Cor~\ref{downward brp} \\ \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & ?    \\ \hline
-
- \end{tabular}
-  }
-\end{center}
-\fi
-
-\iffalse
-\begin{center}
- \resizebox{\textwidth}{!}{
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \
-		& {\bf WBTS model (with $\Safe = \mathop{\uparrow} \Safe$) }~ 
-		 & ~ {\bf Decidability}  ~ 
- \\ \hline
-   \multirow{3}{*}{ {\bf RP }} 
-  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
-   		 &    \multirow{2}{*}{Decidable}  \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		&  \\\cline{2-3}
-~	& WSTS ($\post$-effective) Prop~\ref{indec WSTS}  & Undecidable    \\ \hline
-   \multirow{3}{*}{{\bf BRP }} 
-  	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up}
-   		 & \multirow{2}{*}{Decidable}  \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		&   \\\cline{2-3}
-~	&  WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS}  & Undecidable    \\ \hline
-   \multirow{3}{*}{ {\bf kRP } } 
- 	& WBTS ($\post$-ideal-effective) Thm~\ref{down-up} 
-   		 & \multirow{2}{*}{Decidable}  \\ 
-~	& WSTS ($\pred$-basis-effective) Thm~\ref{xxx}
-		& \\\cline{2-3}
-~	& ? & Undecidable   \\ \hline
-
-   \multirow{3}{*}{ {\bf sRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective)  
-   		 & \multirow{2}{*}{Decidable} \\
-   	& Thm~\ref{post srp} &	 \\ \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable   \\ \hline
-   \multirow{3}{*}{ {\bf BsRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective) 
-   		 & \multirow{2}{*}{Decidable} \\
-   	& Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
-   	& \\ 	  \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
-   \multirow{3}{*}{ {\bf ksRP } } 
-  	& WSTS ($\pred$-basis-effective + $\post^*$-basis-effective)
-   		 & \multirow{2}{*}{Decidable} \\
-   	&  Thm~\ref{ref ozkan} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}  		& \\	  \cline{2-3}
-~	& WSTS ($\pred$-basis-effective+$\post$-effective) Prop~\ref{srp up down} & Undecidable    \\ \hline
-
- \end{tabular}
-  }
-\end{center}
-
-\begin{center}
- \resizebox{\textwidth}{!}{
-\begin{tabular}{ | l | c | c | c | r |}
-\hline   \ 
-		& {\bf WBTS model (with $\Safe = \mathop{\downarrow} \Safe$)} ~ 
-		 & ~ {\bf Decidability}  ~ 
- \\ \hline
-   \multirow{3}{*}{ {\bf RP }} 
-   		 & Downard-WBTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
-   		 	& \multirow{2}{*}{Decidable}  \\
-
-		& VASS & \\\cline{2-3}
-		& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{indec WSTS with dcs} 
-		& Undecidable   \\ \hline
-   \multirow{2}{*}{{\bf BRP }} 
-   		 & VASS 
-   		 & Decidable   \\\cline{2-3}
-~	&  ? & Undecidable    \\ \hline
-   \multirow{2}{*}{{\bf kRP }} 
-~	& VASS
-		& Decidable \\\cline{2-3}
-~	& ? & Undecidable    \\ \hline
-
-   \multirow{2}{*}{{\bf sRP ~ ~ }} 
-   		 & WBTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
-~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp down up}  & Undecidable  \\ \hline
-   \multirow{2}{*}{{\bf BsRP }} 
-   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
-~	&  ? & Undecidable   \\ \hline
-   \multirow{2}{*}{{\bf ksRP }} 
-   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
-~	 & ? & Undecidable    \\ \hline
-
- \end{tabular}
-  }
-\end{center}
-\fi
 
@@
 
-\iffalse
- \begin{center}
-	\begin{figure}
-%	\hspace{1.2cm}
-\includegraphics[width=0.70\textwidth]{Safe_up_R}
-	\end{figure}
-\end{center}
-
- \begin{center}
-	\begin{figure}
-\includegraphics[width=0.70\textwidth]{Safe_up_State_R}
-	\end{figure}
-\end{center}
-
- \begin{center}
-	\begin{figure}
-\includegraphics[width=0.70\textwidth]{Safe_down_R}
-	\end{figure}
-\end{center}
-\fi
-
-% \alain{je pense qu'il faut faire deux tableaux pour que ce soit lisible. attention, la décidabilité n'est pas encore indiquée}
-
 \noindent
@@
 
-% Section 2 introduit les notions, Section 3 et 4 les résultats de résilience et state résilience pour les WBTS, Section 5 sur les VASS et variantes
 Section~2 introduces the necessary definitions of WSTS and variations as well as the different notions of effectivity.
@@
 %
-%		Finkel and Schnoebelen, 
-%
 \begin{definition}
@@
 
-\iffalse\begin{proof}
-Let us prove $1 \implies 2$. 
-Let $D$ be a downward-closed subset of $S$
-and let $x \in \mathop{\downarrow} \pred^*(D)$.
-By downward closure, there exists
-$y \in \pred^*(D)$ 
-such that $x \leq y$.
-By definition of $\pred^*(D)$, there exists 
-$d \in D$ such that: (1) either $y=d$ and then $x \in D \subseteq \pred^*(D)$ or (2) there exist $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
-$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.
-%
-By downward compatibility $a_0 \to a_1$
-implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
-$x \to^* a'_1$.
-More generally $a_i \to a_{i+1}$ and
-$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
-$a'_i \to^* a'_{i+1}$,
-and, by induction,
- $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
-with $d' \leq d$.
-Since
-$d'$ 
-belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$ and then $\mathop{\downarrow}\pred^*(D) \subseteq \pred^*(D)$.
-Since $\pred^*(D) \subseteq \mathop{\downarrow} \pred^*(D)$ always hold, we deduce that $\pred^*(D)=\mathop{\downarrow}\pred^*(D)$ is downward-closed.
-
-Let us prove $2 \implies 1$.
-Let $s_1, t_1 , s_2 \in S$ with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, Since $s_1 \rightarrow t_1$ we deduce that $s_1 \in \pred^*(\mathop{\downarrow} t_1)$, and since $s_2 \leq s_1$ and $\pred^*(\mathop{\downarrow} t_1)$ is downward-closed, we also have $s_2 \in \pred^*(\mathop{\downarrow} t_1)$.
-This means that there exists $t_2 \leq t_1$ such that
-$s_2 \rightarrow^{*} t_2$. Hence, we proved that $\mathscr{S}$ is downward compatible.
-\qed
-\end{proof}\fi
-
-%A WSTS $\mathscr{S}=(S,\rightarrow,\leq)$ is {\em completion-$\post$-effective} if
-%it is $\post$-effective, and there exists an algorithm $M_\downarrow$ that computes 
-% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
-% $s_1, s_2, \ldots, s_m \in S$.
-
 To obtain decidability results, we must introduce some notions of \emph{effectiveness}.
@@
 
-%Most of usual WSTS are $\post$-effective and ideally-effective.
-
-% \alain{le rappel ci-dessous est-il entièrement utile ? Je pense qu'on peut l'enlever, et toi ?}
-% Now, we may recall a simple condition that insures that a finite basis of $\pred^*(\mathop{\uparrow} s )$ is computable for every $s \in S$.
-% We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
-% For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$.
-When $\mathscr{S}$ has strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges.
-Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t.
-$U_k = U_{k_0}$ for all $k \geq k_0$.
-We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$.
-When $\mathscr{S}$ have upward-compatibility but not strong compatibility, we can similarly compute $\pred^*(U)$ by studying the sequence $(\mathop{\uparrow} U_k)_k$ instead.
-
 Recall the \emph{coverability problem} for ordered transition systems.
@@
 
-% \alain{il manque des refs dans la fig 1 qui sont déjà dans la biblio: WBTS, omega2,on peut mettre les front lossy avec les LCM et LCS}
 %
@@
 
-%Again $Q \times \N^d$
-% is the set of states of $V$.
 For every states $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
@@
 
-\iffalse
-\problemx{Resilience (RP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
-{$S \longrightarrow^{*} \Safe$ ?\newline}
-
-\problemx{$k$-resilience (kRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a set $\Safe \subseteq S$.}
-{$S \longrightarrow^{\leq k} \Safe$ ?\newline}
-
-\problemx{Bounded resilience (BRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
-{$\exists k \geq 0$ such that {$S \longrightarrow^{\leq k} \Safe$ ?\newline}}
-\fi
-
 \problemx{Resilience problems}
@@
 %
-% \alain{pourquoi n'a-t-on pas prouvé la même chose pour la BR et la k-R ?: où est la difficulté ?}
 \begin{proposition}\label{indec WSTS}
@@
 \end{proposition}
-% \alain{peut-on ajouter des hypothèses d'effectivity ?
-peut-on montrer que l'exemple n'est pas pred effectif ?
-l'exemple est-il  ideal-effective (oui pour les propriétes 1 et 2 mais pas 3 car ça permettrait de décider l'arrêt ?)?
-on voit que l'exemple n'est pas completion-$\post$-effective vu le theoreme qui suit}
 \begin{proof}
@@
 
-% \mathieu{Reviewer 1: the example at the bottom of \textbf{ page~\pageref{indec WSTS}}, even if it is a WSTS, is a very particular case [...] wonder if classes containing such  systems are interesting per se.
-=> citer un autre résultat où les increasing recursive functions ont été utilisées ?}
-% \alain{ma réponse: termination and boundedness are decidable for increasing recursive (functional) WSTS \cite{FMP-ic04}; the decidability of coverability and place-boundedness depends on the effectivity hypothesis on recursive functions.}
-
 We now prove that {\sc Resilience} is undecidable for reset-VASS when 
@@
 
-%\alain{a-t-on le même rsultat pour les WSTS avec pred* calculable ? J'ai enlevé l'hypothèse omega2 car j'ai l'impression qu'elle n'est pas utile...à vérifier}
-
 \begin{proof}
@@
 %%%%%%%
-%%\alain{prouver, si possible, les deux autres résilience}
 Strong compatibility implies furthermore the decidability
@@
 
-%				The completion-$\post$-effective hypothesis is needed to decide {\sc Resilience}.
-%%%
-%%%
-%Indeed, consider the family $\{ f_j: \N^2 \to \N^2\}$ of 
-%increasing recursive functions from \cite{FMP-ic04} defined as
-%
-%$ f_j(n,k) = \begin{dcases*} (n,0) & \text{ if } k=0 \text{ and } TM$_j$ \text{ runs for more than } n \text{ steps} \\
-%		(n,n+k) &\text{ otherwise,} \end{dcases*} $
-%		
-%	\noindent	
-%	where TM$_j$ is the $j$-th Turing machine (in a classical enumeration)
-%	which moreover begins
-%by writing the integer $j$ on its tape,	
-%	and consider additionnally the function $g : \N^2 \to \N^2$ defined by $g(n,k) = (n+1,k)$. 
-%	The
-%	transition system 
-%	$S_j = (\N^2, \{f_j, g\}, \leq)$ 
-%	is a WSTS
-%	and
-%	has the property that
-%	it is
-%	$\mathop{\uparrow} (1,1)$-resilient
-%	iff TM$_j$ halts on input $j$,
-%	since $(1,1)$ is coverable from $(0,0)$
-%	iff TM$_j$ halts on input $j$. 
-%	Hence there is no Turing machine which correctly determines 
-%	resilience and halts whenever its input is a
-%	WSTS.
-%Hence resilience for WSTS in general is undecidable.
-
 \subsection{Case: $\Safe=\mathop{\downarrow} \Safe$.}\label{safe-down}
@@
 The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
-%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
 Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
@@
 
-\iffalse
-
-\problemx{State-resilience (SRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
-{$\post^*(s)  \rightarrow^{*} \Safe $ ? \newline}
-
-\problemx{$k$-state-resilience (kSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, $k \geq 0$ and $\Safe \subseteq S$.}
-{ $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-
-\problemx{bounded-state-resilience (BSRP)}
-{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$ and $\Safe \subseteq S$.}
-{$\exists k \geq 0$ such that $\post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
-
-\fi
-
 \problemx{State resilience problems}
@@
 is undecidable~\cite{araki1976PN}.
-\iffalse
-
-\problemx{zero-reachability}
-{A reset-VASS $V=(Q,T)$ of dimension $d$, $q_f \in Q$, $p(\textbf{u}) \in Q \times \N^d$.} 
-{  $%\exists q \in Q, ~ 
-p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}
-
-% 
-\fi
-\iffalse {\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
-{\sc zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
-which could be unreachable from other reachable states. 
-%
-Despite this it is possible to reduce 
- {\sc zero-reachability} 
-to
-{\sc $t$-liveness}.
-\fi
-% The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets,
 The following Proposition is a variation  to reset-VASS 
@@
 
-% We may now prove the undecidability of the three state-resilience problems.
 Since {\sc zero-reachability} for reset-VASS is undecidable \cite{dufourd1998reset}, the reduction implies 
@@
 
-%\alain{énoncer pour les reset-VASS c'est plus fort}
 \begin{proposition}\label{srp up down}
@@
 %
-%		(\"Ozkan,   \"Ozkan and W{\"{u}}rdemann,  
 %
@@
 \end{theorem}
-%
-%
-%\alain{ce qui suit est à supprimer car subsumé par Theo 8}
-%%
-%We may immediately  \emph{generalize} this last result by strengthening to \emph{unbounded} {\sc State-resilience}. The proof is essentially the same than the previous one.
-%
-%\begin{proposition}\label{postcomputable}
-%{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
-%when
-%$\Safe=\mathop{\uparrow} \Safe$.
-%\end{proposition}
-%
-%\begin{proof}
-%\alain{où utilise-t-on la strong compatibility ? et si on ne l'utilise pas, c'est le theo 8...}
-%Since $\mathscr{S}$ is a WSTS there exists an $n_0 \in \N$ such that
-%$\pred^*(\Safe) =  \pred^{\leq n_0}(\Safe)$. We can compute the least $n_0$ by iteratively computing 
-%$\pred^{\leq n+1}(\Safe)$ from $\pred^{\leq n}(\Safe)$, checking 
-%$\pred^{\leq n+1}(\Safe) = \pred^{\leq n}(\Safe)$, 
-%returning $n$ if that is the case.
-%Then, because {\sc $n_0$-state-resilience} is decidable, 
-%checking $\mathop{\uparrow} \post^*(s) 
-% \subseteq \pred^{\leq n}(\Safe) = \pred^*(\Safe)$ is,
-%and {\sc State-resilience} is decidable. \qed
-%\end{proof}
-%%
-%\alain{ce qui précède est à supprimer car subsumé par Theo 8}
-%
 The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
@@
 \end{lemma}
-
-\iffalse \begin{proof}
-Let us suppose that $A \cap D \subseteq U$. Then ${\mathop{\uparrow} (A \cap D)} \subseteq {\mathop{\uparrow} U} = U$.
-Let us show that $({\mathop{\uparrow} A}) \cap D \subseteq {\mathop{\uparrow} (A \cap D)}$.
-Let $x \in ({\mathop{\uparrow} A}) \cap D$, then there exists $a \in A$ such that $x \geq a$.
-Since $x \in D$ and $D$ is downward-closed, we also have $a \in D$.
-Hence $a \in A \cap D$ and then $x \in { \mathop{\uparrow} (A \cap D)}$.
-In the other direction,
-since $A \subseteq {\mathop{\uparrow} A}$, the inclusion
-$({\mathop{\uparrow}  A}) \cap D \subseteq U$ implies
-$A \cap D \subseteq ({\mathop{\uparrow}  A}) \cap D \subseteq U$. \qed
-\end{proof} \fi
 
@@
 Lemma~\ref{Lemma intersection}. For the converse direction, one computes the sequence of upward-closed sets
-$U_n = \mathop{\uparrow} \post^{\leq n}(s)$ %\alain{il manque l'hypothèse $\post$-effective}
+$U_n = \mathop{\uparrow} \post^{\leq n}(s)$ 
 until it becomes stationnary. 
@@
 
-%\alain{Theo 8 est plus fort que prop 6}
 \begin{theorem}\label{post srp}
@@
 %
-%Recall that $\mathscr{S}$ is state-resilient iff  $\post^*(s) 
-% \subseteq \pred^*(\Safe)$ iff 
 Now the last inclusion is equivalent to:
@@
 Since both $B_{\mathop{\uparrow} \post^*(s)}$ and $B_{\pred^*(\Safe)}$ are finite, {\sc State-resilience} is decidable.\qed
-%
-%
-%			Since $B_{\mathop{\uparrow} \post^*(s)}$ is finite
-%\alain{comment calcule-t-on une telle base sans l'hypothèse $\pred$-basis-effective qui n'est pas dans les hypothèses du théorème ? Est-elle chez Ozkan ?} 
-%			we can check that $B_{\mathop{\uparrow} \post^*(s)} 
-%			 \setminus \pred^*(\Safe) = \emptyset$.  \qed
 \end{proof}
@@
 Unfortunately {\sc State-resilience} is undecidable in this case. 
-%\mathieu{GOTO}
 This stems from undeciability of {\sc zero-reachability} in reset-VASS,
@@
 
-\iffalse
-\begin{proof}
-Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
-we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
-$q_0({\bf 0})$
-by a transition reseting every counter and
-changing the control-state to $q_0$.
-Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
-iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $V'$,
-which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
-and $s_0 = q_0({\bf 0})$.
-\end{proof}
-\fi
-
 Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \mathop{\downarrow} \Safe$
@@
 %
-% For instance one can write the following lemma:
-%  Page 14: "For instance one can write the following lemma" => Rewrite this: we will need the following lemma for this and that.
 We will need the following lemma in order to obtain symmetrical results.
@@
 {\sc State-resilience} is decidable for $\post$-effective, 
-%$\pred$-basis-effective, 
 $\post$-ideal-effective and $\pred^*$-basis-effective WBTS with downward and upward compatibilities
@@
 
-In order to decide whether %$\post^{\leq n}(s)
+In order to decide whether 
 $\post^{ *}(s)  
@@
 The procedure then checks for every downward-closed subset $D_i$ whether $\mathop{\downarrow} \post(D_i) \subseteq D_i$ (this inclusion is decidable because $\mathscr{S}$ is $\post$-ideal-effective). 
-%\alain{on ne peut pas tester $\downarrow \post^*(D_i) \subseteq D_i$ sans expliquer pourquoi.
-L'hypothèse "$\post$-ideal-effective" permet de calculer $\downarrow \post(D_i)$.
-Il manquait la $\post$-effective dans les hypothèses}
 Every subset $D_i$ such that $\mathop{\downarrow} \post(D_i) \subseteq D_i$ is an “over-approximation” of $\mathop{\downarrow} \post^*(s)$ if it contains $s$.
@@
 $D   \subseteq \pred^*(\Safe)$.
-%\mathieu{ici $\pred^*(\Safe)$ est downward-closed.
-du coup pas sur que la $\pred$-basis-effectivité telle qu'elle est actuellement nous permette de trancher ...
-à moins que les hypothèses d'effectivté sur les idéaux nous permettent de guarantir qu'on puisse calculer le $\pred$ d'un idéal ?}
 
@@
 $\post^{\leq n}(s) $ (this is effective because $\mathscr{S}$ is $\post$-effective)
-%			\alain{ici on utilise $\post$-effective que j'ai rajoutée}
 until it finds an element
@@
 
-% \mathieu{Donner des exemples dans cette classe.}
-% \alain{je pense aux LCM avec insertions, elle  devrait avoir toutes les propriétés mais faut vérifier}
-
-$\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
-Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96%,DBLP:journals/iandc/CeceFI96
-%
-%% GOTO bibliographic reference when doing the bibtex final.aux did not work
-}.
+Note that $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility include for instance
+Lossy Channel Systems with insertions~\cite{DBLP:journals/iandc/AbdullaJ96}.
 
@@
 
-% The above results concern only unbounded resilience.
-We consider now  {\sc Bounded resilience} and {\sc $k$-resilience}, and focus on the length of the path from $\Bad$ to $\Safe$ when $\Safe$ is downward-closed.
-
 \section{Conclusion and perspectives}
@@
 up until now since these are fairly straightforward technical proofs.
-% \mathieu{mettre ici les preuves suivantes (?): lemme1, et lemme 2}
 We provide now both. \\
@@
 \end{proof}
-
-% Additionnally ...
 
@@
 
-	\iffalse
-\section{Reset-VASS and Minsky-machines}\label{minsk}
-
-We now delve into more details into the undecidability results 
-upon which rely
-Proposition~\ref{indec WSTS with dcs},
-% and
-% Proposition~\ref{srp down up},
-namely, that
-{\sc Resilience} %and {\sc State-resilience} 
-when $\Safe = \mathop{\downarrow}~\Safe$
-%are
-is
-%
- undecidable for Minsky machines
-and 
-for reset-VASS.
-
-The 
-% problems are
-problem is undecidable for a Minsky machine with at least $2$ counters when
-$\Safe$ is downward-closed, since it is possible to consider $\Safe = \{q_f(0,0)\}$. We then use Reset-VASS to simulate the Minsky machines~\cite{araki1976PN,dufourd1998reset}, albeit in a weak way. 
-% The simulation however is 
-Since by a particular choice of $\Safe$ and $\Bad$ we are only interested in the reachability of $q_f(0,0)$ in the Minsky machine, 
-this weak simulation will be enough to obtain undecidability of both problems.
-
-\begin{proposition}\label{both minsk indec}
-{\sc Resilience} and {\sc State-resilience} when $\Safe = \mathop{\downarrow}~\Safe$ are undecidable for reset-VASS, hence also undecidable for $\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
-\end{proposition}
-
-\begin{proof}
-First, note that deciding whether from all inputs, all runs are terminating
- is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well. Hence 
-  deciding whether from all inputs, there exists at least one terminating run
- is undecidable for Minsky machines with at least $2$ counters, since,
- for deterministic Minsky machines, the two problems are one and the same.
-
-Now, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $2$ counters.
-%	This is due to the undecidability of $2$-counter Minsky machine structural termination.
-From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $2$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
-such that there exists a terminating run in $M$ for all inputs iff machine $M'$ can reach $q_f'(0,0)$ from any input. 
-%			with at least $1$ on its third counter. 
-We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the two counters until they reach $0$, then, and only then, go in $q_f'$.
-Based on this construction, the problem of deciding whether the downward-closed set 
-$\{q_f'(0,0)\}$ is reachable in the $2$-counter Minsky machine $M'$
-from any input is equivalent to the asking whether there exists a terminating run from every input of machine $M$, that is undecidable. 
-Note that by chosing $s_0 \in Q_{M'} \times \N^2 \setminus \{q_f'(0,0) \}$,
-deciding whether the downward-closed set 
-$ \{q_f'(0,0) \}$ is reachable
-from every element of $\post^*(s_0)$ 
-is again equivalent to asking whether $M$ terminate on every input.
-Hence {\sc Resilience} and {\sc State-resilience} are undecidable for Minsky machines.
-
-Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
-% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
-%Reset-VASS extend the basic VASS model with special “reset
-%transitions” that set to $0$ some coordinates in the vector. 
-%			Let us recall their definition here.
-%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
-The simulation operates by building a reset-VASS 
-which mimick the behaviors of the Minsky machines' counters by using two mirror counters 
-for each of them,
-and where instead of being tested for zero before a transition, one of the the counters in the pair is being reset; if the counters were zero before the transition then this was a faithful simulation,
-otherwise, there is a gap between the two counters which cannot be undone otherwise than by letting the machine backtrack.
-The construction is such that
-there exists
- a sequence from 
-an initial state %$(s,x_1 \ldots x_n)$
-	with equal mirror counters
-to
-a state $ q({\bf 0})$ if and only if
-there exists
-a sequence
- in the Minsky machine
- from a corresponding initial state
- to
- the state $q_f(0,0,0)$
-  (from Theorem~$5$ in \cite{araki1976PN}).
-  [... from initial states with unequal mirror counters ... -> go back to a state with equal mirror counters -> then ...]
-  [ initial states with unequal mirror counters should be allowed to have some kind of backtracking to them]
-Hence {\sc Resilience} (resp. {\sc State resilience}) holds in the Minsky machine $M'$ iff 
-{\sc Resilience} (resp. {\sc State resilience}) holds in the reset-VASS simulating $M'$.
-Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} 
-and  {\sc State-resilience} are undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
-$\omega^2$-WSTS with strong compatibility as well.\qed
-\end{proof}
-
-	\fi
-
 \section{Decision procedure for VASS when $\Safe = \mathop{\downarrow} \Safe$}\label{appendix vass}
-
-% \mathieu{mettre ici le détail du Theo 13}
 
@@
 is stationnary. 
-%\mathieu{
 We check the finiteness of the set by comparing the elements from the finite basis of $\Bad$ and

### 2023-12-11 diapo/presentation.tex
up

\documentclass{beamer}
%  \usepackage[utf8]{inputenc}
  \usetheme{Warsaw}
  \usepackage{dsfont}
  \usepackage{xcolor}
%  \usepackage[dvipsnames]{xcolor}
  \usepackage{tabto}

\title[Resilience and Home-Space for WSTS]{Resilience and Home-Space for WSTS}

\author[Alain Finkel ~ Mathieu Hilaire]{Alain Finkel \inst{1} \and Mathieu Hilaire \inst{2}}
\institute[Alain Finkel ~ Mathieu Hilaire]{\inst{1} Université Paris-Saclay, CNRS, ENS Paris-Saclay, LMF, Gif-sur-Yvette, France, Institut Universitaire de France \and %
                      \inst{2} Université Lyon 1, LIRIS, France}

\date{}

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[latin1]{inputenc}

\bibliographystyle{alpha}% the recommnded bibstyle

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}

\colorlet{Mycolor1}{green!10!orange!90!}
\colorlet{Mycolor3}{blue!20!magenta!90!}
\definecolor{Mycolor2}{HTML}{00F9DE}

\begin{document}

\maketitle

%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}{Outline}

\tableofcontents
  \end{frame}
  \section{Introduction}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}

TODO

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
  \begin{frame}

TODO

  \end{frame}
%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv

\end{document}

### 2023-12-18 diapo/presentation.tex
up

@@
 %  \usepackage[utf8]{inputenc}
+\usepackage[utf8]{inputenc}
   \usetheme{Warsaw}
@@
 \usetikzlibrary{arrows,automata}
-\usepackage[latin1]{inputenc}
+%\usepackage[latin1]{inputenc}
 
@@
 \definecolor{Mycolor2}{HTML}{00F9DE}
+
+\newcommand{\Bad}{\textsf{Bad}}
+\newcommand{\Safe}{\textsf{Safe}}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}
+  \begin{frame}{Motivation}
 
@@
 
+Exemple de motivation
+
+Fridge \quad	Bad state is when temperature is too high \\
+\phantom{Fridge}	\quad Good state is temperature low enough \\
+
+Speaking: We are usually focused on how to avoid bad states in general but you can see when your fridge the system stops providing refrigerating temperature it's not over yet
+If you leave your fridge off for a minute it's not going to have Terrible impact but if you leave it off for more than one hour you're gonna have to throw the things in it away
+
+And I've been talking like your frige at home but it apply to any type of cooling system it can be like refrigerating medical supplies or nuclear reactor.
+
+And this is interesting Bcz now the stuff is not necessarily avoid bad states completely
+
+This lead to concept of resilience
+
   \end{frame}
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}
+  \begin{frame}{Resilience(s)}
 
 TODO
+
+Define the Six different problems we study/focus on.
+
+(Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{WSTS}
+
+TODO
+
+Resilience(s) "obviously"  undecidable in general so restriction
+
+restriction already to WSTS
+
+so we need to define WSTS properly
+
+This is what this slide could be about
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{... on effectivity}
+
+TODO
+
+... and effectivity as well
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Results}
+
+Now that everything is somewhat properly defined maybe the picture with all the results ?
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result A}
+
+   Now we won't have time to detail all the result. 
+   Result blabla are consequences of the paper    by Okzan so let's not delve into it.\\
+   ( should still be quite precise about Okzan result)
+
+    Let's delve into [This Result] in particular.
+
+   Now recall the definitions ...
+
+   \pause
+
+   And let's give an intuition behind the proof
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Conclusion}
+
+How to expand current results ?
+
+- detailed complexity analysis
+
+- controller/environment framework
+
+- other classes of $\Safe / \Bad$
 

### 2023-12-22 diapo/presentation.tex
up

@@
   \usepackage{tabto}
+
+\usepackage{mathtools}
+\usepackage{todonotes}
+\usepackage{microtype}
+
+\usepackage{complexity}
+\usepackage{amsmath}
+
+\usepackage{stmaryrd}
+\usepackage{dsfont}
+
+\usepackage{mathrsfs}
+\usepackage{mathalpha}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{multirow}
+
+\usepackage{ textcomp }
+
+\usepackage{stmaryrd}
+\usepackage{wrapfig}
 
@@
 
+\newcommand{\pred}{\textsf{pred}}
+\newcommand{\post}{\textsf{post}}
+
 \newcommand{\Bad}{\textsf{Bad}}
@@
 
-TODO
+% TODO
 
-Exemple de motivation
+% Exemple de motivation
 
-Fridge \quad	Bad state is when temperature is too high \\
-\phantom{Fridge}	\quad Good state is temperature low enough \\
+Fridge.png \quad	{\sf Bad} state is when temperature is too high \\
+\phantom{Fridge.png}	\quad {\sf Good} state is temperature low enough \\
 
-Speaking: We are usually focused on how to avoid bad states in general but you can see when your fridge the system stops providing refrigerating temperature it's not over yet
+\todo[inline,color=blue!20]{ {\it À l'oral~} 
+\tiny We are usually focused on how to avoid bad states in general but you can see when your fridge the system stops providing refrigerating temperature it's not over yet
 If you leave your fridge off for a minute it's not going to have Terrible impact but if you leave it off for more than one hour you're gonna have to throw the things in it away
@@
 This lead to concept of resilience
+}
 
@@
 
-TODO
+% TODO
 
-Define the Six different problems we study/focus on.
+% Define the Six different problems we study/focus on.
 
-(Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
+% (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
+
+ {\small
+
+{\sc Resilience problems}
+
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$.}
+
+\hspace{-0.5cm}  {\bf QUESTION:\ }{ ({\sc resilience problem (RP)}) $\Bad \rightarrow^{*} \Safe$ ?
+
+\hspace{-0.5cm}  ({\sc $k$-resilience problem (kRP)})		$\Bad \rightarrow^{\leq k} \Safe$ ?
+
+\hspace{-0.5cm}  ({\sc bounded resilience problem (BRP)})	$\exists k \in \N^+ ~ S \rightarrow^{\leq k} \Safe$ ?\newline}
+
+  \hspace{-0.2cm}
+{\sc State resilience problems}
+
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, and a set $\Safe \subseteq S$.}
+
+\hspace{-0.5cm}  {\bf QUESTION:\ }
+
+\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\post^*(s) \rightarrow^{*} \Safe$ ?
+
+\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\post^*(s) \rightarrow^{\leq k} \Safe$ ?
+
+\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \post^*(s) \rightarrow^{\leq k} \Safe$?\newline}
+
+ }
+
+\todo[inline,color=blue!20]{{\bf 
+\tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
+}}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+	\section{Definitions}
   \begin{frame}{WSTS}
 
-TODO
+% TODO
 
+  \todo[inline,color=blue!20]{ {\it À l'oral~} 
 Resilience(s) "obviously"  undecidable in general so restriction
+$\to$
+restriction already to WSTS
+$\to$
+so we need to define WSTS properly
+}
 
-restriction already to WSTS
+\begin{definition}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
+A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
+is a transition system $(S, \rightarrow)$
+equipped with a wqo ${\leq} \subseteq S \times S$ such that   
+the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+	$t_2 \in S$ with 
+	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
+\end{definition}
 
-so we need to define WSTS properly
-
-This is what this slide could be about
+  \todo[inline,color=blue!20]{ 
+  \tiny Definition right now is very formal for a slideshow, maybe make it less verbose? 
+}
 
@@
 
-Now that everything is somewhat properly defined maybe the picture with all the results ?
+% Now that everything is somewhat properly defined maybe the picture with all the results ?
+
+   \begin{center}
+ 	\begin{figure}
+ 	\hspace{-2.cm}
+\includegraphics[width=1.00\textwidth]{All_results}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
 

### 2023-12-22 diapo/presentation.tex
up

@@
   \todo[inline,color=blue!20]{ {\it À l'oral~}
-\tiny We are usually focused on how to avoid bad states in general but you can see when your fridge the system stops providing refrigerating temperature it's not over yet
+We are usually focused on how to avoid bad states in general but you can see when your fridge the system stops providing refrigerating temperature it's not over yet
 If you leave your fridge off for a minute it's not going to have Terrible impact but if you leave it off for more than one hour you're gonna have to throw the things in it away
@@
 
+  \todo[inline,color=blue!20]{ {\it À l'oral~} 
    Now we won't have time to detail all the result. 
-   Result blabla are consequences of the paper    by Okzan so let's not delve into it.\\
-   ( should still be quite precise about Okzan result)
+   Result blabla are consequences of the paper    by Okzan so let's not delve into it.
+   }
 
+    \todo[inline,color=blue!20]{  ( should still be quite precise about Okzan result and references)}
+
+Result A
+
+  \todo[inline,color=blue!20]{ {\it À l'oral~}   
     Let's delve into [This Result] in particular.
 
-   Now recall the definitions ...
+   And let's give an intuition behind the proof
 
-   \pause
-
-   And let's give an intuition behind the proof
+	...}
 

### 2023-12-31 diapo/presentation.tex
up

@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Example}
+
+   \begin{center}
+ 	\begin{figure}
+ 	\vspace{.06cm}
+\includegraphics[width=.50\textwidth]{FigA}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Example}
+
+   \begin{center}
+ 	\begin{figure}
+% 	\hspace{-2.cm}
+\includegraphics[width=.50\textwidth]{FigB}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{Resilience(s)}
@@
 
-\hspace{-0.5cm}  ({\sc bounded resilience problem (BRP)})	$\exists k \in \N^+ ~ S \rightarrow^{\leq k} \Safe$ ?\newline}
+\hspace{-0.5cm}  ({\sc bounded resilience problem (BRP)})	$\exists k ~ S \rightarrow^{\leq k} \Safe$ ?\newline}
+
+}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience(s)}
+
+ {\small
 
   \hspace{-0.2cm}
-{\sc State resilience problems}
+{\sc \textcolor{red}{State} resilience problems}
 
-\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, $s \in S$, and a set $\Safe \subseteq S$.}
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, \textcolor{red}{$s \in S$}, and a set $\Safe \subseteq S$.}
 
@@
 
-\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\post^*(s) \rightarrow^{*} \Safe$ ?
+\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
 
-\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\post^*(s) \rightarrow^{\leq k} \Safe$ ?
+\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
 
-\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \post^*(s) \rightarrow^{\leq k} \Safe$?\newline}
+\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
 
@@
 
-\todo[inline,color=blue!20]{{\bf 
-\tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
-}}
+% \todo[inline,color=blue!20]{{\bf  \tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)}}
 
@@
 
-  \todo[inline,color=blue!20]{ 
-  \tiny Definition right now is very formal for a slideshow, maybe make it less verbose? 
+\todo[inline,color=blue!20]{    \tiny
+diapo 5: théorème: c'est indécidable dans le cas général des systèmes de transitions
+diapo 5: déjà pour les bien structurés c'est indécidable
+résilience indécidable pour les post effectif WSTS quelque soit safe clos par le haut ou par le bas
 }
+
+%  \todo[inline,color=blue!20]{    \tiny Definition right now is very formal for a slideshow, maybe make it less verbose? }
 
@@
 ... and effectivity as well
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}
+
+   \begin{center}
+ 	\begin{figure}
+% 	\hspace{-2.cm}
+\includegraphics[width=.50\textwidth]{WSTS_taxonomy_large}
+	\end{figure}
+\end{center}
 
@@
 
-Result A
+Result A: Théorème 4 ou 5
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result A}
+
+   \begin{center}
+ 	\begin{figure}
+ 	\hspace{-2.cm}
+\includegraphics[width=1.00\textwidth]{resultA}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result A}
 
@@
 	...}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result B}
+
+ résultat sur la deuxième figure ?
+ on signale juste le thm 8 en passant
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result C}
+
+On On parle des VASS quand Safe downward closed
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result C}
+
+   \begin{center}
+ 	\begin{figure}
+ 	\hspace{-2.cm}
+\includegraphics[width=1.00\textwidth]{resultC}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
 

### 2024-01-02 diapo/presentation.tex
up

@@
 
-% TODO
+\begin{columns}[T]
+ \column{0.35\textwidth}
+ \begin{figure}
+ 	\hspace{-0.5cm}
+\includegraphics[width=1.\textwidth]{fridge}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+\end{figure}
 
-% Exemple de motivation
+ \column{0.65\textwidth}
+\phantom{Fridge.png} \\
+{\sf Bad} state is when temperature is too high \\
+ {\sf Good} state is temperature low enough \\
 
-Fridge.png \quad	{\sf Bad} state is when temperature is too high \\
-\phantom{Fridge.png}	\quad {\sf Good} state is temperature low enough \\
+\end{columns}
 
+\iffalse
   \todo[inline,color=blue!20]{ {\it À l'oral~}
@@
 }
+\fi
 
@@
    Now we won't have time to detail all the result. 
-   Result blabla are consequences of the paper    by Okzan so let's not delve into it.
 }
 
-    \todo[inline,color=blue!20]{  ( should still be quite precise about Okzan result and references)}
+Result A: Théorème 4 ou 5
 
-Result A: Théorème 4 ou 5
+\setcounter{theorem}{3}
+
+\begin{theorem}\label{down-up}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
+{\sc Resilience}, {\sc Bounded resilience} 
+and {\sc $k$-resilience} are decidable.
+\end{theorem}
+
+ou
+
+\begin{theorem}\label{xxx}
+Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
+ {\sc Resilience} is decidable.
+\end{theorem}
 
@@
  résultat sur la deuxième figure ?
- on signale juste le thm 8 en passant
+ on signale juste le thm 8 en passant ?
+
+     \todo[inline,color=blue!20]{    Result blabla are consequences of the paper    by Okzan so let's not delve into it.
+    ( should still be quite precise about Okzan result and references)}
 

### 2024-01-02 diapo/presentation.tex
up

@@
 \end{columns}
+
+\pause
+
+\phantom{Fridge.png} \\
+Other applications e.g. refrigerating medical supplies or nuclear reactor
 
@@
 $\to$
-restriction already to WSTS
+restriction to WSTS
 $\to$
@@
 
+\begin{theorem}
+{\sc Resilience} undecidable for WSTS
+\end{theorem}
+
 \todo[inline,color=blue!20]{    \tiny
@@
 
-TODO
-
-... and effectivity as well
+\begin{enumerate}\label{on effectivity}
+%\item {\em effective} if there exists a pair of algorithms
+%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
+%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
+%$M_\leq$ the ordering relation “$\leq$”.
+%
+\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
+exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
+on input $s \in S$.
+%
+%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
+%to the encoding of the ideal $\downarrow s$ is computable.
+%
+\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
+%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
+%
+\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
+%
+%and there exist an algorithm $M_\downarrow$ that computes 
+% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
+% $s_1, s_2, \ldots, s_m \in S$.
+%
+\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
+%
+\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
+\end{enumerate}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}
+  \begin{frame}{WBTS Taxonomy}
 
@@
 % 	\hspace{-2.cm}
-\includegraphics[width=.50\textwidth]{WSTS_taxonomy_large}
+\includegraphics[width=.80\textwidth]{WSTS_taxonomy_large}
 	\end{figure}
@@
 
+  \todo[inline,color=blue!20]{ 
+Refaire les références dans l'image (pas les même défintions/ref) 
+}
+
   \end{frame}
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+	\section{Contributions}
   \begin{frame}{Results}

### 2024-01-02 diapo/presentation.tex
up

@@
 
+\hspace{-2cm}
 \begin{enumerate}\label{on effectivity}
-%\item {\em effective} if there exists a pair of algorithms
-%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
-%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
-%$M_\leq$ the ordering relation “$\leq$”.
-%
-\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
-exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
-on input $s \in S$.
+\item {\em $\post$-effective}  = effective +
+  $ s \in S \to |\post(s)| \in \N \cup \{\infty\}$ computable
 %
@@
 %
-\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
+\item {\em $\post$-ideal-effective} = effective +
+ $I \in Ideals(S) \to \mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
 %\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
 %
-\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
+\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} = effective + there exists an algorithm accepting
 any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.

### 2024-01-03 diapo/presentation.tex
up

@@
 % TODO
-
-  \todo[inline,color=blue!20]{ {\it À l'oral~} 
-Resilience(s) "obviously"  undecidable in general so restriction
-$\to$
-restriction to WSTS
-$\to$
-so we need to define WSTS properly
-}
+%  \todo[inline,color=blue!20]{ {\it À l'oral~} 
+% Resilience(s) "obviously"  undecidable in general so restriction $\to$ restriction to WSTS
+% $\to$ so we need to define WSTS properly }
 
@@
 
+\pause
+
 \begin{theorem}
@@
 
-\todo[inline,color=blue!20]{    \tiny
-diapo 5: théorème: c'est indécidable dans le cas général des systèmes de transitions
-diapo 5: déjà pour les bien structurés c'est indécidable
-résilience indécidable pour les post effectif WSTS quelque soit safe clos par le haut ou par le bas
-}
+% \todo[inline,color=blue!20]{    \tiny
+% diapo 5: théorème: c'est indécidable dans le cas général des systèmes de transitions
+% diapo 5: déjà pour les bien structurés c'est indécidable
+% résilience indécidable pour les post effectif WSTS quelque soit safe clos par le haut ou par le bas
+% }
 
@@
  	\begin{figure}
- 	\hspace{-2.cm}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
 \includegraphics[width=1.00\textwidth]{All_results}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Results}
+
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{resultA}
 % \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
@@
 
-  \todo[inline,color=blue!20]{ {\it À l'oral~} 
-   Now we won't have time to detail all the result. 
-}
-
-Result A: Théorème 4 ou 5
+% Result A: Théorème 4
 
@@
 \end{theorem}
-
-ou
-
-\begin{theorem}\label{xxx}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
- {\sc Resilience} is decidable.
-\end{theorem}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result A}
-
-   \begin{center}
- 	\begin{figure}
- 	\hspace{-2.cm}
-\includegraphics[width=1.00\textwidth]{resultA}
-% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
-Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
-	\end{figure}
-\end{center}
 
@@
  	\begin{figure}
- 	\hspace{-2.cm}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
 \includegraphics[width=1.00\textwidth]{resultC}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{References}
+
+      {\tiny
+\bibliography{bibliography}
+}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
 

### 2024-01-03 diapo/presentation.tex
up

@@
 \phantom{Fridge.png} \\
-{\sf Bad} state is when temperature is too high \\
- {\sf Good} state is temperature low enough \\
+\phantom{Fridge.png} \\
+{\sf Bad} state = temperature is too high \newline
+
+ {\sf Good} state = temperature low enough \\
 
@@
 
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+\includegraphics[width=.25\textwidth]{WSTS_def}
+	\end{figure}
+\end{center}
+
 \pause
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{... on effectivity}
+  \begin{frame}{WSTS with strong compatibility}
+
+% TODO
+%  \todo[inline,color=blue!20]{ {\it À l'oral~} 
+% Resilience(s) "obviously"  undecidable in general so restriction $\to$ restriction to WSTS
+% $\to$ so we need to define WSTS properly }
+
+\begin{definition}
+A  Well-Structured Transition System {\em with \textcolor{red}{strong} compatibility}  $\mathscr{S}=(S, \rightarrow, \leq)$
+is a transition system $(S, \rightarrow)$
+equipped with a wqo ${\leq} \subseteq S \times S$ such that   
+the transition relation $ \rightarrow$ is \textcolor{red}{strongly} (upward) compatible with $\leq$, i.e., for all 
+$s_1, t_1 , s_2 \in S$
+	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+	$t_2 \in S$ with 
+	$t_1 \leq t_2$ and $s_2 \textcolor{red}{\rightarrow} t_2$.
+\end{definition}
+
+\begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+\includegraphics[width=.25\textwidth]{WSTS_strong}
+	\end{figure}
+\end{center}
+
+\phantom{fridge}\\
+\phantom{ undecidable for WSTS} \\
+\phantom{ undecidable for WSTS} \\
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{On effectivity}
+
+Ideal =
+\begin{itemize}
+ \item downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+ \item directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
+\end{itemize}
+
+  \todo[inline,color=blue!20]{ 
+Définir Ideals and ideals.
+}
+
+\hspace{-2cm}
+
+ effective = 
+ \begin{itemize}
+\item s $\to$ t decidable 
+\item membership decidable 
+\item $\leq$ decidable 
+\item membership in $Ideals(S)$ decidable 
+\item inclusion of ideals decidable
+\end{itemize}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{On effectivity}
 
@@
 
+Définir rapidement les VASS
+
+ \begin{definition} 
+A {\em vector addition system with (control-)states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
+ \end{definition}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Result C}
+
    \begin{center}
@@
       {\tiny
-\bibliography{bibliography}
+\bibliography{bib}
 }

### 2024-01-05 diapo/presentation.tex
up

@@
 
+% Define WBSTS also ?
+
   \end{frame}
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{On effectivity}
+  \begin{frame}{Definitions}
 
@@
 \begin{itemize}
- \item downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
- \item directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
+ \item Downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+ ($y \leq x  \wedge x \in D \to y \in D$) 
+ \item Directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
+ ($D \neq \emptyset \wedge \forall a,b \in D, \exists c \in D, a \leq c \vee b \leq c$ )
 \end{itemize}
 
-  \todo[inline,color=blue!20]{ 
-Définir Ideals and ideals.
-}
+\pause
 
@@
 
- effective = 
+$S$ effective = 
  \begin{itemize}
-\item s $\to$ t decidable 
+\item $s \to t$ decidable 
 \item membership decidable 
@@
 
+	$\to$ link between runs and runs of the completion (must define completion)
+
+	$\to$ ideal decomposition of Bad
+
+	$\to$ basis of Safe
+
+	$\to$ coverability questions
+
+	$\to$  resilience(s) can be reduced to coverability
+
   \end{frame}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result C}
+  \begin{frame}{Resilience for VASS}
 
-On On parle des VASS quand Safe downward closed
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result C}
-
-Définir rapidement les VASS
+% Définir rapidement les VASS
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result C}
+  \begin{frame}{Resilience for VASS}
 

### 2024-01-05 diapo/presentation.tex
up

@@
 
-Ideal =
+$I$ Ideal =
 \begin{itemize}
  \item Downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
- ($y \leq x  \wedge x \in D \to y \in D$) 
+ ($y \leq x  \wedge x \in I \to y \in I$) 
  \item Directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
- ($D \neq \emptyset \wedge \forall a,b \in D, \exists c \in D, a \leq c \vee b \leq c$ )
+ ($I \neq \emptyset \wedge \forall a,b \in I, \exists c \in I, a \leq c \vee b \leq c$ )
 \end{itemize}
@@
 \begin{enumerate}\label{on effectivity}
-\item {\em $\post$-effective}  = effective +
+\item {\em $\post$-effective}  =
+
+effective +
   $ s \in S \to |\post(s)| \in \N \cup \{\infty\}$ computable
@@
 %
-\item {\em $\post$-ideal-effective} = effective +
- $I \in Ideals(S) \to \mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
+\item {\em $\post$-ideal-effective} =
+
+effective +
+ $I \in Ideals(S) \to \mathop{\downarrow} \post(I)$ % expressed as a finite union of (maximal) ideals.
+ computable
+
 %\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
 %
-\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} = effective + there exists an algorithm accepting
-any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
+\item {\em $\pred$-basis-effective}  =
+
+effective +  $s \in S \to$ a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$ computable
+
+\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
 %
@@
 %
-\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
+\item {\em $\post^*$-basis-effective} =
+
+effective +  $s \in S \to$ a finite basis of $\mathop{\uparrow} \post^*(s)$ computable
 %
-\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
-any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
+
+\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
+\item {\em $\pred^*$-basis-effective} =
+
+effective +  $I \in Ideals(S) \to$ a finite basis of $\mathop{\downarrow} \pred^*(I)$ computable
 \end{enumerate}

### 2024-01-05 diapo/presentation.tex
up

@@
 
-\setcounter{theorem}{3}
-
-\begin{theorem}\label{down-up}
+\begin{theorem}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result A}
+  \begin{frame}{General Idea}
 
@@
 	$\to$  resilience(s) can be reduced to coverability
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Ingredient 1: the completion}
+
+ \begin{definition}[\cite{BFM-ic17}]
+The \emph{completion}   of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
+%$Ideals(S)$ is the set of ideals of $S$ and
+ $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\mathop{\downarrow} \post_{\mathscr{S}}(I)$. 
+% The completion is always \emph{finitely branching} but it is not necessarly a WSTS since $\subseteq$ is not necessarly a wqo. 
+%It is proved in  \cite{BFM-ic17} that $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS. \\
+\end{definition}
+
+\begin{exampleblock}{\cite[Proposition 30]{BFM-ic17}}
+If $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
+\end{exampleblock}
+
+\begin{exampleblock}{\cite[Proposition 29]{BFM-ic17}}
+If $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
+Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+\end{exampleblock}
 

### 2024-01-05 diapo/presentation.tex
up

@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Ingredient 2: link with coverability}
+
+  \begin{block}{ }
+   $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}} $ 
+ \end{block}
+
+\pause
+
+  \begin{block}{ }
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
+Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+ \end{block}
+
+    \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Ingredient 3 (only for $k$-resilience)}
+
+     \todo[inline,color=blue!20]{ 
+  ... actually it is generally over now maybe we don't have to go into all the details here
+}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{Result B}

### 2024-01-05 diapo/presentation.tex
up

@@
 
+ \begin{theorem}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}
+{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
+when
+$\Safe=\mathop{\uparrow} \Safe$.
+\end{theorem}
+
+\pause
+
+% Keeping the $\mathop{\uparrow} \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
+Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
+
+\begin{theorem}\label{post srp}
+{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
+when
+$\Safe=\mathop{\uparrow} \Safe$. 
+\end{theorem}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience for VASS}
+
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{resultC}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
   \end{frame}
@@
 
-   \begin{center}
- 	\begin{figure}
- 	% \hspace{-2.cm}
- 	\vspace{-0.25cm}
-\includegraphics[width=1.00\textwidth]{resultC}
-% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
-Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
-	\end{figure}
-\end{center}
+- Resilience $\to$ inclusion d'idéaux
+
+- Bounded resilience and $k$-resilience $\to$ more complex
+
+- Remaque sur les VAS : it does not hold
+
+- Remarque additionnelle du coup
+
+-
 

### 2024-01-08 diapo/presentation.tex
up

@@
 
-% Define WBSTS also ?
+ \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{WBTS}
+
+ Define WBSTS also ?
+And $\omega^2$-WSTS ?
 
@@
 
-  \todo[inline,color=blue!20]{ 
-Refaire les références dans l'image (pas les même défintions/ref) 
-}
-
   \end{frame}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result A}
+  \begin{frame}{Resilience when $\Safe = {\uparrow \Safe}$}
 
@@
 
+% \begin{theorem}
+% Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
+% {\sc Resilience}, {\sc Bounded resilience} 
+% and {\sc $k$-resilience} are decidable.
+% \end{theorem}
+
 \begin{theorem}
-Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
+WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ 
+\begin{itemize}
+\item $\post$-ideal-effective 
+\item strong compatibility
+\item $\Safe = \mathop{\uparrow} \Safe$
+\end{itemize}
 {\sc Resilience}, {\sc Bounded resilience} 
-and {\sc $k$-resilience} are decidable.
+and {\sc $k$-resilience} decidable.
 \end{theorem}
@@
   \begin{frame}{General Idea}
-
-  \todo[inline,color=blue!20]{ {\it À l'oral~}   
-    Let's delve into [This Result] in particular.
-
-   And let's give an intuition behind the proof
-
-	...}
 
@@
 
+\pause
+
 \begin{exampleblock}{\cite[Proposition 30]{BFM-ic17}}
@@
 \end{exampleblock}
+
+\pause
 
@@
 
-  \begin{block}{ }
+  \begin{block}{Proposition}
    $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}} $ 
@@
 The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
-Let us show how this infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+ \end{block}
+
+ \pause
+
+  \begin{block}{ }
+This infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
  \end{block}

### 2024-01-09 diapo/presentation.tex
up

@@
 
-	$\to$ link between runs and runs of the completion (must define completion)
+% 	$\to$ ideal decomposition of Bad
 
-	$\to$ ideal decomposition of Bad
+%	$\to$ basis of Safe
 
-	$\to$ basis of Safe
+ 	$\to$ resilience =$>$ infinite number of coverability questions
 
-	$\to$ coverability questions
+	$\to$ introduce the completion
 
-	$\to$  resilience(s) can be reduced to coverability
+	$\to$  here resilience =$>$ finite number of coverability questions
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Ingredient 3 (only for $k$-resilience)}
+  \begin{frame}{Ingredient 3}
 
+    \begin{block}{ }
+This infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+ \end{block}
+
+ $\to$ consider max length of such coverability questions
+
+  $\to$ answers {\sc Bounded resilience} and {\sc $k$-resilience} as well
      \todo[inline,color=blue!20]{ 
@@
   %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Result B}
+    \begin{frame}{Results}
 
- résultat sur la deuxième figure ?
- on signale juste le thm 8 en passant ?
+% Now that everything is somewhat properly defined maybe the picture with all the results ?
 
-     \todo[inline,color=blue!20]{    Result blabla are consequences of the paper    by Okzan so let's not delve into it.
-    ( should still be quite precise about Okzan result and references)}
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{All_results}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{State Resilience when $\Safe = {\uparrow \Safe}$}
+
+ % résultat sur la deuxième figure ?
+ % on signale juste le thm 8 en passant ?
+
+ %    \todo[inline,color=blue!20]{    Result blabla are consequences of the paper    by Okzan so let's not delve into it.
+ %   ( should still be quite precise about Okzan result and references)}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience for VASS}
+  \begin{frame}{Resilience $\ldots$}
+
+% Now that everything is somewhat properly defined maybe the picture with all the results ?
+
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{All_results}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience $\ldots$ for VASS}
 

### 2024-01-09 diapo/presentation.tex
up

@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience for VASS}
+  \begin{frame}{Vector Addition System with States}
 
@@
 
+ \begin{definition} 
+A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
+ \end{definition}
+
   \end{frame}
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience for VASS}
+  \begin{frame}{Resilience(s) for VASS}
+
+\begin{theorem}\label{SL VASS}
+{\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
+\end{theorem}
 
@@
 
+   \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience(s) for VASS}
+
+- Bounded resilience and $k$-resilience $\to$ more complex
+
 - Remaque sur les VAS : it does not hold
 
-- Remarque additionnelle du coup
+\begin{block}{Proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
+\end{block}
 
--
+   \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience(s) for VASS}
+
+\begin{theorem}\label{vass down}
+{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+$\Safe = \mathop{\downarrow} \Safe$.
+\end{theorem}
 

### 2024-01-09 diapo/presentation.tex
diapos

@@
 
-\begin{definition}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
+\begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
 A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
-is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ such that   
-the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
-$s_1, t_1 , s_2 \in S$
-	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
-	$t_2 \in S$ with 
-	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
+is a well-ordered transition system $(S, \rightarrow)$ such that   
+%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
+%$s_1, t_1 , s_2 \in S$
+%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+%	$t_2 \in S$ with 
+%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
 \end{definition}

### 2024-01-09 diapo/presentation.tex

@@
 \begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
-A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
-is a well-ordered transition system $(S, \rightarrow)$ such that   
+A {\em Well-Structured Transition System} (WSTS) 
+is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
 %the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 

### 2024-01-09 diapo/presentation.tex
up

@@
 \date{}
+
+\setbeamertemplate{page number in head/foot}[totalframenumber]
 
@@
 \phantom{Fridge.png} \\
-Other applications e.g. refrigerating medical supplies or nuclear reactor
+Other applications e.g.  nuclear reactor
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Example}
+  \begin{frame}{Example: Reset-VASS $V_1$}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Example}
+  \begin{frame}{Example: Reset-VASS $V_2$}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-	\section{Definitions}
-  \begin{frame}{WSTS}
+	\section{WBTS, WSTS}
+  \begin{frame}{}
 
+  \begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
+A {\em Well-Behaved Transition System} (WBTS) 
+is a FAC\footnote{FAC = all anti-chains are finite}-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
+%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
+%$s_1, t_1 , s_2 \in S$
+%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+%	$t_2 \in S$ with 
+%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
+\end{definition}
+
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+\includegraphics[width=.25\textwidth]{WSTS_def}
+	\end{figure}
+\end{center}
+
+    \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}
 % TODO
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{WSTS with strong compatibility}
-
-% TODO
-%  \todo[inline,color=blue!20]{ {\it À l'oral~} 
-% Resilience(s) "obviously"  undecidable in general so restriction $\to$ restriction to WSTS
-% $\to$ so we need to define WSTS properly }
-
-\begin{definition}
-A  Well-Structured Transition System {\em with \textcolor{red}{strong} compatibility}  $\mathscr{S}=(S, \rightarrow, \leq)$
-is a transition system $(S, \rightarrow)$
-equipped with a wqo ${\leq} \subseteq S \times S$ such that   
-the transition relation $ \rightarrow$ is \textcolor{red}{strongly} (upward) compatible with $\leq$, i.e., for all 
-$s_1, t_1 , s_2 \in S$
-	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
-	$t_2 \in S$ with 
-	$t_1 \leq t_2$ and $s_2 \textcolor{red}{\rightarrow} t_2$.
-\end{definition}
-
-\begin{center}
- 	\begin{figure}
- 	% \hspace{-2.cm}
-\includegraphics[width=.25\textwidth]{WSTS_strong}
-	\end{figure}
-\end{center}
-
-\phantom{fridge}\\
-\phantom{ undecidable for WSTS} \\
-\phantom{ undecidable for WSTS} \\
-
- \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{WBTS}
-
- Define WBSTS also ?
-And $\omega^2$-WSTS ?
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Definitions}
-
-$I$ Ideal =
-\begin{itemize}
- \item Downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
- ($y \leq x  \wedge x \in I \to y \in I$) 
- \item Directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
- ($I \neq \emptyset \wedge \forall a,b \in I, \exists c \in I, a \leq c \vee b \leq c$ )
-\end{itemize}
-
-\pause
-
-\hspace{-2cm}
-
-$S$ effective = 
- \begin{itemize}
-\item $s \to t$ decidable 
-\item membership decidable 
-\item $\leq$ decidable 
-\item membership in $Ideals(S)$ decidable 
-\item inclusion of ideals decidable
-\end{itemize}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{On effectivity}
-
-\hspace{-2cm}
-\begin{enumerate}\label{on effectivity}
-\item {\em $\post$-effective}  =
-
-effective +
-  $ s \in S \to |\post(s)| \in \N \cup \{\infty\}$ computable
-%
-%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
-%to the encoding of the ideal $\downarrow s$ is computable.
-%
-\item {\em $\post$-ideal-effective} =
-
-effective +
- $I \in Ideals(S) \to \mathop{\downarrow} \post(I)$ % expressed as a finite union of (maximal) ideals.
- computable
-
-%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
-%
-\item {\em $\pred$-basis-effective}  =
-
-effective +  $s \in S \to$ a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$ computable
-
-\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
-%
-%and there exist an algorithm $M_\downarrow$ that computes 
-% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
-% $s_1, s_2, \ldots, s_m \in S$.
-%
-\item {\em $\post^*$-basis-effective} =
-
-effective +  $s \in S \to$ a finite basis of $\mathop{\uparrow} \post^*(s)$ computable
-%
-
-\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} 
-\item {\em $\pred^*$-basis-effective} =
-
-effective +  $I \in Ideals(S) \to$ a finite basis of $\mathop{\downarrow} \pred^*(I)$ computable
-\end{enumerate}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{WBTS Taxonomy}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-	\section{Contributions}
+	\section{Resiliences}
   \begin{frame}{Results}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Definitions}
+
+$I$ Ideal =
+\begin{itemize}
+ \item Downward-closed % is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
+ ($y \leq x  \wedge x \in I \to y \in I$) 
+ \item Directed % i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
+ ($I \neq \emptyset \wedge \forall a,b \in I, \exists c \in I, a \leq c \vee b \leq c$ )
+\end{itemize}
+
+\hspace{-2cm}
+
+$S$ effective = 
+ \begin{itemize}
+\item $s \to t$ decidable 
+\item membership decidable 
+\item $\leq$ decidable 
+\item membership in $Ideals(S)$ decidable 
+\item inclusion of ideals decidable
+\end{itemize}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{Resilience when $\Safe = {\uparrow \Safe}$}
@@
 \begin{itemize}
-\item $\post$-ideal-effective 
-\item strong compatibility
+\item $\mathop{\downarrow} \post(I)$ computable %$\post$-ideal-effective 
+\item \textcolor{red}{strong} compatibility
 \item $\Safe = \mathop{\uparrow} \Safe$
@@
 
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{General Idea}
-
-% 	$\to$ ideal decomposition of Bad
-
-%	$\to$ basis of Safe
-
- 	$\to$ resilience =$>$ infinite number of coverability questions
-
-	$\to$ introduce the completion
-
-	$\to$  here resilience =$>$ finite number of coverability questions
+\begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+\includegraphics[width=.25\textwidth]{WSTS_strong}
+	\end{figure}
+\end{center}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Ingredient 1: the completion}
+  \begin{frame}{Ingredient 1: link with coverability}
+
+  \begin{block}{Proposition}
+   $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}} $ 
+ \end{block}
+
+\pause
+
+  \begin{block}{ }
+The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
+ \end{block}
+
+ \pause
+
+  \begin{block}{ }
+This infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
+ \end{block}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Ingredient 2: the completion}
 
@@
   %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Ingredient 2: link with coverability}
-
-  \begin{block}{Proposition}
-   $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}} $ 
- \end{block}
-
-\pause
-
-  \begin{block}{ }
-The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
- \end{block}
-
- \pause
-
-  \begin{block}{ }
-This infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
- \end{block}
-
-    \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Ingredient 3}
-
-    \begin{block}{ }
-This infinite set of coverability questions can be reduced to a \emph{finite} set of coverability questions in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.
- \end{block}
-
- $\to$ consider max length of such coverability questions
-
-  $\to$ answers {\sc Bounded resilience} and {\sc $k$-resilience} as well
-     \todo[inline,color=blue!20]{ 
-  ... actually it is generally over now maybe we don't have to go into all the details here
-}
-
-  \end{frame}
-  %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
     \begin{frame}{Results}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{State Resilience when $\Safe = {\uparrow \Safe}$}
+  \begin{frame}{State resilience}
 
- % résultat sur la deuxième figure ?
- % on signale juste le thm 8 en passant ?
-
- %    \todo[inline,color=blue!20]{    Result blabla are consequences of the paper    by Okzan so let's not delve into it.
- %   ( should still be quite precise about Okzan result and references)}
-
- \begin{theorem}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}
-{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
-when
-$\Safe=\mathop{\uparrow} \Safe$.
-\end{theorem}
-
-\pause
-
-% Keeping the $\mathop{\uparrow} \post^*$ effectiveness hypothesis but \emph{loosening} the strong compatibility one still yields some decidability result for the general {\sc State-resilience}.
-Using the same proof structure as Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889} we obtain:
-
-\begin{theorem}\label{post srp}
-{\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
-when
-$\Safe=\mathop{\uparrow} \Safe$. 
-\end{theorem}
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{resultB}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
 

### 2024-01-11 diapo/presentation.tex
up

@@
 \usepackage{dsfont}
+
+ \usepackage{cancel}
 
@@
 \phantom{Fridge.png} \\
-{\sf Bad} state = temperature is too high \newline
+{\sf Bad} state = temperature $\geq 6^\circ C$ \newline
 
- {\sf Good} state = temperature low enough \\
+ {\sf Good} state = temperature $\leq 6^\circ C$ \\
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Example: Reset-VASS $V_2$}
-
-   \begin{center}
- 	\begin{figure}
-% 	\hspace{-2.cm}
-\includegraphics[width=.50\textwidth]{FigB}
-	\end{figure}
-\end{center}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{Resilience(s)}
@@
 
-\hspace{-0.5cm}  {\bf QUESTION:\ }{ ({\sc resilience problem (RP)}) $\Bad \rightarrow^{*} \Safe$ ?
+\hspace{-0.5cm}  {\bf QUESTION:\ }
+
+\hspace{-0.5cm}  {({\sc resilience problem (RP)}) $\Bad \rightarrow^{*} \Safe$ ?
 
@@
 
-\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
+\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
 
-\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
+\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\cancel{\Bad}~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
 
-\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
+\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
 
@@
 % \todo[inline,color=blue!20]{{\bf  \tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)}}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Example: Reset-VASS $V_1$}
+
+   \begin{center}
+ 	\begin{figure}
+ 	\vspace{.06cm}
+\includegraphics[width=.50\textwidth]{FigA}
+	\end{figure}
+\end{center}
+
+$\Safe = \{q_0(n) \mid n \in \N\}$
+
+{resilience}, 
+{$k$-resilience} and {bounded resilience} are not satisfied
+
+$\Safe = \{q_1(0)\} $
+
+ {resilience}
+  However, {bounded resilience} does not hold
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Example: Reset-VASS $V_2$}
+
+   \begin{center}
+ 	\begin{figure}
+% 	\hspace{-2.cm}
+\includegraphics[width=.50\textwidth]{FigB}
+	\end{figure}
+\end{center}
+
+  $\Safe = \{q_1(0)\} $
+   {bounded resilience} hold, $6$-resilience hold
 
@@
 
-\begin{theorem}
+\begin{block}{Theorem~4}
 WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ 
 \begin{itemize}
-\item $\mathop{\downarrow} \post(I)$ computable %$\post$-ideal-effective 
+\item $\post$-ideal-effective =effective + $\mathop{\downarrow} \post(I)$ computable %$\post$-ideal-effective 
 \item \textcolor{red}{strong} compatibility
@@
 and {\sc $k$-resilience} decidable.
-\end{theorem}
+\end{block}
 
@@
 
-  \begin{block}{Proposition}
-   $s_j$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} s_j$ is coverable (for inclusion) from $\mathop{\downarrow} x$ in $\hat{\mathscr{S}} $ 
+  \begin{block}{Reformulation}
+Resilience $\equiv$ $\forall x \in \Bad$, $\exists y \in \Safe$, $y$ is coverable from $x$ ?
  \end{block}
@@
 
-  \begin{block}{ }
-The resilience problem can be reduced to the following infinite number of instances of the coverability problem in $\mathscr{S}$: for all $x \in \Bad$ does there exist an $j$ such that $s_j$ is coverable from $x$.
+  \begin{block}{Proposition}
+   $y$ is coverable from $x$ in $\mathscr{S}$ if and only if $\mathop{\downarrow} y$ is coverable from $\mathop{\downarrow} x$ in the completion $\hat{\mathscr{S}} $ 
  \end{block}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Vector Addition System with States}
-
-% Définir rapidement les VASS
-
- \begin{definition} 
-A {\em vector addition system with (control-)states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
- \end{definition}
-
- \begin{definition} 
-A {\em vector addition system (VAS)} in dimension $d$ ($d$-VAS for short) is a $d$-VASS where the set of control-states is a singleton.
- \end{definition}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}{Resilience(s) for VASS}
@@
 
-- Resilience $\to$ inclusion d'idéaux
-
-- Bounded resilience and $k$-resilience $\to$ more complex
+$\to$ it is decidable whether $\post^*(\Bad) \subseteq \pred^*(\Safe)$ when 
+$\Bad$ and $\Safe$ are both semilinear~\cite{DBLP:journals/corr/abs-2207-02697}
 
@@
 
-- detailed complexity analysis
+- detailed complexity analysis (in VASS \& variations)
 
@@
 
-- other classes of $\Safe / \Bad$
+- other classes of $\Safe / \Bad$ (e.g. semilinear)
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}
+
+  \begin{center}
+  Thanks for your attention
+  \end{center}
 

### 2024-01-11 diapo/presentation.tex
up

@@
 
-\hspace{-0.5cm}  {({\sc resilience problem (RP)}) $\Bad \rightarrow^{*} \Safe$ ?
+\hspace{-0.5cm}  ({\sc resilience problem}) $\Bad \rightarrow^{*} \Safe$ ?
 
-\hspace{-0.5cm}  ({\sc $k$-resilience problem (kRP)})		$\Bad \rightarrow^{\leq k} \Safe$ ?
+\hspace{-0.5cm}  ({\sc $k$-resilience problem})		$\Bad \rightarrow^{\leq k} \Safe$ ?
 
-\hspace{-0.5cm}  ({\sc bounded resilience problem (BRP)})	$\exists k ~ S \rightarrow^{\leq k} \Safe$ ?\newline}
+\hspace{-0.5cm}  ({\sc bounded resilience problem})	$\exists k ~ S \rightarrow^{\leq k} \Safe$ ?\newline
 
 }
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s)}
-
- {\small
-
-  \hspace{-0.2cm}
-{\sc \textcolor{red}{State} resilience problems}
-
-\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, \textcolor{red}{$s \in S$}, and a set $\Safe \subseteq S$.}
-
-\hspace{-0.5cm}  {\bf QUESTION:\ }
-
-\hspace{-0.5cm}{ ({\sc state-resilience problem (SRP)}) $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
-
-\hspace{-0.5cm}  ({\sc $k$-state-resilience problem (kSRP)})		$\cancel{\Bad}~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
-
-\hspace{-0.5cm} ({\sc bounded-state-resilience problem (BSRP)})	$\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
-
- }
-
-% \todo[inline,color=blue!20]{{\bf  \tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)}}
 
@@
 
- {resilience}
-  However, {bounded resilience} does not hold
+ {resilience} hold, {bounded resilience} does not hold
 
@@
    {bounded resilience} hold, $6$-resilience hold
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience(s)}
+
+ {\small
+
+  \hspace{-0.2cm}
+{\sc \textcolor{red}{State} resilience problems}
+
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, \textcolor{red}{$s \in S$}, and a set $\Safe \subseteq S$.}
+
+\hspace{-0.5cm}  {\bf QUESTION:\ }
+
+\hspace{-0.5cm}{ ({\sc state-resilience problem}) $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
+
+\hspace{-0.5cm}  ({\sc $k$-state-resilience problem})		$\cancel{\Bad}~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
+
+\hspace{-0.5cm} ({\sc bounded-state-resilience problem})	$\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
+
+ }
+
+% \todo[inline,color=blue!20]{{\bf  \tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)}}
 
@@
 
-   \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s) for VASS}
-
-- Bounded resilience and $k$-resilience $\to$ more complex
-
-- Remaque sur les VAS : it does not hold
+\pause
 

### 2024-01-11 diapo/presentation.tex
ip
up

@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Outline}
+%  \begin{frame}{Outline}
 
-\tableofcontents
-  \end{frame}
+%\tableofcontents
+%  \end{frame}
   \section{Introduction}
@@
  	\vspace{.06cm}
-\includegraphics[width=.50\textwidth]{FigA}
+\includegraphics[width=.47\textwidth]{FigA}
 	\end{figure}
 \end{center}
+
+$\Safe = \{q_0(n) \mid n \in \N\}$, $\Bad$ complement of $\Safe$
+
+\phantom{ {resilience},  {$k$-resilience} and {bounded resilience} are not satisfied }
+
+$\Safe = \{q_1(0)\} $, $\Bad$ complement of $\Safe$
+
+\phantom{ {resilience} hold, {bounded resilience} does not hold }
 
@@
  	\vspace{.06cm}
-\includegraphics[width=.50\textwidth]{FigA}
+\includegraphics[width=.47\textwidth]{FigA}
 	\end{figure}
@@
 
-$\Safe = \{q_0(n) \mid n \in \N\}$
+$\Safe = \{q_0(n) \mid n \in \N\}$, $\Bad$ complement of $\Safe$
 
@@
 
-$\Safe = \{q_1(0)\} $
+$\Safe = \{q_1(0)\} $, $\Bad$ complement of $\Safe$
 
@@
 % 	\hspace{-2.cm}
-\includegraphics[width=.50\textwidth]{FigB}
+\includegraphics[width=.47\textwidth]{FigB}
 	\end{figure}
@@
 A {\em Well-Behaved Transition System} (WBTS) 
-is a FAC\footnote{FAC = all anti-chains are finite}-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
+is a FAC$^1$-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
 %the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
@@
 \end{center}
+
+\phantom{line}
+
+{\small $^1${FAC = all anti-chains are finite}}
 

### 2024-01-11 diapo/presentation.tex
up
uo

@@
 
-$\Safe = \{q_0(n) \mid n \in \N\}$, $\Bad$ complement of $\Safe$
+$\Safe_1 = \{q_0(n) \mid n \in \N\}$
 
@@
 
-$\Safe = \{q_1(0)\} $, $\Bad$ complement of $\Safe$
+$\Safe_2 = \{q_1(0)\} $
 
@@
 
-  $\Safe = \{q_1(0)\} $
+  $\Safe_2 = \{q_1(0)\} $
    {bounded resilience} hold, $6$-resilience hold
@@
   \begin{frame}{Resilience(s)}
-
- {\small
 
@@
 
- }
-
 % \todo[inline,color=blue!20]{{\bf  \tiny (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)}}
@@
 
-  \begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
+  \begin{definition}[\cite{DBLP:journals/lmcs/BlondinFM17}]
 A {\em Well-Behaved Transition System} (WBTS) 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Definitions}
+  \begin{frame}{Ideals and effectivity}
 

### 2024-01-11 diapo/presentation.tex
up

@@
 
-\begin{exampleblock}{\cite[Proposition 30]{BFM-ic17}}
+\begin{exampleblock}{Relation with paths in the completion}
 If $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
@@
 
-\begin{exampleblock}{\cite[Proposition 29]{BFM-ic17}}
+\begin{exampleblock}{Relation with paths in the completion}
 If $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
-Moreover, if $\mathscr{S}$ has transitive compatibility then $k’ \geq k$; if $\mathscr{S}$ has strong compatibility then $k’ = k$.
+If $\mathscr{S}$ has strong compatibility then $k’ = k$.
 \end{exampleblock}
@@
 
+  Remettre un déssin pour un VASS
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{Resilience(s) for VASS and VAS}
+
 \begin{theorem}\label{SL VASS}
@@
 
-\begin{block}{Proposition}
-{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
-\end{block}
-
-   \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s) for VASS}
-
 \begin{theorem}\label{vass down}
-{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
+{\sc Bounded resilience} and {\sc $k$-resilience } are decidable for VASS when 
 $\Safe = \mathop{\downarrow} \Safe$.
 \end{theorem}
+
+\pause
+
+\begin{block}{Proposition}
+{\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe \neq \N^d$.
+\end{block}
 
@@
 
-How to expand current results ?
+   \begin{center}
+ 	\begin{figure}
+ 	% \hspace{-2.cm}
+ 	\vspace{-0.25cm}
+\includegraphics[width=1.00\textwidth]{All_results}
+% \caption{Hasse diagram of some classes of transition systems, together with the decidability (in green) or undecidability (in red) of the resilience problems.
+Decidability of bounded resilience and $k-$resilience variants are indicated in blue.}
+	\end{figure}
+\end{center}
 
-- detailed complexity analysis (in VASS \& variations)
+% How to expand current results ?
 
-- controller/environment framework
+% - detailed complexity analysis (in VASS \& variations)
 
-- other classes of $\Safe / \Bad$ (e.g. semilinear)
+% - other classes of $\Safe / \Bad$ (e.g. semilinear)
 

### 2024-01-11 diapo/presentation.tex
up

@@
 
+\begin{columns}[T]
+ \column{0.6\textwidth}
+
 $\Safe_1 = \{q_0(n) \mid n \in \N\}$
 
-\phantom{ {resilience},  {$k$-resilience} and {bounded resilience} are not satisfied }
+\vspace{0.1cm}
 
@@
 
-\phantom{ {resilience} hold, {bounded resilience} does not hold }
+ \column{0.4\textwidth}
+
+\phantom{ {resilience} }
+
+\vspace{0.1cm}
+
+\phantom{ {resilience} }
+
+\end{columns}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s)}
+  \begin{frame}{Resilience problems}
 
@@
 
- {\small
-
-{\sc Resilience problems}
-
-\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$.}
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$, \\ $\Bad$ complement of $\Safe$.}
 
@@
 
-\hspace{-0.5cm}  ({\sc resilience problem}) $\Bad \rightarrow^{*} \Safe$ ?
+\begin{columns}[T]
+ \column{0.3\textwidth}
 
-\hspace{-0.5cm}  ({\sc $k$-resilience problem})		$\Bad \rightarrow^{\leq k} \Safe$ ?
+  $\Bad \rightarrow^{*} \Safe$  ?
 
-\hspace{-0.5cm}  ({\sc bounded resilience problem})	$\exists k ~ S \rightarrow^{\leq k} \Safe$ ?\newline
+\vspace{0.15cm}
 
-}
+   $\Bad \rightarrow^{\leq k} \Safe$  ?
+
+\vspace{0.15cm}
+
+  $\exists k ~ \Bad \rightarrow^{\leq k} \Safe$   ?\newline
+
+ \column{0.7\textwidth}
+
+\hspace{0.6cm} {\sc  \phantom{bounded} resilience problem}
+
+\vspace{0.16cm}
+
+\hspace{0.6cm} {\sc \phantom{boundi} $k$-resilience problem}
+
+\vspace{0.175cm}
+
+\hspace{0.6cm} {\sc bounded resilience problem}
+
+\end{columns}
 
@@
 
+\only<1>{
+\begin{columns}[T]
+ \column{0.6\textwidth}
+
 $\Safe_1 = \{q_0(n) \mid n \in \N\}$
 
-{resilience}, 
-{$k$-resilience} and {bounded resilience} are not satisfied
+ \column{0.4\textwidth}
+
+ Resilience(s) not satisfied
+
+\vspace{0.1cm}
+
+\phantom{bla} $q_1(m) ~ \cancel{\to} ~ q_0(n)$
+
+\end{columns}
+
+}
+
+\only<2>{
+\begin{columns}[T]
+ \column{0.6\textwidth}
 
@@
 
- {resilience} hold, {bounded resilience} does not hold
+ \column{0.4\textwidth}
+
+ Resilience hold,
+
+ \vspace{0.1cm}
+
+ \hspace{-0.6cm} Bounded resilience does not
+
+\end{columns}
+
+}
 
@@
 
+\begin{columns}[T]
+ \column{0.6\textwidth}
+
 $\Safe_2 = \{q_1(0)\} $
-   {bounded resilience} hold, $6$-resilience hold
+
+ \column{0.4\textwidth}
+
+ Bounded resilience hold,
+
+ \vspace{0.1cm}
+
+ \phantom{bla} $6$-resilience hold
+
+\end{columns}
+
+$q_1(9) \to q_0(10) \to q_2(0) \to q_1(0)$
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s)}
+  \begin{frame}{{State} resilience problems}
 
-  \hspace{-0.2cm}
-{\sc \textcolor{red}{State} resilience problems}
+% TODO
 
-\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, \textcolor{red}{$s \in S$}, and a set $\Safe \subseteq S$.}
+% Define the Six different problems we study/focus on.
+
+% (Takes some room but I think mb having the six on the same page could be better than having three then three state-* on two separate pages ?)
+
+\hspace{-0.5cm}  {\bf INPUT:\ }{A transition system $\mathscr{S}=(S,\rightarrow)$, and a set $\Safe \subseteq S$, \\ $\Bad$ complement of $\Safe$, \textcolor{red}{$s \in S$}.}
 
@@
 
-\hspace{-0.5cm}{ ({\sc state-resilience problem}) $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$ ?
+\begin{columns}[T]
+ \column{0.45\textwidth}
 
-\hspace{-0.5cm}  ({\sc $k$-state-resilience problem})		$\cancel{\Bad}~ \textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$ ?
+  $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{*} \Safe$  ?
 
-\hspace{-0.5cm} ({\sc bounded-state-resilience problem})	$\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s)} \rightarrow^{\leq k} \Safe$?\newline}
+\vspace{0.15cm}
+
+   $\cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{\leq k} \Safe$  ?
+
+\vspace{0.15cm}
+
+  $\exists k ~ \cancel{\Bad}~\textcolor{red}{\post^*(s) } \rightarrow^{\leq k} \Safe$   ?\newline
+
+ \column{0.7\textwidth}
+
+\hspace{0.6cm} {\sc  \phantom{bounded} state-resilience problem}
+
+\vspace{0.19cm}
+
+\hspace{0.6cm} {\sc \phantom{boundi} $k$-state-resilience problem}
+
+\vspace{0.20cm}
+
+\hspace{0.6cm} {\sc bounded-state-resilience problem}
+
+\end{columns}
 
@@
 
-  \begin{definition}[\cite{DBLP:journals/lmcs/BlondinFM17}]
+  \begin{block}{Definition~\cite{DBLP:journals/lmcs/BlondinFM17}}
 A {\em Well-Behaved Transition System} (WBTS) 
-is a FAC$^1$-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
+is an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
+all antichains of $(S, \leq)$ are finite, and
+%the transition relation 
+$ \rightarrow$ is (upward) compatible with $\leq$. 
+%$s_1, t_1 , s_2 \in S$
+%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+%	$t_2 \in S$ with 
+%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
+\end{block}
+
+\begin{block}{Definition~\cite{DBLP:journals/iandc/Finkel90}}
+A {\em Well-Structured Transition System} (WSTS) 
+is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ with 
+ $ \rightarrow$ is (upward) compatible with $\leq$. 
 %the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
@@
 %	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
-\end{definition}
-
-   \begin{center}
- 	\begin{figure}
- 	% \hspace{-2.cm}
-\includegraphics[width=.25\textwidth]{WSTS_def}
-	\end{figure}
-\end{center}
-
-\phantom{line}
-
-{\small $^1${FAC = all anti-chains are finite}}
-
-    \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}
-% TODO
-%  \todo[inline,color=blue!20]{ {\it À l'oral~} 
-% Resilience(s) "obviously"  undecidable in general so restriction $\to$ restriction to WSTS
-% $\to$ so we need to define WSTS properly }
-
-\begin{definition}[\cite{DBLP:journals/iandc/Finkel90}]
-A {\em Well-Structured Transition System} (WSTS) 
-is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ such that   
-%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
-%$s_1, t_1 , s_2 \in S$
-%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
-%	$t_2 \in S$ with 
-%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
-\end{definition}
+\end{block}
 
@@
 \begin{itemize}
-\item $\post$-ideal-effective =effective + $\mathop{\downarrow} \post(I)$ computable %$\post$-ideal-effective 
+\item effective + $\mathop{\downarrow} \post(I)$ computable$^1$ %$\post$-ideal-effective 
 \item \textcolor{red}{strong} compatibility
@@
 \end{center}
+
+\phantom{Footnote}
+
+{\small $^1$called $\post$-ideal-effective }
 
@@
  \begin{definition}[\cite{BFM-ic17}]
-The \emph{completion}   of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
+The \emph{completion}   of a WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
 %$Ideals(S)$ is the set of ideals of $S$ and
@@
 
-\begin{exampleblock}{Relation with paths in the completion}
+\begin{exampleblock}{From a WBTS to its completion}
 If $x \xrightarrow{k} y$ in $\mathscr{S}$ then for every ideal $I \supseteq \mathop{\downarrow} x$, there exists an ideal $J \supseteq \mathop{\downarrow} y$ such that $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$.
@@
 
-\begin{exampleblock}{Relation with paths in the completion}
+\begin{exampleblock}{From the completion to the WBTS}
 If $I \xrightarrow{k} J$ in $\hat{\mathscr{S}}$ then for every $y \in J$, there exists $x \in I$ and $y' \geq y$ such that $x \xrightarrow{k'} y'$ in $\mathscr{S}$.
@@
 
-\begin{theorem}\label{SL VASS}
+\begin{block}{Theorem}
 {\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
-\end{theorem}
+\end{block}
 
@@
 
-\begin{theorem}\label{vass down}
+\begin{block}{Theorem}
 {\sc Bounded resilience} and {\sc $k$-resilience } are decidable for VASS when 
 $\Safe = \mathop{\downarrow} \Safe$.
-\end{theorem}
+\end{block}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
+  \begin{frame}{References}
+
+      {\tiny
+\bibliography{bib}
+}
+
+  \end{frame}
+%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
   \begin{frame}
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{References}
-
-      {\tiny
-\bibliography{bib}
-}
-
-  \end{frame}
-%vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
 

### 2024-01-11 diapo/presentation.tex
up

@@
  	\vspace{.06cm}
-\includegraphics[width=.47\textwidth]{FigA}
+\includegraphics[width=.4\textwidth]{FigA}
 	\end{figure}
@@
 % 	\hspace{-2.cm}
-\includegraphics[width=.47\textwidth]{FigB}
+\includegraphics[width=.4\textwidth]{FigB}
 	\end{figure}
@@
 
-$q_1(9) \to q_0(10) \to q_2(0) \to q_1(0)$
+ \vspace{0.1cm}
+
+$q_1(7) \to q_0(8) \to q_0(9) \to q_0(10) \to q_2(0) \to q_1(0)$
+
+$q_1(6) \to q_1(5) \to q_1(4) \to q_1(3) \to q_1(2) \to q_1(1) \to q_1(0)$
 
@@
 %the transition relation 
-$ \rightarrow$ is (upward) compatible with $\leq$. 
-%$s_1, t_1 , s_2 \in S$
-%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
-%	$t_2 \in S$ with 
-%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
-\end{block}
-
-\begin{block}{Definition~\cite{DBLP:journals/iandc/Finkel90}}
-A {\em Well-Structured Transition System} (WSTS) 
-is a well-ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ with 
- $ \rightarrow$ is (upward) compatible with $\leq$. 
-%the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
+$ \rightarrow$ is upward-compatible with $\leq$. 
 %$s_1, t_1 , s_2 \in S$
@@
 \end{center}
+
+\begin{block}{Definition~\cite{DBLP:journals/iandc/Finkel90}}
+A {\em Well-Structured Transition System} (WSTS) 
+is a well-ordered WBTS.
+% ward) compatible with $\leq$, i.e., for all 
+%$s_1, t_1 , s_2 \in S$
+%	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
+%	$t_2 \in S$ with 
+%	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
+\end{block}
 
@@
 
+\nocite{DBLP:conf/gg/Ozkan22}
+
   \end{frame}

### 2024-01-12 diapo/presentation.tex
up

@@
 
- \vspace{0.1cm}
+ \vspace{0.25cm}
 
+ \only<2>{
 $q_1(7) \to q_0(8) \to q_0(9) \to q_0(10) \to q_2(0) \to q_1(0)$
+}
 
+\only<1>{
 $q_1(6) \to q_1(5) \to q_1(4) \to q_1(3) \to q_1(2) \to q_1(1) \to q_1(0)$
+}
 
@@
 
- \begin{definition}[\cite{BFM-ic17}]
+ \begin{block}{Definition~\cite{BFM-ic17}}
 The \emph{completion}   of a WBTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where 
@@
 %It is proved in  \cite{BFM-ic17} that $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS. \\
-\end{definition}
+\end{block}
 
@@
 %vvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvvv
-  \begin{frame}{Resilience(s) for VASS}
+  \begin{frame}{Vector Addition System with States}
 
-  Remettre un déssin pour un VASS
+   \begin{center}
+ 	\begin{figure}
+ 	\vspace{.06cm}
+\includegraphics[width=.4\textwidth]{FigC}
+	\end{figure}
+\end{center}
 

### 2024-05-30 LMCS/final.tex
Adding version LMCS

--- 2023-11-03 VMCAI-revised-version/final.tex
@@
 %
-\documentclass[runningheads]{llncs}
+\documentclass{lmcs}
 %
@@
 
-\usepackage{complexity}
+% \usepackage{complexity}
 \usepackage{amsmath}
@@
 \newtheorem{subclaim}{Subclaim}{}
-\newtheorem{fact}{Fact}{}
+% \newtheorem{fact}{Fact}{}
 
@@
 
-\renewcommand{\poly}{\mathrm{poly}}
+\newcommand{\poly}{\mathrm{poly}}
+\newcommand{\EXPSPACE}{\mathsf{EXPSPACE}}
 
@@
 %
-\title{Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}} 
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{Alain Finkel\inst{1,2} \and Mathieu Hilaire\inst{1}}
-% This research has been supported by ANR programme BraVAS (ANR-17-CE40-0028).
-%
-\authorrunning{A. Finkel, M. Hilaire.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Université Paris-Saclay, CNRS, ENS Paris-Saclay, LMF, Gif-sur-Yvette, France %\email{hilaire@lsv.fr}
-\and
-Institut Universitaire de France }
-
-%
-\maketitle              % typeset the header of the contribution
-%
+\author{Alain Finkel%\inst{1,2} \and 
+Mathieu Hilaire%\inst{1}
+}
+
+\title[Resilience and Home-Space for WSTS]{Resilience and Home-Space for WSTS}
+% \titlecomment{{\lsuper*}OPTIONAL comment concerning the title, \eg, if a variant or an extended abstract of the paper has appeared elsewhere.}
+\thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}	%optional
+
+% affiliations are numbered automatically with a, b, c (see below)
+% use the optional argument to indicate the affiliation(s) of each author
+% omit the argument if there is only one author, or only one affiliation
+\author[A.~Finkel]{Alain Finkel}[a,b]
+\author[M.~Hilaire]{Mathieu Hilaire}[a]
+
+% affiliation 1 (automatically numbered a)
+\address{University 1, address1}	%optional
+% write emails for all authors having that affiliation
+\email{name1@email1, name2@email1, name3@email1}  %optional
+
+% affiliation 2 (automatically numbered b)
+\address{University 2, address2}	%optional
+\email{name2@email2}  %optional
+
+%% etc.
 
@@
 
+\maketitle
+
 \newcommand{\LCM}{\mathsf{LCM}}
@@
 
+%% start the paper here:
 \section{Introduction}\label{section introduction}
@@
 %
-\item In particular, the three resilience problems are decidable for post-ideal-effective (Definition \ref{eff}) Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem \ref{down-up}).
+\item In particular, the three resilience problems are decidable for post-ideal-effective (Definition~\ref{eff}) Well Behaved Transition Systems (a generalisation of WSTS where the quasi-ordering is not necessarly well founded) with strong (upward) compatibility and when $\Safe$ is upward-closed (Theorem~\ref{down-up}).
 Resilience is also decidable, when
-$\Safe$ is upward-closed, for effective pred-basis (Definition \ref{eff}) WSTS with strong (upward) compatibility (Theorem \ref{xxx}).
+$\Safe$ is upward-closed, for effective pred-basis (Definition \ref{eff}) WSTS with strong (upward) compatibility (Theorem~\ref{xxx}).
 %
@@
 %
-\item We generalize with Theorem \ref{post srp} the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability (Proposition \ref{srp up down}).
-
-\item We show that the three state-resilience problems are decidable for post-ideal-effective WBTS with downward and upward compatibilities and $\Safe$ downward-closed (Theorem \ref{downward srp} extends \cite[Theorem 1]{DBLP:journals/corr/PrasadZ16}).
+\item We generalize with Theorem~\ref{post srp} the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we show that relaxing some hypothesis leads to undecidability (Proposition \ref{srp up down}).
+
+\item We show that the three state-resilience problems are decidable for post-ideal-effective WBTS with downward and upward compatibilities and $\Safe$ downward-closed (Theorem~\ref{downward srp} extends \cite[Theorem 1]{DBLP:journals/corr/PrasadZ16}).
 The two other types of resilience, {$k$-state-resilience} and {bounded-state-resilience}, are decidable for post-ideal-effective WBTS with strong downward compatibility.
@@
 Let us now recall the (most general) definition of well-structured transition systems.
-\begin{definition}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
+\begin{defi}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
 A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
-\end{definition}
+\end{defi}
 
@@
 
-\begin{definition}
+\begin{defi}
 A wqo $(X, \leq)$ is an \emph{$\omega^2$-wqo} if $(Ideals(X), \subseteq)$ forms a wqo.
-\end{definition}
+\end{defi}
 
@@
 
-\begin{definition}(Finkel and Goubault-Larrecq \cite{DBLP:journals/corr/abs-1208-4549})
+\begin{defi}(Finkel and Goubault-Larrecq \cite{DBLP:journals/corr/abs-1208-4549})
 A WSTS  $\mathscr{S}=(S, \rightarrow, \leq)$
 is an $\omega^2$-WSTS if $(S, \leq)$ is an $\omega^2$-wqo.
-\end{definition}
+\end{defi}
 Since almost of usual wqo are $\omega^2$-wqo, all naturally occurring WSTS are in fact $\omega^2$-WSTS: for example, reset/transfer VASS are $\omega^2$-WSTS.\\
@@
 %
-\begin{definition}
+\begin{defi}
 A {\em downward (compatible) WBTS (short, downward-WBTS)} $\mathscr{S}=(S, \rightarrow, \leq)$
@@
 A  {\em downward (compatible) WSTS (short, downward-WSTS)} \cite[Definition 5.1]{DBLP:journals/tcs/FinkelS01} is a downward-WBTS where $\leq$ is a wqo.
-\end{definition}
+\end{defi}
 
@@
 
-\begin{lemma}\label{downward compatible}
+\begin{lem}\label{downward compatible}
 For an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ (not necessarly a WSTS), the two following properties are equivalent: (1) $\mathscr{S}=(S, \rightarrow, \leq)$ is downward compatible ; (2) for every downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
-\end{lemma}
+\end{lem}
 
@@
 %
-\begin{definition}\label{eff}
+\begin{defi}\label{eff}
 %
@@
 %
-\end{definition}
+\end{defi}
 %
@@
 %
-\begin{theorem}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
+\begin{thm}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
  A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis.
 Hence coverability is decidable.
-\end{theorem}
+\end{thm}
 
@@
 %
-\begin{theorem}[Corollaire 4.4, ~\cite{DBLP:journals/lmcs/BlondinFM17}]
+\begin{thm}[Corollaire 4.4, ~\cite{DBLP:journals/lmcs/BlondinFM17}]
 Coverability is decidable for any $\post$-ideal-effective WBTS.
-\end{theorem}
+\end{thm}
 
@@
 %
-\begin{theorem}[Proposition 5.4, \cite{DBLP:journals/tcs/FinkelS01}]
+\begin{thm}[Proposition 5.4, \cite{DBLP:journals/tcs/FinkelS01}]
 Finitely branching downward-WSTS with reflexive compatibility and $\post$-effective are $\post^*$-basis-effective.
-\end{theorem}
+\end{thm}
 %
@@
 %
-%\begin{theorem} \cite{BFM-ic17}
+%\begin{thm} \cite{BFM-ic17}
 %Coverability is decidable for ideally-effective WSTS.
-%\end{theorem}
+%\end{thm}
 
 Let us recall the  definition of vector addition system with (control-)states. 
- \begin{definition} 
+ \begin{defi} 
 A {\em vector addition system with (control-)states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}.
- \end{definition} 
+ \end{defi} 
 %
@@
 
-\begin{definition}
+\begin{defi}
 A {\em reset-VASS} in dimension $d$ 
@@
 		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
-\end{definition}
+\end{defi}
 
@@
 %
-\begin{proposition}\label{indec WSTS}
+\begin{prop}\label{indec WSTS}
 {\sc Resilience} and {\sc Bounded resilience} for $\post$-effective $\omega^2$-WSTS with strong compatibility, for upward closed sets $\Safe$ are undecidable.
-\end{proposition}
+\end{prop}
 \begin{proof}
@@
 
-\begin{proposition}\label{indec WSTS with dcs}
+\begin{prop}\label{indec WSTS with dcs}
 {\sc Resilience} is undecidable for reset-VASS, hence it is also undecidable for $\post$-effective and $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility and with 
 $\Safe=\mathop{\downarrow} \Safe$.
-\end{proposition}
+\end{prop}
 
@@
 %
-\begin{theorem}\label{down-up}
+\begin{thm}\label{down-up}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\post$-ideal-effective WBTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.
@@
 and {\sc $k$-resilience} are decidable.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{theorem}\label{xxx}
+\begin{thm}\label{xxx}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS and a set $\Safe = \mathop{\uparrow} \Safe$.
  {\sc Resilience} is decidable.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{corollary}
+\begin{cor}
 Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a $\pred$-basis-effective WSTS with strong compatibility and a set $\Safe = \mathop{\uparrow} \Safe$.  {\sc Bounded resilience} 
 and {\sc $k$-resilience} are decidable.
-\end{corollary}
+\end{cor}
 
@@
 %
-\begin{theorem}\label{up-down}
+\begin{thm}\label{up-down}
 {\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WBTS with 
 $\Safe=\mathop{\downarrow} \Safe$.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{proposition}\label{reductions}
+\begin{prop}\label{reductions}
 {\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
-\end{proposition}
+\end{prop}
 
@@
 .
-\begin{proposition}\label{liveness reset}
+\begin{prop}\label{liveness reset}
 {\sc zero-reachability} can be reduced to {\sc $t$-liveness}.
-\end{proposition}
+\end{prop}
 
@@
 
-\begin{proposition}\label{srp up down}
+\begin{prop}\label{srp up down}
 {\sc State-resilience},
@@
 $\Safe=\mathop{\uparrow} \Safe$.
-\end{proposition}
+\end{prop}
 
@@
 %
-\begin{theorem}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}
+\begin{thm}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]\label{ref ozkan}
 {\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
@@
 $\Safe=\mathop{\uparrow} \Safe$.
-\end{theorem}
+\end{thm}
 The proof of Theorem~\ref{ref ozkan} rely on the computability of $\mathop{\uparrow} \post^*(s)$ and on the following lemma.
 
-\begin{lemma}\label{Lemma intersection}
+\begin{lem}\label{Lemma intersection}
 Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
 Then $A \cap D \subseteq U$  iff $ (\mathop{\uparrow}  A) \cap D \subseteq U$.
-\end{lemma}
+\end{lem}
 
@@
 
-\begin{proposition}[Proposition 1, \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
+\begin{prop}[Proposition 1, \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
 For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of 
 $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
-\end{proposition}
+\end{prop}
 
@@
 
-\begin{theorem}\label{post srp}
+\begin{thm}\label{post srp}
 {\sc State-resilience} is decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS
@@
 $\Safe=\mathop{\uparrow} \Safe$. 
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{proposition}\label{srp down up}
+\begin{prop}\label{srp down up}
 {\sc State-resilience} is undecidable for reset-VASS when
@@
  $\pred$-basis-effective $\omega^2$-WSTS with strong compatibility.
- \end{proposition}
+ \end{prop}
 
@@
 
-\begin{lemma}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
+\begin{lem}(Symmetrical from Lemma~\ref{Lemma intersection})\label{Lemma intersection 2}
 Let $A \subseteq S$, $D \subseteq S$ be a downward-closed set and $U \subseteq S$ be an upward-closed set. 
 Then $A \cap U \subseteq D$  iff $ (\mathop{\downarrow}  A) \cap U \subseteq D$.
-\end{lemma}
+\end{lem}
 
@@
 
-\begin{theorem}\label{downward srp}
+\begin{thm}\label{downward srp}
 {\sc State-resilience} is decidable for $\post$-effective, 
@@
 and $\Safe = \mathop{\downarrow} \Safe$.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{corollary}\label{downward brp}
+\begin{cor}\label{downward brp}
 {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WBTS, strong downward compatibility and upward compatibility when
@@
  $\mathop{\downarrow} \pred(I)$.
-\end{corollary}
+\end{cor}
 
@@
 Surprisingly, when $\Safe$ is downward-closed, we have the following result for VAS (not VASS):
-\begin{proposition}
+\begin{prop}
 {\sc Bounded resilience} and {\sc $k$-resilience} never hold for $d$-VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Safe \neq \N^d$.
-\end{proposition}
+\end{prop}
 
@@
  which lead to the following decidability result:
-\begin{theorem}\label{vass down}
+\begin{thm}\label{vass down}
 {\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
 $\Safe = \mathop{\downarrow} \Safe$.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{theorem}\label{SL VASS}
+\begin{thm}\label{SL VASS}
 {\sc Resilience} is decidable for VASS, and $\mathbb{Z}-$VASS, when $\Safe$  is a semilinear set.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{theorem}\label{SL Lossy}
+\begin{thm}\label{SL Lossy}
 {\sc Resilience} is decidable for lossy counter machines when $\Safe$  is a semilinear set.
-\end{theorem}
+\end{thm}
 
@@
 
-\begin{theorem}\label{RP VASS}
+\begin{thm}\label{RP VASS}
 {\sc Resilience} is decidable for continuous VASS when $\Safe$  is
 definable in the existential theory of the rationals with addition and order.
-\end{theorem}
+\end{thm}
 
